\NormalFont\ShortTitle{1 Corinthians}
{\MT 1 Corinthians

\par }\ChapOne{1}{\SH Salutation
\par }{\PP \VerseOne{1}From Paul,
called
to be an apostle
of Christ
Jesus
by
the will
of God,
and
Sosthenes,
our brother,
\VS{2}to the church
of God
that is
in
Corinth,
to those who are sanctified
in
Christ
Jesus,
and called
to be saints,
with
all
those in
every
place
who call
on the name
of our
Lord
Jesus
Christ,
their
Lord and
ours.
\VS{3}Grace
and
peace
to you
from
God
our
Father
and
the Lord
Jesus
Christ!
\par }{\SH Thanksgiving
\par }{\PP \VS{4}I
always
thank
my God
for
you
because of
the grace
of God
that was given
to you
in
Christ
Jesus.
\VS{5}For
you were made rich
in
every way
in
him,
in
all
your speech
and
in every kind
of knowledge –
\VS{6}just as
the testimony
about Christ
has been confirmed
among
you –
\VS{7}so that
you
do
not
lack
any
spiritual gift
as you wait
for the revelation
of our
Lord
Jesus
Christ.
\VS{8}He will
also
strengthen
you
to
the end,
so that you will be blameless
on
the day
of our
Lord
Jesus
Christ.
\VS{9}God
is faithful,
by
whom
you were called
into
fellowship
with his
son,
Jesus
Christ
our
Lord.
\par }{\SH Divisions in the Church
\par }{\PP \VS{10}I urge
you,
brothers and sisters,
by
the name
of our
Lord
Jesus
Christ,
to
agree together, to end your
divisions,
and
to be united
by
the same
mind
and
purpose.
\VS{11}For
members of
Chloe’s household
have made
it clear
to me,
my
brothers and sisters,
that
there are
quarrels
among
you.
\VS{12}Now
I mean
this,
that
each
of you
is saying,
“I
am
with Paul,”
or “I
am with Apollos,”
or
“I
am with Cephas,”
or
“I
am with Christ.”
\VS{13}Is Christ
divided? Paul
wasn’t
crucified
for
you,
was he? Or
were you
in fact baptized
in
the name
of Paul?
\VS{14}I thank God
that
I did
not
baptize
any of you except
Crispus
and
Gaius,
\VS{15}so that
no
one
can say
that
you were baptized
in
my
name!
\VS{16}(I
also
baptized
the household
of Stephanus.
Otherwise,
I do
not
remember
whether
I baptized
anyone
else.)
\VS{17}For
Christ
did
not
send
me
to baptize,
but
to preach the gospel
– and not
with
clever
speech,
so that
the cross
of Christ
would
not
become useless.
\par }{\SH The Message of the Cross
\par }{\PP \VS{18}For
the message
about the cross
is
foolishness
to those who are perishing,
but
to us
who are being saved
it is
the power
of God.
\VS{19}For
it is written, “{\QT{
I will destroy
the wisdom
of the wise, and
I will thwart
the cleverness
of the intelligent}}.”
\VS{20}Where
is the wise
man? Where
is the expert in the Mosaic law? Where
is the debater
of this
age? Has God
not
made
the wisdom
of the world
foolish?
\VS{21}For
since
in
the wisdom
of God
the world
by
its wisdom
did not
know
God,
God
was pleased
to save
those who believe
by
the foolishness
of preaching.
\VS{22}For
Jews
demand
miraculous signs
and
Greeks
ask for
wisdom,
\VS{23}but
we
preach
about a crucified
Christ,
a stumbling block
to Jews
and
foolishness
to Gentiles.
\VS{24}But
to those
who are called,
both
Jews
and
Greeks,
Christ
is the power
of God
and
the wisdom
of God.
\VS{25}For
the foolishness
of God
is
wiser than
human wisdom,
and
the weakness
of God
is stronger than
human strength.
\par }{\PP \VS{26}Think about the circumstances of your
call,
brothers and sisters.
Not
many
were wise
by
human standards,
not
many
were powerful,
not
many
were born to a privileged position.
\VS{27}But
God
chose
what the world
thinks foolish
to
shame
the wise,
and
God
chose
what the world
thinks weak
to
shame
the strong.
\VS{28}God
chose
what is low
and
despised
in the world,
what is regarded as nothing, to set aside what is regarded as something,
\VS{29}so that
no
one can boast
in his presence.
\VS{30}He
is the reason you
have a relationship with
Christ
Jesus,
who
became
for us
wisdom
from
God,
and
righteousness
and
sanctification
and
redemption,
\VS{31}so that,
as
it is written, “{\QT{
Let the one who boasts, boast
in
the Lord}}.”

\par }\Chap{2}{\PP \VerseOne{1}When
I
came
to
you,
brothers and sisters,
I did
not
come
with
superior
eloquence
or
wisdom
as
I proclaimed
the testimony
of God.
\VS{2}For
I decided
to be concerned
about nothing
among
you
except
Jesus
Christ,
and
him
crucified.
\VS{3}And I
was
with
you
in
weakness
and
in
fear
and
with
much
trembling.
\VS{4}My
conversation
and
my
preaching
were not
with
persuasive
words
of wisdom,
but
with
a demonstration
of the Spirit
and
of power,
\VS{5}so
that your
faith
would
not
be based on
human
wisdom
but
on
the power
of God.
\par }{\SH Wisdom from God
\par }{\PP \VS{6}Now
we do speak
wisdom
among
the mature,
but
not
a wisdom
of this
age
or
of the rulers
of this
age,
who are perishing.
\VS{7}Instead
we speak
the wisdom
of God,
hidden
in
a mystery,
that
God
determined
before
the ages
for
our
glory.
\VS{8}None
of the rulers
of this
age
understood
it. If
they had known
it, they would
not
have crucified
the Lord
of glory.
\VS{9}But
just as
it is written, “{\QT{
Things that
no
eye
has seen, or
ear
heard, or
mind
imagined,
are the things God
has prepared
for those who love
him}}.”
\VS{10}God
has revealed
these to us
by
the Spirit.
For
the Spirit
searches
all things,
even
the deep
things of God.
\VS{11}For
who
among men
knows
the things of a man
except
the man’s
spirit
within
him? So
too,
no one
knows
the things of God
except
the Spirit
of God.
\VS{12}Now
we
have
not
received
the spirit
of the world,
but
the Spirit
who is from
God,
so that
we may know
the things that are freely given
to us
by
God.
\VS{13}And
we speak
about these things, not
with words taught
us by human
wisdom,
but
with
those taught
by the Spirit,
explaining
spiritual
things to spiritual people.
\VS{14}The unbeliever
does
not
receive
the things of the Spirit
of God,
for
they are foolishness
to him.
And
he cannot
understand
them, because
they are
spiritually
discerned.
\VS{15}The one who is spiritual
discerns
all things,
yet
he
himself
is understood
by
no one.
\VS{16}{\QT{
For
who
has known
the mind
of the Lord, so as to advise
him?}}
But
we
have
the mind
of Christ.

\par }\Chap{3}{\PP \VerseOne{1}So, brothers and sisters,
I could
not
speak
to you
as
spiritual
people, but
instead as
people of the flesh,
as
infants
in
Christ.
\VS{2}I fed
you
milk,
not
solid food,
for
you were
not yet
ready.
In fact, you are
still
not
ready,
\VS{3}for
you are
still
influenced by the flesh.
For
since
there is still jealousy
and
dissension
among
you,
are you
not
influenced by the flesh
and
behaving
like unregenerate people?
\VS{4}For
whenever
someone
says,
“I
am
with Paul,”
or
“I
am with Apollos,”
are you
not
merely human?
\par }{\PP \VS{5}What
is
Apollos,
really? Or
what
is
Paul? Servants
through
whom
you came to believe,
and
each
of us in the ministry the Lord
gave us.
\VS{6}I
planted,
Apollos
watered,
but
God
caused
it to grow.
\VS{7}So
neither
the one who plants
counts
for anything,
nor
the one who waters,
but
God
who causes the growth.
\VS{8}The one who plants
and
the one who waters
work as one,
but
each
will receive
his
reward
according
to his
work.
\VS{9}We are
coworkers
belonging
to God.
You are
God’s
field,
God’s
building.
\VS{10}According to
the grace
of God
given
to me,
like
a skilled
master-builder
I laid
a foundation,
but
someone else
builds
on it. And
each one
must be careful
how
he builds.
\VS{11}For
no one
can
lay
any foundation
other
than
what is being laid,
which
is
Jesus
Christ.
\VS{12}If
anyone
builds
on
the foundation
with gold,
silver,
precious
stones,
wood,
hay,
or straw,
\VS{13}each
builder’s work
will be
plainly seen,
for
the Day
will make
it clear,
because
it will be revealed
by
fire.
And
the fire
will test
what kind
of work
each
has done.
\VS{14}If
what someone
has built
survives,
he will receive
a reward.
\VS{15}If
someone’s
work
is burned up,
he will suffer loss.
He
himself
will be saved,
but
only
as
through
fire.
\par }{\PP \VS{16}Do you
not
know
that
you are
God’s
temple
and
that God’s
Spirit
lives
in
you?
\VS{17}If
someone
destroys
God’s
temple,
God
will destroy
him.
For
God’s
temple
is
holy,
which
is what you
are.
\par }{\PP \VS{18}Guard against self-deception,
each of you. If
someone
among
you
thinks
he is
wise
in
this age,
let
him
become
foolish
so that
he can become
wise.
\VS{19}For
the wisdom
of this
age
is
foolishness
with
God.
As it is written, “{\QT{
He catches
the wise
in
their
craftiness}}.”
\VS{20}And
again, “{\QT{
The Lord
knows
that the thoughts
of the wise
are
futile}}.”
\VS{21}So then,
no more
boasting
about
mere mortals! For
everything
belongs
to you,
\VS{22}whether
Paul
or
Apollos
or
Cephas
or
the world
or
life
or
death
or
the present
or
the future.
Everything belongs to you,
\VS{23}and
you
belong to Christ,
and
Christ
belongs to God.

\par }\Chap{4}{\PP \VerseOne{1}One
should think
about us
this way
– as
servants
of Christ
and
stewards
of the mysteries
of God.
\VS{2}Now
what is sought
in
stewards
is that
one
be found
faithful.
\VS{3}So
for me,
it is
a minor matter
that
I am judged
by
you
or
by
any human court.
In fact, I do
not even
judge
myself.
\VS{4}For
I am
not
aware
of anything against myself,
but
I am
not
acquitted
because of
this.
The one who judges
me
is
the Lord.
\VS{5}So then,
do
not
judge
anything
before
the time.
Wait until
the Lord
comes.
He will bring to light
the hidden things
of darkness
and
reveal
the motives
of hearts.
Then
each
will receive
recognition
from
God.
\par }{\PP \VS{6}I have applied
these things
to
myself
and
Apollos
because of
you,
brothers and sisters,
so that
through
us
you may learn
“not
to go beyond
what
is written,”
so that
none
of you will be puffed up
in
favor
of the one
against
the other.
\VS{7}For
who
concedes
you
any superiority? What
do you have
that
you did
not
receive? And
if
you received
it, why
do you boast
as though
you did
not?
\VS{8}Already
you are
satisfied! Already
you are rich! You have become kings
without
us! I wish
you had become kings
so that
we
could reign
with you!
\VS{9}For,
I think,
God
has exhibited
us
apostles
last
of all, as
men condemned to die,
because
we have become
a spectacle
to the world,
both
to angels
and
to people.
\VS{10}We
are fools
for
Christ,
but
you
are wise
in
Christ! We
are weak,
but
you
are strong! You
are distinguished,
we
are dishonored!
\VS{11}To
the present
hour
we are hungry
and
thirsty,
poorly clothed,
brutally treated,
and
without a roof over
our heads.
\VS{12}We do hard work,
toiling
with our own
hands.
When we are verbally abused,
we respond with a blessing,
when persecuted,
we endure,
\VS{13}when people lie
about us, we answer in a friendly manner.
We are
the world’s
dirt
and scum,
even
now.
\par }{\SH A Father’s Warning
\par }{\PP \VS{14}I am
not
writing
these things
to shame
you,
but
to correct
you as
my
dear
children.
\VS{15}For
though
you may have
ten thousand
guardians
in
Christ,
you do not
have many
fathers,
because
I
became
your
father
in
Christ
Jesus
through
the gospel.
\VS{16}I encourage
you,
then,
be
imitators
of me.
\VS{17}For
this reason,
I have sent
Timothy
to you,
who
is
my
dear
and
faithful
son
in
the Lord.
He will remind
you
of my
ways
in
Christ,
as
I teach
them everywhere
in
every
church.
\VS{18}Some
have become arrogant,
as
if I
were
not
coming
to
you.
\VS{19}But
I will come
to
you
soon,
if
the Lord
is willing,
and
I will find out
not
only the talk
of these arrogant people,
but
also their power.
\VS{20}For
the kingdom
of God
is demonstrated not
in
idle talk
but
with
power.
\VS{21}What
do you want? Shall I come
to
you
with
a rod
of discipline or
with
love
and
a spirit
of gentleness?

\par }\Chap{5}{\PP \VerseOne{1}It is
actually
reported
that sexual immorality
exists among
you,
the kind
of immorality
that
is not permitted even
among
the Gentiles,
so that
someone
is cohabiting
with his father’s
wife.
\VS{2}And
you
are
proud! Shouldn’t you have been deeply sorrowful
instead
and
removed
the one who did
this
from
among
you?
\VS{3}For
even though
I
am absent
physically,
I am present
in spirit.
And
I have
already
judged
the one who did
this,
just as
though I were present.
\VS{4}When you
gather together
in
the name
of our
Lord
Jesus,
and
I
am with you in spirit,
along
with the power
of our
Lord
Jesus,
\VS{5}turn
this man
over
to Satan
for
the destruction
of the flesh,
so that
his spirit
may be saved
in
the day
of the Lord.
\par }{\PP \VS{6}Your
boasting
is not
good.
Don’t
you know
that
a little
yeast
affects
the whole
batch of dough?
\VS{7}Clean out
the old
yeast
so that
you may be
a new
batch of dough
– you are,
in fact, without yeast.
For
Christ,
our
Passover lamb,
has been sacrificed.
\VS{8}So then,
let us celebrate the festival,
not
with
the old
yeast,
the yeast
of vice
and
evil,
but
with
the bread without yeast,
the bread of sincerity
and
truth.
\par }{\PP \VS{9}I wrote
you
in
my letter
not
to associate with
sexually immoral people.
\VS{10}In no way did I mean
the immoral people
of this
world,
or
the greedy
and
swindlers
and
idolaters,
since
you would
then
have to
go out
of
the world.
\VS{11}But
now
I am writing
to you
not
to associate with
anyone
who calls
himself a Christian
who is sexually immoral,
or
greedy,
or
an idolater,
or
verbally abusive,
or
a drunkard,
or
a swindler.
Do
not
even eat with
such a person.
\VS{12}For
what
do I
have to do with judging
those outside? Are you
not
to judge
those inside?
\VS{13}But
God
will judge
those outside.
{\IT{
Remove
the evil person
from among
you}}.

\par }\Chap{6}{\PP \VerseOne{1}When any
of you
has
a legal dispute
with
another,
does he dare
go to court
before
the unrighteous
rather
than before
the saints?
\VS{2}Or
do you
not
know
that
the saints
will judge
the world? And
if
the world
is to be judged
by
you,
are you
not competent
to settle
trivial suits?
\VS{3}Do you
not
know
that
we will judge
angels? Why not
ordinary matters!
\VS{4}So
if
you have
ordinary lawsuits,
do you appoint as judges
those
who have no standing
in
the church?
\VS{5}I say
this to
your
shame! Is there no one
among
you
wise
enough to settle disputes
between
fellow Christians?
\VS{6}Instead,
does a Christian
sue
a Christian,
and
do
this
before
unbelievers?
\VS{7}The fact that you have
lawsuits
among
yourselves
demonstrates that
you
have already
been defeated.
Why
not
rather
be wronged? Why
not
rather
be cheated?
\VS{8}But
you
yourselves
wrong
and
cheat,
and
you do this
to your brothers and sisters!
\par }{\PP \VS{9}Do you
not
know
that
the unrighteous
will
not
inherit
the kingdom
of God? Do
not
be deceived! The sexually immoral,
idolaters,
adulterers,
passive homosexual partners,
practicing homosexuals,
\VS{10}thieves,
the greedy,
drunkards,
the verbally abusive,
and swindlers
will
not
inherit
the kingdom
of God.
\VS{11}Some
of you once lived
this way.
But
you were washed,
you were sanctified,
you were justified
in
the name
of the Lord
Jesus
Christ
and
by
the Spirit
of our
God.
\par }{\SH Flee Sexual Immorality
\par }{\PP \VS{12}“All things
are lawful
for me” –
but
not
everything
is beneficial.
“All things
are lawful
for me”
– but
I will
not
be controlled
by
anything.
\VS{13}“Food
is for the stomach
and
the stomach
is for food,
but
God
will do away
with both.” The body
is not
for sexual immorality,
but
for the Lord,
and
the Lord
for the
body.
\VS{14}Now
God
indeed raised
the Lord
and
he will raise
us
by
his
power.
\VS{15}Do you
not
know
that
your
bodies
are
members
of Christ? Should I take
the members
of Christ
and make
them members
of a prostitute? Never!
\VS{16}Or
do you
not
know
that
anyone who is united
with a prostitute
is
one
body
with her? For
it is said, “{\QT{
The two
will become
one
flesh}}.”
\VS{17}But
the one united
with the Lord
is
one
spirit with him.
\VS{18}Flee
sexual immorality! “Every
sin
a person
commits
is
outside
of the body” –
but
the immoral person
sins
against
his own
body.
\VS{19}Or
do you
not
know
that
your
body
is the temple
of the Holy
Spirit
who is
in
you,
whom
you have
from
God,
and
you are
not
your own?
\VS{20}For
you were bought
at
a price. Therefore glorify
God
with
your
body.

\par }\Chap{7}{\PP \VerseOne{1}Now with regard to the issues you wrote
about: “It is good
for a man
not
to have sexual relations
with a woman.”
\VS{2}But
because of
immoralities,
each man
should have
relations with his own
wife
and
each woman
with her own
husband.
\VS{3}A husband
should give
to his wife
her sexual rights,
and
likewise
a wife
to her husband.
\VS{4}It is not
the wife
who has
the rights
to her own
body,
but
the husband.
In the same way,
it is not
the husband
who has
the rights
to his own
body,
but
the wife.
\VS{5}Do
not
deprive
each other,
except
by mutual agreement
for
a specified time,
so that
you may devote
yourselves to prayer.
Then
resume your relationship, so that
Satan
may
not
tempt
you
because of
your
lack of self-control.
\VS{6}I say
this
as
a concession,
not
as
a command.
\VS{7}I wish
that everyone
was
as
I am.
But
each
has
his own
gift
from
God,
one
this way,
another that.
\par }{\PP \VS{8}To the unmarried
and
widows
I say
that it is best
for them
to remain
as
I am.
\VS{9}But
if
they do
not
have self-control,
let them get married.
For
it is
better
to marry
than
to burn with sexual desire.
\par }{\PP \VS{10}To the married
I give
this command
– not
I,
but
the Lord –
a wife
should
not
divorce
a husband
\VS{11}(but
if
she does, let her remain
unmarried,
or
be reconciled
to her husband), and
a husband
should
not
divorce
his wife.
\par }{\PP \VS{12}To the rest
I say
– I,
not
the Lord –
if
a brother
has
a wife
who is not a believer
and
she
is happy
to live
with
him,
he should
not
divorce
her.
\VS{13}And
if a woman
has
a husband
who is not a believer
and
he
is happy
to live
with
her,
she should
not
divorce
him.
\VS{14}For
the unbelieving
husband
is sanctified
because of
the wife,
and the unbelieving
wife
because of
her husband.
Otherwise
your
children
are
unclean,
but
now
they are
holy.
\VS{15}But
if
the unbeliever
wants a divorce,
let it take place.
In
these circumstances
the brother
or
sister
is
not
bound.
God
has called
you
in
peace.
\VS{16}For
how
do you know,
wife,
whether
you will bring
your husband
to salvation? Or
how
do you know,
husband,
whether
you will bring
your wife
to salvation?
\par }{\SH The Circumstances of Your Calling
\par }{\PP \VS{17}Nevertheless,
as
the Lord
has assigned
to each one,
as
God
has called
each person, so
must he live.
I give
this sort of direction
in
all
the churches.
\VS{18}Was anyone
called
after
he had been circumcised? He should
not
try to undo his circumcision.
Was anyone
called
who is uncircumcised? He should
not
get circumcised.
\VS{19}Circumcision
is
nothing
and
uncircumcision
is
nothing.
Instead,
keeping
God’s
commandments is what counts.
\VS{20}Let
each one
remain
in
that
situation
in life in which
he was called.
\VS{21}Were you called
as a slave? Do
not
worry
about it. But
if
indeed you are able
to be
free,
make
the most
of the opportunity.
\VS{22}For
the one who was called
in
the Lord
as a slave
is
the Lord’s
freedman.
In the same way,
the one who was called
as a free person
is
Christ’s
slave.
\VS{23}You were bought
with a price.
Do
not
become
slaves
of men.
\VS{24}In
whatever situation
someone
was called,
brothers and sisters,
let him remain
in
it
with
God.
\par }{\SH Remaining Unmarried
\par }{\PP \VS{25}With regard
to the question about people who have never married,
I have
no
command
from
the Lord,
but
I give
my opinion
as
one shown mercy
by
the Lord
to be
trustworthy.
\VS{26}Because of
the impending crisis
I think
it best
for you to remain as you are.
\VS{27}The one bound
to a wife
should
not
seek
divorce.
The one released
from
a wife
should
not
seek
marriage.
\VS{28}But
if
you marry,
you have
not
sinned.
And
if
a virgin
marries,
she has
not
sinned.
But
those who marry will face
difficult circumstances,
and I
am trying to spare
you
such problems.
\VS{29}And
I say
this, brothers and sisters: The time
is
short.
So then
those who have
wives
should be
as
those who have
none,
\VS{30}those with tears
like
those not
weeping,
those who rejoice
like
those not
rejoicing,
those who buy
like
those without
possessions,
\VS{31}those who use
the world
as though
they were
not
using
it to the full. For
the present shape
of this
world
is passing away.
\par }{\PP \VS{32}And
I want
you
to be
free from concern.
An unmarried man
is concerned
about the things of the Lord,
how
to please
the Lord.
\VS{33}But
a married man
is concerned
about the things of the world,
how
to please
his wife,
\VS{34}and
he is divided.
An unmarried
woman
or
a virgin
is concerned
about the things of the Lord,
to be holy
both
in body
and
spirit.
But
a married woman
is concerned
about the things of the
world,
how
to please
her husband.
\VS{35}I am saying
this
for your
benefit,
not
to
place
a limitation
on you,
but
so that without distraction
you may give notable
and
constant service
to the Lord.
\par }{\PP \VS{36}If
anyone
thinks he is acting inappropriately
toward
his
virgin,
if
she is
past the bloom of youth
and
it seems
necessary,
he should do what
he wishes;
he does
not
sin.
Let them marry.
\VS{37}But
the man who is firm
in
his
commitment,
and is under
no
necessity
but
has
control
over
his
will,
and
has decided
in
his own
mind
to keep
his own
virgin,
does
well.
\VS{38}So then,
the one who marries
his own
virgin
does
well,
but
the one who does
not,
does
better.
\par }{\PP \VS{39}A wife
is bound
as long as
her
husband
is living.
But
if
her husband
dies,
she is
free
to marry
anyone
she wishes
(only
someone in
the Lord).
\VS{40}But
in my
opinion,
she will be
happier
if
she remains
as she is
– and I think
that I too
have
the Spirit
of God!

\par }\Chap{8}{\PP \VerseOne{1}With regard to
food sacrificed to idols,
we know
that
“we
all
have
knowledge.”
Knowledge
puffs up,
but
love
builds up.
\VS{2}If
someone
thinks
he knows
something,
he does not yet
know
to the degree that
he needs
to know.
\VS{3}But
if
someone
loves
God,
he
is known
by
God.
\par }{\PP \VS{4}With regard
then
to eating food
sacrificed
to idols, we know
that
“an idol
in
this world
is nothing,”
and
that
“there is no
God
but
one.”
\VS{5}If after all
there are
so-called
gods,
whether
in
heaven
or
on
earth
(as
there are
many
gods
and
many
lords),
\VS{6}yet
for us
there is one
God,
the Father,
from
whom
are all things
and
for whom we
live, and
one
Lord,
Jesus
Christ,
through
whom
are all things
and
through
whom we live.
\par }{\PP \VS{7}But
this knowledge
is not
shared by
all.
And
some,
by being accustomed
to
idols
in former times, eat
this food as
an idol sacrifice,
and
their conscience,
because it is weak,
is
defiled.
\VS{8}Now food
will
not
bring
us
close
to God.
We are
no
worse
if
we do
not
eat
and no
better
if
we do.
\VS{9}But
be careful
that this
liberty
of yours
does
not
become
a hindrance
to the weak.
\VS{10}For
if
someone
weak
sees
you
who possess
knowledge
dining
in
an idol’s temple,
will
not
his
conscience
be
“strengthened”
to eat
food offered to idols?
\VS{11}So
by
your
knowledge
the weak
brother or sister,
for
whom
Christ
died,
is destroyed.
\VS{12}If you sin
against
your brothers or sisters
in this way
and
wound
their
weak
conscience,
you sin
against
Christ.
\VS{13}For this reason,
if
food
causes
my
brother or sister
to sin,
I will
never
eat
meat
again,
so that
I may
not
cause
one of them
to sin.

\par }\Chap{9}{\PP \VerseOne{1}Am I
not
free? Am I
not
an apostle? Have
I
not seen
Jesus
our
Lord? Are
you
not
my
work
in
the Lord?
\VS{2}If
I am
not
an apostle
to others,
at least
I am
to you,
for
you
are
the confirming sign
of my
apostleship
in
the Lord.
\VS{3}This
is
my
defense
to those who examine
me.
\VS{4}Do
we
not
have
the right
to financial support?
\VS{5}Do
we
not
have
the right
to the company
of a believing
wife,
like
the other
apostles
and
the Lord’s
brothers
and
Cephas?
\VS{6}Or
do only
Barnabas
and
I
lack
the right
not
to work?
\VS{7}Who
ever serves in the army
at his own
expense? Who
plants
a vineyard
and
does
not
eat
its
fruit? Who
tends
a flock
and
does
not
consume
its
milk?
\VS{8}Am I saying
these things
only on the basis of
common sense,
or
does the law
not
say
this as well?
\VS{9}For
it is written
in the law
of Moses, “{\QT{
Do
not
muzzle
an ox
while
it is treading out the grain}}.” God is not
concerned
here about oxen, is he?
\VS{10}Or
is he
not surely speaking
for
our benefit? It was written
for
us,
because the one plowing
and
threshing
ought
to work in
hope
of enjoying the harvest.
\VS{11}If
we
sowed
spiritual blessings
among you,
is it too much
to reap
material things
from you?
\VS{12}If
others
receive this right
from you,
are we
not
more
deserving?
\par }{\PP But
we have
not
made use
of this
right.
Instead
we endure
everything
so that
we may
not
be
a hindrance
to the gospel
of Christ.
\VS{13}Don’t
you know
that
those who serve
in the temple
eat food
from
the temple,
and those who serve
at the altar
receive a part of the offerings?
\VS{14}In the same way
the Lord
commanded
those who proclaim
the gospel
to receive
their living
by
the gospel.
\VS{15}But
I
have
not
used
any of these
rights. And I am
not
writing
these things
so that
something will be done
for
me.
In fact, it would be better
for
me
to die
than
– no one
will deprive
me of my
reason for boasting!
\VS{16}For
if
I preach
the gospel, I
have
no
reason
for boasting,
because
I
am compelled
to do this. Woe
to me
if
I do
not
preach the gospel!
\VS{17}For
if
I do
this
voluntarily,
I have
a reward.
But
if
I do it unwillingly,
I am entrusted with a responsibility.
\VS{18}What
then
is
my
reward? That
when
I preach the gospel
I may offer
the gospel
free of charge,
and so
not
make full use
of my
rights
in
the gospel.
\par }{\PP \VS{19}For
since I am
free
from
all
I can make
myself
a slave
to all,
in order
to gain
even more people.
\VS{20}To the Jews
I became
like
a Jew
to gain
the Jews.
To those under
the law
I became like
one under
the law
(though I myself am not
under
the law) to gain
those under
the law.
\VS{21}To those free from the law
I became like
one free from the law
(though I am not
free from
God’s
law
but
under the law
of Christ) to
gain
those free from the law.
\VS{22}To the weak
I became
weak
in order
to gain
the weak.
I have become
all things
to all people,
so that
by all means
I may save
some.
\par }{\PP \VS{23}I do
all these things
because of
the gospel,
so that
I can be
a participant
in it.
\par }{\PP \VS{24}Do you
not
know
that
all
the runners in
a stadium
compete,
but
only one
receives
the prize? So
run
to
win.
\VS{25}Each competitor
must exercise
self-control
in everything.
They do it to
receive
a perishable
crown,
but
we
an imperishable one.
\par }{\PP \VS{26}So
I
do
not
run
uncertainly
or
box
like
one who hits
only
air.
\VS{27}Instead
I subdue
my
body
and
make it
my slave,
so that
after preaching
to others
I
myself
will
not
be
disqualified.

\par }\Chap{10}{\PP \VerseOne{1}For
I do
not
want
you
to be unaware,
brothers and sisters,
that
our
fathers
were
all
under
the cloud
and
all
passed
through
the sea,
\VS{2}and
all
were baptized
into
Moses
in
the cloud
and
in
the sea,
\VS{3}and
all
ate
the same
spiritual
food,
\VS{4}and
all
drank
the same
spiritual
drink.
For
they were
all drinking
from
the spiritual
rock
that followed
them, and the rock
was
Christ.
\VS{5}But
God
was
not
pleased
with
most
of them,
for
they were cut down
in
the wilderness.
\VS{6}These things
happened
as examples
for us,
so that
we
will
not
crave
evil things
as
they did.
\VS{7}So do
not
be
idolaters,
as
some
of them
were. As
it is written, “{\QT{
The people
sat down
to eat
and
drink
and
rose up
to play}}.”
\VS{8}And
let us
not
be immoral,
as
some
of them
were, and
twenty-three
thousand
died
in a single
day.
\VS{9}And
let us not
put
Christ
to the test,
as
some
of them
did,
and
were destroyed
by
snakes.
\VS{10}And
do
not
complain,
as
some
of them
did,
and
were killed
by
the destroying angel.
\VS{11}These things
happened
to them
as examples
and
were written
for
our
instruction,
on
whom
the ends
of the ages
have come.
\VS{12}So
let the one who thinks
he is standing
be careful
that he does
not
fall.
\VS{13}No
trial
has overtaken
you
that is not
faced
by others. And God
is faithful: He will
not
let
you
be tried
beyond
what
you are able
to bear,
but
with
the trial
will also provide a way out
so that you may be able
to endure it.
\par }{\SH Avoid Idol Feasts
\par }{\PP \VS{14}So then,
my
dear friends,
flee
from
idolatry.
\VS{15}I am speaking
to thoughtful people.
Consider
what
I say.
\VS{16}Is
not
the cup
of blessing
that
we bless
a sharing
in the blood
of Christ? Is
not
the bread
that
we break
a sharing
in the body
of Christ?
\VS{17}Because
there is one
bread,
we
who are
many
are one
body,
for
we
all
share
the one
bread.
\VS{18}Look
at the people of Israel.
Are not
those who eat
the sacrifices
partners
in the altar?
\VS{19}Am I saying
that
idols
or
food sacrificed to them
amount
to anything?
\VS{20}No,
I mean that
what
the pagans
sacrifice
is to demons
and
not
to God.
I do
not
want
you
to be
partners
with demons.
\VS{21}You cannot
drink
the cup
of
the Lord
and
the cup
of demons.
You cannot
take part
in the table
of the Lord
and
the table
of demons.
\VS{22}Or
are we trying to provoke
the Lord
to jealousy? Are we really stronger than
he is?
\par }{\SH Live to Glorify God
\par }{\PP \VS{23}“Everything
is lawful,”
but
not
everything
is beneficial.
“Everything
is lawful,”
but
not
everything
builds others up.
\VS{24}Do
not
seek
your own
good, but
the good of the other person.
\VS{25}Eat
anything
that is sold
in
the marketplace
without
questions
of
conscience,
\VS{26}for
{\IT{
the earth
and
its
abundance
are the Lord’s}}.
\VS{27}If
an unbeliever
invites
you
to dinner and
you want
to go,
eat
whatever
is served
without
asking questions
of
conscience.
\VS{28}But
if
someone
says
to you,
“This
is
from a sacrifice,”
do
not
eat,
because of
the one
who told
you and
because of conscience –
\VS{29}I do
not
mean
yours
but
the other person’s.
For
why
is
my
freedom
being judged
by
another’s
conscience?
\VS{30}If
I
partake
with thankfulness,
why
am I blamed
for
the food that
I
give thanks for?
\VS{31}So
whether
you eat
or
drink,
or
whatever
you do,
do
everything
for
the glory
of God.
\VS{32}Do
not give
offense
to Jews
or
Greeks
or
to the church
of God,
\VS{33}just as
I also
try to please
everyone
in all things.
I do
not
seek
my own
benefit,
but
the benefit of many,
so that
they may be saved.

\par }\Chap{11}{\PP \VerseOne{1}Be
imitators
of me,
just as
I also
am of Christ.
\par }{\SH Women’s Head Coverings
\par }{\PP \VS{2}I praise
you
because
you remember
me
in everything
and
maintain
the traditions
just as
I passed
them on
to you.
\VS{3}But
I want
you
to know
that
Christ
is
the head
of every
man,
and
the man
is the head
of a woman,
and
God
is the head
of Christ.
\VS{4}Any
man
who prays
or
prophesies
with
his head
covered
disgraces
his
head.
\VS{5}But
any
woman
who prays
or
prophesies
with her head
uncovered
disgraces
her
head,
for
it is
one
and
the same thing
as having a shaved head.
\VS{6}For
if
a woman
will
not
cover
her head, she should cut off her hair.
But
if
it is disgraceful
for a woman
to have
her hair cut off
or
her head shaved,
she should cover
her head.
\VS{7}For
a man
should
not
have
his head
covered,
since he is
the image
and
glory
of God.
But
the woman
is
the glory
of the man.
\VS{8}For
man
did not
come
from
woman,
but
woman
from
man.
\VS{9}Neither
was
man
created
for the sake
of woman,
but
woman
for
man.
\VS{10}For
this reason
a woman
should
have
a symbol of authority
on
her head,
because of
the angels.
\VS{11}In any case,
in
the Lord
woman
is not independent
of man,
nor
is man
independent
of woman.
\VS{12}For
just as
woman
came from
man,
so
man
comes through
woman.
But
all things
come from
God.
\VS{13}Judge
for
yourselves: Is it
proper
for a woman
to pray
to God
with her head uncovered?
\VS{14}Does
not
nature
itself
teach
you
that
if
a man
has long hair,
it is
a disgrace
for him,
\VS{15}but
if
a woman
has long hair,
it is
her
glory? For
her hair
is given
to her
for
a covering.
\VS{16}If
anyone
intends
to quarrel
about this, we
have
no
other
practice,
nor
do the churches
of God.
\par }{\SH The Lord’s Supper
\par }{\PP \VS{17}Now in giving the following instruction
I do
not
praise
you, because
you come together
not
for
the better
but
for
the worse.
\VS{18}For
in the first place,
when
you
come together
as
a church
I hear
there are
divisions
among
you,
and
in part
I believe it.
\VS{19}For
there must
in fact be
divisions
among
you,
so that
those of
you
who are approved
may be
evident.
\VS{20}Now
when
you
come together
at
the same place,
you are
not
really eating
the Lord’s
Supper.
\VS{21}For
when it is time
to eat,
everyone
proceeds with his own
supper.
One is hungry
and
another becomes drunk.
\VS{22}Do you not
have
houses
so that
you can eat
and
drink? Or
are you trying to show contempt
for the church
of God
by shaming
those who have
nothing? What
should I say
to you? Should I praise
you? I will
not
praise
you for
this!
\par }{\PP \VS{23}For
I
received
from
the Lord
what
I
also
passed on
to you,
that
the Lord
Jesus
on
the night
in which
he was betrayed
took
bread,
\VS{24}and
after
he had given thanks
he broke
it and
said,
“This
is
my
body,
which is for
you.
Do
this
in
remembrance
of me.”
\VS{25}In the same way,
he also took the cup
after
supper,
saying,
“This
cup
is
the new
covenant
in
my
blood.
Do
this,
every time
you drink
it, in
remembrance
of me.”
\VS{26}For
every time
you eat
this
bread
and
drink
the cup,
you proclaim
the Lord’s
death
until
he comes.
\par }{\PP \VS{27}For this reason,
whoever
eats
the bread
or
drinks
the cup
of the Lord
in an unworthy manner
will be
guilty
of the body
and
blood
of the Lord.
\VS{28}A person
should examine
himself
first, and
in this way
let him eat
the bread
and
drink
of
the cup.
\VS{29}For
the one who eats
and
drinks
without
careful regard
for the body
eats and drinks
judgment
against himself.
\VS{30}That
is why many
of you
are weak
and
sick,
and
quite a few
are dead.
\VS{31}But
if
we examined
ourselves,
we would
not
be judged.
\VS{32}But
when
we are judged
by
the Lord,
we are disciplined
so that
we may
not
be condemned
with
the world.
\VS{33}So then,
my
brothers and sisters,
when
you come together
to eat,
wait for
one another.
\VS{34}If
anyone
is hungry,
let him eat
at
home,
so
that when you assemble
it does not
lead to
judgment.
I will give directions
about other matters
when
I come.

\par }\Chap{12}{\PP \VerseOne{1}With regard to
spiritual gifts,
brothers and sisters,
I do
not
want
you
to be uninformed.
\VS{2}You know
that
when
you were
pagans
you were
often led astray
by speechless
idols,
however
you were led.
\VS{3}So
I want
you
to understand that
no one
speaking
by
the Spirit
of God
says,
“Jesus
is cursed,”
and
no one
can
say,
“Jesus
is Lord,”
except
by
the Holy
Spirit.
\par }{\PP \VS{4}Now there are
different
gifts,
but
the same
Spirit.
\VS{5}And
there are
different
ministries,
but
the same
Lord.
\VS{6}And
there are
different
results,
but
the same
God
who produces
all
of them in
everyone.
\VS{7}To each
person the manifestation
of the Spirit
is given
for
the benefit of all.
\VS{8}For
one person is given
through
the Spirit
the message
of wisdom,
and
another
the message
of knowledge
according
to the same
Spirit,
\VS{9}to another
faith
by
the same
Spirit,
and
to another
gifts
of healing
by
the one
Spirit,
\VS{10}to another
performance
of miracles,
to another
prophecy,
and
to another
discernment
of spirits,
to another different
kinds
of tongues,
and
to another
the interpretation
of tongues.
\VS{11}It is one
and
the same
Spirit,
distributing
as
he decides
to each
person, who produces
all
these things.
\par }{\SH Different Members in One Body
\par }{\PP \VS{12}For
just as
the body
is
one
and
yet has
many
members,
and
all
the members
of the body
– though many
– are
one
body,
so
too
is
Christ.
\VS{13}For
in
one
Spirit
we
were all
baptized
into
one
body.
Whether
Jews
or
Greeks
or
slaves
or
free,
we were
all
made to drink
of the one
Spirit.
\VS{14}For
in fact the body
is
not
a single
member,
but
many.
\VS{15}If
the foot
says,
“Since
I am
not
a hand,
I am
not
part of
the body,”
it does not
lose its membership in the body
because of
that.
\VS{16}And
if
the ear
says,
“Since
I am
not
an eye,
I am
not
part of
the body,”
it does
not lose its membership in the body
because of
that.
\VS{17}If
the whole
body
were an eye,
what part would do
the hearing? If
the whole
were an ear,
what part would
exercise the sense of smell?
\VS{18}But
as a matter of fact, God
has placed
each
of the members
in
the body
just as
he decided.
\VS{19}If
they were
all
the same member,
where
would the body be?
\VS{20}So
now
there are many
members,
but
one
body.
\VS{21}The eye
cannot
say
to the hand,
“I do not
need
you,”
nor in turn
can the head
say to the foot,
“I do
not
need
you.”
\VS{22}On the contrary,
those members
that seem
to be weaker
are
essential,
\VS{23}and
those members
we consider
less honorable
we clothe
with greater
honor,
and
our
unpresentable
members are clothed
with dignity,
\VS{24}but
our
presentable
members do
not
need
this. Instead,
God
has blended together
the body,
giving
greater
honor
to the lesser member,
\VS{25}so that
there may be no
division
in
the body,
but
the members
may have mutual concern
for
one another.
\VS{26}If
one
member
suffers,
everyone
suffers
with it. If
a member
is honored,
all
rejoice with it.
\par }{\PP \VS{27}Now
you
are
Christ’s
body,
and
each of you is a member
of
it.
\VS{28}And
God
has placed
in
the church
first
apostles,
second
prophets,
third
teachers,
then
miracles,
gifts
of healing,
helps,
gifts of leadership,
different kinds
of tongues.
\VS{29}Not
all
are apostles,
are they? Not
all
are prophets,
are they? Not
all
are teachers,
are they? Not
all
perform miracles, do they?
\VS{30}Not
all
have
gifts
of healing,
do they? Not
all
speak
in tongues,
do they? Not
all
interpret, do they?
\VS{31}But
you should be eager
for the greater
gifts.
\par }{\PP And
now
I will show
you
a way
that is beyond
comparison.

\par }\Chap{13}{\PP \VerseOne{1}If
I speak
in the tongues
of men
and
of angels,
but
I do
not
have
love,
I am
a noisy gong
or
a clanging
cymbal.
\VS{2}And if
I have
prophecy,
and
know
all
mysteries
and
all
knowledge,
and if
I have
all
faith
so that
I can remove
mountains,
but
do
not
have
love,
I am
nothing.
\VS{3}If
I give away
everything
I
own,
and if
I give over
my
body
in order to
boast,
but
do
not
have
love,
I receive
no
benefit.
\par }{\PP \VS{4}Love
is patient,
love
is kind,
it is
not
envious.
Love does
not
brag,
it is
not
puffed up.
\VS{5}It is
not
rude,
it is
not
self-serving,
it is
not
easily angered
or resentful.
\VS{6}It is
not
glad
about
injustice,
but
rejoices
in the truth.
\VS{7}It bears
all things,
believes
all things,
hopes
all things,
endures
all things.
\par }{\PP \VS{8}Love
never
ends.
But
if
there are prophecies,
they will be set aside;
if
there are tongues,
they will cease;
if
there is knowledge,
it will be set aside.
\VS{9}For
we know
in
part,
and
we prophesy
in
part,
\VS{10}but
when
what is
perfect
comes,
the partial
will be set aside.
\VS{11}When
I was
a child,
I talked
like
a child,
I thought
like
a child,
I reasoned
like
a child.
But when
I became
an adult,
I set aside
childish ways.
\VS{12}For
now
we see
in
a mirror
indirectly,
but
then
we will see face
to
face.
Now
I know
in
part,
but
then
I will know fully,
just as
I have been fully known.
\VS{13}And
now
these
three
remain: faith,
hope,
and love.
But
the greatest
of these
is love.

\par }\Chap{14}{\PP \VerseOne{1}Pursue
love
and
be eager
for the spiritual gifts,
especially
that
you may prophesy.
\VS{2}For
the one speaking
in a tongue
does
not
speak
to people
but
to God,
for
no one
understands;
he is speaking
mysteries
by the Spirit.
\VS{3}But
the one who prophesies
speaks
to people
for their strengthening,
encouragement,
and
consolation.
\VS{4}The one who speaks
in a tongue
builds
himself
up,
but
the one who prophesies
builds up
the church.
\VS{5}I wish
you
all
spoke
in tongues,
but
even more
that
you would prophesy.
The one who prophesies
is greater than
the one who speaks
in tongues,
unless
he interprets
so that
the church
may be
strengthened.
\par }{\PP \VS{6}Now,
brothers and sisters,
if
I come
to
you
speaking
in tongues,
how
will I help
you
unless
I speak
to you
with
a revelation
or
with
knowledge
or
prophecy
or
teaching?
\VS{7}It is similar
for lifeless
things that make
a sound,
like a flute
or
harp.
Unless
they make
a distinction
in the notes,
how
can what is played on the flute
or
harp
be understood?
\VS{8}If,
for example, the trumpet
makes
an unclear
sound,
who
will get ready
for
battle?
\VS{9}It
is the same for you. If
you do not
speak clearly
with
your tongue,
how
will
anyone know
what is being said? For
you will be
speaking
into
the air.
\VS{10}There are
probably many
kinds
of languages
in
the world,
and
none
is without meaning.
\VS{11}If
then
I do
not
know
the meaning
of a language,
I will be
a foreigner
to the speaker
and
the speaker
a foreigner
to
me.
\VS{12}It is the same
with you.
Since
you are
eager for manifestations
of the Spirit,
seek
to
abound
in order to
strengthen
the church.
\par }{\PP \VS{13}So then,
one who speaks
in a tongue
should pray
that
he may interpret.
\VS{14}If
I pray
in a tongue,
my
spirit
prays,
but
my
mind
is
unproductive.
\VS{15}What
should I do? I will pray
with my spirit,
but
I will
also
pray
with my mind.
I will sing praises
with my spirit,
but
I will
also
sing praises
with my mind.
\VS{16}Otherwise,
if
you are praising God
with
your spirit,
how can
someone without the gift
say
“Amen”
to
your
thanksgiving,
since
he does
not
know
what
you are saying?
\VS{17}For
you
are
certainly
giving thanks
well,
but
the other person
is
not
strengthened.
\VS{18}I thank
God
that I speak
in tongues
more than
all
of you,
\VS{19}but
in
the church
I want
to speak
five
words
with my
mind
to instruct
others,
rather than
ten thousand
words
in
a tongue.
\par }{\PP \VS{20}Brothers and sisters,
do
not
be
children
in your thinking.
Instead,
be infants
in evil,
but
in your thinking
be
mature.
\VS{21}It is written
in
the law: “{\QT{
By people with
strange tongues
and
by
the lips
of strangers
I will speak
to this
people, yet not even
in this way
will they listen
to me}},” says
the Lord.
\VS{22}So then,
tongues
are
a sign
not
for believers
but
for unbelievers.
Prophecy,
however,
is not
for unbelievers
but
for believers.
\VS{23}So
if
the whole
church
comes together
and
all
speak
in tongues,
and
unbelievers
or
uninformed
people enter,
will they
not
say
that
you have lost
your minds?
\VS{24}But
if
all
prophesy,
and
an unbeliever
or
uninformed person
enters,
he will be convicted
by
all,
he will be called to account
by
all.
\VS{25}The secrets
of his
heart
are
disclosed,
and
in this way
he will fall down
with
his face
to the ground and worship
God,
declaring,
“God
is
really
among
you.”
\par }{\SH Church Order
\par }{\PP \VS{26}What
should you do then,
brothers and sisters? When
you come together,
each one
has
a song,
has
a lesson,
has
a revelation,
has
a tongue,
has
an interpretation.
Let all these things
be done
for
the strengthening of the church.
\VS{27}If
someone
speaks
in
a tongue,
it should be two,
or
at the most
three,
one after the other,
and
someone
must interpret.
\VS{28}But
if
there is
no
interpreter,
he should be silent
in
the church.
Let him speak
to himself
and
to God.
\VS{29}Two
or
three
prophets
should speak
and
the others
should evaluate what is said.
\VS{30}And
if
someone
sitting down
receives a revelation,
the person who is speaking
should conclude.
\VS{31}For
you can
all
prophesy
one after another,
so
all
can learn
and
be encouraged.
\VS{32}Indeed,
the spirits
of the prophets
are subject
to the prophets,
\VS{33}for
God
is
not
characterized by disorder
but
by peace.
\par }{\PP As
in
all
the churches
of the saints,
\VS{34}the women
should be silent
in
the churches,
for
they
are not
permitted
to speak.
Rather,
let them be in submission,
as
in fact the law
says.
\VS{35}If
they want
to find out
about something,
they should ask
their
husbands
at
home,
because
it is
disgraceful
for a woman
to speak
in
church.
\VS{36}Did
the word
of God
begin
with
you,
or
did it come
to
you
alone?
\par }{\PP \VS{37}If
anyone
considers
himself a prophet
or
spiritual person,
he should acknowledge
that
what
I write
to you
is
the Lord’s
command.
\VS{38}If
someone
does not recognize
this, he is not recognized.
\VS{39}So then,
brothers and sisters,
be eager
to prophesy,
and
do
not
forbid
anyone from speaking
in tongues.
\VS{40}And
do everything
in a decent
and
orderly manner.

\par }\Chap{15}{\PP \VerseOne{1}Now I want
to make clear for you,
brothers and sisters,
the gospel
that
I preached
to you,
that
you received
and
on
which
you stand,
\VS{2}and
by
which
you are being saved,
if
you hold firmly
to the message
I preached
to you
– unless
you believed
in vain.
\VS{3}For
I passed on
to you
as of first importance
what
I
also
received
– that
Christ
died
for
our
sins
according to
the scriptures,
\VS{4}and
that
he was buried,
and
that
he was raised
on the third
day
according to
the scriptures,
\VS{5}and
that
he appeared
to Cephas,
then
to the twelve.
\VS{6}Then
he appeared
to more than
five hundred
of the brothers and sisters
at one time,
most
of
whom
are
still
alive,
though
some
have fallen asleep.
\VS{7}Then
he appeared
to James,
then
to all
the apostles.
\VS{8}Last
of all,
as though
to one born at the wrong time,
he appeared
to me also.
\VS{9}For
I
am
the least
of the apostles,
unworthy
to be called
an apostle,
because
I persecuted
the church
of God.
\VS{10}But
by the grace
of God
I am
what
I am,
and
his
grace
to
me
has
not
been
in vain.
In fact,
I worked
harder than
all
of them
– yet
not
I,
but
the grace
of God
with
me.
\VS{11}Whether
then
it was I
or
they,
this is the way
we preach
and
this is the way
you believed.
\par }{\SH No Resurrection?
\par }{\PP \VS{12}Now
if
Christ
is being preached
as raised
from
the dead,
how
can some
of
you
say
there is
no
resurrection
of the dead?
\VS{13}But
if
there is
no
resurrection
of the dead,
then not even
Christ
has been raised.
\VS{14}And
if
Christ
has
not
been raised,
then
our
preaching
is futile
and
your
faith
is empty.
\VS{15}Also,
we are found
to be false witnesses
about God,
because
we have testified
against
God
that
he raised
Christ
from the dead, when in reality he did
not
raise
him, if indeed
the dead
are
not
raised.
\VS{16}For
if
the dead
are
not
raised,
then not even
Christ
has been raised.
\VS{17}And
if
Christ
has
not
been raised,
your
faith
is
useless;
you are
still
in
your
sins.
\VS{18}Furthermore,
those who have fallen asleep
in
Christ
have
also
perished.
\VS{19}For if
only
in
this
life
we have hope
in
Christ,
we should be
pitied
more than anyone.
\par }{\PP \VS{20}But
now
Christ
has been raised
from
the dead,
the firstfruits
of those who have fallen asleep.
\VS{21}For
since
death
came through
a man,
the resurrection
of the dead
also came through
a man.
\VS{22}For
just as
in
Adam
all
die,
so
also
in
Christ
all
will be made alive.
\VS{23}But
each
in
his own
order: Christ,
the firstfruits;
then
when Christ
comes,
those who belong to him.
\VS{24}Then
comes the end,
when
he hands over
the kingdom
to God
the Father,
when
he has brought to an end
all
rule
and
all
authority
and
power.
\VS{25}For
he must
reign
until
he has put
all
his enemies
under
his
feet.
\VS{26}The last
enemy
to be eliminated
is death.
\VS{27}For
{\BD{
{\QT{
{\IT{he has put}}}}
everything
in subjection
under
his
feet
everything
in subjection
under
his
feet}}. But
when
it says
“everything”
has been put in subjection,
it is clear
that
this does not include
the one who put
everything
in subjection
to him.
\VS{28}And
when
all things
are subjected
to him,
then
the
Son
himself
will be subjected
to the one who subjected
everything
to him,
so that
God
may be all
in
all.
\par }{\PP \VS{29}Otherwise,
what
will
those do
who are baptized
for
the dead? If
the dead
are
not
raised
at all,
then
why
are they baptized
for
them?
\VS{30}Why
too
are
we
in danger
every
hour?
\VS{31}Every
day
I am in danger of death! This is as sure as my boasting
in you,
which
I have
in
Christ
Jesus
our
Lord.
\VS{32}If
from
a human
point of view I fought with wild beasts
at
Ephesus,
what
did it benefit
me? If
the dead
are
not
raised,
{\IT{
let us eat
and
drink,
for
tomorrow
we die}}.
\VS{33}Do
not
be deceived: “Bad
company
corrupts
good
morals.”
\VS{34}Sober up
as you should, and
stop
sinning! For
some
have
no knowledge
of God
– I say
this to
your
shame!
\par }{\SH The Resurrection Body
\par }{\PP \VS{35}But
someone
will say,
“How
are
the dead
raised? With what kind
of body
will they come?”
\VS{36}Fool! What
you
sow
will
not
come to life
unless
it dies.
\VS{37}And
what
you sow
is not
the body
that is to be,
but
a bare
seed –
perhaps
of wheat
or
something
else.
\VS{38}But
God
gives
it
a body
just
as he planned,
and
to each
of the seeds
a body
of its own.
\VS{39}All
flesh
is not
the same: People
have one flesh, animals
have another, birds
and
fish another.
\VS{40}And
there are heavenly
bodies
and
earthly
bodies. The glory
of the heavenly
body is one sort and
the earthly
another.
\VS{41}There is one
glory
of the sun,
and
another
glory
of the moon
and
another
glory
of the stars,
for
star
differs
from star
in
glory.
\par }{\PP \VS{42}It is the same with
the resurrection
of the dead.
What is sown
is perishable,
what is raised
is imperishable.
\VS{43}It is sown
in
dishonor,
it is raised
in
glory;
it is sown
in
weakness,
it is raised
in
power;
\VS{44}it is sown
a natural
body,
it is raised
a spiritual
body.
If
there is
a natural
body,
there is
also
a spiritual body.
\VS{45}So
also
it is written, “{\QT{
The first
man, Adam, became
a living
person}}”; the last
Adam
became a life-giving
spirit.
\VS{46}However,
the spiritual
did not
come first,
but
the natural,
and then
the spiritual.
\VS{47}The first
man
is from
the earth,
made of dust;
the second
man
is from
heaven.
\VS{48}Like
the one made of dust,
so too
are those made of dust,
and
like
the one from heaven,
so too
those who are heavenly.
\VS{49}And
just as
we have borne
the image
of the man of dust,
let us
also
bear
the image
of the man of heaven.
\par }{\PP \VS{50}Now this
is what I am saying,
brothers and sisters: Flesh
and
blood
cannot
inherit
the kingdom
of God,
nor
does the perishable
inherit
the imperishable.
\VS{51}Listen,
I will tell
you
a mystery: We will
not
all
sleep,
but
we will
all
be changed –
\VS{52}in
a moment,
in
the blinking
of an eye,
at
the last
trumpet.
For
the trumpet will sound,
and
the dead
will be raised
imperishable,
and
we
will be changed.
\VS{53}For
this
perishable body
must put on
the imperishable,
and
this
mortal body
must put on
immortality.
\VS{54}Now
when
this
perishable
puts on
the imperishable, and this mortal puts on immortality,
then
the saying
that is written
will happen,
\par }{\Q “{\IT{
{\BD{Death}}
has been swallowed up
in
victory.}}”
\par }{\Q \VS{55}“{\QT{
Where, O death, is your
victory?}}
\par }{\Q {\QT{
Where, O death, is your
sting?}}”
\par }{\PI \VS{56}The
sting
of death
is sin,
and
the
power
of sin
is the law.
\VS{57}But
thanks
be to God,
who gives
us
the victory
through
our
Lord
Jesus
Christ!
\VS{58}So then,
dear
brothers and sisters,
be
firm.
Do not be moved! Always
be outstanding
in
the work
of the Lord,
knowing
that
your
labor
is
not
in vain
in
the Lord.

\par }\Chap{16}{\PP \VerseOne{1}With regard
to the collection
for
the saints,
please follow
the directions that
I gave
to the churches
of Galatia:
\VS{2}On
the first
day of the week,
each
of you
should set aside
some income and save it to the extent that God has blessed
you, so that
a collection
will
not
have to be made
when
I come.
\VS{3}Then,
when
I arrive,
I will send
those
whom you approve
with
letters
of explanation to carry
your
gift
to
Jerusalem.
\VS{4}And
if
it seems advisable
that I
should go
also,
they will go
with
me.
\par }{\SH Paul’s Plans to Visit
\par }{\PP \VS{5}But
I will come
to
you
after
I have gone through
Macedonia
– for
I will be going through
Macedonia –
\VS{6}and perhaps I will stay
with
you,
or
even spend the winter,
so that
you
can send
me
on
my journey,
wherever
I go.
\VS{7}For
I do
not
want
to see
you
now
in
passing,
since I
hope
to spend
some time
with
you,
if
the Lord
allows.
\VS{8}But
I will stay
in
Ephesus
until
Pentecost,
\VS{9}because
a door
of great
opportunity stands wide
open
for me,
but
there are many
opponents.
\par }{\PP \VS{10}Now
if
Timothy
comes,
see
that
he has
nothing to fear
among
you,
for
he is doing
the Lord’s
work,
as
I am too.
\VS{11}So then,
let
no
one
treat
him
with contempt.
But
send
him
on
his way
in
peace
so that
he may come
to
me.
For
I am expecting
him
with
the brothers.
\par }{\PP \VS{12}With regard
to our brother
Apollos: I
strongly
encouraged
him
to
visit
you
with
the other brothers,
but
it was
simply
not
his intention
to
come
now.
He will come
when
he has
the opportunity.
\par }{\SH Final Challenge and Blessing
\par }{\PP \VS{13}Stay alert,
stand firm
in
the faith,
show courage,
be strong.
\VS{14}Everything
you
do should be done
in
love.
\par }{\PP \VS{15}Now,
brothers and sisters,
you know
about the household
of Stephanus,
that
as the first converts
of Achaia,
they devoted
themselves
to
ministry
for the saints.
I urge
you
\VS{16}also
to
submit
to people like this,
and
to everyone
who cooperates in the work
and
labors hard.
\VS{17}I was glad
about
the arrival
of Stephanus,
Fortunatus,
and
Achaicus
because
they have supplied
the fellowship with you
that
I lacked.
\VS{18}For
they refreshed
my
spirit
and
yours.
So then,
recognize
people like this.
\par }{\PP \VS{19}The churches
in the province of Asia
send greetings
to you.
Aquila
and
Prisca
greet
you
warmly
in
the Lord,
with
the church
that meets in
their
house.
\VS{20}All
the brothers and sisters
send greetings.
Greet
one another
with
a holy
kiss.
\par }{\PP \VS{21}I, Paul,
send this greeting
with my own
hand.
\par }{\PP \VS{22}Let anyone
who has
no
love
for the Lord
be
accursed.
Our Lord, come!
\par }{\PP \VS{23}The grace
of the Lord
Jesus
be with
you.
\par }{\PP \VS{24}My
love
be with
all
of you
in
Christ
Jesus.
\par }