\NormalFont\ShortTitle{2 John}
{\MT 2 John

\par }\OneChap {\SH Introduction and Thanksgiving
\par }{\PP \VerseOne{1}From the elder,
to an elect
lady
and
her
children,
whom
I
love
in
truth
(and
not
I
alone,
but
also
all
those who know
the truth),
\VS{2}because of
the truth
that resides
in
us
and
will be
with
us
forever.
\VS{3}Grace,
mercy,
and peace
will be
with
us
from
God
the Father
and
from
Jesus
Christ
the Son
of the Father,
in
truth
and
love.
\par }{\PP \VS{4}I rejoiced
greatly
because
I have found
some
of your
children
living
according to
the truth,
just as
the Father
commanded us.
\par }{\SH Warning Against False Teachers
\par }{\PP \VS{5}But
now
I ask
you,
lady
(not
as
if I were writing
a new
commandment
to you,
but
the one
we have had
from
the beginning), that
we love
one another.
\VS{6}(Now
this
is
love: that
we walk
according to
his
commandments.) This
is
the commandment,
just as
you have heard
from
the beginning;
thus
you should walk
in
it.
\VS{7}For
many
deceivers
have gone out
into
the world,
people who do
not
confess
Jesus
as Christ
coming
in
the flesh.
This person
is
the deceiver
and
the antichrist!
\VS{8}Watch out,
so that
you do
not
lose
the things
we have worked for,
but
receive
a full
reward.
\par }{\PP \VS{9}Everyone
who goes on ahead
and
does
not
remain
in
the teaching
of Christ
does
not
have
God.
The one who remains
in
this
teaching
has
both
the Father
and
the Son.
\VS{10}If
anyone
comes
to
you
and
does not
bring
this
teaching,
do
not
receive
him
into
your house
and
do
not
give
him
any greeting,
\VS{11}because
the person who gives
him
a greeting
shares
in his
evil
deeds.
\par }{\SH Conclusion
\par }{\PP \VS{12}Though I have
many
other things to write
to you,
I do
not
want
to do so
with paper
and
ink,
but
I hope
to come
visit you
and
speak
face
to
face,
so that
our
joy
may be complete.
\par }{\PP \VS{13}The children
of your
elect
sister
greet
you.
\par }