\NormalFont\ShortTitle{2 Thessalonians}
{\MT 2 Thessalonians

\par }\ChapOne{1}{\SH Salutation
\par }{\PP \VerseOne{1}From Paul
and
Silvanus
and
Timothy,
to the church
of the Thessalonians
in
God
our
Father
and
the Lord
Jesus
Christ.
\VS{2}Grace
and
peace
to you
from
God
the Father
and
the Lord
Jesus
Christ!
\par }{\SH Thanksgiving
\par }{\PP \VS{3}We ought
to thank
God
always
for
you,
brothers and sisters,
and rightly
so,
because
your
faith
flourishes more and more
and
the love
of each
one
of you
all
for
one another
is ever greater.
\VS{4}As a result
we
ourselves
boast
about
you
in
the churches
of God
for
your
perseverance
and
faith
in
all
the persecutions
and
afflictions
you are enduring.
\par }{\SH Encouragement in Persecution
\par }{\PP \VS{5}This is evidence
of God’s
righteous
judgment,
to
make
you
worthy
of the kingdom
of God,
for
which
in fact you are suffering.
\VS{6}For
it is right
for
God
to repay with affliction
those who afflict
you,
\VS{7}and
to you
who are being afflicted
to give rest
together with
us
when the Lord
Jesus
is revealed
from
heaven
with
his
mighty
angels.
\VS{8}{\IT{
With
flaming
fire
he will mete out}}
{\IT{
punishment
on those who do
not
know
God}}
and
do
not
obey
the gospel
of our
Lord
Jesus.
\VS{9}They
will undergo
the penalty
of eternal
destruction,
{\IT{
away from
the presence
of the Lord
and
from
the glory
of his
strength}},
\VS{10}when
he comes
to be glorified
among
his
saints
and
admired
on
that
day
among
all
who have believed
– and you
did in fact
believe
our
testimony.
\VS{11}And
in this regard we pray
for
you
always,
that
our
God
will make
you
worthy
of his calling
and
fulfill
by
his power
your every
desire
for goodness
and
every work
of faith,
\VS{12}that the name
of our
Lord
Jesus
may
be glorified
in
you,
and
you
in
him,
according to
the grace
of our
God
and
the Lord
Jesus
Christ.

\par }\Chap{2}{\PP \VerseOne{1}Now regarding
the arrival
of our
Lord
Jesus
Christ
and
our
being gathered
to be with him,
we ask
you,
brothers and sisters,
\VS{2}not
to be
easily
shaken
from
your composure
or
disturbed
by
any kind of spirit
or
message
or
letter
allegedly from
us,
to the effect that
the day
of the Lord
is already here.
\VS{3}Let
no
one
deceive
you
in
any
way.
For
that day will
not
arrive
until the rebellion
comes and
the man
of lawlessness
is revealed,
the son
of destruction.
\VS{4}He opposes
{\IT{
and
exalts
himself above
every}}
so-called
{\IT{
god}}
or
object of worship,
and as a result
{\IT{
he takes
his
seat}}
in
God’s
temple,
displaying
himself
as
God.
\VS{5}Surely you recall
that
I used to tell
you
these things
while I
was still
with
you.
\VS{6}And
so
you know
what holds
him back, so that
he will be revealed
in
his own
time.
\VS{7}For
the hidden
power of lawlessness
is already
at work.
However,
the one who holds
him back will do so until he is taken out of the way,
\VS{8}and
then
the lawless one
will be revealed,
whom
the Lord
will destroy
by the breath
of his
mouth
and
wipe out
by the manifestation
of his
arrival.
\VS{9}The arrival
of the lawless one
will be
by
Satan’s
working
with
all kinds
of miracles
and
signs
and
false
wonders,
\VS{10}and
with
every kind
of evil
deception
directed against those who are perishing,
because
they found
no
place in
their hearts
for the truth
so as to be saved.
\VS{11}Consequently
God
sends
on them
a deluding
influence
so that
they
will believe
what is false.
\VS{12}And so
all of them
who have
not
believed
the truth
but
have delighted
in evil
will be condemned.
\par }{\SH Call to Stand Firm
\par }{\PP \VS{13}But
we
ought
to thank
God
always
for
you,
brothers and sisters
loved
by
the Lord,
because
God
chose
you
from
the beginning
for
salvation
through
sanctification
by
the Spirit
and
faith
in
the truth.
\VS{14}He called
you
to this salvation through
our
gospel,
so that
you may possess
the glory
of our
Lord
Jesus
Christ.
\VS{15}Therefore,
brothers and sisters,
stand firm
and
hold on
to the traditions
that
we
taught
you, whether
by
speech
or
by
letter.
\VS{16}Now
may our
Lord
Jesus
Christ
himself
and
God
our
Father,
who loved
us
and
by
grace
gave
us eternal
comfort
and
good
hope,
\VS{17}encourage
your
hearts
and
strengthen
you in
every
good thing
you do
or
say.

\par }\Chap{3}{\PP \VerseOne{1}Finally,
pray
for
us,
brothers and sisters,
that
the Lord’s
message
may spread quickly
and
be honored
as
in fact it was among
you,
\VS{2}and
that
we may be delivered
from
perverse
and
evil
people.
For
not
all
have faith.
\VS{3}But
the Lord
is
faithful,
and he will strengthen
you
and
protect
you from
the evil one.
\VS{4}And
we are confident
about
you
in
the Lord
that
you are
both
doing
– and
will do
– what
we are commanding.
\VS{5}Now
may the Lord
direct
your
hearts
toward
the love
of God
and
the endurance
of Christ.
\par }{\SH Response to the Undisciplined
\par }{\PP \VS{6}But
we command
you,
brothers and sisters,
in
the name
of our Lord
Jesus
Christ,
to keep away
from
any
brother
who lives
an undisciplined life
and
not
according to
the tradition
they received
from
us.
\VS{7}For
you know
yourselves
how
you must
imitate
us,
because
we did
not
behave without discipline
among
you,
\VS{8}and we did
not
eat
anyone’s
food
without paying.
Instead,
in
toil
and
drudgery
we worked
night
and
day
in order
not
to burden
any
of you.
\VS{9}It was not
because
we do
not
have
that right,
but
to give
ourselves
as an example
for you
to
imitate.
\VS{10}For
even when
we were
with
you,
we used to give
you
this command: “If
anyone
is
not
willing
to work,
neither
should he eat.”
\VS{11}For
we hear
that some
among
you
are living
an undisciplined life,
not
doing
their own
work
but
meddling in the work of others.
\VS{12}Now
such people
we command
and
urge
in
the Lord
Jesus
Christ
to work
quietly
and so provide
their own
food
to eat.
\VS{13}But
you,
brothers and sisters,
do
not
grow weary
in doing what is right.
\VS{14}But
if
anyone
does
not
obey
our
message
through
this
letter,
take note
of him and do
not
associate closely
with him,
so that
he may be ashamed.
\VS{15}Yet
do
not
regard
him as
an enemy,
but
admonish
him as
a brother.
\par }{\SH Conclusion
\par }{\PP \VS{16}Now
may
the Lord
of peace
himself
give
you
peace
at
all times
and in
every
way.
The Lord
be with
you
all.
\VS{17}I, Paul,
write this greeting
with my own
hand,
which
is
how
I write
in
every
letter.
\VS{18}The grace
of our
Lord
Jesus
Christ
be with
you
all.
\par }