\NormalFont\ShortTitle{Amos}
{\MT Amos

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \VerseOne{1}The following is a record of what
Amos
prophesied. He was one of the herdsmen
from Tekoa.
These prophecies
about Israel
were revealed
to him
during the time
of King
Uzziah
of Judah
and King
Jeroboam
son
of Joash
of Israel,
two years
before
the earthquake.
\par }{\SH God Will Judge the Surrounding Nations
\par }{\Q \VS{2}Amos said:
\par }{\Q “The
{\ND{Lord}}
comes roaring out
of Zion;
\par }{\Q from Jerusalem
he comes
bellowing!

\par }{\Q The shepherds’
pastures
wilt;
\par }{\Q the summit
of Carmel
withers.”
\par }{\Q \VS{3}This is what
the {\ND{Lord}}
says:
\par }{\Q “Because Damascus
has committed
three
crimes –
\par }{\Q make that four! – I will not revoke my
\par }{\Q decree of judgment.
\par }{\Q They ripped through Gilead like threshing sledges with iron teeth.
\par }{\Q \VS{4}So I will set
Hazael’s
house
on fire;
\par }{\Q fire will consume
Ben Hadad’s
fortresses.
\par }{\Q \VS{5}I will break
the bar on the gate
of Damascus.
\par }{\Q I will remove
the ruler
from Wicked
Valley,
\par }{\Q the one who holds
the royal scepter
from Beth
Eden.
\par }{\Q The people
of Aram
will be deported
to Kir.”
\par }{\Q The
{\ND{Lord}}
has spoken!
\par }{\Q \VS{6}This is what
the {\ND{Lord}}
says:
\par }{\Q “Because
Gaza
has committed three
crimes –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They deported a whole community and sold them to Edom.
\par }{\Q \VS{7}So
I will set
Gaza’s
city wall
on fire;
\par }{\Q fire will consume
her fortresses.
\par }{\Q \VS{8}I will remove
the ruler
from Ashdod,
\par }{\Q the one who holds
the royal scepter
from Ashkelon.
\par }{\Q I will strike Ekron
with my hand;
\par }{\Q the rest
of the Philistines
will also die.”
\par }{\Q The sovereign

{\ND{Lord}}
has spoken!
\par }{\Q \VS{9}This is what
the {\ND{Lord}}
says:
\par }{\Q “Because
Tyre
has committed three crimes –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They sold a whole community to Edom;
\par }{\Q they failed to observe a treaty of brotherhood.
\par }{\Q \VS{10}So
I will set
fire
to Tyre’s
city wall;
\par }{\Q fire will consume
her fortresses.”
\par }{\Q \VS{11}This is what
the
{\ND{Lord}}
says:
\par }{\Q “Because Edom
has committed three
crimes –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q He chased his brother with a sword;
\par }{\Q he wiped out his allies.
\par }{\Q In his anger he tore them apart without stopping to rest;
\par }{\Q in his fury he relentlessly attacked them.
\par }{\Q \VS{12}So I will set
Teman
on fire;
\par }{\Q fire will consume
Bozrah’s
fortresses.”
\par }{\Q \VS{13}This is what the
{\ND{Lord}}
says:
\par }{\Q “Because the Ammonites
have committed three
crimes –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They ripped open Gilead’s pregnant women
\par }{\Q so they could expand their territory.
\par }{\Q \VS{14}So I will set
fire
to Rabbah’s
city wall;
\par }{\Q fire will consume
her fortresses.
\par }{\Q War cries
will be heard on the day
of battle;
\par }{\Q a strong gale
will blow on the day
of the windstorm.
\par }{\Q \VS{15}Ammon’s
king
will be deported;
\par }{\Q he
and his officials
will be carried off together.”
\par }{\Q The
{\ND{Lord}} has spoken!

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}This is what
the {\ND{Lord}}
says:
\par }{\Q “Because
Moab
has committed three
crimes –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They burned the bones of Edom’s king into lime.
\par }{\Q \VS{2}So I will set
Moab
on fire,
\par }{\Q and it will consume
Kerioth’s
fortresses.
\par }{\Q Moab
will perish
in the heat of battle
\par }{\Q amid war cries
and the blaring
of the ram’s horn.
\par }{\Q \VS{3}I will remove
Moab’s leader;
\par }{\Q I will kill
all
Moab’s
officials
with
him.”
\par }{\Q The
{\ND{Lord}} has spoken!
\par }{\Q \VS{4}This is what
the
{\ND{Lord}}
says:
\par }{\Q “Because Judah
has committed three
covenant transgressions –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They rejected the
{\ND{Lord}}’s law;
\par }{\Q they did not obey his commands.
\par }{\Q Their false gods,
\par }{\Q to which their fathers were loyal,
\par }{\Q led them astray.
\par }{\Q \VS{5}So I will set
Judah
on fire,
\par }{\Q and it will consume
Jerusalem’s
fortresses.”
\par }{\SH God Will Judge Israel
\par }{\Q \VS{6}This
is what the
{\ND{Lord}}
says:
\par }{\Q “Because
Israel
has committed three
covenant transgressions –
\par }{\Q make that four! – I will not revoke my decree of judgment.
\par }{\Q They sold the innocent for silver,
\par }{\Q the needy for a pair of sandals.
\par }{\Q \VS{7}They trample
on
the dirt-covered
heads
of the poor;
\par }{\Q they push
the destitute away.
\par }{\Q A man
and his father
go
to
the same girl;
\par }{\Q in this
way
they
show
disrespect for my
moral purity.
\par }{\Q \VS{8}They stretch
out on
clothing
seized as collateral;
\par }{\Q they do so right beside
every
altar!
\par }{\Q They drink
wine
bought with the fines
they have levied;
\par }{\Q they do so right in the temple
of their God!
\par }{\Q \VS{9}For Israel’s sake I
destroyed
the Amorites.
\par }{\Q They were as tall
as cedars
\par }{\Q and as strong
as oaks,
\par }{\Q but I destroyed
the fruit
on their branches
\par }{\Q and their roots
in the ground.
\par }{\Q \VS{10}I
brought you up
from the land
of Egypt;
\par }{\Q I led
you through the wilderness
for forty
years
\par }{\Q so you could take
the
Amorites’
land as your own.
\par }{\Q \VS{11}I made some
of your sons
prophets
\par }{\Q and some of your young men
Nazirites.
\par }{\Q Is this
not
true, you Israelites?”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{12}“But you made the Nazirites
drink
wine;
\par }{\Q you commanded
the prophets,
‘Do not
prophesy!’
\par }{\Q \VS{13}Look! I
will press
you down,
\par }{\Q like a cart
loaded down
with grain
presses down.
\par }{\Q \VS{14}Fast runners
will find no place
to hide;
\par }{\Q strong
men will have no
strength
left;
\par }{\Q warriors
will not
be able to save
their lives.
\par }{\Q \VS{15}Archers
will not
hold their ground;
\par }{\Q fast
runners
will not
save
their lives,
\par }{\Q nor
will those who ride
horses.
\par }{\Q \VS{16}Bravehearted
warriors
will run away
naked
in that
day.”
\par }{\Q The
{\ND{Lord}}
is speaking!

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}Listen,
you Israelites,
to this
message
which
the {\ND{Lord}}
is
proclaiming
against
you! This
message is for the entire
clan
I brought up
from the land
of Egypt:
\VS{2}“I have chosen
you alone
from all
the clans
of the earth.
Therefore
I will punish
you for all
your sins.”
\par }{\Q \VS{3}Do two
walk
together
without
having met?
\par }{\Q \VS{4}Does a lion
roar
in the woods
if he has not
cornered
his prey?

\par }{\Q Does
a young lion
bellow from his den
if
he has not
caught something?
\par }{\Q \VS{5}Does a bird
swoop down into
a trap
on the ground
if
there is no
bait?
\par }{\Q Does a trap
spring up
from
the ground
unless
it
has surely caught something?
\par }{\Q \VS{6}If
an alarm
sounds
in a city,
do people
not
fear?

\par }{\Q If
disaster
overtakes
a city,
is the
{\ND{Lord}}
not
responsible?
\par }{\Q \VS{7}Certainly
the sovereign

{\ND{Lord}}
does nothing
without first revealing
his plan to
his servants
the prophets.
\par }{\Q \VS{8}A lion
has roared! Who
is not
afraid?

\par }{\Q The sovereign

{\ND{Lord}}
has spoken! Who can
refuse
to prophesy?
\par }{\SH Samaria Will Fall
\par }{\Q \VS{9}Make this announcement
in
the fortresses
of Ashdod
\par }{\Q and in
the fortresses
in the land
of Egypt.
\par }{\Q Say
this:
\par }{\Q “Gather
on
the hills
around Samaria!

\par }{\Q Observe
the many
acts of violence
taking place within
the city,

\par }{\Q the oppressive
deeds occurring in it.”
\par }{\Q \VS{10}“They do not
know
how to do
what is right.”
(The
{\ND{Lord}}
is speaking.)
\par }{\Q “They store up
the spoils of destructive
violence
in their fortresses.
\par }{\Q \VS{11}Therefore,”
says
the sovereign

{\ND{Lord}}, “an enemy
will encircle
the land.
\par }{\Q He will take
away your power;
\par }{\Q your fortresses
will be looted.”
\par }{\Q \VS{12}This is what
the {\ND{Lord}}
says:
\par }{\Q “Just
as a shepherd
salvages
from the lion’s
mouth
a couple
of leg
bones or
a piece
of an ear,
\par }{\Q so
the Israelites
who live
in Samaria
will be salvaged.
\par }{\Q They will be left with just a corner
of a bed,
\par }{\Q and a part
of a couch.”
\par }{\Q \VS{13}Listen
and warn
the family
of Jacob!

\par }{\Q The sovereign

{\ND{Lord}}, the God
who commands armies,
is speaking!
\par }{\Q \VS{14}“Certainly
when
I punish
Israel
for their covenant transgressions,
\par }{\Q I will destroy Bethel’s
altars.
\par }{\Q The horns
of the altar
will be cut off
and fall
to the ground.
\par }{\Q \VS{15}I will destroy
both
the winter
and summer
houses.
\par }{\Q The houses
filled with ivory
will be ruined,
\par }{\Q the great
houses
will be swept away.”
\par }{\Q The
{\ND{Lord}}
is speaking!

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}Listen
to this
message,
you cows
of Bashan
who
live on Mount
Samaria!
\par }{\Q You oppress
the poor;
\par }{\Q you crush
the needy.
\par }{\Q You say
to your husbands,
\par }{\Q “Bring
us more to drink!”
\par }{\Q \VS{2}The sovereign

{\ND{Lord}}
confirms this oath
by his own holy
character:

\par }{\Q “Certainly the time
is approaching

\par }{\Q when you will be carried away
in baskets,
\par }{\Q every last one
of you in fishermen’s
pots.
\par }{\Q \VS{3}Each of you will go
straight
through the gaps
in the walls;

\par }{\Q you will be thrown
out toward Harmon.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\SH Israel has an Appointment with God
\par }{\Q \VS{4}“Go
to Bethel
and rebel!

\par }{\Q At Gilgal
rebel
some more!
\par }{\Q Bring
your sacrifices
in the morning,
\par }{\Q your tithes
on the third
day!
\par }{\Q \VS{5}Burn
a thank offering
of bread made with yeast!

\par }{\Q Make a public
display
of your voluntary
offerings!

\par }{\Q For
you love
to do this,
you Israelites.”
\par }{\Q The sovereign

{\ND{Lord}}
is speaking!
\par }{\Q \VS{6}“But surely I
gave
you no food to eat
in any
of your cities;
\par }{\Q you lacked
food
everywhere
you live.
\par }{\Q Still you did not
come back
to
me.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{7}“I
withheld
rain
from
you three
months
before
the harvest.
\par }{\Q I gave rain
to one
city,
but not
to another.
\par }{\Q One
field
would get rain,
but the field
that
received no
rain
dried up.
\par }{\Q \VS{8}People from two
or three
cities
staggered
into
one
city
to get
water,
\par }{\Q but remained thirsty.
\par }{\Q Still you did not
come back
to me.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{9}“I destroyed
your crops with blight
and disease.
\par }{\Q Locusts
kept devouring
your orchards,
vineyards,
fig trees,
and olive
trees.
\par }{\Q Still you did not
come back
to me.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{10}“I sent
against you a plague
like
one of the Egyptian
plagues.
\par }{\Q I killed
your young men
with
the sword,
\par }{\Q along with
the horses
you had captured.
\par }{\Q I made the stench
from the corpses
rise up
into
your nostrils.
\par }{\Q Still you did not
come back
to me.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{11}“I overthrew
some of you the way God
overthrew
Sodom
and Gomorrah.
\par }{\Q You were
like a burning stick snatched
from
the flames.
\par }{\Q Still you did not
come back
to me.”
\par }{\Q The
{\ND{Lord}}
is speaking!
\par }{\Q \VS{12}“Therefore
this is what
I will do
to you, Israel.
\par }{\Q Because
I will do
this
to you,
\par }{\Q prepare
to meet
your God,
Israel!
\par }{\Q \VS{13}For
here
he is!
\par }{\Q He formed
the mountains
and created
the wind.
\par }{\Q He reveals
his plans to men.
\par }{\Q He turns the dawn
into darkness
\par }{\Q and marches
on
the heights
of the earth.
\par }{\Q The
{\ND{Lord}}, the God
who commands armies,
is his name!”

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}Listen
to this
funeral song
I am
ready to sing
about you, family
of Israel:
\par }{\Q \VS{2}“The virgin
Israel
has fallen
down and will not
get
up again.
\par }{\Q She is abandoned
on
her own land
\par }{\Q with no
one to help her get up.”
\par }{\Q \VS{3}The
sovereign

{\ND{Lord}}
says this:
\par }{\Q “The city
that marches out
with a thousand
soldiers will have only a hundred
left;
\par }{\Q the town that marches out
with a hundred
soldiers will have only ten
left
for the family
of Israel.”
\par }{\Q \VS{4}The
{\ND{Lord}}
says
this to the family
of Israel:
\par }{\Q “Seek
me so you can live!
\par }{\Q \VS{5}Do not
seek
Bethel!

\par }{\Q Do not visit Gilgal!
\par }{\Q Do not
journey down
to Beer Sheba!
\par }{\Q For
the people of Gilgal
will certainly
be carried
into exile;
\par }{\Q and Bethel
will become
a place where disaster abounds.”
\par }{\Q \VS{6}Seek
the
{\ND{Lord}}
so you can live!
\par }{\Q Otherwise
he will break
out like fire
against Joseph’s
family;
\par }{\Q the fire will consume
\par }{\Q and no
one will be able to quench
it and save Bethel.
\par }{\Q \VS{7}The Israelites turn
justice
into bitterness;
\par }{\Q they throw what is fair
and right
to the ground.
\par }{\Q \VS{8}(But there is one who made
the constellations Pleiades
and Orion;
\par }{\Q he can turn
the darkness
into morning
\par }{\Q and daylight
into night.
\par }{\Q He summons
the water
of the seas
\par }{\Q and pours
it out on
the earth’s
surface.
\par }{\Q The
{\ND{Lord}}
is his name!
\par }{\Q \VS{9}He flashes
destruction
down upon
the strong
\par }{\Q so that destruction
overwhelms
the fortified places.)
\par }{\Q \VS{10}The Israelites hate
anyone who arbitrates
at the city gate;
\par }{\Q they despise
anyone who speaks
honestly.
\par }{\Q \VS{11}Therefore,
because
you make the poor
pay taxes
on
their crops

\par }{\Q and exact
a grain
tax
from
them,
\par }{\Q you will not
live
in the houses
you built
with chiseled stone,
\par }{\Q nor
will you drink
the
wine
from the fine
vineyards
you planted.
\par }{\Q \VS{12}Certainly
I am aware
of your many
rebellious acts
\par }{\Q and your numerous
sins.
\par }{\Q You torment
the innocent,
you take
bribes,
\par }{\Q and you deny justice
to the needy
at the city gate.
\par }{\Q \VS{13}For this reason
whoever
is
smart
keeps quiet in such a time,
\par }{\Q for
it is an evil
time.
\par }{\Q \VS{14}Seek
good
and not
evil
so
you can live!
\par }{\Q Then
the {\ND{Lord}}, the God
who commands armies,
just
might be with
you,
\par }{\Q as you claim he is.
\par }{\Q \VS{15}Hate
what is wrong,
love
what is right!
\par }{\Q Promote
justice
at the city gate!

\par }{\Q Maybe
the {\ND{Lord}}, the God
who commands armies,
will have mercy
on those who are left
from Joseph.
\par }{\Q \VS{16}Because
of Israel’s sins this is what
the {\ND{Lord}}, the God
who commands armies,
the sovereign One,
says:
\par }{\Q “In all
the squares
there will be wailing,
\par }{\Q in all
the streets
they will mourn the dead.
\par }{\Q They will tell
the field workers
to lament
\par }{\Q and the professional
mourners
to wail.
\par }{\Q \VS{17}In all
the vineyards
there will be wailing,
\par }{\Q for
I will pass
through your midst,”
says
the {\ND{Lord}}.
\par }{\SH The Lord Demands Justice
\par }{\Q \VS{18}Woe
to those who wish
for the
day
of the
{\ND{Lord}}

\par }{\Q Why
do you want
the
{\ND{Lord}}’s
day
of judgment to come?
\par }{\Q It will bring darkness,
not
light.
\par }{\Q \VS{19}Disaster will be inescapable,
\par }{\Q as if a man
ran from a lion
only to meet
a bear,
\par }{\Q then escaped
into a house,
\par }{\Q leaned
his hand
against
the wall,
\par }{\Q and was bitten
by a poisonous snake.
\par }{\Q \VS{20}Don’t
you realize the
{\ND{Lord}}’s
day
of judgment will bring darkness,
not
light
–
\par }{\Q gloomy
blackness, not
bright light?
\par }{\Q \VS{21}“I absolutely
despise
your festivals!
\par }{\Q I get no
pleasure from your religious
assemblies!
\par }{\Q \VS{22}Even
if
you offer
me burnt
and grain offerings,
I will not
be satisfied;
\par }{\Q I will not
look
with favor
on your peace offerings
of fattened calves.
\par }{\Q \VS{23}Take away
from me your noisy
songs;
\par }{\Q I don’t
want to hear
the music
of your stringed instruments.
\par }{\Q \VS{24}Justice
must flow like torrents
of water,
\par }{\Q righteous
actions like a stream
that never dries up.
\par }{\Q \VS{25}You did not bring me sacrifices
and grain offerings
during the forty
years
you spent in the wilderness,
family
of Israel.
\par }{\Q \VS{26}You will pick up
your images
of Sikkuth,
your king,
\par }{\Q and Kiyyun,
your star
god,
which
you made for yourselves,
\par }{\Q \VS{27}and I will drive you into exile
beyond
Damascus,”
says
the {\ND{Lord}}.
\par }{\Q He is called
the God
who commands armies!

\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}Woe
to those who live in ease
in Zion,
\par }{\Q to those who feel secure
on Mount
Samaria.
\par }{\Q They think of themselves as the elite class of the best
nation.
\par }{\Q The family
of Israel
looks to
them for leadership.
\par }{\Q \VS{2}They say to the people:

\par }{\Q “Journey over
to Calneh
and look
at it!
\par }{\Q Then go
from there
to Hamath-Rabbah!

\par }{\Q Then go down
to Gath
of the Philistines!
\par }{\Q Are they superior
to
our two kingdoms?
\par }{\Q Is their territory
larger
than
yours?”
\par }{\Q \VS{3}You refuse to believe
a day
of disaster
will come,
\par }{\Q but you establish a reign
of violence.
\par }{\Q \VS{4}They lie
around on
beds
decorated with ivory,
\par }{\Q and sprawl
out on
their couches.
\par }{\Q They eat
lambs
from the flock,
\par }{\Q and calves
from the middle
of the pen.
\par }{\Q \VS{5}They sing to the tune
of stringed instruments;
\par }{\Q like David
they invent
musical
instruments.
\par }{\Q \VS{6}They drink
wine
from sacrificial bowls,
\par }{\Q and pour
the very best
oils
on themselves.

\par }{\Q Yet they are not
concerned
over
the ruin
of Joseph.
\par }{\Q \VS{7}Therefore
they will now
be the first
to go into exile,
\par }{\Q and the religious
banquets
where they sprawl
on couches will end.
\par }{\Q \VS{8}The sovereign

{\ND{Lord}}
confirms this oath by
his very own life.
\par }{\Q The
{\ND{Lord}}, the God
who commands armies,
is speaking:
\par }{\Q “I
despise
Jacob’s
arrogance;
\par }{\Q I hate
their fortresses.
\par }{\Q I will hand over
to their enemies the city
of Samaria and everything in it.”
\par }{\Q \VS{9}If
ten
men
are left
in one
house,
they too will die.
\VS{10}When
their close relatives,
the ones who will burn the corpses, pick
up their bodies
to remove
the bones
from
the house,
they will say
to
anyone who
is in the inner rooms
of the house,
“Is anyone else
with
you?” He will respond,
“Be quiet! Don’t
invoke
the
{\ND{Lord}}’s
name!”
\par }{\Q \VS{11}Indeed,
look! The
{\ND{Lord}}
is giving the command.
\par }{\Q He will smash
the large
house
to bits,
\par }{\Q and the small
house
into little pieces.
\par }{\Q \VS{12}Can horses
run
on rocky cliffs?
\par }{\Q Can
one plow
the sea with oxen?

\par }{\Q Yet
you have turned
justice
into a poisonous
plant,
\par }{\Q and the fruit
of righteous
actions into a bitter plant.
\par }{\Q \VS{13}You are happy
because you conquered Lo-Debar.
\par }{\Q You say,
“Did we not
conquer Karnaim
by our own power?”
\par }{\Q \VS{14}“Look! I am about to bring
a nation
against
you, family
of Israel.”
\par }{\Q The
{\ND{Lord}}, the God
who commands armies,
is speaking.
\par }{\Q “They will oppress
you all the way
from Lebo-Hamath
to the Stream
of the Arabah.”

\par }\Chap{7}{\PP \VerseOne{1}The sovereign

{\ND{Lord}}
showed
me this: I saw
him making
locusts
just as the crops planted late
were beginning
to sprout.
(The crops planted late
sprout after
the royal
harvest. )
\VS{2}When
they had completely
consumed
the earth’s
vegetation,
I said,
\par }{\Q “Sovereign

{\ND{Lord}}, forgive
Israel!

\par }{\Q How can
Jacob
survive?

\par }{\Q He is too weak!”
\par }{\Q \VS{3}The
{\ND{Lord}}
decided
not to do this. “It will not
happen,”
the {\ND{Lord}}
said.
\par }{\PP \VS{4}The sovereign

{\ND{Lord}}
showed me this: I saw
the sovereign

{\ND{Lord}}
summoning
a shower
of fire.
It consumed
the great
deep
and devoured
the fields.
\par }{\Q \VS{5}I said,
“Sovereign

{\ND{Lord}}, stop!
\par }{\Q How
can
Jacob
survive?

\par }{\Q He is too weak!”
\par }{\Q \VS{6}The
{\ND{Lord}}
decided
not to do this.
The sovereign

{\ND{Lord}}
said,
“This will not
happen
either.”
\par }{\PP \VS{7}He showed
me this: I saw
the sovereign
One standing
by a tin
wall
holding tin
in his hand.
\VS{8}The
{\ND{Lord}}
said
to
me, “What
do you
see,
Amos?” I said,
“Tin.”
The sovereign
One then said,
\par }{\Q “Look,
I am about to place
tin
among
my people
Israel.
\par }{\Q I will no
longer
overlook their sin.
\par }{\Q \VS{9}Isaac’s
centers of worship
will become desolate;
\par }{\Q Israel’s
holy places
will be in
ruins.
\par }{\Q I will attack
Jeroboam’s
dynasty
with the sword.”
\par }{\SH Amos Confronts a Priest
\par }{\PP \VS{10}Amaziah
the priest
of Bethel
sent
this message to
King
Jeroboam
of Israel: “Amos
is conspiring against
you in
the very heart
of the kingdom
of Israel! The land
cannot
endure
all
his prophecies.
\VS{11}As a matter of fact,
Amos
is saying
this: ‘Jeroboam
will die
by the sword
and Israel
will certainly be carried
into exile
away from its land.’ ”
\par }{\PP \VS{12}Amaziah
then said
to
Amos,
“Leave,
you visionary! Run away
to
the land
of Judah! Earn
your living
and prophesy
there!
\VS{13}Don’t
prophesy
at Bethel
any longer,
for
a royal
temple
and palace are here!”
\par }{\PP \VS{14}Amos
replied
to Amaziah,
“I
was not
a prophet
by profession.
No,
I
was a herdsman
who also took care
of sycamore fig trees.
\VS{15}Then the
{\ND{Lord}}
took
me
from tending
flocks
and gave me this
commission, ‘Go! Prophesy
to
my people
Israel!’
\VS{16}So now
listen
to the
{\ND{Lord}}’s
message! You
say,
‘Don’t
prophesy
against
Israel! Don’t
preach
against
the family
of Isaac!’
\par }{\Q \VS{17}“Therefore
this is what
the {\ND{Lord}}
says:

\par }{\Q ‘Your wife
will become a prostitute
in the streets

\par }{\Q and your sons
and daughters
will die violently.
\par }{\Q Your land
will be given to others

\par }{\Q and you
will die
in a foreign
land.
\par }{\Q Israel
will certainly be carried
into exile
away from its land.’ ”

\par }\Chap{8}{\PP \par }{\Q \VerseOne{1}The sovereign

{\ND{Lord}}
showed
me this: I saw
a basket
of summer fruit.
\par }{\PP \VS{2}He said,
“What
do you
see,
Amos?” I replied,
“A basket
of summer fruit.”
Then the
{\ND{Lord}}
said to me,
“The end
has come
for my people
Israel! I will no
longer
overlook their sins.
\par }{\Q \VS{3}The women singing
in the temple
will wail
in that day.”
\par }{\Q The sovereign

{\ND{Lord}}
is
speaking.
\par }{\Q “There will be many
corpses
littered
everywhere! Be quiet!”
\par }{\Q \VS{4}Listen
to this,
you who trample
the needy,
\par }{\Q and do away
with the destitute in the land.
\par }{\Q \VS{5}You say,
\par }{\Q “When
will the new moon
festival be over, so we can sell
grain?
\par }{\Q When will the Sabbath
end, so we can open
up the grain
bins?

\par }{\Q We’re eager to sell less
for a higher
price,
\par }{\Q and to cheat
the buyer with rigged
scales!
\par }{\Q \VS{6}We’re eager to trade
silver
for the poor,
\par }{\Q a pair of sandals
for the needy!
\par }{\Q We want to
mix in some chaff
with the grain!”
\par }{\Q \VS{7}The
{\ND{Lord}}
confirms this oath by
the arrogance
of Jacob:

\par }{\Q “I swear
I will never
forget
all
you have done!
\par }{\Q \VS{8}Because
of this
the earth
will quake,
\par }{\Q and all
who live
in it will mourn.
\par }{\Q The whole
earth will rise
like the River
Nile,

\par }{\Q it will surge upward
and then grow calm,
like the Nile
in Egypt.
\par }{\Q \VS{9}In that day,”
says
the sovereign

{\ND{Lord}}, “I will make the sun
set
at noon,
\par }{\Q and make the earth
dark
in the middle
of the day.
\par }{\Q \VS{10}I will turn
your festivals
into funerals,
\par }{\Q and all
your songs
into funeral dirges.
\par }{\Q I will make
everyone
wear funeral clothes
\par }{\Q and cause every
head
to be shaved bald.
\par }{\Q I will make
you mourn
as if you had lost your only son;
\par }{\Q when it ends
it will indeed have been a bitter
day.
\par }{\Q \VS{11}Be certain of this,
the time
is coming,”
says
the sovereign

{\ND{Lord}},
\par }{\Q “when I will send
a famine
through the land
–
\par }{\Q not
a shortage
of food
or
water
\par }{\Q but
an end to divine revelation!
\par }{\Q \VS{12}People will stagger
from sea
to
sea,
\par }{\Q and from the north
around to
the east.
\par }{\Q They will wander about
looking
for a revelation
from the

{\ND{Lord}},
\par }{\Q but they will not
find any.
\par }{\Q \VS{13}In that day
your beautiful
young women
and your young men
will faint
from thirst.
\VS{14}These are the ones who now take oaths
in the name of the sinful
idol goddess of Samaria.
\par }{\Q They vow, ‘As surely
as your god
lives, O Dan,’
or ‘As surely as your beloved
one lives,
O Beer Sheba!’
\par }{\Q But they will fall down
and not
get up
again.”

\par }\Chap{9}{\PP \par }{\Q \VerseOne{1}I saw
the
sovereign
One standing
by the altar
and he said,
“Strike
the tops
of the support pillars,
so the thresholds
shake!
\par }{\Q Knock them
down on the heads
of all
the people,

\par }{\Q and I will kill
the survivors
with the sword.
\par }{\Q No
one will be able to run
away;
\par }{\Q no
one will be able to escape.
\par }{\Q \VS{2}Even if
they could dig
down into the netherworld,
\par }{\Q my hand
would pull
them up
from there.
\par }{\Q Even if
they could climb up
to heaven,
\par }{\Q I would drag them down
from there.
\par }{\Q \VS{3}Even if
they were to hide
on the top
of Mount Carmel,
\par }{\Q I would hunt
them down
and take
them from
there.
\par }{\Q Even if
they tried
to hide
from
me
at the bottom
of the sea,
\par }{\Q from
there
I
would command
the
Sea Serpent
to bite them.
\par }{\Q \VS{4}Even
when
their enemies
drive them into captivity,
\par }{\Q from there
I will command
the sword
to kill
them.
\par }{\Q I
will not let
them out of my sight;
\par }{\Q they will experience disaster,
not
prosperity.”
\par }{\Q \VS{5}The sovereign

{\ND{Lord}}
who commands armies
will do this.

\par }{\Q He touches
the earth
and it dissolves;
\par }{\Q all
who live
on it mourn.
\par }{\Q The whole
earth rises
like the River Nile,
\par }{\Q and then grows calm
like the Nile
in Egypt.
\par }{\Q \VS{6}He builds
the upper rooms
of his palace in heaven
\par }{\Q and sets
its foundation
supports
on
the earth.
\par }{\Q He summons
the water
of the sea
\par }{\Q and pours
it out
on
the earth’s
surface.
\par }{\Q The
{\ND{Lord}}
is his name.
\par }{\Q \VS{7}“You Israelites
are just like the Ethiopians
in my sight,” says
the {\ND{Lord}}.
\par }{\Q “Certainly I brought Israel
up
from the land
of Egypt,
\par }{\Q but I also brought the Philistines
from Caphtor
and the Arameans
from Kir.
\par }{\Q \VS{8}Look,
the sovereign

{\ND{Lord}}
is watching the sinful
nation,
\par }{\Q and I will destroy
it from the
face
of the earth.
\par }{\Q But
I will not
completely
destroy
the
family
of Jacob,”
says
the {\ND{Lord}}.
\par }{\Q \VS{9}“For
look,
I am
giving a command
\par }{\Q and I will shake
the
family
of Israel
together with all
the nations.
\par }{\Q It will resemble
a sieve
being shaken,
\par }{\Q when not
even a pebble
falls
to the ground.
\par }{\Q \VS{10}All
the sinners
among my people
will die
by the sword
–
\par }{\Q the ones who say,
‘Disaster
will not
come near,
it will not confront us.’
\par }{\SH The Restoration of the Davidic Dynasty
\par }{\Q \VS{11}“In that day
I will rebuild the collapsing
hut
of David.
\par }{\Q I will seal
its gaps,
\par }{\Q repair
its ruins,
\par }{\Q and restore
it to what it was like in days gone by.
\par }{\Q \VS{12}As a result
they will conquer
those left
in Edom
\par }{\Q and all
the nations
subject
to my rule.”
\par }{\Q The
{\ND{Lord}}, who is about to do
this,
is speaking!
\par }{\Q \VS{13}“Be sure of this,
the time
is coming,”
says
the {\ND{Lord}},
\par }{\Q “when the plowman
will catch up
to the reaper
\par }{\Q and the one who stomps
the grapes
will overtake
the planter.
\par }{\Q Juice
will run down
the slopes,
\par }{\Q it will flow down
all
the hillsides.
\par }{\Q \VS{14}I will bring back
my people,
Israel;
\par }{\Q they will rebuild
the cities
lying in rubble
and settle down.
\par }{\Q They will plant
vineyards
and drink
the
wine
they produce;
\par }{\Q they will grow orchards
and eat
the
fruit they produce.
\par }{\Q \VS{15}I will plant
them on
their land
\par }{\Q and they will never
again
be uprooted
from the land
I have
given
them,”
\par }{\PP says
the
{\ND{Lord}}
your God.
\par }