\PeriphTitle{ }
{\TINYIMT NET Bible\par}
{\TINYIMTB NET Bible®, New English Translation (NET)\par}
{\TINYIMTB \par}
{\TINY copyright © 1996-2016 \href{http://netbible.com}{Biblical Studies Press, L. L. C.}\vskip 1ex 
Please see \href{https://netbible.com/net-bible-copyright}{https://netbible.com/net-bible-copyright} for full NET Bible copyright and permissions information. You can also read the \href{http://netbible.org}{NET} Bible at \href{http://netbible.org}{http://NETBible.org}.\vskip 1ex 
Language: \href{http://www.ethnologue.org/language/eng}{English }\vskip 1ex 
Dialect (if applicable): American\vskip 1ex 
Language in English: English\vskip 1ex 
Translation by:\vskip 1ex 
\vskip 1ex 
Please see \href{https://netbible.com/net-bible-copyright}{https://netbible.com/net-bible-copyright} for full NET Bible copyright and permissions information. You can also read the \href{http://netbible.org}{NET} Bible at \href{http://netbible.org}{http://NETBible.org}.\vskip 1ex 
\par}{\TINY The \href{http://netbible.org}{NET Bible} (New English Translation) is a completely new translation of the Bible with \href{http://netbible.org}{58,506 translators’ notes}! It was completed by more than 25 scholars – experts in the original biblical languages – who worked directly from the best currently available Hebrew, Aramaic, and Greek texts. Turn the pages and see the breadth of the translators’ notes, documenting their decisions and choices as they worked.\par}
{\TINY The translators’ notes make the original languages far more accessible, allowing you to look over the translator’s shoulder at the very process of translation. This level of documentation is a first for a Bible translation, making transparent the textual basis and the rationale for key renderings (including major interpretive options and alternative translations).\par}
{\TINY This unparalleled level of detail helps connect people to the Bible in the original languages in a way never before possible without years of study of Hebrew, Aramaic, and Greek. It unlocks the riches of the Bible’s truth from entirely new perspectives.\par}
{\TINY A donation is appreciated. Donate at \href{https://bible.org/donations\_form}{Bible.org/donations\_form} so we can continue to provide our material for free for our 16,000+ articles on \href{https://bible.org/}{Bible.org} and our \href{https://lumina.bible.org/}{world class Bible Study tool} at \href{https://lumina.bible.org/}{lumina.bible.org}.\par}
{\TINY 2018-12-21\par}
\par\vskip 1ex\hrule\vskip 0.5ex\par {\TINY PDF generated using Haiola and XeLaTeX on 12 Mar 2019 from source files dated 22 Dec 2018\par  6910647c-3b09-5fc2-8919-336c085b8702}
\vfill\eject
