\NormalFont\ShortTitle{Deuteronomy}
{\MT Deuteronomy

\par }\ChapOne{1}{\SH The Covenant Setting
\par }{\PP \VerseOne{1}This
is what
Moses
said
to
the assembly
of Israel
in the Transjordanian
wastelands,
the arid country
opposite
Suph,
between
Paran
and Tophel,
Laban,
Hazeroth,
and Di Zahab
\VS{2}Now it is ordinarily an eleven-day
journey from Horeb
to Kadesh Barnea
by way
of Mount
Seir.
\VS{3}However, it was not until the first
day
of the eleventh
month
of the fortieth
year
that Moses
addressed
the Israelites
just
as the
{\ND{Lord}}
had instructed
him to do.
\VS{4}This took place after
the defeat
of King
Sihon
of the Amorites,
whose
capital
was in
Heshbon,
and King
Og
of Bashan,
whose
capital
was in
Ashtaroth,
specifically in Edrei.
\VS{5}So it was in the Transjordan,
in Moab,
that Moses
began
to deliver
these
words:
\par }{\SH Events at Horeb
\par }{\PP \VS{6}The
{\ND{Lord}}
our God
spoke
to
us at Horeb
and said,
“You have stayed
in the area of this
mountain
long enough.
\VS{7}Get up now, resume
your journey,
heading
for the Amorite
hill country,
to
all
its areas
including the arid country,
the highlands,
the Shephelah,
the Negev,
and the coastal plain
– all
of Canaan
and Lebanon
as far as
the Great
River,
that is, the Euphrates.
\VS{8}Look! I have already given
the
land
to you. Go,
occupy
the territory
that
I, the
{\ND{Lord}}, promised
to give to your ancestors
Abraham,
Isaac,
and Jacob,
and to their descendants.”
\VS{9}I also said
to
you at that time,
“I am no
longer able
to sustain
you by myself.
\VS{10}The
{\ND{Lord}}
your God
has increased
your population
to the point that
you are now
as numerous
as the very stars
of the sky.
\VS{11}Indeed, may the
{\ND{Lord}}, the God
of your ancestors,
make
you a thousand
times
more numerous than you are now, blessing
you just
as he said he would!
\VS{12}But how
can I alone
bear
up under the burden
of your hardship
and strife?
\VS{13}Select wise
and practical
men,
those known
among your tribes,
whom
I may
appoint
as your leaders.”
\VS{14}You replied
to me that what
I had said
to you was good.
\VS{15}So I chose
as your tribal
leaders
wise
and well-known
men,
placing
them over you as administrators
of groups of thousands,
hundreds,
fifties,
and tens,
and also as other tribal
officials.
\VS{16}I
furthermore admonished
your judges
at that time
that
they should
pay attention
to issues among
your fellow
citizens and judge
fairly,
whether
between
one
citizen and another
or a citizen and a resident foreigner.
\VS{17}They must not
discriminate
in judgment,
but hear
the lowly
and the great
alike. Nor
should they be intimidated
by human beings,
for
judgment
belongs to God.
If
the matter
being adjudicated is too
difficult
for them,
they should bring it before me
for a hearing.
\par }{\SH Instructions at Kadesh Barnea
\par }{\PP \VS{18}So I instructed
you at that time
regarding everything
you should do.
\VS{19}Then we left
Horeb
and passed through
all
that immense,
forbidding
wilderness
that
you saw
on the way
to the Amorite
hill country
as the
{\ND{Lord}}
our God
had commanded
us to do, finally
arriving
at Kadesh Barnea.
\VS{20}Then I said
to
you, “You have come
to
the Amorite
hill country
which
the {\ND{Lord}}
our God
is about to give us.
\VS{21}Look,
he has placed
the land
in front
of you! Go up,
take possession
of it, just as
the {\ND{Lord}}, the God
of your ancestors,
said
to do. Do
not
be afraid
or
discouraged!”
\VS{22}So all
of you approached
me
and said,
“Let’s send
some men
ahead
of us to scout
out the land
and bring
us back
word
as to how we should
attack
it and
what the
cities
are like
there.”
\VS{23}I thought this was a good
idea, so
I sent twelve
men
from among you, one
from each
tribe.
\VS{24}They left
and went up
to the hill country,
coming
to the Eshcol
Valley,
which they scouted out.
\VS{25}Then they took
some of the produce
of the land
and carried it back
down
to
us. They also brought a report
to us, saying,
“The land
that
the {\ND{Lord}}
our God
is about to give
us is good.”
\par }{\SH Disobedience at Kadesh Barnea
\par }{\PP \VS{26}You were not
willing
to go up,
however, but instead rebelled
against the
{\ND{Lord}}
your God.
\VS{27}You complained
among yourselves privately
and said,
“Because the
{\ND{Lord}}
hates
us he brought
us from Egypt
to deliver
us over
to the Amorites
so they could destroy us!
\VS{28}What
is going to happen to us? Our
brothers
have drained
away our courage
by describing
people
who are more
numerous and taller
than
we are, and great
cities
whose defenses
appear to be as high
as heaven
itself! Moreover,
they said
they saw
Anakites
there.”
\VS{29}So I responded to
you, “Do not
be
terrified
of them!
\VS{30}The
{\ND{Lord}}
your God
is about to go
ahead
of you; he
will fight
for you, just
as you saw him
do
in Egypt
\VS{31}and in the desert,
where
you saw
him
carrying
you along like
a man
carries
his son.
This he did
everywhere
you went
until
you came
to
this
very
place.”
\VS{32}However, through all this
you did not
have confidence
in the
{\ND{Lord}}
your God,
\VS{33}the one who was constantly
going
before
you to find places
for you to set up camp.
He appeared
by fire
at night
and cloud
by day,
to show
you the way
you ought to go.
\par }{\SH Judgment at Kadesh Barnea
\par }{\PP \VS{34}When
the {\ND{Lord}}
heard
you, he became
angry
and made this vow:
\VS{35}“Not a single
person
of this
evil
generation
will see
the good
land
that
I promised
to give
to your ancestors!
\VS{36}The exception
is Caleb
son
of Jephunneh;
he
will see
it and I will give
him and his descendants
the territory
on which
he has
walked,
because
he has
wholeheartedly
followed me.”
\VS{37}As for me, the
{\ND{Lord}}
was also
angry
with me on your account.
He said,
“You
also
will not
be able to go
there.
\VS{38}However, Joshua
son
of Nun,
your assistant,
will go.
Encourage
him, because
he will enable Israel
to inherit the land.
\VS{39}Also, your infants,
who
you thought
would die on the way,
and your children,
who
as yet
do not
know
good
from bad,
will go
there;
I will give
them the land and they
will possess it.
\VS{40}But as for you,
turn
back and head
for the desert
by the way
to the Red
Sea.”
\par }{\SH Unsuccessful Conquest of Canaan
\par }{\PP \VS{41}Then you responded
to me
and admitted,
“We have sinned
against the
{\ND{Lord}}. We
will now go up
and fight
as
the {\ND{Lord}}
our God
has told
us to do.” So you each
put on your
battle
gear
and prepared to go up
to the hill country.
\VS{42}But the
{\ND{Lord}}
told me: “Tell
them this: ‘Do not
go up
and fight,
because
I will not
be with you and you will be defeated
by your enemies.’ ”
\VS{43}I spoke
to
you, but you did not
listen.
Instead you rebelled
against the
{\ND{Lord}}
and recklessly
went up
to the hill country.
\VS{44}The Amorite
inhabitants
of that
area
confronted
you and chased
you like a swarm
of bees,
striking
you down from Seir
as far
as Hormah.
\VS{45}Then you came back
and wept
before
the {\ND{Lord}}, but he
paid
no
attention
to you whatsoever.
\VS{46}Therefore, you remained
at Kadesh
for a long
time
– indeed, for the full
time.

\par }\Chap{2}{\PP \VerseOne{1}Then we turned
and set
out toward the desert
land on the way
to the Red
Sea
just
as the
{\ND{Lord}}
told
me
to do, detouring
around Mount
Seir
for a long time.
\VS{2}At this point the
{\ND{Lord}}
said
to me,
\VS{3}“You have circled
around
this
mountain
long enough; now
turn
north.
\VS{4}Instruct these people
as follows: ‘You
are about to cross
the border
of your relatives
the descendants
of Esau,
who inhabit
Seir.
They will be afraid
of
you, so watch
yourselves carefully.
\VS{5}Do not
be hostile
toward them, because
I am not
giving
you any of their land,
not even
a footprint,
for
I have given
Mount
Seir
as an inheritance
for Esau.
\VS{6}You may purchase
food
to eat
and water
to drink from them.
\VS{7}All
along
the way
I,
the {\ND{Lord}}
your God,
have blessed
your
every
effort.
I have been attentive to your travels
through
this
great
wasteland.
These
forty
years
I have been with
you; you have lacked
for nothing.’ ”
\par }{\PP \VS{8}So we turned away
from
our relatives
the descendants
of Esau,
the inhabitants
of Seir,
turning
from the desert
route,
from Elat
and Ezion Geber,
and traveling
the way
of the Moab
wastelands.
\VS{9}Then the
{\ND{Lord}}
said to me,
“Do not
harass
Moab
and provoke
them to war,
for
I will not
give
you any of their land
as your territory.
This is because
I have given
Ar
to the descendants
of Lot
as their possession.
\VS{10}(The Emites
used to live
there, a people
as powerful,
numerous,
and tall
as the Anakites.
\VS{11}These
people, as well as
the Anakites,
are also
considered
Rephaites;
the Moabites
call
them Emites.
\VS{12}Previously the Horites
lived
in Seir
but the descendants of
Esau
dispossessed
and destroyed
them and settled
in their place,
just
as Israel
did
to the land
it came to possess,
the land
the {\ND{Lord}}
gave them.)
\VS{13}Now,
get up
and cross
the Wadi
Zered.” So we did so.
\VS{14}Now the length of time
it took for us to go
from Kadesh Barnea
to
the
crossing
of Wadi
Zered
was thirty-eight
years,
time
for all
the military
men
of that generation
to die,
just as
the {\ND{Lord}}
had vowed to them.
\VS{15}Indeed,
it was
the very hand
of the {\ND{Lord}}
that eliminated
them from within
the camp
until they were all gone.
\par }{\SH Instructions Concerning Ammon
\par }{\PP \VS{16}So it was
that
after all
the military men had been eliminated from the community,
\VS{17}the {\ND{Lord}}
said
to me,
\VS{18}“Today
you
are going to cross
the border
of Moab,
that is, of Ar.
\VS{19}But when you come
close
to the Ammonites,
do not
harass
or
provoke
them because
I am not
giving
you any of the Ammonites’
land
as your possession;
I have already given
it to Lot’s
descendants
as their possession.
\par }{\PP \VS{20}(That also
is considered
to be a land
of the Rephaites.
The Rephaites
lived
there originally;
the Ammonites
call
them Zamzummites.
\VS{21}They are a people
as powerful,
numerous,
and tall
as the Anakites.
But the
{\ND{Lord}}
destroyed
the Rephaites in advance of the Ammonites, so they dispossessed
them and settled
down in their place.
\VS{22}This
is exactly what
he did
for the descendants
of Esau
who lived
in Seir
when he destroyed
the
Horites
before
them so that
they could dispossess
them and settle in
their area
to
this very
day.
\VS{23}As for the Avvites
who lived
in settlements
as far
west as Gaza,
Caphtorites
who came from
Crete
destroyed
them and settled down
in their place.)
\par }{\PP \VS{24}Get
up, make
your way across
Wadi
Arnon.
Look! I have already delivered
over
to you Sihon
the Amorite,
king
of Heshbon,
and his land.
Go
ahead! Take
it! Engage
him in war!
\VS{25}This
very day
I will begin
to fill
all
the people
of the earth
with dread
and to terrify
them when they hear
about
you. They will shiver
and shake in anticipation
of your approach.”
\par }{\SH Defeat of Sihon, King of Heshbon
\par }{\PP \VS{26}Then I sent
messengers
from the Kedemoth
Desert
to
King
Sihon
of Heshbon
with an offer
of peace:
\VS{27}“Let me pass
through your land;
I will keep strictly
to the roadway. I will not
turn aside
to the right
or the left.
\VS{28}Sell me food
for cash
so that I can eat
and sell
me water
to drink.
Just allow
me to go through
on foot,
\VS{29}just
as the descendants
of Esau
who live
at Seir
and the Moabites
who live
in Ar
did
for me, until
I cross
the Jordan
to
the land
the {\ND{Lord}}
our God
is giving us.”
\VS{30}But King
Sihon
of Heshbon
was unwilling
to allow us to pass
near him because
the {\ND{Lord}}
our God
had made him obstinate
and stubborn
so that
he might deliver
him over to you
this
very day.
\VS{31}The
{\ND{Lord}}
said
to
me, “Look! I have already begun
to give
over Sihon
and his land
to you. Start
right now to take
his land
as your possession.”
\VS{32}When
Sihon
and all
his troops
emerged
to encounter us in battle
at Jahaz,
\VS{33}the {\ND{Lord}}
our God
delivered
him over to us and we struck
him down, along
with his sons
and everyone else.
\VS{34}At that time
we
seized
all
his cities
and put
every
one of them under divine
judgment, including even the women
and children;
we left
no survivors.
\VS{35}We kept
only
the livestock
and plunder
from the cities
for ourselves.
\VS{36}From Aroer,
which
is at the edge
of Wadi
Arnon
(it is the city
in the wadi), all the way
to Gilead
there was
not
a town
able to resist
us
– the

{\ND{Lord}}
our God
gave
them all to us.
\VS{37}However,
you did not
approach
the land
of the Ammonites,
the Wadi
Jabbok,
the cities
of the hill country,
or any
place else forbidden
by the
{\ND{Lord}}
our God.

\par }\Chap{3}{\PP \VerseOne{1}Next
we set out
on
the route
to Bashan,
but King
Og
of Bashan
and his whole
army
came out
to meet us
in battle
at Edrei.
\VS{2}The
{\ND{Lord}}, however, said
to
me, “Don’t
be afraid
of him
because
I have already
given
him,
his whole
army,
and his land
to you. You will do
to him exactly what
you did
to King
Sihon
of the Amorites
who
lived
in Heshbon.”
\VS{3}So
the {\ND{Lord}}
our God
did indeed
give over
to us King
Og
of Bashan
and his whole
army
and we struck
them down until
not
a single survivor
was left.
\VS{4}We captured
all
his cities
at that time
– there was
not
a town
we did not
take from them – sixty cities, all the region of Argob, the dominion of Og in Bashan.
\VS{5}All
of these
cities
were fortified
by high
walls,
gates,
and locking bars;
in addition
there were a great many
open villages.
\VS{6}We put
all of these under divine
judgment just
as we
had
done
to King
Sihon
of Heshbon
– every
occupied
city,
including women
and children.
\VS{7}But all
the livestock
and plunder
from the cities
we kept for ourselves.
\VS{8}So at that time
we took
the land
of the two
Amorite
kings
in the Transjordan
from Wadi
Arnon
to Mount
Hermon
\VS{9}(the Sidonians
call
Hermon
Sirion
and the Amorites
call
it Senir),
\VS{10}all
the cities
of the plateau,
all
of Gilead
and Bashan
as far
as Salecah
and Edrei,
cities
of the kingdom
of Og
in Bashan.
\VS{11}Only
King
Og
of Bashan
was left
of the remaining
Rephaites.
(It is noteworthy
that his sarcophagus
was made of iron.
Does it not,
indeed, still remain in Rabbath
of the Ammonites? It is thirteen and a half feet
long
and six feet
wide
according to standard measure.)
\par }{\SH Distribution of the Transjordanian Allotments
\par }{\PP \VS{12}This
is the land
we brought under our control
at that time: The territory extending from Aroer
by
the Wadi
Arnon
and half
the Gilead
hill country
with its cities
I gave
to the Reubenites
and Gadites.
\VS{13}The rest
of Gilead
and all
of Bashan,
the kingdom
of Og,
I gave
to half
the tribe
of Manasseh.
(All
the region of Argob,
that is, all
Bashan,
is called
the land
of Rephaim.
\VS{14}Jair,
son
of Manasseh,
took
all
the Argob
region as far
as the border
with the Geshurites
and Maacathites
(namely Bashan) and called
it by his name,
Havvoth-Jair,
which it retains to
this
very
day.)
\VS{15}I gave
Gilead
to Machir.
\VS{16}To the Reubenites
and Gadites
I allocated
the territory extending from
Gilead
as far
as Wadi
Arnon
(the exact middle
of the wadi
was a boundary) all
the way
to the Wadi
Jabbok,
the Ammonite
border.
\VS{17}The Arabah
and the Jordan River
were also a border,
from the sea of Chinnereth
to
the sea
of the Arabah
(that is, the Salt
Sea), beneath
the watershed
of Pisgah
to the east.
\par }{\SH Instructions to the Transjordanian Tribes
\par }{\PP \VS{18}At that time
I instructed
you as follows: “The
{\ND{Lord}}
your God
has given
you this
land
for your possession.
You warriors
are to cross
over before
your fellow
Israelites
equipped for battle.
\VS{19}But
your wives,
children,
and livestock
(of which I know
you have many) may remain
in the cities
I have
given you.
\VS{20}You must fight until
the {\ND{Lord}}
gives
your countrymen
victory
as he did
you and they
take possession
of the land
that
the {\ND{Lord}}
your God
is giving
them
on the other side
of the Jordan River.
Then each
of you may return
to his own territory
that
I have given you.”
\VS{21}I also
commanded
Joshua
at the same time,
“You have seen
everything
the {\ND{Lord}}
your God
did
to these
two
kings;
he will do
the same
to all
the kingdoms
where
you
are going.
\VS{22}Do not
be afraid
of them, for
the {\ND{Lord}}
your God
will personally fight for you.”
\par }{\SH Denial to Moses of the Promised Land
\par }{\PP \VS{23}Moreover, at that time
I pleaded
with the
{\ND{Lord}},
\VS{24}“O, Lord

{\ND{God}}, you
have begun
to show
me
your greatness
and strength.
(What
god
in heaven
or earth
can
rival your works
and mighty deeds?)
\VS{25}Let me please
cross
over to see
the good
land
on the other side
of the Jordan River
– this
good
hill country
and the Lebanon!”
\VS{26}But the
{\ND{Lord}}
was angry
at me because
of you and would not
listen
to me.
Instead, he said
to
me, “Enough
of that! Do not
speak
to me
anymore
about this
matter.
\VS{27}Go up
to the top
of Pisgah
and take
a good look
to the west,
north,
south,
and east,
for
you will not
be allowed to cross
the
Jordan.
\VS{28}Commission
Joshua,
and encourage
and strengthen
him, because
he will lead
these
people
over and will enable them to inherit
the land
you will see.”
\VS{29}So we settled
down in the valley
opposite
Beth Peor.

\par }\Chap{4}{\PP \VerseOne{1}Now,
Israel,
pay attention
to
the statutes
and ordinances
I am
about to teach
you, so that you might
live
and go on to enter
and take possession
of the
land
that
the {\ND{Lord}}, the God
of your ancestors,
is giving you.
\VS{2}Do not
add
a thing
to what
I
command
you nor
subtract
from
it, so that you may keep
the
commandments
of the {\ND{Lord}}
your God
that
I am
delivering to you.
\VS{3}You have witnessed
what
the
{\ND{Lord}}
did
at Baal Peor,
how he eradicated
from your midst
everyone
who
followed
Baal Peor.
\VS{4}But you
who remained
faithful to the
{\ND{Lord}}
your God
are still alive
to this very day,
every one of you.
\VS{5}Look! I have taught
you statutes
and ordinances
just
as the
{\ND{Lord}}
my God
told
me to do,
so
that
you might carry
them out in the land
you
are about to enter
and possess.
\VS{6}So be sure
to do
them, because
this
will testify of your wise
understanding
to the people
who
will learn
of all
these
statutes
and say,
“Indeed, this
great
nation
is a very
wise
people.”
\VS{7}In fact, what other
great
nation
has
a god
so near
to
them like the
{\ND{Lord}}
our God
whenever
we call
on him?
\VS{8}And what other
great
nation
has
statutes
and ordinances
as just
as this
whole
law
that
I am
about to share
with you today?
\par }{\SH Reminder of the Horeb Covenant
\par }{\PP \VS{9}Again, however,
pay
very
careful
attention,
lest
you forget
the things
you have
seen
and disregard
them for the rest
of your life;
instead teach
them to your children
and grandchildren.
\VS{10}You
stood
before
the {\ND{Lord}}
your God
at Horeb
and he said
to
me, “Assemble
the
people
before me so
that
I can tell them my commands. Then they will learn
to revere
me all
the days
they
live
in the land,
and
they will instruct
their children.”
\VS{11}You approached
and stood
at the foot
of the mountain,
a mountain
ablaze
to
the sky
above
it and yet dark
with a thick
cloud.
\VS{12}Then
the {\ND{Lord}}
spoke to
you from the middle
of the fire;
you
heard
speech
but you could not
see anything – only a voice was heard.
\VS{13}And he revealed
to you the covenant
he has
commanded
you to keep,
the ten
commandments,
writing
them on
two
stone
tablets.
\VS{14}Moreover, at that same time
the {\ND{Lord}}
commanded
me to teach
you statutes
and ordinances
for you to keep
in the land
which
you
are about
to enter
and possess.
\par }{\SH The Nature of Israel’s God
\par }{\PP \VS{15}Be very
careful,
then, because
you saw
no
form
at the time
the {\ND{Lord}}
spoke
to
you at Horeb
from the middle
of the fire.
\VS{16}I say
this so
you will not corrupt
yourselves by making
an image
in the form
of any
kind of figure.
This includes the likeness
of a human male
or
female,
\VS{17}any
kind
of land
animal,
any
bird
that
flies
in the sky,
\VS{18}anything
that crawls
on the ground,
or any
fish
in the deep waters
of the earth.
\VS{19}When you look up
to the sky
and see
the sun,
moon,
and stars
– the whole
heavenly
creation –
you must not be seduced
to worship
and serve
them, for the
{\ND{Lord}}
your God
has assigned
them to all
the people
of the world.
\VS{20}You, however, the
{\ND{Lord}}
has selected
and brought
from Egypt,
that iron-smelting
furnace,
to be
his special
people
as you are today.
\VS{21}But the
{\ND{Lord}}
became angry
with me because
of you and vowed
that I would never
cross
the Jordan
nor
enter
the good
land
that
he
is about to give
you.
\VS{22}So I
must die
here
in this
land;
I will not
cross
the Jordan.
But you
are going
over and will possess
that good
land.
\VS{23}Be on guard
so
that you do not forget
the covenant
of the {\ND{Lord}}
your God
that
he has made
with
you, and that you do
not make an image
of any kind,
just as
he
has forbidden you.
\VS{24}For
the {\ND{Lord}}
your God
is a consuming
fire;
he is
a jealous
God.
\par }{\SH Threat and Blessing following Covenant Disobedience
\par }{\PP \VS{25}After
you have produced
children
and grandchildren
and have been in the land
a long time,
if you become corrupt
and make
an image
of any kind
and do
other evil things
before
the {\ND{Lord}}
your God
that enrage him,
\VS{26}I invoke
heaven
and earth
as
witnesses against you today
that
you will surely
and swiftly
be removed
from the very land
you
are about to cross
the Jordan
to possess.
You will not
last long
there
because
you will surely be annihilated.
\VS{27}Then the
{\ND{Lord}}
will scatter
you among the
peoples
and there
will be very few
of you among the nations
where
the {\ND{Lord}}
will drive you.
\VS{28}There
you will worship
gods
made
by human
hands
– wood
and stone
that
can neither
see,
hear,
eat,
nor
smell.
\VS{29}But if you seek
the {\ND{Lord}}
your God
from there,
you will find
him, if, indeed,
you seek
him with all
your heart
and soul.
\VS{30}In your distress
when all
these
things
happen
to you in the latter
days,
if you return
to
the {\ND{Lord}}
your God
and obey
him
\VS{31}(for
he is a merciful
God), he will not
let
you down or
destroy
you, for he cannot
forget
the
covenant
with your ancestors
that
he confirmed by oath to them.
\par }{\SH The Uniqueness of Israel’s God
\par }{\PP \VS{32}Indeed,
ask
about the distant
past,
starting
from
the day
God
created
humankind
on
the earth,
and ask from one end
of heaven
to
the other,
whether there has ever been such a great
thing
as this,
or
even a rumor of it.
\VS{33}Have a people
ever heard
the voice
of God
speaking
from the middle
of fire,
as
you yourselves
have, and lived
to tell about it?
\VS{34}Or
has God
ever before tried
to deliver
a nation
from the middle
of another nation,
accompanied by judgments,
signs,
wonders,
war,
strength, power,
and other very
terrifying
things
like the
{\ND{Lord}}
your God
did
for you in Egypt
before your very eyes?
\VS{35}You have been
taught
that
the {\ND{Lord}}
alone
is
God
– there is no
other
besides him.
\VS{36}From
heaven
he spoke
to you in order
to teach
you, and on
earth
he showed
you his great
fire
from
which you also
heard
his words.
\VS{37}Moreover, because
he loved
your ancestors,
he chose
their descendants
who followed
them and personally brought
you out
of Egypt
with his great
power
\VS{38}to dispossess
nations
greater
and stronger
than
you and brought
you here this
day
to give
you their land
as your property.
\VS{39}Today
realize
and carefully
consider
that
the {\ND{Lord}}
is God
in heaven
above
and on
earth
below
– there is no
other!
\VS{40}Keep
his statutes
and commandments
that
I
am setting forth
today
so that
it may go well
with you and your descendants
and that
you may enjoy longevity
in
the land
that
the {\ND{Lord}}
your God
is about to give
you as a permanent possession.
\par }{\SH The Narrative Concerning Cities of Refuge
\par }{\PP \VS{41}Then
Moses
selected
three
cities
in the Transjordan,
toward the east.
\VS{42}Anyone
who accidentally
killed someone without
hating
him at the time of the accident
could flee
to
one
of those
cities
and be safe.
\VS{43}These cities are Bezer,
in the desert
plateau,
for the Reubenites;
Ramoth
in Gilead
for the Gadites;
and Golan
in Bashan
for the Manassehites.
\par }{\SH The Setting and Introduction of the Covenant
\par }{\PP \VS{44}This
is the law
that
Moses
set
before
the Israelites.
\VS{45}These
are the stipulations,
statutes,
and ordinances
that
Moses
spoke
to
the Israelites
after he had brought them out
of Egypt,
\VS{46}in the Transjordan,
in the valley
opposite
Beth Peor,
in the land
of King
Sihon
of the Amorites,
who
lived
in Heshbon.
(It is he whom
Moses
and the Israelites
attacked
after they came out
of Egypt.
\VS{47}They possessed
his land
and that of King
Og
of Bashan
– both
of whom
were Amorite
kings
in the Transjordan,
to the east.
\VS{48}Their territory extended from Aroer
at the edge
of the Arnon
valley
as far
as Mount
Siyon –
that
is, Hermon –
\VS{49}including all
the Arabah
of the Transjordan
in the east
to
the sea
of the Arabah,
beneath
the watershed
of Pisgah.)

\par }\Chap{5}{\PP \VerseOne{1}Then Moses
called
all
the people of Israel
together and said
to
them: “Listen,
Israel,
to the statutes
and ordinances
that
I
am about to deliver
to you today;
learn
them and be careful
to keep them!
\VS{2}The
{\ND{Lord}}
our God
made
a covenant
with us at Horeb.
\VS{3}He did not
make
this
covenant
with
our ancestors
but
with
us, we
who are
here
today,
all
of us living now.
\VS{4}The
{\ND{Lord}}
spoke
face
to face
with
you at the mountain,
from the middle
of the fire.
\VS{5}(I
was standing
between
the {\ND{Lord}}
and you at that time
to reveal
to you the
message
of the {\ND{Lord}}, because
you were afraid
of the fire
and would not
go up
the mountain.) He said:
\par }{\SH The Ten Commandments
\par }{\PP \VS{6}“I am
the {\ND{Lord}}
your God,
he who
brought
you from the land
of Egypt,
from the place
of slavery.
\VS{7}You must not
have any other
gods
besides me.
\VS{8}You must not
make
for yourself an image
of anything
in heaven
above,
on earth
below,
or in the waters
beneath.
\VS{9}You must not
worship
or serve
them, for
I,
the {\ND{Lord}}
your God,
am a jealous
God.
I punish
the sons,
grandsons, and great-grandsons
for the sin
of the fathers
who reject
me,
\VS{10}but I show
covenant faithfulness
to the thousands
who choose
me and keep
my commandments.
\VS{11}You must not
make use
of the name
of the {\ND{Lord}}
your God
for worthless
purposes, for
the {\ND{Lord}}
will not
exonerate
anyone who
abuses
his name
that way.
\VS{12}Be careful
to observe
the Sabbath
day
just
as the
{\ND{Lord}}
your God
has commanded you.
\VS{13}You are to work
and do
all
your tasks
in six
days,
\VS{14}but the seventh
day
is the Sabbath
of the {\ND{Lord}}
your God.
On that day you must not
do
any
work,
you,
your son,
your daughter,
your male slave,
your female slave,
your ox,
your donkey,
any other
animal,
or the foreigner
who
lives with you, so that
your male
and female
slaves,
like yourself,
may have rest.
\VS{15}Recall
that
you were slaves
in the land
of Egypt
and that the
{\ND{Lord}}
your God
brought
you out
of there
by
strength
and power.
That is why
the {\ND{Lord}}
your God
has
commanded
you to observe
the Sabbath
day.
\VS{16}Honor
your father
and your mother
just
as the
{\ND{Lord}}
your God
has commanded
you to do, so that
your days
may be extended
and that
it may go well
with you in
the land
that
he
is about to give you.
\VS{17}You must not
murder.
\VS{18}You must not
commit adultery.
\VS{19}You must not
steal.
\VS{20}You must not
offer
false
testimony
against another.
\VS{21}You must not
desire another
man’s wife,
nor
should you crave
his
house,
his
field,
his male
and female
servants,
his ox,
his donkey,
or anything
else
he owns.”
\par }{\SH The Narrative of the Sinai Revelation and Israel’s Response
\par }{\PP \VS{22}The

{\ND{Lord}}
said
these
things
to
your entire
assembly
at the mountain
from the middle
of the fire,
the cloud,
and the darkness
with a loud
voice,
and that was all
he said.
Then
he inscribed
the words
on
two
stone
tablets
and gave
them to me.
\VS{23}Then, when
you heard
the voice
from the midst
of the darkness
while the mountain
was ablaze,
all
your tribal
leaders
and elders
approached
me.
\VS{24}You said,
“The
{\ND{Lord}}
our God
has shown
us his great
glory
and we have heard
him
speak from the middle
of the fire.
It is now
clear
to us
that
God
can
speak
to human
beings
and they can keep
on living.
\VS{25}But now,
why
should we die,
because
this
intense
fire
will consume
us! If
we
keep
hearing
the voice
of the {\ND{Lord}}
our God
we will
die!
\VS{26}Who
is there from the entire
human race
who has
heard
the voice
of the living
God
speaking
from the middle
of the fire
as
we have, and has lived?
\VS{27}You go near
so that you
can hear
everything
the {\ND{Lord}}
our God
is saying
and then you
can tell
us
whatever
he says
to
you; then we will pay attention
and do it.”
\VS{28}When
the {\ND{Lord}}
heard
you speaking
to me,
he said
to me,
“I have heard
what
these
people
have
said
to you – they have spoken well.
\VS{29}If only
it would
really be
their desire
to fear
me and obey
all
my commandments
in the future, so that
it may
go well
with them and their descendants
forever.
\VS{30}Go
and tell
them, ‘Return
to your tents!’
\VS{31}But as for you,
remain
here
with me
so I can declare
to
you all
the commandments,
statutes,
and ordinances
that
you are to teach
them, so
that they can
carry them out
in the land
I am
about to give
them.”
\VS{32}Be careful,
therefore, to do
exactly
what
the {\ND{Lord}}
your God
has commanded you; do not
turn
right
or left!
\VS{33}Walk
just as he
has commanded
you so that
you may live,
that it may go well
with you, and that you may live long
in the land
you are going to possess.

\par }\Chap{6}{\PP \VerseOne{1}Now these
are the commandments,
statutes,
and ordinances
that
the {\ND{Lord}}
your God
instructed
me to teach
you so that
you may carry
them out in the land
where
you
are headed
\VS{2}and that
you may so revere
the {\ND{Lord}}
your God
that you will keep
all
his statutes
and commandments
that
I
am giving you – you, your children, and your grandchildren – all your lives, to prolong your days.
\VS{3}Pay attention,
Israel,
and be careful
to do
this so that
it may go well
with you and that
you may increase
greatly
in number –
as
the {\ND{Lord}}, God
of your ancestors,
said
to you, you will have a land
flowing
with milk
and honey.
\par }{\SH The Essence of the Covenant Principles
\par }{\PP \VS{4}Listen,
Israel: The
{\ND{Lord}}
is our God,
the {\ND{Lord}}
is one!
\VS{5}You must love
the {\ND{Lord}}
your God
with your whole
mind,
your whole
being,
and all
your strength.
\par }{\SH Exhortation to Teach the Covenant Principles
\par }{\PP \VS{6}These
words
I am
commanding
you today
must be kept in mind,
\VS{7}and you must teach
them to your children
and speak
of them as you sit
in your house,
as you walk
along the road,
as you lie
down, and as you get up.
\VS{8}You should tie
them as a reminder
on
your forearm
and fasten them as symbols
on your forehead.
\VS{9}Inscribe
them on
the doorframes
of your houses
and gates.
\par }{\SH Exhortation to Worship the
{\ND{Lord}} Exclusively
\par }{\PP \VS{10}Then when
the {\ND{Lord}}
your God
brings you to
the land
he promised
your ancestors
Abraham,
Isaac,
and Jacob
to give you – a land with large, fine cities you did not build,
\VS{11}houses
filled
with choice
things
you did not
accumulate,
hewn out
cisterns
you did not
dig,
and vineyards
and olive groves
you did not
plant
– and you eat
your fill,
\VS{12}be careful
not
to forget
the {\ND{Lord}}
who
brought
you out of Egypt,
that place
of slavery.
\VS{13}You must
revere
the {\ND{Lord}}
your God,
serve
him, and take oaths
using only his name.
\VS{14}You must not
go
after
other
gods,
those
of the surrounding
peoples,
\VS{15}for
the {\ND{Lord}}
your God,
who is present among
you,
is a jealous
God
and his anger
will erupt
against
you and remove
you from the land.
\par }{\SH Exhortation to Obey the
{\ND{Lord}} Exclusively
\par }{\PP \VS{16}You must not
put
the {\ND{Lord}}
your God
to the test
as you did at Massah.
\VS{17}Keep
his commandments
very carefully,
as well
as the stipulations
and statutes
he commanded you to observe.
\VS{18}Do
whatever is proper
and good
before
the {\ND{Lord}}
so that it may
go well
with you and that you may enter
and occupy
the good
land
that
he promised
your ancestors,
\VS{19}and that you may drive
out all
your enemies
just
as the
{\ND{Lord}}
said.
\par }{\SH Exhortation to Remember the Past
\par }{\PP \VS{20}When
your children
ask
you later
on, “What
are the stipulations,
statutes,
and ordinances
that
the {\ND{Lord}}
our God
commanded you?”
\VS{21}you must say
to them, “We were Pharaoh’s
slaves
in Egypt,
but the
{\ND{Lord}}
brought us out
of Egypt
in a powerful way.
\VS{22}And he brought signs
and great,
devastating
wonders
on
Egypt,
on Pharaoh,
and on his whole
family
before our very eyes.
\VS{23}He delivered
us from there
so that
he could give
us the land
he had
promised
our ancestors.
\VS{24}The
{\ND{Lord}}
commanded
us
to obey
all
these
statutes
and to revere
him so that
it may always
go well
for us and he may preserve
us, as he has to this
day.
\VS{25}We will be
innocent
if
we carefully
keep
all
these commandments
before
the {\ND{Lord}}
our God,
just
as he demands.”

\par }\Chap{7}{\PP \VerseOne{1}When
the {\ND{Lord}}
your God
brings
you to
the land
that
you
are going
to occupy
and forces out
many
nations
before you – Hittites, Girgashites, Amorites, Canaanites, Perizzites, Hivites, and Jebusites, seven nations more numerous and powerful than you –
\VS{2}and he delivers them over
to you and you attack
them, you must utterly
annihilate
them. Make
no
treaty
with them and show them no
mercy!
\VS{3}You must not
intermarry
with them. Do not
give
your daughters
to their sons
or
take
their daughters
for your sons,
\VS{4}for
they will turn
your sons
away
from
me to worship
other
gods.
Then the anger
of the {\ND{Lord}}
will erupt
against you and he will quickly
destroy you.
\VS{5}Instead,
this is what
you must do
to them: You must tear
down their altars, shatter
their sacred pillars,
cut down
their sacred Asherah poles,
and burn up
their idols.
\VS{6}For
you
are a people
holy
to the
{\ND{Lord}}
your God.
He has chosen
you to be
his people,
prized
above all
others
on
the face
of the earth.
\par }{\SH The Basis of Israel’s Election
\par }{\PP \VS{7}It is not
because
you were more numerous
than all
the other peoples
that
the {\ND{Lord}}
favored
and chose you – for in fact you were the least numerous of all peoples.
\VS{8}Rather
it is because of his love
for
you and his faithfulness to the promise he solemnly
vowed to your ancestors
that the
{\ND{Lord}}
brought you out
with great power,
redeeming
you from the place of slavery,
from the power
of Pharaoh
king
of Egypt.
\VS{9}So realize
that
the {\ND{Lord}}
your God
is the true God,
the faithful
God
who keeps
covenant
faithfully
with those who love
him and keep
his commandments,
to a thousand
generations,
\VS{10}but who pays
back those who hate
him as they deserve
and destroys
them. He will not
ignore
those who hate
him but will repay
them as they deserve!
\VS{11}So keep
the commandments,
statutes,
and ordinances
that
I today
am
commanding
you to do.
\par }{\SH Promises of Good for Covenant Obedience
\par }{\PP \VS{12}If
you obey
these
ordinances
and are careful
to do
them, the
{\ND{Lord}}
your God
will faithfully
keep
covenant
with you as
he promised
your ancestors.
\VS{13}He will love
and bless
you, and make you numerous.
He will bless
you with many children,
with the produce
of your soil,
your grain,
your new wine,
your oil,
the offspring
of your oxen,
and the young
of your flocks
in
the land
which
he promised
your ancestors
to give you.
\VS{14}You will be blessed
beyond
all
peoples;
there will be
no
barrenness
among you or your livestock.
\VS{15}The
{\ND{Lord}}
will protect
you from
all
sickness,
and you will not experience any
of the terrible
diseases
that
you knew
in Egypt;
instead he will inflict
them
on
all
those who hate you.
\par }{\SH Exhortation to Destroy Canaanite Paganism
\par }{\PP \VS{16}You must destroy
all
the people
whom the
{\ND{Lord}}
your God
is about to deliver
over to you; you must not
pity
them or
worship
their gods,
for
that will be a snare to you.
\VS{17}If
you think,
“These
nations
are more numerous
than I – how can I dispossess them?”
\VS{18}you must not
fear
them.
You must
carefully recall
what
the {\ND{Lord}}
your God
did
to Pharaoh
and all
Egypt,
\VS{19}the great
judgments
you saw,
the signs
and wonders,
the strength
and power
by which
he brought you out
– thus
the {\ND{Lord}}
your God
will do
to all
the people
you
fear.
\VS{20}Furthermore,
the {\ND{Lord}}
your God
will release
hornets
among them until
the very last ones who hide
from you perish.
\VS{21}You must not
tremble
in their presence,
for
the {\ND{Lord}}
your God,
who is present among
you, is a great
and awesome
God.
\VS{22}He, the God
who leads you, will expel
the
nations
little
by little.
You will not
be allowed
to destroy
them all at once
lest
the wild
animals
overrun you.
\VS{23}The
{\ND{Lord}}
your God
will give
them over
to you; he will throw
them into a great
panic
until
they are destroyed.
\VS{24}He will hand
over
their kings
to you and you will erase
their very names
from
memory.
Nobody
will be able to resist
you until
you destroy them.
\VS{25}You must burn
the images
of their gods,
but do not
covet
the silver
and gold
that covers them
so
much that you take
it for yourself and thus become ensnared
by it; for
it is abhorrent
to the
{\ND{Lord}}
your God.
\VS{26}You must not
bring
any abhorrent
thing into
your house
and thereby become an object
of divine
wrath along with it. You must absolutely detest
and abhor
it, for
it is an object of divine wrath.

\par }\Chap{8}{\PP \VerseOne{1}You must keep
carefully
all
these commandments
I am
giving
you today
so that you may
live,
increase
in number, and go
in and occupy
the land
that
the {\ND{Lord}}
promised
to your ancestors.
\VS{2}Remember
the
whole
way
by which
he
has brought
you these
forty
years
through the desert
so that
he might, by humbling
you, test
you to see if
you have
it within
you to keep
his commandments
or
not.
\VS{3}So he humbled
you by making you hungry
and then feeding
you with unfamiliar
manna.
He did this to teach
you that
humankind
cannot
live
by
bread
alone,
but also
by
everything
that comes
from the
{\ND{Lord}}’s
mouth.
\VS{4}Your clothing
did not
wear out
nor did your feet
swell
all these
forty
years.
\VS{5}Be keenly
aware
that
just as
a parent
disciplines
his child,
the {\ND{Lord}}
your God
disciplines you.
\VS{6}So you must keep
his commandments,
live according to his standards, and revere him.
\VS{7}For
the {\ND{Lord}}
your God
is bringing
you to
a good
land,
a land
of brooks,
springs,
and fountains
flowing forth
in valleys
and hills,
\VS{8}a land
of wheat,
barley,
vines,
fig trees,
and pomegranates,
of olive trees
and honey,
\VS{9}a land
where
you may eat
food
in plenty
and find no
lack
of anything,
a land
whose
stones
are iron
and from whose hills
you can mine
copper.
\VS{10}You will eat
your fill
and then praise
the {\ND{Lord}}
your God
because
of the good
land
he has
given you.
\par }{\SH Exhortation to Remember That Blessing Comes from God
\par }{\PP \VS{11}Be sure
you do not
forget
the {\ND{Lord}}
your God
by not
keeping
his commandments,
ordinances,
and statutes
that
I am
giving
you today.
\VS{12}When
you eat
your fill,
when you build
and occupy
good
houses,
\VS{13}when
your cattle
and flocks
increase,
when you have plenty of silver
and gold,
and when you have abundance
of everything,
\VS{14}be sure you do not feel self-important
and forget
the {\ND{Lord}}
your God
who brought
you from the land
of Egypt,
the place
of slavery,
\VS{15}and who brought
you through the great,
fearful
desert
of venomous serpents
and scorpions,
an arid place
with no
water.
He made water
flow from a flint
rock and
\VS{16}fed
you in the desert
with manna
(which
your ancestors
had never
before known) so that
he might by humbling
you test
you and eventually
bring good to you.
\VS{17}Be careful not to say,
“My own ability
and skill
have gotten
me this
wealth.”
\VS{18}You must remember
the {\ND{Lord}}
your God,
for
he is the one who gives
ability
to get wealth;
if you do this
he will confirm
his covenant
that
he made by oath
to your ancestors,
even as he has to this
day.
\VS{19}Now
if
you forget
the {\ND{Lord}}
your God
at all and follow
other
gods,
worshiping
and prostrating
yourselves before them, I testify
to you today
that
you will surely be annihilated.
\VS{20}Just
like the nations
the {\ND{Lord}}
is about to destroy
from your sight,
so
he will do to you because
you would not
obey
him.

\par }\Chap{9}{\PP \VerseOne{1}Listen,
Israel: Today
you
are about to cross
the Jordan
so you can dispossess
the nations
there, people greater
and stronger
than
you who live in large cities
with extremely high
fortifications.
\VS{2}They include
the Anakites,
a numerous and tall people
whom
you
know
about
and of whom
it is said, “Who is able to resist
the Anakites?”
\VS{3}Understand
today
that
the {\ND{Lord}}
your God
who goes
before
you is a devouring
fire;
he
will defeat
and subdue
them before
you. You will dispossess
and destroy
them quickly
just
as he
has told you.
\VS{4}Do not
think
to yourself after the
{\ND{Lord}}
your God
has driven
them out
before
you, “Because of my own righteousness
the {\ND{Lord}}
has brought
me here to possess
this
land.”
It is because of the wickedness
of these
nations
that the Lord
is driving
them out
ahead of you.
\VS{5}It is not
because of your righteousness,
or even your inner
uprightness,
that you
have come
here to possess
their land.
Instead, because
of the wickedness
of these
nations
the {\ND{Lord}}
your God
is driving
them out
ahead
of you in order to
confirm
the promise
he
made on oath
to your ancestors,
to Abraham,
Isaac,
and Jacob.
\VS{6}Understand,
therefore, that
it is not
because of your righteousness
that the
{\ND{Lord}}
your God
is about to give
you this
good
land
as a possession,
for
you
are a stubborn
people!
\par }{\SH The History of Israel’s Stubbornness
\par }{\PP \VS{7}Remember
– don’t
ever forget –
how
you provoked
the {\ND{Lord}}
your God
in the desert;
from
the time
you left
the land
of Egypt
until
you came
to
this
place
you were constantly rebelling
against him.
\VS{8}At Horeb
you provoked
him
and he
was angry
enough with you to destroy you.
\VS{9}When I went up
the mountain
to receive
the stone
tablets,
the tablets
of the covenant
that
the {\ND{Lord}}
made
with
you, I remained
there
forty
days
and nights,
eating
and drinking
nothing.
\VS{10}The
{\ND{Lord}}
gave
me
the
two
stone
tablets,
written
by the very finger
of God,
and on
them was everything
he said
to you
at the mountain
from the midst
of the fire
at the time
of that assembly.
\VS{11}Now
at the end
of the forty
days
and nights
the {\ND{Lord}}
presented
me
with the
two
stone
tablets,
the tablets
of the covenant.
\VS{12}And he said
to me,
“Get
up, go down
at once
from here
because
your people
whom
you brought out
of Egypt
have sinned! They have quickly
turned from
the way
I commanded
them and have made
for themselves a cast metal image.”
\VS{13}Moreover, he said
to
me, “I have taken note
of these
people;
they are a stubborn lot!
\VS{14}Stand
aside
and I will destroy
them, obliterating
their very name
from
memory,
and I will make
you into
a stronger
and more numerous
nation
than they are.”
\par }{\PP \VS{15}So I turned
and went down
the mountain
while it
was blazing
with fire;
the two
tablets
of the covenant
were in my hands.
\VS{16}When I looked,
you had indeed
sinned
against the
{\ND{Lord}}
your God
and had cast for yourselves a metal
calf;
you had quickly
turned aside
from
the way
he had
commanded you!
\VS{17}I grabbed
the two
tablets,
threw
them down, and shattered
them before your very eyes.
\VS{18}Then I again fell
down before
the {\ND{Lord}}
for forty
days
and nights;
I ate
and drank
nothing
because
of all
the sin
you had
committed,
doing
such evil
before
the {\ND{Lord}}
as to enrage him.
\VS{19}For
I was terrified
at the
{\ND{Lord}}’s
intense
anger
that
threatened
to destroy
you. But he
listened
to me
this time
as well.
\VS{20}The
{\ND{Lord}}
was also angry
enough
at Aaron
to kill
him, but at
that time
I prayed
for him too.
\VS{21}As for your sinful
thing
that you had
made,
the calf,
I took
it,
melted it down,
ground
it up
until
it was as fine as dust,
and tossed
the dust
into
the
stream
that flows
down
the mountain.
\VS{22}Moreover, you continued to provoke the
{\ND{Lord}}
at Taberah,
Massah,
and Kibroth-Hattaavah.
\VS{23}And when he
sent
you from Kadesh-Barnea
and told
you, “Go up
and possess
the land
I have
given
you,” you rebelled
against the
{\ND{Lord}}
your God
and would neither
believe
nor
obey him.
\VS{24}You have been rebelling
against him
from the very first day
I knew you!
\par }{\SH Moses’ Plea on Behalf of the
{\ND{Lord}}’s Reputation
\par }{\PP \VS{25}I lay flat on
the ground
before
the {\ND{Lord}}
for forty
days
and nights,
for
he had
said
he would destroy you.
\VS{26}I prayed
to him: O, Lord

{\ND{God}}, do not
destroy
your people,
your valued
property that
you have powerfully
redeemed,
whom
you brought out
of Egypt
by
your strength.
\VS{27}Remember
your servants
Abraham,
Isaac,
and Jacob;
ignore the stubbornness,
wickedness,
and sin
of these
people.
\VS{28}Otherwise
the people of the land
from which
you brought
us will say,
“The
{\ND{Lord}}
was unable
to bring
them to
the land
he promised
them, and because of his hatred
for them he has brought them out
to kill
them in the desert.”
\VS{29}They
are your people,
your valued
property, whom
you brought out
with great
strength
and power.

\par }\Chap{10}{\PP \VerseOne{1}At that same time
the {\ND{Lord}}
said
to me,
“Carve
out for yourself two
stone
tablets
like the first
ones and come up
the mountain
to me; also make
for yourself a wooden
ark.
\VS{2}I will write
on
the tablets
the same words
that
were on
the first
tablets
you broke,
and you must put
them into the ark.”
\VS{3}So I made
an ark
of acacia
wood
and carved out
two
stone
tablets
just like the first
ones. Then I went up
the mountain
with the two
tablets
in my hands.
\VS{4}The
{\ND{Lord}}
then wrote
on the
tablets
the same
words,
the ten
commandments,
which
he had spoken
to
you at the mountain
from the middle
of the fire
at the time
of that assembly,
and he gave
them to me.
\VS{5}Then I turned,
went down
the mountain,
and placed
the tablets
into the ark
I had
made
– they are still there,
just
as the
{\ND{Lord}}
commanded me.
\par }{\SH Conclusion of the Historical Resume
\par }{\PP \VS{6}“During those days the Israelites
traveled
from Beeroth Bene-Yaaqan
to Moserah.
There
Aaron
died
and was buried,
and his son
Eleazar
became priest
in his place.
\VS{7}From there
they traveled
to Gudgodah,
and from
Gudgodah
to Jotbathah,
a place of flowing streams.
\VS{8}At that time
the {\ND{Lord}}
set apart
the
tribe
of Levi
to carry
the
ark
of the
{\ND{Lord}}’s
covenant,
to stand
before
the {\ND{Lord}}
to serve
him, and to formulate
blessings in his name,
as
they do
to this
very
day.
\VS{9}Therefore
Levi
has no
allotment
or inheritance
among his brothers;
the {\ND{Lord}}
is
his inheritance
just
as the
{\ND{Lord}}
your God
told him.
\VS{10}As for me, I
stayed
at the mountain
as I
did the first
time,
forty
days
and nights.
The
{\ND{Lord}}
listened
to
me that
time
as well
and decided
not
to destroy you.
\VS{11}Then he said
to
me, “Get
up, set
out leading
the people
so they may go
and possess
the land
I promised
to give
to their ancestors.”
\par }{\SH An Exhortation to Love Both God and People
\par }{\PP \VS{12}Now,
Israel,
what
does the
{\ND{Lord}}
your God
require
of you
except
to revere
him, to obey
all
his commandments,
to love
him, to serve
him with all
your mind
and being,
\VS{13}and to keep
the Lord’s
commandments
and statutes
that
I am
giving
you today
for your own good?
\VS{14}The heavens
– indeed
the highest heavens
– belong to the
{\ND{Lord}}
your God,
as does the earth
and everything in it.
\VS{15}However,
only to your ancestors
did he
show his loving favor,
and he chose
you, their descendants,
from all
peoples
– as is apparent
today.
\VS{16}Therefore,
cleanse
your heart
and stop being so stubborn!
\VS{17}For
the {\ND{Lord}}
your God
is
God
of gods
and Lord
of lords,
the great,
mighty,
and awesome
God
who
is unbiased
and takes
no
bribe,
\VS{18}who justly
treats
the orphan
and widow,
and who loves
resident foreigners,
giving
them food
and clothing.
\VS{19}So you must
love
the resident foreigner
because
you were foreigners
in the land
of Egypt.
\VS{20}Revere
the {\ND{Lord}}
your God,
serve
him, be loyal
to him and take oaths
only in his name.
\VS{21}He is
the one you should praise;
he is
your God,
the one who has
done
these great
and awesome
things
for you that
you have seen.
\VS{22}When your ancestors
went down
to Egypt,
they numbered
only seventy,
but now
the {\ND{Lord}}
your God
has made you as numerous
as the stars
of the sky.

\par }\Chap{11}{\PP \VerseOne{1}You must love
the {\ND{Lord}}
your God
and do
what he requires;
keep
his statutes,
ordinances,
and commandments
at
all
times.
\VS{2}Bear in mind
today
that
I am
not
speaking
to your children
who
have not
personally experienced
the judgments
of the {\ND{Lord}}
your God,
which
revealed his greatness,
strength,
and power.
\VS{3}They did not see the awesome
deeds
he performed
in the midst
of Egypt
against Pharaoh
king
of Egypt
and his whole
land,
\VS{4}or what
he did
to the army
of Egypt,
including their horses
and chariots,
when he made the waters
of the Red
Sea
overwhelm
them while they were pursuing
you and he
annihilated them.
\VS{5}They did not see what he did
to you in the desert
before
you reached
this
place,
\VS{6}or what
he did
to Dathan
and Abiram,
sons
of Eliab
the Reubenite,
when
the earth
opened
its mouth
in the middle
of the Israelite
camp and swallowed
them, their families,
their tents,
and all
the property
they brought with them.
\VS{7}I am speaking
to you because
you are the ones who
saw
all
the great
deeds
of the {\ND{Lord}}!
\par }{\SH The Abundance of the Land of Promise
\par }{\PP \VS{8}Now pay attention
to all
the commandments
I am
giving
you today,
so that you may
be strong
enough to enter
and possess
the land
where
you
are headed,
\VS{9}and that
you may enjoy long
life
in
the land
the {\ND{Lord}}
promised
to give
to your ancestors
and their descendants,
a land
flowing
with milk
and honey.
\VS{10}For the
land
where
you
are headed
is not
like the land
of Egypt
from which
you came,
a land where you planted
seed
and which
you irrigated
by hand
like a vegetable
garden.
\VS{11}Instead, the land
you
are crossing
the Jordan to occupy
is one of hills
and valleys,
a land
that drinks
in water
from the rains,
\VS{12}a land
the {\ND{Lord}}
your God
looks after.
He is constantly
attentive
to it from the beginning
to
the end
of the year.
\VS{13}Now,
if
you pay close
attention
to
my commandments
that
I am
giving
you today
and love
the {\ND{Lord}}
your God
and serve
him with all
your mind
and being,
\VS{14}then he promises, “I will send
rain
for your land
in its season,
the autumn
and the spring rains,
so that you may gather
in your grain,
new wine,
and olive oil.
\VS{15}I will provide
pasture
for your livestock
and you will eat
your fill.”
\par }{\SH Exhortation to Instruction and Obedience
\par }{\PP \VS{16}Make sure
you do not turn
away to serve
and worship
other
gods!
\VS{17}Then the anger
of the {\ND{Lord}}
will erupt
against you and he will close up
the sky
so that
it does not
rain.
The land
will not
yield
its produce,
and you will soon
be removed
from
the good
land
that
the Lord
is about to give you.
\VS{18}Fix
these
words
of mine into your mind
and being,
and tie
them as a reminder
on
your hands
and let them be
symbols
on your forehead.
\VS{19}Teach
them
to your children
and speak
of them as you sit
in your house,
as you walk
along the road,
as you lie
down, and as you get up.
\VS{20}Inscribe
them on
the doorframes
of your houses
and on your gates
\VS{21}so
that your
days
and those
of your
descendants
may be extended
in the land
which
the {\ND{Lord}}
promised
to give
to your ancestors,
like the days
of heaven itself.
\VS{22}For
if
you carefully
observe
all
of these commandments
I am
giving
you and love
the {\ND{Lord}}
your God,
live
according to his standards,
and remain loyal to him,
\VS{23}then
he will drive
out all
these
nations
ahead
of you, and you will dispossess
nations
greater
and stronger
than you.
\VS{24}Every
place
you set
your foot
will be
yours; your border
will extend
from
the desert
to Lebanon
and from
the River
(that is, the Euphrates) as far as
the Mediterranean
Sea.
\VS{25}Nobody
will be able to resist
you;
the {\ND{Lord}}
your God
will spread
the fear
and terror
of you over
the whole
land
on which
you walk,
just
as he promised you.
\par }{\SH Anticipation of a Blessing and Cursing Ceremony
\par }{\PP \VS{26}Take note
– I am
setting
before
you today
a blessing
and a curse:
\VS{27}the blessing
if you take to
heart
the commandments
of the {\ND{Lord}}
your God
that
I am
giving
you today,
\VS{28}and the curse
if
you pay no
attention
to
his commandments
and turn
from
the way
I am
setting
before you today
to pursue
other
gods
you have
not
known.
\VS{29}When
the {\ND{Lord}}
your God
brings
you into
the land
you
are to possess,
you must pronounce
the blessing
on
Mount
Gerizim
and the curse
on
Mount
Ebal.
\VS{30}Are they
not
across
the Jordan River,
toward
the west,
in the land
of the Canaanites
who live
in the Arabah
opposite
Gilgal
near
the oak
of Moreh?
\VS{31}For
you
are about to cross
the Jordan
to possess
the land
the {\ND{Lord}}
your God
is giving
you, and you will possess
and inhabit it.
\VS{32}Be certain to keep
all
the statutes
and ordinances
that
I am
presenting
to you
today.

\par }\Chap{12}{\PP \VerseOne{1}These
are the statutes
and ordinances
you must be careful
to obey
as long
as you
live
in the land
the {\ND{Lord}}, the God
of your ancestors,
has
given
you to possess.
\VS{2}You must by all means destroy
all
the places
where
the nations
you
are about to dispossess
worship
their gods
– on
the high
mountains
and hills
and under
every
leafy
tree.
\VS{3}You must tear down
their altars,
shatter
their sacred pillars,
burn up
their sacred Asherah poles,
and cut down
the images
of their gods;
you must eliminate
their very memory
from
that
place.
\VS{4}You must not
worship
the {\ND{Lord}}
your God the way they worship.
\VS{5}But
you must seek
only the place
he
chooses
from all
your tribes
to establish
his name
as his place
of residence,
and you must go
there.
\VS{6}And there
you must take
your burnt offerings,
your sacrifices,
your tithes,
the personal offerings
you have prepared,
your votive offerings,
your freewill offerings,
and the firstborn
of your herds
and flocks.
\VS{7}Both you and your families must feast
there
before
the {\ND{Lord}}
your God
and rejoice
in all
the output
of your
labor
with which he
has
blessed you.
\VS{8}You must not
do
like
we
are doing
here
today,
with everyone
doing what seems best to him,
\VS{9}for
you have not
yet
come
to
the final stop
and inheritance
the
{\ND{Lord}}
your God
is giving you.
\VS{10}When
you do go across
the Jordan River
and settle
in the land
he is granting you as an inheritance
and you find relief
from all
the enemies
who surround
you, you will live
in safety.
\VS{11}Then you must come to the place
the {\ND{Lord}}
your God
chooses
for his name
to reside,
bringing
everything
I am
commanding you – your burnt offerings, sacrifices, tithes, the personal offerings you have prepared, and all your choice votive offerings which you devote to him.
\VS{12}You shall rejoice
in the presence
of the {\ND{Lord}}
your God,
along with your
sons,
daughters,
male
and female
servants,
and the Levites
in your villages
(since
they have no
allotment
or inheritance
with you).
\VS{13}Make sure
you do not offer
burnt offerings
in any
place
you wish,
\VS{14}for
you may do so only
in the place
the {\ND{Lord}}
chooses
in one
of your tribal
areas
– there
you may do
everything
I am
commanding you.
\par }{\SH Regulations for Profane Slaughter
\par }{\PP \VS{15}On the other hand,
you may slaughter
and eat
meat
as you please
when the
{\ND{Lord}}
your God
blesses
you in all
your villages.
Both the ritually
pure
and impure
may eat
it, whether it is a gazelle
or an ibex.
\VS{16}However,
you must not
eat
blood
– pour
it out on
the ground
like water.
\VS{17}You will not
be allowed
to eat
in your villages
your tithe
of grain,
new wine,
olive oil,
the firstborn
of your herd
and flock,
any
votive offerings
you have
vowed,
or your freewill
and personal
offerings.
\VS{18}Only
in the presence
of the {\ND{Lord}}
your God
may you eat
these, in the place
he chooses.
This applies to you,
your son,
your daughter,
your male
and female
servants,
and the Levites
in your villages.
In that
place you will rejoice
before
the {\ND{Lord}}
your God
in all
the output of your labor.
\VS{19}Be careful
not
to overlook
the
Levites
as long
as you live in the land.
\par }{\SH The Sanctity of Blood
\par }{\PP \VS{20}When
the {\ND{Lord}}
your God
extends
your borders
as
he said
he would do and you say,
“I want
to eat
meat
just as I please,”
you may do so as you wish.
\VS{21}If
the place
he chooses
to locate
his name
is too far
for you, you may slaughter
any
of your herd
and flock
he has
given
you just
as I have stipulated;
you may eat
them in your
villages
just
as you wish.
\VS{22}Like
you eat
the gazelle
or ibex,
so
you may eat
these; the ritually impure
and pure
alike
may eat them.
\VS{23}However,
by no means
eat
the blood,
for
the blood
is life itself – you must not eat the life with the meat!
\VS{24}You must not
eat
it! You must pour
it out on
the ground
like water.
\VS{25}You must not
eat
it so that it may
go well
with you and your children
after
you; you will be doing
what is right
in the
{\ND{Lord}}’s
sight.
\VS{26}Only
the holy
things and votive offerings
that
belong
to you, you must pick up
and take
to
the place
the {\ND{Lord}}
will choose.
\VS{27}You must offer
your burnt offerings,
both meat
and blood,
on
the altar
of the {\ND{Lord}}
your God;
the blood
of your other sacrifices
you must pour
out on
his altar
while you eat
the meat.
\VS{28}Pay careful
attention
to all
these
things
I am
commanding
you so that it may
always
go well
with you and your children
after
you when
you do
what is good
and right
in the sight
of the {\ND{Lord}}
your God.
\par }{\SH The Abomination of Pagan Gods
\par }{\PP \VS{29}When the
{\ND{Lord}}
your God
eliminates
the nations
from the place where you
are headed
and you dispossess
them, you will settle down
in their land.
\VS{30}After
they have been destroyed
from your presence,
be careful
not
to be
ensnared
like they are; do not
pursue
their gods
and say,
“How
do these
nations
serve
their gods? I
will do
the same.”
\VS{31}You must not
worship
the {\ND{Lord}}
your God
the way
they do! For
everything
that is abhorrent
to him, everything
he hates,
they have done
when worshiping their gods.
They even
burn up
their sons
and daughters
before their gods!
\par }{\SH Idolatry and False Prophets
\par }{\PP \VS{32} You must be careful
to do
everything
I am
commanding
you. Do
not
add
to it or
subtract
from it!

\par }\Chap{13}{\PP \VerseOne{1}Suppose
a prophet
or
one who foretells
by dreams
should appear among you and show you a sign
or
wonder,
\VS{2}and the sign
or wonder
should come
to pass concerning what
he said
to
you, namely,
“Let us
follow
other
gods”
– gods whom
you have not
previously known
– “and let us serve them.”
\VS{3}You must not
listen
to
the words
of that prophet
or
dreamer,
for
the {\ND{Lord}}
your God
will be testing
you to see if
you love
him with all
your mind
and being.
\VS{4}You must follow
the
{\ND{Lord}}
your God
and revere
only him; and you must observe
his commandments,
obey
him, serve
him, and remain loyal to him.
\VS{5}As for that prophet
or
dreamer,
he must be executed
because
he encouraged
rebellion
against
the {\ND{Lord}}
your God
who brought
you from the
land
of Egypt,
redeeming
you from that place of slavery,
and because he has tried to entice
you from
the way
the {\ND{Lord}}
your God
has
commanded
you to go.
In this way you must purge
out evil
from within.
\par }{\SH False Prophets in the Family
\par }{\PP \VS{6}Suppose
your own full brother,
your son,
your daughter,
your beloved wife,
or
your closest
friend
should seduce
you secretly
and encourage you to go
and serve
other
gods
that neither
you
nor your ancestors
have previously known,
\VS{7}the gods
of the surrounding
people
(whether near
you or
far
from you, from
one end
of the earth
to
the other).
\VS{8}You must not
give in
to him or
even listen
to
him; do not
feel
sympathy
for
him or
spare
him or cover up
for him.
\VS{9}Instead,
you must kill
him without fail! Your own hand
must be
the first
to strike
him, and then the hands
of the whole community.
\VS{10}You must stone
him to death
because
he tried to entice
you away from the
{\ND{Lord}}
your God,
who delivered
you from
the land
of Egypt,
that place
of slavery.
\VS{11}Thus all
Israel
will hear
and be afraid;
no
longer
will they continue to do
evil
like this
among you.
\par }{\SH Punishment of Community Idolatry
\par }{\PP \VS{12}Suppose
you should hear
in one
of your cities,
which
the {\ND{Lord}}
your God
is giving
you as a place
to live, that
\VS{13}some evil
people
have departed
from among
you to entice
the inhabitants
of their cities,
saying,
“Let’s
go and serve
other
gods”
(whom
you have not
known before).
\VS{14}You must
investigate
thoroughly and inquire
carefully. If it
is indeed
true
that such a disgraceful
thing
is being done
among you,
\VS{15}you must by
all
means
slaughter
the inhabitants
of that city
with the sword;
annihilate
with the sword
everyone
in it, as well as the livestock.
\VS{16}You must gather
all
of its plunder
into
the middle
of the plaza
and burn
the city
and all
its plunder
as a whole burnt offering
to the
{\ND{Lord}}
your God.
It will be
an abandoned ruin
forever
– it must never
be rebuilt
again.
\VS{17}You must not
take for yourself anything
that has been placed under judgment.
Then
the {\ND{Lord}}
will relent from his intense
anger,
show
you compassion,
have mercy
on you, and multiply
you as
he promised
your ancestors.
\VS{18}Thus
you must
obey
the {\ND{Lord}}
your God,
keeping
all
his commandments
that
I am
giving
you today
and doing
what is right
before him.

\par }\Chap{14}{\PP \VerseOne{1}You
are children
of the {\ND{Lord}}
your God.
Do not
cut yourselves
or
shave
your forehead
bald
for the sake of the dead.
\VS{2}For
you
are a people
holy
to the
{\ND{Lord}}
your God.
He has chosen
you to be
his people,
prized
above all
others
on
the face
of the earth.
\VS{3}You must not
eat
any
forbidden thing.
\VS{4}These
are the animals
you may eat: the ox,
the sheep,
the goat,
\VS{5}the ibex,
the gazelle,
the deer,
the wild goat,
the antelope,
the wild oryx,
and the mountain sheep.
\VS{6}You may eat
any
animal
that has hooves
divided
into two
parts
and that chews
the cud.
\VS{7}However,
you may not
eat
the following animals among those that
chew
the cud
or those that have divided
hooves: the
camel,
the
hare,
and the rock badger.
(Although
they chew
the cud,
they
do not
have divided hooves
and are
therefore ritually impure to you).
\VS{8}Also the pig
is ritually impure
to you; though
it has divided
hooves,
it does not
chew the cud.
You may not
eat
their meat
or
even touch
their remains.
\VS{9}These
you may eat
from among
water
creatures: anything
with fins
and scales
you may eat,
\VS{10}but whatever
does not
have fins
and scales
you may not
eat;
it is
ritually impure to you.
\VS{11}All
ritually clean
birds
you may eat.
\VS{12}These
are the ones you may not
eat: the
eagle,
the vulture,
the black vulture,
\VS{13}the kite,
the black kite,
the dayyah
after its species,
\VS{14}every
raven
after its species,
\VS{15}the ostrich,
the owl,
the seagull,
the falcon
after its species,
\VS{16}the little owl,
the long-eared owl,
the white owl,
\VS{17}the jackdaw,
the carrion vulture,
the cormorant,
\VS{18}the stork,
the heron
after
its species,
the hoopoe,
the bat,
\VS{19}and any
winged
thing on the ground
are impure to you – they may not be eaten.
\VS{20}You may eat
any
clean
bird.
\VS{21}You may not
eat
any
corpse,
though you may give
it to the resident foreigner
who is living in your villages
and he may eat
it, or
you may sell
it to a foreigner.
You
are a people
holy
to the
{\ND{Lord}}
your God.
Do not
boil
a young goat
in its mother’s
milk.
\par }{\SH The Offering of Tribute
\par }{\PP \VS{22}You must be certain
to tithe
all
the produce
of your seed
that comes from
the field
year
after year.
\VS{23}In the presence
of the
{\ND{Lord}}
your God
you must eat
from the tithe
of your grain,
your new wine,
your olive oil,
and the firstborn
of your herds
and flocks
in the place
he chooses
to locate
his name,
so that
you may learn
to revere
the
{\ND{Lord}}
your God
always.
\VS{24}When
he blesses
you, if
the place
where
he chooses
to locate
his name
is distant,
\VS{25}you may
convert
the tithe into money,
secure the money,
and travel
to
the place
the {\ND{Lord}}
your God
chooses for himself.
\VS{26}Then you
may spend the money
however you
wish
for cattle,
sheep,
wine,
beer,
or whatever
you desire.
You
and your household
may eat
there
in the presence
of the {\ND{Lord}}
your God
and enjoy it.
\VS{27}As for the Levites
in your villages,
you must not
ignore
them, for
they have no
allotment
or inheritance
along with you.
\VS{28}At the end
of every three
years
you must bring
all
the tithe
of your produce,
in that
very
year,
and you must store
it up in your villages.
\VS{29}Then
the Levites
(because
they have no
allotment
or inheritance
with
you), the resident foreigners,
the orphans,
and the widows
of your villages
may come and eat
their fill
so that
the {\ND{Lord}}
your God
may bless
you in all
the work
you do.

\par }\Chap{15}{\PP \VerseOne{1}At the end
of every seven
years
you must declare
a cancellation of debts.
\VS{2}This
is the nature
of the cancellation: Every
creditor
must remit
what
he has loaned
to another person;
he must not
force payment
from his fellow Israelite,
for
it is to be recognized
as “the
{\ND{Lord}}’s
cancellation of debts.”
\VS{3}You may exact payment
from a foreigner,
but whatever
your fellow Israelite
owes
you, you must remit.
\VS{4}However,
there should
not
be
any poor
among you, for
the {\ND{Lord}}
will surely bless
you in the land
that
he is giving
you as an inheritance,
\VS{5}if
you carefully obey
him by keeping
all
these
commandments
that
I am
giving
you today.
\VS{6}For
the {\ND{Lord}}
your God
will bless
you just
as he has promised;
you will lend
to many
nations
but
will not
borrow
from any, and you will rule
over many
nations
but they will not
rule
over you.
\par }{\SH The Spirit of Liberality
\par }{\PP \VS{7}If
a fellow Israelite
from one
of your villages
in the land
that
the {\ND{Lord}}
your God
is giving
you should be poor,
you must not
harden
your heart
or be insensitive
to his impoverished
condition.
\VS{8}Instead,
you must be
sure to open
your hand
to him and generously
lend him whatever
he needs.
\VS{9}Be careful
lest
you entertain
the wicked
thought
that the seventh
year,
the year
of cancellation of debts,
has almost arrived, and your attitude be wrong
toward your impoverished
fellow
Israelite and you do not
lend
him anything; he will cry out
to
the {\ND{Lord}}
against
you and you will be
regarded as having sinned.
\VS{10}You
must by
all means lend
to him and not
be upset
by doing
it,
for because
of this
the
{\ND{Lord}}
your God
will bless
you in all
your work
and in everything
you attempt.
\VS{11}There will never
cease
to be some poor
people in
the land;
therefore,
I
am commanding
you to make sure
you open
your hand
to your fellow
Israelites who
are needy
and poor
in your land.
\par }{\SH Release of Debt Slaves
\par }{\PP \VS{12}If
your fellow
Hebrew –
whether male or
female –
is sold
to you and serves
you for six
years,
then in the seventh
year
you must let that servant go free.
\VS{13}If
you set
them free,
you must not
send
them away empty-handed.
\VS{14}You must supply
them generously
from your flock,
your threshing floor,
and your winepress
– as the
{\ND{Lord}}
your God
has
blessed
you, you must give to them.
\VS{15}Remember
that
you were a slave
in the land
of Egypt
and the
{\ND{Lord}}
your God
redeemed
you; therefore,
I am
commanding
you to do this
thing
today.
\VS{16}However, if
the servant says
to
you, “I do not
want to leave
you,”
because
he loves
you and your household,
since
he is well
off with you,
\VS{17}you shall take
an awl
and pierce a hole
through his ear
to the door.
Then he will become
your servant
permanently
(this
applies
to your female servant as well).
\VS{18}You should not
consider
it difficult
to let him go
free,
for
he will have served
you for six
years,
twice
the time of a hired
worker;
the {\ND{Lord}}
your God
will bless
you in everything
you do.
\par }{\SH Giving God the Best
\par }{\PP \VS{19}You must set apart
for the
{\ND{Lord}}
your God
every
firstborn
male
born
to your herds
and flocks.
You must not
work
the firstborn
of your bulls
or
shear
the firstborn
of your flocks.
\VS{20}You
and your household
must eat
them annually
before
the {\ND{Lord}}
your God
in the place
he chooses.
\VS{21}If
they have any kind of blemish
– lameness,
blindness,
or
anything else –
you may not
offer
them as a sacrifice
to the
{\ND{Lord}}
your God.
\VS{22}You may eat
it in your villages,
whether you are ritually impure
or clean,
just as you would eat a gazelle
or an ibex.
\VS{23}However,
you must not
eat
its blood;
you must pour
it out on
the ground
like water.

\par }\Chap{16}{\PP \VerseOne{1}Observe
the month
Abib
and keep
the Passover
to the
{\ND{Lord}}
your God,
for
in that
month
he
brought you out
of Egypt
by night.
\VS{2}You must sacrifice
the Passover animal
(from the flock
or the herd) to the
{\ND{Lord}}
your God
in the place
where
he chooses
to locate
his name.
\VS{3}You must not
eat
any yeast
with it; for seven
days
you must eat
bread
made without yeast,
symbolic of affliction,
for
you came out
of Egypt
hurriedly.
You must do this so
you will remember
for the rest
of your life
the day
you came out
of the land
of Egypt.
\VS{4}There must not
be a scrap of yeast
within your land
for seven
days,
nor
can any
of the meat
you sacrifice
on the evening
of the first
day
remain until the next morning.
\VS{5}You may
not
sacrifice
the Passover
in just any
of your villages
that
the {\ND{Lord}}
your God
is giving you,
\VS{6}but
you must sacrifice
it
in the evening
in
the place
where
he
chooses
to locate
his name,
at sunset,
the time of day
you came out
of Egypt.
\VS{7}You must cook
and eat
it in the place
the {\ND{Lord}}
your God
chooses;
you may return
the next
morning
to your tents.
\VS{8}You must eat
bread made without yeast
for six
days.
The seventh
day
you are to hold an assembly
for the
{\ND{Lord}}
your God;
you must not
do
any work
on that day.
\par }{\SH The Festival of Weeks
\par }{\PP \VS{9}You must count
seven
weeks;
you must begin
to count
them from the time you begin
to harvest
the standing grain.
\VS{10}Then you are to celebrate
the Festival
of Weeks
before the
{\ND{Lord}}
your God
with the voluntary
offering
that
you will bring,
in proportion to how
he has blessed you.
\VS{11}You shall rejoice
before
him –
you,
your son,
your daughter,
your male and female
slaves,
the Levites
in your villages,
the resident foreigners,
the orphans,
and the widows
among you – in the place where the
{\ND{Lord}} chooses to locate his name.
\VS{12}Furthermore, remember
that
you were a slave
in Egypt,
and so be careful
to observe
these
statutes.
\par }{\SH The Festival of Temporary Shelters
\par }{\PP \VS{13}You must celebrate
the Festival
of Temporary Shelters
for seven
days,
at the time of the grain
and grape
harvest.
\VS{14}You are to rejoice
in your festival,
you, your son,
your daughter,
your male
and female
slaves,
the Levites,
the resident foreigners,
the orphans,
and the widows
who
are in your villages.
\VS{15}You are to celebrate
the festival seven
days
before the
{\ND{Lord}}
your God
in the place
he
chooses,
for
he
will bless
you in all
your productivity
and in whatever
you do;
so you will indeed
rejoice!
\VS{16}Three
times
a year
all
your males
must appear
before
the {\ND{Lord}}
your God
in the place
he chooses
for the Festival
of Unleavened
Bread, the Festival
of Weeks,
and the Festival
of Temporary Shelters;
and they must not
appear
before
him empty-handed.
\VS{17}Every one
of you must give
as you are able, according
to the blessing
of the {\ND{Lord}}
your God
that
he has given you.
\par }{\SH Provision for Justice
\par }{\PP \VS{18}You must appoint judges
and civil
servants for each tribe
in all
your villages
that
the {\ND{Lord}}
your God
is giving
you, and they must judge
the
people
fairly.
\VS{19}You must not
pervert
justice
or
show
favor. Do not
take
a bribe,
for
bribes
blind
the eyes
of the wise
and distort
the words
of the righteous.
\VS{20}You must
pursue
justice
alone so that you may
live
and inherit
the land
the {\ND{Lord}}
your God
is giving you.
\par }{\SH Examples of Legal Cases
\par }{\PP \VS{21}You must not
plant
any
kind of tree
as a sacred Asherah pole
near
the altar
of the {\ND{Lord}}
your God
which
you build
for yourself.
\VS{22}You must not
erect
a sacred pillar,
a thing the
{\ND{Lord}}
your God
detests.

\par }\Chap{17}{\PP \VerseOne{1}You must not
sacrifice
to him
a bull
or sheep
that
has a blemish
or any
other defect,
because
that is considered offensive
to the
{\ND{Lord}}
your God.
\VS{2}Suppose
a man
or
woman
is discovered
among you – in one of your villages that the
{\ND{Lord}} your God is giving you – who sins before the Lord your God and breaks his covenant
\VS{3}by serving
other
gods
and worshiping them – the sun, moon, or any other heavenly bodies which I have not permitted you to worship.
\VS{4}When it is reported
to you and you hear
about it, you must
investigate
carefully. If it is indeed
true
that such a disgraceful
thing
is being done
in Israel,
\VS{5}you must bring
to
your city gates
that man
or
woman
who has
done
this
wicked
thing
– that very man
or
woman
– and you must stone
that person
to death.
\VS{6}At the testimony of two
or
three
witnesses
they must
be executed.
They cannot
be put to death
on
the testimony
of only one
witness.
\VS{7}The witnesses
must be
first
to begin the execution,
and then
all
the people
are to join
in afterward. In this way you will purge
evil
from among you.
\par }{\SH Appeal to a Higher Court
\par }{\PP \VS{8}If
a matter
is too difficult
for you to judge
– bloodshed,
legal claim, or assault – matters of controversy in your villages – you must leave there and go up to the place the
{\ND{Lord}} your God chooses.
\VS{9}You will go
to
the Levitical
priests
and the judge
in office in those
days
and seek
a solution; they
will render
a verdict.
\VS{10}You must then do
as they have determined
at that
place
the {\ND{Lord}}
chooses.
Be careful
to do
just
as you are taught.
\VS{11}You must do what
you are instructed,
and the verdict
they pronounce to you, without fail. Do
not
deviate
right
or left
from
what
they tell you.
\VS{12}The person
who
pays no
attention
to
the priest
currently
serving
the
{\ND{Lord}}
your God
there,
or
to
the verdict
– that person
must die,
so that you may
purge
evil
from Israel.
\VS{13}Then all
the people
will hear
and be afraid,
and not
be so presumptuous
again.
\par }{\SH Provision for Kingship
\par }{\PP \VS{14}When
you come
to
the land
the {\ND{Lord}}
your God
is giving
you and take
it over and live
in it and then say,
“I will select
a king
like all
the nations
surrounding me,”
\VS{15}you must select
without fail a king
whom
the {\ND{Lord}}
your God
chooses.
From among
your fellow
citizens you must appoint
a king
– you may
not
designate
a foreigner
who
is not
one of your fellow Israelites.
\VS{16}Moreover,
he must not
accumulate
horses
for himself or
allow the
people
to return
to Egypt
to
do so, for the
{\ND{Lord}}
has said
you must never
again
return
that way.
\VS{17}Furthermore, he must not
marry many
wives
lest his affections turn
aside,
and he must not
accumulate
much silver
and gold.
\VS{18}When
he sits
on
his royal
throne
he must
make a copy
of this
law
on
a scroll
given to him by the Levitical
priests.
\VS{19}It must be
with
him constantly and he must read
it as long
as he lives,
so that he may
learn
to revere
the

{\ND{Lord}}
his God
and observe
all
the words
of this
law
and these
statutes
and carry them out.
\VS{20}Then he will not
exalt
himself
above his fellow citizens
or
turn
from
the commandments
to the right
or left,
and he and his descendants
will enjoy
many years
ruling over
his kingdom
in
Israel.

\par }\Chap{18}{\PP \VerseOne{1}The Levitical
priests –
indeed, the entire
tribe
of Levi
– will have no allotment
or inheritance
with
Israel;
they may eat
the burnt offerings
of the {\ND{Lord}}
and of his inheritance.
\VS{2}They will have no
inheritance
in the midst
of their fellow
Israelites; the
{\ND{Lord}}
alone
is their inheritance,
just
as he had told them.
\VS{3}This
shall be
the priests’
fair allotment
from the
people
who offer
sacrifices,
whether
bull
or
sheep
– they must give
to the priest
the shoulder,
the jowls,
and the stomach.
\VS{4}You must give
them the best
of your grain,
new wine,
and olive oil,
as well as the best
of your wool when you shear
your flocks.
\VS{5}For
the {\ND{Lord}}
your God
has chosen
them and their sons
from all
your tribes
to stand
and serve
in his name permanently.
\VS{6}Suppose
a Levite
comes
by his own free will
from one
of your villages,
from any
part of Israel
where he
is living,
to
the place
the {\ND{Lord}}
chooses
\VS{7}and serves
in the name
of the {\ND{Lord}}
his God
like his fellow
Levites
who stand
there
before
the {\ND{Lord}}.
\VS{8}He must eat
the same
share
they do, despite
any profits he may gain from the sale
of his family’s inheritance.
\par }{\SH Provision for Prophetism
\par }{\PP \VS{9}When
you
enter
the land
the {\ND{Lord}}
your God
is giving
you, you must not
learn
the abhorrent
practices
of those
nations.
\VS{10}There must never
be found
among you anyone who sacrifices
his son
or
daughter
in the fire,
anyone who practices
divination,
an
omen reader, a
soothsayer,
a sorcerer,
\VS{11}one
who casts spells,
one who conjures
up spirits,
a practitioner of the occult,
or a necromancer.
\VS{12}Whoever
does
these things
is abhorrent
to the
{\ND{Lord}}
and because
of these
detestable
things
the {\ND{Lord}}
your God
is about to drive
them out from before you.
\VS{13}You must be blameless
before the
{\ND{Lord}}
your God.
\VS{14}Those
nations
that
you
are about
to
dispossess
listen
to omen
readers
and diviners, but
the {\ND{Lord}}
your God
has not
given
you
permission
to do such things.
\par }{\PP \VS{15}The
{\ND{Lord}}
your God
will raise
up for you a prophet
like
me from among you – from your fellow Israelites; you must listen to him.
\VS{16}This accords
with
what happened at Horeb
in the day
of the assembly.
You asked
the {\ND{Lord}}
your God: “Please do not
make
us hear
the
voice
of the {\ND{Lord}}
our God
any more or
see
this
great
fire
any more
lest we die.”
\VS{17}The
{\ND{Lord}}
then said
to me,
“What they have
said
is good.
\VS{18}I will raise
up a prophet
like you for them
from among
their fellow
Israelites. I will put
my words
in his mouth
and he will speak
to them
whatever
I command.
\VS{19}I
will personally
hold responsible anyone
who then pays no
attention
to
the words
that
prophet speaks
in my name.
\par }{\PP \VS{20}“But
if any prophet
presumes
to speak
anything
in my name
that
I have not
authorized him
to speak,
or speaks
in the name
of other
gods,
that prophet
must die.
\VS{21}Now if
you say
to yourselves, ‘How
can we tell that
a message
is not
from the
{\ND{Lord}}?’ –
\VS{22}whenever
a prophet
speaks
in my
name
and the prediction
is not
fulfilled,
then I
have not
spoken
it; the prophet
has presumed
to speak
it, so you need not
fear him.”

\par }\Chap{19}{\PP \VerseOne{1}When
the {\ND{Lord}}
your God
destroys
the nations
whose
land
he is about to give
you and you dispossess
them and settle
in their cities
and houses,
\VS{2}you must set apart
for yourselves three
cities
in the middle
of your land
that
the {\ND{Lord}}
your God
is giving
you as a possession.
\VS{3}You shall
build a roadway
and divide into thirds
the whole extent
of your
land
that
the {\ND{Lord}}
your God
is providing as your inheritance;
anyone
who kills
another person should flee
to the closest of these cities.
\VS{4}Now this
is the law pertaining
to one who flees
there
in order to live,
if he has
accidentally
killed
another
without
hating
him at the time of the accident.
\VS{5}Suppose
he goes
with
someone else
to the forest
to cut
wood
and when he raises
the ax
to cut
the tree,
the ax head
flies loose
from
the handle
and strikes
his fellow worker
so hard that
he dies.
The person responsible may then flee
to
one
of these
cities
to save himself.
\VS{6}Otherwise
the blood
avenger
will chase
after
the killer
in the heat
of his anger, eventually
overtake
him,
and kill
him, though this is not
a capital case
since
he
did not
hate
him at the time of the accident.
\VS{7}Therefore,
I am
commanding
you to set apart
for yourselves three
cities.
\VS{8}If
the {\ND{Lord}}
your God
enlarges
your borders
as
he promised
your ancestors
and gives
you all
the land
he pledged
to them,
\VS{9}and then you are careful
to observe all
these commandments
I am
giving
you today
(namely, to love
the
{\ND{Lord}}
your God
and to always
walk
in his ways), then you must add
three
more
cities
to these
three.
\VS{10}You must not
shed
innocent
blood
in your
land
that
the {\ND{Lord}}
your God
is giving
you as an inheritance,
for that would make you guilty.
\VS{11}However, suppose
a person
hates
someone else
and stalks
him, attacks
him, kills
him,
and then flees
to
one
of these
cities.
\VS{12}The elders
of his own city
must
send
for him and remove
him
from
there
to deliver
him over
to the blood
avenger
to die.
\VS{13}You must not
pity
him,
but purge
out the blood
of the innocent
from Israel,
so that it may go well with you.
\par }{\SH Laws Concerning Witnesses
\par }{\PP \VS{14}You must not
encroach
on
your neighbor’s
property, which
will have
been defined
in the inheritance
you will obtain
in the land
the {\ND{Lord}}
your God
is giving you.
\par }{\PP \VS{15}A single
witness
may not
testify
against another person
for any
trespass
or sin
that he commits.
A matter
may be legally established
only on
the testimony
of two
or
three
witnesses.
\VS{16}If
a false
witness
testifies
against another person
and accuses him
of a crime,
\VS{17}then
both
parties
to the controversy
must stand
before
the {\ND{Lord}}, that is, before
the priests
and judges
who
will be in office in those
days.
\VS{18}The judges
will thoroughly
investigate
the matter, and if the witness
should prove to be false
and to have given
false
testimony
against the accused,
\VS{19}you must do
to him what he had
intended
to do
to the accused. In this way you will purge
evil
from among you.
\VS{20}The rest
of the people will hear
and become afraid
to keep
doing
such evil
among you.
\VS{21}You must not
show pity;
the principle
will be a life
for a life,
an eye
for an eye,
a tooth
for a tooth,
a hand
for a hand,
and a foot
for a foot.

\par }\Chap{20}{\PP \VerseOne{1}When
you go
to war
against
your enemies
and see
chariotry
and troops
who outnumber
you, do not
be afraid
of them,
for
the {\ND{Lord}}
your God,
who brought
you up
out of the land
of Egypt,
is with you.
\VS{2}As you
move forward
for battle,
the priest
will approach
and say
to
the soldiers,
\VS{3}“Listen,
Israel! Today
you
are moving
forward to do battle
with your enemies.
Do not
be
fainthearted.
Do not
fear
and tremble or be terrified because of them,
\VS{4}for
the {\ND{Lord}}
your God
goes
with
you to fight
on your behalf
against your enemies
to give you victory.”
\VS{5}Moreover,
the officers
are to say
to the troops, “Who
among you has
built
a new
house
and not
dedicated
it? He may go
home,
lest
he die
in battle
and someone
else
dedicate it.
\VS{6}Or who
among you has
planted
a vineyard
and not
benefited
from it? He may go
home,
lest
he die
in battle
and someone
else
benefit from it.
\VS{7}Or who
among you
has become engaged
to a woman
but has not
married
her? He may go
home,
lest
he die
in battle
and someone
else
marry her.”
\VS{8}In addition,
the officers
are to say
to
the troops,
“Who
among you is afraid
and fainthearted? He may go
home
so that he will not
make
his fellow
soldier’s heart
as fearful
as his own.”
\VS{9}Then, when
the officers
have finished
speaking,
they must appoint
unit commanders
to lead
the troops.
\par }{\PP \VS{10}When
you approach
a city
to
wage war
against
it, offer it
terms of
peace.
\VS{11}If
it accepts
your terms
and submits
to you, all
the people
found
in it will become
your slaves.
\VS{12}If
it does not
accept terms of peace
but makes
war
with
you, then you are to lay siege to it.
\VS{13}The
{\ND{Lord}}
your God
will deliver
it over to you and you must kill
every
single male
by the sword.
\VS{14}However,
the women,
little children,
cattle,
and anything
else in the city
– all
its plunder
– you may take for yourselves as spoil.
You may take
from your enemies
the plunder
that
the {\ND{Lord}}
your God
has given you.
\VS{15}This
is how you are to deal
with all
those cities
located
far
from
you, those that
do not
belong to these
nearby
nations.
\par }{\SH Laws Concerning War with Canaanite Nations
\par }{\PP \VS{16}As
for the cities
of these
peoples
that
the {\ND{Lord}}
your God
is going to give
you as an inheritance,
you must not
allow a single living thing to survive.
\VS{17}Instead
you must utterly
annihilate them –
the Hittites,
Amorites,
Canaanites,
Perizzites,
Hivites,
and Jebusites –
just
as the
{\ND{Lord}}
your God
has commanded you,
\VS{18}so
that
they cannot
teach
you all
the abhorrent
ways they worship
their gods,
causing you to sin
against the
{\ND{Lord}}
your God.
\VS{19}If
you besiege
a city
for a long
time
while attempting to
capture
it, you must not
chop down
its trees,
for
you may eat
fruit from
them
and should not
cut
them down.
A tree
in the field
is not human
that
you should besiege it!
\VS{20}However,
you may chop down
any tree
you know
is not
suitable for food,
and you may use it
to build
siege
works against
the city
that is
making
war
with
you until
that city falls.

\par }\Chap{21}{\PP \VerseOne{1}If
a homicide
victim should be found
lying
in a field
in the land
the {\ND{Lord}}
your God
is giving
you, and no
one knows
who
killed him,
\VS{2}your elders
and judges
must go out
and measure
how far it is to
the cities
in the vicinity
of the corpse.
\VS{3}Then
the elders
of the city
nearest
to
the corpse
must take
from the herd
a heifer
that
has not
been worked
– that
has never
pulled
with the yoke –
\VS{4}and bring
the heifer
down
to
a wadi
with flowing water,
to a valley that
is neither
plowed
nor
sown.
There
at the wadi
they are to break the
heifer’s
neck.
\VS{5}Then the Levitical
priests
will approach
(for
the {\ND{Lord}}
your God
has chosen
them to serve
him and to pronounce blessings
in his name,
and to decide every
judicial verdict )
\VS{6}and all
the elders
of that city
nearest
the corpse
must wash
their hands
over
the heifer
whose neck
was broken in the valley.
\VS{7}Then
they must proclaim, “Our hands
have not
spilled
this
blood,
nor
have
we witnessed
the crime.
\VS{8}Do not blame
your people
Israel
whom
you redeemed,
O
{\ND{Lord}}, and do not
hold them accountable
for the bloodshed
of an innocent
person.” Then atonement
will be made for the bloodshed.
\VS{9}In this manner you
will purge
out the guilt of innocent
blood
from among
you, for
you must do
what is right
before
the {\ND{Lord}}.
\par }{\SH Laws Concerning Wives
\par }{\PP \VS{10}When
you go out
to do battle
with your enemies
and the
{\ND{Lord}}
your God
allows
you to prevail and you take prisoners,
\VS{11}if you should see
among them
an attractive
woman
whom you wish
to take
as a wife,
\VS{12}you may bring
her back to
your house.
She must shave
her head,
trim
her nails,
\VS{13}discard
the clothing
she was wearing when captured,
and stay
in
your house,
lamenting
for her father
and mother
for a full month.
After
that you may have sexual relations
with
her and become her husband
and she your wife.
\VS{14}If
you are not
pleased
with her, then
you must let
her go where she pleases.
You cannot in any case sell
her; you must not take
advantage
of her, since you have
already humiliated her.
\par }{\SH Laws Concerning Children
\par }{\PP \VS{15}Suppose
a man
has two
wives,
one
whom he loves
more than the other,
and they both bear
him sons,
with the firstborn
being the child
of the less
loved wife.
\VS{16}In the day
he divides his inheritance
he must not
appoint
as firstborn
the son
of the favorite
wife in place of the other wife’s son
who is actually the firstborn.
\VS{17}Rather,
he must acknowledge
the son
of the less
loved wife as firstborn
and give
him the double
portion of all
he has,
for that
son is the beginning
of his father’s procreative power –
to him should go the right
of the firstborn.
\par }{\PP \VS{18}If
a person
has a stubborn,
rebellious
son
who pays no
attention
to his father
or mother,
and they discipline
him to no
avail,
\VS{19}his father
and mother
must seize
him and bring
him to
the elders
at the gate
of his city.
\VS{20}They must declare
to
the elders
of his city,
“Our son
is
stubborn
and rebellious
and pays no
attention to what we say – he is a glutton and drunkard.”
\VS{21}Then
all
the men
of his city
must stone
him to death.
In this way you will purge
out wickedness
from among
you, and all
Israel
will hear
about it and be afraid.
\par }{\SH Disposition of a Criminal’s Remains
\par }{\PP \VS{22}If
a person
commits
a sin punishable by
death
and is executed,
and you hang
the corpse on
a tree,
\VS{23}his body
must not
remain all night
on
the tree;
instead
you must make certain you bury
him
that
same day,
for
the one who is left exposed
on a tree is cursed
by God.
You must not
defile
your land
which
the {\ND{Lord}}
your God
is giving
you as an inheritance.

\par }\Chap{22}{\PP \VerseOne{1}When you see
your neighbor’s
ox
or
sheep
going astray,
do not ignore
it; you must return
it without fail
to your neighbor.
\VS{2}If
the owner
does not
live near
you or you do not
know
who the owner is, then you must corral
the animal at your house
and let it stay
with
you until
the owner
looks for
it; then you must return it to him.
\VS{3}You shall do
the same
to his donkey,
his clothes,
or anything
else your neighbor
has lost
and you have found;
you must not
refuse
to get involved.
\VS{4}When you see
your neighbor’s
donkey
or
ox
fallen
along the road,
do not ignore
it; instead, you must be sure to help
him
get
the animal on its feet again.
\par }{\PP \VS{5}A woman
must not
wear
men’s
clothing,
nor should a man dress up in women’s
clothing, for
anyone
who does
this
is offensive
to the
{\ND{Lord}}
your God.
\par }{\PP \VS{6}If
you happen to notice
a bird’s
nest
along
the road,
whether in a tree
or
on
the ground,
and there are chicks
or
eggs
with the mother
bird
sitting
on
them, you must not
take
the mother
from the young.
\VS{7}You must be sure
to let
the mother
go,
but you may take
the young
for yourself. Do this so that
it may go well
with you and you may have a long
life.
\par }{\PP \VS{8}If
you build
a new
house,
you must construct
a guard rail
around your roof
to avoid
being culpable
in the event
someone
should fall
from it.
\par }{\SH Illustrations of the Principle of Purity
\par }{\PP \VS{9}You must not
plant
your vineyard
with two kinds
of seed; otherwise
the entire yield,
both of the seed
you plant
and the produce
of the vineyard, will be defiled.
\VS{10}You must not
plow
with an ox
and a donkey
harnessed together.
\VS{11}You must not
wear
clothing
made with wool
and linen
meshed together.
\VS{12}You shall make
yourselves tassels
for the four
corners
of the clothing you wear.
\par }{\SH Purity in the Marriage Relationship
\par }{\PP \VS{13}Suppose
a man
marries
a woman,
has sexual relations
with her, and then
rejects her,
\VS{14}accusing
her of impropriety
and defaming
her reputation
by saying,
“I married
this
woman
but when I had sexual
relations
with her I discovered
she was not
a virgin!”
\VS{15}Then
the father
and mother
of the
young woman
must produce the evidence
of virginity
for the elders
of the city
at the gate.
\VS{16}The young
woman’s
father
must say
to
the elders,
“I gave
my daughter
to this
man
and he has rejected
her.
\VS{17}Moreover,
he has raised accusations
of impropriety by
saying,
‘I discovered
your daughter
was not
a virgin,’
but this
is the evidence
of my daughter’s
virginity!” The cloth
must then be spread
out before
the city’s
elders.
\VS{18}The elders
of that city
must then seize
the man
and punish him.
\VS{19}They will fine
him one hundred
shekels of silver
and give
them to the young woman’s
father,
for
the man who made
the accusation
ruined the reputation
of an Israelite
virgin.
She will then become
his wife
and he may never
divorce
her as long as
he lives.
\par }{\PP \VS{20}But if
the accusation
is true
and the young woman
was
not
a virgin,
\VS{21}the men
of her city
must bring
the young
woman to
the door
of her father’s
house
and stone
her to
death,
for
she has done
a disgraceful thing
in Israel
by behaving like a prostitute
while living in her father’s
house.
In this way you will purge
evil
from among you.
\par }{\PP \VS{22}If
a man
is caught
having sexual
relations with
a married
woman
both
the man
who had relations
with
the woman
and the woman
herself
must die;
in this way you will purge
evil
from Israel.
\par }{\PP \VS{23}If
a virgin
is engaged
to a man
and another man
meets
her in the city
and has sexual relations
with her,
\VS{24}you must bring
the two
of them to
the gate
of that city
and stone
them to death,
the young woman
because she did not
cry out
though in the city
and the man
because he violated
his neighbor’s
fiancée;
in this way you will purge
evil
from among you.
\VS{25}But if
the man
came across
the engaged
woman
in the field
and overpowered
her and raped
her, then only the rapist
must die.
\VS{26}You must not
do
anything
to the young woman – she has done nothing deserving of death. This case is the same as when someone attacks another person and murders him,
\VS{27}for
the man met
her in the field
and the engaged
woman
cried out,
but there was no
one to rescue her.
\par }{\PP \VS{28}Suppose
a man
comes across
a virgin
who
is not
engaged
and overpowers
and rapes
her and they are discovered.
\VS{29}The man
who has raped
her must pay
her
father
fifty
shekels of silver
and she must become
his wife
because
he has violated
her; he may
never
divorce
her as long as he lives.
\par }{\PP \VS{30}A man
may not
marry
his father’s
former wife
and in this way dishonor
his father.

\par }\Chap{23}{\PP \VerseOne{1}A man
with crushed
or severed genitals
may not
enter
the assembly
of the {\ND{Lord}}.
\VS{2}A person of illegitimate
birth may not
enter
the assembly
of the {\ND{Lord}}; to the tenth
generation
no one related to him may do so.
\par }{\PP \VS{3}An Ammonite
or Moabite
may not
enter
the assembly
of the {\ND{Lord}}; to the tenth
generation
none
of their
descendants shall ever do so,
\VS{4}for they did not
meet
you with food
and water
on the way
as you came
from Egypt,
and furthermore,
they hired
Balaam
son
of Beor
of Pethor
in Aram Naharaim
to curse you.
\VS{5}But the
{\ND{Lord}}
your God
refused
to listen
to
Balaam
and changed
the curse
to a blessing,
for
the {\ND{Lord}}
your God
loves you.
\VS{6}You must not
seek
peace
and prosperity
for them through all
the ages to come.
\VS{7}You must not
hate
an Edomite,
for
he is
your relative;
you must not
hate
an Egyptian,
for
you lived as a foreigner
in his land.
\VS{8}Children
of the third
generation
born
to them may enter
the assembly
of the {\ND{Lord}}.
\par }{\SH Purity in Personal Hygiene
\par }{\PP \VS{9}When
you go out
as an army
against
your enemies,
guard
yourselves against anything
impure.
\VS{10}If
there
is someone
among you who
is
impure
because of some nocturnal
emission,
he must leave
the camp;
he may not
reenter
it immediately.
\VS{11}When
evening
arrives
he must wash
himself with water
and then at sunset
he may reenter
the camp.
\par }{\PP \VS{12}You are to have a place outside
the camp
to serve
as a latrine.
\VS{13}You must have a spade
among
your other equipment
and when
you relieve yourself
outside
you must dig a hole
with the spade and then turn
and cover
your excrement.
\VS{14}For
the {\ND{Lord}}
your God
walks
about in the middle
of your camp
to deliver
you
and defeat
your enemies
for you. Therefore your camp
should be holy,
so that he does not
see
anything indecent
among you and turn
away from you.
\par }{\SH Purity in the Treatment of the Nonprivileged
\par }{\PP \VS{15}You must not
return an escaped
slave
to
his master
when he has
run
away
to
you.
\VS{16}Indeed,
he may live
among
you in any place
he chooses,
in whichever of your villages
he prefers;
you must not
oppress him.
\par }{\SH Purity in Cultic Personnel
\par }{\PP \VS{17}There must never
be
a sacred prostitute
among the young women
of Israel
nor
a sacred male prostitute
among the young
men of Israel.
\VS{18}You must never
bring
the pay
of a female prostitute
or the wage
of a male prostitute
into the temple
of the {\ND{Lord}}
your God
in fulfillment of any
vow,
for
both
of these are abhorrent
to the
{\ND{Lord}}
your God.
\par }{\SH Respect for Others’ Property
\par }{\PP \VS{19}You must not
charge interest on
a loan
to your fellow
Israelite, whether on
money,
food,
or anything
else that has been loaned
with interest.
\VS{20}You may lend
with interest
to a foreigner,
but not
to your fellow
Israelite;
if you keep
this command the
{\ND{Lord}}
your God
will bless
you in all
you undertake
in the land
you
are about
to enter
to possess.
\VS{21}When
you make
a vow
to the
{\ND{Lord}}
your God
you must not
delay
in fulfilling
it, for otherwise he will surely
hold
you accountable
as a sinner.
\VS{22}If
you refrain
from making a vow,
it will not
be
sinful.
\VS{23}Whatever you vow,
you must be careful
to do
what
you have promised,
such as what you have vowed
to the
{\ND{Lord}}
your God
as a freewill offering.
\VS{24}When
you enter
the vineyard
of your neighbor
you may eat
as many grapes
as you please,
but you must not
take away
any in
a container.
\VS{25}When
you go
into the ripe grain
fields of your neighbor
you may pluck
off the kernels
with your hand,
but you must not
use
a sickle
on
your neighbor’s
ripe grain.

\par }\Chap{24}{\PP \VerseOne{1}If
a man
marries
a woman
and she does
not
please him
because
he has found
something offensive
in her, then he may draw up a divorce
document, give
it
to her, and evict
her from his
house.
\VS{2}When she has left
him she may go
and become
someone
else’s wife.
\VS{3}If the second husband
rejects
her and then divorces
her, gives
her the papers,
and evicts her from
his house,
or
if
the second husband
who
married
her
dies,
\VS{4}her first
husband
who divorced
her is not
permitted
to remarry her
after
she has become ritually impure,
for
that is
offensive
to the
{\ND{Lord}}. You must not
bring guilt
on the land
which
the {\ND{Lord}}
your God
is giving
you as an inheritance.
\par }{\PP \VS{5}When
a man
is newly
married,
he need not
go
into the army
nor
be obligated
in any
way;
he must be free
to stay at home
for a full
year
and bring joy
to the wife
he has
married.
\par }{\PP \VS{6}One must not
take
either lower
or upper
millstones
as security
on a loan, for that
is like taking
a life
itself
as security.
\par }{\PP \VS{7}If
a man
is found
kidnapping
a person
from among his fellow
Israelites,
and regards
him as mere property and sells
him, that kidnapper
must die.
In this way you will purge
evil
from among you.
\par }{\SH Respect for Human Dignity
\par }{\PP \VS{8}Be careful
during an outbreak
of leprosy
to follow precisely all
that
the Levitical
priests
instruct
you; as
I have commanded
them, so
you should
do.
\VS{9}Remember
what
the
{\ND{Lord}}
your God
did
to Miriam
along the way
after you left
Egypt.
\par }{\PP \VS{10}When
you make any kind of loan
to your neighbor,
you may not
go
into his house
to claim
what he is offering as security.
\VS{11}You must
stand
outside
and the person
to whom
you
are making the loan
will bring out
to
you what he is offering as security.
\VS{12}If
the person
is poor
you may not
use what he gives you as security for a covering.
\VS{13}You must by all means return
to him at sunset
the item he gave you as security
so that he may sleep
in his outer garment
and bless
you for it; it will be
considered a just deed
by the
{\ND{Lord}}
your God.
\par }{\PP \VS{14}You must not
oppress
a lowly and poor
servant,
whether one from among your fellow
Israelites or
from the resident foreigners
who are living in your land
and villages.
\VS{15}You must pay
his wage
that very day
before the sun
sets,
for
he is poor
and his life
depends
on
it. Otherwise he will cry out
to
the {\ND{Lord}}
against you, and you will be
guilty of sin.
\par }{\PP \VS{16}Fathers
must not
be put to death
for what their children
do, nor
children
for what their fathers
do; each
must be put to death
for his own sin.
\par }{\PP \VS{17}You must not
pervert
justice
due a resident foreigner
or an orphan,
or
take
a widow’s
garment
as security for a loan.
\VS{18}Remember
that
you were slaves
in Egypt
and that the
{\ND{Lord}}
your God
redeemed
you from there;
therefore
I am
commanding
you to do
all
this.
\VS{19}Whenever
you reap
your harvest
in your field
and leave some unraked
grain there,
you must not
return
to get
it; it should
go to the resident foreigner,
orphan,
and widow
so
that the
{\ND{Lord}}
your God
may bless
all
the work you do.
\VS{20}When
you beat
your olive
tree you must not
repeat
the procedure;
the remaining olives belong to the resident foreigner,
orphan,
and widow.
\VS{21}When
you gather the grapes
of your vineyard
you must not
do so
a second time;
they should go
to the resident foreigner,
orphan,
and widow.
\VS{22}Remember
that
you were slaves
in the land
of Egypt;
therefore,
I am
commanding
you to do
all
this.

\par }\Chap{25}{\PP \VerseOne{1}If
controversy
arises
between
people,
they should go to
court for judgment.
When the judges
hear
the case, they shall exonerate
the innocent
but
condemn
the guilty.
\VS{2}Then,
if
the guilty person is sentenced
to a beating,
the judge
shall force him
to lie down
and be
beaten
in his presence
with the number
of blows
his wicked behavior deserves.
\VS{3}The judge may
sentence him to forty
blows,
but no
more.
If
he is struck
with more than
these,
you might view
your fellow Israelite with contempt.
\par }{\PP \VS{4}You must not
muzzle
your ox
when it is treading grain.
\par }{\SH Respect for the Sanctity of Others
\par }{\PP \VS{5}If
brothers
live
together
and one
of them
dies
without
having a son,
the dead
man’s wife
must not
remarry
someone
outside
the family. Instead, her
late husband’s brother
must
go
to her, marry
her, and perform the duty of a brother-in-law.
\VS{6}Then
the first son
she bears
will continue
the name
of the dead
brother,
thus preventing
his name
from being blotted out
of Israel.
\VS{7}But if
the man
does not
want
to marry
his brother’s widow,
then she
must go
to
the elders
at the town gate
and say,
“My husband’s brother
refuses
to preserve
his brother’s
name
in Israel;
he is unwilling
to perform the duty of a brother-in-law to me!”
\VS{8}Then the elders
of his city
must summon
him and speak
to him.
If he persists,
saying,
“I don’t want
to marry her,”
\VS{9}then his sister-in-law
must approach
him
in view
of the elders,
remove his sandal
from his foot,
and spit
in his face.
She will then respond,
“Thus
may it be done
to any man
who
does not
maintain
his brother’s
family line!”
\VS{10}His family name
will be referred
to in Israel
as “the family
of the one
whose sandal was removed.”
\par }{\PP \VS{11}If
two men
get into
a hand-to-hand
fight,
and the wife
of one
of them gets involved
to help her husband
against his attacker,
and she reaches out
her hand
and grabs
his genitals,
\VS{12}then you must cut
off her hand
– do not
pity her.
\par }{\PP \VS{13}You must not
have in your bag
different stone
weights,
a heavy
and a light one.
\VS{14}You must not
have in your house
different measuring containers,
a large
and a small one.
\VS{15}You must have
an accurate
and correct
stone weight
and an accurate
and correct
measuring container,
so that
your life
may be extended
in
the land
the {\ND{Lord}}
your God
is about to give you.
\VS{16}For
anyone
who acts
dishonestly
in these
ways is abhorrent
to the
{\ND{Lord}}
your God.
\par }{\SH Treatment of the Amalekites
\par }{\PP \VS{17}Remember
what
the Amalekites
did
to you on your way
from
Egypt,
\VS{18}how
they met
you along the way
and cut off
all
your stragglers
in the rear
of the march when you
were exhausted
and tired;
they were unafraid
of God.
\VS{19}So when
the {\ND{Lord}}
your God
gives you relief
from all
the enemies
who surround
you in the land
he is giving
you as an inheritance,
you must wipe out
the
memory
of the Amalekites
from under
heaven –
do not
forget!

\par }\Chap{26}{\PP \VerseOne{1}When
you enter
the land
that
the {\ND{Lord}}
your God
is giving
you as an inheritance,
and you occupy
it and live in it,
\VS{2}you must take
the first
of all
the ground’s
produce
you harvest
from the land
the {\ND{Lord}}
your God
is giving
you, place
it in a basket,
and go
to
the place
where
he
chooses
to locate
his name.
\VS{3}You must go
to
the priest
in office at that time
and say
to him,
“I declare
today
to the
{\ND{Lord}}
your God
that
I have come
into
the land
that
the {\ND{Lord}}
promised
to our ancestors
to give us.”
\VS{4}The priest
will then take
the basket
from you and set
it before
the altar
of the {\ND{Lord}}
your God.
\VS{5}Then
you must affirm
before
the {\ND{Lord}}
your {\ND{God}}, “A wandering
Aramean
was my ancestor,
and he went down
to Egypt
and lived
there
as a foreigner with a household few
in number,
but there
he became
a great,
powerful,
and numerous
people.
\VS{6}But
the Egyptians
mistreated
and oppressed
us, forcing
us to do burdensome
labor.
\VS{7}So we cried out
to
the {\ND{Lord}}, the God
of our ancestors,
and he
heard
us and saw
our humiliation,
toil,
and oppression.
\VS{8}Therefore the
{\ND{Lord}}
brought us out
of Egypt
with tremendous
strength
and power,
as well as with great
awe-inspiring
signs
and wonders.
\VS{9}Then he brought
us to
this
place
and gave
us this
land,
a land
flowing
with milk
and honey.
\VS{10}So now,
look! I have brought
the first
of the ground’s
produce
that
you,
{\ND{Lord}},
have given
me.” Then you must set
it down before
the {\ND{Lord}}
your God
and worship
before
him.
\VS{11}You will celebrate
all
the good
things that
the {\ND{Lord}}
your God
has given you
and your family,
along with the Levites
and the resident foreigners
among you.
\par }{\SH Presentation of the Third-year Tithe
\par }{\PP \VS{12}When
you finish
tithing
all
your income
in the third
year
(the year
of tithing), you must give
it to the Levites,
the resident foreigners,
the orphans,
and the widows
so that they may eat
to their satisfaction
in your villages.
\VS{13}Then you shall say
before
the {\ND{Lord}}
your God,
“I have removed
the sacred
offering from
my house
and given
it to the Levites,
the resident foreigners,
the orphans,
and the widows
just
as you have commanded
me. I have not
violated
or
forgotten
your commandments.
\VS{14}I have not
eaten
anything when I was in mourning,
or
removed
any
of it while ceremonially unclean,
or offered
any
of it
to the dead;
I have obeyed
you and have done
everything
you have
commanded me.
\VS{15}Look
down from
your holy
dwelling place
in heaven
and bless
your people
Israel
and the land
you have
given
us, just
as you promised
our ancestors
– a land
flowing
with milk
and honey.”
\par }{\SH Narrative Interlude
\par }{\PP \VS{16}Today
the {\ND{Lord}}
your God
is commanding
you to keep
these
statutes
and ordinances,
something
you must do
with all
your heart
and soul.
\VS{17}Today
you have declared
the {\ND{Lord}}
to be
your God,
and that you will walk
in his ways,
keep
his statutes,
commandments,
and ordinances,
and obey
him.
\VS{18}And today
the {\ND{Lord}}
has declared
you to be
his special
people
(as
he already promised
you) so you may keep
all
his commandments.
\VS{19}Then he will elevate
you above
all
the nations
he has
made
and you will receive praise,
fame,
and honor.
You will be
a people
holy
to the
{\ND{Lord}}
your God,
as he has said.

\par }\Chap{27}{\PP \VerseOne{1}Then Moses
and the elders
of Israel
commanded
the people: “Pay attention
to all
the commandments
I am
giving
you today.
\VS{2}When
you cross
the Jordan River
to
the land
the {\ND{Lord}}
your God
is giving
you, you must erect
great
stones
and cover them with plaster.
\VS{3}Then you must inscribe
on
them
all
the words
of this
law
when you cross
over, so
that
you may enter
the land
the {\ND{Lord}}
your God
is giving
you, a land
flowing
with milk
and honey
just
as the
{\ND{Lord}}, the God
of your ancestors,
said to you.
\VS{4}So when
you cross
the Jordan
you must erect
on Mount
Ebal
these
stones
about which
I am
commanding
you today,
and you must cover them with plaster.
\VS{5}Then you must build
an altar
there
to the
{\ND{Lord}}
your God,
an altar
of stones
– do not
use
an iron
tool on them.
\VS{6}You must build
the altar
of the {\ND{Lord}}
your God
with whole
stones
and offer
burnt offerings
on
it to the
{\ND{Lord}}
your God.
\VS{7}Also you must offer fellowship offerings
and eat
them there,
rejoicing
before
the {\ND{Lord}}
your God.
\VS{8}You must inscribe
on
the stones
all
the words
of this
law,
making
them clear.”
\par }{\PP \VS{9}Then
Moses
and the Levitical
priests
spoke to
all
Israel: “Be quiet
and pay attention,
Israel.
Today
you have become
the people
of the {\ND{Lord}}
your God.
\VS{10}You must obey
him and keep
his commandments
and statutes
that
I
am giving
you today.”
\VS{11}Moreover, Moses
commanded
the people
that day:
\VS{12}“The following
tribes must stand
to bless
the people
on
Mount
Gerizim
when you cross
the Jordan: Simeon,
Levi,
Judah,
Issachar,
Joseph,
and Benjamin.
\VS{13}And these
other tribes must stand
for the curse
on Mount
Ebal: Reuben,
Gad,
Asher,
Zebulun,
Dan,
and Naphtali.
\par }{\SH The Covenant Curses
\par }{\PP \VS{14}“The Levites
will call
out to
every
Israelite
with a loud
voice:
\VS{15}‘Cursed
is the one
who makes
a carved
or metal image
– something abhorrent
to the
{\ND{Lord}}, the work
of the craftsman –
and sets it up
in a secret place.’
Then all
the people
will say,
‘Amen!’
\VS{16}‘Cursed
is the one who disrespects
his father
and mother.’
Then all
the people
will say, ‘Amen!’
\VS{17}‘Cursed
is the one who moves
his neighbor’s
boundary
marker.’ Then all
the people
will say, ‘Amen!’
\VS{18}‘Cursed
is the one who misleads
a blind
person on the road.’
Then all
the people
will say, ‘Amen!’
\VS{19}‘Cursed
is the one who perverts
justice
for the resident foreigner,
the orphan,
and the widow.’
Then all
the people
will say, ‘Amen!’
\VS{20}‘Cursed
is the one who has sexual relations
with
his father’s
former wife,
for
he dishonors
his father.’
Then all
the people
will say, ‘Amen!’
\VS{21}‘Cursed
is the one
who commits bestiality.’
Then all
the people
will say, ‘Amen!’
\VS{22}‘Cursed
is the one who has sexual
relations with
his sister,
the daughter
of either his father
or
mother.’
Then all
the people
will say, ‘Amen!’
\VS{23}‘Cursed
is the one who has sexual
relations with
his mother-in-law.’
Then all
the people
will say, ‘Amen!’
\VS{24}‘Cursed
is the one who kills
his neighbor
in private.’
Then all
the people
will say, ‘Amen!’
\VS{25}‘Cursed
is the one who takes
a bribe
to kill
an innocent
person.’
Then all
the people
will say,
‘Amen!’
\VS{26}‘Cursed
is the one who refuses
to keep the words
of this
law.’
Then all
the people
will say, ‘Amen!’

\par }\Chap{28}{\PP \VerseOne{1}“If
you indeed
obey
the {\ND{Lord}}
your God
and are careful
to observe
all
his commandments
I am
giving
you today,
the {\ND{Lord}}
your God
will elevate
you above
all
the nations
of the earth.
\VS{2}All
these
blessings
will come
to you in abundance
if
you obey
the {\ND{Lord}}
your God:
\VS{3}You
will be blessed
in the city
and blessed
in the field.
\VS{4}Your children
will be blessed,
as well as the produce
of your soil,
the offspring
of your livestock,
the calves of your herds,
and the lambs
of your flocks.
\VS{5}Your basket
and your mixing bowl
will be blessed.
\VS{6}You will be blessed
when you
come
in and blessed
when you
go out.
\VS{7}The

{\ND{Lord}}
will cause your enemies
who attack
you to be struck
down before
you; they will attack you from one
direction
but flee
from you in seven
different directions.
\VS{8}The
{\ND{Lord}}
will decree
blessing
for you with respect to your barns
and in everything
you do – yes, he will bless you in the land he is giving you.
\VS{9}The
{\ND{Lord}}
will designate
you as his holy
people
just
as he promised
you, if
you keep
his commandments
and obey him.
\VS{10}Then all
the peoples
of the earth
will see
that
you belong to the
{\ND{Lord}},
and they will respect
you.
\VS{11}The

{\ND{Lord}}
will greatly
multiply
your children,
the offspring
of your livestock,
and the produce
of your soil
in the land
which
he promised
your ancestors
he would give you.
\VS{12}The
{\ND{Lord}}
will open
for you his good
treasure
house, the
heavens,
to give
you rain
for the land
in its season
and to bless
all
you do;
you will lend
to many
nations
but you
will not
borrow from any.
\VS{13}The
{\ND{Lord}}
will make you
the head
and not
the tail,
and you will
always
end up at the top and not
at the bottom,
if
you obey
his commandments
which
I am
urging
you today
to be careful
to do.
\VS{14}But you must not
turn
away from all
the commandments
I am
giving
you today,
to either the right
or left,
nor pursue
other
gods
and worship them.
\par }{\SH Curses as Reversal of Blessings
\par }{\PP \VS{15}“But if
you ignore
the {\ND{Lord}}
your God
and are not careful to keep
all
his commandments
and statutes
I
am giving
you today,
then all
these
curses
will come
upon
you in full force:
\VS{16}You
will be cursed
in the city
and cursed
in the field.
\VS{17}Your basket
and your mixing bowl
will be cursed.
\VS{18}Your children
will be cursed,
as well as the produce
of your soil,
the calves of your herds,
and the lambs
of your flocks.
\VS{19}You will be cursed
when you
come
in and cursed
when you
go out.
\par }{\SH Curses by Disease and Drought
\par }{\PP \VS{20}“The
{\ND{Lord}}
will send
on you a curse,
confusing
you and opposing you
in everything
you undertake
until
you are destroyed
and quickly
perish
because
of the evil
of your deeds,
in that
you have forsaken me.
\VS{21}The
{\ND{Lord}}
will plague
you with deadly diseases
until
he has completely removed
you from
the land
you
are about
to possess.
\VS{22}He will afflict
you with weakness,
fever,
inflammation,
infection,
sword,
blight,
and mildew;
these will attack
you until
you perish.
\VS{23}The sky
above
your heads
will be
bronze
and the earth
beneath
you iron.
\VS{24}The
{\ND{Lord}}
will make
the rain
of your land
powder
and dust;
it will come down
on
you from
the sky
until
you are destroyed.
\par }{\SH Curses by Defeat and Deportation
\par }{\PP \VS{25}“The
{\ND{Lord}}
will allow
you to be struck
down before
your enemies;
you will attack them from one
direction
but flee
from them
in seven
directions
and will become an object of terror
to all
the kingdoms
of the earth.
\VS{26}Your carcasses
will be
food
for every
bird
of the sky
and wild animal
of the earth,
and there will be no
one to chase them off.
\VS{27}The
{\ND{Lord}}
will afflict
you with the boils
of Egypt
and with tumors,
eczema,
and scabies,
all of which
cannot
be healed.
\VS{28}The
{\ND{Lord}}
will also subject
you to madness,
blindness,
and confusion
of mind.
\VS{29}You will feel
your way along
at noon
like
the blind
person does in darkness
and you will not
succeed
in anything you do;
you will be
constantly
oppressed
and continually
robbed,
with no
one to save you.
\VS{30}You will be engaged
to a woman
and another
man
will rape
her. You will build
a house
but not
live in it. You will plant
a vineyard
but not
even begin to use it.
\VS{31}Your ox
will be slaughtered
before your very eyes
but you will not
eat
of it. Your donkey
will be stolen
from
you as you watch and will not
be returned
to you. Your flock
of sheep will be given
to your enemies
and there will be no
one to save you.
\VS{32}Your sons
and daughters
will be given
to another
people
while
you look
on in vain
all
day,
and you will be powerless
to do anything about it.
\VS{33}As for the produce
of your land
and all
your labor,
a people
you do not
know
will consume
it, and you will be
nothing but
oppressed
and crushed
for the rest of your lives.
\VS{34}You will
go insane
from seeing
all this.
\VS{35}The
{\ND{Lord}}
will afflict
you in your knees
and on
your legs
with painful, incurable
boils
– from the soles
of your feet
to
the top of
your head.
\VS{36}The
{\ND{Lord}}
will force you
and your king
whom
you will appoint
over
you to
go
away to
a people
whom
you
and your ancestors
have not
known,
and you
will serve
other
gods
of wood
and stone
there.
\VS{37}You will become
an occasion of horror,
a proverb,
and an object of ridicule
to all
the peoples
to whom
the {\ND{Lord}}
will drive you.
\par }{\SH The Curse of Reversed Status
\par }{\PP \VS{38}“You will take
much
seed
to the field
but gather
little
harvest, because
locusts
will consume it.
\VS{39}You will plant
vineyards
and cultivate
them, but you will not
drink
wine
or
gather
in grapes, because
worms
will eat them.
\VS{40}You will have olive
trees throughout
your territory
but you will not
anoint
yourself with olive oil,
because
the olives will drop off the trees while still unripe.
\VS{41}You will bear
sons
and daughters
but not
keep them, because
they will be taken into captivity.
\VS{42}Whirring locusts
will take over every
tree
and all
the produce
of your soil.
\VS{43}The foreigners
who reside among
you will become higher
and higher
over
you and you
will become lower
and lower.
\VS{44}They will lend
to you
but you will not
lend
to them; they will become
the head
and you
will become
the tail!
\par }{\PP \VS{45}All
these
curses
will fall
on
you, pursuing
and overtaking
you until
you are destroyed,
because
you would not
obey
the {\ND{Lord}}
your God
by keeping
his commandments
and statutes
that
he has given you.
\VS{46}These curses will be
a perpetual
sign
and wonder
with reference
to you and your descendants.
\par }{\SH The Curse of Military Siege
\par }{\PP \VS{47}“Because
you have not
served
the {\ND{Lord}}
your God
joyfully
and wholeheartedly
with the abundance
of everything you have,
\VS{48}instead in hunger,
thirst,
nakedness,
and poverty
you will serve
your enemies
whom
the {\ND{Lord}}
will send
against you. They will place
an iron
yoke
on
your neck
until
they have destroyed you.
\VS{49}The
{\ND{Lord}}
will raise
up a distant
nation
against
you, one from the other side of the earth
as
the eagle
flies,
a nation
whose
language
you will not
understand,
\VS{50}a nation
of stern
appearance that
will have no
regard
for the elderly
or
pity
for the young.
\VS{51}They will devour
the offspring
of your livestock
and the produce
of your soil
until
you are destroyed.
They will not
leave
you with any grain,
new wine,
olive oil,
calves of your herds,
or lambs
of your flocks
until
they have destroyed you.
\VS{52}They will besiege
all
of your villages
until
all of your high
and fortified
walls
collapse
– those in which
you
put your confidence
throughout
the land.
They will besiege
all
your villages
throughout
the land
the {\ND{Lord}}
your God
has
given you.
\VS{53}You will then
eat
your own offspring,
the flesh
of the sons
and daughters
the {\ND{Lord}}
your God
has
given you, because of the severity
of the siege
by which
your enemies
will constrict you.
\VS{54}The man
among you who is by nature
tender
and sensitive
will turn against
his brother,
his beloved
wife,
and his remaining
children.
\VS{55}He will
withhold
from all
of them
his children’s
flesh
that
he is
eating
(since there is
nothing else left), because of the severity
of the siege
by which
your enemy
will constrict
you in your villages.
\VS{56}Likewise, the most tender
and delicate
of your women, who
would never
think
of putting
even the sole
of her foot
on
the ground
because of her daintiness,
will turn against
her beloved husband,
her sons
and daughters,
\VS{57}and will secretly eat
her afterbirth
and her newborn
children
(since
she has nothing else), because
of the severity
of the siege
by which
your enemy
will constrict
you in your villages.
\par }{\SH The Curse of Covenant Termination
\par }{\PP \VS{58}“If
you refuse
to obey
all
the words
of this
law,
the things written
in this scroll,
and refuse to fear
this
glorious
and awesome
name,
the {\ND{Lord}}
your God,
\VS{59}then the
{\ND{Lord}}
will increase
your punishments
and those of your descendants
– great
and long-lasting
afflictions
and severe,
enduring
illnesses.
\VS{60}He will infect you with all
the diseases
of Egypt
that
you dreaded,
and they will persistently
afflict you.
\VS{61}Moreover,
the {\ND{Lord}}
will bring upon
you every
kind of sickness
and plague
not
mentioned in
this scroll
of commandments,
until
you have perished.
\VS{62}There
will be very few
of you left,
though at one time you were
as numerous
as the stars
in the sky,
because
you will have
disobeyed
the {\ND{Lord}}
your God.
\VS{63}This is what will happen: Just
as the
{\ND{Lord}}
delighted
to do good
for you and make you numerous,
he
will take delight
in
destroying
and decimating
you. You will be uprooted
from the land
you
are about
to possess.
\VS{64}The
{\ND{Lord}}
will scatter
you among all
nations,
from one end
of the earth
to
the other.
There
you will worship
other
gods
that
neither you
nor your ancestors
have known,
gods of wood
and stone.
\VS{65}Among those
nations
you will have no
rest
nor
will there be
a place of peaceful rest
for the soles
of your feet,
for there
the {\ND{Lord}}
will give
you an anxious
heart,
failing
eyesight,
and a spirit
of despair.
\VS{66}Your life
will hang
in doubt before
you; you will be terrified
by night
and day
and will have no
certainty
of surviving from one day to the next.
\VS{67}In the morning
you will say,
‘If only
it
were evening!’ And in the evening
you will say,
‘I wish
it
were
morning!’ because
of the things
you will fear
and the things
you will see.
\VS{68}Then the
{\ND{Lord}}
will make you return
to Egypt
by
ship,
over a route
I said
to you that you would never
see
again.
There
you will sell
yourselves to your enemies
as male and female
slaves,
but no
one will buy you.”

\par }\Chap{29}{\PP \VerseOne{1} These
are the words
of the covenant
that
the {\ND{Lord}}
commanded
Moses
to make
with
the people of Israel
in the land
of Moab,
in addition
to the covenant
he had
made
with
them at Horeb.
\par }{\SH The Exodus, Wandering, and Conquest Reviewed
\par }{\PP \VS{2}Moses
proclaimed
to
all
Israel
as follows: “You
have seen
all
that
the {\ND{Lord}}
did
in the land
of Egypt
to Pharaoh,
all
his servants,
and his land.
\VS{3}Your eyes
have
seen
the great
judgments,
those
signs
and mighty
wonders.
\VS{4}But to this
very
day
the {\ND{Lord}}
has not
given
you an understanding
mind,
perceptive
eyes,
or discerning ears!
\VS{5}I have led
you through the desert
for forty
years.
Your clothing
has not
worn out
nor
have your sandals
deteriorated.
\VS{6}You have eaten
no
bread
and drunk
no
wine
or beer
– all so that you might
know
that
I am
the {\ND{Lord}}
your God!
\VS{7}When you came
to
this
place
King
Sihon
of Heshbon
and King
Og
of Bashan
came out
to make war
and we defeated them.
\VS{8}Then we took
their land
and gave
it as an inheritance
to Reuben,
Gad,
and half
the tribe
of Manasseh.
\par }{\SH The Present Covenant Setting
\par }{\PP \VS{9}“Therefore, keep
the terms
of this
covenant
and obey
them so that you may
be successful
in everything
you do.
\VS{10}You
are standing
today,
all
of you, before
the {\ND{Lord}}
your God
– the heads
of your tribes,
your elders,
your officials,
every
Israelite
man,
\VS{11}your infants,
your wives,
and the foreigners
living in your
encampment,
those who chop wood
and those who carry water –
\VS{12}so that you may enter
by oath
into the covenant
the {\ND{Lord}}
your God
is making
with
you today.
\VS{13}Today
he will affirm that you
are his people
and that
he is your God,
just
as he promised
you and as
he swore
by oath to your ancestors
Abraham,
Isaac,
and Jacob.
\VS{14}It is not
with
you alone
that I am
making
this
covenant
by oath,
\VS{15}but
with
whoever
stands
with us
here
today
before
the {\ND{Lord}}
our God
as well
as those
not
with
us here
today.
\par }{\SH The Results of Disobedience
\par }{\PP \VS{16}“(For
you
know
how we lived
in the land
of Egypt
and how
we crossed
through
the nations
as
we traveled.
\VS{17}You have seen
their detestable things
and idols
of wood,
stone,
silver,
and gold.)
\VS{18}Beware that
the heart
of no man,
woman,
clan,
or
tribe
among you turns away
from the
{\ND{Lord}}
our God
today
to
pursue
and serve
the gods
of those
nations;
beware
that there is
among you no root
producing
poisonous
and bitter fruit.
\VS{19}When such a person hears
the words
of this
oath
he secretly blesses
himself
and says,
“I will
have peace
though I continue to walk
with a stubborn
spirit.”
This
will destroy
the watered
ground with the parched.
\VS{20}The
{\ND{Lord}}
will
be unwilling
to forgive
him, and his
intense
anger
will rage
against that
man;
all
the curses
written
in this
scroll
will fall
upon him and the
{\ND{Lord}}
will obliterate
his name
from
memory.
\VS{21}The
{\ND{Lord}}
will single
him out for judgment
from all
the tribes
of Israel
according to all
the curses
of the covenant
written
in this
scroll
of the law.
\VS{22}The generation to come – your descendants who will rise up after you, as well as the foreigner who will come from distant places – will see the afflictions of that land and the illnesses that the
{\ND{Lord}} has brought on it.
\VS{23}The whole
land
will be covered with brimstone,
salt,
and burning
debris; it will not
be planted
nor
will it sprout
or
produce grass.
It will resemble the destruction
of Sodom
and Gomorrah,
Admah
and Zeboiim,
which
the {\ND{Lord}}
destroyed
in his intense
anger.
\VS{24}Then all
the nations
will ask, “Why
has the
{\ND{Lord}}
done
all this
to this
land? What
is this fierce,
heated
display of anger
all about?”
\VS{25}Then people will say,
“Because they abandoned
the
covenant
of the {\ND{Lord}}, the God
of their ancestors,
which
he made
with
them when he brought them out
of the
land
of Egypt.
\VS{26}They went
and served
other
gods
and worshiped
them, gods
they did not
know
and that he did not
permit them to worship.
\VS{27}That is
why the
{\ND{Lord}}’s
anger
erupted
against this land,
bringing
on
it all
the curses
written
in this
scroll.
\VS{28}So the
{\ND{Lord}}
has uprooted
them
from their land
in anger,
wrath,
and great
rage
and has deported them
to
another
land,
as is clear
today.”
\VS{29}Secret
things belong to the
{\ND{Lord}}
our God,
but those that are revealed
belong to us and our descendants
forever,
so
that we
might obey
all
the words
of this
law.

\par }\Chap{30}{\PP \VerseOne{1}“When you have experienced
all
these
things,
both the blessings
and the curses
I have
set
before
you, you will reflect
upon them
in all
the nations
where
the {\ND{Lord}}
your God
has banished you.
\VS{2}Then if you and your descendants
turn
to
the {\ND{Lord}}
your God
and obey
him
with your whole
mind
and being
just
as I am
commanding
you today,
\VS{3}the

{\ND{Lord}}
your God
will reverse
your captivity
and have pity
on you. He will turn
and gather
you from all
the peoples
among
whom
he
has scattered you.
\VS{4}Even if
your exiles
are in the most distant
land,
from there
the {\ND{Lord}}
your God
will gather
you and bring
you back.
\VS{5}Then he will bring
you to
the land
your ancestors
possessed
and you also
will possess
it; he will do better for you and multiply
you more than he did your ancestors.
\VS{6}The
{\ND{Lord}}
your God
will also cleanse
your heart
and the hearts
of your descendants
so that you may love
him with all
your mind
and being
and so that you may
live.
\VS{7}Then
the {\ND{Lord}}
your God
will put all
these
curses
on
your enemies,
on
those who hate
you and persecute you.
\VS{8}You
will return
and obey
the {\ND{Lord}}, keeping
all
his commandments
I am
giving
you today.
\VS{9}The
{\ND{Lord}}
your God
will make the labor
of your hands
abundantly successful
and multiply your children,
the offspring
of your cattle,
and the produce
of your soil.
For
the
{\ND{Lord}} your God
will once more
rejoice
over
you to make you prosperous
just
as he rejoiced
over
your ancestors,
\VS{10}if
you obey
the {\ND{Lord}}
your God
and keep
his commandments
and statutes
that are written
in this
scroll
of the law.
But
you must turn
to him
with your whole
mind
and being.
\par }{\SH Exhortation to Covenant Obedience
\par }{\PP \VS{11}“This
commandment
I am
giving
you today
is not
too difficult
for you, nor
is it
too
remote.
\VS{12}It is not
in heaven,
as though
one must say,
“Who
will go up
to heaven
to get
it for us and proclaim
it to us so we may obey it?”
\VS{13}And it is not
across
the sea,
as though
one must say,
“Who
will cross
over to
the other side
of the sea
and get
it for us and proclaim
it to us so we may obey it?”
\VS{14}For
the thing
is very
near you – it is in your mouth and in your mind so that you can do it.
\par }{\PP \VS{15}“Look! I have set
before
you today
life
and prosperity
on the one hand, and death
and disaster on the other.
\VS{16}What
I
am commanding
you today
is to love
the {\ND{Lord}}
your God,
to walk
in his ways,
and to obey
his commandments,
his statutes,
and his ordinances.
Then you will live
and become numerous
and the
{\ND{Lord}}
your God
will bless
you in the land
which
you
are about to possess.
\VS{17}However, if
you turn
aside
and do not
obey,
but are lured
away to worship
and serve
other
gods,
\VS{18}I declare
to you this very day
that
you will certainly
perish! You will not
extend
your time
in
the land
you
are crossing
the Jordan
to possess.
\VS{19}Today
I invoke
heaven
and earth
as a witness against you that I
have set
life
and death,
blessing
and curse,
before
you. Therefore choose
life
so
that you
and your descendants
may live!
\VS{20}I also call on you to love
the

{\ND{Lord}}
your God,
to obey
him and be loyal
to him, for
he
gives you life
and enables you to live
continually
in
the
land
the {\ND{Lord}}
promised
to give
to your ancestors
Abraham,
Isaac,
and Jacob.”

\par }\Chap{31}{\PP \VerseOne{1}Then Moses
went
and spoke
these
words
to
all
Israel.
\VS{2}He said
to
them, “Today
I am a hundred
and twenty
years
old. I
am no
longer
able
to get
about,
and the
{\ND{Lord}}
has said
to
me, ‘You will not
cross
the Jordan.’
\VS{3}As for the
{\ND{Lord}}
your God,
he is
about to cross
over before
you; he
will destroy
these
nations
before
you and dispossess
them. As for Joshua,
he is
about to cross
before
you just
as the
{\ND{Lord}}
has said.
\VS{4}The
{\ND{Lord}}
will do
to them just
what he did
to Sihon
and Og,
the Amorite
kings,
and to their land,
which
he destroyed.
\VS{5}The
{\ND{Lord}}
will deliver them over
to you and you will
do
to them according to the whole
commandment
I have
given you.
\VS{6}Be strong
and courageous! Do not
fear
or
tremble
before
them, for
the {\ND{Lord}}
your God
is the one who is going
with
you. He will not
fail
you or
abandon you!”
\VS{7}Then Moses
called out
to Joshua
in the presence
of all
Israel,
“Be strong
and courageous,
for
you
will accompany
these
people
to
the land
that
the {\ND{Lord}}
promised
to give
their ancestors,
and you
will enable them to inherit it.
\VS{8}The
{\ND{Lord}}
is indeed going
before you – he will be with you; he will not fail you or abandon you. Do not be afraid or discouraged!”
\par }{\SH The Deposit of the Covenant Text
\par }{\PP \VS{9}Then Moses
wrote
down this
law
and gave
it to
the Levitical
priests,
who carry
the ark
of the
{\ND{Lord}}’s
covenant,
and to
all
Israel’s
elders.
\VS{10}He
commanded
them: “At the end
of seven
years,
at the appointed
time
of the cancellation
of debts,
at the Feast
of Temporary Shelters,
\VS{11}when all
Israel
comes
to appear
before
the {\ND{Lord}}
your God
in the place
he chooses,
you must read
this
law
before
them
within their hearing.
\VS{12}Gather
the
people
– men,
women,
and children,
as well as
the resident foreigners
in your villages
– so they may
hear
and thus
learn
about and fear
the

{\ND{Lord}}
your God
and carefully
obey
all
the words
of this
law.
\VS{13}Then their children,
who
have not
known
this law, will also hear
about and learn
to fear
the

{\ND{Lord}}
your God
for as long
as you
live
in the land
you
are crossing
the
Jordan
to possess.”
\par }{\SH The Commissioning of Joshua
\par }{\PP \VS{14}Then the
{\ND{Lord}}
said to
Moses,
“The day
of your death
is near.
Summon
Joshua
and present
yourselves in the tent
of meeting
so that I can
commission
him.” So
Moses
and Joshua
presented themselves
in the tent
of meeting.
\VS{15}The
{\ND{Lord}}
appeared
in the tent
in a pillar
of cloud
that stood
above
the door
of the tent.
\VS{16}Then the
{\ND{Lord}}
said
to
Moses,
“You are about to die,
and then these
people
will begin
to prostitute
themselves with the foreign
gods
of the land
into which
they are going.
They will reject
me and break
my covenant
that
I have made
with them.
\VS{17}At that time
my anger
will erupt
against them and I will abandon
them and hide
my face
from them until they
are devoured.
Many
disasters
and distresses
will overcome them so that they will say
at that time,
‘Have not
these
disasters
overcome
us because
our God
is not
among us ?’
\VS{18}But I
will certainly
hide
myself at that time
because
of all
the wickedness
they will have
done
by turning
to
other
gods.
\VS{19}Now
write
down for yourselves the following
song
and teach
it to the Israelites.
Put
it into their very mouths
so
that this
song
may serve as my witness
against the Israelites!
\VS{20}For
after I have brought
them to
the land
I promised
to their ancestors
– one flowing
with milk
and honey
– and they eat
their fill
and become fat,
then they will turn
to
other
gods
and worship
them; they will reject
me and break
my covenant.
\VS{21}Then
when
many
disasters
and distresses
overcome them this
song
will testify
against them, for
their descendants
will not
forget
it. I know
the
intentions
they have
in mind today,
even before
I bring
them to
the land
I have
promised.”
\VS{22}So on that day
Moses
wrote
down this
song
and taught
it to the Israelites,
\VS{23}and the
{\ND{Lord}} commissioned
Joshua
son
of Nun,
“Be strong
and courageous,
for
you
will take
the Israelites
to
the land
I have
promised
them, and I
will be
with you.”
\par }{\SH Anticipation of Disobedience
\par }{\PP \VS{24}When
Moses
finished
writing
on
a scroll
the words
of this
law in their entirety,
\VS{25}he
commanded
the Levites
who carried
the ark
of the
{\ND{Lord}}’s
covenant,
\VS{26}“Take
this
scroll
of the law
and place
it beside
the ark
of the covenant
of the {\ND{Lord}}
your God.
It will
remain there
as a witness against you,
\VS{27}for
I
know
about your rebellion
and stubbornness.
Indeed,
even while
I have been
living
among you
to this very day,
you have rebelled
against
the {\ND{Lord}}; you will be
even
more rebellious after
my death!
\VS{28}Gather
to
me all
your tribal
elders
and officials
so I can speak
to them directly
about these
things
and call the heavens
and the earth
to witness against them.
\VS{29}For
I know
that after
I die
you will totally corrupt
yourselves
and turn
away from
the path
I have
commanded
you to walk.
Disaster
will confront you in the
days
to come
because
you will act
wickedly
before
the {\ND{Lord}}, inciting
him to anger
because of your actions.”
\VS{30}Then
Moses
recited
the words
of this
song
from
start to finish in the hearing
of the whole
assembly
of Israel.

\par }\Chap{32}{\PP \par }{\Q \VerseOne{1}Listen,
O heavens,
and I will speak;
\par }{\Q hear,
O earth,
the words
of my mouth.
\par }{\Q \VS{2}My teaching
will drop
like the rain,
\par }{\Q my sayings
will drip
like the dew,
\par }{\Q as rain drops
upon
the grass,
\par }{\Q and showers
upon
new growth.
\par }{\Q \VS{3}For
I will proclaim
the name
of the {\ND{Lord}};
\par }{\Q you must acknowledge
the greatness
of our God.
\par }{\Q \VS{4}As for the Rock,
his work
is perfect,
\par }{\Q for
all
his ways
are just.
\par }{\Q He is a reliable
God
who is never
unjust,
\par }{\Q he is fair
and upright.
\par }{\Q \VS{5}His people have been unfaithful
to him;
\par }{\Q they have
not
acted like his children –
this is their sin.
\par }{\Q They are a perverse
and deceitful
generation.
\par }{\Q \VS{6}Is this how you repay
the {\ND{Lord}},
\par }{\Q you foolish,
unwise
people?
\par }{\Q Is he
not
your father,
your creator?
\par }{\Q He
has made
you and established you.
\par }{\Q \VS{7}Remember
the ancient
days;
\par }{\Q bear in mind
the years
of past
generations.
\par }{\Q Ask
your father
and he will inform
you,
\par }{\Q your elders,
and they will tell you.
\par }{\Q \VS{8}When the Most
High gave the nations
their inheritance,
\par }{\Q when he divided
up
humankind,
\par }{\Q he set
the boundaries
of the peoples,
\par }{\Q according to the number of the heavenly assembly.
\par }{\Q \VS{9}For
the
{\ND{Lord}}’s
allotment
is his people,
\par }{\Q Jacob
is his special
possession.
\par }{\Q \VS{10}The
{\ND{Lord}} found
him in a desolate
land,
\par }{\Q in an empty wasteland
where animals howl.
\par }{\Q He continually guarded him and taught him;
\par }{\Q he continually protected him
like
the pupil
of his eye.
\par }{\Q \VS{11}Like an eagle
that stirs
up its nest,
\par }{\Q that hovers over
its young,
\par }{\Q so the
{\ND{Lord}}
spread
out his wings
and took
him,

\par }{\Q he lifted
him up
on
his pinions.
\par }{\Q \VS{12}The
{\ND{Lord}}
alone
was guiding
him,

\par }{\Q no
foreign
god
was with him.
\par }{\Q \VS{13}He enabled
him to travel over
the high terrain
of the land,
\par }{\Q and he ate
of the produce
of the fields.
\par }{\Q He provided
honey
for him from the cliffs,
\par }{\Q and olive oil
from the hardest
of rocks,
\par }{\Q \VS{14}butter
from the herd
\par }{\Q and milk
from the flock,
\par }{\Q along with
the fat
of lambs,
\par }{\Q rams
and goats
of Bashan,
\par }{\Q along with
the best
of the kernels
of wheat;
\par }{\Q and from the juice
of grapes
you drank
wine.
\par }{\SH Israel’s Rebellion
\par }{\Q \VS{15}But Jeshurun
became fat
and kicked,
\par }{\Q you got fat,
thick,
and stuffed!
\par }{\Q Then he deserted
the God
who made
him,
\par }{\Q and treated
the Rock
who saved
him with contempt.
\par }{\Q \VS{16}They made him jealous
with other gods,
\par }{\Q they enraged
him with abhorrent idols.
\par }{\Q \VS{17}They sacrificed
to demons,
not
God,
\par }{\Q to gods
they had not
known;
\par }{\Q to new
gods who had recently
come along,
\par }{\Q gods your ancestors
had not
known about.
\par }{\Q \VS{18}You have forgotten
the Rock
who fathered
you,
\par }{\Q and put out of mind the God
who gave you birth.
\par }{\SH A Word of Judgment
\par }{\Q \VS{19}But
the {\ND{Lord}}
took
note
and despised
them
\par }{\Q because his sons
and daughters
enraged him.
\par }{\Q \VS{20}He said,
“I will reject
them,
\par }{\Q I will see
what
will happen
to them;
\par }{\Q for
they are
a perverse
generation,
\par }{\Q children
who show no
loyalty.
\par }{\Q \VS{21}They
have made me jealous
with false gods,
\par }{\Q enraging
me with their worthless
gods;

\par }{\Q so I
will make them jealous
with a people
they do not
recognize,

\par }{\Q with a nation
slow to learn
I will enrage them.
\par }{\Q \VS{22}For
a fire
has been kindled
by my anger,
\par }{\Q and it burns
to
lowest
Sheol;
\par }{\Q it consumes
the earth
and its produce,
\par }{\Q and ignites
the foundations
of the mountains.
\par }{\Q \VS{23}I will increase
their disasters,
\par }{\Q I will use up my arrows
on
them.
\par }{\Q \VS{24}They will be starved
by famine,
\par }{\Q eaten
by plague,
and bitterly
stung;
\par }{\Q I will send
the teeth
of wild animals
against them,
\par }{\Q along with
the poison
of creatures that crawl
in the dust.
\par }{\Q \VS{25}The sword
will make people childless
outside,
\par }{\Q and terror
will do so inside;
\par }{\Q they will destroy both
the young man
and the virgin,
\par }{\Q the infant
and the gray-haired
man.
\par }{\SH The Weakness of Other Gods
\par }{\Q \VS{26}“I said,
‘I want
to cut them
in pieces.

\par }{\Q I want
to make
people
forget
they ever
existed.
\par }{\Q \VS{27}But I fear
the reaction
of their enemies,
\par }{\Q for their adversaries
would misunderstand
\par }{\Q and say,
“Our power
is great,

\par }{\Q and the
{\ND{Lord}}
has not
done
all
this!” ’
\par }{\Q \VS{28}They are a nation
devoid
of wisdom,
\par }{\Q and there is no
understanding among them.
\par }{\Q \VS{29}I wish
that they were wise
and could understand
this,
\par }{\Q and that they could comprehend what
will happen to them.”
\par }{\Q \VS{30}How
can one
man chase
a thousand
of them,

\par }{\Q and two
pursue
ten thousand;
\par }{\Q unless
their Rock
had delivered them
up,

\par }{\Q and the
{\ND{Lord}}
had handed them over?
\par }{\Q \VS{31}For
our enemies’ rock
is not
like our Rock,
\par }{\Q as even our enemies
concede.
\par }{\Q \VS{32}For
their vine
is from the stock
of Sodom,
\par }{\Q and from the fields of Gomorrah.
\par }{\Q Their grapes
contain venom,
\par }{\Q their clusters
of grapes
are bitter.
\par }{\Q \VS{33}Their wine
is snakes’
poison,
\par }{\Q the deadly
venom
of cobras.
\par }{\Q \VS{34}“Is this not
stored
up with me?” says the
{\ND{Lord}},

\par }{\Q “Is it not sealed
up in my storehouses?
\par }{\Q \VS{35}I will get revenge
and pay them back
\par }{\Q at the time
their foot
slips;
\par }{\Q for
the day
of their disaster
is near,
\par }{\Q and the impending
judgment is rushing upon them!”
\par }{\Q \VS{36}The
{\ND{Lord}}
will judge
his people,
\par }{\Q and will change
his plans concerning
his servants;
\par }{\Q when
he sees
that
their power
has disappeared,
\par }{\Q and that no one
is left, whether confined
or set free.
\par }{\Q \VS{37}He will say,
“Where
are their gods,
\par }{\Q the rock
in whom they sought security,
\par }{\Q \VS{38}who
ate
the best
of their sacrifices,
\par }{\Q and drank
the wine
of their drink offerings?
\par }{\Q Let them rise
and help
you;
\par }{\Q let them be
your refuge!
\par }{\SH The Vindication of the
{\ND{Lord}}
\par }{\Q \VS{39}“See
now
that
I,
indeed
I, am
he!” says the
{\ND{Lord}},

\par }{\Q “and there is no
other god
besides
me.
\par }{\Q I
kill
and give life,
\par }{\Q I smash
and I
heal,
\par }{\Q and none
can resist
my power.
\par }{\Q \VS{40}For
I raise
up my hand
to
heaven,
\par }{\Q and say,
‘As surely
as I
live forever,
\par }{\Q \VS{41}I will sharpen
my lightning-like
sword,
\par }{\Q and my hand
will grasp hold
of the weapon of judgment;
\par }{\Q I will execute
vengeance
on my foes,
\par }{\Q and repay
those who hate me!
\par }{\Q \VS{42}I will make my arrows
drunk
with blood,
\par }{\Q and my sword
will devour
flesh
–
\par }{\Q the blood
of the slaughtered
and captured,
\par }{\Q the chief
of the enemy’s
leaders!’ ”
\par }{\Q \VS{43}Cry
out, O nations,
with his people,
\par }{\Q for
he will avenge
his servants’
blood;
\par }{\Q he will take
vengeance
against his enemies,
\par }{\Q and make atonement
for his land
and people.
\par }{\SH Narrative Interlude
\par }{\PP \VS{44}Then Moses
went
with Joshua
son
of Nun
and recited
all
the words
of this
song
to the people.
\VS{45}When Moses
finished
reciting
all
these
words
to all
Israel
\VS{46}he said
to
them, “Keep
in mind
all
the words
I am
solemnly
proclaiming to you today;
you must
command
your children
to observe
carefully
all
the words
of this
law.
\VS{47}For
this is no
idle
word
for
you – it is your life! By this word you will live a long time in the land you are about to cross the Jordan to possess.”
\par }{\SH Instructions about Moses’ Death
\par }{\PP \VS{48}Then
the {\ND{Lord}}
said
to
Moses
that same
day,
\VS{49}“Go up
to
this
Abarim
hill country,
to Mount
Nebo
(which
is in the land
of Moab
opposite
Jericho
) and look
at the
land
of Canaan
that
I am
giving
to the Israelites
as a possession.
\VS{50}You will die
on the mountain
that
you
ascend
and join
your deceased ancestors,
just
as Aaron
your brother
died
on Mount
Hor
and joined
his deceased ancestors,
\VS{51}for both
of you rebelled
against me among
the Israelites
at the waters
of Meribah
Kadesh
in the desert
of Zin
when you did not
show me proper respect
among
the Israelites.
\VS{52}You will see
the land
before
you, but you will not
enter
the land
that
I am
giving
to the Israelites.”

\par }\Chap{33}{\PP \VerseOne{1}This
is the blessing
Moses
the man
of God
pronounced
upon the Israelites
before
his death.
\VS{2}He said:
\par }{\SH A Historical Review
\par }{\Q The
{\ND{Lord}} came from Sinai
\par }{\Q and revealed himself to Israel from Seir.
\par }{\Q He appeared in splendor from Mount Paran,
\par }{\Q and came forth with ten thousand holy ones.
\par }{\Q With his right hand he gave a fiery law to them.
\par }{\Q \VS{3}Surely
he loves
the people;
\par }{\Q all
your holy ones
are in your power.
\par }{\Q And they
sit
at your feet,
\par }{\Q each receiving
your words.
\par }{\Q \VS{4}Moses
delivered
to us a law,
\par }{\Q an inheritance
for the assembly
of Jacob.
\par }{\Q \VS{5}The
{\ND{Lord}} was
king
over Jeshurun,
\par }{\Q when the leaders
of the people
assembled,
\par }{\Q the tribes
of Israel
together.
\par }{\SH Blessing on Reuben
\par }{\Q \VS{6}May Reuben
live
and not
die,
\par }{\Q and may
his people multiply.
\par }{\SH Blessing on Judah
\par }{\Q \VS{7}And this
is the blessing to Judah.
He said,
\par }{\Q Listen,
O
{\ND{Lord}}, to Judah’s
voice,
\par }{\Q and bring
him to
his people.
\par }{\Q May
his power
be
great,
\par }{\Q and may you help
him against his foes.
\par }{\SH Blessing on Levi
\par }{\Q \VS{8}Of Levi
he said:
\par }{\Q Your Thummim
and Urim
belong to your godly
one,

\par }{\Q whose
authority you challenged
at Massah,
\par }{\Q and with
whom you argued
at the waters
of Meribah.
\par }{\Q \VS{9}He said
to his father
and mother,
“I have not
seen
him,”
\par }{\Q and he did not
acknowledge
his own brothers
\par }{\Q or
know
his own children,
\par }{\Q for
they kept
your word,
\par }{\Q and guarded
your covenant.
\par }{\Q \VS{10}They will teach
Jacob
your ordinances
\par }{\Q and Israel
your law;
\par }{\Q they will offer
incense
as a pleasant odor,
\par }{\Q and a whole offering
on
your altar.
\par }{\Q \VS{11}Bless,
O
{\ND{Lord}}, his
goods,
\par }{\Q and be pleased with
his efforts;
\par }{\Q undercut
the legs of any who attack
him,
\par }{\Q and of those who hate
him, so that they cannot stand.
\par }{\SH Blessing on Benjamin
\par }{\Q \VS{12}Of Benjamin
he said:
\par }{\Q The beloved
of the {\ND{Lord}}
will live
safely
by him;
\par }{\Q he protects
him all
the time,
\par }{\Q and the
{\ND{Lord}} places
him on his chest.
\par }{\SH Blessing on Joseph
\par }{\Q \VS{13}Of Joseph
he said:
\par }{\Q May the
{\ND{Lord}}
bless
his land
\par }{\Q with the harvest produced
by the sky,
by the dew,
\par }{\Q and by the depths
crouching
beneath;
\par }{\Q \VS{14}with the harvest
produced
by the daylight
\par }{\Q and by the moonlight;
\par }{\Q \VS{15}with the best
of the ancient
mountains
\par }{\Q and the harvest produced
by the age-old
hills;
\par }{\Q \VS{16}with the harvest
of the earth
and its fullness
\par }{\Q and the pleasure
of him who resided
in the burning bush.
\par }{\Q May blessing rest on
Joseph’s
head,
\par }{\Q and on the top of
the head
of the one set apart
from his brothers.
\par }{\Q \VS{17}May
the firstborn
of his
bull
bring him honor,
\par }{\Q and may his
horns
be those of a wild ox;
\par }{\Q with
them may he gore
all peoples,
\par }{\Q all the far reaches of the earth.
\par }{\Q They
are the ten thousands
of Ephraim,
\par }{\Q and they
are the thousands
of Manasseh.
\par }{\SH Blessing on Zebulun and Issachar
\par }{\Q \VS{18}Of Zebulun
he said:
\par }{\Q Rejoice,
Zebulun,
when
you go outside,
\par }{\Q and Issachar,
when you are in your tents.
\par }{\Q \VS{19}They will summon
peoples
to the mountain,
\par }{\Q there
they will sacrifice
proper
sacrifices;
\par }{\Q for
they will enjoy the abundance
of the seas,
\par }{\Q and the hidden treasures of the shores.
\par }{\SH Blessing on Gad
\par }{\Q \VS{20}Of Gad
he said:
\par }{\Q Blessed
be the one who enlarges
Gad.
\par }{\Q Like a lioness
he will dwell;
\par }{\Q he will tear
at an arm
– indeed,
a scalp.
\par }{\Q \VS{21}He has selected
the best
part for himself,
\par }{\Q for the portion
of the ruler
is set aside
there;
\par }{\Q he came
with the leaders
of the people,
\par }{\Q he obeyed the righteous
laws of the
{\ND{Lord}}
\par }{\Q and his ordinances
with
Israel.
\par }{\SH Blessing on Dan
\par }{\Q \VS{22}Of Dan
he said:
\par }{\Q Dan
is a lion’s
cub;
\par }{\Q he will leap
forth from
Bashan.
\par }{\SH Blessing on Naphtali
\par }{\Q \VS{23}Of Naphtali
he said:
\par }{\Q O Naphtali,
overflowing
with favor,
\par }{\Q and full
of the
{\ND{Lord}}’s
blessing,
\par }{\Q possess
the west
and south.
\par }{\SH Blessing on Asher
\par }{\Q \VS{24}Of Asher
he said:
\par }{\Q Asher
is blessed
with children,
\par }{\Q may he be favored by
his brothers
\par }{\Q and may he dip
his foot
in olive oil.
\par }{\Q \VS{25}The bars of your gates
will be made of iron
and bronze,
\par }{\Q and may you have lifelong
strength.
\par }{\SH General Praise and Blessing
\par }{\Q \VS{26}There is no
one like God,
O Jeshurun,
\par }{\Q who rides
through the sky
to help
you,
\par }{\Q on the clouds
in majesty.
\par }{\Q \VS{27}The everlasting
God
is a refuge,
\par }{\Q and underneath
you are his eternal
arms;
\par }{\Q he has driven out
enemies
before
you,
\par }{\Q and has said,
“Destroy!”
\par }{\Q \VS{28}Israel
lives
in safety,
\par }{\Q the fountain
of Jacob
is quite secure,
\par }{\Q in
a land
of grain
and new wine;
\par }{\Q indeed,
its heavens
rain down
dew.
\par }{\Q \VS{29}You have joy,
Israel! Who
is like
you?
\par }{\Q You are a people
delivered
by the
{\ND{Lord}},
\par }{\Q your protective
shield
\par }{\Q and your exalted
sword.
\par }{\Q May your enemies
cringe
before you;
\par }{\Q may you trample
on their backs.

\par }\Chap{34}{\PP \VerseOne{1}Then
Moses
ascended
from the deserts
of Moab
to
Mount
Nebo,
to the summit
of Pisgah,
which
is opposite
Jericho.
The
{\ND{Lord}}
showed
him
the whole
land
– Gilead
to Dan,
\VS{2}and all
of Naphtali,
the land
of Ephraim
and Manasseh,
all
the land
of Judah
as far
as the distant sea,
\VS{3}the Negev,
and the plain
of the valley
of Jericho,
the city
of the date
palm trees, as far
as Zoar.
\VS{4}Then the
{\ND{Lord}}
said to him,
“This
is the land
I promised
to Abraham,
Isaac,
and Jacob
when I said,
‘I will give it to your descendants.’
I have let
you see
it, but you will not
cross
over there.”
\par }{\PP \VS{5}So Moses,
the servant
of the {\ND{Lord}}, died
there
in the land
of Moab
as the
{\ND{Lord}}
had said.
\VS{6}He buried
him in the land
of Moab
near Beth Peor,
but no
one
knows
his exact
burial
place to this very
day.
\VS{7}Moses
was 120
years
old
when he died,
but his eye
was not
dull
nor
had his vitality
departed.
\VS{8}The Israelites
mourned
for Moses
in the deserts
of Moab
for thirty
days;
then the days
of mourning
for Moses
ended.
\par }{\SH The Epitaph of Moses
\par }{\PP \VS{9}Now Joshua
son
of Nun
was full
of the spirit
of wisdom,
for
Moses
had
placed
his hands
on
him; and the Israelites
listened
to him
and did
just
what the
{\ND{Lord}}
had commanded
Moses.
\VS{10}No
prophet
ever again
arose
in Israel
like Moses,
who
knew
the
{\ND{Lord}}
face
to
face.
\VS{11}He did all
the signs
and wonders
the {\ND{Lord}}
had
sent
him to do
in the land
of Egypt,
to Pharaoh,
all
his servants,
and the whole
land,
\VS{12}and he
displayed
great
power
and awesome
might
in view
of all
Israel.
\par }