\NormalFont\ShortTitle{Ecclesiastes}
{\MT Ecclesiastes

\par }\ChapOne{1}{\SH Title
\par }{\PP \VerseOne{1}The words
of the Teacher,
the son
of David,
king
in Jerusalem:
\par }{\SH Introduction: Utter Futility
\par }{\Q \VS{2}“Futile! Futile!” laments
the Teacher,
\par }{\Q “Absolutely
futile! Everything
is futile!”
\par }{\SH Futility Illustrated from Nature
\par }{\Q \VS{3}What
benefit
do people
get from all
the effort
\par }{\Q which they expend
on
earth?
\par }{\Q \VS{4}A generation
comes
and a generation
goes,

\par }{\Q but
the earth
remains
the same through
the ages.
\par }{\Q \VS{5}The sun
rises
and the sun
sets;
\par }{\Q it hurries away
to
a place
from which it rises again.
\par }{\Q \VS{6}The wind goes
to the south
and circles around
to
the north;
\par }{\Q round
and round
the wind
goes
and on
its rounds it
returns.
\par }{\Q \VS{7}All
the streams
flow
into the sea,
but the sea
is not
full,
\par }{\Q and to
the place
where the streams
flow,
there
they
will flow
again.
\par }{\Q \VS{8}All
this monotony
is tiresome;
no
one
can bear
to describe
it:

\par }{\Q The eye is never
satisfied
with seeing,
nor
is the ear ever
content
with hearing.
\par }{\Q \VS{9}What
exists
now is what will be,
\par }{\Q and what
has been done
is what will be done;
\par }{\Q there is nothing
truly new
on earth.
\par }{\Q \VS{10}Is there
anything about
which
someone can say,
“Look
at this! It is new!”?

\par }{\Q It was already
done long ago,
before
our time.
\par }{\Q \VS{11}No
one remembers
the former
events,

\par }{\Q nor
will anyone remember the events that are yet to happen;
\par }{\Q they will not
be
remembered
by the future generations.
\par }{\SH Futility of Secular Accomplishment
\par }{\Q \VS{12}I,
the Teacher,
have been
king
over
Israel
in Jerusalem.
\par }{\Q \VS{13}I decided
to carefully
and thoroughly examine
\par }{\Q all
that
has been accomplished
on earth.
\par }{\Q I concluded: God
has given
people
a burdensome
task
\par }{\Q that keeps them occupied.
\par }{\Q \VS{14}I reflected
on everything
that is accomplished
by man on earth,
\par }{\Q and I concluded: Everything
he has accomplished is futile –
like chasing
the wind!
\par }{\Q \VS{15}What is bent
cannot
be straightened,
\par }{\Q and what is missing
cannot
be supplied.
\par }{\SH Futility of Secular Wisdom
\par }{\Q \VS{16}I thought
to myself,
\par }{\Q “I
have
become much
wiser
than any
of my predecessors
who ruled
over
Jerusalem;
\par }{\Q I
have acquired
much
wisdom
and knowledge.”
\par }{\Q \VS{17}So I decided
to discern
the benefit of wisdom
and knowledge
over foolish
behavior and ideas;
\par }{\Q however, I concluded
that even
this
endeavor is like trying to chase
the wind!
\par }{\Q \VS{18}For
with great
wisdom
comes great
frustration;
\par }{\Q whoever increases
his knowledge
merely increases
his heartache.
\par }{\Q \par }
\Chap{2}{\PP \VerseOne{1}\par }{\SH Futility of Self-Indulgent Pleasure
I
thought to
myself,
\par }{\PP \par }{\Q “Come
now, I will try
self-indulgent
pleasure to
see
if it is
worthwhile.”

\par }{\Q But I
found that it
also is futile.
\par }{\Q \VS{2}I said
of partying, “It is folly,”
\par }{\Q and of self-indulgent pleasure, “It accomplishes nothing!”
\par }{\Q \VS{3}I thought
deeply
about the effects
of indulging
myself with wine
\par }{\Q (all the while my
mind
was guiding
me with wisdom)

\par }{\Q and the effects of behaving
foolishly,
\par }{\Q so that
I might discover
what
is profitable
\par }{\Q for people
to do
on earth
during the few
days
of their lives.
\par }{\SH Futility of Materialism
\par }{\Q \VS{4}I increased
my possessions:

\par }{\Q I built
houses
for myself;

\par }{\Q I planted
vineyards for myself.
\par }{\Q \VS{5}I designed
royal gardens
and parks
for myself,
\par }{\Q and I planted
all
kinds of fruit
trees in them.
\par }{\Q \VS{6}I constructed
pools
of water
for myself,
\par }{\Q to irrigate
my grove
of flourishing
trees.
\par }{\Q \VS{7}I purchased
male and female slaves,
\par }{\Q and I owned slaves
who were born in my house;
\par }{\Q I also
possessed more livestock
– both herds
and flocks
–
\par }{\Q than
any
of my predecessors
in Jerusalem.
\par }{\Q \VS{8}I also
amassed
silver
and gold
for myself,
\par }{\Q as
well as valuable treasures
taken from kingdoms
and provinces.
\par }{\Q I acquired male singers
and female singers
for myself,
\par }{\Q and what gives a man
sensual delight –
a harem
of beautiful concubines!
\par }{\Q \VS{9}So I was far wealthier
than
all
my predecessors
in Jerusalem,
\par }{\Q yet
I maintained my objectivity:
\par }{\Q \VS{10}I did not
restrain
myself from getting whatever
I wanted;
\par }{\Q I did not
deny
myself
anything
that would bring me pleasure.
\par }{\Q So all
my accomplishments
gave
me joy;
\par }{\Q this
was
my reward
for all
my effort.
\par }{\Q \VS{11}Yet when I
reflected
on everything
I had accomplished
\par }{\Q and on all the effort
that I had expended
to accomplish
it,

\par }{\Q I concluded: “All these
achievements and possessions are ultimately profitless –
\par }{\Q like chasing
the wind!
\par }{\Q There is nothing
gained
from
them on earth.”
\par }{\SH Wisdom is Better than Folly
\par }{\Q \VS{12}Next, I
decided
to consider
wisdom,
as well as foolish behavior
and ideas.

\par }{\Q For
what
more can the king’s successor
do
than what the king
has already
done?
\par }{\Q \VS{13}I
realized
that wisdom
is
preferable
to folly,
\par }{\Q just
as light
is preferable
to darkness:
\par }{\Q \VS{14}The wise
man can see
where he is going, but the fool
walks
in darkness.
\par }{\Q Yet I
also
realized
that the same
fate
happens to them both.
\par }{\Q \VS{15}So I thought
to myself,
“The fate
of the fool
will happen
even
to me!

\par }{\Q Then what
did I
gain by becoming
so excessively wise?”

\par }{\Q So I
lamented to myself,

\par }{\Q “The benefits
of wisdom
are ultimately meaningless!”
\par }{\Q \VS{16}For
the wise
man, like the fool,
will not
be
remembered
for very long,
\par }{\Q because in the days
to come,
both will already
have been forgotten.
\par }{\Q Alas,
the wise
man dies
– just like the fool!
\par }{\Q \VS{17}So I loathed
life
because
what
\par }{\Q happens
on
earth
seems awful
to me;
\par }{\Q for
all
the benefits of wisdom are futile
– like chasing
the wind.
\par }{\SH Futility of Being a Workaholic
\par }{\Q \VS{18}So I
loathed
all
the fruit of my effort,
\par }{\Q for which I
worked
so hard on earth,
\par }{\Q because I must leave
it behind
in the hands of my successor.
\par }{\Q \VS{19}Who
knows
if he will be a wise man
or
a fool?
\par }{\Q Yet he will be master over
all
the fruit of my labor
\par }{\Q for which I worked
so wisely
on
earth!

\par }{\Q This
also
is futile!
\par }{\Q \VS{20}So
I
began to despair
about all
the fruit
of my labor
\par }{\Q for which I worked
so hard
on earth.
\par }{\Q \VS{21}For
a man
may do
his work
with wisdom,
knowledge,
and skill;
\par }{\Q however, he must hand over
the fruit of his labor
as an inheritance

\par }{\Q to someone
else who did not
work for it.
\par }{\Q This
also
is futile,
and an awful
injustice!
\par }{\SH Painful Days and Restless Nights
\par }{\Q \VS{22}What
does a man
acquire
from all
his labor
\par }{\Q and from the anxiety
that accompanies
his
toil
on earth?
\par }{\Q \VS{23}For
all
day
long his work produces pain
and frustration,
\par }{\Q and even
at night
his mind
cannot relax!

\par }{\Q This
also
is futile!
\par }{\SH Enjoy Work and its Benefits
\par }{\Q \VS{24}There is nothing
better
for people
than
to eat
and drink,
\par }{\Q and to find enjoyment
in their work.
\par }{\Q I also
perceived
that
this ability
to find enjoyment comes from
God.
\par }{\Q \VS{25}For
no one can
eat
and drink
\par }{\Q or experience
joy apart
from him.
\par }{\Q \VS{26}For
to the one
who pleases
him, God gives
wisdom,
knowledge,
and joy,
\par }{\Q but to the sinner,
he gives
the task
of amassing wealth – ‘span class=”s 05414” title=”05414” ’
\par }{\Q only to give
it to the one who pleases
God.
\par }{\PP \par }{\Q This
task of the wicked is futile
– like chasing
the wind!

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}For everything
there is an appointed time,
\par }{\Q and an appropriate time
for every
activity
on
earth:
\par }{\Q \VS{2}A time
to be born,
and a time
to die;
\par }{\Q a time
to plant,
and a time
to uproot
what was planted;
\par }{\Q \VS{3}A time
to kill,
and a time
to heal;
\par }{\Q a time
to break down,
and a time
to build up;
\par }{\Q \VS{4}A time
to weep,
and a time
to laugh;
\par }{\Q a time
to mourn,
and a time
to dance.
\par }{\Q \VS{5}A time
to throw
away stones,
and a time
to gather
stones;
\par }{\Q a time
to embrace,
and a time
to refrain from embracing;
\par }{\Q \VS{6}A time
to search,
and a time
to give something up as lost;
\par }{\Q a time
to keep,
and a time
to throw away;
\par }{\Q \VS{7}A time
to rip,
and a time
to sew;
\par }{\Q a time
to keep silent,
and a time
to speak.
\par }{\Q \VS{8}A time
to love,
and a time
to hate;
\par }{\Q a time
for war,
and a time
for peace.
\par }{\SH Man is Ignorant of God’s Timing
\par }{\Q \VS{9}What
benefit
can a worker
gain from his toil?
\par }{\Q \VS{10}I have observed
the burden
\par }{\Q that
God
has given
to people
to keep them occupied.
\par }{\Q \VS{11}God has made
everything
fit beautifully
in its appropriate time,
\par }{\Q but he has also
placed
ignorance in the human heart
\par }{\Q so that
people
cannot
discover
what
God
has ordained,
\par }{\Q from the beginning
to
the end of their lives.
\par }{\SH Enjoy Life in the Present
\par }{\Q \VS{12}I have concluded
that
there is nothing
better
for
people

\par }{\Q than
to be happy
and to enjoy
\par }{\Q themselves
as long as they live,
\par }{\Q \VS{13}and also
that everyone
should eat
and drink,
and find
enjoyment
in all
his toil,
\par }{\Q for these things are a gift
from God.
\par }{\SH God’s Sovereignty
\par }{\Q \VS{14}I also know
that
whatever
God
does
will endure forever;
\par }{\Q nothing
can be added
to it,
and nothing
taken
away from it.
\par }{\Q God
has made
it this way, so that men will fear him.
\par }{\Q \VS{15}Whatever
exists now has already
been, and whatever
will be
has already
been;
\par }{\Q for God
will seek
to do again what has occurred in the past.
\par }{\SH The Problem of Injustice and Oppression
\par }{\Q \VS{16}I saw
something else on earth:

\par }{\Q In the place
of justice,
there
was wickedness,
\par }{\Q and in the place
of fairness,
there
was wickedness.
\par }{\Q \VS{17}I thought
to myself,
“God
will judge
both the righteous
and the
wicked;
\par }{\Q for
there is an appropriate time
for every
activity,
\par }{\Q and there
is a time of judgment for every
deed.
\par }{\Q \VS{18}I
also thought
to myself,
“It is for the sake
of people,
\par }{\Q so God
can clearly
show
them that they
are like animals.
\par }{\Q \VS{19}For
the fate
of humans
and the fate
of animals
are the same:
\par }{\Q As one
dies,
so
dies
the other; both have the same
breath.
\par }{\Q There is no
advantage
for humans
over animals,
\par }{\Q for
both
are fleeting.
\par }{\Q \VS{20}Both
go
to
the same
place,
\par }{\Q both
come from
the dust,
\par }{\Q and to dust
both
return.
\par }{\Q \VS{21}Who
really knows
if the human
spirit
ascends
upward,
\par }{\Q and the animal’s
spirit
descends
into the earth?
\par }{\Q \VS{22}So I perceived
there is nothing
better
than for people
to enjoy
their work,
\par }{\Q because
that
is their reward;
\par }{\Q for
who
can show
them what
the future holds?

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}So I
again
considered
all
the oppression
that
continually occurs
on earth.
\par }{\Q This is what I saw:

\par }{\Q The oppressed
were in tears,
but no
one was comforting
them;
\par }{\Q no
one delivers
them from the power
of their oppressors.
\par }{\Q \VS{2}So I
considered
those who are dead
and gone
\par }{\Q more
fortunate
than
those
who are still
alive.
\par }{\Q \VS{3}But better
than both
is the one who
has not
been
born

\par }{\Q and has not
seen
the evil
things
that
are done
on earth.
\par }{\SH Labor Motivated by Envy
\par }{\Q \VS{4}Then I
considered
all
the skillful
work
that is done:
\par }{\Q Surely
it
is nothing more than competition between
one person
and another.
\par }{\Q This
also
is profitless
– like chasing
the wind.
\par }{\Q \VS{5}The fool
folds
his hands
and does no work,

\par }{\Q so he has nothing to eat
but his own flesh.
\par }{\Q \VS{6}Better
is one handful
with some rest
\par }{\Q than two hands
full
of toil
and chasing
the wind.
\par }{\SH Labor Motivated by Greed
\par }{\Q \VS{7}So I
again
considered
another futile
thing on
earth:
\par }{\Q \VS{8}A man
who is all alone
with no companion,
\par }{\Q he has no
children
nor siblings;
\par }{\Q yet
there is no
end
to all
his toil,
\par }{\Q and he
is never
satisfied
with riches.
\par }{\Q He laments, “For whom
am I
toiling
and depriving
myself
of pleasure?”

\par }{\Q This
also
is futile
and a burdensome
task!
\par }{\SH Labor is Beneficial When Its Rewards Are Shared
\par }{\Q \VS{9}Two
people are better
than one,
\par }{\Q because they can reap
more benefit
from their labor.
\par }{\Q \VS{10}For
if
they fall,
one
will help
his companion
up,
\par }{\Q but pity
the person
who falls down
and has no
one
to help
him up.
\par }{\Q \VS{11}Furthermore,
if
two lie
down together, they can keep each other
warm,
\par }{\Q but how
can one
person keep warm by himself?
\par }{\Q \VS{12}Although
an assailant
may overpower one
person,
\par }{\Q two
can withstand
him.
\par }{\Q Moreover, a three-stranded
cord
is not
quickly
broken.
\par }{\SH Labor Motivated by Prestige-Seeking
\par }{\Q \VS{13}A poor
but wise
youth
is better
than an old
and foolish
king
\par }{\Q who
no
longer
knows
how to receive advice.
\par }{\Q \VS{14}For
he came out
of prison
to become king,
\par }{\Q even
though he had been born
poor
in what would become his kingdom.
\par }{\Q \VS{15}I considered
all
the living
who walk
on
earth,
\par }{\Q as well
as the successor
who
would arise
in his place.
\par }{\Q \VS{16}There is no
end
to all
the people
nor to the past generations,

\par }{\Q yet
future generations will not
rejoice
in him.
\par }{\Q This
also
is profitless
and like chasing
the wind.

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}Be careful
what
you do
when you go
to
the temple
of God;
\par }{\Q draw near
to listen
rather
than to offer a sacrifice
like fools,
\par }{\Q for
they do
not
realize
that they are doing
wrong.
\par }{\Q \VS{2}Do not
be rash
with your mouth
or hasty
in your heart
to bring up
a matter
before
God,
\par }{\Q for
God
is in heaven
and you
are on
earth!
\par }{\Q Therefore,
let your words
be
few.
\par }{\Q \VS{3}Just as dreams
come
when there are many
cares,
\par }{\Q so the rash vow of a fool
occurs
when there are many
words.
\par }{\Q \VS{4}When you make
a vow
to God,
do not
delay
in paying it.
\par }{\Q For
God takes no
pleasure
in fools:
\par }{\Q Pay
what you vow!
\par }{\Q \VS{5}It is better
for you not
to vow
\par }{\Q than to vow
and not
pay it.
\par }{\Q \VS{6}Do not
let
your mouth
cause you to sin,
\par }{\Q and do not
tell
the priest, “It was a mistake!”

\par }{\Q Why
make God
angry
at you

\par }{\Q so that he would destroy
the
work
of your hands?”
\par }{\Q \VS{7}Just as there is futility
in many
dreams,
\par }{\Q so also in many
words.
\par }{\Q Therefore, fear
God!
\par }{\SH Government Corruption
\par }{\Q \VS{8}If
you see
the extortion
of the poor,
\par }{\Q or the perversion
of justice
and fairness
in the government,
\par }{\Q do not
be astonished
by
the matter.
\par }{\Q For
the high
official is watched
by
a higher
official,

\par }{\Q and there are higher
ones over them!
\par }{\Q \VS{9}The produce
of the land
is seized by all
of them,
\par }{\Q even the king
is served
by the fields.
\par }{\SH Covetousness
\par }{\Q \VS{10}The one who loves
money
will never
be satisfied
with money,
\par }{\Q he who
loves
wealth
will never
be satisfied with his income.
\par }{\Q This
also
is futile.
\par }{\Q \VS{11}When someone’s prosperity
increases,
those who consume
it also increase;
\par }{\Q so what
does its owner
gain,
except
that
he gets to see
it with his eyes?
\par }{\Q \VS{12}The sleep
of the laborer
is pleasant
– whether
he eats
little
or much
–
\par }{\Q but the wealth
of the rich
will not
allow
him to sleep.
\par }{\SH Materialism Thwarts Enjoyment of Life
\par }{\Q \VS{13}Here is
a misfortune
on earth
that I have seen:
\par }{\Q Wealth
hoarded
by its owner
to his own misery.
\par }{\Q \VS{14}Then that
wealth
was lost
through bad
luck;
\par }{\Q although he fathered
a son,
he has nothing left to give him.
\par }{\Q \VS{15}Just
as he came forth
from his mother's
womb,
naked
will he return
as he came,
\par }{\Q and he will take nothing
in his hand
that he may carry
away from his toil.
\par }{\Q \VS{16}This
is another
misfortune:

\par }{\Q Just
as
he came,
so
will he go.
\par }{\Q What
did he gain from toiling
for
the wind?
\par }{\Q \VS{17}Surely,
he ate
in darkness
every
day
of his life,
\par }{\Q and he suffered
greatly
with sickness
and anger.
\par }{\SH Enjoy the Fruit of Your Labor
\par }{\Q \VS{18}I
have seen
personally what
is the only beneficial
and
appropriate
course of action for people:

\par }{\Q to eat
and drink,
and find
enjoyment
in all
their hard
work
on earth
\par }{\Q during the few
days
of their life
which
God
has given
them,
\par }{\Q for
this is
their reward.
\par }{\Q \VS{19}To every
man
whom
God
has given
wealth,
and possessions,
\par }{\Q he has also
given him the ability
\par }{\Q to eat
from
them,
to receive
his reward
and to find enjoyment
in his toil;
\par }{\Q these things are the gift
of God.
\par }{\Q \VS{20}For
he does not
think much
about the fleeting
days
of his life
\par }{\Q because
God
keeps him preoccupied
with the joy
he derives from his activity.


\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}Here is
another misfortune
that
I have seen
on
earth,
\par }{\Q and it weighs
heavily on
people:
\par }{\Q \VS{2}God
gives
a man
riches,
property, and wealth
\par }{\Q so that he lacks
nothing
that
his heart
desires,
\par }{\Q yet God
does not
enable
him
to enjoy the fruit of his labor –
\par }{\Q instead, someone else enjoys it!
\par }{\Q This is fruitless and a grave misfortune.
\par }{\Q \VS{3}Even if
a man
fathers
a hundred
children
and lives
many
years
–
\par }{\Q even if he lives
a long,
long time,
but cannot
enjoy
his prosperity
–
\par }{\Q even if
he were
to live forever –
\par }{\Q I would say, “A stillborn child is better off than he is!”
\par }{\Q \VS{4}Though
the stillborn child came
into the world for no reason
and departed
into darkness,
\par }{\Q though its
name
is shrouded
in darkness,
\par }{\Q \VS{5}though
it never
saw
the light of day
nor
knew
anything,

\par }{\Q yet it has more rest than that man –
\par }{\Q \VS{6}if
he should live
a thousand
years
twice,
yet does not
enjoy
his prosperity.
\par }{\Q For both
of them die!
\par }{\Q \VS{7}All
of man’s
labor
is for nothing more than to fill his stomach – ‘span class=”s 01571” title=”01571” ’
\par }{\Q yet
his appetite
is never
satisfied!
\par }{\PP \par }{\Q \VS{8}So what
advantage
does a wise
man have over a fool?

\par }{\Q And what
advantage does a pauper
gain by knowing how
to survive?
\par }{\Q \VS{9}It is better
to be content with what the eyes
can see
\par }{\Q than
for one’s heart
always to crave
more.

\par }{\Q This
continual longing is futile
– like chasing
the wind.
\par }{\SH The Futile Way Life Works
\par }{\Q \VS{10}Whatever
has happened
was foreordained,
\par }{\Q and what happens to a person
was
also
foreknown.
\par }{\Q It is useless for him to argue with
God about his fate
\par }{\Q because God is more
powerful than he is.
\par }{\Q \VS{11}The more one
argues with words,
the less he accomplishes.
\par }{\Q How
does that benefit him?
\par }{\Q \VS{12}For
no one
knows
what
is best
for a person
during his life –
\par }{\Q during the few
days
of his fleeting
life –
\par }{\Q for they pass away like a shadow.
\par }{\Q Nor can anyone tell him what the future will hold for him on earth.

\par }\Chap{7}{\PP \par }{\Q \VerseOne{1}A good
reputation
is better
than precious perfume;
\par }{\Q likewise, the day
of one’s death
is better than the day
of one’s birth.
\par }{\Q \VS{2}It is better
to go
to
a funeral
\par }{\Q than a feast.
\par }{\Q For death is the destiny
of every
person,
\par }{\Q and the living
should take this to
heart.
\par }{\Q \VS{3}Sorrow is better
than laughter,
\par }{\Q because
sober reflection
is good
for
the heart.
\par }{\Q \VS{4}The heart
of the wise
is in the house
of mourning,
\par }{\Q but the heart
of fools
is in the house
of merrymaking.
\par }{\SH Frivolous Living Versus Wisdom
\par }{\Q \VS{5}It is better
for a person
to receive
a rebuke
from those who are wise
\par }{\Q than to listen
to the song
of fools.
\par }{\Q \VS{6}For
like the crackling
of quick-burning thorns
under
a cooking pot,
\par }{\Q so
is the laughter
of the fool.
\par }{\Q This
kind of folly also
is useless.
\par }{\SH Human Wisdom Overturned by Adversity
\par }{\Q \VS{7}Surely oppression
can turn a wise person
into a fool;
\par }{\Q likewise, a bribe
corrupts
the heart.
\par }{\Q \VS{8}The end
of a matter
is better
than its beginning;
\par }{\Q likewise, patience
is better
than pride.
\par }{\Q \VS{9}Do not
let yourself be quickly
provoked,
\par }{\Q for
anger resides
in the lap
of fools.
\par }{\Q \VS{10}Do not
say,
“Why
were the old
days
better
than these
days?”

\par }{\Q for
it is not
wise
to ask that.
\par }{\SH Wisdom Can Lengthen One’s Life
\par }{\Q \VS{11}Wisdom,
like an inheritance,
is a good
thing;
\par }{\Q it benefits
those who see
the light of day.
\par }{\Q \VS{12}For
wisdom
provides protection,
\par }{\Q just as money
provides protection.
\par }{\Q But the advantage
of knowledge
is this:
\par }{\Q Wisdom
preserves
the life
of its owner.
\par }{\SH Wisdom Acknowledges God’s Orchestration of Life
\par }{\Q \VS{13}Consider
the work
of God:
\par }{\Q For
who can
make straight what he has bent?
\par }{\Q \VS{14}In times
of prosperity
be
joyful,
\par }{\Q but in times
of adversity
consider
this:
\par }{\Q God
has made
one as
well as the other,
\par }{\Q so that no
one
can discover
what the future
holds.
\par }{\SH Exceptions to the Law of Retribution
\par }{\Q \VS{15}During the days
of my fleeting life
I have seen
both of these things:
\par }{\Q Sometimes
a righteous person
dies prematurely
in spite of his righteousness,
\par }{\Q and sometimes
a wicked person
lives long
in spite of his evil deeds.
\par }{\Q \VS{16}So do not
be
excessively righteous
or
excessively
wise;
\par }{\Q otherwise
you might be disappointed.
\par }{\Q \VS{17}Do not
be excessively
wicked and do not
be a fool;
\par }{\Q otherwise
you might die
before your time.
\par }{\Q \VS{18}It is best
to take hold
of one warning
without
letting
go
of the other warning;
\par }{\Q for
the one who fears
God
will follow
both warnings.
\par }{\SH Wisdom Needed Because No One is Truly Righteous
\par }{\Q \VS{19}Wisdom
gives
a wise
person more protection
\par }{\Q than ten
rulers
in a city.
\par }{\Q \VS{20}For
there is not
one truly righteous
person
on the earth
\par }{\Q who
continually does
good
and never
sins.
\par }{\Q \VS{21}Also,
do not
pay
attention
to everything
that
people say;
\par }{\Q otherwise, you might even hear
your servant
cursing you.
\par }{\Q \VS{22}For
you know
in your own heart
\par }{\Q that
you also
have
cursed
others
many
times.
\par }{\SH Human Wisdom is Limited
\par }{\Q \VS{23}I have examined
all
this by
wisdom;
\par }{\Q I said,
“I am determined to comprehend this” – but it was beyond my grasp.
\par }{\Q \VS{24}Whatever
has happened
is beyond human understanding;

\par }{\Q it is far
deeper
than anyone
can fathom.
\par }{\SH True Righteousness and Wisdom are Virtually Nonexistent
\par }{\Q \VS{25}I
tried
to understand,
examine,
and comprehend
\par }{\Q the role of wisdom
in the scheme of things,
\par }{\Q and to understand
the stupidity
of wickedness
and the insanity
of folly.
\par }{\Q \VS{26}I
discovered
this:

\par }{\Q More bitter
than death
is the kind of woman
who
is like a hunter’s snare;
\par }{\Q her heart
is like a hunter’s net
and her hands
are like prison chains.
\par }{\Q The man who pleases
God
escapes
her,
\par }{\Q but the sinner
is captured by her.
\par }{\Q \VS{27}The Teacher
says:
\par }{\Q I discovered
this
while trying to discover
the scheme
of things, item by item.
\par }{\Q \VS{28}What
I have continually
sought,
I
have not
found;
\par }{\Q I have found only one upright
man
among a thousand,
\par }{\Q but I have
not
found
one upright
woman
among
all of them.
\par }{\Q \VS{29}This
alone
have
I discovered: God
made
humankind
upright,
\par }{\Q but they
have sought
many
evil schemes.

\par }\Chap{8}{\PP \par }{\Q \VerseOne{1}Who
is a wise person? Who
knows
the solution
to a problem?

\par }{\Q A person’s
wisdom
brightens
his appearance,
and softens
his harsh
countenance.
\par }{\Q \VS{2}Obey
the king’s
command,
\par }{\Q because you took an oath
before God to be loyal to him.
\par }{\Q \VS{3}Do not
rush
out of the king’s presence
in haste
– do not
delay
when the matter
is unpleasant,
\par }{\Q for
he can do
whatever
he pleases.
\par }{\Q \VS{4}Surely
the king’s
authority
is absolute;

\par }{\Q no one
can say
to him, “What
are you doing?”
\par }{\Q \VS{5}Whoever
obeys
his command will not
experience
harm,
\par }{\Q and a wise
person knows
the proper
time
and procedure.
\par }{\Q \VS{6}For
there is a proper time
and procedure
for
every
matter,
\par }{\Q for the oppression
of the king
is severe
upon his victim.
\par }{\Q \VS{7}Surely
no
one knows
the future,
\par }{\Q and no one can
tell
another person what
will happen.
\par }{\Q \VS{8}Just as no
one
has power
over the wind
to restrain
it,
\par }{\Q so no one
has power
over the day
of his death.
\par }{\Q Just as no one
can be discharged
during the battle,
\par }{\Q so wickedness
cannot
rescue
the wicked.
\par }{\Q \VS{9}While applying
my mind
to everything
that
happens in this world,
I have seen
all
this:
\par }{\Q Sometimes
one person
dominates
other people
to their harm.
\par }{\SH Contradictions to the Law of Retribution
\par }{\Q \VS{10}Not only that, but I have seen
the wicked
approaching
and entering
the temple,
\par }{\Q and as they left
the holy temple,
they
\par }{\Q boasted
in the city
that
they had done
so.
\par }{\Q This
also
is an enigma.
\par }{\Q \VS{11}When a sentence
is not
executed at once against a crime,
\par }{\Q the human
heart
is encouraged to do
evil.
\par }{\Q \VS{12}Even though
a sinner
might commit
a hundred
crimes
and still live a long time,
\par }{\Q yet I
know
that
it will go
well
with God-fearing people – for they stand in fear before him.
\par }{\Q \VS{13}But it will not
go
well
with the wicked,
\par }{\Q nor
will they prolong
their days
like a shadow,
\par }{\Q because
they do not
stand in fear
before
God.
\par }{\Q \VS{14}Here is another
enigma
that
occurs
on
earth:
\par }{\Q Sometimes
there are righteous
people who
get
what
the wicked
deserve,

\par }{\Q and sometimes
there are wicked
people who get
what
the righteous
deserve.

\par }{\Q I said,
“This
also
is an enigma.”
\par }{\SH Enjoy Life In Spite of Its Injustices
\par }{\Q \VS{15}So I
recommend
the enjoyment
of life,

\par }{\Q for there is nothing
better
on earth
for
a person
to do except
to eat,
drink,
and enjoy life.
\par }{\Q So joy will accompany
him in his toil
\par }{\Q during the days
of his life
which
God
gives
him on earth.
\par }{\SH Limitations of Human Wisdom
\par }{\Q \VS{16}When I tried
to gain
wisdom
\par }{\Q and to observe
the activity
on
earth
–
\par }{\Q even
though
it prevents
anyone from sleeping
day
or night –
\par }{\Q \VS{17}then I discerned
all
that God
has done:

\par }{\Q No
one
really comprehends
what happens
on earth.
\par }{\Q Despite
all human
efforts
to discover
it, no
one can ever grasp
it.

\par }{\Q Even
if
a wise person
claimed
that he understood,
\par }{\Q he would
not
really comprehend it.

\par }\Chap{9}{\PP \par }{\Q \VerseOne{1}So I reflected
on all
this,
attempting to
clear it all
up.
\par }{\Q I concluded
that
the righteous
and the wise,
as well as their works,
are in the hand
of God;
\par }{\Q whether a person will be loved
or hated –
\par }{\Q no
one
knows
what lies
ahead.
\par }{\Q \VS{2}Everyone
shares the same
fate –
\par }{\Q the righteous
and the wicked,
\par }{\Q the good
and the bad,

\par }{\Q the ceremonially clean
and unclean,
\par }{\Q those who
offer
sacrifices and those who
do
not.
\par }{\Q What
happens to the good
person, also happens to the sinner;
\par }{\Q what happens
to those who
make vows, also happens to those who
are afraid
to make vows.
\par }{\Q \VS{3}This
is the unfortunate fact
about everything
that
happens
on earth:

\par }{\Q the same
fate
awaits everyone.
\par }{\Q In addition to this, the hearts
of all people
are full
of evil,
\par }{\Q and there is folly
in their hearts
during their lives
– then they die.
\par }{\SH Better to Be Poor but Alive than Rich but Dead
\par }{\Q \VS{4}But
whoever
is among the living
has hope;
\par }{\Q a live
dog
is better
than
a dead
lion.
\par }{\Q \VS{5}For
the living
know
that they will die,
but the dead
do not
know
anything;
\par }{\Q they have no
further
reward
– and even the memory
of them disappears.
\par }{\Q \VS{6}What they loved,
as well as
what they hated
and envied,
perished
long ago,
\par }{\Q and they
no
longer
have a part
in anything
that
happens
on earth.
\par }{\SH Life is Brief, so Cherish its Joys
\par }{\Q \VS{7}Go,
eat
your food
with joy,
\par }{\Q and drink
your wine
with a happy
heart,
\par }{\Q because
God
has already
approved
your works.
\par }{\Q \VS{8}Let
your clothes
always
be
white,
\par }{\Q and do not
spare precious
ointment
on
your head.
\par }{\Q \VS{9}Enjoy
life
with
your beloved
wife
during all
the days
of your fleeting
life
\par }{\Q that God has
given
you on
earth
during all
your fleeting
days;
\par }{\Q for
that
is your reward
in life
and in your
burdensome work
on earth.
\par }{\Q \VS{10}Whatever
you find
to do
with your hands,
\par }{\Q do
it with all
your might,
\par }{\Q because
there is neither
work
nor planning
nor knowledge
nor wisdom
in the grave,
\par }{\Q the place where
you
will eventually go.
\par }{\SH Wisdom Cannot Protect against Seemingly Chance Events
\par }{\Q \VS{11}Again,
I observed
this on
the earth:

\par }{\Q the race
is not
always won by the swiftest,
\par }{\Q the battle
is not
always won by the strongest;
\par }{\Q prosperity
does not
always belong to those who are the wisest,
\par }{\Q wealth
does not
always belong to those who are the most discerning,
\par }{\Q nor
does success
always come to those with the most knowledge
–
\par }{\Q for
time
and chance
may overcome
them all.
\par }{\Q \VS{12}Surely,
no
one
knows
his appointed time!

\par }{\Q Like fish
that are caught
in a deadly
net,
and like birds
that are caught
in a snare
–
\par }{\Q just like them,
all people
are ensnared
at an unfortunate
time
that falls
upon
them suddenly.
\par }{\SH Most People Are Not Receptive to Wise Counsel
\par }{\Q \VS{13}This
is what I also
observed
about wisdom
on earth,
\par }{\Q and it is a great
burden to me:
\par }{\Q \VS{14}There was once a small
city
with a few
men
in it,
\par }{\Q and a mighty king
attacked
it, besieging
it and building
strong
siege works
against it.
\par }{\Q \VS{15}However,
a poor
but wise
man
lived in the city,

\par }{\Q and he could have delivered
the city
by his wisdom,
\par }{\Q but no
one
listened
to that poor
man.
\par }{\Q \VS{16}So I
concluded that wisdom
is better
than might,
\par }{\Q but a poor
man’s wisdom
is despised;
no
one ever listens
to his advice.
\par }{\SH Wisdom versus Fools, Sin, and Folly
\par }{\Q \VS{17}The words
of the wise
are heard
in quiet,
\par }{\Q more than the shouting
of a ruler
is heard among fools.
\par }{\Q \VS{18}Wisdom
is better
than weapons
of war,
\par }{\Q but one
sinner
can destroy
much
that is good.

\par }\Chap{10}{\PP \par }{\Q \VerseOne{1}One dead
fly
makes the perfumer’s
ointment
give off
a rancid stench,
\par }{\Q so a little
folly
can outweigh
much wisdom.
\par }{\SH Wisdom Can Be Nullified By the Caprice of Rulers
\par }{\Q \VS{2}A wise
person’s
good sense
protects
him,

\par }{\Q but a
fool’s lack of sense
leaves him vulnerable.
\par }{\Q \VS{3}Even
when a fool
walks
along the road
he lacks
sense,
\par }{\Q and shows everyone
what a fool
he is.
\par }{\Q \VS{4}If
the anger
of the ruler
flares up
against
you, do not
resign
from your position,
\par }{\Q for
a calm
response can undo
great
offenses.
\par }{\Q \VS{5}I have seen
another misfortune
on the earth:

\par }{\Q It is an error
a ruler
makes.
\par }{\Q \VS{6}Fools
are placed
in many
positions of authority,
\par }{\Q while wealthy
men sit
in lowly positions.
\par }{\Q \VS{7}I have seen
slaves
on
horseback
\par }{\Q and princes
walking on
foot like slaves.
\par }{\SH Wisdom is Needed to Avert Dangers in Everyday Life
\par }{\Q \VS{8}One who digs
a pit
may fall
into it,
\par }{\Q and one who breaks through
a wall
may be bitten
by a snake.
\par }{\Q \VS{9}One who quarries
stones
may be injured by
them;
\par }{\Q one who splits
logs
may be endangered by them.
\par }{\Q \VS{10}If
an iron
axhead
is blunt
and a workman does not
sharpen its edge,

\par }{\Q he must
exert
a great deal of effort;

\par }{\Q so wisdom
has the advantage of giving success.
\par }{\Q \VS{11}If
the snake
should bite
before it is charmed,
\par }{\Q the snake charmer is
in trouble.
\par }{\SH Words and Works of Wise Men and Fools
\par }{\Q \VS{12}The words
of a wise
person win him favor,
\par }{\Q but the words
of a fool
are self-destructive.
\par }{\Q \VS{13}At the beginning
his words
are foolish
\par }{\Q and at the end
his talk
is wicked
madness,
\par }{\Q \VS{14}yet a fool
keeps on babbling.
\par }{\Q No
one
knows
what
will happen;
\par }{\Q who can
tell
him what will happen
in the future?
\par }{\Q \VS{15}The toil
of a stupid fool
wears
him out,

\par }{\Q because he does not
even know
the way
to
the city.
\par }{\SH The Problem with Foolish Rulers
\par }{\Q \VS{16}Woe
to you, O land,
when your king
is childish,
\par }{\Q and your princes
feast
in the morning!
\par }{\Q \VS{17}Blessed
are you, O land,
when your king
is the son
of nobility,
\par }{\Q and your princes
feast
at the proper time –
with self-control
and not
in drunkenness.
\par }{\Q \VS{18}Because of laziness
the roof
caves in,
\par }{\Q and because of idle
hands
the house
leaks.
\par }{\Q \VS{19}Feasts
are made
for laughter,
\par }{\Q and wine
makes
life
merry,

\par }{\Q but money
is the answer
for everything.
\par }{\Q \VS{20}Do not
curse
a king
even
in your thoughts,
\par }{\Q and do not
curse
the rich
while in your bedroom;
\par }{\Q for
a bird
might report
what you are thinking,
\par }{\Q or some winged creature
might repeat
your words.

\par }\Chap{11}{\PP \par }{\Q \VerseOne{1}Send
your grain
overseas,
\par }{\Q for
after many
days
you will get a return.
\par }{\Q \VS{2}Divide
your merchandise among
seven
or
even
eight
investments,
\par }{\Q for
you do not
know
what
calamity
may happen
on
earth.
\par }{\Q \VS{3}If
the clouds
are full
of rain,
they will empty
themselves on
the earth,
\par }{\Q and whether
a tree
falls
to the south
or
to the north,
the tree
will lie wherever
it falls.
\par }{\Q \VS{4}He who watches
the wind
will not
sow,
\par }{\Q and he who observes
the clouds
will not
reap.
\par }{\Q \VS{5}Just
as you do not
know
the path
of the wind,
\par }{\Q or how
the bones
form
in the womb
of a pregnant
woman,

\par }{\Q so you do not
know
the work
of God
who
makes
everything.
\par }{\Q \VS{6}Sow
your seed
in the morning,
\par }{\Q and do not
stop working
until the evening;
\par }{\Q for
you do not
know
which activity
will succeed – ‘span class=”s 02088” title=”02088” ’
\par }{\Q whether this
one
or
that one,
or whether
both
will prosper
equally.
\par }{\SH Life Should Be Enjoyed Because Death is Inevitable
\par }{\Q \VS{7}Light
is sweet,
\par }{\Q and it is pleasant
for a person
to see
the sun.
\par }{\Q \VS{8}So, if
a man
lives
many
years,
let him rejoice
in them all,
\par }{\Q but let him remember
that the days
of darkness
will be
many
– all
that is about to come
is obscure.
\par }{\SH Enjoy Life to the Fullest under the Fear of God
\par }{\Q \VS{9}Rejoice,
young man,
while you are young,
\par }{\Q and let your heart
cheer
you in the days
of your youth.
\par }{\Q Follow
the impulses
of your heart
and the desires
of your eyes,
\par }{\Q but know
that
God
will judge your motives and actions.
\par }{\Q \VS{10}Banish emotional
stress
from your mind.
\par }{\Q and put away
pain
from your body;
\par }{\Q for
youth
and the prime of life
are fleeting.

\par }\Chap{12}{\PP \par }{\Q \VerseOne{1}So remember
your Creator
in the days
of your youth
–
\par }{\Q before
the difficult days
come,
\par }{\Q and the years
draw
near when
you will say,
“I have no
pleasure in them”;
\par }{\Q \VS{2}before the sun
and the light
of the moon
and the stars
grow dark,
\par }{\Q and the clouds
disappear
after
the rain;
\par }{\Q \VS{3}when
those who keep watch
over the house
begin to tremble,
\par }{\Q and the virile
men
begin to stoop
over,

\par }{\Q and the grinders
begin to cease
because
they grow few,
\par }{\Q and those who look
through the windows
grow dim,
\par }{\Q \VS{4}and the doors
along the street
are shut;
\par }{\Q when the sound
of the grinding
mill
grows low,
\par }{\Q and one is awakened
by the sound
of a bird,
\par }{\Q and all
their songs grow faint,
\par }{\Q \VS{5}and
they are afraid
of heights
and the dangers
in the street;
\par }{\Q the almond blossoms
grow white,
\par }{\Q and the grasshopper
drags itself along,
\par }{\Q and the caper berry
shrivels up –
\par }{\Q because man goes to his eternal home,
\par }{\Q and the mourners go about in the streets –
\par }{\Q \VS{6}before
the silver
cord
is removed,
\par }{\Q or the golden
bowl
is broken,
\par }{\Q or the pitcher
is shattered
at
the well,
\par }{\Q or the water wheel
is broken
at
the cistern –
\par }{\Q \VS{7}and the dust
returns
to the earth
as it was,
\par }{\Q and the life’s breath
returns
to
God
who
gave it.
\par }{\SH Concluding Refrain: Qoheleth Restates His Thesis
\par }{\Q \VS{8}“Absolutely
futile!” laments
the Teacher,
\par }{\Q “All
of these things are futile!”
\par }{\SH Concluding Epilogue: Qoheleth’s Advice is Wise
\par }{\Q \VS{9}Not only
was
the Teacher
wise,
\par }{\Q but he also
taught
knowledge
to the people;
\par }{\Q he carefully evaluated
and arranged
many
proverbs.
\par }{\Q \VS{10}The Teacher
sought
to find
delightful
words,
\par }{\Q and to write
accurately
truthful
sayings.
\par }{\Q \VS{11}The words
of the sages
are like prods,
\par }{\Q and the collected
sayings are like firmly
fixed nails;
\par }{\Q they are given
by one
shepherd.
\par }{\SH Concluding Exhortation: Fear God and Obey His Commands!
\par }{\Q \VS{12}Be
warned,
my son,
of anything in addition
to them.

\par }{\Q There is no
end
to the making
of many
books,
\par }{\Q and much
study
is exhausting
to the body.
\par }{\Q \VS{13}Having heard
everything,
I have reached this conclusion:

\par }{\Q Fear
God
and keep
his commandments,
\par }{\Q because
this
is the whole
duty of man.
\par }{\Q \VS{14}For
God
will evaluate
every
deed,
\par }{\PP including every
secret
thing, whether
good
or
evil.
\par }