\NormalFont\ShortTitle{Ezekiel}
{\MT Ezekiel

\par }\ChapOne{1}{\SH A Vision of God’s Glory
\par }{\PP \VerseOne{1}In the thirtieth
year,
on the fifth day of the fourth
month,
while I
was among
the exiles
at the Kebar
River,
the heavens
opened
and I saw
a divine
vision.
\VS{2}(On the fifth
day of the month
– it was
the fifth
year
of King
Jehoiachin’s
exile –
\VS{3}the word
of the {\ND{Lord}}
came to
the priest
Ezekiel
the son
of Buzi,
at the Kebar
River
in the land
of the Babylonians.
The hand
of the {\ND{Lord}}
came on
him there).
\par }{\PP \VS{4}As I watched, I noticed
a windstorm
coming
from
the north
– an enormous
cloud,
with lightning flashing,
such that bright light
rimmed
it
and came from it like glowing
amber
from the middle
of a fire.
\VS{5}In
the fire
were what looked
like four
living
beings.
In their appearance
they had human
form,
\VS{6}but each
had four
faces
and four
wings.
\VS{7}Their legs
were straight,
but the soles
of their feet
were like calves’
feet.
They gleamed
like polished
bronze.
\VS{8}They had human
hands
under
their wings
on
their four sides.
As for the faces
and wings
of the four of them,
\VS{9}their wings
touched
each other;
they did not
turn
as they moved,
but went
straight
ahead.
\par }{\PP \VS{10}Their faces
had this appearance: Each of the four
had the face
of a man,
with the face
of a lion
on the right,
the face
of an ox
on the left
and also the face
of an eagle.
\VS{11}Their wings
were spread
out above
them; each
had two wings touching
the wings of one of the other beings
on either side and two
wings covering
their bodies.
\VS{12}Each
moved
straight
ahead –
wherever
the spirit
would go,
they would go,
without
turning
as they went.
\VS{13}In the middle
of the living
beings was something
like burning
coals
of fire
or like
torches.
It moved
back and forth among
the living
beings. It was
bright,
and lightning
was flashing out
of the fire.
\VS{14}The living
beings moved backward
and forward as quickly as flashes
of lightning.
\par }{\PP \VS{15}Then I looked,
and I saw
one
wheel
on the ground
beside
each of the four
beings.
\VS{16}The appearance
of the wheels
and their construction
was like gleaming
jasper,
and all four
wheels looked
alike.
Their structure
was
like a wheel
within
a wheel.
\VS{17}When they moved they would go
in any of the four
directions
they faced without
turning
as they moved.
\VS{18}Their rims
were high
and awesome,
and the rims
of all four
wheels were full
of eyes
all around.
\par }{\PP \VS{19}When
the living
beings moved,
the wheels
beside
them moved; when
the living
beings rose up from the ground,
the wheels
rose up too.
\VS{20}Wherever
the spirit
would go,
they would
go,
and the wheels
would rise up
beside
them because
the spirit
of the living
being was in the wheel.
\VS{21}When
the living beings moved,
the wheels moved,
and when
they stopped
moving, the wheels stopped.
When
they rose up
from the ground,
the wheels rose
up from the ground; the wheels
rose up beside
them because
the spirit
of the living
being was
in
the wheel.
\par }{\PP \VS{22}Over
the heads
of the living
beings
was something like a platform,
glittering awesomely
like ice,
stretched
out over
their heads.
\VS{23}Under
the platform
their wings
were
stretched out,
each
toward the other.
Each of the beings
also had two
wings covering
its body.
\VS{24}When
they moved, I heard
the
sound
of their wings
– it was like the sound
of rushing
waters,
or the voice
of the Almighty,
or the tumult
of an army.
When they stood
still, they lowered their wings.
\par }{\PP \VS{25}Then there was
a voice
from above
the platform
over
their heads
when they stood
still.
\VS{26}Above
the platform
over
their heads
was something like
a sapphire
shaped like
a throne.
High above on
the throne
was a form
that appeared
to be a man.
\VS{27}I saw
an amber
glow
like a fire
enclosed
all around
from his waist
up.
From his waist
down
I saw
something
that looked like fire.
There was a brilliant light
around it,
\VS{28}like the appearance
of a rainbow
in the clouds
after
the rain.
This
was the appearance
of the surrounding
brilliant light;
it
looked
like
the glory
of the {\ND{Lord}}. When I saw
it, I threw
myself face
down, and I heard
a voice
speaking.

\par }\Chap{2}{\PP \VerseOne{1}He said
to me,
“Son
of man,
stand
on
your feet
and I will speak with you.”
\VS{2}As
he spoke
to me,
a wind
came
into
me and stood
me on
my feet,
and I heard
the one speaking
to me.
\par }{\PP \VS{3}He said
to me,
“Son
of man,
I am
sending
you to
the house
of Israel,
to
rebellious
nations
who
have rebelled
against me; both they
and their fathers
have revolted
against me to
this
very
day.
\VS{4}The people to whom
I am
sending
you are obstinate and hard-hearted,
and you must
say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says.’
\VS{5}And as for them,
whether
they listen
or
not
– for
they are
a rebellious
house –
they
will know
that
a prophet
has been
among them.
\VS{6}But you,
son
of man,
do not
fear
them,
and do not
fear
their words
– even though
briers
and thorns
surround you and you
live among scorpions – do not fear their words and do not be terrified of the looks they give you, for they are a rebellious house!
\VS{7}You must speak
my words
to
them whether
they listen
or
not,
for
they are
rebellious.
\VS{8}As for you,
son
of man,
listen
to what
I am
saying
to
you: Do
not
rebel
like that rebellious
house! Open
your mouth
and eat
what
I am
giving you.”
\par }{\PP \VS{9}Then I looked
and realized
a hand
was stretched out
to
me, and in it was a written scroll.
\VS{10}He unrolled
it before
me, and it had writing
on the front
and back;
written
on
it were laments,
mourning,
and woe.

\par }\Chap{3}{\PP \VerseOne{1}He said
to
me, “Son
of man,
eat
what
you see in front of you – eat this scroll – and then go and speak to the house of Israel.”
\VS{2}So I opened
my mouth
and he fed
me the scroll.
\par }{\PP \VS{3}He said
to
me, “Son
of man,
feed
your stomach
and fill
your belly
with
this
scroll
I
am giving
to
you.” So I ate
it, and it was
sweet
like honey
in my mouth.
\par }{\PP \VS{4}He said
to
me, “Son
of man,
go
to
the house
of Israel
and speak
my words
to them.
\VS{5}For
you are
not
being sent
to
a people
of unintelligible
speech
and difficult
language,
but to
the house
of Israel –
\VS{6}not
to
many
peoples
of unintelligible
speech
and difficult
language,
whose
words
you cannot
understand –
surely if
I had sent
you to them, they
would listen
to you!
\VS{7}But the house
of Israel
is unwilling
to listen
to
you, because
they
are not
willing
to listen
to
me, for
the whole
house
of Israel
is hard-headed
and hard-hearted.
\par }{\PP \VS{8}“I
have made
your face
adamant
to match
their faces,
and your forehead
hard to match
their foreheads.
\VS{9}I have made
your forehead
harder
than flint
– like diamond! Do not
fear
them or
be terrified
of the looks they give
you, for
they are
a rebellious
house.”
\par }{\PP \VS{10}And he said
to me,
“Son
of man,
take
all
my words
that
I speak
to
you to heart
and listen carefully.
\VS{11}Go
to
the exiles,
to
your fellow countrymen,
and speak
to them – say to them, ‘This is what the sovereign
{\ND{Lord}} says,’ whether they pay attention or not.”
\par }{\SH Ezekiel Before the Exiles
\par }{\PP \VS{12}Then a wind
lifted
me up and I heard
a great
rumbling
sound
behind
me
as the glory
of the {\ND{Lord}}
rose from its place,
\VS{13}and the sound
of the living
beings’ wings
brushing
against
each
other,
and the sound
of the wheels
alongside
them, a great
rumbling
sound.
\VS{14}A wind
lifted
me up and carried
me away.
I went
bitterly,
my spirit
full of fury,
and the hand
of the {\ND{Lord}}
rested powerfully
on
me.
\VS{15}I came
to
the exiles
at Tel Abib,
who lived
by
the Kebar
River.
I sat
dumbfounded
among
them
there,
where
they were living,
for seven
days.
\par }{\PP \VS{16}At the end
of seven
days
the word
of the {\ND{Lord}}
came to me:
\VS{17}“Son
of man,
I have appointed you a watchman
for the house
of Israel.
Whenever you hear
a word
from my mouth,
you must
give them
a warning
from me.
\VS{18}When I say
to the wicked,
“You will certainly
die,”
and you do not
warn him – you do not speak out to warn the wicked to turn from his wicked deed and wicked lifestyle so that he may live – that wicked person will die for his iniquity, but I will hold you accountable for his death.
\VS{19}But as for
you,
if
you warn
the wicked
and he does not
turn
from his wicked
deed and from his wicked
lifestyle,
he
will die
for his iniquity
but you
will have saved
your own life.
\par }{\PP \VS{20}“When
a righteous
person turns
from his righteousness
and commits
iniquity,
and I set
an obstacle
before
him, he
will die.
If
you have not
warned
him, he will die
for his sin.
The righteous
deeds he performed
will not
be considered, but I will hold
you accountable for his death.
\VS{21}However, if you
warn
the righteous
person not
to sin,
and he
does not
sin,
he will certainly
live
because
he was warned,
and you
will have saved
your own life.”
\par }{\SH Isolated and Silenced
\par }{\PP \VS{22}The hand
of the {\ND{Lord}}
rested on
me there,
and he said
to me,
“Get
up, go out
to
the valley,
and I will speak
with
you there.”
\VS{23}So I got
up and went out
to
the valley,
and the glory
of the {\ND{Lord}}
was standing
there, just like the glory
I had
seen
by the Kebar
River,
and I threw
myself face
down.
\par }{\PP \VS{24}Then a wind
came
into me and stood
me on
my feet.
The
{\ND{Lord}} spoke
to
me and said,
“Go
shut
yourself in
your house.
\VS{25}As for you,
son
of man,
they will put
ropes
on
you and tie
you up
with them, so you cannot
go out
among them.
\VS{26}I will make your tongue
stick
to
the roof of your mouth
so that you will be
silent
and unable
to reprove
them, for
they
are a rebellious
house.
\VS{27}But when I speak
with
you, I will loosen
your tongue
and you must say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says.’ Those who listen
will listen,
but the indifferent
will refuse, for
they are
a rebellious
house.

\par }\Chap{4}{\PP \VerseOne{1}“And you,
son
of man,
take
a brick
and set it
in front
of you. Inscribe
a city
on it – Jerusalem.
\VS{2}Lay
siege
to it! Build
siege works against
it. Erect
a siege
ramp
against it! Post soldiers outside it
and station battering
rams around it.
\VS{3}Then
for your
part take
an iron
frying
pan and set
it up as an iron
wall
between
you and the city.
Set
your face
toward
it. It is to
be
under siege;
you are to besiege
it. This is a sign
for the house
of Israel.
\par }{\PP \VS{4}“Also for your
part lie
on
your left
side
and place
the
iniquity
of the house
of Israel
on
it. For the number
of days
you
lie
on
your side you will bear
their iniquity.
\VS{5}I
have determined
that the number of the years
of their iniquity
are to be the number
of days for you – 390 days. So bear the iniquity of the house of Israel.
\par }{\PP \VS{6}“When you have completed
these
days, then lie
down a second
time, but on your right
side,
and bear
the iniquity
of the house
of Judah
40
days –
I have assigned one day
for each
year.
\VS{7}You must turn your face toward
the siege
of Jerusalem
with your arm
bared
and prophesy
against it.
\VS{8}Look
here, I will
tie you up with ropes,
so you cannot
turn
from one side
to
the other
until
you complete
the days
of your siege.
\par }{\PP \VS{9}“As for you,
take
wheat,
barley,
beans,
lentils,
millet,
and spelt,
put
them in
a single
container,
and make
food
from them for yourself. For the same number
of days
that
you
lie
on
your side
– 390
days –
you will eat it.
\VS{10}The food
you eat
will be eight ounces
a day
by weight; you must eat
it at fixed times.
\VS{11}And you must drink
water
by measure,
a pint and a half;
you must drink
it at fixed times.
\VS{12}And you must eat
the food like you would a barley
cake.
You must bake
it in front of them
over a fire made with dried human
excrement.”
\VS{13}And the
{\ND{Lord}}
said,
“This
is how the people of Israel
will eat
their unclean
food
among the nations
where
I will banish them.”
\par }{\PP \VS{14}And I said,
“Ah,
sovereign

{\ND{Lord}}, I
have never
been ceremonially
defiled
before. I have never
eaten
a carcass
or an animal torn
by wild beasts; from my youth
up, unclean
meat
has never
entered
my mouth.”
\par }{\PP \VS{15}So he said
to
me, “All right then, I will substitute cow’s manure
instead
of human
excrement.
You will cook
your food
over it.”
\par }{\PP \VS{16}Then he said
to
me, “Son
of man,
I am about to remove
the bread
supply
in Jerusalem.
They will eat
their bread
ration
anxiously,
and they will drink
their water
ration
in terror
\VS{17}because
they will lack
bread
and water.
Each
one
will be terrified,
and they will rot
for their iniquity.

\par }\Chap{5}{\PP \VerseOne{1}“As for you,
son
of man,
take
a sharp
sword
and use it as a barber’s
razor.
Shave off
some of the hair from
your head
and your beard.
Then take
scales
and divide up the hair you cut off.
\VS{2}Burn
a third
of it in the fire
inside
the city
when the days
of your siege
are completed.
Take
a third
and slash
it with a sword
all around
the city. Scatter
a third
to the wind,
and I will unleash
a sword
behind them.
\VS{3}But take
a few
strands of hair from those and tie
them
in the ends of your garment.
\VS{4}Again,
take
more of them and throw
them into
the fire,
and burn
them up.
From
there a fire
will spread
to
all
the house
of Israel.
\par }{\PP \VS{5}“This is what
the sovereign

{\ND{Lord}}
says: This
is Jerusalem;
I placed
her in the center
of the nations
with countries
all around her.
\VS{6}Then she defied
my regulations
and my statutes,
becoming more wicked
than
the nations
and the countries
around
her. Indeed,
they have rejected
my regulations,
and they do not
follow
my statutes.
\par }{\PP \VS{7}“Therefore
this is what
the sovereign

{\ND{Lord}}
says: Because
you are more arrogant
than
the nations
around
you, you have not followed my statutes
and have not
carried
out my regulations.
You have not
even carried
out the regulations
of the nations
around you!
\par }{\PP \VS{8}“Therefore
this is what
the sovereign

{\ND{Lord}}
says: I
– even I – am against you, and I will execute judgment among you while the nations watch.
\VS{9}I will do
to you what
I have never
done
before and will never
do
again
because
of all
your abominable practices.
\VS{10}Therefore
fathers
will eat
their sons
within
you, Jerusalem, and sons
will eat
their fathers.
I will execute
judgments
on you, and I will scatter
any
survivors
to the winds.
\par }{\PP \VS{11}“Therefore,
as surely as I
live,
says
the sovereign

{\ND{Lord}}, because
you defiled
my sanctuary
with all
your detestable idols
and with all
your abominable practices,
I
will withdraw;
my eye will not
pity
you, nor
will I
spare you.
\VS{12}A third
of your people will die
of plague
or be overcome
by the famine
within
you. A third
of your people will fall
by the sword
surrounding
you, and a third
I will scatter
to the winds.
I will unleash
a sword
behind them.
\VS{13}Then my anger
will be fully vented;
I will exhaust
my rage
on them, and I will be appeased.
Then they will know
that
I,
the {\ND{Lord}}, have spoken
in my jealousy
when
I have
fully vented
my rage against them.
\par }{\PP \VS{14}“I will make
you desolate
and an object of scorn
among the nations
around
you, in the sight
of everyone
who passes by.
\VS{15}You will be
an object of scorn
and taunting,
a prime example of destruction
among the nations
around
you when I execute
judgments
against you in anger
and raging
fury.
I,
the {\ND{Lord}},
have spoken!
\VS{16}I will shoot against
them deadly,
destructive
arrows
of famine,
which
I will shoot
to destroy
you. I will prolong
a famine
on
you and will remove
the bread
supply.
\VS{17}I will send
famine
and wild beasts
against you and they will take your children from you. Plague
and bloodshed
will overwhelm
you, and I will bring
a sword
against
you. I,
the {\ND{Lord}},
have spoken!”

\par }\Chap{6}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
the mountains
of Israel
and prophesy
against them:
\VS{3}Say,
‘Mountains
of Israel,
Hear
the word
of the sovereign

{\ND{Lord}}! This is what
the sovereign

{\ND{Lord}}
says
to the mountains
and the hills,
to the ravines
and the valleys: I am
bringing
a sword
against
you, and I will destroy
your high places.
\VS{4}Your altars will be ruined
and your incense altars
will be broken.
I will
throw down
your slain
in front
of your idols.
\VS{5}I will place
the corpses
of the people of Israel
in front
of their idols,
and I will scatter
your bones
around
your altars.
\VS{6}In all
your dwellings,
the cities
will be laid waste
and the high places
ruined
so that
your altars
will be laid waste
and ruined,
your idols
will be shattered
and demolished,
your incense altars
will be broken down,
and your works
wiped out.
\VS{7}The slain
will fall
among
you and then you will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{8}“‘But
I will
spare
some of you. Some will escape
the sword
when you are scattered
in foreign
lands.
\VS{9}Then your survivors
will remember
me among the
nations
where
they are exiled.
They will
realize how
I
was crushed
by their unfaithful
heart
which
turned
from me
and by their eyes
which lusted
after
their idols.
They will loathe
themselves because of the evil
they have
done
and
because of all
their abominable practices.
\VS{10}They will know
that
I am
the {\ND{Lord}};
my threats
to bring this
catastrophe
on them were not empty.’
\par }{\PP \VS{11}“‘This is what
the sovereign

{\ND{Lord}}
says: Clap
your hands,
stamp
your feet,
and say,
“Ah!” because
of all
the evil,
abominable practices
of the house
of Israel,
for they will fall
by the sword,
famine,
and pestilence.
\VS{12}The one far
away will die
by pestilence,
the one close
by will fall
by the sword,
and whoever
is left
and has escaped
these will die
by
famine.
I will fully vent
my rage against them.
\VS{13}Then you will know
that
I
am the
{\ND{Lord}} –
when
their dead
lie among
their idols
around
their altars,
on
every
high
hill
and all
the mountaintops,
under
every
green
tree
and every
leafy
oak,
the places
where
they have
offered
fragrant
incense
to all
their idols.
\VS{14}I will stretch out
my hand
against
them and make
the
land
a desolate
waste
from the wilderness
to Riblah,
in all
the places where they live.
Then they will know
that
I
am the
{\ND{Lord}}!”

\par }\Chap{7}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“You,
son
of man
– this is what
the sovereign

{\ND{Lord}}
says to the land
of Israel: An end! The end
is coming
on
the four
corners
of the land!
\VS{3}The end
is now
upon
you, and I will release
my anger
against you; I will judge
you according to your behavior,
I will
hold you accountable
for all
your abominable practices.
\VS{4}My eye will not
pity
you; I will not
spare
you. For
I will hold you responsible for your behavior,
and you will suffer
the consequences of your abominable practices.
Then
you will
know
that
I
am the
{\ND{Lord}}!
\par }{\PP \VS{5}“This is what
the sovereign

{\ND{Lord}}
says: A disaster –
a one-of-a-kind
disaster
– is
coming!
\VS{6}An end
comes –
the end comes! It has awakened
against you – the end is upon you! Look, it is coming!
\VS{7}Doom
is coming
upon you who live
in the land! The time
is coming,
the day
is near.
There are sounds of tumult,
not
shouts
of joy, on the mountains.
\VS{8}Soon
now I will pour
out my rage
on
you; I will fully vent
my anger
against you. I will judge
you according to your behavior.
I will
hold you accountable
for all
your abominable practices.
\VS{9}My eye will not
pity
you; I will not
spare
you. For your behavior
I will
hold you accountable,
and you will suffer the consequences
of your abominable
practices. Then you will
know
that
it is I,
the {\ND{Lord}},
who is striking you.
\par }{\PP \VS{10}“Look,
the day! Look,
it is coming! Doom
has gone out! The staff
has budded,
pride
has blossomed!
\VS{11}Violence
has grown into
a staff
that supports wickedness.
Not
one of them will be left – not from their crowd, not from their wealth, not from their prominence.
\VS{12}The time
has come;
the day
has struck! The customer
should not
rejoice,
nor
the seller
mourn;
for
divine wrath
comes against
their whole
crowd.
\VS{13}The customer will no longer
pay
the seller
while
both
parties are alive,
for
the vision
against
their whole
crowd
will not
be revoked.
Each person,
for his iniquity,
will fail
to preserve
his life.
\par }{\PP \VS{14}“They have blown
the trumpet
and everyone
is ready,
but no
one goes
to battle,
because
my anger
is against their whole
crowd.
\VS{15}The sword
is outside;
pestilence
and famine
are inside the house.
Whoever
is in the open field
will die
by the sword,
and famine
and pestilence
will consume
everyone in the city.
\VS{16}Their survivors
will escape
to
the mountains
and become like doves
of the valleys;
all
of them will moan
– each
one for his iniquity.
\VS{17}All
of their hands
will hang limp;
their knees
will be wet with urine.
\VS{18}They will wear sackcloth,
terror
will cover
them; shame
will be on all
their faces,
and all
of their heads
will be shaved bald.
\VS{19}They will discard
their silver
in the streets,
and their gold
will be
treated like filth.
Their silver
and gold
will not
be able
to deliver
them on the day
of the
{\ND{Lord}}’s
fury.
They will not satisfy
their hunger
or
fill
their stomachs
because
their wealth was the obstacle
leading to their iniquity.
\VS{20}They rendered
the beauty
of his ornaments
into pride,
and with it they made
their abominable
images
– their detestable idols.
Therefore
I will render
it filthy to them.
\VS{21}I will give
it to foreigners
as loot,
to the world’s
wicked
ones as plunder,
and they will desecrate it.
\VS{22}I will turn
my face
away from them and they
will desecrate
my treasured
place. Vandals
will enter
it and desecrate it.
\VS{23}(Make
the chain,
because
the land
is full
of murder
and the city
is full
of violence.)
\VS{24}I will bring
the most wicked
of the nations
and they will take possession
of their houses.
I will put an end
to the arrogance
of the strong, and their sanctuaries
will be
desecrated.
\VS{25}Terror
is coming! They will seek
peace,
but find none.
\VS{26}Disaster
after
disaster
will come,
and one rumor
after
another.
They will seek
a vision
from a prophet;
priestly
instruction
will disappear,
along with counsel
from the elders.
\VS{27}The king
will mourn
and the prince
will be clothed
with shuddering;
the hands
of the people
of the land
will tremble.
Based on their behavior
I will deal
with them, and by their standard of justice
I will judge
them. Then they will know
that
I
am the
{\ND{Lord}}!”

\par }\Chap{8}{\PP \VerseOne{1}In the sixth
year,
in the sixth
month, on the fifth
of the month,
as I
was sitting
in my house
with the elders
of Judah
sitting
in front
of me, the hand
of the sovereign

{\ND{Lord}}
seized me.
\VS{2}As I watched,
I noticed
a form
that appeared
to be a man. From his waist
downward
was
something
like fire,
and from his waist
upward
something
like a brightness,
like an amber
glow.
\VS{3}He stretched out
the form
of a hand
and grabbed
me by a lock
of hair on my head.
Then
a wind
lifted
me up between
the earth
and sky
and brought
me to
Jerusalem
by means of divine
visions,
to the door
of the inner
gate
which faces
north
where
the statue
which provokes
to
jealousy was located.
\VS{4}Then I
perceived
that the glory
of the God
of Israel
was there, as
in the vision
I had
seen
earlier in the valley.
\par }{\PP \VS{5}He said
to
me, “Son
of man,
look up
toward
the north.”
So I looked up
toward
the north,
and I noticed
to the north
of the altar
gate
was this
statue
of jealousy
at the entrance.
\par }{\PP \VS{6}He said
to me,
“Son
of man,
do you
see
what they
are doing
– the great
abominations
that
the people
of Israel
are practicing
here,
to drive me far
from my sanctuary? But you will
see
greater
abominations
than these!”
\par }{\PP \VS{7}He brought
me to
the entrance
of the court,
and as I watched, I noticed
a hole
in the wall.
\VS{8}He said
to me,
“Son
of man,
dig
into the wall.”
So I dug into
the wall
and discovered
a doorway.
\par }{\PP \VS{9}He said
to
me, “Go
in and see
the evil
abominations
they are practicing
here.”
\VS{10}So I went
in and looked.
I noticed
every
figure
of creeping thing
and beast
– detestable images – and every idol of the house of Israel, engraved on the wall all around.
\VS{11}Seventy
men
from the elders
of the house
of Israel
(with Jaazaniah
son
of Shaphan
standing
among
them) were standing
in front of
them, each
with a censer
in his hand,
and fragrant vapors
from a cloud
of incense
were swirling upward.
\par }{\PP \VS{12}He said
to
me, “Do you see,
son
of man,
what
the elders
of the house
of Israel
are doing
in the dark,
each
in the chamber
of his idolatrous images? For
they think,
‘The
{\ND{Lord}}
does not
see
us! The
{\ND{Lord}}
has abandoned
the
land!’ ”
\VS{13}He said
to me,
“You will
see
them practicing
even greater
abominations!”
\par }{\PP \VS{14}Then he brought
me to
the entrance
of the north
gate
of the
{\ND{Lord}}’s
house.
I noticed
women
sitting
there
weeping
for Tammuz.
\VS{15}He said
to me,
“Do you see
this, son
of man? You will
see
even greater
abominations
than
these!”
\par }{\PP \VS{16}Then he brought
me to
the inner
court
of the
{\ND{Lord}}’s
house.
Right there
at the entrance
to the
{\ND{Lord}}’s
temple,
between
the porch
and the altar,
were about twenty-five men
with their backs
to
the
{\ND{Lord}}’s
temple,
facing
east
– they
were worshiping
the sun
toward the east!
\par }{\PP \VS{17}He said
to me,
“Do you see,
son
of man? Is it a trivial
thing that the house
of Judah
commits
these abominations
they are practicing
here? For
they have filled
the
land
with violence
and provoked me to anger
still further. Look,
they are putting
the branch
to
their nose!
\VS{18}Therefore
I
will act
with fury! My eye will
not
pity
them nor will I spare
them. When they have shouted
in my ears,
I will not
listen to them.”

\par }\Chap{9}{\PP \VerseOne{1}Then he shouted
in my ears,
“Approach,
you who are to visit destruction
on the city,
each
with
his destructive
weapon in his hand!”
\VS{2}Next, I noticed
six
men
coming
from the direction
of the upper
gate
which
faces
north,
each
with his war club
in his hand.
Among
them was a man
dressed
in linen
with a writing kit
at his side.
They
came
and stood
beside
the bronze
altar.
\par }{\PP \VS{3}Then the glory
of the God
of Israel
went up
from the cherub
where
it had rested to
the threshold
of the temple.
He called
to
the man
dressed
in linen
who had
the writing kit at his side.
\VS{4}The
{\ND{Lord}}
said
to
him, “Go
through
the city
of Jerusalem
and put
a mark
on
the foreheads
of the people
who moan
and groan
over
all
the abominations
practiced
in it.”
\par }{\PP \VS{5}While
I listened,
he said
to the others, “Go through
the city
after
him and strike
people down; do
no let
your eye
pity nor spare anyone!
\VS{6}Old men,
young men,
young women,
little children,
and women
– wipe
them out! But do not
touch
anyone
who has
the mark.
Begin
at my sanctuary!” So they began
with the elders
who
were at the front
of the temple.
\par }{\PP \VS{7}He said
to them,
“Defile
the temple
and fill
the courtyards
with corpses.
Go!” So
they went
out
and struck
people down throughout the city.
\VS{8}While
they were striking
them down, I was left
alone, and I
threw
myself face
down and cried
out, “Ah,
sovereign

{\ND{Lord}}! Will you
destroy
the
entire
remnant
of Israel
when you pour
out your fury
on
Jerusalem?”
\par }{\PP \VS{9}He said
to
me, “The sin
of the house
of Israel
and Judah
is extremely
great;
the land
is full
of murder,
and the city
is full
of corruption,
for
they say,
‘The
{\ND{Lord}}
has abandoned
the
land,
and the
{\ND{Lord}}
does not
see!’
\VS{10}But as
for me,
my eye will not
pity
them nor will I spare
them; I hereby
repay them for what they have done.”
\par }{\PP \VS{11}Next I noticed
the man
dressed
in linen
with the writing kit
at his side bringing back
word: “I have done
just
as you commanded me.”

\par }\Chap{10}{\PP \VerseOne{1}As I watched,
I saw
on
the platform
above
the top
of the cherubim
something like a sapphire,
resembling
the shape
of a throne,
appearing
above them.
\VS{2}The
{\ND{Lord}} said
to
the man
dressed
in linen,
“Go
between
the wheelwork
underneath
the cherubim.
Fill
your hands
with burning coals
from among
the cherubim
and scatter
them over
the city.”
He went
as I watched.
\par }{\PP \VS{3}(The cherubim
were standing
on the south side
of the temple
when the man
went
in, and a cloud
filled
the inner
court.)
\VS{4}Then
the glory
of the
{\ND{Lord}}
arose from the cherub
and moved to the threshold
of the temple.
The temple
was filled
with
the cloud
while the court
was filled
with
the brightness
of the
{\ND{Lord}}’s
glory.
\VS{5}The sound
of the wings
of the cherubim
could be heard
from
the outer
court,
like the sound
of the sovereign
God
when he speaks.
\par }{\PP \VS{6}When
the
{\ND{Lord}} commanded
the
man
dressed
in linen,
“Take
fire
from within the wheelwork,
from among
the cherubim,”
the man went
in and stood
by
one of the wheels.
\VS{7}Then one of the cherubim
stretched out
his hand
toward
the fire
which
was among
the cherubim.
He took
some and put
it into
the hands
of the man dressed
in linen,
who took
it and left.
\VS{8}(The cherubim
appeared
to have the form
of human
hands
under
their wings.)
\par }{\PP \VS{9}As I watched,
I noticed
four
wheels
by
the cherubim,
one
wheel
beside
each
cherub;
the wheels
gleamed
like jasper.
\VS{10}As for their appearance,
all
four
of them looked the same,
something like a wheel
within
a wheel.
\VS{11}When they moved,
they would go
in any of the four
directions
they faced
without
turning
as
they moved;
in the direction
the head
would turn they would follow without turning
as they moved,
\VS{12}along with their entire
bodies,
their backs,
their hands,
and their wings.
The wheels
of the four
of them were full
of eyes
all around.
\VS{13}As for their wheels,
they were called
“the wheelwork”
as I listened.
\VS{14}Each of the cherubim had four
faces: The first
was the face
of a
cherub,
the second
that of a man,
the third
that of a lion,
and the fourth
that of an eagle.
\par }{\PP \VS{15}The cherubim
rose
up; these were the living
beings I saw
at the Kebar
River.
\VS{16}When
the cherubim
moved,
the wheels
moved beside
them; when
the cherubim
spread their wings
to rise
from the ground,
the wheels
did not
move
from their
side.
\VS{17}When the cherubim stood
still, the wheels stood
still, and when they rose up,
the wheels
rose
up with them, for
the spirit
of the living beings was in the wheels.
\par }{\PP \VS{18}Then
the glory
of the {\ND{Lord}}
moved away from the threshold
of the temple
and stopped
above
the cherubim.
\VS{19}The
cherubim
spread
their wings,
and they rose up
from
the earth
while I watched
(when they went
the wheels
went alongside
them). They stopped
at the entrance
to the east
gate
of the
{\ND{Lord}}’s
temple
as the glory
of the God
of Israel
hovered
above them.
\par }{\PP \VS{20}These were the living
creatures which
I saw
at the Kebar
River
underneath
the God
of Israel;
I knew
that
they were cherubim.
\VS{21}Each
had four
faces;
each
had four
wings
and the form
of human
hands
under
the wings.
\VS{22}As for the form
of their faces,
they
were the faces
whose
appearance
I had
seen
at the Kebar
River.
Each
one moved
straight
ahead.

\par }\Chap{11}{\PP \VerseOne{1}A wind
lifted
me up
and brought
me to
the east
gate
of the Lord’s
temple
that faces
the east.
There, at the entrance
of the gate,
I noticed twenty-five
men.
Among
them I saw
Jaazaniah
son
of Azzur
and Pelatiah
son
of Benaiah,
officials
of the people.
\VS{2}The
{\ND{Lord}} said
to me,
“Son
of man,
these
are the men
who plot
evil
and give
wicked
advice
in this
city.
\VS{3}They say, ‘The time is not
near
to build
houses;
the city is a cooking
pot and we
are the meat in it.’
\VS{4}Therefore,
prophesy
against
them! Prophesy,
son
of man!”
\par }{\PP \VS{5}Then
the Spirit
of the {\ND{Lord}}
came upon me and said
to me,
“Say: This is what
the {\ND{Lord}}
says: ‘This
is what you are thinking,
O house
of Israel;
I
know
what goes through
your minds.
\VS{6}You have killed
many
people in this
city;
you have filled
its streets
with corpses.’
\VS{7}Therefore,
this is what
the sovereign

{\ND{Lord}}
says: ‘The corpses
you have
dumped
in the midst
of the city are the meat,
and this city is the cooking pot,
but I will take
you out
of it.
\VS{8}You fear
the sword,
so the sword
I will bring
against
you,’ declares
the sovereign

{\ND{Lord}}.
\VS{9}‘But I will take you out
of the city.
And I will
hand
you over to foreigners.
I will execute
judgments on you.
\VS{10}You will die
by the sword;
I will
judge
you at the border
of Israel.
Then you will know
that
I am
the {\ND{Lord}}.
\VS{11}This city will not
be a cooking pot
for you, and you
will not be
meat
within
it; I will judge
you at the border
of Israel.
\VS{12}Then you will know
that
I am
the {\ND{Lord}}, whose
statutes
you have not
followed
and whose regulations
you have not
carried out. Instead you have behaved
according to the regulations
of the nations
around
you!’ ”
\par }{\PP \VS{13}Now, while
I was prophesying,
Pelatiah
son
of Benaiah
died.
Then
I threw
myself face
down
and cried
out with a loud
voice,
“Alas,
sovereign

{\ND{Lord}}! You
are completely
wiping out
the remnant
of Israel!”
\par }{\PP \VS{14}Then
the word
of the {\ND{Lord}}
came to me:
\VS{15}“Son
of man,
your brothers,
your relatives,
and the whole
house
of Israel,
all
of them are those to whom
the inhabitants
of Jerusalem
have said,
‘They have gone far away
from
the {\ND{Lord}}; to us this
land
has been given
as a possession.’
\par }{\PP \VS{16}“Therefore
say: ‘This is what
the sovereign

{\ND{Lord}}
says: Although
I have removed them far away
among the nations
and have dispersed
them among the countries,
I have been
a little
sanctuary
for them among the lands
where
they have gone.’
\par }{\PP \VS{17}“Therefore
say: ‘This is what
the sovereign

{\ND{Lord}}
says: When I regather
you from
the peoples
and assemble
you from
the lands
where
you have been dispersed,
I will give
you back the country
of Israel.’
\par }{\PP \VS{18}“When they return
to it, they will remove
from it all
its detestable
things and all
its abominations.
\VS{19}I will give
them one
heart
and I will put
a new
spirit
within
them; I will remove
the hearts
of stone
from their bodies
and I will give
them tender
hearts,
\VS{20}so that they may
follow
my statutes
and observe
my regulations
and carry
them out.
Then
they will be
my people,
and I
will be
their God.
\VS{21}But those whose hearts
are devoted
to detestable things
and abominations,
I hereby repay
them for what they have done,
says
the sovereign

{\ND{Lord}}.”
\par }{\PP \VS{22}Then
the cherubim
spread their wings
with their wheels
alongside
them while the glory
of the God
of Israel
hovered
above them.
\VS{23}The glory
of the {\ND{Lord}}
rose up
from within
the city
and stopped
over
the mountain
east of it.
\VS{24}Then a wind
lifted
me up
and carried
me to
the exiles
in Babylonia,
in the vision
given to me by the Spirit
of God.
\par }{\PP Then the vision
I had
seen
went up
from me.
\VS{25}So I told
the exiles
everything
the {\ND{Lord}}
had
shown me.

\par }\Chap{12}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
you
are living
in the midst
of a rebellious
house.
They have
eyes
to see,
but do not
see,
and ears
to hear,
but do not
hear,
because
they are
a rebellious
house.
\par }{\PP \VS{3}“Therefore,
son
of man,
pack
up your belongings
as if for
exile.
During the day,
while they are watching,
pretend to
go into exile.
Go from where
you live to another
place.
Perhaps
they
will understand, although
they are
a rebellious
house.
\VS{4}Bring out
your belongings
packed
for exile
during the day
while they are watching.
And go out
at evening,
while they are watching,
as if for exile.
\VS{5}While they are watching,
dig a hole
in the wall
and carry your belongings out through it.
\VS{6}While they are watching,
raise your baggage onto your shoulder
and carry
it out
in the dark.
You must cover
your face
so that you cannot
see
the
ground
because
I have made you an object lesson
to the house
of Israel.”
\par }{\PP \VS{7}So
I did
just
as I was commanded.
I carried
out
my belongings packed
for exile
during the day,
and at evening
I dug myself a hole
through the wall
with my hands.
I went out
in the darkness,
carrying
my baggage on
my shoulder
while they watched.
\par }{\PP \VS{8}The word
of the {\ND{Lord}}
came to
me in the morning:
\VS{9}“Son
of man,
has not
the house
of Israel,
that rebellious
house,
said
to
you, ‘What
are you
doing?’
\VS{10}Say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says: The prince
will raise this
burden
in Jerusalem,
and all
the house
of Israel
within it.’
\VS{11}Say,
‘I am
an object lesson
for you. Just
as I have done,
it will be done
to them; they will go into exile
and captivity.’
\par }{\PP \VS{12}“The prince
who
is among
them will raise
his belongings onto his shoulder
in darkness,
and will go out.
He will dig
a hole in the wall
to leave
through. He will cover
his face
so that
he cannot
see
the
land
with his eyes.
\VS{13}But I
will throw
my net
over
him, and he will be caught
in my snare.
I will bring
him to Babylon,
the land
of the Chaldeans
(but he will not
see
it), and there
he will die.
\VS{14}All
his retinue
– his attendants
and his troops
– I will scatter
to every
wind;
I will unleash
a sword
behind them.
\par }{\PP \VS{15}“Then they will know
that
I am
the {\ND{Lord}}
when I disperse them among
the nations
and scatter
them among foreign countries.
\VS{16}But
I will let a small
number
of them
survive
the sword,
famine,
and pestilence,
so
that they can confess
all
their abominable practices
to the nations
where
they go.
Then they will know
that
I am
the {\ND{Lord}}.”
\par }{\PP \VS{17}The word
of the {\ND{Lord}}
came to me:
\VS{18}“Son
of man,
eat
your bread
with trembling,
and drink
your water
with anxious
shaking.
\VS{19}Then say
to
the people
of the land,
‘This is what
the sovereign

{\ND{Lord}}
says
about the inhabitants
of Jerusalem
and of the land
of Israel: They will eat
their bread
with anxiety
and drink
their water
in fright,
for
their land
will be stripped
bare of all it contains
because of the violence
of all
who live in it.
\VS{20}The inhabited
towns
will be
left in ruins
and the land
will be devastated.
Then you will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\PP \VS{21}The word
of the {\ND{Lord}}
came to me:
\VS{22}“Son
of man,
what
is this
proverb
you have in
the land
of Israel,
‘The days
pass slowly,
and every
vision
fails’?
\VS{23}Therefore
tell
them,
‘This is what
the sovereign

{\ND{Lord}}
says: I hereby end
this
proverb;
they will not
recite
it in Israel
any longer.’
But
say
to them,
‘The days
are at hand when every
vision
will be fulfilled.
\VS{24}For
there will no
longer
be any
false
visions
or flattering
omens
amidst
the house
of Israel.
\VS{25}For
I,
the {\ND{Lord}}, will speak.
Whatever
word
I speak
will be accomplished. It
will not
be delayed
any longer.
Indeed
in your days,
O
rebellious
house,
I will speak
the word
and accomplish it,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\PP \VS{26}The word
of the {\ND{Lord}}
came to me:
\VS{27}“Take note, son
of man,
the house
of Israel
is saying,
‘The vision
that
he sees
is for distant
days;
he is
prophesying about the far future.’
\VS{28}Therefore
say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says: None
of my words
will be delayed
any
longer! The word
I speak
will come to pass, declares
the sovereign

{\ND{Lord}}.’ ”

\par }\Chap{13}{\PP \VerseOne{1}Then
the word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
prophesy
against the prophets
of Israel
who are now prophesying.
Say
to the prophets
who prophesy from their imagination: ‘Hear
the word
of the {\ND{Lord}}!
\VS{3}This is what
the sovereign

{\ND{Lord}}
says: Woe
to the foolish
prophets
who
follow
their own
spirit
but have seen
nothing!
\VS{4}Your prophets
have become
like jackals
among the ruins,
O Israel.
\VS{5}You have not
gone up
in the breaks
in the wall, nor repaired
a wall
for the house
of Israel
that it would stand
strong in the battle
on the day
of the {\ND{Lord}}.
\VS{6}They see
delusion
and their omens
are a lie.
They say,
“the
{\ND{Lord}}
declares,”
though the
{\ND{Lord}}
has not
sent
them; yet they expect
their word
to be confirmed.
\VS{7}Have you not
seen
a false
vision
and announced
a lying
omen
when you say,
“the
{\ND{Lord}}
declares,”
although I myself
never
spoke?
\par }{\PP \VS{8}“‘Therefore,
this is what
the sovereign

{\ND{Lord}}
says: Because
you have spoken
false
words and forecast
delusion, look,
I am against you, declares
the sovereign

{\ND{Lord}}.
\VS{9}My hand
will be
against
the prophets
who see
delusion
and announce
lying
omens.
They will not
be included
in the council
of my people,
nor
be written
in the registry
of the house
of Israel,
nor
enter
the land
of Israel.
Then you will know
that
I
am the sovereign

{\ND{Lord}}.
\par }{\PP \VS{10}“‘This
is because
they have led
my people
astray saying,
“All is well,”
when things are not
well.
When anyone
builds
a wall
without mortar, they coat
it with whitewash.
\VS{11}Tell
the ones who coat
it with whitewash
that it will fall.
When
there is a deluge
of rain,
hailstones
will fall
and a violent
wind
will break out.
\VS{12}When the wall
has collapsed,
people will ask
you, “Where
is the whitewash
you coated it with?”
\par }{\PP \VS{13}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says: In my rage
I will make a violent
wind
break out.
In my anger
there will be
a deluge
of rain
and hailstones
in destructive
fury.
\VS{14}I will break
down the
wall
you coated
with whitewash
and knock
it to
the ground
so that its foundation
is exposed.
When it falls
you will be destroyed
beneath
it, and you will know
that
I am
the {\ND{Lord}}.
\VS{15}I will vent
my rage
against the wall,
and against those who coated
it with whitewash.
Then I will say
to you, “The wall
is no
more and those who whitewashed
it are no more –
\VS{16}those prophets
of Israel
who would prophesy
about Jerusalem
and would see
visions
of peace
for it, when there was no
peace,”
declares
the sovereign

{\ND{Lord}}.’
\par }{\PP \VS{17}“As for you,
son
of man,
turn
toward
the daughters
of your people
who
are prophesying
from their imagination.
Prophesy
against them
\VS{18}and say
‘This is what
the sovereign

{\ND{Lord}}
says: Woe
to those who sew
bands
on
all
their wrists
and make
headbands
for heads
of every
size to entrap
people’s lives! Will you entrap
my people’s
lives,
yet preserve
your own lives?
\VS{19}You have profaned
me
among my people
for handfuls
of barley
and scraps
of bread.
You have put to death
people
who
should not
die
and kept alive
those
who
should not
live
by your lies
to my people,
who listen
to lies!
\par }{\PP \VS{20}“‘Therefore,
this is what
the sovereign

{\ND{Lord}}
says: Take note
that I am
against your wristbands
with which
you
entrap
people’s lives
like birds.
I will
tear
them
from your arms
and will release
the
people’s lives,
which
you
hunt
like birds.
\VS{21}I will tear
off your headbands
and rescue
my people
from your power;
they will
no
longer
be
prey
in your hands.
Then you will know
that
I am
the {\ND{Lord}}.
\VS{22}This is because
you have disheartened
the righteous
person with lies
(although I
have not
grieved
him), and because you have encouraged
the wicked person
not
to turn
from his evil
conduct
and preserve his life.
\VS{23}Therefore
you will no
longer
see
false
visions
and practice divination.
I will rescue
my people
from your power,
and you will know
that
I am
the {\ND{Lord}}.’ ”

\par }\Chap{14}{\PP \VerseOne{1}Then some men
from Israel’s
elders
came
to
me and sat down
in front of me.
\VS{2}The word
of the {\ND{Lord}}
came to me:
\VS{3}“Son
of man,
these
men
have erected
their idols
in their hearts
and placed
the obstacle
leading to their iniquity
right
before
their faces. Should
I really
allow
them to seek me?
\VS{4}Therefore
speak
to them and say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says: When any one
from the house
of Israel
erects
his idols
in his heart
and sets
the obstacle
leading to his iniquity
before his face,
and then consults
a prophet,
I
the {\ND{Lord}}
am determined to
answer
him personally according to the enormity
of his idolatry.
\VS{5}I will do this
in order
to capture
the
hearts
of the house
of Israel,
who have
alienated
themselves from me on
account of all
their idols.’
\par }{\PP \VS{6}“Therefore
say
to
the house
of Israel,
‘This is what
the sovereign

{\ND{Lord}}
says: Return! Turn
from your idols,
and turn
your faces
away from your abominations.
\VS{7}For
when anyone
from the house
of Israel,
or the foreigner
who
lives
in Israel,
separates
himself from me and erects
his idols
in
his heart
and sets the obstacle
leading to his iniquity
before
his face,
and then consults
a prophet
to
seek
something from me, I
the {\ND{Lord}}
am determined to answer him personally.
\VS{8}I will set
my face
against that
person
and will make
him an object lesson
and a byword
and will cut
him off
from among
my people.
Then you will know
that
I
am the
{\ND{Lord}}.
\par }{\PP \VS{9}“‘As for
the prophet,
if
he is made a fool by being deceived
into speaking
a prophetic word
– I,
the {\ND{Lord}}, have made a fool
of that prophet,
and I will stretch
out my hand
against
him and destroy
him from among
my people
Israel.
\VS{10}They will bear
their punishment;
the punishment
of the one who sought
an oracle will be the same as the punishment
of the prophet who gave it
\VS{11}so
that the house
of Israel
will no
longer
go astray
from me,
nor
continue
to defile
themselves
by all
their sins.
They will be
my people
and I
will be
their God,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\PP \VS{12}The word
of the {\ND{Lord}}
came to me:
\VS{13}“Son
of man,
suppose
a country
sins
against me by being
unfaithful,
and I stretch
out my hand
against
it,
cut off its bread
supply,
cause famine
to come on it, and kill both people
and animals.
\VS{14}Even if these
three
men,
Noah,
Daniel,
and Job,
were
in it, they
would save
only their own lives
by their righteousness,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{15}“Suppose
I were to send wild animals
through
the land
and kill its children, leaving
it desolate,
without
travelers
due
to the wild animals.
\VS{16}Even if these three
men
were
in it,
as surely
as I
live, declares
the sovereign

{\ND{Lord}}, they could
not save their own sons
or
daughters;
they would save
only their own lives,
and the land
would become
desolate.
\par }{\PP \VS{17}“Or
suppose I were to bring
a sword
against
that land
and say,
‘Let a sword
pass
through the land,’
and I
were to kill
both people
and animals.
\VS{18}Even if these three
men
were
in it,
as surely
as I
live, declares
the sovereign

{\ND{Lord}}, they could not
save
their own sons
or daughters
– they
would save
only their own lives.
\par }{\PP \VS{19}“Or
suppose I were to send
a plague
into
that land,
and pour
out my rage
on
it with bloodshed,
killing
both people
and animals.
\VS{20}Even if Noah,
Daniel,
and Job
were in it,
as surely
as I
live, declares
the sovereign

{\ND{Lord}}, they could not
save their own son
or
daughter;
they
would save
only their own
lives
by their righteousness.
\par }{\PP \VS{21}“For
this is what
the sovereign

{\ND{Lord}}
says: How much
worse will it be when
I send
my four
terrible
judgments
– sword,
famine,
wild animals,
and plague
– to
Jerusalem
to kill
both people
and animals!
\VS{22}Yet some
survivors
will be left
in it, sons
and daughters
who will be brought
out.
They will come out
to
you, and when you see
their behavior
and their deeds,
you will be consoled
about the catastrophe
I have
brought
on
Jerusalem
– for everything
I brought
on it.
\VS{23}They will console
you when
you see
their behavior
and their deeds,
because
you will know
that
it was not
without reason
that I have done
everything
which
I have done
in it, declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{15}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
of all
the woody
branches
among the trees
of the forest,
what
happens
to the wood
of the vine?
\VS{3}Can wood
be taken
from
it to make
anything useful? Or
can
anyone make a peg
from
it to hang
things on?
\VS{4}No! It is thrown
in the fire
for fuel;
when the fire
has burned up
both
ends
of it and it is charred
in the middle,
will it be useful
for anything?
\VS{5}Indeed! If it was
not
made
into anything useful
when it was whole,
how much less
can it be made
into anything
when the fire
has burned
it up
and it is charred?
\par }{\PP \VS{6}“Therefore,
this is what
the sovereign

{\ND{Lord}}
says: Like
the wood
of the vine
is among the trees
of the forest
which
I
have provided
as fuel
for the fire
– so
I will provide
the
residents
of Jerusalem as fuel.
\VS{7}I will set
my face against them – although they have escaped from the fire, the fire will still consume them! Then you will know that I am the
{\ND{Lord}}, when I set my face against them.
\VS{8}I will make
the land
desolate
because
they have acted unfaithfully,
declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{16}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
confront
Jerusalem
with her abominable practices
\VS{3}and say,
‘This is what
the sovereign

{\ND{Lord}}
says to Jerusalem: Your origin
and your birth
were in the land
of the Canaanites;
your father
was an Amorite
and your mother
a Hittite.
\VS{4}As for your birth,
on the day
you were born
your umbilical cord
was not
cut,
nor
were you washed
in water;
you were certainly not
rubbed down
with salt,
nor
wrapped
with blankets.
\VS{5}No
eye
took pity
on
you to do
even one
of these
things for you to spare
you; you were thrown out
into
the open
field
because you were detested
on the day
you were born.
\par }{\PP \VS{6}“‘I passed
by you and saw
you kicking around helplessly
in your blood.
I said
to you as you lay there in your blood,
“Live!” I said
to you as you lay there in your blood,
“Live!”
\VS{7}I made
you plentiful
like sprouts
in a field;
you grew
tall
and came
of age so that
you could wear jewelry.
Your breasts
had formed
and your hair
had grown,
but you
were still naked
and bare.
\par }{\PP \VS{8}“‘Then I passed
by you and watched
you, noticing that
you had reached the age for love. I spread
my cloak
over
you and covered
your nakedness.
I swore a solemn
oath to you and entered
into a marriage covenant
with
you, declares
the sovereign

{\ND{Lord}}, and you became mine.
\par }{\PP \VS{9}“‘Then I bathed
you in water,
washed
the blood
off you, and anointed you with fragrant
oil.
\VS{10}I dressed
you in embroidered clothing
and put fine leather
sandals
on your feet. I wrapped
you with fine linen
and covered
you with silk.
\VS{11}I adorned
you with jewelry.
I put
bracelets
on
your hands
and a necklace around your neck.
\VS{12}I put
a ring
in
your nose,
earrings
on
your ears,
and a beautiful
crown
on your head.
\VS{13}You were adorned
with gold
and silver,
while your clothing
was of fine linen,
silk,
and embroidery.
You ate
the finest flour,
honey,
and olive oil.
You became extremely
beautiful
and attained the position
of royalty.
\VS{14}Your fame
spread among the nations
because
of your beauty; your beauty
was perfect
because
of the splendor
which
I bestowed
on
you, declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{15}“‘But you trusted
in your beauty
and capitalized
on
your fame
by becoming a prostitute.
You offered your sexual favors
to
every
man who passed by so that your beauty became his.
\VS{16}You took
some of your clothing
and made
for yourself decorated
high places;
you engaged
in prostitution
on
them. You went
to him to become his.
\VS{17}You also took
your beautiful
jewelry,
made of my gold
and my silver
I had given
to you, and made
for yourself male
images
and engaged in prostitution with them.
\VS{18}You took
your embroidered
clothing
and used it to cover
them; you offered my olive oil
and my incense
to them.
\VS{19}As for my food
that
I gave you – the fine flour, olive oil, and honey I fed you – you placed it before them as a soothing aroma. That is exactly what happened, declares the sovereign
{\ND{Lord}}.
\par }{\PP \VS{20}“‘You took
your sons
and your daughters
whom
you bore
to me and you sacrificed
them as food for the idols to eat.
As if your prostitution
not enough,
\VS{21}you slaughtered
my children
and sacrificed them to the idols.
\VS{22}And with all
your abominable practices
and prostitution
you did not
remember
the days
of your youth
when
you were naked
and bare,
kicking around
in your blood.
\par }{\PP \VS{23}“‘After
all
of your evil
– “Woe! Woe
to you!” declares
the sovereign

{\ND{Lord}} –
\VS{24}you built
yourself a chamber
and put up a pavilion
in every
public square.
\VS{25}At
the head
of every
street
you erected
your pavilion
and you disgraced
your beauty
when you spread
your legs
to every
passerby
and multiplied
your promiscuity.
\VS{26}You engaged in prostitution
with
the Egyptians,
your sexually aroused neighbors,
multiplying
your promiscuity
and provoking
me to anger.
\VS{27}So see here,
I have stretched
out my hand
against
you and cut off your rations.
I have delivered
you into the power of those
who hate
you, the daughters
of the Philistines,
who were ashamed
by
your obscene
conduct.
\VS{28}You engaged in prostitution
with
the Assyrians
because your sexual desires were insatiable;
you prostituted
yourself with them and yet
you were still not
satisfied.
\VS{29}Then you multiplied
your promiscuity
to
the land
of merchants,
Babylonia,
but you were not
satisfied
there
either.
\par }{\PP \VS{30}“‘How
sick
is your heart,
declares
the sovereign

{\ND{Lord}}, when you perform
all
of these
acts, the deeds
of a bold
prostitute.
\VS{31}When you built
your chamber
at the head
of every
street
and put up your pavilion
in every
public square,
you were not
like a prostitute,
because you scoffed
at payment.
\par }{\PP \VS{32}“‘Adulterous
wife,
who prefers
strangers
instead
of her
own husband!
\VS{33}All
prostitutes
receive
payment,
but instead you
give
gifts
to every one
of your lovers.
You bribe
them to come
to
you from all around
for your sexual favors!
\VS{34}You were
different
from
other prostitutes
because no one
solicited
you. When you gave
payment
and no
payment
was given
to you, you became
the opposite!
\par }{\PP \VS{35}“‘Therefore
O prostitute,
hear
the word
of the {\ND{Lord}}:
\VS{36}This is what
the sovereign

{\ND{Lord}}
says: Because
your lust
was poured out
and your nakedness
was uncovered
in your prostitution
with
your lovers,
and because of all
your detestable
idols,
and because of the blood
of your children
you have given to them,
\VS{37}therefore,
take note: I am about
to gather
all
your lovers
whom
you enjoyed,
both all
those
you loved
and all
those you hated.
I will gather
them against
you from all around,
and I will expose
your nakedness
to
them, and they will see
all
your nakedness.
\VS{38}I will punish you as an adulteress
and murderer
deserves. I will
avenge
your bloody
deeds with furious
rage.
\VS{39}I will give
you into their hands
and they will destroy
your chambers
and tear down
your pavilions.
They will strip
you of your clothing
and take
your beautiful
jewelry
and leave
you naked
and bare.
\VS{40}They will summon a mob
who will stone
you and hack
you in pieces with
their swords.
\VS{41}They will burn down
your houses
and execute
judgments
on you in front of many
women.
Thus I will put
a stop
to your prostitution,
and you will no
longer
give
gifts to your clients.
\VS{42}I will exhaust
my rage
on you, and then my fury
will turn
from
you. I will calm
down and no
longer
be angry.
\par }{\PP \VS{43}“‘Because
you did not
remember
the
days
of your youth
and have enraged
me with all
these
deeds, I
hereby
repay
you for what you have done,
declares
the sovereign

{\ND{Lord}}. Have you not
engaged
in prostitution
on top
of all
your other abominable practices?
\par }{\PP \VS{44}“ ‘Observe – everyone who quotes proverbs will quote this proverb about you: “Like mother, like daughter.”
\VS{45}You are
the daughter
of your mother,
who detested
her husband
and her sons,
and you are
the sister
of your sisters
who
detested
their husbands
and their sons.
Your mother
was a Hittite
and your father
an Amorite.
\VS{46}Your older
sister
was Samaria,
who lived
north
of you with her daughters,
and your younger
sister,
who lived
south
of you, was Sodom
with her daughters.
\VS{47}Have you not
copied
their behavior
and practiced
their abominable deeds? In a short
time you became even more
depraved
in all
your conduct
than they were!
\VS{48}As surely
as I
live, declares
the sovereign

{\ND{Lord}}, your sister
Sodom
and her daughters
never behaved
as
wickedly as you
and your daughters
have
behaved.
\par }{\PP \VS{49}“‘See here – this was the iniquity of your sister Sodom: She and her daughters had majesty, abundance of food, and enjoyed carefree ease, but they did not help the poor and needy.
\VS{50}They were haughty
and practiced
abominable deeds
before
me. Therefore when I saw
it I removed them.
\VS{51}Samaria
has not
committed
half
the sins
you have; you have done more
abominable deeds
than they did. You have made your sisters
appear righteous
with all
the abominable things
you have
done.
\VS{52}So now,
bear
your disgrace,
because you
have
given your sisters
reason
to justify
their behavior. Because
the sins
you have committed were more abominable than
those of your sisters; they have become more
righteous than
you.
So
now, be ashamed
and bear
the disgrace
of making your sisters
appear righteous.
\par }{\PP \VS{53}“‘I will restore
their fortunes,
the fortunes
of Sodom
and her daughters,
and the fortunes
of Samaria
and her daughters
(along with your fortunes
among them),
\VS{54}so that you may
bear
your disgrace
and be ashamed
of all
you have
done
in consoling
them.
\VS{55}As for your sisters,
Sodom
and her daughters
will be restored
to their former status,
Samaria
and her daughters
will be restored
to their former status,
and you
and your daughters
will be restored
to your former status.
\VS{56}In your days
of majesty,
was not
Sodom
your sister
a byword
in your mouth,
\VS{57}before
your evil
was exposed? Now
you have become an object
of scorn
to the daughters
of Aram
and all
those around
her and to the daughters
of the
Philistines
– those all around
you who despise you.
\VS{58}You
must bear
your punishment for your obscene
conduct and your abominable practices,
declares
the {\ND{Lord}}.
\par }{\PP \VS{59}“‘For
this is what
the sovereign

{\ND{Lord}}
says: I will deal
with
you according to what
you have done
when
you despised
your oath
by breaking
your covenant.
\VS{60}Yet I
will remember
the
covenant
I made with
you in the days
of your youth,
and I will establish
a lasting
covenant with you.
\VS{61}Then you will remember
your conduct,
and be ashamed
when you receive
your older
and younger
sisters.
I will give
them
to you as daughters,
but not
on account of my covenant with you.
\VS{62}I
will establish
my covenant
with
you, and then you will know
that
I am
the {\ND{Lord}}.
\VS{63}Then
you will remember,
be ashamed,
and remain
silent
when I make atonement
for all
you have
done,
declares
the sovereign

{\ND{Lord}}.’ ”

\par }\Chap{17}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
offer a riddle,
and tell a parable
to
the house
of Israel.
\VS{3}Say
to them: ‘This is what
the sovereign

{\ND{Lord}}
says:

\par }{\Q “ ‘A great
eagle
with broad
wings,
long
feathers,
\par }{\Q with full
plumage
which
was multi-hued,
\par }{\Q came
to
Lebanon
and took
the top
of the cedar.
\par }{\Q \VS{4}He plucked
off its topmost
shoot;
\par }{\Q he brought
it to
a land
of merchants
\par }{\Q and planted
it in a city
of traders.
\par }{\Q \VS{5}He took
one
of the seedlings
of the land,
\par }{\Q placed
it in a cultivated plot;
\par }{\Q a shoot by abundant
water,
\par }{\Q like
a willow
he planted it.
\par }{\Q \VS{6}It sprouted
and became
a vine,
\par }{\Q spreading
low
to the ground;
\par }{\Q its branches
turning
toward
him, its roots
were under
itself.

\par }{\Q So it became
a vine;
it produced
shoots
and sent out
branches.
\par }{\Q \VS{7}“‘There was
another
great
eagle
\par }{\Q with broad
wings
and thick
plumage.
\par }{\Q Now
this
vine
twisted
its roots
toward him
\par }{\Q and sent
its branches
toward him
\par }{\Q to be watered
from the soil
where it was planted.
\par }{\Q \VS{8}In
a good
field,
by
abundant
waters,
it was
planted
\par }{\Q to grow
branches,
bear
fruit,
and become
a beautiful
vine.
\par }{\PP \VS{9}“‘Say
to them: This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Will it prosper?
\par }{\Q Will he not
rip out
its roots
\par }{\Q and cause its fruit
to rot
and wither?
\par }{\Q All
its foliage
will wither.
\par }{\Q No
strong
arm
or large
army
\par }{\Q will be needed to pull
it out
by its roots.
\par }{\Q \VS{10}Consider! It is planted,
but will it prosper?
\par }{\Q Will it not
wither completely
when the east
wind
blows
on
it?
\par }{\Q Will it not wither
in the soil
where it sprouted?’ ”
\par }{\PP \VS{11}Then
the word
of the {\ND{Lord}}
came to me:
\VS{12}“Say
to the rebellious
house
of Israel: ‘Don’t
you know
what
these
things mean?’ Say: ‘See here,
the king
of Babylon
came
to Jerusalem
and took
her king
and her officials
prisoner and brought
them to
himself in Babylon.
\VS{13}He took
one from the royal
family,
made
a treaty
with
him, and put
him under oath.
He then took
the leaders
of the land
\VS{14}so it would be a lowly
kingdom
which could not
rise
on its own but must keep
its treaty
with him in order to stand.
\VS{15}But this one from Israel’s royal family rebelled
against the king of Babylon by sending
his emissaries
to Egypt
to obtain
horses
and a large
army.
Will he prosper? Will the one
doing
these
things escape? Can he break
the covenant
and escape?
\par }{\PP \VS{16}“‘As surely as I
live,
declares
the sovereign

{\ND{Lord}}, surely
in the city
of the
king
who crowned
him, whose
oath
he despised
and whose
covenant
he broke
– in
the middle
of Babylon
he will die!
\VS{17}Pharaoh
with his great
army
and mighty
horde
will not
help him in battle,
when siege
ramps
are erected and siege-walls
are built
to kill
many people.
\VS{18}He despised
the oath
by breaking
the covenant.
Take note –
he gave
his
promise and did
all
these things – he will not escape!
\par }{\PP \VS{19}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says: As surely
as I
live, I will certainly
repay him for despising
my oath
and breaking
my covenant!
\VS{20}I
will throw
my net
over him and he will be caught
in my snare;
I will bring
him to Babylon
and judge
him
there
because of the unfaithfulness he committed against me.
\VS{21}All
the choice men
among his troops
will die
by the sword
and the survivors
will be scattered
to every
wind.
Then you will know
that
I,
the {\ND{Lord}},
have spoken!
\par }{\PP \VS{22}“‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘I
will take
a sprig from the lofty
top
of the cedar
and plant
it.

\par }{\Q I will pluck
from the top
one of its tender
twigs;
\par }{\Q I myself
will plant
it on
a high
and lofty
mountain.
\par }{\Q \VS{23}I will plant
it on a high
mountain
of Israel,
\par }{\Q and it will raise
branches
and produce
fruit
and become
a beautiful
cedar.
\par }{\Q Every
bird
will live
under
it;
\par }{\Q Every
winged creature
will live
in the shade
of its branches.
\par }{\Q \VS{24}All
the trees
of the field
will know
that
I am
the {\ND{Lord}}.
\par }{\Q I make the high
tree
low;
I raise up the low
tree.
\par }{\Q I make the green
tree
wither,
and I make the dry
tree
sprout.
\par }{\Q I,
the {\ND{Lord}},
have spoken,
and I will do it!’ ”

\par }\Chap{18}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“What
do you
mean by quoting
this
proverb
concerning
the land
of Israel,
\par }{\Q “‘The fathers
eat
sour grapes
\par }{\Q And the children’s
teeth
become numb?’
\par }{\PP \VS{3}“As surely
as I
live, declares
the sovereign

{\ND{Lord}}, you will not
quote
this
proverb
in Israel
anymore!
\VS{4}Indeed! All
lives
are mine
– the life
of the father
as well as the life
of the son
is mine.
The one who
sins
will die.
\par }{\PP \VS{5}“Suppose
a man
is righteous.
He practices
what is just
and right,
\VS{6}does not
eat
pagan
sacrifices on the mountains
or
pray
to the idols
of the house
of Israel,
does not
defile
his neighbor’s
wife,
does not
have sexual
relations with a woman
during her period,
\VS{7}does not
oppress
anyone, but gives the debtor
back
whatever was given in pledge,
does not
commit robbery,
but gives his bread
to the hungry
and clothes
the naked,
\VS{8}does not
engage
in usury
or
charge interest,
but refrains
from wrongdoing,
promotes true
justice
between
men,
\VS{9}and follows
my statutes
and observes
my regulations
by carrying
them out.
That man is righteous;
he will certainly
live,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{10}“Suppose such a man has
a violent
son
who sheds
blood
and does
any of these
things mentioned previously
\VS{11}(though
the father
did not
do
any
of them). He eats
pagan
sacrifices on the mountains,
defiles
his neighbor’s
wife,
\VS{12}oppresses
the poor
and the needy,
commits
robbery,
does not
give back
what was given in pledge,
prays to
idols,
performs
abominable
acts,
\VS{13}engages
in usury
and charges
interest.
Will he live? He will not! Because he has done
all
these
abominable deeds
he will certainly die.
He will bear
the responsibility
for his own death.
\par }{\PP \VS{14}“But suppose
he in turn has
a son
who notices
all
the sins
his father
commits,
considers
them, and does not
follow his father’s example.
\VS{15}He does not
eat
pagan
sacrifices on
the mountains,
does not
pray
to the idols
of the house
of Israel,
does not
defile
his neighbor’s
wife,
\VS{16}does not
oppress
anyone or
keep what has
been given in pledge,
does not
commit robbery, gives
his food
to the hungry,
and clothes
the naked,
\VS{17}refrains
from wrongdoing, does
not engage
in usury
or charge interest,
carries
out my regulations
and follows
my statutes.
He will not
die
for his father’s
iniquity;
he will surely
live.
\VS{18}As for his father,
because
he practices
extortion,
robs
his brother,
and does what is not
good
among
his people,
he will die
for his iniquity.
\par }{\PP \VS{19}“Yet you say,
‘Why
should the son
not
suffer
for his father’s
iniquity?’ When the son
does what is just
and right,
and observes
all
my statutes
and carries
them out, he will surely
live.
\VS{20}The person
who sins
is
the one
who will die.
A son
will not
suffer
for his father’s
iniquity,
and a father
will not
suffer
for his son’s
iniquity;
the righteous
person will be
judged according to his righteousness,
and the wicked
person according to his wickedness.
\par }{\PP \VS{21}“But if
the wicked
person turns from
all
the sin
he has
committed
and observes
all
my statutes
and does
what is just
and right,
he will surely
live;
he will not
die.
\VS{22}None
of the sins
he has committed
will be held against
him; because of the righteousness
he has
done,
he will live.
\VS{23}Do I actually
delight
in the death
of the wicked,
declares
the sovereign

{\ND{Lord}}? Do I not
prefer that he turn
from his wicked conduct
and live?
\par }{\PP \VS{24}“But if a righteous
man turns away
from his righteousness
and practices
wrongdoing
according to all
the abominable
practices
the wicked
carry out,
will he live? All
his righteous
acts
will not
be remembered;
because
of the unfaithful acts
he has done
and the sin
he has committed,
he will die.
\par }{\PP \VS{25}“Yet you say,
‘The Lord’s
conduct
is unjust!’ Hear,
O house
of Israel: Is my conduct
unjust? Is it not
your conduct
that is unjust?
\VS{26}When a righteous
person turns back from his righteousness
and practices
wrongdoing,
he will die
for it;
because of the wrongdoing
he has
done,
he will die.
\VS{27}When
a wicked
person turns
from the wickedness
he has
committed
and does
what is just
and right,
he
will preserve
his life.
\VS{28}Because he considered
and turned
from all
the sins
he had
done,
he will surely
live;
he will not
die.
\VS{29}Yet the house
of Israel
says,
‘The Lord’s
conduct
is unjust!’ Is my conduct
unjust,
O house
of Israel? Is it not
your conduct
that is unjust?
\par }{\PP \VS{30}“Therefore
I will judge
each
person according to his conduct,
O house
of Israel,
declares
the sovereign

{\ND{Lord}}. Repent
and turn
from all
your wickedness;
then it will not
be
an obstacle
leading to iniquity.
\VS{31}Throw
away all
your sins
you have committed and fashion
yourselves a new
heart
and a new
spirit! Why
should you die,
O house
of Israel?
\VS{32}For
I take no
delight
in the death
of anyone, declares
the sovereign

{\ND{Lord}}. Repent
and live!

\par }\Chap{19}{\PP \VerseOne{1}“And you,
sing
a lament
for the princes
of Israel,
\VS{2}and say:
\par }{\Q “ ‘What
a lioness
was your mother
among
the lions!
\par }{\Q She lay
among
young lions;
she reared
her cubs.
\par }{\Q \VS{3}She reared
one
of her cubs;
he became a young lion.
\par }{\Q He learned
to tear
prey;
he devoured
people.
\par }{\Q \VS{4}The nations
heard
about him;
he was trapped
in their pit.
\par }{\Q They brought
him
with hooks
to
the land
of Egypt.
\par }{\Q \VS{5}“‘When she realized
that
she waited
in vain, her hope
was lost.
\par }{\Q She took
another
of her cubs
and made
him a young lion.
\par }{\Q \VS{6}He walked
about among
the lions;
he became
a young lion.
\par }{\Q He learned
to tear
prey;
he devoured
people.
\par }{\Q \VS{7}He broke down
their strongholds
and devastated
their cities.
\par }{\Q The land
and everything
in it was frightened
at the sound
of his roaring.
\par }{\Q \VS{8}The nations
– the surrounding
regions
– attacked him.
\par }{\Q They threw their net
over him; he was caught
in their pit.
\par }{\Q \VS{9}They put
him in a collar
with hooks;
\par }{\Q they brought
him to
the king
of Babylon;
\par }{\Q they brought
him to prison
\par }{\Q so
that his voice
would not
be heard
\par }{\Q any longer
on
the mountains
of Israel.
\par }{\Q \VS{10}“‘Your mother
was like a vine
in your vineyard,
planted
by water.
\par }{\Q It was fruitful
and full of branches because it was well-watered.
\par }{\Q \VS{11}Its boughs
were
strong,
fit for
rulers’
scepters;
it reached up
into
the clouds.
\par }{\Q It stood out
because of its height
and its many
branches.
\par }{\Q \VS{12}But
it was plucked
up in anger;
it was thrown
down to the ground.
\par }{\Q The east
wind
dried
up its fruit;
\par }{\Q its strong
branches broke off
and withered
–
\par }{\Q a fire
consumed them.
\par }{\Q \VS{13}Now
it is planted
in the wilderness,
\par }{\Q in a dry
and thirsty
land.
\par }{\Q \VS{14}A fire
has gone out
from its branch;
it has consumed
its shoot and its fruit.
\par }{\Q No
strong
branch
was left in it, nor a scepter
to rule.’
\par }{\PP This is a lament
song, and has become
a lament song.”

\par }\Chap{20}{\PP \VerseOne{1}In the seventh
year,
in the fifth
month, on the tenth
of the month,
some
of the elders
of Israel
came
to seek
the

{\ND{Lord}}, and they sat
down in front of me.
\VS{2}The word
of the {\ND{Lord}}
came to me:
\VS{3}“Son
of man,
speak
to the
elders
of Israel,
and tell
them: ‘This is what
the sovereign

{\ND{Lord}}
says: Are you
coming
to seek
me? As surely
as I
live, I will not
allow you to seek
me, declares
the sovereign

{\ND{Lord}}.’
\VS{4}“Are you willing to pronounce
judgment? Are you willing to pronounce judgment,
son
of man? Then confront
them with the abominable practices
of their fathers,
\VS{5}and say
to them:
\par }{\PP “ ‘This is what
the sovereign
{\SC{
Lord}}
says: On the day
I chose
Israel
I swore
to the descendants
of the house
of Jacob
and made myself known
to them in the land
of Egypt.
I swore
to them, “I am
the
{\ND{Lord}}
your God.”
\VS{6}On that day
I swore
to bring them out
of the land
of Egypt
to
a land
which
I had picked
out for them, a land flowing
with milk
and honey,
the most beautiful
of all
lands.
\VS{7}I said
to
them, “Each
of you must get rid of
the detestable idols
you keep before you,
and do not
defile
yourselves with the idols
of Egypt;
I
am the
{\ND{Lord}}
your God.”
\VS{8}But they rebelled
against me, and refused
to listen
to
me; no one
got rid of
their detestable
idols,
nor
did they abandon
the idols
of Egypt.
Then I decided to pour out
my rage
on
them and fully vent
my anger
against them in the midst
of the land
of Egypt.
\VS{9}I acted
for the sake
of my reputation,
so that I would not
be profaned
before
the nations
among
whom
they
lived,
before
whom
I revealed
myself by bringing
them
out
of the land
of Egypt.
\par }{\PP \VS{10}“‘So I brought
them out
of the land
of Egypt
and led
them to
the wilderness.
\VS{11}I gave
them my statutes
and revealed
my regulations
to them. The one
who
carries
them out will live by them!
\VS{12}I also
gave
them my Sabbaths
as a reminder
of our relationship,
so that they would know
that
I,
the {\ND{Lord}},
sanctify them.
\VS{13}But the house
of Israel
rebelled
against me in the wilderness;
they did not
follow
my statutes
and they rejected
my regulations
(the one
who obeys
them will live
by them), and they utterly
desecrated
my Sabbaths.
So I decided
to pour out
my rage
on
them in the wilderness
and destroy them.
\VS{14}I acted
for the sake
of my reputation,
so that I would not
be profaned
before
the nations
in whose
sight
I had brought them out.
\VS{15}I
also
swore
to them in the wilderness
that I would not
bring
them to
the land
I had
given them – a land flowing with milk and honey, the most beautiful of all lands.
\VS{16}I did this because
they rejected
my regulations,
did not
follow
my statutes,
and desecrated
my Sabbaths;
for
their
hearts
followed
their idols.
\VS{17}Yet I had pity
on
them and did not destroy
them, so I did not
make
an end
of them in the wilderness.
\par }{\PP \VS{18}“‘But I said
to
their children
in the wilderness,
“Do not
follow
the practices
of your fathers;
do
not
observe
their regulations,
nor defile
yourselves with their idols.
\VS{19}I am
the {\ND{Lord}}
your God;
follow
my statutes,
observe
my regulations,
and carry
them out.
\VS{20}Treat my Sabbaths
as holy
and they will be
a reminder
of our relationship,
and then you will know
that
I am
the {\ND{Lord}}
your God.”
\VS{21}“ ‘But the children
rebelled
against me, did not
follow
my statutes,
did not
observe
my regulations
by carrying
them out
(the one
who
obeys
them will live
by them), and desecrated
my Sabbaths.
I decided
to pour out
my rage
on
them and fully vent
my anger
against them in the wilderness.
\VS{22}But I refrained from
doing
so, and acted
instead for the sake
of my reputation,
so that I would not
be profaned
before
the nations
in whose
sight
I had brought them out.
\VS{23}I
also
swore
to them in the wilderness
that I would scatter
them among the nations
and disperse
them throughout the lands.
\VS{24}I did this because
they did not
observe
my regulations,
they rejected
my statutes,
they desecrated
my Sabbaths,
and their eyes
were fixed
on their fathers’
idols.
\VS{25}I
also
gave
them decrees
which were not
good
and regulations
by which they could not
live.
\VS{26}I declared them to be defiled
because of their sacrifices –
they caused
all
their first born
to pass through the fire – so that I would devastate them, so that they will know that I am the
{\ND{Lord}}.’
\par }{\PP \VS{27}“Therefore,
speak
to
the house
of Israel,
son
of man,
and tell
them, ‘This is what
the sovereign

{\ND{Lord}}
says: In this
way too
your fathers
blasphemed
me when they were unfaithful to me.
\VS{28}I brought
them to
the land
which
I swore
to give
them, but whenever they saw
any
high
hill
or leafy
tree,
they offered their sacrifices there
and presented
the
offerings that provoke
me to
anger. They
offered their soothing
aroma
there
and poured
out their drink offerings.
\VS{29}So I said
to
them, What
is this high place
you
go to?’ ” (So
it is called
“High Place”
to
this
day.)
\par }{\PP \VS{30}“Therefore
say
to
the house
of Israel,
‘This is what
the sovereign

{\ND{Lord}}
says: Will you
defile
yourselves
like your fathers
and engage in prostitution
with detestable idols?
\VS{31}When you present
your sacrifices –
when you make your sons
pass
through the fire
– you
defile
yourselves with all
your idols
to
this very
day.
Will I
allow you to seek
me, O house
of Israel? As surely
as I
live, declares
the sovereign

{\ND{Lord}}, I will not
allow you to seek me!
\par }{\PP \VS{32}“‘What
you plan
will never
happen.
You
say,
“We will be
like the nations,
like the clans
of the lands,
who serve
gods of wood
and stone.”
\VS{33}As surely
as I
live, declares
the sovereign

{\ND{Lord}}, with a powerful
hand
and an outstretched
arm,
and with an outpouring
of rage,
I will be king over you.
\VS{34}I will bring you out
from
the nations,
and will gather
you from
the lands
where
you are scattered,
with
a powerful
hand
and an outstretched
arm
and with
an outpouring
of rage!
\VS{35}I will bring
you into
the wilderness
of the nations,
and there
I will enter into judgment
with
you face
to
face.
\VS{36}Just
as I entered into judgment
with
your fathers
in the wilderness
of the land
of Egypt,
so
I will enter into judgment
with
you, declares
the sovereign

{\ND{Lord}}.
\VS{37}I will make you pass
under
the shepherd’s staff,
and I will bring
you into the bond
of the covenant.
\VS{38}I will eliminate
from
among
you the rebels
and those who revolt against me. I will bring
them out from
the land
where they have been residing,
but they will not
come
to
the land
of Israel.
Then you will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{39}“‘As for you,
O house
of Israel,
this is what
the sovereign

{\ND{Lord}}
says: Each
of you go
and serve
your idols,
if
you will not
listen
to me.
But my holy
name
will not
be profaned
again
by your sacrifices
and your idols.
\VS{40}For there
on my holy
mountain,
the high
mountain
of Israel,
declares
the sovereign

{\ND{Lord}}, all
the house
of Israel
will serve
me, all
of them in the land.
I will accept
them there,
and there
I will seek
your contributions
and your choice
gifts,
with all
your holy things.
\VS{41}When I bring you out
from
the nations
and gather
you from
the lands
where
you are scattered,
I will accept
you along with your soothing
aroma.
I will display my holiness among
you in the sight
of the nations.
\VS{42}Then you will know
that
I am
the {\ND{Lord}}
when I bring
you to
the land
of Israel,
to
the land
I swore
to give
to your fathers.
\VS{43}And there
you will remember
your conduct
and all
your deeds
by which
you defiled
yourselves. You will despise
yourselves because of all
the evil
deeds you have
done.
\VS{44}Then you will know
that
I am
the {\ND{Lord}}, when I deal
with
you for the sake
of my reputation
and not
according to your wicked
conduct
and corrupt
deeds,
O house
of Israel,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\SH Prophecy Against the South
\par }{\PP \VS{45} The word
of the {\ND{Lord}}
came to me:
\VS{46}“Son
of man,
turn
toward
the south,
and speak
out against the south.
Prophesy
against the open scrub land
of the Negev,
\VS{47}and say
to the scrub land
of the Negev,
‘Hear
the word
of the {\ND{Lord}}: This is what the sovereign

{\ND{Lord}}
says: Look
here, I am about to start a fire
in you,
and it will devour
every
green
tree
and every
dry
tree
in you. The flaming
fire
will not
be extinguished,
and the whole
surface
of the ground from the Negev
to the north
will be scorched by it.
\VS{48}And everyone
will see
that
I,
the {\ND{Lord}},
have burned
it; it will not
be extinguished.’ ”
\par }{\PP \VS{49}Then I said,
“O
sovereign

{\ND{Lord}}! They
are saying
of me, ‘Does he not
simply speak
in eloquent figures of speech?’ ”

\par }\Chap{21}{\PP \VerseOne{1} The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
Jerusalem
and speak
out against the sanctuaries.
Prophesy
against the land
of Israel
\VS{3}and say
to them, ‘This is what
the {\ND{Lord}}
says: Look,
I am against
you. I will draw
my sword
from its sheath
and cut off
from
you both the righteous
and the wicked.
\VS{4}Because
I will cut off
from
you both the righteous
and the wicked,
my sword
will go out
from its sheath
against
everyone
from the south
to the north.
\VS{5}Then everyone
will know
that
I
am the
{\ND{Lord}},
who drew
my sword
from its sheath
– it will not
be sheathed
again!’
\par }{\PP \VS{6}“And you,
son
of man,
groan
with an aching
heart
and bitterness;
groan
before their eyes.
\VS{7}When
they ask
you, ‘Why
are you
groaning?’ you will reply,
‘Because of the report
that
has come.
Every
heart
will melt
with fear
and every
hand
will be limp;
everyone
will faint
and every
knee
will be wet with urine.’ Pay attention – it is coming and it will happen, declares the sovereign
{\ND{Lord}}.”
\par }{\PP \VS{8}The word
of the {\ND{Lord}}
came to me:
\VS{9}“Son
of man,
prophesy
and say: ‘This is what
the Lord
says:
\par }{\Q “ ‘A sword,
a sword
is sharpened,
\par }{\Q and also
polished.
\par }{\Q \VS{10}It is sharpened
for
slaughter,
\par }{\Q it is polished
to
flash like lightning!
\par }{\PP “ ‘Should we rejoice
in the scepter
of my son? No! The sword despises
every
tree!
\par }{\Q \VS{11}“‘He gave
it to be polished,
\par }{\Q to be grasped
in the hand
–
\par }{\Q the sword
is sharpened,
it is
polished
–
\par }{\Q giving
it into the hand
of the executioner.
\par }{\Q \VS{12}Cry
out and moan,
son
of man,
\par }{\Q for
it is
wielded
against my people;
\par }{\Q against all
the princes
of Israel.
\par }{\Q They are delivered up to
the sword,
along with
my people.
\par }{\Q Therefore,
strike
your thigh.
\par }{\PP \VS{13}“‘For
testing
will come, and what
will happen
when the scepter,
which the sword despises,
is no
more? declares
the sovereign

{\ND{Lord}}.’
\par }{\Q \VS{14}“And you,
son
of man,
prophesy,
\par }{\Q and clap your hands
together.
\par }{\Q Let
the sword
strike
twice,
even three times!
\par }{\Q It is
a sword
for slaughter,
\par }{\Q a sword
for the great
slaughter
surrounding them.
\par }{\Q \VS{15}So
hearts
melt
with fear and many
stumble.
\par }{\Q At
all
their gates
I have stationed
the sword
for slaughter.
\par }{\Q Ah! It is made
to flash,
it is drawn
for slaughter!
\par }{\Q \VS{16}Cut sharply
on the right!
\par }{\Q Swing
to the left,
\par }{\Q wherever your edge
is appointed to strike.
\par }{\Q \VS{17}I
too
will clap my hands
together,
\par }{\Q I will exhaust
my rage;
\par }{\Q I
the {\ND{Lord}}
have spoken.”
\par }{\PP \VS{18}The word
of the {\ND{Lord}}
came to me:
\VS{19}“You,
son
of man,
mark out
two
routes
for the king
of Babylon’s
sword
to take;
both
of them will originate
in a single
land.
Make
a signpost
and put it at the beginning
of the road
leading to the city.
\VS{20}Mark out
the routes
for the sword
to take: “Rabbah
of the Ammonites”
and “Judah
with Jerusalem
in it.”
\VS{21}For
the king
of Babylon
stands
at
the fork
in the road
at the head
of the two
routes.
He looks for
omens: He shakes
arrows,
he consults
idols,
he examines
animal livers.
\VS{22}Into his right hand
comes
the portent
for Jerusalem
– to set up
battering rams,
to give
the signal
for slaughter,
to shout out
the battle cry,
to set up
battering rams
against
the gates,
to erect a siege
ramp,
to build
a siege wall.
\VS{23}But
those
in Jerusalem
will view
it as a false omen.
They have sworn
solemn
oaths, but the king
of Babylon
will accuse
them of violations
in order to seize them.
\par }{\PP \VS{24}“Therefore
this is what
the sovereign

{\ND{Lord}}
says: ‘Because
you have brought up
your own guilt
by uncovering
your transgressions
and revealing
your sins
through all
your actions,
for this
reason
you will be taken
by force.
\par }{\Q \VS{25}“‘As for you,
profane
and wicked
prince
of Israel,
\par }{\Q whose
day
has come,
the time
of final
punishment,
\par }{\Q \VS{26}this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q Tear off
the turban,
\par }{\Q take off the crown!
\par }{\Q Things must
change!

\par }{\Q Exalt
the lowly,
\par }{\Q bring down
the proud!
\par }{\Q \VS{27}A total
ruin
I will make
it!

\par }{\Q It will come
to an end
\par }{\Q when the one arrives
to whom
I have assigned
judgment.’
\par }{\PP \VS{28}“As for you,
son
of man,
prophesy
and say,
‘This is what
the sovereign

{\ND{Lord}}
says
concerning the Ammonites
and their coming humiliation;
say:
\par }{\Q “ ‘A sword,
a sword
drawn
for slaughter,
\par }{\Q polished
to consume,
to flash
like lightning –
\par }{\Q \VS{29}while seeing
false visions
for you
\par }{\Q and reading
lying
omens for you –
\par }{\Q to place that sword on the necks of the profane wicked,
\par }{\Q whose day has come,
\par }{\Q the time of final punishment.
\par }{\Q \VS{30}Return
it to
its sheath!

\par }{\Q In the place
where
you were created,
\par }{\Q in your native
land,
I will judge you.
\par }{\Q \VS{31}I will pour
out my anger
on
you;
\par }{\Q the fire
of my fury
I will blow
on
you.
\par }{\Q I will
hand
you over to brutal men,
\par }{\Q who are skilled
in destruction.
\par }{\Q \VS{32}You will become
fuel
for the fire
–
\par }{\Q your blood
will stain the middle
of the land;
\par }{\Q you will no
longer be remembered,
\par }{\Q for
I,
the {\ND{Lord}}, have spoken.’ ”

\par }\Chap{22}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“As for you,
son
of man,
are you willing to pronounce judgment, are you willing to pronounce
judgment
on the bloody
city? Then confront
her with all
her abominable deeds!
\VS{3}Then say,
‘This is what
the sovereign
{\SC{
Lord}}
says: O city,
who spills
blood
within
herself (which brings
on her doom), and who makes
herself idols
(which
results in impurity),
\VS{4}you are guilty
because of the blood
you shed
and defiled
by the idols
you
made.
You have hastened
the day
of your doom;
the end
of your years
has come.
Therefore
I will make
you an object of scorn
to the nations,
an object to be mocked
by all
lands.
\VS{5}Those both near
and far
from
you will mock
you, you with your bad
reputation,
full
of turmoil.
\par }{\PP \VS{6}“‘See
how each
of the princes
of Israel
living within
you has used his
authority
to shed
blood.
\VS{7}They have treated father
and mother
with contempt
within you; they have oppressed the foreigner
among
you; they have wronged
the orphan
and the widow
within you.
\VS{8}You have despised
my holy
things and desecrated
my Sabbaths!
\VS{9}Slanderous
men
shed
blood
within you. Those who live within you eat
pagan sacrifices on the mountains;
they commit obscene
acts
among you.
\VS{10}They have sex
with their father’s
wife within you; they violate
women during their menstrual period within you.
\VS{11}One commits
an abominable act
with
his neighbor’s
wife;
another obscenely
defiles
his daughter-in-law;
another violates
his sister
– his father’s
daughter –
within you.
\VS{12}They take
bribes
within you to shed
blood.
You engage in usury
and charge
interest;
you extort
money from your neighbors.
You have forgotten
me, declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{13}“‘See,
I strike
my hands
together at the dishonest profit
you have
made,
and at the bloodshed
they have
done among you.
\VS{14}Can your heart
endure,
or
can your hands
be strong
when
I
deal
with
you? I,
the {\ND{Lord}},
have spoken,
and I will do it!
\VS{15}I will scatter
you among the nations
and disperse
you among various countries;
I will remove your impurity
from you.
\VS{16}You will be profaned
within yourself in the sight
of the nations;
then you will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\PP \VS{17}The word
of the {\ND{Lord}}
came to me:
\VS{18}“Son
of man,
the house
of Israel
has become slag
to me. All
of them are like bronze,
tin,
iron,
and lead
in
the furnace;
they are the worthless slag
of silver.
\VS{19}Therefore
this is what
the sovereign

{\ND{Lord}}
says: ‘Because
all
of you have become
slag,
look out! – I am about to gather you in the middle of Jerusalem.
\VS{20}As silver,
bronze,
iron,
lead,
and tin
are gathered
in
a furnace
so that the fire
can melt
them, so
I will gather
you in my anger
and in my rage.
I will deposit
you there and melt you.
\VS{21}I will gather
you and blow
on
you with the fire
of my fury,
and you will be melted
in it.
\VS{22}As silver
is melted
in
a furnace,
so
you will be melted
in
it, and you will know
that
I,
the {\ND{Lord}},
have poured out
my anger
on you.’ ”
\par }{\PP \VS{23}The word
of the {\ND{Lord}}
came to me:
\VS{24}“Son
of man,
say
to her: ‘You
are a land
that receives no
rain or showers
in the day
of my anger.’
\VS{25}Her princes
within
her are like a roaring
lion
tearing
its prey;
they have devoured
lives.
They take
away riches
and valuable things;
they have made many
women widows
within it.
\VS{26}Her priests
abuse
my law
and have desecrated
my holy
things. They do not
distinguish
between
the holy
and the profane,
or
recognize
any distinction between
the unclean
and the clean.
They ignore
my Sabbaths
and I am profaned
in their midst.
\VS{27}Her officials
are like wolves
in her midst
rending
their prey
– shedding
blood
and destroying
lives
– so
they can get dishonest profit.
\VS{28}Her
prophets
coat
their messages with whitewash.
They see
false
visions
and announce lying omens
for them, saying,
‘This is what
the sovereign

{\ND{Lord}}
says,’ when the
{\ND{Lord}}
has not
spoken.
\VS{29}The people
of the land
have practiced extortion
and committed robbery.
They have wronged
the poor
and needy;
they have oppressed
the foreigner
who lives among them and denied
them justice.
\par }{\PP \VS{30}“I looked
for a man
from among them who would repair
the wall
and stand
in the gap
before
me on behalf
of the land,
so that I would not
destroy
it, but I found no one.
\VS{31}So I have poured
my anger
on
them,
and destroyed
them with the fire
of my fury.
I hereby
repay
them for what they have done, declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{23}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
there were two
women
who were daughters
of the same
mother.
\VS{3}They engaged in prostitution
in Egypt;
in their youth
they engaged in prostitution.
Their breasts
were squeezed there;
lovers fondled their virgin
nipples
there.
\VS{4}Oholah
was the name
of the older
and Oholibah
the name
of her younger sister.
They became
mine, and gave birth
to sons
and daughters.
Oholah
is Samaria
and Oholibah
is Jerusalem.
\par }{\PP \VS{5}“Oholah
engaged in prostitution
while she was mine.
She lusted
after her lovers,
the Assyrians –
warriors
\VS{6}clothed
in blue,
governors
and officials,
all
of them desirable
young men,
horsemen
riding
on horses.
\VS{7}She bestowed
her sexual favors
on
them; all
of them were the choicest
young men
of Assyria.
She defiled
herself with all
whom
she desired –
with all
their idols.
\VS{8}She did not
abandon
the prostitution
she had practiced in Egypt;
for
in
her youth men
had sex
with her, fondled
her virgin
breasts,
and ravished her.
\VS{9}Therefore
I handed
her over
to her lovers,
the Assyrians
for
whom
she lusted.
\VS{10}They
exposed
her nakedness,
seized
her sons
and daughters,
and killed
her
with the sword.
She became
notorious
among women,
and they executed
judgments against her.
\par }{\PP \VS{11}“Her sister
Oholibah
watched
this, but she became more corrupt
in her lust
than
her sister had been, and her acts of prostitution
were more numerous than those of her sister.
\VS{12}She lusted
after
the Assyrians
– governors
and officials,
warriors in full
armor,
horsemen
riding
on horses,
all
of them desirable
young men.
\VS{13}I saw
that
she was defiled;
both
of them followed the same
path.
\VS{14}But she increased
her prostitution.
She saw
men
carved
on
the wall,
images
of the Chaldeans
carved
in bright red,
\VS{15}wearing
belts
on their waists
and flowing
turbans
on their heads,
all
of them looking like
officers,
the image
of Babylonians
whose native
land
is Chaldea.
\VS{16}When she saw them, she lusted
after them and sent
messengers
to them
in Chaldea.
\VS{17}The Babylonians
crawled into
bed
with her. They defiled
her with their lust; after
she was defiled
by them, she became disgusted
with them.
\VS{18}When she lustfully
exposed
her nakedness,
I was disgusted
with her, just
as I
had been disgusted
with her sister.
\VS{19}Yet she increased
her prostitution,
remembering
the days
of her youth
when she engaged in prostitution
in the land
of Egypt.
\VS{20}She lusted
after
their genitals
– as large as those of donkeys,
and their seminal emission
was as strong
as that of stallions.
\VS{21}This is how you assessed
the obscene
conduct of your youth,
when the Egyptians
fondled your nipples
and squeezed your young
breasts.
\par }{\PP \VS{22}“Therefore,
Oholibah,
this is what
the sovereign

{\ND{Lord}}
says: Look
here, I am about to stir
up against
you the
lovers
with whom
you were disgusted;
I will bring
them against
you from every side:
\VS{23}the Babylonians
and all
the Chaldeans,
Pekod,
Shoa,
and Koa,
and all
the Assyrians
with
them, desirable
young men,
all
of them governors
and officials,
officers
and nobles,
all
of them riding
on horses.
\VS{24}They will attack
you with weapons,
chariots,
wagons,
and with a huge army;
they will array
themselves against
you on every side
with large shields,
small shields,
and helmets.
I will assign
them the task
of judgment;
they will punish
you according to their laws.
\VS{25}I will direct
my jealous
anger against you, and they will deal
with
you in rage.
They will cut off
your nose
and your ears,
and your survivors
will die
by the sword.
They
will seize
your sons
and daughters,
and your survivors
will be consumed
by fire.
\VS{26}They will strip
your clothes
off you and take
away your beautiful
jewelry.
\VS{27}So I will put an end
to your obscene
conduct
and your prostitution
which you have practiced
in the land
of Egypt.
You will not
seek their help or remember
Egypt
anymore.
\par }{\PP \VS{28}“For
this is what
the sovereign

{\ND{Lord}}
says: Look
here, I am about to deliver
you over
to those whom
you hate,
to those with whom
you were disgusted.
\VS{29}They will treat
you with
hatred,
take away
all
you have labored
for, and leave
you naked
and bare.
Your nakedness
will be exposed,
just as when you engaged in prostitution
and obscene
conduct.
\VS{30}I will do
these
things to you because you engaged in prostitution
with the nations,
polluting
yourself with their idols.
\VS{31}You have followed
the ways
of your sister,
so
I will place
her cup
of judgment in your hand.
\VS{32}“This is what
the sovereign

{\ND{Lord}}
says: “You will drink
your sister’s
deep
and wide
cup;
you will be
scorned
and derided,
for it holds
a great deal.
\VS{33}You will be overcome
by drunkenness
and sorrow.
The cup
of your sister
Samaria
is a cup
of horror
and desolation.
\VS{34}You will drain
it dry, gnaw
its pieces, and tear out your breasts,
for
I
have spoken,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{35}“Therefore
this is what
the sovereign

{\ND{Lord}}
says: Because
you have forgotten
me and completely disregarded
me, you
must bear
now the punishment for your obscene conduct
and prostitution.”
\par }{\PP \VS{36}The
{\ND{Lord}}
said
to
me: “Son
of man,
are you willing
to pronounce
judgment on Oholah
and Oholibah? Then declare
to them their abominable deeds!
\VS{37}For
they have committed adultery
and blood
is on their hands.
They have committed adultery
with
their idols,
and their
sons,
whom
they bore
to me, they have passed
through the fire as food to their idols.
\VS{38}Moreover,
they have done
this
to me: In the very same day
they desecrated
my sanctuary
and profaned
my Sabbaths.
\VS{39}On the same
day
they slaughtered
their sons
for their idols,
they came
to
my sanctuary
to desecrate
it. This is what
they have done
in the middle
of my house.
\par }{\PP \VS{40}“They even
sent
for
men
from far away;
when the messenger
arrived,
those men set out.
For
them you bathed,
painted
your eyes,
and decorated yourself
with jewelry.
\VS{41}You sat
on
a magnificent
couch,
with a table
arranged
in front
of it where you placed
my incense
and my olive oil.
\VS{42}The sound
of a carefree
crowd
accompanied her, including all kinds of men;
even Sabeans
were brought
from the desert.
The sisters put
bracelets
on
their wrists
and beautiful
crowns
on
their heads.
\VS{43}Then I said
about the one worn out
by adultery,
‘Now they will commit
immoral acts
with her.’
\VS{44}They had sex
with
her
as one does with a prostitute.
In this way
they had sex
with
Oholah
and Oholibah,
promiscuous
women.
\VS{45}But
upright men
will punish them
appropriately
for their adultery
and bloodshed,
because
they are adulteresses
and blood
is on their hands.
\par }{\PP \VS{46}“For
this is what
the sovereign

{\ND{Lord}}
says: Bring up
an army against
them and subject
them to terror
and plunder.
\VS{47}That army will pelt
them with stones
and slash them
with their swords;
they will kill
their sons
and daughters
and burn
their houses.
\VS{48}I will put an end
to the obscene
conduct
in the land;
all
the women
will learn a lesson from
this and not
engage
in obscene
conduct.
\VS{49}They will repay
you for your obscene
conduct, and you will be punished
for idol
worship. Then
you will know
that
I am
the sovereign

{\ND{Lord}}.”

\par }\Chap{24}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me
in the ninth
year,
in the tenth
month,
on the tenth
day of the month :
\VS{2}“Son
of man,
write
down the name
of this
day,
this
very
day.
The king
of Babylon
has laid
siege to
Jerusalem
this
very
day.
\VS{3}Recite
a proverb
to this rebellious
house
and say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Set on
the pot,
set
it on,
\par }{\Q pour
water
in it too;
\par }{\Q \VS{4}add
the pieces
of meat to
it,
\par }{\Q every
good
piece,
\par }{\Q the thigh
and the shoulder;
\par }{\Q fill
it with choice
bones.
\par }{\Q \VS{5}Take
the choice
bone of the flock,
\par }{\Q heap up
bones
under
it;
\par }{\Q boil
rapidly,
\par }{\Q and
boil
its bones
in it.
\par }{\Q \VS{6}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q Woe
to the city
of bloodshed,
\par }{\Q the pot
whose
rot
is in it,
\par }{\Q whose rot
has not
been removed
from
it!
\par }{\Q Empty it piece
by piece.
\par }{\Q No
lot
has fallen
on it.
\par }{\Q \VS{7}For
her blood
was in it;
\par }{\Q she poured
it on
an exposed
rock;
\par }{\Q she did not
pour
it on
the ground
to cover
it up with dust.
\par }{\Q \VS{8}To arouse
anger,
to take vengeance,
\par }{\Q I have placed
her blood
on
an exposed
rock
so that it cannot be covered up.
\par }{\Q \VS{9}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q Woe
to the city
of bloodshed!
\par }{\Q I
will also
make
the pile high.
\par }{\Q \VS{10}Pile up
the bones,
kindle
the fire;
\par }{\Q cook
the meat
well, mix in the spices,
\par }{\Q let the bones
be charred.
\par }{\Q \VS{11}Set
the empty
pot on
the coals,
\par }{\Q until it becomes hot
and its copper
glows,
\par }{\Q until its uncleanness
melts
within
it and its rot
is consumed.
\par }{\Q \VS{12}It has tried
my patience;
\par }{\Q yet its thick
rot
is not
removed
from
it.
\par }{\Q Subject its rot
to the fire!
\par }{\Q \VS{13}You mix uncleanness
with obscene
conduct.
\par }{\Q I tried to cleanse
you, but you are not clean.
\par }{\Q You will not
be cleansed
from your uncleanness
\par }{\Q until
I have exhausted
my anger on you.
\par }{\PP \VS{14}“‘I
the {\ND{Lord}}
have spoken;
judgment is coming
and I will act! I will not
relent,
or
show pity,
or
be sorry! I will judge
you according to your conduct
and your deeds,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\SH Ezekiel’s Wife Dies
\par }{\PP \VS{15}The word
of
{\ND{Lord}}
came to me:
\VS{16}“Son
of man,
realize
that I am about to take
the delight
of your eyes
away from you with a jolt,
but you must not
mourn
or
weep
or
shed
tears.
\VS{17}Groan
in silence
for the dead,
but do not
perform
mourning
rites.
Bind on
your turban
and put
your sandals
on your feet.
Do not
cover
your lip
and do not
eat
food
brought by others.”
\par }{\PP \VS{18}So I spoke
to
the people
in the morning,
and my wife
died
in the evening.
In the morning
I acted
just
as I was commanded.
\VS{19}Then the people
said
to me,
“Will you not
tell
us what
these
things you
are doing
mean for us?”
\par }{\PP \VS{20}So I said
to
them: “The word
of the {\ND{Lord}}
came to me:
\VS{21}Say
to the house
of Israel,
‘This is what
the sovereign

{\ND{Lord}}
says: Realize
I am about to desecrate
my sanctuary
– the source of your confident
pride,
the object in which
your eyes
delight,
and your life’s passion.
Your very own sons
and daughters
whom
you have left
behind will die
by the sword.
\VS{22}Then you will do
as
I have done: You will not
cover your lip
or
eat
food brought by others.
\VS{23}Your turbans
will be on
your heads
and your sandals
on your feet;
you will not
mourn
or
weep,
but you will rot
for your iniquities
and groan
among
yourselves.
\VS{24}Ezekiel
will be
an object lesson
for you; you will do
all
that
he has done.
When
it happens,
then you will know
that
I am
the sovereign

{\ND{Lord}}.’
\par }{\PP \VS{25}“And you,
son
of man,
this is what will happen on the day
I take
from them
their stronghold
– their beautiful
source of joy,
the object in which their eyes
delight,
and the main concern
of their lives,
as well as their sons
and daughters:
\VS{26}On that day
a fugitive
will come
to you to report
the news.
\VS{27}On that day
you will be able
to speak
again; you will talk
with the fugitive
and be silent
no
longer.
You will be
an object lesson
for them, and they will know
that I
am the
{\ND{Lord}}.”

\par }\Chap{25}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
the Ammonites
and prophesy
against them.
\VS{3}Say
to the Ammonites,
‘Hear
the word
of the sovereign

{\ND{Lord}}: This is what
the sovereign

{\ND{Lord}}
says: You said
“Aha!” about my sanctuary
when
it was desecrated,
about the land
of Israel
when
it was made desolate,
and about the house
of Judah
when
they went
into exile.
\VS{4}So
take note,
I am about to make
you slaves of
the tribes
of the east.
They
will make camps
among you and pitch
their tents
among you. They
will eat
your fruit
and drink
your milk.
\VS{5}I will make
Rabbah
a pasture
for camels
and Ammon
a resting place
for sheep.
Then you will know
that
I
am the
{\ND{Lord}}.
\VS{6}For
this is what
the sovereign

{\ND{Lord}}
says: Because
you clapped
your hands,
stamped
your feet,
and rejoiced
with intense
scorn
over the land
of Israel,
\VS{7}take note, I have stretched
out my hand
against
you, and I will hand
you over as plunder
to the nations.
I will cut
you off
from
the peoples
and make you perish
from
the lands.
I will destroy
you; then you will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\SH A Prophecy Against Moab
\par }{\PP \VS{8}“This is what
the sovereign

{\ND{Lord}}
says: ‘Moab
and Seir
say, “Look,
the house
of Judah
is like all
the other nations.”
\VS{9}So
look,
I am about to open up
Moab’s
flank,
eliminating the cities,
including its frontier
cities,
the beauty
of the land
– Beth Jeshimoth,
Baal Meon,
and Kiriathaim.
\VS{10}I will hand
it
over,
along
with the Ammonites,
to the tribes
of the east,
so that
the Ammonites
will no longer
be remembered among
the nations.
\VS{11}I will execute
judgments
against Moab.
Then they will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\SH A Prophecy Against Edom
\par }{\PP \VS{12}“This is what the sovereign

{\ND{Lord}}
says: ‘Edom
has taken vengeance
against the house
of Judah;
they have made
themselves fully culpable
by taking vengeance on them.
\VS{13}So
this is what
the sovereign

{\ND{Lord}}
says: I will stretch
out my hand
against
Edom,
and I will kill
the people
and animals
within her, and I will make
her desolate;
from Teman
to Dedan
they will die
by the sword.
\VS{14}I will exact
my vengeance
upon Edom
by the hand
of my people
Israel.
They will carry
out in Edom
my anger
and rage;
they will experience
my vengeance,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\SH A Prophecy Against Philistia
\par }{\PP \VS{15}“This is what
the sovereign

{\ND{Lord}}
says: ‘The Philistines
have exacted
merciless
revenge,
showing intense
scorn
in their effort to destroy
Judah with unrelenting
hostility.
\VS{16}So
this
is what
the sovereign

{\ND{Lord}}
says: Take note, I am about to stretch
out my hand
against
the Philistines.
I will kill
the Cherethites
and destroy
those who remain
on the seacoast.
\VS{17}I will exact
great
vengeance
upon them with angry
rebukes.
Then they will know
that
I am
the {\ND{Lord}}, when I exact
my vengeance upon them.’ ”

\par }\Chap{26}{\PP \VerseOne{1}In the eleventh
year,
on the first
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
because
Tyre
has
said
about Jerusalem, ‘Aha,
the gateway
of the peoples
is broken;
it has swung
open to me.
I will become
rich, now that she has been destroyed,’
\VS{3}therefore
this is what
the sovereign

{\ND{Lord}}
says: Look,
I am against
you, O Tyre! I will bring up
many
nations
against you, as the sea
brings up
its waves.
\VS{4}They will destroy
the walls
of Tyre
and break
down her towers.
I will scrape
her soil
from
her and make
her a bare
rock.
\VS{5}She will be
a place where fishing
nets
are spread,
surrounded
by the sea.
For
I
have spoken,
declares
the sovereign

{\ND{Lord}}. She will become
plunder
for the nations,
\VS{6}and her daughters
who
are in the field
will be slaughtered
by the sword.
Then they will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{7}“For
this is what
the sovereign

{\ND{Lord}}
says: Take note
that I am about
to bring
King
Nebuchadrezzar
of Babylon,
king
of kings,
against Tyre
from the north,
with horses,
chariots,
and horsemen,
an army
and hordes
of people.
\VS{8}He will kill
your daughters
in the field
with the sword.
He will
build a siege
wall against
you, erect
a siege
ramp
against you, and raise
a great shield against you.
\VS{9}He will direct
the blows
of his battering rams
against your walls
and tear down
your towers
with his weapons.
\VS{10}He will cover
you with the dust
kicked up by his many horses.
Your walls
will shake
from the noise
of the horsemen,
wheels,
and chariots
when he enters
your gates
like those who invade
through a city’s
broken walls.
\VS{11}With
his horses’
hoofs
he will trample
all
your streets.
He will kill
your people
with the sword,
and your strong
pillars
will tumble down
to the ground.
\VS{12}They
will steal
your wealth
and loot
your merchandise.
They will tear
down your walls
and destroy your luxurious homes.
Your stones,
your trees,
and your soil
he will throw into
the water.
\VS{13}I will silence
the noise
of your songs;
the sound
of your harps
will be heard
no
more.
\VS{14}I will make
you a bare
rock;
you will be a place where fishing
nets
are spread.
You will
never
be
built
again,
for
I,
the
{\ND{Lord}},
have spoken,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{15}“This is what the sovereign

{\ND{Lord}}
says
to Tyre: Oh, how the coastlands
will shake
at the sound
of your fall,
when the wounded
groan,
at the massive
slaughter
in your midst!
\VS{16}All
the princes
of the sea
will vacate
their thrones.
They will remove
their robes
and
strip
off their embroidered
clothes;
they will clothe
themselves with trembling.
They will sit
on
the ground;
they will tremble
continually
and be shocked at what has happened to you.
\VS{17}They will sing this lament
over
you:

\par }{\Q “ ‘How
you have perished
– you have
vanished
from the seas,
\par }{\Q O renowned
city,
once mighty
in the sea,
\par }{\Q she
and her
inhabitants,
who spread
their terror!
\par }{\Q \VS{18}Now
the coastlands
will tremble
on the day
of your fall;
\par }{\Q the coastlands
by the sea
will be terrified
by your passing.’
\par }{\PP \VS{19}“For
this is what
the sovereign

{\ND{Lord}}
says: When I make
you desolate
like the uninhabited
cities,
when
I bring up
the deep
over
you and the surging
waters
overwhelm you,
\VS{20}then I will bring
you down
to bygone
people,
to
be with
those who descend
to the pit.
I will make you live
in the lower parts
of the earth,
among the primeval
ruins,
with
those who descend
to the pit,
so
that you will not
be inhabited
or stand
in the land
of the living.
\VS{21}I will bring terrors
on you, and you will
be no
more! Though you are sought
after, you will never
be found
again,
declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{27}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“You,
son
of man,
sing
a lament
for Tyre.
\VS{3}Say
to Tyre,
who sits
at the entrance
of the sea,
merchant
to the peoples
on many
coasts,
‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘O
Tyre,
you
have said,
“I am
perfectly
beautiful.”
\VS{4}\par }{\Q Your borders
are in the heart
of the seas;
\par }{\Q your builders
have perfected
your beauty.
\par }{\Q \VS{5}They crafted
all
your planks
out of fir trees
from Senir;
\par }{\Q they took
a cedar
from Lebanon
to make
your mast.
\par }{\Q \VS{6}They made
your oars
from oaks
of Bashan;
\par }{\Q they made
your deck
with cypresses
from the Kittean
isles.
\par }{\Q \VS{7}Fine linen
from Egypt,
woven with patterns,
was used for
your sail
\par }{\Q to serve
as your banner;
\par }{\Q blue
and purple
from the coastlands
of Elishah
was used for
your deck’s awning.
\par }{\Q \VS{8}The leaders
of Sidon
and Arvad
were
your rowers;
\par }{\Q your skilled men,
O Tyre,
were
your captains.
\par }{\Q \VS{9}The elders
of Gebal
and her skilled men
were
within you, mending
cracks;
\par }{\Q all
the ships
of the sea
and their mariners
were
within you to trade for
your merchandise.
\par }{\Q \VS{10}Men of Persia,
Lud,
and Put
were
in your army,
men
of war.
\par }{\Q They hung
shield
and helmet
on you; they
gave
you your splendor.
\par }{\Q \VS{11}The Arvadites
joined your army
on
your walls
all around,
\par }{\Q and the Gammadites
were in your towers.
\par }{\Q They hung
their quivers
on
your walls
all around;
\par }{\Q they
perfected
your beauty.
\par }{\PP \VS{12}“‘Tarshish
was your trade
partner because of your abundant
wealth;
they exchanged silver,
iron,
tin,
and lead
for your products.
\VS{13}Javan,
Tubal,
and Meshech
were your clients;
they exchanged
slaves
and bronze
items
for your merchandise.
\VS{14}Beth
Togarmah
exchanged horses,
chargers,
and mules
for your products.
\VS{15}The Dedanites
were your clients.
Many
coastlands
were your customers;
they paid
you with ivory
tusks
and ebony.
\VS{16}Edom
was your trade
partner because of the abundance
of your goods;
they exchanged turquoise,
purple,
embroidered work,
fine linen,
coral,
and rubies
for your products.
\VS{17}Judah
and the land
of Israel
were your clients;
they traded
wheat
from Minnith,
millet,
honey,
olive oil,
and balm
for your merchandise.
\VS{18}Damascus
was your trade
partner
because of the abundance
of your goods and of all
your wealth: wine
from Helbon,
white
wool from Zahar,
\VS{19}and casks of wine from Izal they exchanged
for your products.
Wrought
iron,
cassia,
and sweet cane
were
among your merchandise.
\VS{20}Dedan
was your client
in saddlecloths
for riding.
\VS{21}Arabia
and all
the princes
of Kedar
were
your trade
partners;
for lambs,
rams,
and goats
they traded with you.
\VS{22}The merchants
of Sheba
and Raamah
engaged in trade
with you; they
traded
the best
kinds
of spices
along with precious
stones
and gold
for your products.
\VS{23}Haran,
Kanneh,
Eden,
merchants
from Sheba,
Asshur,
and Kilmad
were your clients.
\VS{24}They
traded
with you choice
garments,
purple
clothes
and embroidered work,
and multicolored
carpets,
bound
and reinforced with cords;
these were among your merchandise.
\VS{25}The ships
of Tarshish
were the transports
for your merchandise.
\par }{\Q “‘So you were filled
and weighed down in the heart
of the seas.
\par }{\Q \VS{26}Your rowers
have brought
you into surging
waters.
\par }{\Q The east
wind
has wrecked
you in the heart
of the seas.
\par }{\Q \VS{27}Your wealth,
products,
and merchandise,
your sailors
and captains,
\par }{\Q your ship’s carpenters,
your merchants,
\par }{\Q and all
your fighting
men
within you,
\par }{\Q along with
all
your crew
who
are in
you,
\par }{\Q will fall
into the heart
of the seas
on the day
of your downfall.
\par }{\Q \VS{28}At
the sound
of your captains’
cry the waves will surge;
\par }{\Q \VS{29}They will descend
from their ships
– all
who handle
the oar,
\par }{\Q the sailors
and all
the sea
captains
– they will stand
on the land.
\par }{\Q \VS{30}They will lament
loudly
over
you and cry
bitterly.
\par }{\Q They will throw
dust
on
their heads
and roll
in the ashes;
\par }{\Q \VS{31}they will tear
out their hair
because of you and put on
sackcloth,
\par }{\Q and they will weep
bitterly
over you with intense
mourning.
\par }{\Q \VS{32}As they wail
they will lament
over you, chanting:
\par }{\Q “Who
was like Tyre,
like a tower
in the midst
of the sea?”
\par }{\Q \VS{33}When
your products
went out from
the seas,
\par }{\Q you satisfied
many
peoples;
\par }{\Q with the abundance
of your wealth
and merchandise
\par }{\Q you enriched
the kings
of the earth.
\par }{\Q \VS{34}Now you are wrecked
by the seas,
in the depths
of the waters;
\par }{\Q your merchandise
and all
your company
have sunk along with you.
\par }{\Q \VS{35}All
the inhabitants
of the coastlands
are shocked
at you,
\par }{\Q and their kings
are horribly afraid – their faces are troubled.
\par }{\Q \VS{36}The traders
among the peoples
hiss
at you;
\par }{\Q you have become a horror,
and will be no
more.’ ”

\par }\Chap{28}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
say
to the prince
of Tyre,
‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Your heart
is proud
and you said,
“I am a god;
\par }{\Q I
sit
in the seat
of gods,
in the heart
of the seas”
–
\par }{\Q yet you
are a man
and not
a god,
\par }{\Q though you
think
you are godlike.
\par }{\Q \VS{3}Look,
you
are wiser
than Daniel;
\par }{\Q no
secret
is hidden from you.
\par }{\Q \VS{4}By your wisdom
and understanding
you have gained
wealth
for yourself;
\par }{\Q you have amassed
gold
and silver
in your treasuries.
\par }{\Q \VS{5}By your great
skill
in trade
you have increased
your wealth,
\par }{\Q and your heart
is proud
because of your wealth.
\par }{\Q \VS{6}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q Because
you think
you are godlike,
\par }{\Q \VS{7}I am
about to bring
foreigners
against
you, the most terrifying
of nations.
\par }{\Q They will draw
their swords
against
the grandeur
made by your wisdom,
\par }{\Q and they will defile
your splendor.
\par }{\Q \VS{8}They will bring
you down
to the pit,
and you will die
violently
in the heart
of the seas.
\par }{\Q \VS{9}Will you still say,
“I am
a god,”
before
the one who kills you –
\par }{\Q though you are a man and not a god –
\par }{\Q when you are in the power of those who wound you?
\par }{\Q \VS{10}You will die
the death
of the uncircumcised
by the hand
of foreigners;
\par }{\Q for
I
have spoken,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\PP \VS{11}The word
of the {\ND{Lord}}
came to me:
\VS{12}“Son
of man,
sing
a lament
for the king
of Tyre,
and say
to him, ‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘You
were the sealer
of perfection,
\par }{\Q full
of wisdom,
and perfect
in beauty.
\par }{\Q \VS{13}You were in Eden,
the garden
of God.
\par }{\Q Every
precious
stone
was your covering,
\par }{\Q the ruby,
topaz,
and emerald,
\par }{\Q the chrysolite,
onyx,
and jasper,
\par }{\Q the sapphire,
turquoise,
and beryl;
\par }{\Q your settings
and mounts
were made
of gold.
\par }{\Q On the day
you were created
they were prepared.
\par }{\Q \VS{14}I placed you
there with an anointed
guardian
cherub;
\par }{\Q you were on the holy
mountain
of God;
\par }{\Q you walked about
amidst
fiery
stones.
\par }{\Q \VS{15}You were blameless
in your behavior
from the day
you were created,
\par }{\Q until
sin
was discovered in you.
\par }{\Q \VS{16}In the abundance
of your trade
you were filled
with violence,
and you sinned;
\par }{\Q so I defiled
you and banished
you from the mountain
of God
–
\par }{\Q the guardian cherub
expelled
you from the midst
of the stones
of fire.
\par }{\Q \VS{17}Your heart
was proud
because of your beauty;
\par }{\Q you corrupted
your wisdom
on
account of your splendor.
\par }{\Q I threw
you down to the ground;
\par }{\Q I placed you before
kings,
that they might see you.
\par }{\Q \VS{18}By
the multitude
of your iniquities,
through the sinfulness
of your trade,
\par }{\Q you desecrated
your sanctuaries.
\par }{\Q So I drew
fire
out from within
you;
\par }{\Q it consumed
you,
\par }{\Q and I
turned you
to ashes
on
the earth
\par }{\Q before the eyes
of all
who saw you.
\par }{\Q \VS{19}All
who know
you among the peoples
are shocked
at you;
\par }{\Q you have become terrified
and will be no
more.’ ”
\par }{\SH A Prophecy Against Sidon
\par }{\PP \VS{20}The word
of the {\ND{Lord}}
came to me:
\VS{21}“Son
of man,
turn
toward
Sidon
and prophesy
against it.
\VS{22}Say,
‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Look,
I am against
you, Sidon,
\par }{\Q and I will magnify
myself in your midst.
\par }{\Q Then they will know
that
I am
the {\ND{Lord}}
\par }{\Q when I execute
judgments
on her
\par }{\Q and reveal my sovereign power in her.
\par }{\Q \VS{23}I will send
a plague
into the city and bloodshed
into its streets;
\par }{\Q the slain
will fall
within
it, by the sword
that attacks it from every side.
\par }{\Q Then they will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{24}“‘No
longer
will Israel
suffer
from the sharp briers
or painful
thorns
of all
who surround
and scorn
them. Then they will know
that
I am
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{25}“‘This is what
the sovereign

{\ND{Lord}}
says: When I regather
the
house
of Israel
from
the peoples
where
they are dispersed,
I will reveal
my sovereign power over them in the sight
of the nations,
and they will live
in their land
that
I gave
to my servant
Jacob.
\VS{26}They will live
securely
in it; they will build
houses
and plant
vineyards.
They will live
securely
when I execute
my judgments
on all
those who scorn
them and surround
them. Then they will know
that
I am
the {\ND{Lord}}
their God.’ ”

\par }\Chap{29}{\PP \VerseOne{1}In the tenth
year,
in the tenth
month, on the twelfth
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
Pharaoh
king
of Egypt,
and prophesy
against him
and against
all
Egypt.
\VS{3}Tell
them, ‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Look,
I am against
you, Pharaoh
king
of Egypt,
\par }{\Q the great
monster
lying
in the midst
of its waterways,
\par }{\Q who
has said,
“My Nile
is my own, I
made it for myself.”
\par }{\Q \VS{4}I will put
hooks
in your jaws
\par }{\Q and stick
the fish
of your waterways
to your scales.
\par }{\Q I will haul you up
from the midst
of your waterways,
\par }{\Q and all
the fish
of your waterways will
stick
to your scales.
\par }{\Q \VS{5}I will leave
you in the wilderness,
\par }{\Q you and all
the fish
of your waterways;
\par }{\Q you will fall
in
the open
field
and will not
be gathered up
or
collected.
\par }{\Q I have given
you as food
to the beasts
of the earth
and the birds
of the skies.
\par }{\Q \VS{6}Then all
those living
in Egypt
will know
that
I am
the {\ND{Lord}}
\par }{\Q because
they were a reed
staff
for the house
of Israel;
\par }{\Q \VS{7}when they grasped
you with their hand,
you broke
and tore their shoulders,
\par }{\Q and when they leaned
on
you, you splintered
and caused their legs to be
unsteady.
\par }{\PP \VS{8}“‘Therefore,
this is what
the sovereign

{\ND{Lord}}
says: Look,
I am about to bring
a sword
against you, and I will kill
every person
and every animal.
\VS{9}The land
of Egypt
will become a desolate
ruin.
Then they will know
that
I am
the {\ND{Lord}}.
\par }{\PP Because
he said,
“The Nile
is mine and I
made it,”
\VS{10}I am against
you and your waterways.
I will turn
the land
of Egypt
into an utter
desolate
ruin
from Migdol
to Syene,
as far
as the border
with Ethiopia.
\VS{11}No
human
foot
will pass
through it, and no
animal’s
foot
will pass
through it; it will be uninhabited
for forty
years.
\VS{12}I will turn
the
land
of Egypt
into a desolation
in the midst
of desolate
lands;
for forty
years
her cities
will lie desolate
in the midst
of ruined
cities.
I will scatter
Egypt
among the nations
and disperse
them among foreign countries.
\par }{\PP \VS{13}“‘For
this is what
the sovereign

{\ND{Lord}}
says: At the end
of forty
years
I will gather
Egypt
from
the peoples
where
they were scattered.
\VS{14}I will restore
the fortunes
of Egypt,
and will bring them back
to the land
of Pathros,
to the land
of their origin;
there
they will be
an insignificant
kingdom.
\VS{15}It will be
the most insignificant
of the kingdoms;
it will never
again
exalt
itself over
the nations.
I will make them so small
that they will not
rule
over the nations.
\VS{16}It will
never
again
be Israel’s
source
of confidence,
but a reminder
of how
they sinned
by turning
to Egypt for help. Then
they will know
that
I am
the sovereign

{\ND{Lord}}.’ ”
\par }{\PP \VS{17}In the twenty-seventh
year,
in the first
month, on the first
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{18}“Son
of man,
King
Nebuchadrezzar
of Babylon
made
his army
labor
hard
against
Tyre.
Every
head
was rubbed bald
and every
shoulder
rubbed
bare; yet he and his army
received
no
wages
from Tyre
for
the work
he carried out
against it.
\VS{19}Therefore
this is what
the sovereign

{\ND{Lord}}
says: Look,
I am about to give
the land
of Egypt
to King
Nebuchadrezzar
of Babylon.
He will carry off
her wealth,
capture
her loot,
and seize
her plunder;
it will be
his army’s
wages.
\VS{20}I have
given
him the land
of Egypt
as
his compensation
for attacking Tyre , because they did
it for me, declares
the sovereign

{\ND{Lord}}.
\VS{21}On that day
I will make
Israel
powerful, and I will give
you the right to be heard
among
them. Then they will know
that
I am
the {\ND{Lord}}.”

\par }\Chap{30}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
prophesy
and say,
‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Wail,
“Alas,
the day is here!”
\par }{\Q \VS{3}For
the day
is near,
\par }{\Q the day
of the {\ND{Lord}}
is near;
\par }{\Q it will be a day
of storm clouds,
\par }{\Q it will be a time
of judgment for the nations.
\par }{\Q \VS{4}A sword
will come
against Egypt
\par }{\Q and panic
will overtake
Ethiopia
\par }{\Q when the slain
fall
in Egypt
\par }{\Q and they carry away
her wealth
\par }{\Q and dismantle
her foundations.
\par }{\PP \VS{5}Ethiopia,
Put,
Lud,
all
the foreigners, Libya,
and the people of the covenant
land
will die
by the sword
along with them.
\par }{\Q \VS{6}“‘This is what
the
{\ND{Lord}}
says:
\par }{\Q Egypt’s
supporters
will fall;
\par }{\Q her confident pride
will crumble.
\par }{\Q From Migdol
to Syene
they will die
by the sword
within her,
\par }{\Q declares
the sovereign

{\ND{Lord}}.
\par }{\Q \VS{7}They will be desolate
among
desolate
lands,
\par }{\Q and their cities
will be
among
ruined
cities.
\par }{\Q \VS{8}They will know
that
I am
the {\ND{Lord}}
\par }{\Q when I ignite
a fire
in Egypt
\par }{\Q and all
her allies are defeated.
\par }{\PP \VS{9}On that
day
messengers
will go out
from
me in ships
to frighten
overly confident
Ethiopia;
panic
will overtake
them on the day
of Egypt’s
doom; for
beware
– it is coming!
\par }{\Q \VS{10}“‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q I will put an end
to the hordes
of Egypt,
\par }{\Q by the hand
of King
Nebuchadrezzar
of Babylon.
\par }{\Q \VS{11}He
and his people
with
him,
\par }{\Q the most terrifying
of the nations,
\par }{\Q will be brought
there to destroy
the land.
\par }{\Q They will draw
their swords
against
Egypt,
\par }{\Q and fill
the
land
with corpses.
\par }{\Q \VS{12}I will
dry up
the waterways
\par }{\Q and hand
the land
over to
evil men.
\par }{\Q I will make
the land
and everything
in it desolate
by the hand
of foreigners.
\par }{\Q I,
the {\ND{Lord}}, have spoken!
\par }{\Q \VS{13}“‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q I will destroy
the idols,
\par }{\Q and put an end
to the gods
of Memphis.
\par }{\Q There will no
longer
be a prince
from the land
of Egypt;
\par }{\Q so I will make
the land
of Egypt
fearful.
\par }{\Q \VS{14}I will desolate
Pathros,
\par }{\Q I
will ignite
a fire
in Zoan,
\par }{\Q and I will execute
judgments
on Thebes.
\par }{\Q \VS{15}I will pour out
my anger
upon
Pelusium,
\par }{\Q the stronghold
of Egypt;
\par }{\Q I will cut off
the hordes
of Thebes.
\par }{\Q \VS{16}I will ignite
a fire
in Egypt;
\par }{\Q Syene
will writhe
in agony,
\par }{\Q Thebes
will be
broken down,
\par }{\Q and Memphis
will face enemies
every day.
\par }{\Q \VS{17}The young men
of On
and of Pi-beseth
will die
by the sword;
\par }{\Q and the cities will go
into captivity.
\par }{\Q \VS{18}In Tahpanhes
the day
will be dark
\par }{\Q when I break
the yoke
of Egypt
there.
\par }{\Q Her confident
pride
will cease
within her;
\par }{\Q a cloud
will cover
her, and her daughters
will go
into captivity.
\par }{\Q \VS{19}I will execute
judgments
on Egypt.
\par }{\Q Then they will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\PP \VS{20}In the eleventh
year,
in the first
month, on the seventh
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{21}“Son
of man,
I have broken
the arm
of Pharaoh
king
of Egypt.
Look,
it has not
been bandaged
for healing
or set
with a dressing so that it might become strong
enough to grasp
a sword.
\VS{22}Therefore
this is what
the sovereign

{\ND{Lord}}
says: Look,
I am against Pharaoh
king
of Egypt,
and I will break
his arms,
the strong
arm and the broken
one, and I will make
the sword
drop
from his hand.
\VS{23}I will scatter
the Egyptians
among the nations,
and disperse
them among foreign countries.
\VS{24}I will strengthen
the arms
of the king
of Babylon,
and I will place
my sword
in his hand,
but I will break
the arms
of Pharaoh,
and he will groan
like the fatally
wounded
before the king of Babylon.
\VS{25}I will strengthen
the
arms
of the king
of Babylon,
but the arms
of Pharaoh
will fall
limp. Then they will know
that
I am
the {\ND{Lord}}
when I place
my sword
in the hand
of the king
of Babylon
and he extends
it against the land
of Egypt.
\VS{26}I will scatter
the Egyptians
among the nations
and disperse
them among foreign countries.
Then they will know
that
I am
the {\ND{Lord}}.”

\par }\Chap{31}{\PP \VerseOne{1}In the eleventh
year,
in the third
month, on the first
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
say
to
Pharaoh
king
of Egypt
and his hordes:
\par }{\Q “ ‘Who
are you like
in your greatness?
\par }{\Q \VS{3}Consider
Assyria,
a cedar
in Lebanon,
\par }{\Q with beautiful
branches,
like a forest
giving shade,
\par }{\Q and extremely
tall;
\par }{\Q its top
reached
into
the clouds.
\par }{\Q \VS{4}The water
made it grow;
\par }{\Q underground
springs
made it grow tall.
\par }{\Q Rivers
flowed
all around
the place it was planted,
\par }{\Q while smaller channels
watered all
the trees
of the field.
\par }{\Q \VS{5}Therefore
it grew
taller
than all
the trees
of the field;
\par }{\Q its boughs
grew large
and its branches
grew long,
\par }{\Q because of the plentiful
water
in its shoots.
\par }{\Q \VS{6}All
the birds
of the sky
nested in its boughs;
\par }{\Q under
its branches
all
the beasts
of the field
gave birth,
\par }{\Q in its shade
all
the great
nations
lived.
\par }{\Q \VS{7}It was beautiful
in its loftiness,
in the length
of its branches;
\par }{\Q for
its roots
went down
deep to
plentiful
waters.
\par }{\Q \VS{8}The cedars
in the garden
of God
could not
eclipse
it,
\par }{\Q nor
could the fir trees
match
its boughs;
\par }{\Q the plane trees
were
as nothing
compared to its branches;
\par }{\Q no
tree
in the garden
of God
could rival
its beauty.
\par }{\Q \VS{9}I made
it beautiful
with its many
branches;
\par }{\Q all
the trees
of Eden,
in the garden
of God,
envied it.
\par }{\PP \VS{10}“‘Therefore
this is what
the sovereign

{\ND{Lord}}
says: Because
it
was tall
in stature,
and its top reached into the clouds, and it was proud
of its height,
\VS{11}I gave
it over
to the leader
of the nations.
He has
judged it thoroughly, as its sinfulness
deserves. I have thrown it out.
\VS{12}Foreigners
from the most terrifying
nations
have
cut
it down
and left it to
lie
there on the mountains.
In all
the valleys
its branches
have fallen,
and its boughs
lie broken
in the ravines
of the land.
All
the peoples
of the land
have departed
from its shade and left it.
\VS{13}On
its ruins
all
the birds
of the sky
will live,
and all
the wild
animals
will walk on its branches.
\VS{14}For
this reason no
watered
trees
will grow
so tall;
their tops
will not
reach
into the clouds, nor will the well-watered
ones
grow that high.
For
all
of them
have been appointed
to
die
in the lower
parts of the earth;
they will be among
mere mortals,
with those who descend
to
the pit.
\par }{\PP \VS{15}“‘This is what
the sovereign

{\ND{Lord}}
says: On the day
it went down
to Sheol
I caused observers to lament.
I covered
it with the deep
and held back
its rivers;
its plentiful
water
was restrained.
I clothed Lebanon
in
black
for it, and all
the trees
of the field
wilted
because of it.
\VS{16}I made the nations
shake
at the sound
of its fall,
when I threw it down
to Sheol,
along with
those who descend
to the pit.
Then
all
the trees
of Eden,
the choicest
and the best
of Lebanon,
all
that were well-watered,
were comforted
in the earth
below.
\VS{17}Those
who lived
in its shade,
its allies
among
the nations,
also
went down
with it to Sheol,
to
those killed
by the sword.
\VS{18}Which
of the trees
of Eden
was like
you in majesty
and loftiness? You will be brought down
with
the trees
of Eden
to
the lower
parts of the earth;
you will lie
among
the uncircumcised,
with
those killed
by the sword! This is
what will happen to Pharaoh
and all
his hordes,
declares
the sovereign

{\ND{Lord}}.’ ”

\par }\Chap{32}{\PP \VerseOne{1}In the twelfth
year,
in the twelfth
month,
on the first
of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
sing
a lament
for
Pharaoh
king
of Egypt,
and say
to
him:
\par }{\Q “ ‘You were like a lion
among the nations,
\par }{\Q but you
are a monster
in the seas;
\par }{\Q you thrash
about in your streams,
\par }{\Q stir up
the water
with your feet,
\par }{\Q and muddy
your streams.
\par }{\PP \VS{3}“‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘I
will throw
my net
over you in the assembly
of many
peoples;
\par }{\Q and they will haul
you up
in my dragnet.
\par }{\Q \VS{4}I will leave
you on
the ground,
\par }{\Q I will fling
you on the open field,
\par }{\Q I will allow
all
the birds
of the sky
to settle on
you,
\par }{\Q and I will permit
all
the wild animals
to gorge
themselves on you.
\par }{\Q \VS{5}I will put
your flesh
on
the mountains,
\par }{\Q and fill
the valleys
with your maggot-infested carcass.
\par }{\Q \VS{6}I will drench
the land
with the flow
\par }{\Q of your blood
up to
the mountains,
\par }{\Q and the ravines
will be full
of your blood.
\par }{\Q \VS{7}When I extinguish
you, I will cover
the sky;
\par }{\Q I will darken
its stars.
\par }{\Q I will cover
the sun
with a cloud,
\par }{\Q and the moon
will not
shine.
\par }{\Q \VS{8}I will darken
all
the lights
in the sky
over
you,
\par }{\Q and I will darken
your land,
\par }{\Q declares
the sovereign

{\ND{Lord}}.
\par }{\Q \VS{9}I will disturb
many
peoples,
\par }{\Q when I bring
about your destruction
among the nations,
\par }{\Q among countries
you do not
know.
\par }{\Q \VS{10}I will shock
many
peoples
with you,
\par }{\Q and their kings
will shiver
with horror
because of you.
\par }{\Q When
I brandish
my sword
before
them,
\par }{\Q every moment
each
one will tremble
for his life,
on the day
of your fall.
\par }{\PP \VS{11}“‘For
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘The sword
of the king
of Babylon
will attack you.
\par }{\Q \VS{12}By the swords
of the mighty warriors
I will cause your hordes
to fall
–
\par }{\Q all
of them are the most terrifying
among the nations.
\par }{\Q They will devastate
the pride
of Egypt,
\par }{\Q and all
its hordes
will be destroyed.
\par }{\Q \VS{13}I will destroy
all
its cattle
beside
the plentiful
waters;
\par }{\Q and no
human
foot
will disturb
the waters again,
\par }{\Q nor
will the hooves
of cattle
disturb them.
\par }{\Q \VS{14}Then
I will make their waters
calm,
\par }{\Q and will make their streams
flow
like olive oil,
declares
the sovereign

{\ND{Lord}}.
\par }{\Q \VS{15}When
I turn
the
land
of Egypt
into desolation
\par }{\Q and the land
is destitute of everything
that fills it,
\par }{\Q when I strike
all
those who live
in it,
\par }{\Q then they will know
that
I am
the {\ND{Lord}}.’
\par }{\Q \VS{16}This is
a lament;
they will chant it.
\par }{\Q The daughters
of the nations
will chant
it.
\par }{\Q They will chant
it over
Egypt
and over
all
her hordes,
\par }{\Q declares
the sovereign

{\ND{Lord}}.”
\par }{\PP \VS{17}In the twelfth
year,
on the fifteenth
day of the month,
the word
of the {\ND{Lord}}
came to me:
\VS{18}“Son
of man,
wail
over
the horde
of Egypt.
Bring it down;
bring her and the daughters
of powerful
nations
down to
the lower
parts of the earth,
along with
those who descend
to the pit.
\VS{19}Say to them, ‘Whom
do you surpass
in beauty? Go down
and be laid
to rest with
the uncircumcised!’
\VS{20}They will fall
among
those killed
by
the sword.
The sword
is drawn;
they carry her and all
her hordes away.
\VS{21}The bravest
of the warriors
will speak
to him from the midst
of Sheol
along with
his allies,
saying: ‘The uncircumcised
have come down;
they lie
still, killed
by the sword.’
\par }{\PP \VS{22}“Assyria
is there
with all
her assembly
around
her grave,
all
of them struck
down
by the sword.
\VS{23}Their graves
are located in the remote
slopes of the pit.
Her assembly
is around
her grave,
all
of them struck
down
by the sword,
those who
spread
terror
in the land
of the living.
\par }{\PP \VS{24}“Elam
is there
with all
her hordes
around
her grave;
all
of them struck
down
by the sword.
They went down
uncircumcised
to
the lower
parts of the earth,
those who
spread
terror
in the land
of the living.
Now they will bear
their shame
with
those who descend
to the pit.
\VS{25}Among
the dead
they have made
a bed
for her, along with all
her hordes
around
her grave.
All
of them are uncircumcised,
killed
by the sword,
for
their terror
had spread in the land
of the living.
They bear
their shame
along with
those who descend
to the pit;
they are placed
among
the dead.
\par }{\PP \VS{26}“Meshech-Tubal
is there,
along with all
her hordes
around
her grave.
All
of them are uncircumcised,
killed
by the sword,
for
they spread
their terror
in the land
of the living.
\VS{27}They do not
lie
with
the fallen
warriors
of ancient times,
who
went down
to Sheol
with their weapons
of war,
having their swords
placed
under
their heads
and their shields
on
their bones,
when
the terror
of these warriors
was in the land
of the living.
\par }{\PP \VS{28}“But as for you,
in the midst
of the uncircumcised
you will be broken,
and you will lie
with
those killed
by the sword.
\par }{\PP \VS{29}“Edom
is there
with
her kings
and all
her princes.
Despite
their might
they are laid with
those killed
by the sword;
they
lie
with
the uncircumcised
and those who descend
to the pit.
\par }{\PP \VS{30}“All
the leaders
of the north
are there,
along with all
the Sidonians;
despite their might
they have
gone down
in shameful
terror
with
the dead.
They lie
uncircumcised
with
those killed
by the sword,
and bear
their shame
with
those who descend
to the pit.
\par }{\PP \VS{31}“Pharaoh
will see
them and be consoled
over
all
his hordes
who were killed
by the sword,
Pharaoh
and all
his army,
declares
the sovereign

{\ND{Lord}}.
\VS{32}Indeed,
I
terrified
him in the land
of the living,
yet he will lie
in the midst
of the uncircumcised
with
those killed
by the sword,
Pharaoh
and all
his hordes,
declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{33}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
speak
to
your people,
and say
to
them, ‘Suppose
I bring
a sword
against the land,
and the people
of the land
take
one
man
from their borders
and make
him their watchman.
\VS{3}He sees
the sword
coming
against
the land,
blows
the trumpet,
and warns
the people,
\VS{4}but there is one who hears
the sound
of the trumpet
yet does not
heed the warning.
Then the sword
comes
and sweeps
him away.
He will be
responsible
for his own death.
\VS{5}He
heard
the sound
of the trumpet
but
did not
heed the warning,
so
he is responsible
for himself. If he had heeded
the warning,
he would have saved
his life.
\VS{6}But suppose
the watchman
sees
the
sword
coming
and does not
blow
the trumpet
to warn
the people.
Then
the sword
comes
and takes
one
of their
lives.
He is
swept away
for his iniquity,
but I will hold
the watchman
accountable for that person’s death.’
\par }{\PP \VS{7}“As for you,
son
of man,
I
have made you
a watchman
for the house
of Israel.
Whenever you hear
a word
from my mouth,
you must warn
them on my behalf.
\VS{8}When I say
to the wicked,
‘O wicked
man, you must certainly
die,’
and you do not
warn
the wicked
about
his behavior,
the wicked man
will die
for his iniquity,
but
I will hold you accountable for his death.
\VS{9}But
if
you
warn
the wicked
man to change
his behavior,
and he refuses
to change,
he will die
for his iniquity,
but you
have saved
your own life.
\par }{\PP \VS{10}“And you,
son
of man,
say
to
the house
of Israel,
‘This
is what you have said: “Our rebellious
acts and our sins
have caught up with us, and we
are wasting
away because of them. How
then can we live?” ’
\VS{11}Say
to them,
‘As surely
as I
live, declares
the sovereign

{\ND{Lord}}, I take no pleasure
in the death
of the wicked,
but prefer that
the wicked
change
his behavior
and live.
Turn back,
turn back
from your evil
deeds! Why
should you die,
O house
of Israel?’
\par }{\PP \VS{12}“And you,
son
of man,
say
to
your people, ‘The righteousness
of the righteous
will not
deliver
him if he rebels.
As for the wicked,
his wickedness
will not
make him stumble
if he turns from
it.
The righteous
will not
be able
to live
by his righteousness
if he sins.’
\VS{13}Suppose I tell
the righteous
that he
will certainly
live,
but he
becomes confident
in his righteousness
and commits
iniquity.
None
of his righteous
deeds will be remembered;
because of the iniquity
he has
committed
he will die.
\VS{14}Suppose I say
to the wicked,
‘You must certainly
die,’
but he turns from
his sin
and does
what is just
and right.
\VS{15}He returns
what was taken in pledge,
pays
back what he has stolen,
and follows the statutes
that give life,
committing
no iniquity.
He will certainly
live
– he will not
die.
\VS{16}None
of the sins
he has
committed
will be counted
against him. He has done
what is just
and right;
he will certainly
live.
\par }{\PP \VS{17}“Yet
your people
say,
‘The behavior
of the Lord
is
not right,’ when it is
their
behavior that is
not
right.
\VS{18}When a righteous
man turns from his godliness
and commits
iniquity,
he will die for it.
\VS{19}When
the wicked
turns from his sin and does
what is just
and right,
he
will live because of it.
\VS{20}Yet you say,
‘The behavior
of the Lord
is not
right.’ House
of Israel,
I will judge
each
of you according to his behavior.”
\par }{\SH The Fall of Jerusalem
\par }{\PP \VS{21}In the twelfth
year
of our exile,
in the tenth
month, on the fifth
of the month,
a refugee
came
to me
from Jerusalem
saying,
“The city
has been defeated!”
\VS{22}Now the hand
of the {\ND{Lord}}
had been
on
me the evening
before
the refugee
reached
me, but the
{\ND{Lord}} opened
my mouth
by the time the refugee arrived
in
the morning;
he opened
my mouth
and I was no
longer
unable to speak.
\VS{23}The word
of the {\ND{Lord}}
came to me:
\VS{24}“Son
of man,
the ones living
in these
ruins
in the land
of Israel
are saying,
‘Abraham
was
only one
man, yet he possessed
the
land,
but we
are many;
surely the land
has been given
to us for a possession.’
\VS{25}Therefore
say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says: You eat
the meat with the blood
still in it, pray
to
your idols,
and shed
blood.
Do you really think you will possess
the land?
\VS{26}You rely
on
your swords
and commit
abominable deeds;
each
of you defiles
his neighbor’s
wife.
Will you possess
the land?’
\par }{\PP \VS{27}“This is what
you must say
to
them, ‘This is what
the sovereign

{\ND{Lord}}
says: As surely
as I
live, those
living in the ruins
will die
by the sword,
those
in the open
field
I will give
to the wild beasts
for food,
and those who
are in the strongholds
and caves
will die
of disease.
\VS{28}I
will turn
the land
into a desolate
ruin;
her confident
pride
will come to an end.
The mountains
of Israel
will be so desolate
no
one will pass through them.
\VS{29}Then they will know
that
I am
the {\ND{Lord}}
when I turn
the land
into a desolate
ruin
because
of all
the abominable deeds
they have
committed.’
\par }{\PP \VS{30}“But as for you,
son
of man,
your people
(who are talking
about you by
the walls
and at the doors
of the houses) say
to one
another, ‘Come
hear
the word
that
comes
from the
{\ND{Lord}}.’
\VS{31}They come
to
you in crowds,
and they sit
in front
of you as my people.
They hear
your words,
but do not
obey
them. For
they
talk
lustfully,
and their
heart
is set on their own advantage.
\VS{32}Realize
that to them you are like a sensual
song,
a beautiful
voice
and skilled musician.
They hear
your words,
but they do
not obey them.
\VS{33}When all this comes true – and it certainly will – then they will know that a prophet was among them.”

\par }\Chap{34}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
prophesy
against
the shepherds
of Israel;
prophesy,
and say to them – to the shepherds: ‘This is what the sovereign
{\ND{Lord}} says: Woe to the shepherds of Israel who have been feeding themselves! Should not shepherds feed the flock?
\VS{3}You eat
the fat,
you clothe
yourselves with the wool,
you slaughter
the choice
animals, but you do not
feed
the sheep!
\VS{4}You have not
strengthened
the weak,
healed
the sick,
bandaged
the injured,
brought back
the strays,
or
sought
the lost,
but with force
and harshness you have ruled over them.
\VS{5}They were scattered
because
they had no
shepherd,
and they became
food
for every
wild
beast.
\VS{6}My sheep
wandered
over all
the mountains
and on
every
high
hill.
My sheep
were scattered
over
the entire
face
of the earth
with no one
looking
or
searching for them.
\par }{\PP \VS{7}“‘Therefore,
you shepherds,
hear
the word
of the {\ND{Lord}}:
\VS{8}As surely
as I
live, declares
the sovereign

{\ND{Lord}}, my
sheep
have
become
prey
and have become
food
for all
the wild
beasts.
There was no
shepherd,
and my shepherds did not
search
for my
flock,
but fed themselves and did not
feed
my sheep,
\VS{9}Therefore,
you shepherds,
hear
the word
of the {\ND{Lord}}:
\VS{10}This is what
the sovereign

{\ND{Lord}}
says: Look,
I am against the shepherds,
and I will demand
my sheep
from their hand.
I will
no longer
let them be shepherds;
the shepherds
will not
feed
themselves anymore.
I will rescue
my
sheep
from their
mouth,
so that they will no
longer be
food for them.
\par }{\PP \VS{11}“‘For
this is what
the sovereign

{\ND{Lord}}
says: Look,
I myself
will search
for my sheep
and seek them out.
\VS{12}As a shepherd
seeks
out his flock
when
he is among
his scattered
sheep,
so
I will seek out
my flock.
I will rescue
them from all
the places
where
they have
been scattered
on a cloudy,
dark
day.
\VS{13}I will bring them out
from
among the peoples
and gather
them from
foreign countries;
I will bring
them to
their own land.
I will feed
them
on the mountains
of Israel,
by the streams
and all
the inhabited places
of the land.
\VS{14}In a good
pasture
I will feed
them; the
mountain
heights
of Israel
will be
their pasture.
There
they will lie
down in a lush
pasture,
and they will feed
on rich
grass on the mountains
of Israel.
\VS{15}I myself
will feed
my sheep
and I myself
will make them lie
down, declares
the sovereign

{\ND{Lord}}.
\VS{16}I will seek
the lost
and bring back
the strays;
I will bandage
the injured
and strengthen
the sick,
but the fat
and the strong
I will destroy.
I will feed them – with judgment!
\par }{\PP \VS{17}“‘As for you,
my sheep,
this is what
the sovereign

{\ND{Lord}}
says: Look,
I am about to judge
between
one sheep
and another,
between rams
and goats.
\VS{18}Is it not enough
for you to feed
on the good
pasture,
that you must trample
the rest
of your pastures
with your feet? When you drink
clean
water,
must you muddy
the rest
of the water by trampling it with your feet?
\VS{19}As for my sheep,
they must eat what you trampled
with your feet,
and drink
what you have muddied
with your feet!
\par }{\PP \VS{20}“‘Therefore,
this is what
the sovereign

{\ND{Lord}}
says
to
them: Look,
I myself
will judge
between
the fat
sheep
and
the lean
sheep.
\VS{21}Because
you push with your side
and your shoulder,
and thrust
your horns
at
all
the weak
sheep
until
you scatter
them abroad,
\VS{22}I will save
my sheep;
they will
no
longer
be
prey.
I will judge
between
one sheep
and another.
\par }{\PP \VS{23}I will set
one
shepherd
over
them, and he will feed them – namely, my servant David. He will feed them and will be their shepherd.
\VS{24}I,
the {\ND{Lord}}, will be
their God,
and my servant
David
will be prince
among
them; I,
the {\ND{Lord}},
have spoken!
\par }{\PP \VS{25}“‘I will make
a covenant
of peace
with them and will rid
the land
of wild beasts,
so that they can live
securely
in the wilderness
and even sleep
in the woods.
\VS{26}I
will turn
them and the regions around
my hill
into a blessing.
I will make showers
come down
in their season;
they will be
showers
that
bring blessing.
\VS{27}The trees
of the field
will
yield their fruit
and the earth
will yield
its crops. They will live securely
on
their land;
they will know
that
I am
the {\ND{Lord}},
when I break
the
bars
of their yoke
and rescue
them from the hand
of those who enslaved them.
\VS{28}They will
no
longer
be prey
for the nations
and the wild beasts
will not
devour
them. They will live
securely
and no
one will make them afraid.
\VS{29}I will prepare
for them a healthy
planting.
They will
no
longer
be victims
of famine
in the land
and will no
longer
bear
the insults
of the nations.
\VS{30}Then they will know
that
I,
the {\ND{Lord}}
their God,
am with
them, and that they
are my people,
the house
of Israel,
declares
the sovereign

{\ND{Lord}}.
\VS{31}And you,
my sheep,
the sheep
of my pasture,
are my people,
and I am
your God,
declares
the sovereign

{\ND{Lord}}.’ ”

\par }\Chap{35}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
Mount
Seir,
and prophesy
against it.
\VS{3}Say
to it, ‘This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “ ‘Look,
I am against you, Mount
Seir;
\par }{\Q I will stretch
out my hand
against
you
\par }{\Q and turn
you into a desolate
ruin.
\par }{\Q \VS{4}I will lay
waste
your cities;
\par }{\Q and you
will become desolate.
\par }{\Q Then
you will
know
that
I am
the {\ND{Lord}}!
\par }{\PP \VS{5}“‘You have shown
unrelenting
hostility
and poured
the people
of Israel
onto
the blades
of a sword
at the time
of their calamity,
at the time
of their final
punishment.
\VS{6}Therefore,
as surely
as I
live, declares
the sovereign

{\ND{Lord}}, I will subject
you to bloodshed,
and bloodshed
will pursue
you. Since
you did not
hate
bloodshed,
bloodshed
will pursue you.
\VS{7}I will turn
Mount
Seir
into a desolate
ruin;
I will cut off
from
it the one who passes
through or returns.
\VS{8}I will fill
its
mountains
with its dead;
on your hills
and in your valleys
and in all
your ravines,
those killed
by the sword
will fall.
\VS{9}I will turn you into a perpetual
desolation,
and your cities
will not
be inhabited.
Then you will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{10}“‘You said,
“These
two
nations,
these two
lands
will be
mine, and we will possess them,” – although the
{\ND{Lord}} was there –
\VS{11}therefore,
as surely
as I
live, declares
the sovereign

{\ND{Lord}}, I will deal
with you according to your anger,
and according to your envy,
by which
you acted
spitefully
against them. I will reveal
myself to them when
I judge you.
\VS{12}Then you will know
that
I,
the {\ND{Lord}}, have heard
all
the insults
you spoke
against
the mountains
of Israel,
saying,
“They are desolate,
they have been given
to us for food.”
\VS{13}You exalted
yourselves against
me with your speech
and hurled
many insults
against me – I have heard them all!
\VS{14}This is what
the sovereign

{\ND{Lord}}
says: While
the whole
earth
rejoices,
I will turn you into a desolation.
\VS{15}As
you rejoiced
over
the inheritance
of the house
of Israel
because it was desolate,
so
will I deal with you – you will be desolate, Mount Seir, and all of Edom – all of it! Then they will know that I am the
{\ND{Lord}}.’ ”

\par }\Chap{36}{\PP \VerseOne{1}“As for you,
son
of man,
prophesy
to
the mountains
of Israel,
and say: ‘O mountains
of Israel,
hear
the word
of the {\ND{Lord}}!
\VS{2}This is what
the sovereign

{\ND{Lord}}
says: The enemy
has spoken against
you, saying “Aha!” and, “The ancient
heights
have become our property!” ’
\VS{3}So
prophesy
and say: ‘This is what
the sovereign

{\ND{Lord}}
says: Surely because
they have made you desolate
and crushed
you from all directions,
so that you have become
the property
of the rest
of the nations,
and have become
the subject of gossip
and slander
among the people,
\VS{4}therefore,
O mountains
of Israel,
hear
the word
of the sovereign

{\ND{Lord}}: This is what
the sovereign

{\ND{Lord}}
says to the mountains
and hills,
the ravines
and valleys,
and to the desolate
ruins
and the abandoned
cities
that have
become
prey
and an object of derision
to the rest
of the nations
round about –
\VS{5}therefore
this is what
the sovereign

{\ND{Lord}}
says: Surely
I have spoken
in the fire
of my zeal
against
the rest
of the nations,
and against
all
Edom,
who
with great joy
and utter
contempt
have made
my land
their
property
and prey,
because
of its pasture.’
\par }{\PP \VS{6}“Therefore
prophesy
concerning
the land
of Israel,
and say
to the mountains
and hills,
the ravines
and valleys,
‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I have spoken
in my zeal
and in my anger,
because
you have endured
the insults
of the nations.
\VS{7}So
this is what
the sovereign

{\ND{Lord}}
says: I
vow
that the nations
around
you will endure
insults as well.
\par }{\PP \VS{8}“‘But you,
mountains
of Israel,
will grow
your branches,
and bear
your fruit
for my people
Israel;
for
they will arrive soon.
\VS{9}For indeed,
I am
on your side; I will turn
to
you, and you will be plowed
and planted.
\VS{10}I will multiply
your people –
the whole
house
of Israel,
all
of it. The cities
will be populated
and the ruins
rebuilt.
\VS{11}I will increase
the number of people
and animals
on
you; they will increase
and be fruitful.
I will cause you to be inhabited
as in ancient times,
and will do more good
for you than at the beginning
of your history. Then you will know
that
I
am the
{\ND{Lord}}.
\VS{12}I will lead
people,
my people
Israel,
across you; they will possess
you and you will become
their inheritance.
No
longer
will you bereave them of their children.
\par }{\PP \VS{13}“‘This is what
the sovereign

{\ND{Lord}}
says: Because
they are saying
to you, “You are a devourer
of men,
and bereave
your nation
of children,”
\VS{14}therefore
you will no
longer
devour
people
and no longer
bereave
your nation
of children, declares
the sovereign

{\ND{Lord}}.
\VS{15}I
will no
longer
subject
you to the nations’
insults;
no longer
will you bear
the shame
of the peoples,
and no
longer will
you bereave
your nation,
declares
the sovereign

{\ND{Lord}}.’ ”
\par }{\PP \VS{16}The word
of the {\ND{Lord}}
came to me:
\VS{17}“Son
of man,
when the house
of Israel
was living
on
their own land,
they defiled
it by their behavior
and their deeds.
In my sight
their behavior
was like the uncleanness
of a woman having her monthly period.
\VS{18}So I poured
my anger
on
them because
of the blood
they shed
on
the land
and because of the idols
with which they defiled it.
\VS{19}I scattered
them among the nations;
they were dispersed throughout foreign countries.
In accordance with their behavior
and their deeds
I judged them.
\VS{20}But when they arrived
in the nations
where
they went,
they profaned
my holy
name.
It was said
of them, ‘These
are the people
of the {\ND{Lord}}, yet they have departed
from his land.’
\VS{21}I was concerned
for my holy
reputation
which
the house
of Israel
profaned
among the nations
where
they went.
\par }{\PP \VS{22}“Therefore
say
to the house
of Israel,
‘This is what
the sovereign

{\ND{Lord}}
says: It is not
for your sake
that I am
about to act,
O house
of Israel,
but
for
the sake of my
holy
reputation
which
you profaned
among the nations
where
you went.
\VS{23}I will magnify
my great
name
that has been
profaned
among the nations,
that
you have
profaned
among
them. The nations
will know
that
I am
the {\ND{Lord}},
declares
the sovereign

{\ND{Lord}}, when
I magnify
myself among you in their sight.
\par }{\PP \VS{24}“‘I will take
you from
the nations
and gather
you from all
the countries;
then I will bring
you to
your land.
\VS{25}I will sprinkle
you with pure
water
and you will be clean
from all
your impurities.
I will purify
you from all
your idols.
\VS{26}I will give
you a new
heart,
and I will put
a new
spirit
within
you. I will
remove
the heart
of stone
from your body
and give
you a heart
of flesh.
\VS{27}I will put
my Spirit
within
you; I will take the initiative
and you will obey
my statutes
and carefully
observe
my regulations.
\VS{28}Then you will live
in the land
I gave
to your fathers;
you will be
my people,
and I
will be
your God.
\VS{29}I will save
you from all
your uncleanness.
I will call
for the grain
and multiply
it; I will
not
bring a famine
on
you.
\VS{30}I will multiply
the
fruit
of the trees
and the produce
of the fields,
so
that you will never
again
suffer
the disgrace
of famine
among the nations.
\VS{31}Then you will remember
your evil
behavior
and your deeds
which
were not
good;
you will loathe
yourselves on
account of your sins
and your abominable deeds.
\VS{32}Understand
that it is not
for your sake
I am
about to act,
declares
the sovereign

{\ND{Lord}}. Be ashamed
and embarrassed
by your behavior,
O house
of Israel.
\par }{\PP \VS{33}“‘This is what
the sovereign

{\ND{Lord}}
says: In the day
I cleanse
you from all
your sins,
I will populate
the cities
and the ruins
will be rebuilt.
\VS{34}The desolate
land
will be plowed,
instead
of being desolate
in the sight
of everyone
who passes by.
\VS{35}They will say,
“This desolate
land
has become
like the garden
of Eden;
the ruined,
desolate,
and destroyed
cities
are now fortified
and inhabited.”
\VS{36}Then the nations
which
remain
around
you will know
that
I,
the {\ND{Lord}},
have rebuilt
the ruins and replanted
what
was desolate.
I,
the {\ND{Lord}}, have spoken
– and I will do it!’
\par }{\PP \VS{37}“This is what
the sovereign

{\ND{Lord}}
says: I will allow
the house
of Israel
to ask
me to do
this
for them: I will multiply
their people
like sheep.
\VS{38}Like the sheep
for offerings,
like the sheep
of Jerusalem
during her appointed feasts,
so
will the ruined
cities
be filled
with flocks
of people.
Then they will know
that
I am
the {\ND{Lord}}.”

\par }\Chap{37}{\PP \VerseOne{1}The
hand
of the {\ND{Lord}}
was on
me, and he brought me out
by the Spirit
of the {\ND{Lord}}
and placed
me in the midst
of the valley,
and it was full
of bones.
\VS{2}He made me walk
all
around
among them. I realized there
were a great
many
bones in the valley
and they were very
dry.
\VS{3}He said
to me,
“Son
of man,
can these
bones
live?” I said
to him, “Sovereign
Lord,
you
know.”
\VS{4}Then he said
to me,
“Prophesy
over
these
bones,
and tell
them: ‘Dry
bones,
hear
the word
of the {\ND{Lord}}.
\VS{5}This is what
the sovereign

{\ND{Lord}}
says
to these
bones: Look,
I
am about to infuse
breath
into you and you will live.
\VS{6}I will put
tendons
on
you and muscles over
you and will cover
you with skin;
I will put
breath
in you and you will live.
Then you will know
that
I am
the {\ND{Lord}}.’ ”
\par }{\PP \VS{7}So I prophesied
as I was commanded.
There was
a sound
when I prophesied
– I
heard a rattling,
and the bones
came
together, bone
to bone.
\VS{8}As I watched,
I saw
tendons
on
them, then muscles appeared, and skin
covered
over them from above,
but there was no
breath in them.
\par }{\PP \VS{9}He said
to me,
“Prophesy
to
the breath, –
prophesy,
son
of man
– and say
to
the breath: ‘This is what
the sovereign

{\ND{Lord}}
says: Come
from the four
winds,
O breath,
and breathe
on these
corpses so that they may live.’ ”
\VS{10}So I prophesied
as
I was commanded,
and the breath
came into
them; they lived
and stood
on
their feet,
an extremely
great
army.
\par }{\PP \VS{11}Then he said
to
me, “Son
of man,
these
bones
are all
the house
of Israel.
Look,
they
are saying,
‘Our bones
are dry,
our hope
has perished;
we are cut off.’
\VS{12}Therefore
prophesy,
and tell
them,
‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I am
about to open
your graves
and will raise you from
your graves,
my people.
I will bring
you to
the land
of Israel.
\VS{13}Then you will know
that
I am
the {\ND{Lord}}, when I open
your graves
and raise you from
your graves,
my people.
\VS{14}I will place
my breath
in you and you will live;
I will give you rest
in your own land.
Then you will know
that
I am
the {\ND{Lord}} –
I have spoken
and I will act,
declares
the {\ND{Lord}}.’ ”
\par }{\PP \VS{15}The word
of the {\ND{Lord}}
came to me:
\VS{16}“As for you,
son
of man,
take
one
branch,
and write
on
it, ‘For Judah,
and for the Israelites
associated
with him.’ Then take
another
branch
and write
on
it, ‘For Joseph,
the branch
of Ephraim
and all
the house
of Israel
associated with him.’
\VS{17}Join
them as one
stick;
they will be
as one
in your hand.
\VS{18}When your people
say
to you, ‘Will you not
tell
us what
these things mean?’
\VS{19}tell
them,
‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I am
about to take
the
branch
of Joseph
which
is in the hand
of Ephraim
and the tribes
of Israel
associated
with him, and I will place
them
on
the stick
of Judah,
and make
them into one
stick
– they will be
one
in my hand.’
\VS{20}The sticks
you write
on
will be
in your hand
in front of them.
\VS{21}Then tell
them,
‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I am
about to take
the Israelites
from among
the nations
where
they have
gone.
I will gather
them from round
about and bring
them to
their land.
\VS{22}I will make
them
one
nation
in the land,
on the mountains
of Israel,
and one
king
will
rule
over them all.
They will never
again be
two
nations
and never
again
be divided
into two
kingdoms.
\VS{23}They will not
defile
themselves with their idols,
their detestable things,
and all
their rebellious
deeds. I will save
them from all
their unfaithfulness
by which
they sinned.
I will purify
them; they will become
my people
and I
will become
their God.
\par }{\PP \VS{24}“‘My servant
David
will be king
over
them; there will be
one
shepherd
for all
of them. They will follow
my regulations
and carefully
observe
my statutes.
\VS{25}They will live
in the land
I gave
to my servant
Jacob,
in which
your fathers
lived;
they will live
in it
– they
and their children
and their grandchildren
forever.
David
my servant
will be prince
over
them forever.
\VS{26}I will make
a covenant
of peace
with them; it will be
a perpetual
covenant
with
them. I will
establish them, increase
their numbers,
and place my sanctuary
among
them forever.
\VS{27}My dwelling place
will be with them; I will be
their God,
and they
will be
my people.
\VS{28}Then, when
my sanctuary
is among
them forever,
the nations
will know
that
I,
the {\ND{Lord}},
sanctify
Israel.’ ”

\par }\Chap{38}{\PP \VerseOne{1}The word
of the {\ND{Lord}}
came to me:
\VS{2}“Son
of man,
turn
toward
Gog,
of the land
of Magog,
the chief
prince
of Meshech
and Tubal.
Prophesy
against him
\VS{3}and say: ‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I am against you, Gog,
chief
prince
of Meshech
and Tubal.
\VS{4}I will turn
you around,
put
hooks
into your jaws,
and bring
you out
with all
your army,
horses
and horsemen,
all
of them fully
armed,
a great
company
with shields
of different types,
all
of them armed
with swords.
\VS{5}Persia,
Ethiopia,
and Put
are with
them, all
of them with shields
and helmets.
\VS{6}They are joined by Gomer
with all
its troops,
and by Beth
Togarmah
from the remote
parts of the north
with all
its troops
– many
peoples
are with you.
\par }{\PP \VS{7}“‘Be ready
and stay ready,
you
and all
your companies
assembled
around you, and be
a guard for them.
\VS{8}After many
days
you will be summoned;
in the latter
years
you will come
to
a land
restored
from the ravages of war,
with many
peoples
gathered
on
the mountains
of Israel
that
had long
been
in ruins.
Its people
were brought out
from the peoples, and all
of them will be living
securely.
\VS{9}You will advance;
you will come
like a storm.
You will be
like a cloud
covering
the earth,
you,
all
your troops,
and the many
other peoples
with you.
\par }{\PP \VS{10}“‘This is what
the sovereign

{\ND{Lord}}
says: On that day
thoughts
will come
into your mind,
and you will devise
an evil
plan.
\VS{11}You will say,
“I will invade
a land
of unwalled
towns; I will advance
against those living
quietly
in security
– all
of them living
without
walls
and barred
gates –
\VS{12}to loot
and plunder,
to attack
the inhabited
ruins
and the people
gathered
from the nations,
who are acquiring
cattle
and goods,
who live
at
the center
of the earth.”
\VS{13}Sheba
and Dedan
and the traders
of Tarshish
with all
its young warriors
will say
to you, “Have you come
to loot? Have you
assembled
your armies
to plunder,
to carry away
silver
and gold,
to take away
cattle
and goods,
to haul away
a great amount
of spoils?” ’
\par }{\PP \VS{14}“Therefore,
prophesy,
son
of man,
and say
to Gog: ‘This is what
the sovereign

{\ND{Lord}}
says: On that day
when
my people
Israel
are living
securely,
you will take notice
\VS{15}and come
from your place,
from the remote
parts of the north,
you
and many
peoples
with
you, all
of them riding
on horses,
a great
company
and a vast
army.
\VS{16}You will advance
against
my people
Israel
like a cloud
covering
the earth.
In the latter
days
I will
bring
you against
my land
so
that the nations
may acknowledge
me, when before their eyes
I magnify
myself through you, O Gog.
\par }{\PP \VS{17}“‘This is what
the sovereign

{\ND{Lord}}
says: Are you
the one of whom
I spoke
in former
days
by
my servants
the prophets
of Israel,
who prophesied
in those
days
that I would bring
you against them?
\VS{18}On
that day, when
Gog
invades
the land
of Israel,
declares
the sovereign

{\ND{Lord}}, my rage
will mount up
in my anger.
\VS{19}In my zeal,
in the fire
of my fury,
I declare
that on that day
there will be
a great
earthquake
in the land
of Israel.
\VS{20}The fish
of the sea,
the birds
of the sky,
the wild
beasts,
all
the things
that creep
on
the ground,
and all
people
who
live on
the face
of the earth
will shake
at my presence.
The mountains
will topple,
the cliffs
will fall,
and every
wall
will fall
to the ground.
\VS{21}I will call
for a sword
to attack Gog on
all
my mountains,
declares
the sovereign

{\ND{Lord}}; every man’s
sword
will be
against his brother.
\VS{22}I will judge
him
with plague
and bloodshed.
I will rain
down
on him,
his troops
and the many
peoples
who
are with
him a torrential
downpour, hailstones,
fire,
and brimstone.
\VS{23}I will exalt and magnify
myself;
I will reveal myself
before
many
nations.
Then they will know
that
I am
the {\ND{Lord}}.’

\par }\Chap{39}{\PP \VerseOne{1}“As for you,
son
of man,
prophesy
against
Gog,
and say: ‘This is what
the sovereign

{\ND{Lord}}
says: Look,
I am against you, O Gog,
chief
prince
of Meshech
and Tubal!
\VS{2}I will turn
you around and drag you along; I will lead you up
from the remotest
parts
of the north
and bring
you against
the mountains
of Israel.
\VS{3}I will knock
your bow
out of your left hand
and make your arrows
fall
from your right hand.
\VS{4}You will fall dead
on
the mountains
of Israel,
you
and all
your troops
and the people
who
are with
you. I give
you as food
to every
kind
of bird
and every wild
beast.
\VS{5}You will fall
dead in the open
field;
for
I
have spoken,
declares
the sovereign

{\ND{Lord}}.
\VS{6}I will send
fire
on Magog
and those who live
securely
in the coastlands;
then they will know
that
I am
the {\ND{Lord}}.
\par }{\PP \VS{7}“‘I will make my holy
name
known
in the midst
of my people
Israel;
I will not
let my holy
name
be profaned
anymore.
Then the nations
will know
that
I am
the {\ND{Lord}}, the Holy One
of Israel.
\VS{8}Realize
that it is coming
and it will be
done, declares
the sovereign

{\ND{Lord}}. It is the day
I have
spoken about.
\par }{\PP \VS{9}“‘Then those who live
in the cities
of Israel
will go out
and use the weapons
for kindling –
the shields,
bows
and arrows,
war clubs
and spears
– they will burn
them for seven
years.
\VS{10}They will not
need to take
wood
from
the field
or
cut down
trees from
the forests,
because
they
will make fires
with the weapons.
They
will take
the loot
from those who looted
them
and seize the plunder
of those who plundered them, declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{11}“‘On that day
I will assign
Gog
a grave
in Israel.
It will be the valley
of those who travel
east
of the sea;
it will block
the way of the travelers.
There
they will bury
Gog
and all
his horde;
they will call
it the valley
of Hamon-Gog.
\VS{12}For seven
months
Israel
will bury
them, in order
to cleanse
the land.
\VS{13}All
the people
of the land
will bury
them, and it will be
a memorial
for them on the day
I magnify
myself, declares
the sovereign

{\ND{Lord}}.
\VS{14}They will designate
men
to scout continually
through
the land,
burying
those who remain
on
the surface
of the ground,
in order to cleanse
it. They will search
for seven
full months.
\VS{15}When
the scouts
survey the land
and see
a human
bone,
they will place a sign
by
it, until
those assigned to
burial
duty have buried
it in the valley
of Hamon-Gog.
\VS{16}(A city
by the name
of Hamonah
will also be there.) They will cleanse
the land.’
\par }{\PP \VS{17}“As for you,
son
of man,
this is what
the sovereign

{\ND{Lord}}
says: Tell every kind
of bird
and every
wild
beast: ‘Assemble
and come! Gather
from all around
to my slaughter
which
I am
going to make for you, a great
slaughter
on
the mountains
of Israel! You will eat
flesh
and drink
blood.
\VS{18}You will eat
the flesh
of warriors
and drink
the blood
of the princes
of the earth
– the rams,
lambs,
goats,
and bulls,
all
of them fattened animals
of Bashan.
\VS{19}You will eat
fat
until you are full,
and drink
blood
until you are drunk,
at my slaughter
which I have made for you.
\VS{20}You will fill
up at my table
with horses
and charioteers,
with warriors
and all
the soldiers,’
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{21}“I will display
my majesty
among the nations.
All
the nations
will witness
the judgment
I have executed,
and the power
I have
exhibited among them.
\VS{22}Then the house
of Israel
will know
that
I am
the {\ND{Lord}}
their God,
from
that day
forward.
\VS{23}The nations
will know
that
the house
of Israel
went into exile
due to their iniquity,
for they were unfaithful
to me. So I hid
my face
from them
and handed
them over
to their enemies;
all
of them died
by the sword.
\VS{24}According to their uncleanness
and rebellion
I have dealt
with them,
and I hid
my face
from them.
\par }{\PP \VS{25}“Therefore
this is what
the sovereign

{\ND{Lord}}
says: Now
I will restore
the fortunes
of Jacob,
and I will have mercy
on the entire
house
of Israel.
I will be zealous
for my holy
name.
\VS{26}They will bear
their shame
for all
their unfaithful
acts against me, when
they live
securely
on
their land
with no
one to make them afraid.
\VS{27}When I have brought them
back
from
the peoples
and gathered
them from the countries
of their enemies,
I will magnify
myself among them in the sight
of many
nations.
\VS{28}Then they will know
that
I am
the {\ND{Lord}}
their God,
because I sent them
into exile
among the nations,
and then gathered
them into their own land.
I will not
leave any
of them
in exile
any longer.
\VS{29}I will no
longer
hide
my face
from them,
when I pour
out my Spirit
on
the house
of Israel,
declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{40}{\PP \VerseOne{1}In the twenty-fifth
year
of our exile,
at the beginning
of the year,
on the tenth
day of the month,
in the fourteenth
year
after
the city
was struck
down, on
this very
day,
the hand
of the {\ND{Lord}}
was
on
me, and he brought
me there.
\VS{2}By means of divine
visions
he brought
me to
the land
of Israel
and placed
me on
a very
high
mountain,
and on
it was a structure
like a city,
to the south.
\VS{3}When he brought
me there,
I saw
a man
whose appearance
was like bronze,
with a linen
cord
and a measuring
stick
in his hand.
He was
standing
in the gateway.
\VS{4}The man
said
to me,
“Son
of man,
watch
closely,
listen
carefully, and pay
attention
to everything
I
show
you, for
you have been brought
here
so
that I can show
it to you. Tell
the
house
of Israel
everything
you
see.”
\par }{\PP \VS{5}I saw
a wall
all
around
the outside
of the temple.
In the man’s
hand
was a measuring
stick
10½ feet
long.
He measured
the
thickness
of the wall
as 10½ feet,
and its height
as 10½ feet.
\VS{6}Then he went
to
the gate
facing
east.
He climbed
its steps
and measured
the threshold
of the gate
as 10½
feet deep.
\VS{7}The alcoves
were 10½
feet
long
and 10½
feet
wide;
between
the alcoves
were 8¾
feet.
The threshold
of the gate
by
the porch
of the gate
facing inward
was 10½
feet.
\VS{8}Then he measured
the porch
of the gate
facing inward as 10½ feet.
\VS{9}He measured
the porch
of the gate
as 14 feet,
and its jambs
as 3½ feet;
the porch
of the gate faced inward.
\VS{10}There were three
alcoves
on each side
of the east
gate;
the three
had the same
measurement,
and the jambs
on either side
had the same
measurement.
\VS{11}He measured
the width
of the entrance
of the gateway
as 17½ feet,
and the length
of the gateway
as 22¾ feet.
\VS{12}There was a barrier
in front
of the alcoves,
1¾ feet
on either side;
the alcoves
were 10½ feet
on either side.
\VS{13}He measured
the gateway
from the roof
of one alcove
to the roof
of the other, a width
of 43¾
feet
from one entrance
to the opposite
one.
\VS{14}He measured
the porch
at 105 feet
high; the gateway
went all around
to
the jamb
of the courtyard.
\VS{15}From the front
of the entrance
gate
to the porch
of the inner
gate
was 87½ feet.
\VS{16}There were closed
windows
toward
the alcoves
and toward
their jambs
within
the gate
all
around,
and likewise
for the porches.
There were windows
all
around
the inside,
and on
each jamb
were decorative palm trees.
\par }{\PP \VS{17}Then he brought
me to
the outer
court.
I saw
chambers
there, and a pavement
made
for the court
all
around;
thirty
chambers
faced the pavement.
\VS{18}The pavement
was beside
the gates,
corresponding
to the length
of the gates;
this was the lower
pavement.
\VS{19}Then he measured
the width
from before
the lower
gate
to the front
of the exterior
of the inner
court
as
175 feet
on the east
and on the north.
\par }{\PP \VS{20}He measured
the length
and width
of the gate
of the outer
court
which
faces
north.
\VS{21}Its alcoves,
three
on each side,
and its jambs
and porches
had the same measurement
as the first
gate;
87½ feet
long
and 43¾ feet
wide.
\VS{22}Its windows,
its porches,
and its decorative palm trees
had the same measurement
as the gate
which
faced
east.
Seven
steps
led up
to it, and its porch
was in front of them.
\VS{23}Opposite
the gate
on the north
and the east
was a gate
of the inner
court;
he measured
the distance from gate
to
gate
at 175 feet.
\par }{\PP \VS{24}Then
he led
me toward
the south.
I saw
a gate
on
the south.
He measured
its jambs
and its porches;
they had the same dimensions
as the others.
\VS{25}There were windows
all around
it and its porches,
like the windows
of the others;
87½ feet
long
and 43¾ feet
wide.
\VS{26}There were seven
steps
going up
to it; its porches
were in front
of them. It had decorative palm trees
on its jambs,
one
on
either side.
\VS{27}The inner
court
had a gate
toward
the south;
he measured
it from gate
to
gate
toward
the south
as 175 feet.
\par }{\PP \VS{28}Then he brought
me to
the inner
court
by the south
gate.
He measured
the south
gate;
it had the same dimensions
as the others.
\VS{29}Its alcoves,
its jambs,
and its porches
had the same dimensions
as the others, and there were windows
all
around
it and its porches;
its length
was 87½ feet
and its width
43¾
feet.
\VS{30}There were porches
all
around,
43¾ feet
long
and 8¾ feet
wide.
\VS{31}Its porches
faced
the outer
court,
and decorative palm trees
were on
its jambs,
and its stairway
had eight
steps.
\par }{\PP \VS{32}Then he brought
me to
the inner
court
on
the east side.
He measured
the gate;
it had the same dimensions
as the others.
\VS{33}Its alcoves,
its jambs,
and its porches
had the same dimensions
as the others, and there were windows
all
around
it and its porches;
its length
was 87½ feet
and its width
43¾
feet.
\VS{34}Its porches
faced the outer
court,
it had decorative palm trees
on
its jambs,
and its stairway
had eight
steps.
\par }{\PP \VS{35}Then he brought
me to
the north
gate,
and he measured
it; it had the same dimensions
as the others –
\VS{36}its alcoves,
its jambs,
and its porches.
It had windows
all
around
it; its length
was 87½ feet
and its width
43¾
feet.
\VS{37}Its jambs
faced the outer
court,
and it had decorative palm trees
on
its jambs,
on either side,
and its stairway
had eight
steps.
\par }{\PP \VS{38}There was a chamber
with its door
by the porch
of the gate;
there
they washed
the burnt offering.
\VS{39}In the porch
of the gate
were two
tables
on either side
on which to slaughter
the burnt offering,
the sin offering,
and the guilt offering.
\VS{40}On
the outside
of the porch as one goes up
at the entrance
of the north
gate
were two
tables,
and on
the other
side
of the porch
of the gate
were two
tables.
\VS{41}Four
tables
were on each side
of the gate,
eight
tables
on which the sacrifices were to
be slaughtered.
\VS{42}The four
tables
for the burnt offering
were of carved
stone,
32 inches
long,
32 inches
wide,
and 21 inches
high.
They would put
the
instruments
which
they used to slaughter
the
burnt offering
and the sacrifice on them.
\VS{43}There were hooks
three inches long,
fastened
in the house
all around,
and on
the tables
was the flesh
of the offering.
\par }{\PP \VS{44}On the outside
of the inner
gate
were chambers
for the singers
of the inner
court,
one at
the side
of the north
gate
facing
south,
and the other
at
the side
of the south
gate
facing
north.
\VS{45}He said
to me,
“This
chamber
which
faces
south
is for the priests
who keep
charge
of the temple,
\VS{46}and the chamber
which
faces
north
is for the priests
who keep
charge
of the altar.
These
are the descendants
of Zadok,
from the descendants
of Levi,
who may approach
the {\ND{Lord}}
to minister to him.”
\VS{47}He measured
the court
as a square
175
feet long
and 175 feet
wide;
the altar
was in front
of the temple.
\par }{\PP \VS{48}Then he brought
me to
the porch
of the temple
and measured
the jambs
of the porch
as 8¾ feet
on either side,
and the width
of the gate
was 24½ feet
and the sides
were 5¼ feet
on each side.
\VS{49}The length
of the porch
was 35 feet
and the width
19¼ feet;
steps
led up
to
it, and there were pillars
beside
the jambs
on either side.

\par }\Chap{41}{\PP \VerseOne{1}Then he brought
me to
the outer sanctuary,
and measured
the jambs;
the jambs were 10½
feet
wide
on each side.
\VS{2}The width
of the entrance
was 17½ feet,
and the sides
of the entrance
were 8¾ feet
on each side.
He measured
the length
of the outer sanctuary as 70 feet,
and its width
as 35 feet.
\par }{\PP \VS{3}Then he went
into the inner
sanctuary and measured
the jambs
of the entrance
as 3½ feet,
the entrance
as 10½ feet,
and the width
of the entrance
as 12¼ feet
\VS{4}Then he measured
its length
as 35 feet,
and its width
as 35 feet,
before
the outer sanctuary.
He said
to me,
“This
is the most
holy place.”
\par }{\PP \VS{5}Then he measured
the wall
of the temple
as 10½ feet,
and the width
of the side chambers
as 7 feet,
all
around
the temple.
\VS{6}The side
chambers
were in three
stories,
one above the other, thirty
in each
story. There were offsets
in the wall
all
around
to serve
as supports
for the side chambers,
so that the supports
were not
in the wall
of the temple.
\VS{7}The side chambers
surrounding
the temple were wider
at each successive
story;
for the structure surrounding the temple
went up
story
by story
all around
the temple.
For this reason
the width
of the temple
increased as it went up,
and one went up
from the lowest
story to the highest
by the way of the middle story.
\par }{\PP \VS{8}I saw
that the temple
had a raised platform all
around;
the foundations
of the side chambers
were a full
measuring
stick of 10½ feet
high.
\VS{9}The width
of the outer
wall
of the side
chambers
was 8¾ feet,
and the open area between the side chambers
of the temple
\VS{10}and the chambers
of the court was 35 feet
in width
all around
the temple
on
every side.
\VS{11}There were entrances
from the side
chambers
toward the open area, one
entrance
toward
the north,
and another entrance
toward
the south;
the width
of the open area
was 8¾ feet
all
around.
\par }{\PP \VS{12}The building
that
was facing
the temple courtyard
at the west
side
was 122½ feet
wide;
the wall
of the building
was 8¾ feet
all around,
and its length
157½ feet.
\par }{\PP \VS{13}Then he measured
the temple
as 175 feet
long,
the courtyard
of the temple and the building
and its walls
as 175 feet
long,
\VS{14}and also the width
of the front
of the temple
and the courtyard
on the east
as 175 feet.
\par }{\PP \VS{15}Then he measured
the length
of the building
facing
the courtyard
at the rear
of the temple, with its galleries
on either side
as 175 feet.
\par }{\PP The interior
of the outer sanctuary
and the porch
of the court,
\VS{16}as well as the thresholds,
narrow
windows
and galleries
all around
on three
sides facing
the threshold
were paneled
with wood
all
around,
from the ground
up
to the windows
(now the windows
were covered),
\VS{17}to the space above
the entrance,
to
the inner
room,
and on the outside,
and on
all
the walls
in the inner
room and outside,
by measurement.
\VS{18}It was made
with cherubim
and decorative palm trees,
with a palm tree
between
each cherub.
Each cherub
had two
faces:
\VS{19}a human
face
toward
the palm tree
on one side
and a lion’s
face
toward
the palm tree
on the other side.
They were carved
on
the whole
temple
all
around;
\VS{20}from the ground
to the area above
the entrance,
cherubim
and decorative palm trees
were carved
on the wall
of the outer sanctuary.
\VS{21}The doorposts
of the outer sanctuary
were square.
In front
of the sanctuary
one doorpost
looked just like the other.
\VS{22}The altar
was of wood,
5¼ feet
high,
with its length
3½ feet;
its corners,
its length,
and its walls
were of wood.
He said
to me,
“This
is the table
that
is before
the
{\ND{Lord}}.”
\VS{23}The outer sanctuary
and the inner sanctuary
each had a double
door.
\VS{24}Each
of the doors
had two
leaves, two
swinging
leaves;
two
leaves
for one
door and two
leaves
for the other.
\VS{25}On the doors
of the outer sanctuary
were carved
cherubim
and palm trees,
like those carved
on the walls,
and there was a canopy
of wood
on the front
of the outside
porch.
\VS{26}There were narrow
windows
and decorative palm trees
on either side
of the side
walls of the porch;
this is what the side chambers
of the temple
and the canopies were like.

\par }\Chap{42}{\PP \VerseOne{1}Then he led me out
to
the outer
court,
toward
the north,
and brought
me to
the chamber
which
was opposite
the courtyard
and opposite
the building
on the north.
\VS{2}Its length
was 175 feet
on the north
side, and its width
87½ feet.
\VS{3}Opposite
the 35 feet
that belonged
to the inner
court,
and opposite
the pavement
which
belonged to the outer
court,
gallery
faced
gallery
in the three stories.
\VS{4}In front
of the chambers
was a walkway
on
the inner
side, 17½ feet
wide
at a distance
of 1¾ feet,
and their entrances
were on the north.
\VS{5}Now the upper
chambers
were
narrower,
because
the galleries
took more space
from them
than from the lower
and middle
chambers of the building.
\VS{6}For
they
were in three
stories and had no
pillars
like the pillars
of the courts;
therefore
the upper chambers were set back
from the ground
more than the lower
and upper ones.
\VS{7}As
for the outer
wall
by the
side of the chambers,
toward
the outer
court
facing
the chambers,
it was 87½ feet
long.
\VS{8}For
the chambers
on the outer
court
were 87½ feet
long, while
those facing
the temple
were 175 feet
long.
\VS{9}Below
these
chambers
was a passage
on the east side
as one enters
from the outer
court.
\par }{\PP \VS{10}At the beginning
of the wall
of the court
toward
the south,
facing
the courtyard
and the building,
were chambers
\VS{11}with a passage
in front
of them. They looked
like the chambers
on
the north.
Of the same
length
and width,
and all
their exits
according to their arrangements
and entrances
\VS{12}were the chambers
which
were toward
the south.
There was an opening
at the head
of the passage,
the passage
in front
of the corresponding
wall
toward
the east
when one enters.
\par }{\PP \VS{13}Then he said
to me,
“The north
chambers
and the south
chambers
which
face
the courtyard
are
holy
chambers
where
the priests
who
approach
the
{\ND{Lord}}
will eat
the most
holy
offerings. There
they will place
the most
holy offerings – the grain offering, the sin offering, and the guilt offering, because the place is holy.
\VS{14}When the priests
enter,
then they will not
go out
from the sanctuary
to
the outer
court
without taking
off their garments
in which
they minister,
for
these are
holy;
they
will put on
other
garments,
then they
will go
near
the places where
the people are.”
\par }{\PP \VS{15}Now when he had finished
measuring
the interior
of the temple,
he led
me out
by
the gate
which
faces
east
and measured
all
around.
\VS{16}He measured
the east
side with the measuring
stick
as 875 feet
by
the measuring
stick.
\VS{17}He measured
the north
side as 875
feet by the measuring
stick.
\VS{18}He measured
the south
side as 875
feet by the measuring stick.
\VS{19}He turned
to
the west
side
and measured
875
feet by the measuring stick.
\VS{20}He measured
it on all
four
sides. It
had a wall
around
it, 875
feet long
and 875
feet wide,
to separate
the holy
and common places.

\par }\Chap{43}{\PP \VerseOne{1}Then
he brought
me to
the gate
that
faced
toward
the east.
\VS{2}I saw
the glory
of the God
of Israel
coming
from the east;
the sound
was
like that of rushing
water;
and the earth
radiated
his glory.
\VS{3}It was like
the vision
I saw
when he came
to destroy
the city,
and the vision
I saw
by the Kebar
River.
I threw
myself face
down.
\VS{4}The glory
of the {\ND{Lord}}
came
into the temple
by way
of the gate
that
faces
east.
\VS{5}Then a wind
lifted
me up and brought
me to
the inner
court;
I watched
the glory
of the {\ND{Lord}}
filling
the temple.
\par }{\PP \VS{6}I heard
someone speaking
to
me from the temple,
while the man
was standing
beside me.
\VS{7}He said
to me: “Son
of man,
this is the place
of my throne
and the place
for the soles
of my feet,
where
I will live
among
the people of
Israel
forever.
The house
of
Israel
will no
longer
profane
my holy
name,
neither they
nor their kings,
by their spiritual prostitution
or by the pillars
of their kings set up when they die.
\VS{8}When they placed
their threshold
by my threshold
and their doorpost
by my doorpost,
with only the wall
between
me and them, they profaned
my holy
name
by the abominable deeds
they committed. So
I consumed
them
in my anger.
\VS{9}Now
they must
put away
their spiritual
prostitution
and the pillars
of their kings
far
from
me,
and then I will live
among
them forever.
\par }{\PP \VS{10}“As for you,
son
of man,
describe
the
temple
to the house
of Israel,
so that they will be ashamed
of their sins
and measure
the
pattern.
\VS{11}When
they are ashamed
of all
that
they have done,
make known to them the design
of the temple,
its pattern,
its exits
and entrances,
and its whole
design
– all
its statutes,
its entire
design,
and all
its laws;
write
it all down in their sight,
so that they may observe
its entire
design
and all
its statutes
and do them.
\par }{\PP \VS{12}“This
is the law
of the temple: The entire
area
on
top
of the mountain
all
around
will be most
holy.
Indeed,
this
is the law
of the temple.
\par }{\SH The Altar
\par }{\PP \VS{13}“And these
are the measurements
of the altar: Its base
is 1¾ feet
high, and 1¾ feet
wide,
and its border
nine inches
on
its edge.
This
is to be the height
of the altar.
\VS{14}From the base
of the ground
to
the lower
edge
is 3½ feet,
and the width
1¾ feet;
and from the smaller
ledge
to
the larger
edge,
7 feet,
and the width
1¾ feet;
\VS{15}and the altar hearth,
7 feet,
and from the altar hearth
four
horns
projecting upward.
\VS{16}Now the altar hearth
is a perfect
square,
21 feet
long
and 21 feet
wide.
\VS{17}The ledge
is 24½ feet
long
and 24½ feet
wide
on
four
sides;
the border
around
it is 10½ inches,
and its surrounding
base
1¾ feet.
Its steps
face
east.”
\par }{\PP \VS{18}Then he said
to me: “Son
of man,
this is what
the sovereign

{\ND{Lord}}
says: These
are the statutes
of the altar: On the day
it
is built
to offer up
burnt offerings
on it
and to sprinkle
blood
on it,
\VS{19}you will give
a young
bull
for a sin offering
to
the Levitical
priests
who
are
descended
from Zadok,
who approach
me to minister
to me,
declares
the sovereign

{\ND{Lord}}.
\VS{20}You will take
some of its blood,
and place
it on
the four
horns
of the altar, on the four
corners
of the ledge,
and on the border
all around;
you will cleanse
it and make atonement for it.
\VS{21}You will also take
the bull
for the sin offering,
and it will be burned
in the appointed place
in the temple,
outside
the sanctuary.
\par }{\PP \VS{22}“On the second
day,
you will offer
a male
goat
without blemish
for a sin offering.
They will purify
the
altar
just
as they purified
it with the bull.
\VS{23}When you have finished
purifying
it,
you will offer
an unblemished young
bull
and an unblemished
ram
from
the flock.
\VS{24}You will present
them
before
the {\ND{Lord}}, and the priests
will scatter
salt
on
them
and offer them
up
as a burnt offering
to the
{\ND{Lord}}.
\par }{\PP \VS{25}“For seven
days
you will provide
every day
a goat
for a sin offering;
a young bull
and a ram
from
the flock,
both without blemish,
will be provided.
\VS{26}For seven
days
they will make atonement
for the altar
and cleanse
it, so they will consecrate it.
\VS{27}When
the
prescribed
period
is over, on
the eighth
day
and thereafter
the priests
will offer up on
the altar
your burnt offerings
and your peace offerings;
I will accept
you, declares
the sovereign

{\ND{Lord}}.”

\par }\Chap{44}{\PP \VerseOne{1}Then he brought me back
by way
of the outer
gate
of the sanctuary
which faces
east,
but it was
shut.
\VS{2}The
{\ND{Lord}}
said
to
me: “This
gate
will be shut;
it will not
be opened,
and no
one will enter
by it. For
the {\ND{Lord}}, the God
of Israel,
has entered
by it; therefore it will remain shut.
\VS{3}Only the prince
may sit
in it to eat
a sacrificial meal
before
the {\ND{Lord}}; he will enter
by way
of the porch
of the gate
and will go out
by the same way.”
\par }{\PP \VS{4}Then he brought
me by way
of the north
gate
to
the front
of the temple.
As I watched,
I noticed
the glory
of the {\ND{Lord}}
filling
the
{\ND{Lord}}’s
temple,
and I threw
myself face down.
\VS{5}The
{\ND{Lord}}
said
to me: “Son
of man,
pay
attention,
watch
closely
and listen
carefully to everything
I
tell
you concerning all
the statutes
of the
{\ND{Lord}}’s
house
and all
its laws.
Pay
attention
to the entrances
to the temple
with all
the exits
of the sanctuary.
\VS{6}Say
to
the rebellious,
to
the house
of Israel,
‘This is what
the sovereign

{\ND{Lord}}
says: Enough
of all
your abominable practices,
O house
of Israel!
\VS{7}When you bring
foreigners,
those uncircumcised
in heart
and in flesh,
into my sanctuary,
you desecrate it – even my house – when you offer my food, the fat and the blood. You have broken my covenant by all your abominable practices.
\VS{8}You have not
kept
charge
of my holy
things, but you have assigned
foreigners to keep
charge
of my sanctuary for you.
\VS{9}This is what
the sovereign

{\ND{Lord}}
says: No
foreigner,
who is uncircumcised
in heart
and flesh
among all
the foreigners
who
are among
the people
of Israel,
will enter
into
my sanctuary.
\par }{\PP \VS{10}“‘But
the Levites
who
went
far
from me, straying
off from me after
their idols
when Israel
went astray,
will be responsible
for their sin.
\VS{11}Yet they will be
ministers
in my sanctuary,
having oversight
at the gates
of the temple,
and serving
the
temple.
They
will slaughter
the
burnt offerings
and the
sacrifices
for the people,
and they
will stand
before
them to minister to them.
\VS{12}Because
they used to minister
to them
before
their idols,
and became
a sinful
obstacle
to the house
of Israel,
consequently
I have made a vow concerning
them,
declares
the sovereign

{\ND{Lord}}, that they
will be responsible
for their sin.
\VS{13}They will not
come near
me to
serve
me as priest, nor will they come near
any
of my holy
things, the things which
are
most sacred.
They
will bear
the shame
of the abominable deeds
they have committed.
\VS{14}Yet I will appoint
them
to keep
charge
of the temple,
all
of its service
and all
that
will be done in it.
\par }{\SH The Levitical Priests
\par }{\PP \VS{15}“‘But
the Levitical
priests,
the descendants
of Zadok
who
kept
the
charge
of my sanctuary
when the people of Israel
went astray
from me, will approach
me to
minister
to me; they will stand
before
me to offer
me the fat
and the blood,
declares
the sovereign

{\ND{Lord}}.
\VS{16}They
will enter
my sanctuary,
and approach
my table
to minister
to me;
they will keep
my charge.
\par }{\PP \VS{17}“‘When
they enter
the gates
of the inner
court,
they must wear
linen
garments;
they must not
have any wool
on them when they minister
in the inner
gates
of the court
and in the temple.
\VS{18}Linen
turbans
will be
on
their heads
and linen
undergarments
will be
around their waists;
they must not
bind themselves with anything that causes sweat.
\VS{19}When they go out
to
the outer
court
to
the people,
they must remove
the
garments
they
were ministering
in, and place them in
the
holy
chambers;
they must put on
other
garments
so that they will not
transmit holiness
to the people
with their garments.
\par }{\PP \VS{20}“‘They must not
shave
their heads
nor
let their hair
grow long; they must only trim
their heads.
\VS{21}No
priest
may drink
wine
when he enters
the inner
court.
\VS{22}They must not
marry
a widow
or a divorcee,
but
they may
marry a virgin
from the house
of Israel
or a widow
who is a priest’s
widow.
\VS{23}Moreover, they will teach
my people
the difference
between
the holy
and the common, and show them how to distinguish
between
the ceremonially unclean
and the clean.
\par }{\PP \VS{24}“‘In a controversy
they
will act as judges;
they will judge
according to my ordinances.
They will keep my laws
and my statutes
regarding all
my appointed festivals
and will observe
my Sabbaths.
\par }{\PP \VS{25}“‘They must not
come
near a dead
person
or
they will be defiled;
however,
for
father,
mother,
son,
daughter,
brother
or
sister,
they may
defile themselves.
\VS{26}After
a priest has become ceremonially clean,
they must count
off a period of seven
days for him.
\VS{27}On the day
he enters
the sanctuary,
into
the inner
court
to serve
in the sanctuary,
he must offer
his sin offering,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{28}“‘This will be
their inheritance: I am
their inheritance,
and you must give
them no
property
in Israel;
I am
their property.
\VS{29}They
may eat
the grain offering,
the sin offering,
and the guilt offering,
and every
devoted
thing in Israel
will be theirs.
\VS{30}The first
of all
the first fruits
and all
contributions
of any
kind
will be
for the priests;
you will also give
to the priest
the first portion
of your dough,
so that a blessing
may rest
on
your house.
\VS{31}The priests
will not
eat
any
bird
or animal
that has died a natural death
or was torn
to pieces by a wild animal.

\par }\Chap{45}{\PP \VerseOne{1}“‘When
you allot
the land
as an inheritance,
you will offer
an allotment
to the
{\ND{Lord}}, a holy
portion from
the land;
the length
will be eight and a quarter miles
and the width
three and one-third miles.
This
entire
area
will be holy.
\VS{2}Of this
area
a square
875
feet by 875 feet will be designated for the sanctuary,
with 87½ feet
set aside for its open space
round about.
\VS{3}From
this
measured
area you will measure
a length
of eight and a quarter miles
and a width
of three and one-third miles;
in it will be
the sanctuary,
the most
holy place.
\VS{4}It will be
a holy
portion
of the land;
it
will be for the priests,
the ministers
of the sanctuary
who approach
the

{\ND{Lord}}
to minister
to him. It will be
a place
for their houses
and a holy place
for the sanctuary.
\VS{5}An area eight and a quarter miles
in length
and three and one-third miles
in width
will be
for the Levites,
who minister
at the temple,
as the place for the cities in which they will live.
\par }{\PP \VS{6}“‘Alongside
the portion set apart
as the holy
allotment,
you will allot
for the city
an area one and two-thirds miles
wide
and eight and a quarter miles
long;
it will be
for the whole
house
of Israel.
\par }{\PP \VS{7}“‘For the prince
there will be land
on both sides
of the holy
allotment
and the allotted
city,
alongside
the holy
allotment
and the allotted
city,
on
the west
side
and on
the east
side;
it will be comparable
in length
to one
of the portions,
from
the west
border
to
the east
border
\VS{8}of the land.
This will be
his property
in Israel.
My princes
will no
longer
oppress
my people,
but the land
will be allotted
to the house
of Israel
according to their tribes.
\par }{\PP \VS{9}“‘This is what
the sovereign

{\ND{Lord}}
says: Enough,
you princes
of Israel! Put away
violence
and destruction,
and do
what is just
and right.
Put an end
to your evictions
of my people,
declares
the sovereign

{\ND{Lord}}.
\VS{10}You must use just
balances,
a just
dry measure (an ephah), and a just liquid measure (a bath).
\VS{11}The dry
and liquid
measures
will be
the same,
the bath
will contain
a tenth
of a homer,
and the ephah
a tenth
of a homer;
the homer
will be
the standard measure.
\VS{12}The shekel
will be twenty
gerahs.
Sixty shekels
will be
a mina for you.
\par }{\PP \VS{13}“‘This
is the offering
you must offer: a sixth
of an ephah
from a homer
of wheat;
a sixth
of an ephah
from a homer
of barley,
\VS{14}and as the prescribed portion
of olive oil,
one tenth
of a bath
from
each kor
(which is ten
baths
or a homer,
for
ten
baths
make a homer);
\VS{15}and one
sheep
from
each flock
of two hundred,
from
the watered places
of Israel,
for a grain offering,
burnt offering,
and peace offering,
to make atonement
for
them, declares
the sovereign

{\ND{Lord}}.
\VS{16}All
the people
of the land
will
contribute to this
offering
for the prince
of Israel.
\VS{17}It
will be
the duty of the prince
to provide the burnt offerings,
the grain offering,
and the drink offering
at festivals,
on the new moons
and Sabbaths,
at all
the appointed feasts
of the house
of Israel;
he
will provide
the
sin offering,
the
grain offering,
the
burnt offering,
and the
peace offerings
to make atonement
for the house
of Israel.
\par }{\PP \VS{18}“‘This is what
the sovereign

{\ND{Lord}}
says: In the first
month, on the first
day of the month,
you must take
an unblemished
young
bull
and purify
the sanctuary.
\VS{19}The priest
will take
some of the blood
of the sin offering
and place
it on the doorpost
of the temple,
on the four
corners
of the ledge
of the altar,
and on
the doorpost
of the gate
of the inner
court.
\VS{20}This
is what you must do
on the seventh
day of the month
for anyone
who sins inadvertently
or through ignorance;
so you will make atonement
for the temple.
\par }{\PP \VS{21}“‘In the first
month, on the fourteenth
day
of the month,
you will
celebrate
the Passover,
and for seven
days
bread made without yeast
will be eaten.
\VS{22}On that day
the prince
will provide
for himself
and for all
the people
of the land
a bull
for a sin offering.
\VS{23}And during
the seven
days
of the feast
he
will provide
as a burnt offering
to the
{\ND{Lord}}
seven
bulls
and seven
rams,
all without blemish,
on each
of the seven
days,
and a male
goat
daily
for a sin offering.
\VS{24}He will provide as a grain offering
an ephah
for each bull,
an ephah
for each ram,
and a gallon
of olive oil
for each ephah of grain.
\VS{25}In the seventh
month, on the fifteenth
day
of the month,
at the feast,
he will make
the same provisions
for the sin offering,
burnt offering,
and grain offering,
and for the olive oil,
for the seven
days.

\par }\Chap{46}{\PP \VerseOne{1}“‘This is what
the sovereign

{\ND{Lord}}
says: The gate
of the inner
court
that faces
east
will be
closed
six
working
days,
but on the Sabbath
day
it will be opened
and on the day
of the new moon
it will be opened.
\VS{2}The prince
will enter
by way
of the porch
of the gate
from the outside,
and will stand
by the doorpost
of the gate.
The priests
will provide
his burnt offering
and his peace offerings,
and he will bow
down at the threshold
of the gate
and then go out.
But the gate
will not
be closed
until
evening.
\VS{3}The people
of the land
will bow
down at the entrance
of that gate
before
the {\ND{Lord}}
on the Sabbaths
and on the new moons.
\VS{4}The burnt offering
which
the prince
will offer
to the
{\ND{Lord}}
on the Sabbath
day
will be six
unblemished
lambs
and one unblemished
ram.
\VS{5}The grain offering
will be an ephah
with the ram,
and the grain offering
with the lambs
will be as much as he is able to give,
and a gallon
of olive oil
with an ephah.
\VS{6}On the day
of the new moon
he will offer an unblemished
young
bull,
and six
lambs
and a ram,
all without blemish.
\VS{7}He will provide
a grain offering: an ephah
with the bull
and an ephah
with the ram,
and with the lambs
as much as he wishes, and a gallon
of olive oil
with each ephah of grain.
\VS{8}When the prince
enters,
he will come
by way
of the porch
of the gate
and will go out
the same way.
\par }{\PP \VS{9}“‘When the people
of the land
come
before
the {\ND{Lord}}
at the appointed feasts,
whoever enters
by way
of the north
gate
to worship
will go out
by way
of the south
gate;
whoever enters
by way
of the south
gate
will go out
by way
of the north
gate.
No
one will return
by way
of the gate
they entered
but will go out
straight ahead.
\VS{10}When they come
in, the prince
will come
in with them, and when they go out,
he will go out.
\par }{\PP \VS{11}“‘At
the festivals
and at the appointed feasts
the grain offering
will be
an ephah
with the bull
and an ephah
with the ram,
and with the lambs
as much as one is able, and a gallon
of olive oil
with each ephah of grain.
\VS{12}When
the prince
provides a freewill
offering, a burnt offering,
or
peace offerings
as a voluntary
offering to the
{\ND{Lord}}, the
gate
facing
east
will be opened
for him, and he will provide
his burnt offering
and his peace offerings
just
as he did
on
the Sabbath.
Then he will go out,
and the
gate
will be closed
after
he goes out.
\par }{\PP \VS{13}“‘You will provide a lamb
a year
old without blemish
for a burnt offering
daily
to the
{\ND{Lord}}; morning
by morning
he will provide it.
\VS{14}And you will provide
a grain offering
with it morning
by morning,
a sixth
of an ephah,
and a third
of a gallon
of olive oil
to moisten
the choice flour,
as a grain offering
to the
{\ND{Lord}}; this is a perpetual
statute.
\VS{15}Thus they will provide
the lamb,
the grain offering,
and the olive oil
morning
by morning,
as a perpetual
burnt offering.
\par }{\PP \VS{16}“‘This is what
the sovereign

{\ND{Lord}}
says: If
the prince
should give
a gift
to one
of his sons
as his inheritance,
it
will belong to his sons,
it is
their property
by inheritance.
\VS{17}But if
he gives
a gift
from his inheritance
to one
of his servants,
it will be
his until
the year
of liberty;
then it will revert
to the prince.
His inheritance
will only
remain with his sons.
\VS{18}The prince
will not
take
away any of the people’s
inheritance
by oppressively
removing them from their property.
He will give his sons
an inheritance
from his own possessions
so that
my people
will not
be scattered,
each
from his own property.’ ”
\par }{\PP \VS{19}Then he brought
me through the entrance,
which
was at the side
of the gate,
into
the holy
chambers
for the priests
which faced
north.
There
I saw
a place
at the extreme
western end.
\VS{20}He said
to
me, “This
is the place
where
the priests
will boil
the guilt offering
and the sin offering,
and where
they will bake
the
grain offering,
so that they do not
bring
them out
to
the outer
court
to transmit holiness
to the
people.”
\par }{\PP \VS{21}Then he brought
me out
to
the outer
court
and led
me past
the four
corners
of the court,
and I noticed
that in every corner
of the court
there was a court.
\VS{22}In the four
corners
of the court
were small courts,
70 feet
in length
and 52½ feet
in width;
the four
were all the same
size.
\VS{23}There was a row
of masonry around
each
of the four
courts, and places for boiling
offerings were made
under
the rows
all around.
\VS{24}Then he said
to me,
“These
are the houses
for boiling,
where
the ministers
of the temple
boil
the sacrifices
of the people.”

\par }\Chap{47}{\PP \VerseOne{1}Then he brought me back
to
the entrance
of the temple.
I noticed
that water
was flowing from under
the threshold
of the temple
toward the east
(for
the temple
faced east). The water
was flowing down
from under
the right
side
of the temple,
from south
of the altar.
\VS{2}He led me out
by way
of the north
gate
and brought me around
the outside
of the outer
gate
that faces
toward
the east;
I noticed
that the water
was trickling
out from
the south
side.
\par }{\PP \VS{3}When the man
went out
toward the east
with a measuring line
in his hand,
he measured
1,750 feet,
and then he led
me through
water,
which was ankle deep.
\VS{4}Again he measured
1,750
feet and led me through
the water,
which was now knee
deep. Once more he measured
1,750
feet and led me through
the water,
which was waist deep.
\VS{5}Again he measured
1,750
feet and it was a river
I could
not
cross,
for
the water
had risen;
it was deep
enough to swim
in, a river
that
could not
be crossed.
\VS{6}He said
to
me, “Son
of man,
have you seen
this?”
\par }{\PP Then he led
me back
to the bank
of the river.
\VS{7}When I had returned,
I noticed
a vast
number of trees
on the banks
of the river,
on both sides.
\VS{8}He said
to
me, “These
waters
go out
toward
the eastern
region
and flow down
into
the Arabah;
when they enter
the Dead Sea,
where the sea
is stagnant,
the waters
become fresh.
\VS{9}Every
living
creature
which
swarms
where the river
flows
will live;
there will be
many
fish,
for
these
waters
flow
there.
It will become fresh
and everything
will live
where
the river
flows.
\VS{10}Fishermen
will stand
beside
it; from Engedi
to
En-eglaim
they will spread
nets.
They will catch many kinds
of fish,
like the fish
of the Great
Sea.
\VS{11}But its swamps
and its marshes
will not
become fresh; they will remain salty.
\VS{12}On
both sides of the river’s
banks,
every
kind of tree
will grow for food.
Their leaves
will not
wither
nor
will their fruit
fail,
but they will bear fruit
every month,
because
their water
source flows
from
the sanctuary.
Their fruit
will be
for food
and their leaves
for healing.”
\par }{\SH Boundaries for the Land
\par }{\PP \VS{13}This is what
the sovereign

{\ND{Lord}}
says: “Here are the borders
you will observe as you allot
the land
to the twelve
tribes
of Israel.
(Joseph
will have two
portions.)
\VS{14}You must divide
it equally just as I vowed
to give
it to your forefathers;
this
land
will be assigned
as your inheritance.
\par }{\PP \VS{15}“This
will be the border
of the land: On the north
side,
from
the Great
Sea
by way
of Hethlon
to the entrance
of Zedad;
\VS{16}Hamath,
Berothah,
Sibraim,
which
is between
the border
of Damascus
and the border
of Hamath,
as far as Hazer-hattikon,
which
is on the border
of Hauran.
\VS{17}The border
will
run from
the sea
to Hazar-enan,
at the border
of Damascus,
and on the north
is the border
of Hamath.
This is the north
side.
\VS{18}On
the east
side,
between
Hauran
and Damascus,
and between
Gilead
and the land
of Israel,
will be the Jordan.
You will measure
from
the border
to the eastern
sea.
This is the east
side.
\VS{19}On the south
side it will run from Tamar
to
the waters
of Meribath
Kadesh,
the river,
to
the Great
Sea.
This is the south
side.
\VS{20}On the west
side
the Great
Sea
will be the boundary
to
a point opposite
Lebo-hamath.
This
is the west
side.
\par }{\PP \VS{21}“This is how you will divide
this
land
for yourselves among the tribes
of Israel.
\VS{22}You must allot
it as an inheritance
among yourselves and for the foreigners
who reside
among
you, who have
fathered
sons
among
you. You must treat them as native-born
among the people of Israel;
they will be allotted
an inheritance
with you
among
the tribes
of Israel.
\VS{23}In whatever
tribe
the foreigner
resides,
there
you will give
him his inheritance,”
declares
the sovereign

{\ND{Lord}}.

\par }\Chap{48}{\PP \VerseOne{1}“These
are the names
of the tribes: From the northern
end
beside
the road
of Hethlon
to Lebo-hamath,
as far as Hazar-enan
(which is on the border
of Damascus,
toward the north
beside
Hamath), extending
from the east
side
to the west,
Dan
will have one portion.
\VS{2}Next to the border
of Dan,
from
the east
side
to the west,
Asher
will have one portion.
\VS{3}Next to the border
of Asher
from
the east
side
to the west,
Naphtali
will have one portion.
\VS{4}Next to the border
of Naphtali
from
the east
side
to the west,
Manasseh
will have one portion.
\VS{5}Next to the border
of Manasseh
from
the east
side
to the west,
Ephraim
will have one portion.
\VS{6}Next to the border
of Ephraim
from
the east
side
to the west,
Reuben
will have one portion.
\VS{7}Next to the border
of Reuben
from
the east
side
to the west,
Judah
will have one portion.
\par }{\PP \VS{8}“Next to the border
of Judah
from
the east
side
to the west
will be
the allotment
you must set apart. It
is to be
eight and a quarter miles
wide,
and the same length
as one
of the tribal
portions,
from
the east
side
to the west;
the sanctuary
will be in the middle of it.
\VS{9}The allotment
you set apart
to the
{\ND{Lord}}
will be eight and a quarter miles
in length
and three and one-third miles
in width.
\VS{10}These
will be
the allotments
for the holy
portion: for the priests,
toward the north
eight and a quarter miles
in length, toward the west
three and one-third miles
in width,
toward the east
three and one-third miles
in width,
and toward the south
eight and a quarter miles
in length;
the sanctuary
of the
{\ND{Lord}}
will be
in the middle.
\VS{11}This will be for the priests
who are set apart
from the descendants
of Zadok
who
kept
my charge
and did not
go
astray
when the people of Israel
strayed
off, like the Levites did.
\VS{12}It will be
their portion
from the allotment
of the land,
a most
holy
place, next to the border
of the Levites.
\par }{\PP \VS{13}“Alongside
the border
of the priests,
the Levites
will have an allotment eight and a quarter miles
in length
and three and one-third miles
in width.
The whole
length
will be eight and a quarter miles
and the width
three and one-third miles.
\VS{14}They must not
sell
or
exchange
any of it; they must not
transfer
this choice portion
of land,
for
it is set apart
to the
{\ND{Lord}}.
\par }{\PP \VS{15}“The remainder,
one and two-thirds
miles in width
and eight and a quarter miles
in
length, will be for common
use
by the city,
for houses and for open space.
The city
will be in the middle of it;
\VS{16}these
will be its measurements: The north
side
will be one and one-half miles,
the south
side
one and one-half miles,
the east
side
one and one-half miles,
and the west
side
one and one-half miles.
\VS{17}The city
will have open spaces: On the north
there will be
437½ feet,
on the south
437½ feet,
on the east
437½ feet,
and on the west
437½ feet.
\VS{18}The remainder
of the length
alongside
the holy
allotment
will be three and one-third miles
to the east
and three and one-third miles
toward the west,
and it will be
beside
the holy
allotment.
Its produce
will be
for food
for the workers
of the city.
\VS{19}The workers
of the city
from all
the tribes
of Israel
will cultivate it.
\VS{20}The whole
allotment
will be eight and a quarter miles
square,
you must set apart
the holy
allotment
with
the possession
of the city.
\par }{\PP \VS{21}“The rest,
on both sides
of the holy
allotment
and the property
of the city,
will belong to
the prince.
Extending from
the eight and a quarter miles
of the holy allotment
to
the east
border,
and westward
from
the eight and a quarter miles
to
the west
border,
alongside
the portions,
it will belong
to the prince.
The holy
allotment
and the sanctuary
of the temple
will be in the middle of it.
\VS{22}The property
of the Levites
and of the city
will be
in the middle
of that which
belongs to the prince.
The portion between
the border
of Judah
and the border
of Benjamin
will be
for the prince.
\par }{\PP \VS{23}“As for the rest
of the tribes: From
the east
side
to the west
side, Benjamin
will have one portion.
\VS{24}Next to the border
of Benjamin,
from
the east
side
to the west
side, Simeon
will have one portion.
\VS{25}Next to the border
of Simeon,
from
the east
side
to the west
side, Issachar
will have one portion.
\VS{26}Next to the border
of Issachar,
from
the east
side
to the west
side, Zebulun
will have one portion.
\VS{27}Next to the border
of Zebulun,
from
the east
side
to the west
side, Gad
will have one portion.
\VS{28}Next to
the border
of Gad,
at the south
side,
the border
will run
from Tamar
to the waters
of Meribath
Kadesh,
to the Stream
of Egypt and on
to the Great
Sea.
\VS{29}This
is the land
which
you will
allot
to the tribes
of Israel,
and these
are their portions,
declares
the sovereign

{\ND{Lord}}.
\par }{\PP \VS{30}“These
are the exits
of the city: On the north
side,
one and one-half miles
by measure,
\VS{31}the gates
of the city
will be named
for the tribes
of Israel;
there will be three
gates
to the north: one
gate
for Reuben,
one
gate
for Judah,
and one
gate
for Levi.
\VS{32}On the east
side,
one and one-half miles
in length, there will be three
gates: one
gate
for Joseph,
one
gate
for Benjamin,
and one
gate
for Dan.
\VS{33}On the south
side,
one and one-half miles
by measure,
there will be three
gates: one
gate
for Simeon,
one
gate
for Issachar,
and one
gate
for Zebulun.
\VS{34}On the west
side,
one and one-half miles
in length, there will be three
gates: one
gate
for Gad,
one
gate
for Asher,
and one
gate
for Naphtali.
\VS{35}The circumference
of the city will be six miles.
The name
of the city
from that day
forward will be: ‘The
{\ND{Lord}}
Is There.’ ”
\par }