\NormalFont\ShortTitle{Habakkuk}
{\MT Habakkuk

\par }\ChapOne{1}{\SH Habakkuk Complains to the Lord
\par }{\PP \VerseOne{1}The following is the message
which
God revealed
to Habakkuk
the prophet:
\par }{\Q \VS{2}How
long,

{\ND{Lord}}, must I cry for help?
\par }{\Q But you do not
listen!
\par }{\Q I call out
to
you, “Violence!”
\par }{\Q But you do not
intervene!
\par }{\Q \VS{3}Why
do you force me to witness
injustice?

\par }{\Q Why do you put up
with wrongdoing?

\par }{\Q Destruction
and violence
confront
me;
\par }{\Q conflict
is present
and one must endure
strife.
\par }{\Q \VS{4}For
this reason
the law
lacks power,
\par }{\Q and justice
is never
carried out.
\par }{\Q Indeed,
the wicked
intimidate
the innocent.
\par }{\Q For
this reason
justice
is perverted.
\par }{\SH The Lord Reveals Some Startling News
\par }{\Q \VS{5}“Look
at the nations
and pay attention!

\par }{\Q You will be shocked
and amazed!

\par }{\Q For
I will do
something
in your lifetime
\par }{\Q that you will not
believe
even though
you are forewarned.
\par }{\Q \VS{6}Look,
I am about to empower
the Babylonians,
\par }{\Q that ruthless
and greedy
nation.
\par }{\Q They sweep across
the surface
of the earth,
\par }{\Q seizing
dwelling places
that do not belong to them.
\par }{\Q \VS{7}They are frightening
and terrifying;
\par }{\Q they
decide
for themselves what is right.
\par }{\Q \VS{8}Their horses
are faster
than leopards
\par }{\Q and more alert
than wolves
in the desert.
\par }{\Q Their horses
gallop,
\par }{\Q their horses
come
a great distance;
\par }{\Q like a vulture
they swoop down
quickly
to devour their prey.
\par }{\Q \VS{9}All
of them intend to do
violence;
\par }{\Q every face
is determined.
\par }{\Q They take
prisoners
as easily as one scoops up
sand.
\par }{\Q \VS{10}They mock
kings
\par }{\Q and laugh
at rulers.
\par }{\Q They laugh at every
fortified
city;
\par }{\Q they build siege ramps
and capture them.
\par }{\Q \VS{11}They sweep
by like the wind
and pass on.
\par }{\Q But the one who
considers himself a god
will be held guilty.”
\par }{\SH Habakkuk Voices Some Concerns
\par }{\Q \VS{12}{\ND{Lord}}, you
have been active from ancient times;
\par }{\Q my sovereign
God,
you are immortal.
\par }{\Q {\ND{Lord}}, you have made them your instrument of judgment.
\par }{\Q Protector,
you have appointed
them as your instrument
of punishment.
\par }{\Q \VS{13}You are too just
to tolerate
evil;
\par }{\Q you are unable
to
condone
wrongdoing.
\par }{\Q So why
do you put up
with such treacherous people?

\par }{\Q Why do you say nothing
when the wicked
devour those more righteous
than they are?
\par }{\Q \VS{14}You made
people
like fish
in the sea,
\par }{\Q like animals in the sea
that have no
ruler.
\par }{\Q \VS{15}The Babylonian tyrant pulls
them all
up
with a fishhook;
\par }{\Q he hauls
them in
with his throw net.
\par }{\Q When he catches
them in
his dragnet,
\par }{\Q he is very happy.
\par }{\Q \VS{16}Because of
his success he offers sacrifices
to his throw net
\par }{\Q and burns incense
to his dragnet;
\par }{\Q for because
of them
he has plenty
of food,
\par }{\Q and more than enough
to eat.
\par }{\Q \VS{17}Will he then continue
to fill and empty
his throw net?

\par }{\Q Will he always
destroy
nations
and spare
none?

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}I will stand
at
my watch post;
\par }{\Q I will remain stationed
on
the city wall.
\par }{\Q I will keep watching,
so I can see
what
he says
to me
\par }{\Q and can know how
I should answer
\par }{\Q when he counters
my argument.
\par }{\SH The Lord Assures Habakkuk
\par }{\Q \VS{2}The
{\ND{Lord}}
responded:


\par }{\Q “Write down
this message! Record it legibly
on
tablets,
\par }{\Q so
the one who announces
it may read it easily.
\par }{\Q \VS{3}For
the message
is a witness
to what is decreed;
\par }{\Q it gives reliable testimony
about how matters will turn out.
\par }{\Q Even if
the message is not fulfilled right away,
wait patiently;
\par }{\Q for
it will certainly come
to pass
– it will not
arrive late.
\par }{\Q \VS{4}Look,
the one whose desires
are not
upright will faint from exhaustion,
\par }{\Q but the person
of integrity
will live because of his faithfulness.
\par }{\Q \VS{5}Indeed,
wine
will betray
the proud,
restless
man!

\par }{\Q His appetite
is as big
as Sheol’s;
\par }{\Q like death,
he is never
satisfied.
\par }{\Q He gathers
all
the nations;
\par }{\Q he seizes
all
peoples.
\par }{\SH The Proud Babylonians are as Good as Dead
\par }{\Q \VS{6}“But
all
these
nations will someday taunt
him

\par }{\Q and ridicule
him with proverbial sayings:

\par }{\Q ‘The one who accumulates
what does not
belong to him is as good as
dead

\par }{\Q (How long
will this go on?) –
\par }{\Q he who gets rich by extortion!’
\par }{\Q \VS{7}Your creditors
will suddenly
attack;

\par }{\Q those who terrify
you will spring into action,
\par }{\Q and they will rob you.
\par }{\Q \VS{8}Because
you
robbed
many
countries,
\par }{\Q all
who are left among the nations
will rob
you.
\par }{\Q You have shed human
blood
\par }{\Q and committed violent
acts against lands,
cities,
and those who live in them.
\par }{\Q \VS{9}The one who builds
his house
by
unjust gain
is as good as dead.

\par }{\Q He does
this so he can
build his nest
way up high
\par }{\Q and escape the clutches
of disaster.
\par }{\Q \VS{10}Your schemes
will bring shame
to your house.
\par }{\Q Because you destroyed many
nations,
you will self-destruct.
\par }{\Q \VS{11}For
the stones
in the walls
will cry
out,
\par }{\Q and the wooden
rafters
will answer back.
\par }{\Q \VS{12}The one who builds
a city
by bloodshed is as good as dead –
\par }{\Q he who starts a town by unjust deeds.
\par }{\Q \VS{13}Be sure
of this! The
{\ND{Lord}}
who commands armies
has decreed:
\par }{\Q The nations’
efforts
will go up in smoke;
\par }{\Q their
exhausting
work will be for nothing.
\par }{\Q \VS{14}For
recognition
of the
{\ND{Lord}}’s
sovereign majesty
will fill
the earth
\par }{\Q just as the waters
fill up
the sea.
\par }{\Q \VS{15}“You who force your neighbor
to drink wine are as good as dead –
\par }{\Q you who make others intoxicated by forcing them to drink from the bowl of your furious anger,
\par }{\Q so you can look at their genitals.
\par }{\Q \VS{16}But you will become drunk
with shame,
not majesty.
\par }{\Q Now
it is your
turn
to drink
and expose
your uncircumcised foreskin!

\par }{\Q The cup
of wine in the
{\ND{Lord}}’s
right hand
is coming to you,
\par }{\Q and disgrace
will replace your majestic glory!
\par }{\Q \VS{17}For
you will pay in full for your violent
acts against Lebanon;
\par }{\Q terrifying judgment will come upon
you because of the way you destroyed
the wild animals
living there.

\par }{\Q You have shed
human
blood
\par }{\Q and committed violent
acts against lands,
cities,
and those who live in them.
\par }{\Q \VS{18}What
good
is an idol? Why would
a craftsman
make
it?

\par }{\Q What good is a metal image
that
gives misleading
oracles?

\par }{\Q Why would its creator
place his trust
in
it
\par }{\Q and make
such mute,
worthless things?
\par }{\Q \VS{19}The one
who says
to wood,
‘Wake
up!’ is as good as dead –
\par }{\Q he who
says to speechless
stone,
‘Awake!’
\par }{\Q Can it give reliable guidance?

\par }{\Q It is
overlaid
with gold
and silver;
\par }{\Q it has no
life’s
breath
inside it.
\par }{\Q \VS{20}But the
{\ND{Lord}}
is in his majestic
palace.
\par }{\Q The whole
earth
is speechless
in his presence!”

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}This is a prayer
of Habakkuk
the prophet:
\par }{\Q \VS{2}{\ND{Lord}}, I have heard
the report of what you did;
\par }{\Q I am awed,

{\ND{Lord}}, by what you accomplished.
\par }{\Q In
our time
repeat those deeds;
\par }{\Q in
our time
reveal
them again.

\par }{\Q But when you cause turmoil,
remember
to show
us mercy!
\par }{\Q \VS{3}God
comes
from Teman,
\par }{\Q the sovereign one
from Mount
Paran.
{\IT{
Selah}}.

\par }{\Q His splendor covers
the skies,
\par }{\Q his glory
fills
the earth.
\par }{\Q \VS{4}He is as bright
as lightning;
\par }{\Q a two-pronged lightning bolt
flashes from his hand.
\par }{\Q This is the outward
display of his power.
\par }{\Q \VS{5}Plague
goes
before
him;
\par }{\Q pestilence
marches right behind him.
\par }{\Q \VS{6}He
takes
his battle position and shakes
the earth;
\par }{\Q with a mere look
he frightens
the nations.
\par }{\Q The ancient
mountains
disintegrate;
\par }{\Q the primeval hills
are
flattened.
\par }{\Q He travels on the ancient roads.
\par }{\Q \VS{7}I see
the tents
of Cushan
overwhelmed by trouble;
\par }{\Q the tent curtains
of the land
of Midian are shaking.
\par }{\Q \VS{8}Is the
{\ND{Lord}}
mad
at the rivers?
\par }{\Q Are you
angry
with the rivers?
\par }{\Q Are you
enraged
at the sea?

\par }{\Q Is this why
you climb
into
your horse-drawn
chariots,

\par }{\Q your victorious
chariots?
\par }{\Q \VS{9}Your bow
is ready
for action;
\par }{\Q you commission
your arrows.
{\IT{
Selah}}.
\par }{\Q You cause flash floods
on the earth’s surface.
\par }{\Q \VS{10}When the mountains
see
you, they shake.
\par }{\Q The torrential downpour
sweeps through.
\par }{\Q The great deep
shouts
out;
\par }{\Q it lifts
its hands
high.
\par }{\Q \VS{11}The sun
and moon
stand
still in their courses;
\par }{\Q the flash
of your arrows
drives
them away,
\par }{\Q the bright light
of your lightning-quick
spear.
\par }{\Q \VS{12}You furiously
stomp
on the earth,
\par }{\Q you angrily
trample
down the nations.
\par }{\Q \VS{13}You march out
to deliver your
people,
\par }{\Q to deliver your
special servant.
\par }{\Q You strike
the leader
of the wicked
nation,
\par }{\Q laying
him open from the lower
body to
the neck.
{\IT{
Selah}}.
\par }{\Q \VS{14}You pierce
the heads
of his warriors with a spear.
\par }{\Q They storm
forward to scatter
us;

\par }{\Q they shout with joy
as
if they were plundering
the poor
with no opposition.
\par }{\Q \VS{15}But you trample
on the sea
with your horses,
\par }{\Q on the surging,
raging
waters.
\par }{\SH Habakkuk Declares His Confidence
\par }{\Q \VS{16}I listened
and my stomach
churned;
\par }{\Q the sound
made my lips
quiver.
\par }{\Q My frame went
limp, as if my bones
were decaying,
\par }{\Q and I shook
as
I tried to walk.

\par }{\Q I long
for the day
of distress
\par }{\Q to come upon
the people
who attack us.
\par }{\Q \VS{17}When
the fig tree
does not
bud,
\par }{\Q and there are no
grapes
on the vines;
\par }{\Q when the olive trees
do not produce,
\par }{\Q and the fields
yield
no
crops;
\par }{\Q when the sheep
disappear
from the pen,
\par }{\Q and there are no
cattle
in the stalls,
\par }{\Q \VS{18}I
will rejoice
because of the
{\ND{Lord}};
\par }{\Q I will be happy
because of the God
who delivers me!
\par }{\Q \VS{19}The sovereign

{\ND{Lord}}
is my source of strength.
\par }{\Q He gives
me the agility
of a deer;
\par }{\Q he enables me to negotiate
the rugged terrain.
\par }{\PP (This prayer is for the song leader.
It is to be accompanied by stringed instruments.)
\par }