\NormalFont\ShortTitle{Haggai}
{\MT Haggai

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \VerseOne{1}On the first
day
of the sixth
month
of King
Darius’
second
year,
the {\ND{Lord}}
spoke this message
through
the prophet
Haggai
to
Zerubbabel
son
of Shealtiel,
governor
of Judah,
and to
the high priest
Joshua
son
of Jehozadak:
\par }{\SH The Indifference of the People
\par }{\PP \VS{2}The

{\ND{Lord}}
who rules over all
says
this: “These people
have said,
‘The time
for rebuilding
the
{\ND{Lord}}’s
temple
has not
yet
come.’ ”
\VS{3}So the

{\ND{Lord}}
spoke
through
the prophet
Haggai as follows:
\VS{4}“Is
it right for you
to live
in richly
paneled
houses
while my temple
is in ruins?
\VS{5}Here
then
is what
the {\ND{Lord}}
who rules over all
says: ‘Think
carefully
about what you are doing.
\VS{6}You have planted
much,
but have harvested
little.
You eat,
but are never
filled.
You drink,
but are still
thirsty.
You put on clothes,
but are not
warm.
Those who earn
wages
end up with
holes
in their money bags.’ ”
\par }{\SH The Instruction of the People
\par }{\PP \VS{7}“Moreover,
the {\ND{Lord}}
who rules over all
says: ‘Pay
close
attention to these things also.
\VS{8}Go up
to the hill country
and bring
back timber
to build
the temple.
Then I will be pleased
and honored,’
says
the {\ND{Lord}}.
\VS{9}‘You expected
a large harvest, but
instead there
was little,
and when you
brought
it
home
it
disappeared right away.
Why?’ asks
the
{\ND{Lord}}
who rules over all.
‘Because
my temple
remains in ruins,
thanks to each
of you
favoring
his own
house!
\VS{10}This is why
the sky
has held back
its dew
and the earth
its produce.
\VS{11}Moreover, I have called
for a drought
that will affect the fields,
the hill country,
the grain,
new wine,
fresh olive oil,
and everything
that grows
from the ground;
it also will harm people,
animals,
and everything
they produce.’ ”
\par }{\SH The Response of the People
\par }{\PP \VS{12}Then
Zerubbabel
son
of Shealtiel
and the high priest
Joshua
son
of Jehozadak,
along with the whole
remnant
of the people,
obeyed
the {\ND{Lord}}
their God.
They responded favorably to the message
of the prophet
Haggai,
who
spoke just as
the {\ND{Lord}}
their God
had instructed him,
and the people
began to respect
the {\ND{Lord}}.
\VS{13}Then Haggai,
the
{\ND{Lord}}’s
messenger,
spoke
the
{\ND{Lord}}’s
word
to the people: “I am
with
you!” says
the {\ND{Lord}}.
\VS{14}So
the {\ND{Lord}}
energized
and encouraged Zerubbabel
son
of Shealtiel,
governor
of Judah,
the
high priest
Joshua
son
of Jehozadak,
and the
whole
remnant
of the people.
They came
and worked
on the temple
of their God,
the {\ND{Lord}}
who rules over all.
\VS{15}This took place on the twenty-fourth
day
of the sixth
month
of King
Darius’
second
year.

\par }\Chap{2}{\PP \VerseOne{1}On the twenty-first
day of the seventh
month,
the {\ND{Lord}}
spoke
again through
the prophet
Haggai:
\VS{2}“Ask
the following questions
to
Zerubbabel
son
of Shealtiel,
governor
of Judah,
the high priest
Joshua
son
of Jehozadak,
and the remnant
of the people:
\VS{3}‘Who
among you survivors
saw
the former
splendor
of this
temple? How
does it look
to you
now? Isn’t
it nothing
by comparison?
\VS{4}Even so,
take heart,
Zerubbabel,’
says
the {\ND{Lord}}. ‘Take heart,
Joshua
son
of Jehozadak,
the high priest,
and all
you citizens
of the land,’
says
the {\ND{Lord}}, ‘and begin to work.
For
I am
with
you,’ says
the {\ND{Lord}}
who rules over all.
\VS{5}‘Do not
fear,
because I made
a promise
to your ancestors when they left
Egypt,
and my spirit
even now testifies to you.’
\VS{6}Moreover,
the {\ND{Lord}}
who rules over all
says: ‘In just
a little
while
I
will once
again
shake
the sky
and the earth,
the sea
and the dry ground.
\VS{7}I will also shake up
all
the nations,
and they
will offer
their treasures;
then I will fill
this
temple
with glory,’
says
the {\ND{Lord}}
who rules over
all.
\VS{8}‘The silver
and gold
will be mine,’ says
the {\ND{Lord}}
who rules over all.
\VS{9}‘The future
splendor
of this
temple
will be
greater
than
that of former
times,’ the
{\ND{Lord}}
who rules over all
declares,
‘and in this
place
I will give
peace.’ ”
\par }{\SH The Promised Blessing
\par }{\PP \VS{10}On the twenty-fourth
day of the ninth
month of Darius’
second
year,
the
{\ND{Lord}}
spoke
again to the prophet
Haggai:
\VS{11}“The

{\ND{Lord}}
who rules over all
says, ‘Ask
the priests
about the law.
\VS{12}If
someone
carries
holy
meat
in a fold
of his garment
and that fold
touches
bread,
a boiled dish,
wine,
olive oil,
or any
other food,
will that item become holy?’ ” The priests
answered,
“It will not.”
\VS{13}Then Haggai
asked, “If
a person who
is ritually unclean
because of touching
a dead body comes in contact
with one
of these
items, will it become unclean?” The priests
answered,
“It will be unclean.”
\par }{\PP \VS{14}Then
Haggai
responded,
“ ‘The people
of this
nation
are unclean in my sight,’
says
the {\ND{Lord}}. ‘And so
is all
their effort;
everything
they offer
is also
unclean.
\VS{15}Now
therefore reflect carefully
on
the recent
past,
before
one stone
was laid
on
another
in the
{\ND{Lord}}’s
temple.
\VS{16}From that time
when one came
expecting a heap
of twenty
measures, there were
only ten;
when one came
to
the wine vat
to draw out
fifty
measures from it,
there were
only twenty.
\VS{17}I struck
all
the products of your labor
with blight,
disease,
and hail,
and yet you brought
nothing
to
me,’ says
the {\ND{Lord}}.
\VS{18}‘Think
carefully
about the past: from
today,
the twenty-fourth
day
of the ninth
month, to the day
work
on
the temple
of the {\ND{Lord}}
was resumed,
think
about it.
\VS{19}The seed
is still
in the storehouse,
isn’t
it? And the vine,
fig tree,
pomegranate,
and olive
tree
have not
produced. Nevertheless,
from
today
on I will bless you.’ ”
\par }{\SH Zerubbabel the Chosen One
\par }{\PP \VS{20}Then
the {\ND{Lord}}
spoke
again
to
Haggai
on the twenty-fourth
day of the month:
\VS{21}Tell
Zerubbabel
governor
of Judah: ‘I am
ready to shake
the sky
and the earth.
\VS{22}I will overthrow
royal
thrones
and shatter
the might of earthly
kingdoms.
I will overthrow
chariots
and those
who ride
them, and horses
and their riders
will
fall as people
kill
one
another.
\VS{23}On that day,’
says
the {\ND{Lord}}
who rules over all,
‘I will take
you, Zerubbabel
son
of Shealtiel,
my servant,’
says
the {\ND{Lord}}, ‘and I will make
you like a signet
ring, for
I have chosen
you,’ says
the {\ND{Lord}}
who rules over all.”
\par }