\NormalFont\ShortTitle{Hebrews}
{\MT Hebrews

\par }\ChapOne{1}{\SH Introduction: God Has Spoken Fully and Finally in His Son
\par }{\PP \VerseOne{1}After God
spoke
long ago
in various portions
and
in various ways
to our ancestors
through
the prophets,
\VS{2}in these last days he has spoken to us in a son, whom
he appointed
heir
of all things,
and through
whom
he created
the world.
\VS{3}The Son
is
the radiance
of his glory
and
the representation
of his
essence,
and
he sustains
all things
by his
powerful
word,
and so when
he had accomplished
cleansing
for sins,
{\IT{
he sat down
at
the right hand
of the Majesty
on
high}}.
\VS{4}Thus he became
so far
better than
the angels
as
he has inherited
a name
superior
to
theirs.
\par }{\SH The Son Is Superior to Angels
\par }{\PP \VS{5}For
to which of the angels
did God ever
say, “{\QT{
You
are
my
son! Today
I
have fathered
you}}”? And
in another
place he says, “{\QT{
I
will be
his
father
and
he
will be
my
son}}.”
\VS{6}But
when
he again
brings
his firstborn
into
the world,
he says, “{\QT{
Let
all
the angels
of God
worship
him!}}”
\VS{7}And
he says
of the angels, “{\QT{
He makes
his
angels
spirits
and
his
ministers
a flame
of fire}},”
\VS{8}but
of the Son
he says,

\par }{\Q “{\QT{
Your
throne, O God, is forever
and ever,}}
\par }{\Q {\QT{
and
a righteous
scepter
is the scepter
of your kingdom}}.
\par }{\Q \VS{9}{\QT{
You have loved
righteousness
and
hated
lawlessness.}}
\par }{\Q {\QT{
So
God, your
God, has anointed
you
over
your
companions
with the oil
of rejoicing}}.”
\par }{\PP \VS{10}And,
\par }{\Q “{\QT{
You
founded
the earth
in
the beginning, Lord}},
\par }{\Q {\QT{
and
the heavens
are
the works
of your
hands}}.
\par }{\Q \VS{11}{\QT{
They
will perish,
but
you
continue}}.
\par }{\Q {\QT{
And
they will
all
grow old
like
a garment}},
\par }{\Q \VS{12}{\QT{
and
like
a robe
you will fold
them
up}}
\par }{\Q {\QT{
and}}
like
a garment
{\QT{
they will be changed,}}
\par }{\Q {\QT{
but
you
are
the same
and
your
years
will
never
run out}}.”
\par }{\PP \VS{13}But
to
which
of the angels
has he
ever
said, “{\QT{
Sit
at
my
right hand
until
I make
your
enemies
a footstool
for your
feet}}”?
\VS{14}Are they
not
all
ministering
spirits,
sent out
to
serve
those who will
inherit
salvation?

\par }\Chap{2}{\PP \VerseOne{1}Therefore
we
must
pay
closer
attention
to what we have heard,
so that we do
not
drift away.
\VS{2}For
if
the message
spoken
through
angels
proved to be
so firm
that every
violation
or
disobedience
received
its just
penalty,
\VS{3}how
will
we
escape
if we neglect
such a great
salvation? It was first
communicated
through
the Lord
and was confirmed
to
us
by
those who heard him,
\VS{4}while
God
confirmed
their witness
with signs
and
wonders
and
various
miracles
and
gifts of the Holy
Spirit
distributed
according to
his
will.
\par }{\SH Exposition of Psalm 8: Jesus and the Destiny of Humanity
\par }{\PP \VS{5}For
he did not
put the world
to come,
about
which
we are speaking,
under the control
of angels.
\VS{6}Instead
someone
testified
somewhere:
\par }{\Q “{\QT{
What
is
man
that
you think
of him
or
the son
of man
that
you care
for him?}}
\par }{\Q \VS{7}{\QT{
You made
him
lower
than
the angels
for a little while}}.
\par }{\Q {\QT{
You crowned
him
with glory
and
honor}}.
\par }{\Q \VS{8}{\QT{
You put
all things
under
his
control}}.”
\par }{\PP For
when he
{\QT{
put
all things
under
his control}}, he left
nothing
outside of his
control.
At present
we do
not yet
see
{\QT{
all things
under
his
control}},
\VS{9}but
we see
Jesus,
who was made
{\QT{
lower
than
the angels
for a little while}}, now crowned
with glory
and
honor
because he suffered
death,
so that
by God’s
grace
he would experience
death
on behalf
of everyone.
\VS{10}For
it was fitting
for him,
for
whom
and
through
whom
all things
exist, in bringing
many
sons
to
glory,
to make
the pioneer
of their
salvation
perfect
through
sufferings.
\VS{11}For
indeed he who makes holy
and
those being made holy
all
have the same origin,
and so
he is
not
ashamed
to call
them
brothers and sisters,
\VS{12}saying, “{\QT{
I will proclaim
your
name
to my
brothers;
in
the midst
of
the assembly
I will praise
you}}*.”
\VS{13}Again
he says, “I
will be
confident
in
him,”
and
again, “{\QT{
Here
I
am, with the children
God
has given
me}}*.”
\VS{14}Therefore,
since
the children
share
in flesh
and
blood,
he
likewise
shared in
their
humanity,
so that
through
death
he could destroy
the one who holds
the power
of death
(that
is,
the devil),
\VS{15}and
set free
those
who were
held in slavery
all
their lives
by
their fear
of death.
\VS{16}For
surely
his concern
is not
for angels,
but
he is concerned
for Abraham’s
descendants.
\VS{17}Therefore
he had
to be made like
his brothers and sisters
in
every respect,
so that
he could become
a merciful
and
faithful
high priest
in things relating to
God,
to
make atonement
for the sins
of the people.
\VS{18}For
since he
himself suffered
when
he
was tempted,
he is able
to help
those who are tempted.

\par }\Chap{3}{\PP \VerseOne{1}Therefore,
holy
brothers and sisters,
partners
in a heavenly
calling,
take note
of Jesus,
the apostle
and
high priest
whom we
confess,
\VS{2}who is
faithful
to the one who appointed
him,
as
Moses
was also
in
God’s
house.
\VS{3}For
he
has come to deserve
greater
glory
than
Moses,
just as
the builder of a house
deserves
greater
honor
than
the house
itself!
\VS{4}For
every
house
is built
by
someone,
but
the builder
of all things
is God.
\VS{5}Now
Moses
was
{\QT{
faithful
in
all
God’s
house}}
as
a servant,
to
testify
to the things that would be spoken.
\VS{6}But
Christ
is faithful as
a son
over
God’s
house.
We
are
of his house,
if
in fact we hold firmly
to our confidence
and
the hope
we take pride in.
\par }{\SH Exposition of Psalm 95: Hearing God’s Word in Faith
\par }{\PP \VS{7}Therefore,
as
the Holy
Spirit
says,
\par }{\Q “{\QT{
Oh, that today
you would listen
as he
speaks!}}
\par }{\Q \VS{8}“{\QT{
Do
not
harden
your
hearts
as
in
the rebellion, in
the day
of testing
in
the wilderness}}.
\par }{\Q \VS{9}“{\QT{
There
your
fathers
tested
me and tried
me, and
they saw
my
works
for forty
years}}.
\par }{\Q \VS{10}“{\QT{
Therefore, I became provoked at
that
generation
and
said, ‘Their
hearts
are
always
wandering
and
they have
not
known
my
ways}}.’
\par }{\Q \VS{11}“{\QT{
As
I swore
in
my
anger, ‘They will
never
enter
my
rest}}!’ ”
\par }{\PP \VS{12}See
to it, brothers and sisters,
that none
of you
has
an evil,
unbelieving
heart
that forsakes
the living
God.
\VS{13}But
exhort
one another
each
day,
as long as
it is called
“Today,”
that
none
of
you
may become hardened
by sin’s
deception.
\VS{14}For
we have become
partners
with Christ,
if
in fact we hold
our initial
confidence
firm
until
the end.
\VS{15}As
it says, “{\QT{
O, that today
you would listen
as he
speaks! Do
not
harden
your
hearts
as
in
the rebellion}}.”
\VS{16}For
which ones
heard
and
rebelled? Was it not
all
who came out
of
Egypt
under
Moses’ leadership?
\VS{17}And
against whom
was God provoked
for forty
years? Was it not
those who sinned,
{\IT{
whose
dead bodies
fell
in
the wilderness}}?
\VS{18}And
to whom
did he swear
they would
never
enter
into
his
rest,
except
those who were disobedient?
\VS{19}So
we see
that
they could
not
enter
because of
unbelief.

\par }\Chap{4}{\PP \VerseOne{1}Therefore
we must be wary
that, while the promise
of entering
his
rest
remains
open, none
of
you
may seem
to have come short of it.
\VS{2}For
we had
good news proclaimed
to us just as
they
did. But
the message
they heard
did
them no
good,
since they did
not
join in with
those
who heard
it in faith.
\VS{3}For
we who have believed
enter
that rest,
as
he has said, “{\QT{
As
I swore
in
my
anger, ‘They will never enter
my
rest!}}’ ”
And yet
God’s works
were accomplished
from
the foundation
of the world.
\VS{4}For
he has spoken
somewhere
about
the seventh
day in this way: “{\QT{
And
God
rested
on
the seventh
day
from
all
his
works}},”
\VS{5}but
to repeat
the text cited earlier: “{\QT{
They will
never
enter
my
rest!}}”
\VS{6}Therefore
it remains
for some
to enter
it, yet those to whom it was
previously
proclaimed
did
not
enter
because of
disobedience.
\VS{7}So God again
ordains
a certain
day,
“Today,”
speaking
through
David
after
so long
a time,
as
in the words quoted before, “{\QT{
O, that today
you would listen
as he
speaks! Do
not
harden
your
hearts}}.”
\VS{8}For
if
Joshua
had given
them
rest,
God would
not
have spoken
afterward
about
another
day.
\VS{9}Consequently
a Sabbath rest
remains
for the people
of God.
\VS{10}For
the one who enters
God’s
rest
has
also
rested
from
his
works,
just
as God
did from
his own works.
\VS{11}Thus
we must make every effort
to enter
that
rest,
so that
no
one
may fall
by following the same pattern
of disobedience.
\VS{12}For
the word
of God
is living
and
active
and
sharper
than
any
double-edged
sword,
piercing
even
to
the point of dividing
soul
from spirit,
and
joints
from marrow;
it is able to judge
the desires
and
thoughts
of the heart.
\VS{13}And
no
creature
is
hidden
from
God,
but everything
is naked
and
exposed
to the eyes
of him
to
whom
we
must render an account.
\par }{\SH Jesus Our Compassionate High Priest
\par }{\PP \VS{14}Therefore
since
we have
a great
high priest
who has passed through
the heavens,
Jesus
the Son
of God,
let us hold fast
to our confession.
\VS{15}For
we do
not
have
a high priest
incapable
of sympathizing
with our
weaknesses,
but
one who has been tempted
in
every way
just as we are,
yet without
sin.
\VS{16}Therefore
let us
confidently
approach
the throne
of grace
to
receive
mercy
and
find
grace
whenever we need
help.

\par }\Chap{5}{\PP \VerseOne{1}For
every
high priest
is taken
from among
the people
and appointed to represent
them
before
God,
to
offer
both
gifts
and
sacrifices
for
sins.
\VS{2}He is able
to deal compassionately
with those who are ignorant
and
erring,
since
he
also
is subject
to weakness,
\VS{3}and
for
this reason
he is obligated
to make
sin
offerings
for
himself
as well as
for
the people.
\VS{4}And
no
one
assumes
this honor
on his own initiative,
but
only when called
to it by
God,
as
in fact Aaron was.
\VS{5}So
also
Christ
did
not
glorify
himself
in becoming
high priest,
but
the one who glorified him was God, who said
to
him, “{\QT{
You
are
my
Son! Today
I
have fathered
you}},”
\VS{6}as
also
in
another place
God says, “{\QT{
You
are a priest
forever
in
the order
of Melchizedek}}.”
\VS{7}During
his
earthly
life Christ offered
both
requests
and
supplications,
with
loud
cries
and
tears,
to
the one who was able
to save
him
from
death
and
he was heard
because of
his devotion.
\VS{8}Although
he was
a son,
he learned
obedience
through
the things
he suffered.
\VS{9}And
by being perfected
in this way, he became
the source
of eternal
salvation
to all
who obey
him,
\VS{10}and he was designated
by
God
as high priest
{\QT{
in
the order
of Melchizedek}}.
\par }{\SH The Need to Move on to Maturity
\par }{\PP \VS{11}On
this topic
we
have much
to say
and it is difficult to explain,
since
you have become
sluggish
in hearing.
\VS{12}For
though
you should
in fact be
teachers
by
this time,
you need
someone to teach
you
the beginning
elements
of God’s
utterances.
You have gone back
to needing
milk,
not
solid
food.
\VS{13}For
everyone
who lives
on milk
is inexperienced
in the message
of righteousness,
because
he is
an infant.
\VS{14}But
solid
food
is
for the mature,
whose perceptions
are trained
by
practice
to
discern
both
good
and
evil.

\par }\Chap{6}{\PP \VerseOne{1}Therefore
we must progress
beyond the elementary
instructions
about Christ
and move on
to
maturity,
not
laying
this foundation
again: repentance
from
dead
works
and
faith
in
God,
\VS{2}teaching
about baptisms,
laying on
of hands,
resurrection
of the dead,
and
eternal
judgment.
\VS{3}And
this
is what we intend to do,
if
God
permits.
\VS{4}For
it is impossible
in the case of those who have
once
been enlightened,
tasted
the heavenly
gift,
become
partakers
of the Holy
Spirit,
\VS{5}tasted
the good
word
of God
and
the miracles
of the coming
age,
\VS{6}and then
have committed apostasy,
to renew
them again
to
repentance,
since they are crucifying
the Son
of God
for themselves
all over again and
holding
him up to contempt.
\VS{7}For
the ground
that has soaked up
the rain
that frequently
falls
on
it
and
yields
useful
vegetation
for those
who
tend
it receives a
blessing
from
God.
\VS{8}But
if it produces
thorns
and
thistles,
it is useless
and
about
to be cursed;
its fate
is to be burned.
\VS{9}But
in
your
case, dear friends,
even though
we speak
like this,
we are convinced
of better things
relating
to salvation.
\VS{10}For
God
is not
unjust
so as to forget
your
work
and
the love
you have demonstrated
for
his
name,
in having served
and
continuing to serve
the saints.
\VS{11}But
we passionately want
each
of you
to demonstrate
the same
eagerness
for
the fulfillment
of your hope
until
the end,
\VS{12}so that
you may
not
be
sluggish,
but
imitators
of those who through
faith
and
perseverance
inherit
the promises.
\par }{\PP \VS{13}Now when
God
made
his promise
to Abraham,
since
he could swear
by
no one
greater,
he swore
by
himself,
\VS{14}saying, “{\QT{
Surely
I will bless
you
greatly and
multiply
your descendants
abundantly}}.”
\VS{15}And
so
by persevering,
Abraham inherited
the promise.
\VS{16}For
people
swear
by
something greater than
themselves, and
the oath
serves as a confirmation
to end
all
dispute.
\VS{17}In
the same way God
wanted
to demonstrate
more clearly
to the heirs
of the promise
that his
purpose
was unchangeable,
and so he intervened
with an oath,
\VS{18}so that
we who have
found refuge
in him may find strong
encouragement
to hold fast
to the hope
set before us
through
two
unchangeable
things,
since it is impossible
for God
to lie.
\VS{19}We have
this hope as
an anchor
for the soul,
sure
and
steadfast,
which reaches
inside behind
the curtain,
\VS{20}where
Jesus
our forerunner
entered
on
our
behalf,
since
he became
{\BD{
{\QT{
{\IT{a priest}}}}
forever
in
the order
of Melchizedek
forever
in
the order
of Melchizedek}}.

\par }\Chap{7}{\PP \VerseOne{1}Now this
{\QT{
Melchizedek, king
of Salem, priest
of the most high
God, met
Abraham
as he was returning
from
defeating
the kings}}
and
{\QT{
blessed
him}}.
\VS{2}To him
also
{\BD{
{\QT{t}}
{\IT{Abraham}}
apportioned
a tithe
apportioned
a tithe}}
{\QT{of}}
everything.
His name first
means
king
of righteousness,
then
{\BD{
{\QT{
{\IT{king}}}}
of Salem
of Salem}}, that
is,
king
of peace.
\VS{3}Without father,
without mother,
without genealogy,
he has
neither
beginning
of days
nor
end
of life
but
is like
the son
of God,
and he remains
a priest
for
all time.
\VS{4}But
see
how great
he
must be, if Abraham
the patriarch
gave
him a tithe
of his plunder.
\VS{5}And
those of the sons
of Levi
who receive
the priestly office
have
authorization
according to
the law
to collect a tithe
from the people,
that
is,
from their
fellow countrymen,
although
they too are descendants
of Abraham.
\VS{6}But
Melchizedek who does
not
share
their
ancestry
collected a tithe
from Abraham
and
blessed
the one who possessed
the promise.
\VS{7}Now
without
dispute
the inferior
is blessed
by
the superior,
\VS{8}and
in one case
tithes
are received
by mortal
men,
while in the other
by him who is affirmed
to be alive.
\VS{9}And
it could be said
that Levi
himself, who receives
tithes,
paid a tithe
through
Abraham.
\VS{10}For
he was
still
in
his ancestor
Abraham’s loins
when
Melchizedek
met
him.
\par }{\SH Jesus and the Priesthood of Melchizedek
\par }{\PP \VS{11}So
if
perfection
had in fact been possible through
the Levitical
priesthood
– for
on
that basis
the people
received the law
– what
further
need
would there have been for another
priest
to arise,
said
to be in
the order
of Melchizedek
and
not
in
Aaron’s
order?
\VS{12}For
when
the priesthood
changes,
a change
in the law
must come as well.
\VS{13}Yet
the one
these things
are spoken about
belongs
to a different
tribe,
and no one
from
that tribe
has ever officiated
at the altar.
\VS{14}For
it is clear
that
our
Lord
is descended
from
Judah,
yet Moses
said
nothing
about
priests
in
connection with that
tribe.
\VS{15}And
this is
even
clearer
if
another
priest
arises
in
the likeness
of Melchizedek,
\VS{16}who
has become a priest not
by
a legal
regulation
about physical
descent
but
by
the power
of an indestructible
life.
\VS{17}For
here is the testimony
about him: “{\QT{
You
are a priest
forever
in
the order
of Melchizedek}}.”
\VS{18}On the one hand
a former
command
is
set aside
because
it
is weak
and
useless,
\VS{19}for
the law
made
nothing
perfect.
On the other hand
a better
hope
is introduced,
through
which
we draw near
to God.
\VS{20}And
since
this was not
done without
a sworn affirmation
– for
the others
have become
priests
without
a sworn affirmation,
\VS{21}but
Jesus did so with
a sworn affirmation
by
the one who said
to
him, “{\QT{
The Lord
has sworn
and
will
not
change
his mind, ‘You
are a priest
forever}}’ ” –
\VS{22}accordingly
Jesus
has become
the guarantee
of a better
covenant.
\VS{23}And
the others
who became
priests
were
numerous,
because
death
prevented
them from continuing in office,
\VS{24}but
he holds
his priesthood
permanently
since
he
lives
forever.
\VS{25}So
he is able
to save
completely
those who come
to God
through
him,
because he always
lives
to
intercede
for
them.
\VS{26}For
it is indeed fitting
for us
to have such
a high priest: holy,
innocent,
undefiled,
separate
from
sinners,
and
exalted
above
the heavens.
\VS{27}He has
no
need
to do every
day
what those priests
do, to offer
sacrifices
first
for
their own
sins
and then
for the sins of the people,
since
he did
this
in offering
himself
once for all.
\VS{28}For
the law
appoints
as high priests
men
subject
to weakness,
but
the word
of solemn affirmation
that came after
the law
appoints a son
made perfect
forever.

\par }\Chap{8}{\PP \VerseOne{1}Now the main point
of
what we are saying
is this: We have
such
a high priest,
one who
{\IT{
sat down
at
the right hand
of the throne
of the Majesty
in
heaven}},
\VS{2}a minister
in the sanctuary
and
the true
tabernacle
that
the Lord,
not
man,
set up.
\VS{3}For
every
high priest
is appointed
to offer
both
gifts
and
sacrifices.
So
this
one
too had to have
something
to offer.
\VS{4}Now if
he were
on
earth,
he would
not
be
a priest,
since there are
already priests who offer
the gifts
prescribed by
the law.
\VS{5}The place where they serve
is a sketch
and
shadow
of the heavenly sanctuary,
just as
Moses
was warned by God
as he was about
to complete
the tabernacle.
For
he says, “{\QT{
See
that you make
everything
according to
the design
shown
to you
on
the mountain}}.”
\VS{6}But
now
Jesus has obtained
a superior
ministry,
since the covenant
that he
mediates
is also better and is enacted
on
better
promises.
\par }{\PP \VS{7}For
if
that
first
covenant had been
faultless,
no
one
would have looked for
a second
one.
\VS{8}But
showing
its fault,
God says
to them,
\par }{\Q “{\QT{
Look, the days
are coming, says
the Lord, when I will complete
a new
covenant
with
the house
of Israel
and
with
the house
of Judah}}.
\par }{\Q \VS{9}“{\QT{
It will not
be like
the covenant
that
I made
with their
fathers, on
the day
when
I
took
them
by the hand
to lead
them
out of
Egypt, because
they
did
not
continue
in
my
covenant
and I
had no regard
for them, says
the Lord}}*.
\par }{\Q \VS{10}“{\QT{
For
this
is the covenant
that
I will establish
with the house
of Israel
after
those
days, says
the Lord. I will put
my
laws
in
their
minds
and
I will inscribe
them
on
their
hearts. And
I will be
their
God
and
they
will be
my
people}}.
\par }{\Q \VS{11}“{\QT{
And
there will be
no
need
at all for each one
to teach
his
countryman
or
each one
to teach his
brother
saying, ‘Know
the Lord,’ since
they will
all
know
me, from
the least
to
the greatest}}.
\par }{\Q \VS{12}“{\QT{
For
I will be
merciful
toward their
evil deeds, and
their
sins
I will remember
no
longer}}.”
\par }{\PP \VS{13}When he speaks
of a new covenant,
he makes
the first
obsolete.
Now
what is growing obsolete
and
aging
is about to
disappear.

\par }\Chap{9}{\PP \VerseOne{1}Now the first
covenant, in fact,
had
regulations
for worship
and
its earthly
sanctuary.
\VS{2}For
a tent
was prepared,
the outer
one, which
contained the lampstand,
the table,
and
the presentation
of the loaves;
this is called
the holy place.
\VS{3}And
after
the second
curtain
there was
a tent
called
the holy
of holies.
\VS{4}It contained
the golden
altar of incense
and
the ark
of the covenant
covered
entirely
with gold.
In
this ark were
the golden
urn
containing the manna,
Aaron’s
rod
that budded,
and
the stone tablets
of the covenant.
\VS{5}And
above
the ark
were the cherubim
of glory
overshadowing
the mercy seat.
Now
is
not
the time to speak
of
these things
in
detail.
\VS{6}So
with these things
prepared
like this, the priests
enter continually
into
the outer
tent
as they perform
their duties.
\VS{7}But
only
the high priest
enters once
a year
into the inner tent, and not
without
blood
that
he offers
for
himself
and
for the sins of the people
committed in ignorance.
\VS{8}The Holy
Spirit
is making clear
that
the way
into the holy place
had
not yet
appeared
as long as the old
tabernacle
was
standing.
\VS{9}This was a symbol
for
the time
then present,
when
gifts
and
sacrifices
were offered
that could
not
perfect
the conscience
of the worshiper.
\VS{10}They served only
for matters of
food
and
drink
and
various
washings;
they are external
regulations
imposed
until
the new order came.
\par }{\SH Christ’s Service in the Heavenly Sanctuary
\par }{\PP \VS{11}But
now Christ
has come
as the high priest
of the good things
to come.
He passed through
the greater
and
more perfect
tent
not
made with hands,
that
is,
not
of this
creation,
\VS{12}and he entered
once for all
into
the most holy place
not
by
the blood
of goats
and
calves
but
by
his own
blood,
and so he himself secured
eternal
redemption.
\VS{13}For
if
the blood
of goats
and
bulls
and
the ashes
of a young cow
sprinkled
on those who are defiled
consecrated
them and provided
ritual
purity,
\VS{14}how much
more
will the blood
of Christ,
who
through
the eternal
Spirit
offered
himself
without blemish
to God,
purify
our
consciences
from
dead
works
to
worship
the living
God.
\par }{\PP \VS{15}And
so
he is
the mediator
of a new
covenant,
so that
those who are called
may receive
the eternal
inheritance
he has promised,
since he died
to
set
them free
from the violations committed
under
the first
covenant.
\VS{16}For
where
there is a will,
the death
of the one who made it
must
be proven.
\VS{17}For
a will
takes effect
only at
death,
since
it carries no
force
while
the one who made
it is alive.
\VS{18}So even
the first covenant
was inaugurated
with
blood.
\VS{19}For
when
Moses
had spoken
every
command
to all
the people
according to
the law,
he took
the blood
of calves
and
goats
with
water
and
scarlet
wool
and
hyssop
and sprinkled
both
the book
itself and
all
the people,
\VS{20}and said, “{\QT{
This
is the blood
of the covenant
that
God
has commanded
you
to keep}}.”
\VS{21}And
both
the tabernacle
and
all
the utensils
of worship
he
likewise
sprinkled
with blood.
\VS{22}Indeed according to
the law
almost
everything
was purified
with
blood,
and
without
the shedding of blood
there is
no
forgiveness.
\VS{23}So
it was necessary
for the sketches of the things
in
heaven
to be purified
with these sacrifices,
but
the heavenly things
themselves
required better
sacrifices
than
these.
\VS{24}For
Christ
did
not
enter
a sanctuary
made with hands
– the representation
of the true sanctuary –
but
into
heaven
itself,
and he appears
now
in God’s
presence
for
us.
\VS{25}And
he did not
enter to offer
himself
again and again,
the way the high priest
enters
the sanctuary
year after year
with
blood
that is not his own,
\VS{26}for then
he
would have had
to suffer
again and again
since
the foundation
of
the world.
But
now
he has appeared
once for all
at
the consummation
of the ages
to
put away
sin
by
his sacrifice.
\VS{27}And
just as
people
are appointed
to die
once,
and
then
to face judgment,
\VS{28}so
also,
after
Christ
was offered
once
to
{\IT{
bear
the sins
of many}}, to those who eagerly await
him he will appear
a second time,
not
to bear sin
but to
bring salvation.

\par }\Chap{10}{\PP \VerseOne{1}For
the law
possesses
a shadow
of the good things
to come
but not
the reality
itself,
and is therefore completely unable,
by the same
sacrifices
offered
continually,
year after year,
to perfect
those who come to worship.
\VS{2}For otherwise
would they
not
have ceased
to be offered,
since
the worshipers
would have been purified
once for all
and so have
no
further
consciousness
of sin?
\VS{3}But
in
those sacrifices
there is a reminder
of sins
year after year.
\VS{4}For
the blood
of bulls
and
goats
cannot
take away
sins.
\VS{5}So
when
he came
into
the world,
he said,
\par }{\Q “{\QT{
Sacrifice
and
offering
you did
not
desire, but
a body
you prepared
for me}}.
\par }{\Q \VS{6}“{\IT{
{\BD{Whole burnt offerings}}
and
sin-offerings
you took
no
delight in.}}
\par }{\Q \VS{7}“{\QT{
Then
I said, ‘Here I am: I have come
– it is written
of
me
in
the scroll
of the book
– to do
your
will, O God}}*.’ ”
\par }{\PP \VS{8}When
he says
above, “{\QT{
Sacrifices
and
offerings}}
and
{\QT{
whole burnt offerings
and
sin-offeringsyou did
not
desire
nor
did you take delight}}
in them” (which
are offered
according
to the law),
\VS{9}then
he says, “{\QT{
Here
I am: I have come
to do
your
will}}.” He does away
with the first
to
establish
the second.
\VS{10}By
his will
we have been
made holy
through
the offering
of the body
of Jesus
Christ
once for all.
\VS{11}And
every
priest
stands
day after day
serving
and offering
the same
sacrifices
again and again
– sacrifices that
can
never
take away
sins.
\VS{12}But
when this
priest had offered
one
sacrifice
for
sins
for
all time,
{\IT{
he sat down
at
the right hand}}
of God,
\VS{13}where he is now
waiting
{\IT{
until
his
enemies
are made
a footstool
for his
feet}}.
\VS{14}For
by one
offering
he has perfected
for
all time
those who are made holy.
\VS{15}And
the Holy
Spirit
also
witnesses
to us,
for
after
saying,
\VS{16}“{\QT{
This
is the covenant
that
I will establish
with
them
after
those
days, says
the Lord. I will put
my
laws
on
their
hearts
and
I will inscribe
them
on
their
minds}},”
\VS{17}then
he says, “{\QT{
Their
sins
and
their
lawless
deeds I will remember
no
longer}}.”
\VS{18}Now
where
there is forgiveness
of these,
there is no longer
any offering
for
sin.
\par }{\SH Drawing Near to God in Enduring Faith
\par }{\PP \VS{19}Therefore,
brothers and sisters,
since we have
confidence
to
enter
the sanctuary
by
the blood
of Jesus,
\VS{20}by the fresh
and
living
way
that
he inaugurated
for us
through
the curtain,
that
is,
through his
flesh,
\VS{21}and
since we have a great
priest
over
the house
of God,
\VS{22}let us draw near
with
a sincere
heart
in
the assurance
that faith
brings, because we have had our hearts
sprinkled clean
from
an evil
conscience
and
our bodies
washed
in pure
water.
\VS{23}And let us hold
unwaveringly
to the hope
that we confess,
for
the one who made the promise
is trustworthy.
\VS{24}And
let us take thought
of how to spur
one another
on to
love
and
good
works,
\VS{25}not
abandoning
our own
meetings,
as
some
are in the habit
of doing, but
encouraging
each other, and
even
more
so because you see
the day
drawing near.
\par }{\PP \VS{26}For
if we deliberately
keep on sinning
after
receiving
the knowledge
of the truth,
no further
sacrifice
for
sins
is left
for us,
\VS{27}but
only a certain
fearful
expectation
of judgment
and
{\IT{
a fury}}
{\IT{
of fire
that will
consume
God’s enemies}}.
\VS{28}Someone
who rejected
the law
of Moses
was put to death
without
mercy
{\IT{
on
the testimony of two
or
three
witnesses}}.
\VS{29}How much
greater
punishment
do you think
that person deserves
who has contempt
for the Son
of God,
and
profanes
the blood
of the covenant
that made
him holy,
and
insults
the Spirit
of grace?
\VS{30}For
we know
the one who said, “{\QT{
Vengeance
is mine, I
will repay}},” and
again, “{\QT{
The Lord
will judge
his
people}}.”
\VS{31}It is a terrifying
thing to fall
into
the hands
of the living
God.
\par }{\PP \VS{32}But
remember
the former
days
when you endured
a harsh
conflict
of suffering
after
you were enlightened.
\VS{33}At times you were publicly exposed
to abuse and
afflictions,
and at other times you came
to share
with others who were treated
in that way.
\VS{34}For
in fact you shared the sufferings
of those in prison,
and
you accepted
the confiscation
of your
belongings
with
joy,
because you knew
that you
certainly had
a better
and
lasting
possession.
\VS{35}So
do
not
throw away
your
confidence,
because it has
great
reward.
\VS{36}For
you need
endurance
in order
to do
God’s
will
and so receive
what is promised.
\VS{37}For
{\QT{
just a little
longer}}
and
{\QT{
he who is coming
will arrive
and
not
delay}}.
\VS{38}{\QT{
But
my
righteous one
will live
by
faith, and
if
he shrinks back, I
take
no
pleasure
in
him}}.
\VS{39}But
we
are
not
among those who shrink back
and thus
perish,
but
are among those who have faith
and preserve
their souls.

\par }\Chap{11}{\PP \VerseOne{1}Now faith
is
being sure
of what
we hope for,
being convinced
of what we do
not
see.
\VS{2}For
by
it
the people of old
received God’s commendation.
\VS{3}By faith
we understand
that the worlds
were set in order
at God’s
command,
so that the visible
has its origin
in
the invisible.
\VS{4}By faith
Abel
offered
God
a greater
sacrifice
than
Cain,
and through
his faith
he was commended
as righteous,
because God
commended
him for
his
offerings.
And
through
his
faith he
still
speaks,
though he is dead.
\VS{5}By faith
Enoch
was taken up
so that he did not
see
death,
and
he was
not
to be found
because
God
took
him
up.
For
before
his removal
he had been commended
as having pleased
God.
\VS{6}Now without
faith
it is impossible
to please
him, for
the one who approaches
God
must
believe
that
he exists
and
that he rewards
those who seek
him.
\VS{7}By faith
Noah,
when
he was warned
about
things not yet
seen,
with reverent regard
constructed
an ark
for
the deliverance
of his
family.
Through
faith
he condemned
the world
and
became
an heir
of the righteousness
that comes by
faith.
\par }{\PP \VS{8}By faith
Abraham
obeyed
when
he was called
to go out
to
a place
he would
later receive
as
an inheritance,
and
he went out
without
understanding
where
he was going.
\VS{9}By faith
he lived as a foreigner
in
the promised
land
as though
it were a foreign country,
living
in
tents
with
Isaac
and
Jacob,
who were fellow heirs
of the same
promise.
\VS{10}For
he was looking forward
to the city
with firm
foundations,
whose
architect
and
builder
is God.
\VS{11}By faith,
even though Sarah
herself
was barren and he was too old,
he received
the ability
to
procreate,
because
he regarded
the one who had given the promise
to be trustworthy.
\VS{12}So
in fact children were fathered
by
one man
– and
this one
as good as
dead –
{\IT{
like the number
of stars
in the sky
and
like
the innumerable
grains of sand
on
the seashore}}.
\VS{13}These
all
died
in
faith
without
receiving
the things promised,
but
they saw
them
in the distance
and
welcomed
them and
acknowledged
that
they were
strangers
and
foreigners
on
the earth.
\VS{14}For
those who speak
in such a way
make
it clear
that
they are seeking
a homeland.
\VS{15}In fact,
if
they had been thinking
of
the land
that
they had left,
they would have had
opportunity
to return.
\VS{16}But
as it is, they aspire
to a better
land, that
is,
a heavenly one.
Therefore,
God
is not
ashamed
to be called
their
God,
for he has prepared
a city
for
them.
\VS{17}By faith
Abraham,
when
he was tested,
offered up
Isaac.
He had received
the promises,
yet
he was ready to offer up
his only son.
\VS{18}God
had told
him, “{\QT{
Through
Isaac
descendants
will carry on
your
name}},”
\VS{19}and
he reasoned
that
God
could
even raise
him from
the dead,
and
in
a sense
he received
him
back
from there.
\VS{20}By faith
also
Isaac
blessed
Jacob
and
Esau
concerning
the future.
\VS{21}By faith
Jacob,
as he was dying,
blessed
each
of the sons
of Joseph
and
{\BD{
{\QT{
{\IT{worshiped}}}}
as he leaned
on
his
staff
as he leaned
on
his
staff}}.
\VS{22}By faith
Joseph,
at the end of
his life,
mentioned
the exodus
of the sons
of Israel
and
gave instructions
about
his
burial.
\par }{\PP \VS{23}By faith,
when Moses
was born,
his
parents
hid
him for three months,
because
they saw
the child
was beautiful
and
they were
not
afraid
of the king’s
edict.
\VS{24}By faith,
when he grew up,
Moses
refused
to be called
the son
of Pharaoh’s
daughter,
\VS{25}choosing
rather
to be ill-treated
with the people
of God
than
to enjoy
sin’s
fleeting pleasure.
\VS{26}He regarded
abuse suffered
for Christ
to be greater
wealth
than
the treasures
of Egypt,
for
his eyes were fixed
on
the reward.
\VS{27}By faith
he left
Egypt
without
fearing
the king’s
anger,
for
he persevered
as though
he could see
the one who is invisible.
\VS{28}By faith
he kept
the Passover
and
the sprinkling
of the blood,
so that
the one who destroyed
the firstborn
would
not
touch
them.
\VS{29}By faith
they crossed
the Red
Sea
as
if on
dry
ground,
but when the Egyptians
tried
it, they were swallowed up.
\VS{30}By faith
the walls
of Jericho
fell
after the people marched around
them for
seven
days.
\VS{31}By faith
Rahab
the prostitute
escaped the destruction
of the disobedient,
because she welcomed
the spies
in
peace.
\par }{\PP \VS{32}And
what
more
shall I say? For
time
will fail
me
if I tell
of
Gideon,
Barak,
Samson,
Jephthah,
of David
and
Samuel
and
the prophets.
\VS{33}Through
faith
they conquered
kingdoms,
administered
justice,
gained
what was promised,
shut
the mouths
of lions,
\VS{34}quenched
raging
fire,
escaped
the edge
of
the sword,
gained strength
in
weakness,
became
mighty
in
battle,
put
foreign
armies
to flight,
\VS{35}and women
received back
their
dead
raised to life.
But
others
were tortured,
not
accepting
release,
to obtain
resurrection
to
a better life.
\VS{36}And
others
experienced
mocking
and
flogging,
and
even
chains
and
imprisonment.
\VS{37}They were stoned,
sawed apart,
murdered
with
the sword;
they went about
in
sheepskins
and goatskins;
they were destitute,
afflicted,
ill-treated
\VS{38}(the world
was
not
worthy
of them); they wandered
in
deserts
and
mountains
and
caves
and
openings
in the earth.
\VS{39}And
these
all
were commended
for
their faith,
yet they did
not
receive
what was promised.
\VS{40}For God
had provided
something
better
for
us,
so that
they would be made perfect
together with
us.

\par }\Chap{12}{\PP \VerseOne{1}Therefore,
since we
are surrounded
by such a great
cloud
of witnesses,
we must get rid
of every
weight
and
the sin
that clings so closely,
and run
with
endurance
the race
set out
for us,
\VS{2}keeping
our eyes fixed
on Jesus,
the pioneer
and
perfecter
of our faith.
For
the joy
set out
for him
he endured
the cross,
disregarding
its shame,
and
{\IT{
has taken
his seat
at
the right hand
of the throne}}
of God.
\VS{3}Think
of him who endured
such
opposition
against
himself
by
sinners,
so that
you may
not
grow weary
in your
souls
and give up.
\VS{4}You have
not yet
resisted
to the point
of bloodshed
in your struggle
against
sin.
\VS{5}And
have you forgotten
the exhortation
addressed
to you
as
sons?
\par }{\Q “{\QT{
My
son, do
not
scorn
the Lord’s
discipline}}
\par }{\Q {\QT{
or
give up
when
he
corrects you}}*.
\par }{\Q \VS{6}“{\IT{
{\BD{For}}
the Lord
disciplines
the one
he loves
and
chastises
every
son
he accepts.}}”
\par }{\PP \VS{7}Endure
your suffering
as
discipline;
God
is treating
you
as
sons.
For
what
son
is there that
a father
does
not
discipline?
\VS{8}But
if
you do not
experience discipline,
something all sons
have shared
in, then
you are illegitimate
and
are
not
sons.
\VS{9}Besides,
we have experienced
discipline
from our
earthly
fathers
and
we respected
them; shall we
not
submit
ourselves all the more
to the Father
of spirits
and
receive life?
\VS{10}For
they disciplined
us for
a little while
as
seemed
good to them,
but
he does so for
our benefit,
that we may share
his
holiness.
\VS{11}Now
all
discipline
seems
painful
at the time,
not
joyful.
But
later
it produces
the fruit
of peace
and righteousness
for those trained
by
it.
\VS{12}Therefore,
{\BD{
{\QT{
{\IT{strengthen}}}}}}
{\QT{your listless}}
hands
and
your weak
knees,
\VS{13}and
{\BD{
{\QT{
{\IT{make}}}}
straight
paths
for your
feet
straight
paths
for your
feet}}, so that
what is lame
may
not
be put out of joint
but
be healed.
\par }{\SH Do Not Reject God’s Warning
\par }{\PP \VS{14}Pursue
peace
with
everyone,
and
holiness,
for without
it no one
will see
the Lord.
\VS{15}See
to it that no
one
comes short
of
the grace
of God,
that no
one
be like
{\IT{
a bitter
root
springing up}}
and causing trouble,
and
through
him
many
become defiled.
\VS{16}And see to it that no
one
becomes an immoral
or
godless person
like
Esau,
who
{\IT{
sold
his own
birthright
for
a single
meal}}.
\VS{17}For
you know
that
later
when
he wanted
to inherit
the blessing,
he was rejected,
for
he found
no
opportunity
for repentance,
although
he sought
the blessing
with
tears.
\VS{18}For
you have
not
come to
something that can be touched,
to a burning
fire
and
darkness
and
gloom
and
a whirlwind
\VS{19}and
the blast
of a trumpet
and
a voice
uttering words
such that those who heard
begged
to hear
no more.
\VS{20}For
they could
not
bear
what was commanded: “{\QT{
If
even an animal
touches
the mountain, it must be stoned}}.”
\VS{21}In fact, the scene
was
so
terrifying
that Moses
said, “{\QT{
I
shudder
with fear}}.”
\VS{22}But
you have come
to Mount
Zion,
the city
of the living
God,
the heavenly
Jerusalem,
and
to myriads
of angels, to the assembly
\VS{23}and
congregation
of the firstborn,
who are enrolled
in
heaven,
and
to God,
the judge
of all,
and
to the spirits
of the righteous,
who have been made perfect,
\VS{24}and
to Jesus,
the mediator
of a new
covenant,
and
to the sprinkled
blood
that speaks
of something better
than
Abel’s does.
\par }{\PP \VS{25}Take care
not
to refuse
the one who is speaking! For
if
they did
not
escape
when
they refused
the one who warned
them on
earth,
how much less
shall we,
if we reject
the one who warns from
heaven?
\VS{26}Then
his voice
shook
the earth,
but
now
he has promised, “{\QT{
I
will
once
more
shake
not
only
the earth
but
heaven
too}}.”
\VS{27}Now this phrase “{\QT{
once
more}}”
indicates
the removal
of what is shaken,
that is,
of created things,
so that
what is unshaken
may remain.
\VS{28}So
since
we are receiving
an unshakable
kingdom,
let us give
thanks,
and through
this
let us offer worship
pleasing
to God
in
devotion
and
awe.
\VS{29}For
our
{\QT{
God
is indeed a devouring
fire}}.

\par }\Chap{13}{\PP \VerseOne{1}Brotherly love
must continue.
\VS{2}Do
not
neglect
hospitality,
because
through
it
some
have entertained
angels
without knowing it.
\VS{3}Remember
those in prison
as though
you were in prison
with them, and those ill-treated
as though
you too
felt
their
torment.
\VS{4}Marriage
must be honored
among
all
and
the marriage bed
kept undefiled,
for
God
will judge
sexually immoral people
and
adulterers.
\VS{5}Your conduct
must be free from the love of money
and you must be content
with what you have,
for
he
has said, “{\QT{
I will
never
leave
you
and
I will
never
abandon
you}}.”
\VS{6}So
we
can say
with confidence, “{\QT{
The Lord
is my
helper, and I will
not
be afraid. What
can
man
do
to me?}}*”
\VS{7}Remember
your
leaders,
who
spoke
God’s
message
to you;
reflect
on the outcome
of their lives
and imitate
their faith.
\VS{8}Jesus
Christ
is the same
yesterday
and
today
and
forever!
\VS{9}Do
not
be carried away
by all sorts
of strange
teachings.
For
it is good
for the heart
to be strengthened
by grace,
not
ritual meals,
which
have
never
benefited
those who participated in them.
\VS{10}We have
an altar
that those who serve
in the tabernacle
have
no
right
to eat
from.
\VS{11}For
the bodies
of those animals
whose blood
the high priest
brings into
the sanctuary
as an offering for
sin
are burned
outside
the camp.
\VS{12}Therefore,
to
sanctify
the people
by
his own
blood,
Jesus
also
suffered
outside
the camp.
\VS{13}We must go out
to
him,
then, outside
the camp,
bearing
the abuse
he experienced.
\VS{14}For
here
we have
no
lasting
city,
but
we seek
the city that is to come.
\VS{15}Through
him
then let us
continually
offer up
a sacrifice
of praise
to God,
that
is,
the fruit
of our lips,
acknowledging
his
name.
\VS{16}And
do
not
neglect
to do good
and
to share
what you have, for
God
is pleased
with such
sacrifices.
\par }{\PP \VS{17}Obey
your
leaders
and
submit
to them,
for
they keep watch
over
your
souls
and will give
an account
for their work. Let them do
this
with
joy
and
not
with complaints,
for
this
would be no advantage
for you.
\VS{18}Pray
for
us,
for
we are sure
that
we have
a clear
conscience
and desire
to conduct
ourselves rightly
in
every respect.
\VS{19}I
especially
ask
you to pray
that
I may be restored
to you
very soon.
\par }{\SH Benediction and Conclusion
\par }{\PP \VS{20}Now may the God
of peace
who by
the blood
of the eternal
covenant
brought back
from
the dead
the great
shepherd
of the sheep,
our
Lord
Jesus Christ,
\VS{21}equip
you
with
every
good thing
to
do
his
will,
working
in
us
what is pleasing
before
him
through
Jesus
Christ,
to whom
be glory
forever.
Amen.
\par }{\PP \VS{22}Now
I urge
you,
brothers and sisters,
bear with
my message
of exhortation,
for
in fact I have written
to you
briefly.
\VS{23}You should know
that our
brother
Timothy
has been released.
If
he comes
soon,
he will be with me when I see
you.
\VS{24}Greetings
to all
your
leaders
and
all
the saints.
Those from
Italy
send
you
greetings.
\VS{25}Grace
be with
you
all.
\par }