\NormalFont\ShortTitle{Hosea}
{\MT Hosea

\par }\ChapOne{1}{\SH Superscription
\par }{\PP \VerseOne{1}This is the word
of the {\ND{Lord}}
which
was
revealed to
Hosea
son
of Beeri
during the time
when Uzziah,
Jotham,
Ahaz,
and Hezekiah
ruled
Judah,
and during the time
when Jeroboam
son
of Joash
ruled
Israel.
\par }{\SH Symbols of Sin and Judgment: The Prostitute and Her Children
\par }{\PP \VS{2}When the
{\ND{Lord}}
first
spoke
through Hosea,
he said
to
him, “Go
marry
a prostitute
who will bear illegitimate
children
conceived through prostitution,
because
the nation
continually commits spiritual prostitution
by turning away from the
{\ND{Lord}}.”
\VS{3}So
Hosea
married
Gomer,
the daughter
of Diblaim.
Then she conceived
and gave birth
to a son for him.
\VS{4}Then the
{\ND{Lord}}
said to
Hosea, “Name
him ‘Jezreel,’
because
in a little
while
I will punish
the dynasty
of Jehu
on
account of the bloodshed
in the valley of Jezreel,
and I will put an end
to the kingdom
of Israel.
\VS{5}At that time,
I will
destroy the military
power of Israel
in the valley
of Jezreel.”
\par }{\PP \VS{6}She conceived
again
and gave birth
to a daughter.
Then the
{\ND{Lord}} said
to him, “Name
her ‘No Pity’ (Lo-Ruhamah) because
I will no longer
have pity
on the nation of Israel.
For
I will certainly
not forgive their guilt.
\VS{7}But I will have pity
on the nation
of Judah.
I will deliver
them by the
{\ND{Lord}}
their God;
I will not
deliver
them by the warrior’s bow,
by sword,
by military
victory, by chariot horses,
or by chariots.”
\par }{\PP \VS{8}When she had weaned
‘No Pity’ (Lo-Ruhamah) she conceived
again and gave birth
to another son.
\VS{9}Then the
{\ND{Lord}} said: “Name
him ‘Not My People’ (Lo-Ammi), because
you
are not
my people
and I am
not your God.”
\par }{\SH The Restoration of Israel
\par }{\PP \VS{10} However, in the future the number
of the people of
Israel
will be like the sand
of the sea
which
can be neither
measured
nor
numbered.
Although it was said
to them,
“You are not
my people,”
it will be said
to them, “You
are
children
of the living
God!”
\VS{11}Then
the people of Judah
and the people of Israel
will be gathered
together.
They will appoint
for themselves one
leader,
and will flourish
in the land.
Certainly,
the day
of Jezreel
will be great!

\par }\Chap{2}{\PP \VerseOne{1}Then you will call your brother,
“My People”
(Ammi)! You will call your sister,
“Pity” (Ruhamah)!
\par }{\SH Idolatrous Israel Will Be Punished Like a Prostitute
\par }{\Q \VS{2}Plead
earnestly
with your mother
\par }{\Q (for
she
is not
my wife,
and I
am not
her husband),
\par }{\Q so that she might put an end to
her adulterous
lifestyle,
\par }{\Q and turn away from her sexually immoral behavior.
\par }{\Q \VS{3}Otherwise,
I will strip
her naked,
\par }{\Q and expose
her like she was when
she was born.
\par }{\Q I will turn
her land
into a wilderness
\par }{\Q and make
her country
a parched
land,
\par }{\Q so that I might kill
her with thirst.
\par }{\Q \VS{4}I will have no
pity
on her children,
\par }{\Q because
they are children
conceived in adultery.
\par }{\Q \VS{5}For
their mother
has committed adultery;
\par }{\Q she who conceived
them has acted shamefully.
\par }{\Q For
she said,
“I will seek out my lovers;
\par }{\Q they are the ones who give
me my bread
and my water,
\par }{\Q my wool,
my flax,
my olive oil,
and my wine.
\par }{\SH The Lord{\SC{’}}s Discipline Will Bring Israel Back
\par }{\Q \VS{6}Therefore,
I will soon
fence her in
with thorns;
\par }{\Q I will wall
her in so that she cannot
find
her way.
\par }{\Q \VS{7}Then she will pursue
her lovers,
but she will not
catch
them;
\par }{\Q she will seek
them, but she will not
find
them.

\par }{\Q Then she will say,
\par }{\Q “I will go
back
to
my husband,
\par }{\Q because
I was better
off then
than I am now.”
\par }{\SH Agricultural Fertility Withdrawn from Israel
\par }{\Q \VS{8}Yet until now she has refused
to acknowledge
that
I
was the one
\par }{\Q who gave
her the grain,
the new wine,
and the olive oil;
\par }{\Q and that it was I who lavished
on her the silver
and gold
–
\par }{\Q which they used
in worshiping Baal!
\par }{\Q \VS{9}Therefore,
I will take
back
my grain
during the harvest time
\par }{\Q and my new wine
when it ripens;
\par }{\Q I will take away
my wool
and my flax
\par }{\Q which I had provided in order to clothe
her.
\par }{\Q \VS{10}Soon
I will expose
her lewd
nakedness in front
of her lovers,
\par }{\Q and no
one will be able to rescue her from me!
\par }{\Q \VS{11}I will put an end
to all
her celebration:
\par }{\Q her annual
religious festivals,
\par }{\Q monthly new moon
celebrations,
\par }{\Q and weekly
Sabbath festivities –
\par }{\Q all her appointed festivals.
\par }{\Q \VS{12}I will destroy
her vines
and fig trees,
\par }{\Q about which
she said,
“These are my wages
for prostitution

\par }{\Q that
my lovers
gave
to me!”
\par }{\Q I will turn
her cultivated vines and fig trees into an uncultivated thicket,
\par }{\Q so that wild
animals
will devour them.
\par }{\Q \VS{13}“I will punish
her for the festival days
\par }{\Q when she burned incense
to the Baal
idols;

\par }{\Q she adorned
herself with
earrings
and jewelry,
\par }{\Q and went
after
her lovers,
\par }{\Q but she forgot
me!” says
the {\ND{Lord}}.
\par }{\SH Future Repentance and Restoration of Israel
\par }{\Q \VS{14}However, in the future I
will allure
her;

\par }{\Q I
will lead
her back
into the wilderness,
\par }{\Q and speak
tenderly to her.
\par }{\Q \VS{15}From there
I will give back
her vineyards
to her,
\par }{\Q and turn the “Valley
of Trouble”
into an “Opportunity
for Hope.”
\par }{\Q There
she will sing
as she did when
she was young,
\par }{\Q when
she came up
from the land
of Egypt.
\par }{\Q \VS{16}“At that time,”
declares
the {\ND{Lord}},
\par }{\Q “you will call, ‘My husband’;

\par }{\Q you will never
again
call
me, ‘My master.’
\par }{\Q \VS{17}For I will remove
the
names
of the Baal
idols from your lips,
\par }{\Q so that you will never
again
utter
their names!”
\par }{\SH New Covenant Relationship with Repentant Israel
\par }{\Q \VS{18}“At that time
I will make
a covenant
for them with
the wild
animals,
\par }{\Q the birds
of the air,
and the creatures that crawl
on the ground.
\par }{\Q I will abolish
the warrior’s bow
and sword
\par }{\Q – that is, every weapon of warfare – from the land,
\par }{\Q and I will allow them to live securely.”
\par }{\Q \VS{19}I will commit myself to you
forever;
\par }{\Q I will commit myself to you
in righteousness
and justice,
\par }{\Q in steadfast love
and tender compassion.
\par }{\Q \VS{20}I will commit myself to you
in faithfulness;
\par }{\Q then you will acknowledge
the {\ND{Lord}}.”
\par }{\SH Agricultural Fertility Restored to the Repentant Nation
\par }{\Q \VS{21}“At that time,
I will willingly respond,”
declares
the {\ND{Lord}}.
\par }{\Q “I will respond
to the
sky,
\par }{\Q and the sky
will respond
to the
ground;
\par }{\Q \VS{22}then the ground
will respond
to the grain,
the new wine,
and the olive oil;
\par }{\Q and they
will respond
to ‘God Plants’ (Jezreel)!
\par }{\Q \VS{23}Then I will plant
her as my own in the land.
\par }{\Q I will have pity
on ‘No
Pity’ (Lo-Ruhamah).
\par }{\Q I will say
to ‘Not
My People’
(Lo-Ammi), ‘You are my people!’
\par }{\Q And he
will say,
‘You are my God!’ ”

\par }\Chap{3}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to me,
“Go,
show love
to your wife
again,
even though she loves
another man
and continually commits adultery.
Likewise, the
{\ND{Lord}}
loves
the
Israelites
although they
turn
to
other
gods
and love
to offer raisin
cakes to idols.”
\VS{2}So I paid
fifteen
shekels of silver
and about seven bushels
of barley to purchase her.
\VS{3}Then I told
her, “You must live
with me many
days;
you must not
commit adultery
or
have sexual intercourse with another man,
and I
also will wait for you.”
\VS{4}For
the Israelites
must live
many
days
without
a king
or
prince,
without
sacrifice
or
sacred fertility pillar,
without
ephod
or idols.
\VS{5}Afterward,
the Israelites
will turn
and seek
the

{\ND{Lord}}
their God
and their Davidic
king.
Then they will submit to
the
{\ND{Lord}}
in
fear
and receive his blessings
in the future.

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}Hear
the word
of the {\ND{Lord}}, you Israelites!

\par }{\Q For
the {\ND{Lord}}
has a covenant lawsuit
against the people
of Israel.
\par }{\Q For
there is neither
faithfulness
nor loyalty
in the land,
\par }{\Q nor do they acknowledge
God.
\par }{\Q \VS{2}There is only cursing,
lying,
murder,
stealing,
and adultery.
\par }{\Q They resort
to violence
and bloodshed.
\par }{\Q \VS{3}Therefore
the land
will mourn,
\par }{\Q and all
its inhabitants
will perish.
\par }{\Q The wild
animals,
the birds
of the sky,
\par }{\Q and even
the fish
in the sea
will perish.
\par }{\SH The Lord’s Dispute against the Sinful Priesthood
\par }{\Q \VS{4}Do not
let
anyone
accuse
or
contend
against anyone else:

\par }{\Q for my case is against you priests!
\par }{\Q \VS{5}You stumble
day
and night,
\par }{\Q and the false prophets
stumble
with
you;
\par }{\Q You have destroyed
your own people!
\par }{\Q \VS{6}You have destroyed
my people
\par }{\Q by failing
to acknowledge
me!
\par }{\Q Because
you
refuse to acknowledge
me,

\par }{\Q I
will reject
you as my priests.
\par }{\Q Because
you reject
the law
of your God,
\par }{\Q I
will reject
your descendants.
\par }{\Q \VS{7}The more
the priests increased
in numbers,
\par }{\Q the more they rebelled
against
me.
\par }{\Q They
have turned their glorious calling
\par }{\Q into
a shameful
disgrace!
\par }{\Q \VS{8}They feed
on
the sin offerings
of my people;
\par }{\Q their appetites
long for their iniquity!
\par }{\Q \VS{9}I will deal with the people
and priests
together:

\par }{\Q I will punish
them both for their ways,
\par }{\Q and I will repay
them for their deeds.
\par }{\Q \VS{10}They will eat,
but not
be satisfied;
\par }{\Q they will engage in prostitution,
but not
increase in numbers;
\par }{\Q because
they have abandoned
the {\ND{Lord}}
\par }{\Q by pursuing other gods.
\par }{\SH Judgment of Pagan Idolatry and Cultic Prostitution
\par }{\Q \VS{11}Old
and new wine
\par }{\Q take away
the understanding of my people.
\par }{\Q \VS{12}They consult their wooden
idols,
\par }{\Q and their diviner’s
staff
answers
with an oracle.
\par }{\Q The wind
of prostitution
blows them astray;
\par }{\Q they commit spiritual
adultery
against their God.
\par }{\Q \VS{13}They sacrifice
on
the mountaintops,
\par }{\Q and burn offerings
on
the hills;
\par }{\Q they sacrifice under
oak,
poplar,
and terebinth,
\par }{\Q because
their shade
is so pleasant.
\par }{\Q As a result,
your daughters
have become cult prostitutes,
\par }{\Q and your daughters-in-law
commit adultery!
\par }{\Q \VS{14}I will not
punish
your daughters
when
they commit prostitution,
\par }{\Q nor your daughters-in-law
when
they commit adultery.
\par }{\Q For
the men
consort
with
harlots,
\par }{\Q they sacrifice
with
temple prostitutes.
\par }{\Q It is true: “A people
that lacks
understanding
will come to ruin!”
\par }{\SH Warning to Judah: Do Not Join in Israel’s Apostasy!
\par }{\Q \VS{15}Although
you,
O Israel,
commit adultery,
\par }{\Q do not
let Judah
become guilty!
\par }{\Q Do not
journey
to Gilgal!
\par }{\Q Do not
go up
to Beth Aven!

\par }{\Q Do not
swear,
“As surely as the
{\ND{Lord}}
lives!”
\par }{\Q \VS{16}Israel
has
rebelled like
a stubborn
heifer!
\par }{\Q Soon
the {\ND{Lord}}
will put them out to pasture
\par }{\Q like a lamb
in a broad field!
\par }{\Q \VS{17}Ephraim
has attached
himself to idols;
\par }{\Q Do not go near him!
\par }{\SH The Shameful Sinners Will Be Brought to Shame
\par }{\Q \VS{18}They consume their alcohol,
\par }{\Q then engage in cult prostitution;
\par }{\Q they
dearly
love
their shameful behavior.
\par }{\Q \VS{19}A whirlwind
has wrapped
them in its wings;
\par }{\Q they will be brought to shame because of their idolatrous worship.

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}Hear
this,
you priests!
\par }{\Q Pay attention,
you Israelites!

\par }{\Q Listen closely,
O king!

\par }{\Q For
judgment
is about to overtake you!

\par }{\Q For
you were like a trap
to Mizpah,
\par }{\Q like a net
spread
out to catch Tabor.
\par }{\Q \VS{2}Those who revolt
are knee-deep
in slaughter,
\par }{\Q but I
will discipline
them all.
\par }{\Q \VS{3}I
know
Ephraim
all too well;

\par }{\Q the evil of Israel
is not
hidden
from
me.
\par }{\Q For
you have engaged in prostitution,
O Ephraim;
\par }{\Q Israel
has defiled itself.
\par }{\Q \VS{4}Their wicked deeds do not
allow
them to return
to
their God;
\par }{\Q because
a spirit
of idolatry
controls
their heart,

\par }{\Q and they do not
acknowledge
the {\ND{Lord}}.
\par }{\Q \VS{5}The arrogance
of Israel
testifies
against it;
\par }{\Q Israel
and Ephraim
will be overthrown
because of their iniquity.
\par }{\Q Even
Judah
will be brought down
with them.
\par }{\SH The Futility of Sacrificial Ritual without Moral Obedience
\par }{\Q \VS{6}Although they
bring their
flocks
and herds
\par }{\Q to seek
the favor of the

{\ND{Lord}},

\par }{\Q They will not
find him –
\par }{\Q he has withdrawn himself from them!
\par }{\Q \VS{7}They have committed treason
against the
{\ND{Lord}},
\par }{\Q because
they bore
illegitimate
children.
\par }{\Q Soon
the new moon
festival will devour
them and their fields.
\par }{\SH The Prophet’s Declaration of Judgment
\par }{\Q \VS{8}Blow
the ram’s horn
in Gibeah!
\par }{\Q Sound the trumpet
in Ramah!
\par }{\Q Sound the alarm
in Beth Aven!

\par }{\Q Tremble
in fear, O Benjamin!
\par }{\Q \VS{9}Ephraim
will be
ruined
in the day
of judgment!

\par }{\Q What I am declaring
to the tribes
of Israel
will certainly take place!
\par }{\SH The Oppressors of the Helpless Will Be Oppressed
\par }{\Q \VS{10}The princes
of Judah
are like those who move
boundary markers.
\par }{\Q I will pour
out my rage
on them like a torrential flood!
\par }{\Q \VS{11}Ephraim
will be oppressed,
crushed
under judgment,
\par }{\Q because
he was determined
to pursue
worthless idols.
\par }{\SH The Curse of the Incurable Wound
\par }{\Q \VS{12}I
will be like a moth
to Ephraim,
\par }{\Q like wood rot
to the house
of Judah.
\par }{\Q \VS{13}When Ephraim
saw
his sickness
\par }{\Q and Judah
saw his wound,
\par }{\Q then
Ephraim
turned to
Assyria,
\par }{\Q and begged
its great king
for
help.
\par }{\Q But he
will not
be able
to heal
you!
\par }{\Q He cannot
cure
your wound!
\par }{\SH The Lion Will Carry Israel Off Into Exile
\par }{\Q \VS{14}I
will be like a lion
to Ephraim,
\par }{\Q like a young lion
to the house
of Judah.
\par }{\Q I
myself
will tear
them to pieces,
\par }{\Q then
I will carry
them off, and no
one will be able to rescue them!
\par }{\Q \VS{15}Then
I will return
again to
my lair
\par }{\Q until
they have
suffered their punishment.
\par }{\Q Then
they will seek
me;

\par }{\Q in their distress
they
will earnestly seek me.

\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}“Come
on! Let’s return
to
the {\ND{Lord}}!
\par }{\Q He himself
has torn
us to pieces,
\par }{\Q but he will heal
us!
\par }{\Q He has injured
us,
\par }{\Q but he will bandage our wounds!
\par }{\Q \VS{2}He will restore
us in a very short time;
\par }{\Q he will heal
us in a little
while,

\par }{\Q so that we may live
in his presence.
\par }{\Q \VS{3}So let us acknowledge
him!

\par }{\Q Let us seek to acknowledge
the

{\ND{Lord}}!
\par }{\Q He will come to our rescue as certainly as the appearance of the dawn,
\par }{\Q as certainly
as the winter rain
comes,
\par }{\Q as certainly as the spring rain
that waters
the land.”
\par }{\SH Transitory Faithfulness and Imminent Judgment
\par }{\Q \VS{4}What
am I going to do
with you, O Ephraim?
\par }{\Q What
am I going to do
with you, O Judah?
\par }{\Q For your faithfulness
is as fleeting as the morning
mist;
\par }{\Q it disappears
as quickly
as dawn’s
dew!
\par }{\Q \VS{5}Therefore,
I will certainly cut you into pieces
at the hands of the prophets;
\par }{\Q I will certainly kill
you in fulfillment
of my oracles
of judgment;
\par }{\Q for my judgment will come forth
like the light of the dawn.
\par }{\Q \VS{6}For
I delight
in faithfulness,
not
simply in sacrifice;
\par }{\Q I delight in acknowledging
God,
not simply in whole burnt offerings.
\par }{\SH Indictments Against the Cities of Israel and Judah
\par }{\Q \VS{7}At Adam
they
broke
the covenant;
\par }{\Q Oh how they were unfaithful to me!
\par }{\Q \VS{8}Gilead
is a city
full of evildoers;
\par }{\Q its streets are stained
with bloody footprints!
\par }{\Q \VS{9}The company
of priests
is like a gang
of robbers,
\par }{\Q lying
in ambush
to pounce on
a victim.
\par }{\Q They commit murder
on
the road
to Shechem;
\par }{\Q they have
done
heinous crimes!
\par }{\Q \VS{10}I have seen
a disgusting thing
in the temple
of Israel:
\par }{\Q there
Ephraim
practices temple prostitution
\par }{\Q and Judah
defiles itself.
\par }{\Q \VS{11}I have appointed
a time
to reap judgment for you also,
O Judah!
\par }{\SH If Israel Would Repent of Sin, God Would Relent of Judgment
\par }{\Q Whenever I want to restore the fortunes of my people,

\par }\Chap{7}{\PP \par }{\Q \VerseOne{1}whenever I want to heal
Israel,
\par }{\Q the sin
of Ephraim
is revealed,
\par }{\Q and the evil
deeds of Samaria
are exposed.
\par }{\Q For
they do
what is wrong;
\par }{\Q thieves
break into
houses,
\par }{\Q and gangs
rob people out in the streets.
\par }{\Q \VS{2}They do not realize
\par }{\Q that I remember
all
of their wicked deeds.
\par }{\Q Their evil
deeds have now
surrounded
them;
\par }{\Q their sinful
deeds
are always before me.
\par }{\SH Political Intrigue and Conspiracy in the Palace
\par }{\Q \VS{3}The royal advisers delight the king
with their evil
schemes,
\par }{\Q the princes
make him glad with their lies.
\par }{\Q \VS{4}They are all
like bakers,
\par }{\Q they are like
a smoldering
oven;
\par }{\Q they are like a baker
who does not
stoke
the fire
\par }{\Q until
the kneaded
dough
is ready for baking.
\par }{\Q \VS{5}At the celebration
of their king,
\par }{\Q his princes
become inflamed
with wine;
\par }{\Q they conspire
with
evildoers.
\par }{\Q \VS{6}They approach
him, all the while plotting
against him.
\par }{\Q Their hearts
are like an oven;
\par }{\Q their anger
smolders
all
night
long,
\par }{\Q but in the morning
it bursts
into a flaming
fire.
\par }{\Q \VS{7}All
of them are blazing
like an oven;
\par }{\Q they devour
their rulers.
\par }{\Q All
of their kings
fall
–
\par }{\Q and none
of them call
on me!
\par }{\SH Israel Lacks Discernment and Refuses to Repent
\par }{\Q \VS{8}Ephraim
has mixed
itself
like flour among the nations;
\par }{\Q Ephraim
is like a ruined cake
of bread that is scorched
on one side.
\par }{\Q \VS{9}Foreigners
are consuming
what his strenuous
labor produced,

\par }{\Q but he does not
recognize
it!
\par }{\Q His head is filled
with gray hair,
\par }{\Q but he does not
realize it!
\par }{\Q \VS{10}The arrogance
of Israel
testifies
against him,
\par }{\Q yet they refuse
to return
to
the {\ND{Lord}}
their God!
\par }{\Q In spite of all
this
they refuse
to seek him!
\par }{\SH Israel Turns to Assyria and Egypt for Help
\par }{\Q \VS{11}Ephraim
has been like a dove,
\par }{\Q easily deceived
and lacking
discernment.
\par }{\Q They called
to Egypt
for help;
\par }{\Q they turned to Assyria
for protection.
\par }{\Q \VS{12}I will throw
my bird net
over
them while they are flying,
\par }{\Q I will bring
them down
like birds
in the sky;
\par }{\Q I will discipline
them when I hear
them flocking together.
\par }{\SH Israel Has Turned Away from the Lord
\par }{\Q \VS{13}Woe
to them! For
they have fled
from
me!
\par }{\Q Destruction
to them! For
they have rebelled
against me!
\par }{\Q I
want to deliver
them,
\par }{\Q but they
have
lied to me.
\par }{\Q \VS{14}They do not
pray
to
me,
\par }{\Q but howl
in distress on
their beds;
\par }{\Q They slash
themselves for
grain
and new wine,
\par }{\Q but turn away from me.
\par }{\Q \VS{15}Although
I
trained
and strengthened
them,
\par }{\Q they plot
evil against me!
\par }{\Q \VS{16}They turn
to Baal;
\par }{\Q they are
like an unreliable
bow.
\par }{\Q Their leaders
will fall
by the sword
\par }{\Q because their prayers to Baal
have made me angry.
\par }{\Q So
people will disdain
them in the land
of Egypt.

\par }\Chap{8}{\PP \par }{\Q \VerseOne{1}Sound
the alarm!

\par }{\Q An eagle
looms over
the temple
of the {\ND{Lord}}!
\par }{\Q For they have
broken
their covenant
with me,

\par }{\Q and have rebelled
against my law.
\par }{\Q \VS{2}Israel
cries
out to me,
\par }{\Q “My God,
we acknowledge you!”
\par }{\Q \VS{3}But Israel
has rejected
what is morally good;
\par }{\Q so an enemy
will pursue him.
\par }{\SH The Political and Cultic Sin of Israel
\par }{\Q \VS{4}They
enthroned kings
without
my consent!

\par }{\Q They appointed princes without
my approval!

\par }{\Q They made
idols
out of their silver
and gold,
\par }{\Q but they will
be destroyed!
\par }{\Q \VS{5}O Samaria,
he has rejected
your calf
idol!
\par }{\Q My anger
burns
against them!
\par }{\Q They will not
survive much longer
without being punished,
\par }{\Q even though they are Israelites!
\par }{\Q \VS{6}That
idol
was made
by a workman
– it is not
God!
\par }{\Q The calf
idol of Samaria
will be
broken to bits.
\par }{\SH The Fertility Cultists Will Become Infertile
\par }{\Q \VS{7}They sow
the wind,
\par }{\Q and so they will reap
the whirlwind!
\par }{\Q The stalk
does not have
any standing grain;
\par }{\Q it will not
produce
any flour.
\par }{\Q Even if
it were to yield
grain,
\par }{\Q foreigners
would swallow it all up.
\par }{\Q \VS{8}Israel
will be swallowed
up among the nations;
\par }{\Q they will be like a worthless piece of pottery.
\par }{\SH The Willful Donkey and the Wanton Harlot
\par }{\Q \VS{9}They
have gone up
to Assyria,
\par }{\Q like a wild donkey
that wanders off.
\par }{\Q Ephraim
has hired
prostitutes as lovers.
\par }{\Q \VS{10}Even
though
they have hired
lovers among the nations,
\par }{\Q I will soon
gather
them together for judgment.

\par }{\Q Then they will begin
to waste away
\par }{\Q under the oppression
of a mighty king.
\par }{\SH Sacrifices Ineffective without Moral Obedience
\par }{\Q \VS{11}Although
Ephraim
has built many
altars
for sin
offerings,
\par }{\Q these have become
altars
for sinning!
\par }{\Q \VS{12}I spelled out
my law
for him in great detail,
\par }{\Q but they regard
it as something
totally unknown to them!
\par }{\Q \VS{13}They offer up sacrificial
gifts
to me,
\par }{\Q and eat
the meat,
\par }{\Q but the
{\ND{Lord}}
does not
accept
their sacrifices.

\par }{\Q Soon
he will remember
their wrongdoing,
\par }{\Q he will punish
their sins,
\par }{\Q and they
will return
to Egypt.
\par }{\Q \VS{14}Israel
has forgotten
his Maker
and built
royal palaces,
\par }{\Q and Judah
has built many
fortified
cities.
\par }{\Q But I will send
fire
on their cities;
\par }{\Q it will consume
their royal citadels.

\par }\Chap{9}{\PP \par }{\Q \VerseOne{1}O Israel,
do not
rejoice
jubilantly
like the nations,
\par }{\Q for
you are unfaithful
to
your God.
\par }{\Q You love
to receive a prostitute's wages
\par }{\Q on
all
the floors where you thresh
your grain.
\par }{\Q \VS{2}Threshing floors
and wine vats
will not
feed
the people,

\par }{\Q and new wine
only deceives them.
\par }{\SH Assyrian Exile Will Reverse the Egyptian Exodus
\par }{\Q \VS{3}They will not
remain
in the
{\ND{Lord}}’s
land.
\par }{\Q Ephraim
will return
to Egypt;
\par }{\Q they will eat
ritually unclean
food in Assyria.
\par }{\Q \VS{4}They will not
pour
out drink offerings of wine
to the
{\ND{Lord}};
\par }{\Q they will not
please
him with their sacrifices.
\par }{\Q Their sacrifices will be like bread
eaten while in mourning;
\par }{\Q all
those who eat
them will make themselves ritually unclean.
\par }{\Q For
their bread
will be only to satisfy their appetite;
\par }{\Q it will not
come
into the temple
of the {\ND{Lord}}.
\par }{\Q \VS{5}So what
will you do
on
the festival
day,
\par }{\Q on the festival
days
of the {\ND{Lord}}?
\par }{\SH No Escape for the Israelites This Time!
\par }{\Q \VS{6}Look! Even if
they flee
from the destruction,
\par }{\Q Egypt
will take hold of them,
\par }{\Q and Memphis
will bury
them.
\par }{\Q The weeds
will inherit the silver they treasure –
\par }{\Q thorn bushes will occupy their homes.
\par }{\Q \VS{7}The time
of judgment
is about to arrive!

\par }{\Q The time
of retribution
is imminent!

\par }{\Q Let Israel
know!
\par }{\SH Israel Rejects Hosea’s Prophetic Exhortations
\par }{\Q The prophet is considered a fool –
\par }{\Q the inspired man is viewed as a madman –
\par }{\Q because of the multitude of your sins
\par }{\Q and your intense animosity.
\par }{\Q \VS{8}The prophet
is a watchman
over
Ephraim
on
behalf
of God,
\par }{\Q yet traps
are laid for him along all
of his paths;
\par }{\Q animosity
rages against him
in the land of his God.
\par }{\SH The Best of Times, the Worst of Times
\par }{\Q \VS{9}They have sunk deep
into corruption
\par }{\Q as in the days
of Gibeah.
\par }{\Q He will remember
their wrongdoing.
\par }{\Q He will repay them
for their sins.
\par }{\Q \VS{10}When I found
Israel,
it was like finding
grapes
in the wilderness.
\par }{\Q I viewed
your ancestors
like an early fig on a fig tree
in its first season.
\par }{\Q Then they
came
to Baal-Peor
and they dedicated
themselves to shame
–
\par }{\Q they became
as detestable
as what they loved.
\par }{\SH The Fertility Worshipers Will Become Infertile
\par }{\Q \VS{11}Ephraim
will be like a bird;
\par }{\Q what they value
will fly
away.
\par }{\Q They will not bear children –
\par }{\Q they will not enjoy pregnancy –
\par }{\Q they will not even conceive!
\par }{\Q \VS{12}Even
if
they raise
their children,
\par }{\Q I will take away
every last one of them.
\par }{\Q Woe
to them!
\par }{\Q For
I will turn away
from them.
\par }{\Q \VS{13}Just as lion cubs are born predators,

\par }{\Q so Ephraim
will bear
his sons
for slaughter.
\par }{\Q \VS{14}Give
them, O
{\ND{Lord}} –
\par }{\Q what
will you give
them?
\par }{\Q Give
them
wombs
that miscarry,
\par }{\Q and breasts
that cannot nurse!
\par }{\Q \VS{15}Because
of all
their evil
in Gilgal,
\par }{\Q I hate
them there.
\par }{\Q On
account of their evil
deeds,
\par }{\Q I will drive
them out of my land.

\par }{\Q I will no
longer
love
them;
\par }{\Q all
their rulers
are rebels.
\par }{\Q \VS{16}Ephraim
will be struck down –
\par }{\Q their root
will be dried
up;
\par }{\Q they will not
yield
any fruit.
\par }{\Q Even if
they do bear children,
\par }{\Q I will kill
their precious
offspring.
\par }{\Q \VS{17}My God
will reject
them,
\par }{\Q for
they have not
obeyed
him;
\par }{\Q so they will be
fugitives
among the nations.

\par }\Chap{10}{\PP \par }{\Q \VerseOne{1}Israel
was a fertile vine
\par }{\Q that yielded
fruit.
\par }{\Q As
his fruit
multiplied,
\par }{\Q he multiplied altars
to Baal.

\par }{\Q As his land
prospered,
\par }{\Q they adorned
the fertility pillars.
\par }{\Q \VS{2}Their heart
is slipping;
\par }{\Q soon they will be punished
for their guilt.
\par }{\Q The
{\ND{Lord}} will break
their altars;
\par }{\Q he will completely destroy
their fertility pillars.
\par }{\SH The Lord Will Punish Israel by Removing Its Kings
\par }{\Q \VS{3}Very soon
they will say,
“We have no
king
\par }{\Q since
we did not
fear
the {\ND{Lord}}.
\par }{\Q But what can
a king
do
for us anyway?”
\par }{\Q \VS{4}They utter
empty words,
\par }{\Q taking false
oaths
and making
empty agreements.
\par }{\Q Therefore legal disputes
sprout up
\par }{\Q like poisonous weeds
in
the furrows
of a plowed field.
\par }{\SH The Calf Idol and Idolaters of Samaria Will Be Exiled
\par }{\Q \VS{5}The inhabitants
of Samaria
will lament
over the calf
idol of Beth Aven.
\par }{\Q Its people
will mourn
over
it;
\par }{\Q its idolatrous priests
will wail
over
it,

\par }{\Q because
its splendor
will be taken from
them into exile.
\par }{\Q \VS{6}Even
the calf idol will be carried
to Assyria,
\par }{\Q as tribute
for the great king.
\par }{\Q Ephraim
will be disgraced;
\par }{\Q Israel
will be put
to shame
because of its wooden idol.
\par }{\Q \VS{7}Samaria
and its
king
will be carried
off
\par }{\Q like
a twig
on
the surface
of the waters.
\par }{\Q \VS{8}The high places
of the “House of Wickedness”
will be destroyed;
\par }{\Q it is the place where Israel
sins.
\par }{\Q Thorns
and thistles
will grow up
over
its altars.
\par }{\Q Then they will say
to the mountains,
“Cover
us!”
\par }{\Q and to the hills,
“Fall on us!”
\par }{\SH Failure to Learn from the Sin and Judgment of Gibeah
\par }{\Q \VS{9}O Israel,
you have sinned
since the time
of Gibeah,
\par }{\Q and there
you have remained.
\par }{\Q Did not
war
overtake
the evildoers
in Gibeah?
\par }{\Q \VS{10}When I please,
I will discipline
them;

\par }{\Q I will gather
nations
together to attack them,

\par }{\Q to bind
them in chains for their two sins.
\par }{\SH Fertility Imagery: Plowing, Sowing, and Reaping
\par }{\Q \VS{11}Ephraim
was a well-trained
heifer
who loved
to thresh
grain;
\par }{\Q I myself
put a fine yoke on her neck.
\par }{\Q I will harness Ephraim.
\par }{\Q Let Judah plow!
\par }{\Q Let Jacob break up the unplowed ground for himself!
\par }{\Q \VS{12}Sow
righteousness
for yourselves,
\par }{\Q reap
unfailing love.
\par }{\Q Break up
the unplowed ground
for yourselves,
\par }{\Q for it is time
to seek
the
{\ND{Lord}},
\par }{\Q until
he comes
and showers
deliverance on you.
\par }{\Q \VS{13}But
you have plowed
wickedness;
\par }{\Q you have reaped
injustice;
\par }{\Q you have eaten
the fruit
of deception.
\par }{\Q Because
you have depended on
your chariots;

\par }{\Q you have relied on
your many
warriors.
\par }{\SH Bethel Will Be Destroyed Like Beth Arbel
\par }{\Q \VS{14}The roar
of battle will rise against
your people;
\par }{\Q all
your fortresses
will be devastated,
\par }{\Q just as Shalman
devastated
Beth Arbel
on the day
of battle,
\par }{\Q when mothers
were dashed to the ground
with their children.
\par }{\Q \VS{15}So
will it happen
to you, O Bethel,
\par }{\Q because
of your great wickedness!
\par }{\Q When that day dawns,
\par }{\Q the king
of Israel
will be destroyed.

\par }\Chap{11}{\PP \par }{\Q \VerseOne{1}When
Israel
was a young man,
I loved
him like a son,

\par }{\Q and I summoned
my son
out of Egypt.
\par }{\Q \VS{2}But the more I summoned
them,
\par }{\Q the farther
they departed
from me.
\par }{\Q They sacrificed
to the Baal idols
\par }{\Q and burned incense
to images.
\par }{\Q \VS{3}Yet it was I
who led
Ephraim,
\par }{\Q I took
them by
the arm;
\par }{\Q but they did not
acknowledge
\par }{\Q that
I had
healed them.
\par }{\Q \VS{4}I led
them with leather
cords,
\par }{\Q with leather
ropes;
\par }{\Q I lifted
the yoke
from
their neck,
\par }{\Q and gently
fed them.
\par }{\Q \VS{5}They will return
to
Egypt!

\par }{\Q Assyria
will rule over
them

\par }{\Q because
they refuse
to repent!
\par }{\Q \VS{6}A sword
will flash
in their cities,
\par }{\Q it will destroy
the bars of
their city gates,
\par }{\Q and will devour
them in their fortresses.
\par }{\Q \VS{7}My people
are obsessed
with turning away
from me;
\par }{\Q they call
to Baal,
but he will never
exalt them!
\par }{\SH The Divine Dilemma: Judgment or Mercy?
\par }{\Q \VS{8}How
can I give
you up, O Ephraim?
\par }{\Q How can I surrender
you, O Israel?
\par }{\Q How
can I treat
you like Admah?
\par }{\Q How can I make
you like Zeboiim?
\par }{\Q I have had a change of heart!

\par }{\Q All
my tender compassions
are aroused!
\par }{\Q \VS{9}I cannot
carry
out my fierce
anger!
\par }{\Q I cannot
totally
destroy
Ephraim!
\par }{\Q Because
I am
God,
and not
man
– the Holy One
among you –
\par }{\Q I will not come in wrath!
\par }{\SH God Will Restore the Exiles to Israel
\par }{\Q \VS{10}He will roar
like a lion,
\par }{\Q and they will follow
the {\ND{Lord}};
\par }{\Q when
he
roars,
\par }{\Q his children
will come trembling
from the west.
\par }{\Q \VS{11}They will return in fear and trembling
\par }{\Q like birds
from Egypt,
\par }{\Q like doves
from Assyria,
\par }{\Q and I will settle
them in their homes,”
declares
the {\ND{Lord}}.
\par }{\SH God’s Lawsuit against Israel: Breach of Covenant
\par }{\Q \VS{12} Ephraim
has surrounded
me with lies;
\par }{\Q the house
of Israel
has surrounded me with deceit.
\par }{\Q But Judah
still
roams
about with
God;
\par }{\Q he remains faithful
to the Holy One.

\par }\Chap{12}{\PP \par }{\Q \VerseOne{1}Ephraim
continually feeds
on the wind;
\par }{\Q he chases
the east wind
all
day;
\par }{\Q he multiplies
lies
and violence.
\par }{\Q They make treaties
with
Assyria,
\par }{\Q and send
olive oil
as tribute
to Egypt.
\par }{\Q \VS{2}The
{\ND{Lord}}
also has a covenant lawsuit
against
Judah;
\par }{\Q he will punish
Jacob
according to his ways
\par }{\Q and repay
him according to his deeds.
\par }{\SH Israel Must Return to the God of Jacob
\par }{\Q \VS{3}In the womb
he attacked
his brother;
\par }{\Q in his manly vigor
he struggled
with
God.
\par }{\Q \VS{4}He struggled
with an angel
and prevailed;
\par }{\Q he wept
and begged
for his favor.
\par }{\Q He found
God at Bethel,
\par }{\Q and there
he spoke
with him!
\par }{\Q \VS{5}As for the
{\ND{Lord}}
God
Almighty,
\par }{\Q the {\ND{Lord}}
is the name by which he is remembered!
\par }{\Q \VS{6}But you
must return
to
your God,
\par }{\Q by maintaining
love
and justice,
\par }{\Q and by waiting
for your God to return to you.
\par }{\SH The Lord Refutes Israel’s False Claim of Innocence
\par }{\Q \VS{7}The businessmen
love
to cheat;
\par }{\Q they use
dishonest
scales.
\par }{\Q \VS{8}Ephraim
boasts, “I am very
rich!
\par }{\Q I have become
wealthy!

\par }{\Q In all
that I have done
to gain my wealth,

\par }{\Q no
one can accuse
me of any offense
that
is actually sinful.”
\par }{\Q \VS{9}“I am
the {\ND{Lord}}
your God
who brought you out of Egypt;
\par }{\Q I will make you live
in tents
again
as in the days
of old.
\par }{\Q \VS{10}I spoke
to
the prophets;
\par }{\Q I
myself revealed many
visions;
\par }{\Q I spoke in parables
through
the prophets.”
\par }{\Q \VS{11}Is there idolatry
in Gilead?

\par }{\Q Certainly
its inhabitants will come to nothing!

\par }{\Q Do they sacrifice
bulls
in Gilgal?
\par }{\Q Surely
their altars
will be like stones heaped
up on
a plowed
field!
\par }{\SH Jacob in Aram, Israel in Egypt, and Ephraim in Trouble
\par }{\Q \VS{12}Jacob
fled
to the country
of Aram,
\par }{\Q then Israel
worked
to acquire
a wife;
\par }{\Q he tended
sheep to pay for her.
\par }{\Q \VS{13}The
{\ND{Lord}}
brought
Israel
out of Egypt
by a prophet,
\par }{\Q and due to a prophet
Israel was preserved alive.
\par }{\Q \VS{14}But Ephraim
bitterly
provoked
him to anger;
\par }{\Q so he will hold him accountable for the blood
he has
shed,

\par }{\Q his Lord
will repay
him for the contempt
he has shown.

\par }\Chap{13}{\PP \par }{\Q \VerseOne{1}When Ephraim
spoke,
there was terror;
\par }{\Q he
was exalted
in Israel,
\par }{\Q but he became guilty
by worshiping Baal
and died.
\par }{\Q \VS{2}Even now
they persist in sin!

\par }{\Q They make
metal images
for themselves,
\par }{\Q idols
that they skillfully
fashion from their own silver;
\par }{\Q all
of them are nothing but the work
of craftsmen!
\par }{\Q There is a saying
about them:

\par }{\Q “Those who sacrifice
to the calf
idol are calf kissers!”
\par }{\Q \VS{3}Therefore
they will disappear
like the morning mist,
\par }{\Q like early
morning
dew
that evaporates,
\par }{\Q like chaff
that is blown away
from a threshing floor,
\par }{\Q like smoke
that disappears through an open window.
\par }{\SH Well-Fed Israel Will Be Fed to Wild Animals
\par }{\Q \VS{4}But I am
the {\ND{Lord}}
your God,
\par }{\Q who brought you out of Egypt.
\par }{\Q Therefore, you must not
acknowledge
any God
but me;
\par }{\Q except
me there is no
Savior.
\par }{\Q \VS{5}I
cared
for you in the wilderness,
\par }{\Q in the dry desert
where no water was.
\par }{\Q \VS{6}When they were fed,
they became satisfied;
\par }{\Q when they were satisfied,
they became proud;
\par }{\Q as a result,
they forgot me!
\par }{\Q \VS{7}So I will pounce on them like
a lion;
\par }{\Q like a leopard
I will lurk
by the path.
\par }{\Q \VS{8}I will attack them
like a bear
robbed
of her cubs
–
\par }{\Q I will rip
open their chests.
\par }{\Q I will devour
them there
like a lion
–
\par }{\Q like a wild
animal
would tear them apart.
\par }{\SH Israel’s King Unable to Deliver the Nation
\par }{\Q \VS{9}I will destroy
you, O Israel!
\par }{\Q Who is there to help you?
\par }{\Q \VS{10}Where
then is
your king,
\par }{\Q that
he may save
you in all
your cities?
\par }{\Q Where are your rulers
for whom
you asked, saying,
\par }{\Q “Give
me a king
and princes”?
\par }{\Q \VS{11}I granted
you a king
in my anger,
\par }{\Q and I will take
him away in my wrath!
\par }{\SH Israel’s Punishment Will Not Be Withheld Much Longer
\par }{\Q \VS{12}The punishment
of Ephraim
has been decreed;

\par }{\Q his punishment
is being stored up for the future.
\par }{\Q \VS{13}The labor
pains
of a woman will overtake
him,
\par }{\Q but the baby
will lack
wisdom;
\par }{\Q when
the time
arrives,
\par }{\Q he will not
come out
of the womb!
\par }{\SH The Lord Will Not Relent from the Threatened Judgment
\par }{\Q \VS{14}Will I deliver
them from the power
of Sheol? No, I will not!

\par }{\Q Will
I redeem
them from death? No, I will
not!
\par }{\Q O Death,
bring on your plagues!

\par }{\Q O Sheol,
bring on your destruction!

\par }{\Q My eyes
will not show
any compassion!
\par }{\SH The Capital of the Northern Empire Will Be Destroyed
\par }{\Q \VS{15}Even though
he flourishes
like a reed plant,

\par }{\Q a scorching east wind
will come,
\par }{\Q a wind
from the
{\ND{Lord}}
rising up
from the desert.
\par }{\Q As a result, his spring
will dry up;
\par }{\Q his well
will become dry.
\par }{\Q That
wind will spoil
all
his delightful
foods
\par }{\Q in the containers
in his storehouse.
\par }{\Q \VS{16}Samaria
will be held guilty,
\par }{\Q because
she rebelled
against her God.
\par }{\Q They will fall
by the sword,
\par }{\Q their infants
will be dashed to the ground –
\par }{\Q their pregnant women will be ripped open.

\par }\Chap{14}{\PP \par }{\Q \VerseOne{1}Return,
O Israel,
to the
{\ND{Lord}}
your God,
\par }{\Q for
your sin
has been your downfall!
\par }{\Q \VS{2}Return
to
the {\ND{Lord}}
and repent!

\par }{\Q Say
to him: “Completely
forgive
our iniquity;
\par }{\Q accept
our penitential
prayer,

\par }{\Q that we may offer the praise of our lips
as sacrificial bulls.
\par }{\Q \VS{3}Assyria
cannot
save
us;
\par }{\Q we will not
ride
warhorses.
\par }{\Q We will never
again
say,
‘Our gods’
\par }{\Q to what our own hands
have made.
\par }{\Q For
only you will show compassion
to Orphan Israel!”
\par }{\SH Divine Promise to Relent from Judgment and to Restore Blessings
\par }{\Q \VS{4}“I will heal
their waywardness
\par }{\Q and love
them freely,
\par }{\Q for
my anger
will turn
away from them.
\par }{\Q \VS{5}I will be
like the dew
to Israel;
\par }{\Q he will blossom
like a lily,
\par }{\Q he will send down
his roots
like a cedar of Lebanon.
\par }{\Q \VS{6}His young shoots
will grow;
\par }{\Q his splendor
will be like
an olive
tree,
\par }{\Q his fragrance
like a cedar of Lebanon.
\par }{\Q \VS{7}People
will reside
again
in
his shade;
\par }{\Q they will plant
and harvest grain
in abundance.

\par }{\Q They will blossom
like a vine,
\par }{\Q and his fame
will be like the wine
from Lebanon.
\par }{\Q \VS{8}O Ephraim,
I
do
not want
to have anything to do with idols
anymore!
\par }{\Q I
will answer
him and care
for him.
\par }{\Q I am
like
a luxuriant
cypress tree;

\par }{\Q your fruitfulness comes from me!
\par }{\SH Concluding Exhortation
\par }{\Q \VS{9}Who
is wise?
\par }{\Q Let him discern
these
things!
\par }{\Q Who is discerning?
\par }{\Q Let him understand
them!
\par }{\Q For
the ways
of the
{\ND{Lord}}
are right;
\par }{\Q the godly
walk
in them,
\par }{\PP but in them the rebellious
stumble.
\par }