\NormalFont\ShortTitle{Isaiah}
{\MT Isaiah

\par }\ChapOne{1}{\SH Heading
\par }{\PP \VerseOne{1}Here is the message
about Judah
and Jerusalem
that
was revealed
to Isaiah
son
of Amoz
during the time
when Uzziah,
Jotham,
Ahaz,
and Hezekiah
reigned over
Judah.
\par }{\SH Obedience, not Sacrifice
\par }{\Q \VS{2}Listen,
O heavens,
\par }{\Q pay attention,
O earth!

\par }{\Q For
the {\ND{Lord}}
speaks:
\par }{\Q “I raised
children,
I brought
them up,
\par }{\Q but they
have rebelled against me!
\par }{\Q \VS{3}An ox
recognizes
its owner,
\par }{\Q a donkey
recognizes where its owner
puts its food;
\par }{\Q but Israel
does not
recognize
me,

\par }{\Q my people
do not
understand.”
\VS{4}\par }{\Q The sinful
nation
is as good
as dead,

\par }{\Q the people
weighed down
by evil deeds.
\par }{\Q They are offspring
who do wrong,
\par }{\Q children
who do wicked
things.
\par }{\Q They have abandoned
the

{\ND{Lord}},
\par }{\Q and rejected
the
Holy
One of Israel.
\par }{\Q They are alienated from him.
\VS{5}\par }{\Q Why
do you insist on
being battered?
\par }{\Q Why do you continue
to rebel?

\par }{\Q Your head
has a massive wound,
\par }{\Q your whole
body
is weak.
\par }{\Q \VS{6}From the soles
of your feet
to your head,
\par }{\Q there is no
spot that is unharmed.
\par }{\Q There are only bruises,
cuts,
\par }{\Q and open
wounds.
\par }{\Q They have not
been cleansed
or
bandaged,
\par }{\Q nor
have they been treated
with olive oil.
\par }{\Q \VS{7}Your land
is devastated,
\par }{\Q your cities
burned
with fire.
\par }{\Q Right before
your eyes your crops
\par }{\Q are being destroyed
by foreign invaders.
\par }{\Q They leave behind
devastation
and destruction.
\par }{\Q \VS{8}Daughter
Zion
is left isolated,
\par }{\Q like a hut
in a vineyard,
\par }{\Q or a shelter
in a cucumber field;
\par }{\Q she is a besieged
city.
\par }{\Q \VS{9}If
the {\ND{Lord}}
who commands armies
had not left
us a few
survivors,
\par }{\Q we would have quickly
become like Sodom,
\par }{\Q we
would have become
like Gomorrah.
\par }{\Q \VS{10}Listen
to the
{\ND{Lord}}’s
word,
\par }{\Q you leaders
of Sodom!

\par }{\Q Pay attention
to our God’s
rebuke,
\par }{\Q people
of Gomorrah!
\par }{\Q \VS{11}“Of what importance
to me are your many
sacrifices?”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q “I am stuffed
with burnt sacrifices
\par }{\Q of rams
and the fat
from steers.
\par }{\Q The blood
of bulls,
lambs,
and goats
\par }{\Q I do not
want.
\par }{\Q \VS{12}When
you enter
my presence,
\par }{\Q do you actually
think I want
this
–
\par }{\Q animals trampling
on my courtyards?
\par }{\Q \VS{13}Do not
bring
any more
meaningless offerings;
\par }{\Q I consider
your incense
detestable!

\par }{\Q You observe new moon
festivals, Sabbaths,
and convocations,
\par }{\Q but I cannot
tolerate sin-stained
celebrations!
\par }{\Q \VS{14}I hate
your new moon
festivals
and assemblies;
\par }{\Q they are a burden
\par }{\Q that I am tired
of carrying.
\par }{\Q \VS{15}When you spread
out your hands
in prayer,
\par }{\Q I look
the other way;
\par }{\Q when
you offer your many
prayers,
\par }{\Q I do
not
listen,
\par }{\Q because
your hands
are covered
with blood.
\VS{16}\par }{\Q Wash! Cleanse
yourselves!
\par }{\Q Remove
your sinful
deeds
\par }{\Q from my sight.
\par }{\Q Stop
sinning!
\par }{\Q \VS{17}Learn
to do what is right!
\par }{\Q Promote
justice!
\par }{\Q Give
the oppressed
reason to celebrate!

\par }{\Q Take up the cause
of the orphan!
\par }{\Q Defend
the rights of the widow!
\VS{18}\par }{\Q Come,
let’s
consider
your options,” says
the {\ND{Lord}}.
\par }{\Q “Though
your sins
have stained you like the color red,
\par }{\Q you can become white
like snow;
\par }{\Q though
they are as easy to see as the color
scarlet,
\par }{\Q you can become
white like wool.
\par }{\Q \VS{19}If
you have a willing
attitude and obey,
\par }{\Q then you will again eat
the good crops
of the land.
\par }{\Q \VS{20}But if
you refuse
and rebel,
\par }{\Q you will be devoured
by the sword.”
\par }{\Q Know for
certain
that the
{\ND{Lord}}
has spoken.
\par }{\SH Purifying Judgment
\par }{\Q \VS{21}How
tragic that the once-faithful
city
\par }{\Q has become
a prostitute!

\par }{\Q She was once a center
of justice,
\par }{\Q fairness
resided
in her,
\par }{\Q but now
only murderers.
\par }{\Q \VS{22}Your silver
has become
scum,
\par }{\Q your beer
is diluted
with water.
\par }{\Q \VS{23}Your officials
are rebels,
\par }{\Q they associate
with thieves.
\par }{\Q All
of them love
bribery,
\par }{\Q and look for
payoffs.
\par }{\Q They do not
take up
the cause
of the orphan,
\par }{\Q or
defend
the rights of
the widow.
\par }{\Q \VS{24}Therefore,
the sovereign

{\ND{Lord}}
who commands armies,
\par }{\Q the powerful ruler
of Israel,
says
this:
\par }{\Q “Ah,
I will seek vengeance
against my adversaries,
\par }{\Q I will take revenge
against my enemies.
\par }{\Q \VS{25}I will attack
you;

\par }{\Q I will purify
your metal
with flux.
\par }{\Q I will remove
all
your slag.
\par }{\Q \VS{26}I will reestablish
honest judges
as in former
times,
\par }{\Q wise advisers
as in earlier
days.

\par }{\Q Then
you will be called,
‘The Just
City,
\par }{\Q Faithful
Town.’ ”
\VS{27}\par }{\Q Zion
will be freed by justice,
\par }{\Q and her returnees
by righteousness.
\par }{\Q \VS{28}All
rebellious
sinners
will be shattered,
\par }{\Q those who abandon
the {\ND{Lord}}
will perish.
\par }{\Q \VS{29}Indeed,
they will be ashamed
of the sacred trees
\par }{\Q you find so desirable;
\par }{\Q you will be embarrassed
because of the sacred orchards
\par }{\Q where
you choose to worship.
\par }{\Q \VS{30}For
you will be
like a tree
whose leaves
wither,
\par }{\Q like an orchard that is unwatered.
\par }{\Q \VS{31}The powerful
will be
like a thread
of yarn,
\par }{\Q their deeds
like a spark;
\par }{\Q both
will burn
together,
\par }{\Q and no
one will put out the fire.

\par }\Chap{2}{\PP \VerseOne{1}Here is the message
about Judah
and Jerusalem
that
was revealed
to Isaiah
son
of Amoz.
\par }{\Q \VS{2}In the future
\par }{\Q the mountain
of the
{\ND{Lord}}’s
temple
will endure
\par }{\Q as the most important
of mountains,
\par }{\Q and will be the most prominent
of hills.
\par }{\Q All
the nations
will stream
to it,
\par }{\Q \VS{3}many
peoples
will come
and say,
\par }{\Q “Come, let
us go
up
to
the
{\ND{Lord}}’s
mountain,
\par }{\Q to
the temple
of the God
of Jacob,
\par }{\Q so he can teach
us his requirements,
\par }{\Q and we can follow
his standards.”
\par }{\Q For
Zion
will be the center for moral instruction;
\par }{\Q the
{\ND{Lord}}
will issue
edicts
from
Jerusalem.
\par }{\Q \VS{4}He will judge
disputes between
nations;
\par }{\Q he will settle
cases for many
peoples.
\par }{\Q They will beat
their swords
into plowshares,
\par }{\Q and their spears
into pruning hooks.
\par }{\Q Nations
will not
take up
the sword
against
other nations,
\par }{\Q and they will no
longer
train
for war.
\par }{\Q \VS{5}O descendants
of Jacob,
\par }{\Q come,
let us walk
in the
{\ND{Lord}}’s
guiding light.
\par }{\SH The Lord’s Day of Judgment
\par }{\Q \VS{6}Indeed,
O
{\ND{Lord}}, you have abandoned
your people,
\par }{\Q the descendants
of Jacob.
\par }{\Q For
diviners from the east
are everywhere;
\par }{\Q they consult omen readers
like the Philistines
do.

\par }{\Q Plenty
of foreigners
are around.
\par }{\Q \VS{7}Their land
is full
of gold
and silver;
\par }{\Q there is no
end
to their wealth.
\par }{\Q Their land
is full
of horses;
\par }{\Q there is no
end
to their chariots.
\par }{\Q \VS{8}Their land
is full
of worthless idols;
\par }{\Q they worship
the product
of their own hands,
\par }{\Q what their own fingers
have
fashioned.
\par }{\Q \VS{9}Men
bow down
to them in homage,
\par }{\Q they lie flat on the ground in worship.
\par }{\Q Don’t
spare them!
\par }{\Q \VS{10}Go
up into the rocky
cliffs,
\par }{\Q hide
in the ground.
\par }{\Q Get away
from the dreadful
judgment of the
{\ND{Lord}},

\par }{\Q from his royal
splendor!
\par }{\Q \VS{11}Proud
men
will be brought low,
\par }{\Q arrogant
men
will be humiliated;
\par }{\Q the {\ND{Lord}}
alone
will be exalted
\par }{\Q in that day.
\par }{\Q \VS{12}Indeed,
the {\ND{Lord}}
who commands armies
has planned a day
of judgment,

\par }{\Q for all
the high
and mighty,
\par }{\Q for all
who are proud
– they will be humiliated;
\par }{\Q \VS{13}for all
the cedars
of Lebanon,
\par }{\Q that are so high
and mighty,
\par }{\Q for all
the oaks
of Bashan;
\par }{\Q \VS{14}for all
the tall mountains,
\par }{\Q for all
the high
hills,
\par }{\Q \VS{15}for every
high
tower,
\par }{\Q for every
fortified
wall,
\par }{\Q \VS{16}for
all
the large ships,
\par }{\Q for
all
the impressive
ships.
\par }{\Q \VS{17}Proud
men
will be humiliated,
\par }{\Q arrogant
men
will be brought low;
\par }{\Q the {\ND{Lord}}
alone
will be exalted
\par }{\Q in that day.
\par }{\Q \VS{18}The worthless idols
will be completely
eliminated.
\par }{\Q \VS{19}They will go
into caves
in the rocky
cliffs
\par }{\Q and into holes
in the ground,
\par }{\Q trying to escape the dreadful
judgment of the
{\ND{Lord}}
\par }{\Q and his royal
splendor,
\par }{\Q when he rises
up to terrify
the earth.
\par }{\Q \VS{20}At that
time
men
will throw
\par }{\Q their silver
and gold
idols,
\par }{\Q which
they made
for themselves to worship,
\par }{\Q into the caves where rodents
and bats live,
\par }{\Q \VS{21}so they themselves can go into
the crevices
of the rocky
cliffs
\par }{\Q and the openings under the rocky
overhangs,

\par }{\Q trying to escape
the dreadful judgment
of the {\ND{Lord}}
\par }{\Q and his royal
splendor,
\par }{\Q when he rises up
to terrify
the earth.
\par }{\Q \VS{22}Stop
trusting
in human
beings,
\par }{\Q whose
life’s breath
is in their nostrils.
\par }{\Q For
why
should they be
given special consideration?

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}Look,
the sovereign

{\ND{Lord}}
who commands armies
\par }{\Q is about to remove
from Jerusalem
and Judah
\par }{\Q every source
of security,
including

\par }{\Q all
the food
and water,
\par }{\Q \VS{2}the mighty
men
and warriors,
\par }{\Q judges
and prophets,
\par }{\Q omen readers
and leaders,
\par }{\Q \VS{3}captains
of groups of fifty,
\par }{\Q the respected
citizens,

\par }{\Q advisers
and those skilled
in magical arts,
\par }{\Q and those who
know incantations.
\par }{\Q \VS{4}The
{\ND{Lord}} says, “I will make
youths
their officials;
\par }{\Q malicious young men
will rule over them.
\par }{\Q \VS{5}The people
will treat
each other
harshly;
\par }{\Q men
will oppose each other;
\par }{\Q neighbors
will fight.

\par }{\Q Youths
will proudly defy
the elderly
\par }{\Q and riffraff
will challenge those who were once respected.
\par }{\Q \VS{6}Indeed,
a man
will grab
his brother
\par }{\Q right in his father’s
house
and say,

\par }{\Q ‘You own a coat
–
\par }{\Q you be
our leader!
\par }{\Q This
heap of ruins
will be under
your control.’
\par }{\Q \VS{7}At
that time
the brother will shout,

\par }{\Q ‘I am no
doctor,
\par }{\Q I have no
food
or coat
in my
house;
\par }{\Q don’t make
me a leader of the people!’ ”
\par }{\Q \VS{8}Jerusalem
certainly
stumbles,
\par }{\Q Judah
falls,
\par }{\Q for
their words
and their actions
offend
the {\ND{Lord}};

\par }{\Q they rebel against his royal authority.
\par }{\Q \VS{9}The look
on their faces
testifies
to their guilt;

\par }{\Q like the people of Sodom
they openly
boast of their sin.
\par }{\Q Too bad for
them!

\par }{\Q For they
bring disaster on themselves.
\par }{\Q \VS{10}Tell
the innocent
it will go well
with them,

\par }{\Q for
they will be rewarded
for what they have done.
\par }{\Q \VS{11}Too bad
for the wicked
sinners!
\par }{\Q For
they will get exactly what they deserve.
\par }{\Q \VS{12}Oppressors
treat my people
cruelly;
\par }{\Q creditors
rule over
them.

\par }{\Q My people’s leaders
mislead
them;
\par }{\Q they give you confusing
directions.
\par }{\Q \VS{13}The
{\ND{Lord}}
takes
his position
to judge;
\par }{\Q he stands
up to pass sentence
on his people.
\par }{\Q \VS{14}The
{\ND{Lord}}
comes
to pronounce judgment
\par }{\Q on the leaders
of his people
and their officials.
\par }{\Q He says, “It is you
who have ruined
the vineyard!

\par }{\Q You have stashed
in your houses
what you have stolen from the poor.
\par }{\Q \VS{15}Why do you crush
my people
\par }{\Q and grind
the faces
of the poor?”

\par }{\Q The sovereign

{\ND{Lord}}
who commands armies has spoken.
\par }{\SH Washing Away Impurity
\par }{\Q \VS{16}The
{\ND{Lord}}
says,
\par }{\Q “The women
of Zion
are proud.
\par }{\Q They walk
with their
heads high

\par }{\Q and flirt
with their
eyes.
\par }{\Q They skip
along

\par }{\Q and the jewelry on their ankles
jingles.
\par }{\Q \VS{17}So the sovereign master
will afflict
the foreheads
of Zion’s
women
with skin diseases,

\par }{\Q the {\ND{Lord}}
will make the front
of their heads
bald.”
\VS{18}\par }{\PP At that
time
the sovereign master
will remove
their beautiful
ankle
jewelry, neck
ornaments,
crescent shaped ornaments,
\VS{19}earrings,
bracelets,
veils,
\VS{20}headdresses,
ankle ornaments,
sashes,
sachets, amulets,
\VS{21}rings,
nose
rings,
\VS{22}festive dresses,
robes,
shawls,
purses,
\VS{23}garments,
vests,
head coverings,
and gowns.
\par }{\Q \VS{24}A putrid stench
will replace
the smell
of spices,
\par }{\Q a rope
will replace
a belt,
\par }{\Q baldness
will replace
braided
locks
of hair,
\par }{\Q a sackcloth
garment
will replace
a fine robe,
\par }{\Q and a prisoner’s brand
will replace
beauty.
\par }{\Q \VS{25}Your men
will fall
by the sword,
\par }{\Q your strong men
will die in battle.
\par }{\Q \VS{26}Her gates
will mourn
and lament;
\par }{\Q deprived
of her people, she will sit
on the ground.

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}Seven
women
will grab hold
of
\par }{\Q one
man
at that time.
\par }{\Q They will say,
“We will provide
our own food,
\par }{\Q we will provide
our own clothes;
\par }{\Q but
let us belong to you –
\par }{\Q take away our shame!”
\par }{\SH The Branch of the Lord
\par }{\Q \VS{2}At that time
\par }{\Q the crops
given by the
{\ND{Lord}}
will bring admiration
and honor;
\par }{\Q the produce
of the land
will be a source of pride
and delight
\par }{\Q to those who remain
in Israel.
\par }{\Q \VS{3}Those remaining
in Zion,
those left
in Jerusalem,
\par }{\Q will be called “holy,”
\par }{\Q all
in Jerusalem
who are destined
to live.
\par }{\Q \VS{4}At that time
the sovereign master
will wash
the
excrement
from Zion’s
women,
\par }{\Q he will rinse
the
bloodstains
from Jerusalem’s
midst,
\par }{\Q as he comes to judge
\par }{\Q and to bring devastation.
\par }{\Q \VS{5}Then the
{\ND{Lord}}
will create
\par }{\Q over
all
of Mount
Zion
\par }{\Q and over
its
convocations
\par }{\Q a cloud
and smoke
by day
\par }{\Q and a bright
flame
of fire
by night;
\par }{\Q indeed
a canopy
will accompany the
{\ND{Lord}}’s glorious presence.
\par }{\Q \VS{6}By day
it will be a shelter
to provide
shade
from the heat,
\par }{\Q as well as safety
and protection
from the heavy
downpour.

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}I will sing
to my love
–
\par }{\Q a song
to my lover
about his vineyard.
\par }{\Q My love
had a vineyard
\par }{\Q on a fertile hill.
\par }{\Q \VS{2}He built a hedge
around it, removed its stones,
\par }{\Q and planted
a vine.
\par }{\Q He built
a tower
in the middle
of it,
\par }{\Q and constructed
a winepress.
\par }{\Q He waited for
it to produce
edible grapes,
\par }{\Q but it produced
sour ones instead.
\par }{\Q \VS{3}So now,
residents
of Jerusalem,
\par }{\Q people
of Judah,
\par }{\Q you decide
between
me and my vineyard!
\par }{\Q \VS{4}What
more
can I do
for my vineyard
\par }{\Q beyond what I have already done?
\par }{\Q When I waited
for it to produce
edible grapes,
\par }{\Q why
did
it produce sour ones instead?
\par }{\Q \VS{5}Now
I will inform
you
\par }{\Q what
I am
about to do
to my vineyard:
\par }{\Q I will remove
its hedge
and turn
it into
pasture,
\par }{\Q I will break
its wall
and allow
animals to graze there.
\par }{\Q \VS{6}I will make
it a wasteland;
\par }{\Q no
one will prune
its vines or
hoe
its ground,

\par }{\Q and thorns
and briers
will grow
there.
\par }{\Q I will order
the clouds
\par }{\Q not to drop
any rain
on it.
\par }{\Q \VS{7}Indeed
Israel
is the vineyard
of the {\ND{Lord}}
who commands armies,
\par }{\Q the people
of Judah
are the cultivated place
in which he took delight.
\par }{\Q He waited
for justice,
but look what he got – disobedience!
\par }{\Q He waited for fairness, but look what he got – cries for help!
\par }{\SH Disaster is Coming
\par }{\Q \VS{8}Those who accumulate
houses
are as good as dead,

\par }{\Q those who also accumulate
landed
property

\par }{\Q until
there is no land left,

\par }{\Q and you are the only
landowners
remaining within
the land.
\par }{\Q \VS{9}The
{\ND{Lord}}
who commands armies
told
me this:

\par }{\Q “Many
houses
will certainly become desolate,
\par }{\Q large,
impressive
houses will have no
one living in them.
\par }{\Q \VS{10}Indeed,
a large
vineyard
will produce
just a few
gallons,
\par }{\Q and enough seed
to yield several bushels
will produce
less than a bushel.”
\par }{\Q \VS{11}Those who get up
early
to drink beer
are as good as dead,
\par }{\Q those who keep
drinking long after
dark
\par }{\Q until they are intoxicated
with wine.
\par }{\Q \VS{12}They have stringed instruments,
tambourines,
flutes,
\par }{\Q and wine
at their parties.
\par }{\Q So they do not
recognize
what the
{\ND{Lord}}
is doing,
\par }{\Q they do not
perceive what he is bringing about.
\par }{\Q \VS{13}Therefore
my people
will be deported
\par }{\Q because of their lack
of understanding.
\par }{\Q Their leaders
will have nothing to eat,
\par }{\Q their masses
will have nothing to drink.
\par }{\Q \VS{14}So
Death
will open
up
its throat,
\par }{\Q and open wide
its mouth;
\par }{\Q Zion’s dignitaries
and masses will descend
into it,
\par }{\Q including those who revel and celebrate within her.
\par }{\Q \VS{15}Men
will be humiliated,
\par }{\Q they
will be brought low;
\par }{\Q the proud
will be brought low.
\par }{\Q \VS{16}The
{\ND{Lord}}
who commands armies
will be exalted
when he punishes,

\par }{\Q the sovereign
God’s
authority will be recognized when he judges.
\par }{\Q \VS{17}Lambs
will graze
as if in their pastures,
\par }{\Q amid the ruins
the rich sojourners
will graze.
\par }{\Q \VS{18}Those
who pull
evil
along using cords
of emptiness
are as good as dead,

\par }{\Q who pull sin
as with cart
ropes.
\par }{\Q \VS{19}They say,
“Let him hurry,
let him act
quickly,
\par }{\Q so
we can see;
\par }{\Q let the plan
of the Holy One
of Israel
take shape
and come
to pass,
\par }{\Q then we will know it!”
\par }{\Q \VS{20}Those
who call evil
good
and good
evil
are as good as dead,

\par }{\Q who turn
darkness
into light
and light
into darkness,
\par }{\Q who turn
bitter
into sweet
and sweet
into bitter.
\par }{\Q \VS{21}Those
who think they are wise
are as good as dead,

\par }{\Q those who think
they possess
understanding.
\par }{\Q \VS{22}Those who are champions
at drinking
wine
are as good as dead,

\par }{\Q who display
great courage when mixing
strong drinks.
\par }{\Q \VS{23}They pronounce the guilty
innocent
for
a payoff,
\par }{\Q they ignore
the just
cause of the innocent.
\par }{\Q \VS{24}Therefore,
as flaming
fire
devours
straw,
\par }{\Q and dry grass
disintegrates
in the flames,
\par }{\Q so their root
will rot,
\par }{\Q and their flower
will blow away
like dust.
\par }{\Q For
they have rejected
the
law
of the {\ND{Lord}}
who commands armies,
\par }{\Q they have spurned
the
commands
of the Holy One
of Israel.
\par }{\Q \VS{25}So
the {\ND{Lord}}
is furious
with his people;
\par }{\Q he lifts
his hand
and strikes
them.
\par }{\Q The mountains
shake,
\par }{\Q and corpses
lie
like manure
in the middle
of the streets.
\par }{\Q Despite all
this,
his anger
does not
subside,
\par }{\Q and his hand
is ready to strike
again.
\par }{\Q \VS{26}He lifts a signal flag
for a distant
nation,
\par }{\Q he whistles
for it to come from the far regions
of the earth.
\par }{\Q Look,
they come
quickly
and swiftly.
\par }{\Q \VS{27}None
tire
or stumble,
\par }{\Q they don’t
stop to nap
or sleep.
\par }{\Q They don’t
loosen
their belts,
\par }{\Q or unstrap
their sandals
to rest.
\par }{\Q \VS{28}Their arrows
are sharpened,
\par }{\Q and all
their bows
are prepared.
\par }{\Q The hooves
of their horses
are hard as flint,
\par }{\Q and their chariot wheels
are like a windstorm.
\par }{\Q \VS{29}Their roar
is like a lion’s;
\par }{\Q they roar
like young lions.
\par }{\Q They growl
and seize
their prey;
\par }{\Q they drag
it away
and no
one can come to the rescue.
\par }{\Q \VS{30}At that time
they will growl
over
their prey,

\par }{\Q it
will sound
like sea
waves crashing against
rocks.

\par }{\Q One will look
out over the land
and see
the darkness
of disaster,
\par }{\Q clouds
will turn the light
into darkness.

\par }\Chap{6}{\PP \VerseOne{1}In the year
of King
Uzziah’s
death,
I saw
the sovereign
master seated
on
a high,
elevated
throne.
The hem
of his robe filled
the temple.
\VS{2}Seraphs
stood
over
him; each one
had six
wings.
With two
wings they covered
their faces,
with two
they covered
their feet,
and they used the remaining two
to fly.
\VS{3}They called out
to
one
another,
“Holy,
holy,
holy is the Lord who commands armies! His majestic splendor fills the entire earth!”
\VS{4}The sound of their voices
shook the door frames,
and the temple
was filled
with smoke.
\par }{\PP \VS{5}I said,
“Too bad
for
me! I am
destroyed,
for
my lips
are contaminated
by sin, and I
live
among
people
whose lips
are contaminated
by sin. My eyes
have seen
the king,
the {\ND{Lord}}
who commands armies.”
\VS{6}But then one
of the seraphs
flew
toward
me. In his hand
was a hot coal
he had taken
from the altar
with tongs.
\VS{7}He touched
my mouth
with it and said,
“Look,
this
coal has touched
your lips.
Your evil
is removed;
your sin
is forgiven.”
\VS{8}I heard
the voice
of the sovereign
master say,
“Whom
will I send? Who
will go
on our behalf?” I answered,
“Here I am,
send me!”
\VS{9}He said,
“Go
and tell
these
people:
\par }{\Q ‘Listen
continually, but don’t
understand!
\par }{\Q Look
continually, but don’t
perceive!’
\par }{\Q \VS{10}Make the hearts
of these
people
calloused;
\par }{\Q make their ears
deaf
and their eyes
blind!
\par }{\Q Otherwise
they might see
with their eyes
and hear
with their ears,
\par }{\Q their hearts
might understand
and they might repent
and be healed.”
\par }{\Q \VS{11}I replied,
“How
long,
sovereign master?” He said,
\par }{\Q “Until
cities
are in ruins
and unpopulated,
\par }{\Q and houses
are uninhabited,
\par }{\Q and the land
is ruined
and devastated,
\par }{\Q \VS{12}and the
{\ND{Lord}}
has sent
the
people
off to a distant
place,
\par }{\Q and the very heart
of the land
is completely abandoned.
\par }{\Q \VS{13}Even
if only a tenth
of the people remain in the land, it will again
be destroyed,
like one of the large sacred trees
or an Asherah pole, when a sacred pillar
on a high place
is thrown down.
That sacred pillar
symbolizes the special chosen family.”

\par }\Chap{7}{\PP \VerseOne{1}During
the reign
of Ahaz
son
of Jotham,
son
of Uzziah,
king
of Judah,
King
Rezin
of Syria
and King
Pekah
son
of Remaliah
of Israel
marched up to Jerusalem
to do battle,
but
they were unable
to prevail against
it.
\par }{\PP \VS{2}It was reported
to the family
of David,
“Syria
has allied
with
Ephraim.”
They and their people
were emotionally
shaken,
just as the trees
of the forest
shake
before
the wind.
\VS{3}So the
{\ND{Lord}}
told
Isaiah,
“Go out
with your son
Shear-jashub
and meet
Ahaz
at the end
of the conduit
of the upper
pool
which is located on
the road
to the field
where they wash and dry cloth.
\VS{4}Tell
him,
‘Make sure
you stay calm! Don’t
be afraid! Don’t
be intimidated
by these
two
stubs
of smoking
logs,
or by the raging
anger
of Rezin,
Syria,
and the son
of Remaliah.
\VS{5}Syria
has plotted
with Ephraim
and the son
of Remaliah to bring about your demise.
\VS{6}They say, “Let’s attack
Judah,
terrorize
it, and conquer it. Then we’ll set up the son
of Tabeel
as
its king.”
\VS{7}For this
reason the sovereign
master, the
{\ND{Lord}}, says:
\par }{\Q “It will not
take place;
\par }{\Q it will not
happen.
\par }{\Q \VS{8}For
Syria’s
leader
is Damascus,
\par }{\Q and the leader
of Damascus
is Rezin.
\par }{\Q Within sixty-five
years
Ephraim
will no longer exist as a nation.
\par }{\Q \VS{9}Ephraim’s
leader
is Samaria,
\par }{\Q and Samaria’s
leader
is the son
of Remaliah.
\par }{\Q If
your faith does not
remain firm,
\par }{\Q then you will not
remain secure.”
\par }{\PP \VS{10}The
{\ND{Lord}}
again
spoke
to
Ahaz:
\VS{11}“Ask
for a confirming sign
from
the {\ND{Lord}}
your God.
You can even
ask for something miraculous.”
\VS{12}But Ahaz
responded, “I don’t
want to ask;
I don’t
want to put the
{\ND{Lord}}
to a test.”
\VS{13}So Isaiah replied, “Pay attention,
family
of David.
Do you consider it too insignificant
to try the patience
of men? Is that
why you are
also
trying the patience
of my God?
\VS{14}For this reason
the sovereign
master himself
will give you a confirming sign.
Look,
this young woman
is about to conceive
and will give birth
to a son.
You, young woman, will name
him Immanuel.
\VS{15}He will eat
sour milk
and honey,
which will help him know
how to reject
evil
and choose
what is right.
\VS{16}Here is why this will be so: Before
the child
knows
how to reject
evil
and choose
what is right,
the land
whose
two
kings
you
fear will be desolate.
\VS{17}The
{\ND{Lord}}
will bring
on
you, your people,
and your father’s
family
a time
unlike
any since
Ephraim
departed
from
Judah
– the king
of Assyria!”
\par }{\PP \VS{18}At that
time
the {\ND{Lord}}
will whistle
for flies
from the distant
streams
of Egypt
and for bees
from the land
of Assyria.
\VS{19}All
of them will come
and make their home
in the ravines
between the cliffs,
and in the crevices
of the cliffs,
in all
the thorn bushes,
and in all
the watering holes.
\VS{20}At that time
the sovereign
master will use a razor
hired
from the banks
of the Euphrates River,
the king
of Assyria,
to shave
the head
and the pubic hair;
it will also
shave off
the beard.
\VS{21}At that time
a man
will keep alive
a young cow
from the herd
and a couple
of goats.
\VS{22}From the abundance
of milk
they produce,
he will have sour milk
for
his meals.
Indeed, everyone
left
in the heart
of the land
will eat
sour milk
and honey.
\VS{23}At that
time
every
place
where
there
had been a thousand
vines
worth a thousand
shekels
will be overrun
with thorns
and briers.
\VS{24}With bow
and arrow
men will hunt
there,
for
the whole
land
will be
covered with thorns
and briers.
\VS{25}They will stay away from all
the hills
that
were cultivated,
for fear
of the thorns
and briers.
Cattle
will graze
there and sheep
will trample on them.

\par }\Chap{8}{\PP \VerseOne{1}The
{\ND{Lord}}
told
me,
“Take
a large
tablet
and inscribe
these words on
it with an ordinary stylus: ‘Maher-Shalal-Hash-Baz.’
\VS{2}Then I will summon
as my reliable
witnesses
Uriah
the priest
and Zechariah
son
of Jeberekiah.”
\VS{3}I then had sexual relations
with the prophetess;
she conceived
and gave birth
to a son.
The
{\ND{Lord}}
told
me,
“Name
him Maher-Shalal-Hash-Baz,
\VS{4}for
before
the child
knows
how to cry
out, ‘My father’
or ‘My mother,’
the
wealth
of Damascus
and the
plunder
of Samaria
will be carried
off by the king
of Assyria.”
\par }{\PP \VS{5}The
{\ND{Lord}}
spoke
to me
again:
\VS{6}“These people
have rejected
the gently
flowing
waters
of Shiloah
and melt in fear
over Rezin
and the son
of Remaliah.
\VS{7}So
look,
the sovereign master
is bringing up
against
them the turbulent
and mighty
waters
of the Euphrates River –
the king
of Assyria
and all
his majestic
power. It will reach
flood stage
and overflow
its banks.
\VS{8}It will
spill
into Judah,
flooding
and engulfing, as
it reaches
to the necks
of its victims. He will spread
his wings
out over
your entire
land,
O Immanuel.”
\par }{\Q \VS{9}You will be broken,
O nations;
\par }{\Q you will be shattered!

\par }{\Q Pay attention,
all
you distant
lands of the earth!
\par }{\Q Get ready
for battle, and you will be shattered!
\par }{\Q Get ready
for battle, and you will be shattered!
\par }{\Q \VS{10}Devise
your strategy,
but it will be thwarted!
\par }{\Q Issue
your orders,
but they will not
be executed!

\par }{\Q For
God
is with us!
\par }{\SH The Lord Encourages Isaiah
\par }{\PP \VS{11}Indeed
this is what
the {\ND{Lord}}
told
me. He took
hold of me firmly
and warned
me not to act
like these
people:
\par }{\Q \VS{12}“Do not
say,
‘Conspiracy,’
every
time these people
say
the word.

\par }{\Q Don’t be afraid
of what
scares
them; don’t be terrified.
\par }{\Q \VS{13}You must
recognize
the authority
of the {\ND{Lord}}
who commands armies.
\par }{\Q He is the one you must respect;
\par }{\Q he is the one you must fear.
\par }{\Q \VS{14}He will become
a sanctuary,
\par }{\Q but a stone
that makes a person trip,
\par }{\Q and a rock
that makes one stumble
–
\par }{\Q to the two
houses
of Israel.
\par }{\Q He will become a trap
and a snare
\par }{\Q to the residents
of Jerusalem.
\par }{\Q \VS{15}Many
will stumble
over the stone and the rock,

\par }{\Q and will fall
and be
seriously injured,
\par }{\Q and will be
ensnared
and captured.”
\par }{\Q \VS{16}Tie up
the scroll as legal evidence,
\par }{\Q seal
the official record of
God’s instructions
and give it to my followers.
\par }{\Q \VS{17}I will wait patiently
for the
{\ND{Lord}},
\par }{\Q who has rejected
the family
of Jacob;
\par }{\Q I will wait for him.
\par }{\Q \VS{18}Look,
I
and the sons
whom
the {\ND{Lord}}
has given
me are reminders
and object lessons
in Israel,
sent from
the {\ND{Lord}}
who commands armies,
who lives
on Mount
Zion.
\par }{\SH Darkness Turns to Light as an Ideal King Arrives
\par }{\PP \VS{19}They will say
to
you, “Seek
oracles
at the pits used to
conjure
up underworld
spirits, from the magicians
who chirp
and mutter
incantations. Should people
not
seek
oracles
from their gods,
by asking the dead
about
the destiny
of the living?”
\VS{20}Then you must recall the
{\ND{Lord}}’s instructions
and the prophetic testimony
of what would happen. Certainly they say
such things
because their minds are spiritually darkened.
\VS{21}They will pass
through the land destitute
and starving.
Their hunger
will
make them angry,
and they will curse
their king
and their God
as they look
upward.
\VS{22}When one looks out
over the land,
he sees
distress
and darkness,
gloom
and anxiety,
darkness
and people forced from the land.

\par }\Chap{9}{\PP \VerseOne{1}The gloom
will be dispelled
for
those who
were anxious.
\par }{\PP \par }{\Q In earlier
times
he humiliated
\par }{\Q the land
of Zebulun,
\par }{\Q and the land
of Naphtali;
\par }{\Q but now
he brings honor
\par }{\Q to the way
of the sea,
\par }{\Q the region beyond
the Jordan,
\par }{\Q and Galilee
of the nations.
\par }{\Q \VS{2}The people
walking
in darkness
\par }{\Q see
a bright
light;
\par }{\Q light
shines
\par }{\Q on
those who live in
a land
of deep darkness.
\par }{\Q \VS{3}You have enlarged
the nation;
\par }{\Q you give them great
joy.
\par }{\Q They rejoice
in your presence
\par }{\Q as
harvesters
rejoice;
\par }{\Q as warriors celebrate when
they divide
up the plunder.
\par }{\Q \VS{4}For
their oppressive
yoke
\par }{\Q and the club
that strikes their shoulders,
\par }{\Q the cudgel
the oppressor
uses on them,

\par }{\Q you have shattered,
as in the day
of Midian’s defeat.
\par }{\Q \VS{5}Indeed
every
boot
that marches
and shakes the earth
\par }{\Q and every garment
dragged through
blood
\par }{\Q is used
as fuel
for the fire.
\par }{\Q \VS{6}For
a child
has been born
to us,
\par }{\Q a son
has been given
to us.
\par }{\Q He shoulders
responsibility
\par }{\Q and is called:

\par }{\Q Extraordinary
Strategist,
\par }{\Q Mighty
God,
\par }{\Q Everlasting
Father,

\par }{\Q Prince
of Peace.
\par }{\Q \VS{7}His dominion
will be vast
\par }{\Q and he will bring immeasurable
prosperity.
\par }{\Q He will rule on
David’s
throne
\par }{\Q and over
David’s kingdom,
\par }{\Q establishing
it and strengthening
it
\par }{\Q by promoting justice
and fairness,
\par }{\Q from this time
forward and forevermore.
\par }{\Q The
{\ND{Lord}}’s
intense
devotion
to his people will accomplish
this.
\par }{\SH God’s Judgment Intensifies
\par }{\Q \VS{8}The sovereign
master
decreed judgment on Jacob,
\par }{\Q and it fell
on Israel.
\par }{\Q \VS{9}All
the people were aware
of it,
\par }{\Q the people
of Ephraim
and those living in
Samaria.
\par }{\Q Yet with pride
and an arrogant
attitude,
they said,
\par }{\Q \VS{10}“The bricks
have fallen,
\par }{\Q but we will rebuild
with chiseled
stone;
\par }{\Q the sycamore fig trees
have been cut down,
\par }{\Q but we will replace
them with cedars.”
\par }{\Q \VS{11}Then the
{\ND{Lord}}
provoked
their adversaries
to attack them,

\par }{\Q he stirred up their enemies –
\par }{\Q \VS{12}Syria
from the east,
\par }{\Q and the Philistines
from the west,
\par }{\Q they gobbled up
Israelite
territory.

\par }{\Q Despite all
this,
his anger
does not
subside,
\par }{\Q and his hand
is ready to strike
again.
\par }{\Q \VS{13}The people
did not
return
to
the one who struck
them,
\par }{\Q they did not
seek
reconciliation
with the
{\ND{Lord}}
who commands armies.
\par }{\Q \VS{14}So the
{\ND{Lord}}
cut off
Israel’s
head
and tail,
\par }{\Q both the shoots
and stalk
in one
day.
\par }{\Q \VS{15}The leaders
and the highly
respected people are the head,
\par }{\Q the prophets
who teach
lies
are the tail.
\par }{\Q \VS{16}The leaders
of this
nation
were misleading
people,
\par }{\Q and the people being led
were destroyed.
\par }{\Q \VS{17}So
the sovereign
master was not pleased
with their young men,
\par }{\Q he took
no
pity
on their orphans
and widows;
\par }{\Q for
the whole
nation was godless
and did wicked
things,

\par }{\Q every
mouth
was speaking
disgraceful words.
\par }{\Q Despite all
this,
his anger
does not
subside,
\par }{\Q and his hand
is ready to strike
again.
\par }{\Q \VS{18}For
evil
burned
like a fire,
\par }{\Q it consumed
thorns
and briers;
\par }{\Q it burned up
the thickets
of the forest,
\par }{\Q and they went up
in smoke.
\par }{\Q \VS{19}Because of the anger
of the
{\ND{Lord}}
who commands armies,
the land
was scorched,
\par }{\Q and the people
became fuel
for the fire.
\par }{\Q People
had no
compassion
on
one another.
\par }{\Q \VS{20}They devoured
on
the right,
but were still hungry,
\par }{\Q they ate
on
the left,
but were not
satisfied.
\par }{\Q People
even ate
the flesh
of their own arm!
\par }{\Q \VS{21}Manasseh
fought against Ephraim,
\par }{\Q and Ephraim
against Manasseh;
\par }{\Q together
they
fought against
Judah.
\par }{\Q Despite all
this,
his anger
does not
subside,
\par }{\Q and his hand
is ready to strike
again.

\par }\Chap{10}{\PP \par }{\Q \VerseOne{1}Those who enact
unjust
policies
are as good as dead,

\par }{\Q those who are always instituting
unfair
regulations,
\par }{\Q \VS{2}to keep the poor
from getting
fair treatment,
\par }{\Q and to deprive
the oppressed
among my people
of justice,
\par }{\Q so they can steal what
widows
own,
\par }{\Q and loot
what belongs
to orphans.
\par }{\Q \VS{3}What
will you do
on judgment
day,
\par }{\Q when destruction
arrives
from
a distant place?
\par }{\Q To whom
will you run
for help?
\par }{\Q Where
will you leave
your wealth?
\par }{\Q \VS{4}You will have no place
to go, except to kneel
with the prisoners,
\par }{\Q or to fall
among those who have been killed.
\par }{\Q Despite
all
this,
his anger
does not
subside,
\par }{\Q and his hand
is ready to strike
again.
\par }{\SH The Lord Turns on Arrogant Assyria
\par }{\Q \VS{5}Assyria,
the club
I use to vent my anger,
is as good as dead,
\par }{\Q a cudgel
with which I angrily punish.
\par }{\Q \VS{6}I sent
him against
a godless
nation,
\par }{\Q I ordered
him to attack the people
with whom I was angry,
\par }{\Q to take
plunder
and to carry
away loot,
\par }{\Q to trample them down
like dirt
in the streets.
\par }{\Q \VS{7}But he does not
agree
with this,
\par }{\Q his mind
does not
reason
this way,

\par }{\Q for
his goal
is to destroy,
\par }{\Q and to eliminate
many nations.
\par }{\Q \VS{8}Indeed,
he says:
\par }{\Q “Are not
my officials
all
kings?
\par }{\Q \VS{9}Is not
Calneh
like Carchemish?
\par }{\Q Hamath
like Arpad?
\par }{\Q Samaria
like Damascus?
\par }{\Q \VS{10}I overpowered
kingdoms
ruled by idols,
\par }{\Q whose carved images
were more impressive than Jerusalem’s
or Samaria’s.
\par }{\Q \VS{11}As
I have
done
to Samaria
and its idols,
\par }{\Q so
I will do
to Jerusalem
and its idols.”
\par }{\PP \VS{12}But
when
the sovereign master
finishes
judging
Mount
Zion
and Jerusalem,
then I will punish
the king
of Assyria
for
what
he has proudly
planned
and for
the arrogant
attitude
he displays.
\VS{13}For
he says:
\par }{\Q “By my strong
hand
I have accomplished
this,
\par }{\Q by my strategy
that
I devised.
\par }{\Q I invaded
the territory
of nations,
\par }{\Q and looted
their storehouses.
\par }{\Q Like a mighty
conqueror, I brought down
rulers.
\par }{\Q \VS{14}My hand
discovered
the wealth
of the nations,
as if it were in a nest,
\par }{\Q as one gathers up
abandoned
eggs,
\par }{\Q I
gathered up
the whole
earth.
\par }{\Q There was
no
wing
flapping,
\par }{\Q or open
mouth
chirping.”
\par }{\Q \VS{15}Does an ax
exalt itself over
the one who wields it,
\par }{\Q or a saw
magnify
itself over
the one who cuts with it?

\par }{\Q As
if
a scepter
should brandish the
one who raises
it,
\par }{\Q or a staff
should lift up
what is not
made of wood!
\par }{\Q \VS{16}For
this reason the sovereign master,
the {\ND{Lord}}
who commands armies,
will make his healthy
ones emaciated.
His majestic glory
will go up in smoke.
\par }{\Q \VS{17}The light
of Israel
will become
a fire,
\par }{\Q their Holy One
will become a flame;
\par }{\Q it will burn
and consume
the Assyrian king’s briers
\par }{\Q and his thorns
in one
day.
\par }{\Q \VS{18}The splendor
of his forest
and his orchard
\par }{\Q will be
completely destroyed,

\par }{\Q as when a sick man’s life
ebbs
away.
\par }{\Q \VS{19}There will be so few
trees
left
in his forest,
\par }{\Q a child
will be
able to count them.
\par }{\PP \VS{20}At that time
those left
in Israel,
those who remain
of the family
of Jacob,
will no
longer
rely
on
a foreign leader that abuses them.
Instead they will truly
rely
on
the {\ND{Lord}}, the Holy
One of Israel.
\VS{21}A remnant
will come back,
a remnant
of Jacob,
to
the mighty
God.
\VS{22}For
though
your people,
Israel,
are as numerous as the sand
on the seashore,
only a remnant
will come back.
Destruction
has been decreed;
just punishment is about to engulf you.
\VS{23}The sovereign
master, the
{\ND{Lord}}
who commands armies,
is certainly ready to carry
out the decreed
destruction
throughout
the land.
\par }{\PP \VS{24}So
here is what
the sovereign
master, the
{\ND{Lord}}
who commands armies,
says: “My people
who live
in Zion,
do not
be afraid
of Assyria,
even though they beat
you with a club
and lift
their cudgel
against
you as Egypt did.
\VS{25}For
very
soon
my fury
will subside,
and my anger
will be directed toward their destruction.”
\VS{26}The

{\ND{Lord}}
who commands armies
is about to beat them with a whip,
similar
to the way
he struck down Midian
at the rock
of Oreb.
He will use his staff
against
the sea,
lifting
it up as he did in Egypt.
\par }{\Q \VS{27}At that time
\par }{\Q the
{\ND{Lord}} will remove
their burden
from
your shoulders,
\par }{\Q and their yoke
from
your neck;
\par }{\Q the yoke
will be taken off
because
your neck will be too large.
\VS{28}\par }{\Q They attacked
Aiath,
\par }{\Q moved through
Migron,
\par }{\Q depositing
their supplies
at Micmash.
\par }{\Q \VS{29}They went through
the pass,
\par }{\Q spent the night
at Geba.
\par }{\Q Ramah
trembled,
\par }{\Q Gibeah
of Saul
ran away.
\par }{\Q \VS{30}Shout out,
daughter
of Gallim!
\par }{\Q Pay attention,
Laishah!
\par }{\Q Answer her, Anathoth!
\par }{\Q \VS{31}Madmenah
flees,
\par }{\Q the residents
of Gebim have hidden.
\par }{\Q \VS{32}This very
day,
standing
in Nob,
\par }{\Q they shake their fist
at Daughter
Zion’s
mountain – ‘span class=”s 01389” title=”01389” ’
\par }{\Q at the hill
of Jerusalem.
\par }{\PP \par }{\Q \VS{33}Look,
the sovereign master,
the {\ND{Lord}}
who commands armies,
\par }{\Q is ready to cut off
the branches
with terrifying power.
\par }{\Q The tallest
trees will be cut down,
\par }{\Q the loftiest
ones will be brought low.
\par }{\Q \VS{34}The thickets
of the forest
will be chopped down
with an ax,
\par }{\Q and mighty
Lebanon
will fall.

\par }\Chap{11}{\PP \par }{\Q \VerseOne{1}A shoot
will grow out
of Jesse’s
root stock,
\par }{\Q a bud
will sprout
from his roots.
\par }{\Q \VS{2}The
{\ND{Lord}}’s
spirit
will rest
on him –
\par }{\Q a spirit that gives extraordinary wisdom,
\par }{\Q a spirit that provides the ability to execute plans,
\par }{\Q a spirit that produces absolute loyalty to the
{\ND{Lord}}.
\par }{\Q \VS{3}He will take delight
in obeying
the {\ND{Lord}}.

\par }{\Q He will not
judge
by mere appearances,
\par }{\Q or make decisions
on the basis
of hearsay.
\par }{\Q \VS{4}He will treat
the poor
fairly,
\par }{\Q and make
right decisions
for the downtrodden
of the earth.
\par }{\Q He will strike
the earth
with the rod
of his mouth,
\par }{\Q and order the wicked
to be executed.
\par }{\Q \VS{5}Justice
will be
like a belt
around his waist,
\par }{\Q integrity
will be like a belt
around his hips.
\par }{\Q \VS{6}A wolf
will reside
with
a lamb,
\par }{\Q and a leopard
will lie down
with
a young goat;
\par }{\Q an ox
and a young lion
will graze
together,
\par }{\Q as a small
child
leads them along.
\par }{\Q \VS{7}A cow
and a bear
will graze
together,
\par }{\Q their young
will lie
down together.

\par }{\Q A lion,
like an ox,
will eat
straw.
\par }{\Q \VS{8}A baby
will play
\par }{\Q over
the hole
of a snake;
\par }{\Q over
the nest
of a serpent
\par }{\Q an infant
will put
his hand.
\par }{\Q \VS{9}They will no
longer injure
or
destroy
\par }{\Q on my entire
royal mountain.
\par }{\Q For
there will be universal
submission
to the

{\ND{Lord}}’s
sovereignty,
\par }{\Q just as the waters
completely cover
the sea.
\par }{\SH Israel is Reclaimed and Reunited
\par }{\PP \VS{10}At that time
a root
from Jesse
will stand
like a signal flag
for the nations.
Nations
will look to him for guidance,
and his residence
will be
majestic.
\VS{11}At that
time
the sovereign master
will
again
lift
his hand
to reclaim
the remnant
of his people
from Assyria,
Egypt,
Pathros,
Cush,
Elam,
Shinar,
Hamath,
and the seacoasts.
\par }{\Q \VS{12}He will lift
a signal flag
for the nations;
\par }{\Q he will gather
Israel’s
dispersed
people

\par }{\Q and assemble
Judah’s
scattered
people
\par }{\Q from the four
corners
of the earth.
\par }{\Q \VS{13}Ephraim’s
jealousy
will end,

\par }{\Q and Judah’s
hostility
will be eliminated.
\par }{\Q Ephraim
will no
longer be jealous
of Judah,
\par }{\Q and Judah
will no
longer be hostile
toward Ephraim.
\par }{\Q \VS{14}They will swoop down
on the Philistine
hills
to the west;
\par }{\Q together
they will loot
the
people
of the east.
\par }{\Q They will take over
Edom
and Moab,
\par }{\Q and the Ammonites
will be their subjects.
\par }{\Q \VS{15}The
{\ND{Lord}}
will divide
the gulf
of the Egyptian
Sea;
\par }{\Q he will wave
his hand
over
the Euphrates River
and send a strong
wind,
\par }{\Q he will turn it into seven
dried-up
streams,
\par }{\Q and enable them to walk across
in their sandals.
\par }{\Q \VS{16}There will be
a highway
leading
out
of Assyria
\par }{\Q for the remnant
of his people,
\par }{\Q just as
there was
for Israel,
\par }{\Q when
they went up
from the land
of Egypt.

\par }\Chap{12}{\PP \par }{\Q \VerseOne{1}At that time
you will say:
\par }{\Q “I praise
you, O
{\ND{Lord}},
\par }{\Q for
even though you were angry
with me,
\par }{\Q your anger
subsided,
and you consoled me.
\par }{\Q \VS{2}Look,
God
is my deliverer!

\par }{\Q I will trust
in him and not
fear.
\par }{\Q For
the
{\ND{Lord}}
gives me strength
and protects
me;

\par }{\Q he
has become
my deliverer.”
\par }{\Q \VS{3}Joyfully
you will draw
water
\par }{\Q from the springs
of deliverance.
\par }{\Q \VS{4}At that time
you will say:
\par }{\Q “Praise
the
{\ND{Lord}}!
\par }{\Q Ask him for
help!

\par }{\Q Publicize his
mighty acts among
the nations!
\par }{\Q Make
it known
that
he is unique!
\par }{\Q \VS{5}Sing
to the
{\ND{Lord}},
for
he has done
magnificent
things,
\par }{\Q let this
be known
throughout
the earth!
\par }{\Q \VS{6}Cry out
and shout
for joy, O citizens
of Zion,
\par }{\Q for
the Holy One
of Israel
acts mightily
among you!”

\par }\Chap{13}{\PP \par }{\Q \VerseOne{1}This is a message
about Babylon
that
God revealed
to Isaiah
son
of Amoz:
\VS{2}\par }{\Q On
a bare
hill
raise
a signal flag,
\par }{\Q shout
to them,
\par }{\Q wave
your hand,
\par }{\Q so they might enter
the gates
of the princes!
\par }{\Q \VS{3}I
have given orders
to my chosen
soldiers;

\par }{\Q I have
summoned
the warriors
through whom I will vent my anger,
\par }{\Q my boasting,
arrogant ones.
\VS{4}\par }{\Q There is a loud
noise
on the mountains
–
\par }{\Q it sounds like
a large
army!

\par }{\Q There is great
commotion
among the kingdoms –
\par }{\Q nations
are being assembled!
\par }{\Q The
{\ND{Lord}}
who commands armies
is mustering
\par }{\Q forces
for battle.
\par }{\Q \VS{5}They come
from a distant
land,
\par }{\Q from the horizon.
\par }{\Q It is the
{\ND{Lord}}
with his instruments
of judgment,
\par }{\Q coming to destroy
the whole
earth.
\par }{\Q \VS{6}Wail,
for
the
{\ND{Lord}}’s
day
of judgment is near;
\par }{\Q it comes
with all the destructive
power of the sovereign judge.
\par }{\Q \VS{7}For this reason
all
hands
hang limp,
\par }{\Q every
human heart
loses
its courage.
\par }{\Q \VS{8}They panic
–
\par }{\Q cramps
and pain
seize hold
of them
\par }{\Q like those of a woman who is straining
to give birth.
\par }{\Q They look at
one
another
in astonishment;
\par }{\Q their faces
are flushed red.
\par }{\Q \VS{9}Look,
the
{\ND{Lord}}’s
day
of judgment is coming;
\par }{\Q it is a day of cruelty
and savage,
raging
anger,
\par }{\Q destroying
the earth
\par }{\Q and annihilating
its sinners.
\par }{\Q \VS{10}Indeed
the stars
in the sky
and their constellations
\par }{\Q no
longer give out
their light;
\par }{\Q the sun
is darkened
as soon as it rises,
\par }{\Q and the moon
does not
shine.
\VS{11}\par }{\Q I will punish
the world
for its evil,
\par }{\Q and wicked
people for their sin.
\par }{\Q I will put an end
to the pride
of the insolent,
\par }{\Q I will bring down
the arrogance
of tyrants.
\par }{\Q \VS{12}I will make human beings more scarce than pure gold,
\par }{\Q and people
more scarce than gold
from Ophir.
\par }{\Q \VS{13}So
I will shake
the heavens,
\par }{\Q and the earth
will shake
loose
from its
foundation,

\par }{\Q because of the fury
of the {\ND{Lord}}
who commands armies,
\par }{\Q in the day
he vents his raging
anger.
\par }{\Q \VS{14}Like a frightened gazelle
\par }{\Q or a sheep
with no
shepherd,
\par }{\Q each
will turn toward home,
\par }{\Q each
will run
to
his homeland.
\par }{\Q \VS{15}Everyone
who
is caught
will be stabbed;
\par }{\Q everyone
who is seized will
die
by the sword.
\par }{\Q \VS{16}Their children
will be smashed
to pieces before their very eyes;
\par }{\Q their houses
will be looted
\par }{\Q and their wives
raped.
\par }{\Q \VS{17}Look,
I am stirring
up the Medes
to attack them;

\par }{\Q they are not concerned about
silver,
\par }{\Q nor are they interested
in gold.
\par }{\Q \VS{18}Their arrows
will cut young men
to ribbons;
\par }{\Q they have no compassion on
a person’s offspring,
\par }{\Q they will not
look
with pity
on
children.
\par }{\Q \VS{19}Babylon,
the most admired
of kingdoms,
\par }{\Q the Chaldeans’
source of honor
and pride,
\par }{\Q will be destroyed
by God
\par }{\Q just as Sodom
and Gomorrah were.
\par }{\Q \VS{20}No one
will live
there again;
\par }{\Q no one
will ever
reside
there again.
\par }{\Q No
bedouin
will camp
there,
\par }{\Q no
shepherds
will rest
their flocks there.
\par }{\Q \VS{21}Wild animals
will rest
there,
\par }{\Q the ruined houses
will be full
of hyenas.
\par }{\Q Ostriches
will live
there,
\par }{\Q wild goats
will skip among the ruins.
\par }{\Q \VS{22}Wild dogs
will yip
in her ruined fortresses,
\par }{\Q jackals
will yelp in the once-splendid
palaces.
\par }{\Q Her time
is almost
up,
\par }{\Q her days
will not
be prolonged.


\par }\Chap{14}{\PP \VerseOne{1}The

{\ND{Lord}}
will certainly
have compassion
on Jacob;
he will again
choose
Israel
as his special
people and restore them
to their land.
Resident foreigners
will join
them and unite
with the family
of Jacob.
\VS{2}Nations
will take
them and bring
them back to
their own
place.
Then the family
of Jacob
will make foreigners
their servants
as they settle
in the
{\ND{Lord}}’s
land.
They will make their captors
captives and rule
over the ones who oppressed
them.
\VS{3}When
the {\ND{Lord}}
gives you relief
from your suffering
and anxiety,
and from
the hard
labor
which
you were made to perform,
\VS{4}you will taunt
the king
of Babylon
with these
words:

\par }{\Q “Look how
the oppressor
has met his end!
\par }{\Q Hostility
has ceased!
\par }{\Q \VS{5}The
{\ND{Lord}}
has broken
the club
of the wicked,
\par }{\Q the scepter
of rulers.
\par }{\Q \VS{6}It furiously
struck down
nations
\par }{\Q with unceasing
blows.
\par }{\Q It angrily
ruled over
nations,
\par }{\Q oppressing
them without
restraint.
\par }{\Q \VS{7}The whole
earth
rests
and is quiet;
\par }{\Q they break into song.
\par }{\Q \VS{8}The evergreens
also
rejoice
over your demise,

\par }{\Q as do the cedars
of Lebanon,
singing,

\par }{\Q ‘Since
you fell asleep,
\par }{\Q no
woodsman
comes up
to chop us down!’
\par }{\Q \VS{9}Sheol
below
is stirred up
about you,
\par }{\Q ready to meet
you when you arrive.
\par }{\Q It rouses
the spirits of the dead
for you,
\par }{\Q all
the former leaders
of the earth;
\par }{\Q it makes all
the former kings
of the nations
\par }{\Q rise
from their thrones.
\par }{\Q \VS{10}All
of them respond
to you, saying:
\par }{\Q ‘You too
have
become weak
like
us!
\par }{\Q You
have become just like us!
\par }{\Q \VS{11}Your splendor
has been brought down
to Sheol,
\par }{\Q as well as the sound
of your stringed instruments.
\par }{\Q You lie on a bed
of maggots,
\par }{\Q with a blanket
of worms over you.
\par }{\Q \VS{12}Look how
you have fallen
from the sky,
\par }{\Q O shining
one, son
of the dawn!

\par }{\Q You have been cut down
to the ground,
\par }{\Q O conqueror
of the nations!
\par }{\Q \VS{13}You
said
to yourself,
\par }{\Q “I will climb up
to the sky.
\par }{\Q Above
the stars
of El
\par }{\Q I will set up
my throne.
\par }{\Q I will rule
on the mountain
of assembly
\par }{\Q on the remote slopes
of Zaphon.
\par }{\Q \VS{14}I will climb up
to the tops
of the clouds;
\par }{\Q I will make myself
like the Most High!”
\par }{\Q \VS{15}But
you were brought down
to
Sheol,
\par }{\Q to
the remote
slopes of the pit.
\par }{\Q \VS{16}Those who see
you stare
at you,
\par }{\Q they look at you carefully,
thinking:

\par }{\Q “Is this
the man
who shook
the earth,
\par }{\Q the one who made kingdoms
tremble?
\par }{\Q \VS{17}Is this the one who
made
the world
like a desert,
\par }{\Q who ruined
its cities,
\par }{\Q and refused
to free
his prisoners
so they could return home?” ’
\VS{18}\par }{\Q As for all
the kings
of the nations,
\par }{\Q all
of them lie
down in splendor,
\par }{\Q each
in his own tomb.
\par }{\Q \VS{19}But you
have been thrown out
of your grave
\par }{\Q like a shoot
that is thrown away.
\par }{\Q You lie among
the slain,
\par }{\Q among those who have been slashed
by the sword,
\par }{\Q among those headed for
the stones
of the pit,
\par }{\Q as if you were a mangled
corpse.
\par }{\Q \VS{20}You will not
be buried
with
them,

\par }{\Q because
you destroyed
your land
\par }{\Q and killed
your people.
\par }{\Q The offspring
of the wicked
\par }{\Q will never
be mentioned again.
\par }{\Q \VS{21}Prepare
to execute
his sons
\par }{\Q for the sins
their ancestors
have committed.

\par }{\Q They must not
rise up
and take possession
of the earth,
\par }{\Q or fill
the surface
of the world
with cities.”
\par }{\Q \VS{22}“I will rise
up against
them,”
\par }{\Q says
the
{\ND{Lord}}
who commands armies.
\par }{\Q “I will blot out
all remembrance
of Babylon
and destroy all her people,
\par }{\Q including the offspring
she produces,”

\par }{\Q says
the
{\ND{Lord}}.
\par }{\Q \VS{23}“I will turn
her into
a place that is overrun
with wild animals
\par }{\Q and covered with pools
of stagnant water.
\par }{\Q I will get rid of
her, just as one sweeps
away dirt with a broom,”
\par }{\Q says
the {\ND{Lord}}
who commands armies.
\VS{24}\par }{\Q The
{\ND{Lord}}
who commands armies
makes
this solemn
vow:
\par }{\Q “Be sure of this:
\par }{\Q Just
as I have intended,
so
it will be;
\par }{\Q just
as I have planned,
it will happen.
\par }{\Q \VS{25}I will break
Assyria
in
my land,
\par }{\Q I will trample
them underfoot on
my hills.
\par }{\Q Their yoke
will be removed from
my people,
\par }{\Q the burden
will be lifted from
their shoulders.
\par }{\Q \VS{26}This
is the plan
I have devised
for the whole
earth;
\par }{\Q my hand
is ready to strike
all
the nations.”
\par }{\Q \VS{27}Indeed,
the {\ND{Lord}}
who commands armies
has a plan,
\par }{\Q and who can
possibly frustrate
it?
\par }{\Q His hand
is ready to strike,
\par }{\Q and who can
possibly stop it?
\par }{\SH The Lord Will Judge the Philistines
\par }{\PP \VS{28}In the year
King
Ahaz
died,
this
message
was revealed:
\par }{\Q \VS{29}Don’t
be so happy,
all
you Philistines,
\par }{\Q just because
the club
that beat
you has been broken!

\par }{\Q For
a viper
will grow out
of the serpent’s
root,
\par }{\Q and its fruit
will be a darting
adder.
\par }{\Q \VS{30}The poor
will graze
in my pastures;

\par }{\Q the needy
will rest
securely.
\par }{\Q But I will kill
your root
by famine;
\par }{\Q it will put to death
all your survivors.
\par }{\Q \VS{31}Wail,
O city gate!
\par }{\Q Cry out,
O city!
\par }{\Q Melt
with fear, all
you Philistines!
\par }{\Q For
out of the north
comes
a cloud of smoke,
\par }{\Q and there are no
stragglers
in its ranks.
\par }{\Q \VS{32}How
will they respond
to the messengers
of this nation?

\par }{\Q Indeed,
the {\ND{Lord}}
has made
Zion
secure;
\par }{\Q the oppressed
among his people
will find safety in her.

\par }\Chap{15}{\PP \par }{\Q \VerseOne{1}Here is a message
about Moab:
\par }{\Q Indeed,
in a night
it is devastated,
\par }{\Q Ar
of Moab
is destroyed!
\par }{\Q Indeed,
in a night
it is devastated,
\par }{\Q Kir
of Moab
is destroyed!
\par }{\Q \VS{2}They went up
to the temple,
\par }{\Q the people of Dibon
went up to the high places
to lament.
\par }{\Q Because of what happened to Nebo
and Medeba,
Moab
wails.
\par }{\Q Every
head
is shaved bare,
\par }{\Q every
beard
is trimmed off.
\par }{\Q \VS{3}In their streets
they wear sackcloth;
\par }{\Q on
their roofs
and in their town squares
\par }{\Q all
of them wail,
\par }{\Q they fall down
weeping.
\par }{\Q \VS{4}The people of Heshbon
and Elealeh
cry
out,
\par }{\Q their voices
are heard
as far
away as Jahaz.
\par }{\Q For this reason
Moab’s
soldiers shout
in distress;
\par }{\Q their courage wavers.
\par }{\Q \VS{5}My heart
cries
out because of Moab’s
plight,

\par }{\Q and for the fugitives
stretched out as far as
Zoar
and Eglath Shelishiyah.
\par }{\Q For
they weep
as they make their way up
the ascent
of Luhith;
\par }{\Q they loudly
lament
their demise
on the road
to Horonaim.
\par }{\Q \VS{6}For
the waters
of Nimrim
are gone;
\par }{\Q the grass
is dried
up,
\par }{\Q the vegetation
has disappeared,
\par }{\Q and there are
no
plants.
\par }{\Q \VS{7}For this reason
what they have
made
and stored up,
\par }{\Q they carry over
the Stream
of the Poplars.
\par }{\Q \VS{8}Indeed,
the cries of distress
echo throughout
Moabite
territory;
\par }{\Q their wailing
can be heard in
Eglaim
and Beer Elim.
\par }{\Q \VS{9}Indeed,
the waters
of Dimon
are full
of blood!
\par }{\Q Indeed,
I will heap even more trouble on Dimon.
\par }{\Q A lion
will
attack the Moabite
fugitives
\par }{\Q and the people left
in the land.

\par }\Chap{16}{\PP \par }{\Q \VerseOne{1}Send
rams
as tribute to the ruler
of the land,
\par }{\Q from Sela
in the desert
\par }{\Q to
the hill
of Daughter
Zion.
\par }{\Q \VS{2}At the fords
of the Arnon
\par }{\Q the Moabite
women
are like a bird
\par }{\Q that flies
about when forced
from its nest.
\par }{\Q \VS{3}“Bring
a plan,
make
a decision!

\par }{\Q Provide
some shade
in the middle
of the day!

\par }{\Q Hide
the fugitives! Do not
betray
the one who tries to escape!
\par }{\Q \VS{4}Please let the Moabite
fugitives
live
among you.
\par }{\Q Hide
them
from
the destroyer!”
\par }{\Q Certainly the one who applies pressure
will cease,
\par }{\Q the destroyer
will come to an end,
\par }{\Q those who trample
will disappear
from
the earth.
\par }{\Q \VS{5}Then a trustworthy
king
will be established;
\par }{\Q he will rule
in
a reliable manner,
\par }{\Q this one from David’s
family.
\par }{\Q He will be sure
to make
just decisions
\par }{\Q and will be experienced
in executing justice.
\par }{\Q \VS{6}We have heard
about Moab’s
pride,
\par }{\Q their great
arrogance,
\par }{\Q their boasting,
pride,
and excess.
\par }{\Q But their boastful claims are empty!
\par }{\Q \VS{7}So
Moab
wails over its demise –
\par }{\Q they all wail!
\par }{\Q Completely devastated, they moan
\par }{\Q about what has happened to the raisin cakes of Kir Hareseth.
\par }{\Q \VS{8}For
the fields
of Heshbon
are dried up,
\par }{\Q as well as the vines
of Sibmah.
\par }{\Q The rulers
of the nations
trample
all over
its vines,
\par }{\Q which reach
Jazer
and spread
to the desert;
\par }{\Q their shoots
spread out
and cross
the sea.
\par }{\Q \VS{9}So
I weep
along with
Jazer
\par }{\Q over the vines
of Sibmah.
\par }{\Q I will saturate
you with my tears,
Heshbon
and Elealeh,
\par }{\Q for
the conquering invaders shout triumphantly
\par }{\Q over
your fruit
and crops.
\par }{\Q \VS{10}Joy
and happiness
disappear
from
the orchards,
\par }{\Q and in the vineyards
no
one rejoices
or
shouts;
\par }{\Q no
one treads out
juice in the wine
vats – ‘span class=”s 01959” title=”01959” ’
\par }{\Q I have brought the joyful shouts
to an end.
\par }{\PP \par }{\Q \VS{11}So
my heart
constantly sighs for Moab,
like the strumming
of a harp,
\par }{\Q my inner
being sighs for Kir Hareseth.
\par }{\Q \VS{12}When
the Moabites
plead
with all
their might
at their high places,
\par }{\Q and enter
their temples
to
pray,
their prayers will be ineffective!
\par }{\PP \VS{13}This
is the message
the
{\ND{Lord}}
previously
announced
about Moab.
\VS{14}Now
the {\ND{Lord}}
makes
this announcement: “Within exactly three
years
Moab’s
splendor
will disappear,
along with all
her many
people;
there will be just
a few,
insignificant
survivors left.”

\par }\Chap{17}{\PP \par }{\Q \VerseOne{1}Here is a message
about Damascus:
\par }{\Q “Look,
Damascus
is no longer
a city,
\par }{\Q it is
a heap
of ruins!
\par }{\Q \VS{2}The cities
of Aroer
are abandoned.
\par }{\Q They will be
used for herds,
\par }{\Q which will lie
down there
in peace.
\par }{\Q \VS{3}Fortified
cities will disappear
from Ephraim,
\par }{\Q and Damascus
will lose its kingdom.
\par }{\Q The survivors
in Syria
\par }{\Q will end
up like the splendor
of the Israelites,”
\par }{\Q says
the {\ND{Lord}}
who commands armies.
\par }{\Q \VS{4}“At that
time
\par }{\Q Jacob’s
splendor
will be
greatly diminished,
\par }{\Q and he
will become skin and bones.
\par }{\Q \VS{5}It will be
as when one gathers
the grain
harvest,
\par }{\Q and his hand
gleans
the ear of grain.
\par }{\Q It will be
like one gathering
the ears of grain
\par }{\Q in the Valley
of Rephaim.
\par }{\Q \VS{6}There will be some left behind,
\par }{\Q like when an olive tree
is beaten
–
\par }{\Q two
or three
ripe olives
remain
toward the very top,
\par }{\Q four
or five
on its fruitful
branches,”
\par }{\Q says
the {\ND{Lord}}
God
of Israel.
\par }{\Q \VS{7}At that time
men
will trust
in
their creator;
\par }{\Q they will depend on
the Holy One
of Israel.
\par }{\Q \VS{8}They will no
longer trust
in the altars
their hands
made,
\par }{\Q or
depend
on the Asherah poles
and incense altars
their fingers made.
\par }{\Q \VS{9}At that
time
their fortified
cities
will be
\par }{\Q like the abandoned
summits
of the Amorites,

\par }{\Q which
they abandoned
because of the Israelites;
\par }{\Q there will be
desolation.
\par }{\Q \VS{10}For
you ignore
the God
who rescues
you;
\par }{\Q you pay
no
attention
to
your strong
protector.
\par }{\Q So this is
what happens:
\par }{\Q You cultivate
beautiful
plants
\par }{\Q and plant
exotic
vines.
\par }{\Q \VS{11}The day
you begin cultivating,
you do what you can to make
it grow;
\par }{\Q the morning
you begin planting,
you do what you can to make it sprout.
\par }{\Q Yet the harvest
will disappear
in the day
of disease
\par }{\Q and incurable
pain.
\par }{\Q \VS{12}The many
nations
massing together
are as good as dead,
\par }{\Q those who make a commotion
as loud as the roaring
of the sea’s waves.
\par }{\Q The people
making such an uproar
are as good as dead,

\par }{\Q those who make an uproar as loud as
the roaring
of powerful
waves.
\par }{\Q \VS{13}Though these people
make an uproar as loud as
the roaring
of powerful
waves,
\par }{\Q when he shouts at
them, they will flee
to a distant land,
\par }{\Q driven
before
the wind
like dead weeds
on the hills,
\par }{\Q or like dead thistles
before
a strong gale.
\par }{\Q \VS{14}In the evening
there is
sudden terror;
\par }{\Q by morning
they vanish.
\par }{\Q This
is the fate of those who try to plunder us,
\par }{\Q the destiny
of those who try
to loot us!

\par }\Chap{18}{\PP \par }{\Q \VerseOne{1}The land
of buzzing
wings
is as good as dead,
\par }{\Q the one
beyond
the rivers
of Cush,
\par }{\Q \VS{2}that sends
messengers
by sea,
\par }{\Q who glide over
the water’s
surface
in boats
made of papyrus.
\par }{\Q Go,
you swift
messengers,
\par }{\Q to
a nation
of tall,
smooth-skinned
people,

\par }{\Q to
a people
that
are feared
far and wide,
\par }{\Q to a nation
strong
and victorious,
\par }{\Q whose land
rivers
divide.
\par }{\Q \VS{3}All
you who live
in the world,
\par }{\Q who reside
on the earth,
\par }{\Q you will see
a signal flag
raised on the mountains;
\par }{\Q you will hear
a trumpet
being blown.
\par }{\Q \VS{4}For
this is what
the {\ND{Lord}}
has told
me:
\par }{\Q “I will wait and watch from my place,
\par }{\Q like
scorching
heat
produced by
the sunlight,
\par }{\Q like a cloud
of mist
in the heat
of harvest.”
\par }{\Q \VS{5}For
before
the harvest,
when the bud
has sprouted,
\par }{\Q and the ripening
fruit
appears,
\par }{\Q he will cut off
the unproductive shoots
with pruning knives;
\par }{\Q he will prune
the tendrils.
\par }{\Q \VS{6}They will all
be left
for the birds
of the hills
\par }{\Q and the wild animals;
\par }{\Q the birds
will eat them during the summer,
\par }{\Q and all
the wild animals
will eat them during the winter.
\par }{\Q \VS{7}At that
time
\par }{\Q tribute
will be brought
to the
{\ND{Lord}}
who commands armies,
\par }{\Q by a people
that are tall
and smooth-skinned,
\par }{\Q a people
that
are feared
far and wide,
\par }{\Q a nation
strong
and victorious,
\par }{\Q whose land
rivers
divide.
\par }{\Q The tribute will be brought to
the place
where the
{\ND{Lord}}
who commands armies
has chosen to reside, on Mount
Zion.

\par }\Chap{19}{\PP \par }{\Q \VerseOne{1}Here is a message
about Egypt:
\par }{\Q Look,
the {\ND{Lord}}
rides
on
a swift-moving
cloud
\par }{\Q and approaches
Egypt.
\par }{\Q The idols
of Egypt
tremble
before
him;
\par }{\Q the Egyptians
lose their courage.
\par }{\Q \VS{2}“I will provoke
civil strife
in Egypt,
\par }{\Q brothers
will fight
with each other,
\par }{\Q as will neighbors,
\par }{\Q cities,
and kingdoms.
\par }{\Q \VS{3}The Egyptians
will panic,
\par }{\Q and I will confuse
their strategy.
\par }{\Q They will seek guidance
from
the idols
and from
the spirits of the dead,
\par }{\Q from
the pits used to conjure up underworld spirits,
and from
the magicians.
\par }{\Q \VS{4}I will hand
Egypt
over to a harsh
master;
\par }{\Q a powerful
king
will rule
over them,”
\par }{\Q says
the sovereign master,
the {\ND{Lord}}
who commands armies.
\par }{\Q \VS{5}The water
of the sea
will be dried up,
\par }{\Q and the river
will dry up and be empty.
\par }{\Q \VS{6}The canals
will stink;
\par }{\Q the streams
of Egypt
will trickle
and then dry up;
\par }{\Q the bulrushes
and reeds
will decay,
\par }{\Q \VS{7}along with
the plants
by
the mouth
of the river.
\par }{\Q All
the cultivated land
near the river
\par }{\Q will turn to dust
and be blown away.
\par }{\Q \VS{8}The fishermen
will mourn
and lament,
\par }{\Q all
those who cast
a fishhook
into the river,
\par }{\Q and those who spread out
a net
on
the water’s
surface
will grieve.
\par }{\Q \VS{9}Those who make
clothes from combed
flax
will be embarrassed;
\par }{\Q those who weave
will turn pale.
\par }{\Q \VS{10}Those who make cloth
will be
demoralized;
\par }{\Q all
the hired
workers
will be depressed.
\par }{\Q \VS{11}The officials
of Zoan
are nothing but
fools;
\par }{\Q Pharaoh’s
wise
advisers
give stupid
advice.
\par }{\Q How
dare you say
to
Pharaoh,
\par }{\Q “I
am one
of the sages,
\par }{\Q one well-versed in the writings of the ancient
kings?”
\par }{\Q \VS{12}But where,
oh
where, are your wise
men?

\par }{\Q Let them tell
you, let
them find
out
\par }{\Q what
the
{\ND{Lord}}
who commands armies
has planned
for Egypt.
\par }{\Q \VS{13}The officials
of Zoan
are fools,
\par }{\Q the officials
of Memphis
are misled;
\par }{\Q the rulers of her
tribes
lead Egypt
astray.
\par }{\Q \VS{14}The
{\ND{Lord}}
has made them undiscerning;
\par }{\Q they lead Egypt
astray
in all
she does,
\par }{\Q so that she is like a drunk
sliding around
in his own vomit.
\par }{\Q \VS{15}Egypt
will not
be
able
to do
a thing,
\par }{\Q head
or tail,
shoots
and stalk.
\par }{\PP \VS{16}At that
time
the Egyptians
will be like
women.
They will tremble
and fear
because the
{\ND{Lord}}
who commands armies
brandishes
his fist
against them.
\VS{17}The land
of Judah
will humiliate
Egypt.
Everyone
who
hears
about Judah will be
afraid
because of what the
{\ND{Lord}}
who commands armies
is
planning
to do to them.
\par }{\PP \VS{18}At that
time
five
cities
in the land
of Egypt
will speak
the language
of Canaan
and swear
allegiance to the
{\ND{Lord}}
who commands armies.
One
will be called
the City
of the Sun.
\VS{19}At that time
there will be
an altar
for the
{\ND{Lord}}
in the middle
of the land
of Egypt,
as well as a sacred pillar
dedicated
to the
{\ND{Lord}}
at its border.
\VS{20}It will become
a visual
reminder
in the land
of Egypt
of the
{\ND{Lord}}
who commands armies.
When
they cry out
to
the {\ND{Lord}}
because
of oppressors,
he will send
them a deliverer
and defender
who will rescue them.
\VS{21}The
{\ND{Lord}}
will reveal
himself to the Egyptians,
and they will acknowledge
the
{\ND{Lord}}’s
authority
at that time.
They will present
sacrifices
and offerings;
they
will make
vows
to the
{\ND{Lord}}
and fulfill them.
\VS{22}The

{\ND{Lord}}
will strike
Egypt,
striking
and then healing
them. They will turn
to
the {\ND{Lord}}
and he will listen to their prayers
and heal them.
\par }{\PP \VS{23}At that time
there will be
a highway
from Egypt
to Assyria.
The Assyrians
will visit
Egypt,
and the Egyptians
will visit Assyria.
The Egyptians
and Assyrians
will worship together.
\VS{24}At that time
Israel
will be the third
member of the group, along with Egypt
and Assyria,
and will be a recipient
of blessing
in the earth.
\VS{25}The
{\ND{Lord}}
who commands armies
will pronounce a blessing
over the earth, saying, “Blessed
be my people,
Egypt,
and the work
of my hands,
Assyria,
and my special
possession, Israel!”

\par }\Chap{20}{\PP \VerseOne{1}The
{\ND{Lord}} revealed
the following message during the year
in which King
Sargon
of Assyria
sent
his commanding general
to Ashdod,
and he fought
against it and captured it.
\VS{2}At that time
the
{\ND{Lord}}
announced
through
Isaiah
son
of Amoz: “Go,
remove
the sackcloth
from your waist
and take your sandals
off your feet.”
He did
as instructed
and walked
around in undergarments
and barefoot.
\VS{3}Later the
{\ND{Lord}}
explained,
“In the same way that
my servant
Isaiah
has walked around
in undergarments
and barefoot
for the past three
years,
as an object lesson
and omen
pertaining
to Egypt
and Cush,
\VS{4}so
the king
of Assyria
will lead
away the
captives
of Egypt
and the
exiles
of Cush,
both young
and old.
They will
be in undergarments
and barefoot,
with
the buttocks
exposed; the Egyptians
will be publicly humiliated.
\VS{5}Those who put their hope
in Cush
and took pride
in Egypt
will be afraid
and embarrassed.
\VS{6}At that time
those who live
on this
coast
will say,
‘Look
what
has happened to our source of hope
to whom
we fled
for help,
expecting to be rescued
from the king
of Assyria! How
can we
escape now?’ ”

\par }\Chap{21}{\PP \par }{\Q \VerseOne{1}Here is a message
about the Desert
by the Sea:

\par }{\Q Like strong winds
blowing
in the south,
\par }{\Q one invades
from the desert,
\par }{\Q from a land
that is feared.
\par }{\Q \VS{2}I have received
a distressing
message:

\par }{\Q “The deceiver
deceives,
\par }{\Q the destroyer
destroys.
\par }{\Q Attack,
you Elamites!
\par }{\Q Lay siege,
you Medes!
\par }{\Q I will put an end
to all
the groaning!”
\par }{\Q \VS{3}For this reason
my stomach
churns;
\par }{\Q cramps
overwhelm
me
\par }{\Q like the contractions
of a woman in labor.
\par }{\Q I am disturbed
by what I hear,
\par }{\Q horrified
by what I see.
\par }{\Q \VS{4}My
heart
palpitates,
\par }{\Q I shake in fear;
\par }{\Q the twilight
I desired
\par }{\Q has brought
me terror.
\par }{\Q \VS{5}Arrange
the table,
\par }{\Q lay out the carpet,
\par }{\Q eat
and drink!

\par }{\Q Get
up, you officers,
\par }{\Q smear
oil on the shields!
\par }{\Q \VS{6}For
this is what
the sovereign
master has told
me:
\par }{\Q “Go,
post
a guard!
\par }{\Q He must report what
he sees.
\par }{\Q \VS{7}When he sees
chariots,
\par }{\Q teams
of horses,
\par }{\Q riders
on donkeys,
\par }{\Q riders
on camels,
\par }{\Q he must
be alert,
\par }{\Q very
alert.”
\par }{\Q \VS{8}Then the guard
cries out:
\par }{\Q “On
the watchtower,
O sovereign master,
\par }{\Q I
stand
all
day
long;
\par }{\Q at
my post
\par }{\Q I
am stationed
every
night.
\par }{\Q \VS{9}Look
what’s
coming!
\par }{\Q A charioteer,
\par }{\Q a team
of horses.”
\par }{\Q When
questioned, he replies,
\par }{\Q “Babylon
has fallen,
fallen!
\par }{\Q All
the idols
of her gods
lie shattered
on the ground!”
\par }{\Q \VS{10}O my downtrodden
people,
crushed like stalks on the threshing floor,
\par }{\Q what
I have heard
\par }{\Q from the

{\ND{Lord}}
who commands armies,
\par }{\Q the God
of Israel,
\par }{\Q I have reported to you.
\par }{\SH Bad News for Seir
\par }{\Q \VS{11}Here is a message
about Dumah:

\par }{\Q Someone calls
to me
from Seir,
\par }{\Q “Watchman,
what
is left of the night?
\par }{\Q Watchman,
what
is left of the night?”
\par }{\Q \VS{12}The watchman
replies,
\par }{\Q “Morning
is coming,
but then night.
\par }{\Q If
you want to ask,
ask;
\par }{\Q come
back again.”
\par }{\SH The Lord Will Judge Arabia
\par }{\Q \VS{13}Here is a message
about Arabia:
\par }{\Q In the thicket
of Arabia
you spend
the night,
\par }{\Q you Dedanite
caravans.
\par }{\Q \VS{14}Bring out
some water
for the thirsty.
\par }{\Q You who live
in the land
of Tema,
\par }{\Q bring
some food
for the fugitives.
\par }{\Q \VS{15}For
they flee from the swords
–
\par }{\Q from the drawn sword
\par }{\Q and from the battle-ready
bow
\par }{\Q and from the severity
of the battle.
\par }{\PP \VS{16}For
this is what
the sovereign
master has told me: “Within
exactly one year
all
the splendor
of Kedar
will come to an end.
\VS{17}Just a handful
of archers,
the warriors
of Kedar,
will be left.”
Indeed,
the {\ND{Lord}}
God
of Israel
has spoken.

\par }\Chap{22}{\PP \par }{\Q \VerseOne{1}Here is a message
about the Valley
of Vision:

\par }{\Q What
is the reason
\par }{\Q that
all
of you go up
to the rooftops?
\par }{\Q \VS{2}The noisy
city
is full
of raucous
sounds;
\par }{\Q the town
is filled with revelry.
\par }{\Q Your slain
were not
cut down by the sword;
\par }{\Q they did not
die
in battle.
\VS{3}\par }{\Q All
your leaders
ran away
together
–
\par }{\Q they fled
to a distant
place;
\par }{\Q all
your refugees were captured
together
–
\par }{\Q they were captured
without a single arrow being shot.
\par }{\Q \VS{4}So
I say:
\par }{\Q “Don’t look
at me!

\par }{\Q I am weeping
bitterly.
\par }{\Q Don’t
try
to console
me
\par }{\Q concerning
the destruction
of my defenseless
people.”
\par }{\Q \VS{5}For
the sovereign
master, the
{\ND{Lord}}
who commands armies,
\par }{\Q has planned a day
of panic,
defeat,
and confusion.
\par }{\Q In the Valley
of Vision
people shout

\par }{\Q and cry out
to
the hill.
\par }{\Q \VS{6}The Elamites
picked up
the quiver,
\par }{\Q and came with chariots
and horsemen;
\par }{\Q the men
of Kir
prepared
the shield.
\par }{\Q \VS{7}Your very best
valleys
were full
of chariots;
\par }{\Q horsemen
confidently
took their positions
at the gate.
\par }{\Q \VS{8}They removed
the
defenses
of Judah.
\par }{\Q At that time
you looked
\par }{\Q for the weapons
in the House
of the Forest.
\par }{\Q \VS{9}You saw
the many
breaks
\par }{\Q in the walls of the city
of David;
\par }{\Q you stored
up water
in the lower
pool.
\par }{\Q \VS{10}You counted
the houses
in Jerusalem,
\par }{\Q and demolished
houses
so you could have material to reinforce
the wall.
\par }{\Q \VS{11}You made a reservoir
between
the two walls
\par }{\Q for the water
of the old
pool
–
\par }{\Q but you did not
trust in the one who made
it;

\par }{\Q you did not
depend
on
the one who formed
it long ago!
\par }{\Q \VS{12}At that time
the sovereign
master, the
{\ND{Lord}}
who commands armies,
called
for weeping
and mourning,
\par }{\Q for shaved
heads and sackcloth.
\par }{\Q \VS{13}But look,
there is outright celebration!

\par }{\Q You say, “Kill
the ox
and slaughter
the sheep,
\par }{\Q eat
meat
and drink
wine.
\par }{\Q Eat
and drink,
for
tomorrow
we die!”
\par }{\PP \VS{14}The
{\ND{Lord}}
who commands armies
told me this: “Certainly
this
sin
will not
be forgiven
as long
as you live,”
says
the sovereign
master, the
{\ND{Lord}}
who commands armies.
\par }{\Q \VS{15}This is what
the sovereign
master, the
{\ND{Lord}}
who commands armies,
says:
\par }{\Q “Go
visit
this
administrator,
Shebna,
who
supervises
the palace, and tell him:
\par }{\Q \VS{16}‘What
right do you have to be here? What
relatives do you have buried here?

\par }{\Q Why
do you chisel out
a tomb for yourself here?
\par }{\Q He chisels
out his burial site
in an elevated
place,
\par }{\Q he carves
out his tomb
on a cliff.
\par }{\Q \VS{17}Look,
the {\ND{Lord}}
will throw
you far away,
you mere man!

\par }{\Q He will wrap
you up
tightly.
\par }{\Q \VS{18}He will wind
you up
tightly
into a ball
\par }{\Q and throw
you into
a wide,
open land.
\par }{\Q There
you will die,
\par }{\Q and there
with you will be your impressive
chariots,
\par }{\Q which bring disgrace
to the house
of your master.
\par }{\Q \VS{19}I will remove
you from
your office;
\par }{\Q you will be thrown down
from your position.
\par }{\PP \VS{20}“At that time
I will summon
my servant
Eliakim,
son
of Hilkiah.
\VS{21}I will put
your robe
on him, tie
your belt
around him, and transfer
your authority
to him.
He will become
a protector
of the residents
of Jerusalem
and of the people
of Judah.
\VS{22}I will place
the key
to the house
of David
on
his shoulder.
When he opens
the door, no
one can close
it; when he closes
the door,
no
one can open it.
\VS{23}I will fasten
him like a peg
into a solid
place;
he will
bring honor
and respect to his father’s
family.
\VS{24}His father’s
family
will gain increasing
prominence
because of him, including the offspring
and the offshoots.
All
the small
containers,
including the bowls
and all
the jars
will hang from this peg.’
\par }{\PP \VS{25}“At that
time,”
says
the {\ND{Lord}}
who commands armies,
“the peg
fastened
into a solid
place
will come loose. It will be cut off
and fall,
and the load
hanging on
it will be cut off.”
Indeed,
the {\ND{Lord}}
has spoken.

\par }\Chap{23}{\PP \par }{\Q \VerseOne{1}Here is a message
about Tyre:
\par }{\Q Wail,
you large ships,
\par }{\Q for
the port
is too devastated
to enter!

\par }{\Q From the land
of Cyprus
this news is announced to them.
\par }{\Q \VS{2}Lament,
you residents
of the coast,
\par }{\Q you merchants
of Sidon
who travel over
the sea,
\par }{\Q whose agents
sail over
\VS{3}the deep
waters!

\par }{\Q Grain
from the Shihor
region,

\par }{\Q crops
grown near the Nile
she receives;
\par }{\Q she is
the trade center
of the nations.
\par }{\Q \VS{4}Be ashamed,
O Sidon,
\par }{\Q for
the sea
says
this, O fortress
of the sea:
\par }{\Q “I have not
gone into labor
\par }{\Q or
given birth;
\par }{\Q I have not
raised
young men
\par }{\Q or brought up
young women.”
\par }{\Q \VS{5}When the news
reaches Egypt,
\par }{\Q they will be shaken
by what has happened
to Tyre.
\par }{\Q \VS{6}Travel
to Tarshish!
\par }{\Q Wail,
you residents
of the coast!
\par }{\Q \VS{7}Is this
really your boisterous
city

\par }{\Q whose origins
are in the distant past,
\par }{\Q and whose feet
led
her to a distant land
to reside?
\par }{\Q \VS{8}Who
planned
this
for royal
Tyre,
\par }{\Q whose merchants
are princes,
\par }{\Q whose
traders
are the dignitaries
of the earth?
\par }{\Q \VS{9}The
{\ND{Lord}}
who commands armies
planned
it
–
\par }{\Q to dishonor
the pride
that comes from all
her beauty,
\par }{\Q to humiliate
all
the dignitaries
of the earth.
\par }{\Q \VS{10}Daughter
Tarshish,
travel
back to your land,
as one crosses the Nile;
\par }{\Q there is no
longer
any marketplace in Tyre.
\par }{\Q \VS{11}The
{\ND{Lord}}
stretched out
his hand
over
the sea,
\par }{\Q he shook
kingdoms;
\par }{\Q he gave the order
\par }{\Q to
destroy
Canaan’s
fortresses.
\par }{\Q \VS{12}He said,
\par }{\Q “You will no
longer
celebrate,
\par }{\Q oppressed
virgin
daughter
Sidon!
\par }{\Q Get
up, travel
to Cyprus,
\par }{\Q but you will find no
relief
there.”
\par }{\Q \VS{13}Look
at the land
of the Chaldeans,
\par }{\Q these
people
who have lost their identity!

\par }{\Q The Assyrians
have made it a home for wild animals.
\par }{\Q They erected
their siege towers,
\par }{\Q demolished
its fortresses,
\par }{\Q and turned
it into a heap of ruins.
\par }{\Q \VS{14}Wail,
you large ships,
\par }{\Q for
your fortress
is destroyed!
\par }{\PP \VS{15}At that time
Tyre
will be
forgotten
for seventy
years,
the typical
life span of a king.
At the end
of seventy
years
Tyre
will try to attract attention again, like the prostitute
in the popular song:
\par }{\Q \VS{16}“Take
the harp,
\par }{\Q go
through the city,
\par }{\Q forgotten
prostitute!
\par }{\Q Play it well,
\par }{\Q play
lots of songs,
\par }{\Q so you’ll
be noticed!”
\par }{\PP \VS{17}At the end
of seventy
years
the {\ND{Lord}}
will revive
Tyre.
She will start making money again
by selling
her services
to all
the earth’s
kingdoms.
\VS{18}Her profits
and earnings
will be
set apart
for the
{\ND{Lord}}. They will not
be stored up
or
accumulated,
for
her profits
will be given to those who live
in the
{\ND{Lord}}’s
presence
and will be
used to purchase large quantities
of food
and beautiful
clothes.

\par }\Chap{24}{\PP \par }{\Q \VerseOne{1}Look,
the {\ND{Lord}}
is ready to devastate
the earth
\par }{\Q and leave
it in ruins;
\par }{\Q he will mar
its surface
\par }{\Q and scatter
its inhabitants.
\par }{\Q \VS{2}Everyone will suffer – the priest as well as the people,
\par }{\Q the master as well as the servant,
\par }{\Q the elegant lady as well as the female attendant,
\par }{\Q the seller as well as the buyer,
\par }{\Q the borrower as well as the lender,
\par }{\Q the creditor as well as the debtor.
\par }{\Q \VS{3}The earth
will be completely devastated
\par }{\Q and thoroughly
ransacked.
\par }{\Q For
the {\ND{Lord}}
has decreed
this
judgment.
\par }{\Q \VS{4}The earth
dries up
and withers,
\par }{\Q the world
shrivels up
and withers;
\par }{\Q the prominent
people
of the earth
fade away.
\par }{\Q \VS{5}The earth
is defiled
by its inhabitants,
\par }{\Q for
they have violated
laws,
\par }{\Q disregarded
the regulation,
\par }{\Q and broken
the permanent
treaty.
\par }{\Q \VS{6}So
a treaty curse
devours
the earth;
\par }{\Q its inhabitants
pay for their guilt.
\par }{\Q This is why
the inhabitants
of the earth
disappear,
\par }{\Q and are reduced
to just a handful
of people.
\par }{\Q \VS{7}The new wine
dries up,
\par }{\Q the vines
shrivel up,
\par }{\Q all
those who like to celebrate
groan.
\par }{\Q \VS{8}The happy sound
of the tambourines
stops,
\par }{\Q the revelry
of those who celebrate
comes to a halt,
\par }{\Q the happy sound
of the harp
ceases.
\par }{\Q \VS{9}They no
longer sing
and drink
wine;
\par }{\Q the beer
tastes bitter
to those who drink it.
\par }{\Q \VS{10}The ruined
town
is shattered;
\par }{\Q all
of the houses
are shut up
tight.
\par }{\Q \VS{11}They howl
in
the streets
because of what happened to the wine;
\par }{\Q all
joy
turns to sorrow;
\par }{\Q celebrations
disappear
from the earth.
\par }{\Q \VS{12}The city
is left
in ruins;
\par }{\Q the gate
is reduced
to rubble.
\par }{\Q \VS{13}This
is what
will happen
throughout
the earth,
\par }{\Q among
the nations.
\par }{\Q It will be like when they beat
an olive tree,
\par }{\Q and just a few olives
are left
at the end
of the harvest.
\par }{\Q \VS{14}They
lift
their voices
and shout
joyfully;
\par }{\Q they praise
the majesty
of the
{\ND{Lord}}
in the west.
\par }{\Q \VS{15}So
in the east
extol the
{\ND{Lord}},
\par }{\Q along the seacoasts
extol the fame
of the {\ND{Lord}}
God
of Israel.
\par }{\Q \VS{16}From the ends
of the earth
we hear
songs
–
\par }{\Q the Just One
is majestic.
\par }{\Q But I say,
“I’m wasting away! I’m wasting away! I’m doomed!
\par }{\Q Deceivers
deceive,
deceivers thoroughly deceive!”
\par }{\Q \VS{17}Terror,
pit,
and snare
\par }{\Q are ready to overtake you inhabitants
of the earth!
\par }{\Q \VS{18}The one who runs away from
the sound
of the terror
\par }{\Q will fall
into the pit;
\par }{\Q the one who climbs out of the pit,
\par }{\Q will be
trapped by
the snare.
\par }{\Q For
the floodgates
of the heavens
are opened
up

\par }{\Q and the foundations
of the earth
shake.
\par }{\Q \VS{19}The earth
is broken
in pieces,
\par }{\Q the earth
is ripped to shreds,
\par }{\Q the earth
shakes violently.
\par }{\Q \VS{20}The earth
will stagger
around
like a drunk;
\par }{\Q it
will sway back and forth like a hut
in a windstorm.

\par }{\Q Its sin
will weigh it down,
\par }{\Q and it will fall
and never
get
up again.
\par }{\SH The Lord Will Become King
\par }{\Q \VS{21}At that time
the {\ND{Lord}}
will punish
\par }{\Q the heavenly forces
in the heavens
\par }{\Q and the earthly
kings
on
the earth.
\par }{\Q \VS{22}They will be imprisoned
in a pit,
\par }{\Q locked up
in a prison,
\par }{\Q and after staying there for a long
time,
they will be punished.
\par }{\Q \VS{23}The full moon
will be covered up,
\par }{\Q the bright sun
will be darkened;

\par }{\Q for
the {\ND{Lord}}
who commands armies
will rule
\par }{\Q on Mount
Zion
in Jerusalem
\par }{\Q in the presence
of his assembly, in majestic
splendor.

\par }\Chap{25}{\PP \par }{\Q \VerseOne{1}O
{\ND{Lord}}, you are
my God!

\par }{\Q I will exalt
you
in praise,
I will extol your fame.
\par }{\Q For
you have done
extraordinary
things,
\par }{\Q and executed plans
made long ago exactly as you decreed.
\par }{\Q \VS{2}Indeed,
you have made
the city
into a heap
of rubble,
\par }{\Q the fortified
town
into a heap of ruins;
\par }{\Q the fortress
of foreigners
is no longer a
city,
\par }{\Q it will
never
be
rebuilt.
\par }{\Q \VS{3}So
a strong
nation
will extol
you;
\par }{\Q the towns
of powerful nations
will fear you.
\par }{\Q \VS{4}For
you are a protector
for the poor,
\par }{\Q a protector
for the needy
in their distress,
\par }{\Q a shelter
from the rainstorm,
\par }{\Q a shade
from the heat.
\par }{\Q Though
the breath
of tyrants
is like a winter rainstorm,
\par }{\Q \VS{5}like heat
in a dry land,
\par }{\Q you humble
the boasting
foreigners.
\par }{\Q Just as the shadow
of a cloud
causes the heat
to subside,

\par }{\Q so he causes the song
of tyrants
to cease.
\par }{\Q \VS{6}The
{\ND{Lord}}
who commands armies
will hold
a banquet
for all
the nations
on this mountain.
\par }{\Q At this
banquet
there
will be plenty of meat
and aged wine
–
\par }{\Q tender meat
and choicest wine.
\par }{\Q \VS{7}On this
mountain
he will swallow up
\par }{\Q the shroud
that is over
all
the peoples,
\par }{\Q the woven
covering
that is over
all
the nations;
\par }{\Q \VS{8}he will swallow
up death
permanently.
\par }{\Q The sovereign

{\ND{Lord}}
will wipe away
the tears
from every
face,
\par }{\Q and remove
his people’s
disgrace
from all
the earth.
\par }{\Q Indeed,
the {\ND{Lord}}
has announced it!
\par }{\Q \VS{9}At that time
they will say,
\par }{\Q “Look,
here
is our God!
\par }{\Q We waited
for him and he delivered
us.
\par }{\Q Here
is the
{\ND{Lord}}! We waited
for him.
\par }{\Q Let’s
rejoice
and celebrate his deliverance!”
\par }{\Q \VS{10}For
the
{\ND{Lord}}’s
power
will make this
mountain
secure.
\par }{\Q Moab
will be trampled
down where it stands,
\par }{\Q as a heap of straw
is trampled
down in a manure pile.
\par }{\Q \VS{11}Moab will spread
out its hands
in the middle
of it,

\par }{\Q just as
a swimmer
spreads
his hands to swim;
\par }{\Q the
{\ND{Lord}} will bring down
Moab’s pride
as it spreads
its hands.
\par }{\Q \VS{12}The fortified city
(along with the very tops
of your walls) he will knock down,
\par }{\Q he will bring
it down,
he will throw
it down
to
the dusty
ground.

\par }\Chap{26}{\PP \par }{\Q \VerseOne{1}At that time
this
song
will be sung
in the land
of Judah:
\par }{\Q “We have a strong
city!
\par }{\Q The
{\ND{Lord}}’s deliverance,
like walls
and a rampart, makes it secure.
\par }{\Q \VS{2}Open
the gates
so
a righteous
nation
can enter
–
\par }{\Q one that remains
trustworthy.
\par }{\Q \VS{3}You keep completely
safe
the people who
maintain
their faith,
\par }{\Q for
they trust in you.
\par }{\Q \VS{4}Trust
in the
{\ND{Lord}}
from this time forward,
\par }{\Q even in Yah,
the {\ND{Lord}}, an enduring
protector!
\par }{\Q \VS{5}Indeed,
the
{\ND{Lord}} knocks down
those who live in
a high place,
\par }{\Q he brings down
an elevated
town;
\par }{\Q he brings
it down
to
the ground,
\par }{\Q he throws
it down
to
the dust.
\par }{\Q \VS{6}It is trampled
underfoot
\par }{\Q by the feet
of the oppressed,
\par }{\Q by
the soles
of the poor.”
\par }{\SH God’s People Anticipate Vindication
\par }{\Q \VS{7}The way
of the righteous
is level,
\par }{\Q the path
of the righteous
that you make is straight.
\par }{\Q \VS{8}Yes,
as your judgments
unfold,
\par }{\Q O
{\ND{Lord}}, we wait for
you.
\par }{\Q We desire
your fame
and reputation to grow.
\par }{\Q \VS{9}I
look for
you during the night,
\par }{\Q my spirit
within
me seeks
you at dawn,
\par }{\Q for
when your judgments
come upon the earth,
\par }{\Q those who live
in the world
learn
about justice.
\par }{\Q \VS{10}If
the wicked
are shown
mercy,
\par }{\Q they do not
learn
about justice.
\par }{\Q Even
in a land
where right is rewarded,
they act unjustly;
\par }{\Q they do not
see
the
{\ND{Lord}}’s
majesty revealed.
\par }{\Q \VS{11}O
{\ND{Lord}}, you are
ready
to act,

\par }{\Q but they don’t even notice.
\par }{\Q They will see
and be put to shame
by your angry
judgment against humankind,
\par }{\Q yes,
fire
will consume
your enemies.
\par }{\Q \VS{12}O
{\ND{Lord}}, you make us secure,
\par }{\Q for
even
all
we have accomplished,
you have done for us.
\par }{\Q \VS{13}O
{\ND{Lord}}, our God,
\par }{\Q masters
other than
you have ruled
us,
\par }{\Q but we praise
your name
alone.
\par }{\Q \VS{14}The dead
do not
come back to life,
\par }{\Q the spirits of the dead
do not
rise.
\par }{\Q That is because
you came in judgment
and destroyed
them,
\par }{\Q you wiped out
all
memory of them.
\par }{\Q \VS{15}You have made the nation
larger,
O
{\ND{Lord}},
\par }{\Q you have made the nation
larger
and revealed your splendor,

\par }{\Q you have extended
all
the borders
of the land.
\par }{\Q \VS{16}O
{\ND{Lord}}, in distress
they looked for
you;
\par }{\Q they uttered
incantations
because of your discipline.
\par }{\Q \VS{17}As when a pregnant woman
gets ready to deliver
\par }{\Q and strains
and cries
out because of her labor
pains,
\par }{\Q so
were we because
of you, O
{\ND{Lord}}.
\par }{\Q \VS{18}We were pregnant,
we strained,
\par }{\Q we gave birth,
as it were,
to wind.
\par }{\Q We cannot
produce deliverance
on the earth;
\par }{\Q people
to populate
the world
are not
born.
\VS{19}\par }{\Q Your dead
will come back to life;
\par }{\Q your corpses
will rise up.
\par }{\Q Wake up
and shout joyfully,
you who live
in the ground!

\par }{\Q For
you will grow like plants drenched
with the morning
dew,
\par }{\Q and the earth
will bring forth
its dead spirits.
\par }{\Q \VS{20}Go,
my people! Enter
your inner rooms!
\par }{\Q Close
your doors
behind
you!
\par }{\Q Hide
for a little
while,
\par }{\Q until
his angry
judgment is over!
\par }{\Q \VS{21}For
look,
the {\ND{Lord}}
is coming out
of the place
where he lives,

\par }{\Q to punish
the sin
of those who live
on
the earth.
\par }{\Q The earth
will display
the
blood
shed on
it;
\par }{\Q it will no
longer
cover
up its slain.

\par }\Chap{27}{\PP \par }{\Q \VerseOne{1}At that time
the {\ND{Lord}}
will punish
\par }{\Q with his destructive,
great,
and powerful
sword
\par }{\Q Leviathan
the fast-moving
serpent,
\par }{\Q Leviathan
the squirming
serpent;
\par }{\Q he will kill
the sea
monster.
\par }{\Q \VS{2}When that time
comes,

\par }{\Q sing about a delightful
vineyard!
\par }{\Q \VS{3}I,
the {\ND{Lord}}, protect
it;

\par }{\Q I water
it regularly.
\par }{\Q I guard
it night
and day,
\par }{\Q so no one
can harm it.
\par }{\Q \VS{4}I am not
angry.
\par }{\Q I wish
I could confront
some thorns
and briers!
\par }{\Q Then I would march
against them for battle;
\par }{\Q I would set
them all
on fire,
\par }{\Q \VS{5}unless
they became
my subjects
\par }{\Q and made
peace
with me;
\par }{\Q let them make
peace with me.
\par }{\Q \VS{6}The time is coming
when Jacob
will take root;
\par }{\Q Israel
will blossom
and grow branches.
\par }{\Q The produce
will fill
the surface
of the world.
\par }{\Q \VS{7}Has the
{\ND{Lord}}
struck
down Israel like he did their oppressors?

\par }{\Q Has Israel been killed
like their enemies?
\par }{\Q \VS{8}When you summon
her for divorce,
you prosecute
her;

\par }{\Q he drives her away
with
his strong
wind
in the day
of the east
wind.
\par }{\Q \VS{9}So
in this
way Jacob’s
sin
will be forgiven,
\par }{\Q and this
is how they
will show they are finished
sinning:

\par }{\Q They will make
all
the stones
of the altars
\par }{\Q like crushed
limestone,
\par }{\Q and the Asherah poles
and the incense altars
will no
longer stand.
\par }{\Q \VS{10}For
the fortified
city
is left
alone;
\par }{\Q it is a deserted settlement
\par }{\Q and abandoned
like the desert.
\par }{\Q Calves
graze
there;
\par }{\Q they lie
down there
\par }{\Q and eat its branches
bare.
\par }{\Q \VS{11}When its branches
get brittle,
they break;
\par }{\Q women
come
and use them for kindling.
\par }{\Q For
these people
lack understanding,
\par }{\Q therefore
the one who made
them has no
compassion
on them;
\par }{\Q the one who formed
them has no
mercy on them.
\par }{\PP \VS{12}At that time
the
{\ND{Lord}}
will shake
the tree,
from the Euphrates
River to
the Stream
of Egypt.
Then you
will be gathered
up one
by one,
O Israelites.
\VS{13}At that
time
a large
trumpet
will be blown,
and the ones lost
in the land
of Assyria
will come,
as well as the refugees
in the land
of Egypt.
They will worship
the {\ND{Lord}}
on the holy
mountain
in Jerusalem.

\par }\Chap{28}{\PP \par }{\Q \VerseOne{1}The splendid
crown
of Ephraim’s
drunkards
is doomed,
\par }{\Q the withering
flower,
its beautiful
splendor,
\par }{\Q situated
at
the head
of a rich
valley,
\par }{\Q the crown of those overcome
with wine.
\par }{\Q \VS{2}Look,
the sovereign master
sends a strong,
powerful
one.

\par }{\Q With the force of a hailstorm
or a destructive
windstorm,
\par }{\Q with the might of a driving,
torrential
rainstorm,
\par }{\Q he will knock that crown
to the ground
with his hand.
\par }{\Q \VS{3}The splendid
crown
of Ephraim’s
drunkards
\par }{\Q will be trampled
underfoot.
\par }{\Q \VS{4}The withering
flower,
its beautiful
splendor,
\par }{\Q situated at the head
of a rich
valley,
\par }{\Q will be like an early fig
before
harvest
–
\par }{\Q as
soon as someone notices
it,
\par }{\Q he grabs
it and swallows it.
\par }{\Q \VS{5}At that time
the {\ND{Lord}}
who commands armies
will become
a beautiful
crown
\par }{\Q and a splendid
diadem
for the remnant
of his people.
\par }{\Q \VS{6}He will give discernment to the one who
makes judicial decisions,
\par }{\Q and strength to those who defend the city
from attackers.
\par }{\Q \VS{7}Even
these
men stagger
because of wine,
\par }{\Q they stumble around
because of beer
–
\par }{\Q priests
and prophets
stagger
because of beer,
\par }{\Q they are confused
because
of wine,
\par }{\Q they stumble around
because
of beer;
\par }{\Q they stagger
while seeing prophetic visions,
\par }{\Q they totter
while making legal decisions.
\par }{\Q \VS{8}Indeed,
all
the tables
are covered
with vomit;
\par }{\Q no
place
is untouched.
\par }{\Q \VS{9}Who
is the
{\ND{Lord}} trying to teach?
\par }{\Q To whom
is he explaining
a message?

\par }{\Q Those just weaned
from milk!
\par }{\Q Those just taken from their mother’s
breast!
\par }{\Q \VS{10}Indeed,
they will hear meaningless
gibberish,
\par }{\Q senseless
babbling,
\par }{\Q a syllable here,
a syllable there.
\par }{\Q \VS{11}For
with mocking
lips
and a foreign
tongue
\par }{\Q he will speak
to
these
people.
\par }{\Q \VS{12}In the past he said
to
them,

\par }{\Q “This
is where security
can be found.
\par }{\Q Provide security
for the one who is exhausted!
\par }{\Q This
is where rest
can be found.”

\par }{\Q But they refused
to listen.
\par }{\Q \VS{13}So the
{\ND{Lord}}’s
word
to them
will sound like
\par }{\Q meaningless
gibberish,
\par }{\Q senseless
babbling,
\par }{\Q a syllable here,
a syllable there.
\par }{\Q As a result, they will fall
on their backsides
when they try to walk,
\par }{\Q and be injured,
ensnared,
and captured.
\par }{\SH The Lord Will Judge Jerusalem
\par }{\Q \VS{14}Therefore,
listen
to the
{\ND{Lord}}’s
word,
\par }{\Q you who mock,
\par }{\Q you rulers
of these
people
\par }{\Q who
reside in Jerusalem!
\par }{\Q \VS{15}For
you say,
\par }{\Q “We have made
a treaty
with
death,
\par }{\Q with
Sheol
we have made
an agreement.
\par }{\Q When the overwhelming judgment
sweeps by
\par }{\Q it will not
reach
us.
\par }{\Q For
we have made
a lie
our refuge,
\par }{\Q we have hidden
ourselves in a deceitful word.”
\par }{\Q \VS{16}Therefore,
this is what
the sovereign
master, the
{\ND{Lord}},
says:
\par }{\Q “Look,
I am laying a stone
in Zion,
\par }{\Q an approved
stone,
\par }{\Q set in place as a precious
cornerstone
for the foundation.
\par }{\Q The one who maintains
his faith
will not
panic.
\par }{\Q \VS{17}I will make
justice
the measuring line,
\par }{\Q fairness
the plumb
line;
\par }{\Q hail
will sweep away
the unreliable
refuge,
\par }{\Q the floodwaters
will overwhelm
the hiding place.
\par }{\Q \VS{18}Your treaty
with
death
will be dissolved;
\par }{\Q your agreement
with
Sheol
will not
last.
\par }{\Q When the overwhelming
judgment
sweeps
by,

\par }{\Q you will be
overrun by it.
\par }{\Q \VS{19}Whenever
it sweeps
by, it will overtake
you;
\par }{\Q indeed,
every morning
it will sweep by,
\par }{\Q it will come through
during the day
and the night.”
\par }{\Q When
this announcement
is understood,
\par }{\Q it will cause nothing but terror.
\par }{\Q \VS{20}For
the bed
is too short
to stretch out
on,
\par }{\Q and the blanket
is too narrow
to wrap around oneself.
\par }{\Q \VS{21}For
the {\ND{Lord}}
will rise up,
as he did at Mount
Perazim,
\par }{\Q he will rouse
himself, as
he did
in the Valley
of
Gibeon,
\par }{\Q to accomplish
his work,
\par }{\Q his peculiar
work,
\par }{\Q to perform
his task,
\par }{\Q his
strange task.
\par }{\Q \VS{22}So now,
do not
mock,
\par }{\Q or
your chains will become heavier!
\par }{\Q For
I
have heard
a message about decreed
destruction,
\par }{\Q from the sovereign
master, the
{\ND{Lord}}
who commands armies,
against
the entire
land.
\par }{\Q \VS{23}Pay attention
and listen
to my message!

\par }{\Q Be attentive
and listen
to what I have to say!
\par }{\Q \VS{24}Does a farmer
just keep
on plowing
at planting
time?

\par }{\Q Does he keep breaking up and harrowing
his ground?
\par }{\Q \VS{25}Once he has leveled
its surface,
\par }{\Q does
he not scatter
the seed
of the caraway plant,
\par }{\Q sow
the seed of the cumin plant,
\par }{\Q and plant
the wheat,
barley,
and grain
in their designated
places?
\par }{\Q \VS{26}His God
instructs
him;
\par }{\Q he teaches
him the principles of agriculture.
\par }{\Q \VS{27}Certainly
caraway seed
is not
threshed with
a sledge,
\par }{\Q nor is the wheel
of
a cart
rolled over
cumin seed.
\par }{\Q Certainly
caraway seed
is beaten with
a stick,
\par }{\Q and cumin seed
with a flail.
\par }{\Q \VS{28}Grain
is crushed,
\par }{\Q though one certainly
does not
thresh
it forever.
\par }{\Q The wheel
of one’s wagon
rolls over
it,
\par }{\Q but his horses
do not
crush it.
\par }{\Q \VS{29}This
also
comes from
the {\ND{Lord}}
who commands armies,
\par }{\Q who gives
supernatural
guidance
and imparts great
wisdom.

\par }\Chap{29}{\PP \par }{\Q \VerseOne{1}Ariel is as good as dead –
\par }{\Q Ariel, the town David besieged!
\par }{\Q Keep observing your annual rituals,
\par }{\Q celebrate your festivals on schedule.
\par }{\Q \VS{2}I will threaten
Ariel,
\par }{\Q and she will mourn
intensely
\par }{\Q and become
like an altar hearth before me.
\par }{\Q \VS{3}I will lay siege
to you on
all sides;
\par }{\Q I will besiege
you with
troops;

\par }{\Q I will raise
siege works
against you.
\par }{\Q \VS{4}You will fall;
\par }{\Q while lying
on the ground
you will speak;
\par }{\Q from the dust
where you lie,
your words
will be
heard.

\par }{\Q Your voice
will sound like a spirit
speaking from the underworld;
\par }{\Q from the dust
you will chirp
as if muttering an incantation.
\par }{\Q \VS{5}But the horde
of invaders
will be
like fine
dust,
\par }{\Q the horde
of tyrants
like chaff
that is blown away.
\par }{\Q It will happen
suddenly,
in a flash.
\par }{\Q \VS{6}Judgment
will come from
the {\ND{Lord}}
who commands armies,
\par }{\Q accompanied by thunder,
earthquake,
and a loud noise,
\par }{\Q by a strong
gale,
a windstorm,
and a consuming
flame
of fire.
\par }{\Q \VS{7}It will be
like a dream,
a night
vision.
\par }{\Q There will be a horde
from all
the nations
that fight against
Ariel,
\par }{\Q those who
attack
her and her stronghold
and besiege her.
\par }{\Q \VS{8}It will be
like a hungry
man dreaming
that
he is
eating,
\par }{\Q only to awaken
and find that
his
stomach
is
empty.
\par }{\Q It will be like a
thirsty
man dreaming
that
he is
drinking,
\par }{\Q only to awaken
and find
that he is
still weak
and his
thirst unquenched.
\par }{\Q So
it will be
for the horde from
all
the
nations
\par }{\Q that fight against
Mount
Zion.
\par }{\SH God’s People are Spiritually Insensitive
\par }{\Q \VS{9}You will be shocked
and amazed!

\par }{\Q You are totally blind!

\par }{\Q They are drunk,
but not
because of wine;
\par }{\Q they stagger,
but not
because of beer.
\par }{\Q \VS{10}For
the
{\ND{Lord}}
has poured out
on
you
\par }{\Q a strong urge
to sleep deeply.
\par }{\Q He has shut
your eyes
(the prophets),
\par }{\Q and covered
your heads
(the seers).
\par }{\PP \VS{11}To you this entire
prophetic revelation
is like words
in a sealed
scroll.
When they hand
it to
one who
can
read
and say,
“Read
this,”
he responds,
“I can’t,
because
it is
sealed.”
\VS{12}Or when they hand
the scroll
to
one who
can’t
read
and say,
“Read
this,”
he says,
“I can’t
read.”
\par }{\Q \VS{13}The sovereign master
says,
\par }{\Q “These
people
say
they are loyal
to me;

\par }{\Q they say
wonderful things
about me,

\par }{\Q but they are not
really loyal
to me.

\par }{\Q Their worship
consists
of
\par }{\Q nothing but man-made
ritual.
\par }{\Q \VS{14}Therefore
I
will again
do an amazing thing
for
these
people
–
\par }{\Q an absolutely extraordinary deed.
\par }{\Q Wise men
will have nothing to say,
\par }{\Q the sages
will have no
explanations.”
\par }{\Q \VS{15}Those
who try to hide
their plans
from the
{\ND{Lord}}
are as good as dead,

\par }{\Q who do their work
in secret
and boast,
\par }{\Q “Who
sees
us? Who
knows what we’re doing?”
\par }{\Q \VS{16}Your thinking is perverse!

\par }{\Q Should
the potter
be regarded
as clay?

\par }{\Q Should the thing made
say
about its maker,
“He didn’t
make
me”?
\par }{\Q Or should the pottery
say
about the potter,
“He doesn’t
understand”?
\par }{\SH Changes are Coming
\par }{\Q \VS{17}In just a very
short time
\par }{\Q Lebanon
will turn into
an orchard,
\par }{\Q and the orchard
will be considered
a forest.
\par }{\Q \VS{18}At that time
the deaf
will be able to hear
words
read from a scroll,
\par }{\Q and the eyes
of the blind
will be able to see
through deep
darkness.
\par }{\Q \VS{19}The downtrodden
will again
rejoice
in the
{\ND{Lord}};
\par }{\Q the poor
among humankind
will take delight
in the Holy One
of Israel.
\par }{\Q \VS{20}For
tyrants
will disappear,
\par }{\Q those who taunt
will vanish,
\par }{\Q and all
those who love to do
wrong
will be eliminated –
\par }{\Q \VS{21}those
who bear false
testimony
against a person,

\par }{\Q who entrap
the one who arbitrates
at the city gate
\par }{\Q and deprive
the innocent
of justice by making false charges.
\par }{\Q \VS{22}So
this is what
the {\ND{Lord}}, the one who
delivered
Abraham,
says
to
the family
of Jacob:

\par }{\Q “Jacob
will no
longer be ashamed;
\par }{\Q their faces will no
longer show their embarrassment.
\par }{\Q \VS{23}For
when they see
their children,
\par }{\Q whom I will produce
among
them,

\par }{\Q they
will honor
my name.
\par }{\Q They
will honor
the
Holy One
of Jacob;
\par }{\Q they will respect
the God
of Israel.
\par }{\Q \VS{24}Those who stray
morally
will gain
understanding;
\par }{\Q those who complain
will acquire insight.

\par }\Chap{30}{\PP \par }{\Q \VerseOne{1}“The rebellious
children
are as good as dead,”
says
the {\ND{Lord}},
\par }{\Q “those who make
plans
without
consulting me,

\par }{\Q who form
alliances
without
consulting my Spirit,
\par }{\Q and thereby
compound
their sin.
\par }{\Q \VS{2}They travel
down
to Egypt
\par }{\Q without
seeking
my will,
\par }{\Q seeking
Pharaoh’s
protection,
\par }{\Q and looking for safety
in Egypt’s
protective shade.
\par }{\Q \VS{3}But Pharaoh’s
protection
will bring you nothing but shame,
\par }{\Q and the safety of Egypt’s
protective
shade
nothing but humiliation.
\par }{\Q \VS{4}Though
his officials
are in Zoan
\par }{\Q and his messengers
arrive at Hanes,
\par }{\Q \VS{5}all
will be put to shame
\par }{\Q because of a nation
that cannot
help
them,
\par }{\Q who cannot
give
them aid
or
help,
\par }{\Q but
only shame
and disgrace.”
\par }{\Q \VS{6}This is a message
about the animals
in the Negev:
\par }{\Q Through a land
of distress
and danger,
\par }{\Q inhabited by lionesses
and roaring lions,
\par }{\Q by snakes
and darting
adders,
\par }{\Q they
transport
their wealth
on
the backs
of donkeys,
\par }{\Q their riches
on
the humps
of camels,
\par }{\Q to a nation
that cannot
help them.
\par }{\Q \VS{7}Egypt
is totally incapable of helping.
\par }{\Q For this
reason
I call
her
\par }{\Q ‘Proud
one who is silenced.’ ”
\par }{\Q \VS{8}Now
go,
write
it down on
a tablet
in their presence,

\par }{\Q inscribe
it on
a scroll,
\par }{\Q so that it might be preserved
for a future time
\par }{\Q as an enduring
witness.
\par }{\Q \VS{9}For
these
are rebellious
people
–
\par }{\Q they are lying
children,
\par }{\Q children
unwilling
to obey
the
{\ND{Lord}}’s
law.
\par }{\Q \VS{10}They say
to the visionaries,
“See
no
more visions!”
\par }{\Q and to the seers,
“Don’t
relate
messages
to us about what is right!

\par }{\Q Tell
us nice things,
\par }{\Q relate deceptive messages.
\par }{\Q \VS{11}Turn aside
from
the way,
\par }{\Q stray
off the path.
\par }{\Q Remove
from our
presence
the Holy One
of Israel.”
\par }{\Q \VS{12}For this reason
this is what
the Holy One
of Israel
says:
\par }{\Q “You have rejected
this
message;
\par }{\Q you trust
instead in your ability
to oppress and trick,
\par }{\Q and rely
on that kind of behavior.
\par }{\Q \VS{13}So
this
sin
will become
your downfall.
\par }{\Q You will be like a high
wall
\par }{\Q that bulges and cracks
and is ready to collapse;
\par }{\Q it crumbles
suddenly,
in a flash.
\par }{\Q \VS{14}It shatters
in pieces
like a clay
jar,
\par }{\Q so shattered to bits
that none
of it can be salvaged.
\par }{\Q Among its fragments
one cannot
find
a shard
large enough

\par }{\Q to scoop
a hot coal
from a fire
\par }{\Q or to skim off
water
from a cistern.”
\par }{\Q \VS{15}For
this is what
the master,
the
{\ND{Lord}}, the Holy
One of Israel
says:
\par }{\Q “If you repented
and patiently waited
for me, you would be
delivered;
\par }{\Q if you calmly
trusted
in me you would find strength,
\par }{\Q but you are unwilling.
\par }{\Q \VS{16}You say,
‘No,
we
will flee
on
horses,’
\par }{\Q so
you will indeed flee.
\par }{\Q You say, ‘We will ride
on
fast horses,’
\par }{\Q so
your pursuers
will be fast.
\par }{\Q \VS{17}One
thousand
will scurry
at the battle cry of one
enemy soldier;

\par }{\Q at
the battle cry
of five
enemy soldiers you will all run
away,

\par }{\Q until
the remaining
few are as isolated
\par }{\Q as a flagpole
on a mountaintop
\par }{\Q or a signal flag
on
a hill.”
\par }{\SH The Lord Will Not Abandon His People
\par }{\Q \VS{18}For this reason
the {\ND{Lord}}
is ready
to show you mercy;
\par }{\Q he sits on his throne,
ready
to have compassion
on you.

\par }{\Q Indeed,
the {\ND{Lord}}
is a just
God;
\par }{\Q all
who wait
for him in faith will be blessed.
\par }{\Q \VS{19}For
people
will live
in Zion;
\par }{\Q in Jerusalem
you will weep
no
more.

\par }{\Q When he hears your cry
of despair,
he will indeed show
you mercy;
\par }{\Q when
he hears
it, he will respond to you.
\par }{\Q \VS{20}The sovereign
master will give
you distress
to eat
\par }{\Q and suffering to drink;
\par }{\Q but your teachers
will no
longer
be
hidden;
\par }{\Q your eyes
will see them.
\par }{\Q \VS{21}You will hear
a word
spoken behind
you, saying,
\par }{\Q “This
is the correct way,
walk in
it,”
\par }{\Q whether you are heading to the right
or the left.
\par }{\Q \VS{22}You will desecrate
your silver-plated
idols
\par }{\Q and your gold-plated
images.
\par }{\Q You will throw them away as
if they were a menstrual rag,
\par }{\Q saying
to them, “Get out!”
\par }{\Q \VS{23}He will water
the seed
you plant
in the ground,
\par }{\Q and the ground
will
produce crops
in abundance.
\par }{\Q At that time
your cattle
will graze
in wide
pastures.
\par }{\Q \VS{24}The oxen
and donkeys
used in plowing
\par }{\Q will eat
seasoned
feed
winnowed
with
a shovel
and pitchfork.
\par }{\Q \VS{25}On
every
high
mountain
\par }{\Q and every
high
hill
\par }{\Q there will be streams
flowing
with water,
\par }{\Q at the time
of great
slaughter
when
the fortified towers
collapse.
\par }{\Q \VS{26}The light
of the full moon
will be
like the sun’s
glare
\par }{\Q and the sun’s
glare
will be
seven times
brighter,
\par }{\Q like the light
of seven
days,
\par }{\Q when
the {\ND{Lord}}
binds up
his people’s
fractured
bones

\par }{\Q and heals
their severe wound.
\par }{\Q \VS{27}Look,
the name
of the {\ND{Lord}}
comes
from a distant place
\par }{\Q in raging
anger
and awesome
splendor.
\par }{\Q He speaks
angrily
\par }{\Q and his word
is like destructive
fire.
\par }{\Q \VS{28}His battle cry
overwhelms like a flooding
river
\par }{\Q that reaches
one’s neck.
\par }{\Q He shakes
the nations
in a sieve
that isolates
the chaff;
\par }{\Q he puts a bit
into
the mouth
of the nations
and leads
them to destruction.
\par }{\Q \VS{29}You will sing
\par }{\Q as you do in the evening
when
you are celebrating
a festival.
\par }{\Q You will be happy
like one who plays
a flute
\par }{\Q as he goes
to
the mountain
of the {\ND{Lord}}, the Rock
who shelters Israel.
\par }{\Q \VS{30}The
{\ND{Lord}}
will give a mighty
shout
\par }{\Q and intervene
in power,
\par }{\Q with furious
anger
and flaming,
destructive
fire,
\par }{\Q with a driving
rainstorm
and hailstones.
\par }{\Q \VS{31}Indeed,
the
{\ND{Lord}}’s
shout
will shatter
Assyria;
\par }{\Q he will beat
them with a club.
\par }{\Q \VS{32}Every
blow
from his punishing
cudgel,
\par }{\Q with which
the {\ND{Lord}}
will beat
them,

\par }{\Q will be accompanied by music from
the tambourine
and harp,
\par }{\Q and he will attack
them with his weapons.
\par }{\Q \VS{33}For
the burial place
is already prepared;
\par }{\Q it has been made
deep
and wide
for the king.
\par }{\Q The firewood
is piled
high
on it.

\par }{\Q The
{\ND{Lord}}’s
breath,
like a stream
flowing with brimstone,
\par }{\Q will ignite it.

\par }\Chap{31}{\PP \par }{\Q \VerseOne{1}Those
who go down
to Egypt
for help
are as good as dead,

\par }{\Q those who rely
on war horses,
\par }{\Q and trust
in Egypt’s many chariots
\par }{\Q and in their many,
many horsemen.
\par }{\Q But
they do not
rely
on
the Holy One
of Israel
\par }{\Q and do not
seek
help from the
{\ND{Lord}}.
\par }{\Q \VS{2}Yet he too
is wise
and he
will bring
disaster;
\par }{\Q he does not
retract his decree.
\par }{\Q He will attack
the wicked
nation,
\par }{\Q and the nation that helps
those who commit sin.
\par }{\Q \VS{3}The Egyptians
are mere humans,
not
God;
\par }{\Q their horses
are made of flesh,
not
spirit.
\par }{\Q The
{\ND{Lord}}
will strike
with his hand;
\par }{\Q the one who helps
will stumble
\par }{\Q and the one being helped
will fall.
\par }{\Q Together
they will perish.
\par }{\SH The Lord Will Defend Zion
\par }{\Q \VS{4}Indeed,
this is what
the {\ND{Lord}}
says
to me:
\par }{\Q “The
{\ND{Lord}}
will be like a growling
lion,
\par }{\Q like a young lion
growling
over
its prey.
\par }{\Q Though a whole group
of shepherds
gathers against it,
\par }{\Q it is not afraid
of their shouts
\par }{\Q or
intimidated
by
their yelling.
\par }{\Q In this same
way the
{\ND{Lord}}
who commands armies
will descend
\par }{\Q to do battle on
Mount
Zion
and on
its hill.
\par }{\Q \VS{5}Just as birds
hover
over a nest,

\par }{\Q so the
{\ND{Lord}}
who commands armies
will protect Jerusalem.
\par }{\Q He will protect
and deliver
it;
\par }{\Q as he passes over he
will rescue it.
\par }{\PP \VS{6}You Israelites! Return
to the one against whom
you have so blatantly
rebelled!
\VS{7}For
at that
time
everyone
will get rid
of the silver
and gold
idols
your hands
sinfully
made.
\par }{\Q \VS{8}Assyria
will fall
by a sword,
but not
one human-made;
\par }{\Q a sword
not
made by humankind
will destroy
them.

\par }{\Q They will run away
from
this sword
\par }{\Q and their young men
will be
forced to do hard labor.
\par }{\Q \VS{9}They will surrender
their stronghold
because of fear;
\par }{\Q their officers
will be afraid
of the
{\ND{Lord}}’s battle flag.”
\par }{\Q This is what
the
{\ND{Lord}}
says
–
\par }{\Q the one whose fire
is in Zion,
\par }{\Q whose firepot
is in Jerusalem.

\par }\Chap{32}{\PP \par }{\Q \VerseOne{1}Look,
a king
will promote
fairness;
\par }{\Q officials
will promote
justice.
\par }{\Q \VS{2}Each
of them will be
like a shelter
from the wind
\par }{\Q and a refuge
from a rainstorm;
\par }{\Q like streams
of water
in a dry region
\par }{\Q and like the shade
of a large
cliff
in a parched
land.
\par }{\Q \VS{3}Eyes
will no longer be blind
\par }{\Q and ears
will be attentive.
\par }{\Q \VS{4}The mind
that acts rashly
will possess discernment
\par }{\Q and the tongue
that stutters
will speak
with ease and clarity.
\par }{\Q \VS{5}A fool
will no
longer be called
honorable;
\par }{\Q a deceiver
will no
longer
be called
principled.
\par }{\Q \VS{6}For
a fool
speaks
disgraceful things;
\par }{\Q his mind
plans out
sinful deeds.
\par }{\Q He commits
godless deeds
\par }{\Q and says
misleading things
about
the
{\ND{Lord}};
\par }{\Q he gives
the hungry
nothing
to satisfy their appetite
\par }{\Q and gives
the thirsty
nothing
to drink.
\par }{\Q \VS{7}A deceiver’s
methods
are evil;
\par }{\Q he
dreams up evil
plans

\par }{\Q to ruin
the poor with
lies,
\par }{\Q even when the needy
are in the right.
\par }{\Q \VS{8}An honorable
man makes honorable
plans;
\par }{\Q his honorable
character gives him security.
\par }{\SH The Lord Will Give True Security
\par }{\Q \VS{9}You complacent
women,
\par }{\Q get
up and listen
to me!
\par }{\Q You carefree
daughters,
\par }{\Q pay attention
to what I say!
\par }{\Q \VS{10}In
a year’s
time
\par }{\Q you carefree ones
will shake
with fear,
\par }{\Q for
the grape harvest
will fail,
\par }{\Q and the fruit harvest
will not
arrive.
\par }{\Q \VS{11}Tremble,
you complacent
ones!
\par }{\Q Shake
with fear, you carefree
ones!
\par }{\Q Strip
off your clothes and expose yourselves –
\par }{\Q put sackcloth on your waist!
\par }{\Q \VS{12}Mourn
over
the field,
\par }{\Q over
the delightful
fields
\par }{\Q and the fruitful
vine!
\par }{\Q \VS{13}Mourn over
the land
of my people,
\par }{\Q which is overgrown
with thorns
and briers,
\par }{\Q and over
all
the once-happy
houses
\par }{\Q in the city
filled with revelry.
\par }{\Q \VS{14}For
the fortress
is neglected;
\par }{\Q the once-crowded
city
is abandoned.
\par }{\Q Hill
and watchtower
\par }{\Q are permanently
uninhabited.
\par }{\Q Wild donkeys
love
to go there,
\par }{\Q and flocks
graze there.
\par }{\Q \VS{15}This desolation will continue until
new life
is poured
out on
us from heaven.
\par }{\Q Then the desert
will become
an orchard
\par }{\Q and the orchard
will be considered
a forest.
\par }{\Q \VS{16}Justice
will settle
down in the desert
\par }{\Q and fairness
will live
in the orchard.
\par }{\Q \VS{17}Fairness
will produce
peace
\par }{\Q and result
in lasting
security.
\par }{\Q \VS{18}My people
will live
in peaceful
settlements,
\par }{\Q in secure
homes,
\par }{\Q and in safe, quiet
places.
\par }{\Q \VS{19}Even if the forest
is destroyed
\par }{\Q and the city
is annihilated,
\par }{\Q \VS{20}you will be blessed,
\par }{\Q you who plant
seed by
all
the banks
of the streams,
\par }{\Q you who let
your
ox
and donkey graze.

\par }\Chap{33}{\PP \par }{\Q \VerseOne{1}The destroyer
is as good as dead,

\par }{\Q you
who have not
been destroyed!
\par }{\Q The deceitful
one is as good as dead,

\par }{\Q the one whom others have not
deceived!
\par }{\Q When you are through
destroying,
you will be destroyed;
\par }{\Q when you finish
deceiving,
others will deceive you!
\par }{\Q \VS{2}{\ND{Lord}}, be merciful
to us! We wait
for you.
\par }{\Q Give us strength
each morning!

\par }{\Q Deliver
us when
distress comes.
\par }{\Q \VS{3}The nations
run away
when they hear a loud noise;
\par }{\Q the nations scatter when you spring into action!
\par }{\Q \VS{4}Your plunder
disappears
as if locusts
were eating
it;

\par }{\Q they swarm over
it like locusts!
\par }{\Q \VS{5}The
{\ND{Lord}}
is exalted,
\par }{\Q indeed,
he lives
in heaven;
\par }{\Q he fills
Zion
with justice
and fairness.
\par }{\Q \VS{6}He is your constant source of stability;
\par }{\Q he abundantly
provides safety
and great wisdom;
\par }{\Q he gives all this
to those who fear him.
\par }{\Q \VS{7}Look,
ambassadors
cry out
in the streets;
\par }{\Q messengers
sent to make peace
weep
bitterly.
\par }{\Q \VS{8}Highways
are empty,

\par }{\Q there are no travelers.
\par }{\Q Treaties
are broken,
\par }{\Q witnesses are despised,
\par }{\Q human life is treated with disrespect.
\par }{\Q \VS{9}The land
dries up
and withers away;
\par }{\Q the forest of Lebanon
shrivels up
and decays.
\par }{\Q Sharon
is like
the desert;
\par }{\Q Bashan
and Carmel
are parched.
\par }{\Q \VS{10}“Now
I will rise
up,” says
the {\ND{Lord}}.
\par }{\Q “Now
I will exalt
myself;
\par }{\Q now
I will magnify myself.
\par }{\Q \VS{11}You conceive
straw,
\par }{\Q you give birth
to chaff;
\par }{\Q your breath
is a fire
that destroys you.
\par }{\Q \VS{12}The nations
will be
burned
to ashes;
\par }{\Q like thorn bushes
that have been cut down,
they will be set
on fire.
\par }{\Q \VS{13}You who are far
away, listen
to what
I have
done!
\par }{\Q You
who are close
by, recognize
my strength!”
\par }{\Q \VS{14}Sinners
are afraid
in Zion;
\par }{\Q panic
grips
the godless.
\par }{\Q They say, ‘Who
among us can coexist
with destructive
fire?
\par }{\Q Who
among us can coexist
with unquenchable
fire?’
\par }{\Q \VS{15}The one who lives
uprightly
\par }{\Q and speaks
honestly;
\par }{\Q the one who refuses
to profit
from oppressive
measures
\par }{\Q and rejects
a bribe;
\par }{\Q the one who does not
plot
violent crimes
\par }{\Q and does not
seek
to harm others –
\par }{\Q \VS{16}This is the person who will live
in a secure place;
\par }{\Q he will find safety
in the rocky,
mountain strongholds;
\par }{\Q he will have
food
\par }{\Q and a constant
supply of water.
\par }{\Q \VS{17}You will see
a king
in his splendor;
\par }{\Q you will see
a wide
land.
\par }{\Q \VS{18}Your mind
will recall
the terror
you experienced,

\par }{\Q and you will ask yourselves, “Where
is the scribe?
\par }{\Q Where is
the one who weighs
the money?
\par }{\Q Where is
the one
who counts
the
towers?”
\par }{\Q \VS{19}You
will no longer see
a defiant
people
\par }{\Q whose language
you do not comprehend,
\par }{\Q whose derisive
speech
you do not
understand.
\par }{\Q \VS{20}Look
at Zion,
the city
where we hold religious festivals!
\par }{\Q You will see
Jerusalem,
\par }{\Q a peaceful settlement,
\par }{\Q a tent
that stays put;

\par }{\Q its stakes
will never
be pulled up;
\par }{\Q none
of its ropes
will snap
in two.
\par }{\Q \VS{21}Instead
the {\ND{Lord}}
will rule there
as our mighty
king.

\par }{\Q Rivers
and wide
streams
will flow through
it;

\par }{\Q no
war
galley
will enter;
\par }{\Q no
large
ships
will sail through.
\par }{\Q \VS{22}For
the {\ND{Lord}}, our ruler,
\par }{\Q the {\ND{Lord}}, our commander,
\par }{\Q the {\ND{Lord}}, our king
–
\par }{\Q he will deliver us.
\par }{\Q \VS{23}Though at this time your ropes
are slack,
\par }{\Q the mast
is not
secured,
\par }{\Q and the sail
is not
unfurled,
\par }{\Q at that time
you will divide up
a great quantity
of loot;
\par }{\Q even the lame
will drag off
plunder.
\par }{\Q \VS{24}No resident
of Zion
will say,
“I am ill”;
\par }{\Q the people
who live
there will have their sin
forgiven.

\par }\Chap{34}{\PP \par }{\Q \VerseOne{1}Come near,
you nations,
and listen!
\par }{\Q Pay attention,
you people!
\par }{\Q The earth
and everything
it contains
must listen,
\par }{\Q the world
and everything that lives in it.
\par }{\Q \VS{2}For
the {\ND{Lord}}
is angry
at all
the nations
\par }{\Q and furious
with all
their armies.
\par }{\Q He will annihilate
them and slaughter them.
\par }{\Q \VS{3}Their slain
will be left
unburied,
\par }{\Q their corpses
will stink;
\par }{\Q the hills
will soak up
their blood.
\par }{\Q \VS{4}All
the stars
in the sky
will fade away,
\par }{\Q the sky
will roll up
like a scroll;
\par }{\Q all
its stars
will wither,
\par }{\Q like a leaf
withers
and falls from a vine
\par }{\Q or a fig
withers and falls from a tree.
\par }{\Q \VS{5}He says, “Indeed,
my sword
has slaughtered heavenly
powers.

\par }{\Q Look,
it now descends
on
Edom,
\par }{\Q on
the people
I will annihilate
in judgment.”
\par }{\Q \VS{6}The
{\ND{Lord}}’s
sword
is dripping with blood,
\par }{\Q it is covered with fat;
\par }{\Q it drips with the blood of young rams and goats
\par }{\Q and is covered with the fat of rams’ kidneys.
\par }{\Q For the
{\ND{Lord}} is holding a sacrifice in Bozrah,
\par }{\Q a bloody slaughter in the land of Edom.
\par }{\Q \VS{7}Wild oxen
will be slaughtered
along with
them,
\par }{\Q as well as strong
bulls.
\par }{\Q Their land
is drenched
with blood,
\par }{\Q their soil
is covered
with fat.
\par }{\Q \VS{8}For
the {\ND{Lord}}
has planned a day
of revenge,
\par }{\Q a time
when he will repay
Edom for her hostility
toward Zion.
\par }{\Q \VS{9}Edom’s
streams
will be
turned into
pitch
\par }{\Q and her soil
into brimstone;
\par }{\Q her land
will become burning
pitch.
\par }{\Q \VS{10}Night
and day
it will burn;
\par }{\Q its smoke
will
ascend
continually.
\par }{\Q Generation
after generation
it will be a wasteland
\par }{\Q and no
one will ever
pass through it again.
\par }{\Q \VS{11}Owls
and wild animals
will live
there,

\par }{\Q all kinds of wild birds
will settle in
it.
\par }{\Q The
{\ND{Lord}} will stretch out
over
her
\par }{\Q the measuring line
of ruin
\par }{\Q and the plumb line
of destruction.
\par }{\Q \VS{12}Her nobles
will have nothing
left
to call
a kingdom
\par }{\Q and all
her officials
will disappear.
\par }{\Q \VS{13}Her fortresses
will be overgrown
with thorns;
\par }{\Q thickets
and weeds
will grow in her fortified cities.
\par }{\Q Jackals
will settle
there;
\par }{\Q ostriches will live there.
\par }{\Q \VS{14}Wild animals
and wild dogs
will congregate
there;

\par }{\Q wild goats
will bleat
to
one another.
\par }{\Q Yes,
nocturnal animals
will rest
there
\par }{\Q and make
for themselves a nest.
\par }{\Q \VS{15}Owls
will make nests
and lay eggs
there;
\par }{\Q they will hatch
them and protect
them.

\par }{\Q Yes, hawks
will gather
there,
\par }{\Q each with its mate.
\par }{\Q \VS{16}Carefully
read
the scroll
of the {\ND{Lord}}!

\par }{\Q Not
one
of these creatures
will be missing,
\par }{\Q none
will lack a mate.
\par }{\Q For
the
{\ND{Lord}}
has issued
the decree,
\par }{\Q and his own spirit
gathers them.
\par }{\Q \VS{17}He assigns
them
their allotment;
\par }{\Q he measures out
their assigned
place.

\par }{\Q They
will live
there permanently;
\par }{\Q they will settle
in it through successive generations.

\par }\Chap{35}{\PP \par }{\Q \VerseOne{1}Let the desert
and dry
region be happy;
\par }{\Q let the wilderness
rejoice
and bloom
like a lily!
\par }{\Q \VS{2}Let
it richly
bloom;
\par }{\Q let it rejoice
and shout
with delight!

\par }{\Q It is given
the grandeur
of Lebanon,
\par }{\Q the splendor
of Carmel
and Sharon.
\par }{\Q They
will see
the grandeur
of the {\ND{Lord}},
\par }{\Q the splendor
of our God.
\par }{\Q \VS{3}Strengthen
the hands
that have gone limp,
\par }{\Q steady
the knees
that shake!
\par }{\Q \VS{4}Tell
those who panic,
\par }{\Q “Be strong! Do not
fear!
\par }{\Q Look,
your God
comes
to avenge!
\par }{\Q With divine
retribution
he comes
to deliver you.”
\par }{\Q \VS{5}Then
blind
eyes
will open,
\par }{\Q deaf
ears will hear.
\par }{\Q \VS{6}Then
the lame
will leap
like a deer,
\par }{\Q the mute
tongue
will shout
for joy;
\par }{\Q for
water
will flow
in the desert,
\par }{\Q streams
in the wilderness.
\par }{\Q \VS{7}The dry soil
will become
a pool
of water,
\par }{\Q the parched ground
springs
of water.
\par }{\Q Where jackals
once lived and sprawled out,
\par }{\Q grass,
reeds,
and papyrus will grow.
\par }{\Q \VS{8}A thoroughfare
will be there
–
\par }{\Q it will be called
the Way
of Holiness.
\par }{\Q The unclean
will not
travel
on
it;
\par }{\Q it is reserved for those authorized to use it –
\par }{\Q fools will not stray into it.
\par }{\Q \VS{9}No
lions
will be
there,
\par }{\Q no ferocious
wild animals will be on it –
\par }{\Q they will not be found there.
\par }{\Q Those delivered from bondage will travel on it,
\par }{\Q \VS{10}those whom the
{\ND{Lord}}
has ransomed
will return
that way.

\par }{\Q They will enter
Zion
with a happy
shout.
\par }{\Q Unending
joy will crown
them,

\par }{\Q happiness
and joy
will overwhelm
them;
\par }{\Q grief
and suffering
will disappear.

\par }\Chap{36}{\PP \VerseOne{1}In the fourteenth
year
of King
Hezekiah’s
reign, King
Sennacherib
of Assyria
marched up
against
all
the fortified
cities
of Judah
and captured them.
\VS{2}The king
of Assyria
sent
his chief adviser
from Lachish
to
King
Hezekiah
in Jerusalem,
along with a large
army.
The chief adviser stood
at the conduit
of the upper
pool
which is located on the road
to the field
where they wash and dry cloth.
\VS{3}Eliakim
son
of Hilkiah,
the palace
supervisor,
accompanied by Shebna
the scribe
and Joah
son
of Asaph,
the secretary,
went out
to meet
him.
\par }{\PP \VS{4}The chief adviser
said
to
them, “Tell
Hezekiah: ‘This is what
the great
king,
the king
of Assyria,
says: “What
is your source of confidence?
\VS{5}Your claim
to have a strategy
and military
strength
is just empty talk.
In
whom
are you trusting,
that
you would dare to rebel against me?
\VS{6}Look,
you must be trusting
in
Egypt,
that splintered
reed
staff.
If someone
leans
on
it for support, it punctures
his hand
and wounds
him. That is what
Pharaoh
king
of Egypt
does to all
who trust in him!
\VS{7}Perhaps
you will tell
me,
‘We are trusting
in the
{\ND{Lord}}
our God.’
But
Hezekiah
is
the one
who
eliminated
his high places
and altars
and then told
the people of Judah
and Jerusalem,
‘You must worship
at this
altar.’
\VS{8}Now
make a deal
with
my master
the king
of Assyria,
and I will give
you two thousand
horses,
provided you can
find enough riders for them.
\VS{9}Certainly
you will not refuse
one
of my master’s
minor
officials
and trust
in
Egypt
for chariots
and horsemen.
\VS{10}Furthermore
it was by the command of the
{\ND{Lord}}
that I marched up
against
this
land
to destroy
it. The
{\ND{Lord}}
told
me, ‘March up
against
this
land
and destroy it!’ ” ’ ”
\par }{\PP \VS{11}Eliakim,
Shebna,
and Joah
said
to
the chief adviser,
“Speak
to
your servants
in Aramaic,
for
we
understand
it. Don’t
speak
with us in
the Judahite dialect
in the hearing
of the people
who
are on
the wall.”
\VS{12}But the chief adviser
said,
“My master
did not
send
me to
speak
these
words
only to your master
and to you. His message is also for the men
who sit
on
the wall,
for they will eat
their own excrement
and drink
their own urine
along with you!”
\par }{\PP \VS{13}The chief adviser
then stood
there and called out
loudly
in the Judahite dialect, “Listen
to the message
of the great
king,
the king
of Assyria.
\VS{14}This is what
the king
says: ‘Don’t
let Hezekiah
mislead
you, for
he is not
able
to rescue you!
\VS{15}Don’t
let Hezekiah
talk you into
trusting
in the
{\ND{Lord}}
by saying,
“The
{\ND{Lord}}
will certainly rescue
us; this
city
will not
be handed
over to the king
of Assyria.”
\VS{16}Don’t
listen
to
Hezekiah!’ For
this is what
the king
of Assyria
says,
‘Send
me a token of your submission
and surrender
to
me. Then each
of you may eat
from his own
vine
and fig tree
and drink
water
from his own
cistern,
\VS{17}until
I come
and take
you to
a land just like your own – a land of grain and new wine, a land of bread and vineyards.
\VS{18}Hezekiah
is misleading
you when he says,
“The

{\ND{Lord}}
will rescue
us.”
Has any
of the gods
of the nations
rescued
his land
from the power
of the king
of Assyria?
\VS{19}Where
are the gods
of Hamath
and Arpad? Where
are the gods
of Sepharvaim? Indeed,
did any gods rescue
Samaria
from my power?
\VS{20}Who
among all
the gods
of these
lands
have
rescued
their lands
from my power? So how can the
{\ND{Lord}}
rescue
Jerusalem
from my power?’ ”
\VS{21}They were silent
and did not
respond,
for
the king
had ordered,
“Don’t
respond to him.”
\par }{\PP \VS{22}Eliakim
son
of Hilkiah,
the palace supervisor,
accompanied
by Shebna
the scribe
and Joah
son
of Asaph,
the secretary,
went to
Hezekiah
with their clothes
torn
in grief and reported
to him what
the chief adviser had said.

\par }\Chap{37}{\PP \VerseOne{1}When
King
Hezekiah
heard
this, he tore
his clothes,
put on sackcloth,
and went
to the
{\ND{Lord}}’s
temple.
\VS{2}Eliakim
the palace
supervisor,
Shebna
the scribe,
and the leading
priests,
clothed in sackcloth,
sent
this message to
the prophet
Isaiah
son
of Amoz:
\VS{3}“This is what
Hezekiah
says: ‘This is a day
of distress,
insults,
and humiliation,
as when
a baby
is ready to leave
the birth
canal,
but the mother lacks
the strength to push it through.
\VS{4}Perhaps
the
{\ND{Lord}}
your God
will hear
all these things
the chief adviser
has spoken on
behalf of his master,
the king
of Assyria,
who sent
him to taunt
the living
God.
When the
{\ND{Lord}}
your God
hears,
perhaps
he will punish
him for the things
he has said. So
pray
for this remnant
that remains.’ ”
\par }{\PP \VS{5}When King
Hezekiah’s
servants
came
to
Isaiah,
\VS{6}Isaiah
said
to
them, “Tell
your master
this: ‘This is what
the {\ND{Lord}}
says: “Don’t
be afraid
because
of the things
you have heard
– these insults
the king
of Assyria’s
servants have hurled against me.
\VS{7}Look,
I will
take control
of his mind; he will receive
a report
and return
to
his own land.
I will cut him down
with a sword
in his own land.” ’ ”
\par }{\PP \VS{8}When the chief adviser
heard
the king
of Assyria
had departed from Lachish,
he left and went to Libnah, where the king was campaigning.
\VS{9}The king heard
that King
Tirhakah
of Ethiopia
was marching out
to fight
him.
He again sent
messengers
to
Hezekiah,
ordering them:
\VS{10}“Tell
King
Hezekiah
of Judah
this: ‘Don’t
let your God
in whom
you
trust
mislead
you when he says,
“Jerusalem
will not
be handed
over to the king
of Assyria.”
\VS{11}Certainly
you
have heard
how
the kings
of Assyria
have annihilated
all
lands.
Do you really
think you will be rescued?
\VS{12}Were the nations
whom
my predecessors
destroyed
– the nations of Gozan,
Haran,
Rezeph,
and the people
of Eden
in Telassar
– rescued
by their gods?
\VS{13}Where
are the king
of Hamath,
the king
of Arpad,
and the kings
of Lair, Sepharvaim,
Hena,
and Ivvah?’ ”
\par }{\PP \VS{14}Hezekiah
took
the letter
from
the messengers
and read
it. Then Hezekiah
went up
to the
{\ND{Lord}}’s
temple
and spread
it out before
the {\ND{Lord}}.
\VS{15}Hezekiah
prayed
before
the {\ND{Lord}}:
\VS{16}“O
{\ND{Lord}}
who commands armies,
O God
of Israel,
who
is enthroned
on the cherubim! You
alone
are God
over all
the kingdoms
of the earth.
You
made
the sky
and the earth.
\VS{17}Pay
attention,

{\ND{Lord}}, and hear! Open
your eyes,

{\ND{Lord}}, and observe! Listen
to this
entire
message
Sennacherib
sent
and how he taunts
the living
God!
\VS{18}It is true,

{\ND{Lord}}, that the kings
of Assyria
have destroyed
all
the nations
and their lands.
\VS{19}They
have burned
the gods
of the nations,
for
they
are not
really gods,
but only
the product
of human
hands
manufactured from wood
and stone.
That
is why the Assyrians could destroy
them.
\VS{20}Now,
O
{\ND{Lord}}
our God,
rescue
us from his power,
so all
the kingdoms
of the earth
may know
that
you
alone
are the
{\ND{Lord}}.”
\par }{\PP \VS{21}Isaiah
son
of Amoz
sent
this message to
Hezekiah: “This is what
the {\ND{Lord}}
God
of Israel
says: ‘Because
you prayed
to me
concerning King
Sennacherib
of Assyria,
\VS{22}this
is what
the {\ND{Lord}}
says
about him:

\par }{\Q “The virgin
daughter
Zion
\par }{\Q despises you – she makes fun of you;
\par }{\Q daughter Jerusalem
\par }{\Q shakes her head after you.
\par }{\Q \VS{23}Whom
have you taunted
and hurled insults
at?
\par }{\Q At whom
have you shouted
\par }{\Q and looked
so arrogantly?

\par }{\Q At the Holy One
of Israel!
\par }{\Q \VS{24}Through
your messengers
you taunted
the sovereign master,
\par }{\Q ‘With my many
chariots
I
climbed up
\par }{\Q the high
mountains,
\par }{\Q the slopes
of Lebanon.
\par }{\Q I cut down
its tall
cedars
\par }{\Q and its best
evergreens.
\par }{\Q I invaded
its most remote regions,
\par }{\Q its thickest
woods.
\par }{\Q \VS{25}I
dug wells
\par }{\Q and drank
water.
\par }{\Q With the soles
of my feet I dried up
\par }{\Q all
the rivers
of Egypt.’
\VS{26}\par }{\Q Certainly
you must have heard!

\par }{\Q Long
ago
I worked
it out,
\par }{\Q in ancient
times I planned
it,
\par }{\Q and now
I am bringing
it to pass.
\par }{\Q The plan is this:
\par }{\Q Fortified
cities
will crash
\par }{\Q into heaps
of ruins.
\par }{\Q \VS{27}Their residents
are powerless;
\par }{\Q they are terrified
and ashamed.
\par }{\Q They are as short-lived as plants
in the field
\par }{\Q or green
vegetation.
\par }{\Q They are as short-lived as grass
on the rooftops
\par }{\Q when it is scorched
by
the east wind.
\par }{\Q \VS{28}I know
where you live
\par }{\Q and everything you do
\par }{\Q and how you rage
against me.
\par }{\Q \VS{29}Because
you rage
against me
\par }{\Q and the uproar
you create
has reached my ears,
\par }{\Q I
will put
my hook
in your nose,
\par }{\Q and my bridle
between your lips,
\par }{\Q and I will lead you back
\par }{\Q the way
you came.”
\VS{30}\par }{\PP “This
will be your reminder
that I have spoken the truth: This year
you will eat
what grows
wild, and next
year
what grows on its own.
But the year
after
that you will plant seed
and harvest
crops; you will plant
vines
and consume
their produce.
\VS{31}Those who remain
in Judah
will take root
in the ground
and bear
fruit.
\par }{\Q \VS{32}“For
a remnant
will leave Jerusalem;
\par }{\Q survivors will
come out
of Mount
Zion.
\par }{\Q The intense
devotion of the
{\ND{Lord}}
who commands armies
will accomplish
this.
\par }{\Q \VS{33}So
this is what
the {\ND{Lord}}
says about the king
of Assyria:
\par }{\Q ‘He will not
enter
this
city,
\par }{\Q nor
will he shoot
an
arrow
here.

\par }{\Q He will not
attack
it with his shielded
warriors,

\par }{\Q nor
will he build siege
works against it.
\par }{\Q \VS{34}He will go back
the way
he came
–
\par }{\Q he will not
enter
this
city,’
says
the {\ND{Lord}}.
\par }{\Q \VS{35}I will shield
this
city
and rescue
it for
the sake
of my reputation
and because of my promise to David
my servant.” ’ ”
\par }{\PP \VS{36}The
{\ND{Lord}}’s
messenger
went out
and killed
185,000
troops in the Assyrian
camp.
When they got up early
the next morning,
there
were all
the corpses!
\VS{37}So King
Sennacherib
of Assyria
broke
camp and went
on his way. He went home
and stayed
in Nineveh.
\VS{38}One day, as he was worshiping
in the temple
of his god
Nisroch,
his sons
Adrammelech
and Sharezer
struck
him down
with the sword.
They
ran
away to the land
of Ararat;
his son
Esarhaddon
replaced
him as king.

\par }\Chap{38}{\PP \VerseOne{1}In those
days
Hezekiah
was stricken
with a terminal illness.
The prophet
Isaiah
son
of Amoz
visited
him and told
him, “This is what
the {\ND{Lord}}
says,
‘Give instructions
to your household,
for
you are about to die;
you
will not
get well.’ ”
\VS{2}Hezekiah
turned
his face
to
the wall
and prayed
to
the {\ND{Lord}},
\VS{3}“Please,

{\ND{Lord}}. Remember
how I have served you faithfully
and with wholehearted
devotion,
and how I have carried
out your will.” Then Hezekiah
wept bitterly.
\par }{\PP \VS{4}The
{\ND{Lord}}
told
Isaiah,
\VS{5}“Go
and tell
Hezekiah: ‘This is what
the
{\ND{Lord}}
God
of your ancestor
David
says: “I have heard
your prayer;
I have seen
your tears.
Look,
I will add
fifteen
years
to your life,
\VS{6}and rescue
you and this
city
from the king
of Assyria.
I will shield
this
city.” ’ ”
\VS{7}Isaiah replied, “This
is your sign
from the
{\ND{Lord}}
confirming that
the {\ND{Lord}}
will do
what
he has
said:
\VS{8}Look,
I will make
the shadow
go
back
ten
steps
on the stairs
of Ahaz.”
And then
the shadow went back
ten
steps.
\par }{\SH Hezekiah’s Song of Thanks
\par }{\PP \VS{9}This is the prayer
of King
Hezekiah
of Judah
when he was sick
and then recovered
from his illness:
\par }{\Q \VS{10}“I
thought,
\par }{\Q ‘In the middle
of my life
I must walk
through the gates
of Sheol,
\par }{\Q I am deprived
of the rest
of my years.’
\par }{\Q \VS{11}“I thought,
\par }{\Q ‘I will no
longer see
the
{\ND{Lord}}
in the land
of the living,
\par }{\Q I will no
longer
look
on humankind
with
the inhabitants
of the world.
\par }{\Q \VS{12}My dwelling place
is removed
and taken away
from
me
\par }{\Q like a shepherd’s
tent.
\par }{\Q I rolled up
my life
like a weaver
rolls cloth;

\par }{\Q from the loom
he cuts me off.
\par }{\Q You turn day
into night
and end my life.
\par }{\Q \VS{13}I cry
out until
morning;
\par }{\Q like a lion
he shatters
all
my bones;
\par }{\Q you turn day
into night
and end my life.
\par }{\Q \VS{14}Like
a swallow
or a thrush
I chirp,
\par }{\Q I coo
like a dove;
\par }{\Q my eyes
grow tired
from looking up to the sky.
\par }{\Q O sovereign master,
I am oppressed;
\par }{\Q help me!
\par }{\Q \VS{15}What
can I say?
\par }{\Q He has
decreed
and acted.
\par }{\Q I will walk slowly
all
my years
because I am overcome
with grief.
\par }{\Q \VS{16}O sovereign
master, your decrees can give men life;
\par }{\Q may years of life
be restored
to me.

\par }{\Q Restore my health
and preserve my life.’
\par }{\Q \VS{17}“Look,
the grief
I experienced
was for my benefit.
\par }{\Q You
delivered
me
from the pit
of oblivion.
\par }{\Q For
you removed
all
my sins
from
your sight.
\par }{\Q \VS{18}Indeed
Sheol
does not
give you thanks;
\par }{\Q death
does not praise
you.
\par }{\Q Those who descend
into
the pit
do not
anticipate
your faithfulness.
\par }{\Q \VS{19}The living
person, the living
person, he gives you thanks,
\par }{\Q as
I do
today.
\par }{\Q A father
tells
his sons
about
your faithfulness.
\par }{\Q \VS{20}The
{\ND{Lord}}
is about to deliver
me,

\par }{\Q and we will celebrate
with music
\par }{\Q for the rest
of our
lives
in the
{\ND{Lord}}’s
temple.”
\VS{21}Isaiah
ordered,
“Let them take
a fig
cake
and apply
it to
the ulcerated sore
and he will get well.”
\VS{22}Hezekiah
said,
“What
is the confirming sign
that
I will go up
to the
{\ND{Lord}}’s
temple?”

\par }\Chap{39}{\PP \VerseOne{1}At that time
Merodach-Baladan
son
of Baladan,
king
of Babylon,
sent letters
and a gift
to
Hezekiah,
for he heard
that
Hezekiah had been ill
and had recovered.
\VS{2}Hezekiah
welcomed
them and showed
them his storehouse
with its
silver,
gold,
spices,
and high-quality olive oil,
as well
as
his whole
armory
and everything
in his treasuries.
Hezekiah
showed
them everything
in his palace
and in his whole
kingdom.
\VS{3}Isaiah
the prophet
visited
King
Hezekiah
and asked
him,
“What
did these
men
say? Where
do they come
from?” Hezekiah
replied,
“They come
from the distant
land
of Babylon.”
\VS{4}Isaiah asked,
“What
have they seen
in your palace?” Hezekiah
replied,
“They have
seen
everything
in my palace.
I showed
them everything
in my treasuries.”
\VS{5}Isaiah
said
to
Hezekiah,
“Listen
to the word
of the {\ND{Lord}}
who commands armies:
\VS{6}‘Look,
a time
is coming
when everything
in your palace
and the things your ancestors
have accumulated
to this
day
will be carried
away to Babylon;
nothing
will be left,’
says
the {\ND{Lord}}.
\VS{7}‘Some
of your very own descendants
whom
you father
will be taken
away and will be
made eunuchs
in the palace
of the king
of Babylon.’ ”
\VS{8}Hezekiah
said
to
Isaiah,
“The
{\ND{Lord}}’s
word
which
you have announced
is appropriate.”
Then he thought, “For
there will be
peace
and stability
during my lifetime.”

\par }\Chap{40}{\PP \par }{\Q \VerseOne{1}“Comfort,
comfort
my people,”
\par }{\Q says
your God.
\par }{\Q \VS{2}“Speak
kindly
to Jerusalem,
and tell her
\par }{\Q that
her time of warfare
is over,

\par }{\Q that
her
punishment
is completed.
\par }{\Q For
the
{\ND{Lord}}
has made her pay double
for all
her sins.”
\par }{\Q \VS{3}A voice
cries out,
\par }{\Q “In the wilderness
clear
a way
for the
{\ND{Lord}};
\par }{\Q construct
in the desert
a road
for our God.
\par }{\Q \VS{4}Every
valley
must be elevated,
\par }{\Q and every
mountain
and hill
leveled.
\par }{\Q The rough terrain
will become
a level plain,
\par }{\Q the rugged landscape
a wide valley.
\par }{\Q \VS{5}The splendor
of the {\ND{Lord}}
will be revealed,
\par }{\Q and all
people
will see
it at the same time.
\par }{\Q For
the {\ND{Lord}}
has decreed it.”
\par }{\Q \VS{6}A voice
says,
“Cry
out!”
\par }{\Q Another asks, “What
should I cry
out?”
\par }{\Q The first voice responds: “All
people
are like grass,
\par }{\Q and all
their promises
are like the flowers
in the field.
\par }{\Q \VS{7}The grass
dries up,
\par }{\Q the flowers
wither,
\par }{\Q when
the wind
sent by the
{\ND{Lord}}
blows
on them.
\par }{\Q Surely
humanity is like grass.
\par }{\Q \VS{8}The grass
dries up,
\par }{\Q the flowers
wither,
\par }{\Q but the decree
of our God
is forever
reliable.”
\par }{\Q \VS{9}Go up
on
a high
mountain,
O herald
Zion!
\par }{\Q Shout out
loudly,
O herald
Jerusalem!

\par }{\Q Shout, don’t
be afraid!
\par }{\Q Say
to the towns
of Judah,
\par }{\Q “Here
is your God!”
\par }{\Q \VS{10}Look,
the sovereign

{\ND{Lord}}
comes
as a victorious
warrior;

\par }{\Q his military power
establishes his rule.
\par }{\Q Look,
his reward
is with
him;
\par }{\Q his prize
goes before him.
\par }{\Q \VS{11}Like a shepherd
he tends his flock;
\par }{\Q he gathers up
the lambs
with his arm;
\par }{\Q he carries
them close to his heart;

\par }{\Q he leads
the ewes along.
\par }{\SH The Lord is Incomparable
\par }{\Q \VS{12}Who
has measured
out the waters
in the hollow of his hand,
\par }{\Q or carefully measured
the sky,
\par }{\Q or carefully weighed
the soil
of the earth,
\par }{\Q or weighed
the mountains
in a balance,
\par }{\Q or the hills
on scales?
\par }{\Q \VS{13}Who
comprehends
the mind
of the {\ND{Lord}},
\par }{\Q or gives him
instruction as his counselor?
\par }{\Q \VS{14}From whom
does he receive directions?

\par }{\Q Who teaches
him the correct
way
to do things,

\par }{\Q or imparts
knowledge
to him,
\par }{\Q or instructs
him in skillful
design?
\par }{\Q \VS{15}Look,
the nations
are like a drop
in a bucket;
\par }{\Q they are regarded
as dust
on the scales.
\par }{\Q He lifts
the coastlands
as
if they were dust.
\par }{\Q \VS{16}Not
even Lebanon
could supply
enough
firewood
for a sacrifice;

\par }{\Q its
wild animals
would not
provide enough
burnt offerings.
\par }{\Q \VS{17}All
the nations
are insignificant
before
him;
\par }{\Q they are regarded as
absolutely
nothing.
\par }{\Q \VS{18}To
whom
can you compare
God?
\par }{\Q To what
image
can you liken him?
\par }{\Q \VS{19}A craftsman
casts
an idol;
\par }{\Q a metalsmith
overlays
it with gold
\par }{\Q and forges
silver
chains for it.
\par }{\Q \VS{20}To make a contribution
one selects
wood
that will not
rot;
\par }{\Q he then seeks
a skilled
craftsman
\par }{\Q to make
an idol
that will not
fall over.
\par }{\Q \VS{21}Do you not
know?
\par }{\Q Do you not
hear?
\par }{\Q Has it not
been told
to you since the very beginning?
\par }{\Q Have you not
understood
from the time the earth’s
foundations were made?
\par }{\Q \VS{22}He is the one who sits
on
the earth’s
horizon;
\par }{\Q its inhabitants
are like grasshoppers
before him.

\par }{\Q He
is the one who stretches out
the sky
like a thin curtain,
\par }{\Q and spreads
it out like a pitched
tent.
\par }{\Q \VS{23}He is
the one
who reduces
rulers
to nothing;
\par }{\Q he makes the earth’s
leaders insignificant.
\par }{\Q \VS{24}Indeed,
they are barely
planted;
\par }{\Q yes,
they are barely
sown;
\par }{\Q yes,
they barely
take root
in the earth,
\par }{\Q and then
he blows on
them, causing them to dry up,
\par }{\Q and the wind
carries
them away
like straw.
\par }{\Q \VS{25}“To whom
can you compare
me? Whom do I resemble?”
\par }{\Q says
the Holy One.
\par }{\Q \VS{26}Look up
at the sky!

\par }{\Q Who
created
all these
heavenly lights?

\par }{\Q He is the one who leads out
their ranks;
\par }{\Q he calls
them all
by name.
\par }{\Q Because of his absolute
power
and awesome
strength,
\par }{\Q not
one
of them is missing.
\par }{\Q \VS{27}Why
do you say,
Jacob,
\par }{\Q Why do you say,
Israel,
\par }{\Q “The
{\ND{Lord}}
is not aware
of what is happening
to me,

\par }{\Q My God
is not concerned
with my vindication”?
\par }{\Q \VS{28}Do you not
know?
\par }{\Q Have you not
heard?
\par }{\Q The
{\ND{Lord}}
is an eternal
God,
\par }{\Q the creator
of the whole earth.
\par }{\Q He does not
get tired
or
weary;
\par }{\Q there is no
limit
to his wisdom.
\par }{\Q \VS{29}He gives
strength
to those who are tired;
\par }{\Q to the ones who lack
power,
he gives renewed
energy.
\par }{\Q \VS{30}Even youths
get tired
and weary;
\par }{\Q even strong young men
clumsily
stumble.
\par }{\Q \VS{31}But those who wait
for the
{\ND{Lord}}’s
help find renewed
strength;
\par }{\Q they rise up
as if they had eagles’
wings,
\par }{\Q they run
without
growing weary,
\par }{\Q they walk
without
getting tired.

\par }\Chap{41}{\PP \par }{\Q \VerseOne{1}“Listen
to me
in silence,
you coastlands!

\par }{\Q Let the nations
find renewed
strength!
\par }{\Q Let them approach
and then
speak;
\par }{\Q let us come together
for debate!
\par }{\Q \VS{2}Who
stirs
up this one from the east?

\par }{\Q Who officially
commissions
him for service?

\par }{\Q He hands nations
over
to him,

\par }{\Q and enables him to subdue
kings.
\par }{\Q He
makes
them like dust
with his sword,
\par }{\Q like windblown
straw
with his bow.
\par }{\Q \VS{3}He pursues
them and passes by
unharmed;
\par }{\Q he advances
with great speed.
\par }{\Q \VS{4}Who
acts and carries
out decrees?

\par }{\Q Who summons
the successive
generations
from the beginning?
\par }{\Q I,
the {\ND{Lord}}, am present at the very beginning,
\par }{\Q and at the very end
– I am the one.
\par }{\Q \VS{5}The coastlands
see
and are afraid;
\par }{\Q the whole
earth
trembles;
\par }{\Q they approach
and come.
\par }{\Q \VS{6}They help
one
another;
\par }{\Q one says
to the other,
‘Be strong!’
\par }{\Q \VS{7}The craftsman
encourages
the metalsmith,
\par }{\Q the one who wields
the hammer
encourages the one who pounds
on the anvil.
\par }{\Q He approves
the quality
of the welding,
\par }{\Q and nails
it down
so it won’t
fall over.”
\par }{\SH The Lord Encourages His People
\par }{\Q \VS{8}“You,
my servant
Israel,
\par }{\Q Jacob
whom
I have chosen,
\par }{\Q offspring
of Abraham
my friend,
\par }{\Q \VS{9}you whom
I am bringing back
from the earth’s
extremities,
\par }{\Q and have summoned
from the remote regions
–
\par }{\Q I told
you, “You
are my servant.”
\par }{\Q I have chosen
you and not
rejected you.
\par }{\Q \VS{10}Don’t
be afraid,
for
I
am with you!
\par }{\Q Don’t
be frightened,
for I
am your God!

\par }{\Q I strengthen you –
\par }{\Q yes, I help you –
\par }{\Q yes, I uphold you with my saving right hand!
\par }{\Q \VS{11}Look,
all
who were angry
at you will be
ashamed
and humiliated;
\par }{\Q your adversaries
will be reduced to nothing
and perish.
\par }{\Q \VS{12}When you will look for
your opponents,
you will not
find
them;
\par }{\Q your enemies
will be
reduced to absolutely
nothing.
\par }{\Q \VS{13}For
I am
the
{\ND{Lord}}
your God,
\par }{\Q the one who takes hold
of your right hand,
\par }{\Q who says
to you, ‘Don’t
be afraid,
I am
helping you.’
\par }{\Q \VS{14}Don’t
be afraid,
despised insignificant
Jacob,
\par }{\Q men
of Israel.
\par }{\Q I
am helping
you,” says
the {\ND{Lord}},
\par }{\Q your protector,
the Holy One
of Israel.
\par }{\Q \VS{15}“Look,
I am making
you like a sharp
threshing sledge,
\par }{\Q new
and double-edged.
\par }{\Q You will thresh
the mountains
and crush
them;
\par }{\Q you will make
the hills
like straw.
\par }{\Q \VS{16}You will winnow
them and the wind
will blow them away;
\par }{\Q the wind
will scatter
them.
\par }{\Q You
will rejoice
in the
{\ND{Lord}};
\par }{\Q you will boast
in the Holy One
of Israel.
\par }{\Q \VS{17}The oppressed
and the poor
look
for water,
but there is none;
\par }{\Q their tongues
are parched
from thirst.
\par }{\Q I,
the {\ND{Lord}}, will respond
to their prayers;

\par }{\Q I, the God
of Israel,
will not
abandon them.
\par }{\Q \VS{18}I will make
streams
flow
down
the slopes
\par }{\Q and produce springs
in the middle
of the valleys.
\par }{\Q I will turn
the desert
into a pool
of water
\par }{\Q and the arid
land
into springs.
\par }{\Q \VS{19}I will make
cedars,
acacias,
myrtles,
and olive
trees
grow in the wilderness;
\par }{\Q I will make
evergreens,
firs,
and cypresses
grow together
in the desert.
\par }{\Q \VS{20}I
will do this so
people will observe
and recognize,
\par }{\Q so
they will pay attention
and understand
\par }{\Q that
the
{\ND{Lord}}’s
power
has accomplished
this,
\par }{\Q and that the Holy One
of Israel
has brought it into being.”
\par }{\SH The Lord Challenges the Pagan Gods
\par }{\Q \VS{21}“Present
your argument,”
says
the {\ND{Lord}}.
\par }{\Q “Produce your evidence,”
says
Jacob’s
king.
\par }{\Q \VS{22}“Let them produce evidence! Let them tell
us what
will happen!
\par }{\Q Tell us about your earlier
predictive oracles,
\par }{\Q so we may examine
them and see how
they were fulfilled.
\par }{\Q Or
decree
for us some
future
events!
\par }{\Q \VS{23}Predict
how future events will turn out,
\par }{\Q so we might know
you
are gods.
\par }{\Q Yes, do something good
or bad,
\par }{\Q so we might be frightened
and in awe.
\par }{\Q \VS{24}Look,
you
are nothing,
and your accomplishments
are nonexistent;
\par }{\Q the one who chooses
to worship you is disgusting.
\par }{\Q \VS{25}I have stirred up
one out of the north
and he advances,
\par }{\Q one from the eastern
horizon
who prays
in my name.
\par }{\Q He steps on
rulers
as if
they were clay,
\par }{\Q like
a potter
treading
the clay.
\par }{\Q \VS{26}Who
decreed
this from the beginning,
so we could know?
\par }{\Q Who announced it ahead
of time, so we could say,
‘He’s
correct’?
\par }{\Q Indeed,
none
of them decreed it!
\par }{\Q Indeed,
none
of them announced
it!
\par }{\Q Indeed,
no
one heard
you say anything!
\par }{\Q \VS{27}I first
decreed to Zion,
‘Look,
here’s
what will happen!’

\par }{\Q I
sent a herald
to Jerusalem.
\par }{\Q \VS{28}I look,
but there is no
one,
\par }{\Q among them there is no
one
who serves
as an adviser,
\par }{\Q that I might ask
questions
and receive answers.
\par }{\Q \VS{29}Look,
all
of them are nothing,
\par }{\Q their accomplishments
are nonexistent;
\par }{\Q their metal images
lack any real substance.

\par }\Chap{42}{\PP \par }{\Q \VerseOne{1}“Here
is my servant
whom I support,
\par }{\Q my chosen one
in whom I take pleasure.
\par }{\Q I have placed
my spirit
on
him;
\par }{\Q he will make
just decrees
for the nations.
\par }{\Q \VS{2}He will not
cry out
or shout;
\par }{\Q he will not publicize
himself in the streets.
\par }{\Q \VS{3}A crushed
reed
he will not
break,
\par }{\Q a dim wick
he will not
extinguish;
\par }{\Q he will faithfully
make just decrees.
\par }{\Q \VS{4}He will not
grow dim
or
be crushed
\par }{\Q before establishing
justice
on the earth;
\par }{\Q the coastlands
will wait in anticipation for his decrees.”
\par }{\Q \VS{5}This is what
the true God,
the {\ND{Lord}}, says –
\par }{\Q the one who created the sky and stretched it out,
\par }{\Q the one who fashioned the earth and everything that lives on it,
\par }{\Q the one who gives breath to the people on it,
\par }{\Q and life to those who live on it:
\par }{\Q \VS{6}“I,
the {\ND{Lord}}, officially
commission
you;

\par }{\Q I take hold
of your hand.
\par }{\Q I protect
you and make
you a covenant
mediator for people,
\par }{\Q and a light
to the nations,
\par }{\Q \VS{7}to open
blind
eyes,
\par }{\Q to release
prisoners
from dungeons,
\par }{\Q those who live
in darkness
from prisons.
\par }{\SH The Lord Intervenes
\par }{\Q \VS{8}I am
the {\ND{Lord}}! That is
my name!
\par }{\Q I will not share my glory
with anyone else,
\par }{\Q or the praise
due me with idols.
\par }{\Q \VS{9}Look,
my earlier
predictive oracles have come
to pass;

\par }{\Q now I
announce
new
events.
\par }{\Q Before
they begin to occur,
\par }{\Q I reveal them to you.”
\par }{\Q \VS{10}Sing
to the
{\ND{Lord}}
a brand new
song!
\par }{\Q Praise
him from the horizon
of the earth,
\par }{\Q you who go down
to the sea,
and everything
that lives in it,

\par }{\Q you coastlands
and those who live there!
\par }{\Q \VS{11}Let the desert
and its cities
shout out,
\par }{\Q the towns
where the nomads of Kedar
live!
\par }{\Q Let the residents
of Sela
shout joyfully;
\par }{\Q let them shout loudly
from the mountaintops.
\par }{\Q \VS{12}Let them give
the {\ND{Lord}}
the honor
he deserves;

\par }{\Q let them praise
his deeds in the coastlands.
\par }{\Q \VS{13}The
{\ND{Lord}}
emerges like
a hero,
\par }{\Q like
a warrior
he inspires himself for battle;
\par }{\Q he shouts, yes,
he yells,
\par }{\Q he shows his enemies
his power.
\par }{\Q \VS{14}“I have been inactive
for a long time;
\par }{\Q I kept quiet
and held back.
\par }{\Q Like a woman in labor
I groan;
\par }{\Q I pant
and gasp.
\par }{\Q \VS{15}I will make the trees on the mountains
and hills
wither
up;

\par }{\Q I will dry up
all
their vegetation.
\par }{\Q I will turn
streams
into islands,
\par }{\Q and dry up
pools of water.
\par }{\Q \VS{16}I will lead
the blind
along an unfamiliar
way;
\par }{\Q I will guide
them down paths
they have never
traveled.
\par }{\Q I will turn
the darkness
in front
of them into light,
\par }{\Q and level
out the rough ground.
\par }{\Q This
is what
I will do
for them.
\par }{\Q I will not
abandon them.
\par }{\Q \VS{17}Those who trust
in idols
\par }{\Q will turn
back
and be utterly humiliated,
\par }{\Q those who say
to metal images,
‘You
are our gods.’ ”
\par }{\SH The Lord Reasons with His People
\par }{\Q \VS{18}“Listen,
you deaf
ones!
\par }{\Q Take notice,
you blind ones!
\par }{\Q \VS{19}My servant
is truly
blind,
\par }{\Q my messenger
is truly
deaf.
\par }{\Q My covenant partner, the servant
of the {\ND{Lord}}, is truly
blind.
\par }{\Q \VS{20}You see
many
things, but don’t
comprehend;
\par }{\Q their ears
are open,
but do not
hear.”
\par }{\Q \VS{21}The
{\ND{Lord}}
wanted
to exhibit
his justice
\par }{\Q by magnifying
his law
and displaying it.
\par }{\Q \VS{22}But
these people
are looted
and plundered;
\par }{\Q all
of them are trapped
in pits
\par }{\Q and held captive
in prisons.
\par }{\Q They were carried away
as loot
with no one
to rescue
them;
\par }{\Q they were carried away as plunder,
and no one
says,
“Bring
that back!”
\par }{\Q \VS{23}Who
among you will pay attention
to this?
\par }{\Q Who will listen
attentively
in the future?
\par }{\Q \VS{24}Who
handed
Jacob
over to the robber?
\par }{\Q Who handed Israel
over to the looters?

\par }{\Q Was it not
the
{\ND{Lord}}, against whom
we sinned?
\par }{\Q They refused
to follow
his commands;
\par }{\Q they disobeyed
his law.
\par }{\Q \VS{25}So he poured
out his fierce
anger
on
them,
\par }{\Q along with the devastation of war.
\par }{\Q Its flames
encircled
them, but they did not
realize
it;

\par }{\Q it burned
against
them, but they did notice.


\par }\Chap{43}{\PP \par }{\Q \VerseOne{1}Now,
this is what
the {\ND{Lord}}
says,
\par }{\Q the one who created
you, O Jacob,
\par }{\Q and formed
you, O Israel:
\par }{\Q “Don’t
be afraid,
for
I will protect
you.
\par }{\Q I call
you by name,
you are mine.
\par }{\Q \VS{2}When
you pass
through the waters,
I am
with you;
\par }{\Q when you pass through the streams,
they will not
overwhelm you.
\par }{\Q When you walk through
the fire,
you
will not
be burned;
\par }{\Q the flames
will not
harm you.
\par }{\Q \VS{3}For
I am
the
{\ND{Lord}}
your God,
\par }{\Q the Holy One
of Israel,
your deliverer.
\par }{\Q I
have handed
over Egypt
as a ransom
price,
\par }{\Q Ethiopia
and Seba
in place of you.
\par }{\Q \VS{4}Since you are precious
and special
in my sight,
\par }{\Q and I
love
you,
\par }{\Q I will hand
over people
in place
of you,
\par }{\Q nations
in place
of your life.
\par }{\Q \VS{5}Don’t
be afraid,
for
I am
with you.
\par }{\Q From the east
I will bring
your descendants;
\par }{\Q from the west
I will gather you.
\par }{\Q \VS{6}I will say
to the north,
‘Hand
them over!’
\par }{\Q and to the south,
‘Don’t
hold
any back!’
\par }{\Q Bring
my sons
from distant
lands,
\par }{\Q and my daughters
from the remote
regions of the earth,
\par }{\Q \VS{7}everyone
who belongs
to me,
\par }{\Q whom I created
for my
glory,
\par }{\Q whom I formed
– yes,
whom I made!
\par }{\SH The Lord Declares His Sovereignty
\par }{\Q \VS{8}Bring out
the people
who are blind,
even though they have eyes,
\par }{\Q those who are deaf,
even though they have ears!
\par }{\Q \VS{9}All
nations
gather
together,
\par }{\Q the peoples
assemble.
\par }{\Q Who
among them announced
this?
\par }{\Q Who predicted
earlier
events for us?
\par }{\Q Let them produce
their witnesses
to testify they were right;
\par }{\Q let them listen
and affirm,
‘It is true.’
\par }{\Q \VS{10}You
are my witnesses,”
says
the {\ND{Lord}},
\par }{\Q “my servant
whom
I have chosen,
\par }{\Q so that
you may consider
and believe
in me,
\par }{\Q and understand
that
I
am he.
\par }{\Q No
god
was formed
before
me,
\par }{\Q and none
will outlive me.
\par }{\Q \VS{11}I,
I am
the {\ND{Lord}},
\par }{\Q and there is no
deliverer
besides me.
\par }{\Q \VS{12}I
decreed
and delivered
and proclaimed,
\par }{\Q and there was no
other
god among you.
\par }{\Q You
are my witnesses,”
says
the {\ND{Lord}}, “that I am
God.
\par }{\Q \VS{13}From this day
forward

\par }\Chap{44}{\PP \par }{\Q \VerseOne{1}“Now,
listen,
Jacob
my servant,
\par }{\Q Israel
whom I have chosen!”
\par }{\Q \VS{2}This is what
the {\ND{Lord}}, the one who made you, says –
\par }{\Q the one who formed you in the womb and helps you:
\par }{\Q “Don’t be afraid, my servant Jacob,
\par }{\Q Jeshurun, whom I have chosen!
\par }{\Q \VS{3}For
I will pour
water
on
the parched
ground

\par }{\Q and cause streams
to flow on
the dry land.
\par }{\Q I will pour
my spirit
on
your offspring
\par }{\Q and my blessing
on
your children.
\par }{\Q \VS{4}They will sprout
up like a tree in the grass,
\par }{\Q like poplars
beside
channels
of water.
\par }{\Q \VS{5}One
will say,
‘I belong
to the
{\ND{Lord}},’
\par }{\Q and another
will use the name
‘Jacob.’
\par }{\Q One
will write
on his hand,
‘The
{\ND{Lord}}’s,’
\par }{\Q and use
the name
‘Israel.’ ”
\par }{\SH The Absurdity of Idolatry
\par }{\Q \VS{6}This is what
the {\ND{Lord}}, Israel’s
king,
says,
\par }{\Q their protector,
the {\ND{Lord}}
who commands armies:
\par }{\Q “I am
the first
and I am
the last,
\par }{\Q there is no
God
but me.
\par }{\Q \VS{7}Who
is like
me? Let him make his claim!

\par }{\Q Let him announce
it and explain it to me –
\par }{\Q since I established an ancient people –

\par }{\Q let them announce future events!
\par }{\Q \VS{8}Don’t
panic! Don’t
be afraid!

\par }{\Q Did I
not
tell
you beforehand
and decree it?
\par }{\Q You
are my witnesses! Is there
any God
but
me?
\par }{\Q There is no
other sheltering
rock;
I know of none.
\par }{\Q \VS{9}All
who form
idols
are nothing;
\par }{\Q the things in which they delight
are
worthless.
\par }{\Q Their witnesses
cannot
see;
\par }{\Q they
recognize
nothing, so
they are put to shame.
\par }{\Q \VS{10}Who
forms
a god
and casts
an idol
\par }{\Q that
will prove worthless?
\par }{\Q \VS{11}Look,
all
his associates
will be put to shame;
\par }{\Q the craftsmen
are mere humans.
\par }{\Q Let them all
assemble
and take their stand!
\par }{\Q They will panic
and be put to shame.
\par }{\Q \VS{12}A blacksmith
works
with his tool
\par }{\Q and forges metal over the coals.
\par }{\Q He forms
it with hammers;
\par }{\Q he makes
it with his strong
arm.
\par }{\Q He gets hungry
and loses
his energy;
\par }{\Q he drinks
no
water
and gets tired.
\par }{\Q \VS{13}A carpenter
takes
measurements;
\par }{\Q he marks out
an outline
of its form;

\par }{\Q he scrapes
it with chisels,
\par }{\Q and marks
it with a compass.
\par }{\Q He patterns
it after the human
form,
\par }{\Q like a well-built
human
being,
\par }{\Q and puts
it in a shrine.
\par }{\Q \VS{14}He cuts down
cedars
\par }{\Q and acquires
a cypress
or an oak.
\par }{\Q He gets
trees
from the forest;
\par }{\Q he plants
a cedar
and the rain
makes it grow.
\par }{\Q \VS{15}A man
uses it to make a fire;
\par }{\Q he takes
some of it and warms
himself.
\par }{\Q Yes,
he kindles
a fire and bakes
bread.
\par }{\Q Then he makes
a god
and worships
it;
\par }{\Q he makes
an idol
and bows down to it.
\par }{\Q \VS{16}Half
of it he burns
in
the fire
–
\par }{\Q over
that half
he cooks
meat;
\par }{\Q he roasts
a meal
and fills
himself.
\par }{\Q Yes,
he warms
himself and says,
\par }{\Q ‘Ah! I am warm
as I look at
the fire.’
\par }{\Q \VS{17}With the rest
of it he makes a
god,
his idol;
\par }{\Q he bows down to it and worships
it.
\par }{\Q He prays
to
it, saying,
\par }{\Q ‘Rescue
me, for
you
are my god!’
\par }{\Q \VS{18}They do not
comprehend
or
understand,
\par }{\Q for
their eyes
are blind
and cannot see;
\par }{\Q their minds
do not discern.
\par }{\Q \VS{19}No
one thinks
to
himself,
\par }{\Q nor
do they comprehend
or
understand
and say
to themselves:
\par }{\Q ‘I burned
half
of it in
the fire
–
\par }{\Q yes,
I baked
bread
over
the coals;
\par }{\Q I roasted
meat
and ate
it.
\par }{\Q With the rest
of it should I make
a disgusting
idol?
\par }{\Q Should I bow down
to dry wood?’
\par }{\Q \VS{20}He feeds
on ashes;
\par }{\Q his deceived
mind
misleads
him.
\par }{\Q He cannot
rescue
himself,
\par }{\Q nor
does he say,
‘Is this not
a false
god I hold in my right hand?’
\par }{\Q \VS{21}Remember
these
things, O Jacob,
\par }{\Q O Israel,
for
you are my servant.
\par }{\Q I formed
you to be my servant;
\par }{\Q O
Israel,
I will not
forget you!
\par }{\Q \VS{22}I remove
the guilt
of your rebellious
deeds as if they were a cloud,
\par }{\Q the guilt of your sins
as if they were a cloud.

\par }{\Q Come back
to me,
for
I protect you.”
\par }{\Q \VS{23}Shout
for joy, O sky,
for
the
{\ND{Lord}}
intervenes;
\par }{\Q shout
out, you subterranean
regions of the earth.
\par }{\Q O mountains,
give a joyful
shout;
\par }{\Q you too, O forest
and all
your trees!

\par }{\Q For
the
{\ND{Lord}}
protects
Jacob;
\par }{\Q he reveals his splendor
through Israel.
\par }{\SH The Lord Empowers Cyrus
\par }{\Q \VS{24}This is what
the {\ND{Lord}}, your protector,
says,
\par }{\Q the one who formed
you in the womb:
\par }{\Q “I am
the {\ND{Lord}}, who made
everything,
\par }{\Q who alone
stretched
out the sky,
\par }{\Q who fashioned
the earth all by myself,
\par }{\Q \VS{25}who frustrates
the omens
of the empty talkers
\par }{\Q and humiliates
the omen readers,
\par }{\Q who overturns
the counsel
of the wise men
\par }{\Q and makes their advice seem foolish,
\par }{\Q \VS{26}who
fulfills
the oracles of his prophetic
servants
\par }{\Q and brings to pass the announcements
of his messengers,
\par }{\Q who says
about Jerusalem, ‘She will be inhabited,’
\par }{\Q and about the towns
of Judah,
‘They will be rebuilt,
\par }{\Q her ruins
I will raise up,’
\par }{\Q \VS{27}who says
to the deep
sea, ‘Be dry!
\par }{\Q I will dry up
your sea currents,’
\par }{\Q \VS{28}who commissions Cyrus,
the one I appointed as shepherd
\par }{\Q to carry out
all
my wishes

\par }{\Q and to decree concerning
Jerusalem,
‘She will be rebuilt,’
\par }{\Q and concerning the temple,
‘It will be reconstructed.’ ”

\par }\Chap{45}{\PP \par }{\Q \VerseOne{1}This is what
the {\ND{Lord}}
says to his chosen one,
\par }{\Q to Cyrus,
whose
right hand
I hold
\par }{\Q in order to subdue
nations
before
him,
\par }{\Q and disarm
kings,
\par }{\Q to open
doors
before
him,
\par }{\Q so gates
remain
unclosed:
\par }{\Q \VS{2}“I
will go
before
you
\par }{\Q and level
mountains.

\par }{\Q Bronze
doors
I will shatter
\par }{\Q and iron
bars
I will hack through.
\par }{\Q \VS{3}I will give
you hidden
treasures,
\par }{\Q riches stashed away
in secret places,
\par }{\Q so
you may recognize
that
I am
the {\ND{Lord}},
\par }{\Q the one who calls
you by name,
the God
of Israel.
\par }{\Q \VS{4}For the sake
of my servant
Jacob,
\par }{\Q Israel,
my chosen one,
\par }{\Q I call
you by name
\par }{\Q and give you a title of respect,
even though you do not
recognize me.
\par }{\Q \VS{5}I
am the
{\ND{Lord}}, I have no
peer,
\par }{\Q there is no
God
but me.
\par }{\Q I arm
you for battle, even though you do not
recognize me.
\par }{\Q \VS{6}I do this so
people will recognize
from east
to west
\par }{\Q that
there is no
God but
me;
\par }{\Q I
am the
{\ND{Lord}}, I have no
peer.
\par }{\Q \VS{7}I am the one who forms
light
\par }{\Q and creates
darkness;
\par }{\Q the one who brings
about peace
\par }{\Q and creates
calamity.
\par }{\Q I
am the
{\ND{Lord}}, who accomplishes
all
these things.
\par }{\Q \VS{8}O sky,
rain
down from above!
\par }{\Q Let the clouds
send down showers
of deliverance!
\par }{\Q Let the earth
absorb
it so salvation may grow,

\par }{\Q and deliverance
may sprout
up along with it.
\par }{\Q I,
the {\ND{Lord}}, create it.
\par }{\SH The Lord Gives a Warning
\par }{\Q \VS{9}One who argues
with his creator
is in grave danger,

\par }{\Q one who is like a mere shard
among the
other shards
on the ground!
\par }{\Q The clay
should not
say
to the potter,
\par }{\Q “What
in the world are
you doing?
\par }{\Q Your work
lacks skill!”
\par }{\Q \VS{10}Danger awaits
one who says
to his father,
\par }{\Q “What
in the world are you fathering?”
\par }{\Q and to his mother,
\par }{\Q “What
in the world are you bringing forth?”
\par }{\Q \VS{11}This is what
the {\ND{Lord}}
says,
\par }{\Q the Holy One
of Israel,
the one who formed
him,
\par }{\Q concerning things to come:

\par }{\Q “How dare you question
me about
my children!
\par }{\Q How dare
you tell
me what to do
with the work
of my own hands!
\par }{\Q \VS{12}I
made
the earth,
\par }{\Q I created
the people
who live on
it.
\par }{\Q It was me
– my hands
stretched
out the sky,
\par }{\Q I give orders
to all
the heavenly
lights.
\par }{\Q \VS{13}It is me – I stir him up and commission him;
\par }{\Q I will make all his ways level.
\par }{\Q He will rebuild my city;
\par }{\Q he will send my exiled people home,
\par }{\Q but not for a price or a bribe,”
\par }{\Q says the
{\ND{Lord}} who commands armies.
\par }{\SH The Lord is the Nations’ Only Hope
\par }{\Q \VS{14}This is what
the
{\ND{Lord}}
says:
\par }{\Q “The profit
of Egypt
and the revenue
of Ethiopia,
\par }{\Q along with the Sabeans,
those tall men,
\par }{\Q will be brought to you and become
yours.
\par }{\Q They will walk
behind
you, coming along in chains.
\par }{\Q They will bow
down to
you
\par }{\Q and pray
to
you:

\par }{\Q ‘Truly
God
is with you; he has no
peer;
\par }{\Q there is no other God!’ ”
\par }{\Q \VS{15}Yes,
you
are a God
who keeps hidden,
\par }{\Q O God
of Israel,
deliverer!
\par }{\Q \VS{16}They will all be ashamed
and
embarrassed;
\par }{\Q those who fashion
idols
will all
be humiliated.
\par }{\Q \VS{17}Israel
will be delivered
once and for all
by the
{\ND{Lord}};

\par }{\Q you will never again
be ashamed
or
humiliated.
\par }{\Q \VS{18}For
this is what
the {\ND{Lord}}
says,
\par }{\Q the one who created
the sky
–
\par }{\Q he is the true God,
\par }{\Q the one who formed
the earth
and made
it;
\par }{\Q he established
it,
\par }{\Q he did not
create
it without order,
\par }{\Q he formed
it to be inhabited
–
\par }{\Q “I am
the {\ND{Lord}}, I have no
peer.
\par }{\Q \VS{19}I have not
spoken
in secret,
\par }{\Q in some hidden
place.
\par }{\Q I did not
tell
Jacob’s
descendants,
\par }{\Q ‘Seek
me in vain!’

\par }{\Q I am
the {\ND{Lord}},
\par }{\Q the one who speaks
honestly,
\par }{\Q who makes reliable announcements.
\par }{\Q \VS{20}Gather
together and come!
\par }{\Q Approach
together,
you refugees
from the nations!
\par }{\Q Those who carry
wooden
idols
know
nothing,
\par }{\Q those who pray
to
a god
that cannot
deliver.
\par }{\Q \VS{21}Tell
me! Present the evidence!

\par }{\Q Let
them consult
with one another!
\par }{\Q Who
predicted
this
in the past?
\par }{\Q Who announced
it beforehand?
\par }{\Q Was it not
I,
the {\ND{Lord}}?
\par }{\Q I have no
peer,
there is no God
but
me,
\par }{\Q a God
who vindicates
and delivers;
\par }{\Q there is none
but me.
\par }{\Q \VS{22}Turn
to me
so you can be delivered,
\par }{\Q all
you who live in the earth’s
remote regions!
\par }{\Q For
I am
God,
and I have no
peer.
\par }{\Q \VS{23}I solemnly make this oath –
\par }{\Q what I say is true and reliable:
\par }{\Q ‘Surely every knee will bow to me,
\par }{\Q every tongue will solemnly affirm;
\par }{\Q \VS{24}they will say
about me,
\par }{\Q “Yes, the
{\ND{Lord}}
is a powerful
deliverer.” ’”

\par }{\Q All
who are angry
at him will cower before him.
\par }{\Q \VS{25}All
the descendants
of Israel
will be vindicated
by the
{\ND{Lord}}
\par }{\Q and will boast in him.

\par }\Chap{46}{\PP \par }{\Q \VerseOne{1}Bel
kneels down,
\par }{\Q Nebo
bends low.
\par }{\Q Their images
weigh down
animals
and beasts.
\par }{\Q Your heavy images
are burdensome
to tired animals.
\par }{\Q \VS{2}Together
they bend
low and kneel
down;
\par }{\Q they are unable
to rescue
the images;
\par }{\Q they themselves
head off into captivity.
\par }{\Q \VS{3}“Listen
to me,
O family
of Jacob,
\par }{\Q all
you who are left
from the family
of Israel,
\par }{\Q you who have been carried
from
birth,
\par }{\Q you who have been supported
from
the time you left the womb.
\par }{\Q \VS{4}Even
when you are old,
I
will take care of you,

\par }{\Q even
when you have gray hair,
I
will carry
you.
\par }{\Q I
made
you and I
will support
you;
\par }{\Q I
will carry
you and rescue you.
\par }{\Q \VS{5}To whom
can you compare
and liken
me?
\par }{\Q Tell me whom you think I resemble,
so we can be compared!
\par }{\Q \VS{6}Those who empty
out gold
from a purse
\par }{\Q and weigh
out silver
on the scale

\par }{\Q hire
a metalsmith,
who makes it into a
god.
\par }{\Q They then bow down and worship it.
\par }{\Q \VS{7}They put
it on
their shoulder
and carry
it;
\par }{\Q they put
it in its place
and it just stands
there;
\par }{\Q it does not
move
from its place.
\par }{\Q Even
when someone cries out
to
it, it does not
reply;
\par }{\Q it does not
deliver
him from his distress.
\par }{\Q \VS{8}Remember
this,
so you can be brave!

\par }{\Q Think about it, you rebels!
\par }{\Q \VS{9}Remember
what I accomplished
in antiquity!

\par }{\Q Truly I am
God,
I have no peer;
\par }{\Q I am God,
and there is none
like me,
\par }{\Q \VS{10}who announces
the end
from
the beginning
\par }{\Q and reveals beforehand
what
has not
yet occurred,
\par }{\Q who says,
‘My plan
will be realized,
\par }{\Q I will accomplish
what I desire,’
\par }{\Q \VS{11}who summons
an eagle
from the east,
\par }{\Q from a distant
land,
one
who carries out
my plan.
\par }{\Q Yes,
I have decreed,
\par }{\Q yes,
I will bring
it to pass;
\par }{\Q I have formulated a plan,
\par }{\Q yes,
I will carry it out.
\par }{\Q \VS{12}Listen
to
me, you stubborn
people,

\par }{\Q you who distance
yourself from doing what is right.
\par }{\Q \VS{13}I am bringing
my deliverance
near,
it is not
far away;
\par }{\Q I am bringing my salvation
near, it does not
wait.
\par }{\Q I will save
Zion;
\par }{\Q I will adorn Israel
with my splendor.”

\par }\Chap{47}{\PP \par }{\Q \VerseOne{1}“Fall down! Sit
in the dirt,
\par }{\Q O virgin
daughter
Babylon!
\par }{\Q Sit
on the ground,
not
on a throne,
\par }{\Q O daughter
of the Babylonians!
\par }{\Q Indeed,
you will no
longer
be called
delicate
and pampered.
\par }{\Q \VS{2}Pick
up millstones
and grind
flour!

\par }{\Q Remove your veil,
\par }{\Q strip off
your skirt,
\par }{\Q expose
your legs,
\par }{\Q cross
the streams!
\par }{\Q \VS{3}Let your private parts
be exposed!
\par }{\Q Your genitals
will be on display!

\par }{\Q I will get
revenge;
\par }{\Q I will not
have pity
on anyone,”
\par }{\Q \VS{4}says our protector
–
\par }{\Q the {\ND{Lord}}
who commands armies
is his name,
\par }{\Q the Holy One
of Israel.
\par }{\Q \VS{5}“Sit
silently! Go
to a hiding
place,

\par }{\Q O daughter
of the Babylonians!
\par }{\Q Indeed,
you will no
longer
be called
‘Queen
of kingdoms.’
\par }{\Q \VS{6}I was angry
at my people;
\par }{\Q I defiled
my special
possession
\par }{\Q and handed
them over to you.
\par }{\Q You showed them
no
mercy;
\par }{\Q you even
placed
a very
heavy
burden
on
old people.
\par }{\Q \VS{7}You said,
\par }{\Q ‘I will rule forever
as permanent
queen!’

\par }{\Q You did not think
about these
things;

\par }{\Q you did not consider
how it
would turn
out.
\par }{\Q \VS{8}So now,
listen
to this,
\par }{\Q O one who lives so lavishly,
\par }{\Q who lives
securely,
\par }{\Q who says
to herself,
\par }{\Q ‘I am
unique! No one can compare
to me!

\par }{\Q I will never
have to live
as a widow;
\par }{\Q I will never
lose my children.’
\par }{\Q \VS{9}Both
of these
will come
upon you
\par }{\Q suddenly,
in one
day!
\par }{\Q You will lose your children
and be widowed.
\par }{\Q You will be overwhelmed
by
these tragedies,
\par }{\Q despite your many
incantations
\par }{\Q and your numerous
amulets.
\par }{\Q \VS{10}You were complacent
in your evil
deeds;

\par }{\Q you thought, ‘No
one sees
me.’
\par }{\Q Your self-professed
wisdom
and knowledge
lead you astray,
\par }{\Q when you say,
‘I am
unique! No
one can compare to me!’
\par }{\Q \VS{11}Disaster
will overtake
you;
\par }{\Q you will not
know
how to charm
it away.

\par }{\Q Destruction
will fall
on
you;
\par }{\Q you will not
be able
to appease
it.
\par }{\Q Calamity
will strike
you suddenly,
\par }{\Q before you recognize it.
\par }{\Q \VS{12}Persist
in trusting
your amulets
\par }{\Q and your many
incantations,
\par }{\Q which
you have faithfully recited
since your youth!
\par }{\Q Maybe
you will be successful –
\par }{\Q maybe
you will scare away disaster.
\par }{\Q \VS{13}You are tired
out from listening to so much
advice.
\par }{\Q Let them take their stand
–
\par }{\Q the ones who see omens
in the sky,
\par }{\Q who gaze
at the stars,
\par }{\Q who make
monthly
predictions
–
\par }{\Q let them rescue
you from the disaster
that
is about to overtake you!
\par }{\Q \VS{14}Look,
they are like straw,
\par }{\Q which the fire
burns up;
\par }{\Q they cannot
rescue
themselves
\par }{\Q from the heat of the flames.
\par }{\Q There are no
coals
to warm
them,
\par }{\Q no firelight to enjoy.
\par }{\Q \VS{15}They will disappoint you,

\par }{\Q those you have so faithfully dealt
with since your youth.
\par }{\Q Each
strays off
in his own direction,
\par }{\Q leaving no one
to rescue you.”

\par }\Chap{48}{\PP \par }{\Q \VerseOne{1}Listen
to this,
O family
of Jacob,
\par }{\Q you who are called
by the name
‘Israel,’
\par }{\Q and are descended
from Judah,
\par }{\Q who take oaths
in the name
of the {\ND{Lord}},
\par }{\Q and invoke
the God
of Israel
–
\par }{\Q but not
in an honest
and just manner.
\par }{\Q \VS{2}Indeed,
they live
in the holy
city;
\par }{\Q they trust
in the God
of Israel,
\par }{\Q whose
name
is the
{\ND{Lord}}
who commands armies.
\par }{\Q \VS{3}“I announced events
beforehand,
\par }{\Q I issued
the decrees
and made
the predictions;
\par }{\Q suddenly
I acted
and they came to pass.
\par }{\Q \VS{4}I did this because
I know
how stubborn
you
are.
\par }{\Q Your neck
muscles
are like iron
\par }{\Q and your forehead
like bronze.
\par }{\Q \VS{5}I announced
them to you beforehand;
\par }{\Q before
they happened,
I predicted
them for you,
\par }{\Q so you could never say,
\par }{\Q ‘My image
did
these things,
\par }{\Q my idol,
my cast image,
decreed them.’
\par }{\Q \VS{6}You have heard;
now look
at all
the evidence!

\par }{\Q Will you
not
admit
that what I say
is true?

\par }{\Q From this point on I am announcing to you new
events
\par }{\Q that are previously unrevealed
and you do not
know about.
\par }{\Q \VS{7}Now
they come into being,
not
in the past;
\par }{\Q before
today
you did not
hear
about them,
\par }{\Q so you could
not say,
\par }{\Q ‘Yes,
I know about them.’
\par }{\Q \VS{8}You did not
hear,
\par }{\Q you do not
know,
\par }{\Q you were not
told
beforehand.
\par }{\Q For
I know
that you are very deceitful;
\par }{\Q you were labeled
a rebel
from birth.
\par }{\Q \VS{9}For the sake
of my reputation
I hold back
my anger;
\par }{\Q for the sake of my prestige
I restrain
myself from destroying you.
\par }{\Q \VS{10}Look,
I have refined
you, but not
as silver;
\par }{\Q I have purified
you in the furnace
of misery.
\par }{\Q \VS{11}For
my sake
alone I will act,
\par }{\Q for
how
can I allow my name to be defiled?

\par }{\Q I will not
share
my glory
with anyone else!
\par }{\Q \VS{12}Listen
to me,
O Jacob,
\par }{\Q Israel,
whom I summoned!
\par }{\Q I am
the one;
\par }{\Q I
am
present at the very beginning
\par }{\Q and at the very end.
\par }{\Q \VS{13}Yes,
my hand
founded
the earth;
\par }{\Q my right hand
spread
out the sky.
\par }{\Q I
summon
them;
\par }{\Q they stand
together.
\par }{\Q \VS{14}All
of you, gather
together and listen!
\par }{\Q Who
among them announced
these
things?
\par }{\Q The
{\ND{Lord}}’s
ally
will carry
out his desire
against Babylon;
\par }{\Q he will exert
his power against the Babylonians.
\par }{\Q \VS{15}I,
I
have spoken
–
\par }{\Q yes,
I have summoned
him;
\par }{\Q I lead
him and he will succeed.
\par }{\Q \VS{16}Approach
me! Listen
to this!
\par }{\Q From the very first
I have not
spoken
in secret;
\par }{\Q when
it happens, I am
there.”
\par }{\Q So now,
the sovereign

{\ND{Lord}}
has sent
me, accompanied by his spirit.
\par }{\Q \VS{17}This is what
the {\ND{Lord}}, your protector,
says,
\par }{\Q the Holy One
of Israel:

\par }{\Q “I am
the {\ND{Lord}}
your God,
\par }{\Q who teaches
you how
to succeed,
\par }{\Q who
leads
you in the way
you should go.
\par }{\Q \VS{18}If only
you had obeyed my
commandments,
\par }{\Q prosperity
would have flowed to you like a river,
\par }{\Q deliverance
would have come to you like the waves
of the sea.
\par }{\Q \VS{19}Your descendants
would have been
as numerous as sand,
\par }{\Q and your children
like its granules.
\par }{\Q Their name
would not
have been cut off
\par }{\Q and eliminated
from my presence.
\par }{\Q \VS{20}Leave
Babylon!
\par }{\Q Flee
from the Babylonians!
\par }{\Q Announce it with a shout
of joy!
\par }{\Q Make this
known!
\par }{\Q Proclaim
it throughout the earth!

\par }{\Q Say,
‘The
{\ND{Lord}}
protects
his servant
Jacob.
\par }{\Q \VS{21}They do not
thirst
as he leads them through
dry regions;
\par }{\Q he makes water
flow
out of a rock
for them;
\par }{\Q he splits
open a rock
and water
flows out.’
\par }{\Q \VS{22}There will be no
prosperity
for the wicked,”
says
the {\ND{Lord}}.

\par }\Chap{49}{\PP \par }{\Q \VerseOne{1}Listen
to me, you coastlands!

\par }{\Q Pay attention,
you people
who live far away!
\par }{\Q The
{\ND{Lord}}
summoned
me from birth;
\par }{\Q he commissioned me when my mother
brought me into the world.
\par }{\Q \VS{2}He made
my mouth
like a sharp
sword,
\par }{\Q he hid
me in the hollow
of his hand;
\par }{\Q he made
me like a sharpened
arrow,
\par }{\Q he hid
me in his quiver.
\par }{\Q \VS{3}He said
to me, “You
are my servant,
\par }{\Q Israel,
through whom
I will reveal my splendor.”
\par }{\Q \VS{4}But I
thought, “I have worked
in vain;
\par }{\Q I have expended
my energy
for absolutely nothing.”
\par }{\Q But
the {\ND{Lord}}
will vindicate
me;
\par }{\Q my God
will reward me.
\par }{\Q \VS{5}So now
the
{\ND{Lord}}
says,
\par }{\Q the one who formed
me from birth
to be his servant
–
\par }{\Q he did this to restore
Jacob
to
himself,
\par }{\Q so that Israel
might be gathered
to him;
\par }{\Q and I will be honored
in the
{\ND{Lord}}’s
sight,
\par }{\Q for my God
is my source
of strength –
\par }{\Q \VS{6}he says,
“Is it too insignificant
a task for you to be
my servant,
\par }{\Q to reestablish
the tribes
of Jacob,
\par }{\Q and restore
the remnant
of Israel?

\par }{\Q I will make
you a light
to the nations,
\par }{\Q so
you can bring my deliverance
to
the remote
regions of the earth.”
\par }{\Q \VS{7}This is what
the {\ND{Lord}},
\par }{\Q the protector
of Israel,
their Holy One,
says
\par }{\Q to the one
who is despised
and rejected
by nations,
\par }{\Q a servant
of rulers:
\par }{\Q “Kings
will see
and rise
in respect,

\par }{\Q princes
will bow
down,
\par }{\Q because
of the faithful

{\ND{Lord}},
\par }{\Q the Holy One
of Israel
who has chosen you.”
\par }{\Q \VS{8}This is what
the {\ND{Lord}}
says:
\par }{\Q “At the time
I decide to show my favor,
I will respond
to you;
\par }{\Q in the day
of deliverance
I will help
you;
\par }{\Q I will protect
you and make
you a covenant
mediator
for people,
\par }{\Q to rebuild the land
\par }{\Q and to reassign
the desolate
property.
\par }{\Q \VS{9}You will say
to the prisoners,
‘Come out,’
\par }{\Q and to those who
are in dark dungeons, ‘Emerge.’
\par }{\Q They will graze
beside the roads;
\par }{\Q on
all
the slopes
they will find pasture.
\par }{\Q \VS{10}They will not
be hungry
or thirsty;
\par }{\Q the sun’s
oppressive heat
will not
beat down
on them,

\par }{\Q for
one who has compassion
on
them will guide
them;
\par }{\Q he will lead
them to springs
of water.
\par }{\Q \VS{11}I will make
all
my mountains
into a road;
\par }{\Q I will construct
my roadways.”
\par }{\Q \VS{12}Look,
they come
from far
away!
\par }{\Q Look,
some come from the north
and west,
\par }{\Q and others
from the land
of Sinim!
\par }{\Q \VS{13}Shout
for joy, O sky!

\par }{\Q Rejoice,
O earth!
\par }{\Q Let the mountains
give a joyful
shout!
\par }{\Q For
the {\ND{Lord}}
consoles
his people
\par }{\Q and shows compassion
to the oppressed.
\par }{\SH The Lord Remembers Zion
\par }{\Q \VS{14}“Zion
said,
‘The
{\ND{Lord}}
has abandoned
me,
\par }{\Q the sovereign
master has forgotten me.’
\par }{\Q \VS{15}Can a woman
forget
her baby who nurses at her breast?

\par }{\Q Can she withhold compassion
from the child
she has borne?

\par }{\Q Even if
mothers were to forget,
\par }{\Q I
could never
forget you!
\par }{\Q \VS{16}Look,
I have inscribed
your name on
my palms;
\par }{\Q your walls
are constantly
before me.
\par }{\Q \VS{17}Your children
hurry
back,
\par }{\Q while those who destroyed
and devastated
you depart.
\par }{\Q \VS{18}Look
all around
you!

\par }{\Q All
of them gather
to you.
\par }{\Q As surely as I
live,”
says
the {\ND{Lord}},
\par }{\Q “you will certainly
wear all
of them like jewelry;
\par }{\Q you will put
them on as if you were a bride.
\par }{\Q \VS{19}Yes,
your land
lies in ruins;
\par }{\Q it is desolate
and devastated.
\par }{\Q But now
you will be too small to hold your residents,
\par }{\Q and those who devoured
you will be far away.
\par }{\Q \VS{20}Yet the children born during your time of bereavement
\par }{\Q will say
within your hearing,
\par }{\Q ‘This place
is too cramped
for us,

\par }{\Q make room for us so we can live here.’
\par }{\Q \VS{21}Then you will think
to yourself,

\par }{\Q ‘Who
bore
these
children
for me?
\par }{\Q I
was bereaved
and barren,
\par }{\Q dismissed
and divorced.
\par }{\Q Who
raised
these
children?
\par }{\Q Look,
I
was left
all alone;
\par }{\Q where
did these children come from?’ ”
\par }{\Q \VS{22}This is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “Look
I will raise
my hand
to
the nations;
\par }{\Q I will raise
my signal flag
to
the peoples.
\par }{\Q They will bring
your sons
in their arms
\par }{\Q and carry
your daughters
on
their shoulders.
\par }{\Q \VS{23}Kings
will be
your children’s guardians;
\par }{\Q their princesses
will nurse
your children.

\par }{\Q With their faces
to the ground
they will bow down
to you
\par }{\Q and they will lick
the dirt
on your feet.
\par }{\Q Then you will recognize
that
I
am the
{\ND{Lord}};
\par }{\Q those who wait
patiently for me are not
put to shame.
\par }{\Q \VS{24}Can spoils
be taken
from a warrior,
\par }{\Q or
captives
be rescued
from a conqueror?
\par }{\Q \VS{25}Indeed,”
says
the {\ND{Lord}},
\par }{\Q “captives
will be taken
from a warrior;
\par }{\Q spoils
will be rescued
from
a conqueror.
\par }{\Q I
will oppose
your adversary
\par }{\Q and I
will rescue
your children.
\par }{\Q \VS{26}I will make your oppressors
eat
their own flesh;
\par }{\Q they will get drunk
on their own blood,
as if it were wine.
\par }{\Q Then all
humankind
will recognize
that
\par }{\Q I am
the {\ND{Lord}}, your deliverer,
\par }{\Q your protector,
the powerful ruler
of Jacob.”

\par }\Chap{50}{\PP \par }{\Q \VerseOne{1}This is what
the {\ND{Lord}}
says:
\par }{\Q “Where
is
your mother’s
divorce
certificate
\par }{\Q by which
I divorced
her?
\par }{\Q Or
to which
of my creditors
did I sell
you?

\par }{\Q Look,
you were sold
because of your sins;
\par }{\Q because of your rebellious
acts I divorced
your mother.
\par }{\Q \VS{2}Why
does no
one
challenge me when I come?
\par }{\Q Why does no
one respond
when I call?

\par }{\Q Is my hand
too weak
to deliver
you?
\par }{\Q Do I lack
the power
to rescue
you?
\par }{\Q Look,
with a mere shout
I can dry up
the sea;
\par }{\Q I can turn
streams
into a desert,
\par }{\Q so the fish
rot away
and die
\par }{\Q from lack of water.
\par }{\Q \VS{3}I can clothe
the sky
in darkness;
\par }{\Q I can
cover
it with sackcloth.”
\par }{\SH The Servant Perseveres
\par }{\Q \VS{4}The sovereign

{\ND{Lord}}
has given
me the capacity to be his spokesman,
\par }{\Q so that I know
how to help
the weary.
\par }{\Q He wakes
me up
every morning;
\par }{\Q he makes
me alert
so I can listen attentively
as disciples do.
\par }{\Q \VS{5}The sovereign

{\ND{Lord}}
has spoken
to me clearly;
\par }{\Q I
have not
rebelled,
\par }{\Q I have not
turned
back.
\par }{\Q \VS{6}I offered
my back
to those who attacked,
\par }{\Q my jaws
to those who tore out
my beard;
\par }{\Q I did not
hide
my face
\par }{\Q from insults
and spitting.
\par }{\Q \VS{7}But the sovereign

{\ND{Lord}}
helps
me,
\par }{\Q so
I
am not
humiliated.
\par }{\Q For that reason
I am steadfastly
resolved;
\par }{\Q I know
I will not
be put to shame.
\par }{\Q \VS{8}The one who
vindicates me
is close
by.
\par }{\Q Who dares
to argue
with
me? Let us confront each other!

\par }{\Q Who is
my accuser? Let him challenge
me!
\par }{\Q \VS{9}Look,
the sovereign

{\ND{Lord}}
helps
me.
\par }{\Q Who dares
to condemn
me?
\par }{\Q Look,
all
of them will wear out
like clothes;
\par }{\Q a moth
will eat away at them.
\par }{\Q \VS{10}Who
among you fears
the {\ND{Lord}}?
\par }{\Q Who
obeys
his servant?
\par }{\Q Whoever
walks
in deep darkness,
\par }{\Q without
light,
\par }{\Q should trust
in the name
of the {\ND{Lord}}
\par }{\Q and rely
on his God.
\par }{\Q \VS{11}Look,
all
of you who start
a fire
\par }{\Q and who equip
yourselves with flaming arrows,
\par }{\Q walk
in the light
of the fire
you started
\par }{\Q and among the flaming arrows
you ignited!

\par }{\Q This
is what you will receive
from me:

\par }{\Q you will lie down
in a place of pain.

\par }\Chap{51}{\PP \par }{\Q \VerseOne{1}“Listen
to
me, you who pursue
godliness,
\par }{\Q who seek
the {\ND{Lord}}!
\par }{\Q Look
at the rock
from
which you were chiseled,
\par }{\Q at the quarry
from
which you were dug!
\par }{\Q \VS{2}Look
at Abraham,
your father,
\par }{\Q and Sarah,
who gave you birth.
\par }{\Q When
I summoned
him, he was a lone
individual,

\par }{\Q but I blessed
him and gave him numerous descendants.
\par }{\Q \VS{3}Certainly
the {\ND{Lord}}
will console
Zion;
\par }{\Q he will console
all
her ruins.
\par }{\Q He will make
her wilderness
like Eden,
\par }{\Q her desert
like the Garden
of the {\ND{Lord}}.
\par }{\Q Happiness
and joy
will be restored
to her,
\par }{\Q thanksgiving
and the sound
of music.
\par }{\Q \VS{4}Pay attention
to me,
my people!
\par }{\Q Listen
to me,
my people!
\par }{\Q For
I will issue
a decree,
\par }{\Q I will make my
justice
a light
to the nations.
\par }{\Q \VS{5}I am ready
to vindicate,
\par }{\Q I am ready
to deliver,
\par }{\Q I will establish
justice
among
the nations.
\par }{\Q The coastlands
wait patiently
for me;
\par }{\Q they wait in anticipation
for
the revelation of my power.
\par }{\Q \VS{6}Look
up at the sky!
\par }{\Q Look
at the earth
below!
\par }{\Q For
the sky
will dissipate
like smoke,
\par }{\Q and the earth
will wear out
like clothes;
\par }{\Q its residents
will die
like
gnats.
\par }{\Q But the deliverance
I give is permanent;
\par }{\Q the vindication
I provide will
not
disappear.
\par }{\Q \VS{7}Listen
to me,
you who know
what is right,
\par }{\Q you people
who are aware
of my
law!

\par }{\Q Don’t
be afraid
of the insults
of men;
\par }{\Q don’t
be discouraged
because of their abuse!
\par }{\Q \VS{8}For
a moth
will eat
away at them like clothes;
\par }{\Q a clothes
moth
will devour
them like wool.
\par }{\Q But the vindication
I provide will be
permanent;
\par }{\Q the deliverance I give will last.”
\par }{\Q \VS{9}Wake up! Wake up!
\par }{\Q Clothe
yourself with strength,
O arm
of the
{\ND{Lord}}!

\par }{\Q Wake up
as in former
times,
as in antiquity!
\par }{\Q Did you
not
smash
the Proud One?

\par }{\Q Did you not wound
the sea monster?
\par }{\Q \VS{10}Did you
not
dry
up the sea,
\par }{\Q the waters
of the great
deep?
\par }{\Q Did you not make
a path
through the depths
of the sea,
\par }{\Q so
those delivered from bondage
could cross over?
\par }{\Q \VS{11}Those whom the
{\ND{Lord}}
has ransomed
will return;
\par }{\Q they will enter
Zion
with a happy
shout.
\par }{\Q Unending
joy will crown
them,

\par }{\Q happiness
and joy
will overwhelm
them;
\par }{\Q grief
and suffering
will disappear.
\par }{\Q \VS{12}“I,
I am
the one who
consoles
you.
\par }{\Q Why are you
afraid
of mortal
men,
\par }{\Q of mere human
beings who are as short-lived
as grass?
\par }{\Q \VS{13}Why do you forget
the {\ND{Lord}}, who made
you,
\par }{\Q who stretched
out the sky
\par }{\Q and founded
the earth?
\par }{\Q Why do you constantly
tremble
all
day
long

\par }{\Q at the anger
of the oppressor,
\par }{\Q when he makes
plans to destroy?
\par }{\Q Where
is the anger
of the oppressor?
\par }{\Q \VS{14}The one who suffers
will soon
be released;
\par }{\Q he will not
die
in prison,
\par }{\Q he will not
go hungry.
\par }{\Q \VS{15}I
am the
{\ND{Lord}}
your God,
\par }{\Q who churns
up the sea
so that its waves
surge.
\par }{\Q The
{\ND{Lord}}
who commands armies
is his name!
\par }{\SH Zion’s Time to Celebrate
\par }{\Q \VS{16}I commission
you as my spokesman;
\par }{\Q I cover
you with the palm
of my hand,
\par }{\Q to establish
the sky
and to found
the earth,
\par }{\Q to say
to Zion,
‘You
are my people.’ ”
\par }{\Q \VS{17}Wake
up! Wake
up!
\par }{\Q Get
up, O Jerusalem!
\par }{\Q You drank
from
the cup
the {\ND{Lord}}
passed to you,
\par }{\Q which was full of his anger!

\par }{\Q You drained dry
\par }{\Q the goblet
full of intoxicating wine.
\par }{\Q \VS{18}There was no
one to lead
her
\par }{\Q among all
the children
she bore;
\par }{\Q there was no
one to take
her by the hand
\par }{\Q among all
the children
she raised.
\par }{\Q \VS{19}These
double
disasters confronted
you.
\par }{\Q But who
feels sorry
for you?
\par }{\Q Destruction
and devastation,
\par }{\Q famine
and sword.
\par }{\Q But who
consoles you?
\par }{\Q \VS{20}Your children
faint;
\par }{\Q they lie
at the head
of every
street
\par }{\Q like an antelope
in a snare.
\par }{\Q They are left in a stupor
by the
{\ND{Lord}}’s
anger,
\par }{\Q by the battle cry
of your God.
\par }{\Q \VS{21}So
listen
to this,
oppressed
one,
\par }{\Q who is drunk,
but not
from wine!
\par }{\Q \VS{22}This is what
your sovereign master,
the {\ND{Lord}}
your God,
says:
\par }{\Q “Look,
I have removed
from your hand
\par }{\Q the cup
of intoxicating wine,
\par }{\Q the goblet
full of my anger.
\par }{\Q You will no
longer
have to drink it.
\par }{\Q \VS{23}I will put
it into the hand
of your tormentors
\par }{\Q who
said
to you, ‘Lie down,
so we can walk over
you.’
\par }{\Q You made
your back
like the ground,
\par }{\Q and like the street
for those who walked over you.”

\par }\Chap{52}{\PP \par }{\Q \VerseOne{1}Wake up! Wake up!
\par }{\Q Clothe
yourself with strength,
O Zion!
\par }{\Q Put on
your beautiful
clothes,
\par }{\Q O Jerusalem,
holy
city!
\par }{\Q For
uncircumcised
and unclean
pagans
\par }{\Q will no
longer
invade you.
\par }{\Q \VS{2}Shake
off the dirt!

\par }{\Q Get
up, captive
Jerusalem!
\par }{\Q Take off the iron
chains
around your neck,
\par }{\Q O captive
daughter
Zion!
\par }{\Q \VS{3}For
this is what
the {\ND{Lord}}
says:
\par }{\Q “You were sold
for nothing,
\par }{\Q and you will not
be redeemed
for money.”
\par }{\Q \VS{4}For
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “In the beginning
my people
went
to live temporarily
in Egypt;
\par }{\Q Assyria
oppressed them for no good reason.
\par }{\Q \VS{5}And now,
what
do we have here?” says
the {\ND{Lord}}.
\par }{\Q “Indeed my people
have been carried away
for nothing,
\par }{\Q those who rule
over them taunt,”
says
the {\ND{Lord}},
\par }{\Q “and my name
is constantly
slandered
all
day long.
\par }{\Q \VS{6}For this reason
my people
will know
my name,
\par }{\Q for this reason
they will know at that time
that
I am
the one
who says,
\par }{\Q ‘Here I am.’ ”
\par }{\Q \VS{7}How
delightful
it is to see approaching over
the mountains
\par }{\Q the feet
of a messenger
who announces
peace,
\par }{\Q a messenger
who brings good
news, who announces
deliverance,
\par }{\Q who says
to Zion,
“Your God
reigns!”
\par }{\Q \VS{8}Listen,
your watchmen
shout;
\par }{\Q in unison
they shout
for joy,
\par }{\Q for
they see
with their very
own eyes
\par }{\Q the
{\ND{Lord}}’s
return
to Zion.
\par }{\Q \VS{9}In unison
give a joyful
shout,
\par }{\Q O ruins
of Jerusalem!
\par }{\Q For
the {\ND{Lord}}
consoles
his people;
\par }{\Q he protects
Jerusalem.
\par }{\Q \VS{10}The
{\ND{Lord}}
reveals
his royal power
\par }{\Q in the sight
of all
the nations;
\par }{\Q the entire
earth
sees
\par }{\Q our God
deliver.
\par }{\Q \VS{11}Leave! Leave! Get out
of there!
\par }{\Q Don’t
touch
anything unclean!
\par }{\Q Get out
of it!
\par }{\Q Stay pure,
you who carry
the
{\ND{Lord}}’s
holy items!
\par }{\Q \VS{12}Yet
do not
depart
quickly
\par }{\Q or
leave
in a panic.
\par }{\Q For
the
{\ND{Lord}}
goes
before
you;
\par }{\Q the God
of Israel
is your rear guard.
\par }{\SH The Lord Will Vindicate His Servant
\par }{\Q \VS{13}“Look,
my servant
will succeed!

\par }{\Q He will be elevated,
lifted
high,
and greatly
exalted –
\par }{\Q \VS{14}(just
as many
were horrified
by
the sight of you)

\par }{\Q he was so
disfigured
he no longer looked like
a man;
\par }{\Q \VS{15}his form was so
marred he no longer looked human –
\par }{\Q so now he will startle many nations.
\par }{\Q Kings will be shocked by his exaltation,
\par }{\Q for they will witness something unannounced to them,
\par }{\Q and they will understand something they had not heard about.

\par }\Chap{53}{\PP \par }{\Q \VerseOne{1}Who
would have believed
what we just heard?

\par }{\Q When was the
{\ND{Lord}}’s
power
revealed through him?
\par }{\Q \VS{2}He sprouted up
like a twig
before
God,

\par }{\Q like a root
out of parched
soil;
\par }{\Q he had no
stately form
or
majesty
that might catch our attention,
\par }{\Q no
special appearance
that we should want to follow him.
\par }{\Q \VS{3}He was despised
and rejected
by people,
\par }{\Q one
who experienced pain
and was acquainted
with illness;
\par }{\Q people hid
their faces
from
him;

\par }{\Q he was despised,
and we considered him insignificant.
\par }{\Q \VS{4}But
he lifted up our illnesses,
\par }{\Q he carried
our pain;
\par }{\Q even though we
thought
he was
being punished,
\par }{\Q attacked
by God,
and afflicted for something he had done.
\par }{\Q \VS{5}He was
wounded
because of our rebellious
deeds,
\par }{\Q crushed
because of our sins;
\par }{\Q he endured
punishment that made us well;
\par }{\Q because of his wounds
we have been healed.
\par }{\Q \VS{6}All
of us had wandered
off like sheep;
\par }{\Q each
of us had strayed
off on his own path,
\par }{\Q but
the {\ND{Lord}}
caused
the sin
of all
of us to attack him.
\par }{\Q \VS{7}He was treated harshly
and afflicted,
\par }{\Q but he did not
even open
his mouth.
\par }{\Q Like a lamb
led
to the slaughtering
block,
\par }{\Q like a sheep
silent
before
her shearers,
\par }{\Q he did not
even open
his mouth.
\par }{\Q \VS{8}He was led away
after an unjust trial –
\par }{\Q but who
even cared?

\par }{\Q Indeed,
he was cut
off from the land
of the living;
\par }{\Q because of the rebellion
of his own people
he was wounded.
\par }{\Q \VS{9}They intended to bury
him with criminals,
\par }{\Q but he ended up in a rich
man’s tomb,
\par }{\Q because
he had committed no
violent
deeds,
\par }{\Q nor
had he spoken
deceitfully.
\par }{\Q \VS{10}Though the
{\ND{Lord}}
desired
to crush
him and make him ill,
\par }{\Q once restitution
is made,
\par }{\Q he will see
descendants
and enjoy long
life,
\par }{\Q and the
{\ND{Lord}}’s
purpose
will be accomplished through him.
\par }{\Q \VS{11}Having
suffered,
he will reflect
on his work,
\par }{\Q he will be satisfied
when he understands
what he has done.

\par }{\Q “My servant
will acquit many,
\par }{\Q for he carried
their sins.
\par }{\Q \VS{12}So
I will assign
him a portion
with
the multitudes,
\par }{\Q he will divide
the spoils
of victory with the powerful,
\par }{\Q because
he willingly submitted
to death
\par }{\Q and was numbered
with
the rebels,
\par }{\Q when he
lifted up
the sin
of many
\par }{\Q and intervened on behalf
of the rebels.”

\par }\Chap{54}{\PP \par }{\Q \VerseOne{1}“Shout for joy,
O barren
one who has not
given birth!
\par }{\Q Give
a joyful shout
and cry out,
you who have not
been in labor!
\par }{\Q For
the children
of the desolate
one are more numerous
\par }{\Q than the children
of the married woman,”
says
the {\ND{Lord}}.
\par }{\Q \VS{2}Make
your tent
larger,
\par }{\Q stretch
your tent
curtains
farther out!

\par }{\Q Spare
no
effort,
\par }{\Q lengthen
your ropes,
\par }{\Q and pound
your stakes
deep.
\par }{\Q \VS{3}For
you will spread out
to the right
and to the left;
\par }{\Q your children
will conquer
nations
\par }{\Q and will resettle
desolate
cities.
\par }{\Q \VS{4}Don’t
be afraid,
for
you will not
be put to shame!
\par }{\Q Don’t
be intimidated,
for
you will not
be humiliated!
\par }{\Q You will forget
about the shame
you experienced in your youth;
\par }{\Q you will no
longer
remember
the disgrace
of your abandonment.
\par }{\Q \VS{5}For
your husband
is the one who made you –
\par }{\Q the
{\ND{Lord}} who commands armies is his name.
\par }{\Q He is your protector, the Holy One of Israel.
\par }{\Q He is called “God of the entire earth.”
\par }{\Q \VS{6}“Indeed, the
{\ND{Lord}}
will call
you back
\par }{\Q like a wife
who
has been abandoned
and suffers
from depression,
\par }{\Q like a young
wife
when
she has been rejected,”
says
your God.
\par }{\Q \VS{7}“For a short
time
I abandoned
you,
\par }{\Q but with great
compassion
I will gather you.
\par }{\Q \VS{8}In a burst
of anger
I rejected
you momentarily,
\par }{\Q but with
lasting
devotion
I will have compassion
on you,”
\par }{\Q says
your protector,
the {\ND{Lord}}.
\par }{\Q \VS{9}“As
far as I am concerned,
this
is like
in Noah’s
time,
\par }{\Q when I vowed
that
the waters
of Noah’s
flood would never again
cover the earth.
\par }{\Q In the same
way I have vowed
that I will not be angry
at you or shout at you.
\par }{\Q \VS{10}Even if
the mountains
are removed
\par }{\Q and the hills
displaced,
\par }{\Q my devotion
will not
be
removed
from you,
\par }{\Q nor will my covenant
of friendship
be displaced,”
\par }{\Q says
the {\ND{Lord}}, the one who has compassion on you.
\par }{\Q \VS{11}“O afflicted
one, driven away, and unconsoled!
\par }{\Q Look,
I am
about to set your stones
in antimony
\par }{\Q and I lay
your foundation
with lapis-lazuli.
\par }{\Q \VS{12}I will make
your pinnacles
out of gems,
\par }{\Q your gates out
of beryl,
\par }{\Q and your outer
wall out
of beautiful
stones.
\par }{\Q \VS{13}All
your children
will be followers
of the {\ND{Lord}},
\par }{\Q and your children will enjoy great
prosperity.
\par }{\Q \VS{14}You will be reestablished when I
vindicate
you.

\par }{\Q You will not experience oppression;
\par }{\Q indeed,
you will not
be afraid.
\par }{\Q You will not be terrified,

\par }{\Q for
nothing frightening
will come near you.
\par }{\Q \VS{15}If
anyone dares
to challenge
you, it will not
be my doing!
\par }{\Q Whoever
tries
to challenge
you
will
be defeated.
\par }{\Q \VS{16}Look,
I
create
the craftsman,
\par }{\Q who fans
the coals
into a fire
\par }{\Q and forges
a weapon.
\par }{\Q I
create
the destroyer
so he might devastate.
\par }{\Q \VS{17}No
weapon
forged
to be used
against
you will succeed;
\par }{\Q you will refute
everyone
who tries
to accuse
you.

\par }{\Q This
is what the
{\ND{Lord}}
will do for his servants
–
\par }{\Q I will vindicate
them,”

\par }{\Q says
the {\ND{Lord}}.

\par }\Chap{55}{\PP \par }{\Q \VerseOne{1}“Hey,
all
who are thirsty,
come
to the water!
\par }{\Q You who have
no
money,
come!
\par }{\Q Buy
and eat!
\par }{\Q Come! Buy
wine
and milk
\par }{\Q without
money
and without
cost!
\par }{\Q \VS{2}Why
pay
money
for something that will not
nourish
you?

\par }{\Q Why spend your hard-earned money
on something that will not
satisfy?
\par }{\Q Listen
carefully
to
me and eat
what is nourishing!

\par }{\Q Enjoy
fine food!
\par }{\Q \VS{3}Pay attention
and come
to
me!
\par }{\Q Listen,
so you
can live!

\par }{\Q Then I will make
an unconditional covenantal promise
to you,
\par }{\Q just like the reliable
covenantal
promises
I made to David.
\par }{\Q \VS{4}Look,
I made him a witness
to nations,
\par }{\Q a ruler
and commander
of nations.”
\par }{\Q \VS{5}Look,
you will summon
nations
you did not
previously know;
\par }{\Q nations
that
did not
previously know
you will run
to you,
\par }{\Q because
of the {\ND{Lord}}
your God,
\par }{\Q the Holy One
of Israel,
\par }{\Q for
he bestows honor on you.
\par }{\Q \VS{6}Seek
the {\ND{Lord}}
while he makes himself available;
\par }{\Q call
to him while
he is nearby!
\par }{\Q \VS{7}The wicked
need to abandon
their lifestyle
\par }{\Q and sinful
people their plans.
\par }{\Q They should return
to
the
{\ND{Lord}}, and he will show mercy
to
them,

\par }{\Q and to their God,
for
he will freely
forgive them.
\par }{\Q \VS{8}“Indeed,
my plans are not
like your plans,
\par }{\Q and my deeds are not
like your deeds,
\par }{\Q \VS{9}for
just as
the sky
is higher than
the earth,
\par }{\Q so
my deeds are superior
to your deeds
\par }{\Q and my plans superior
to your plans.
\VS{10}\par }{\Q The rain
and snow
fall from
the sky
\par }{\Q and do not
return,
\par }{\Q but
instead water
the
earth
\par }{\Q and make it produce
and yield
crops,
\par }{\Q and provide
seed
for
the planter
and food
for those who must eat.
\par }{\Q \VS{11}In the same way,
the promise
that
I make
\par }{\Q does not
return
to
me, having accomplished
nothing.
\par }{\Q No,
it is realized as
I desire
\par }{\Q and is fulfilled
as
I intend.”
\par }{\Q \VS{12}Indeed
you will go out
with joy;
\par }{\Q you will be led
along in peace;
\par }{\Q the mountains
and hills
will give
a joyful shout
before
you,
\par }{\Q and all
the trees
in the field
will clap
their hands.
\par }{\Q \VS{13}Evergreens
will grow
in place of
thorn bushes,
\par }{\Q firs
will grow
in place of
nettles;
\par }{\Q they will be
a monument
to the
{\ND{Lord}},

\par }{\Q a permanent
reminder
that will remain.

\par }\Chap{56}{\PP \par }{\Q \VerseOne{1}This is what
the {\ND{Lord}}
says,
\par }{\Q “Promote
justice! Do
what is right!
\par }{\Q For
I am ready
to deliver you;
\par }{\Q I am ready
to vindicate
you openly.
\par }{\Q \VS{2}The people
who do
this
will be blessed,
\par }{\Q the people
who commit themselves to obedience,
\par }{\Q who observe
the Sabbath
and do not defile
it,
\par }{\Q who refrain
from
doing
anything
that is wrong.
\par }{\Q \VS{3}No
foreigner
who becomes a follower
of the
{\ND{Lord}}
should say,
\par }{\Q ‘The
{\ND{Lord}}
will certainly exclude
me from his people.’
\par }{\Q The eunuch
should not
say,
\par }{\Q ‘Look,
I am
like
a dried-up
tree.’ ”
\par }{\Q \VS{4}For
this is what
the {\ND{Lord}}
says:

\par }{\Q “For the eunuchs
who
observe
my Sabbaths
\par }{\Q and choose
what
pleases
me
\par }{\Q and are faithful
to my covenant,
\par }{\Q \VS{5}I will set up
within my temple
and my walls
a monument
\par }{\Q that will be better
than sons
and daughters.
\par }{\Q I will set up
a permanent
monument
for them that
will remain.
\par }{\Q \VS{6}As for foreigners
who become followers
of the
{\ND{Lord}}
and serve
him,
\par }{\Q who love
the
name
of the {\ND{Lord}}
and want to be
his servants
–
\par }{\Q all
who observe
the Sabbath
and do not defile
it,
\par }{\Q and who are faithful
to my covenant –
\par }{\Q \VS{7}I will bring
them to
my holy
mountain;
\par }{\Q I will make them happy
in the temple
where people pray
to me.

\par }{\Q Their burnt offerings
and sacrifices
will be accepted
on
my altar,
\par }{\Q for
my temple
will be known as a temple
where all
nations
may pray.”
\par }{\Q \VS{8}The sovereign
{\SC{
Lord}}
says
this,
\par }{\Q the one who gathers
the dispersed
of Israel:
\par }{\Q “I will still
gather
them up.”
\par }{\SH The Lord Denounces Israel’s Paganism
\par }{\Q \VS{9}All
you wild animals
in the fields,
come
and devour,
\par }{\Q all
you wild animals
in the forest!
\par }{\Q \VS{10}All
their watchmen
are blind,
\par }{\Q they are unaware.
\par }{\Q All
of them are like mute
dogs,
\par }{\Q unable
to bark.
\par }{\Q They pant,
lie down,
\par }{\Q and love
to snooze.
\par }{\Q \VS{11}The dogs
have big appetites;
\par }{\Q they are never
full.
\par }{\Q They
are shepherds
who have no
understanding;
\par }{\Q they all
go their own way,
\par }{\Q each
one looking for monetary
gain.
\par }{\Q \VS{12}Each one says,

\par }{\Q ‘Come
on, I’ll get
some wine!
\par }{\Q Let’s guzzle
some beer!
\par }{\Q Tomorrow
will be
just like today!
\par }{\Q We’ll have everything we want!’

\par }\Chap{57}{\PP \par }{\Q \VerseOne{1}The godly
perish,
\par }{\Q but no
one
cares.
\par }{\Q Honest
people
disappear,
\par }{\Q when no
one minds
\par }{\Q that
the godly
disappear
because of evil.
\par }{\Q \VS{2}Those who live
uprightly
enter
a place of peace;
\par }{\Q they rest
on
their beds.
\par }{\Q \VS{3}But
approach,
you
sons
of omen
readers,
\par }{\Q you offspring
of adulteresses
and prostitutes!
\par }{\Q \VS{4}At
whom
are you laughing?
\par }{\Q At
whom
are you opening
your mouth
\par }{\Q and sticking out
your tongue?
\par }{\Q You
are the children
of rebels,
\par }{\Q the offspring
of liars,
\par }{\Q \VS{5}you who practice ritual sex
under
the oaks and every
green
tree,
\par }{\Q who slaughter
children
near the streams
under
the rocky
overhangs.
\par }{\Q \VS{6}Among the smooth
stones of the stream
are
the idols
you love;
\par }{\Q they,
they
are
the object
of your devotion.

\par }{\Q You pour
out liquid offerings
to them,
\par }{\Q you make an offering.
\par }{\Q Because
of these
things I will seek vengeance.
\par }{\Q \VS{7}On
every high,
elevated
hill
you prepare
your bed;
\par }{\Q you go up
there
to offer
sacrifices.
\par }{\Q \VS{8}Behind
the door
and doorpost
you put
your symbols.
\par }{\Q Indeed,
you depart from
me and go up
\par }{\Q and invite
them into bed
with you.

\par }{\Q You purchase favors from
them,
\par }{\Q you love
their bed,
\par }{\Q and gaze longingly
on their genitals.
\par }{\Q \VS{9}You take
olive oil
as tribute
to your king,
\par }{\Q along with many
perfumes.
\par }{\Q You send
your messengers
to
a distant
place;
\par }{\Q you go
all the way to
Sheol.
\par }{\Q \VS{10}Because
of the long distance
you must travel,
you get tired,
\par }{\Q but you do not
say,
‘I give up.’

\par }{\Q You get renewed
energy,
\par }{\Q so
you don’t
collapse.
\par }{\Q \VS{11}Whom
are you worried
about?
\par }{\Q Whom do you fear,
that
you would act so deceitfully
\par }{\Q and not
remember
me
\par }{\Q or
think
about me?

\par }{\Q Because
I
have been silent
for so long,
\par }{\Q you are not
afraid of me.
\par }{\Q \VS{12}I
will denounce
your so-called righteousness
and your deeds,
\par }{\Q but they will not
help you.
\par }{\Q \VS{13}When you cry out
for help,
let your idols
help you!
\par }{\Q The wind
blows
them all
away,
\par }{\Q a breeze
carries
them away.
\par }{\Q But the one who looks
to me for help
will inherit
the land
\par }{\Q and will have access
to my holy
mountain.”
\par }{\Q \VS{14}He says,
\par }{\Q “Build it! Build it! Clear
a way!
\par }{\Q Remove
all the obstacles
out of the way
of my people!”
\par }{\Q \VS{15}For
this is what
the high
and exalted
one says,
\par }{\Q the one who rules
forever,
whose name
is holy:
\par }{\Q “I dwell
in an exalted
and holy
place,
\par }{\Q but also with
the discouraged
and humiliated,
\par }{\Q in order to cheer
up the humiliated
\par }{\Q and to encourage
the discouraged.
\par }{\Q \VS{16}For
I will not
be hostile forever
\par }{\Q or
perpetually
angry,
\par }{\Q for
then man’s spirit
would grow faint before
me,

\par }{\Q the life-giving
breath
I
created.
\par }{\Q \VS{17}I was
angry because
of their sinful
greed;
\par }{\Q I attacked
them and angrily rejected
them,

\par }{\Q yet they remained disobedient
and stubborn.
\par }{\Q \VS{18}I have seen
their behavior,
\par }{\Q but
I will heal
them
and give them rest,
\par }{\Q and I will once again console
those who mourn.
\par }{\Q \VS{19}I am the one who gives
them reason to celebrate.
\par }{\Q Complete prosperity
is available both to those who are far
away and those who are nearby,”
\par }{\Q says
the {\ND{Lord}}, “and I will heal them.
\par }{\Q \VS{20}But the wicked
are like a surging
sea
\par }{\Q that
is unable
to be quiet;
\par }{\Q its waves
toss up
mud
and sand.
\par }{\Q \VS{21}There will be no
prosperity,”
says
my God,
“for the wicked.”

\par }\Chap{58}{\PP \par }{\Q \VerseOne{1}“Shout loudly! Don’t
be quiet!
\par }{\Q Yell
as loud as a trumpet!
\par }{\Q Confront
my people
with their rebellious
deeds;

\par }{\Q confront Jacob’s
family
with their sin!
\par }{\Q \VS{2}They seek
me day
after day;
\par }{\Q they want
to know
my requirements,

\par }{\Q like
a nation
that
does what is right
\par }{\Q and does
not
reject
the law of their God.
\par }{\Q They ask
me for just
decrees;
\par }{\Q they want
to be near
God.
\par }{\Q \VS{3}They lament, ‘Why
don’t
you notice
when we fast?
\par }{\Q Why don’t
you pay attention when
we humble
ourselves?’

\par }{\Q Look,
at the same time
you fast,
you satisfy
your selfish desires,

\par }{\Q you oppress
your workers.
\par }{\Q \VS{4}Look,
your fasting
is accompanied
by arguments,
brawls,
\par }{\Q and fistfights.
\par }{\Q Do not
fast
as you do today,
\par }{\Q trying to make your voice
heard
in heaven.
\par }{\Q \VS{5}Is this
really the kind of fasting
I want?

\par }{\Q Do I want
a day
when people
merely humble
themselves,
\par }{\Q bowing
their heads
like a reed
\par }{\Q and stretching
out on sackcloth
and ashes?
\par }{\Q Is this
really what you call
a fast,
\par }{\Q a day
that is pleasing
to the
{\ND{Lord}}?
\par }{\Q \VS{6}No,
this
is the kind of fast
I want.
\par }{\Q I want you to remove the sinful chains,
\par }{\Q to tear away the ropes of the burdensome yoke,
\par }{\Q to set free the oppressed,
\par }{\Q and to break every burdensome yoke.
\par }{\Q \VS{7}I want you to share
your food
with the hungry
\par }{\Q and to provide
shelter
for
homeless,
oppressed
people.

\par }{\Q When you see
someone naked,
clothe
him!
\par }{\Q Don’t
turn
your back
on your own flesh and blood!
\par }{\Q \VS{8}Then
your light
will shine
like the sunrise;
\par }{\Q your restoration
will quickly
arrive;
\par }{\Q your godly
behavior will go
before
you,
\par }{\Q and the
{\ND{Lord}}’s
splendor
will be your rear guard.
\par }{\Q \VS{9}Then
you will call
out, and the
{\ND{Lord}}
will respond;
\par }{\Q you will cry out,
and he will reply,
‘Here I am.’
\par }{\Q You must
remove
the burdensome yoke
from among
you
\par }{\Q and stop pointing
fingers
and speaking
sinfully.
\par }{\Q \VS{10}You must actively
help the hungry
\par }{\Q and feed the oppressed.
\par }{\Q Then your
light
will dispel
the darkness,
\par }{\Q and your darkness will be transformed
into noonday.
\par }{\Q \VS{11}The
{\ND{Lord}}
will continually
lead
you;
\par }{\Q he will feed
you even in parched regions.
\par }{\Q He will give
you renewed strength,
\par }{\Q and you will be
like a well-watered
garden,
\par }{\Q like a spring
that
continually produces water.
\par }{\Q \VS{12}Your perpetual
ruins
will be rebuilt;
\par }{\Q you will reestablish
the ancient
foundations.
\par }{\Q You will be called,
‘The one
who
repairs broken walls,
\par }{\Q the one
who makes the streets inhabitable again.’
\par }{\Q \VS{13}You must
observe
the Sabbath
\par }{\Q rather than doing
anything you please
on my holy
day.
\par }{\Q You must look forward
to the Sabbath
\par }{\Q and treat the
{\ND{Lord}}’s
holy day
with respect.
\par }{\Q You must treat it with respect
by refraining
from your normal activities,
\par }{\Q and by refraining
from your selfish pursuits
and from making
business deals.
\par }{\Q \VS{14}Then
you will find joy
in
your relationship to the
{\ND{Lord}},

\par }{\Q and I will give you great prosperity,
\par }{\Q and cause crops
to grow on the land
I gave to your ancestor
Jacob.”
\par }{\Q Know for
certain that the
{\ND{Lord}}
has spoken.

\par }\Chap{59}{\PP \par }{\Q \VerseOne{1}Look,
the
{\ND{Lord}}’s
hand
is not
too weak
to deliver
you;
\par }{\Q his ear
is not
too deaf
to hear you.
\par }{\Q \VS{2}But
your sinful
acts have alienated
you from your God;
\par }{\Q your sins
have caused him to reject
you and not listen to your prayers.
\par }{\Q \VS{3}For
your hands
are stained
with blood
\par }{\Q and your fingers
with sin;
\par }{\Q your lips
speak
lies,
\par }{\Q your tongue
utters
malicious words.
\par }{\Q \VS{4}No
one is concerned
about justice;
\par }{\Q no
one sets forth
his case truthfully.
\par }{\Q They depend
on
false words
and tell
lies;
\par }{\Q they conceive
of oppression
\par }{\Q and give birth
to sin.
\par }{\Q \VS{5}They hatch
the eggs
of a poisonous snake
\par }{\Q and spin
a spider’s
web.
\par }{\Q Whoever eats
their eggs
will die,
\par }{\Q a poisonous snake
is hatched.
\par }{\Q \VS{6}Their webs
cannot
be used for clothing;
\par }{\Q they cannot
cover
themselves with what
they make.
\par }{\Q Their deeds
are sinful;
\par }{\Q they commit violent crimes.
\par }{\Q \VS{7}They are eager
to do evil,
\par }{\Q quick
to shed
innocent
blood.
\par }{\Q Their thoughts
are sinful;
\par }{\Q they crush
and destroy.
\par }{\Q \VS{8}They are
unfamiliar
with peace;
\par }{\Q their deeds are unjust.

\par }{\Q They use deceitful
methods,
\par }{\Q and whoever deals
with them
is unfamiliar
with peace.
\par }{\SH Israel Confesses its Sin
\par }{\Q \VS{9}For this
reason deliverance is far
from
us

\par }{\Q and salvation
does not
reach
us.
\par }{\Q We wait
for light,
but see
only darkness;
\par }{\Q we wait for a bright
light, but live in deep darkness.
\par }{\Q \VS{10}We grope
along the wall
like the blind,
\par }{\Q we grope
like those who cannot see;
\par }{\Q we stumble
at noontime
as if it were evening.
\par }{\Q Though others are strong,
we are like dead men.
\par }{\Q \VS{11}We all
growl
like bears,
\par }{\Q we coo
mournfully
like doves;
\par }{\Q we wait
for deliverance,
but there is none,
\par }{\Q for salvation,
but it is far
from us.
\par }{\Q \VS{12}For
you are aware
of our
many
rebellious
deeds,
\par }{\Q and our sins
testify
against us;
\par }{\Q indeed,
we are aware of our rebellious
deeds;
\par }{\Q we know
our sins
all too well.
\par }{\Q \VS{13}We have rebelled
and tried to deceive
the {\ND{Lord}};
\par }{\Q we turned back
from following
our God.
\par }{\Q We stir up oppression
and rebellion;
\par }{\Q we tell
lies
we concocted
in our minds.
\par }{\Q \VS{14}Justice
is driven
back;
\par }{\Q godliness
stands
far
off.
\par }{\Q Indeed,
honesty stumbles
in the city square
\par }{\Q and morality
is not
even able
to enter.
\par }{\Q \VS{15}Honesty
has disappeared;
\par }{\Q the one who tries
to avoid evil
is robbed.
\par }{\Q The
{\ND{Lord}}
watches
and is displeased,
\par }{\Q for
there is no
justice.
\par }{\SH The Lord Intervenes
\par }{\Q \VS{16}He sees
there is no
advocate;
\par }{\Q he is shocked
that
no
one intervenes.
\par }{\Q So he takes matters
into his
own hands;

\par }{\Q his desire
for justice
drives him on.
\par }{\Q \VS{17}He wears
his desire for justice
like body armor,
\par }{\Q and his desire to deliver
is like a helmet
on his head.
\par }{\Q He puts on
the garments
of vengeance
\par }{\Q and wears
zeal
like a robe.
\par }{\Q \VS{18}He repays
them for what they have done,
\par }{\Q dispensing
angry
judgment
to his adversaries
\par }{\Q and punishing
his enemies.
\par }{\Q He repays
the coastlands.
\par }{\Q \VS{19}In the west,
people respect
the

{\ND{Lord}}’s
reputation;
\par }{\Q in the east
they recognize his splendor.
\par }{\Q For
he comes
like a rushing
stream
\par }{\Q driven
on by wind
sent from the
{\ND{Lord}}.
\par }{\Q \VS{20}“A protector
comes
to Zion,
\par }{\Q to those in Jacob
who repent
of their rebellious
deeds,” says
the {\ND{Lord}}.
\par }{\PP \VS{21}“As for me,
this
is my promise
to them,”
says
the {\ND{Lord}}. “My spirit,
who
is upon
you, and my words,
which
I have placed
in your mouth,
will not
depart
from your
mouth
or from the mouths of your
children
and descendants
from this time forward,”
says
the {\ND{Lord}}.

\par }\Chap{60}{\PP \par }{\Q \VerseOne{1}“Arise! Shine! For
your light
arrives!
\par }{\Q The splendor
of the {\ND{Lord}}
shines
on you!
\par }{\Q \VS{2}For,
look,
darkness
covers
the earth
\par }{\Q and deep darkness
covers the nations,
\par }{\Q but the
{\ND{Lord}}
shines
on
you;
\par }{\Q his splendor
appears
over you.
\par }{\Q \VS{3}Nations
come
to your light,
\par }{\Q kings
to your bright
light.
\par }{\Q \VS{4}Look
all around
you!

\par }{\Q They all
gather
and come to you –
\par }{\Q your sons come from far away
\par }{\Q and your daughters are escorted by guardians.
\par }{\Q \VS{5}Then
you will look
and smile,

\par }{\Q you will be excited
and your
heart
will swell with pride.

\par }{\Q For
the riches of distant lands
will belong to you
\par }{\Q and the wealth
of nations
will come to you.
\par }{\Q \VS{6}Camel
caravans
will cover
your roads,

\par }{\Q young camels
from Midian
and Ephah.
\par }{\Q All
the merchants of Sheba
will come,
\par }{\Q bringing gold
and incense
\par }{\Q and singing
praises
to the
{\ND{Lord}}.
\par }{\Q \VS{7}All
the sheep
of Kedar
will be gathered
to you;
\par }{\Q the rams
of Nebaioth
will be available
to you as sacrifices.

\par }{\Q They will go up
on
my altar
acceptably,
\par }{\Q and I will bestow honor
on my majestic
temple.
\par }{\Q \VS{8}Who
are these
who float along like a cloud,
\par }{\Q who fly
like doves
to
their shelters?
\par }{\Q \VS{9}Indeed,
the coastlands
look
eagerly for me,
\par }{\Q the large ships
are in the lead,
\par }{\Q bringing
your sons
from far away,
\par }{\Q along with
their silver
and gold,
\par }{\Q to honor
the {\ND{Lord}}
your God,
\par }{\Q the Holy One
of Israel,
for
he has bestowed honor on you.
\par }{\Q \VS{10}Foreigners
will rebuild
your walls;
\par }{\Q their kings
will serve
you.
\par }{\Q Even though
I struck
you down in my anger,
\par }{\Q I will restore my favor
and have compassion on you.
\par }{\Q \VS{11}Your gates
will remain open
at all times;
\par }{\Q they will not
be shut
during the day
or at night,
\par }{\Q so that the wealth
of nations
may be delivered,
\par }{\Q with their kings
leading the way.
\par }{\Q \VS{12}Indeed,
nations
or kingdoms
that
do not
serve
you will perish;
\par }{\Q such nations
will be totally destroyed.
\par }{\Q \VS{13}The splendor
of Lebanon
will come
to you,
\par }{\Q its evergreens,
firs,
and cypresses
together,
\par }{\Q to beautify
my palace;
\par }{\Q I will bestow honor on my throne room.
\par }{\Q \VS{14}The children of
your oppressors
will come
bowing
to
you;
\par }{\Q all
who treated
you with disrespect
will bow down
at your feet.
\par }{\Q They will call
you, ‘The City
of the
{\ND{Lord}},
\par }{\Q Zion
of the Holy One
of Israel.’
\par }{\Q \VS{15}You were once
abandoned
\par }{\Q and despised,
with no
one passing
through,
\par }{\Q but I will make
you a permanent
source of
pride
\par }{\Q and joy
to coming
generations.
\par }{\Q \VS{16}You will drink
the milk
of nations;
\par }{\Q you will nurse
at the breasts
of kings.
\par }{\Q Then you will recognize
that
I,
the {\ND{Lord}}, am your deliverer,
\par }{\Q your protector,
the powerful ruler
of Jacob.
\par }{\Q \VS{17}Instead
of bronze,
I will bring
you gold,
\par }{\Q instead
of iron,
I will bring
you silver,
\par }{\Q instead
of wood,
I will bring you bronze,
\par }{\Q instead
of stones,
I
will bring
you iron.
\par }{\Q I
will make
prosperity
your overseer,
\par }{\Q and vindication
your sovereign ruler.
\par }{\Q \VS{18}Sounds of violence
will no
longer
be heard
in your land,
\par }{\Q or the sounds of destruction
and devastation
within your borders.
\par }{\Q You will name
your walls,
‘Deliverance,’
\par }{\Q and your gates,
‘Praise.’
\par }{\Q \VS{19}The sun
will no
longer
supply light
for you by day,
\par }{\Q nor
will the moon’s
brightness
shine
on you;
\par }{\Q the {\ND{Lord}}
will be your permanent
source of light
–
\par }{\Q the splendor
of your God will shine upon you.
\par }{\Q \VS{20}Your sun
will no
longer
set;
\par }{\Q your moon
will not
disappear;
\par }{\Q the {\ND{Lord}}
will be
your permanent
source
of light;
\par }{\Q your time
of sorrow will be over.
\par }{\Q \VS{21}All
of your people
will be godly;
\par }{\Q they will possess
the land
permanently.
\par }{\Q I will plant
them like a shoot;
\par }{\Q they will be the product
of my labor,
\par }{\Q through
whom I reveal my splendor.
\par }{\Q \VS{22}The least
of you will multiply into a thousand;
\par }{\Q the smallest
of you will become
a large
nation.
\par }{\Q When the right time
comes, I
the {\ND{Lord}}
will quickly do this!”

\par }\Chap{61}{\PP \par }{\Q \VerseOne{1}The spirit
of the sovereign

{\ND{Lord}}
is upon
me,
\par }{\Q because
the {\ND{Lord}}
has chosen
me.

\par }{\Q He has commissioned
me to encourage
the poor,
\par }{\Q to help the brokenhearted,
\par }{\Q to decree
the release
of captives,
\par }{\Q and the freeing of
prisoners,
\par }{\Q \VS{2}to announce
the year
when the
{\ND{Lord}}
will show his favor,
\par }{\Q the day
when our God
will seek vengeance,
\par }{\Q to console
all
who mourn,
\par }{\Q \VS{3}to strengthen
those who mourn
in Zion,
\par }{\Q by giving
them a turban,
instead
of ashes,
\par }{\Q oil
symbolizing joy,
instead
of mourning,
\par }{\Q a garment
symbolizing praise,
instead
of discouragement.
\par }{\Q They will be called
oaks
of righteousness,
\par }{\Q trees planted
by the
{\ND{Lord}}
to reveal
his splendor.
\par }{\Q \VS{4}They will rebuild
the perpetual
ruins
\par }{\Q and restore the places that were desolate;
\par }{\Q they will reestablish
the ruined
cities,
\par }{\Q the places that have been desolate since ancient times.
\par }{\Q \VS{5}“Foreigners
will take care
of your sheep;
\par }{\Q foreigners
will work in your fields
and vineyards.
\par }{\Q \VS{6}You
will be called,
‘the
{\ND{Lord}}’s
priests,
\par }{\Q servants
of our God.’
\par }{\Q You will enjoy
the wealth
of nations
\par }{\Q and boast
about the riches you receive from them.
\par }{\Q \VS{7}Instead
of shame,
you will get a double
portion;

\par }{\Q instead of humiliation,
they will rejoice
over the land
they receive.
\par }{\Q Yes, they will possess
a double
portion
in their land
\par }{\Q and experience lasting
joy.
\par }{\Q \VS{8}For
I,
the {\ND{Lord}}, love
justice
\par }{\Q and hate
robbery
and sin.
\par }{\Q I will repay
them because
of my faithfulness;
\par }{\Q I will make
a permanent
covenant with them.
\par }{\Q \VS{9}Their descendants
will be known
among the nations,
\par }{\Q their offspring
among
the peoples.
\par }{\Q All
who see
them will recognize
that
\par }{\Q the {\ND{Lord}}
has blessed them.”
\par }{\Q \VS{10}I will
greatly
rejoice in
the {\ND{Lord}};
\par }{\Q I will be overjoyed because
of my God.
\par }{\Q For
he clothes
me
in garments
of deliverance;
\par }{\Q he puts on me
a robe
symbolizing
vindication.
\par }{\Q I look like a bridegroom
when
he wears
a turban
as
a priest
would;
\par }{\Q I look like a bride
when
she puts
on her jewelry.
\par }{\Q \VS{11}For just as
the ground
produces
its crops
\par }{\Q and a garden
yields
its produce,
\par }{\Q so
the sovereign

{\ND{Lord}}
will cause deliverance
to grow,
\par }{\Q and give his people reason to praise
him in the sight
of all
the nations.

\par }\Chap{62}{\PP \par }{\Q \VerseOne{1}“For the sake
of Zion
I will not
be silent;
\par }{\Q for the sake
of Jerusalem
I will not
be quiet,
\par }{\Q until
her vindication shines
brightly
\par }{\Q and her deliverance
burns
like a torch.”
\par }{\Q \VS{2}Nations
will see
your vindication,
\par }{\Q and all
kings
your splendor.
\par }{\Q You will be called
by a new
name
\par }{\Q that
the {\ND{Lord}}
himself will give you.
\par }{\Q \VS{3}You will be
a majestic
crown
in the hand
of the {\ND{Lord}},
\par }{\Q a royal
turban
in the hand
of your God.
\par }{\Q \VS{4}You will no
longer
be called,
“Abandoned,”
\par }{\Q and your land
will no
longer
be called
“Desolate.”
\par }{\Q Indeed,
you will be called
“My Delight is in Her,”

\par }{\Q and your land
“Married.”
\par }{\Q For
the {\ND{Lord}}
will take delight
in you,
\par }{\Q and your land
will be married to him.
\par }{\Q \VS{5}As
a young man
marries
a young woman,
\par }{\Q so your sons
will marry
you.
\par }{\Q As a bridegroom
rejoices
over
a bride,
\par }{\Q so your God
will rejoice
over you.
\par }{\Q \VS{6}I post watchmen on
your walls,
O Jerusalem;
\par }{\Q they should keep
praying
all day and all night.
\par }{\Q You who pray
to the
{\ND{Lord}}, don’t
be silent!
\par }{\Q \VS{7}Don’t
allow
him to rest
until
he reestablishes
Jerusalem,

\par }{\Q until
he makes
Jerusalem
the pride
of the earth.
\par }{\Q \VS{8}The
{\ND{Lord}}
swears
an oath by his right hand,
\par }{\Q by his strong
arm:

\par }{\Q “I will never
again
give
your grain
\par }{\Q to your enemies
as food,
\par }{\Q and foreigners
will not
drink
your wine,
\par }{\Q which
you worked hard to produce.
\par }{\Q \VS{9}But
those who harvest the grain
will eat
it,
\par }{\Q and will praise
the {\ND{Lord}}.
\par }{\Q Those who pick the grapes
will drink
the wine

\par }{\Q in the courts
of my holy sanctuary.”
\par }{\Q \VS{10}Come through! Come through
the gates!
\par }{\Q Prepare
the way
for the people!
\par }{\Q Build
it! Build
the roadway!
\par }{\Q Remove
the stones!
\par }{\Q Lift
a signal flag
for
the nations!
\par }{\Q \VS{11}Look,
the {\ND{Lord}}
announces
to
the entire earth:

\par }{\Q “Say
to Daughter
Zion,
\par }{\Q ‘Look,
your deliverer
comes!
\par }{\Q Look,
his reward
is with
him
\par }{\Q and his reward
goes before him!’ ”
\par }{\Q \VS{12}They will be called,
“The Holy
People,
\par }{\Q the Ones Protected
by the
{\ND{Lord}}.”
\par }{\Q You will be called,
“Sought
After,
\par }{\Q City
Not
Abandoned.”

\par }\Chap{63}{\PP \par }{\Q \VerseOne{1}Who
is this
who comes
from Edom,
\par }{\Q dressed
in bright red, coming from Bozrah?

\par }{\Q Who is this one
wearing
royal attire,
\par }{\Q who marches confidently because
of his great
strength?
\par }{\Q “It is I,
the one who announces
vindication,
\par }{\Q and who is able
to deliver!”
\par }{\Q \VS{2}Why
are your clothes
red?
\par }{\Q Why do you look like someone who has stomped
on grapes in a vat?
\par }{\Q \VS{3}“I have stomped
grapes in the winepress
all by myself;
\par }{\Q no
one
from the nations
joined
me.
\par }{\Q I stomped
on them in my anger;
\par }{\Q I trampled
them down in my rage.
\par }{\Q Their juice
splashed
on
my garments,
\par }{\Q and stained
all
my clothes.
\par }{\Q \VS{4}For
I looked forward to the day
of vengeance,
\par }{\Q and then payback
time arrived.
\par }{\Q \VS{5}I looked,
but there was no
one to help;
\par }{\Q I was shocked
because there was no
one offering support.
\par }{\Q So my right arm
accomplished deliverance;
\par }{\Q my raging anger
drove me on.
\par }{\Q \VS{6}I trampled
nations
in my anger,
\par }{\Q I made them drunk
in my rage,
\par }{\Q I splashed
their blood
on the ground.”
\par }{\SH A Prayer for Divine Intervention
\par }{\Q \VS{7}I
will tell of the faithful
acts
of the {\ND{Lord}},
\par }{\Q of the {\ND{Lord}}’s praiseworthy deeds.
\par }{\Q I will tell about all the
{\ND{Lord}} did for us,
\par }{\Q the many good things he did for the family of Israel,
\par }{\Q because of his compassion and great faithfulness.
\par }{\Q \VS{8}He said,
“Certainly
they will be my people,
\par }{\Q children
who are
not
disloyal.”
\par }{\Q He became their
deliverer.
\par }{\Q \VS{9}Through
all
that they suffered,
he suffered
too.

\par }{\Q The messenger
sent from his very presence
delivered
them.
\par }{\Q In his love
and mercy
he protected
them;
\par }{\Q he lifted them
up and carried
them throughout
ancient
times.
\par }{\Q \VS{10}But they
rebelled
and offended
his holy
Spirit,
\par }{\Q so he turned
into an enemy
\par }{\Q and fought against them.
\par }{\Q \VS{11}His people
remembered
the ancient
times.
\par }{\Q Where
is the one who brought
them up out
of the sea,
\par }{\Q along with
the shepherd
of his flock?
\par }{\Q Where
is the one who placed
his holy
Spirit
among them,
\par }{\Q \VS{12}the one who made his majestic
power
available to Moses,
\par }{\Q who divided
the water
before
them,
\par }{\Q gaining
for himself a lasting
reputation,
\par }{\Q \VS{13}who led
them through the deep
water?
\par }{\Q Like a horse
running on flat land
they did not
stumble.
\par }{\Q \VS{14}Like an animal
that goes down
into a valley
to graze,

\par }{\Q so the Spirit
of the {\ND{Lord}}
granted them rest.
\par }{\Q In this way
you guided
your people,
\par }{\Q gaining
for yourself an honored reputation.
\par }{\Q \VS{15}Look down
from heaven
and take notice,
\par }{\Q from your holy,
majestic
palace!
\par }{\Q Where
are your zeal
and power?
\par }{\Q Do not hold back
your tender
compassion!
\par }{\Q \VS{16}For
you
are our father,
\par }{\Q though
Abraham
does not
know
us
\par }{\Q and Israel
does not
recognize
us.
\par }{\Q You,

{\ND{Lord}}, are our father;
\par }{\Q you have been called
our protector
from ancient times.
\par }{\Q \VS{17}Why,

{\ND{Lord}}, do you make us stray
from your ways,
\par }{\Q and make our minds
stubborn so that we do not obey
you?

\par }{\Q Return
for the sake
of your servants,
\par }{\Q the tribes
of your inheritance!
\par }{\Q \VS{18}For a short
time your special
nation
possessed
a land,

\par }{\Q but then our adversaries
knocked down
your holy sanctuary.
\par }{\Q \VS{19}We existed
from ancient
times,

\par }{\Q but you did not
rule
over them,
\par }{\Q they were not
your subjects.

\par }\Chap{64}{\PP \par }{\Q \VerseOne{1}If only you would tear
apart the sky
and come down!
\par }{\Q The mountains
would tremble
before you!
\par }{\Q \VS{2}As when fire
ignites
dry wood,
\par }{\Q or fire
makes water
boil,
\par }{\Q let your adversaries
know
who you
are,

\par }{\Q and may the nations
shake
at your presence!
\par }{\Q \VS{3}When
you performed
awesome
deeds that took us by surprise,
\par }{\Q you came down,
and the mountains
trembled
before you.
\par }{\Q \VS{4}Since ancient
times no
one has heard
or
perceived,
\par }{\Q no
eye
has seen
any God
besides
you,
\par }{\Q who intervenes
for those who wait for him.
\par }{\Q \VS{5}You assist
those who delight
in doing what
is right,
\par }{\Q who observe
your commandments.
\par }{\Q Look,
you
were angry
because we violated
them continually.
\par }{\Q How then
can we be saved?
\par }{\Q \VS{6}We are
all
like one who is unclean,
\par }{\Q all
our so-called righteous
acts are like a menstrual
rag
in your sight.

\par }{\Q We all
wither
like a leaf;
\par }{\Q our sins
carry
us away
like the wind.
\par }{\Q \VS{7}No
one invokes
your name,
\par }{\Q or
makes an effort
to take hold
of you.
\par }{\Q For
you have rejected
us

\par }{\Q and handed
us over
to our own sins.
\par }{\Q \VS{8}Yet,

{\ND{Lord}}, you
are our father.
\par }{\Q We
are the clay,
and you
are our potter;
\par }{\Q we are all
the product of your labor.
\par }{\Q \VS{9}{\ND{Lord}}, do not
be too angry!
\par }{\Q Do not
hold
our sins
against
us
continually!

\par }{\Q Take a good look
at your people,
at all
of us!
\par }{\Q \VS{10}Your chosen
cities
have become
a desert;
\par }{\Q Zion
has become a desert,
\par }{\Q Jerusalem
is a desolate ruin.
\par }{\Q \VS{11}Our holy
temple,
our pride
and joy,

\par }{\Q the place where
our ancestors
praised
you,
\par }{\Q has been
burned
with fire;
\par }{\Q all
our prized possessions
have been
destroyed.
\par }{\Q \VS{12}In
light of all this,
how can you still hold back,

{\ND{Lord}}?
\par }{\Q How can you be silent
and continue
to humiliate us?

\par }\Chap{65}{\PP \par }{\Q \VerseOne{1}“I made myself available
to those who did not
ask
for me;

\par }{\Q I appeared
to those who did not
look
for me.

\par }{\Q I said,
‘Here I am! Here I am!’
\par }{\Q to
a nation
that did not
invoke
my name.
\par }{\Q \VS{2}I spread
out my hands
all
day
long
\par }{\Q to
my rebellious
people,
\par }{\Q who lived in a way
that is morally
unacceptable,
\par }{\Q and who did what
they desired.
\par }{\Q \VS{3}These people
continually
and blatantly offend
me
\par }{\Q as they sacrifice
in their sacred orchards
\par }{\Q and burn incense
on
brick altars.
\par }{\Q \VS{4}They sit
among the tombs
\par }{\Q and keep
watch all night long.
\par }{\Q They eat
pork,
\par }{\Q and broth
from unclean sacrificial
meat
is in their pans.
\par }{\Q \VS{5}They say,
‘Keep
to
yourself!
\par }{\Q Don’t
get near
me, for
I am holier
than you!’
\par }{\Q These
people are like smoke
in my nostrils,
\par }{\Q like a fire
that keeps burning
all
day long.
\par }{\Q \VS{6}Look,
I have decreed:

\par }{\Q I will not
keep silent,
but
will pay
them back;
\par }{\Q I will pay
them back exactly what they deserve,
\par }{\Q \VS{7}for your sins
and your ancestors’
sins,”
says
the {\ND{Lord}}.
\par }{\Q “Because
they burned incense
on
the mountains
\par }{\Q and offended
me on
the hills,
\par }{\Q I will punish them in full measure.”
\par }{\Q \VS{8}This is what
the {\ND{Lord}}
says:
\par }{\Q “When juice
is discovered
in a cluster
of grapes,
\par }{\Q someone says,
‘Don’t
destroy
it, for
it contains juice.’
\par }{\Q So
I will do
for the sake
of my servants
–
\par }{\Q I will not
destroy
everyone.
\par }{\Q \VS{9}I will bring forth
descendants
from Jacob,
\par }{\Q and from Judah
people to take possession
of my mountains.
\par }{\Q My chosen ones
will take possession
of the land;

\par }{\Q my servants
will live
there.
\par }{\Q \VS{10}Sharon
will become
a pasture
for sheep,
\par }{\Q and the Valley
of Achor
a place
where cattle
graze;

\par }{\Q they will belong to my people,
who
seek me.
\par }{\Q \VS{11}But as for you
who abandon
the {\ND{Lord}}
\par }{\Q and forget
about worshiping at my holy
mountain,
\par }{\Q who prepare
a feast
for the god called ‘Fortune,’
\par }{\Q and fill
up wine jugs
for the god called ‘Destiny’ –
\par }{\Q \VS{12}I predestine
you to die by the sword,
\par }{\Q all
of you will kneel down
at the slaughtering block,
\par }{\Q because
I called
to you, and you did not
respond,
\par }{\Q I spoke
and you did not
listen.
\par }{\Q You did
evil
before me;
\par }{\Q you chose
to do what displeases me.”
\par }{\Q \VS{13}So
this is what
the sovereign

{\ND{Lord}}
says:
\par }{\Q “Look,
my servants
will eat,
but you
will be hungry!
\par }{\Q Look,
my servants
will drink,
but you
will be thirsty!
\par }{\Q Look,
my servants
will rejoice,
but you
will be humiliated!
\par }{\Q \VS{14}Look,
my servants
will shout for joy
as happiness fills
their hearts!

\par }{\Q But you
will cry out
as sorrow
fills your hearts;
\par }{\Q you will wail
because your spirits
will be crushed.
\par }{\Q \VS{15}Your names
will live on
in the curse formulas
of my chosen ones.
\par }{\Q The sovereign

{\ND{Lord}}
will kill
you,
\par }{\Q but he will give
his servants
another
name.
\par }{\Q \VS{16}Whoever
pronounces a blessing
in the earth
\par }{\Q will do so
in the name of the faithful
God;
\par }{\Q whoever makes an oath
in the earth
\par }{\Q will do so
in the name of the faithful
God.
\par }{\Q For
past
problems
will be forgotten;
\par }{\Q I will no longer think about them.
\par }{\Q \VS{17}For
look,
I am ready to create
\par }{\Q new
heavens
and a new
earth!

\par }{\Q The former
ones will not be remembered;
\par }{\Q no one will think
about them anymore.
\par }{\Q \VS{18}But
be happy
and rejoice
forevermore
\par }{\Q over what I
am about to create!
\par }{\Q For
look,
I am ready to create
Jerusalem
to be a source of joy,
\par }{\Q and her people
to be a source of happiness.
\par }{\Q \VS{19}Jerusalem
will bring me joy,
\par }{\Q and my people
will bring me
happiness.
\par }{\Q The sound
of weeping
or cries
of sorrow
\par }{\Q will never
be heard
in her again.
\par }{\Q \VS{20}Never
again
will one of her infants
live just a few days
\par }{\Q or an old man
die before
his time.
\par }{\Q Indeed,
no one will die before the age
of a hundred,
\par }{\Q anyone who fails to reach
the age
of a hundred
will be considered cursed.
\par }{\Q \VS{21}They will build
houses
and live
in them;
\par }{\Q they will plant
vineyards
and eat
their fruit.
\par }{\Q \VS{22}No
longer will they build
a house only to have another
live in it,
\par }{\Q or
plant
a vineyard only to have another
eat
its fruit,

\par }{\Q for
my people
will live as long as
trees,
\par }{\Q and my chosen ones
will enjoy to the fullest
what they have produced.
\par }{\Q \VS{23}They will not
work
in vain,
\par }{\Q or
give birth
to children that will experience disaster.
\par }{\Q For
the
{\ND{Lord}}
will bless
their
children
\par }{\Q and their descendants.
\par }{\Q \VS{24}Before
they even call
out, I
will respond;
\par }{\Q while they are
still
speaking,
I
will hear.
\par }{\Q \VS{25}A wolf
and a lamb
will graze
together;
\par }{\Q a lion,
like an ox,
will eat
straw,
\par }{\Q and a snake’s
food
will be dirt.
\par }{\Q They will no
longer injure
or
destroy
\par }{\Q on
my entire
royal
mountain,”
says
the {\ND{Lord}}.

\par }\Chap{66}{\PP \par }{\Q \VerseOne{1}This is what
the {\ND{Lord}}
says:
\par }{\Q “The heavens
are my throne
\par }{\Q and the earth
is my footstool.
\par }{\Q Where
then is the house
you will build
for me?
\par }{\Q Where
is the place
where I will rest?
\par }{\Q \VS{2}My hand
made
them;
\par }{\Q that is how they came to be,”
says
the {\ND{Lord}}.
\par }{\Q I show special favor to the humble and contrite,
\par }{\Q who respect
what I have to say.
\par }{\Q \VS{3}The one who slaughters
a bull
also strikes down
a man;
\par }{\Q the one who sacrifices
a lamb
also breaks
a dog’s
neck;
\par }{\Q the one who presents
an offering
includes pig’s
blood
with it;

\par }{\Q the one who offers
incense
also
praises
an idol.
\par }{\Q They
have decided
to behave this way;
\par }{\Q they enjoy
these disgusting practices.
\par }{\Q \VS{4}So
I
will choose
severe punishment
for them;
\par }{\Q I will bring
on them what they
dread,
\par }{\Q because
I called,
and no
one responded,
\par }{\Q I spoke
and they did not
listen.
\par }{\Q They did
evil
before
me;

\par }{\Q they chose
to do
what displeases me.”
\par }{\Q \VS{5}Hear
the word
of the {\ND{Lord}},
\par }{\Q you who respect
what
he has to say!

\par }{\Q Your countrymen,
who hate
you
\par }{\Q and exclude you, supposedly
for the sake
of my name,
\par }{\Q say, “May
the {\ND{Lord}}
be glorified,
\par }{\Q then
we will witness
your joy.”
\par }{\Q But they
will be put to shame.
\par }{\Q \VS{6}The sound of battle
comes from the city;
\par }{\Q the sound
comes from the temple!
\par }{\Q It is the sound
of the {\ND{Lord}}
paying
back
his enemies.
\par }{\Q \VS{7}Before
she goes into labor,
she gives birth!
\par }{\Q Before
her contractions
begin, she delivers
a boy!
\par }{\Q \VS{8}Who
has ever heard
of such a thing?
\par }{\Q Who
has ever seen
this?
\par }{\Q Can a country
be brought forth
in one
day?
\par }{\Q Can
a nation
be born
in a single
moment?
\par }{\Q Yet
as soon as Zion
goes into labor she gives birth
to sons!
\par }{\Q \VS{9}“Do I
bring a baby
to the birth
opening and then not
deliver it?”
\par }{\Q asks the
{\ND{Lord}}.
\par }{\Q “Or
do I
bring a baby
to the point of delivery
and then hold it back?”
\par }{\Q asks your God.
\par }{\Q \VS{10}Be happy
for
Jerusalem
\par }{\Q and rejoice
with
her, all
you who love
her!
\par }{\Q Share
in her great joy,
\par }{\Q all
you who have mourned
over her!
\par }{\Q \VS{11}For
you will nurse
from her satisfying
breasts
and be nourished;
\par }{\Q you will feed
with joy
from her milk-filled
breasts.
\par }{\Q \VS{12}For
this is what
the {\ND{Lord}}
says:

\par }{\Q “Look,
I am ready to extend
to her
prosperity
that will flow
like a river,
\par }{\Q the riches
of nations
will flow into her like a stream that floods its banks.
\par }{\Q You will nurse
from her breast and be carried
at her side;
\par }{\Q you will play
on
her knees.
\par }{\Q \VS{13}As
a mother
consoles
a child,

\par }{\Q so
I
will console
you,
\par }{\Q and you will be consoled
over Jerusalem.”
\par }{\Q \VS{14}When you see
this, you will be happy,
\par }{\Q and you will be revived.
\par }{\Q The
{\ND{Lord}}
will reveal
his power
to
his servants
\par }{\Q and his anger
to his enemies.
\par }{\Q \VS{15}For
look,
the {\ND{Lord}}
comes
with fire,
\par }{\Q his chariots
come like a windstorm,
\par }{\Q to reveal
his raging
anger,
\par }{\Q his battle cry,
and his flaming
arrows.
\par }{\Q \VS{16}For
the {\ND{Lord}}
judges
all
humanity
\par }{\Q with fire
and his sword;
\par }{\Q the {\ND{Lord}}
will kill many.
\par }{\PP \VS{17}“As for those who consecrate
and ritually purify
themselves so they can follow
their leader and worship in the sacred orchards,
those who eat
the flesh
of pigs
and other disgusting
creatures, like mice –
they will all be destroyed
together,”
says
the {\ND{Lord}}.
\VS{18}“I
hate their deeds
and thoughts! So I am coming
to gather
all
the nations
and ethnic
groups; they will come
and witness
my splendor.
\VS{19}I will perform
a mighty act
among them and then send
some of those who remain
to
the nations
– to Tarshish,
Pul,
Lud
(known
for its archers
), Tubal,
Javan,
and to the distant
coastlands
that
have not
heard
about me or
seen
my splendor.
They will tell
the nations
of my splendor.
\VS{20}They will bring
back all
your countrymen
from all
the nations
as an offering
to the
{\ND{Lord}}. They will bring them on horses,
in chariots,
in wagons,
on mules,
and on
camels
to my holy
hill
Jerusalem,”
says
the {\ND{Lord}}, “just as the Israelites
bring
offerings
to the
{\ND{Lord}}’s
temple
in ritually pure
containers.
\VS{21}And I will choose some
of them
as
priests
and Levites,”
says
the {\ND{Lord}}.
\VS{22}“For
just
as the new
heavens
and the new
earth
I am
about
to make
will remain standing
before
me,” says
the {\ND{Lord}}, “so
your descendants
and your name
will remain.
\VS{23}From one
month
to the next
and from
one Sabbath
to the next, all
people
will come
to worship
me,”
says
the {\ND{Lord}}.
\VS{24}“They will go out
and observe
the corpses
of those
who rebelled
against me, for
the maggots
that eat them will not
die,
and the fire
that consumes
them will not
die out. All
people
will find the sight abhorrent.”
\par }