\NormalFont\ShortTitle{Jeremiah}
{\MT Jeremiah

\par }\ChapOne{1}{\SH The Superscription
\par }{\PP \VerseOne{1}The following
is a record of what
Jeremiah
son
of Hilkiah
prophesied.
He was one of the priests
who
lived at Anathoth
in the territory
of the tribe of Benjamin.
\VS{2}The
{\ND{Lord}}
began
to speak
to
him in the thirteenth
year
that Josiah
son
of Amon
ruled over
Judah.
\VS{3}The
{\ND{Lord}} also spoke to him when
Jehoiakim
son
of Josiah
ruled over
Judah,
and he continued to speak to him until
the fifth
month
of the eleventh
year
that Zedekiah
son
of Josiah
ruled over
Judah.
That was when the people of Jerusalem
were taken into exile.
\par }{\SH Jeremiah’s Call and Commission
\par }{\PP \VS{4}The
{\ND{Lord}}
said
to me,
\par }{\Q \VS{5}“Before
I formed
you in your mother’s womb
I chose
you.

\par }{\Q Before
you were born
I set you apart.
\par }{\Q I appointed you to be a prophet
to the nations.”
\par }{\PP \VS{6}I answered,
“Oh,
Lord

{\ND{God}}, I really
do not
know
how to speak
well enough for
that, for I am
too young.”
\VS{7}The
{\ND{Lord}}
said
to
me, “Do not
say,
‘I am
too young.’
But
go
to whomever
I send
you and say whatever
I
tell you.
\VS{8}Do not
be afraid
of those to whom I send you, for
I
will be with
you to protect
you,” says
the
{\ND{Lord}}.
\VS{9}Then the
{\ND{Lord}}
reached
out his hand
and touched
my mouth
and said
to me,
“I
will most assuredly
give
you the words
you are to speak for me.
\VS{10}Know for
certain that I hereby
give you the authority to announce to nations
and kingdoms
that they will be uprooted
and torn
down, destroyed
and demolished,
rebuilt
and firmly planted.”
\par }{\SH Visions Confirming Jeremiah’s Call and Commission
\par }{\PP \VS{11}Later
the {\ND{Lord}}
asked
me, “What
do you
see,
Jeremiah?” I answered,
“I
see
a branch
of an almond tree.”
\VS{12}Then the
{\ND{Lord}}
said,
“You have observed
correctly.
This means
I am
watching
to make
sure my threats
are carried out.”
\par }{\PP \VS{13}The
{\ND{Lord}}
again
asked
me, “What
do you
see?” I answered,
“I
see
a pot
of boiling
water; it is tipped toward
us from the north.”
\VS{14}Then the
{\ND{Lord}}
said,
“This means
destruction
will break out from the north
on
all
who live
in the land.
\VS{15}For
I will soon
summon
all
the peoples
of the kingdoms
of the north,”
says
the {\ND{Lord}}. “They will come
and their kings will set
up their thrones
near the entrances
of the gates
of Jerusalem.
They will attack
all
the walls
surrounding
it, and all
the towns
in Judah.
\VS{16}In this way
I will pass sentence
on
the people of Jerusalem and Judah because
of all
their wickedness.
For they rejected
me and offered
sacrifices to other
gods,
worshiping
what they made
with their own hands.”
\par }{\PP \VS{17}“But you,
Jeremiah, get
yourself ready! Go
and tell
these people everything
I
instruct
you to
say. Do not
be terrified
of them, or
I will give you good reason to be terrified of them.
\VS{18}I,
the
{\ND{Lord}}, hereby
promise to make
you as strong as a fortified
city,
an iron
pillar,
and a bronze
wall.
You will be able to stand up against
all
who live in the land,
including the kings
of Judah,
its officials,
its priests
and all the people
of the land.
\VS{19}They will attack
you but they will not
be able
to overcome you, for
I
will be with
you to rescue
you,” says
the {\ND{Lord}}.

\par }\Chap{2}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
me. He said:
\VS{2}“Go
and declare
in the hearing
of the people of Jerusalem: ‘This is what
the
{\ND{Lord}}
says: “I have fond memories
of you, how devoted
you were to me in your early
years. I remember
how you loved
me like a new bride;
you followed
me
through the
wilderness,
through a land
that had
never
been planted.
\VS{3}Israel
was set apart
to the
{\ND{Lord}}; they were like the first fruits
of a harvest
to him. All
who tried to devour
them were punished;
disaster
came
upon them,”
says
the {\ND{Lord}}.’ ”
\par }{\SH The Lord Reminds Them of the Unfaithfulness of Their Ancestors
\par }{\Q \VS{4}Now listen
to what
the Lord
has to say, you descendants of Jacob,
\par }{\Q all
you family
groups from the nation
of Israel.
\par }{\Q \VS{5}This is what
the Lord
says:
\par }{\Q “What
fault
could your ancestors
have possibly
found in me
\par }{\Q that
they strayed so far
from me?

\par }{\Q They paid allegiance
to worthless idols,
and so became worthless to me.
\par }{\Q \VS{6}They did not
ask:
\par }{\Q ‘Where
is the
{\ND{Lord}}
who delivered
us out of Egypt,
\par }{\Q who brought us through
the wilderness,
\par }{\Q through a land
of desert
sands and rift
valleys,
\par }{\Q through a land
of drought
and deep darkness,
\par }{\Q through a land
in which no
one
travels,
\par }{\Q and where
no
one
lives?’
\par }{\Q \VS{7}I brought
you into
a fertile
land
\par }{\Q so you could enjoy
its fruits
and its rich bounty.
\par }{\Q But when you entered
my land, you defiled
it;

\par }{\Q you made
the land
I call my own
loathsome to me.
\par }{\Q \VS{8}Your priests
did not
ask,
‘Where
is the
{\ND{Lord}}?’

\par }{\Q Those responsible
for teaching my law
did not
really know
me.

\par }{\Q Your rulers
rebelled against
me.
\par }{\Q Your prophets
prophesied
in the name of the god Baal.
\par }{\Q They all worshiped
idols that could not
help them.
\par }{\SH The Lord Charges Contemporary Israel with Spiritual Adultery
\par }{\Q \VS{9}“So,
once more
I will state
my case
against you,”
says
the {\ND{Lord}}.
\par }{\Q “I will also state
it against your children
and grandchildren.
\par }{\Q \VS{10}Go
west across
the sea to the coasts
of Cyprus
and see.
\par }{\Q Send
someone east to Kedar
and have them look carefully.
\par }{\Q See
if
such a thing as this has ever happened:
\par }{\Q \VS{11}Has a nation
ever changed
its gods
\par }{\Q (even though they
are not
really gods
at all)?
\par }{\Q But my people
have exchanged
me, their glorious
God,

\par }{\Q for a god
that cannot help them at all!
\par }{\Q \VS{12}Be amazed
at this,
O heavens!

\par }{\Q Be shocked
and utterly
dumbfounded,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{13}“Do
so because
my people
have committed
a double
wrong:

\par }{\Q they have rejected
me,
\par }{\Q the fountain
of life-giving
water,
\par }{\Q and they have dug
cisterns
for themselves,
\par }{\Q cracked
cisterns
which
cannot
even hold
water.”
\par }{\SH Israel’s Reliance on Foreign Alliances (not on God)
\par }{\Q \VS{14}“Israel
is not a slave,
is he?
\par }{\Q He was not born
into slavery, was he?

\par }{\Q If not, why
then is he being
carried off?
\par }{\Q \VS{15}Like lions
his enemies roar
victoriously over
him;
\par }{\Q they raise
their voices
in triumph.

\par }{\Q They have laid his land
waste;
\par }{\Q his cities
have been burned
down and deserted.
\par }{\Q \VS{16}Even
the soldiers from Memphis
and Tahpanhes
\par }{\Q have cracked
your skulls, people of Israel.
\par }{\Q \VS{17}You have
brought all this
on yourself, Israel,

\par }{\Q by deserting
the {\ND{Lord}}
your God
when
he was leading
you along
the right path.
\par }{\Q \VS{18}What
good will it do you then to go down to
Egypt
\par }{\Q to seek help
from the Egyptians?

\par }{\Q What
good will it do you to go over
to Assyria
\par }{\Q to seek help
from the Assyrians?
\par }{\Q \VS{19}Your own wickedness
will bring about
your punishment.
\par }{\Q Your unfaithful acts
will bring down discipline
on you.

\par }{\Q Know,
then, and realize
how
utterly harmful
\par }{\Q it was for you to reject
me, the

{\ND{Lord}}
your God,
\par }{\Q to show
no
respect
for
me,”

\par }{\Q says
the Lord

{\ND{God}}
who rules over all.
\par }{\SH The Lord Expresses His Exasperation at Judah’s Persistent Idolat
\par }{\Q \VS{20}“Indeed,
long ago
you threw off
my authority
\par }{\Q and refused
to be subject to me.

\par }{\Q You said,
‘I will not
serve
you.’
\par }{\Q Instead,
you
gave yourself to other gods on
every
high
hill
\par }{\Q and under
every
green
tree,
\par }{\Q like a prostitute
sprawls out before her lovers.
\par }{\Q \VS{21}I
planted
you in the land
\par }{\Q like a special vine
of the very best stock.
\par }{\Q Why in the world have you turned
into something like a wild vine
\par }{\Q that produces rotten,
foul-smelling grapes?
\par }{\Q \VS{22}You can try to wash
away your guilt with a strong detergent.
\par }{\Q You can use as much
soap
as you want.
\par }{\Q But the stain
of your guilt
is still there for me to see,”

\par }{\Q says
the Lord

{\ND{God}}.
\par }{\Q \VS{23}“How
can you say,
‘I have not
made myself unclean.
\par }{\Q I have not
paid allegiance
to the gods called Baal.’
\par }{\Q Just look
at the way
you
have behaved in the Valley
of Hinnom!

\par }{\Q Think about the things you have
done
there!
\par }{\Q You are like a flighty,
young female camel
\par }{\Q that rushes here and there, crisscrossing
its path.
\par }{\Q \VS{24}You are like a wild female donkey
brought up
in the wilderness.
\par }{\Q In her lust
she sniffs
the wind
to get the scent of a male.

\par }{\Q No
one can
hold her back
when she is
in heat.
\par }{\Q None
of the males need wear themselves out chasing after her.
\par }{\Q At mating time
she is easy
to find.
\par }{\Q \VS{25}Do not chase after
other gods until your shoes wear out
\par }{\Q and your throats
become dry.
\par }{\Q But you say,
‘It is useless
for you to try and stop
me
\par }{\Q because
I love
those foreign
gods and want to pursue them!’
\par }{\Q \VS{26}Just as a thief
has to suffer dishonor
when
he is caught,
\par }{\Q so
the people
of Israel
will suffer
dishonor
for
what they
have done.

\par }{\Q So will their kings
and officials,
\par }{\Q their priests
and their prophets.
\par }{\Q \VS{27}They say
to a wooden
idol, ‘You
are my father.’
\par }{\Q They say to a stone
image, ‘You
gave birth
to me.’

\par }{\Q Yes,
they have
turned
away from me instead of turning
to me.

\par }{\Q Yet when they are in trouble,
they say,
‘Come and save us!’
\par }{\Q \VS{28}But where
are the gods
you made
for yourselves?
\par }{\Q Let them save
you when
you are in trouble.
\par }{\Q The sad fact is that
you have as many
gods
\par }{\Q as you have towns,
Judah.
\par }{\Q \VS{29}“Why
do you try to refute
me?

\par }{\Q All
of you have rebelled
against me,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{30}“It did no good
for me to punish
your people.
\par }{\Q They did not
respond
to such correction.
\par }{\Q You slaughtered
your prophets
\par }{\Q like a voracious
lion.”
\par }{\Q \VS{31}You
people
of this generation,
\par }{\Q listen to what
the {\ND{Lord}}
says.
\par }{\Q “Have I been like a wilderness
to you, Israel?
\par }{\Q Have I been like a dark
and dangerous land
to you?

\par }{\Q Why
then do you say,
‘We are free to wander.
\par }{\Q We will not
come
to you any more?’
\par }{\Q \VS{32}Does a young woman
forget
to put on her jewels?
\par }{\Q Does a bride
forget to put on her bridal attire?
\par }{\Q But my people
have forgotten
me
\par }{\Q for more days
than can even be counted.
\par }{\Q \VS{33}“My, how
good
you have become
\par }{\Q at chasing
after your lovers!

\par }{\Q Why,
you could even
teach
prostitutes
a thing
or two!
\par }{\Q \VS{34}Even
your clothes
are stained
with
\par }{\Q the lifeblood
of the poor
who had not done anything wrong;
\par }{\Q you did not catch
them breaking
into your homes.

\par }{\Q Yet,
in spite of all
these things you have done,
\par }{\Q \VS{35}you say,
‘I have not done anything wrong,
\par }{\Q so the
{\ND{Lord}}
cannot really be angry
with me any
more.’
\par }{\Q But, watch out! I
will bring down judgment
on
you
\par }{\Q because
you say,
‘I have not
committed any sin.’
\par }{\Q \VS{36}Why
do you constantly
go
about
\par }{\Q changing
your political allegiances?

\par }{\Q You will get no help
from Egypt
\par }{\Q just as
you got no help
from Assyria.
\par }{\Q \VS{37}Moreover,
you will come
away from Egypt
\par }{\Q with your hands
covering your faces
in sorrow and shame

\par }{\Q because
the {\ND{Lord}}
will not allow your reliance
on
them to be successful
\par }{\Q and you will not gain any help from them.

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}“If
a man
divorces
his wife
\par }{\Q and she leaves
him and becomes
another
man’s
wife,
\par }{\Q he may not
take her
back
again.
\par }{\Q Doing
that would utterly defile
the land.
\par }{\Q But you,
Israel,
have given yourself as a prostitute
to many
gods.

\par }{\Q So what makes you think
you can
return
to
me?”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{2}“Look
up
at
the hilltops
and consider
this.

\par }{\Q You have had sex with other gods
on
every one of them.

\par }{\Q You waited for
those gods like a thief lying in wait
in the desert.
\par }{\Q You defiled
the land
by your wicked
prostitution to other gods.
\par }{\Q \VS{3}That is why the rains
have
been withheld,
\par }{\Q and the spring rains
have not
come.
\par }{\Q Yet in spite of this you are obstinate
as a prostitute.
\par }{\Q You refuse
to be ashamed of what you have done.
\par }{\Q \VS{4}Even now
you say
to me, ‘You are my father!

\par }{\Q You
have been my faithful companion
ever since I was young.
\par }{\Q \VS{5}You will not always
be angry
with me, will you?
\par }{\Q You will
not be mad
at me forever,
will you?’

\par }{\Q That is what
you say,
\par }{\Q but you continually do
all the evil
that you can.”
\par }{\PP \VS{6}When
Josiah
was king
of Judah, the
{\ND{Lord}}
said
to
me, “Jeremiah, you have no doubt seen what
wayward
Israel
has
done.
You have seen how she went
up to
every
high
hill
and under
every
green
tree
to give herself like a prostitute
to other gods.
\VS{7}Yet even after
she had done
all
that, I thought
that she might come back
to me. But she did not.
Her sister,
unfaithful
Judah,
saw what she did.
\VS{8}She also saw
that
I gave
wayward
Israel
her divorce
papers
and sent
her away because
of her adulterous
worship of other gods. Even after her unfaithful
sister
Judah
had seen this, she still was not
afraid,
and she too
went
and gave herself like a prostitute to other gods.
\VS{9}Because she took her prostitution
so lightly,
she defiled
the land
through her adulterous
worship of gods made of wood
and stone.
\VS{10}In spite
of all
this,
Israel’s sister,
unfaithful
Judah,
has not
turned back
to me
with any
sincerity;
she has only
pretended
to do
so,” says
the {\ND{Lord}}.
\VS{11}Then the
{\ND{Lord}}
said
to
me, “Under the circumstances, wayward
Israel
could even be considered
less guilty than unfaithful
Judah.
\par }{\SH The Lord Calls on Israel and Judah to Repent
\par }{\PP \VS{12}“Go
and shout
this message
to my people in the countries in the north.
Tell
them,
\par }{\Q ‘Come back
to me, wayward
Israel,’
says
the {\ND{Lord}}.
\par }{\Q ‘I will not
continue to look
on you with displeasure.
\par }{\Q For
I am
merciful,’
says
the {\ND{Lord}}.
\par }{\Q ‘I will not
be angry
with you forever.
\par }{\Q \VS{13}However,
you must confess
that
you have done wrong,

\par }{\Q and that
you have rebelled
against the
{\ND{Lord}}
your God.
\par }{\Q You must
confess
that you have given yourself
to foreign
gods under
every
green
tree,
\par }{\Q and have not
obeyed
my commands,’ says
the {\ND{Lord}}.
\par }{\PP \VS{14}“Come back
to me, my wayward
sons,”
says
the {\ND{Lord}}, “for
I am
your true
master. If you do, I will take
one
of you from each town
and two
of you from each family group,
and I will bring
you back to Zion.
\VS{15}I will give
you leaders
who will be faithful to me. They will lead you with knowledge and insight.
\VS{16}In those days,
your population
will greatly
increase
in the land.
At that time,”
says
the {\ND{Lord}}, “people
will no
longer
talk about having
the ark
that contains the
{\ND{Lord}}’s
covenant
with us.
They will not
call it to mind,
remember
it, or
miss
it. No, that will not
be done
any more!
\VS{17}At that time
the city of Jerusalem
will be called
the
{\ND{Lord}}’s
throne.
All
nations
will gather
there in
Jerusalem
to honor
the
{\ND{Lord}}’s
name. They will no
longer
follow
the stubborn inclinations
of their own evil
hearts.
\VS{18}At that time
the nation
of Judah
and the nation
of Israel
will be reunited.
Together
they
will come back
from a land
in the north
to the land
that
I gave
to your ancestors as a permanent possession. ”
\par }{\Q \VS{19}“I
thought
to myself,

\par }{\Q ‘Oh
what a joy
it would
be for me to treat
you like a son!

\par }{\Q What a joy
it would
be for me to give
you a pleasant
land,
\par }{\Q the most
beautiful
piece
of property
there is in all the world!’

\par }{\Q I thought
you would
call
me, ‘Father’
\par }{\Q and would never
cease being
loyal to me.
\par }{\Q \VS{20}But,
you have been unfaithful
to me, nation
of Israel,
\par }{\Q like an unfaithful
wife
who has left her husband,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{21}“A noise
is heard
on
the hilltops.
\par }{\Q It is the sound of the people
of Israel
crying
and pleading
to their gods.
\par }{\Q Indeed
they have followed sinful
ways;
\par }{\Q they have forgotten
to be true to the

{\ND{Lord}}
their God.
\par }{\Q \VS{22}Come back
to me, you wayward
people.
\par }{\Q I want to cure
your waywardness.
\par }{\Q Say, ‘Here
we are. We come
to you
\par }{\Q because
you
are the
{\ND{Lord}}
our God.
\par }{\Q \VS{23}We know
our noisy
worship of false gods
\par }{\Q on the hills
and mountains
did not help us.

\par }{\Q We know
that the
{\ND{Lord}}
our God
\par }{\Q is the only one who can deliver
Israel.
\par }{\Q \VS{24}From earliest times our worship of that shameful
god, Baal,
\par }{\Q has taken away
all that our ancestors
worked
for.
\par }{\Q It has taken away our flocks
and our herds,
\par }{\Q and even our sons
and daughters.
\par }{\Q \VS{25}Let us acknowledge
our shame.
\par }{\Q Let us bear the disgrace
that we
deserve.

\par }{\Q For
we have sinned
against the
{\ND{Lord}}
our God,
\par }{\Q both we
and our ancestors.
\par }{\Q From earliest
times to this very
day
\par }{\Q we have not
obeyed
the {\ND{Lord}}
our God.’

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}“If
you, Israel,
want to come back,”
says
the {\ND{Lord}},
\par }{\Q “if
you want to come back
to
me

\par }{\Q you must get
those disgusting idols
out of my sight
\par }{\Q and must no
longer go astray.
\par }{\Q \VS{2}You must be truthful,
honest
and upright
\par }{\Q when you take an oath
saying, ‘As surely as the
{\ND{Lord}}
lives!’

\par }{\Q If you do, the nations
will pray to be as blessed
by him as you are
\par }{\Q and will make him the object of their boasting.”
\par }{\Q \VS{3}Yes,
the {\ND{Lord}}
has this
to say
\par }{\Q to the people
of Judah
and Jerusalem:
\par }{\Q “Like a farmer breaking up
hard unplowed ground,
\par }{\Q you must break your rebellious will and make a new beginning;
\par }{\Q just as a farmer must clear away thorns
lest the seed
is wasted,
\par }{\Q you must get rid of the sin that is ruining your lives.
\par }{\Q \VS{4}Just as ritual circumcision
cuts away
the foreskin
\par }{\Q as an external symbol
of dedicated covenant commitment,
\par }{\Q you must genuinely dedicate yourselves to the
{\ND{Lord}}
\par }{\Q and get rid of everything
that hinders
your commitment to me,

\par }{\Q people of Judah
and inhabitants
of Jerusalem.
\par }{\Q If you do not,
my anger
will blaze up
like a flaming
fire
against you
\par }{\Q that no one
will be able to extinguish.
\par }{\Q That will happen because of the evil
you have done.”
\par }{\SH Warning of Coming Judgment
\par }{\PP \VS{5}The
{\ND{Lord}} said,
\par }{\Q “Announce
this in Judah
and proclaim
it in Jerusalem:

\par }{\Q ‘Sound
the trumpet
throughout the land!’
\par }{\Q Shout out
loudly,
\par }{\Q ‘Gather together! Let us flee
into
the fortified
cities!’
\par }{\Q \VS{6}Raise
a signal flag
that tells people to go to Zion.
\par }{\Q Run for safety! Do not
delay!
\par }{\Q For
I
am about to bring
disaster
out of the north.
\par }{\Q It will bring great
destruction.
\par }{\Q \VS{7}Like a lion
that has come up
from its lair
\par }{\Q the one who destroys
nations
has set
out
from his home
base.

\par }{\Q He is coming out to lay
your land
waste.
\par }{\Q Your cities
will become ruins
and lie uninhabited.
\par }{\Q \VS{8}So put
on
sackcloth!
\par }{\Q Mourn
and wail,
saying,
\par }{\Q ‘The fierce
anger
of the {\ND{Lord}}
\par }{\Q has not
turned
away from us!’ ”
\par }{\Q \VS{9}“When
this
happens,”
says
the {\ND{Lord}},
\par }{\Q “the king
and his officials
will lose
their courage.
\par }{\Q The priests
will be struck with horror,
\par }{\Q and the prophets
will be speechless in astonishment.”
\par }{\PP \VS{10}In response to all this I said,
“Ah,
Lord

{\ND{God}}, you have surely
allowed
the people
of Judah and Jerusalem
to be deceived by those who
say,
‘You will be safe!’ But in fact
a sword
is already at our throats.”
\par }{\Q \VS{11}“At that time
the people
of Judah and Jerusalem
will be told,
\par }{\Q ‘A scorching
wind
will sweep down
\par }{\Q from the hilltops
in the desert
on
my dear
people.
\par }{\Q It will
not
be
a gentle
breeze
\par }{\Q for winnowing
the grain
and blowing away
the chaff.
\par }{\Q \VS{12}No, a wind
too strong
for that will come
at my bidding.
\par }{\Q Yes, even
now
I,
myself, am calling down
judgment on them.’
\par }{\Q \VS{13}Look! The enemy is approaching
like gathering
clouds.

\par }{\Q The roar of his chariots
is like that of a whirlwind.
\par }{\Q His horses
move more swiftly
than eagles.”
\par }{\Q I cry out, “We are doomed,
for
we will be destroyed!”
\par }{\Q \VS{14}“Oh people of Jerusalem,
purify
your hearts
from evil
\par }{\Q so that you may
yet be delivered.
\par }{\Q How long
will you continue
to harbor up
\par }{\Q wicked schemes
within you?
\par }{\Q \VS{15}For
messengers are coming, heralding
disaster,
\par }{\Q from the city of Dan
and from the hills
of Ephraim.
\par }{\Q \VS{16}They are saying,

\par }{\Q ‘Announce
to the surrounding nations,
\par }{\Q “The enemy is coming!”

\par }{\Q Proclaim this
message to Jerusalem:
\par }{\Q “Those who besiege
cities are coming
from a distant
land.
\par }{\Q They are ready to raise
the battle cry
against
the towns
in Judah.” ’
\par }{\Q \VS{17}They will surround
Jerusalem

\par }{\Q like men guarding
a field
\par }{\Q because
they have rebelled
against me,”
\par }{\Q says
the
{\ND{Lord}}.
\par }{\Q \VS{18}“The way
you have lived and the things
you have done
\par }{\Q will bring this
on you.
\par }{\Q This
is the punishment you deserve, and it will be painful indeed.
\par }{\Q The pain will be so
bad it
will pierce
your heart.”
\par }{\PP \VS{19}I said,

\par }{\Q “Oh, the feeling in the pit
of my stomach!

\par }{\Q I writhe in anguish.
\par }{\Q Oh, the pain
in my heart!

\par }{\Q My heart
pounds
within me.
\par }{\Q I cannot
keep silent.
\par }{\Q For
I hear
the sound
of the trumpet;
\par }{\Q the sound of the battle
cry
pierces my soul!
\par }{\Q \VS{20}I see one destruction
after another
taking place,
\par }{\Q so that
the whole
land
lies in ruins.
\par }{\Q I see our tents
suddenly
destroyed,
\par }{\Q their curtains torn
down in a mere instant.
\par }{\Q \VS{21}“How long
must I see
the enemy’s battle flags
\par }{\Q and hear
the military signals
of their bugles?”
\par }{\PP \VS{22}The
{\ND{Lord}} answered,

\par }{\Q “This will happen because
my people
are foolish.
\par }{\Q They do not
know
me.
\par }{\Q They are like children
who have no sense.
\par }{\Q They have no
understanding.
\par }{\Q They
are skilled
at doing evil.
\par }{\Q They do not
know
how to do good.”
\par }{\Q \VS{23}“I looked
at the land
and saw
that it was an empty
wasteland.
\par }{\Q I looked up
at the sky,
and its light
had vanished.
\par }{\Q \VS{24}I looked
at the mountains
and saw
that they were shaking.
\par }{\Q All
the hills
were swaying
back and forth!
\par }{\Q \VS{25}I looked
and saw
that there
were no
more people,
\par }{\Q and that all
the birds
in the sky
had flown away.
\par }{\Q \VS{26}I looked
and saw
that the fruitful land
had become a desert
\par }{\Q and that all
of the cities
had been laid
in ruins.
\par }{\Q The
{\ND{Lord}}
had brought this all about
\par }{\Q because
of his blazing
anger.
\par }{\Q \VS{27}All this
will happen because
the

{\ND{Lord}}
said,
\par }{\Q “The whole
land
will be desolate;
\par }{\Q however, I will not
completely destroy it.
\par }{\Q \VS{28}Because
of this
the land
will mourn
\par }{\Q and the sky
above
will grow black.
\par }{\Q For
I have made
my purpose
known

\par }{\Q and I will not
relent
or
turn back
from carrying it out.”
\par }{\Q \VS{29}At the sound
of the approaching
horsemen
and archers
\par }{\Q the people of every
town
will flee.
\par }{\Q Some of them will hide
in the thickets.
\par }{\Q Others will climb
up
among the rocks.
\par }{\Q All
the cities
will be deserted.
\par }{\Q No
one will remain in them.
\par }{\Q \VS{30}And you,
Zion, city doomed to destruction,
\par }{\Q you accomplish
nothing by wearing a beautiful
dress,
\par }{\Q decking
yourself out
in jewels
of gold,
\par }{\Q and putting on
eye
shadow!

\par }{\Q You are making yourself beautiful
for nothing.
\par }{\Q Your lovers
spurn
you.
\par }{\Q They want
to kill you.
\par }{\Q \VS{31}In fact, I hear
a cry
like that
of a woman in labor,
\par }{\Q a cry
of anguish
like that of a woman giving birth
to her first
baby.
\par }{\Q It is the cry of Daughter
Zion
gasping
for
breath,
\par }{\Q reaching out for
help, saying, “I am done in!

\par }{\Q My life
is ebbing away before these murderers!”

\par }\Chap{5}{\PP \VerseOne{1}The
{\ND{Lord}} said,

\par }{\Q “Go up and down through
the streets
of Jerusalem.
\par }{\Q Look
around and see
for yourselves.
\par }{\Q Search
through its public
squares.
\par }{\Q See if
any
of you can find
a single
person
\par }{\Q who deals honestly
and tries
to be truthful.
\par }{\Q If you can, then I will not punish this city.
\par }{\Q \VS{2}These people make promises in the name of the
{\ND{Lord}}.

\par }{\Q But the fact is, what
they swear
to is really a lie.”
\par }{\Q \VS{3}{\ND{Lord}}, I know you look
for faithfulness.
\par }{\Q But even when
you punish these people, they feel no
remorse.
\par }{\Q Even
when you nearly
destroy
them, they refuse
to be corrected.
\par }{\Q They have become as hardheaded
as a rock.
\par }{\Q They refuse
to change their ways.
\par }{\Q \VS{4}I
thought,
“Surely
it is only the ignorant poor
who act this way.
\par }{\Q They
act like fools
because
they do not
know
what the
{\ND{Lord}}
demands.
\par }{\Q They do
not know what their God
requires of them.
\par }{\Q \VS{5}I will go
to
the leaders
\par }{\Q and speak
with them.
\par }{\Q Surely
they
know what the
{\ND{Lord}}
demands.
\par }{\Q Surely
they
know
what their God
requires
of them.”
\par }{\Q Yet all
of them,
too,
have rejected his authority
\par }{\Q and refuse to submit to him.
\par }{\Q \VS{6}So
like a lion
from the thicket
their enemies will kill
them.
\par }{\Q Like a wolf
from the desert
they will destroy
them.
\par }{\Q Like a leopard
they will lie in wait
outside their cities
\par }{\Q and totally
destroy
anyone
who ventures out.
\par }{\Q For
they have rebelled
so much
\par }{\Q and done so many
unfaithful things.
\par }{\PP \VS{7}The
{\ND{Lord}} asked,
\par }{\Q “How can
I leave you unpunished, Jerusalem?

\par }{\Q Your people have rejected
me
\par }{\Q and have
worshiped
gods
that are
not
gods
at all.

\par }{\Q Even though
I supplied all
their
needs,
they
were like an unfaithful
wife to me.

\par }{\Q They
went flocking
to the houses
of prostitutes.
\par }{\Q \VS{8}They are
like lusty,
well-fed
stallions.
\par }{\Q Each
of them lusts
after his neighbor’s
wife.
\par }{\Q \VS{9}I will surely
punish
them for doing such things!” says
the {\ND{Lord}}.
\par }{\Q “I will surely bring retribution
on
such
a nation
as
this!”
\par }{\Q \VS{10}The
{\ND{Lord}} commanded
the enemy,

\par }{\Q “March
through the vineyards of Israel and Judah and ruin
them.

\par }{\Q But do not
destroy
them completely.
\par }{\Q Strip off
their branches
\par }{\Q for
these people do not
belong to the
{\ND{Lord}}.
\par }{\Q \VS{11}For
the nations
of Israel
and Judah
\par }{\Q have been very unfaithful
to me,”
\par }{\PP says
the {\ND{Lord}}.
\par }{\Q \VS{12}“These people have denied
what the
{\ND{Lord}}
says.

\par }{\Q They have said,
‘That is
not
so!

\par }{\Q No
harm
will come
to us.
\par }{\Q We will not
experience
war
and famine.
\par }{\Q \VS{13}The prophets
will prove to be
full of wind.
\par }{\Q The
{\ND{Lord}} has not
spoken through
them.

\par }{\Q So, let
what they say happen to them.’ ”
\par }{\PP \VS{14}Because of
that,
the {\ND{Lord}}, the God
who rules over all,
said
to me,

\par }{\Q “Because
these people have spoken
like this,
\par }{\Q I will make
the words
that I put in your mouth
like fire.
\par }{\Q And I will make this
people
like wood
\par }{\Q which the fiery judgments you speak will burn up.”
\par }{\Q \VS{15}The
{\ND{Lord}}
says, “Listen,
nation
of Israel!

\par }{\Q I am about to bring
a nation
from far away
to attack you.
\par }{\Q It will be a nation
that was founded
long ago
\par }{\Q and has lasted for a long time.
\par }{\Q It will be a nation
whose language
you will not
know.
\par }{\Q Its people will speak
words that you will not
be able to understand.
\par }{\Q \VS{16}All
of its soldiers
are strong
and mighty.
\par }{\Q Their arrows
will send
you to your grave.
\par }{\Q \VS{17}They will eat up
your crops
and your food.
\par }{\Q They will kill off
your sons
and your daughters.
\par }{\Q They will eat up
your sheep
and your cattle.
\par }{\Q They will destroy
your vines
and your fig trees.
\par }{\Q Their weapons
will batter down
\par }{\Q the fortified
cities
you
trust in.
\par }{\PP \VS{18}Yet even
then I will not
completely
destroy you,” says
the {\ND{Lord}}.
\VS{19}“So
then, Jeremiah, when
your people ask, ‘Why
has the
{\ND{Lord}}
our God
done
all
this
to us?’ tell
them,
‘It is because you rejected
me and served
foreign
gods
in your own land.
So
you must serve
foreigners
in a land
that does not
belong to you.’
\par }{\Q \VS{20}“Proclaim
this
message among the descendants
of Jacob.
\par }{\Q Make it known
throughout Judah.
\par }{\Q \VS{21}Tell
them: ‘Hear
this,
\par }{\Q you foolish
people
who have no
understanding,
\par }{\Q who have eyes
but do not
discern,
\par }{\Q who have
ears
but do not
perceive:
\par }{\Q \VS{22}“You should fear
me!” says
the {\ND{Lord}}.
\par }{\Q “You should
tremble
in awe before
me!

\par }{\Q I made
the sand
to be a boundary
for the sea,
\par }{\Q a permanent
barrier
that it can never
cross.
\par }{\Q Its
waves
may roll,
but they
can never
prevail.
\par }{\Q They
may roar,
but they can
never
cross beyond that boundary.”
\par }{\Q \VS{23}But these
people
have stubborn
and rebellious
hearts.
\par }{\Q They have
turned aside
and gone
their own way.
\par }{\Q \VS{24}They do not
say
to themselves,
\par }{\Q “Let
us revere
the

{\ND{Lord}}
our God.
\par }{\Q It
is he who gives
us the autumn
rains
and the spring rains
at the proper time.
\par }{\Q It is he who assures
us of the regular weeks
of harvest.”
\par }{\Q \VS{25}Your misdeeds
have stopped these
things from
coming.

\par }{\Q Your sins
have deprived
you of my bounty.’
\par }{\Q \VS{26}“Indeed,
there
are wicked
scoundrels among
my people.
\par }{\Q They lie in wait like bird catchers
hiding in ambush.

\par }{\Q They set
deadly traps
to catch
people.
\par }{\Q \VS{27}Like a cage
filled
with the birds
that
have been
caught,

\par }{\Q their houses
are filled
with the gains of their fraud and deceit.
\par }{\Q That is
how
they have
gotten so
rich and powerful.
\par }{\Q \VS{28}That is how they
have
grown fat
and sleek.
\par }{\Q There
is no limit to the evil
things they do.

\par }{\Q They do not
plead
the cause
of the fatherless
in such a way as to win
it.
\par }{\Q They do not
defend
the rights
of the poor.
\par }{\Q \VS{29}I will certainly
punish
them for doing such things!” says
the {\ND{Lord}}.
\par }{\Q “I will certainly bring retribution
on
such
a nation
as
this!
\par }{\Q \VS{30}“Something horrible
and shocking
\par }{\Q is going on
in the land of Judah:
\par }{\Q \VS{31}The prophets
prophesy
lies.
\par }{\Q The priests
exercise
power
by their own authority.

\par }{\Q And my people
love
to have it this
way.
\par }{\Q But they will not be able to help you when the time of judgment comes!

\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}“Run for safety,
people
of Benjamin!
\par }{\Q Get out of
Jerusalem!

\par }{\Q Sound
the trumpet
in
Tekoa!
\par }{\Q Light
the signal fires
at Beth Hakkerem!
\par }{\Q For
disaster
lurks out
of the north;
\par }{\Q it will bring great
destruction.
\par }{\Q \VS{2}I will destroy
Daughter
Zion,
\par }{\Q who is as delicate
and defenseless as
a young
maiden.
\par }{\Q \VS{3}Kings will come
against
it with their armies.

\par }{\Q They will encamp in siege all around
it.

\par }{\Q Each of them will devastate the portion assigned to him.
\par }{\Q \VS{4}They will say, ‘Prepare
to do battle
against it!
\par }{\Q Come on! Let’s
attack
it at noon!’
\par }{\Q But
later they will say, ‘Oh, oh! Too bad!

\par }{\Q The day
is almost
over
\par }{\Q and the shadows
of evening
are getting long.
\par }{\Q \VS{5}So come
on, let’s
go
ahead and attack
it by night
\par }{\Q and destroy
all its fortified buildings.’
\par }{\Q \VS{6}All
of this is because
the {\ND{Lord}}
who rules over all
has said:
\par }{\Q ‘Cut
down the trees
around Jerusalem
\par }{\Q and build up a siege
ramp
against its walls.

\par }{\Q This is
the city
which is to be punished.
\par }{\Q Nothing but oppression
happens in it.
\par }{\Q \VS{7}As a well
continually pours out
fresh water
\par }{\Q so
it continually pours out
wicked
deeds.

\par }{\Q Sounds of violence
and destruction
echo
throughout it.

\par }{\Q All I see are sick
and wounded people.’
\par }{\Q \VS{8}So take warning,
Jerusalem,
\par }{\Q or
I will abandon you in disgust
\par }{\Q and make you
desolate,
\par }{\Q a place
where no
one can live.”
\par }{\PP \VS{9}This is what
the {\ND{Lord}}
who rules over all
said
to me:

\par }{\Q “Those who remain
in Israel
will be
\par }{\Q like the grapes thoroughly gleaned
from a vine.
\par }{\Q So go over them again, as though you were a grape harvester
\par }{\Q passing your hand
over
the branches one last time.”
\par }{\PP \VS{10}I answered,

\par }{\Q “Who
would listen
\par }{\Q if I spoke
to them and warned
them?

\par }{\Q Their ears
are so closed
\par }{\Q that they cannot
hear!
\par }{\Q Indeed,
what
the {\ND{Lord}}
says
is offensive
to them.
\par }{\Q They do not
like it at all.
\par }{\Q \VS{11}I am as full
of anger
as you are,
{\ND{Lord}},

\par }{\Q I am tired
of trying to hold
it in.”
\par }{\PP The
{\ND{Lord}} answered,

\par }{\Q “Vent it,
then, on
the children who play
in the street
\par }{\Q and on
the young men
who are gathered together.
\par }{\Q Husbands
and wives
are to be
included,
\par }{\Q as well as
the old
and those
who are advanced
in years.
\par }{\Q \VS{12}Their houses
will be turned
over to others
\par }{\Q as will their fields
and their wives.
\par }{\Q For
I will unleash
my power
\par }{\Q against
those who live
in this land,”
\par }{\PP says
the {\ND{Lord}}.
\par }{\Q \VS{13}“That is because,
from the least
important
to the most important
of them,
\par }{\Q all
of them are greedy
for dishonest gain.
\par }{\Q Prophets
and priests
alike,
\par }{\Q all
of them practice
deceit.
\par }{\Q \VS{14}They offer only superficial
help
\par }{\Q for the harm my people
have suffered.
\par }{\Q They say,
‘Everything
will be all right!’
\par }{\Q But everything is not
all right!
\par }{\Q \VS{15}Are they ashamed
because
they have done
such shameful things?
\par }{\Q No, they are not
at all ashamed.
\par }{\Q They do not
even
know
how to blush!
\par }{\Q So
they will die,
just like others have died.
\par }{\Q They will be brought
to ruin
when I punish
them,”
\par }{\PP says
the {\ND{Lord}}.
\par }{\PP \VS{16}The

{\ND{Lord}}
said
to his people:

\par }{\Q “You are standing
at the crossroads. So consider
your path.
\par }{\Q Ask
where the old, reliable
paths
are.
\par }{\Q Ask where
the path
is
that leads to blessing
and follow
it.
\par }{\Q If you do, you will find
rest
for your souls.”
\par }{\Q But they said,
“We will not
follow it!”
\par }{\PP \VS{17}The
{\ND{Lord}} said,

\par }{\Q “I appointed
prophets as watchmen
to warn
you, saying:
\par }{\Q ‘Pay attention
to the warning
sound
of the trumpet!’ ”

\par }{\Q But they said,
“We will not
pay attention!”
\par }{\PP \VS{18}So
the
{\ND{Lord}} said,

\par }{\Q “Hear,
you nations!
\par }{\Q Be witnesses
and take note of what will happen to these people.
\par }{\Q \VS{19}Hear
this,
you peoples of the earth:

\par }{\Q ‘Take note! I am
about to bring
disaster
on these people.
\par }{\Q It will come as punishment for their scheming.
\par }{\Q For
they have paid
no
attention
to what I have said,

\par }{\Q and they have rejected
my law.
\par }{\Q \VS{20}I take no delight
when they offer up to me

\par }{\Q frankincense
that
comes
from Sheba
\par }{\Q or sweet-smelling
cane imported
from a faraway
land.
\par }{\Q I cannot accept the burnt offerings
they bring me.
\par }{\Q I get no
pleasure
from the sacrifices they offer to me.’
\par }{\Q \VS{21}So,
this
is what the
{\ND{Lord}}
says:
\par }{\Q ‘I will assuredly
make
these
people
stumble
to
their doom.

\par }{\Q Parents
and children
will stumble
and fall
to their destruction.

\par }{\Q Friends
and neighbors
will die.’
\par }{\Q \VS{22}“This is what
the {\ND{Lord}}
says:
\par }{\Q ‘Beware! An army
is coming
from a land
in the north.
\par }{\Q A mighty nation
is stirring into action
in faraway parts
of the earth.
\par }{\Q \VS{23}Its soldiers are armed
with bows
and spears.
\par }{\Q They are
cruel
and show no
mercy.
\par }{\Q They sound
like the roaring
sea
\par }{\Q as they ride
forth on
their
horses.
\par }{\Q Lined
up
in formation
like men
going
into battle
\par }{\Q to attack
you, Daughter
Zion.’ ”
\par }{\Q \VS{24}The people cry out, “We have heard
reports
about them!
\par }{\Q We have become helpless
with fear!

\par }{\Q Anguish
grips
us,
\par }{\Q agony
like that of a woman giving birth to a baby!
\par }{\Q \VS{25}Do not
go out
into the countryside.
\par }{\Q Do not
travel
on the roads.
\par }{\Q For
the enemy
is there with sword
in hand.

\par }{\Q They are spreading terror
everywhere.”
\par }{\Q \VS{26}So I said, “Oh, my dear
people,
put on
sackcloth
\par }{\Q and roll
in ashes.
\par }{\Q Mourn
with painful
sobs
\par }{\Q as though you had lost your only child.
\par }{\Q For any moment now
that
destructive army
\par }{\Q will come
against us.”
\par }{\PP \VS{27}The
{\ND{Lord}} said to me,

\par }{\Q “I have made
you like a metal assayer
\par }{\Q to test
my people
like ore.

\par }{\Q You are to observe
them
\par }{\Q and evaluate
how they behave.”
\par }{\PP \VS{28}I reported,
\par }{\Q “All
of them are
the most stubborn of rebels!

\par }{\Q They are
as hard as
bronze
or
iron.
\par }{\Q They go about telling lies.
\par }{\Q They all
deal corruptly.
\par }{\Q \VS{29}The fiery
bellows
of judgment burn fiercely.
\par }{\Q But there is too much dross
to be removed.

\par }{\Q The process
of refining
them has proved useless.
\par }{\Q The wicked
have not
been purged.
\par }{\Q \VS{30}They are regarded
as ‘rejected
silver’
\par }{\Q because
the {\ND{Lord}}
rejects them.”

\par }\Chap{7}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
Jeremiah:
\VS{2}“Stand
in the gate
of the
{\ND{Lord}}’s
temple
and proclaim
this
message: ‘Listen,
all
you people of Judah
who have passed through
these
gates
to worship
the {\ND{Lord}}. Hear what
the {\ND{Lord}}
has to say.
\VS{3}The

{\ND{Lord}}
God
of Israel
who rules over all
says: Change
the way
you have been living
and do what is right. If you do, I will allow
you to continue
to live in
this land.
\VS{4}Stop
putting your confidence
in the false
belief
that says, “We are safe! The temple
of the {\ND{Lord}}
is here! The temple
of the {\ND{Lord}}
is here! The temple
of the {\ND{Lord}} is here!”
\VS{5}You must
change
the way
you have been living
and do what is right. You must
treat
one
another
fairly.
\VS{6}Stop
oppressing
foreigners
who live in your land, children who have lost their fathers,
and women who have lost their husbands.
Stop
killing
innocent
people
in this
land.
Stop
paying allegiance to other
gods.
That will only bring about
your ruin.
\VS{7}If you stop doing these things, I will allow
you to continue to live in this
land
which
I gave
to your ancestors
as
a lasting possession.
\par }{\PP \VS{8}“‘But just look
at you! You
are putting your confidence
in a false
belief
that will not deliver you.
\VS{9}You steal.
You murder.
You commit adultery.
You lie
when you swear
on oath. You sacrifice
to the god
Baal.
You pay allegiance
to other
gods
whom
you have not
previously known.
\VS{10}Then you come
and stand
in my presence
in this
temple
I have
claimed
as my own
and say,
“We are
safe!” You think you are
so
safe that you go
on
doing
all
those hateful sins!
\VS{11}Do you think this
temple
I have claimed
as
my own
is to be
a hideout
for robbers? You had better take note! I
have seen
for myself what you have done! says
the
{\ND{Lord}}.
\VS{12}So, go
to
the place
in Shiloh
where
I allowed
myself to be worshiped
in the early days.
See
what
I did
to it because
of the wicked
things my people
Israel did.
\VS{13}You also have
done
all
these
things,
says
the {\ND{Lord}}, and I have spoken
to
you over and over again.
But
you have not
listened! You have refused
to respond
when I called
you to repent!
\VS{14}So
I will destroy this temple
which
I have claimed
as my own,
this temple that
you
are trusting
to protect you. I will destroy this place
that
I gave
to you and your ancestors,
just
like I destroyed Shiloh.
\VS{15}And I will drive
you out
of my sight
just
like I drove
out your relatives,
the people
of Israel.’ ”
\par }{\PP \VS{16}Then the
{\ND{Lord}} said, “As for you,
Jeremiah, do not
pray
for
these
people! Do not
cry out
to me or petition
me on their behalf! Do not
plead
with me to save them, because
I will not
listen to you.
\VS{17}Do
you see
what
they
are doing
in the towns
of Judah
and in the streets
of Jerusalem?
\VS{18}Children
are gathering
firewood,
fathers
are building
fires
with it, and
women
are mixing
dough
to bake cakes
to offer
to the goddess they call the Queen
of Heaven.
They are
also pouring
out drink offerings
to other
gods.
They seem to do
all this
just
to trouble me.
\VS{19}But
I am not
really
the one being troubled!” says
the {\ND{Lord}}. “Rather
they are bringing trouble on themselves to their own shame!
\VS{20}So,”
the Lord

{\ND{God}}
says,
“my raging
fury
will be poured
out on
this
land.
It will be poured
out on
human
beings and animals,
on
trees
and crops.
And it will burn
like a fire which cannot
be extinguished.”
\par }{\PP \VS{21}The
{\ND{Lord}} said
to the people of Judah, “The
{\ND{Lord}}
God
of Israel
who rules over all
says: ‘You might as well go ahead and add
the meat
of your burnt offerings
to that of the other sacrifices
and eat it, too!
\VS{22}Consider this: When
I spoke
to your ancestors
after
I brought them out
of Egypt,
I did not
merely
give them commands about
burnt offerings
and sacrifices.
\VS{23}I also
explicitly
commanded
them: “Obey
me.
If
you do, I will be
your God
and you
will be
my people.
Live
exactly
the way
I tell you
and things will
go well with you.”
\VS{24}But they did not
listen
to me or
pay
any attention
to me. They followed
the stubborn inclinations
of their own wicked
hearts.
They acted
worse and worse
instead
of better.
\VS{25}From
the time
your ancestors
departed
the land
of Egypt
until
now, I sent
my servants
the prophets
to you again
and again,
day
after day.
\VS{26}But your ancestors
did not
listen
to
me nor
pay
attention
to me. They became obstinate
and were more wicked
than even their own forefathers.’ ”
\par }{\PP \VS{27}Then the
{\ND{Lord}} said
to
me, “When you tell them
all
this,
they will not
listen
to
you. When you call
out to
them, they will not
respond to you.
\VS{28}So tell
them: ‘This
is a nation
that has not
obeyed
the {\ND{Lord}}
their God
and has not
accepted
correction.
Faithfulness
is nowhere to be found
in it. These people do not even profess it anymore.
\VS{29}So, mourn, you people of this nation. Cut off
your hair
and throw
it away.
Sing
a song of mourning
on
the hilltops.
For
the {\ND{Lord}}
has decided to reject
and forsake
this generation
that has provoked his wrath!’ ”
\par }{\PP \VS{30}The
{\ND{Lord}}
says,
“I have rejected them because
the people
of Judah
have done
what I consider evil.
They have set up
their disgusting idols
in the temple
which
I have claimed
for my own
and have defiled it.
\VS{31}They have
also built
places of worship
in a place called Topheth
in the Valley
of Ben Hinnom
so that
they can sacrifice
their sons
and daughters
by fire.
That
is something I never commanded
them to do! Indeed, it never
even entered
my mind to command such a thing!
\VS{32}So,
watch
out!” says
the {\ND{Lord}}. “The time
will soon come
when people will no
longer
call those places Topheth
or the Valley
of Ben Hinnom.
But
they will call that valley
the Valley of Slaughter
and they will bury
so many people in Topheth
they will run
out of room.
\VS{33}Then the dead bodies
of these
people
will be
left on the ground for the birds
and wild
animals
to eat.
There will not
be any
survivors to scare them away.
\VS{34}I will put an end
to the sounds
of joy
and gladness,
or the glad celebration
of brides
and grooms
throughout the towns
of Judah
and the streets
of Jerusalem.
For
the whole land
will become
a desolate wasteland.”

\par }\Chap{8}{\PP \VerseOne{1}The
{\ND{Lord}}
says,
“When
that time
comes,
the
bones
of the kings
of Judah
and its
leaders,
the
bones
of the priests
and prophets
and of all the other people who lived
in Jerusalem
will be dug up
from their graves.
\VS{2}They will be spread out
and exposed to the sun,
the moon
and the stars.
These are things they adored
and served,
things to which
they paid
allegiance,
from which
they sought
guidance, and worshiped.
The bones of these people will never
be regathered
and reburied.
They will
be like manure
used to fertilize the ground.
\VS{3}However, I will leave
some of
these wicked
people
alive
and banish
them to other
places.
But wherever
these people who survive
may go,
they will wish
they had died
rather than lived,”
says
the {\ND{Lord}}
who rules over all.
\par }{\SH Willful Disregard of God Will Lead to Destruction
\par }{\PP \VS{4}The
{\ND{Lord}} said
to me,
\par }{\Q “Tell
them, ‘The
{\ND{Lord}}
says,
\par }{\Q Do people not
get
back
up when they fall
down?
\par }{\Q Do they not
turn around when they go the wrong way?
\par }{\Q \VS{5}Why,
then, do these
people
of Jerusalem
\par }{\Q continually
turn away
from me in apostasy?
\par }{\Q They hold fast
to their deception.
\par }{\Q They refuse
to turn back to me.
\par }{\Q \VS{6}I have listened
to them very carefully,

\par }{\Q but they do not
speak
honestly.
\par }{\Q None
of them regrets
the evil
he has done.
\par }{\Q None of them says,
“I have
done
wrong!”

\par }{\Q All
of them persist
in their own wayward course
\par }{\Q like a horse
charging recklessly into battle.
\par }{\Q \VS{7}Even
the stork
knows
\par }{\Q when it is time to move on.
\par }{\Q The turtledove,
swallow,
and crane
\par }{\Q recognize
the normal times
for their migration.
\par }{\Q But my people
pay
no
attention
\par }{\Q to what I, the

{\ND{Lord}}, require of them.
\par }{\Q \VS{8}How
can you say,
“We are wise!
\par }{\Q We
have the law
of the {\ND{Lord}}”?
\par }{\Q The truth is, those who teach
it have used their writings
\par }{\Q to make
it say what it does not really mean.
\par }{\Q \VS{9}Your wise
men will be put to shame.
\par }{\Q They will be dumbfounded
and be brought to judgment.

\par }{\Q Since
they
have rejected
the word
of the {\ND{Lord}},
\par }{\Q what
wisdom do they really have?
\VS{10}\par }{\Q So
I will give
their wives
to other
men
\par }{\Q and their
fields
to new owners.
\par }{\Q For
from the least
important
to the most important
of them,
\par }{\Q all
of them are greedy
for dishonest gain.
\par }{\Q Prophets
and priests
alike,
\par }{\Q all
practice
deceit.
\par }{\Q \VS{11}They offer only superficial
help
\par }{\Q for the hurt my dear
people
have suffered.
\par }{\Q They say,
“Everything will be all right!”
\par }{\Q But everything
is not
all right!
\par }{\Q \VS{12}Are they ashamed
because
they have done
such
disgusting things?
\par }{\Q No, they are not
at all ashamed!
\par }{\Q They do not
even know
how to blush!
\par }{\Q So
they will die
just like
others have died.
\par }{\Q They will be brought to ruin
when
I punish
them,
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{13}I will take away
their harvests, says
the
{\ND{Lord}}.
\par }{\Q There will be no
grapes
on their vines.
\par }{\Q There will be no
figs
on their fig trees.
\par }{\Q Even the leaves
on their trees will wither.
\par }{\Q The crops that I gave
them
will be taken away.’ ”
\par }{\SH Jeremiah Laments over the Coming Destruction
\par }{\PP \VS{14}The people say,

\par }{\Q “Why
are we
just sitting
here?
\par }{\Q Let us gather together
inside
the fortified
cities.
\par }{\Q Let us at least die
there
fighting,

\par }{\Q since
the {\ND{Lord}}
our God
has condemned us to die.
\par }{\Q He has condemned us to drink
the poison
waters
of judgment

\par }{\Q because
we have sinned
against him.
\par }{\Q \VS{15}We hoped
for good fortune,
but nothing
good
has come of it.
\par }{\Q We hoped for a time
of relief,
but instead we experience terror.
\par }{\Q \VS{16}The snorting
of the enemy’s horses
\par }{\Q is already being heard
in the city of Dan.
\par }{\Q The sound
of the neighing
of their stallions
\par }{\Q causes the whole
land
to tremble
with fear.
\par }{\Q They are coming
to destroy
the land
and everything
in it!
\par }{\Q They are coming to destroy the cities
and everyone who lives in them!”
\par }{\PP \VS{17}The

{\ND{Lord}}
says,
\par }{\Q “Yes
indeed,
I am sending
an enemy against you
\par }{\Q that will be like poisonous snakes
which
cannot
be charmed
away.

\par }{\Q And they will inflict fatal wounds on you.”
\par }{\PP \VS{18}Then I said,

\par }{\Q “There is no cure
for my grief!
\par }{\Q I am sick
at heart!
\par }{\Q \VS{19}I
hear
my dear
people
crying out
\par }{\Q throughout the length and breadth
of the land.
\par }{\Q They are crying, ‘Is the
{\ND{Lord}}
no
longer in Zion?
\par }{\Q Is her divine King
no
longer there?’ ”
\par }{\Q The
{\ND{Lord}} answers,

\par }{\Q “Why
then do they provoke
me to anger
with their images,
\par }{\Q with their worthless
foreign idols?”
\par }{\Q \VS{20}“They cry, ‘Harvest
time has
come and gone,
and the summer
is over,

\par }{\Q and still we
have not
been delivered.’
\par }{\Q \VS{21}My heart is crushed
because my dear
people
are being crushed.
\par }{\Q I go about crying and grieving.
I am overwhelmed
with dismay.
\par }{\Q \VS{22}There
is still medicinal ointment
available
in Gilead!
\par }{\Q There is
still a physician
there!

\par }{\Q Why
then have my dear
people
\par }{\Q not
been restored to health?

\par }\Chap{9}{\PP \par }{\Q \VerseOne{1}I
wish
that my head
were a well full of water
\par }{\Q and my eyes
were a fountain
full of tears!
\par }{\Q If they were, I could cry
day
and night
\par }{\Q for those of my dear
people
who have been killed.
\par }{\Q \VS{2}I wish
I had
a lodging place
in the desert
\par }{\Q where I could spend some time like a weary traveler.
\par }{\Q Then I would desert
my people
\par }{\Q and walk away
from them
\par }{\Q because
they are all
unfaithful
to God,
\par }{\Q a congregation
of people that has been disloyal to him.
\par }{\SH The Lord Laments That He Has No Choice But to Judge Them
\par }{\PP \VS{3}The
{\ND{Lord}}
says,
\par }{\Q “These people are like soldiers who
have readied
their bows.
\par }{\Q Their tongues
are always ready
to shoot
out lies.
\par }{\Q They have become powerful
in the land,
\par }{\Q but they have not
done
so by
honest
means.
\par }{\Q Indeed,
they do one evil
thing after another
\par }{\Q and do not pay attention to me.
\par }{\Q \VS{4}Everyone
must be on
his guard
around his friends.
\par }{\Q He must not
even trust
any of his relatives.
\par }{\Q For
every one
of them
will find some way
to cheat
him.

\par }{\Q And all
of his friends
will tell lies about him.
\par }{\Q \VS{5}One
friend
deceives
another
\par }{\Q and no
one tells
the truth.
\par }{\Q These people have
trained
themselves to tell
lies.
\par }{\Q They do wrong
and are unable
to repent.
\par }{\Q \VS{6}They do one act of violence
after another,
\par }{\Q and one deceitful
thing after another.

\par }{\Q They refuse
to pay attention
to me,”

\par }{\Q says
the {\ND{Lord}}.
\par }{\PP \VS{7}Therefore
the

{\ND{Lord}}
who rules over all
says,
\par }{\Q “I will now purify them in the fires of affliction
and test
them.
\par }{\Q The wickedness of my dear
people
has left me no choice.
\par }{\Q What
else can I do?
\par }{\Q \VS{8}Their tongues
are like deadly
arrows.
\par }{\Q They are always telling
lies.
\par }{\Q Friendly
words
for
their neighbors
come from their mouths.
\par }{\Q But their minds are thinking up ways to trap them.
\par }{\Q \VS{9}I will certainly
punish
them for doing such things!” says
the {\ND{Lord}}.
\par }{\Q “I will certainly bring retribution
on such a nation
as
this!”
\par }{\SH The Coming Destruction Calls For Mourning
\par }{\PP \VS{10}I said,

\par }{\Q “I will weep
and mourn
for the grasslands on
the mountains,
\par }{\Q I will sing
a mournful song
for the pastures
in the wilderness
\par }{\Q because
they are so scorched
no one travels through
them.
\par }{\Q The sound
of livestock
is no longer
heard
there.
\par }{\Q Even the birds
in the sky
and the wild animals
in the fields
\par }{\Q have fled
and are gone.”
\par }{\PP \VS{11}The
{\ND{Lord}} said,

\par }{\Q “I will make
Jerusalem
a heap
of ruins.
\par }{\Q Jackals
will make their home
there.

\par }{\Q I will destroy
the towns
of Judah
\par }{\Q so that no one
will be able to live in them.”
\par }{\PP \VS{12}I said,

\par }{\Q “Who
is wise
enough to understand
why this
has happened?

\par }{\Q Who
has a word
from the
{\ND{Lord}}
that can explain
it?

\par }{\Q Why
does the land
lie in ruins?
\par }{\Q Why
is it as scorched
as a desert
through which no
one travels?”
\par }{\PP \VS{13}The
{\ND{Lord}}
answered,
“This
has happened
because these people have rejected
my laws
which
I gave
them. They have not
obeyed
me
or
followed those laws.
\VS{14}Instead they have followed
the stubborn inclinations
of their own hearts.
They have paid allegiance
to the gods called Baal,
as their fathers
taught them to do.
\VS{15}So
then, listen to what
I, the
{\ND{Lord}}
God
of Israel
who rules over all,
say. ‘I
will make these
people
eat
the
bitter food of suffering
and drink
the poison
water of judgment.
\VS{16}I will scatter
them among nations
that
neither
they
nor their ancestors
have known
anything
about. I will send
people chasing after
them with swords
until
I have destroyed them.’ ”
\par }{\PP \VS{17}The
{\ND{Lord}}
who rules over all
told me to say
to this people,

\par }{\Q “Take note
of what I say.

\par }{\Q Call
for the women who mourn
for the dead!
\par }{\Q Summon those who are the most skilled at it!”
\par }{\Q \VS{18}I said, “Indeed, let them come quickly
and sing
a song of mourning for us.
\par }{\Q Let them wail
loudly
until tears
stream from our own eyes
\par }{\Q and our eyelids
overflow
with water.
\par }{\Q \VS{19}For
the sound
of wailing
is soon to be heard
in Zion.
\par }{\Q They will wail, ‘We are utterly ruined! We are completely
disgraced!
\par }{\Q For
our houses
have been torn down
\par }{\Q and we must leave
our land.’ ”
\par }{\PP \VS{20}I said,

\par }{\Q “So now, you wailing women,
hear
what
the {\ND{Lord}}
says.

\par }{\Q Open
your ears
to the words
from his mouth.
\par }{\Q Teach
your daughters
this mournful
song,
\par }{\Q and each of you teach your neighbor
this lament.
\par }{\Q \VS{21}‘Death
has climbed
in through our windows.
\par }{\Q It has entered
into our fortified houses.
\par }{\Q It has taken
away our children who play
in the streets.
\par }{\Q It has taken away our young men
who gather in the city squares.’
\par }{\Q \VS{22}Tell
your daughters and neighbors, ‘The
{\ND{Lord}}
says,
\par }{\Q “The dead bodies
of people
will lie scattered
everywhere
\par }{\Q like manure
scattered on
a field.
\par }{\Q They will lie scattered on the ground
\par }{\Q like grain
that has been cut down
but has not
been gathered.” ’ ”
\VS{23}\par }{\PP The
{\ND{Lord}}
says,
\par }{\Q “Wise people
should not
boast
that they are wise.
\par }{\Q Powerful people
should not
boast
that they are powerful.
\par }{\Q Rich people
should not
boast
that they are rich.
\par }{\Q \VS{24}If
people want to boast,
they should boast about this:
\par }{\Q They should boast
that they understand
and know
me.
\par }{\Q They should boast that they know and understand
\par }{\Q that
I,
the {\ND{Lord}}, act
out of faithfulness,
fairness,
and justice
in the earth
\par }{\Q and that
I desire
people to do these
things,”

\par }{\PP says
the {\ND{Lord}}.
\par }{\PP \VS{25}The
{\ND{Lord}}
says,
“Watch
out! The time
is soon coming when
I will punish
all
those who are circumcised only in the flesh.
\VS{26}That is, I will punish the Egyptians,
the Judeans,
the Edomites,
the Ammonites,
the Moabites,
and all
the desert
people
who cut
their hair
short
at the temples.
I will do so because
none
of the people
of those nations
are really circumcised
in the
{\ND{Lord}}’s sight. Moreover, none
of the people
of Israel
are circumcised
when it comes to their hearts.”

\par }\Chap{10}{\PP \VerseOne{1}You people
of Israel,
listen
to what
the {\ND{Lord}}
has
to say to you.
\par }{\PP \VS{2}The

{\ND{Lord}}
says,
\par }{\Q “Do not
start
following pagan religious practices.
\par }{\Q Do
not
be in awe
of signs
that
occur in the sky
\par }{\Q even
though
the nations
hold them
in awe.
\par }{\Q \VS{3}For
the religion
of these people
is worthless.
\par }{\Q They cut down
a tree
in the forest,
\par }{\Q and a craftsman
makes it into an idol with his tools.
\par }{\Q \VS{4}He decorates
it with overlays of silver
and gold.
\par }{\Q He uses hammer
and nails
to fasten
it together
\par }{\Q so that it will not
fall over.
\par }{\Q \VS{5}Such idols are like scarecrows
in a cucumber
field.
\par }{\Q They
cannot talk.
\par }{\Q They must be carried
\par }{\Q because
they cannot
walk.
\par }{\Q Do not
be afraid
of them
\par }{\Q because
they cannot
hurt
you.
\par }{\Q And they do not
have any power to help you.”
\par }{\PP \VS{6}I said,

\par }{\Q “There is no
one like
you,
{\ND{Lord}}.

\par }{\Q You are great.
\par }{\Q And you
are renowned
for your power.
\par }{\Q \VS{7}Everyone should revere
you, O King
of all nations,
\par }{\Q because
you deserve to be revered.
\par }{\Q For
there is no one
like
you
\par }{\Q among any
of the wise people
of the nations
nor among any
of their kings.
\par }{\Q \VS{8}The people of those nations are both
stupid
and foolish.
\par }{\Q Instruction
from a wooden
idol is worthless!
\par }{\Q \VS{9}Hammered-out
silver
is brought
from Tarshish
\par }{\Q and gold
is brought from Uphaz
to cover those idols.

\par }{\Q They are the handiwork
of carpenters
and goldsmiths.
\par }{\Q They are clothed in blue
and purple
clothes.
\par }{\Q They are all
made
by skillful workers.
\par }{\Q \VS{10}The
{\ND{Lord}}
is the only true
God.
\par }{\Q He
is the living
God
and the everlasting
King.
\par }{\Q When he shows his anger
the earth
shakes.
\par }{\Q None of the nations
can stand up to his fury.
\par }{\Q \VS{11}You people of Israel should tell
those nations this:
\par }{\Q ‘These
gods
did not
make
heaven
and earth.
\par }{\Q They will disappear
from the earth
and from
under
the heavens.’
\par }{\Q \VS{12}The
{\ND{Lord}} is the one who by his power
made
the earth.
\par }{\Q He is the one who by his wisdom
established
the world.
\par }{\Q And by his understanding
he spread
out the skies.
\par }{\Q \VS{13}When his voice
thunders,
the heavenly
ocean
roars.
\par }{\Q He makes the clouds rise from
the far-off horizons.
\par }{\Q He
makes the lightning
flash out
in the midst of the rain.
\par }{\Q He unleashes
the wind
from the
places where he stores it.
\par }{\Q \VS{14}All
these idolaters
will prove to be stupid and ignorant.
\par }{\Q Every
goldsmith
will be disgraced
by the idol
he made.
\par }{\Q For
the image
he forges is merely a sham.
\par }{\Q There is no
breath in any of those idols.
\par }{\Q \VS{15}They
are worthless,
mere objects
to be mocked.
\par }{\Q When the time
comes to punish
them, they will be destroyed.
\par }{\Q \VS{16}The
{\ND{Lord}}, who is the inheritance
of Jacob’s
descendants, is not
like them.
\par }{\Q He is the one who created
everything.
\par }{\Q And the people
of Israel
are those he claims as his own.
\par }{\Q He is known
as the
{\ND{Lord}}
who rules over all.”
\par }{\SH Jeremiah Laments for and Prays for the Soon-to-be-Judged People
\par }{\Q \VS{17}Gather
your belongings
together
and prepare to leave the land,
\par }{\Q you people
of Jerusalem who are being besieged.
\par }{\Q \VS{18}For
the
{\ND{Lord}}
says,
“I
will now
throw
out
\par }{\Q those who live
in this
land.
\par }{\Q I will bring so much
trouble
on them
\par }{\Q that
they
will actually feel it.”
\par }{\Q \VS{19}And I cried out, “We are doomed!

\par }{\Q Our wound
is severe!
\par }{\Q We once thought,
‘This
is only
an illness.
\par }{\Q And we will be able to bear it!’
\par }{\Q \VS{20}But our tents
have been destroyed.
\par }{\Q The ropes
that held them
in place have
been ripped apart.
\par }{\Q Our children
are gone
and are not
coming back.

\par }{\Q There is no
survivor
to put
our tents
back up,
\par }{\Q no one left
to hang their tent curtains in place.
\par }{\Q \VS{21}For
our leaders
are stupid.
\par }{\Q They have not
sought
the
{\ND{Lord}}’s
advice.

\par }{\Q So
they do not
act wisely,
\par }{\Q and the people
they are responsible for have all
been scattered.
\par }{\Q \VS{22}Listen! News
is coming even now.
\par }{\Q The rumble
of a great
army is heard approaching from a land
in the north.
\par }{\Q It is coming
to turn
the towns
of Judah
into rubble,
\par }{\Q places where only
jackals
live.
\par }{\Q \VS{23}{\ND{Lord}}, we know
that
people
do not
control their own destiny.
\par }{\Q It is not
in their
power
to determine
what will happen to them.
\par }{\Q \VS{24}Correct
us,
{\ND{Lord}}, but only
in due
measure.

\par }{\Q Do not
punish us in anger
or
you will reduce us to nothing.
\par }{\Q \VS{25}Vent
your anger
on
the nations
that
do not
acknowledge
you.

\par }{\Q Vent
it on
the peoples
who
do not
worship
you.

\par }{\Q For
they have destroyed
the people of Jacob.
\par }{\Q They have completely destroyed
them

\par }{\Q and left their homeland
in utter ruin.

\par }\Chap{11}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
Jeremiah:
\VS{2}“Hear
the
terms
of the covenant
I made with Israel and pass them
on
to
the people
of Judah
and the citizens
of Jerusalem.
\VS{3}Tell
them
that the

{\ND{Lord}},
the God
of Israel,
says,
‘Anyone
who
does not
keep the terms
of the covenant
will be under a curse.
\VS{4}Those are the terms that
I charged
your ancestors
to keep when
I brought
them out of Egypt,
that place which was like an iron-smelting furnace.
I said
at
that
time, “Obey
me and carry
out the terms of the agreement exactly
as I commanded
you. If you do, you will be my people
and I
will be
your God.
\VS{5}Then
I will keep
the promise
I swore on oath
to your ancestors
to give
them
a land
flowing
with milk
and honey.”
That is
the very land
that you
still live in today.’ ”
And I responded,
“Amen! Let it be so,
{\ND{Lord}}!”
\par }{\PP \VS{6}The
{\ND{Lord}}
said
to
me, “Announce
all
the following
words
in the towns
of Judah
and in the streets
of Jerusalem: ‘Listen
to the
terms
of my covenant
with you and carry them out!
\VS{7}For
I solemnly
warned
your ancestors
to obey
me.
I warned
them again and again,
ever since
I delivered
them out of Egypt
until
this
very day.
\VS{8}But they did not
listen
to me or
pay
any attention
to me! Each
one of them followed
the stubborn inclinations
of his own wicked
heart.
So I brought
on
them all
the punishments
threatened in the covenant
because they did not
carry out
its terms as I commanded
them to do.’ ”
\par }{\PP \VS{9}The
{\ND{Lord}}
said
to me,
“The people
of Judah
and the citizens
of Jerusalem
have
plotted rebellion against me!
\VS{10}They have gone back
to the evil ways
of their ancestors
of old
who
refused
to obey
what I told
them. They,
too, have paid allegiance
to other
gods
and worshiped
them. Both the nation
of Israel
and the nation
of Judah
have violated
the covenant
I made
with
their ancestors.
\VS{11}So
I,
the {\ND{Lord}}, say
this: ‘I will soon
bring
disaster
on them which
they will not
be able
to escape! When they cry
out to me
for help, I will not
listen
to them.
\VS{12}Then
those living
in the towns
of Judah
and in Jerusalem
will go and cry out
for help to
the gods
to whom
they
have been sacrificing.
However, those gods will by no
means be able to save
them when
disaster strikes them.
\VS{13}This
is in spite of the fact
that
the people of Judah
have as many
gods
as they have towns and the citizens of Jerusalem
have set
up as many
altars
to sacrifice
to that disgusting
god, Baal,
as they have streets
in the city!’
\VS{14}So, Jeremiah, do not
pray
for
these
people.
Do not
cry out
to me or petition
me on their behalf. Do not plead with me to save them. For
I will not
listen
to them when
they call out
to
me for
help when disaster strikes them.”
\par }{\PP \VS{15}The
{\ND{Lord}} says to the people of Judah,

\par }{\Q “What
right do you have to be in my temple,
my beloved
people?


\par }{\Q Many
of you have done
wicked things.

\par }{\Q Can your acts of treachery be so
easily canceled
by
sacred
offerings

\par }{\Q that
you take joy in doing evil
even while you make them?
\par }{\Q \VS{16}I, the
{\ND{Lord}}, once called
you a thriving
olive tree,
\par }{\Q one that produced beautiful
fruit.
\par }{\Q But I will set
you on fire,
\par }{\Q fire that will blaze with a mighty roar.
\par }{\Q Then all your branches
will be good for nothing.
\par }{\Q \VS{17}For though I, the
{\ND{Lord}}
who rules over all,
planted
you in the
land,

\par }{\Q I now decree
that disaster
will come on you

\par }{\Q because the nations of Israel
and Judah
have
done
evil
\par }{\Q and have made me angry
by offering sacrifices
to the god Baal.”
\par }{\SH A Plot Against Jeremiah is Revealed and He Complains of Injustice
\par }{\Q \VS{18}The
{\ND{Lord}}
gave me knowledge,
that I might have understanding.
\par }{\Q Then
he showed
me what the people were doing.
\par }{\Q \VS{19}Before this I
had been like a docile
lamb
ready to be led
to the slaughter.
\par }{\Q I did not
know
they were making
plans
to kill me.
\par }{\Q I did not know they were saying,

\par }{\Q “Let’s destroy
the tree
along with its fruit!

\par }{\Q Let’s remove
Jeremiah from the world
of the living
\par }{\Q so people will not
even be reminded
of him
any more.”
\par }{\Q \VS{20}So I said to the
{\ND{Lord}},

\par }{\Q “O
{\ND{Lord}}
who rules over all,
you are a just
judge!
\par }{\Q You examine
people’s hearts
and minds.
\par }{\Q I want to see
you pay
them back
for what they
have done
\par }{\Q because
I trust you to vindicate
my cause.”
\par }{\PP \VS{21}Then the

{\ND{Lord}}
told me about some men
from Anathoth
who were threatening to kill
me. They had threatened, “Stop prophesying
in the name
of the {\ND{Lord}}
or
we will kill you!”
\VS{22}So
the

{\ND{Lord}}
who rules over all
said,
“I
will surely
punish
them! Their young men
will be killed
in battle.
Their sons
and daughters
will die
of starvation.
\VS{23}Not
one of them will survive.
I will bring
disaster
on those men
from Anathoth
who threatened you. A day
of reckoning
is coming for them.”

\par }\Chap{12}{\PP \par }{\Q \VerseOne{1}{\ND{Lord}}, you
have always been fair
\par }{\Q whenever I have complained
to
you.

\par }{\Q However,
I would like to speak
with
you about the disposition
of justice.
\par }{\Q Why
are wicked
people successful?

\par }{\Q Why do all
dishonest
people have such easy lives?
\par }{\Q \VS{2}You plant
them like trees and they put down their roots.

\par }{\Q They grow prosperous
and are very fruitful.
\par }{\Q They always talk
about
you,
\par }{\Q but they really
care nothing
about you.
\par }{\Q \VS{3}But you,

{\ND{Lord}}, know
all about me.
\par }{\Q You watch
me and test
my devotion
to you.

\par }{\Q Drag
these wicked men away
like sheep
to be slaughtered!
\par }{\Q Appoint
a time
when they will be killed!
\par }{\Q \VS{4}How long
must the land
be parched
\par }{\Q and the grass
in every
field
be withered?
\par }{\Q How long must the animals
and the birds
die
\par }{\Q because
of the wickedness
of the people who live
in this land?

\par }{\Q For these people boast,
\par }{\Q “God will not
see
what happens to us.”
\par }{\PP \VS{5}The
{\ND{Lord}} answered,

\par }{\Q “If
you have raced
on foot against men and they
have worn
you out,
\par }{\Q how
will you be able to compete
with
horses?
\par }{\Q And if you
feel secure
only in safe
and open country,
\par }{\Q how
will you
manage
in the thick undergrowth
along the Jordan River?
\par }{\Q \VS{6}As a matter of fact, even
your own brothers
\par }{\Q and the members
of your own family
have betrayed
you too.
\par }{\Q Even they
have plotted
to do away with
you.

\par }{\Q So do not
trust
them even when they say
kind things
to you.
\par }{\Q \VS{7}“I will abandon
my nation.
\par }{\Q I will forsake
the
people
I call my own.

\par }{\Q I will turn
my beloved
people

\par }{\Q over to the power
of their enemies.
\par }{\Q \VS{8}The people I call my own
have turned on
me
\par }{\Q like a lion
in the forest.
\par }{\Q They have roared
defiantly at me.
\par }{\Q So I will
treat them
as though I
hate them.
\par }{\Q \VS{9}The people
I call my own attack me like birds of prey
or like hyenas.
\par }{\Q But other birds of prey
are all around
them.

\par }{\Q Let
all
the nations gather together
like wild
beasts.
\par }{\Q Let them come
and destroy these people I call my own.
\par }{\Q \VS{10}Many
foreign rulers
will ruin
the land where I planted my people.
\par }{\Q They will trample all over
my chosen land.
\par }{\Q They will turn
my beautiful
land
\par }{\Q into a desolate
wasteland.
\par }{\Q \VS{11}They will lay
it waste.
\par }{\Q It will lie parched
and empty
before me.
\par }{\Q The whole
land
will be laid waste.
\par }{\Q But
no one
living
in it will pay
any heed.
\par }{\Q \VS{12}A destructive army
will come
marching
\par }{\Q over the hilltops
in the desert.
\par }{\Q For
the {\ND{Lord}}
will use them as his destructive
weapon

\par }{\Q against
everyone
from one end
of the land
to
the other.
\par }{\Q No
one
will be safe.
\par }{\Q \VS{13}My people will sow
wheat,
but will harvest
weeds.
\par }{\Q They will work until
they are exhausted,
but will get nothing
from it.
\par }{\Q They will be disappointed
in their harvests
\par }{\Q because the
{\ND{Lord}}
will take them away in his fierce
anger.
\par }{\PP \VS{14}“I, the
{\ND{Lord}}, also
have something
to say
concerning
the wicked
nations who surround
my land
and have
attacked and plundered
the land
that
I gave
to my people
as
a permanent possession.
I say: ‘I
will uproot
the
people of those nations
from
their lands
and I will free
the people of Judah
who have been
taken there.
\VS{15}But
after
I have uprooted
the people of those nations, I will relent
and have pity
on them. I will restore
the people
of each
of those nations to their own
lands and to their own
country.
\VS{16}But they must make sure you learn
to follow the religious
practices of my people.
Once
they taught
my people
to swear
their oaths using
the name of the god Baal.
But then,
they must swear
oaths using my name,
saying, “As surely
as the
{\ND{Lord}}
lives, I swear.” If they do these things,
then they will be included among
the people I call my own.
\VS{17}But I will completely uproot
and destroy
any of those nations
that
will not
pay heed,’ ”
says
the {\ND{Lord}}.

\par }\Chap{13}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
me, “Go
and buy
some linen
shorts
and put
them on.
Do not
put
them in water.”
\VS{2}So I bought
the shorts
as
the {\ND{Lord}}
had told
me to do and put
them on.
\VS{3}Then
the {\ND{Lord}}
spoke
to
me again
and said,
\VS{4}“Take
the shorts
that
you bought
and are wearing
and go
at once to Perath.
Bury
the shorts there in
a crack
in the rocks.”
\VS{5}So I went
and buried
them at Perath
as
the {\ND{Lord}}
had ordered me to do.
\VS{6}Many
days
later
the {\ND{Lord}}
said
to
me, “Go
at once to Perath
and get
the shorts
I ordered
you to bury
there.”
\VS{7}So I went
to Perath
and dug
up the
shorts
from
the place
where
I had
buried
them. I found
that they were ruined; they were good for nothing.
\par }{\PP \VS{8}Then
the {\ND{Lord}}
said
to me,
\VS{9}“I,
the {\ND{Lord}}, say: ‘This shows
how I will ruin
the highly exalted
position in which Judah
and Jerusalem
take pride.
\VS{10}These
wicked
people
refuse
to obey
what I have said.
They follow
the stubborn inclinations
of their own hearts
and pay allegiance
to other
gods
by worshiping
and serving
them. So they
will become
just like these
linen shorts
which
are good
for nothing.
\VS{11}For,’
I say, ‘just
as shorts
cling
tightly to a person’s
body,
so
I bound
the whole
nation
of Israel
and the whole
nation
of Judah
tightly to me.’ I
intended for them to be my special people
and to bring me fame,
honor,
and praise.
But they would not
obey me.
\par }{\PP \VS{12}“So tell
them, ‘The

{\ND{Lord}},
the
God
of Israel,
says,
“Every
wine jar
is made to be filled
with wine.” ’
And they will probably
say
to
you, ‘Do
you not
think we know
that
every
wine jar is supposed
to be filled
with wine?’
\VS{13}Then tell
them, ‘The
{\ND{Lord}}
says,
“I will soon
fill
all the people
who live
in this
land
with stupor. I will also fill the kings
from David’s
dynasty,
the priests,
the prophets,
and the
citizens
of Jerusalem
with stupor.
\VS{14}And I will smash
them like wine bottles against
one
another,
children
and parents
alike.
I will not
show any pity,
mercy, or
compassion.
Nothing will keep
me from destroying
them,’ says
the {\ND{Lord}}.”
\par }{\PP \VS{15}Then I said to the people of Judah,

\par }{\Q “Listen
and pay attention! Do not
be arrogant!
\par }{\Q For
the {\ND{Lord}}
has spoken.
\par }{\Q \VS{16}Show
the {\ND{Lord}}
your God
the respect
that is due him.

\par }{\Q Do it before
he brings the darkness
of disaster.

\par }{\Q Do it before
you stumble
into distress
\par }{\Q like a traveler on
the mountains
at twilight.
\par }{\Q Do it before he turns
the light
of deliverance you hope
for
\par }{\Q into the darkness
and gloom of exile.
\par }{\Q \VS{17}But if
you will not
pay attention
to this warning,

\par }{\Q I will weep
alone
because
of your arrogant pride.
\par }{\Q I will weep bitterly
and my eyes
will overflow
with tears
\par }{\Q because
you, the
{\ND{Lord}}’s
flock,
will be carried into exile.”
\par }{\PP \VS{18}The
{\ND{Lord}} told
me,

\par }{\Q “Tell the king
and the queen mother,
\par }{\Q ‘Surrender
your thrones,
\par }{\Q for
your glorious
crowns
\par }{\Q will be removed
from your heads.
\par }{\Q \VS{19}The gates of the towns
in southern
Judah will be shut
tight.

\par }{\Q No
one will be able
to go
in or out of them.

\par }{\Q All
Judah
will be carried off into exile.
\par }{\Q They will be completely carried off into exile.’ ”
\par }{\PP \VS{20}Then I said,

\par }{\Q “Look
up,
Jerusalem, and see
\par }{\Q the enemy that is coming
from the north.
\par }{\Q Where
now is the flock
of people that were entrusted
to your care?

\par }{\Q Where now are the ‘sheep’
that you take such pride in?
\par }{\Q \VS{21}What
will you say
when the
{\ND{Lord}} appoints
as rulers over
you
those allies
\par }{\Q that you, yourself, had actually prepared
as such?

\par }{\Q Then anguish
and agony
will grip
you
\par }{\Q like
that of a woman
giving birth to a baby.
\par }{\Q \VS{22}You will probably
ask yourself,

\par }{\Q ‘Why have these
things happened
to me?
\par }{\Q Why
have I been treated like a disgraced adulteress
\par }{\Q whose skirt
has been torn off and her limbs
exposed?’

\par }{\Q It is because
you have sinned so much.
\par }{\Q \VS{23}But there is little hope for you ever doing good,
\par }{\Q you
who are so
accustomed
to doing evil.
\par }{\Q Can
an Ethiopian
change
the color of his skin?
\par }{\Q Can a leopard
remove
its spots?
\par }{\PP \VS{24}“The
{\ND{Lord}} says,

\par }{\Q ‘That is why I will scatter
your people like chaff
\par }{\Q that is blown
away by
a desert
wind.
\par }{\Q \VS{25}This
is your fate,
\par }{\Q the destiny
to which I
have appointed
you,
\par }{\Q because
you have forgotten
me
\par }{\Q and have trusted
in false gods.
\par }{\Q \VS{26}So
I
will pull
your skirt
up
over
your face
\par }{\Q and expose
you to shame like a disgraced adulteress!
\par }{\Q \VS{27}People of Jerusalem,
I have seen your adulterous
worship,
\par }{\Q your shameless
prostitution
to, and your lustful pursuit of, other gods.

\par }{\Q I have seen
your disgusting acts
of worship

\par }{\Q on
the hills
throughout the countryside.
\par }{\Q You are doomed
to destruction!

\par }{\Q How
long
will you continue
to be unclean?’ ”

\par }\Chap{14}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
about
the drought.
\par }{\Q \VS{2}“The people of Judah
are in mourning.
\par }{\Q The people in her cities
are pining away.
\par }{\Q They lie on the ground
expressing their sorrow.
\par }{\Q Cries of distress
come up
to me from Jerusalem.
\par }{\Q \VS{3}The leading men
of the cities send
their servants
for water.
\par }{\Q They go
to the cisterns,
but they do not
find
any water
there.
\par }{\Q They return
with their containers
empty.
\par }{\Q Disappointed
and dismayed, they bury
their faces in their hands.
\par }{\Q \VS{4}They are dismayed because
the ground
is cracked
\par }{\Q because
there
has been no
rain
in the land.
\par }{\Q The farmers,
too, are dismayed
\par }{\Q and bury
their faces in their hands.
\par }{\Q \VS{5}Even
the doe
abandons her newborn
fawn
in the field
\par }{\Q because
there is no
grass.
\par }{\Q \VS{6}Wild donkeys
stand
on
the hilltops
\par }{\Q and pant
for breath
like jackals.
\par }{\Q Their eyes
are strained
looking for food,
\par }{\Q because
there is none to be found.”
\par }{\PP \VS{7}Then I said,

\par }{\Q “O
{\ND{Lord}}, intervene
for
the honor of your name
\par }{\Q even though
our sins
speak out
against us.

\par }{\Q Indeed,
we have turned away
from you many times.
\par }{\Q We have sinned against you.
\par }{\Q \VS{8}You have been the object of Israel’s
hopes.
\par }{\Q You have saved
them when
they were in trouble.
\par }{\Q Why
have you become
like a resident foreigner
in the land?
\par }{\Q Why have you become like a traveler
who only stops
in to spend the night?
\par }{\Q \VS{9}Why
should you be
like someone
who is helpless,
\par }{\Q like a champion
who cannot
save
anyone?
\par }{\Q You
are indeed with us,
\par }{\Q and we belong to you.
\par }{\Q Do not
abandon us!”
\par }{\PP \VS{10}Then
the {\ND{Lord}}
spoke
about these
people.
\par }{\Q “They truly
love
to go astray.
\par }{\Q They cannot
keep
from running
away
from me.

\par }{\Q So I
am not
pleased
with them.
\par }{\Q I will now
call to mind
the wrongs
they have done

\par }{\Q and punish
them for their sins.”
\par }{\SH Judgment for Believing the Misleading Lies of the False Prophets
\par }{\PP \VS{11}Then the
{\ND{Lord}}
said to me, “Do not
pray
for
good
to come to these
people!
\VS{12}Even if
they fast,
I will not
hear
their cries
for
help. Even if
they offer
burnt offerings
and grain offerings,
I will not
accept
them. Instead,
I
will kill them through wars,
famines,
and plagues.”
\par }{\PP \VS{13}Then I said,
“Oh,
Lord

{\ND{God}}, look! The prophets
are telling them that you said, ‘You will not
experience
war
or
suffer
famine.
I will give
you lasting
peace
and prosperity in
this land.’ ”
\par }{\PP \VS{14}Then the
{\ND{Lord}}
said
to
me, “Those prophets
are prophesying
lies
while claiming my authority! I did not
send
them. I did not
commission
them. I did not
speak
to them. They
are prophesying
to
these people false
visions,
worthless predictions,
and the delusions
of their own mind.
\VS{15}I
did not
send
those
prophets,
though they
claim to be prophesying
in my name.
They
may be saying,
‘No
war
or famine
will happen
in this land.’
But I,
the {\ND{Lord}}, say
this
about them: ‘War
and starvation
will kill those
prophets.’
\VS{16}The people
to whom
they
are prophesying
will die through war
and famine.
Their
bodies will be
thrown
out into the streets
of Jerusalem
and there will be no
one to bury
them.
This will happen to the men and their
wives,
their sons,
and their daughters.
For I will pour
out on
them the destruction they deserve.”
\par }{\SH Lament over Present Destruction and Threat of More to Come
\par }{\Q \VS{17}“Tell
these people this,
Jeremiah:

\par }{\Q ‘My eyes
overflow
with tears
\par }{\Q day
and night
without
ceasing.
\par }{\Q For
my people, my dear children,
have suffered
a crushing
blow.
\par }{\Q They have suffered
a serious
wound.
\par }{\Q \VS{18}If
I go out
into the countryside,
\par }{\Q I see
those who have been killed
in battle.
\par }{\Q If
I go
into the city,
\par }{\Q I see
those who are sick
because of starvation.
\par }{\Q For
both
prophet
and priest
go about their own business
\par }{\Q in the land
without
having any real understanding.’ ”
\par }{\PP \VS{19}Then I
said,
\par }{\Q “{\ND{Lord}}, have you completely rejected
the nation of Judah?
\par }{\Q Do you despise the city
of Zion?
\par }{\Q Why
have you struck
us with such force
\par }{\Q that we are beyond recovery?

\par }{\Q We hope
for peace,
but nothing
good
has come of it.
\par }{\Q We hope for a time
of relief
from our troubles, but experience terror.
\par }{\Q \VS{20}{\ND{Lord}}, we confess
that we have been wicked.
\par }{\Q We confess
that our ancestors
have done wrong.

\par }{\Q We have indeed
sinned against you.
\par }{\Q \VS{21}For
the honor of your name,
do not
treat
Jerusalem with contempt.
\par }{\Q Do not
treat
with
disdain the place where your glorious
throne
sits.

\par }{\Q Be mindful
of your covenant
with
us.
Do not
break it!
\par }{\Q \VS{22}Do any of the worthless idols
of the nations
cause rain
to fall?
\par }{\Q Do the skies
themselves send
showers?
\par }{\Q Is it not
you,
O Lord
our God,
who does this?

\par }{\Q So we put
our hopes
in you

\par }{\Q because
you
alone do
all
this.”


\par }\Chap{15}{\PP \VerseOne{1}Then the
{\ND{Lord}}
said to me,
“Even if
Moses
and Samuel
stood
before
me pleading
for these
people,
I would not feel pity for them! Get them away from me! Tell them to go away!
\VS{2}If
they ask
you, ‘Where
should we go?’ tell
them the
{\ND{Lord}}
says
this:
\par }{\Q “Those who
are destined to die of disease
will go to death by disease.
\par }{\Q Those who are destined to die in war
will go to death in war.
\par }{\Q Those who
are destined to die of starvation
will go to death by starvation.
\par }{\Q Those who
are destined to go into exile
will go into exile.”
\par }{\PP \VS{3}“I will punish
them in
four
different ways: I will have war
kill
them. I will have dogs
drag off
their dead bodies. I will have birds
and wild
beasts
devour
and destroy their corpses.
\VS{4}I will make
all the people
in all
the kingdoms
of the world
horrified
at what has happened to them because
of what Hezekiah’s
son
Manasseh,
king
of Judah,
did
in Jerusalem.”
\par }{\PP \VS{5}The
{\ND{Lord}} cried out,

\par }{\Q “Who
in the world will have pity
on
you, Jerusalem?
\par }{\Q Who
will grieve
over you?
\par }{\Q Who
will stop
long enough

\par }{\Q to inquire
about how you are doing?
\par }{\Q \VS{6}I,
the {\ND{Lord}}, say: ‘You
people have deserted
me!
\par }{\Q You keep turning
your back on me.’
\par }{\Q So
I have unleashed
my power
against
you and have begun to destroy
you.

\par }{\Q I have grown tired
of feeling sorry for you!”
\par }{\PP \VS{7}The
{\ND{Lord}} continued,

\par }{\Q “In every town
in the land
I will purge
them
\par }{\Q like straw blown away by the wind.
\par }{\Q I will destroy
my people.
\par }{\Q I will
kill off
their children.
\par }{\Q I will do so because they did not
change
their behavior.
\par }{\Q \VS{8}Their widows
will become in my sight more numerous
\par }{\Q than the grains of sand
on the seashores.
\par }{\Q At noontime
I will bring
a destroyer
\par }{\Q against the mothers
of their
young men.
\par }{\Q I will cause anguish and terror
\par }{\Q to fall
suddenly upon them.
\par }{\Q \VS{9}The mother who had
seven
children
will grow faint.
\par }{\Q All the breath
will go out
of her.

\par }{\Q Her pride and joy will be taken from her in the prime of their life.
\par }{\Q It will seem as if the sun
had set while it was still
day.
\par }{\Q She will suffer shame
and humiliation.
\par }{\Q I will cause
any of them who are still left alive
\par }{\Q to be killed in war by the onslaughts
of
their enemies,”
\par }{\PP says
the {\ND{Lord}}.
\par }{\SH Jeremiah Complains about His Lot and The Lord Responds
\par }{\PP \VS{10}I said,

\par }{\Q “Oh,
mother,
how I regret
that
you ever gave birth
to me!
\par }{\Q I am always starting arguments
and quarrels
with the people
of this land.
\par }{\Q I have not
lent money
to anyone and I have not
borrowed
from anyone.
\par }{\Q Yet all
of these people are treating me with contempt.”
\par }{\PP \VS{11}The
{\ND{Lord}}
said,
\par }{\Q “Jerusalem, I will surely
send
you away
for your own good.
\par }{\Q I will surely
bring
the
enemy
upon
you in a time
of trouble
and distress.
\par }{\Q \VS{12}Can you people who are like iron
and bronze
\par }{\Q break that iron
fist from the north?
\par }{\Q \VS{13}I will give away your wealth
and your treasures
as plunder.
\par }{\Q I will give
it away free
of charge
for the sins
you have committed throughout
your land.
\par }{\Q \VS{14}I will make you serve
your enemies
in a land
that you know
nothing
about.
\par }{\Q For
my anger
is like
a fire
that will burn against you.”
\par }{\PP \VS{15}I said,

\par }{\Q “{\ND{Lord}}, you
know
how I
suffer.

\par }{\Q Take
thought of me and care
for me.
\par }{\Q Pay back
for me those who have been persecuting
me.
\par }{\Q Do not
be so
patient
with them
that you allow them to kill me.
\par }{\Q Be mindful
of how I have put up with their insults for your sake.
\par }{\Q \VS{16}As your words
came
to me I drank
them in,

\par }{\Q and they filled my heart
with joy
and happiness
\par }{\Q because
I belong
to you.
\par }{\Q \VS{17}I did not
spend my time in
the company of other people,
\par }{\Q laughing
and having a good time.
\par }{\Q I stayed
to myself because
I felt obligated
to you

\par }{\Q and because I was filled
with anger at what they had done.
\par }{\Q \VS{18}Why
must I continually
suffer such painful
anguish?
\par }{\Q Why must I endure the sting of their insults like an incurable
wound?
\par }{\Q Will you let me down
when I need you
\par }{\Q like
a brook
one goes to for water,
but that cannot
be relied on?”
\par }{\PP \VS{19}Because
of this, the
{\ND{Lord}}
said,
\par }{\Q “You must repent
of such words and thoughts!
\par }{\Q If
you do, I will restore
you to the privilege of serving
me.

\par }{\Q If
you say
what is worthwhile
instead of what is worthless,
\par }{\Q I will again allow you to be
my spokesman.
\par }{\Q They
must become
as
you have been.
\par }{\Q You
must not
become
like them.
\par }{\Q \VS{20}I will make
you as strong as a wall to these
people,
\par }{\Q a fortified
wall
of bronze.
\par }{\Q They will attack
you,
\par }{\Q but they will not
be able
to overcome you.
\par }{\Q For
I
will be with
you to rescue
you and deliver
you,”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{21}“I will deliver
you from the power
of the wicked.
\par }{\Q I will free you
from the clutches of violent people.”

\par }\Chap{16}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to me,
\VS{2}“Do not
get
married
and do not
have children
here in this
land.
\VS{3}For
I, the
{\ND{Lord}}, tell
you what
will happen to the
children
who are born
here
in this
land
and to the men and women
who
are their mothers
and fathers.
\VS{4}They will die
of deadly
diseases.
No
one will mourn
for them. They will not
be buried.
Their dead bodies will lie like manure
spread on
the ground.
They will be
killed
in war
or die of starvation.
Their corpses
will be
food
for the birds
and wild
animals.
\par }{\PP \VS{5}“Moreover
I, the
{\ND{Lord}}, tell
you: ‘Do not
go
into a house
where they are having a funeral
meal. Do not
go
there to mourn
and express
your sorrow for
them. For I have stopped showing
them my good favor,
my love,
and my compassion.
I, the
{\ND{Lord}}, so affirm it!
\VS{6}Rich
and poor
alike will die
in this
land.
They will not
be buried
or
mourned.
People will not
cut
their
bodies
or
shave off their hair
to show their grief for them.
\VS{7}No one
will take
any food to those who mourn
for
the dead
to comfort
them. No
one will give
them any wine
to drink
to console
them for
the loss of their father
or mother.
\par }{\PP \VS{8}“‘Do not
go
to a house
where people are feasting
and sit
down to eat
and drink
with them either.
\VS{9}For
I,
the {\ND{Lord}}
God
of Israel
who rules over all,
tell
you what
will happen.
I will put an end
to the sounds
of joy
and gladness,
to the glad celebration
of brides
and grooms
in this
land.
You and the rest of the people will live
to see
this happen.’ ”
\par }{\SH The Lord Promises Exile (But Also Restoration)
\par }{\PP \VS{10}“When
you tell
these
people
about all
this,
they will undoubtedly ask
you, ‘Why
has the
{\ND{Lord}}
threatened
us with
such
great
disaster? What
wrong
have we done? What
sin
have we done to offend
the {\ND{Lord}}
our God?’
\VS{11}Then tell
them
that
the {\ND{Lord}}
says, ‘It is because your ancestors
rejected
me
and paid allegiance
to other
gods.
They have served
them and worshiped
them. But they have rejected
me and not
obeyed
my law.
\VS{12}And you
have acted
even more wickedly
than your ancestors! Each
one of you has followed
the stubborn inclinations
of your own wicked
heart
and not
obeyed me.
\VS{13}So I will throw
you out of this
land
into a land
that
neither
you
nor your ancestors
have ever known.
There
you must worship
other
gods
day
and night,
for I will show you no
mercy.’ ”
\par }{\PP \VS{14}Yet I, the
{\ND{Lord}}, say: “A new time
will certainly
come.
People now affirm
their oaths
with ‘I swear as surely
as the
{\ND{Lord}}
lives who
delivered
the
people of Israel
out of Egypt.’
\VS{15}But in that
time they will affirm them with ‘I swear as surely
as the
{\ND{Lord}}
lives who
delivered
the people of Israel
from the land
of the north
and from all
the other lands
where
he
had
banished
them.’ At that
time I will bring them back
to the land
I gave
their ancestors.”
\par }{\PP \VS{16}But for now
I, the
{\ND{Lord}}, say: “I will send
many
enemies
who will catch these people like fishermen.
After
that
I will send
others
who will hunt
them out like hunters
from
all
the mountains,
all
the hills,
and the crevices
in the rocks.
\VS{17}For
I see
everything
they do. Their wicked
ways
are not
hidden
from me. Their sin
is not hidden
away where I cannot
see it.
\VS{18}Before
I restore them I will punish
them in full
for their sins
and the wrongs
they have done. For
they have polluted
my land
with the lifeless statues
of their disgusting idols.
They have filled
the land I have claimed
as my own with their detestable idols.”
\par }{\PP \VS{19}Then I said,

\par }{\Q “{\ND{Lord}}, you give me strength
and protect
me.
\par }{\Q You are the one
I can
run
to for safety when
I am in trouble.
\par }{\Q Nations
from all
over the earth
\par }{\Q will come
to you and say,
\par }{\Q ‘Our
ancestors
had nothing but false gods –
\par }{\Q worthless idols that could not help them at all.
\par }{\Q \VS{20}Can
people
make
their own gods?
\par }{\Q No, what
they
make
are not
gods at all.”
\par }{\PP \VS{21}The
{\ND{Lord}} said,

\par }{\Q “So
I
will now
let
this wicked people know
–
\par }{\Q I will let them know
my
mighty
power
in judgment.
\par }{\Q Then they will know
that
my name
is the
{\ND{Lord}}.”

\par }\Chap{17}{\PP \VerseOne{1}\par }{\Q The sin
of Judah
is engraved
with an iron
chisel
\par }{\Q on their stone-hard
hearts.
\par }{\Q It is inscribed
with a diamond
point
\par }{\Q on
the horns
of their altars.
\par }{\Q \VS{2}Their children
are always thinking about
their altars
\par }{\Q and their sacred poles dedicated to the goddess Asherah,
\par }{\Q set up beside
the green
trees
on
the high
hills
\par }{\Q \VS{3}and on the mountains
and in the fields.
\par }{\Q I will give your wealth
and all
your treasures
away as plunder.
\par }{\Q I will give
it away as the price for the sins
you have committed throughout
your land.
\par }{\Q \VS{4}You will lose your hold
on the land
\par }{\Q which
I gave
to you as a permanent possession.
\par }{\Q I will make you serve
your enemies
in a land
that
you know
nothing
about.
\par }{\Q For
you have made my anger
burn
like a fire
that will never be put out.”
\par }{\SH Individuals Are Challenged to Put Their Trust in the Lord
\par }{\PP \VS{5}The

{\ND{Lord}}
says,
\par }{\Q “I
will put
a curse
on people
\par }{\Q who trust in mere human
beings,
\par }{\Q who depend
on mere
flesh
and blood for their strength,
\par }{\Q and whose hearts
have turned
away from
the {\ND{Lord}}.
\par }{\Q \VS{6}They will be
like a shrub
in the desert.
\par }{\Q They will not
experience
good things
even when they happen.
\par }{\Q It will be as though they were growing
in the desert,
\par }{\Q in a salt
land
where no one
can live.
\par }{\Q \VS{7}My blessing
is on those
people
who trust
in me,
\par }{\Q who put their confidence in me.
\par }{\Q \VS{8}They will be
like a tree
planted
near
a stream
\par }{\Q whose roots
spread out
toward
the water.
\par }{\Q It has nothing
to fear
when
the heat
comes.
\par }{\Q Its leaves
are always green.
\par }{\Q It has no
need to be concerned
in a year
of drought.
\par }{\Q It does not
stop
bearing
fruit.
\par }{\Q \VS{9}The human mind
is more deceitful
than anything else.
\par }{\Q It is incurably
bad. Who
can understand it?
\par }{\Q \VS{10}I,
the {\ND{Lord}}, probe
into people’s minds.
\par }{\Q I examine
people’s hearts.
\par }{\Q I deal with each person
according to how he has behaved.
\par }{\Q I give
them what they deserve
based on what they have done.
\par }{\Q \VS{11}The person who gathers
wealth
by unjust
means
\par }{\Q is like the partridge
that broods over eggs
but does not
hatch
them.

\par }{\Q Before his life
is half
over he will lose
his ill-gotten gains.

\par }{\Q At the end
of his life it will be
clear he was a fool.”
\par }{\SH Jeremiah Appeals to the Lord for Vindication
\par }{\PP \VS{12}Then I said,

\par }{\Q “{\ND{Lord}}, from the very beginning
\par }{\Q you have been seated on your glorious
throne
on high.
\par }{\Q You are the place
where we can find refuge.
\par }{\Q \VS{13}You are the one in whom Israel
may find hope.
\par }{\Q All
who leave
you will suffer shame.
\par }{\Q Those who turn away
from you will be consigned
to the nether world.
\par }{\Q For
they have rejected
you, the

{\ND{Lord}}, the fountain
of life.
\par }{\Q \VS{14}{\ND{Lord}}, grant me relief
from my suffering
\par }{\Q so that
I may have some relief.
\par }{\Q Rescue
me from those who persecute me
\par }{\Q so that I may be rescued.
\par }{\Q \VS{15}Listen
to what
they
are saying
to me.
\par }{\Q They are saying, “Where
are the things
the {\ND{Lord}}
threatens us with?
\par }{\Q Come on! Let’s see them happen!”
\par }{\Q \VS{16}But I
have not
pestered
you to bring
disaster.
\par }{\Q I have not
desired
the time
of irreparable devastation.
\par }{\Q You
know
that.
\par }{\Q You are fully aware of every word
that I have spoken.
\par }{\Q \VS{17}Do not
cause
me dismay!

\par }{\Q You are my source of safety in times
of trouble.
\par }{\Q \VS{18}May those who persecute
me be disgraced.
\par }{\Q Do not
let me
be disgraced.
\par }{\Q May they
be dismayed.
\par }{\Q Do not
let me
be dismayed.
\par }{\Q Bring
days
of disaster
on them.
\par }{\Q Bring on them the destruction they deserve.”
\par }{\SH Observance of the Sabbath Day Is a Key to the Future
\par }{\PP \VS{19}The

{\ND{Lord}}
told
me,
“Go
and stand
in the People’s
Gate
through which
the kings
of Judah
enter
and leave
the city. Then go and stand in all
the other gates
of the city of Jerusalem.
\VS{20}As you stand in those
places announce,
‘Listen,
all
you people who
pass
through these
gates.
Listen, all you kings
of Judah,
all
you people of Judah
and all
you citizens
of Jerusalem.
Listen
to what
the {\ND{Lord}} says.
\VS{21}The
{\ND{Lord}}
says,
‘Be very careful
if you value your lives! Do not
carry
any loads
in through
the gates
of Jerusalem
on the Sabbath
day.
\VS{22}Do not
carry any loads
out
of your houses
or
do
any
work
on the Sabbath
day.
But observe
the Sabbath
day
as
a day set apart
to the
{\ND{Lord}}, as I commanded
your ancestors.
\VS{23}Your ancestors,
however,
did not
listen
to me or pay
any attention
to me. They stubbornly
refused to pay attention
or to respond
to any
discipline.’
\VS{24}The
{\ND{Lord}}
says, ‘You must make sure to
obey
me. You must not
bring
any loads
through the gates
of this
city
on the Sabbath
day.
You must set the Sabbath
day
apart
to me. You must not
do
any
work on that day.
\VS{25}If you do this,
then the kings
and princes
who
follow in
David’s
succession
and ride
in chariots
or on horses
will continue to enter
through these
gates,
as well as their
officials
and the people
of Judah
and the citizens
of Jerusalem.
This
city
will always be filled with people.
\VS{26}Then people will come
here from the towns
in Judah,
from the villages surrounding
Jerusalem,
from the territory
of Benjamin,
from
the western foothills,
from
the southern hill country,
and from
the southern
part of Judah. They will come
bringing
offerings to the temple
of the {\ND{Lord}}: burnt offerings,
sacrifices,
grain offerings,
and incense
along with their thank offerings.
\VS{27}But you
must
obey
me and set
the
Sabbath
day
apart
to me. You must not
carry
any loads
in through the gates
of Jerusalem
on the Sabbath
day.
If you disobey, I will set
the gates
of Jerusalem on fire.
It will burn down
all the fortified dwellings
in Jerusalem
and no one
will be able to put it out.’ ”

\par }\Chap{18}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
Jeremiah:
\VS{2}“Go
down
at once to the potter’s
house.
I will speak to you further
there.”
\VS{3}So I went down
to the potter’s
house
and found
him working
at his wheel.
\VS{4}Now and then there would be something wrong
with the pot
he was
molding
from the clay
with his hands.
So he would rework
the clay
into another
kind of pot
as
he saw fit.
\par }{\PP \VS{5}Then
the {\ND{Lord}}
said
to me,
\VS{6}“I,
the {\ND{Lord}}, say: ‘O nation
of Israel,
can
I
not
deal
with you as this
potter
deals
with the clay? In my hands,
you,
O nation
of Israel,
are just like the clay in this
potter’s
hand.’
\VS{7}There are times,
Jeremiah, when I threaten to uproot,
tear down,
and destroy
a nation
or kingdom.
\VS{8}But if that
nation
I threatened
stops doing wrong,
I will cancel
the destruction
I intended
to do to it.
\VS{9}And there are times
when
I promise
to build
up and establish
a nation
or kingdom.
\VS{10}But
if that nation does what
displeases
me and does not
obey
me,
then I will cancel
the good
I promised
to do to it.
\VS{11}So now,
tell
the people
of Judah
and the citizens
of Jerusalem
this: The
{\ND{Lord}}
says,
‘I am
preparing
to bring disaster
on
you! I am making plans
to punish you. So, every one
of you, stop the
evil
things you have been doing.
Correct the way
you have been living and do what is right.’
\VS{12}But they just keep saying,
‘We do not care what you say! We will do
whatever we want to do! We will continue
to behave wickedly
and stubbornly!’ ”
\par }{\PP \VS{13}Therefore,
the {\ND{Lord}}
says,
\par }{\Q “Ask
the people of other nations
\par }{\Q whether they have heard
of anything like this.
\par }{\Q Israel
should have been like a virgin.
\par }{\Q But she has done
something
utterly
revolting!
\par }{\Q \VS{14}Does
the snow
ever completely vanish
from the rocky
slopes
of Lebanon?
\par }{\Q Do
the cool
waters
from those distant mountains ever cease to flow?
\par }{\Q \VS{15}Yet
my people
have forgotten
me
\par }{\Q and offered sacrifices
to worthless
idols!
\par }{\Q This makes them stumble
along in the way
they live
\par }{\Q and leave the old reliable
path
of their fathers.

\par }{\Q They have left them to walk
in bypaths,
\par }{\Q in roads
that are not
smooth and level.
\par }{\Q \VS{16}So their land
will become
an object of horror.
\par }{\Q People will forever
hiss out
their scorn over
it.
\par }{\Q All
who pass
that way will be filled with horror
\par }{\Q and will shake
their heads in derision.
\par }{\Q \VS{17}I will scatter
them before
their enemies
\par }{\Q like dust blowing
in front
of a
burning east
wind.
\par }{\Q I will turn my back on them and not
look
favorably on them

\par }{\Q when
disaster strikes them.”
\par }{\SH Jeremiah Petitions the Lord to Punish Those Who Attack Him
\par }{\PP \VS{18}Then some people said,
“Come
on! Let us consider
how to deal
with Jeremiah! There will still be priests
to instruct us, wise men
to give us advice,
and prophets
to declare God’s word.
Come on! Let’s bring charges against him and get rid of
him! Then we will not
need to pay attention
to
anything
he says.”
\par }{\PP \VS{19}Then I said,

\par }{\Q “{\ND{Lord}}, pay attention
to
me.
\par }{\Q Listen
to what my enemies
are saying.
\par }{\Q \VS{20}Should
good
be paid
back with evil?
\par }{\Q Yet
they are
virtually
digging a pit
to kill
me.

\par }{\Q Just remember
how I stood
before
you
\par }{\Q pleading
on
their behalf
\par }{\Q to keep you from venting
your anger
on them.
\par }{\Q \VS{21}So
let
their children
die of starvation.
\par }{\Q Let them be cut down
by
the sword.
\par }{\Q Let
their wives
lose
their husbands
and children.
\par }{\Q Let the older men
die
of disease
\par }{\Q and the younger men
die
by the sword
in battle.
\par }{\Q \VS{22}Let cries
of terror be heard
in their houses
\par }{\Q when
you send
bands of raiders
unexpectedly
to plunder them.

\par }{\Q For
they have virtually dug
a pit
to capture
me
\par }{\Q and have hidden
traps
for me to step into.
\par }{\Q \VS{23}But you,

{\ND{Lord}}, know
\par }{\Q all
their plots
to kill me.
\par }{\Q Do not
pardon their crimes!
\par }{\Q Do
not
ignore their sins
as though you had erased
them!

\par }{\Q Let
them be
brought down in defeat
before
you!
\par }{\Q Deal
with them while you are still angry!

\par }\Chap{19}{\PP \VerseOne{1}The
{\ND{Lord}}
told
Jeremiah, “Go
and buy
a clay jar
from a potter.
Take with you some of the leaders
of the people
and some of the leaders
of the priests.
\VS{2}Go out
to
the part of the Hinnom
Valley
which
is near the entrance
of the Potsherd
Gate.
Announce
there
what
I tell you.
\VS{3}Say,
‘Listen
to what
the {\ND{Lord}}
says, you kings
of Judah
and citizens
of Jerusalem! The
{\ND{Lord}}
God
of Israel
who rules over all
says, “I
will bring
a disaster
on
this
place
that
will make the ears
of everyone
who hears
about it ring!
\VS{4}I will do so because
these people have rejected
me and have defiled
this
place.
They have offered
sacrifices in it to other
gods
which
neither they
nor their ancestors
nor the kings
of Judah
knew anything about. They have filled
it with the
blood
of innocent children.
\VS{5}They have built
places
here for worship
of the god Baal
so that they could sacrifice
their children
as burnt offerings
to him in the fire.
Such sacrifices are something I never
commanded
them to make! They are something I never
told
them to do! Indeed, such a thing never
even entered
my mind!
\VS{6}So
I,
the {\ND{Lord}}, say: “The time
will soon come
that people will no
longer
call
this
place
Topheth
or the Hinnom
Valley.
But
they will call this valley the Valley
of Slaughter!
\VS{7}In this
place
I will thwart
the plans
of the people of Judah
and Jerusalem.
I will deliver them over to the power of their enemies
who are seeking
to kill them. They will die by the sword
at the hands
of their enemies. I will make
their dead bodies
food
for the birds
and wild beasts to eat.
\VS{8}I will make
this
city
an object of horror,
a thing to be hissed
at. All
who pass by
it will be filled with horror
and will hiss out their scorn
because
of all
the disasters that have happened to it.
\VS{9}I will reduce
the people of this city to desperate
straits
during the siege
imposed on it by their enemies
who are seeking
to kill
them. I will make them so desperate that they will eat
the flesh
of their own sons
and daughters
and the flesh
of one
another.” ’ ”
\par }{\PP \VS{10}The
{\ND{Lord}} continued, “Now break
the jar
in front
of those who
have come
here with you.
\VS{11}Tell
them
the

{\ND{Lord}}
who rules over all
says, ‘I will do just
as Jeremiah has done.
I will smash
this
nation
and this
city
as
though it were
a potter’s
vessel
which
is broken
beyond repair.
The dead will
be buried
here in Topheth
until there is no
more room
to bury them.’
\VS{12}I, the
{\ND{Lord}}, say: ‘That is how I will deal with this
city
and its citizens.
I will make
it like Topheth.
\VS{13}The houses
in Jerusalem
and the houses
of the kings
of Judah
will be defiled
by dead bodies just like this place,
Topheth.
For they offered sacrifice
to the stars
and poured out
drink offerings
to other
gods
on
the roofs
of those
houses.’ ”
\par }{\PP \VS{14}Then Jeremiah
left
Topheth
where
the {\ND{Lord}}
had sent
him to give that prophecy.
He went to the
{\ND{Lord}}’s
temple
and stood
in its courtyard
and called out
to
all
the people.
\VS{15}“The

{\ND{Lord}}
God
of Israel
who rules over all
says, ‘I will soon
bring
on
this
city
and all
the towns
surrounding it all
the disaster
I threatened
to do to it. I will do so because
they have stubbornly
refused to pay
any attention
to what
I have said!’ ”

\par }\Chap{20}{\PP \VerseOne{1}Now Pashhur
son
of Immer
heard
Jeremiah
prophesy
these
things.
He was
the priest
who was
chief of security in the
{\ND{Lord}}’s
temple.
\VS{2}When he
heard Jeremiah’s
prophecy, he had the prophet
flogged.
Then he put
him in
the stocks
which
were at the Upper
Gate
of Benjamin
in the
{\ND{Lord}}’s
temple.
\VS{3}But the next day
Pashhur
released
Jeremiah
from
the stocks.
When he did, Jeremiah
said
to him,
“The
{\ND{Lord}}’s
name
for
you is not
‘Pashhur’
but
‘Terror is Everywhere.’
\VS{4}For
the {\ND{Lord}}
says,
‘I
will make
both you and your friends
terrified
of what will happen to you. You will see
all
of them die
by the swords
of their enemies.
I will
hand all
the people of Judah
over
to the king
of Babylon.
He will carry some of them away into exile
in Babylon
and he will kill
others of them with the sword.
\VS{5}I will hand over
all
the wealth
of this
city
to their enemies. I will hand over to them all
the fruits of the labor
of the people of this city and all
their prized possessions,
as well as all
the treasures
of the kings
of Judah.
Their enemies
will seize
it all as plunder
and carry
it off to Babylon.
\VS{6}You,
Pashhur,
and all
your household
will go
into exile
in Babylon.
You will die
there
and you
will be buried
there.
The same thing will happen to all
your friends
to whom
you have prophesied
lies.’ ”
\par }{\SH Jeremiah Complains about the Reaction to His Ministry
\par }{\Q \VS{7}{\ND{Lord}}, you coerced
me into being a prophet,
\par }{\Q and I allowed you to do it.
\par }{\Q You overcame
my resistance and prevailed
over me.

\par }{\Q Now I have become
a constant laughingstock.
\par }{\Q Everyone
ridicules me.
\par }{\Q \VS{8}For
whenever
I prophesy,
I must cry out,
\par }{\Q “Violence
and destruction
are coming!”

\par }{\Q This
message
from the
{\ND{Lord}}
has made
me
\par }{\Q an object of
continual insults
and derision.
\par }{\Q \VS{9}Sometimes I think,
“I will make no
mention
of his message.
\par }{\Q I will not
speak
as his messenger any more.”
\par }{\Q But then
his message becomes
like a fire
\par }{\Q locked up inside
of me, burning
in my heart
and soul.
\par }{\Q I grow weary
of trying
to hold it in;
\par }{\Q I cannot
contain it.
\par }{\Q \VS{10}I hear
many
whispering words of intrigue against
me.
\par }{\Q Those who would cause me terror
are everywhere!

\par }{\Q They are saying, “Come on, let’s publicly
denounce
him!”

\par }{\Q All
my so-called
friends
are just watching
for
\par }{\Q something that would lead to my downfall.
\par }{\Q They say, “Perhaps
he can be enticed
into slipping up,
\par }{\Q so we can prevail over
him and get
our revenge
on him.
\par }{\Q \VS{11}But the
{\ND{Lord}}
is with
me to help me like an awe-inspiring
warrior.
\par }{\Q Therefore
those who persecute
me will fail
and will not
prevail
over
me.
\par }{\Q They will be thoroughly
disgraced
because
they did not
succeed.
\par }{\Q Their disgrace
will never
be forgotten.
\par }{\Q \VS{12}O
{\ND{Lord}}
who rules over all,
you test
and prove the righteous.
\par }{\Q You see
into people’s hearts
and minds.
\par }{\Q Pay
them back
for what they
have done
\par }{\Q because
I trust you to vindicate
my cause.
\par }{\Q \VS{13}Sing
to the
{\ND{Lord}}! Praise
the {\ND{Lord}}!
\par }{\Q For
he rescues
the oppressed from the clutches
of evildoers.
\par }{\Q \VS{14}Cursed
be the day
I was born!
\par }{\Q May that day
not
be
blessed
when my mother
gave birth to me.
\par }{\Q \VS{15}Cursed
be the man
\par }{\Q who
made my father
very glad
\par }{\Q when he brought him the news
\par }{\Q that a baby
boy
had been
born to him!
\par }{\Q \VS{16}May that man
be like the cities
\par }{\Q that
the
{\ND{Lord}}
destroyed
without
showing
any mercy.
\par }{\Q May he hear
a cry of distress
in the morning
\par }{\Q and a battle cry
at noon.
\par }{\Q \VS{17}For he did not
kill
me before I came from the womb,
\par }{\Q making my pregnant
mother’s
womb
my grave
forever.
\par }{\Q \VS{18}Why
did I ever
come forth
from my mother’s womb?
\par }{\Q All
I experience
is trouble
and grief,
\par }{\Q and I spend
my days
in shame.

\par }\Chap{21}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
when King
Zedekiah
sent
to him
Pashhur
son
of Malkijah
and the priest
Zephaniah
son
of Maaseiah.
Zedekiah sent them to Jeremiah to ask,
\VS{2}“Please
ask
the

{\ND{Lord}}
to come and help us, because
King
Nebuchadnezzar
of Babylon
is attacking
us. Maybe
the {\ND{Lord}}
will perform
one of his miracles
as
in times past and make him stop attacking us and leave.”
\VS{3}Jeremiah
answered
them, “Tell
Zedekiah
\VS{4}that
the {\ND{Lord}},
the God
of Israel,
says, ‘The
forces
at your disposal
are now outside
the walls
fighting
against King
Nebuchadnezzar of Babylon
and the
Babylonians
who have you under siege.
I will gather
those
forces back inside
the city.
\VS{5}In anger,
in fury,
and in wrath
I myself
will fight
against you with
my
mighty power
and great
strength!
\VS{6}I will kill
everything living
in Jerusalem,
people
and animals
alike! They will die
from terrible
diseases.
\VS{7}Then
I, the
{\ND{Lord}}, promise
that I will hand
over King
Zedekiah
of Judah,
his officials,
and any of the people
who survive
the war,
starvation,
and disease.
I will hand
them over to King
Nebuchadnezzar
of Babylon
and to their enemies
who want
to kill them.
He will slaughter
them with the sword.
He will
not
show them any mercy,
compassion,
or
pity.’
\par }{\PP \VS{8}“But tell the people
of Jerusalem that the

{\ND{Lord}}
says, ‘I
will give
you a choice between two
courses
of action. One will result
in life;
the
other will result
in death.
\VS{9}Those who stay
in this
city
will die
in battle
or of starvation
or disease.
Those who leave
the city and surrender
to the Babylonians
who are besieging
it will live.
They will
escape
with their lives.
\VS{10}For
I, the
{\ND{Lord}}, say that I am determined
not
to deliver this
city
but to bring disaster
on it. It will be handed
over to the king
of Babylon
and he will
destroy
it with fire.’ ”
\par }{\SH Warnings to the Royal Court
\par }{\PP \VS{11}The
{\ND{Lord}} told
me to say to the royal
court
of Judah,
\par }{\Q “Listen
to what
the {\ND{Lord}} says,
\par }{\Q \VS{12}O royal family
descended from David.
\par }{\Q The

{\ND{Lord}}
says:
\par }{\Q ‘See
to it
that people each
day are judged
fairly.
\par }{\Q Deliver
those who have been robbed
from those who oppress
them.
\par }{\Q Otherwise,
my wrath
will blaze
out
against you.
\par }{\Q It will burn like a fire
that cannot be put out
\par }{\Q because of the evil that you have done.
\par }{\Q \VS{13}Listen,
you who sit enthroned
above the valley
on a rocky
plateau.
\par }{\Q I am opposed to you,’ says
the {\ND{Lord}}.

\par }{\Q ‘You boast,
“No one can swoop down
on
us.
\par }{\Q No one can penetrate
into our places of refuge.”
\par }{\Q \VS{14}But I will punish
you as your deeds
deserve,’
\par }{\Q says
the {\ND{Lord}}.

\par }{\Q ‘I will set
fire
to your palace;
\par }{\Q it will burn up
everything
around it.’ ”

\par }\Chap{22}{\PP \VerseOne{1}The

{\ND{Lord}}
told
me, “Go down
to the palace
of the king
of Judah.
Give
him
a message
from me there.
\VS{2}Say: ‘Listen,
O king
of Judah
who follows
in David’s
succession.
You,
your officials,
and your subjects
who pass through
the gates
of this
palace must listen to what
the {\ND{Lord}} says.
\VS{3}The
{\ND{Lord}}
says,
“Do
what is just
and right.
Deliver
those who have been robbed
from those who oppress
them. Do not
exploit
or mistreat
foreigners
who live in your land, children who have no fathers,
or widows.
Do not
kill
innocent
people
in this
land.
\VS{4}If
you are
careful to obey
these
commands, then
the kings
who
follow in
David’s
succession
and ride
in chariots
or on horses
will continue
to come
through the gates
of this
palace,
as will their officials
and their subjects.
\VS{5}But, if
you do not
obey
these
commands, I solemnly
swear that
this
palace
will become
a pile of rubble.
I, the
{\ND{Lord}}, affirm it!”
\par }{\Q \VS{6}“‘For
the {\ND{Lord}}
says
concerning
the palace
of the king
of Judah,
\par }{\Q “This place looks like a veritable
forest of Gilead
to me.
\par }{\Q It is like the wooded heights of Lebanon
in my eyes.
\par }{\Q But I swear that I will make
it like a wilderness
\par }{\Q whose towns
have all been deserted.
\par }{\Q \VS{7}I will send
men
against
it to destroy
it

\par }{\Q with
their axes and hatchets.
\par }{\Q They will hack up its fine
cedar
panels and columns
\par }{\Q and throw them into
the fire.
\par }{\PP \VS{8}“‘People
from other
nations
will pass
by this
city.
They will ask
one
another,
“Why
has
the {\ND{Lord}}
done
such
a thing to this
great
city?”
\VS{9}The answer
will come back,
“It is because they broke
their covenant
with the
{\ND{Lord}}
their God
and worshiped
and served
other
gods.”
\par }{\SH Judgment on Jehoahaz
\par }{\Q \VS{10}“‘Do not
weep
for the king who was killed.
\par }{\Q Do not
grieve
for him.
\par }{\Q But weep mournfully
for the king who has gone
into exile.
\par }{\Q For
he will never
return
to see
his native
land
again.
\par }{\PP \VS{11}“‘For
the {\ND{Lord}}
has spoken
about
Shallum
son
of Josiah,
who succeeded
his father
as
king
of Judah
but was carried off
into exile. He has said, “He will never
return
to this land.
\VS{12}For
he will die
in the country
where
they took him as
a captive.
He will never
see
this
land
again.”
\par }{\SH Judgment on Jehoiakim
\par }{\Q \VS{13}“‘Sure
to be judged
is the king who
builds
his palace
using injustice
\par }{\Q and treats
people unfairly
while
adding
its upper
rooms.

\par }{\Q He makes his
countrymen work
for him for nothing.
\par }{\Q He does not
pay them for their labor.
\par }{\Q \VS{14}He says,
“I will build
myself a large palace
\par }{\Q with spacious
upper
rooms.”
\par }{\Q He cuts
windows
in its walls,
\par }{\Q panels it with
cedar,
and paints
its rooms red.
\par }{\Q \VS{15}Does it make you any more of a king
\par }{\Q that
you
outstrip
everyone else in building with cedar?
\par }{\Q Just think about your father.
\par }{\Q He was content that he had food
and drink.
\par }{\Q He did
what was just and right.
\par }{\Q So
things went well with him.
\par }{\Q \VS{16}He upheld
the cause
of the poor
and needy.
\par }{\Q So things went well for Judah.’

\par }{\Q The
{\ND{Lord}}
says,
\par }{\Q ‘That
is a good
example of what it
means to know me.’
\par }{\Q \VS{17}But
you are always thinking and looking
\par }{\Q for
ways to increase your wealth by dishonest means.
\par }{\Q Your eyes
and your heart
are set
\par }{\Q on
killing
some innocent
person
\par }{\Q and committing fraud
and oppression.
\par }{\Q \VS{18}So
the {\ND{Lord}}
has this
to say
about Josiah’s
son,
King
Jehoiakim
of Judah:
\par }{\Q People will not
mourn
for him, saying,
\par }{\Q “This makes me sad,
my brother!
\par }{\Q This makes me sad,
my sister!”
\par }{\Q They will not
mourn
for him, saying,
\par }{\Q “Poor, poor lord! Poor, poor majesty!”
\par }{\Q \VS{19}He will be left unburied
just like a dead donkey.
\par }{\Q His body will be dragged off
and thrown
outside
the gates
of Jerusalem.’ ”
\par }{\SH Warning to Jerusalem
\par }{\Q \VS{20}People of Jerusalem, go up
to Lebanon
and cry out
in mourning.
\par }{\Q Go to the land of Bashan
and cry out loudly.
\par }{\Q Cry out
in mourning from the mountains of Moab.
\par }{\Q For
your allies
have all
been defeated.
\par }{\Q \VS{21}While you were feeling
secure I gave you warning.

\par }{\Q But
you said,
“I refuse
to
listen
to you.”
\par }{\Q That is
the way
you have acted from your earliest
history
onward.

\par }{\Q Indeed,
you have never
paid attention
to me.
\par }{\Q \VS{22}My judgment will carry off all
your leaders
like a storm
wind!

\par }{\Q Your allies
will go
into captivity.
\par }{\Q Then
you will certainly
be disgraced
and put to shame
\par }{\Q because of all
the wickedness you have done.
\par }{\Q \VS{23}You may feel as secure
as a bird
\par }{\Q nesting
in
the cedars
of Lebanon.
\par }{\Q But oh how you
will groan when the pains
of judgment come
on
you.
\par }{\Q They will be like those of a woman giving birth to a baby.
\par }{\SH Jeconiah Will Be Permanently Exiled
\par }{\PP \VS{24}The
{\ND{Lord}}
says,
\par }{\PP “As surely
as I am
the living God, you, Jeconiah,
king
of Judah,
son
of Jehoiakim,
will not
be the earthly
representative
of my authority.
Indeed,
I will take that
right
away from you.
\VS{25}I will
hand
you over to those who want
to take your life
and of whom
you
are afraid.
I will hand
you over to King
Nebuchadnezzar
of Babylon
and his
Babylonian soldiers.
\VS{26}I will force
you and your mother
who
gave you birth
into exile.
You will be exiled
to a country
where
neither
of you were born,
and you will both die
there.
\VS{27}You will never
come back
to this
land
to which
you will long
to return!”
\par }{\Q \VS{28}This
man,
Jeconiah,
will be like a broken pot someone
threw
away.
\par }{\Q He will be like a clay vessel
that no
one wants.
\par }{\Q Why
will
he
and his children
be forced into exile?
\par }{\Q Why will
they be thrown
out into a country
they know
nothing about?
\par }{\Q \VS{29}O land
of Judah, land
of Judah, land
of Judah!

\par }{\Q Listen
to what
the
{\ND{Lord}} has to say!
\par }{\PP \VS{30}The
{\ND{Lord}}
says,
\par }{\Q “Enroll
this
man
in the register
as though he were childless.
\par }{\Q Enroll him as a man
who will not
enjoy success
during his lifetime.
\par }{\Q For
none of his sons will succeed
in occupying
the throne
of David
\par }{\Q or ever
succeed
in
ruling
over Judah.”

\par }\Chap{23}{\PP \VerseOne{1}The
{\ND{Lord}}
says, “The leaders of my people are sure to be judged.
They were supposed to watch over my people like shepherds
watch over their sheep.
But they are causing my people to be destroyed
and scattered.
\VS{2}So
the

{\ND{Lord}}
God
of Israel
has this
to say about the leaders
who are ruling over
his people: “You
have caused my people to be dispersed
and driven
into exile.
You have not
taken care
of them. So
I
will punish
you for the
evil
that you have done.
I, the
{\ND{Lord}}, affirm it!
\VS{3}Then I myself
will regather
those of my people who are still alive from all
the countries
where
I have driven
them. I will bring them back
to their homeland.
They will greatly increase
in number.
\VS{4}I
will install
rulers
over them who will care
for them.
Then they will no
longer
need to fear
or
be terrified.
None
of them will turn up
missing.
I, the
{\ND{Lord}}, promise it!
\par }{\Q \VS{5}“I,
the {\ND{Lord}}, promise
that a new time
will certainly
come
\par }{\Q when I will raise
up for them a righteous
branch,
a descendant of David.
\par }{\Q He will rule
over them with wisdom and understanding
\par }{\Q and will do
what is just
and right
in the land.
\par }{\Q \VS{6}Under his rule Judah
will enjoy
safety
\par }{\Q and Israel
will live
in security.
\par }{\Q This
is the name
he will go by:
\par }{\Q ‘The
{\ND{Lord}}
has provided
us with justice.’
\par }{\PP \VS{7}“So
I, the
{\ND{Lord}}, say: ‘A new time
will certainly
come.
People now affirm
their oaths
with “I swear as surely
as the
{\ND{Lord}}
lives who
delivered
the
people of Israel
out of Egypt.”
\VS{8}But at that
time they will affirm them with “I swear as surely
as the
{\ND{Lord}}
lives
who delivered
the descendants
of the former nation
of Israel
from the land
of the north
and from all
the other lands
where he had
banished
them.” At that time they will live
in
their own land.’ ”
\par }{\SH Oracles Against the False Prophets
\par }{\Q \VS{9}Here is what the
{\ND{Lord}} says concerning the false prophets:

\par }{\Q My heart
and my mind
are deeply
disturbed.
\par }{\Q I tremble
all over.
\par }{\Q I am
like a drunk
person,
\par }{\Q like a person
who has had too much
wine,
\par }{\Q because
of the way the
{\ND{Lord}}
\par }{\Q and his holy
word are being mistreated.
\par }{\Q \VS{10}For
the land
is full
of people unfaithful
to him.

\par }{\Q They live wicked
lives
and they misuse their power.
\par }{\Q So the land
is dried up
because
it is under
his curse.
\par }{\Q The pastures
in the wilderness
are withered.
\par }{\Q \VS{11}Moreover,
the {\ND{Lord}}
says,
\par }{\Q “Both
the prophets
and priests
are godless.
\par }{\Q I have even
found
them doing evil
in my temple!
\par }{\Q \VS{12}So
the paths
they
follow
will be dark
and slippery.
\par }{\Q They will stumble and fall
headlong.
\par }{\Q For
I will bring
disaster
on
them.
\par }{\Q A day
of reckoning
is coming for them.”

\par }{\Q The
{\ND{Lord}}
affirms it!
\par }{\Q \VS{13}The
{\ND{Lord}} says, “I saw
the prophets
of Samaria
\par }{\Q doing something that was disgusting.
\par }{\Q They prophesied
in the name of the god Baal
\par }{\Q and led my people
Israel
astray.
\par }{\Q \VS{14}But I see
the prophets
of Jerusalem
\par }{\Q doing something
just as shocking.
\par }{\Q They are unfaithful
to me
\par }{\Q and continually prophesy
lies.
\par }{\Q So they give encouragement
to people who are doing evil,
\par }{\Q with the result that they do not
stop their evildoing.
\par }{\Q I consider all
of them as bad as the people of Sodom,
\par }{\Q and the citizens
of Jerusalem as bad as the people of Gomorrah.
\par }{\Q \VS{15}So
then I,
the {\ND{Lord}}
who rules over all,
\par }{\Q have something to say concerning
the prophets
of Jerusalem:

\par }{\Q ‘I will make
these prophets eat
the
bitter food of suffering
\par }{\Q and drink
the poison
water
of judgment.

\par }{\Q For
the prophets
of Jerusalem
are the reason

\par }{\Q that ungodliness
has spread throughout
the land.’ ”
\par }{\Q \VS{16}The
{\ND{Lord}}
who rules over all
says
to the people of Jerusalem:

\par }{\Q “Do not
listen
to
what
\par }{\Q those prophets
are saying
to you.
\par }{\Q They
are filling
you with false hopes.
\par }{\Q They are reporting
visions
of their own imaginations,
\par }{\Q not
something the
{\ND{Lord}}
has given them to say.
\par }{\Q \VS{17}They continually say
to those who reject
what the
{\ND{Lord}}
has said,
\par }{\Q ‘Things will go
well
for you!’

\par }{\Q They say
to all
those who follow
the stubborn inclinations
of their own hearts,
\par }{\Q ‘Nothing
bad
will happen to you!’
\par }{\Q \VS{18}Yet
which of them has ever
stood
in the
{\ND{Lord}}’s
inner circle
\par }{\Q so they could see
and hear
what
he has to say?

\par }{\Q Which of them have ever
paid attention
or listened
to what
he has said?
\par }{\Q \VS{19}But
just watch! The wrath
of the {\ND{Lord}}
\par }{\Q will come like a storm!

\par }{\Q Like a raging
storm
it will rage
down
\par }{\Q on
the heads
of those
who are
wicked.
\par }{\Q \VS{20}The anger
of the {\ND{Lord}}
will not
turn back
\par }{\Q until
he has fully
carried out
his intended
purposes.
\par }{\Q In days
to come
\par }{\Q you people will come to understand
this clearly.
\par }{\Q \VS{21}I did not
send
those prophets.
\par }{\Q Yet they
were in a hurry
to give their message.

\par }{\Q I did not
tell
them anything.
\par }{\Q Yet they
prophesied anyway.
\par }{\Q \VS{22}But if
they had stood
in my inner circle,
\par }{\Q they would have proclaimed
my message
to my people.
\par }{\Q They would have caused my people
to turn
from their wicked ways
\par }{\Q and stop doing the evil
things they are doing.
\par }{\Q \VS{23}Do you people think
that I
am
some local deity
\par }{\Q and not the transcendent
God?”
 the
{\ND{Lord}}
asks.
\par }{\Q \VS{24}“Do you really think anyone
can hide
himself
\par }{\Q where I
cannot
see
him?” the
{\ND{Lord}}
asks.
\par }{\Q “Do you not
know that I
am everywhere?”

\par }{\Q the {\ND{Lord}}
asks.
\par }{\PP \VS{25}The
{\ND{Lord}} says, “I have heard
what
those prophets
who are prophesying
lies
in my name
are saying.
They are saying, ‘I
have had a dream! I have had a dream!’
\VS{26}Those prophets
are just prophesying
lies.
They are prophesying
the delusions
of their own minds.
\VS{27}How
long will they go on plotting
to make my
people
forget
who I am through the dreams
they tell
one
another? That is just
as bad as what their ancestors
did when they forgot
who I am by worshiping
the god Baal.
\VS{28}Let the prophet
who has had
a dream
go ahead and tell
his dream.
Let the person who
has
received my message
report
that message
faithfully.
What
is like straw
cannot compare
to what is like grain! I, the
{\ND{Lord}}, affirm it!
\VS{29}My message
is like a fire
that purges
dross! It is like a hammer
that breaks a rock
in pieces! I, the
{\ND{Lord}}, so affirm it!
\VS{30}So
I, the
{\ND{Lord}}, affirm
that I am opposed
to those prophets
who steal
messages
from one
another that they claim are from me.
\VS{31}I,
the {\ND{Lord}},
affirm
that I am opposed
to those prophets
who are using
their own tongues
to declare,
‘The
{\ND{Lord}} declares….’
\VS{32}I, the
{\ND{Lord}}, affirm
that I am
opposed
to those prophets who
dream
up lies
and report
them. They are misleading my
people
with their reckless
lies.
I
did not send
them. I did not commission
them.
They are not
helping
these
people
at all. I, the
{\ND{Lord}}, affirm it!”
\par }{\PP \VS{33}The
{\ND{Lord}}
said
to me,
“Jeremiah,
when
one of these
people,
or
a prophet,
or
a priest
asks
you, ‘What
burdensome
message do you have from the
{\ND{Lord}}?’ Tell them, ‘You are the burden,
and I
will cast
you away. I,
the

{\ND{Lord}}, affirm it!
\VS{34}I will punish any prophet,
priest,
or other
person who
says
“The
{\ND{Lord}}’s
message is burdensome.”
I will punish
both that person
and his whole family.’ ”
\par }{\PP \VS{35}So I, Jeremiah, tell
you, “Each
of you people should say to his friend
or
his relative,
‘How
did the
{\ND{Lord}}
answer? Or what
did the
{\ND{Lord}}
say?’
\VS{36}You must no longer
say that the
{\ND{Lord}}’s
message
is burdensome.
For
what is ‘burdensome’
really pertains to what
a person
himself says.
You are misrepresenting
the
words
of our God,
the living
God,
the {\ND{Lord}}
who rules over all.
\VS{37}Each
of you should merely
ask the prophet,
‘What
answer
did the
{\ND{Lord}}
give you? Or what
did the
{\ND{Lord}}
say?’
\VS{38}But
just suppose you continue
to say, ‘The message
of the {\ND{Lord}}
is burdensome.’ Here
is what
the {\ND{Lord}}
says
will happen: ‘I sent
word to
you that
you must not
say,
“The
{\ND{Lord}}’s
message is burdensome.”
But you used the words
“The
{\ND{Lord}}’s
message
is burdensome” anyway.
\VS{39}So
I
will carry you far off and throw
you
away. I
will send both you and the
city
I gave
to you and to your ancestors
out
of my sight.
\VS{40}I will bring
on
you lasting
shame
and lasting
disgrace which
will never
be forgotten!’ ”

\par }\Chap{24}{\PP \VerseOne{1}The
{\ND{Lord}}
showed
me two
baskets
of figs
sitting
before
his temple.
This happened after
King
Nebuchadnezzar
of Babylon
deported
Jehoiakim’s
son,
King
Jeconiah
of Judah.
He deported him and the
leaders
of Judah,
along with
the craftsmen
and metal workers,
and took
them to Babylon.
\VS{2}One
basket
had very
good-looking
figs
in it. They
looked like those that had ripened early.
The other
basket
had very
bad-looking
figs
in it, so bad
they could not
be eaten.
\VS{3}The
{\ND{Lord}}
said
to
me, “What
do you
see,
Jeremiah?” I answered,
“I see figs.
The good
ones
look very
good.
But the bad
ones look very
bad,
so bad
that they cannot
be eaten.”
\par }{\PP \VS{4}The
{\ND{Lord}}
said
to me,
\VS{5}“I,
the {\ND{Lord}},
the God
of Israel,
say: ‘The
exiles
whom
I sent away
from
here
to the land
of Babylon
are like
those
good
figs.
I consider
them to be good.
\VS{6}I will look
after their welfare
and will restore
them to this
land.
There I will build
them up and will not
tear
them down. I will plant
them firmly in the land and will not
uproot them.
\VS{7}I will give
them the desire
to acknowledge
that
I am
the {\ND{Lord}}. I
will be
their God
and they will be
my people.
For
they will wholeheartedly
return
to me.’
\par }{\PP \VS{8}“I,
the {\ND{Lord}}, also solemnly assert: ‘King
Zedekiah
of Judah,
his officials,
and the
people who remain
in Jerusalem
or who have gone to live
in Egypt
are like those bad figs. I consider them to be just
like those bad figs
that
are so bad
they cannot be eaten.
\VS{9}I will
bring such disaster
on them that all
the kingdoms
of the earth
will be horrified.
I will make them an object of reproach,
a proverbial example of disaster.
I
will make them an object of ridicule,
an example
to be used in curses. That
is how
they will be remembered wherever
I banish them.
\VS{10}I
will bring war,
starvation,
and disease
on them until
they are completely destroyed from the land
I gave
them
and their ancestors.’ ”

\par }\Chap{25}{\PP \VerseOne{1}In the fourth
year
that Jehoiakim
son
of Josiah
was king
of Judah,
the
{\ND{Lord}} spoke to Jeremiah
concerning
all
the people
of Judah.
(That was the same as the first
year
that Nebuchadnezzar
was king
of Babylon.)
\VS{2}So
the prophet
Jeremiah
spoke
to all
the people
of Judah
and to
all
the people who were living
in Jerusalem.
\VS{3}“For the last twenty-three
years,
from
the thirteenth
year
that Josiah
son
of Amon
was ruling
in Judah
until
now,
the {\ND{Lord}}
has been
speaking
to
me. I told you over and over again
what he said.
But you would not
listen.
\VS{4}Over and over again
the {\ND{Lord}}
has sent
his servants
the prophets
to you. But you have
not
listened
or paid attention.
\VS{5}He said
through them, ‘Each
of you must turn
from your
wicked ways
and stop doing the evil
things you are doing. If you do, I will allow you to continue to live
here in the land
that
I gave
to you and your ancestors
as
a lasting possession.
\VS{6}Do not
pay allegiance
to other
gods
and worship
and serve
them. Do not
make me angry
by the things
that you
do.
Then I will not
cause you any harm.’
\VS{7}So,
now the
{\ND{Lord}}
says, ‘You have not
listened
to me.
But you have made me angry
by the things that you have done.
Thus you have brought harm on yourselves.’
\par }{\PP \VS{8}“Therefore,
the

{\ND{Lord}}
who rules over all
says, ‘You have not
listened
to what
I said.
\VS{9}So I, the
{\ND{Lord}}, affirm
that I will send
for all
the peoples
of the north
and my servant,
King
Nebuchadnezzar
of Babylon.
I will bring
them against
this
land
and its inhabitants
and all
the nations
that surround it. I will utterly destroy
this land, its inhabitants, and all the nations that surround
it and make
them everlasting
ruins.
I will make them objects of horror
and hissing
scorn.
\VS{10}I will put an end
to the sounds
of joy
and gladness,
to the glad celebration
of brides
and grooms
in these lands. I will put an end to the sound
of people grinding meal.
I will put an end to lamps
shining
in their houses.
\VS{11}This
whole
area
will become a desolate wasteland.
These
nations
will be
subject
to the king
of Babylon
for seventy
years.’
\par }{\PP \VS{12}“‘But when
the seventy
years
are over,
I will punish
the king
of Babylon
and his nation
for their sins.
I will make
the land
of Babylon
an everlasting
ruin.
I, the
{\ND{Lord}}, affirm it!
\VS{13}I will bring
on
that land
everything
that
I said
I would. I will bring on
it everything
that is written
in this
book.
I will bring on it everything
that Jeremiah
has prophesied
against
all
the nations.
\VS{14}For
many
nations
and great
kings
will make slaves
of the king of Babylon and his nation too.
I will repay
them for all they have done!’ ”
\par }{\SH Judah and the Nations Will Experience God’s Wrath
\par }{\PP \VS{15}So the

{\ND{Lord}},
the God
of Israel,
spoke to
me in a vision. “Take
this cup
from my hand.
It is filled with the wine
of my wrath.
Take it and make the
nations
to
whom
I
send
you drink it.
\VS{16}When they have drunk
it, they will stagger to and fro
and act insane.
For I
will send
wars
sweeping through them.”
\par }{\PP \VS{17}So I took
the
cup
from the
{\ND{Lord}}’s
hand.
I made all
the nations
to whom
he sent
me
drink the wine of his wrath.
\VS{18}I made Jerusalem
and the
cities
of Judah,
its
kings
and its officials
drink it. I did
it so Judah would become a ruin.
I did it so Judah, its kings, and its officials would become an object of horror
and of hissing scorn,
an example used in curses.
Such is already becoming the case!
\VS{19}I made all
of these other people
drink it: Pharaoh,
king
of Egypt;
his attendants,
his officials, his people,
\VS{20}the
foreigners living in Egypt;
all
the kings
of the land
of Uz;
all
the kings
of the land
of the Philistines,
the
people of Ashkelon,
Gaza,
Ekron,
the
people who had been left alive
from Ashdod;
\VS{21}all the people of Edom,
Moab,
Ammon;
\VS{22}all
the kings
of Tyre,
all the kings
of the coastlands
along the sea;
\VS{23}the people of Dedan,
Tema,
Buz,
all
the desert people who cut their hair short
at the temples;
\VS{24}all
the kings
of Arabia
who
live
in the desert;
\VS{25}all
the kings
of Zimri;
all
the kings
of Elam;
all
the kings
of Media;
\VS{26}all
the kings
of the north,
whether near
or far
from one
another;
and all
the other kingdoms
which
are on
the face
of the earth.
After
all of them have drunk the wine of the
{\ND{Lord}}’s wrath,
the king
of Babylon
must drink it.
\par }{\PP \VS{27}Then the
{\ND{Lord}} said
to me, “Tell
them that the
{\ND{Lord}}
God
of Israel
who rules over all
says, ‘Drink
this cup until you get drunk
and vomit.
Drink until you fall
down and can’t
get
up. For I
will send
wars
sweeping through you.’
\VS{28}If
they refuse
to take
the cup
from your hand
and drink
it, tell
them
that the

{\ND{Lord}}
who rules over all
says ‘You most certainly must
drink it!
\VS{29}For
take note,
I am
already beginning
to bring disaster on the city
that
I call
my own.
So
how can you
possibly avoid
being punished? You
will not
go unpunished! For
I am
proclaiming
war
against
all
who live
on the earth.
I, the
{\ND{Lord}}
who rules over all,
affirm it!’
\par }{\PP \VS{30}“Then, Jeremiah, make the following prophecy
against
them:
\par }{\Q ‘Like a lion about to
attack, the
{\ND{Lord}}
will roar
from the heights
of heaven;
\par }{\Q from his holy
dwelling
on high he will roar loudly.
\par }{\Q He will roar mightily
against
his land.
\par }{\Q He will shout in triumph
like those stomping
juice from the grapes

\par }{\Q against all
those who live
on the earth.
\par }{\Q \VS{31}The sounds of battle
will resound
to
the ends
of the earth.
\par }{\Q For
the {\ND{Lord}}
will bring
charges against the nations.
\par }{\Q He will pass judgment
on all
humankind
\par }{\Q and will hand
the wicked
over
to be killed in war.’
\par }{\Q The
{\ND{Lord}}
so affirms it!
\par }{\Q \VS{32}The

{\ND{Lord}}
who rules over all
says,
\par }{\Q ‘Disaster
will soon come
on one nation
after another.
\par }{\Q A mighty storm
of military
destruction is rising up
\par }{\Q from the distant parts
of the earth.’
\par }{\Q \VS{33}Those who have been
killed
by the
{\ND{Lord}}
at that time
\par }{\Q will be scattered from one end
of the earth
to
the other.
\par }{\Q They will not
be mourned
over, gathered up,
or
buried.
\par }{\Q Their dead bodies will lie scattered over
the ground
like manure.
\par }{\Q \VS{34}Wail
and cry out
in anguish, you rulers!
\par }{\Q Roll
in the dust, you who shepherd
flocks
of people!

\par }{\Q The time
for
you to be slaughtered
has come.
\par }{\Q You will lie scattered
and fallen
like broken pieces of fine
pottery.
\par }{\Q \VS{35}The leaders
will not
be able to run away
and hide.

\par }{\Q The shepherds
of the flocks
will not be able to escape.
\par }{\Q \VS{36}Listen to the cries
of anguish
of the leaders.
\par }{\Q Listen to the wails
of the shepherds
of the flocks.
\par }{\Q They are wailing because
the {\ND{Lord}}
\par }{\Q is about to destroy
their lands.
\par }{\Q \VS{37}Their peaceful dwelling places
will be
laid
waste

\par }{\Q by the fierce
anger
of the {\ND{Lord}}.
\par }{\Q \VS{38}The
{\ND{Lord}} is like a lion
who has left
his lair.
\par }{\Q So their lands
will certainly
be
laid waste
\par }{\Q by the warfare
of the oppressive
nation

\par }{\Q and by the fierce
anger of the
{\ND{Lord}}.”

\par }\Chap{26}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to Jeremiah at the beginning
of the reign
of Josiah’s
son,
King
Jehoiakim
of Judah.
\VS{2}The

{\ND{Lord}}
said,
“Go stand
in the courtyard
of the
{\ND{Lord}}’s
temple.
Speak
out
to all
the people who are coming
from the towns
of Judah
to worship
in the
{\ND{Lord}}’s
temple.
Tell them
everything
I command
you to tell
them.
Do not
leave out a single
word!
\VS{3}Maybe
they will pay attention
and each
of them
will stop
living
the evil
way
they do. If they do that,
then I will forgo
destroying
them as I
had intended
to do
because
of the wicked
things they have been doing.
\VS{4}Tell
them
that the
{\ND{Lord}}
says, ‘You must
obey
me! You must live according to
the way
I have instructed you in my laws.
\VS{5}You must
pay attention
to the exhortations
of my servants
the prophets.
I
have sent
them to you over and over again.
But you have not
paid any attention to them.
\VS{6}If you
do not obey me, then I will
do to this
temple
what
I did to Shiloh.
And I will make
this
city
an example
to be used in curses by people from all
the nations
on the earth.’ ”
\par }{\PP \VS{7}The priests,
the prophets,
and all
the people
heard
Jeremiah
say
these
things
in the
{\ND{Lord}}’s
temple.
\VS{8}Jeremiah
had just barely finished
saying
all
the {\ND{Lord}}
had commanded
him to say
to
all
the people.
All
at once some of the
priests,
the prophets,
and the people
grabbed
him and shouted,
“You deserve to die!
\VS{9}How dare
you claim
the
{\ND{Lord}}’s
authority to prophesy such things! How dare you claim
his authority to prophesy
that this
temple
will become
like Shiloh
and that this
city
will become an uninhabited
ruin!” Then all
the people
crowded
around Jeremiah.
\par }{\PP \VS{10}However, some of the officials
of Judah
heard
about what
was happening
and they rushed up
to the
{\ND{Lord}}’s
temple
from the royal
palace.
They set up court
at the entrance
of the New
Gate
of the
{\ND{Lord}}’s temple.
\VS{11}Then the priests
and the prophets
made their charges before
the officials
and all
the people.
They said, “This
man
should be condemned
to die
because
he prophesied
against this
city.
You have
heard
him do so with your own ears.”
\par }{\PP \VS{12}Then Jeremiah
made his defense before
all
the officials
and all
the people. “The
{\ND{Lord}}
sent
me to prophesy
everything
you have
heard
me say
against this
temple
and against this
city.
\VS{13}But
correct
the way
you have been living
and do what is right. Obey
the {\ND{Lord}}
your God.
If you do, the
{\ND{Lord}}
will forgo
destroying
you as
he threatened he would.
\VS{14}As
to my case, I
am
in your power.
Do
to me what
you deem
fair
and proper.
\VS{15}But
you
should take careful
note
of this: If
you
put me to death,
you
will bring on
yourselves
and this city
and those who live
in it the guilt
of murdering
an innocent
man. For
the {\ND{Lord}}
has sent
me to speak
all
this
where you can hear it. That
is the truth!”
\par }{\PP \VS{16}Then the officials
and all
the people
rendered
their verdict to
the priests
and the prophets.
They said, “This
man
should not
be condemned
to die.
For
he has spoken
to
us under the authority
of the
{\ND{Lord}}
our God.”
\VS{17}Then
some
of the elders
of Judah stepped forward and spoke
to
all
the people
gathered
there. They said,
\VS{18}“Micah
from Moresheth
prophesied
during
the time
Hezekiah
was king
of Judah.
He told
all
the people
of Judah,
\par }{\Q ‘The

{\ND{Lord}}
who rules over all
says,
\par }{\Q “Zion
will become a plowed
field.
\par }{\Q Jerusalem
will become a pile of rubble.
\par }{\Q The temple
mount
will become
a mere wooded
ridge.” ’
\par }{\PP \VS{19}King
Hezekiah
and all
the people of Judah
did not put
him to death,
did they? Did not
Hezekiah show reverence
for the
{\ND{Lord}}
and seek
the
{\ND{Lord}}’s
favor? Did not the
{\ND{Lord}}
forgo
destroying
them as
he threatened
he would? But we
are on
the verge of bringing
great
disaster
on
ourselves.”
\par }{\PP \VS{20}Now
there was
another
man
who prophesied
as the
{\ND{Lord}}’s
representative
against
this city
and this land
just
as Jeremiah
did. His name
was Uriah
son
of Shemaiah
from Kiriath Jearim.
\VS{21}When the king
and all
his bodyguards
and officials
heard
what
he was prophesying, the king
sought
to have him executed.
But Uriah
found out about
it and fled
to Egypt
out of fear.
\VS{22}However, King
Jehoiakim
sent
some men
to Egypt,
including Elnathan
son
of Achbor,
\VS{23}and they brought
Uriah
back
from there.
They took
him to
King
Jehoiakim,
who had him executed
and had his body
thrown
into
the burial place
of the common people.
\par }{\PP \VS{24}However,
Ahikam
son
of Shaphan
used his influence
to keep Jeremiah
from being
handed
over and executed
by the people.


\par }\Chap{27}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to Jeremiah
early
in the reign
of Josiah’s
son,
King
Zedekiah of Judah.
\VS{2}The

{\ND{Lord}}
told
me, “Make
a yoke out of leather straps
and wooden crossbars
and put
it on
your neck.
\VS{3}Use it to send
messages to
the kings
of Edom,
Moab,
Ammon,
Tyre,
to
King
Zedekiah
of Judah.
\VS{4}Charge
them to
give their masters
a message
from me. Tell
them, ‘The
{\ND{Lord}}
God
of Israel
who rules over all
says
to
give your masters this message.
\VS{5}“I
made
the earth
and the people
and animals
on
it by my mighty power
and great
strength,
and I give
it to whomever
I see fit.
\VS{6}I
have
at this time placed
all
these
nations
of yours
under the power
of my servant,
King
Nebuchadnezzar
of Babylon.
I have even
made all the wild
animals
subject to him.
\VS{7}All
nations
must serve
him and his son
and grandson
until
the time
comes
for his own nation
to fall. Then
many
nations and great
kings
will in turn subjugate Babylon.
\VS{8}But suppose
a nation
or a kingdom
will not
be subject
to King
Nebuchadnezzar
of Babylon.
Suppose it will
not
submit
to the yoke of servitude
to him. I, the
{\ND{Lord}}, affirm
that I will punish
that nation.
I will use the king
of Babylon
to punish it with war,
starvation,
and disease
until I have destroyed it.
\VS{9}So
do not
listen
to
your prophets
or to
those who claim to
predict the future by divination,
by
dreams,
by
consulting
the dead, or by practicing magic.
They
keep telling
you, ‘You do not
need to
be subject
to the king
of Babylon.’
\VS{10}Do not listen to them,
because
their
prophecies
are lies.
Listening to them will only cause you to
be taken far
away from your native land.
I will drive
you out
of your country and you will die in exile.
\VS{11}Things will go better for the nation
that
submits
to the yoke
of servitude
to the king
of Babylon
and is subject
to him. I will leave
that nation in
its native land.
Its people can continue to farm
it and live
in it. I, the
{\ND{Lord}}, affirm it!” ’ ”
\par }{\PP \VS{12}I told
King
Zedekiah
of Judah
the same thing.
I said, “Submit
to the yoke
of servitude to the king
of Babylon. Be subject
to him and his people.
Then you will continue
to live.
\VS{13}There is no reason why
you
and your people
should die
in war
or from starvation
or disease! That’s
what
the {\ND{Lord}}
says
will happen to
any nation
that
will not
be subject
to the
king
of Babylon.
\VS{14}Do not
listen
to
the prophets
who are telling
you that you do not
need to serve
the king
of Babylon.
For
they
are prophesying
lies to you.
\VS{15}For
I, the
{\ND{Lord}}, affirm
that I did not
send
them. They
are prophesying
lies
to you. If you listen to them, I will drive
you and the prophets
who are prophesying
lies out of the land and you
will all die in exile.”
\par }{\PP \VS{16}I also told
the priests
and all
the people,
“The
{\ND{Lord}}
says,
‘Do not
listen
to
what
your prophets
are saying.
They are prophesying
to you that the valuable
articles
taken from the
{\ND{Lord}}’s
temple
will be brought back
from Babylon
very
soon.
But
they
are prophesying
a lie to you.
\VS{17}Do not
listen
to them.
Be subject
to the king
of Babylon.
Then you will continue to live.
Why
should this
city
be made a pile of rubble?’ ”
\VS{18}I also told them, “If
they
are really prophets
and the
{\ND{Lord}}
is
speaking
to them, let
them pray earnestly
to the
{\ND{Lord}}
who rules over all.
Let them plead with him not
to let the valuable articles
that are still left
in the
{\ND{Lord}}’s
temple,
in the royal
palace,
and in Jerusalem
be taken away to Babylon.
\VS{19}For
the

{\ND{Lord}}
who rules over all
has already spoken about the two bronze pillars,
the large bronze basin called ‘The Sea,’
and the movable
bronze stands.
He has already spoken about the rest
of the valuable articles
that are left
in this
city.
\VS{20}He has already spoken about these things that King
Nebuchadnezzar
of Babylon
did not
take away
when he carried
Jehoiakim’s
son
King
Jeconiah
of Judah
and the nobles
of Judah
and Jerusalem away as captives.
\VS{21}Indeed,
the

{\ND{Lord}}
God
of Israel
who rules over all
has already spoken about the valuable articles
that are left
in the
{\ND{Lord}}’s
temple,
in the royal
palace
of Judah,
and in Jerusalem.
\VS{22}He has said, ‘They will be
carried off
to Babylon.
They will remain there
until
it is time
for me to show consideration
for them again. Then I will bring
them back and restore
them to
this
place.’
I, the
{\ND{Lord}}, affirm this!”

\par }\Chap{28}{\PP \VerseOne{1}The following events occurred
in that same
year,
early
in the reign
of King
Zedekiah
of Judah.
To
be more precise, it was the fifth
month
of the fourth
year
of his reign. The prophet
Hananiah
son
of Azzur,
who
was from Gibeon,
spoke to
Jeremiah in the
{\ND{Lord}}’s
temple
in the presence
of the priests
and all
the people.
\VS{2}“The

{\ND{Lord}}
God
of Israel
who rules over all
says,
‘I will break
the yoke of servitude
to the king
of Babylon.
\VS{3}Before
two years
are over,
I
will bring back
to
this
place
everything
that
King
Nebuchadnezzar
of Babylon
took
from it
and carried
away to Babylon.
\VS{4}I
will also bring back
to
this
place
Jehoiakim’s
son
King
Jeconiah
of Judah
and all
the exiles
who were taken
to Babylon.’
Indeed,
the {\ND{Lord}}
affirms, ‘I will break
the yoke
of servitude to the king
of Babylon.’ ”
\par }{\PP \VS{5}Then the prophet
Jeremiah
responded to the prophet
Hananiah
in the presence
of the priests
and all
the people
who were standing
in the
{\ND{Lord}}’s
temple.
\VS{6}The prophet
Jeremiah
said,
“Amen! May the
{\ND{Lord}}
do
all this! May the
{\ND{Lord}}
make
your prophecy
come true! May he bring back
to
this
place
from Babylon
all the valuable articles
taken from the
{\ND{Lord}}’s
temple
and the people
who were carried into exile.
\VS{7}But
listen
to what
I
say
to you and to all
these people.
\VS{8}From
earliest times,
the prophets
who
preceded
you and me
invariably
prophesied
war,
disaster,
and plagues
against
many
countries
and great
kingdoms.
\VS{9}So if
a prophet
prophesied
peace
and prosperity, it
was only known
that
the {\ND{Lord}}
truly sent
him when what
he prophesied
came
true.”
\par }{\PP \VS{10}The prophet
Hananiah
then took
the yoke
off the prophet
Jeremiah’s
neck
and broke it.
\VS{11}Then he spoke
up in the presence
of all
the people.
“The
{\ND{Lord}}
says,
‘In the same
way I will break
the
yoke
of servitude
of all
the nations
to King
Nebuchadnezzar
of Babylon
before
two years
are over.’ ”
After he heard this, the prophet
Jeremiah
departed
and went on his way.
\par }{\PP \VS{12}But shortly
after
the prophet
Hananiah
had broken
the yoke
off
the prophet
Jeremiah’s
neck,
the {\ND{Lord}}
spoke
to
Jeremiah.
\VS{13}“Go
and tell
Hananiah
that the
{\ND{Lord}}
says, ‘You have indeed broken
the wooden
yoke.
But you have
only succeeded in replacing
it with an iron one!
\VS{14}For
the

{\ND{Lord}}
God
of Israel
who rules over all
says, “I have put
an irresistible
yoke
of servitude
on
all
these
nations
so
they will serve
King
Nebuchadnezzar
of Babylon.
And they will indeed serve
him. I have even
given
him control over the
wild
animals.” ’ ”
\VS{15}Then the prophet
Jeremiah
told
the prophet
Hananiah,
“Listen,
Hananiah! The
{\ND{Lord}}
did not
send
you! You are
making these
people
trust
in a lie!
\VS{16}So
the {\ND{Lord}}
says,
‘I
will most assuredly
remove you from the face
of the earth.
You will die
this very year
because
you have
counseled
rebellion against the
{\ND{Lord}}.’ ”
\par }{\PP \VS{17}In the seventh
month
of that very same year
the prophet
Hananiah
died.

\par }\Chap{29}{\PP \VerseOne{1}The prophet
Jeremiah
sent
a letter
to
the exiles
Nebuchadnezzar
had carried
off from Jerusalem
to Babylon. It was addressed to the elders
who were left among the exiles, to
the priests,
to
the prophets,
and to
all
the other people
who
were exiled
in Babylon.
\VS{2}He sent it after
King
Jeconiah,
the queen mother,
the palace officials,
the leaders
of Judah
and Jerusalem,
the craftsmen,
and the metal workers
had been exiled from Jerusalem.
\VS{3}He sent it with Elasah
son
of Shaphan
and Gemariah
son
of Hilkiah.
King
Zedekiah
of Judah
had
sent
these men to Babylon
to
King
Nebuchadnezzar
of Babylon.
The letter said:
\par }{\PP \VS{4}“The

{\ND{Lord}}
God
of Israel
who rules over all
says to all
those he sent into exile
to Babylon
from Jerusalem,
\VS{5}‘Build
houses
and settle
down. Plant
gardens
and eat
what they produce.
\VS{6}Marry
and have sons
and daughters.
Find
wives
for your sons
and allow
your daughters
get married so
that they too can have sons
and daughters.
Grow in number;
do not
dwindle away.
\VS{7}Work to see that
the city
where
I sent
you as
exiles enjoys
peace
and prosperity.
Pray
to
the {\ND{Lord}}
for it. For
as it prospers
you will prosper.’
\par }{\PP \VS{8}“For
the

{\ND{Lord}}
God
of Israel
who rules over all
says, ‘Do not
let the prophets
or those
among
you who claim
to be able
to predict
the future
by divination
deceive
you. And do not
pay any attention
to
the dreams
that
you
are encouraging them to dream.
\VS{9}They
are prophesying
lies
to you and claiming my authority
to do so. But I did not
send
them. I, the
{\ND{Lord}}, affirm it!’
\par }{\PP \VS{10}“For
the {\ND{Lord}}
says,
‘Only when
the seventy
years
of Babylonian
rule are over will I again take up consideration
for you. Then
I will fulfill
my gracious
promise
to you and restore
you to
your homeland.
\VS{11}For
I
know
what I have planned
for you,’ says
the {\ND{Lord}}. ‘I
have plans
to prosper
you, not
to harm
you. I have plans to give
you a future
filled with hope.
\VS{12}When you call
out to me and come
to me
in prayer,
I will hear your prayers.
\VS{13}When
you seek
me in prayer and worship, you will find
me available
to you. If
you seek
me with all
your heart and soul,
\VS{14}I
will make myself available
to you,’ says
the {\ND{Lord}}. ‘Then
I will reverse
your plight
and will regather
you from all
the nations
and all
the places
where
I have exiled
you,’ says
the {\ND{Lord}}. ‘I will bring you back
to
the place
from which
I exiled
you.’
\par }{\PP \VS{15}“You say,
‘The
{\ND{Lord}}
has raised
up prophets
of good news for us here in Babylon.’
\VS{16}But just listen to what
the {\ND{Lord}}
has to
say
about the king
who occupies
David’s
throne
and all
your fellow countrymen
who are still living
in this
city
of Jerusalem and were not
carried off
into exile
with you.
\VS{17}The

{\ND{Lord}}
who rules over all
says,
‘I will bring
war,
starvation,
and disease
on them. I will treat
them like figs
that
are so rotten
they cannot
be eaten.
\VS{18}I will chase
after
them with war,
starvation,
and disease.
I will make
all
the kingdoms
of the earth
horrified
at what happens to them. I will make them examples of those who are cursed,
objects of horror,
hissing
scorn,
and ridicule
among all
the nations
where
I exile them.
\VS{19}For
they have
not
paid attention
to
what
I said to them through my servants
the prophets
whom
I sent
to them over and over again,’
says
the {\ND{Lord}}. ‘And you exiles have not
paid any attention
to them either,’ says
the {\ND{Lord}}.
\VS{20}‘So pay attention
to what
I, the
{\ND{Lord}}, have said, all
you exiles
whom
I have sent
to Babylon
from Jerusalem.’
\par }{\PP \VS{21}“The

{\ND{Lord}}
God
of Israel
who rules over all
also has something to
say about Ahab
son
of Kolaiah
and Zedekiah
son
of Maaseiah,
who are prophesying
lies
to
you and claiming
my authority to
do so. ‘I
will
hand
them over to King
Nebuchadnezzar
of Babylon
and he will execute them
before your very eyes.
\VS{22}And all
the exiles
of Judah
who
are in Babylon
will use
them
as
examples
when they
put
a curse
on anyone. They will say,
“May the
{\ND{Lord}}
treat you like Zedekiah
and Ahab
whom
the king
of Babylon
roasted
to death in the fire!”
\VS{23}This will happen to them because
they have
done
what is shameful
in Israel.
They have committed adultery
with
their neighbors’
wives
and have spoken
lies
while claiming
my authority. They have spoken words
that
I did not
command
them to speak. I
know what
they have
done. I
have been
a witness
to it,’ says
the {\ND{Lord}}.”
\par }{\SH A Response to the Letter and a Subsequent Letter
\par }{\PP \VS{24}The
{\ND{Lord}} told Jeremiah, “Tell Shemaiah
the Nehelamite
\VS{25}that
the

{\ND{Lord}}
God
of Israel
who rules over all
has
a message for him. Tell
him, ‘On your own initiative
you
sent
a letter
to
the priest
Zephaniah
son
of Maaseiah
and to
all
the other priests
and to
all
the people
in Jerusalem.
In your letter you said to Zephaniah,
\VS{26}“The
{\ND{Lord}}
has made
you priest
in place
of Jehoiada.
He has put you in charge
in the
{\ND{Lord}}’s
temple
of controlling
any lunatic who pretends to
be a prophet.
And it is your duty to
put
any such person in the stocks
with an iron collar around his neck.
\VS{27}You should have
reprimanded Jeremiah
from Anathoth
who is pretending to be a prophet among you!
\VS{28}For
he has even
sent
a message to
us here in Babylon.
He wrote and told
us, “You will be there a long
time. Build
houses
and settle down.
Plant
gardens
and eat
what they produce.” ’ ”
\par }{\PP \VS{29}Zephaniah
the priest
read
that
letter
to the prophet
Jeremiah.
\VS{30}Then the
{\ND{Lord}}
spoke
to Jeremiah.
\VS{31}“Send
a message to all
the exiles
in Babylon. Tell
them, ‘The
{\ND{Lord}}
has spoken
about
Shemaiah
the Nehelamite.
“Shemaiah
has spoken to you as
a prophet
even though I
did not
send
him. He is making you trust
in
a lie.
\VS{32}Because
he has done this,”
the
{\ND{Lord}}
says,
“I
will punish
Shemaiah
the Nehelamite
and his whole family.
There will not
be
any
of them
left to experience
the good
things that
I
will do
for my people.
I, the
{\ND{Lord}}, affirm
it! For
he counseled
rebellion against
the
{\ND{Lord}}.” ’ ”

\par }\Chap{30}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah.
\VS{2}“The

{\ND{Lord}}
God
of Israel
says, ‘Write
everything
that
I am about
to
tell
you in a scroll.
\VS{3}For
I,
the {\ND{Lord}}, affirm
that the time
will come
when I will reverse the plight
of my people,
Israel
and Judah,’
says
the {\ND{Lord}}. ‘I will bring them back
to
the land
I gave
their ancestors
and they will take possession of it once again.’ ”
\par }{\SH Israel and Judah Will Be Delivered after a Time of Deep Distress
\par }{\PP \VS{4}So here
is what
the {\ND{Lord}}
has
to
say
about Israel
and Judah.
\par }{\PP \VS{5}Yes,
here is what
he says:
\par }{\Q “You hear
cries
of panic
and of terror;
\par }{\Q there is no
peace in sight.
\par }{\Q \VS{6}Ask
yourselves
this
and consider
it carefully:

\par }{\Q Have
you ever
seen
a man
give birth
to a baby?
\par }{\Q Why
then do I see
all
these strong men
\par }{\Q grabbing
their stomachs
in pain like a
woman giving birth?
\par }{\Q And why do their faces
\par }{\Q turn so deathly
pale?
\par }{\Q \VS{7}Alas,
what a terrible
time
of trouble it is!

\par }{\Q There has never
been any like
it.
\par }{\Q It is a time
of trouble
for the descendants of Jacob,
\par }{\Q but some
of them will be rescued out of it.
\par }{\Q \VS{8}When
the time
for them to be rescued comes,”

\par }{\Q says
the {\ND{Lord}}
who rules over all,
\par }{\Q “I will
rescue you from foreign subjugation.
\par }{\Q I will deliver you from captivity.

\par }{\Q Foreigners
will then no
longer
subjugate them.
\par }{\Q \VS{9}But they will be subject
to the

{\ND{Lord}}
their God
\par }{\Q and to the Davidic
ruler whom
I will raise
up as king
over them.
\par }{\Q \VS{10}So I, the
{\ND{Lord}}, tell you
not
to be afraid,
\par }{\Q you descendants of Jacob,
my servants.
\par }{\Q Do not
be terrified,
people of Israel.
\par }{\Q For
I
will rescue
you and your descendants
\par }{\Q from
a faraway
land
where
you are captives.
\par }{\Q The descendants
of Jacob
will return
to their land
and enjoy peace.
\par }{\Q They will be secure
and no
one will terrify them.
\par }{\Q \VS{11}For
I,
the {\ND{Lord}}, affirm
that
\par }{\Q I will be with you
and will rescue
you.
\par }{\Q I will completely
destroy all
the nations
where
I scattered
you.
\par }{\Q But
I will not
completely
destroy you.
\par }{\Q I will indeed discipline
you, but only in due
measure.
\par }{\Q I will not
allow you to go entirely unpunished.”
\par }{\SH The Lord Will Heal the Wounds of Judah
\par }{\PP \VS{12}Moreover,
the {\ND{Lord}}
says
to the people of Zion,

\par }{\Q “Your injuries
are incurable;
\par }{\Q your wounds
are severe.
\par }{\Q \VS{13}There is no
one to plead
your cause.
\par }{\Q There are no remedies
for your wounds.
\par }{\Q There is no
healing for you.
\par }{\Q \VS{14}All
your allies
have abandoned
you.

\par }{\Q They no longer
have any concern
for you.
\par }{\Q For
I have attacked
you like an enemy
would.
\par }{\Q I have chastened
you cruelly.
\par }{\Q For
your wickedness
is so great
\par }{\Q and your sin
is so much.
\par }{\Q \VS{15}Why
do you complain
about your injuries,
\par }{\Q that your pain
is incurable?
\par }{\Q I have done
all this
to
you
\par }{\Q because your wickedness
is so great
\par }{\Q and your sin
is so much.
\par }{\Q \VS{16}But all
who destroyed
you will be destroyed.
\par }{\Q All
your enemies
will go
into exile.
\par }{\Q Those who plundered
you will be
plundered.
\par }{\Q I will cause those who pillaged
you to be pillaged.
\par }{\Q \VS{17}Yes,
I will restore you to health.
\par }{\Q I will heal
your wounds.
\par }{\Q I, the
{\ND{Lord}}, affirm
it!

\par }{\Q For
you have been
called
an outcast,
\par }{\Q Zion,
whom no
one cares for.”
\par }{\SH The Lord Will Restore Israel and Judah
\par }{\PP \VS{18}The
{\ND{Lord}}
says,
\par }{\Q “I will restore
the ruined houses
of the descendants
of Jacob.
\par }{\Q I will show compassion
on their ruined homes.
\par }{\Q Every city
will be rebuilt
on
its former ruins.
\par }{\Q Every fortified dwelling
will occupy
its traditional site.
\par }{\Q \VS{19}Out
of those
places you will hear
songs of thanksgiving
\par }{\Q and the sounds
of laughter and merriment.
\par }{\Q I
will increase
their number and they will not
dwindle
away.

\par }{\Q I will bring them honor
and they will no
longer be despised.
\par }{\Q \VS{20}The descendants
of Jacob will enjoy their former
privileges.
\par }{\Q Their community
will be reestablished
in my favor

\par }{\Q and I will punish
all
who try to oppress them.
\par }{\Q \VS{21}One of their own
people will be
their leader.
\par }{\Q Their ruler
will come
from their own
number.

\par }{\Q I will invite
him to approach
me, and he will do so.

\par }{\Q For
no one
would dare
approach
me on his own.

\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{22}Then you will again be
my people
\par }{\Q and I
will be
your God.
\par }{\Q \VS{23}Just watch! The wrath
of the {\ND{Lord}}
\par }{\Q will come like a storm.
\par }{\Q Like a raging
storm
it will rage
down
\par }{\Q on
the heads
of those
who are
wicked.
\par }{\Q \VS{24}The anger
of the {\ND{Lord}}
will not
turn back
\par }{\Q until
he has fully carried out his intended purposes.
\par }{\Q In days
to come
you will come to understand this.

\par }\Chap{31}{\PP \par }{\Q \VerseOne{1}At that time
I will be
the God
of all
the clans
of Israel
\par }{\Q and they
will be
my people.
\par }{\Q I, the
{\ND{Lord}}, affirm it!”
\par }{\SH Israel Will Be Restored and Join Judah in Worship
\par }{\PP \VS{2}The
{\ND{Lord}}
says,
\par }{\Q “The people
of Israel
who survived
\par }{\Q death at the hands of the enemy
\par }{\Q will find
favor
in the wilderness
\par }{\Q as they journey
to find rest for themselves.
\par }{\Q \VS{3}In a far-off
land the
{\ND{Lord}}
will manifest himself to them.
\par }{\Q He will say to them, ‘I have loved
you with an everlasting
love.
\par }{\Q That is why
I have continued
to be faithful to you.
\par }{\Q \VS{4}I will rebuild
you, my dear children
Israel,
\par }{\Q so that you will once again be built up.
\par }{\Q Once again
you will take up
the tambourine
\par }{\Q and join
in the happy throng
of dancers.
\par }{\Q \VS{5}Once again
you will plant
vineyards
\par }{\Q on the hills
of Samaria.
\par }{\Q Those who plant
them
\par }{\Q will once again enjoy their fruit.
\par }{\Q \VS{6}Yes,
a time is coming
\par }{\Q when
watchmen
will call
out on the mountains
of Ephraim,
\par }{\Q “Come! Let us go
to Zion
\par }{\Q to worship the
{\ND{Lord}}
our God!” ’ ”
\par }{\PP \VS{7}Moreover,
the {\ND{Lord}}
says,
\par }{\Q “Sing
for joy
for the descendants of Jacob.
\par }{\Q Utter glad shouts
for that foremost of the nations.
\par }{\Q Make your praises
heard.
\par }{\Q Then say,
‘{\ND{Lord}}, rescue
your people.
\par }{\Q Deliver those of Israel
who remain alive.’
\par }{\Q \VS{8}Then I will reply, ‘I will bring
them back from the land
of the north.
\par }{\Q I will gather
them in from the distant parts
of the earth.
\par }{\Q Blind
and lame
people will come with them,
\par }{\Q so will pregnant women
and women about to give birth.
\par }{\Q A vast
throng of people
will come back
here.
\par }{\Q \VS{9}They will come back
shedding tears
of contrition.
\par }{\Q I will bring
them back
praying prayers
of repentance.

\par }{\Q I will lead
them besides
streams
of water,
\par }{\Q along smooth
paths
where they will never
stumble.
\par }{\Q I will do this
because
I am Israel’s
father;
\par }{\Q Ephraim
is my firstborn son.’ ”
\par }{\Q \VS{10}Hear
what
the {\ND{Lord}}
has to say, O nations.
\par }{\Q Proclaim
it in the faraway
lands along the sea.
\par }{\Q Say,
“The one who scattered
Israel
will regather
them.
\par }{\Q He will watch
over his people like a shepherd
watches over his flock.”
\par }{\Q \VS{11}For
the {\ND{Lord}}
will rescue
the descendants of Jacob.
\par }{\Q He will secure their release from
those who
had overpowered them.
\par }{\Q \VS{12}They will come
and shout
for joy on Mount
Zion.
\par }{\Q They will be radiant
with
joy over
the good things
the {\ND{Lord}}
provides,
\par }{\Q the grain,
the fresh wine,
the olive oil,
\par }{\Q the young
sheep
and calves
he has given to them.
\par }{\Q They will be
like a well-watered
garden
\par }{\Q and
will not
grow faint
or weary
any more.
\par }{\Q \VS{13}The
{\ND{Lord}} says, “At that time
young women
will dance
and be glad.
\par }{\Q Young men
and old men
will rejoice.

\par }{\Q I will turn
their grief
into gladness.
\par }{\Q I will give them comfort
and joy
in place of their sorrow.
\par }{\Q \VS{14}I will provide
the priests
with abundant provisions.
\par }{\Q My people
will be filled
to the full with
the good things I provide.”
\par }{\PP \VS{15}The
{\ND{Lord}}
says,
\par }{\Q “A sound
is heard
in Ramah,
\par }{\Q a sound
of crying
in bitter
grief.
\par }{\Q It is the sound of Rachel
weeping
for her
children
\par }{\Q and refusing to be comforted,
because
her children
are gone.”
\par }{\Q \VS{16}The
{\ND{Lord}}
says
to her,

\par }{\Q “Stop
crying! Do not shed any more tears!

\par }{\Q For
your heartfelt
repentance will be rewarded.
\par }{\Q Your children will return
from the land
of the enemy.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{17}Indeed, there is
hope
for your posterity.
\par }{\Q Your children
will return
to their own territory.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{18}I have indeed heard
the people of Israel
say mournfully,
\par }{\Q ‘We were like a calf
untrained to the yoke.

\par }{\Q You disciplined
us and we learned
from it.

\par }{\Q Let us come back
to you and we will do so,

\par }{\Q for
you
are the
{\ND{Lord}}
our God.
\par }{\Q \VS{19}For
after
we turned
away from you we
repented.
\par }{\Q After
we came to our senses
we beat our breasts in sorrow.
\par }{\Q We are ashamed
and humiliated
\par }{\Q because
of the disgraceful
things we did previously.’
\par }{\Q \VS{20}Indeed,
the people of Israel
are my dear
children.
\par }{\Q They are the children
I take delight
in.

\par }{\Q For
even though
I must often rebuke
them,
\par }{\Q I still
remember
them
with fondness.
\par }{\Q So
I am deeply
moved
with pity
for them

\par }{\Q and will surely have compassion
on them.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{21}I will say, ‘My dear children of Israel,
keep in mind
\par }{\Q the road
you took
when you were carried off.

\par }{\Q Mark off
in your minds
the landmarks.
\par }{\Q Make
a mental note of telltale signs marking the way
back.
\par }{\Q Return,
my dear children
of Israel.
\par }{\Q Return
to
these cities
of yours.
\par }{\Q \VS{22}How long
will you vacillate,
\par }{\Q you who were once like an unfaithful daughter?

\par }{\Q For
I,
the {\ND{Lord}}, promise
to bring
about something new
on the earth,
\par }{\Q something as unique as a woman
protecting
a man!’ ”
\par }{\SH Judah Will Be Restored
\par }{\PP \VS{23}The

{\ND{Lord}}
God
of Israel
who rules over all
says,
\par }{\Q “I will restore
the people of Judah
to their land
and to their towns.
\par }{\Q When
I do,
they will again
say
of Jerusalem,

\par }{\Q ‘May
the {\ND{Lord}}
bless
you, you holy
mountain,
\par }{\Q the place where righteousness
dwells.’
\par }{\Q \VS{24}The land of Judah
will be inhabited
by people who live
in its towns
\par }{\Q as well as by farmers
and shepherds with their flocks.
\par }{\Q \VS{25}I will fully
satisfy
the needs of those who
are
weary
\par }{\Q and fully
refresh the souls
of those who are faint.
\par }{\Q \VS{26}Then they will say, ‘Under these conditions
I can enjoy
sweet
sleep
\par }{\Q when I wake up
and look around.’ ”
\par }{\SH Israel and Judah Will Be Repopulated
\par }{\PP \VS{27}“Indeed,
a time
is coming,”
says
the {\ND{Lord}}, “when I will cause
people
and animals
to sprout
up in the lands of Israel
and Judah.
\VS{28}In the past I saw
to it that
they were uprooted
and torn down,
that they were destroyed
and demolished.
But now I will see
to it that they are built up
and firmly planted.
I, the
{\ND{Lord}}, affirm it!”
\par }{\SH The Lord Will Make a New Covenant with Israel and Judah
\par }{\PP \VS{29}“When that time
comes, people will no
longer
say,
‘The parents
have eaten
sour grapes,
but the children’s
teeth
have grown numb.’
\VS{30}Rather, each person
will die
for his own sins.
The teeth
of the person
who eats
the sour grapes
will themselves grow numb.
\par }{\PP \VS{31}“Indeed,
a time
is coming,”
says
the {\ND{Lord}}, “when I will make
a new
covenant
with
the people of
Israel
and Judah.
\VS{32}It will not
be like the old covenant
that
I made
with
their ancestors
when
I delivered
them from
Egypt.
For they
violated
that covenant,
even though I
was like a faithful husband
to them,” says
the {\ND{Lord}}.
\VS{33}“But
I will make a new covenant
with
the whole nation
of Israel
after
I plant
them
back in the land,” says
the {\ND{Lord}}. “I will put
my law
within
them and write
it on
their hearts
and minds. I will be
their God
and they
will be
my people.
\par }{\PP \VS{34}“People
will no
longer
need
to teach
their neighbors
and relatives
to know
me. For
all
of them,
from the least
important
to the most important,
will know
me,” says
the {\ND{Lord}}. “For
I will forgive
their sin
and will no
longer
call to mind
the wrong they have done.”
\par }{\SH The Lord Guarantees Israel’s Continuance
\par }{\Q \VS{35}The
{\ND{Lord}}
has made a promise
to Israel.
\par }{\Q He promises
it as the one who fixed
the sun
to give light
by day
\par }{\Q and the moon
and stars
to give light
by night.
\par }{\Q He promises it as the one who stirs
up the sea
so that its waves
roll.
\par }{\Q He promises it as the one who is known as the
{\ND{Lord}}
who rules over all.
\par }{\Q \VS{36}The
{\ND{Lord}}
affirms, “The descendants
of Israel
will
not
\par }{\Q cease forever to be a nation in my sight.
\par }{\Q That could only happen if the fixed ordering of the heavenly lights
\par }{\Q were to cease to operate before me.”
\par }{\Q \VS{37}The
{\ND{Lord}}
says,
“I
will not reject
all
the descendants
of Israel
\par }{\Q because
of all
that
they have
done.
\par }{\Q That
could only happen if
the heavens
above
could be measured
\par }{\Q or the foundations
of the earth
below
could
all be explored,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\SH Jerusalem Will Be Enlarged
\par }{\PP \VS{38}“Indeed
a time
is coming,” says
the {\ND{Lord}}, “when
the city of Jerusalem will be rebuilt
as my special city.
It will be built
from the Tower
of Hananel
westward to the Corner
Gate.
\VS{39}The boundary line
will extend beyond that, straight
west from there
to the Hill
of Gareb
and then turn southward
to Goah.
\VS{40}The whole
valley
where dead bodies
and sacrificial ashes
are thrown and all
the terraced fields
out to the Kidron
Valley
on the east
as
far north as
the Horse
Gate
will be included within this city that is sacred
to the
{\ND{Lord}}. The city will
never
again
be torn
down or
destroyed.” ‘span class=”footnote” id=”footnote-50” ’‘span class=”key” ’50‘a href=”\#note-50” class=”backref” ’31:22‘/a’‘span class=”text” ’
{\IT{sn}}:
{\IT{Heb}} “create.” This word is always used with God as the subject and refers to the production of something new or unique, like the creation of the world and the first man and woman (
Gen 1:1; 2:3; 1:27; 5:1) or the creation of a new heavens and a new earth in a new age (
Isa 65:17), or the bringing about of new and unique circumstances (
Num 16:30). Here reference is made contextually to the new exodus, that marvelous deliverance which will be so great that the old will pale in comparison (see the first note on v.
9).

\par }\Chap{32}{\PP \VerseOne{1}In the tenth
year
that
Zedekiah
was ruling over
Judah
the {\ND{Lord}}
spoke
to
Jeremiah.
That
was the same as the eighteenth
year
of Nebuchadnezzar.
\par }{\PP \VS{2}Now at that time,
the armies
of the king
of Babylon
were besieging
Jerusalem.
The prophet
Jeremiah
was
confined
in the courtyard
of the guardhouse
attached to the royal
palace
of Judah.
\VS{3}For King
Zedekiah
had confined
Jeremiah there after he had reproved him for prophesying as he did. He had asked
Jeremiah, “Why
do you
keep prophesying
these things? Why do you keep saying
that the

{\ND{Lord}}
says, ‘I
will hand
this
city
over
to the king
of Babylon? I will let him capture it.
\VS{4}King
Zedekiah
of Judah
will not
escape
from the Babylonians.
He will certainly
be handed
over
to the king
of Babylon.
He must
answer personally
to the king of Babylon and confront
him face
to face.
\VS{5}Zedekiah
will be
carried
off to Babylon
and will remain there
until
I have fully dealt
with him.
I, the
{\ND{Lord}}, affirm
it! Even if
you continue to fight
against the Babylonians,
you cannot
win.’ ”
\par }{\PP \VS{6}So now, Jeremiah
said,
“The
{\ND{Lord}}
told
me,
\VS{7}‘Hanamel,
the son
of your uncle
Shallum,
will come
to
you soon. He will say
to you, “Buy
my field
at Anathoth
because
you are entitled
as my closest relative
to buy it.” ’
\VS{8}Now it happened just as the
{\ND{Lord}}
had said! My cousin
Hanamel
came
to me
in the courtyard
of the guardhouse.
He said
to me,
‘Buy
my
field
which
is at Anathoth
in the territory
of the tribe of Benjamin.
Buy
it for
yourself since you are entitled
as my closest relative
to take possession
of it for yourself.’ When this happened, I recognized
that
the {\ND{Lord}}
had indeed
spoken to me.
\VS{9}So I bought
the field
at Anathoth
from my cousin
Hanamel.
I weighed out
seven ounces
of silver and gave it to him to pay for it.
\VS{10}I signed
the deed
of purchase,
sealed
it, and had some men serve as witnesses
to the purchase. I weighed
out the silver
for him on a scale.
\VS{11}There were two copies
of the deed
of purchase.
One was sealed
and contained
the order of transfer and the conditions
of purchase. The other was left unsealed.
\VS{12}I
took
both copies
of the deed
of purchase
and gave them to
Baruch
son
of Neriah,
the son
of Mahseiah.
I gave them to him in the presence
of my cousin
Hanamel,
the witnesses
who had signed
the deed
of purchase,
and all
the Judeans
who were housed in
the courtyard
of the guardhouse.
\VS{13}In the presence
of all these people I instructed
Baruch,
\VS{14}‘The

{\ND{Lord}}
God
of Israel
who rules over all
says, “Take
these
documents,
both the
sealed
copy of the deed of purchase
and the unsealed
copy. Put
them in a clay jar
so that they may
be preserved
for a long
time to come.” ’
\VS{15}For
the

{\ND{Lord}}
God
of Israel
who rules over all
says, “Houses,
fields,
and vineyards
will again
be bought
in this
land.” ’
\par }{\SH Jeremiah’s Prayer of Praise and Bewilderment
\par }{\PP \VS{16}“After
I
had given
the copies
of the deed
of purchase
to
Baruch
son
of Neriah,
I prayed
to
the {\ND{Lord}},
\VS{17}‘Oh,
Lord

{\ND{God}}, you
did indeed
make
heaven
and earth
by your mighty power
and great
strength.
Nothing
is too hard for you!
\VS{18}You show
unfailing love
to thousands.
But you also punish children
for the sins
of their parents.
You are the great
and powerful
God
who is known
as the
{\ND{Lord}}
who rules over all.
\VS{19}You plan
great
things and you do mighty
deeds.
You see
everything
people
do.
You
reward
each
of them
for the way
they live
and for the things they do.
\VS{20}You
did miracles
and amazing
deeds in the land
of Egypt
which have had
lasting
effect.
By
this
means you gained both in Israel
and among humankind
a renown
that lasts to this
day.
\VS{21}You used your mighty power
and your great
strength
to perform miracles
and amazing deeds
and to bring
great
terror
on the Egyptians. By this means you brought
your people
Israel
out
of the land
of Egypt.
\VS{22}You kept
the promise
that you swore on oath
to their ancestors.
You gave
them a land
flowing
with milk
and honey.
\VS{23}But when they came
in and took possession
of it, they did not
obey
you or
live
as
you had
instructed
them. They did
not
do
anything
that
you commanded them
to do. So
you brought all
this
disaster on them.
\VS{24}Even now
siege ramps
have been built up
around the city
in order to capture
it. War,
starvation,
and disease
are sure to make
the city
fall into
the hands
of the Babylonians
who are attacking
it.
{\ND{Lord}}, you threatened
that
this would happen.
Now
you can see that it is already taking place.
\VS{25}The city
is sure to fall into the hands
of the Babylonians.
Yet, in spite of this, you,
Lord

{\ND{God}}, have said
to me,
“Buy
that field
with silver
and have the transaction
legally witnessed.” ’ ”
\par }{\SH The Lord Answers Jeremiah’s Prayer
\par }{\PP \VS{26}The
{\ND{Lord}}
answered
Jeremiah.
\VS{27}“I am
the {\ND{Lord}}, the God
of all
humankind.
There is, indeed,
nothing too
difficult for me.
\VS{28}Therefore
I,
the {\ND{Lord}}, say: ‘I will indeed
hand
this
city
over
to King
Nebuchadnezzar
of Babylon
and the Babylonian
army. They will capture it.
\VS{29}The Babylonian soldiers
that are attacking
this
city
will break into
it and set
it on fire.
They will burn
it down
along with
the houses
where
people have made
me angry
by offering sacrifices
to the god Baal
and by pouring out
drink offerings
to other
gods
on their rooftops.
\VS{30}This will happen
because
the people of Israel
and Judah
have repeatedly
done
what displeases
me from their earliest
history until now and because
they have repeatedly
made me angry
by the things they have done.
I, the
{\ND{Lord}}, affirm it!
\VS{31}This will happen because
the people of this city
have aroused my anger
and my wrath
since
the time
they built
it until
now.
They have
made me so angry that
I am determined to remove
it from my sight.
\VS{32}I am determined to do so
because
the people of Israel
and Judah
have made
me angry
with all
their wickedness
– they,
their kings,
their officials,
their priests,
their prophets,
and especially
the people
of Judah
and the citizens
of Jerusalem
have done this wickedness.
\VS{33}They have turned
away from me
instead of turning
to
me.
I tried over
and over again
to instruct
them, but
they did not
listen
and respond to correction.
\VS{34}They set
up their disgusting idols
in the temple
which
I have claimed
for my own
and defiled it.
\VS{35}They built
places of worship
for the god Baal
in the Valley
of Ben
Hinnom
so that they could sacrifice
their sons
and daughters
to the god Molech.
Such
a disgusting practice was not
something I commanded
them to do! It never
even entered
my mind
to command them to do
such a thing! So
Judah
is certainly liable
for punishment.’
\par }{\PP \VS{36}“You
and your people
are right
in saying,
‘War,
starvation,
and disease
are sure to make
this city fall into the hands
of the king
of Babylon.’
But now
I,
the {\ND{Lord}}
God
of Israel,
have something further to
say about this
city:
\VS{37}‘I
will certainly regather
my people from all
the countries
where
I will have exiled
them in my anger,
fury,
and great
wrath.
I will bring them back
to
this
place
and allow them to live
here in safety.
\VS{38}They will be
my people,
and I
will be
their God.
\VS{39}I will give
them a single-minded
purpose
to live in a way
that always
shows respect
for me. They will want to do that
for their own good and the good
of the children
who descend from them.
\VS{40}I will make
a lasting
covenant
with them that
I will never
stop
doing good
to them.
I will
fill
their hearts
and minds with respect for me so that they will never again turn away from me.
\VS{41}I will take delight
in doing good
to them. I will faithfully
and wholeheartedly
plant
them firmly in the land.’
\par }{\PP \VS{42}“For
I,
the {\ND{Lord}}, say: ‘I
will surely bring
on
these
people
all
the good
fortune that
I am
hereby promising
them. I will be just
as sure to do
that as I have been in bringing
all
this
great
disaster
on them.
\VS{43}You and your people are saying
that this land
will become desolate,
uninhabited
by either people
or animals.
You are saying that it will
be handed
over to the Babylonians.
But fields
will again be bought
in this
land.
\VS{44}Fields
will again be bought
with silver,
and deeds of purchase
signed,
sealed,
and witnessed.
This
will happen in the territory
of Benjamin,
the villages surrounding
Jerusalem,
the towns
in Judah,
the southern hill country,
the western foothills,
and southern
Judah. For
I will restore
them
to their land. I, the
{\ND{Lord}}, affirm it!’ ”

\par }\Chap{33}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to Jeremiah
a second
time while
he was still
confined
in the courtyard
of the guardhouse.
\VS{2}“I, the
{\ND{Lord}}, do
these things. I, the
{\ND{Lord}}, form
the plan to bring
them about. I am known
as the
{\ND{Lord}}. I say to you,
\VS{3}‘Call
on me
in prayer and I will answer
you. I will show
you great
and mysterious
things which you still do not
know about.’
\VS{4}For
I,
the {\ND{Lord}}
God
of Israel,
have something more to
say about the houses
in this
city
and the
royal
buildings
which have been torn down
for defenses against the
siege ramps
and military
incursions of the Babylonians:
\VS{5}‘The defenders of the city will go out
and fight
with
the Babylonians.
But they will only fill
those houses and buildings with
the dead bodies
of the people
that
I will kill
in my anger
and my wrath.
That will happen because
I have decided to turn my back on
this
city
on account
of the wicked things they have done.
\VS{6}But I will
most surely
heal the wounds
of this city and restore
it and its people to health.
I will show
them
abundant
peace
and security.
\VS{7}I will restore
Judah
and Israel
and will rebuild
them as they were in days of old.
\VS{8}I will purify
them from all
the sin
that
they committed
against me. I will forgive
all
their sins which
they committed
in rebelling against me.
\VS{9}All
the nations
will hear
about all
the good
things which
I
will do
to them. This city will bring me fame,
honor,
and praise
before them for the joy
that I bring it. The nations will tremble
in awe
at all
the peace
and prosperity
that
I
will provide for it.’
\par }{\PP \VS{10}“I, the
{\ND{Lord}}, say: ‘You
and your people are saying
about this
place, “It lies in ruins.
There are no
people
or
animals
in it.” That
is true. The towns
of Judah
and the streets
of Jerusalem
will soon be desolate,
uninhabited
either by people or by animals.
But happy sounds will again
be heard
in these places.
\VS{11}Once again there will be sounds
of joy
and gladness
and the glad
celebrations
of brides
and grooms.
Once again people will bring
their thank offerings
to the temple
of the
{\ND{Lord}} and will say,
“Give thanks
to the
{\ND{Lord}}
who rules over all.
For
the {\ND{Lord}}
is good
and his unfailing
love lasts forever.”
For
I, the
{\ND{Lord}}, affirm that I will restore
the land to what it was in days of old.’
\par }{\PP \VS{12}“I,
the {\ND{Lord}}
who rules over all,
say: ‘This
place
will indeed lie in ruins.
There will be no
people
or animals
in it. But there will again
be in it and in its
towns
sheepfolds
where shepherds
can rest
their sheep.
\VS{13}I, the
{\ND{Lord}}, say
that shepherds will once again
count
their sheep
as they pass
into
the fold. They will do this in all the towns
in the southern hill country,
the western foothills,
the southern
hill country,
the territory of Benjamin,
the villages surrounding
Jerusalem,
and the towns
of Judah.’
\par }{\SH The Lord Reaffirms His Covenant with David, Israel, and Levi
\par }{\PP \VS{14}“I, the
{\ND{Lord}}, affirm: ‘The time
will certainly
come
when I will fulfill
my gracious
promise concerning
the nations of
Israel
and Judah.
\VS{15}In those
days
and at that time
I
will raise
up for them a righteous
descendant
of David.
\par }{\PP “‘He will do
what is just
and right
in the land.
\VS{16}Under his rule Judah
will enjoy safety
and Jerusalem
will live
in security.
At that
time Jerusalem will be called
“The
{\ND{Lord}}
has provided us with justice.”
\VS{17}For
I,
the {\ND{Lord}}, promise: “David
will never
lack
a successor
to occupy
the throne
over
the nation
of Israel.
\VS{18}Nor
will the Levitical
priests
ever lack
someone
to stand before
me and continually offer up
burnt offerings,
sacrifice cereal offerings,
and offer
the other sacrifices.” ’ ”
\par }{\PP \VS{19}The
{\ND{Lord}}
spoke
further to
Jeremiah.
\VS{20}“I,
{\ND{Lord}}, make the following
promise: ‘I have made a covenant
with the day
and with the night
that
they
will always come
at their proper times.
Only if
you people could break that covenant
\VS{21}could
my covenant
with
my servant
David and my covenant with
the Levites ever be
broken.
So David
will by all means always have a descendant
to occupy
his throne
as king
and the Levites
will by all means always have priests
who will minister before me.
\VS{22}I
will make
the children
who follow one another in the line of my servant
David
very numerous. I
will also make
the
Levites
who minister
before me very numerous.
I will make
them all as numerous
as the stars in the sky
and as the sands
which are on the seashore.’ ”
\par }{\PP \VS{23}The
{\ND{Lord}}
spoke
still further to
Jeremiah.
\VS{24}“You have
surely
noticed
what
these
people
are saying, haven’t
you? They are saying, ‘The
{\ND{Lord}}
has rejected
the
two
families
of Israel
and Judah
that
he chose.’
So they have little regard
that my people
will ever
again
be a nation.
\VS{25}But I,
the {\ND{Lord}}, make the following promise: I have made a covenant
governing the coming of day
and night.
I have established the fixed
laws governing
heaven
and earth.
\VS{26}Just as surely as
I have
done this, so surely will I never reject
the descendants
of Jacob.
Nor will I ever refuse
to choose
one of my servant
David’s
descendants
to rule
over the descendants
of Abraham,
Isaac,
and Jacob.
Indeed,
I will restore
them and show mercy to them.”

\par }\Chap{34}{\PP \VerseOne{1}The

{\ND{Lord}}
spoke
to
Jeremiah
while King
Nebuchadnezzar
of Babylon
was attacking
Jerusalem
and the towns
around it with a large army. This army consisted
of troops
from his own army
and from
the kingdoms
and peoples
of the lands
under his dominion.
\VS{2}The

{\ND{Lord}}
God
of Israel
told
Jeremiah to
go
and give
King
Zedekiah
of Judah
a message. He told Jeremiah to tell
him,
“The
{\ND{Lord}}
says, ‘I am going
to hand
this
city
over
to the
king
of Babylon
and he will burn
it down.
\VS{3}You
yourself will not
escape
his clutches,
but
will certainly
be captured
and handed over
to him. You must confront
the king
of Babylon
face
to face
and answer
to
him personally. Then you must go
to Babylon.
\VS{4}However,
listen
to what
I, the
{\ND{Lord}}, promise you, King
Zedekiah
of Judah.
I,
the {\ND{Lord}}, promise that you will not
die
in battle
or be executed.
\VS{5}You will die
a peaceful
death. They will burn
incense at your burial
just as
they did at the burial of your ancestors,
the former
kings
who
preceded
you. They will mourn
for you, saying, “Poor, poor master!” Indeed,
you have my own word
on this. I,
the {\ND{Lord}}, affirm it!’ ”
\par }{\PP \VS{6}The prophet
Jeremiah
told
all
this
to King
Zedekiah
of Judah
in Jerusalem.
\VS{7}He did this while the army
of the king
of Babylon
was attacking
Jerusalem
and the cities
of Lachish
and Azekah.
He was attacking these
cities because
they were the only fortified
cities
of Judah
which were still holding out.
\par }{\SH The Lord Threatens to Destroy Those Who Wronged Their Slaves
\par }{\PP \VS{8}The
{\ND{Lord}}
spoke
to
Jeremiah
after
King
Zedekiah
had made
a covenant
with
all
the people
in Jerusalem
to grant
their slaves
their freedom.
\VS{9}Everyone
was supposed
to free
their male
and female
Hebrew
slaves.
No one
was supposed to keep
a fellow
Judean
enslaved.
\VS{10}All
the people
and their leaders
had agreed
to this. They had agreed
to free
their male
and female
slaves
and not
keep them enslaved
any longer.
They originally complied
with the covenant and freed them.
\VS{11}But later
they had changed their minds. They had taken back
their male
and female
slaves
that
they had
freed
and forced
them to be slaves again.
\VS{12}That was when
the {\ND{Lord}}
spoke
to
Jeremiah,
\VS{13}“The

{\ND{Lord}}
God
of Israel
has a message for you. ‘I
made
a covenant
with
your ancestors
when
I brought
them out of Egypt
where they had been slaves. It stipulated,
\VS{14}“Every
seven
years
each
of you must
free
any fellow
Hebrews
who have
sold
themselves to you. After they have served
you for six
years,
you shall set
them free.”
But your ancestors
did not
obey
me or pay
any attention to me.
\VS{15}Recently,
however, you yourselves
showed a change of heart and did
what is pleasing
to me. You granted your
fellow countrymen
their freedom
and you made
a covenant
to that effect in my presence
in the house
that
I have
claimed
for my own.
\VS{16}But then you turned right around
and showed that you did not honor
me.
Each of you
took back
your male
and female
slaves
whom
you had freed
as they desired,
and you forced
them
to be
your slaves again.
\VS{17}So
I,
the {\ND{Lord}}, say: “You
have not
really obeyed me
and granted freedom
to
your neighbor
and fellow
countryman.
Therefore, I
will grant you freedom,
the freedom
to
die
in war,
or by
starvation
or disease.
I, the
{\ND{Lord}}, affirm
it! I will make
all
the kingdoms
of the earth
horrified at what happens to you.
\VS{18}I will
punish those people
who have
violated
their covenant
with me. I will make them
like the calf
they cut
in two
and passed
between
its pieces.
I will do so because they did not
keep the
terms
of the covenant
they made in my presence.
\VS{19}I will punish the leaders
of Judah
and Jerusalem,
the court officials,
the priests,
and all
the other people
of the land
who passed
between
the pieces
of the calf.
\VS{20}I will
hand
them over to their enemies
who want
to kill
them. Their dead bodies
will become
food
for the birds
and the wild
animals.
\VS{21}I will
also
hand King
Zedekiah
of Judah
and his officials
over
to their enemies
who want
to kill
them. I will hand
them over to the army
of the king
of Babylon,
even though they have temporarily withdrawn
from attacking you.
\VS{22}For I, the
{\ND{Lord}}, affirm
that I will soon
give the order
and bring them back
to
this
city.
They will fight
against
it and capture
it and burn
it down.
I will also
make
the towns
of Judah
desolate
so that there will be no
one living in them.” ’ ”

\par }\Chap{35}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
when
Jehoiakim
son
of Josiah
was ruling
over Judah.
\VS{2}“Go
to
the Rechabite
community. Invite
them to come
into
one
of the side rooms
of the
{\ND{Lord}}’s
temple
and offer them some wine
to drink.”
\VS{3}So
I went and got
Jaazaniah
son
of Jeremiah
the grandson
of Habazziniah,
his brothers,
all
his sons,
and all
the rest of the Rechabite community.
\VS{4}I took
them to the
{\ND{Lord}}’s
temple.
I took them into
the room
where the disciples
of the prophet Hanan
son
of Igdaliah
stayed. That
room
was next to the one
where the temple officers
stayed and above
the room
where Maaseiah
son
of Shallum,
one of the doorkeepers of the temple, stayed.
\VS{5}Then I set
cups
and pitchers full
of wine
in front
of the members of the Rechabite
community and said
to
them, “Have
some wine.”
\VS{6}But they answered,
“We do not
drink
wine
because
our ancestor
Jonadab
son
of Rechab
commanded
us not
to. He told
us, ‘You
and your children
must never
drink
wine.
\VS{7}Do not
build
houses.
Do not plant crops.
Do not
plant
a vineyard
or
own one. Live
in tents
all
your lives.
If
you do these things
you
will
live
a long
time in the land
that
you
wander
about on.’
\VS{8}We
and our wives
and our sons
and daughters
have obeyed
everything
our ancestor
Jonadab
commanded
us. We have never
drunk
wine.
\VS{9}We have not
built
any houses
to live
in. We do not own any vineyards,
fields,
or crops.
\VS{10}We have lived
in tents.
We have obeyed
our ancestor
Jonadab
and done
exactly
as
he commanded us.
\VS{11}But when King
Nebuchadnezzar
of Babylon
invaded
the land
we said,
‘Let’s get up
and go to
Jerusalem
to get away from
the Babylonian
and Aramean
armies.’
That is why we are staying
here in Jerusalem.”
\par }{\PP \VS{12}Then the
{\ND{Lord}}
spoke
to Jeremiah.
\VS{13}The

{\ND{Lord}}
God
of Israel
who rules over all
told
him, “Go
and speak to
the people
of Judah
and the citizens
of Jerusalem.
Tell them, ‘I, the
{\ND{Lord}}, say: “You must
learn
a lesson
from this about
obeying what
I say!
\VS{14}Jonadab
son
of Rechab
ordered
his descendants
not
to drink
wine.
His orders have been carried out. To this
day
his descendants
have drunk
no
wine because
they have obeyed what
their ancestor
commanded them. But I
have spoken
to
you over and over again,
but
you have not
obeyed
me!
\VS{15}I sent
all
my servants
the prophets
to warn you over and over again.
They said,
“Every one
of you, stop
doing
the evil
things you
have been doing
and do what is right. Do not
pay allegiance
to other
gods
and worship
them. Then you can continue to live
in this
land
that
I gave
to you and your ancestors.”
But you did not
pay
any attention
or
listen
to me.
\VS{16}Yes,
the descendants
of Jonadab
son
of Rechab
have carried out the orders
that
their ancestor
gave them.
But you people
have
not
obeyed
me!
\VS{17}So
I,
the {\ND{Lord}},
the God
who rules over all,
the God
of Israel,
say: “I will soon
bring
on Judah
and all
the citizens
of Jerusalem
all
the disaster
that
I threatened
to bring on
them. I will do this because
I spoke
to them
but they did not
listen.
I called
out to them but they did not
answer.” ’ ”
\par }{\PP \VS{18}Then Jeremiah
spoke to the Rechabite
community, “The

{\ND{Lord}}
God
of Israel
who rules over all
says, ‘You have obeyed
the orders
of your ancestor
Jonadab.
You have followed
all
his
instructions.
You have done
exactly
as he commanded you.’
\VS{19}So
the

{\ND{Lord}}
God
of Israel
who rules over all
says, ‘Jonadab
son
of Rechab
will never
lack
a male
descendant to serve me.’ ”

\par }\Chap{36}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
in the fourth
year
that Jehoiakim
son
of Josiah
was ruling
over Judah.
\VS{2}“Get
a scroll.
Write
on it everything
I have
told
you to
say about Israel,
Judah,
and all
the other nations
since
I began to speak
to
you in the reign
of Josiah
until
now.
\VS{3}Perhaps
when the people
of Judah
hear
about all
the disaster
I
intend
to bring on them, they will
all stop doing
the evil
things they have been
doing.
If
they do,
I will forgive
their sins
and the wicked things
they have done.”
\par }{\PP \VS{4}So Jeremiah
summoned
Baruch
son
of Neriah.
Then Jeremiah
dictated
to Baruch everything
the {\ND{Lord}}
had
told
him
to say
and Baruch
wrote
it all
down in a scroll.
\VS{5}Then
Jeremiah
told
Baruch,
“I am
no
longer allowed
to go
into the
{\ND{Lord}}’s
temple.
\VS{6}So
you
go
there the next time
all
the people of Judah
come
in from their towns
to fast
in the
{\ND{Lord}}’s
temple.
Read
out loud where all of them can hear you what I told
you the
{\ND{Lord}}
said,
which
you wrote
in the scroll.
\VS{7}Perhaps
then
they will ask
the
{\ND{Lord}}
for mercy
and will all stop
doing
the evil
things they have been doing. For
the
{\ND{Lord}}
has threatened
to
bring great
anger
and wrath
against these
people.”
\par }{\PP \VS{8}So
Baruch
son
of Neriah
did
exactly
what the prophet
Jeremiah
had told him to do. He read
what
the {\ND{Lord}}
had said from the scroll
in the temple
of the {\ND{Lord}}.
\VS{9}All
the people
living in Jerusalem
and all
the people
who came
into Jerusalem
from the towns
of Judah
came to observe
a fast
before
the {\ND{Lord}}. The fast took place in the ninth
month
of the fifth
year
that Jehoiakim
son
of Josiah
was ruling
over Judah.
\VS{10}At that time Baruch
went into the temple
of the {\ND{Lord}}. He stood in the entrance of the room
of Gemariah
the son
of Shaphan
who had been the royal secretary.
That room
was in the upper
court
near the entrance
of the New
Gate.
There, where all
the people
could hear
him, he read
from the scroll
what
Jeremiah had said.
\par }{\PP \VS{11}Micaiah,
who was the son
of Gemariah
and the grandson of
Shaphan,
heard
Baruch read from the scroll
everything
the {\ND{Lord}}
had said.
\VS{12}He went down
to the chamber
of the royal secretary
in
the king’s
palace
and found
all
the court officials
in session
there.
Elishama
the royal secretary,
Delaiah
son
of Shemaiah,
Elnathan
son
of Achbor,
Gemariah
son
of Shaphan,
Zedekiah
son
of Hananiah,
and all
the other officials were seated there.
\VS{13}Micaiah
told
them
everything
he had
heard
Baruch
read
from the scroll
in the hearing
of the people.
\VS{14}All
the officials
sent
Jehudi,
who was the son
of Nethaniah
and the grandson
of Cushi,
to Baruch.
They ordered
him to tell Baruch, “Come
here and bring
with you the scroll
you read
in the hearing
of the people.”
So Baruch
son
of Neriah
went
to
them, carrying
the scroll
in his hand.
\VS{15}They said
to
him, “Please
sit down
and read
it to us.” So Baruch
sat down and read
it to them.
\VS{16}When
they had heard
it all,
they expressed
their alarm
to
one
another.
Then they said
to
Baruch,
“We must certainly give the king
a report about everything
you have read!”
\VS{17}Then
they asked
Baruch,
“How
did you come to write
all
these
words? Do they actually come from Jeremiah’s mouth?”
\VS{18}Baruch
answered,
“Yes, they came from his own mouth.
He dictated
all
these
words
to me and I
wrote
them down in ink
on
this scroll.”
\VS{19}Then the officials
said
to Baruch,
“You
and Jeremiah
must go
and hide.
You
must not
let anyone
know
where
you are.”
\par }{\PP \VS{20}The officials put the scroll
in the room
of Elishama,
the royal secretary,
for safekeeping.
Then they went
to
the court
and reported everything
to the
king.
\VS{21}The king
sent
Jehudi
to get
the scroll.
He went and got
it from the room
of Elishama,
the royal secretary.
Then he
himself read
it to the king
and all
the officials
who were standing
around him.
\VS{22}Since it was the ninth
month
of the
year, the king
was sitting
in his winter
quarters.
A fire was burning
in the firepot
in front of him.
\VS{23}As soon as
Jehudi
had read
three
or four
columns
of the scroll, the king would cut
them off
with
a penknife
and throw
them on the fire
in
the firepot.
He kept doing so until
the whole
scroll
was burned up
in
the fire.
\VS{24}Neither
he nor any of his attendants
showed any alarm
when they heard
all
that had
been read. Nor did they tear
their clothes
to show any
grief
or sorrow.
\VS{25}The king
did not even
listen to
Elnathan,
Delaiah,
and Gemariah,
who had urged
him not to burn
the scroll.
\VS{26}He also ordered
Jerahmeel,
who was one of the royal
princes, Seraiah
son
of Azriel,
and Shelemiah
son
of Abdeel
to arrest
the scribe
Baruch
and the
prophet
Jeremiah.
However, the
{\ND{Lord}}
hid them.
\par }{\SH Baruch and Jeremiah Write Another Scroll
\par }{\PP \VS{27}The
{\ND{Lord}}
spoke
to
Jeremiah
after
Jehoiakim had burned
the scroll
containing what
Jeremiah
had spoken
and Baruch
had
written down.
\VS{28}“Get
another
scroll
and write
on
it everything
that
was written on
the original
scroll
that
King
Jehoiakim
of Judah
burned.
\VS{29}Tell King
Jehoiakim
of Judah,
‘The
{\ND{Lord}}
says,
“You
burned
the scroll.
You asked
Jeremiah, ‘How dare you write
in
this
scroll that the king
of Babylon
will certainly
come
and destroy
this
land
and wipe out
all the people
and animals on it?’ ”
\VS{30}So
the {\ND{Lord}}
says
concerning
King
Jehoiakim
of Judah,
“None
of his line will occupy
the throne
of David.
His dead body
will be
thrown
out to be exposed to scorching heat
by day
and frost
by night.
\VS{31}I will punish
him
and his descendants
and the officials
who serve him for the wicked things they have done. I will bring
on
them,
the citizens
of Jerusalem,
and the people
of Judah
all
the disaster
that
I threatened
to do to
them. I will punish them because I threatened them but they still paid no
heed.” ’ ”
\VS{32}Then
Jeremiah
got
another
scroll
and gave
it to
the scribe
Baruch
son
of Neriah.
As Jeremiah
dictated,
Baruch wrote
on
this
scroll
everything
that had
been on the scroll
that
King
Jehoiakim
of Judah
burned
in the fire.
They
also added
on this scroll several
other messages
of the same kind.

\par }\Chap{37}{\PP \VerseOne{1}Zedekiah
son
of Josiah
succeeded
Jeconiah
son
of Jehoiakim
as
king.
He was elevated to the throne
of the land
of Judah
by King
Nebuchadnezzar
of Babylon.
\VS{2}Neither
he nor the officials
who served him nor the people
of Judah paid any attention
to
what
the {\ND{Lord}}
said
through
the prophet
Jeremiah.
\par }{\SH The Lord Responds to Zedekiah’s Hope for Help
\par }{\PP \VS{3}King
Zedekiah
sent
Jehucal
son
of Shelemiah
and the priest
Zephaniah
son
of Maaseiah
to
the prophet
Jeremiah.
He told them to say,
“Please
pray
to
the {\ND{Lord}}
our God on our behalf.”
\VS{4}(Now Jeremiah
had not
yet been put
in prison.
So he was still free to come
and go
among
the people as he pleased.
\VS{5}At that time the Babylonian
forces had temporarily given up their siege
against
Jerusalem.
They had had it under siege, but withdrew
when they heard
that
the army
of Pharaoh
had set out
from Egypt. )
\VS{6}The
{\ND{Lord}}
gave the prophet
Jeremiah
a message
for them. He told him to tell them,
\VS{7}“The

{\ND{Lord}}
God
of Israel
says,
‘Give a message to
the king
of Judah
who sent
you to
ask
me to help him. Tell him, “The army
of Pharaoh
that was on its way to help
you will go back
home
to Egypt.
\VS{8}Then the Babylonian
forces will return.
They will attack
the city
and will capture
it and burn
it down.
\VS{9}Moreover,
I, the
{\ND{Lord}}, warn you not
to deceive
yourselves
into thinking
that the Babylonian
forces will go
away and leave
you alone. For
they will not
go away.
\VS{10}For
even if
you were to defeat
all
the Babylonian
forces fighting
against you
so badly
that only
wounded
men
were left
lying in their tents,
they would get
up and burn
this
city
down.” ’ ”
\par }{\SH Jeremiah is Charged with Deserting, Arrested, and Imprisoned
\par }{\PP \VS{11}The following events also occurred while
the Babylonian
forces had temporarily withdrawn
from Jerusalem
because
the army
of Pharaoh was coming.
\VS{12}Jeremiah
started to leave
Jerusalem
to go
to the territory
of Benjamin.
He wanted to make sure he got his share of the property that was being divided
up among
his family
there.
\VS{13}But he
only got as far as the Benjamin
Gate.
There
an officer
in charge of the guards
named
Irijah,
who was the son
of Shelemiah
and the grandson
of Hananiah,
stopped him. He seized
Jeremiah
and said, “You
are deserting
to
the Babylonians!”
\VS{14}Jeremiah
answered,
“That’s a lie! I am not
deserting
to
the Babylonians.”
But Irijah would not
listen
to
him. Irijah
put
Jeremiah
under arrest and took
him to
the officials.
\VS{15}The officials
were very angry
at Jeremiah.
They had
him flogged
and put
in prison
in the house
of Jonathan,
the royal secretary,
which they had converted
into a place for confining prisoners.
\par }{\PP \VS{16}So
Jeremiah
was put
in
prison
in a cell
in the dungeon
in Jonathan’s house. He
was kept
there
for a long
time.
\VS{17}Then
King
Zedekiah
had him brought
to the palace.
There he questioned him
privately
and asked him, “Is there
any message
from the

{\ND{Lord}}?” Jeremiah
answered,
“Yes, there
is.” Then he announced, “You will be handed
over to the king
of Babylon.”
\VS{18}Then Jeremiah
asked
King
Zedekiah,
“What
crime
have I committed against
you, or the officials who serve
you, or the people
of Judah? What have I done to make you people throw
me into
prison?
\VS{19}Where
now are the prophets
who
prophesied
to you that the king
of Babylon
would not
attack
you or this
land?
\VS{20}But now
please
listen,
your royal
Majesty,
and grant
my plea
for mercy.
Do not
send me back
to the house
of Jonathan,
the royal secretary.
If you do,
I will die
there.”
\VS{21}Then
King
Zedekiah
ordered
that Jeremiah
be committed to the courtyard
of the guardhouse.
He also ordered
that a loaf
of bread
be given
to him every day
from
the baker’s
street
until
all
the bread
in the city
was gone.
So Jeremiah
was kept in the courtyard
of the guardhouse.

\par }\Chap{38}{\PP \VerseOne{1}Now
Shephatiah
son
of Mattan,
Gedaliah
son
of Pashhur,
Jehucal
son
of Shelemiah,
and Pashhur
son
of Malkijah
had heard the things
that
Jeremiah
had been telling
the people.
They had heard him say,
\VS{2}“The
{\ND{Lord}}
says,
‘Those who stay
in this
city
will die
in battle
or of starvation
or disease.
Those who leave the city and surrender
to
the Babylonians
will live.
They will
escape
with their lives.’ ”
\VS{3}They had also heard him say, “The
{\ND{Lord}}
says,
‘This
city
will certainly be handed
over to the army
of the king
of Babylon.
They will capture it.’ ”
\VS{4}So these officials
said to
the king,
“This
man
must be put to death.
For
he is
demoralizing
the
soldiers
who are left
in the city
as well
as all
the other people
there
by these
things
he is saying.
This
man
is not
seeking
to help
these
people
but
is trying to harm them.”
\VS{5}King
Zedekiah
said
to them,
“Very well, you can
do what you want with
him. For
I cannot
do anything to stop you.”
\VS{6}So the
officials took
Jeremiah
and put him
in
the
cistern
of Malkijah,
one of the royal princes,
that
was in the courtyard
of the guardhouse.
There was no
water
in the cistern, only
mud.
So when they lowered
Jeremiah
into the cistern
with ropes
he
sank
in the mud.
\par }{\SH An Ethiopian Official Rescues Jeremiah from the Cistern
\par }{\PP \VS{7}An Ethiopian,
Ebed Melech,
a court official
in the royal
palace,
heard
that
Jeremiah
had been put in the cistern.
While the king
was holding
court at the Benjamin
Gate,
\VS{8}Ebed Melech
departed
the palace
and went to speak
to
the king.
He said to him,
\VS{9}“Your royal
Majesty,
those
men
have been very wicked
in all
that
they have
done
to the prophet
Jeremiah.
They have
thrown
him into
a cistern
and he is sure to die
of starvation
there because
there is no
food
left
in the city.
\VS{10}Then
the king
gave
Ebed Melech
the Ethiopian
the following
order: “Take
thirty
men
with you from
here and go pull
the prophet
Jeremiah
out
of the cistern
before
he dies.”
\VS{11}So
Ebed Melech
took the men
with him
and went to
a room under
the treasure
room in the palace.
He got
some worn-out
clothes
and old
rags
from there and let
them down
by
ropes
to Jeremiah
in
the cistern.
\VS{12}Ebed Melech
called down
to
Jeremiah,
“Put
these rags
and worn-out
clothes
under
your armpits
to pad
the ropes.
Jeremiah
did
as
Ebed Melech instructed.
\VS{13}So they pulled
Jeremiah
up
from
the cistern
with ropes.
Jeremiah,
however, still remained
confined to the courtyard
of the guardhouse.
\par }{\SH Jeremiah Responds to Zedekiah’s Request for Secret Advice
\par }{\PP \VS{14}Some time later Zedekiah
sent
and had Jeremiah
brought
to
him
at the third
entrance
of the
{\ND{Lord}}’s
temple.
The king
said
to
Jeremiah,
“I
would like to ask
you a question.
Do not
hide
anything
from me when you answer.”
\VS{15}Jeremiah
said
to
Zedekiah,
“If
I answer you, you will certainly
kill
me. If
I give you advice,
you will not
listen
to me.”
\VS{16}So
King
Zedekiah
made
a secret
promise
to
Jeremiah
and sealed it with an oath.
He promised, “As surely
as the
{\ND{Lord}}
lives who has
given
us life
and breath, I promise you this: I will not
kill
you or
hand
you over to those
men
who
want
to kill you.”
\par }{\PP \VS{17}Then Jeremiah
said
to
Zedekiah,
“The

{\ND{Lord}}, the God
who rules over all,
the God
of Israel,
says, ‘You must
surrender
to
the officers
of the king
of Babylon.
If you do, your life
will be spared
and this
city
will not
be burned down.
Indeed, you
and your whole family
will be spared.
\VS{18}But if
you do not
surrender
to
the officers
of the king
of Babylon,
this
city
will be handed
over to the Babylonians
and they will burn
it down.
You
yourself will not
escape
from them.’ ”
\VS{19}Then King
Zedekiah
said to
Jeremiah,
“I am afraid
of the Judeans
who have
deserted
to
the Babylonians.
The Babylonians might
hand
me over to them and they will torture me.”
\VS{20}Then Jeremiah
answered, “You will not
be handed
over to them. Please
obey
the {\ND{Lord}}
by doing what
I
have been telling
you. Then all will go well
with you and your life
will be spared.
\VS{21}But if
you
refuse
to surrender,
the {\ND{Lord}}
has shown me a vision of what will happen. Here is
what
I saw:
\VS{22}All
the women
who
are left
in the royal
palace
of Judah
will be led out
to
the officers
of the king
of Babylon.
They
will taunt you saying,
\par }{\Q ‘Your trusted friends
misled
you;
\par }{\Q they have gotten the best
of you.
\par }{\Q Now that your feet
are stuck
in the mud,
\par }{\Q they have turned
their backs on you.’
\par }{\PP \VS{23}“All
your wives
and your children
will be turned
over
to
the Babylonians.
You
yourself will not
escape
from them but will be captured
by the king
of Babylon.
This
city
will be burned down.”
\par }{\PP \VS{24}Then Zedekiah
told Jeremiah,
“Do
not let anyone
know
about the conversation
we have had. If
you do, you will die.
\VS{25}The officials
may hear
that
I have talked
with
you. They may come
to
you and say,
‘Tell
us
what
you said
to
the king
and what
the king
said
to
you. Do
not hide
anything from
us. If you do, we will kill
you.’
\VS{26}If they do this, tell
them,
‘I
was pleading
with the king
not
to send me back
to die
in the dungeon
of Jonathan’s
house.’ ”
\VS{27}All
the officials
did indeed come
and question
Jeremiah.
He told
them
exactly what
the king
had instructed
him to say. They stopped questioning
him any further because
no
one had actually heard
their conversation.
\VS{28}So
Jeremiah
remained
confined in the courtyard
of the guardhouse
until
the day
Jerusalem
was captured.
\par }{\SH The Fall of Jerusalem and Its Aftermath
\par }{\PP The following events occurred
when Jerusalem
was captured.

\par }\Chap{39}{\PP \VerseOne{1}King
Nebuchadnezzar
of Babylon
came
against Jerusalem
with his whole
army
and laid siege to
it. The siege
began in the tenth
month
of the ninth
year
that Zedekiah
ruled over
Judah.
\VS{2}It lasted until the ninth
day
of the fourth
month
of Zedekiah’s
eleventh
year.
On that day they broke
through the city walls.
\VS{3}Then Nergal-Sharezer
of Samgar, Nebo-Sarsekim,
who was a chief officer,
Nergal-Sharezer,
who was a high official,
and all
the other
officers
of the king
of Babylon
came
and set up quarters
in the Middle
Gate.
\VS{4}When
King
Zedekiah
of Judah
and all
his soldiers
saw them, they tried
to escape.
They departed
from
the city
during the night.
They took a path
through the king’s
garden
and passed out
through the gate
between
the two
walls.
Then
they headed for the Jordan Valley.
\VS{5}But the Babylonian
army
chased
after
them. They caught
up with Zedekiah
in the plains
of Jericho
and captured him. They took
him to
King
Nebuchadnezzar
of Babylon
at Riblah
in the territory
of Hamath
and Nebuchadnezzar passed sentence on him there.
\VS{6}There at Riblah
the king
of Babylon
had Zedekiah’s
sons
put to death
while Zedekiah was forced to watch.
The king
of Babylon
also had all
the nobles
of Judah
put to death.
\VS{7}Then he had Zedekiah’s
eyes
put out
and had him bound
in chains
to be led off
to Babylon.
\VS{8}The Babylonians
burned down
the
royal
palace,
the temple
of the
{\ND{Lord}}, and the people’s
homes, and they tore
down the wall
of Jerusalem.
\VS{9}Then Nebuzaradan,
the captain
of the royal guard,
took captive
the rest
of the people
who were left
in the city.
He carried
them off to Babylon
along with
the people
who
had deserted to him.
\VS{10}But he left behind
in the land
of Judah
some
of the poor
people
who
owned nothing.
He gave
them fields
and vineyards
at that
time.
\par }{\PP \VS{11}Now King
Nebuchadnezzar
of Babylon
had issued orders
concerning
Jeremiah.
He had passed them on through
Nebuzaradan,
the captain
of his royal guard,
\VS{12}“Find
Jeremiah
and look
out for him.
Do not
do
anything
to harm
him, but
do
with
him whatever
he tells you.”
\VS{13}So Nebuzaradan,
the captain
of the royal guard,
Nebushazban,
who was a chief officer,
Nergal-Sharezer,
who was a high official,
and all
the other officers
of the king
of Babylon
\VS{14}sent
and had Jeremiah
brought
from the courtyard
of the guardhouse.
They turned him over
to
Gedaliah,
the son
of Ahikam
and the grandson of
Shaphan,
to
take
him home
with him. But Jeremiah stayed
among
the people.
\par }{\SH Ebed Melech Is Promised Deliverance because of His Faith
\par }{\PP \VS{15}Now the
{\ND{Lord}}
had
spoken to Jeremiah
while
he was still confined
in the courtyard
of the guardhouse,
\VS{16}“Go
and tell
Ebed-Melech
the Ethiopian,
‘The
{\ND{Lord}}
God
of Israel
who rules over all
says,
“I
will carry out
against
this
city
what I promised.
It will mean disaster
and not
good fortune
for it. When that
disaster happens, you will be there to see it.
\VS{17}But I will
rescue
you
when
it happens. I, the
{\ND{Lord}}, affirm
it! You will not
be handed
over to those whom
you
fear.
\VS{18}I will certainly
save
you. You will not
fall
victim to violence.
You will escape with your life
because
you trust
in me. I, the
{\ND{Lord}}, affirm it!” ’ ”

\par }\Chap{40}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
after
Nebuzaradan
the captain
of the royal guard
had set him free
at Ramah.
He had taken
him there in chains
along with all
the people from Jerusalem
and Judah
who were being carried off
to exile
to Babylon.
\VS{2}The captain
of the royal guard
took
Jeremiah
aside and said
to him,
“The
{\ND{Lord}}
your God
threatened
this
place
with this
disaster.
\VS{3}Now he has brought
it about.
The
{\ND{Lord}}
has done
just
as he threatened
to do. This
disaster has happened because
you people sinned
against the
{\ND{Lord}}
and did not
obey
him.
\VS{4}But now,
Jeremiah, today
I
will set you free
from
the chains
on
your wrists.
If
you would like
to come
to Babylon
with
me, come
along and I will take care of you. But if you prefer
not to
come
to
Babylon
with
me, you are not
required to
do so.
You are free
to
go anywhere
in the land
you want
to
go.
Go
wherever
you choose.”
\VS{5}Before
Jeremiah could turn to leave,
the captain of the guard added, “Go back
to
Gedaliah,
the son
of Ahikam
and grandson
of Shaphan,
whom
the king
of Babylon
appointed to govern
the towns
of Judah.
Go back and live
with
him among
the people.
Or
go
wherever
else you choose.”
Then the captain
of the guard
gave
Jeremiah some food
and a present
and let
him go.
\VS{6}So Jeremiah
went
to
Gedaliah
son
of Ahikam
at Mizpah
and lived
there with
him. He stayed there to live
among
the people
who had been left
in the land of Judah.
\par }{\SH A Small Judean Province is Established at Mizpah
\par }{\PP \VS{7}Now some
of the officers
of the Judean army
and their troops had been hiding in the countryside.
They
heard
that
the king
of Babylon
had appointed
Gedaliah
son
of Ahikam
to govern
the country.
They also heard that
he had been put in charge over
the men,
women,
and children
from the poorer
classes of the land
who had not
been carried off into exile
in Babylon.
\VS{8}So
all these officers and their troops came
to
Gedaliah
at Mizpah.
The officers who came were
Ishmael
son
of Nethaniah,
Johanan
and Jonathan
the sons
of Kareah,
Seraiah
son
of Tanhumeth,
the sons
of Ephai
the Netophathite,
and Jezaniah
son
of the Maacathite.
\VS{9}Gedaliah,
the son
of Ahikam
and grandson of
Shaphan,
took an oath
so as to give them and their troops some assurance
of safety. “Do not
be afraid
to submit
to the Babylonians.
Settle
down in the land
and submit
to the
king
of Babylon.
Then things will go well for you.
\VS{10}I
for my part will stay
at Mizpah
to represent
you before
the Babylonians
whenever they come
to
us. You
for your part go ahead and harvest
the wine,
the dates, the figs, and the olive oil,
and store
them in jars.
Go ahead and settle down
in the towns
that
you have taken over.”
\VS{11}Moreover,
all
the Judeans
who
were in Moab,
Ammon,
Edom,
and all
the other countries
heard
what had happened.
They heard that
the king
of Babylon
had allowed
some people
to stay
in Judah
and that
he had appointed
Gedaliah,
the son
of Ahikam
and grandson
of Shaphan, to govern them.
\VS{12}So all
these Judeans
returned
to the land
of Judah
from the places
where
they had been scattered.
They came
to
Gedaliah
at Mizpah.
Thus they harvested
a large amount
of wine
and dates and figs.
\par }{\SH Ishmael Murders Gedaliah and Carries the Judeans at Mizpah off as Capt
\par }{\PP \VS{13}Johanan
and all
the officers
of the troops
that had
been hiding in the open country
came
to
Gedaliah
at Mizpah.
\VS{14}They said
to
him, “Are you at all aware
that
King
Baalis
of Ammon
has sent
Ishmael
son
of Nethaniah
to kill
you?” But Gedaliah
son
of Ahikam
would not
believe them.
\VS{15}Then Johanan
son
of Kareah
spoke
privately
to
Gedaliah
there at Mizpah,
“Let
me go
and kill
Ishmael
the son
of Nethaniah
before anyone
knows
about it. Otherwise he will kill
you
and all
the Judeans
who have rallied around
you will be scattered.
Then what remains
of Judah
will disappear.”
\VS{16}But Gedaliah
son
of Ahikam
said to
Johanan
son
of Kareah,
“Do not
do
that because
what
you
are saying
about Ishmael
is not true.”

\par }\Chap{41}{\PP \VerseOne{1}But in the seventh
month
Ishmael,
the son
of Nethaniah
and grandson
of Elishama
who was a member
of the royal
family and had been one of Zedekiah’s
chief officers,
came
with
ten
of his men
to
Gedaliah
son
of Ahikam
at Mizpah.
While they were eating
a meal
together
with him there
at Mizpah,
\VS{2}Ishmael
son
of Nethaniah
and the ten
men
who
were with
him stood up, pulled out their swords, and killed
Gedaliah,
the son
of Ahikam
and grandson of
Shaphan.
Thus Ishmael
killed
the man that
the king
of Babylon
had appointed to govern
the country.
\VS{3}Ishmael
also killed
all
the Judeans
who were
with
Gedaliah
at Mizpah
and the Babylonian
soldiers
who
happened
to be
there.
\par }{\PP \VS{4}On the day
after Gedaliah
had been murdered,
before anyone
even knew about it,
\VS{5}eighty
men
arrived
from Shechem,
Shiloh,
and Samaria.
\VS{6}Ishmael
son
of Nethaniah
went out
from
Mizpah
to meet
them. He was pretending to cry
as he walked
along. When
he met
them, he said
to them,
“Come
with me to
meet Gedaliah
son
of Ahikam.”
\VS{7}But as soon as they were inside
the city,
Ishmael
son
of Nethaniah
and the men
who
were with
him slaughtered
them and threw their bodies in
a cistern.
\VS{8}But there were ten
men
among them who said
to
Ishmael,
“Do not
kill
us. For
we will give you the stores of wheat,
barley,
olive oil,
and honey
we have hidden
in a field.
So he spared
their lives and did not
kill
them along
with the rest.
\VS{9}Now the cistern
where
Ishmael
threw
all
the dead bodies
of those he had
killed
was a large one that
King
Asa
had constructed
as part of his defenses against King
Baasha
of Israel.
Ishmael
son
of Nethaniah
filled
it with dead bodies.
\VS{10}Then Ishmael
took captive
all
the people
who were still left
alive in Mizpah.
This included the royal
princesses
and all
the rest
of the people
in Mizpah
that
Nebuzaradan,
the captain
of the royal guard,
had put under the authority of Gedaliah
son
of Ahikam.
Ishmael
son
of Nethaniah
took
all these people captive
and set out
to cross
over to
the Ammonites.
\par }{\SH Johanan Rescues the People Ishmael Had Carried Off
\par }{\PP \VS{11}Johanan
son
of Kareah
and all
the army
officers
who were
with
him heard about
all
the atrocities
that
Ishmael
son
of Nethaniah
had committed.
\VS{12}So they took
all
their troops
and went
to fight
against
Ishmael
son
of Nethaniah.
They caught
up with him
near the large
pool
at Gibeon.
\VS{13}When
all
the people
that
Ishmael
had taken captive
saw Johanan
son
of Kareah
and all
the army
officers
with
him, they were glad.
\VS{14}All
those
people
that
Ishmael
had taken captive
from
Mizpah
turned
and went
over to Johanan
son
of Kareah.
\VS{15}But Ishmael
son
of Nethaniah
managed to escape
from Johanan
along with eight
of his men,
and he went
on over to
Ammon.
\par }{\PP \VS{16}Johanan
son
of Kareah
and all
the army
officers
who
were with
him led off all
the people
who had been left alive
at Mizpah.
They had rescued
them from
Ishmael
son
of Nethaniah
after
he killed
Gedaliah
son
of Ahikam.
They led off the men,
women,
children,
soldiers,
and court officials
whom
they had brought away
from Gibeon.
\VS{17}They set out
to go to
Egypt
to get away
from the Babylonians, but stopped
at Geruth
Kimham near
Bethlehem.
\VS{18}They were afraid
of what
the Babylonians
might do
because
Ishmael
son
of Nethaniah
had killed
Gedaliah
son
of Ahikam,
whom
the king
of Babylon
had appointed to govern
the country.

\par }\Chap{42}{\PP \VerseOne{1}Then all
the army
officers,
including Johanan
son
of Kareah
and Jezaniah
son
of Hoshaiah
and all
the people
of every class,
went to the prophet Jeremiah.
\VS{2}They said
to
him,
“Please
grant
our request
and pray
to
the {\ND{Lord}}
your God
for
all
those of us who are still left alive
here.
For,
as
you yourself can see,
there are only
a few
of us left out
of the many there were before.
\VS{3}Pray that the
{\ND{Lord}}
your God
will tell
us where
we should
go
and what
we should do.”
\VS{4}The prophet
Jeremiah
answered them, “Agreed! I will indeed
pray
to
the {\ND{Lord}}
your God
as
you have asked. I will
tell you everything
the {\ND{Lord}}
replies
in response to you. I will not
keep anything
back from you.”
\VS{5}They
answered
Jeremiah,
“May the
{\ND{Lord}}
be
a true
and faithful
witness
against us if
we do not
do just as
the {\ND{Lord}}
sends
you to
tell us to do.
\VS{6}We will obey
what
the {\ND{Lord}}
our God
to whom
we
are sending you tells us to
do. It does not matter whether we like what he tells
us
or
not. We
will obey
what
he tells
us to do so
that
things will go well
for us.”
\par }{\PP \VS{7}Ten
days
later
the {\ND{Lord}}
spoke
to Jeremiah.
\VS{8}So Jeremiah summoned
Johanan
son
of Kareah
and all
the army
officers
who
were with
him and all
the people
of every class.
\VS{9}Then Jeremiah said
to them,
“You sent
me to the
{\ND{Lord}}
God
of Israel
to make your request
known to him. Here is what
he says to you:
\VS{10}‘If
you will just
stay
in this
land,
I will build
you up. I will not
tear
you down. I will firmly plant
you. I will not
uproot
you. For
I am filled with sorrow
because of the disaster
that
I have
brought on you.
\VS{11}Do not
be afraid
of the king
of Babylon
whom
you
now fear.
Do not
be afraid
of him
because
I
will be with
you to save
you and to rescue
you from his power.
I, the
{\ND{Lord}}, affirm it!
\VS{12}I will
have compassion
on you so that he in turn will have mercy
on you and allow you to return
to
your land.’
\par }{\PP \VS{13}“You
must
not
disobey
the {\ND{Lord}}
your God
by saying,
‘We will not
stay
in this
land.’
\VS{14}You must not say,
‘No, we will not
stay. Instead we will go
and live
in the land
of Egypt
where
we will not
face
war,
or
hear
the enemy’s trumpet
calls,
or
starve
for lack of food.’
\VS{15}If you people who remain
in Judah
do that,
then listen
to what
the {\ND{Lord}}
says.
The
{\ND{Lord}}
God
of Israel
who rules over all
says, ‘If
you
are so
determined
to go
to Egypt
that you go
and settle
there,
\VS{16}the wars
you
fear
will catch
up with you there
in the
land
of Egypt.
The starvation
you
are worried
about will follow
you
there
to Egypt.
You will die
there.
\VS{17}All
the people
who
are determined
to go
and settle
in Egypt
will die
from war,
starvation,
or disease.
No
one will
survive
or escape
the disaster
I
will bring
on them.’
\VS{18}For
the {\ND{Lord}}
God
of Israel
who rules over all
says,
‘If you go to Egypt, I will pour out
my wrath
on
you just
as
I poured out
my anger and wrath
on
the citizens
of Jerusalem.
You will become
an object of horror
and ridicule,
an example of those who have been cursed
and that people use in pronouncing a curse.
You will never
see
this
place
again.’
\par }{\PP \VS{19}“The
{\ND{Lord}}
has told
you people who remain
in Judah,
‘Do not
go
to Egypt.’
Be very sure
of this: I warn
you here and now.
\VS{20}You are making
a fatal
mistake. For
you
sent
me to
the {\ND{Lord}}
your God
and asked
me, ‘Pray
to
the {\ND{Lord}}
our God
for us. Tell
us what the
{\ND{Lord}}
our God
says
and we will do it.’
\VS{21}This day
I have told you what he said. But you do not
want
to obey
the {\ND{Lord}}
by doing what he sent
me to
tell you.
\VS{22}So now
be very sure
of this: You will die
from war,
starvation,
or disease
in the place
where
you want
to go
and live.”

\par }\Chap{43}{\PP \VerseOne{1}Jeremiah
finished
telling
all
the people
all
these
things
the {\ND{Lord}}
their God
had
sent
him to tell them.
\VS{2}Then Azariah
son
of Hoshaiah,
Johanan
son
of Kareah,
and other
arrogant
men
said
to
Jeremiah,
“You
are telling
a lie! The
{\ND{Lord}}
our God
did not
send
you to tell
us, ‘You must not
go
to Egypt
and settle
there.’
\VS{3}But
Baruch
son
of Neriah
is stirring
you up against us.
He wants
to hand
us over
to the Babylonians
so that
they will
kill
us or carry us off into exile
in Babylon.”
\VS{4}So Johanan
son
of Kareah,
all
the army
officers,
and all
the rest of the people
did not obey
the
{\ND{Lord}}’s
command to stay
in the land.
\VS{5}Instead
Johanan
son
of Kareah
and all
the army
officers
led off all
the Judean
remnant
who had come back
to live
in the land
of Judah
from all
the nations
where
they had been scattered.
\VS{6}They also led off
all
the men,
women,
children,
and royal
princesses
that
Nebuzaradan,
the captain
of the royal guard,
had
left
with
Gedaliah,
the son
of Ahikam
and grandson of
Shaphan.
This included the prophet
Jeremiah
and Baruch
son
of Neriah.
\VS{7}They went
on to Egypt
because
they refused
to obey
the {\ND{Lord}}, and came
to
Tahpanhes.
\par }{\SH Jeremiah Predicts that Nebuchadnezzar Will Plunder Egypt and Its Gods
\par }{\PP \VS{8}At Tahpanhes
the {\ND{Lord}}
spoke
to
Jeremiah.
\VS{9}“Take
some large
stones
and bury
them in the mortar
of the clay pavement
at the entrance
of Pharaoh’s
residence
here in Tahpanhes.
Do it while the people of Judah
present there are
watching.
\VS{10}Then tell
them, ‘The

{\ND{Lord}}
God
of Israel
who rules over all
says, “I
will bring
my servant
King
Nebuchadnezzar
of Babylon.
I will set
his throne
over
these
stones
which
I have buried.
He will pitch his royal tent
over them.
\VS{11}He will come
and attack
Egypt.
Those who
are destined to die of disease will die of disease. Those who
are destined to be carried off into exile
will
be carried off into exile.
Those who
are destined to
die in war
will die in war.
\VS{12}He will set
fire
to the temples
of the gods
of Egypt.
He will burn
their gods or carry them off as captives.
He will pick Egypt
clean like
a shepherd
picks the lice
from his clothing.
He will leave
there
unharmed.
\VS{13}He will demolish
the sacred pillars
in the temple of the sun
in Egypt
and will burn down
the temples
of the gods
of Egypt.” ’ ”

\par }\Chap{44}{\PP \VerseOne{1}The
{\ND{Lord}} spoke
to
Jeremiah
concerning all
the Judeans
who were living
in the land
of Egypt,
those in
Migdol,
Tahpanhes,
Memphis,
and in the region
of southern Egypt.
\VS{2}“The

{\ND{Lord}}
God
of Israel
who rules over all
says, ‘You
have seen
all
the disaster
I brought
on
Jerusalem
and all
the towns
of Judah.
Indeed,
they now lie in ruins
and are deserted.
\VS{3}This happened because
of the wickedness
the people living there did.
They made
me angry
by worshiping
and offering sacrifice
to other
gods
whom
neither they
nor you
nor your ancestors
previously knew.
\VS{4}I sent
my servants
the prophets
to you people over and over again
warning
you not
to do
this
disgusting
thing
I hate.
\VS{5}But the people
of Jerusalem and Judah would not listen
or pay
any attention.
They would not stop
the wickedness
they were doing nor quit
sacrificing
to other
gods.
\VS{6}So my anger
and my wrath
were poured
out and burned
like a fire through the towns
of Judah
and the streets
of Jerusalem.
That is why they have become
the desolate
ruins
that they are today.’
\par }{\PP \VS{7}“So now
the

{\ND{Lord}}, the God
who rules over all,
the God
of Israel,
asks, ‘Why
will you
do
such great
harm
to
yourselves? Why should every man,
woman,
child, and baby
of yours be destroyed from the midst
of Judah? Why should you leave yourselves
without a remnant?
\VS{8}That is what will result from your making me angry
by what you are doing.
You are making me angry by sacrificing
to other
gods
here in the land
of Egypt
where
you
live.
You will be destroyed
for doing that! You will become
an example used in curses
and an object of ridicule
among all
the nations
of the earth.
\VS{9}Have you forgotten
all the wicked
things that have been
done
in the towns
of Judah
and in the streets
of Jerusalem
by your ancestors,
by
the kings
of Judah
and their wives,
by
you and your wives?
\VS{10}To this
day
your people have shown no
contrition! They have not revered
me nor followed
the laws
and statutes
I commanded you
and your ancestors.’
\par }{\PP \VS{11}“Because
of this,
the {\ND{Lord}}
God
of Israel
who rules over all
says, ‘I am
determined
to bring disaster
on you, even to the point of destroying
all
the Judeans here.
\VS{12}I will see
to it that all the Judean
remnant
that
was determined
to go
and live
in the land
of Egypt
will be destroyed.
Here in
the land
of Egypt
they will fall
in battle
or perish
from starvation.
People of every class
will die
in war
or from starvation.
They will become
an object of horror
and ridicule,
an example of those who have been cursed
and that people use in pronouncing a curse.
\VS{13}I will punish
those who live
in the land
of Egypt
with war,
starvation,
and disease
just
as I punished
Jerusalem.
\VS{14}None
of the Judean
remnant
who have come
to live
in the land
of Egypt
will
escape
or survive
to return
to the land
of Judah.
Though they
long
to return
and live
there,
none
of them shall return
except
a few fugitives.’ ”
\par }{\PP \VS{15}Then all
the men
who were aware
that
their wives
were sacrificing
to other
gods,
as well as all
their wives,
answered
Jeremiah.
There was a great
crowd
of them representing
all
the people
who lived
in northern
and southern
Egypt.
They answered,
\VS{16}“We will not
listen
to
what
you claim
the {\ND{Lord}}
has
spoken
to us!
\VS{17}Instead
we will do
everything
we vowed
we would do. We will sacrifice
and pour
out drink offerings
to the goddess
called the Queen
of Heaven
just
as we
and our ancestors,
our kings,
and our leaders
previously did
in the towns
of Judah
and in the streets
of Jerusalem.
For then we had plenty
of food,
were well-off,
and had no
troubles.
\VS{18}But ever since
we stopped
sacrificing
and pouring
out drink offerings
to the Queen
of Heaven,
we have been
in great need. Our people have died in wars
or of starvation.”
\VS{19}The women added, “We
did indeed
sacrifice
and pour out
drink offerings
to the Queen
of Heaven.
But
it was with the full knowledge and approval of our husbands
that we made
cakes
in her image
and poured out
drink offerings to her.”
\par }{\PP \VS{20}Then Jeremiah
replied
to
all
the people,
both men
and women,
who responded
to him in this way.
\VS{21}“The

{\ND{Lord}}
did indeed remember
and call to mind
what you did! He remembered the sacrifices
you
and your ancestors,
your kings,
your leaders,
and all the rest of the people
of the land
offered
to other gods in the towns
of Judah
and in the streets
of Jerusalem.
\VS{22}Finally the
{\ND{Lord}}
could
no
longer
endure
your wicked
deeds
and the disgusting things
you did.
That
is why your land
has become the desolate,
uninhabited
ruin
that it is today.
That
is why it has become a proverbial example used in curses.
\VS{23}You have
sacrificed to other
gods! You have sinned
against the
{\ND{Lord}}! You have not
obeyed
the
{\ND{Lord}}! You have not
followed
his laws,
his statutes,
and his decrees! That is why
this disaster
that is evident
to this
day has happened to you.”
\par }{\PP \VS{24}Then Jeremiah
spoke to
all
the people,
particularly to
all
the women. “Listen
to what
the {\ND{Lord}}
has to say all
you people of Judah
who
are in Egypt.
\VS{25}The

{\ND{Lord}}
God
of Israel
who rules over all
says,
‘You
women
have confirmed
by your actions what
you vowed
with
your
lips! You said,
“We will certainly carry
out our vows
to sacrifice
and pour
out drink offerings
to the Queen
of Heaven.”
Well, then
fulfill
your vows! Carry them out!’
\VS{26}But
listen
to what
the {\ND{Lord}}
has to say, all
you people
of Judah
who are living in the land
of Egypt.
The
{\ND{Lord}}
says, ‘I
hereby swear
by my own great
name
that none
of the people of Judah
who are living
anywhere
in Egypt
will
ever again
invoke
my name
in their oaths! Never again will any of them use it in an oath saying,
“As surely
as the Lord

{\ND{God}} lives….”
\VS{27}I will indeed
see
to it that disaster,
not
prosperity,
happens to them. All
the people
of Judah
who
are in the land
of Egypt
will die
in war
or from starvation
until
not one of them is left.
\VS{28}Some who survive
in battle
will return
to the land
of Judah
from
the land
of Egypt.
But they will be very few
indeed! Then the Judean
remnant
who have come
to live
in the land
of Egypt
will know
whose
word
proves
true, mine or theirs.’
\VS{29}Moreover the
{\ND{Lord}}
says, ‘I will make something happen to prove that
I
will punish
you in
this
place.
I will do it so that
you will know
that
my threats
to bring
disaster on you will prove true.
\VS{30}I, the
{\ND{Lord}}, promise
that I will hand
Pharaoh Hophra
king
of Egypt
over
to his enemies
who are seeking
to kill
him. I will do that just
as surely as I handed
King
Zedekiah
of Judah
over
to King
Nebuchadnezzar
of Babylon,
his enemy
who was seeking
to kill him.’ ”

\par }\Chap{45}{\PP \VerseOne{1}The prophet
Jeremiah
spoke
to
Baruch
son
of Neriah
while he was writing down in
a scroll
the words
that
Jeremiah
spoke
to him. This
happened in the fourth
year
that Jehoiakim
son
of Josiah
was ruling
over Judah.
\VS{2}“The

{\ND{Lord}}
God
of Israel
has a message for you, Baruch.
\VS{3}‘You have said,
“I feel so hopeless! For
the {\ND{Lord}}
has added
sorrow
to my suffering.
I am worn out
from groaning.
I can’t
find
any rest.” ’ ”
\par }{\PP \VS{4}The
{\ND{Lord}} told
Jeremiah, “Tell
Baruch, ‘The
{\ND{Lord}}
says, “I
am
about
to tear
down what I have
built
and to uproot
what
I have planted.
I
will do this throughout the whole
earth.
\VS{5}Are you
looking
for great
things for yourself? Do not
look
for
such things. For
I, the
{\ND{Lord}}, affirm
that I am
about to bring
disaster
on
all
humanity.
But I will allow
you to escape
with your life
wherever
you go.” ’ ”

\par }\Chap{46}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
Jeremiah
about the nations.
\par }{\SH The Prophecy about Egypt’s Defeat at Carchemish
\par }{\PP \VS{2}He spoke about
Egypt
and the army
of Pharaoh Necho
king
of Egypt
which
was encamped
along
the Euphrates
River
at Carchemish.
Now this was the army that
King
Nebuchadnezzar
of Babylon
defeated
in the fourth
year
that Jehoiakim
son
of Josiah
was ruling
over Judah.
\par }{\Q \VS{3}“Fall into ranks
with your shields
ready!

\par }{\Q Prepare to march into battle!
\par }{\Q \VS{4}Harness
the horses
to the chariots!
\par }{\Q Mount your horses!
\par }{\Q Put on your helmets
and take
your positions!
\par }{\Q Sharpen
you spears!
\par }{\Q Put on
your armor!
\par }{\Q \VS{5}What do
I see?” says
the {\ND{Lord}}.

\par }{\Q “The soldiers
are terrified.
\par }{\Q They
are retreating.
\par }{\Q They have been defeated.
\par }{\Q They are overcome with terror;
\par }{\Q they desert quickly
\par }{\Q without
looking back.
\par }{\Q \VS{6}But even
the swiftest
cannot get away.
\par }{\Q Even
the strongest
cannot escape.
\par }{\Q There in the north
by
the Euphrates
River
\par }{\Q they stumble
and fall in defeat.
\par }{\Q \VS{7}“Who
is this
that rises
like the Nile,
\par }{\Q like its streams
turbulent
at flood stage?
\par }{\Q \VS{8}Egypt
rises
like the Nile,
\par }{\Q like its streams
turbulent
at flood
stage.
\par }{\Q Egypt says,
‘I will arise
and cover
the earth.
\par }{\Q I will destroy
cities
and the people who inhabit them.’
\par }{\Q \VS{9}Go ahead and charge into battle,
you horsemen!
\par }{\Q Drive furiously,
you charioteers!
\par }{\Q Let the soldiers
march out
into battle,
\par }{\Q those from Ethiopia
and Libya
who carry
shields,
\par }{\Q and those from Lydia
who are armed
with the bow.
\par }{\Q \VS{10}But that
day
belongs to the Lord
God
who rules over all.
\par }{\Q It is the day
when he will pay back
his enemies.
\par }{\Q His sword
will devour
them until its appetite is satisfied!
\par }{\Q It will drink
their blood
until
it is full!

\par }{\Q For
the Lord

{\ND{God}}
who rules over all
will offer
them up as a sacrifice
\par }{\Q in the land
of the north
by
the Euphrates
River.
\par }{\Q \VS{11}Go up
to Gilead
and get
medicinal ointment,
\par }{\Q you dear poor people
of Egypt.
\par }{\Q But it will prove useless
no matter how much
medicine
you use;

\par }{\Q there will be no
healing for you.
\par }{\Q \VS{12}The nations
will hear
of your devastating defeat.
\par }{\Q your cries of distress
will echo throughout
the earth.
\par }{\Q In the panic of their flight one soldier
will trip over
another
\par }{\Q and both
of them will fall down defeated.”
\par }{\SH The Lord Predicts that Nebuchadnezzar Will Attack and Plunder Egypt
\par }{\PP \VS{13}The
{\ND{Lord}}
spoke
to
the prophet
Jeremiah
about Nebuchadnezzar
coming
to attack
the land
of Egypt.
\par }{\Q \VS{14}“Make an announcement
throughout Egypt.
\par }{\Q Proclaim
it in Migdol,
Memphis,
and Tahpanhes.
\par }{\Q ‘Take your positions
and prepare
to do battle.
\par }{\Q For
the enemy army
is destroying
all the nations around you.’
\par }{\Q \VS{15}Why
will your soldiers be defeated?

\par }{\Q They will not
stand
because
I, the
{\ND{Lord}}, will thrust them down.
\par }{\Q \VS{16}I will make many
stumble.
\par }{\Q They will fall
over one
another
in their hurry
to flee.
\par }{\Q They will say,
‘Get
up!
\par }{\Q Let’s
go back
to our own
people.
\par }{\Q Let’s go
back to
our
homelands
\par }{\Q because the enemy
is coming
to
destroy
us.’
\par }{\Q \VS{17}There
at
home they will say, ‘Pharaoh
king
of Egypt
is just a big noise!

\par }{\Q He has let the most opportune moment pass by.’
\par }{\Q \VS{18}I the King,
whose name
is the
{\ND{Lord}}
who rules over all,
swear this:
\par }{\Q I swear as surely as I
live
that
a conqueror is coming.
\par }{\Q He will be as imposing
as Mount Tabor
is among the mountains,
\par }{\Q as Mount Carmel
is against the backdrop
of the sea.
\par }{\Q \VS{19}Pack your bags
for exile,
\par }{\Q you inhabitants
of poor dear
Egypt.
\par }{\Q For
Memphis
will be laid waste.
\par }{\Q It will lie in ruins
and be uninhabited.
\par }{\Q \VS{20}Egypt
is like a beautiful
young cow.
\par }{\Q But northern
armies will attack
her like swarms of stinging flies.
\par }{\Q \VS{21}Even
her mercenaries
\par }{\Q will prove to be like pampered, well-fed
calves.
\par }{\Q For
they
too
will turn
and run away.
\par }{\Q They will not
stand
their ground
\par }{\Q when the time
for them to be destroyed
comes,
\par }{\Q the time
for them to be punished.
\par }{\Q \VS{22}Egypt will run away,
hissing like a snake,
\par }{\Q as the enemy comes marching up in force.
\par }{\Q They will come
against her with axes
\par }{\Q as if
they were
woodsmen
chopping
down trees.
\par }{\Q \VS{23}The population of Egypt is like a vast, impenetrable
forest.
\par }{\Q But
I, the
{\ND{Lord}}, affirm
that
the enemy will cut them
down.
\par }{\Q For
those who chop them
down will be more numerous
than locusts.
\par }{\Q They
will be too numerous to count.
\par }{\Q \VS{24}Poor dear
Egypt
will be put
to shame.
\par }{\Q She will be handed
over to the people
from the north.”
\par }{\PP \VS{25}The
{\ND{Lord}}
God
of Israel
who rules over all
says,
“I
will punish
Amon,
the god of Thebes.
I will punish Egypt,
its gods,
and its kings.
I will punish Pharaoh
and all who trust in him.
\VS{26}I will hand
them over
to Nebuchadnezzar
and his troops,
who want
to kill
them. But later
on, people will live
in Egypt again as they did in former
times.
I, the
{\ND{Lord}}, affirm it!”
\par }{\SH A Promise of Hope for Israel
\par }{\Q \VS{27}“You
descendants of Jacob,
my servants,
do not
be afraid;
\par }{\Q do not
be terrified,
people of Israel.
\par }{\Q For
I
will rescue
you and your descendants
\par }{\Q from
the faraway lands where
you are captives.
\par }{\Q The descendants
of Jacob
will return
to their land
and enjoy peace.
\par }{\Q They will be secure
and no
one will terrify them.
\par }{\Q \VS{28}I, the
{\ND{Lord}}, tell you
not
to be afraid,
\par }{\Q you descendants of Jacob,
my servant,
\par }{\Q for
I am
with
you.
\par }{\Q Though
I completely
destroy all
the nations
where
I scatter
you,
\par }{\Q I will not
completely
destroy you.
\par }{\Q I will indeed discipline
you but only in due
measure.
\par }{\Q I will not
allow you to go entirely unpunished.”


\par }\Chap{47}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
to
the prophet
Jeremiah
about the Philistines
before
Pharaoh
attacked
Gaza.
\par }{\Q \VS{2}“Look! Enemies are gathering in the north
like water
rising
in a river.

\par }{\Q They will be
like an overflowing
stream.
\par }{\Q They will overwhelm
the whole country
and everything
in it like a flood.
\par }{\Q They will overwhelm the cities
and their inhabitants.
\par }{\Q People
will cry
out in alarm.
\par }{\Q Everyone
living
in the country
will cry out in pain.
\par }{\Q \VS{3}Fathers
will hear
the hoofbeats
of the enemies’ horses,
\par }{\Q the clatter
of their chariots
and the rumbling
of their wheels.
\par }{\Q They will not
turn back
to
save their children
\par }{\Q because they will be paralyzed
with fear.
\par }{\Q \VS{4}For
the time
has come
\par }{\Q to destroy
all
the Philistines.
\par }{\Q The time has come to destroy
all
the help
\par }{\Q that remains
for Tyre
\par }{\Q For
I, the
{\ND{Lord}}, will destroy
the Philistines,
\par }{\Q that remnant
that came from the island
of Crete.
\par }{\Q \VS{5}The people of Gaza
will shave
their heads in mourning.
\par }{\Q The people of Ashkelon
will be struck dumb.
\par }{\Q How
long
will you gash
yourselves to show your sorrow,

\par }{\Q you who remain
of Philistia’s power?
\par }{\Q \VS{6}How long
will you cry out, ‘Oh,
sword
of the
{\ND{Lord}},
\par }{\Q how long
will it be before you stop killing?

\par }{\Q Go back
into
your sheath!
\par }{\Q Stay
there and rest!’
\par }{\Q \VS{7}But how
can it rest
\par }{\Q when I, the
{\ND{Lord}}, have given it orders?
\par }{\Q I have ordered
it to
attack
\par }{\Q the people of Ashkelon
and the seacoast.

\par }\Chap{48}{\PP \VerseOne{1}The

{\ND{Lord}}
God
of Israel
who rules over all
spoke
about Moab.
\par }{\Q “Sure
to
be judged
is Nebo! Indeed,
it will be destroyed!
\par }{\Q Kiriathaim
will suffer disgrace.
It will be captured!
\par }{\Q Its fortress will
suffer disgrace. It will be torn down!
\par }{\Q \VS{2}People will not
praise
Moab
any more.
\par }{\Q The enemy will capture Heshbon
and plot
how to destroy
Moab,

\par }{\Q saying, ‘Come, let’s put an end to that nation!’
\par }{\Q City of Madmen,
you will also
be destroyed.

\par }{\Q A destructive army will march against you.
\par }{\Q \VS{3}Cries
of anguish
will arise in Horonaim,
\par }{\Q ‘Oh, the ruin
and great
destruction!’
\par }{\Q \VS{4}“Moab
will be crushed.
\par }{\Q Her children
will cry
out in distress.
\par }{\Q \VS{5}Indeed
they will climb the slopes
of Luhith,
\par }{\Q weeping
continually as they go.
\par }{\Q For
on the road down
to Horonaim
\par }{\Q they will hear
the cries
of distress
over the destruction.
\par }{\Q \VS{6}They will hear, ‘Run! Save
yourselves!
\par }{\Q Even if you must be
like a lonely shrub
in the desert!’
\par }{\Q \VS{7}“Moab, you trust
in the things
you do and in your riches.
\par }{\Q So you too
will be
conquered.
\par }{\Q Your god Chemosh
will go into exile
\par }{\Q along with his priests
and his officials.
\par }{\Q \VS{8}The destroyer
will come
against
every
town.
\par }{\Q Not
one town
will escape.
\par }{\Q The towns in the valley
will be destroyed.
\par }{\Q The cities on the high plain
will be laid waste.
\par }{\Q I, the
{\ND{Lord}}, have spoken!
\par }{\Q \VS{9}Set up
a gravestone
for Moab,
\par }{\Q for
it will certainly be laid in ruins!

\par }{\Q Its cities
will be laid waste
\par }{\Q and become
uninhabited.”
\par }{\Q \VS{10}A curse
on anyone who is lax
in doing
the
{\ND{Lord}}’s
work!
\par }{\Q A curse
on anyone who keeps from
carrying out his destruction!
\par }{\Q \VS{11}“From its earliest
days Moab
has lived undisturbed.
\par }{\Q It has never been taken into exile.
\par }{\Q Its
people are like
wine allowed to settle undisturbed
on
its
dregs,
\par }{\Q never
poured out
from one jar to another.
\par }{\Q They are like
wine which
tastes like
it always
did,
\par }{\Q whose
aroma
has remained unchanged.
\par }{\Q \VS{12}But
the time
is
coming
when I will send
\par }{\Q men against Moab who
will empty it out.
\par }{\Q They will empty
the towns of their people,
\par }{\Q then will
lay those towns in ruins.

\par }{\Q I,
the {\ND{Lord}}, affirm it!
\par }{\Q \VS{13}The people of Moab
will be disappointed
by their god Chemosh.
\par }{\Q They will be as
disappointed
as the people of Israel
were
\par }{\Q when they put their trust
in the calf god at Bethel.
\par }{\Q \VS{14}How
can you men of Moab say,
‘We
are heroes,
\par }{\Q men
who are mighty
in battle?’
\par }{\Q \VS{15}Moab
will be destroyed.
Its towns
will be invaded.
\par }{\Q Its finest
young men
will be slaughtered.
\par }{\Q I, the King,
the {\ND{Lord}}
who rules over all,
affirm it!
\par }{\Q \VS{16}Moab’s
destruction
is at hand.
\par }{\Q Disaster
will come
on it quickly.
\par }{\Q \VS{17}Mourn
for that nation, all
you nations living around
it,
\par }{\Q all
of you nations that know
of its fame.
\par }{\Q Mourn and say,
‘Alas,
its powerful
influence
has been broken!
\par }{\Q Its glory
and power have been done away!’
\par }{\Q \VS{18}Come down
from your place of honor;
\par }{\Q sit
on the dry ground,
you who live
in Dibon.
\par }{\Q For
the one who will destroy
Moab
will attack
you;
\par }{\Q he will destroy
your fortifications.
\par }{\Q \VS{19}You who live
in Aroer,
\par }{\Q stand
by the road
and watch.
\par }{\Q Question
the man who is fleeing
and the woman who
is escaping.
\par }{\Q Ask
them, ‘What
has happened?’
\par }{\Q \VS{20}They will answer, ‘Moab
is disgraced,
for
it has fallen!
\par }{\Q Wail
and cry out
in mourning!
\par }{\Q Announce
along the Arnon
River
\par }{\Q that
Moab
has been destroyed.’
\par }{\PP \VS{21}“Judgment
will come
on
the cities on the high plain: on Holon,
Jahzah,
and Mephaath,
\VS{22}on
Dibon,
Nebo,
and Beth Diblathaim,
\VS{23}on
Kiriathaim,
Beth Gamul,
and Beth Meon,
\VS{24}on
Kerioth
and Bozrah.
It will come on
all
the towns
of Moab,
both far
and near.
\VS{25}Moab’s
might will be crushed.
Its
power
will be broken.
I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{26}“Moab
has vaunted
itself against
me.
\par }{\Q So make him drunk with the wine of my wrath
\par }{\Q until he splashes
around in his own vomit,
\par }{\Q until others treat him as
a laughingstock.
\par }{\Q \VS{27}For did not
you people of Moab laugh
at the people of Israel?
\par }{\Q Did you think that they were nothing but thieves,

\par }{\Q that
you shook your head in contempt
\par }{\Q every time you talked about them?
\par }{\Q \VS{28}Leave
your towns,
you inhabitants
of Moab.
\par }{\Q Go
and live
in the cliffs.
\par }{\Q Be
like a dove
that makes its nest
\par }{\Q high
on
the sides
of a ravine.
\par }{\Q \VS{29}I have heard
how proud
the people of Moab
are,
\par }{\Q I know how
haughty
they are.
\par }{\Q I have heard how arrogant,
proud,
and haughty
they are,
\par }{\Q what a high
opinion they have of themselves.
\par }{\Q \VS{30}I,
the {\ND{Lord}}, affirm
that
I know
how arrogant they are.
\par }{\Q But their pride is ill-founded.
\par }{\Q Their boastings will prove to be false.
\par }{\Q \VS{31}So
I will weep with sorrow
for
Moab.
\par }{\Q I will cry out
in sadness for all
of Moab.
\par }{\Q I will moan
for
the people
of Kir Heres.
\par }{\Q \VS{32}I will weep
for the grapevines
of Sibmah
\par }{\Q just like the town of Jazer
weeps
over them.

\par }{\Q Their branches
once spread
as far as
the Dead Sea.
\par }{\Q They reached
as far as the town of Jazer.
\par }{\Q The destroyer
will ravage
\par }{\Q her fig, date,
and grape crops.
\par }{\Q \VS{33}Joy
and gladness
will disappear
\par }{\Q from the fruitful
land
of Moab.
\par }{\Q I will stop
the flow of wine
from the winepresses.
\par }{\Q No
one will stomp on
the grapes there and shout for joy.

\par }{\Q The shouts
there will be shouts
of soldiers,
\par }{\Q not
the shouts of those making wine.
\par }{\Q \VS{34}Cries
of anguish
raised from Heshbon
and Elealeh
\par }{\Q will be sounded
as far as
Jahaz.
\par }{\Q They will be sounded
from Zoar
as far as
Horonaim
and Eglath Shelishiyah.
\par }{\Q For
even
the waters
of Nimrim
will be
dried up.
\par }{\Q \VS{35}I will put an end
in Moab
\par }{\Q to those who make offerings
at her places of worship.
\par }{\Q I will put an end to those who sacrifice
to other gods.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{36}So
my heart
moans
for Moab
\par }{\Q like a flute
playing a funeral song.
\par }{\Q Yes, like a flute
playing a funeral song,
\par }{\Q my heart
moans
for
the people
of Kir Heres.
\par }{\Q For
the wealth
they have gained
will perish.
\par }{\Q \VS{37}For
all
of them will shave their heads
in mourning.
\par }{\Q They will all
cut off
their beards
to show their sorrow.
\par }{\Q They will all
make gashes
in their hands.
\par }{\Q They will all
put on
sackcloth.
\par }{\Q \VS{38}On
all
the housetops
in Moab
\par }{\Q and in all
its public squares
\par }{\Q there will be nothing but mourning.
\par }{\Q For
I will break
Moab
like an unwanted
jar.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{39}Oh, how
shattered
Moab
will be!
\par }{\Q Oh, how her people will wail!
\par }{\Q Oh, how
she will turn away
in shame!
\par }{\Q Moab
will become an object of ridicule,
\par }{\Q a terrifying sight
to all
the nations that surround her.”
\par }{\Q \VS{40}For
the {\ND{Lord}}
says,
\par }{\Q “Look! Like an eagle
with outspread
wings
\par }{\Q a nation will swoop down on Moab.
\par }{\Q \VS{41}Her towns will be captured.
\par }{\Q Her fortresses
will be
taken.
\par }{\Q At that time
the soldiers
of Moab
will be frightened
\par }{\Q like a woman
in labor.
\par }{\Q \VS{42}Moab
will be destroyed
and no longer be a nation,
\par }{\Q because
she has vaunted
herself against
the {\ND{Lord}}.
\par }{\Q \VS{43}Terror,
pits,
and traps
are in store
\par }{\Q for the people who live
in Moab.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{44}Anyone who flees
at the sound of terror
\par }{\Q will fall
into
a pit.
\par }{\Q Anyone who climbs out of the pit
\par }{\Q will be caught
in a trap.
\par }{\Q For
the time
is coming
\par }{\Q when I will punish
the people of Moab.
\par }{\Q I, the
{\ND{Lord}}, affirm it!
\par }{\Q \VS{45}In the shadows
of the walls of Heshbon
\par }{\Q those trying to escape
will stand
helpless.
\par }{\Q For
a fire
will burst forth
from Heshbon.
\par }{\Q Flames
will shoot out from
the former territory of Sihon.
\par }{\Q They will burn
the foreheads
of the people
of Moab,
\par }{\Q the skulls
of those war-loving
people.
\par }{\Q \VS{46}Moab,
you are doomed!

\par }{\Q You people
who worship Chemosh
will be destroyed.
\par }{\Q Your sons
will be taken
away captive.
\par }{\Q Your daughters
will be carried away into exile.
\par }{\Q \VS{47}Yet in days
to come
\par }{\Q I will reverse
Moab’s
ill fortune.”
\par }{\Q says
the {\ND{Lord}}.
\par }{\PP The judgment
against Moab
ends
here.

\par }\Chap{49}{\PP \VerseOne{1}The

{\ND{Lord}}
spoke about the Ammonites.
\par }{\Q “Do you think
there are not
any people
of the nation
of Israel
remaining?
\par }{\Q Do you think there are not any of them remaining to reinherit their land?
\par }{\Q Is
that why
you people
who worship the god Milcom

\par }{\Q have taken possession
of the territory of Gad
and live
in his cities?
\par }{\Q \VS{2}Because
you did that,
\par }{\Q I, the
{\ND{Lord}}, affirm
that a time
is coming
\par }{\Q when I will make Rabbah,
the capital city of Ammon,
\par }{\Q hear
the sound
of the battle cry.
\par }{\Q It will become
a mound
covered
with ruins.

\par }{\Q Its villages
will be burned
to the ground.

\par }{\Q Then Israel
will take back
its land
\par }{\Q from those who took their land
from them.
\par }{\Q I, the Lord,
affirm it!
\par }{\Q \VS{3}Wail,
you people in Heshbon,
because
Ai
in Ammon is destroyed.
\par }{\Q Cry out
in anguish, you people in the villages
surrounding Rabbah.
\par }{\Q Put on
sackcloth
and cry out in mourning.
\par }{\Q Run about
covered with gashes.
\par }{\Q For
your god Milcom
will go into
exile
\par }{\Q along with
his priests
and officials.
\par }{\Q \VS{4}Why
do you brag about
your great power?
\par }{\Q Your power
is ebbing away,
you rebellious people of Ammon,
\par }{\Q who trust
in your riches
and say,
\par }{\Q ‘Who
would dare to
attack us?’
\par }{\Q \VS{5}I
will bring
terror
on
you from every
side,”
\par }{\Q says
the Lord

{\ND{God}}
who rules over all.
\par }{\Q “You will be scattered
in every direction.
\par }{\Q No
one will gather
the fugitives back together.
\par }{\Q \VS{6}Yet in days to come
\par }{\Q I will reverse
Ammon’s
ill fortune.”
\par }{\Q says
the {\ND{Lord}}.
\par }{\SH Judgment Against Edom
\par }{\PP \VS{7}The
{\ND{Lord}}
who rules over all
spoke
about Edom.
\par }{\Q “Is wisdom
no
longer
to be found in Teman?

\par }{\Q Can Edom’s counselors not give her any good
advice?

\par }{\Q Has all of their wisdom
turned bad?
\par }{\Q \VS{8}Turn
and flee! Take up refuge in remote places,
\par }{\Q you people who
live
in Dedan.
\par }{\Q For
I will bring
disaster
on the descendants of Esau.
\par }{\Q I have decided
it
is time
for me to punish them.
\par }{\Q \VS{9}If
grape pickers
came
to pick your grapes,
\par }{\Q would they not
leave
a few grapes behind?

\par }{\Q If
robbers
came at night,
\par }{\Q would they not pillage
only what they needed?
\par }{\Q \VS{10}But
I
will strip
everything away
from Esau’s
descendants.
\par }{\Q I will uncover
their hiding places
so they cannot
hide.
\par }{\Q Their children,
relatives,
and neighbors
will all be
destroyed.
\par }{\Q Not one of them will be left!
\par }{\Q \VS{11}Leave
your orphans
behind and I
will keep them alive.
\par }{\Q Your widows
too can depend
on me.”
\par }{\PP \VS{12}For
the {\ND{Lord}}
says,
“If
even those who
did not
deserve
to drink
from the cup
of my wrath must
drink
from it,
do you think you
will go unpunished? You will not
go unpunished,
but must certainly
drink from the cup of my wrath.
\VS{13}For
I solemnly swear,”
says
the {\ND{Lord}}, “that
Bozrah
will become a pile of ruins.
It will become an object of horror
and ridicule,
an example to be
used in curses.
All
the towns
around it will lie
in ruins
forever.”
\par }{\Q \VS{14}I said, “I have heard
a message
from the
{\ND{Lord}}.
\par }{\Q A messenger
has been sent
among the nations
to say,
\par }{\Q ‘Gather
your armies and march out
against
her!
\par }{\Q Prepare to do
battle with her!’ ”
\par }{\Q \VS{15}The
{\ND{Lord}} says to Edom,

\par }{\Q “I
will certainly
make you small
among nations.
\par }{\Q I will make you despised
by all humankind.
\par }{\Q \VS{16}The terror
you inspire in others

\par }{\Q and the arrogance
of your heart
have deceived
you.
\par }{\Q You may make your home
in the clefts
of the rocks;
\par }{\Q you may occupy
the highest
places in the hills.
\par }{\Q But even if
you made your home where the eagles
nest,
\par }{\Q I would bring
you down
from there,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{17}“Edom
will become
an object of horror.
\par }{\Q All
who pass by
it will be filled with horror;
\par }{\Q they will hiss out their scorn
\par }{\Q because
of all
the disasters that have happened to it.
\par }{\Q \VS{18}Edom will be destroyed
like Sodom
and Gomorrah
\par }{\Q and the towns that were around them.
\par }{\Q No
one will live
there.
\par }{\Q No
human
being will settle
in it,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{19}“A lion
coming up
from the thick undergrowth
along the Jordan
\par }{\Q scatters the sheep in the pastureland
around it.

\par }{\Q So too
I will chase
the Edomites off
their land.

\par }{\Q Then I will appoint
over it whomever
I choose.
\par }{\Q For
there is no one
like me,
and there is no one
who can call
me to account.
\par }{\Q There is no
ruler
who
can stand up
against me.
\par }{\Q \VS{20}So
listen
to what I, the
{\ND{Lord}}, have
planned
against Edom,
\par }{\Q what
I intend
to do
to
the people who live
in Teman.
\par }{\Q Their little ones will be dragged
off.
\par }{\Q I will completely destroy their land
because
of what they have done.
\par }{\Q \VS{21}The people of the earth
will quake
when they hear
of their downfall.
\par }{\Q Their cries
of anguish
will be heard
all the way to the Gulf
of Aqaba.
\par }{\Q \VS{22}Look! Like an eagle
with outspread
wings,
\par }{\Q a nation will soar up and swoop down on
Bozrah.
\par }{\Q At that time
the soldiers
of Edom
will be
as fearful
\par }{\Q as a woman
in labor.”
\par }{\SH Judgment Against Damascus
\par }{\PP \VS{23}The
{\ND{Lord}} spoke about Damascus.
\par }{\Q “The people of Hamath
and Arpad
will be dismayed
\par }{\Q because
they have heard
bad
news.
\par }{\Q Their courage will melt
away because of worry.
\par }{\Q Their hearts will not
be able
to rest.
\par }{\Q \VS{24}The people of Damascus
will lose heart
and turn
to flee.
\par }{\Q Panic
will grip
them.
\par }{\Q Pain and anguish
will seize
them
\par }{\Q like a woman in labor.
\par }{\Q \VS{25}How
deserted
will that once-famous
city
be,
\par }{\Q that city
that was once filled with joy!
\par }{\Q \VS{26}For her young men
will fall
in her city squares.
\par }{\Q All
her soldiers
will be destroyed
at that time,”
\par }{\Q says
the {\ND{Lord}}
who rules over all.
\par }{\Q \VS{27}“I will set
fire
to the walls
of Damascus;
\par }{\Q it will burn up
the palaces
of Ben Hadad.”
\par }{\SH Judgment Against Kedar and Hazor
\par }{\PP \VS{28}The
{\ND{Lord}}
spoke
about
Kedar
and the kingdoms
of Hazor
that
King
Nebuchadnezzar
of Babylon conquered.
\par }{\Q “Army of Babylon,
go
and attack
Kedar.
\par }{\Q Lay waste
those who live in
the eastern desert.
\par }{\Q \VS{29}Their tents
and their flocks
will be taken
away.
\par }{\Q Their tent curtains,
equipment,
and camels
will be carried
off.
\par }{\Q People will shout
to them,
\par }{\Q ‘Terror
is all around you!’ ”
\par }{\Q \VS{30}The
{\ND{Lord}}
says, “Flee
quickly,
you who live
in
Hazor.
\par }{\Q Take up refuge in
remote places.
\par }{\Q For
King
Nebuchadnezzar
of Babylon
has laid out plans
to attack you.
\par }{\Q He has formed his strategy
on
how to defeat you.”
\par }{\Q \VS{31}The
{\ND{Lord}}
says, “Army of Babylon,
go
and attack
\par }{\Q a nation
that lives in
peace and security.
\par }{\Q They have no
gates
or
walls
to protect them.

\par }{\Q They live
all alone.
\par }{\Q \VS{32}Their camels
will be
taken as plunder.
\par }{\Q Their vast
herds
will be taken as spoil.
\par }{\Q I will scatter
to the four winds
\par }{\Q those desert peoples who cut their hair short
at the temples.
\par }{\Q I will bring
disaster
against them
\par }{\Q from
every
direction,” says
the {\ND{Lord}}.
\par }{\Q \VS{33}“Hazor
will become
a permanent
wasteland,
\par }{\Q a place where only
jackals
live.
\par }{\Q No
one
will live
there.
\par }{\Q No
human being
will settle in it.”
\par }{\SH Judgment Against Elam
\par }{\PP \VS{34}Early
in the reign
of King
Zedekiah
of Judah,
the {\ND{Lord}}
spoke
to
the prophet
Jeremiah
about
Elam.
\par }{\PP \VS{35}The

{\ND{Lord}}
who rules over all
said,
\par }{\Q “I
will kill
all the archers
of Elam,
\par }{\Q who are the chief source
of her military might.
\par }{\Q \VS{36}I will cause
enemies to blow through Elam
from every direction
\par }{\Q like the winds
blowing in from the four
quarters
of heaven.
\par }{\Q I will scatter
the people
of Elam to the four
winds.
\par }{\Q There will not
be
any nation
where
the refugees
of Elam
will not
go.
\par }{\Q \VS{37}I will make the
people of Elam
terrified
of their enemies,
\par }{\Q who are seeking
to kill
them.
\par }{\Q I will vent my
fierce
anger
\par }{\Q and bring
disaster
upon them,”
says
the {\ND{Lord}}.

\par }{\Q “I will send
armies chasing
after them

\par }{\Q until
I have completely destroyed them.
\par }{\Q \VS{38}I will establish
my sovereignty
over Elam.
\par }{\Q I will destroy
their king
and their leaders,”
says
the {\ND{Lord}}.
\par }{\Q \VS{39}“Yet in days
to come
\par }{\Q I will reverse
Elam’s
ill fortune.”
\par }{\Q says
the {\ND{Lord}}.

\par }\Chap{50}{\PP \VerseOne{1}The
{\ND{Lord}}
spoke
concerning Babylon
and the land
of Babylonia
through
the prophet
Jeremiah.
\par }{\Q \VS{2}“Announce
the news among the nations! Proclaim
it!
\par }{\Q Signal
for people to pay attention!

\par }{\Q Declare the news! Do not
hide
it! Say:
\par }{\Q ‘Babylon
will be captured.
\par }{\Q Bel
will be
put to shame.
\par }{\Q Marduk
will be dismayed.
\par }{\Q Babylon’s idols
will be put to shame.
\par }{\Q Her disgusting images
will be dismayed.
\par }{\Q \VS{3}For
a nation
from the north
will attack
Babylon.
\par }{\Q It will lay
her land
waste.
\par }{\Q People
and animals
will flee
out of it.
\par }{\Q No
one will inhabit it.’
\par }{\Q \VS{4}“When
that time
comes,”
says
the
{\ND{Lord}},

\par }{\Q “the people of
Israel
and Judah
will return
to the land together.
\par }{\Q They will come back
with tears
of repentance
\par }{\Q as they seek
the
{\ND{Lord}}
their God.
\par }{\Q \VS{5}They will ask
the way
to Zion;
\par }{\Q they will turn
their faces
toward it.
\par }{\Q They will come
and bind
themselves to
the {\ND{Lord}}
\par }{\Q in a lasting
covenant
that will never
be forgotten.
\par }{\Q \VS{6}“My people
have been
lost
sheep.
\par }{\Q Their shepherds
have allowed them to go astray.
\par }{\Q They have wandered around
in the mountains.
\par }{\Q They have roamed from
one mountain
and hill
to
another.

\par }{\Q They have forgotten
their resting place.
\par }{\Q \VS{7}All
who encountered
them devoured
them.
\par }{\Q Their enemies
who did this said,
‘We are not
liable
for punishment!
\par }{\Q For those
people have sinned
against the
{\ND{Lord}}, their true pasture.
\par }{\Q They have sinned against the
{\ND{Lord}}
in whom their ancestors
trusted.’
\par }{\Q \VS{8}“People of Judah,
get
out of Babylon
quickly!
\par }{\Q Leave the land
of Babylonia!

\par }{\Q Be
the first to depart!

\par }{\Q Be like
the male goats
that lead
the herd.
\par }{\Q \VS{9}For
I
will rouse
into action
and bring
against
Babylon
\par }{\Q a host
of mighty
nations
from the land
of the north.
\par }{\Q They will set up their battle lines
against her.
\par }{\Q They will come from
the north
and capture
her.

\par }{\Q Their arrows
will be like a skilled soldier
\par }{\Q who
does not
return
from the battle empty-handed.
\par }{\Q \VS{10}Babylonia
will be
plundered.
\par }{\Q Those who plunder
it will take all
they want,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{11}“People of Babylonia, you plundered
my people.
\par }{\Q That
made you happy
and glad.
\par }{\Q You frolic
about like calves
in a pasture.
\par }{\Q Your joyous sounds are like the neighs
of a stallion.
\par }{\Q \VS{12}But Babylonia
will be put to great
shame.
\par }{\Q The land where you were born
will be disgraced.
\par }{\Q Indeed,
Babylonia will become the least important of all nations.
\par }{\Q It will become a dry
and barren
desert.
\par }{\Q \VS{13}After I vent my wrath
on it Babylon
will be uninhabited.
\par }{\Q It will be
totally
desolate.
\par }{\Q All
who pass by
will be filled with horror
and will hiss
out their scorn
\par }{\Q because
of all
the disasters that have happened to it.
\par }{\Q \VS{14}“Take up your battle positions
all around
Babylon,
\par }{\Q all
you soldiers
who are armed with bows.
\par }{\Q Shoot
all your arrows
at her! Do not
hold
any back!

\par }{\Q For
she has sinned
against the
{\ND{Lord}}.
\par }{\Q \VS{15}Shout
the battle cry from all around
the city.
\par }{\Q She will
throw up her hands
in surrender.

\par }{\Q Her towers
will fall.
\par }{\Q Her walls
will be torn down.
\par }{\Q Because
I, the
{\ND{Lord}}, am wreaking revenge,
\par }{\Q take out your vengeance
on her!
\par }{\Q Do
to her as
she
has done!
\par }{\Q \VS{16}Kill
all the farmers who sow
the seed in the land
of Babylon.
\par }{\Q Kill all those who wield
the sickle
at harvest
time.
\par }{\Q Let all the foreigners return to
their own
people.
\par }{\Q Let them hurry back
to
their own
lands
\par }{\Q to
escape
destruction by
that enemy army.
\par }{\Q \VS{17}“The people of Israel
are like scattered
sheep
\par }{\Q which lions
have chased away.
\par }{\Q First
the king
of Assyria
devoured
them.

\par }{\Q Now
last
of all King
Nebuchadnezzar
of Babylon
has gnawed their bones.
\par }{\Q \VS{18}So
I,
the {\ND{Lord}}
God
of Israel
who rules over all,
say:

\par }{\Q ‘I
will punish
the king
of Babylon
and his land
\par }{\Q just
as I punished
the king
of Assyria.
\par }{\Q \VS{19}But
I will restore
the flock of Israel
to
their own pasture.
\par }{\Q They will graze
on Mount Carmel
and the land of Bashan.
\par }{\Q They will eat until they are full
\par }{\Q on the hills
of Ephraim
and the land of Gilead.
\par }{\Q \VS{20}When
that time
comes,
\par }{\Q no
guilt
will be found
in Israel.
\par }{\Q No
sin
will be found
in Judah.
\par }{\Q For
I will forgive
those
of them I have allowed to survive.
\par }{\Q I, the
{\ND{Lord}}, affirm it!’ ”
\par }{\Q \VS{21}The
{\ND{Lord}}
says,
\par }{\Q “Attack
the land
of Merathaim
\par }{\Q and the people who live
in Pekod!

\par }{\Q Pursue,
kill, and completely destroy them!

\par }{\Q Do
just
as I have commanded you!
\par }{\Q \VS{22}The noise
of battle
can be heard in the land
of Babylonia.

\par }{\Q There is the sound of great
destruction.
\par }{\Q \VS{23}Babylon hammered
the whole
world
to pieces.
\par }{\Q But see how that ‘hammer’
has been broken and shattered!

\par }{\Q See what an object of horror
\par }{\Q Babylon
has become among the nations!
\par }{\Q \VS{24}I set a trap
for you, Babylon;
\par }{\Q you
were caught
before you
knew
it.
\par }{\Q You fought
against me.
\par }{\Q So you were found
and captured.
\par }{\Q \VS{25}I have opened
up the
place where my weapons are stored.
\par }{\Q I have brought out
the
weapons
for carrying out my wrath.
\par }{\Q For
I, the Lord

{\ND{God}}
who rules over all,
\par }{\Q have work
to carry out in the land
of Babylonia.
\par }{\Q \VS{26}Come
from far away and attack Babylonia!

\par }{\Q Open
up the places where she stores
her grain!
\par }{\Q Pile her up
in ruins! Destroy
her completely!

\par }{\Q Do not
leave anyone alive!
\par }{\Q \VS{27}Kill
all
her soldiers!

\par }{\Q Let
them be slaughtered!

\par }{\Q They are doomed,
for
their day
of reckoning has come,
\par }{\Q the time
for them to be punished.”
\par }{\Q \VS{28}Listen! Fugitives
and refugees
are coming from the land
of Babylon.
\par }{\Q They are coming to Zion
to declare
there
\par }{\Q how the
{\ND{Lord}}
our God
is getting revenge,
\par }{\Q getting revenge
for what they have done to his temple.
\par }{\Q \VS{29}“Call
for archers
to
come against Babylon!
\par }{\Q Summon against her all
who draw
the bow!
\par }{\Q Set up camp
all around
the city!
\par }{\Q Do not
allow anyone
to escape!
\par }{\Q Pay
her back for what she has done.
\par }{\Q Do
to her what she has done
to others.
\par }{\Q For
she has proudly
defied me,
\par }{\Q the Holy One
of Israel.
\par }{\Q \VS{30}So
her young men
will fall
in her city squares.
\par }{\Q All
her soldiers
will be destroyed
at that time,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{31}“Listen! I am opposed
to
you, you proud
city,”

\par }{\Q says
the Lord

{\ND{God}}
who rules over all.
\par }{\Q “Indeed,
your day
of reckoning has come,
\par }{\Q the time
when I will punish you.
\par }{\Q \VS{32}You will stumble
and fall,
you proud
city;
\par }{\Q no
one will help you get up.
\par }{\Q I will set
fire
to your towns;
\par }{\Q it will burn up
everything
that surrounds you.”
\par }{\Q \VS{33}The

{\ND{Lord}}
who rules over all
says,
\par }{\Q “The people of Israel
are oppressed.
\par }{\Q So too
are the people of Judah.
\par }{\Q All
those who took them captive
are holding
them prisoners.
\par }{\Q They refuse
to set them free.
\par }{\Q \VS{34}But
the one who
will rescue
them is strong.
\par }{\Q He is known as the
{\ND{Lord}}
who rules over all.
\par }{\Q He will strongly champion
their cause.
\par }{\Q As a result
he will bring peace and rest
to the
earth,
\par }{\Q but trouble and turmoil
to the people who inhabit
Babylonia.
\par }{\Q \VS{35}“Destructive
forces will come against
the Babylonians,”
says
the {\ND{Lord}}.

\par }{\Q “They will come against the people
who inhabit
Babylonia,
\par }{\Q against her
leaders
and her
men of wisdom.
\par }{\Q \VS{36}Destructive
forces
will come against her false prophets;

\par }{\Q they will be shown to be fools!

\par }{\Q Destructive
forces
will come against her
soldiers;
\par }{\Q they will be
filled with terror!
\par }{\Q \VS{37}Destructive
forces
will come against her
horses
and her chariots.
\par }{\Q Destructive forces
will come against all
the foreign troops within
her;

\par }{\Q they will be
as frightened as women!

\par }{\Q Destructive
forces
will come
against her treasures;
\par }{\Q they will be taken away as plunder!
\par }{\Q \VS{38}A drought
will come upon her land;
\par }{\Q her rivers and canals
will be dried
up.

\par }{\Q All of this will happen because
her land
is filled with idols.
\par }{\Q Her people act like madmen because of those idols they fear.
\par }{\Q \VS{39}Therefore
desert creatures
and jackals
will live
there.
\par }{\Q Ostriches
will dwell
in it too.

\par }{\Q But no people
will ever
live
there again.
\par }{\Q No one
will dwell
there for
all time to come.
\par }{\Q \VS{40}I will destroy
Babylonia just like I did
\par }{\Q Sodom
and Gomorrah
and the neighboring towns.
\par }{\Q No
one will live
there.
\par }{\Q No
human
being will settle
in it,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{41}“Look! An army
is about to come
from the north.
\par }{\Q A mighty nation
and many
kings
are stirring
into action
\par }{\Q in faraway parts
of the earth.
\par }{\Q \VS{42}Its soldiers are armed
with bows
and spears.
\par }{\Q They are
cruel
and show no
mercy.
\par }{\Q They sound
like the roaring
sea
\par }{\Q as they ride
forth on
their
horses.
\par }{\Q Lined
up
in formation
like men
going
into battle,
\par }{\Q they are coming
against
you, fair
Babylon!
\par }{\Q \VS{43}The king
of Babylon
will become paralyzed
with fear

\par }{\Q when he hears
news
of their coming.
\par }{\Q Anguish
will grip
him,
\par }{\Q agony
like that of a woman giving birth to a baby.
\par }{\Q \VS{44}“A lion
coming up
from the thick undergrowth
along the Jordan
\par }{\Q scatters the sheep in the pastureland
around it.
\par }{\Q So too
I will chase
the Babylonians off
of their land.
\par }{\Q Then I will appoint
over it whomever
I choose.
\par }{\Q For
there is no one
like me.
\par }{\Q There is no one
who can call
me to account.
\par }{\Q There is no
ruler
that
can stand up
against me.
\par }{\Q \VS{45}So
listen
to what I, the
{\ND{Lord}}, have
planned
against
Babylon,
\par }{\Q what
I intend
to do
to
the people
who inhabit the land
of Babylonia.
\par }{\Q Their little ones will be dragged
off.
\par }{\Q I will
completely destroy their land
because
of what they have done.
\par }{\Q \VS{46}The people of the earth
will quake
when they hear Babylon
has been captured.
\par }{\Q Her cries
of anguish
will be heard
by the other nations.”

\par }\Chap{51}{\PP \par }{\Q \VerseOne{1}The
{\ND{Lord}}
says,
\par }{\Q “I
will cause
a destructive
wind
to blow
\par }{\Q against
Babylon
and the people who inhabit
Babylonia.
\par }{\Q \VS{2}I will send
people
to winnow Babylonia
like a wind blowing away
chaff.

\par }{\Q They will winnow
her and strip her land
bare.

\par }{\Q This
will happen when
they come against
her
from every
direction,
\par }{\Q when it
is time
to destroy
her.
\par }{\Q \VS{3}Do not give her archers time to string
their bows
\par }{\Q or to put on their coats
of armor.
\par }{\Q Do not spare
any of her young men.
\par }{\Q Completely
destroy her whole
army.
\par }{\Q \VS{4}Let them fall
slain
in the land
of Babylonia,
\par }{\Q mortally
wounded in the streets of her cities.
\par }{\Q \VS{5}“For
Israel
and Judah
will not
be forsaken
\par }{\Q by their God,
the {\ND{Lord}}
who rules over all.
\par }{\Q For
the land
of Babylonia is full
of guilt
\par }{\Q against the Holy One
of Israel.
\par }{\Q \VS{6}Get
out of Babylonia
quickly, you foreign people.
\par }{\Q Flee
to save
your lives.
\par }{\Q Do not
let yourselves be killed
because
of her sins.
\par }{\Q For
it is
time
for the
{\ND{Lord}}
to wreak
his revenge.
\par }{\Q He will pay
Babylonia back
for what she has done.
\par }{\Q \VS{7}Babylonia
had been a gold
cup
in the
{\ND{Lord}}’s
hand.
\par }{\Q She had made the whole
world
drunk.
\par }{\Q The nations
had drunk
from the wine
of her wrath.

\par }{\Q So
they
have all gone mad.
\par }{\Q \VS{8}But suddenly
Babylonia
will fall
and be destroyed.
\par }{\Q Cry out
in mourning
over
it!
\par }{\Q Get medicine
for her wounds!
\par }{\Q Perhaps
she can be healed!
\par }{\Q \VS{9}Foreigners living there
will say,

\par }{\Q ‘We tried to
heal
her, but
she could not
be healed.
\par }{\Q Let’s
leave
Babylonia
and each
go back to
his own country.
\par }{\Q For
judgment
on her
will be vast
in its proportions.
\par }{\Q It will be like it is piled up to
heaven,
stacked
up into the clouds.’
\par }{\Q \VS{10}The
exiles from
Judah will say,

\par }{\Q ‘The
{\ND{Lord}}
has brought about a great deliverance
for us!

\par }{\Q Come
on, let’s go and proclaim
in Zion
\par }{\Q what the
{\ND{Lord}}
our God
has done!’
\par }{\Q \VS{11}“Sharpen
your arrows!
\par }{\Q Fill
your quivers!

\par }{\Q The
{\ND{Lord}}
will arouse
a spirit
of hostility in the kings
of Media.
\par }{\Q For
he intends
to destroy
Babylonia.
\par }{\Q For
that is how the
{\ND{Lord}}
will get his revenge
–
\par }{\Q how he will get his revenge
for the Babylonians’ destruction of his temple.
\par }{\Q \VS{12}Give the signal to attack
Babylon’s
wall!

\par }{\Q Bring
more guards!

\par }{\Q Post
them all around the city!

\par }{\Q Put
men in ambush!

\par }{\Q For
the {\ND{Lord}}
will do what he has planned.
\par }{\Q He will do
what he said
he would do to
the people
of Babylon.
\par }{\Q \VS{13}“You who live
along the rivers
of Babylon,
\par }{\Q the time of your end
has come.
\par }{\Q You who are rich in plundered treasure,
\par }{\Q it is time for your lives to be cut off.
\par }{\Q \VS{14}The
{\ND{Lord}}
who rules over all
has solemnly sworn,
\par }{\Q ‘I will fill
your land with enemy soldiers.
\par }{\Q They will swarm
over
it like locusts.
\par }{\Q They will raise up shouts of victory over it.’
\par }{\Q \VS{15}He
is the one who by his power
made
the earth.
\par }{\Q He is the one who by his wisdom
fixed
the world
in place,
\par }{\Q by his understanding
he spread
out the heavens.
\par }{\Q \VS{16}When his voice
thunders,
the waters
in the heavens
roar.
\par }{\Q He makes the clouds rise from
the far-off horizons.
\par }{\Q He makes the lightning
flash out in the midst of the rain.
\par }{\Q He unleashes
the wind
from the places where he stores it.
\par }{\Q \VS{17}All
idolaters
will prove to be stupid and ignorant.
\par }{\Q Every
goldsmith
will be disgraced
by the idol
he made.
\par }{\Q For
the image
he forges is merely a sham.
\par }{\Q There is no
breath in any of those idols.
\par }{\Q \VS{18}They
are worthless,
objects
to be ridiculed.
\par }{\Q When the time
comes to punish
them, they will be destroyed.
\par }{\Q \VS{19}The
{\ND{Lord}}, who is the portion
of the descendants of Jacob,
is not
like them.
\par }{\Q For
he is the one who created
everything,
\par }{\Q including the people
of Israel whom he claims as his own.
\par }{\Q He is known
as the
{\ND{Lord}}
who rules over all.
\par }{\Q \VS{20}“Babylon, you
are my war club,
\par }{\Q my weapon
for battle.
\par }{\Q I used you to smash
nations.
\par }{\Q I used you to destroy
kingdoms.
\par }{\Q \VS{21}I used you to smash
horses
and their riders.
\par }{\Q I used you to smash
chariots
and their drivers.
\par }{\Q \VS{22}I used you to smash
men
and women.
\par }{\Q I used you to smash
old men
and young men.
\par }{\Q I used you to smash
young men
and young women.
\par }{\Q \VS{23}I used you to smash
shepherds
and their flocks.
\par }{\Q I used you to smash
farmers
and their teams
of oxen.
\par }{\Q I used you to smash
governors
and leaders.”
\par }{\Q \VS{24}“But I will repay
Babylon
\par }{\Q and all
who live
in Babylonia
\par }{\Q for all
the wicked
things they did
in Zion
\par }{\Q right before the eyes
of you Judeans,”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{25}The
{\ND{Lord}}
says, “Beware! I am opposed
to
you, Babylon!

\par }{\Q You are like a destructive
mountain
that destroys
all
the earth.
\par }{\Q I will unleash
my power
against
you;

\par }{\Q I will roll
you off the cliffs
and make
you like a burned-out
mountain.
\par }{\Q \VS{26}No
one
will use any
of your stones
as a cornerstone.
\par }{\Q No one
will use any
of them in the foundation
of his house.
\par }{\Q For
you will lie desolate
forever,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{27}“Raise
up battle flags
throughout the lands.
\par }{\Q Sound the trumpets
calling the nations
to do battle.
\par }{\Q Prepare
the nations
to do battle against Babylonia.
\par }{\Q Call for these kingdoms
to attack her:
\par }{\Q Ararat,
Minni,
and Ashkenaz.
\par }{\Q Appoint
a commander
to lead the attack.
\par }{\Q Send horses
against her like a swarm
of locusts.
\par }{\Q \VS{28}Prepare
the nations
to do battle against
her.

\par }{\Q Prepare the kings
of the Medes.
\par }{\Q Prepare their governors
and all
their leaders.
\par }{\Q Prepare all
the countries
they rule to do battle against her.
\par }{\Q \VS{29}The earth
will tremble
and writhe
in agony.

\par }{\Q For
the {\ND{Lord}}
will carry
out his plan.
\par }{\Q He plans
to make
the land
of Babylonia
\par }{\Q a wasteland
where no
one lives.
\par }{\Q \VS{30}The soldiers
of Babylonia
will stop
fighting.
\par }{\Q They will remain
in their fortified cities.
\par }{\Q They will lose
their strength
to do battle.

\par }{\Q They will be
as frightened as women.
\par }{\Q The houses in her cities
will be set on fire.
\par }{\Q The gates
of her cities will be broken down.
\par }{\Q \VS{31}One runner
after another will come
to the king of Babylon.
\par }{\Q One
messenger after another will come
bringing news.

\par }{\Q They
will bring news
to the king
of Babylon
\par }{\Q that
his whole city
has been captured.
\par }{\Q \VS{32}They will report that the fords
have been captured,
\par }{\Q the reed marshes
have been burned,
\par }{\Q the soldiers
are terrified.
\par }{\Q \VS{33}For
the

{\ND{Lord}}
God
of Israel
who rules over all
says,
\par }{\Q ‘Fair
Babylon
will
be like a
threshing floor
\par }{\Q which has been
trampled flat for harvest.
\par }{\Q The time
for her to be cut down and harvested
\par }{\Q will come
very soon.’
\par }{\Q \VS{34}“King
Nebuchadnezzar
of Babylon
\par }{\Q devoured
me and drove
my people out.
\par }{\Q Like a monster
from the deep he swallowed
me.
\par }{\Q He filled
his belly
with my riches.
\par }{\Q He made me an empty
dish.
\par }{\Q He completely cleaned
me out.”
\par }{\Q \VS{35}The person who lives in
Zion
says,
\par }{\Q “May Babylon
pay for the violence
done to me
and to
my relatives.”
\par }{\Q Jerusalem
says,
\par }{\Q “May those living
in Babylonia
pay for the bloodshed of my people.”
\par }{\Q \VS{36}Therefore
the {\ND{Lord}}
says,
\par }{\Q “I
will stand up
for your cause.
\par }{\Q I
will pay
the Babylonians back
for what they have done to you.

\par }{\Q I will dry
up their
sea.
\par }{\Q I will make their
springs
run
dry.
\par }{\Q \VS{37}Babylon
will become
a heap
of ruins.
\par }{\Q Jackals
will make their home
there.

\par }{\Q It will become an object of horror
and of hissing
scorn,
\par }{\Q a place where no
one lives.
\par }{\Q \VS{38}The Babylonians are all like lions
roaring
for prey.
\par }{\Q They are like lion
cubs
growling for something to eat.
\par }{\Q \VS{39}When their appetites
are all stirred up,

\par }{\Q I will set
out a banquet
for them.
\par }{\Q I will make them drunk
\par }{\Q so
that they will pass out,
\par }{\Q they will
fall asleep
forever,
\par }{\Q they will never
wake
up,”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{40}“I will lead
them off to be slaughtered
\par }{\Q like lambs,
rams,
and male goats.”
\par }{\Q \VS{41}“See how
Babylon
has been captured!
\par }{\Q See how the pride
of the whole
earth
has been taken!
\par }{\Q See what
an object of horror
\par }{\Q Babylon
has become among the nations!
\par }{\Q \VS{42}The sea
has swept
over
Babylon.
\par }{\Q She has been covered
by a multitude
of its waves.
\par }{\Q \VS{43}The towns
of Babylonia have become
heaps of ruins.
\par }{\Q She has become a dry
and barren desert.
\par }{\Q No
one
lives in those towns any more.
\par }{\Q No
one
even
passes through them.
\par }{\Q \VS{44}I will punish
the god Bel
in
Babylon.
\par }{\Q I will make him spit out
what he has swallowed.
\par }{\Q The nations
will not
come streaming
to
him any longer.
\par }{\Q Indeed,
the walls
of Babylon
will fall.”
\par }{\Q \VS{45}“Get out
of Babylon,
my people!
\par }{\Q Flee to save
your lives
\par }{\Q from the fierce
anger
of the {\ND{Lord}}!
\par }{\Q \VS{46}Do not
lose
your courage
or become
afraid
\par }{\Q because of the reports
that are heard
in the land.
\par }{\Q For a report
will come
in one year.
\par }{\Q Another report
will follow
it in the next.
\par }{\Q There will be violence
in the land
\par }{\Q with ruler
fighting against
ruler.”
\par }{\Q \VS{47}“So
the time
will certainly
come
\par }{\Q when I will punish
the idols
of Babylon.
\par }{\Q Her whole
land
will be put to shame.
\par }{\Q All
her mortally wounded
will collapse
in her midst.
\par }{\Q \VS{48}Then heaven
and earth
and all
that
is in them
\par }{\Q will sing
for joy over
Babylon.
\par }{\Q For
destroyers
from the north
will attack
it,”
\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{49}“Babylon
must fall
\par }{\Q because of the Israelites
she has killed,
\par }{\Q just as
the earth’s
mortally wounded
fell
\par }{\Q because of Babylon.
\par }{\Q \VS{50}You who have escaped
the sword,
\par }{\Q go,
do not
delay.
\par }{\Q Remember
the {\ND{Lord}}
in a faraway
land.
\par }{\Q Think
about Jerusalem.
\par }{\Q \VS{51}‘We are ashamed
because
we
have been insulted.
\par }{\Q Our faces
show our disgrace.
\par }{\Q For
foreigners
have invaded
\par }{\Q the holy rooms
in the
{\ND{Lord}}’s
temple.’
\par }{\Q \VS{52}Yes,
but the time
will certainly come,”
says
the {\ND{Lord}},

\par }{\Q “when I will punish
her idols.
\par }{\Q Throughout
her land
the mortally wounded
will groan.
\par }{\Q \VS{53}Even if
Babylon
climbs
high into the sky
\par }{\Q and fortifies
her elevated
stronghold,
\par }{\Q I will send
destroyers
against her,”

\par }{\Q says
the {\ND{Lord}}.
\par }{\Q \VS{54}Cries
of anguish
will come from Babylon,
\par }{\Q the sound of great
destruction
from the land
of the Babylonians.
\par }{\Q \VS{55}For
the {\ND{Lord}}
is ready to destroy
Babylon,
\par }{\Q and put an end
to her loud
noise.
\par }{\Q Their waves
will roar
like turbulent
waters.
\par }{\Q They will make
a deafening
noise.
\par }{\Q \VS{56}For
a destroyer
is attacking
Babylon.
\par }{\Q Her warriors
will be captured;
\par }{\Q their bows
will be broken.
\par }{\Q For
the {\ND{Lord}}
is a God
who punishes;
\par }{\Q he pays back in full.
\par }{\Q \VS{57}“I will make her officials
and wise
men drunk,
\par }{\Q along with her governors,
leaders,
and warriors.
\par }{\Q They will fall asleep
forever
and never
wake
up,”

\par }{\Q says
the King
whose name
is the
{\ND{Lord}}
who rules over all.
\par }{\Q \VS{58}This is what the
{\ND{Lord}}
who rules over all
says,
\par }{\Q “Babylon’s
thick
wall
will be completely demolished.
\par }{\Q Her high
gates
will be set on fire.
\par }{\Q The peoples
strive
for what does not satisfy.
\par }{\Q The nations
grow weary
trying
to get what will be destroyed.”
\par }{\PP \VS{59}This is the order
Jeremiah
the prophet
gave to Seraiah
son
of Neriah,
son
of Mahseiah,
when he went
to
King
Zedekiah
of Judah
in Babylon
during the fourth
year
of his reign.
(Seraiah
was a quartermaster.)
\VS{60}Jeremiah
recorded
on one
scroll
all
the judgments
that
would
come
upon
Babylon
– all
these
prophecies
written
about
Babylon.
\VS{61}Then Jeremiah
said
to
Seraiah,
“When you arrive
in Babylon,
make sure
you read
aloud all
these
prophecies.
\VS{62}Then say,
‘O
{\ND{Lord}}, you
have announced
that you will destroy
this
place
so that no
people
or animals
live
in it any longer. Certainly
it will lie desolate
forever!’
\VS{63}When
you finish
reading
this
scroll
aloud, tie
a stone
to it and throw
it into
the middle
of the Euphrates River.
\VS{64}Then say,
‘In the same
way Babylon
will sink
and never
rise
again because of the judgments
I am
ready to bring
upon
her; they will grow faint.’ ”
\par }{\PP The prophecies
of Jeremiah
end
here.

\par }\Chap{52}{\PP \VerseOne{1}Zedekiah
was twenty-one
years
old when he became king,
and he ruled
in Jerusalem
for eleven
years.
His mother’s
name
was Hamutal
daughter
of Jeremiah,
from Libnah.
\VS{2}He did
what displeased
the {\ND{Lord}}
just
as Jehoiakim
had done.
\par }{\PP \VS{3}What follows is a record of what happened to Jerusalem
and Judah
because
of the
{\ND{Lord}}’s
anger
when he drove
them out of his sight.
Zedekiah
rebelled
against the king
of Babylon.
\VS{4}King
Nebuchadnezzar
of Babylon
came against
Jerusalem
with his whole
army
and set up camp
outside
it. They built
siege ramps
all around
it. He arrived
on the tenth
day of the tenth
month
in the ninth
year
that Zedekiah ruled over Judah.
\VS{5}The city
remained
under siege
until
Zedekiah’s
eleventh
year.
\VS{6}By the ninth
day
of the fourth
month
the famine
in the city
was so severe
the residents
had no
food.
\VS{7}They broke
through the city
walls, and all
the soldiers tried
to escape.
They left
the city
during the night.
They went through
the gate
between
the two walls
that
is near the king’s
garden.
(The Babylonians
had the city
surrounded.) Then
they headed for the Jordan Valley.
\VS{8}But the Babylonian
army
chased
after
the king.
They caught
up with Zedekiah
in the plains
of Jericho,
and his entire
army
deserted
him.
\VS{9}They captured
him and brought him up
to
the king
of Babylon
at Riblah
in the territory
of Hamath
and he passed sentence on him there.
\VS{10}The king
of Babylon
had Zedekiah’s sons
put to death
while Zedekiah
was forced to watch.
He also
had all
the nobles
of Judah
put to death
there at Riblah.
\VS{11}He had Zedekiah’s
eyes
put out
and had him bound
in chains.
Then the king
of Babylon
had him led
off to Babylon
and he was imprisoned
there until
the day
he died.
\par }{\PP \VS{12}On the tenth
day
of the fifth
month,
in the nineteenth
year
of King
Nebuchadnezzar
of Babylon,
Nebuzaradan,
the captain
of the royal guard
who served
the king
of Babylon,
arrived
in Jerusalem.
\VS{13}He burned down
the
{\ND{Lord}}’s
temple,
the royal
palace,
and all
the houses
in Jerusalem,
including
every
large
house.
\VS{14}The whole
Babylonian
army
that
came with
the captain
of the royal guard
tore
down the walls
that surrounded
Jerusalem.
\VS{15}Nebuzaradan,
the captain
of the royal guard,
took into exile
some
of the poor,
the rest
of the people
who remained
in the city,
those who had deserted
to him, and the
rest
of the craftsmen.
\VS{16}But he left behind
some of the poor
and gave them fields
and vineyards.
\par }{\PP \VS{17}The Babylonians
broke
the
two bronze
pillars
in the temple
of the {\ND{Lord}}, as well as
the
movable stands
and the large bronze
basin called the “The Sea.”
They took
all
the bronze
to Babylon.
\VS{18}They also took
the pots,
shovels,
trimming shears,
basins,
pans,
and all
the bronze
utensils
used by the priests.
\VS{19}The captain
of the royal guard
took
the gold
and silver
bowls,
censers,
basins,
pots,
lampstands,
pans,
and vessels.
\VS{20}The bronze
of the items
that
King
Solomon
made
for the
{\ND{Lord}}’s
temple
(including the two
pillars,
the large bronze basin
called “The Sea,”
the twelve
bronze
bulls
under
“The Sea,” and the movable stands
) was too heavy to be
weighed.
\VS{21}Each of the pillars
was about 27 feet
high,
about 18 feet
in circumference,
three
inches
thick,
and hollow.
\VS{22}The bronze
top of
one
pillar
was about seven and one-half feet
high
and had bronze
latticework
and pomegranate-shaped
ornaments
all
around
it. The second
pillar
with its pomegranate-shaped ornaments was like it.
\VS{23}There were
ninety-six
pomegranate-shaped ornaments
on the sides;
in all
there were one hundred
pomegranate-shaped ornaments
over
the latticework
that went around it.
\par }{\PP \VS{24}The captain
of the royal guard
took
Seraiah
the chief
priest,
Zephaniah
the priest
who was second
in rank, and the three
doorkeepers.
\VS{25}From
the city
he took
an official
who
was in charge of
the soldiers,
seven
of the king’s
advisers
who
were discovered
in the
city,
an official
army
secretary
who drafted
citizens
for military service, and sixty
citizens
who were discovered
in the middle
of the city.
\VS{26}Nebuzaradan,
the captain
of the royal guard,
took
them and brought
them to
the king
of Babylon
at Riblah.
\VS{27}The king
of Babylon
ordered
them to be executed
at Riblah
in the territory
of Hamath.
\par }{\PP So Judah
was taken into exile
away from its land.
\VS{28}Here
is the official record of the number of people
Nebuchadnezzar
carried into exile: In the seventh
year,
3,023
Jews;
\VS{29}in Nebuchadnezzar’s
eighteenth
year,
832
people
from Jerusalem;
\VS{30}in Nebuchadnezzar’s
twenty-third
year,
Nebuzaradan,
the captain
of the royal guard,
carried into exile
745
Judeans.
In all
4,600
people went into exile.
\par }{\SH Jehoiachin in Exile
\par }{\PP \VS{31}In the thirty-seventh
year
of the exile
of King
Jehoiachin
of Judah,
on the twenty-fifth
day of the twelfth
month,
Evil-Merodach,
in the first year
of his reign,
pardoned
King
Jehoiachin
of Judah
and released
him from prison.
\VS{32}He spoke
kindly
to him
and gave
him a more prestigious
position
than the other kings
who
were with
him in Babylon.
\VS{33}Jehoiachin took off
his prison
clothes
and ate
daily
in the king’s presence
for the rest
of his life.
\VS{34}He was given
daily
provisions
by the king
of Babylon
for the rest of his life until
the day
he died.

\par }