\NormalFont\ShortTitle{Job}
{\MT Job
\par }{\SH \par }{\MS I. The Prologue (1:1-2:13)Job’s Good Life


\par }\ChapOne{1}{\PP \VerseOne{1}There was
a man
in the land
of Uz
whose name
was Job.
And that man
was
pure
and upright,
one who feared
God
and turned
away from evil.
\VS{2}Seven
sons
and three
daughters
were born to him.
\VS{3}His possessions
included 7,000
sheep,
3,000
camels,
500
yoke
of oxen,
and 500
female donkeys;
in addition he had a very
great
household.
Thus he was
the greatest
of all
the people in the east.
\par }{\PP \VS{4}Now
his sons
used
to go and hold a feast
in the house
of each
one in turn, and they would send
and invite
their three
sisters
to eat
and to drink
with them.
\VS{5}When
the days
of their feasting
were finished, Job
would send
for them and sanctify
them; he would get up early
in the morning
and offer
burnt offerings
according to the number
of them all.
For
Job
thought,
“Perhaps
my children
have sinned
and cursed
God
in their hearts.”
This
was Job’s
customary
practice.
\par }{\SH Satan’s Accusation of Job
\par }{\PP \VS{6}Now
the day
came
when the sons
of God
came to present
themselves before the
{\ND{Lord}} –
and Satan
also
arrived
among them.
\VS{7}The
{\ND{Lord}}
said
to
Satan,
“Where
have you come
from?” And Satan
answered
the
{\ND{Lord}}, “From roving about
on the earth,
and from walking back and forth across it.”
\VS{8}So the
{\ND{Lord}}
said
to
Satan,
“Have you considered
my servant
Job? There is no
one like
him on the earth,
a pure
and upright
man, one who fears
God
and turns
away from evil.”
\par }{\PP \VS{9}Then
Satan
answered
the {\ND{Lord}}, “Is it for nothing
that Job
fears
God?
\VS{10}Have you
not
made a hedge
around him
and his
household
and all
that
he has
on every side? You have blessed
the work
of his hands,
and his livestock
have increased
in the land.
\VS{11}But extend
your hand
and strike
everything
he has,
and he will no
doubt
curse
you to your face!”
\par }{\PP \VS{12}So the
{\ND{Lord}}
said to
Satan,
“All right then,
everything
he has
is in your power.
Only
do not
extend
your hand
against the man himself!” So Satan
went out
from the presence
of the {\ND{Lord}}.
\par }{\SH Job’s Integrity in Adversity
\par }{\PP \VS{13}Now the day
came when Job’s sons
and daughters
were eating
and drinking
wine
in their oldest
brother’s
house,
\VS{14}and a messenger
came
to
Job,
saying,
“The oxen
were plowing
and the donkeys
were grazing
beside
them,
\VS{15}and the Sabeans
swooped down
and carried
them all away,
and they killed
the servants
with the sword! And I – only I alone – escaped to tell you!”
\par }{\PP \VS{16}While
this
one was still speaking,
another
messenger arrived
and said,
“The fire
of God
has fallen
from
heaven
and has burned
up the sheep
and the servants
– it has consumed them! And I – only I alone – escaped to tell you!”
\par }{\PP \VS{17}While
this
one was still speaking
another
messenger arrived
and said,
“The Chaldeans
formed three
bands
and made a raid
on
the camels
and carried
them all away,
and they killed
the servants
with the sword! And I – only I alone – escaped to tell you!”
\par }{\PP \VS{18}While
this
one was still speaking
another
messenger arrived
and said,
“Your sons
and your daughters
were eating
and drinking
wine
in their oldest
brother’s
house,
\VS{19}and suddenly
a great
wind
swept
across
the wilderness
and struck
the four
corners
of the house,
and it fell
on
the young people,
and they died! And I – only I alone – escaped to tell you!”
\par }{\PP \VS{20}Then Job
got
up and tore
his robe.
He shaved
his head,
and then he threw
himself down with his face
to the ground.
\VS{21}He said,
“Naked
I came
from my mother’s
womb,
and naked
I will return
there.
The
{\ND{Lord}}
gives,
and the
{\ND{Lord}}
takes away.
May
the name
of the {\ND{Lord}}
be blessed!”
\VS{22}In all
this
Job
did not
sin,
nor
did he charge
God
with moral impropriety.

\par }\Chap{2}{\PP \VerseOne{1}Again
the day
came
when the sons
of God
came to present
themselves before the
{\ND{Lord}}, and Satan
also
arrived
among
them to present
himself before the
{\ND{Lord}}.
\VS{2}And the
{\ND{Lord}}
said
to
Satan,
“Where
do
you come
from?” Satan
answered
the {\ND{Lord}}, “From roving about
on the earth,
and from walking back and forth across it.”
\VS{3}Then the
{\ND{Lord}}
said
to
Satan,
“Have you considered
my servant
Job? For
there is no
one like
him on the earth,
a pure
and upright
man,
one who fears
God
and turns away
from evil.
And he still
holds firmly
to his integrity,
so that you stirred
me up
to destroy
him without reason.”
\par }{\PP \VS{4}But Satan
answered
the {\ND{Lord}}, “Skin
for
skin! Indeed,
a man
will give
up all
that
he has to save
his life!
\VS{5}But extend
your hand
and strike
his bone
and his flesh,
and he will no
doubt
curse
you to
your face!”
\par }{\PP \VS{6}So the
{\ND{Lord}}
said
to
Satan,
“All right, he is in your power;
only
preserve
his life.”
\par }{\SH Job’s Integrity in Suffering
\par }{\PP \VS{7}So Satan
went out
from the presence
of the {\ND{Lord}}, and he afflicted
Job
with a malignant
ulcer
from the sole
of his feet
to
the top of
his head.
\VS{8}Job took
a shard
of broken pottery to scrape
himself with while he was sitting
among
the ashes.
\par }{\PP \VS{9}Then his wife
said
to him, “Are you still
holding firmly
to your integrity? Curse
God,
and die!”
\VS{10}But he replied, “You’re talking
like one
of the godless women
would do! Should we receive
what is good
from God,
and not
also receive
what is evil?” In all
this
Job
did not
sin
by what he said.
\par }{\SH The Visit of Job’s Friends
\par }{\PP \VS{11}When Job’s
three
friends
heard
about all
this
calamity
that had happened
to him,
each
of them came
from his own country –
Eliphaz
the Temanite,
Bildad
the Shuhite,
and Zophar
the Naamathite.
They met
together
to come
to show sympathy
for him and to console him.
\VS{12}But when they gazed intently
from a distance
but did not
recognize
him, they began
to weep
loudly.
Each
of them tore
his robes,
and they threw
dust
into the air
over
their heads.
\VS{13}Then they sat
down with
him on the ground
for seven
days
and seven
nights,
yet no
one spoke
a word
to him,
for
they saw
that
his pain
was very
great.

\par }\Chap{3}{\PP \VerseOne{1}After
this
Job
opened
his mouth
and cursed
the day he was born.
\VS{2}Job
spoke up
and said:
\par }{\Q \VS{3}“Let the day
on which I was born
perish,
\par }{\Q and the night
that said,
\par }{\Q ‘A man
has been conceived!’
\par }{\Q \VS{4}That day –
let it be
darkness;
\par }{\Q let not
God
on
high
regard
it,
\par }{\Q nor let
light shine on it!
\par }{\Q \VS{5}Let
darkness
and the deepest
\par }{\Q shadow
claim it;

\par }{\Q let a cloud
settle
on
it;
\par }{\Q let whatever blackens
the day
terrify it!
\par }{\Q \VS{6}That night
– let darkness
seize
it;
\par }{\Q let it not
be included
among the days
of the year;
\par }{\Q let it not
enter
among the number
of the months!
\par }{\Q \VS{7}Indeed,
let that night
be
barren;
\par }{\Q let no
shout of joy
penetrate it!
\par }{\Q \VS{8}Let those who curse
the day
curse it –
\par }{\Q those who are prepared to rouse Leviathan.
\par }{\Q \VS{9}Let its morning
stars
be darkened;
\par }{\Q let it wait
for daylight
but
find none,
\par }{\Q nor let
it see
the first rays
of dawn,
\par }{\Q \VS{10}because
it did not
shut
the doors
of my mother’s womb
on me,

\par }{\Q nor did it hide
trouble
from my eyes!
\par }{\SH Job Wishes He Had Died at Birth
\par }{\Q \VS{11}“Why
did I not
die
at birth,
\par }{\Q and why did I not expire
\par }{\Q as I came out
of the womb?
\par }{\Q \VS{12}Why
did the knees
welcome
me,

\par }{\Q and why
were there two breasts
\par }{\Q that
I might nurse at them?
\par }{\Q \VS{13}For
now
I would be lying
down
\par }{\Q and would be quiet,
\par }{\Q I would be asleep
and then
at peace
\par }{\Q \VS{14}with
kings
and counselors
of the earth
\par }{\Q who built
for themselves places now desolate,
\par }{\Q \VS{15}or
with
princes
who possessed gold,
\par }{\Q who filled
their palaces
with silver.
\par }{\Q \VS{16}Or
why was I not
buried
\par }{\Q like a stillborn
infant,

\par }{\Q like infants
who have never
seen
the light?
\par }{\Q \VS{17}There
the wicked
cease
from turmoil,
\par }{\Q and there
the weary
are at rest.
\par }{\Q \VS{18}There the prisoners
relax
together;
\par }{\Q they do not
hear
the voice
of the oppressor.
\par }{\Q \VS{19}Small
and great
are there,
\par }{\Q and the slave
is free
from his master.
\par }{\SH Longing for Death
\par }{\Q \VS{20}“Why
does God give
light
to one who is in misery,
\par }{\Q and life
to those
whose soul
is bitter,
\par }{\Q \VS{21}to those who wait
for death
that does not
come,
\par }{\Q and search for
it

\par }{\Q more than for hidden treasures,
\par }{\Q \VS{22}who rejoice
even to jubilation,
\par }{\Q and are exultant
when
they find
the grave?
\par }{\Q \VS{23}Why is light given to a man
\par }{\Q whose
way
is hidden,
\par }{\Q and whom God
has
hedged in?
\par }{\Q \VS{24}For
my sighing
comes
in place
of my food,
\par }{\Q and my groanings
flow forth
like water.
\par }{\Q \VS{25}For
the very thing
I dreaded
has happened
to me,
\par }{\Q and what I feared
has
come upon me.
\par }{\Q \VS{26}I have no
ease,
I have no quietness;
\par }{\Q I cannot
rest;
turmoil
has
come upon me.”


\par }\Chap{4}{\PP \VerseOne{1}Then
Eliphaz
the Temanite
answered:
\par }{\Q \VS{2}“If someone
should
attempt
a word
with you,
\par }{\Q will you be impatient?

\par }{\Q But who can
refrain
from speaking ?
\par }{\Q \VS{3}Look,
you have instructed many;
\par }{\Q you have strengthened
feeble
hands.
\par }{\Q \VS{4}Your words
have supported
those
\par }{\Q who stumbled,
\par }{\Q and you have strengthened
the knees
\par }{\Q that gave way.
\par }{\Q \VS{5}But
now
the same thing comes
to you,
\par }{\Q and you are discouraged;
\par }{\Q it strikes
you,
\par }{\Q and you are terrified.
\par }{\Q \VS{6}Is not
your piety
your confidence,
\par }{\Q and your blameless
ways
your hope?
\par }{\Q \VS{7}Call to mind
now:

\par }{\Q Who,
being innocent,
ever perished?

\par }{\Q And where
were upright
people ever destroyed?
\par }{\Q \VS{8}Even as
I have seen,
those who plow
iniquity
\par }{\Q and those who sow
trouble
reap the same.
\par }{\Q \VS{9}By the breath
of God
they perish,
\par }{\Q and by the blast
of his anger
they are consumed.
\par }{\Q \VS{10}There is the roaring
of the lion
\par }{\Q and the growling
of the young lion,
\par }{\Q but the teeth
of the young lions
are broken.
\par }{\Q \VS{11}The mighty lion
perishes
for lack
of prey,
\par }{\Q and the cubs
of the lioness
are scattered.
\par }{\SH Ungodly Complainers Provoke God’s Wrath
\par }{\Q \VS{12}“Now a word
was secretly
brought
to me,
\par }{\Q and my ear
caught a whisper of it.
\par }{\Q \VS{13}In the troubling thoughts
of the dreams
in the night
\par }{\Q when a deep sleep
falls
on
men,
\par }{\Q \VS{14}a trembling
gripped me – and a terror! –
\par }{\Q and made all my bones shake.
\par }{\Q \VS{15}Then a breath
of air passes
by my face;
\par }{\Q it makes
the hair
of my flesh stand up.
\par }{\Q \VS{16}It stands
still,

\par }{\Q but I cannot recognize
its appearance;
\par }{\Q an image
is before
my eyes,
\par }{\Q and I hear
a murmuring
voice:
\par }{\Q \VS{17}“Is a mortal man
righteous
before God?
\par }{\Q Or
a man pure before his Creator?
\par }{\Q \VS{18}If
God puts no
trust
in his servants
\par }{\Q and attributes
folly
to his angels,
\par }{\Q \VS{19}how much
more to those who live
in houses
of clay,
\par }{\Q whose
foundation
is in the dust,
\par }{\Q who are crushed
like a moth?
\par }{\Q \VS{20}They are destroyed
between morning
and evening;
\par }{\Q they perish
forever
without
anyone regarding it.
\par }{\Q \VS{21}Is not
their excess
wealth taken away from
them?

\par }{\Q They die,
yet without
attaining
wisdom.


\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}“Call
now! Is there anyone
who will answer
you?

\par }{\Q To
which
of the holy ones
will you turn?
\par }{\Q \VS{2}For
wrath kills the foolish
person,

\par }{\Q and anger slays
the silly one.
\par }{\Q \VS{3}I myself
have seen
the fool
taking root,
\par }{\Q but suddenly
I cursed
his place
of residence.
\par }{\Q \VS{4}His children
are far
from safety,
\par }{\Q and they are crushed
at the place where judgment is rendered,
\par }{\Q nor is there
anyone to deliver them.
\par }{\Q \VS{5}The hungry
eat up
his harvest,
\par }{\Q and take
it even from behind the thorns,
\par }{\Q and the thirsty
swallow up
their fortune.
\par }{\Q \VS{6}For
evil
does not
come
up from the dust,
\par }{\Q nor does trouble
spring
up from the ground,
\par }{\Q \VS{7}but
people
are born
to trouble,
\par }{\Q as surely as the sparks
fly upward.
\par }{\SH Blessings for the One Who Seeks God
\par }{\Q \VS{8}“But as for
me, I
would seek
God,
\par }{\Q and to
God
I would set
forth my case.
\par }{\Q \VS{9}He does
great
and unsearchable
things,
\par }{\Q marvelous things
without
number;
\par }{\Q \VS{10}he gives
rain
on
the earth,
\par }{\Q and sends
water
on
the fields;
\par }{\Q \VS{11}he sets
the lowly
on high,
\par }{\Q that those who mourn
are raised
to safety.
\par }{\Q \VS{12}He frustrates
the plans
of the crafty
\par }{\Q so that their hands
cannot
accomplish
\par }{\Q what they had planned!
\par }{\Q \VS{13}He catches
the wise
in their own craftiness,
\par }{\Q and the counsel
of the cunning is
brought to a quick end.
\par }{\Q \VS{14}They meet
with darkness
in the daytime,
\par }{\Q and grope
about in the noontime
as if it were night.
\par }{\Q \VS{15}So he saves
from the sword
that comes from their mouth,
\par }{\Q even the poor
from the hand
of the powerful.
\par }{\Q \VS{16}Thus
the poor
have hope,
\par }{\Q and iniquity
shuts
its mouth.
\par }{\Q \VS{17}“Therefore, blessed
is the man
whom God
corrects,
\par }{\Q so do not
despise
the discipline
of the Almighty.
\par }{\Q \VS{18}For
he wounds, but he also bandages;
\par }{\Q he strikes,
but his hands
also heal.
\par }{\Q \VS{19}He will deliver
you from
six
calamities;
\par }{\Q yes, in seven
no
evil
will touch you.
\par }{\Q \VS{20}In time of famine
he will redeem
you from death,
\par }{\Q and in time of war
from the power
of the sword.
\par }{\Q \VS{21}You will be protected
from malicious
gossip,
\par }{\Q and will not
be afraid
of the destruction
when
it comes.
\par }{\Q \VS{22}You will laugh at
destruction
and famine
\par }{\Q and need not
be afraid
of the beasts
of the earth.
\par }{\Q \VS{23}For
you
will have a pact
with
the stones
of the field,
\par }{\Q and the wild
animals
will be at peace
with you.
\par }{\Q \VS{24}And you will know
that
your home
\par }{\Q will be secure,

\par }{\Q and when you inspect
your domains,
\par }{\Q you will not
be missing anything.
\par }{\Q \VS{25}You will also know
that
your children
will be numerous,
\par }{\Q and your descendants
like the grass
of the earth.
\par }{\Q \VS{26}You will come
to
your grave
in a full age,
\par }{\Q As stacks of grain
are harvested in their season.
\par }{\Q \VS{27}Look,
we have investigated
this, so
it is true.
\par }{\Q Hear
it, and apply it for your own good.”

\par }\Chap{6}{\PP \VerseOne{1}Then
Job
responded:
\par }{\Q \VS{2}“Oh, if only
my grief
could be weighed,
\par }{\Q and my misfortune laid on the scales
too!
\par }{\Q \VS{3}But because
it is heavier
than the sand
of the sea,
\par }{\Q that
is why
my words
have been wild.
\par }{\Q \VS{4}For
the arrows
of the Almighty
are within me;
\par }{\Q my
spirit
drinks
their poison;
\par }{\Q God’s
sudden terrors
are arrayed against me.
\par }{\SH Complaints Reflect Suffering
\par }{\Q \VS{5}“Does the wild donkey
bray
when it is near grass?

\par }{\Q Or
does the ox
low near its fodder?
\par }{\Q \VS{6}Can food that is tasteless
be eaten
without
salt?
\par }{\Q Or
is
there any taste
in the white
of an egg?
\par }{\Q \VS{7}I have refused
to touch
such
things;

\par }{\Q they
are like loathsome
food to me.
\par }{\SH A Cry for Death
\par }{\Q \VS{8}“Oh that my request
would be realized,
\par }{\Q and that God
would grant
me what I long for!
\par }{\Q \VS{9}And that God
would be willing
to crush
me,
\par }{\Q that he would let loose
his hand
\par }{\Q and kill me.
\par }{\Q \VS{10}Then I would yet have my comfort,
\par }{\Q then I would rejoice,
\par }{\Q in spite
of pitiless
pain,

\par }{\Q for
I have not
concealed
the words
of the Holy One.
\par }{\Q \VS{11}What
is my strength,
that
I should wait?

\par }{\Q and what
is my end,
\par }{\Q that
I should prolong
my life?
\par }{\Q \VS{12}Is my
strength
like that of stones?

\par }{\Q or
is my flesh
made of bronze?
\par }{\Q \VS{13}Is not
my power to help
myself nothing,
\par }{\Q and has not every resource
been driven
from me?
\par }{\SH Disappointing Friends
\par }{\Q \VS{14}“To the one in despair,
kindness
should come from his friend
\par }{\Q even if he forsakes
the fear
of the Almighty.
\par }{\Q \VS{15}My brothers
have been as treacherous
as
a seasonal stream,
\par }{\Q and as the riverbeds
of the intermittent streams
\par }{\Q that flow away.
\par }{\Q \VS{16}They are dark
because of ice;
\par }{\Q snow
is piled up over them.
\par }{\Q \VS{17}When they are scorched,
they dry up,
\par }{\Q when it is hot,
they vanish
from their place.
\par }{\Q \VS{18}Caravans
turn aside from their routes;
\par }{\Q they go
into the wasteland
and perish.
\par }{\Q \VS{19}The caravans
of Tema
looked intently
for these streams;

\par }{\Q the traveling merchants of Sheba
hoped for them.
\par }{\Q \VS{20}They were distressed,
\par }{\Q because
each one had been so confident;
\par }{\Q they arrived
there, but were disappointed.
\par }{\Q \VS{21}For
now
you have become
like these streams that are no
help;

\par }{\Q you see
a terror,
and are afraid.
\par }{\SH Friends’ Fears
\par }{\Q \VS{22}“Have I ever said, ‘Give
me something,
\par }{\Q and from your fortune make gifts in my favor’?
\par }{\Q \VS{23}Or ‘Deliver me from
the enemy’s
power,
\par }{\Q and from the hand
of tyrants
ransom me’?
\par }{\SH No Sin Discovered
\par }{\Q \VS{24}“Teach
me and I,
for my part, will be silent;
\par }{\Q explain to me
how
I have been mistaken.
\par }{\Q \VS{25}How
painful
are honest
words!
\par }{\Q But what
does
your reproof
prove?
\par }{\Q \VS{26}Do you intend to criticize
mere words,
\par }{\Q and treat
the words
of a despairing
man as
wind?
\par }{\Q \VS{27}Yes,
you would gamble
for the fatherless,
\par }{\Q and auction
off your friend.
\par }{\SH Other Explanation
\par }{\Q \VS{28}“Now
then,
be good enough
to look
at me;

\par }{\Q and I will not
lie
to your face!
\par }{\Q \VS{29}Relent,
let
there be
no falsehood;
\par }{\Q reconsider,
for my righteousness is intact!
\par }{\Q \VS{30}Is there
any falsehood
on my lips?
\par }{\Q Can
my mouth
not
discern
evil things?

\par }\Chap{7}{\PP \par }{\Q \VerseOne{1}“Does not
humanity
have hard service
on
earth?
\par }{\Q Are not their days
also
\par }{\Q like the days
of a hired man?
\par }{\Q \VS{2}Like a servant
longing
for the evening shadow,
\par }{\Q and like a hired man
looking
for his wages,
\par }{\Q \VS{3}thus
I have been made to inherit
\par }{\Q months
of futility,
\par }{\Q and nights
of sorrow
\par }{\Q have been appointed to me.
\par }{\Q \VS{4}If
I lie
down, I say, ‘When
will I arise?’,
\par }{\Q and the night
stretches on
\par }{\Q and I
toss and turn restlessly
\par }{\Q until
the day dawns.
\par }{\Q \VS{5}My body
is clothed
with worms
and dirty
scabs;
\par }{\Q my skin
is broken and festering.
\par }{\Q \VS{6}My days
are swifter
than a weaver’s shuttle
\par }{\Q and they come
to an end
without hope.
\par }{\Q \VS{7}Remember
that
my life
is but a breath,
\par }{\Q that my eyes
will never
again
see
happiness.
\par }{\Q \VS{8}The eye
of him
who sees
me now will see me no
more;

\par }{\Q your eyes
will look for me, but I will be gone.
\par }{\Q \VS{9}As a cloud
is dispersed
and then disappears,
\par }{\Q so
the one who goes down
to the grave
\par }{\Q does not
come up again.
\par }{\Q \VS{10}He returns
no
more
to his house,
\par }{\Q nor
does his place
of residence
know him any more.
\par }{\SH Job Remonstrates with God
\par }{\Q \VS{11}“Therefore,
I
will not
refrain
my mouth;
\par }{\Q I will speak
in the anguish
of my spirit;
\par }{\Q I
will complain
in the bitterness
of my soul.
\par }{\Q \VS{12}Am I
the sea,
or
the creature
of the deep,

\par }{\Q that
you must put
me under
guard?
\par }{\Q \VS{13}If
I say, “My bed
will comfort
me,

\par }{\Q my couch
will ease
my complaint,”
\par }{\Q \VS{14}then you scare
me with dreams
\par }{\Q and terrify
me with visions,
\par }{\Q \VS{15}so that I would prefer strangling,
\par }{\Q and death
more than life.
\par }{\Q \VS{16}I loathe
it; I do not
want to live
forever;
\par }{\Q leave me
alone,
for
my days
are a vapor!
\par }{\SH Insignificance of Humans
\par }{\Q \VS{17}“What
is mankind
that
you make so much
of them,

\par }{\Q and that
you pay
attention
to them?
\par }{\Q \VS{18}And that you visit
them every morning,
\par }{\Q and try
them every moment?
\par }{\Q \VS{19}Will you never
look away
from
me,

\par }{\Q will you not
let
me alone
\par }{\Q long
enough to swallow
my spittle?
\par }{\Q \VS{20}If I have sinned
– what
have I done
to you,

\par }{\Q O watcher
of men?

\par }{\Q Why
have you set
me as your target?

\par }{\Q Have I become
a burden to you?
\par }{\Q \VS{21}And why
do you not
pardon
my transgression,
\par }{\Q and take
away
my iniquity?
\par }{\Q For
now
I will lie
down in the dust,
\par }{\Q and you will seek me diligently,
\par }{\Q but I will be gone.”

\par }\Chap{8}{\PP \VerseOne{1}Then
Bildad
the Shuhite
spoke up and said:
\par }{\Q \VS{2}“How long
will you speak
these
things,

\par }{\Q seeing
that the words
of your mouth
\par }{\Q are like a great wind?
\par }{\Q \VS{3}Does God
pervert
justice?

\par }{\Q Or
does the Almighty
pervert
what is right?
\par }{\Q \VS{4}If
your children
sinned
against him,
\par }{\Q he gave them over
to the penalty of their sin.
\par }{\Q \VS{5}But if
you
will look
to
God,
\par }{\Q and make your supplication
to
the Almighty,
\par }{\Q \VS{6}if
you become pure
and upright,
\par }{\Q even
now
he will rouse
himself
for
you,
\par }{\Q and will restore
your righteous
abode.
\par }{\Q \VS{7}Your beginning
will seem
so small,
\par }{\Q since your future
will flourish.
\par }{\Q \VS{8}“For
inquire
now
of the former
generation,
\par }{\Q and pay attention to the findings
\par }{\Q of their ancestors;
\par }{\Q \VS{9}For
we
were born yesterday
and do not
have knowledge,
\par }{\Q since
our days
on
earth
are but
a shadow.
\par }{\Q \VS{10}Will they
not
instruct
you and speak to you,
\par }{\Q and bring forth
words
\par }{\Q from
their understanding?
\par }{\Q \VS{11}Can the papyrus plant
grow tall
where there is no
marsh?
\par }{\Q Can reeds
flourish
without
water?
\par }{\Q \VS{12}While they are still
beginning to flower
\par }{\Q and not
ripe
for cutting,
\par }{\Q they can wither
away

\par }{\Q faster than any
grass!
\par }{\Q \VS{13}Such is
the destiny
of all
who forget
God;
\par }{\Q the hope
of the godless
perishes,
\par }{\Q \VS{14}whose
trust
is in something futile,
\par }{\Q whose security
is a spider’s
web.
\par }{\Q \VS{15}He leans
against
his house
but it does not
hold up,
\par }{\Q he takes
hold
of it but it does not
stand.
\par }{\Q \VS{16}He is a well-watered
plant in the sun,
\par }{\Q its shoots
spread
over its garden.
\par }{\Q \VS{17}It wraps
its roots
around
a heap
of stones

\par }{\Q and it looks
for a place
among stones.
\par }{\Q \VS{18}If
he is uprooted
from his place,
\par }{\Q then that place
will disown
him, saying,

\par }{\Q ‘I have never
seen you!’
\par }{\Q \VS{19}Indeed,
this is
the joy
of his way,
\par }{\Q and out of the earth
others
spring up.
\par }{\Q \VS{20}“Surely,
God
does not
reject
a blameless
man,

\par }{\Q nor
does he grasp
the hand
\par }{\Q of the evildoers.
\par }{\Q \VS{21}He will yet
fill
your mouth
with laughter,
\par }{\Q and your lips
with gladness.
\par }{\Q \VS{22}Those who hate
you will be clothed
with shame,
\par }{\Q and the tent
of the wicked
will be no more.”


\par }\Chap{9}{\PP \VerseOne{1}Then
Job
answered:
\par }{\Q \VS{2}“Truly,
I know
that
this is so.
\par }{\Q But how
can a human
be just before God?
\par }{\Q \VS{3}If
someone wishes
to contend
with
him,
\par }{\Q he cannot
answer
him one
time in a thousand.
\par }{\Q \VS{4}He is wise
in heart
and mighty
in strength – ‘span class=”s 04310” title=”04310” ’
\par }{\Q who
has resisted
him and remained safe?
\par }{\PP \par }{\Q \VS{5}He who removes
mountains
suddenly,

\par }{\Q who
overturns
them in his anger;
\par }{\Q \VS{6}he who shakes
the earth
out of its place
\par }{\Q so that its pillars
tremble;
\par }{\Q \VS{7}he who commands
the sun
and it does not
shine

\par }{\Q and seals
up
the stars;
\par }{\Q \VS{8}he alone
spreads out
the heavens,
\par }{\Q and treads
on
the waves
of the sea;
\par }{\Q \VS{9}he makes
the Bear,
Orion,
and the Pleiades,
\par }{\Q and the constellations
of the southern sky;
\par }{\Q \VS{10}he does
great
and unsearchable
things,
\par }{\Q and wonderful things
without
number.
\par }{\Q \VS{11}If
he passes
by me, I cannot
see
him,

\par }{\Q if he goes
by, I cannot
perceive him.
\par }{\Q \VS{12}If
he snatches
away, who can
turn him back?

\par }{\Q Who dares
to say
to him,
‘What
are you doing?’
\par }{\Q \VS{13}God
does not
restrain
his anger;
\par }{\Q under
him
the helpers
of Rahab lie crushed.
\par }{\SH The Impossibility of Facing God in Court
\par }{\Q \VS{14}“How much
less,
then, can I
answer
him

\par }{\Q and choose
my words
to argue with him!
\par }{\Q \VS{15}Although
I am innocent,
\par }{\Q I could not
answer
him;

\par }{\Q I could only plead with my judge
for mercy.
\par }{\Q \VS{16}If
I summoned
him, and he answered
me,

\par }{\Q I would not
believe
\par }{\Q that
he would be listening
to my voice –
\par }{\Q \VS{17}he who
crushes
me with a tempest,
\par }{\Q and multiplies
my wounds
for no reason.
\par }{\Q \VS{18}He does not
allow
me to recover
my breath,
\par }{\Q for
he fills
me with bitterness.
\par }{\Q \VS{19}If
it is a matter of strength,
\par }{\Q most certainly
he is the strong one!
\par }{\Q And if
it is a matter of justice,
\par }{\Q he will say, ‘Who
will summon me?’
\par }{\Q \VS{20}Although
I am innocent,
\par }{\Q my mouth
would condemn
me;

\par }{\Q although I am
blameless,
\par }{\Q it would declare me perverse.
\par }{\Q \VS{21}I am blameless.
I
do not
know
myself.
\par }{\Q I despise
my life.
\par }{\SH Accusation of God’s Justice
\par }{\Q \VS{22}“It is all one! That
is why
I say,
\par }{\Q ‘He destroys
the blameless
and the guilty.’
\par }{\Q \VS{23}If
a scourge
brings sudden
death,
\par }{\Q he mocks
at the despair
of the innocent.
\par }{\Q \VS{24}If a land
has been given
\par }{\Q into the hand
of a wicked
man,

\par }{\Q he covers
the faces
of its judges;
\par }{\Q if
it is not
he, then
who is it?
\par }{\SH Renewed Complaint
\par }{\Q \VS{25}“My days
are swifter
than
a runner,
\par }{\Q they speed
by without
seeing
happiness.
\par }{\Q \VS{26}They glide
by like
reed
boats,
\par }{\Q like an eagle
that swoops
down on
its prey.
\par }{\Q \VS{27}If
I say, ‘I will forget
my complaint,
\par }{\Q I will change my expression
and be cheerful,’
\par }{\Q \VS{28}I dread
all
my sufferings,
\par }{\Q for I know
that
you do not
hold me blameless.
\par }{\Q \VS{29}If I am
guilty,
\par }{\Q why
then weary
myself in vain?
\par }{\Q \VS{30}If
I wash
myself with snow water,
\par }{\Q and make my hands
clean with lye,
\par }{\Q \VS{31}then
you plunge
me into a slimy pit
\par }{\Q and my own clothes
abhor me.
\par }{\Q \VS{32}For
he is not
a human
being like
I am,
\par }{\Q that I might answer
him,
\par }{\Q that we might come
together
in judgment.
\par }{\Q \VS{33}Nor is there
an arbiter
between us,
\par }{\Q who
might lay
his hand
on
us both,
\par }{\Q \VS{34}who would
take his rod
away from me
\par }{\Q so that his terror
would not
make me afraid.
\par }{\Q \VS{35}Then would I speak
and not
fear
him,
\par }{\Q but
it is not
so
with me.


\par }\Chap{10}{\PP \par }{\Q \VerseOne{1}“I am weary
of my life;
\par }{\Q I will complain
without restraint;
\par }{\Q I will speak
in the bitterness
of my soul.
\par }{\Q \VS{2}I will say
to
God,
‘Do not
condemn
me;
\par }{\Q tell
me why
you are contending with me.’
\par }{\Q \VS{3}Is it good
for
you to oppress,
\par }{\Q to despise
the work
of your hands,
\par }{\Q while you smile

\par }{\Q on
the schemes
of the wicked?
\par }{\SH Motivations of God
\par }{\Q \VS{4}“Do you have eyes
of flesh,
\par }{\Q or
do you see
as a human
being sees?
\par }{\Q \VS{5}Are
your days
like the days
of a mortal,
\par }{\Q or
your years
like the years
of a mortal,
\par }{\Q \VS{6}that
you must search
out my iniquity,
\par }{\Q and inquire
about my sin,
\par }{\Q \VS{7}although you know
that
I am
not
guilty,
\par }{\Q and that there is no
one who can deliver
\par }{\Q out of your hand?
\par }{\SH Contradictions in God’s Dealings
\par }{\Q \VS{8}“Your hands
have shaped
me and made
me,
\par }{\Q but now you destroy
me completely.
\par }{\Q \VS{9}Remember
that
you have made
me
as with the clay;
\par }{\Q will you return
me to
dust?
\par }{\Q \VS{10}Did you not
pour me out like milk,
\par }{\Q and curdle
me like cheese?
\par }{\Q \VS{11}You clothed
me with skin
and flesh
\par }{\Q and knit me together with bones
and sinews.
\par }{\Q \VS{12}You gave me
life
and favor,
\par }{\Q and your intervention
watched
over my
spirit.
\par }{\Q \VS{13}“But these
things you have concealed
in your heart;
\par }{\Q I know
that
this
is with you:
\par }{\Q \VS{14}If
I sinned,
then you would watch
me
\par }{\Q and you would not
acquit
me of my iniquity.
\par }{\Q \VS{15}If
I am guilty,
woe
to me,
\par }{\Q and if I am innocent,
I cannot
lift
my head;
\par }{\Q I am full
of shame,
\par }{\Q and satiated
with my affliction.
\par }{\Q \VS{16}If
I lift
myself up,

\par }{\Q you hunt
me as a fierce lion,
\par }{\Q and again
you display your power against me.
\par }{\Q \VS{17}You bring new
witnesses
against me,
\par }{\Q and increase
your anger
against me;
\par }{\Q relief troops come
against
me.
\par }{\SH An Appeal for Relief
\par }{\Q \VS{18}“Why
then did you bring me out
from the womb?
\par }{\Q I should have died
\par }{\Q and no
eye
would have seen me!
\par }{\Q \VS{19}I should
have
been
as though
I had never existed;
\par }{\Q I should
have been carried
\par }{\Q right from the womb
to the grave!
\par }{\Q \VS{20}Are not
my days
few?

\par }{\Q Cease,
then, and leave me alone,
\par }{\Q that I may find a little
comfort,
\par }{\Q \VS{21}before
I depart,
never
to return,
\par }{\Q to
the land
of darkness
\par }{\Q and the deepest shadow,
\par }{\Q \VS{22}to the land
of utter darkness,
\par }{\Q like
the deepest darkness,
\par }{\Q and the deepest shadow
and disorder,
\par }{\Q where even the light
is like
darkness.” ‘span class=”footnote” id=”footnote-18” ’‘span class=”key” ’18‘a href=”\#note-18” class=”backref” ’10:7‘/a’‘span class=”text” ’
{\IT{tn}}:
{\IT{Heb}}
{\TL{עַל־דַּעְתְּךָ}} (’al da’tÿkha, “upon your knowledge”). The use of the preposition means basically “in addition to your knowledge,” or “in spite of your knowledge,” i.e., “notwithstanding” or “although” (see GKC 383 §119.{\IT{aa}}, n. 2).

\par }\Chap{11}{\PP \VerseOne{1}Then Zophar
the Naamathite
spoke up and said:
\par }{\Q \VS{2}“Should not
this abundance
of words
be answered,
\par }{\Q or
should this talkative
man
\par }{\Q be vindicated?
\par }{\Q \VS{3}Will your idle
talk reduce
people to silence,
\par }{\Q and will no
one rebuke
you when
you mock?
\par }{\Q \VS{4}For you have said,
‘My teaching
is flawless,
\par }{\Q and I am
pure
in your sight.’
\par }{\Q \VS{5}But if only
God
would speak,
\par }{\Q if only he would open
his lips
against you,
\par }{\Q \VS{6}and reveal
to you the secrets
of wisdom
–
\par }{\Q for
true wisdom
has two sides –
\par }{\Q so that you would know
\par }{\Q that
God
has forgiven
some of your sins.
\par }{\Q \VS{7}“Can you discover
the essence
of God?
\par }{\Q Can you find out
\par }{\Q the perfection
of the Almighty?
\par }{\Q \VS{8}It is higher
than the heavens
– what
can you do?
\par }{\Q It is deeper
than Sheol –
what
can you know?
\par }{\Q \VS{9}Its measure
is longer
than the earth,
\par }{\Q and broader
than
the sea.
\par }{\Q \VS{10}If
he comes by and confines
you

\par }{\Q and convenes
a court,

\par }{\Q then
who can
prevent him?
\par }{\Q \VS{11}For
he knows
deceitful
men;
\par }{\Q when he sees
evil,
will he not
consider it?
\par }{\Q \VS{12}But an empty
man
will become wise,
\par }{\Q when a wild donkey’s
colt
is born
a human being.
\par }{\Q \VS{13}“As for you, if
you prove faithful,

\par }{\Q and if you
stretch
out your hands toward him,
\par }{\Q \VS{14}if
iniquity
is in your hand
– put it far
away,

\par }{\Q and do not
let evil
reside
in your tents.
\par }{\Q \VS{15}For
then
you will lift
up your face
\par }{\Q without blemish;
\par }{\Q you will be
securely
established

\par }{\Q and will not
fear.
\par }{\Q \VS{16}For
you
will forget
your trouble;
\par }{\Q you will remember
it
\par }{\Q like water
that has flowed away.
\par }{\Q \VS{17}And life will be brighter than the noonday;
\par }{\Q though there be darkness,

\par }{\Q it will be
like
the morning.
\par }{\Q \VS{18}And you will be secure,
because
there is
hope;
\par }{\Q you will be protected
\par }{\Q and will take
your rest
in safety.
\par }{\Q \VS{19}You will lie
down with no
one to make you afraid,
\par }{\Q and many
will seek your favor.
\par }{\Q \VS{20}But the eyes
of the wicked
fail,
\par }{\Q and escape
eludes
them;
\par }{\Q their one hope
is to breathe their last.”


\par }\Chap{12}{\PP \VerseOne{1}Then
Job
answered:
\par }{\Q \VS{2}“Without a doubt
you
are the people,
\par }{\Q and wisdom
will die
with you.
\par }{\Q \VS{3}I also
have understanding
as well as
you;
\par }{\Q I am
not
inferior
to you.

\par }{\Q Who
does not
know such
things as these?
\par }{\Q \VS{4}I am a laughingstock
to my friends,
\par }{\Q I, who called
on God
and whom he answered –
\par }{\Q a righteous
and blameless
man
\par }{\Q is a laughingstock!
\par }{\Q \VS{5}For calamity,
there is derision
\par }{\Q (according to the ideas
of the fortunate ) –
\par }{\Q a fate for those whose feet slip!
\par }{\Q \VS{6}But the tents
of robbers
are peaceful,
\par }{\Q and those who provoke
God
are confident – ‘span class=”s 0834” title=”0834” ’
\par }{\Q who
carry
their god
in their hands.
\par }{\SH Knowledge of God’s Wisdom
\par }{\Q \VS{7}“But now,
ask
the animals
and they will teach
you,
\par }{\Q or the birds
of the sky
and they will tell you.
\par }{\Q \VS{8}Or
speak
to the earth
and it will teach
you,
\par }{\Q or let the fish
of the sea
declare to you.
\par }{\Q \VS{9}Which
of all
these
does not
know
\par }{\Q that
the hand
of the
{\ND{Lord}}
has done
this,
\par }{\Q \VS{10}in whose
hand
is the life
of every
creature
\par }{\Q and the breath
of all
the human race.
\par }{\Q \VS{11}Does not
the ear
test
words,
\par }{\Q as the tongue
tastes
food?
\par }{\Q \VS{12}Is not wisdom
found among the aged?

\par }{\Q Does not long
life
bring understanding?
\par }{\Q \VS{13}“With
God are wisdom
and power;
\par }{\Q counsel
and understanding are his.
\par }{\Q \VS{14}If
he tears
down, it cannot
be rebuilt;
\par }{\Q if he imprisons
a person,
there is no
escape.
\par }{\Q \VS{15}If
he holds
back the waters,
then
they dry
up;

\par }{\Q if he releases
them, they destroy
the land.
\par }{\Q \VS{16}With
him are strength
and prudence;
\par }{\Q both the one who goes astray
\par }{\Q and the one who misleads are his.
\par }{\Q \VS{17}He leads
counselors
away stripped
\par }{\Q and makes judges
into fools.
\par }{\Q \VS{18}He loosens
the bonds
of kings
\par }{\Q and binds
a loincloth
around their waist.
\par }{\Q \VS{19}He leads
priests
away stripped
\par }{\Q and overthrows
the potentates.
\par }{\Q \VS{20}He deprives
the trusted
advisers of speech
\par }{\Q and takes
away the discernment
of elders.
\par }{\Q \VS{21}He pours
contempt
on
noblemen
\par }{\Q and disarms
the powerful.
\par }{\Q \VS{22}He reveals
the deep
things of darkness,
\par }{\Q and brings
deep shadows
into the light.
\par }{\Q \VS{23}He makes nations
great,
and destroys
them;
\par }{\Q he extends
the boundaries of nations
\par }{\Q and disperses them.
\par }{\Q \VS{24}He deprives
the leaders
of the earth
\par }{\Q of their understanding;

\par }{\Q he makes them wander
\par }{\Q in a trackless desert waste.
\par }{\Q \VS{25}They grope
about in darkness
without
light;
\par }{\Q he makes them stagger
like drunkards.


\par }\Chap{13}{\PP \par }{\Q \VerseOne{1}“Indeed,
my eyes
have seen
all
this,

\par }{\Q my ears
have heard
and understood it.
\par }{\Q \VS{2}What you know,
I know
also;
\par }{\Q I
am
not
inferior to you!
\par }{\Q \VS{3}But
I
wish to
speak
to
the Almighty,
\par }{\Q and I desire
to argue
my case with God.
\par }{\Q \VS{4}But
you,
however,
are inventors
of lies;
\par }{\Q all
of you are worthless
physicians!
\par }{\Q \VS{5}If only
you would
keep completely silent!

\par }{\Q For you, that would be
wisdom.
\par }{\Q \VS{6}“Listen
now
to my argument,
\par }{\Q and be attentive
to my lips’
contentions.
\par }{\Q \VS{7}Will you speak
wickedly on God’s
behalf?

\par }{\Q Will you speak
deceitfully for him?
\par }{\Q \VS{8}Will you
show him partiality?

\par }{\Q Will
you argue
the case for God?
\par }{\Q \VS{9}Would it
turn out
well
if
he would examine
you?
\par }{\Q Or as one deceives
a man
would you deceive him?
\par }{\Q \VS{10}He would certainly
rebuke
you
\par }{\Q if
you secretly
showed partiality!
\par }{\Q \VS{11}Would not
his
splendor terrify
you
\par }{\Q and the fear
he inspires
fall
on you?
\par }{\Q \VS{12}Your maxims
are proverbs
of ashes;
\par }{\Q your defenses
are defenses
of clay.
\par }{\Q \VS{13}“Refrain
from
talking
with me
so that I
may speak;
\par }{\Q then let
come to me what may.
\par }{\Q \VS{14}Why
do I put myself in peril,
\par }{\Q and take my life
in my hands?
\par }{\Q \VS{15}Even if
he slays
me, I will hope
in him;

\par }{\Q I will surely
defend
my ways
to
his face!
\par }{\Q \VS{16}Moreover,
this
will become my deliverance,
\par }{\Q for
no
godless
person would come
before him.
\par }{\Q \VS{17}Listen
carefully to my words;
\par }{\Q let your ears
be attentive to my explanation.
\par }{\Q \VS{18}See
now,
I have prepared
my case;
\par }{\Q I know
that
I am
right.
\par }{\Q \VS{19}Who
will contend
with me?
\par }{\Q If
anyone can, I will be silent
and die.
\par }{\Q \VS{20}Only
in two
things spare
me, O God,
\par }{\Q and then
I will not
hide
from your face:
\par }{\Q \VS{21}Remove your hand
far
from me
\par }{\Q and stop
making me afraid
with your terror.
\par }{\Q \VS{22}Then call,
and I
will answer,
\par }{\Q or
I will speak,
and you respond to me.
\par }{\Q \VS{23}How
many are my iniquities
and sins?
\par }{\Q Show
me my transgression
and my sin.
\par }{\Q \VS{24}Why
do you hide
your face
\par }{\Q and regard
me as your enemy?
\par }{\Q \VS{25}Do you wish to torment
a windblown
leaf
\par }{\Q and chase
after dry
chaff?
\par }{\Q \VS{26}For
you write
down bitter things
against me
\par }{\Q and cause me to inherit
the sins
of my youth.
\par }{\Q \VS{27}And you put
my feet
in the stocks
\par }{\Q and you watch
all
my movements;
\par }{\Q you put marks
on
the soles
of my feet.
\par }{\Q \VS{28}So I waste away
like something rotten,
\par }{\Q like a garment
eaten
by moths.


\par }\Chap{14}{\PP \par }{\Q \VerseOne{1}“Man,
born
of woman,
\par }{\Q lives but a few
days,
and they are full
of trouble.
\par }{\Q \VS{2}He grows
up like a flower
and then withers
away;
\par }{\Q he flees
like a shadow,
and does not
remain.
\par }{\Q \VS{3}Do
you fix
your eye
on
such
a one?

\par }{\Q And do you bring
me before you for judgment?
\par }{\Q \VS{4}Who can
make
a clean
thing come from an unclean?

\par }{\Q No
one!
\par }{\Q \VS{5}Since
man’s
days
are determined,

\par }{\Q the number
of his months
is under your control;
\par }{\Q you have set his limit
and he
cannot
pass it.
\par }{\Q \VS{6}Look away
from him
and let him desist,
\par }{\Q until
he fulfills his time
like a hired man.
\par }{\SH The Inevitability of Death
\par }{\Q \VS{7}“But there is
hope
for a tree:

\par }{\Q If
it is cut down,
it will sprout
again,
\par }{\Q and its new shoots
will not
fail.
\par }{\Q \VS{8}Although
its roots
may grow old
in the ground
\par }{\Q and its stump
begins to die
in the soil,
\par }{\Q \VS{9}at
the scent
of water
it will flourish
\par }{\Q and put forth shoots
like
a new plant.
\par }{\Q \VS{10}But man
dies
and is powerless;
\par }{\Q he
expires
– and where is he?
\par }{\Q \VS{11}As water
disappears
from
the sea,
\par }{\Q or a river
drains away
and dries up,
\par }{\Q \VS{12}so man
lies
down and does not
rise;
\par }{\Q until
the heavens
are no
more,
\par }{\Q they will not
awake
\par }{\Q nor arise from their sleep.
\par }{\SH The Possibility of Another Life
\par }{\Q \VS{13}“O that you would
hide me
in Sheol,
\par }{\Q and conceal
me till
your anger
has passed!

\par }{\Q O that you would set me a time
\par }{\Q and then remember me!
\par }{\Q \VS{14}If
a man
dies,
will he live
again?

\par }{\Q All
the days
of my hard service
I will wait
\par }{\Q until
my release
comes.
\par }{\Q \VS{15}You will call
and I –
I will answer
you;
\par }{\Q you will long for the creature you have made.
\par }{\SH The Present Condition
\par }{\Q \VS{16}“Surely
now
you count
my steps;
\par }{\Q then you would not
mark my
sin.
\par }{\Q \VS{17}My offenses
would be sealed
up in a bag;
\par }{\Q you would cover over
my sin.
\par }{\Q \VS{18}But as
a mountain
falls
away
and crumbles,
\par }{\Q and as a rock
will be removed from its place,
\par }{\Q \VS{19}as water
wears
away stones,
\par }{\Q and torrents
wash away
the soil,
\par }{\Q so you destroy
man’s
hope.
\par }{\Q \VS{20}You overpower
him once for all,
\par }{\Q and he departs;
\par }{\Q you change his appearance
\par }{\Q and send him away.
\par }{\Q \VS{21}If his sons
are honored,
\par }{\Q he does not
know
it;

\par }{\Q if they are brought low,
\par }{\Q he does not
see it.
\par }{\Q \VS{22}Only
his flesh
has pain
for himself,

\par }{\Q and he mourns
for himself.”

\par }\Chap{15}{\PP \VerseOne{1}Then
Eliphaz
the Temanite
answered:
\par }{\Q \VS{2}“Does a wise
man answer
with blustery
knowledge,
\par }{\Q or fill
his belly
with the east wind?
\par }{\Q \VS{3}Does he argue
with useless
talk,
\par }{\Q with words
that have no
value in them?
\par }{\Q \VS{4}But you
even
break
off
piety,
\par }{\Q and hinder
meditation
before
God.
\par }{\Q \VS{5}Your sin
inspires
your mouth;
\par }{\Q you choose
the language
of the crafty.
\par }{\Q \VS{6}Your own mouth
condemns
you, not
I;
\par }{\Q your own lips
testify against you.
\par }{\Q \VS{7}“Were you the first
man
ever born?
\par }{\Q Were you brought forth
before
the hills?
\par }{\Q \VS{8}Do you listen
in on God’s
secret
council?

\par }{\Q Do you limit
wisdom to yourself?
\par }{\Q \VS{9}What
do you know
that we don’t know?
\par }{\Q What do you understand
that we don’t understand?
\par }{\Q \VS{10}The gray-haired
and the aged
are on our side,

\par }{\Q men far older
than your father.
\par }{\Q \VS{11}Are God’s
consolations
too trivial
for
you;

\par }{\Q or a word
spoken in gentleness
to you?
\par }{\Q \VS{12}Why
has your heart
carried
you away,

\par }{\Q and why
do your eyes
flash,
\par }{\Q \VS{13}when
you turn
your rage against God
\par }{\Q and allow
such words
to escape from
your mouth?
\par }{\Q \VS{14}What
is man
that
he should be pure,
\par }{\Q or one born
of woman,
that
he should be righteous?
\par }{\Q \VS{15}If
God places no
trust
in his holy ones,
\par }{\Q if even the heavens
are not
pure
in his eyes,
\par }{\Q \VS{16}how much less
man,
who is abominable
and corrupt,
\par }{\Q who drinks
in evil
like water!
\par }{\Q \VS{17}“I will explain
to you;
\par }{\Q listen
to me,
\par }{\Q and what I have
seen,
I will declare,
\par }{\Q \VS{18}what
wise
men declare,
\par }{\Q hiding nothing,
\par }{\Q from the tradition of their ancestors,
\par }{\Q \VS{19}to whom alone
the land
was given
\par }{\Q when no
foreigner
passed
among them.
\par }{\Q \VS{20}All
his days
the wicked
man suffers torment,
\par }{\Q throughout the number
of the years
\par }{\Q that are stored
up for the tyrant.
\par }{\Q \VS{21}Terrifying
sounds
fill his ears;
\par }{\Q in a time of peace
marauders
attack him.
\par }{\Q \VS{22}He does not
expect
to escape
from
darkness;
\par }{\Q he is marked
for the sword;
\par }{\Q \VS{23}he wanders about – food for vultures;
\par }{\Q he knows that the day of darkness is at hand.
\par }{\Q \VS{24}Distress
and anguish
terrify
him;
\par }{\Q they prevail against
him
\par }{\Q like a king
ready
to launch an attack,
\par }{\Q \VS{25}for
he stretches
out his hand
against God,
\par }{\Q and vaunts
himself against the Almighty,
\par }{\Q \VS{26}defiantly charging
against him

\par }{\Q with a thick,
strong
shield!
\par }{\Q \VS{27}Because
he covered
his face
with fat,
\par }{\Q and made
his hips
bulge with fat,
\par }{\Q \VS{28}he lived
in ruined
towns
\par }{\Q and in houses
where no
one lives,
\par }{\Q where
they are ready to crumble
into heaps.
\par }{\Q \VS{29}He will not
grow rich,
\par }{\Q and his
wealth
will not
endure,
\par }{\Q nor will his possessions
spread
over the land.
\par }{\Q \VS{30}He will not
escape
the darkness;
\par }{\Q a flame
will wither
his shoots
\par }{\Q and he will depart
\par }{\Q by the breath
of God’s mouth.
\par }{\Q \VS{31}Let him not
trust
in what is worthless,
\par }{\Q deceiving
himself;
\par }{\Q for
worthlessness
will be
his reward.
\par }{\Q \VS{32}Before his time
he will be paid in full,
\par }{\Q and his branches
will not
flourish.
\par }{\Q \VS{33}Like a vine
he will let his sour grapes
fall,

\par }{\Q and like an olive
tree
\par }{\Q he will shed his blossoms.
\par }{\Q \VS{34}For
the company
of the godless
is barren,
\par }{\Q and fire
consumes
the tents
of those who accept bribes.
\par }{\Q \VS{35}They conceive
trouble
and bring forth
evil;
\par }{\Q their belly
prepares
deception.”


\par }\Chap{16}{\PP \VerseOne{1}Then
Job
replied:
\par }{\Q \VS{2}“I have heard
many
things like
these
before.
\par }{\Q What miserable
comforters
are you all!
\par }{\Q \VS{3}Will there be an end
to your windy
words?

\par }{\Q Or
what
provokes
you that
you answer?
\par }{\Q \VS{4}I
also
could speak
like you,
\par }{\Q if
you were in my
place;
\par }{\Q I could
pile
up words
against
you
\par }{\Q and I could shake
my head at you.
\par }{\Q \VS{5}But I would strengthen
you with
my words;
\par }{\Q comfort
from my lips
would bring you relief.
\par }{\SH Abandonment by God and Man
\par }{\Q \VS{6}“But if
I speak,
my pain
is not
relieved,
\par }{\Q and if I refrain from speaking
\par }{\Q – how much of it goes away?
\par }{\Q \VS{7}Surely
now
he has worn
me out,
\par }{\Q you have devastated
my entire
household.
\par }{\Q \VS{8}You have seized
me,

\par }{\Q and it has become
a witness;
\par }{\Q my leanness
has risen up against
me
\par }{\Q and testifies against me.
\par }{\Q \VS{9}His anger
has torn
me and persecuted
me;
\par }{\Q he has gnashed
at
me with his teeth;
\par }{\Q my adversary
locks
his eyes on me.
\par }{\Q \VS{10}People have opened
their mouths
against me,
\par }{\Q they have struck
my cheek
in scorn;
\par }{\Q they unite
together
against me.
\par }{\Q \VS{11}God
abandons
me to evil men,
\par }{\Q and throws
me into
the hands
of wicked men.
\par }{\Q \VS{12}I was
in peace,
and he has shattered
me.

\par }{\Q He has seized
me by the neck
and crushed
me.

\par }{\Q He has made
me his target;
\par }{\Q \VS{13}his archers
surround
me.
\par }{\Q Without pity
he pierces
my kidneys
\par }{\Q and pours out
my gall
on the ground.
\par }{\Q \VS{14}He breaks through
against
me, time
and time
again;
\par }{\Q he rushes
against
me like a warrior.
\par }{\Q \VS{15}I have sewed
sackcloth
on
my skin,
\par }{\Q and buried my horn
in the dust;
\par }{\Q \VS{16}my face
is reddened
because of weeping,
\par }{\Q and on
my eyelids
there is a deep darkness,
\par }{\Q \VS{17}although there is no
violence
in my hands
\par }{\Q and my prayer
is pure.
\par }{\SH An Appeal to God as Witness
\par }{\Q \VS{18}“O earth,
do not
cover
my blood,
\par }{\Q nor let
there be
a secret place
for my cry.
\par }{\Q \VS{19}Even
now
my witness
is
in heaven;
\par }{\Q my advocate
is on high.
\par }{\Q \VS{20}My intercessor
is my friend
\par }{\Q as my eyes
pour out tears to God;
\par }{\Q \VS{21}and he contends
with
God
on behalf of man
\par }{\Q as a man
pleads for his friend.
\par }{\Q \VS{22}For
the years
that lie
ahead
are few,
\par }{\Q and then I will go on the way
of no
return.


\par }\Chap{17}{\PP \par }{\Q \VerseOne{1}My spirit
is broken,

\par }{\Q my days
have faded
out,

\par }{\Q the grave
awaits me.
\par }{\Q \VS{2}Surely
mockery
is with me;
\par }{\Q my eyes must dwell on their hostility.
\par }{\Q \VS{3}Make
then my pledge
with
you.
\par }{\Q Who
else
will put up security for me?
\par }{\Q \VS{4}Because
you have closed
their minds
to understanding,
\par }{\Q therefore
you will not
exalt them.
\par }{\Q \VS{5}If
a man
denounces
his friends
for personal gain,
\par }{\Q the eyes
of his children
will fail.
\par }{\Q \VS{6}He has made me a byword
to people,
\par }{\Q I am the one in whose face
they spit.
\par }{\Q \VS{7}My eyes
have grown dim
with grief;
\par }{\Q my whole
frame
is but a shadow.
\par }{\Q \VS{8}Upright
men are appalled
at this;
\par }{\Q the innocent
man is troubled
with the godless.
\par }{\Q \VS{9}But the righteous
man holds
to his way,
\par }{\Q and the one with clean
hands
grows
stronger.
\par }{\SH Anticipation of Death
\par }{\Q \VS{10}“But
turn,
all
of you, and come
now!

\par }{\Q I will not
find
a wise man among you.
\par }{\Q \VS{11}My days
have passed,
my plans are shattered,
\par }{\Q even the desires
of my heart.
\par }{\Q \VS{12}These men change night
into day;
\par }{\Q they say, ‘The light
is near
\par }{\Q in the face
of darkness.’
\par }{\Q \VS{13}If
I hope
for the grave
to be my home,
\par }{\Q if I spread
out my bed
in darkness,
\par }{\Q \VS{14}If I cry
to corruption, ‘You
are my father,’
\par }{\Q and to the worm,
‘My Mother,’
or ‘My sister,’
\par }{\Q \VS{15}where
then
is my hope?
\par }{\Q And my hope,
who
sees it?
\par }{\Q \VS{16}Will it go down
to the barred gates of death?
\par }{\Q Will we descend
together
into the dust?”

\par }\Chap{18}{\PP \VerseOne{1}Then
Bildad
the Shuhite
answered:
\par }{\Q \VS{2}“How long
until
you make
an end
of words?

\par }{\Q You must consider,
and then we can talk.
\par }{\Q \VS{3}Why
should we be regarded
as beasts,
\par }{\Q and considered stupid
in your sight?
\par }{\Q \VS{4}You who tear
yourself
to pieces
in your
anger,
\par }{\Q will the earth
be abandoned
for your sake?
\par }{\Q Or will a rock
be moved
from its place?
\par }{\Q \VS{5}“Yes,
the lamp
of the wicked
is extinguished;
\par }{\Q his flame
of fire
does not
shine.
\par }{\Q \VS{6}The light
in his tent
grows dark;
\par }{\Q his lamp
above
him is extinguished.
\par }{\Q \VS{7}His vigorous
steps
are restricted,
\par }{\Q and his own counsel
throws
him down.
\par }{\Q \VS{8}For
he has
been thrown into a net
by his feet
\par }{\Q and he wanders into a mesh.
\par }{\Q \VS{9}A trap
seizes
him by the heel;
\par }{\Q a snare
grips him.
\par }{\Q \VS{10}A rope
is hidden
for him on the ground
\par }{\Q and a trap
for him lies on
the path.
\par }{\Q \VS{11}Terrors
frighten
him on all sides
\par }{\Q and dog his every step.
\par }{\Q \VS{12}Calamity is hungry
for him,
\par }{\Q and misfortune
is ready
at his side.
\par }{\Q \VS{13}It eats
away parts
of his skin;
\par }{\Q the most terrible
death
devours
his limbs.
\par }{\Q \VS{14}He is dragged
from the security
of his tent,
\par }{\Q and marched
off to the king
of terrors.
\par }{\Q \VS{15}Fire resides in
his tent;
\par }{\Q over
his residence burning sulfur
is scattered.
\par }{\Q \VS{16}Below
his roots
dry
up,
\par }{\Q and his branches
wither
above.
\par }{\Q \VS{17}His memory
perishes
from
the earth,
\par }{\Q he has no
name
in
the land.
\par }{\Q \VS{18}He is driven
from light
into darkness
\par }{\Q and is banished
from the world.
\par }{\Q \VS{19}He has neither
children
nor
descendants
among his people,
\par }{\Q no
survivor
in those places he once stayed.
\par }{\Q \VS{20}People of the west
are appalled
at his fate;
\par }{\Q people of the east
are seized
with horror, saying,
\par }{\Q \VS{21}‘Surely
such
is the residence
of an evil man;
\par }{\Q and this
is the place
of one who has not
known
God.’ ”


\par }\Chap{19}{\PP \VerseOne{1}Then
Job
answered:
\par }{\Q \VS{2}“How long
will you torment
me

\par }{\Q and crush
me with your words?
\par }{\Q \VS{3}These
ten
times
you have been reproaching
me;

\par }{\Q you are not
ashamed
to attack me!
\par }{\Q \VS{4}But even
if it were true
that I have erred,
\par }{\Q my error
remains solely my concern!
\par }{\Q \VS{5}If
indeed
you would exalt yourselves above
me
\par }{\Q and plead
my disgrace
against me,
\par }{\Q \VS{6}know
then
that
God
has wronged
me

\par }{\Q and encircled
me with his net.
\par }{\SH Job’s Abandonment and Affliction
\par }{\Q \VS{7}“If
I cry out, ‘Violence!’

\par }{\Q I receive no
answer;
\par }{\Q I cry for help,
\par }{\Q but there is no
justice.
\par }{\Q \VS{8}He has blocked
my way
so I cannot
pass,
\par }{\Q and has set
darkness
over
my paths.
\par }{\Q \VS{9}He has stripped
me of my honor
\par }{\Q and has taken the crown
off
my head.
\par }{\Q \VS{10}He tears
me down
on every side
until I perish;
\par }{\Q he uproots
my hope
like one uproots a tree.
\par }{\Q \VS{11}Thus his anger
burns
against
me,
\par }{\Q and he considers
me among his enemies.
\par }{\Q \VS{12}His troops
advance
together;
\par }{\Q they throw up
a siege ramp
against
me,
\par }{\Q and they camp
around
my tent.
\par }{\SH Job’s Forsaken State
\par }{\Q \VS{13}“He has put my relatives
far
from me;
\par }{\Q my acquaintances
only
turn away from me.
\par }{\Q \VS{14}My kinsmen
have failed
me;
\par }{\Q my friends
have forgotten me.
\par }{\Q \VS{15}My guests
and my servant girls
\par }{\Q consider
me a stranger;
\par }{\Q I am a foreigner
in their eyes.
\par }{\Q \VS{16}I summon
my servant,
but he does not
respond,
\par }{\Q even though I implore
him with my own mouth.
\par }{\Q \VS{17}My breath
is repulsive
to my wife;
\par }{\Q I am loathsome
to my brothers.
\par }{\Q \VS{18}Even
youngsters
have scorned
me;
\par }{\Q when I
get
up, they scoff at me.
\par }{\Q \VS{19}All
my closest
friends detest
me;
\par }{\Q and those whom I love
have turned against me.
\par }{\Q \VS{20}My bones
stick
to my skin
and my flesh;
\par }{\Q I have escaped
alive with only the skin
of my teeth.
\par }{\Q \VS{21}Have pity
on me, my friends,
have pity on me,
\par }{\Q for
the hand
of God
has struck me.
\par }{\Q \VS{22}Why
do you pursue
me like
God
does?

\par }{\Q Will you never be satiated
with my flesh?
\par }{\SH Job’s Assurance of Vindication
\par }{\Q \VS{23}“O
that my words
were written
down,
\par }{\Q O that they were written
on a scroll,
\par }{\Q \VS{24}that with an iron
chisel
and with lead
\par }{\Q they were engraved
in a rock
forever!
\par }{\Q \VS{25}As for me, I
know
that my Redeemer
lives,
\par }{\Q and that as the last
\par }{\Q he will stand
upon the earth.
\par }{\Q \VS{26}And after
my skin
has been destroyed,
\par }{\Q yet
in my flesh
I will see
God,
\par }{\Q \VS{27}whom
I
will see
for myself,
\par }{\Q and whom my own eyes
will behold,
\par }{\Q and not
another.
\par }{\Q My heart grows faint within me.
\par }{\Q \VS{28}If
you say,
‘How
we will pursue
him,
\par }{\Q since the root
of the trouble
is found in him!’
\par }{\Q \VS{29}Fear
the sword
yourselves,
\par }{\Q for
wrath
brings the punishment
by the sword,
\par }{\Q so that you may
know
\par }{\Q that there is judgment.”


\par }\Chap{20}{\PP \VerseOne{1}Then Zophar
the Naamathite
answered:
\par }{\Q \VS{2}“This
is why
my troubled thoughts
bring me back – ‘span class=”s 05668” title=”05668” ’
\par }{\Q because of
my feelings within me.
\par }{\PP \par }{\Q \VS{3}When I hear
a
reproof
that dishonors
me,
\par }{\Q then my understanding
prompts
me to answer.
\par }{\Q \VS{4}“Surely
you know
that it has been from
old,
\par }{\Q ever
since
humankind
was placed
on
the earth,
\par }{\Q \VS{5}that
the elation
of the wicked
is brief,
\par }{\Q the joy
of the godless
lasts
but a moment.
\par }{\Q \VS{6}Even though
his stature
reaches
to the heavens
\par }{\Q and his head
touches
the clouds,
\par }{\Q \VS{7}he will perish
forever,
like his own excrement;
\par }{\Q those who used to see
him will say,
‘Where is he?’
\par }{\Q \VS{8}Like a dream
he flies
away,
never
again to be found,
\par }{\Q and like a vision
of the night
he is put to flight.
\par }{\Q \VS{9}People who had seen
him will not
see him again,
\par }{\Q and the place
where he was
\par }{\Q will recognize
him no longer.
\par }{\Q \VS{10}His sons
must recompense
the poor;
\par }{\Q his own hands
must return
his wealth.
\par }{\Q \VS{11}His bones
were full
of his youthful
vigor,
\par }{\Q but that vigor will lie
down with him
in the dust.
\par }{\Q \VS{12}“If
evil
is
sweet
in his mouth
\par }{\Q and he hides
it under
his tongue,
\par }{\Q \VS{13}if he retains
it for
himself
\par }{\Q and does not
let it go,
\par }{\Q and holds
it fast
in
his mouth,
\par }{\Q \VS{14}his food
is turned
sour
in his stomach;
\par }{\Q it becomes the venom of serpents
within him.
\par }{\Q \VS{15}The wealth
that he consumed he vomits
up,
\par }{\Q God
will make him throw it out of his stomach.
\par }{\Q \VS{16}He sucks the poison
of serpents;
\par }{\Q the fangs
of a viper
kill him.
\par }{\Q \VS{17}He will not
look
on the streams,
\par }{\Q the rivers,
which are the torrents
\par }{\Q of honey
and butter.
\par }{\Q \VS{18}He gives back
the ill-gotten gain
\par }{\Q without
assimilating
it;

\par }{\Q he will not
enjoy
the wealth from his commerce.
\par }{\Q \VS{19}For
he has oppressed
the poor
and abandoned
them;

\par }{\Q he has seized
a house
which he did not
build.
\par }{\Q \VS{20}For
he knows
no
satisfaction
in his appetite;
\par }{\Q he does not
let anything he desires
escape.
\par }{\Q \VS{21}“Nothing
is left
for him to devour;
\par }{\Q that is why
his prosperity
does not
last.
\par }{\Q \VS{22}In the fullness
of his sufficiency,
\par }{\Q distress overtakes him.
\par }{\Q the full force
of misery
will come upon him.
\par }{\Q \VS{23}“While
he is filling
his belly,
\par }{\Q God sends
his burning
anger
against him,
\par }{\Q and rains down
his blows upon him.
\par }{\Q \VS{24}If
he flees
from an iron
weapon,
\par }{\Q then an arrow from a bronze
bow
pierces him.
\par }{\Q \VS{25}When he pulls
it out
and it comes out of his back,
\par }{\Q the gleaming
point out of his liver,
\par }{\Q terrors
come
over him.
\par }{\Q \VS{26}Total
darkness
waits
to receive his treasures;
\par }{\Q a fire
which has not
been kindled
\par }{\Q will consume
him
\par }{\Q and devour
what is left
in his tent.
\par }{\Q \VS{27}The heavens
reveal
his iniquity;
\par }{\Q the earth
rises up against him.
\par }{\Q \VS{28}A flood
will carry
off his house,
\par }{\Q rushing
waters on the day
of God’s wrath.
\par }{\Q \VS{29}Such
is the lot God
allots the wicked,
\par }{\Q and the heritage
of his appointment
from God.”


\par }\Chap{21}{\PP \VerseOne{1}Then
Job
answered:
\par }{\Q \VS{2}“Listen
carefully
to my words;
\par }{\Q let this
be the consolation you offer me.
\par }{\Q \VS{3}Bear
with me and I
will speak,
\par }{\Q and after
I have spoken
you may mock.
\par }{\Q \VS{4}Is my
complaint
against a man?

\par }{\Q If
so, why
should I not
be impatient?
\par }{\Q \VS{5}Look
at me
and be appalled;
\par }{\Q put
your hands
over
your mouths.
\par }{\Q \VS{6}For, when I think
about this, I am terrified
\par }{\Q and my body
feels
a shudder.
\par }{\SH The Wicked Prosper
\par }{\Q \VS{7}“Why
do the wicked
go on living,
\par }{\Q grow
old, even
increase
in power?
\par }{\Q \VS{8}Their children
are firmly established
\par }{\Q in their presence,
\par }{\Q their offspring
before their eyes.
\par }{\Q \VS{9}Their houses
are safe
and without fear;
\par }{\Q and no
rod
of punishment from God
is upon them.
\par }{\Q \VS{10}Their bulls
breed without fail;

\par }{\Q their
cows
calve
and do not
miscarry.
\par }{\Q \VS{11}They allow
their children
to run like a flock;
\par }{\Q their little ones
dance about.
\par }{\Q \VS{12}They sing
to the accompaniment of tambourine
and harp,
\par }{\Q and make merry
to the sound
of the flute.
\par }{\Q \VS{13}They live out
their years
in prosperity
\par }{\Q and go down
to the grave
in peace.
\par }{\Q \VS{14}So they say
to God,
‘Turn
away from us!
\par }{\Q We do not
want
to know
your ways.
\par }{\Q \VS{15}Who is the Almighty,
that
we should
serve
him?
\par }{\Q What
would we gain
\par }{\Q if
we were to pray to him?’
\par }{\Q \VS{16}But their prosperity
is not their own doing.
\par }{\Q The counsel
of the wicked
is far
from me!
\par }{\SH How Often Do the Wicked Suffer?
\par }{\Q \VS{17}“How
often is the lamp
of the wicked
extinguished?
\par }{\Q How often does their misfortune
come
upon
them?
\par }{\Q How often does God apportion
pain to them
in his anger?
\par }{\Q \VS{18}How often are they like straw
before
the wind,
\par }{\Q and like chaff
swept away by a whirlwind?
\par }{\Q \VS{19}You may say, ‘God
stores
up a man’s
punishment for
his children!’

\par }{\Q Instead let him repay
the man himself

\par }{\Q so that he may know it!
\par }{\Q \VS{20}Let his own eyes
see
his destruction;
\par }{\Q let him drink
of the anger
of the Almighty.
\par }{\Q \VS{21}For
what
is his interest
in his
home
\par }{\Q after
his death,

\par }{\Q when
the number
of his
months
\par }{\Q has been broken off?
\par }{\Q \VS{22}Can anyone teach
God
knowledge,
\par }{\Q since he judges
those that are on high?
\par }{\SH Death Levels Everything
\par }{\Q \VS{23}“One
man dies
in his full
vigor,
\par }{\Q completely
secure
and prosperous,
\par }{\Q \VS{24}his body
well nourished,
\par }{\Q and the marrow
of his bones
moist.
\par }{\Q \VS{25}And another man
dies
in bitterness
of soul,
\par }{\Q never
having tasted
anything good.
\par }{\Q \VS{26}Together
they lie
down in the dust,
\par }{\Q and worms
cover
over them both.
\par }{\SH Futile Words, Deceptive Answers
\par }{\Q \VS{27}“Yes,
I know
what you are thinking,
\par }{\Q the schemes
by which you would wrong me.
\par }{\Q \VS{28}For
you say,
\par }{\Q ‘Where
now is the nobleman’s
house,
\par }{\Q and where
are the tents
in which
the wicked
lived?’
\par }{\Q \VS{29}Have you never questioned
those who travel
the roads?
\par }{\Q Do you not
recognize
their accounts –
\par }{\Q \VS{30}that
the evil
man is spared
\par }{\Q from the day
of his misfortune,
\par }{\Q that he is delivered

\par }{\Q from the day
of God’s wrath?
\par }{\Q \VS{31}No one
denounces
his conduct
to his face;
\par }{\Q no one
repays
him for what he has done.
\par }{\Q \VS{32}And when he is
carried
to the tombs,
\par }{\Q and watch
is kept over the funeral mound,
\par }{\Q \VS{33}The clods
of the torrent valley
are sweet
to him;
\par }{\Q behind
him everybody
follows in procession,
\par }{\Q and before
him goes a countless throng.
\par }{\Q \VS{34}So how
can you console
me with your futile
words?
\par }{\Q Nothing is left
of your answers
but deception!”


\par }\Chap{22}{\PP \VerseOne{1}Then
Eliphaz
the Temanite
answered:
\par }{\Q \VS{2}“Is it to God
that
a strong man is of benefit?
\par }{\Q Is it to him that
even
a wise
man
is profitable?
\par }{\Q \VS{3}Is
it of any special benefit
to the Almighty
\par }{\Q that
you should be righteous,
\par }{\Q or
is it any gain
to him
\par }{\Q that
you make your ways blameless?
\par }{\Q \VS{4}Is it because of your piety
that he rebukes
you
\par }{\Q and goes
to judgment
with you?
\par }{\Q \VS{5}Is not
your wickedness
great
\par }{\Q and is there no
end
to your iniquity?
\par }{\Q \VS{6}“For
you took pledges
from your brothers
\par }{\Q for no reason,
\par }{\Q and you stripped
the clothing
from the naked.
\par }{\Q \VS{7}You gave the weary
no
water
to drink
\par }{\Q and from the hungry
you withheld
food.
\par }{\Q \VS{8}Although you were a powerful man,
owning
land,
\par }{\Q an honored man living on it,
\par }{\Q \VS{9}you sent
widows
away empty-handed,
\par }{\Q and the arms
of the orphans
you crushed.
\par }{\Q \VS{10}That is why
snares
surround
you,
\par }{\Q and why sudden
fear
terrifies you,
\par }{\Q \VS{11}why it is so
dark
you cannot
see,
\par }{\Q and why a flood
of water
covers you.
\par }{\Q \VS{12}“Is not
God
on high
in heaven?

\par }{\Q And see
the lofty
stars, how high they are!
\par }{\Q \VS{13}But you have said,
‘What
does God
know?
\par }{\Q Does he judge
through
such deep darkness?
\par }{\Q \VS{14}Thick clouds
are a veil
for him, so he does not
see
us,

\par }{\Q as he goes back and forth
\par }{\Q in the vault
of heaven.’
\par }{\Q \VS{15}Will you keep
to the old
path
\par }{\Q that
evil
men
have
walked –
\par }{\Q \VS{16}men who were carried
off before their time,
\par }{\Q when the flood
was poured
out

\par }{\Q on their foundations?
\par }{\Q \VS{17}They were saying
to God,
‘Turn
away from us,’
\par }{\Q and ‘What
can the Almighty
do to us?’
\par }{\Q \VS{18}But it was he who filled
their houses
\par }{\Q with good things –
\par }{\Q yet the counsel of the wicked
\par }{\Q was far from me.
\par }{\Q \VS{19}The righteous
see
their destruction and rejoice;
\par }{\Q the innocent
mock them scornfully, saying,
\par }{\Q \VS{20}‘Surely
our enemies
are destroyed,
\par }{\Q and fire
consumes their wealth.’
\par }{\Q \VS{21}“Reconcile
yourself with
God,

\par }{\Q and be at peace
with him;
\par }{\Q in this way your prosperity will be good.
\par }{\Q \VS{22}Accept
instruction
from his mouth
\par }{\Q and store up
his words
in your heart.
\par }{\Q \VS{23}If
you return
to
the Almighty,
you will be built
up;

\par }{\Q if you remove wicked behavior
far
from your tent,
\par }{\Q \VS{24}and throw
your gold
in
the dust
–
\par }{\Q your gold
of Ophir
\par }{\Q among the rocks
in the ravines –
\par }{\Q \VS{25}then the Almighty
himself will be your gold,
\par }{\Q and the choicest
silver for you.
\par }{\Q \VS{26}Surely
then
you will delight
yourself in the Almighty,
\par }{\Q and will lift
up your face
toward
God.
\par }{\Q \VS{27}You will pray
to him and he will hear
you,
\par }{\Q and you will fulfill
your vows to him.
\par }{\Q \VS{28}Whatever
you decide on a matter,
\par }{\Q it will be established
for you,
\par }{\Q and light
will shine
on
your ways.
\par }{\Q \VS{29}When
people are brought low
and you say
\par }{\Q ‘Lift them up!’

\par }{\Q then he will save
the downcast;
\par }{\Q \VS{30}he will deliver
even someone
who is not innocent,
\par }{\Q who will escape
through the cleanness
of your hands.”


\par }\Chap{23}{\PP \VerseOne{1}Then
Job
answered:
\par }{\Q \VS{2}“Even
today
my complaint
is still bitter;
\par }{\Q his hand
is heavy
despite
my groaning.
\par }{\Q \VS{3}O that I knew
where I might find
him,

\par }{\Q that I could come
to his place of residence!
\par }{\Q \VS{4}I would lay out my case
before
him
\par }{\Q and fill
my mouth
with arguments.
\par }{\Q \VS{5}I would know
with what words
he would answer
me,
\par }{\Q and understand
what
he would say to me.
\par }{\Q \VS{6}Would
he contend
with me with
great
power?
\par }{\Q No,
he would only
pay
attention
to me.
\par }{\Q \VS{7}There
an upright
person
\par }{\Q could present his case
before him,
\par }{\Q and I would be delivered
forever
from my judge.
\par }{\SH The Inaccessibility and Power of God
\par }{\Q \VS{8}“If
I go
to the east,
he is not
there,
\par }{\Q and to the west,
yet I do not
perceive him.
\par }{\Q \VS{9}In the north
when he is at
work,

\par }{\Q I do
not
see
him;

\par }{\Q when he turns to the south,
\par }{\Q I see
no
trace of him.
\par }{\Q \VS{10}But
he knows
the pathway
that I take;
\par }{\Q if he tested
me, I would come forth
like gold.
\par }{\Q \VS{11}My feet
have followed
his steps
closely;
\par }{\Q I have kept
to his way
and have not
turned aside.
\par }{\Q \VS{12}I have not
departed
from the commands
of his lips;
\par }{\Q I have treasured
the words
of his mouth
more than my allotted portion.
\par }{\Q \VS{13}But he is
unchangeable,
and who can
change
him?
\par }{\Q Whatever he has
desired,
he does.
\par }{\Q \VS{14}For
he fulfills
his decree
against me,

\par }{\Q and many
such things are his plans.
\par }{\Q \VS{15}That is why
I am terrified
in his presence;
\par }{\Q when I consider,
I am afraid because of him.
\par }{\Q \VS{16}Indeed, God
has made my heart
faint;
\par }{\Q the Almighty
has terrified me.
\par }{\Q \VS{17}Yet
I have not
been silent
because of the darkness,
\par }{\Q because of the thick darkness
\par }{\Q that covered
my face.


\par }\Chap{24}{\PP \par }{\Q \VerseOne{1}“Why
are times
not
appointed
by the Almighty?

\par }{\Q Why do those
who know
him not
see
his days?
\par }{\Q \VS{2}Men move boundary
stones;
\par }{\Q they seize
the flock
and pasture them.
\par }{\Q \VS{3}They drive away
the orphan’s
donkey;
\par }{\Q they take the widow’s
ox
as a pledge.
\par }{\Q \VS{4}They turn
the needy
from the pathway,
\par }{\Q and the poor
of the land
hide
themselves together.
\par }{\Q \VS{5}Like wild donkeys
in the desert
\par }{\Q they go out
to their labor,
\par }{\Q seeking diligently
for food;
\par }{\Q the wasteland
provides food
for them
\par }{\Q and for their children.
\par }{\Q \VS{6}They reap
fodder
in the field,
\par }{\Q and glean
in the vineyard
of the wicked.
\par }{\Q \VS{7}They spend
the night naked
because they lack
clothing;
\par }{\Q they have no
covering
against the cold.
\par }{\Q \VS{8}They are soaked by mountain
rains
\par }{\Q and huddle
in the rocks
because
they lack shelter.
\par }{\Q \VS{9}The fatherless child
is snatched
from the breast,
\par }{\Q the infant
of the poor
is taken as a pledge.
\par }{\Q \VS{10}They go
about naked,
without
clothing,
\par }{\Q and go hungry
while they carry
the sheaves.
\par }{\Q \VS{11}They press out the olive oil
between
the rows
of olive trees;

\par }{\Q they tread
the winepresses
while they are thirsty.
\par }{\Q \VS{12}From the city
the dying
groan,
\par }{\Q and the wounded
cry out for help,
\par }{\Q but God
charges
no
one with wrongdoing.
\par }{\Q \VS{13}There
are those
who rebel
against the light;
\par }{\Q they do not
know
its ways
\par }{\Q and they do not
stay
on its paths.
\par }{\Q \VS{14}Before daybreak
the murderer
rises
up;
\par }{\Q he kills
the poor
and the needy;
\par }{\Q in the night
he is like a thief.
\par }{\Q \VS{15}And the eye
of the adulterer
watches
for the twilight,
\par }{\Q thinking, ‘No
eye can see me,’
\par }{\Q and covers
his face
with a mask.
\par }{\Q \VS{16}In the dark
the robber
breaks into houses,
\par }{\Q but by day
they shut
themselves in;

\par }{\Q they do not
know
the light.
\par }{\Q \VS{17}For
all
of them, the morning
is to them
\par }{\Q like deep darkness;
\par }{\Q they are friends with the terrors
of darkness.
\VS{18}\par }{\Q “You say, ‘He is
foam
on
the face
of the waters;
\par }{\Q their portion
of the land
is cursed
\par }{\Q so that
no
one goes to their vineyard.
\par }{\Q \VS{19}The drought
as well
as the heat
carry away
\par }{\Q the melted
snow;
\par }{\Q so the grave
takes away those who
have sinned.
\par }{\Q \VS{20}The womb
forgets
him,
\par }{\Q the worm
feasts
on him,
\par }{\Q no
longer
will he be remembered.
\par }{\Q Like a tree,
wickedness
will be broken down.
\par }{\Q \VS{21}He preys
on the barren
and childless
woman,

\par }{\Q and does not
treat the widow
well.
\par }{\Q \VS{22}But God drags off
the mighty
by his power;
\par }{\Q when God rises up
against him, he has no
faith
in his life.
\par }{\Q \VS{23}God may let
them rest in a feeling
of security,
\par }{\Q but he is constantly watching
all their ways.
\par }{\Q \VS{24}They are exalted
for a little
while,
\par }{\Q and then they are gone,
\par }{\Q they are brought low
like all
others,
\par }{\Q and gathered in,
\par }{\Q and like a head
of grain
they are cut off.’
\par }{\Q \VS{25}“If
this is not
so, who can
prove me a liar
\par }{\Q and reduce
my words to nothing?”

\par }\Chap{25}{\PP \VerseOne{1}Then
Bildad
the Shuhite
answered:
\par }{\Q \VS{2}“Dominion
and awesome might
belong to God;
\par }{\Q he establishes
peace
in his heights.
\par }{\Q \VS{3}Can
his armies
be numbered?

\par }{\Q On
whom
does his light
not
rise?
\par }{\Q \VS{4}How
then can a human
being be righteous
before God?
\par }{\Q How
can one born
of a woman
be pure?
\par }{\Q \VS{5}If
even
the moon
is not
bright,
\par }{\Q and the stars
are not
pure
as far as he is concerned,
\par }{\Q \VS{6}how much less
a mortal man,
who is but a maggot – ‘span class=”s 01121” title=”01121” ’
\par }{\Q a son
of man,
who is only a worm!”
\par }{\PP 

\par }\Chap{26}{\PP \VerseOne{1}Then
Job
replied:
\par }{\Q \VS{2}“How
you have helped
the powerless!

\par }{\Q How you have
saved
the person
who has
no
strength!
\par }{\Q \VS{3}How
you have advised
the one without
wisdom,
\par }{\Q and abundantly
revealed
your insight!
\par }{\Q \VS{4}To whom
did you utter
these words?
\par }{\Q And whose
spirit has
come forth
from your mouth?
\par }{\SH A Better Description of God’s Greatness
\par }{\Q \VS{5}“The dead
tremble –
\par }{\Q those beneath
the waters
\par }{\Q and all that live in them.
\par }{\Q \VS{6}The underworld
is naked
before
God;

\par }{\Q the place of destruction
lies uncovered.
\par }{\Q \VS{7}He spreads out
the northern
skies over
empty
space;

\par }{\Q he suspends the earth
on
nothing.
\par }{\Q \VS{8}He locks
the waters
in his clouds,
\par }{\Q and the clouds
do not
burst
with the weight of them.
\par }{\Q \VS{9}He conceals
the face
of the full moon,
\par }{\Q shrouding
it with
his clouds.
\par }{\Q \VS{10}He marks out
the horizon
on
the surface
of the waters
\par }{\Q as
a boundary
between
light
and darkness.
\par }{\Q \VS{11}The pillars
of the heavens
tremble
\par }{\Q and are amazed
at his rebuke.
\par }{\Q \VS{12}By his power
he stills
the sea;
\par }{\Q by his wisdom
he cut
Rahab the great sea monster to pieces.
\par }{\Q \VS{13}By his breath
the skies
became fair;
\par }{\Q his hand
pierced
the fleeing
serpent.
\par }{\Q \VS{14}Indeed,
these
are but the outer fringes
of his ways!

\par }{\Q How
faint is the whisper
we hear
of him!
\par }{\Q But who can
understand
the thunder
of his power?”


\par }\Chap{27}{\PP \VerseOne{1}And Job
took up
his discourse
again:
\par }{\Q \VS{2}“As surely
as God
lives, who has denied
me justice,
\par }{\Q the Almighty,
who has made my life
bitter –
\par }{\Q \VS{3}for
while my spirit is
still
in me,
\par }{\Q and the breath
from God
is in my nostrils,
\par }{\Q \VS{4}my lips
will not
speak
wickedness,
\par }{\Q and my tongue
will whisper
no deceit.
\par }{\Q \VS{5}I will never
declare that you three are in the right;
\par }{\Q until
I die,
I will not
set aside
my integrity!
\par }{\Q \VS{6}I will maintain
my righteousness
\par }{\Q and never
let it go;
\par }{\Q my conscience
will not
reproach
me
\par }{\Q for as
long
as
I live.
\par }{\SH The Condition of the Wicked
\par }{\Q \VS{7}“May
my enemy
be
like the wicked,
\par }{\Q my adversary like the unrighteous.
\par }{\Q \VS{8}For
what
hope
does the godless
have when
he is cut off,
\par }{\Q when
God
takes away
his life?
\par }{\Q \VS{9}Does God
listen
to his cry
\par }{\Q when
distress
overtakes him?
\par }{\Q \VS{10}Will he find delight
in the Almighty?
\par }{\Q Will he call
out to God
at all
times?
\par }{\Q \VS{11}I will teach
you about the power
of God;
\par }{\Q What
is on
the Almighty’s
mind I will not
conceal.
\par }{\Q \VS{12}If
you yourselves
have all
seen
this,
\par }{\Q Why
in the world do you continue this
meaningless talk?
\par }{\Q \VS{13}This
is the portion
of the wicked
man
\par }{\Q allotted
by God,
\par }{\Q the inheritance
that evildoers
receive
\par }{\Q from the Almighty.
\par }{\Q \VS{14}If
his children
increase
– it is for
the sword!

\par }{\Q His offspring
never
have enough
to eat.
\par }{\Q \VS{15}Those who survive
him are buried
by the plague,

\par }{\Q and their widows
do not
mourn for them.
\par }{\Q \VS{16}If
he piles
up silver
like dust
\par }{\Q and stores
up clothing
like mounds of clay,
\par }{\Q \VS{17}what
he stores
up a righteous
man will wear,
\par }{\Q and an innocent
man will inherit
his silver.
\par }{\Q \VS{18}The house
he builds
is as fragile as a moth’s
cocoon,
\par }{\Q like a hut
that a watchman has made.
\par }{\Q \VS{19}He goes to bed
wealthy,
but will do so no
more.
\par }{\Q When he opens
his eyes,
it is all gone.
\par }{\Q \VS{20}Terrors
overwhelm
him like a flood;
\par }{\Q at night
a whirlwind carries him off.
\par }{\Q \VS{21}The east wind
carries
him away, and he is gone;
\par }{\Q it sweeps
him out of his place.
\par }{\Q \VS{22}It hurls itself
against
him without
pity
\par }{\Q as he flees
headlong
from its power.
\par }{\Q \VS{23}It claps
its hands
at him in derision
\par }{\Q and hisses
him away from his place.

\par }\Chap{28}{\PP \par }{\Q \VerseOne{1}“Surely
there
is a mine
for silver,
\par }{\Q and a place
where gold
is refined.
\par }{\Q \VS{2}Iron
is taken
from the ground,
\par }{\Q and rock
is poured out
as copper.
\par }{\Q \VS{3}Man puts
an end
to the darkness;
\par }{\Q he searches
the farthest
recesses
\par }{\Q for the ore
in the deepest darkness.
\par }{\Q \VS{4}Far from where people live he sinks
a shaft,
\par }{\Q in places travelers
have long forgotten,
\par }{\Q far from
other people
he dangles
and sways.
\par }{\Q \VS{5}The earth,
from
which food
comes,
\par }{\Q is overturned
below
as
though by fire;
\par }{\Q \VS{6}a place
whose stones
are sapphires
\par }{\Q and which contains dust
of gold;
\par }{\Q \VS{7}a hidden path
no
bird of prey
knows
–
\par }{\Q no
falcon’s
eye has spotted it.
\par }{\Q \VS{8}Proud
beasts
have not
set foot
on it,
\par }{\Q and no
lion has passed along it.
\par }{\Q \VS{9}On the flinty
rock man has set
to work with his hand;
\par }{\Q he has overturned
mountains
at their bases.
\par }{\Q \VS{10}He has cut out channels
through the rocks;
\par }{\Q his eyes
have spotted
every
precious thing.
\par }{\Q \VS{11}He has searched the sources of the rivers
\par }{\Q and what was hidden
he has brought into
the light.
\par }{\SH No Price Can Buy Wisdom
\par }{\Q \VS{12}“But wisdom
– where
can it be found?
\par }{\Q Where
is
the place
of understanding?
\par }{\Q \VS{13}Mankind
does not
know
its place;
\par }{\Q it cannot
be found
in the land
of the living.
\par }{\Q \VS{14}The deep
says,
‘It is
not
with me.’
\par }{\Q And the sea
says,
‘It is not
with me.’
\par }{\Q \VS{15}Fine gold
cannot
be given
in exchange
for it,
\par }{\Q nor
can its price
be weighed
out in silver.
\par }{\Q \VS{16}It cannot
be measured out
for purchase with the gold
of Ophir,
\par }{\Q with precious
onyx
or sapphires.
\par }{\Q \VS{17}Neither
gold
nor crystal
can be compared
with it,
\par }{\Q nor can a vase
of gold
match its worth.
\par }{\Q \VS{18}Of coral
and jasper
no
mention
will be made;
\par }{\Q the price
of wisdom
is more than pearls.
\par }{\Q \VS{19}The topaz
of Cush
cannot
be compared
with it;
\par }{\Q it cannot
be purchased
with pure
gold.
\par }{\SH God Alone Has Wisdom
\par }{\Q \VS{20}“But wisdom
– where
does it come
from?

\par }{\Q Where
is
the place
of understanding?
\par }{\Q \VS{21}For it has been hidden
\par }{\Q from the eyes
of every
living creature,
\par }{\Q and from the birds
of the sky
it has been concealed.
\par }{\Q \VS{22}Destruction
and Death
say,
\par }{\Q ‘With our ears
we have heard
a rumor about where it can be found.’
\par }{\Q \VS{23}God
understands
the way
to it,
\par }{\Q and he alone knows
its place.
\par }{\Q \VS{24}For
he
looks
to the ends
of the earth
\par }{\Q and observes
everything
under
the heavens.
\par }{\Q \VS{25}When he made
the force
of the wind
\par }{\Q and measured
the waters
with a gauge.
\par }{\Q \VS{26}When
he imposed
a limit
for the rain,
\par }{\Q and a path
for the thunderstorm,
\par }{\Q \VS{27}then
he looked
at wisdom and assessed
its value;

\par }{\Q he established
it and examined
it closely.
\par }{\Q \VS{28}And he said
to mankind,
\par }{\Q ‘The fear
of the
{\ND{Lord}} –
that is
wisdom,
\par }{\Q and to turn
away from evil
is understanding.’ ”


\par }\Chap{29}{\PP \VerseOne{1}Then
Job
continued his speech:
\par }{\Q \VS{2}“O
that I
could be as I was
\par }{\Q in the months
now gone,
\par }{\Q in the days
when God
watched over me,
\par }{\Q \VS{3}when he caused
his lamp
\par }{\Q to shine upon
my head,
\par }{\Q and by his light
\par }{\Q I walked
through darkness;
\par }{\Q \VS{4}just
as I was
in my most productive
time,
\par }{\Q when God’s
intimate friendship
was experienced in
my tent,
\par }{\Q \VS{5}when the Almighty
was still
with me
\par }{\Q and my children
were around me;
\par }{\Q \VS{6}when my steps
were bathed with butter
\par }{\Q and the rock
poured out for me
streams
of olive oil!
\par }{\Q \VS{7}When I went out
to the city gate
\par }{\Q and secured
my seat
in the public
square,
\par }{\Q \VS{8}the young men
would see
me and step aside,
\par }{\Q and the old men
would get up
and remain standing;
\par }{\Q \VS{9}the chief
men refrained
from talking
\par }{\Q and covered their mouths
with their hands;
\par }{\Q \VS{10}the voices
of the nobles fell silent,

\par }{\Q and their tongues
stuck
to the roof of their mouths.
\par }{\SH Job’s Benevolence
\par }{\Q \VS{11}“As soon as the ear
heard
these things, it blessed
me,

\par }{\Q and when the eye
saw
them, it bore witness to me,
\par }{\Q \VS{12}for
I rescued
the poor
who cried out for help,
\par }{\Q and the orphan
who had no
one to assist him;
\par }{\Q \VS{13}the blessing
of the dying man descended
on
me,
\par }{\Q and I made the widow’s
heart
rejoice;
\par }{\Q \VS{14}I put on
righteousness
and it clothed
me,

\par }{\Q my just dealing
was like a robe
and a turban;
\par }{\Q \VS{15}I
was eyes
for the blind
\par }{\Q and feet
for the lame;
\par }{\Q \VS{16}I was a father
to the needy,
\par }{\Q and I
investigated
the case
of the person I did not
know;
\par }{\Q \VS{17}I broke
the fangs
of the wicked,
\par }{\Q and made him drop
his prey
from his teeth.
\par }{\SH Job’s Confidence
\par }{\Q \VS{18}“Then I thought,
‘I will die
in my own home,

\par }{\Q my
days
as numerous
as the grains
of sand.
\par }{\Q \VS{19}My roots
reach
the water,
\par }{\Q and the dew
lies on my branches
all night long.
\par }{\Q \VS{20}My glory
will always be fresh
in me,
\par }{\Q and my bow
ever new
in my hand.’
\par }{\SH Job’s Reputation
\par }{\Q \VS{21}“People listened
to me
and waited
silently;
\par }{\Q they kept silent
for my advice.
\par }{\Q \VS{22}After
I had
spoken,
they did not
respond;
\par }{\Q my words
fell on them drop
by drop.
\par }{\Q \VS{23}They waited
for me as people wait for the rain,
\par }{\Q and they opened
their mouths
\par }{\Q as for the spring rains.
\par }{\Q \VS{24}If I smiled
at
them, they hardly
believed
it;

\par }{\Q and they did not
cause the light
of my face
to darken.
\par }{\Q \VS{25}I chose
the way
for them

\par }{\Q and sat
as their chief;
\par }{\Q I lived
like a king
among his troops;
\par }{\Q I was like one who
comforts
mourners.


\par }\Chap{30}{\PP \par }{\Q \VerseOne{1}“But now
they mock
me, those who are younger
than
I,
\par }{\Q whose
fathers
I disdained
too much

\par }{\Q to put with
my sheep
dogs.
\par }{\Q \VS{2}Moreover,
the strength
of their hands
–
\par }{\Q what
use was it to me?
\par }{\Q Men whose strength
had perished;
\par }{\Q \VS{3}gaunt
with want
and hunger,
\par }{\Q they would gnaw
the parched land,
\par }{\Q in former time
desolate
and waste.
\par }{\Q \VS{4}By the brush
they would gather herbs
from
the salt marshes,
\par }{\Q and the root
of the broom tree
was their food.
\par }{\Q \VS{5}They were banished
from
the community –
\par }{\Q people shouted
at them
\par }{\Q like they would shout
at
thieves –
\par }{\Q \VS{6}so that they had to live
\par }{\Q in the dry stream
beds,
\par }{\Q in the holes
of the ground,
and among the rocks.
\par }{\Q \VS{7}They brayed
like animals among the bushes
\par }{\Q and were huddled
together under
the nettles.
\par }{\Q \VS{8}Sons
of senseless
and nameless
people,
\par }{\Q they were driven out
of the land with whips.
\par }{\SH Job’s Indignities
\par }{\Q \VS{9}“And now
I have become their taunt song;
\par }{\Q I have become
a byword
among them.
\par }{\Q \VS{10}They detest
me
and maintain their distance;
\par }{\Q they do not
hesitate
to spit
in my face.
\par }{\Q \VS{11}Because
God has untied
my tent cord
and afflicted
me,
\par }{\Q people throw off
all restraint
in my presence.
\par }{\Q \VS{12}On
my right
the young rabble
rise up;
\par }{\Q they drive
me from place to place,
\par }{\Q and build up
siege ramps
against me.
\par }{\Q \VS{13}They destroy my path;
\par }{\Q they succeed in destroying me
\par }{\Q without
anyone assisting them.
\par }{\Q \VS{14}They come
in as through a wide
breach;
\par }{\Q amid
the crash
they come rolling in.
\par }{\Q \VS{15}Terrors
are turned
loose on
me;
\par }{\Q they drive
away my honor
like the wind,
\par }{\Q and like a cloud
my deliverance
has passed away.
\par }{\SH Job’s Despondency
\par }{\Q \VS{16}“And now
my soul
pours
itself out within me;

\par }{\Q days
of suffering
take hold of me.
\par }{\Q \VS{17}Night
pierces my bones;
\par }{\Q my gnawing
pains never cease.
\par }{\Q \VS{18}With great
power
God grasps
my clothing;
\par }{\Q he binds
me like the collar
of my tunic.
\par }{\Q \VS{19}He
has flung
me into the mud,
\par }{\Q and I have come to resemble
dust
and ashes.
\par }{\Q \VS{20}I cry out
to
you, but you do not
answer
me;
\par }{\Q I stand
up, and you only
look at me.
\par }{\Q \VS{21}You have become
cruel
to me;

\par }{\Q with the strength
of your hand
you attack me.
\par }{\Q \VS{22}You pick
me up on the wind
and make me ride
on it;

\par }{\Q you toss
me about in the storm.
\par }{\Q \VS{23}I know
that you are bringing
me to death,
\par }{\Q to the meeting
place for all
the living.
\par }{\SH The Contrast With the Past
\par }{\Q \VS{24}“Surely
one does not
stretch out
his hand
\par }{\Q against a broken
man

\par }{\Q when
he cries for help
in his distress.
\par }{\Q \VS{25}Have I not
wept
for the unfortunate?

\par }{\Q Was not my soul
grieved
for the poor?
\par }{\Q \VS{26}But when
I hoped
for good,
trouble
came;
\par }{\Q when I expected
light,
then darkness
came.
\par }{\Q \VS{27}My heart
is in turmoil unceasingly;
\par }{\Q the days
of my affliction
confront me.
\par }{\Q \VS{28}I
go
about blackened, but not
by the sun;
\par }{\Q in the assembly
I stand
up and cry for help.
\par }{\Q \VS{29}I have become
a brother
to jackals
\par }{\Q and a companion
of ostriches.
\par }{\Q \VS{30}My skin
has turned dark
on
me;

\par }{\Q my body
is hot
with
fever.
\par }{\Q \VS{31}My harp
is used for mourning
\par }{\Q and my flute
for the sound
of weeping.

\par }\Chap{31}{\PP \par }{\Q \VerseOne{1}“I made
a covenant
with my eyes;
\par }{\Q how
then could
I entertain
thoughts against
a virgin?
\par }{\Q \VS{2}What
then
would be one’s lot
from God
above,
\par }{\Q one’s heritage
from the Almighty
on high?
\par }{\Q \VS{3}Is it not
misfortune
for the unjust,
\par }{\Q and disaster
for those who work
iniquity?
\par }{\Q \VS{4}Does he not
see
my ways
\par }{\Q and count
all
my steps?
\par }{\Q \VS{5}If
I have walked
in falsehood,
\par }{\Q and if my foot
has hastened
to deceit –
\par }{\Q \VS{6}let him weigh
me with honest
scales;
\par }{\Q then God
will discover
my integrity.
\par }{\Q \VS{7}If
my footsteps have strayed
from
the way,
\par }{\Q if my heart
has gone
after
my eyes,
\par }{\Q or if anything
has defiled my hands,
\par }{\Q \VS{8}then let me sow
and let another
eat,
\par }{\Q and let my crops
be uprooted.
\par }{\Q \VS{9}If
my heart
has been enticed
by
a woman,
\par }{\Q and I have lain in wait
at my neighbor’s
door,
\par }{\Q \VS{10}then let my wife
turn the millstone
for another
man,
\par }{\Q and may other men have sexual relations with her.
\par }{\Q \VS{11}For
I would have committed a shameful
act,

\par }{\Q an iniquity
to be judged.
\par }{\Q \VS{12}For
it is a fire
that
devours
even
to Destruction,
\par }{\Q and it would uproot
all
my harvest.
\par }{\Q \VS{13}“If
I have disregarded
the right
of my male servants
\par }{\Q or my female servants
\par }{\Q when
they disputed with me,
\par }{\Q \VS{14}then
what
will I do
when
God
confronts
me in judgment;

\par }{\Q when
he intervenes,
\par }{\Q how
will I respond to him?
\par }{\Q \VS{15}Did not
the one who made
me in the womb
make
them?

\par }{\Q Did
not the same
one form
us in the womb?
\par }{\Q \VS{16}If
I have refused
to give the poor
what they desired,
\par }{\Q or caused the eyes
of the widow
to fail,
\par }{\Q \VS{17}If I ate
my morsel
of bread myself,
\par }{\Q and did not
share
any of it with
orphans –
\par }{\Q \VS{18}but
from my youth
I raised
the orphan like a father,
\par }{\Q and from my mother’s
womb
\par }{\Q I guided the widow!
\par }{\Q \VS{19}If
I have seen
anyone about to perish
for lack
of clothing,
\par }{\Q or a poor
man without
a coat,
\par }{\Q \VS{20}whose
heart did not
bless
me

\par }{\Q as he warmed
himself with the fleece
of my sheep,
\par }{\Q \VS{21}if
I have raised my hand
to vote
against
the orphan,
\par }{\Q when
I saw
my support
in the court,
\par }{\Q \VS{22}then let my arm
fall
from the shoulder,
\par }{\Q let my arm
be broken
off at the socket.
\par }{\Q \VS{23}For
the calamity
from God
was a terror
to me,
\par }{\Q and by reason of his majesty I was powerless.
\par }{\Q \VS{24}“If
I have put
my confidence
in gold
\par }{\Q or said
to pure gold,
\par }{\Q ‘You are my security!’
\par }{\Q \VS{25}if
I have rejoiced
because
of the extent
of my wealth,
\par }{\Q or because
of the great
wealth
my hand had gained,
\par }{\Q \VS{26}if
I looked
at the sun
when
it was shining,
\par }{\Q and the moon
advancing
as a precious thing,
\par }{\Q \VS{27}so that my heart
was secretly
enticed,
\par }{\Q and my hand
threw them a kiss
from my mouth,
\par }{\Q \VS{28}then this also
would be iniquity
to be judged,
\par }{\Q for
I would have been false
to God
above.
\par }{\Q \VS{29}If
I have rejoiced
over the misfortune
of my enemy

\par }{\Q or exulted because
calamity
found him –
\par }{\Q \VS{30}I have not
even permitted
my mouth
to sin
\par }{\Q by asking
for his life
through a curse –
\par }{\Q \VS{31}if
the members
of my household
have never
said,
\par }{\Q ‘If only there were
someone
\par }{\Q who
has not
been satisfied
from Job’s meat!’ –
\par }{\Q \VS{32}But no
stranger
had to spend
the night outside,
\par }{\Q for I opened
my doors
to the traveler –
\par }{\Q \VS{33}if
I have covered
my transgressions
as men
do,

\par }{\Q by hiding
iniquity
in my heart,
\par }{\Q \VS{34}because
I was terrified
of the great
multitude,
\par }{\Q and the contempt
of families
terrified
me,
\par }{\Q so that I remained silent
\par }{\Q and would not
go
outdoors –

\par }{\SH Job’s Appeal
\par }{\Q \VS{35}“If only
I
had someone
to hear
me!
\par }{\Q Here
is my signature –

\par }{\Q let the Almighty answer me!
\par }{\Q If only I had an indictment
\par }{\Q that my accuser had written.
\par }{\Q \VS{36}Surely
I would wear it proudly on
my shoulder,
\par }{\Q I would bind
it on me like a crown;
\par }{\Q \VS{37}I would give
him an accounting
of my steps;
\par }{\Q like
a prince
I would approach him.
\par }{\SH Job’s Final Solemn Oath
\par }{\Q \VS{38}“If
my land
cried out
against
me

\par }{\Q and all its furrows
wept
together,
\par }{\Q \VS{39}if
I have
eaten
its
produce without
paying,
\par }{\Q or caused the death of its owners,
\par }{\Q \VS{40}then let thorns
sprout up in place
of wheat,
\par }{\Q and in place
of barley,
weeds!”
\par }{\PP The words
of Job are ended.

\par }\Chap{32}{\PP \VerseOne{1}So
these
three
men
refused
to answer
Job
further, because
he was
righteous
in his own eyes.
\VS{2}Then Elihu
son
of Barakel
the Buzite,
of the family
of Ram,
became very angry.
He was angry
with Job
for
justifying
himself
rather than God.
\VS{3}With Job’s three
friends
he was also angry,
because
they could not
find
an answer, and so declared
Job
guilty.
\VS{4}Now Elihu
had waited
before speaking
to Job,
because
the others were
older
than he was.
\VS{5}But when Elihu
saw
that
the three
men
had no
further
reply,
he became very angry.
\par }{\SH Elihu Claims Wisdom
\par }{\PP \VS{6}So Elihu
son
of Barakel
the Buzite
spoke up:

\par }{\Q “I am
young,
but you
are elderly;
\par }{\Q that is why
I was fearful,
\par }{\Q and afraid
to explain
to you what I know.
\par }{\Q \VS{7}I said
to myself, ‘Age
should speak,
\par }{\Q and length
of years
should make wisdom
known.’
\par }{\Q \VS{8}But
it is
a spirit
in people,
\par }{\Q the breath
of the Almighty,
\par }{\Q that makes them understand.
\par }{\Q \VS{9}It is not
the aged
who are wise,
\par }{\Q nor old men
who understand
what is right.
\par }{\Q \VS{10}Therefore
I say,
‘Listen
to me.
\par }{\Q I, even I, will explain
what I
know.’
\par }{\Q \VS{11}Look,
I waited
for you to speak;
\par }{\Q I listened
closely to your wise thoughts, while
you
were searching for
words.
\par }{\Q \VS{12}Now I
was paying
you close attention,

\par }{\Q yet there was no one proving
Job
wrong,

\par }{\Q not
one of you was answering
his statements!
\par }{\Q \VS{13}So do not
say, ‘We have found
wisdom!
\par }{\Q God
will refute
him, not
man!’
\par }{\Q \VS{14}Job has not
directed
his words
to me,
\par }{\Q and so I will not
reply
to him
with your arguments.
\par }{\SH Job’s Friends Failed to Answer
\par }{\Q \VS{15}“They are
dismayed
and cannot
answer
any more;
\par }{\Q they have nothing left
to say.
\par }{\Q \VS{16}And I have waited.
But because
they do not
speak,
\par }{\Q because
they stand
there and answer
no
more,
\par }{\Q \VS{17}I
too will answer
my part,
\par }{\Q I
too will explain
what I
know.
\par }{\Q \VS{18}For
I am full
of words,
\par }{\Q and the spirit
within me constrains me.
\par }{\Q \VS{19}Inside
I am like wine
which has no
outlet,
\par }{\Q like new
wineskins
ready to burst!
\par }{\Q \VS{20}I will speak,
so that I may find relief;
\par }{\Q I will open
my lips,
so that I may answer.
\par }{\Q \VS{21}I will not
show
partiality
to
anyone,
\par }{\Q nor
will I confer a title
on any man.
\par }{\Q \VS{22}for
I do not
know
how to give honorary
titles,
\par }{\Q if I did,
my Creator
would quickly do away with me. ‘span class=”footnote” id=”footnote-37” ’‘span class=”key” ’37‘a href=”\#note-37” class=”backref” ’32:15‘/a’‘span class=”text” ’
{\IT{tn}}:
{\IT{Heb}} “words have moved away from them,” meaning words are gone from them, they have nothing left to say.

\par }\Chap{33}{\PP \par }{\Q \VerseOne{1}“But now,
O
Job,
listen
to my words,
\par }{\Q and hear
everything
I have to say!
\par }{\Q \VS{2}See
now,
I have opened
my mouth;
\par }{\Q my tongue
in my mouth
has spoken.
\par }{\Q \VS{3}My words
come from the uprightness
of my heart,
\par }{\Q and my lips
will utter knowledge
sincerely.
\par }{\Q \VS{4}The Spirit
of God
has made
me,
\par }{\Q and the breath
of the Almighty
gives me life.
\par }{\Q \VS{5}Reply to me, if
you can;
\par }{\Q set your arguments
in order before
me
\par }{\Q and take your stand!
\par }{\Q \VS{6}Look,
I am
just like you in relation
to God;
\par }{\Q I
too
have been molded
from clay.
\par }{\Q \VS{7}Therefore no
fear
of me should terrify
you,
\par }{\Q nor
should my pressure
be heavy on you.
\par }{\SH Elihu Rejects Job’s Plea of Innocence
\par }{\Q \VS{8}“Indeed,
you have said
in my hearing
\par }{\Q (I heard
the sound
of the words!):
\VS{9}\par }{\Q ‘I am
pure,
without
transgression;
\par }{\Q I am
clean
and have no
iniquity.
\VS{10}\par }{\Q Yet God finds
occasions
with me;
\par }{\Q he regards
me as his enemy!
\VS{11}\par }{\Q He puts
my feet
in shackles;
\par }{\Q he watches
closely all
my paths.’
\par }{\Q \VS{12}Now
in this,
you are not
right
– I answer
you,

\par }{\Q for
God
is greater than
a human being.
\par }{\Q \VS{13}Why
do you contend
against him,
\par }{\Q that
he does not
answer
all
a person’s words?
\par }{\SH Elihu Disagrees With Job’s View of God
\par }{\Q \VS{14}“For
God
speaks,
the first
time in one way,
\par }{\Q the second
time in another,
\par }{\Q though a person does not
perceive it.
\par }{\Q \VS{15}In a dream,
a night
vision,
\par }{\Q when deep sleep
falls
on
people
\par }{\Q as they sleep
in
their beds.
\par }{\Q \VS{16}Then
he gives a revelation
to people,
\par }{\Q and terrifies them with warnings,
\par }{\Q \VS{17}to turn
a person
from his sin,

\par }{\Q and to cover
a person’s
pride.
\par }{\Q \VS{18}He spares
a person’s life
from
corruption,
\par }{\Q his very life
from crossing over
the river.
\par }{\Q \VS{19}Or
a person is chastened
by pain
on
his bed,
\par }{\Q and with the continual
strife
of his bones,
\par }{\Q \VS{20}so that his life
loathes
food,
\par }{\Q and his soul
rejects appetizing
fare.
\par }{\Q \VS{21}His flesh
wastes away
from sight,
\par }{\Q and his bones,
which were not
seen,
\par }{\Q are easily visible.
\par }{\Q \VS{22}He draws near
to the place of corruption,
\par }{\Q and his life
to the messengers of death.
\par }{\Q \VS{23}If
there is
an angel
beside
him,
\par }{\Q one
mediator
out of a thousand,
\par }{\Q to tell
a person
what constitutes his uprightness;
\par }{\Q \VS{24}and if God is gracious
to him and says,
\par }{\Q ‘Spare
him from going down
\par }{\Q to the place of corruption,
\par }{\Q I have found
a ransom for him,’
\par }{\Q \VS{25}then his flesh
is restored
like a youth’s;
\par }{\Q he returns
to the days
of his youthful vigor.
\par }{\Q \VS{26}He entreats
God,
and God delights
in him,
\par }{\Q he sees
God’s face
with rejoicing,
\par }{\Q and God restores
to him his righteousness.
\par }{\Q \VS{27}That person sings
to others, saying:
\par }{\Q ‘I have sinned
and falsified
what is right,
\par }{\Q but I was not
punished according to what I deserved.
\par }{\Q \VS{28}He redeemed
my life
\par }{\Q from going down
to the place of corruption,
\par }{\Q and my life
sees
the light!’
\par }{\SH Elihu’s Appeal to Job
\par }{\Q \VS{29}“Indeed,
God
does
all
these
things,
\par }{\Q twice,
three
times, in his dealings with
a person,
\par }{\Q \VS{30}to turn back
his life
from
the place of corruption,
\par }{\Q that he may be enlightened
with the light
of life.
\par }{\Q \VS{31}Pay attention,
Job
– listen
to me;
\par }{\Q be silent,
and I
will speak.
\par }{\Q \VS{32}If
you have
any words,
reply
to me;
\par }{\Q speak,
for
I want
to justify you.
\par }{\Q \VS{33}If
not,
you
listen
to me;
\par }{\Q be silent,
and I will teach
you wisdom.”

\par }\Chap{34}{\PP \VerseOne{1}Elihu
answered:
\par }{\Q \VS{2}“Listen
to my words,
you wise
men;
\par }{\Q hear
me, you learned men.
\par }{\Q \VS{3}For
the ear
assesses
words
\par }{\Q as the mouth
tastes
food.
\par }{\Q \VS{4}Let us evaluate
for ourselves what is right;
\par }{\Q let us come to know
among
ourselves what
is good.
\par }{\Q \VS{5}For
Job
says,
‘I am innocent,
\par }{\Q but God
turns
away my right.
\par }{\Q \VS{6}Concerning
my right,
should I lie?

\par }{\Q My wound
is incurable,
\par }{\Q although I am without
transgression.’
\par }{\Q \VS{7}What
man
is like Job,
\par }{\Q who drinks
derision
like water!
\par }{\Q \VS{8}He goes about
in company
with
evildoers,
\par }{\Q he goes along
with
wicked
men.
\par }{\Q \VS{9}For
he says,
‘It does not
profit
a man
\par }{\Q when he makes his delight
with
God.’
\par }{\SH God is Not Unjust
\par }{\Q \VS{10}“Therefore,
listen
to me, you men
of understanding.
\par }{\Q Far be it
from God
to do wickedness,
\par }{\Q from the Almighty
to do evil.
\par }{\Q \VS{11}For
he repays
a person
for his work,
\par }{\Q and according to the conduct of a person,
\par }{\Q he causes the consequences
to find him.
\par }{\Q \VS{12}Indeed,
in truth,
God
does not
act wickedly,
\par }{\Q and the Almighty
does not
pervert
justice.
\par }{\Q \VS{13}Who
entrusted
to him
the earth?
\par }{\Q And who
put
him over the whole
world?
\par }{\Q \VS{14}If
God were to
set
his heart
on it,

\par }{\Q and gather
in
his spirit
and his breath,
\par }{\Q \VS{15}all
flesh
would perish
together
\par }{\Q and human
beings would return
to dust.
\par }{\SH God Is Impartial and Omniscient
\par }{\Q \VS{16}“If
you have understanding,
listen
to this,
\par }{\Q hear
what
I have to say.
\par }{\Q \VS{17}Do
you really
think

\par }{\Q that one who hates
justice
can govern?

\par }{\Q And will you declare guilty
\par }{\Q the supremely
righteous One,
\par }{\Q \VS{18}who says
to a king, ‘Worthless
man’

\par }{\Q and to nobles,
‘Wicked men,’
\par }{\Q \VS{19}who
shows no
partiality
to princes,
\par }{\Q and does not
take note of the rich more than the poor,
\par }{\Q because
all
of them are the work
of his hands?
\par }{\Q \VS{20}In a moment
they die,
in the middle
of the night,
\par }{\Q people
are shaken
and they pass
away.
\par }{\Q The mighty
are removed
effortlessly.
\par }{\Q \VS{21}For
his eyes
are on
the ways
of an individual,
\par }{\Q he observes
all
a person’s steps.
\par }{\Q \VS{22}There is no
darkness,
and no
deep darkness,
\par }{\Q where
evildoers
can hide
themselves.
\par }{\Q \VS{23}For
he does not
still
consider
a person,
\par }{\Q that he should come
before
God
in judgment.
\par }{\Q \VS{24}He shatters
the great
without
inquiry,
\par }{\Q and sets up
others
in their place.
\par }{\Q \VS{25}Therefore,
he knows
their deeds,
\par }{\Q he overthrows
them in the night
\par }{\Q and they are crushed.
\par }{\Q \VS{26}He strikes
them for their wickedness,
\par }{\Q in a place
where people can see,
\par }{\Q \VS{27}because
they have turned
away from following
him,
\par }{\Q and have not
understood
any
of his ways,
\par }{\Q \VS{28}so that they caused the cry
of the poor
\par }{\Q to come
before him,
\par }{\Q so that he hears
the cry
of the needy.
\par }{\Q \VS{29}But if God is quiet,
who
can condemn
him?
\par }{\Q If he hides
his face,
then who
can see
him?
\par }{\Q Yet he is over
the individual
and the nation
alike,
\par }{\Q \VS{30}so that the godless
man
should not rule,
\par }{\Q and not lay snares
for the people.
\par }{\SH Job Is Foolish to Rebel
\par }{\Q \VS{31}“Has anyone said
to
God,
\par }{\Q ‘I have endured chastisement,
\par }{\Q but I will not
act wrongly any more.
\par }{\Q \VS{32}Teach
me what
I cannot
see.
\par }{\Q If
I have done
evil,
I will do so no
more.’
\par }{\Q \VS{33}Is it
your opinion
that
God should recompense
it,
\par }{\Q because
you reject
this?

\par }{\Q But you
must choose,
and not
I,
\par }{\Q so tell
us what
you know.
\par }{\Q \VS{34}Men
of understanding
say to me –
\par }{\Q any wise man listening to me says –
\par }{\Q \VS{35}that Job
speaks
without
knowledge
\par }{\Q and his words
are without
understanding.
\par }{\Q \VS{36}But Job
will be tested
to
the end,
\par }{\Q because
his answers
are like those of wicked
men.
\par }{\Q \VS{37}For
he adds
transgression
to his sin;
\par }{\Q in our midst he claps
his hands,

\par }{\Q and multiplies
his words
against God.”


\par }\Chap{35}{\PP \VerseOne{1}Then
Elihu
answered:
\par }{\Q \VS{2}“Do you think
this to be just:
\par }{\Q when you say,
‘My right
before God.’
\par }{\Q \VS{3}But
you say,
‘What
will it profit
you,’

\par }{\Q and, ‘What
do I gain
by not sinning?’
\par }{\Q \VS{4}I
will reply
to you,

\par }{\Q and to your friends
with you.
\par }{\Q \VS{5}Gaze
at the heavens
and see;
\par }{\Q consider
the clouds,
which are higher
than you!
\par }{\Q \VS{6}If
you sin,
how
does
it affect God?

\par }{\Q If your transgressions
are many,
\par }{\Q what
does it do to him?
\par }{\Q \VS{7}If
you are righteous,
what
do you give
to God,
\par }{\Q or
what
does he receive
from your hand?
\par }{\Q \VS{8}Your wickedness
affects only a person
like yourself,
\par }{\Q and your righteousness
only other people.
\par }{\Q \VS{9}“People cry out
\par }{\Q because of the excess
of oppression;
\par }{\Q they cry out for help
\par }{\Q because of the power
of the mighty.
\par }{\Q \VS{10}But no
one says,
‘Where
is God,
my Creator,
\par }{\Q who gives
songs
in the night,
\par }{\Q \VS{11}who teaches
us more than the wild animals
of the earth,
\par }{\Q and makes us wiser
than the birds
of the sky?’
\par }{\Q \VS{12}Then they cry out
– but he does not
answer
–
\par }{\Q because
of the arrogance
of the wicked.
\par }{\Q \VS{13}Surely
it is an empty cry –
God
does not
hear
it;
\par }{\Q the Almighty
does not
take notice of it.
\par }{\Q \VS{14}How much
less,
then,
\par }{\Q when
you say
that you do not
perceive
him,
\par }{\Q that the case is before
him
\par }{\Q and you are waiting for him!
\par }{\Q \VS{15}And further,
when you say
\par }{\Q that
his anger
does not
punish,
\par }{\Q and that he does not
know
transgression!
\par }{\Q \VS{16}So Job
opens
his mouth
to no purpose;
\par }{\Q without
knowledge
he multiplies
words.”


\par }\Chap{36}{\PP \VerseOne{1}Elihu
said further:
\par }{\Q \VS{2}“Be patient
with me a little longer
\par }{\Q and I
will instruct
you,
\par }{\Q for
I still
have words
to speak on God’s behalf.
\par }{\Q \VS{3}With my knowledge I will speak comprehensively,
\par }{\Q and to my Creator I
will ascribe righteousness.
\par }{\Q \VS{4}For
in truth,
my words
are not
false;
\par }{\Q it is one complete
in knowledge
\par }{\Q who is with you.
\par }{\Q \VS{5}Indeed,
God
is mighty;
and he does not
despise
people,

\par }{\Q he is mighty,
and firm in his intent.
\par }{\Q \VS{6}He does not
allow the wicked
to live,
\par }{\Q but he gives
justice
to the poor.
\par }{\Q \VS{7}He does not
take
his eyes
off
the righteous;
\par }{\Q but with
kings
on the throne
\par }{\Q he seats
the righteous and exalts them forever.
\par }{\Q \VS{8}But if
they are bound
in chains,
\par }{\Q and held captive
by the cords
of affliction,
\par }{\Q \VS{9}then he reveals to them what they
have done,
\par }{\Q and their transgressions,
\par }{\Q that
they were
behaving proudly.
\par }{\Q \VS{10}And he reveals
this for correction,
\par }{\Q and says
that
they must turn
from evil.
\par }{\Q \VS{11}If
they obey
and serve
him,
\par }{\Q they live
out their days
in prosperity
\par }{\Q and their years
in pleasantness.
\par }{\Q \VS{12}But if
they refuse to listen,
\par }{\Q they pass over
the river of death,
\par }{\Q and expire
without
knowledge.
\par }{\Q \VS{13}The godless
at heart
nourish
anger,
\par }{\Q they do not
cry out
even when
he binds them.
\par }{\Q \VS{14}They die
in their youth,
\par }{\Q and their life
ends among the male cultic prostitutes.
\par }{\Q \VS{15}He delivers
the afflicted
by their afflictions,
\par }{\Q he reveals
himself to them
by their suffering.
\par }{\Q \VS{16}And surely,
he drew
you from the mouth
of distress,
\par }{\Q to a wide place,
unrestricted,
\par }{\Q and to
the comfort
of your table
\par }{\Q filled
with rich food.
\par }{\Q \VS{17}But now you are preoccupied
with
the judgment due the wicked,
\par }{\Q judgment
and justice
take hold of you.
\par }{\Q \VS{18}Be careful
that
no one entices
you with riches;
\par }{\Q do not
let a large
bribe
turn you aside.
\par }{\Q \VS{19}Would your wealth
sustain
you,
\par }{\Q so that you would not
be in distress,

\par }{\Q even all
your mighty
efforts?
\par }{\Q \VS{20}Do not
long for the cover
of night
\par }{\Q to drag
people
away from their homes.
\par }{\Q \VS{21}Take heed,
do not
turn
to evil,
\par }{\Q for because
of this
you have
been tested
by affliction.
\par }{\Q \VS{22}Indeed,
God
is exalted
in his power;
\par }{\Q who
is a teacher
like
him?
\par }{\Q \VS{23}Who
has prescribed
his ways
for him?
\par }{\Q Or
said
to him, ‘You have done
what is wicked’?
\par }{\Q \VS{24}Remember
to extol
his work,
\par }{\Q which
people have
praised in song.
\par }{\Q \VS{25}All
humanity
has seen
it;
\par }{\Q people
gaze on
it from afar.
\par }{\SH The Work and Wisdom of God
\par }{\Q \VS{26}“Yes,
God
is great
– beyond
our knowledge!

\par }{\Q The number
of his years
is unsearchable.
\par }{\Q \VS{27}He draws
up drops
of water;
\par }{\Q they distill
the rain
into its mist,
\par }{\Q \VS{28}which
the clouds
pour
down
\par }{\Q and shower
on
humankind
abundantly.
\par }{\Q \VS{29}Who can understand
the spreading
of the clouds,
\par }{\Q the thunderings
of his pavilion?
\par }{\Q \VS{30}See
how he scattered
his lightning
about him;
\par }{\Q he has covered
the depths
of the sea.
\par }{\Q \VS{31}It is by these that
he judges
the nations
\par }{\Q and supplies
food
in abundance.
\par }{\Q \VS{32}With his hands
he covers
the lightning,
\par }{\Q and directs
it against its target.
\VS{33}\par }{\Q His thunder
announces
the coming storm,
\par }{\Q the cattle
also, concerning the storm’s approach.


\par }\Chap{37}{\PP \par }{\Q \VerseOne{1}At
this
also
my heart
pounds
\par }{\Q and leaps
from its place.
\par }{\Q \VS{2}Listen carefully
to the thunder
of his voice,
\par }{\Q to the rumbling
that proceeds
from his mouth.
\par }{\Q \VS{3}Under
the whole
heaven
he lets
it go,
\par }{\Q even his lightning
to the far corners
of the earth.
\par }{\Q \VS{4}After
that a voice
roars;
\par }{\Q he thunders
with an exalted
voice,
\par }{\Q and he does not
hold back his lightning bolts
\par }{\Q when
his voice
is heard.
\par }{\Q \VS{5}God
thunders
with his voice
in marvelous
ways;

\par }{\Q he does
great
things
beyond our understanding.
\par }{\Q \VS{6}For
to the snow
he says,
‘Fall
to earth,’
\par }{\Q and to the torrential rains, ‘Pour down.’
\par }{\Q \VS{7}He causes
everyone
to stop working,
\par }{\Q so that all
people
may know
his work.
\par }{\Q \VS{8}The wild animals
go
to their lairs,
\par }{\Q and in their dens
they remain.
\par }{\Q \VS{9}A tempest
blows out
from
its chamber,
\par }{\Q icy
cold from the driving winds.
\par }{\Q \VS{10}The breath
of God
produces
ice,
\par }{\Q and the breadth
of the waters
freeze solid.
\par }{\Q \VS{11}He loads
the clouds
with moisture;
\par }{\Q he scatters
his lightning
through
the clouds.
\par }{\Q \VS{12}The clouds go round
in circles,
\par }{\Q wheeling about
according to his plans,
\par }{\Q to carry out
all
that
he commands
them
\par }{\Q over
the face
of the whole inhabited world.
\par }{\Q \VS{13}Whether
it is for punishment
for his land,
\par }{\Q or
whether
it is for mercy,
\par }{\Q he causes it to find its mark.
\par }{\Q \VS{14}“Pay attention
to this,
Job!
\par }{\Q Stand
still and consider
the wonders
God works.
\par }{\Q \VS{15}Do you know
how God
commands them,

\par }{\Q how he makes
lightning
flash
in his storm cloud?
\par }{\Q \VS{16}Do you know
about the balancing
of the clouds,
\par }{\Q that wondrous activity
of him who is perfect
in knowledge?
\par }{\Q \VS{17}You, whose
garments
are hot
\par }{\Q when the earth
is still because of the south wind,
\par }{\Q \VS{18}will you, with him,
spread
out the clouds,
\par }{\Q solid
as a mirror
of molten metal?
\par }{\Q \VS{19}Tell
us what
we should say
to him.
\par }{\Q We cannot
prepare
a case

\par }{\Q because
of the darkness.
\par }{\Q \VS{20}Should
he be informed
that
I want to speak?
\par }{\Q If
a man
speaks, surely
he would be swallowed up!
\par }{\Q \VS{21}But now,
the sun
cannot
be looked at –
\par }{\Q it is bright in the skies –
\par }{\Q after a wind passed and swept the clouds away.
\par }{\Q \VS{22}From the north
he comes
in golden splendor;
\par }{\Q around
God
is awesome
majesty.
\par }{\Q \VS{23}As for the Almighty,
we cannot
attain
to him!
\par }{\Q He is great in power,
\par }{\Q but justice
and abundant
righteousness
he does not
oppress.
\par }{\Q \VS{24}Therefore
people
fear
him,
\par }{\Q for he does not
regard
all
the wise
in heart.”

\par }\Chap{38}{\PP \VerseOne{1}Then
the {\ND{Lord}}
answered
Job
out of the whirlwind:
\par }{\Q \VS{2}“Who
is this
who darkens
counsel
\par }{\Q with words
without
knowledge?
\par }{\Q \VS{3}Get ready
for a difficult
task like a man;
\par }{\Q I will question
you
\par }{\Q and you will inform me!
\par }{\SH God’s questions to Job
\par }{\Q \VS{4}“Where
were you
\par }{\Q when
I laid the foundation
of the earth?
\par }{\Q Tell
me, if
you possess
understanding!
\par }{\Q \VS{5}Who
set
its measurements
– if
you know
–
\par }{\Q or
who
stretched
a measuring line across it?
\par }{\Q \VS{6}On
what
were its bases
set,
\par }{\Q or
who
laid its cornerstone –
\par }{\Q \VS{7}when
the morning
stars
sang
in chorus,
\par }{\Q and all
the sons
of God
shouted for joy?
\par }{\Q \VS{8}“Who shut
up the sea
with
doors
\par }{\Q when it burst
forth,
coming out of the womb,
\par }{\Q \VS{9}when I made
the storm clouds
its garment,
\par }{\Q and thick darkness
its swaddling band,
\par }{\Q \VS{10}when
I prescribed
its limits,
\par }{\Q and set
in place its bolts
and doors,
\par }{\Q \VS{11}when I said,
‘To
here
you may come
\par }{\Q and no
farther,
\par }{\Q here
your proud
waves
will be confined’?
\par }{\Q \VS{12}Have you ever in your life
commanded
the morning,
\par }{\Q or made the dawn
know
its place,
\par }{\Q \VS{13}that it might seize
the corners
of the earth,
\par }{\Q and shake
the wicked
out of it?
\par }{\Q \VS{14}The earth takes shape
like clay
under a seal;
\par }{\Q its features
are dyed like
a garment.
\par }{\Q \VS{15}Then from the wicked
the light
is withheld,
\par }{\Q and the arm
raised
in violence is broken.
\par }{\Q \VS{16}Have you gone
to the springs
that fill the sea,
\par }{\Q or walked
about in the recesses
of the deep?
\par }{\Q \VS{17}Have the gates
of death
been revealed
to you?

\par }{\Q Have you seen
the gates
of deepest darkness?
\par }{\Q \VS{18}Have you considered the vast expanses
of the earth?
\par }{\Q Tell
me, if
you know
it all!
\par }{\Q \VS{19}“In what direction
does light
reside,
\par }{\Q and darkness,
where
is
its place,
\par }{\Q \VS{20}that
you may take
them to
their borders
\par }{\Q and perceive
the pathways
to their homes?
\par }{\Q \VS{21}You know,
for
you were born
before
them;

\par }{\Q and the number
of your days
is great!
\par }{\Q \VS{22}Have you entered
the storehouse
of the snow,
\par }{\Q or seen
the armory
of the hail,
\par }{\Q \VS{23}which
I reserve
for the time
of trouble,
\par }{\Q for
the day
of war
and battle?
\par }{\Q \VS{24}In what direction
is
lightning
dispersed,
\par }{\Q or the east winds
scattered
over
the earth?
\par }{\Q \VS{25}Who
carves
out a channel
for the heavy rains,
\par }{\Q and a path
for the rumble
of thunder,
\par }{\Q \VS{26}to cause it to rain
on
an uninhabited land,
\par }{\Q a desert
where there are no
human
beings,
\par }{\Q \VS{27}to satisfy
a devastated
and desolate
land,
\par }{\Q and to cause it to sprout
with vegetation?
\par }{\Q \VS{28}Does
the rain
have a father,
\par }{\Q or
who
has fathered
the drops
of the dew?
\par }{\Q \VS{29}From whose
womb
does the ice
emerge,
\par }{\Q and the frost
from the sky,
who
gives birth to it,
\par }{\Q \VS{30}when the waters
become hard
like stone,
\par }{\Q when the surface
of the deep
is frozen
solid?
\par }{\Q \VS{31}Can you tie
the bands
of the Pleiades,
\par }{\Q or
release the cords
of Orion?
\par }{\Q \VS{32}Can you lead out
\par }{\Q the constellations
in their seasons,
\par }{\Q or guide
the Bear
with
its cubs?
\par }{\Q \VS{33}Do you know
the laws
of the heavens,
\par }{\Q or
can you set
up their rule
over the earth?
\par }{\Q \VS{34}Can you raise
your voice
to the clouds
\par }{\Q so that a flood
of water
covers you?
\par }{\Q \VS{35}Can you send
out lightning
bolts, and they go?
\par }{\Q Will they say
to you, ‘Here we are’?
\par }{\Q \VS{36}Who
has put wisdom
in the heart,
\par }{\Q or
has imparted
understanding to the mind?
\par }{\Q \VS{37}Who
by wisdom
can count
the clouds,
\par }{\Q and who can
tip
over the water jars
of heaven,
\par }{\Q \VS{38}when the dust
hardens into
a mass,
\par }{\Q and the clumps of earth
stick together?
\par }{\Q \VS{39}“Do you hunt
prey
for the lioness,
\par }{\Q and satisfy the appetite
of the lions,
\par }{\Q \VS{40}when
they crouch
in their dens,
\par }{\Q when they wait in
ambush in the thicket?
\par }{\Q \VS{41}Who
prepares
prey
for the raven,
\par }{\Q when
its young
cry out
to God
\par }{\Q and wander
about for lack
of food?


\par }\Chap{39}{\PP \par }{\Q \VerseOne{1}“Are you acquainted
with the way
\par }{\Q the mountain
goats
give birth?
\par }{\Q Do you watch
as the wild deer
give birth to their young?
\par }{\Q \VS{2}Do you count
the months
they must fulfill,
\par }{\Q and do you know
the time
they give birth?
\par }{\Q \VS{3}They crouch,
they bear their
young,
\par }{\Q they bring forth
the offspring they
have carried.
\par }{\Q \VS{4}Their young
grow
strong, and grow up in the open;
\par }{\Q they go
off, and do not
return to them.
\par }{\Q \VS{5}Who
let
the wild donkey
go free?
\par }{\Q Who
released
the bonds
of the donkey,
\par }{\Q \VS{6}to whom
I appointed
the steppe
for its home,
\par }{\Q the salt wastes
as its dwelling place?
\par }{\Q \VS{7}It scorns
the tumult
in the town;
\par }{\Q it does not
hear
the shouts of a driver.
\par }{\Q \VS{8}It ranges
the hills
as its pasture,
\par }{\Q and searches
after
every
green plant.
\par }{\Q \VS{9}Is the wild ox
willing
to be your servant?
\par }{\Q Will it spend
the night at
your feeding trough?
\par }{\Q \VS{10}Can you bind
the wild ox
to a furrow
with its rope,
\par }{\Q will it till the valleys,
following after you?
\par }{\Q \VS{11}Will you rely
on it because
its strength
is great?
\par }{\Q Will you commit
your labor to it?
\par }{\Q \VS{12}Can you count on
it to bring
in your grain,
\par }{\Q and gather
the grain to your threshing floor?
\VS{13}\par }{\Q “The wings
of the ostrich
flap
with joy,

\par }{\Q but are they the pinions
and plumage
of a stork?
\par }{\Q \VS{14}For
she leaves
her eggs
on the ground,
\par }{\Q and lets them be warmed
on
the soil.
\par }{\Q \VS{15}She forgets
that
a foot
might crush
them,
\par }{\Q or that a wild
animal
might trample them.
\par }{\Q \VS{16}She is harsh
with her young,
\par }{\Q as if they were not
hers;
\par }{\Q she is unconcerned
\par }{\Q about
the uselessness
of her labor.
\par }{\Q \VS{17}For
God
deprived
her of wisdom,
\par }{\Q and did not
impart
understanding to her.
\par }{\Q \VS{18}But as soon as she springs up,

\par }{\Q she
laughs
at the horse
and its
rider.
\par }{\Q \VS{19}“Do you give
the horse
its strength?
\par }{\Q Do you clothe
its neck
with a mane?
\par }{\Q \VS{20}Do you make it leap
like a locust?
\par }{\Q Its proud
neighing
is terrifying!
\par }{\Q \VS{21}It paws
the ground in the valley,
\par }{\Q exulting
mightily,
\par }{\Q it goes out
to meet
the weapons.
\par }{\Q \VS{22}It laughs
at fear
and is not
dismayed;
\par }{\Q it does not
shy away from the sword.
\par }{\Q \VS{23}On
it the quiver
rattles;
\par }{\Q the lance
and javelin
flash.
\par }{\Q \VS{24}In excitement
and impatience
it consumes
the ground;
\par }{\Q it cannot
stand still
when
the trumpet
is blown.
\par }{\Q \VS{25}At the sound of the trumpet,
it says,
‘Aha!’
\par }{\Q And from a distance
it catches the scent
of battle,
\par }{\Q the thunderous
shouting of commanders,
\par }{\Q and the battle cries.
\par }{\Q \VS{26}“Is it by your understanding
that the hawk
soars,
\par }{\Q and spreads
its wings
toward the south?
\par }{\Q \VS{27}Is it at your command
that
the eagle
soars,
\par }{\Q and builds
its nest on high?
\par }{\Q \VS{28}It lives
on a rock
and spends the night
there,
\par }{\Q on
a rocky
crag
and a fortress.
\par }{\Q \VS{29}From there
it spots
its prey,
\par }{\Q its eyes
gaze
intently from a distance.
\par }{\Q \VS{30}And its young
ones devour the blood,
\par }{\Q and where
the dead
carcasses
are,
\par }{\Q there
it is.”

\par }\Chap{40}{\PP \VerseOne{1}Then
the {\ND{Lord}}
answered
Job:
\par }{\Q \VS{2}“Will the one who contends
with
the Almighty
correct
him?

\par }{\Q Let the person who accuses
God
give him an answer!”
\par }{\Q \VS{3}Then
Job
answered
the {\ND{Lord}}:
\par }{\Q \VS{4}“Indeed,
I am completely unworthy –
how
could I reply
to you?
\par }{\Q I put
my hand
over my mouth
to silence myself.
\par }{\Q \VS{5}I have spoken
once,
but I cannot
answer;
\par }{\Q twice,
but I will say no
more.”
\par }{\SH The Lord’s Second Speech
\par }{\PP \VS{6}Then
the {\ND{Lord}}
answered
Job
from the whirlwind:
\par }{\Q \VS{7}“Get ready
for a difficult
task like a man.
\par }{\Q I will question
you and you will inform me!
\par }{\Q \VS{8}Would you indeed
annul
my justice?
\par }{\Q Would you declare me guilty
so that you might
be right?
\par }{\Q \VS{9}Do you have an arm
as powerful as God’s,
\par }{\Q and can you thunder
with a voice
like his?
\par }{\Q \VS{10}Adorn
yourself, then, with majesty
and excellency,
\par }{\Q and clothe
yourself with glory
and honor!
\par }{\Q \VS{11}Scatter
abroad
the abundance
of your anger.
\par }{\Q Look
at every
proud
man and bring him low;
\par }{\Q \VS{12}Look
at every
proud
man and abase
him;
\par }{\Q crush
the wicked
on the spot!
\par }{\Q \VS{13}Hide
them in the dust
together,
\par }{\Q imprison them in the grave.
\par }{\Q \VS{14}Then
I myself
will acknowledge
to you
\par }{\Q that
your own right hand
can save you.
\par }{\SH The Description of Behemoth
\par }{\Q \VS{15}“Look
now
at Behemoth,
which
I made
as I made you;
\par }{\Q it eats
grass
like the ox.
\par }{\Q \VS{16}Look
at its strength
in its loins,
\par }{\Q and its power
in the muscles
of its belly.
\par }{\Q \VS{17}It
makes its tail
stiff
like
a cedar,
\par }{\Q the sinews
of its thighs
are tightly wound.
\par }{\Q \VS{18}Its bones
are tubes
of bronze,
\par }{\Q its limbs
like bars of
iron.
\par }{\Q \VS{19}It ranks first
among the works
of God,
\par }{\Q the One who made
it
\par }{\Q has furnished it with a sword.
\par }{\Q \VS{20}For
the hills
bring
it food,
\par }{\Q where
all
the wild
animals
play.
\par }{\Q \VS{21}Under
the lotus trees
it lies,
\par }{\Q in the secrecy
of the reeds
and the marsh.
\par }{\Q \VS{22}The lotus trees
conceal
it in their shadow;
\par }{\Q the poplars
by the stream conceal it.
\par }{\Q \VS{23}If
the river
rages,
it is not
disturbed,
\par }{\Q it is secure,
though
the Jordan
\par }{\Q should surge
up to
its mouth.
\par }{\Q \VS{24}Can anyone catch
it by its eyes,
\par }{\Q or pierce
its nose
with a snare?

\par }\Chap{41}{\PP \par }{\Q \VerseOne{1} “Can you pull
in Leviathan
with a hook,
\par }{\Q and tie down
its tongue
with a rope?
\par }{\Q \VS{2}Can you put
a cord
through its nose,
\par }{\Q or pierce
its jaw
with a hook?
\par }{\Q \VS{3}Will it make numerous
supplications
to
you,
\par }{\Q will it speak
to
you with tender words?
\par }{\Q \VS{4}Will it make
a pact
with
you,
\par }{\Q so you could take
it as your slave
for life?
\par }{\Q \VS{5}Can you play
with it, like a bird,
\par }{\Q or tie it on a leash
for your girls?
\par }{\Q \VS{6}Will partners
bargain for it?
\par }{\Q Will they divide
it up among
the merchants?
\par }{\Q \VS{7}Can you fill
its hide
with harpoons
\par }{\Q or its head
with fishing spears?
\par }{\Q \VS{8}If
you lay
your hand
on
it,
\par }{\Q you will remember
the fight,
\par }{\Q and you will never do
it again!
\par }{\Q \VS{9}See, his expectation
is wrong,

\par }{\Q he is laid low even
at
the sight of it.
\par }{\Q \VS{10}Is it not
fierce
when
it is awakened?
\par }{\Q Who is he,
then, who can
stand
before it?
\par }{\Q \VS{11}(Who
has confronted
me that I should repay?

\par }{\Q Everything
under
heaven belongs to me!)
\par }{\Q \VS{12}I will not
keep silent
about its limbs,
\par }{\Q and the extent
of its might,
\par }{\Q and the grace
of its arrangement.
\par }{\Q \VS{13}Who can
uncover
its outer
covering?

\par }{\Q Who
can penetrate
to the inside
of its armor?
\par }{\Q \VS{14}Who
can open
the doors
of its mouth?

\par }{\Q Its teeth
all around
are fearsome.
\par }{\Q \VS{15}Its back
has rows
of shields,
\par }{\Q shut up
closely together as with a seal;
\par }{\Q \VS{16}each
one
is so close
to the next

\par }{\Q that no
air
can come
between them.
\par }{\Q \VS{17}They lock tightly
together, one to the next;
\par }{\Q they cling
together and cannot
be separated.
\par }{\Q \VS{18}Its snorting
throws out flashes
of light;
\par }{\Q its eyes
are
like the red glow
of dawn.
\par }{\Q \VS{19}Out of its mouth
go
flames,
\par }{\Q sparks
of fire shoot forth!
\par }{\Q \VS{20}Smoke
streams from its nostrils
\par }{\Q as from a boiling
pot
over burning rushes.
\par }{\Q \VS{21}Its breath
sets
coals
ablaze
\par }{\Q and a flame
shoots
from its mouth.
\par }{\Q \VS{22}Strength
lodges
in its neck,
\par }{\Q and despair
runs before it.
\par }{\Q \VS{23}The folds
of its flesh
are tightly
joined;
\par }{\Q they are firm on
it, immovable.
\par }{\Q \VS{24}Its
heart
is hard as rock,
\par }{\Q hard as a lower
millstone.
\par }{\Q \VS{25}When it rises up,
the mighty
are terrified,
\par }{\Q at its thrashing
about they withdraw.
\par }{\Q \VS{26}Whoever strikes
it with a sword
\par }{\Q will have no
effect,
\par }{\Q nor with the spear,
arrow, or dart.
\par }{\Q \VS{27}It regards
iron
as
straw
\par }{\Q and bronze
as rotten
wood.
\par }{\Q \VS{28}Arrows
do not
make it flee;
\par }{\Q slingstones
become like
chaff to it.
\par }{\Q \VS{29}A club
is counted
as a piece of straw;
\par }{\Q it laughs
at the rattling
of the lance.
\par }{\Q \VS{30}Its underparts
are the sharp
points of potsherds,
\par }{\Q it leaves its mark
in
the mud
\par }{\Q like a threshing sledge.
\par }{\Q \VS{31}It makes the deep
boil
like a cauldron
\par }{\Q and stirs up
the sea
like a pot of ointment,
\par }{\Q \VS{32}It leaves a glistening wake
behind
it;
\par }{\Q one would think
the deep
had a head of white hair.
\par }{\Q \VS{33}The likes
of it is
not
on
earth,
\par }{\Q a creature without
fear.
\par }{\Q \VS{34}It looks
on every
haughty
being;
\par }{\Q it is
king
over
all
that are proud.”

\par }\Chap{42}{\PP \VerseOne{1}Then
Job
answered
the {\ND{Lord}}:
\par }{\Q \VS{2}“I know
that
you can do
all
things;
\par }{\Q no
purpose
of yours can be thwarted;
\par }{\Q \VS{3}you asked,

\par }{\Q ‘Who
is this
who darkens
counsel
\par }{\Q without
knowledge?’
\par }{\Q But I have
declared
without understanding
\par }{\Q things too wonderful
for me
to know.
\par }{\Q \VS{4}You said,

\par }{\Q ‘Pay attention,
and I
will speak;
\par }{\Q I will question you, and you will answer me.’
\par }{\Q \VS{5}I had heard
of you by the hearing
of the ear,
\par }{\Q but now
my eye
has seen you.
\par }{\Q \VS{6}Therefore
I despise
myself,
\par }{\Q and I repent
in dust
and ashes!
\par }{\SH \par }{\MS VII. The Epilogue (42:7-17)
\par }{\PP \VS{7}After
the {\ND{Lord}}
had
spoken
these
things
to
Job,
he said
to
Eliphaz
the Temanite,
“My anger
is stirred up
against you and your two
friends,
because
you have not
spoken
about me
what
is right,
as my servant
Job has.
\VS{8}So now
take
seven
bulls
and seven
rams
and go
to
my servant
Job
and offer
a burnt offering
for
yourselves. And my servant
Job
will intercede
for
you, and I will respect him, so that I do not deal
with
you according to your folly,
because
you have not
spoken
about me
what
is right,
as my servant
Job has.”
\par }{\PP \VS{9}So they went,
Eliphaz
the Temanite,
Bildad
the Shuhite,
and Zophar
the Naamathite,
and did
just
as the
{\ND{Lord}}
had told
them;
and the
{\ND{Lord}}
had respect
for Job.
\par }{\PP \VS{10}So the
{\ND{Lord}}
restored
what Job
had lost after he prayed
for
his friends,
and the
{\ND{Lord}}
doubled
all
that had belonged
to Job.
\VS{11}So they came
to
him, all
his brothers
and sisters
and all
who had known
him before,
and they dined
with
him in his house.
They comforted
him and consoled
him for all
the trouble
the {\ND{Lord}}
had
brought
on
him, and each
one
gave
him a piece of silver
and a
gold
ring.
\par }{\PP \VS{12}So the
{\ND{Lord}}
blessed
the second part
of Job’s
life more than the first.
He had 14,000
sheep,
6,000
camels,
1,000
yoke
of oxen,
and 1,000
female donkeys.
\VS{13}And he also had seven
sons
and three
daughters.
\VS{14}The first
daughter he named
Jemimah,
the second
Keziah,
and the third
Keren-Happuch.
\VS{15}Nowhere in all
the land
could women
be found
who were as beautiful
as Job’s
daughters,
and their father
granted
them
an inheritance
alongside
their brothers.
\par }{\PP \VS{16}After
this
Job
lived
140
years;
he saw
his children
and their children
to the fourth
generation.
\VS{17}And so Job
died,
old
and full
of days.
\par }