\NormalFont\ShortTitle{Joel}
{\MT Joel

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \par }{\Q \VerseOne{1}This
is the
{\ND{Lord}}’s
message
that
was given

\par }{\Q to
Joel
the son
of Pethuel:
\par }{\SH A Locust Plague Foreshadows the Day of the Lord
\par }{\Q \VS{2}Listen
to this,
you elders;
\par }{\Q pay attention,
all
inhabitants
of the land.
\par }{\Q Has anything like
this
ever happened in your whole life
\par }{\Q or
in the lifetime
of your ancestors?
\par }{\Q \VS{3}Tell your children
about it,
\par }{\Q have your children
tell
their children,
\par }{\Q and their children
the following generation.
\par }{\Q \VS{4}What the
{\IT{gazam}}-locust left the
{\IT{‘arbeh}}-locust
consumed,
\par }{\Q {\IT{what the
‘arbeh}}-locust
{\IT{left the
yeleq}}-locust
consumed,
\par }{\Q {\IT{and what the
yeleq}}-locust
{\IT{left the
hasil}}-locust
consumed!
\par }{\Q \VS{5}Wake
up, you drunkards,
and weep!
\par }{\Q Wail,
all
you wine
drinkers,
\par }{\Q because the sweet wine has been taken away from you.
\par }{\Q \VS{6}For
a nation
has invaded
our land.
\par }{\Q There are so many of them they are too numerous
to count.
\par }{\Q Their teeth
are like those
of a lion;
\par }{\Q they tear apart
their prey like a lioness.
\par }{\Q \VS{7}They have destroyed
our vines;
\par }{\Q they have turned our fig trees
into mere splinters.
\par }{\Q They have completely stripped off
the bark
and thrown
them aside;
\par }{\Q the twigs
are stripped bare.
\par }{\SH A Call to Lament
\par }{\Q \VS{8}Wail
like a young virgin
clothed
in sackcloth,
\par }{\Q lamenting the death of her husband-to-be.
\par }{\Q \VS{9}No one brings
grain offerings
or drink offerings
\par }{\Q to the temple
of the {\ND{Lord}}
anymore.

\par }{\Q So the priests,
those who serve
the {\ND{Lord}},
are in mourning.
\par }{\Q \VS{10}The crops of the fields
have been destroyed.
\par }{\Q The ground
is in mourning
because
the grain
has perished.
\par }{\Q The fresh wine
has dried up;
\par }{\Q the olive oil
languishes.
\par }{\Q \VS{11}Be distressed,
farmers;
\par }{\Q wail,
vinedressers,
over
the wheat
and the barley.
\par }{\Q For
the harvest
of the field
has perished.
\par }{\Q \VS{12}The vine
has dried up;
\par }{\Q the fig tree
languishes
–
\par }{\Q the pomegranate,
date,
and apple
as well.
\par }{\Q In fact, all
the trees
of the field
have dried up.
\par }{\Q Indeed,
the joy
of the people
has dried up!
\par }{\Q \VS{13}Get dressed
and lament,
you priests!
\par }{\Q Wail,
you who minister
at the altar!
\par }{\Q Come,
spend
the night in sackcloth,
you servants
of my God,
\par }{\Q because
no one brings
grain offerings
or drink offerings
\par }{\Q to the temple
of your God
anymore.
\par }{\Q \VS{14}Announce a holy
fast;
\par }{\Q proclaim
a sacred assembly.
\par }{\Q Gather
the elders
and all
the inhabitants
of the land
\par }{\Q to the temple
of the {\ND{Lord}}
your God,
\par }{\Q and cry out
to
the {\ND{Lord}}.
\par }{\Q \VS{15}How awful
that day
will be!

\par }{\Q For
the day
of the {\ND{Lord}}
is near;
\par }{\Q it will come as destruction
from the Divine
Destroyer.
\par }{\Q \VS{16}Our food
has been cut off
right before
our eyes!

\par }{\Q There is no
longer any joy
or gladness
in the temple
of our God!
\par }{\Q \VS{17}The grains
of seed
have shriveled beneath
their shovels.
\par }{\Q Storehouses
have been decimated
\par }{\Q and granaries
have been torn down, for
the grain
has dried up.
\par }{\Q \VS{18}Listen to the cattle
groan!


\par }{\Q The herds
of livestock
wander around in confusion
\par }{\Q because
they have no
pasture.
\par }{\Q Even
the flocks
of sheep
are suffering.
\par }{\Q \VS{19}To
you, O
{\ND{Lord}}, I call out
for help,

\par }{\Q for
fire
has burned up
the grassy
pastures,
\par }{\Q flames
have razed
all
the trees
in the fields.
\par }{\Q \VS{20}Even
the wild
animals
cry out
to
you;

\par }{\Q for
the river beds
have dried
up;
\par }{\Q fire
has destroyed
the grassy
pastures.

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}Blow
the trumpet
in Zion;
\par }{\Q sound the alarm
signal on my holy
mountain!
\par }{\Q Let all
the inhabitants
of the land
shake
with fear,
\par }{\Q for
the day
of the {\ND{Lord}}
is about to come.
\par }{\Q Indeed,
it is near!
\par }{\Q \VS{2}It will be a day
of dreadful
darkness,
\par }{\Q a day
of foreboding
storm clouds,
\par }{\Q like blackness
spread
over
the mountains.
\par }{\Q It is a huge
and powerful
army –
\par }{\Q there has never
been
anything like
it ever
before,
\par }{\Q and there will not
be anything like it for many generations
to come!
\par }{\Q \VS{3}Like fire
they devour
everything in their path;

\par }{\Q a flame
blazes
behind
them.
\par }{\Q The land
looks like the Garden
of Eden
before
them,
\par }{\Q but behind
them there is only a desolate
wilderness
–
\par }{\Q for nothing
escapes them!
\par }{\Q \VS{4}They look
like horses;
\par }{\Q they charge ahead
like war horses.
\par }{\Q \VS{5}They sound
like chariots
rumbling over
mountain
tops,
\par }{\Q like the crackling
of blazing
fire
consuming
stubble,
\par }{\Q like the noise of a mighty
army
being drawn up
for battle.
\par }{\Q \VS{6}People
writhe
in fear when they see them.

\par }{\Q All
of their faces
turn
pale with fright.
\par }{\Q \VS{7}They charge
like warriors;
\par }{\Q they scale
walls
like soldiers.
\par }{\Q Each
one proceeds on
his course;
\par }{\Q they do not
alter
their path.
\par }{\Q \VS{8}They
do not
jostle
one another;
\par }{\Q each of them
marches
straight ahead.
\par }{\Q They burst
through
the city defenses
\par }{\Q and do not
break ranks.
\par }{\Q \VS{9}They rush into
the city;
\par }{\Q they scale
its walls.
\par }{\Q They climb up
into the houses;
\par }{\Q they go in
through the windows
like a thief.
\par }{\Q \VS{10}The earth
quakes
before
them;

\par }{\Q the sky
reverberates.
\par }{\Q The sun
and the moon
grow dark;
\par }{\Q the stars
refuse
to shine.
\par }{\Q \VS{11}The voice
of the
{\ND{Lord}}
thunders
as he leads
his army.
\par }{\Q Indeed,
his warriors are innumerable;
\par }{\Q Surely
his command
is carried
out!

\par }{\Q Yes,
the day
of the
{\ND{Lord}}
is awesome
\par }{\Q and very
terrifying
– who can
survive it?
\par }{\SH An Appeal for Repentance
\par }{\Q \VS{12}“Yet even
now,”
the {\ND{Lord}}
says,
\par }{\Q “return
to
me with all
your heart
–
\par }{\Q with fasting,
weeping,
and mourning.
\par }{\Q Tear your hearts,

\par }{\Q not just your garments!”
\par }{\Q \VS{13}Return
to
the {\ND{Lord}}
your God,
\par }{\Q for
he is merciful
and compassionate,
\par }{\Q slow
to anger
and boundless
in loyal love –
often relenting
from
calamitous punishment.
\par }{\Q \VS{14}Who
knows?
\par }{\Q Perhaps
he will be
compassionate
and grant a reprieve,
\par }{\Q and leave blessing in his wake –
\par }{\Q a meal offering and a drink offering for you to offer to the
{\ND{Lord}} your God!
\par }{\Q \VS{15}Blow
the trumpet
in Zion.
\par }{\Q Announce a holy
fast;
\par }{\Q proclaim
a sacred assembly!
\par }{\Q \VS{16}Gather
the people;
\par }{\Q sanctify
an assembly!
\par }{\Q Gather
the elders;
\par }{\Q gather
the children
and the nursing infants.
\par }{\Q Let the bridegroom
come out
from his bedroom
\par }{\Q and the bride
from her private quarters.
\par }{\Q \VS{17}Let the priests,
those who serve
the {\ND{Lord}}, weep
\par }{\Q from the vestibule
all the way back to the altar.
\par }{\Q Let them say,
“Have pity,
O
{\ND{Lord}}, on
your people;
\par }{\Q please do not
turn
over your inheritance
to be mocked,
\par }{\Q to become a proverb
among the nations.
\par }{\Q Why
should it be said
among the peoples,
\par }{\Q “Where
is their God?”
\par }{\SH The
{\ND{Lord}}’s Response
\par }{\Q \VS{18}Then the
{\ND{Lord}}
became zealous
for his land;
\par }{\Q he had compassion
on
his people.
\par }{\Q \VS{19}The
{\ND{Lord}}
responded
to his people,
\par }{\Q “Look! I am about to restore
your grain
\par }{\Q as well as fresh wine
and olive oil.
\par }{\Q You will be fully satisfied.
\par }{\Q I will never
again
make
you an object of mockery
among the nations.
\par }{\Q \VS{20}I will remove
the one from the north
far
from
you.
\par }{\Q I will drive
him out
to
a dry
and desolate
place.
\par }{\Q Those in front
will be driven eastward
into
the Dead Sea,
\par }{\Q and those in back
westward into
the Mediterranean
Sea.
\par }{\Q His stench
will rise up
as a foul smell.”
\par }{\Q Indeed,
the
{\ND{Lord}} has accomplished
great things.
\par }{\Q \VS{21}Do not
fear,
my land!
\par }{\Q Rejoice
and be glad,
\par }{\Q because
the {\ND{Lord}}
has accomplished
great things!
\par }{\Q \VS{22}Do not
fear,
wild
animals!

\par }{\Q For
the pastures
of the wilderness
are again green with grass.
\par }{\Q Indeed,
the trees
bear
their fruit;
\par }{\Q the fig tree
and the vine
yield
to their fullest.
\par }{\Q \VS{23}Citizens
of Zion,
rejoice!
\par }{\Q Be glad
because of what the
{\ND{Lord}}
your God
has done!

\par }{\Q For
he has given
to you the
early rains
as vindication.
\par }{\Q He has sent
to you the rains
–
\par }{\Q both the early
and the late rains
as formerly.
\par }{\Q \VS{24}The threshing floors
are full
of grain;
\par }{\Q the vats
overflow
with fresh wine
and olive oil.
\par }{\Q \VS{25}I will make up
for the years
\par }{\Q that
{\IT{the
‘arbeh}}-locust
consumed your crops –
\par }{\Q the
{\IT{yeleq}}-locust, the
{\IT{hasil}}-locust, and the
{\IT{gazam}}-locust –
\par }{\Q my great army that I sent against you.
\par }{\Q \VS{26}You will have plenty
to eat,
\par }{\Q and your hunger will be fully satisfied;
\par }{\Q you will praise
the name
of the {\ND{Lord}}
your God,
\par }{\Q who
has acted
wondrously
in your behalf.
\par }{\Q My people
will never
again be put to shame.
\par }{\Q \VS{27}You will be convinced
that
I am
in the midst
of Israel.
\par }{\Q I
am
the {\ND{Lord}}
your God;
there is no
other.
\par }{\Q My people
will never
again
be put to shame.
\par }{\SH An Outpouring of the Spirit
\par }{\Q \VS{28} After
all of this
\par }{\Q I will pour
out my Spirit
on
all
kinds
of people.

\par }{\Q Your sons
and daughters
will prophesy.
\par }{\Q Your elderly
will have revelatory
dreams;
\par }{\Q your young men
will see
prophetic visions.
\par }{\Q \VS{29}Even
on
male and female
servants
\par }{\Q I will pour
out my Spirit
in those
days.
\par }{\Q \VS{30}I will produce
portents
both in the sky
and on the earth
–
\par }{\Q blood,
fire,
and columns
of smoke.
\par }{\Q \VS{31}The sunlight
will be turned
to darkness
\par }{\Q and the moon
to the color of blood,
\par }{\Q before
the day
of the {\ND{Lord}}
comes
–
\par }{\Q that great
and terrible day!
\par }{\Q \VS{32}It will so happen
that
\par }{\Q everyone
who
calls
on the name
of the {\ND{Lord}}
will
be delivered.
\par }{\Q For
on Mount
Zion
and in Jerusalem
there will be
those who survive,
\par }{\Q just
as the
{\ND{Lord}}
has
promised;
\par }{\Q the remnant
will be those whom
the {\ND{Lord}}
will call.

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1} For
look! In those
days
and at that time
\par }{\Q I will return
the exiles
to Judah
and Jerusalem.
\par }{\Q \VS{2}Then I will gather
all
the nations,
\par }{\Q and bring them down
to
the valley
of Jehoshaphat.
\par }{\Q I will enter into judgment
against them
there
\par }{\Q concerning
my people
Israel
who are my inheritance,
\par }{\Q whom
they scattered
among the nations.
\par }{\Q They partitioned
my land,
\par }{\Q \VS{3}and they cast
lots
for my people.
\par }{\Q They traded a boy
for a prostitute;
\par }{\Q they sold
a little girl
for wine
so they could drink.
\par }{\Q \VS{4}Why
are you
doing these things
to me,
Tyre
and Sidon?

\par }{\Q Are
you
trying
to get
even
with me, land
of Philistia?

\par }{\Q I will very quickly
repay
you
for what you
have done!
\par }{\Q \VS{5}For you took
my silver
and my gold
\par }{\Q and brought
my precious
valuables
to your own palaces.
\par }{\Q \VS{6}You sold
Judeans
and Jerusalemites
to the Greeks,
\par }{\Q removing
them far
from their own country.
\par }{\Q \VS{7}Look! I am rousing
them from
that place
to which
you sold
them.
\par }{\Q I will repay you for what you have done!
\par }{\Q \VS{8}I will sell
your sons
and daughters
to the people of Judah.
\par }{\Q They will sell
them to the Sabeans,
a nation
far
away.
\par }{\Q Indeed,
the {\ND{Lord}}
has spoken!
\par }{\SH Judgment in the Valley of Jehoshaphat
\par }{\Q \VS{9}Proclaim
this
among the nations:
\par }{\Q “Prepare
for a holy war!
\par }{\Q Call
out the warriors!
\par }{\Q Let all
these fighting
men
approach
and attack!
\par }{\Q \VS{10}Beat
your plowshares
into swords,
\par }{\Q and your pruning hooks
into spears!

\par }{\Q Let the weak
say,
‘I too am
a warrior!’
\par }{\Q \VS{11}Lend your aid
and come,
\par }{\Q all
you surrounding
nations,
\par }{\Q and gather
yourselves to that place.”
\par }{\Q Bring down,
O
{\ND{Lord}}, your warriors!
\par }{\Q \VS{12}Let the nations
be roused
and let them go up
\par }{\Q to
the valley
of Jehoshaphat,
\par }{\Q for
there
I will sit
in judgment
on all
the surrounding
nations.
\par }{\Q \VS{13}Rush forth
with the sickle,
for
the harvest
is ripe!
\par }{\Q Come,
stomp
the grapes, for
the winepress
is full!
\par }{\Q The vats
overflow.
\par }{\Q Indeed,
their evil
is great!
\par }{\Q \VS{14}Crowds,
great crowds
are in the valley
of decision,
\par }{\Q for
the day
of the {\ND{Lord}}
is near
in the valley
of decision!
\par }{\Q \VS{15}The sun
and moon
are darkened;
\par }{\Q the stars
withhold
their brightness.
\par }{\Q \VS{16}The
{\ND{Lord}}
roars
from Zion;
\par }{\Q from Jerusalem
his voice
bellows
out.

\par }{\Q The heavens
and the earth
shake.
\par }{\Q But the
{\ND{Lord}}
is a refuge
for his people;
\par }{\Q he is a stronghold
for the citizens of Israel.
\par }{\SH The
{\ND{Lord}}’s Presence in Zion
\par }{\Q \VS{17}You will be convinced
that
I
the {\ND{Lord}}
am your God,
\par }{\Q dwelling
on Zion,
my holy
mountain.
\par }{\Q Jerusalem
will be
holy
–
\par }{\Q conquering
armies will no
longer
pass through it.
\par }{\Q \VS{18}On that
day
the mountains
will drip
with sweet wine,
\par }{\Q and the hills
will flow
with milk.
\par }{\Q All
the dry stream beds
of Judah
will flow
with water.
\par }{\Q A spring
will flow out
from the temple
of the {\ND{Lord}},
\par }{\Q watering
the Valley
of Acacia Trees.
\par }{\Q \VS{19}Egypt
will be desolate
\par }{\Q and Edom
will be a desolate
wilderness,
\par }{\Q because of the violence
they did to the people of Judah,
\par }{\Q in whose
land
they shed
innocent
blood.
\par }{\Q \VS{20}But Judah
will reside
securely forever,
\par }{\Q and Jerusalem
will be secure from one generation to the next.
\par }{\Q \VS{21}I will avenge
their blood
which I had not
previously acquitted.
\par }{\PP It is the
{\ND{Lord}}
who dwells
in Zion!
\par }