\NormalFont\ShortTitle{Jonah}
{\MT Jonah

\par }\ChapOne{1}{\SH Jonah Tries to Run from the Lord
\par }{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
Jonah
son
of Amittai,
\VS{2}“Go
immediately
to Nineveh,
that large
capital
city,
and announce
judgment against
its people because
their wickedness
has come to my attention.”
\VS{3}Instead, Jonah
immediately headed off
to Tarshish

to escape
from
the commission of the
{\ND{Lord}}. He traveled
to Joppa
and found
a merchant ship
heading
to Tarshish.
So he paid
the fare
and went aboard
it to go
with
them to Tarshish
far away from
the {\ND{Lord}}.
\VS{4}But the
{\ND{Lord}}
hurled
a powerful
wind
on the sea.
Such a violent tempest
arose
on the sea
that the ship
threatened
to break up!
\VS{5}The sailors
were so afraid
that each
cried
out to
his own god
and they flung
the ship’s
cargo
overboard
to
make the ship
lighter. Jonah,
meanwhile, had gone down
into
the hold below deck,
had lain
down, and was sound asleep.
\VS{6}The ship’s
captain
approached
him and said,
“What
are you doing asleep? Get up! Cry
out to
your god! Perhaps
your god
might take notice
of us so that we might not
die!”
\VS{7}The sailors said
to
one
another, “Come on, let’s
cast
lots
to find
out whose fault
it is that this disaster
has overtaken us. ” So they cast
lots,
and Jonah
was singled out.
\VS{8}They said
to him,
“Tell
us,
whose
fault is it that this disaster
has overtaken us? What’s
your occupation? Where
do
you come
from? What’s
your country? And who are
your people?”
\VS{9}He said
to
them, “I am
a Hebrew! And I
worship
the {\ND{Lord}}, the God
of heaven,
who
made
the sea
and the dry land.”
\VS{10}Hearing
this, the men
became even more
afraid
and said
to him,
“What
have you done?” (The men
said this
because
they knew
that
he was
trying to escape
from the
{\ND{Lord}},
because
he had previously told
them. )
\VS{11}Because
the storm
was growing
worse and worse,
they said
to
him, “What
should we do
to you to make
the sea
calm
down for
us?”
\VS{12}He said
to them,
“Pick
me up and throw
me into
the sea
to make the sea
quiet
down, because
I
know
it’s my fault
you are in this
severe
storm.”
\VS{13}Instead, they tried
to row back
to
land,
but they were not
able
to do so
because
the storm
kept growing
worse and worse.
\VS{14}So they cried
out to
the {\ND{Lord}}, “Oh, please,

{\ND{Lord}}, don’t
let us die
on account of this
man! Don’t
hold us guilty of shedding innocent
blood.
After
all, you,

{\ND{Lord}}, have done
just
as you pleased.”
\VS{15}So they picked
Jonah
up and threw
him into
the sea,
and the sea
stopped
raging.
\VS{16}The men
feared
the {\ND{Lord}}
greatly,
and earnestly
vowed
to offer
lavish sacrifices
to the Lord.
\VS{17}\par }{\SH Jonah Prays

\par }{\PP ‘span class=”s H3068” ’The
{\ND{Lord}}
sent a huge
fish
to swallow
Jonah,
and Jonah
was in the stomach
of the fish
three
days
and three
nights.

\par }\Chap{2}{\PP \VerseOne{1}Jonah
prayed
to
the {\ND{Lord}}
his God
from the stomach
of the fish
\VS{2}and said,
\par }{\PP \par }{\Q “I called
out to
the {\ND{Lord}}
from my distress,
\par }{\Q and he answered
me;

\par }{\Q from the belly
of Sheol
I cried out for help,
\par }{\Q and you heard
my prayer.
\par }{\Q \VS{3}You threw
me into the deep
waters,

\par }{\Q into the middle
of the sea;
\par }{\Q the ocean current
engulfed
me;
\par }{\Q all
the mighty waves
you sent swept over me.
\par }{\Q \VS{4}I thought I
had been banished
from your sight,
\par }{\Q that I would never
again
see
your holy
temple!
\par }{\Q \VS{5}Water
engulfed
me up to my neck;
\par }{\Q the deep
ocean surrounded
me;
\par }{\Q seaweed
was wrapped around
my head.
\par }{\Q \VS{6}I went down
to the very bottoms
of the mountains;
\par }{\Q the gates
of the netherworld
barred me in
forever;
\par }{\Q but you brought
me
up
from the Pit,
O
{\ND{Lord}}, my God.
\par }{\Q \VS{7}When my life
was ebbing
away, I called out
to the
{\ND{Lord}},
\par }{\Q and my prayer
came
to
your holy
temple.
\par }{\Q \VS{8}Those who
worship worthless
idols
forfeit the mercy
that could be theirs.
\par }{\Q \VS{9}But as for me,
I
promise to offer a sacrifice
to you with a public
declaration of praise;
\par }{\Q I will surely
do what
I have
promised.
\par }{\Q Salvation
belongs to the
{\ND{Lord}}!”
\par }{\PP \VS{10}Then the
{\ND{Lord}}
commanded the fish
and it disgorged
Jonah
on dry land.

\par }\Chap{3}{\PP \VerseOne{1}The
{\ND{Lord}}
said
to
Jonah
a second time,
\VS{2}“Go
immediately
to
Nineveh,
that large
city,
and proclaim
to
it the message
that
I
tell you.”
\VS{3}So Jonah
went
immediately
to
Nineveh,
as
the {\ND{Lord}}
had said. (Now Nineveh
was
an enormous
city –
it required three
days
to walk through it!)
\VS{4}When Jonah
began
to enter
the city
one
day’s
walk,
he announced,
“At the end of forty
days,
Nineveh
will be overthrown!”
\par }{\PP \VS{5}The people
of Nineveh
believed
in God,
and they declared
a fast
and put on
sackcloth,
from the greatest
to the least of them.
\VS{6}When
the news
reached
the king
of Nineveh,
he got
up from his throne,
took off
his royal robe,
put on
sackcloth,
and sat
on
ashes.
\VS{7}He issued a proclamation
and said, “In Nineveh,
by the decree
of the king
and his nobles: No
human
or animal,
cattle
or sheep,
is to taste
anything;
they must
not
eat and they must not
drink
water.
\VS{8}Every person
and animal
must put on sackcloth
and must cry
earnestly
to
God,
and everyone
must turn
from their evil
way
of living and from
the violence
that they do.
\VS{9}Who
knows? Perhaps God
might be willing to change
his mind
and relent
and turn
from his fierce
anger
so that we might not
die.”
\VS{10}When God
saw
their actions
– they turned
from their evil
way of living! – God relented concerning the judgment he had threatened them with and he did not destroy them.

\par }\Chap{4}{\PP \VerseOne{1}This displeased
Jonah
terribly
and he became very angry.
\VS{2}He prayed
to
the
{\ND{Lord}}
and said,
“Oh,

{\ND{Lord}}, this
is just what
I thought would happen when
I was
in
my own country.
This is what I tried to prevent
by attempting to escape
to Tarshish! – because I knew that you are gracious and compassionate, slow to anger and abounding in mercy, and one who relents concerning threatened judgment.
\VS{3}So now,

{\ND{Lord}}, kill
me instead,
because
I would rather
die
than live!”
\VS{4}The
{\ND{Lord}}
said,
“Are you really
so very angry?”
\par }{\PP \VS{5}Jonah
left
the city
and sat
down east
of it. He made
a shelter
for himself there
and sat
down under
it in the shade
to
see
what
would happen
to the city.
\VS{6}The
{\ND{Lord}}
God
appointed
a little plant
and caused it to grow up
over
Jonah
to be
a shade
over
his head
to rescue
him
from his misery.
Now Jonah
was very delighted
about the little plant.
\par }{\PP \VS{7}So God
sent
a worm
at dawn
the next day,
and it attacked
the little plant
so that it dried up.
\VS{8}When the sun
began
to shine,
God
sent a hot
east
wind.
So the sun
beat down
on
Jonah’s
head,
and he grew faint.
So he despaired
of life,
and said,
“I would rather
die
than
live!”
\VS{9}God
said
to
Jonah,
“Are you really
so very angry
about the little plant?” And he said,
“I am as angry
as
I could possibly be!”
\VS{10}The
{\ND{Lord}}
said,
“You
were upset
about this little plant,
something for
which
you have not
worked
nor
did you do anything to make it grow.
It grew up
overnight
and died
the next day.
\VS{11}Should I
not
be even more concerned
about Nineveh,
this enormous
city? There are
more
than one hundred twenty thousand
people
in it who
do not
know
right
from wrong,
as well as many
animals!”
\par }