\NormalFont\ShortTitle{Jude}
{\MT Jude

\par }\OneChap {\SH Salutation
\par }{\PP \VerseOne{1}From Jude,
a slave
of Jesus
Christ
and
brother
of James,
to those who are called,
wrapped in
the love
of God
the Father
and
kept
for Jesus
Christ.
\VS{2}May
mercy,
peace,
and
love
be lavished
on you!
\par }{\SH Condemnation of the False Teachers
\par }{\PP \VS{3}Dear friends,
although I have been
eager
to write
to you
about
our
common
salvation,
I now feel
compelled
instead to write
to encourage
you
to contend earnestly
for the faith
that was once for all
entrusted
to the saints.
\VS{4}For
certain
men
have secretly slipped in among you – men who long ago were marked out for the condemnation I am about to describe – ungodly men who have turned the grace of our God into a license for evil and who deny our only Master and Lord, Jesus Christ.
\par }{\PP \VS{5}Now I desire
to remind
you
(even though you have been fully informed
of these facts once for all
) that
Jesus,
having saved
the people
out of
the land
of Egypt,
later
destroyed
those who did
not
believe.
\VS{6}You also
know that the angels
who did
not
keep
within their
proper domain
but
abandoned
their own
place of residence,
he has kept
in eternal
chains
in
utter darkness,
locked up for
the judgment
of the great
Day.
\VS{7}So also
Sodom
and
Gomorrah
and
the neighboring
towns,
since they indulged in sexual immorality
and
pursued
unnatural desire
in a way similar
to these angels,
are now displayed
as an example
by suffering
the punishment
of eternal
fire.
\par }{\PP \VS{8}Yet
these men,
as a result
of their dreams,
defile
the flesh,
reject
authority,
and
insult
the glorious ones.
\VS{9}But
even when
Michael
the archangel
was arguing
with the devil
and debating
with him concerning
Moses’
body,
he did
not
dare
to bring
a slanderous
judgment,
but
said,
“May
the Lord
rebuke
you!”
\VS{10}But
these men
do
not
understand
the things
they slander,
and
they are being destroyed
by the very things
that, like
irrational
animals,
they
instinctively
comprehend.
\VS{11}Woe
to them! For
they have traveled down
Cain’s
path,
and
because of greed have abandoned
themselves to Balaam’s
error;
hence,
they will certainly perish
in Korah’s
rebellion.
\VS{12}These men
are
dangerous reefs
at
your
love
feasts, feasting
without reverence,
feeding
only themselves.
They are waterless
clouds,
carried along
by
the winds;
autumn
trees
without fruit –
twice
dead,
uprooted;
\VS{13}wild
sea
waves,
spewing out
the foam of their
shame;
wayward
stars
for whom
the utter depths
of
eternal
darkness
have been reserved.
\par }{\PP \VS{14}Now Enoch,
the seventh
in descent beginning with
Adam,
even
prophesied
of them,
saying,
“Look! The Lord
is coming
with
thousands and thousands
of his
holy ones,
\VS{15}to execute
judgment
on
all,
and
to convict
every person
of
all
their
thoroughly ungodly
deeds
that
they have committed,
and
of
all
the harsh words
that
ungodly
sinners
have spoken
against
him.”
\VS{16}These people
are
grumblers
and fault-finders
who go
wherever
their
desires
lead them, and
they give bombastic
speeches,
enchanting
folks
for their own gain.
\par }{\SH Exhortation to the Faithful
\par }{\PP \VS{17}But
you,
dear friends
– recall
the predictions
foretold
by
the apostles
of our
Lord
Jesus
Christ.
\VS{18}For
they said
to you,
“In
the end
time
there will come
scoffers,
propelled
by
their own
ungodly
desires.”
\VS{19}These people
are
divisive,
worldly,
devoid of the Spirit.
\VS{20}But
you,
dear friends,
by building
yourselves
up
in your
most holy
faith,
by
praying
in the Holy
Spirit,
\VS{21}maintain
yourselves
in
the love
of God,
while anticipating
the mercy
of our
Lord
Jesus
Christ
that brings
eternal
life.
\VS{22}And
have mercy
on those who
waver;
\VS{23}save others by snatching them out of the fire; have mercy
on others, coupled with a fear
of God, hating
even the clothes
stained
by
the flesh.
\par }{\SH Final Blessing
\par }{\PP \VS{24}Now
to the one who is able
to keep
you
from falling,
and
to cause you to stand,
rejoicing,
without blemish
before
his
glorious presence,
\VS{25}to the only
God
our
Savior
through
Jesus
Christ
our
Lord,
be glory,
majesty,
power,
and
authority,
before
all
time,
and
now,
and
for
all
eternity.
Amen.
\par }