\NormalFont\ShortTitle{Lamentations}
{\MT Lamentations

\par }\ChapOne{1}{\SH The Prophet Speaks:
\par }{\SH 
{\BD{
א}}
{\IT{(Alef}})

\par }{\Q \VerseOne{1}Alas! The city
once full of people
\par }{\Q now
sits
all alone!

\par }{\Q The prominent lady
among the nations
\par }{\Q has become
a widow!

\par }{\Q The princess
who once ruled the provinces
\par }{\Q has become
a forced laborer!
\par }{\SH ב ({\IT{Bet}})
\par }{\Q \VS{2}She weeps bitterly at night;
\par }{\Q tears stream down her cheeks.
\par }{\Q She has no one to comfort her
\par }{\Q among all her lovers.
\par }{\Q All her friends have betrayed her;
\par }{\Q they have become her enemies.
\par }{\SH ג ({\IT{Gimel}})
\par }{\Q \VS{3}Judah has departed into exile
\par }{\Q under affliction and harsh oppression.
\par }{\Q She lives among the nations;
\par }{\Q she has found no resting place.
\par }{\Q All who pursued her overtook her
\par }{\Q in narrow straits.
\par }{\SH ד ({\IT{Dalet}})
\par }{\Q \VS{4}The roads to Zion mourn
\par }{\Q because no one travels to the festivals.
\par }{\Q All her city gates are deserted;
\par }{\Q her priests groan.
\par }{\Q Her virgins grieve;
\par }{\Q she is in bitter anguish!
\par }{\SH ה ({\IT{He}})
\par }{\Q \VS{5}Her foes subjugated her;
\par }{\Q her enemies are at ease.
\par }{\Q For the
{\ND{Lord}} afflicted her
\par }{\Q because of her many acts of rebellion.
\par }{\Q Her children went away
\par }{\Q captive before the enemy.
\par }{\SH ו ({\IT{Vav}})
\par }{\Q \VS{6}All of Daughter Zion’s splendor
\par }{\Q has departed.
\par }{\Q Her leaders became like deer;
\par }{\Q they found no pasture,
\par }{\Q so they were too exhausted to escape
\par }{\Q from the hunter.
\par }{\SH ז ({\IT{Zayin}})
\par }{\Q \VS{7}Jerusalem remembers,
\par }{\Q when she became a poor homeless person,
\par }{\Q all her treasures
\par }{\Q that she owned in days of old.
\par }{\Q When her people fell into an enemy’s grip,
\par }{\Q none of her allies came to her rescue.
\par }{\Q Her enemies gloated over her;
\par }{\Q they sneered at her downfall.
\par }{\SH ח ({\IT{Khet}})
\par }{\Q \VS{8}Jerusalem committed terrible sin;
\par }{\Q therefore she became an object of scorn.
\par }{\Q All who admired her have despised her
\par }{\Q because they have seen her nakedness.
\par }{\Q She groans aloud
\par }{\Q and turns away in shame.
\par }{\SH ט ({\IT{Tet}})
\par }{\Q \VS{9}Her menstrual flow has soiled her clothing;
\par }{\Q she did not consider the consequences of her sin.
\par }{\Q Her demise was astonishing,
\par }{\Q and there was no one to comfort her.
\par }{\Q She cried, “Look, O
{\ND{Lord}}, on my affliction
\par }{\Q because my enemy boasts!”
\par }{\SH י ({\IT{Yod}})
\par }{\Q \VS{10}An enemy grabbed
\par }{\Q all her valuables.
\par }{\Q Indeed she watched in horror as Gentiles
\par }{\Q invaded her holy temple –
\par }{\Q those whom you had commanded:
\par }{\Q “They must not enter your assembly place.”
\par }{\SH כ ({\IT{Kaf}})
\par }{\Q \VS{11}All her people groaned
\par }{\Q as they searched for a morsel of bread.
\par }{\Q They exchanged their valuables
\par }{\Q for just enough food
\par }{\Q to stay alive.
\par }{\SH Jerusalem Speaks:
\par }{\Q “Look, O
{\ND{Lord}}! Consider
\par }{\Q that I have become worthless!”
\par }{\SH ל ({\IT{Lamed}})
\par }{\Q \VS{12}Is it nothing to you, all you who pass by on the road?
\par }{\Q Look and see!
\par }{\Q Is there any pain like mine?
\par }{\Q The Lord has afflicted me,
\par }{\Q he has inflicted it on me
\par }{\Q when he burned with anger.
\par }{\SH מ ({\IT{Mem}})
\par }{\Q \VS{13}He sent down fire
\par }{\Q into my bones, and it overcame them.
\par }{\Q He spread out a trapper’s net for my feet;
\par }{\Q he made me turn back.
\par }{\Q He has made me desolate;
\par }{\Q I am faint all day long.
\par }{\SH נ ({\IT{Nun}})
\par }{\Q \VS{14}My sins are bound around my neck like a yoke;
\par }{\Q they are fastened together by his hand.
\par }{\Q He has placed his yoke on my neck;
\par }{\Q he has sapped my strength.
\par }{\Q The Lord has handed me over
\par }{\Q to those whom I cannot resist.
\par }{\SH ס ({\IT{Samek}})
\par }{\Q \VS{15}He rounded up all my mighty ones;
\par }{\Q The Lord did this in my midst.
\par }{\Q He summoned an assembly against me
\par }{\Q to shatter my young men.
\par }{\Q The Lord has stomped like grapes
\par }{\Q the virgin daughter, Judah.
\par }{\SH ע ({\IT{Ayin}})
\par }{\Q \VS{16}I weep because of these things;
\par }{\Q my eyes flow with tears.
\par }{\Q For there is no one in sight who can comfort me
\par }{\Q or encourage me.
\par }{\Q My children are desolated
\par }{\Q because an enemy has prevailed.
\par }{\SH The Prophet Speaks:
\par }{\SH 
{\IT{פ (Pe}})
\par }{\Q \VS{17}Zion
spread out her
hands,
\par }{\Q but there is
no one to
comfort her.
\par }{\Q The
{\ND{Lord}} has issued a
decree against
Jacob;
\par }{\Q his
neighbors have
become his
enemies.
\par }{\Q Jerusalem has become
\par }{\Q like
filthy garbage in their midst.
\par }{\SH Jerusalem Speaks:
\par }{\SH 
{\IT{צ (Tsade}})
\par }{\Q \VS{18}The
{\SC{
Lord}} is
right to judge me!

\par }{\Q Yes, I
rebelled against his
commands.
\par }{\Q Please
listen,
all you
nations,

\par }{\Q and
look at my
suffering!
\par }{\Q My young
women and
men
\par }{\Q have
gone into exile.
\par }{\SH ק ({\IT{Qof}})
\par }{\Q \VS{19}I called for my lovers,
\par }{\Q but they had deceived me.
\par }{\Q My priests and my elders
\par }{\Q perished in the city.
\par }{\Q Truly they had searched for food
\par }{\Q to keep themselves alive.
\par }{\SH ר ({\IT{Resh}})
\par }{\Q \VS{20}Look, O
{\ND{Lord}}! I am distressed;
\par }{\Q my stomach is in knots!
\par }{\Q My heart is pounding inside me.
\par }{\Q Yes, I was terribly rebellious!
\par }{\Q Out in the street the sword bereaves a mother of her children;
\par }{\Q Inside the house death is present.
\par }{\SH ש ({\IT{Sin/Shin}})
\par }{\Q \VS{21}They have heard that I groan,
\par }{\Q yet there is no one to comfort me.
\par }{\Q All my enemies have heard of my trouble;
\par }{\Q they are glad that you have brought it about.
\par }{\Q Bring about the day of judgment that you promised
\par }{\Q so that they may end up like me!
\par }{\SH ת ({\IT{Tav}})
\par }{\Q \VS{22}Let all their wickedness come before you;
\par }{\Q afflict them
\par }{\Q just as you have afflicted me
\par }{\Q because of all my acts of rebellion.
\par }{\Q For my groans are many,
\par }{\Q and my heart is sick with sorrow.

\par }\Chap{2}{\SH 
{\IT{א (Alef}})
\par }{\Q \VerseOne{1}Alas! The
Lord has
covered
\par }{\Q Daughter
Zion with his
anger.

\par }{\Q He has thrown
down the
splendor of
Israel
\par }{\Q from
heaven to
earth;
\par }{\Q he did
not
protect his temple
\par }{\Q when he displayed his anger.
\par }{\SH ב ({\IT{Bet}})
\par }{\Q \VS{2}The Lord destroyed mercilessly
\par }{\Q all the homes of Jacob’s descendants.
\par }{\Q In his anger he tore down
\par }{\Q the fortified cities of Daughter Judah.
\par }{\Q He knocked to the ground and humiliated
\par }{\Q the kingdom and its rulers.
\par }{\SH ג ({\IT{Gimel}})
\par }{\Q \VS{3}In fierce anger he destroyed
\par }{\Q the whole army of Israel.
\par }{\Q He withdrew his right hand
\par }{\Q as the enemy attacked.
\par }{\Q He was like a raging fire in the land of Jacob;
\par }{\Q it consumed everything around it.
\par }{\SH ד ({\IT{Dalet}})
\par }{\Q \VS{4}He prepared his bow like an enemy;
\par }{\Q his right hand was ready to shoot.
\par }{\Q Like a foe he killed everyone,
\par }{\Q even our strong young men;
\par }{\Q he has poured out his anger like fire
\par }{\Q on the tent of Daughter Zion.
\par }{\SH ה ({\IT{He}})
\par }{\Q \VS{5}The Lord, like an enemy,
\par }{\Q destroyed Israel.
\par }{\Q He destroyed all her palaces;
\par }{\Q he ruined her fortified cities.
\par }{\Q He made everyone in Daughter Judah
\par }{\Q mourn and lament.
\par }{\SH ו ({\IT{Vav}})
\par }{\Q \VS{6}He destroyed his temple as if it were a vineyard;
\par }{\Q he destroyed his appointed meeting place.
\par }{\Q The
{\ND{Lord}} has made those in Zion forget
\par }{\Q both the festivals and the Sabbaths.
\par }{\Q In his fierce anger he has spurned
\par }{\Q both king and priest.
\par }{\SH ז ({\IT{Zayin}})
\par }{\Q \VS{7}The Lord rejected his altar
\par }{\Q and abhorred his temple.
\par }{\Q He handed over to the enemy
\par }{\Q her palace walls;
\par }{\Q the enemy shouted in the
{\ND{Lord’s}} temple
\par }{\Q as if it were a feast day.
\par }{\SH ח ({\IT{Khet}})
\par }{\Q \VS{8}The
{\ND{Lord}} was determined to tear down
\par }{\Q Daughter Zion’s wall.
\par }{\Q He prepared to knock it down;
\par }{\Q he did not withdraw his hand from destroying.
\par }{\Q He made the ramparts and fortified walls lament;
\par }{\Q together they mourned their ruin.
\par }{\SH ט ({\IT{Tet}})
\par }{\Q \VS{9}Her city gates have fallen to the ground;
\par }{\Q he smashed to bits the bars that lock her gates.
\par }{\Q Her king and princes were taken into exile;
\par }{\Q there is no more guidance available.
\par }{\Q As for her prophets,
\par }{\Q they no longer receive a vision from the
{\ND{Lord}}.
\par }{\SH י ({\IT{Yod}})
\par }{\Q \VS{10}The elders of Daughter Zion
\par }{\Q sit on the ground in silence.
\par }{\Q They have thrown dirt on their heads;
\par }{\Q They have dressed in sackcloth.
\par }{\Q Jerusalem’s young women stare down at the ground.
\par }{\SH כ ({\IT{Kaf}})
\par }{\Q \VS{11}My eyes are worn out from weeping;
\par }{\Q my stomach is in knots.
\par }{\Q My heart is poured out on the ground
\par }{\Q due to the destruction of my helpless people;
\par }{\Q children and infants faint
\par }{\Q in the town squares.
\par }{\SH ל ({\IT{Lamed}})
\par }{\Q \VS{12}Children say to their mothers,
\par }{\Q “Where are food and drink?”
\par }{\Q They faint like a wounded warrior
\par }{\Q in the city squares.
\par }{\Q They die slowly
\par }{\Q in their mothers’ arms.
\par }{\SH מ ({\IT{Mem}})
\par }{\Q \VS{13}With what can I equate you?
\par }{\Q To what can I compare you, O Daughter Jerusalem?
\par }{\Q To what can I liken you
\par }{\Q so that I might comfort you, O Virgin Daughter Zion?
\par }{\Q Your wound is as deep as the sea.
\par }{\Q Who can heal you?
\par }{\SH נ ({\IT{Nun}})
\par }{\Q \VS{14}Your prophets saw visions for you
\par }{\Q that were worthless lies.
\par }{\Q They failed to expose your sin
\par }{\Q so as to restore your fortunes.
\par }{\Q They saw oracles for you
\par }{\Q that were worthless lies.
\par }{\SH ס ({\IT{Samek}})
\par }{\Q \VS{15}All who passed by on the road
\par }{\Q clapped their hands to mock you.
\par }{\Q They sneered and shook their heads
\par }{\Q at Daughter Jerusalem.
\par }{\Q “Ha! Is this the city they called
\par }{\Q ‘The perfection of beauty,
\par }{\Q the source of joy of the whole earth!’?”
\par }{\SH פ ({\IT{Pe}})
\par }{\Q \VS{16}All your enemies
\par }{\Q gloated over you.
\par }{\Q They sneered and gnashed their teeth;
\par }{\Q they said, “We have destroyed her!
\par }{\Q Ha! We have waited a long time for this day.
\par }{\Q We have lived to see it!”
\par }{\SH ע ({\IT{Ayin}})
\par }{\Q \VS{17}The
{\ND{Lord}} has done what he planned;
\par }{\Q he has fulfilled his promise
\par }{\Q that he threatened long ago:
\par }{\Q He has overthrown you without mercy
\par }{\Q and has enabled the enemy to gloat over you;
\par }{\Q he has exalted your adversaries’ power.
\par }{\SH 
{\BD{צ}} ({\IT{Tsade}})
\par }{\Q \VS{18}Cry out from your heart to the Lord,
\par }{\Q O wall of Daughter Zion!
\par }{\Q Make your tears flow like a river
\par }{\Q all day and all night long!
\par }{\Q Do not rest;
\par }{\Q do not let your tears stop!
\par }{\SH ק ({\IT{Qof}})
\par }{\Q \VS{19}Get up! Cry out in the night
\par }{\Q when the night watches start!
\par }{\Q Pour out your heart like water
\par }{\Q before the face of the Lord!
\par }{\Q Lift up your hands to him
\par }{\Q for your children’s lives;
\par }{\Q they are fainting
\par }{\Q at every street corner.
\par }{\SH Jerusalem Speaks:
\par }{\SH 
{\IT{ר (Resh}})
\par }{\Q \VS{20}Look, O
{\ND{Lord}}! Consider!

\par }{\Q Whom have you ever
afflicted like
this?
\par }{\Q Should
women
eat their
offspring,

\par }{\Q their
healthy
infants?

\par }{\Q Should
priest and
prophet
\par }{\Q be
killed in the
Lord’s sanctuary?
\par }{\SH ש ({\IT{Sin/Shin}})
\par }{\Q \VS{21}The young boys and old men
\par }{\Q lie dead on the ground in the streets.
\par }{\Q My young women and my young men
\par }{\Q have fallen by the sword.
\par }{\Q You killed them when you were angry;
\par }{\Q you slaughtered them without mercy.
\par }{\SH ת ({\IT{Tav}})
\par }{\Q \VS{22}As if it were a feast day, you call
\par }{\Q enemies to terrify me on every side.
\par }{\Q On the day of the Lord’s anger
\par }{\Q no one escaped or survived.
\par }{\Q My enemy has finished off
\par }{\Q those healthy infants whom I bore and raised.

\par }\Chap{3}{\SH 
א ({\IT{Alef}})

\par }{\Q \VerseOne{1}I
am the
man who has
experienced
affliction
\par }{\Q from the
rod of his wrath.
\par }{\Q \VS{2}He drove
me into captivity and made me walk
\par }{\Q in darkness
and not
light.
\par }{\Q \VS{3}He repeatedly
attacks
me,
\par }{\Q he turns
his hand
against me all
day long.
\par }{\SH ב ({\IT{Bet}})
\par }{\Q \VS{4}He has made my mortal skin waste away;
\par }{\Q he has broken my bones.
\par }{\Q \VS{5}He
has besieged
and surrounded
me
\par }{\Q with bitter
hardship.
\par }{\Q \VS{6}He has made me reside
in deepest
darkness

\par }{\Q like those who
died
long ago.
\par }{\SH ג ({\IT{Gimel}})
\par }{\Q \VS{7}He has walled me in so that I cannot get out;
\par }{\Q he has weighted me down with heavy prison chains.
\par }{\Q \VS{8}Also,
when
I cry out
desperately
for help,
\par }{\Q he has shut out
my prayer.
\par }{\Q \VS{9}He has blocked
every road
I take with a wall of hewn stones;
\par }{\Q he has made
every path
impassable.
\par }{\SH ד ({\IT{Dalet}})
\par }{\Q \VS{10}To me he is like a bear lying in ambush,
\par }{\Q like a hidden lion stalking its prey.
\par }{\Q \VS{11}He has obstructed
my paths
and torn me to pieces;

\par }{\Q he has made
me desolate.
\par }{\Q \VS{12}He drew
his bow
and made me

\par }{\Q the target
for his arrow.
\par }{\SH ה ({\IT{He}})
\par }{\Q \VS{13}He shot his arrows
\par }{\Q into my heart.
\par }{\Q \VS{14}I have become
the laughingstock
of all
people,
\par }{\Q their mocking song
all
day long.
\par }{\Q \VS{15}He has given me my fill
of bitter herbs
\par }{\Q and made
me drunk
with bitterness.
\par }{\SH ו ({\IT{Vav}})
\par }{\Q \VS{16}He ground my teeth in gravel;
\par }{\Q he trampled me in the dust.
\par }{\Q \VS{17}I am deprived
of peace;
\par }{\Q I
have forgotten
what happiness is.
\par }{\Q \VS{18}So I said,
“My endurance
has expired;
\par }{\Q I have lost all hope
of deliverance from the
{\ND{Lord}}.”
\par }{\SH ז ({\IT{Zayin}})
\par }{\Q \VS{19}Remember my impoverished and homeless condition,
\par }{\Q which is a bitter poison.
\par }{\Q \VS{20}I continually
think
about this,
\par }{\Q and I am depressed.
\par }{\Q \VS{21}But this
I call
to
mind;
\par }{\Q therefore
I have hope:
\par }{\SH ח ({\IT{Khet}})
\par }{\Q \VS{22}The
{\ND{Lord}}’s loyal kindness never ceases;
\par }{\Q his compassions never end.
\par }{\Q \VS{23}They are fresh
every morning;
\par }{\Q your faithfulness
is abundant!
\par }{\Q \VS{24}“My portion
is the Lord,”
I have said
to myself,
\par }{\Q so
I will put my hope in him.
\par }{\SH ט ({\IT{Tet}})
\par }{\Q \VS{25}The
{\ND{Lord}} is good to those who trust in him,
\par }{\Q to the one who seeks him.
\par }{\Q \VS{26}It is good
to wait patiently
\par }{\Q for deliverance
from the
{\ND{Lord}}.
\par }{\Q \VS{27}It is good
for
a man
\par }{\Q to bear
the yoke
while he is young.
\par }{\SH י ({\IT{Yod}})
\par }{\Q \VS{28}Let a person sit alone in silence,
\par }{\Q when the
{\ND{Lord}} is disciplining him.
\par }{\Q \VS{29}Let
him bury
his face
in the dust;
\par }{\Q perhaps
there is
hope.
\par }{\Q \VS{30}Let him offer
his cheek
to the one who hits
him;

\par }{\Q let him have his fill
of insults.
\par }{\SH כ ({\IT{Kaf}})
\par }{\Q \VS{31}For the Lord will not
\par }{\Q reject us forever.
\par }{\Q \VS{32}Though
he causes
us grief,
he then has compassion
on us

\par }{\Q according to the abundance
of his loyal kindness.
\par }{\Q \VS{33}For
he is not
predisposed
to afflict
\par }{\Q or to grieve
people.
\par }{\SH ל ({\IT{Lamed}})
\par }{\Q \VS{34}To crush underfoot
\par }{\Q all the earth’s prisoners,
\par }{\Q \VS{35}to deprive a person
of his rights

\par }{\Q in
the presence
of the Most High,
\par }{\Q \VS{36}to defraud
a person
in a lawsuit
–
\par }{\Q the Lord
does not
approve of such things!
\par }{\SH מ ({\IT{Mem}})
\par }{\Q \VS{37}Whose command was ever fulfilled
\par }{\Q unless the Lord decreed it?
\par }{\Q \VS{38}Is it
not
from the mouth
of the Most
High that everything comes –
\par }{\Q both calamity and blessing?
\par }{\Q \VS{39}Why
should any living
person
complain
\par }{\Q when punished for his sins?
\par }{\SH נ ({\IT{Nun}})
\par }{\Q \VS{40}Let us carefully examine our ways,
\par }{\Q and let us return to the
{\ND{Lord}}.
\par }{\Q \VS{41}Let us
lift
up our hearts
and our hands
\par }{\Q to
God
in heaven:
\par }{\Q \VS{42}“We have blatantly
rebelled;
\par }{\Q you
have not
forgiven.”
\par }{\SH ס ({\IT{Samek}})
\par }{\Q \VS{43}You shrouded yourself with anger and then pursued us;
\par }{\Q you killed without mercy.
\par }{\Q \VS{44}You shrouded
yourself with a cloud
\par }{\Q so that no prayer
can get through.
\par }{\Q \VS{45}You make
us like filthy scum
\par }{\Q in the estimation
of the nations.
\par }{\SH פ ({\IT{Pe}})
\par }{\Q \VS{46}All our enemies have gloated over us;
\par }{\Q \VS{47}Panic
and pitfall
have come
upon us,
\par }{\Q devastation
and destruction.
\par }{\Q \VS{48}Streams
of tears
flow
from my eyes
\par }{\Q because my people
are destroyed.
\par }{\SH ע ({\IT{Ayin}})
\par }{\Q \VS{49}Tears flow from my eyes and will not stop;
\par }{\Q there will be no break
\par }{\Q \VS{50}until
the {\ND{Lord}}
looks
down from heaven
\par }{\Q and sees what has happened.
\par }{\Q \VS{51}What my eyes
see grieves me –
\par }{\Q all the suffering of the daughters in my city.
\par }{\SH צ ({\IT{Tsade}})
\par }{\Q \VS{52}For no good reason my enemies
\par }{\Q hunted me down like a bird.
\par }{\Q \VS{53}They shut
me up in a pit
\par }{\Q and threw
stones at me.
\par }{\Q \VS{54}The waters
closed
over
my head;
\par }{\Q I thought I was about to die.
\par }{\SH ק ({\IT{Qof}})
\par }{\Q \VS{55}I have called on your name, O
{\ND{Lord}},
\par }{\Q from the deepest pit.
\par }{\Q \VS{56}You heard
my plea:

\par }{\Q “Do not
close
your ears
to my cry
for relief!”
\par }{\Q \VS{57}You came near
on the day
I called
to you;
\par }{\Q you said, “Do not
fear!”
\par }{\SH ר ({\IT{Resh}})
\par }{\Q \VS{58}O Lord, you championed my cause,
\par }{\Q you redeemed my life.
\par }{\Q \VS{59}You have seen
the wrong
done to me, O
{\ND{Lord}};
\par }{\Q pronounce
judgment on my behalf!
\par }{\Q \VS{60}You have seen
all
their vengeance,
\par }{\Q all
their plots against me.
\par }{\SH ש ({\IT{Sin/Shin}})
\par }{\Q \VS{61}You have heard their taunts, O
{\ND{Lord}},
\par }{\Q all their plots against me.
\par }{\Q \VS{62}My assailants revile
and conspire
\par }{\Q against
me all
day long.
\par }{\Q \VS{63}Watch
them
from morning
to evening;

\par }{\Q I
am the object of their mocking songs.
\par }{\SH ת ({\IT{Tav}})
\par }{\Q \VS{64}Pay them back what they deserve, O
{\ND{Lord}},
\par }{\Q according to what they have done.
\par }{\Q \VS{65}Give
them
a distraught
heart;
\par }{\Q may your curse
be on them!
\par }{\Q \VS{66}Pursue
them in anger
and eradicate
them
\par }{\Q from under
the
{\ND{Lord’s}}
heaven.

\par }\Chap{4}{\SH 
{\IT{א (Alef}})

\par }{\Q \VerseOne{1}Alas! Gold
has lost
its luster;
\par }{\Q pure
gold
loses value.

\par }{\Q Jewels
are scattered
\par }{\Q on every
street
corner.
\par }{\SH ב ({\IT{Bet}})
\par }{\Q \VS{2}The precious sons of Zion
\par }{\Q were worth their weight in gold –
\par }{\Q Alas! – but now they are treated like broken clay pots,
\par }{\Q made by a potter.
\par }{\SH ג ({\IT{Gimel}})
\par }{\Q \VS{3}Even the jackals nurse their young
\par }{\Q at their breast,
\par }{\Q but my people are cruel,
\par }{\Q like ostriches in the desert.
\par }{\SH ד ({\IT{Dalet}})
\par }{\Q \VS{4}The infant’s tongue sticks
\par }{\Q to the roof of its mouth due to thirst;
\par }{\Q little children beg for bread,
\par }{\Q but no one gives them even a morsel.
\par }{\SH ה ({\IT{He}})
\par }{\Q \VS{5}Those who once feasted on delicacies
\par }{\Q are now starving to death in the streets.
\par }{\Q Those who grew up wearing expensive clothes
\par }{\Q are now dying amid garbage.
\par }{\SH ו ({\IT{Vav}})
\par }{\Q \VS{6}The punishment of my people
\par }{\Q exceeded that of of Sodom,
\par }{\Q which was overthrown in a moment
\par }{\Q with no one to help her.
\par }{\SH ז ({\IT{Zayin}})
\par }{\Q \VS{7}Her consecrated ones were brighter than snow,
\par }{\Q whiter than milk;
\par }{\Q their bodies more ruddy than corals,
\par }{\Q their hair like lapis lazuli.
\par }{\SH ח ({\IT{Khet}})
\par }{\Q \VS{8}Now their appearance is darker than soot;
\par }{\Q they are not recognized in the streets.
\par }{\Q Their skin has shriveled on their bones;
\par }{\Q it is dried up, like tree bark.
\par }{\SH ט ({\IT{Tet}})
\par }{\Q \VS{9}Those who died by the sword are better off
\par }{\Q than those who die of hunger,
\par }{\Q those who waste away,
\par }{\Q struck down from lack of food.
\par }{\SH י ({\IT{Yod}})
\par }{\Q \VS{10}The hands of tenderhearted women
\par }{\Q cooked their own children,
\par }{\Q who became their food,
\par }{\Q when my people were destroyed.
\par }{\SH כ ({\IT{Kaf}})
\par }{\Q \VS{11}The
{\ND{Lord}} fully vented his wrath;
\par }{\Q he poured out his fierce anger.
\par }{\Q He started a fire in Zion;
\par }{\Q it consumed her foundations.
\par }{\SH ל ({\IT{Lamed}})
\par }{\Q \VS{12}Neither the kings of the earth
\par }{\Q nor the people of the lands ever thought
\par }{\Q that enemy or foe would enter
\par }{\Q the gates of Jerusalem.
\par }{\SH מ ({\IT{Mem}})
\par }{\Q \VS{13}But it happened due to the sins of her prophets
\par }{\Q and the iniquities of her priests,
\par }{\Q who poured out in her midst
\par }{\Q the blood of the righteous.
\par }{\SH נ ({\IT{Nun}})
\par }{\Q \VS{14}They wander blindly through the streets,
\par }{\Q defiled by the blood they shed,
\par }{\Q while no one dares
\par }{\Q to touch their garments.
\par }{\SH ס ({\IT{Samek}})
\par }{\Q \VS{15}People cry to them, “Turn away! You are unclean!
\par }{\Q Turn away! Turn away! Don’t touch us!”
\par }{\Q So they have fled and wander about;
\par }{\Q but the nations say, “They may not stay here any longer.”
\par }{\SH פ ({\IT{Pe}})
\par }{\Q \VS{16}The
{\ND{Lord}} himself has scattered them;
\par }{\Q he no longer watches over them.
\par }{\Q They did not honor the priests;
\par }{\Q they did not show favor to the elders.
\par }{\SH The People of Jerusalem Lament:
\par }{\SH 
{\IT{ע (Ayin}})
\par }{\Q \VS{17}Our
eyes
continually
failed us
\par }{\Q as we
looked in
vain for
help.

\par }{\Q From our watchtowers we
watched
\par }{\Q for a
nation that could
not rescue us.
\par }{\SH צ ({\IT{Tsade}})
\par }{\Q \VS{18}Our enemies hunted us down at every step
\par }{\Q so that we could not walk about in our streets.
\par }{\Q Our end drew near, our days were numbered,
\par }{\Q for our end had come!
\par }{\SH ק ({\IT{Qof}})
\par }{\Q \VS{19}Those who pursued us were swifter
\par }{\Q than eagles in the sky.
\par }{\Q They chased us over the mountains;
\par }{\Q they ambushed us in the wilderness.
\par }{\SH ר ({\IT{Resh}})
\par }{\Q \VS{20}Our very life breath – the
{\ND{Lord}}’s anointed king –
\par }{\Q was caught in their traps,
\par }{\Q of whom we thought,
\par }{\Q “Under his protection we will survive among the nations.”
\par }{\SH The Prophet Speaks:
\par }{\SH 
{\IT{ש (Sin/Shin}})
\par }{\Q \VS{21}Rejoice
and be glad
for now, O people
of Edom,
\par }{\Q who reside in
the land
of Uz.
\par }{\Q But the cup
of judgment will pass
to you also;
\par }{\Q you will get drunk
and take off your clothes.
\par }{\SH ת ({\IT{Tav}})
\par }{\Q \VS{22}O people of Zion, your punishment will come to an end;
\par }{\Q he will not prolong your exile.
\par }{\Q But, O people of Edom, he will punish your sin
\par }{\Q and reveal your offenses!

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}O

{\ND{Lord}}, reflect
on what
has happened
to us;
\par }{\Q consider
and look
at our disgrace.
\par }{\Q \VS{2}Our inheritance
is turned
over to strangers;
\par }{\Q foreigners
now occupy our homes.
\par }{\Q \VS{3}We have become
fatherless
orphans;
\par }{\Q our
mothers
have become widows.
\par }{\Q \VS{4}We must pay money
for our own water;
\par }{\Q we must buy
our own wood
at a steep price.
\par }{\Q \VS{5}We are pursued
– they are breathing down our necks;
\par }{\Q we are weary
and have no
rest.
\par }{\Q \VS{6}We have submitted
to Egypt and Assyria
\par }{\Q in order
to buy
food to eat.
\par }{\Q \VS{7}Our forefathers
sinned
and are dead,
\par }{\Q but we
suffer
their punishment.
\par }{\Q \VS{8}Slaves
rule
over us;
\par }{\Q there is no
one to rescue
us from their power.
\par }{\Q \VS{9}At the risk of our lives
we get
our food
\par }{\Q because robbers lurk
in the countryside.
\par }{\Q \VS{10}Our skin
is hot
as an oven
\par }{\Q due to
a fever
from hunger.
\par }{\Q \VS{11}They raped
women
in Zion,
\par }{\Q virgins
in the towns
of Judah.
\par }{\Q \VS{12}Princes
were hung
by their hands;
\par }{\Q elders
were mistreated.
\par }{\Q \VS{13}The young men
perform
menial labor;
\par }{\Q boys
stagger
from their labor.
\par }{\Q \VS{14}The elders
are gone from the city gate;
\par }{\Q the young men
have stopped
playing their music.
\par }{\Q \VS{15}Our hearts
no longer
have any joy;
\par }{\Q our dancing
is turned
to mourning.
\par }{\Q \VS{16}The crown
has fallen
from our head;
\par }{\Q woe
to us,
for
we have sinned!
\par }{\Q \VS{17}Because
of this,
our hearts
are sick;
\par }{\Q because of these
things, we can hardly see
through our tears.
\par }{\Q \VS{18}For wild animals
are prowling over
Mount
Zion,
\par }{\Q which lies desolate.
\par }{\Q \VS{19}But you,
O
{\ND{Lord}}, reign forever;
\par }{\Q your throne
endures
from generation
to generation.
\par }{\Q \VS{20}Why
do you keep on forgetting us?
\par }{\Q Why do
you forsake
us so long?
\par }{\Q \VS{21}Bring
us
back
to yourself, O
{\ND{Lord}}, so that we may return
to you;
\par }{\Q renew
our life as in days
before,
\par }{\Q \VS{22}unless
you have utterly
rejected
us

\par }{\PP and are angry
with
us beyond
measure.

\par }