\NormalFont\ShortTitle{Malachi}
{\MT Malachi

\par }\ChapOne{1}{\SH Introduction and God’s Election of Israel
\par }{\PP \VerseOne{1}What
follows is divine revelation.
The word
of the {\ND{Lord}}
came to
Israel
through
Malachi:
\par }{\PP \VS{2}“I have shown love
to you,” says
the {\ND{Lord}},
but you say,
“How
have you shown love to us?”
\par }{\PP “Esau
was Jacob’s
brother,”
the {\ND{Lord}}
explains, “yet I chose
Jacob
\VS{3}and rejected
Esau.
I turned
Esau’s mountains
into a deserted
wasteland
and gave his territory
to the wild jackals.”
\par }{\PP \VS{4}Edom
says,
“Though we are devastated,
we will once again
build
the ruined places.”
So the

{\ND{Lord}}
who rules over all
responds,
“They
indeed may build,
but I
will overthrow.
They will be known
as
the land
of evil,
the people
with whom
the {\ND{Lord}}
is permanently
displeased.
\VS{5}Your eyes
will see
it, and then you
will say,
‘May the
{\ND{Lord}}
be magnified
even beyond the border
of Israel!’ ”
\par }{\SH The Sacrilege of Priestly Service
\par }{\PP \VS{6}“A son
naturally honors
his father
and a slave
respects
his master.
If
I am
your father,
where
is my honor? If
I am
your master,
where
is my respect? The
{\ND{Lord}}
who rules over all
asks you this, you priests
who make light
of my name! But you reply,
‘How
have we made light
of your name?’
\VS{7}You are offering
improper
sacrifices
on
my altar,
yet you ask,
‘How
have we offended
you?’ By treating
the table
of the {\ND{Lord}}
as if it is
of no importance!
\VS{8}For
when
you offer
blind
animals as a sacrifice,
is that not
wrong? And when
you offer
the lame
and sick,
is that not
wrong
as well? Indeed, try offering
them to your governor! Will he be pleased
with you or
show
you favor?” asks
the
{\ND{Lord}}
who rules over all.
\VS{9}But now
plead
for God’s
favor that he might be gracious
to us. “With
this
kind
of offering
in your hands,
how can he be
pleased
with you?” asks
the {\ND{Lord}}
who rules over all.
\par }{\PP \VS{10}“I wish
that one of you would close
the temple doors,
so that you no
longer would light
useless fires
on my altar.
I am
not
pleased
with you,” says
the {\ND{Lord}}
who rules over all,
“and I will no
longer accept
an offering from you.
\VS{11}For
from the east
to the west
my name
will be great
among the nations.
Incense
and pure
offerings
will be offered
in my name
everywhere, for
my name
will be great
among the nations,”
says
the {\ND{Lord}}
who rules over all.
\VS{12}“But you
are profaning
it by saying
that the table
of the Lord
is
common and its offerings
despicable.
\VS{13}You also say,
‘How tiresome
it is.’ You turn up your nose
at it,” says
the {\ND{Lord}}
who rules over all,
“and instead bring
what is stolen,
lame,
or sick.
You bring
these things for an offering! Should I accept
this from you?” asks
the {\ND{Lord}}.
\VS{14}“There will be harsh condemnation
for the hypocrite
who has a valuable male
animal in his flock
but vows
and sacrifices
something inferior to the Lord.
For
I
am a great
king,”
says
the {\ND{Lord}}
who rules over all,
“and my name
is awesome
among the nations.”

\par }\Chap{2}{\PP \VerseOne{1}“Now,
you priests,
this
commandment is for you.
\VS{2}If
you do not
listen
and take
seriously
the need
to honor
my name,”
says
the {\ND{Lord}}
who rules over all,
“I will send
judgment
on you and turn your blessings
into curses
– indeed,
I have already done so because
you are not
taking
it to
heart.
\VS{3}I am
about to discipline
your children
and will spread
offal
on
your faces,
the very offal
produced at your festivals,
and you will be carried away along with it.
\VS{4}Then you will know
that
I sent
this
commandment
to you so
that my covenant
may
continue to be with
Levi,”
says
the {\ND{Lord}}
who rules over all.
\VS{5}“My covenant
with him
was designed
to bring life
and peace.
I gave
its statutes to him to fill him with awe,
and he indeed revered
me and stood in awe
before me.
\VS{6}He taught what
was true;
sinful words
were not
found
on his lips.
He walked
with
me in peace
and integrity,
and he turned
many
people away from sin.
\VS{7}For
the lips
of a priest
should preserve
knowledge
of sacred
things, and people
should seek
instruction
from him because
he is
the messenger
of the {\ND{Lord}}
who rules over all.
\VS{8}You,
however, have turned
from
the way.
You have caused many
to violate
the law;
you have corrupted
the covenant
with Levi,”
says
the {\ND{Lord}}
who rules over all.
\VS{9}“Therefore,
I
have caused
you to be ignored
and belittled
before
all
people
to the extent
to which
you are not
following
after me
and are showing partiality
in your instruction.”
\par }{\SH The Rebellion of the People
\par }{\PP \VS{10}Do we not
all
have one
father? Did not
one
God
create
us? Why
do we betray
one
another,
in this way making light
of the covenant
of our ancestors?
\VS{11}Judah
has become disloyal,
and unspeakable sins
have been committed
in Israel
and Jerusalem.
For
Judah
has profaned
the holy things
that
the {\ND{Lord}}
loves
and has turned
to a foreign
god!
\VS{12}May the
{\ND{Lord}}
cut off
from the community of Jacob
every last person
who does
this, as well as
the person who presents
improper
offerings
to the
{\ND{Lord}}
who rules over all!
\par }{\PP \VS{13}You also do this: You cover
the
altar
of the {\ND{Lord}}
with tears
as you weep
and groan,
because he no
longer
pays
any attention to
the offering
nor accepts
it favorably from you.
\VS{14}Yet you ask,
“Why?” The
{\ND{Lord}}
is testifying
against you on
behalf of the wife
you married when you were young,
to whom
you
have become unfaithful
even though she is your companion
and wife
by law.
\VS{15}No
one
who has
even a small portion
of the Spirit
in him does this. What
did our ancestor
do when seeking
a child
from God? Be attentive,
then, to your own spirit,
for one should not
be disloyal
to the wife
he took in his youth.
\VS{16}“I hate
divorce,”
says
the {\ND{Lord}}
God
of Israel,
“and the one who is guilty
of violence,”
says
the {\ND{Lord}}
who rules over all.
“Pay attention
to your conscience,
and do not
be unfaithful.”
\par }{\SH Resistance to the Lord through Self-deceit
\par }{\PP \VS{17}You have wearied
the {\ND{Lord}}
with your words.
But you say,
“How
have we wearied
him?” Because you say,
“Everyone
who does
evil
is good
in
the Lord’s
opinion,
and he
delights
in them,” or
“Where
is the God
of justice?”

\par }\Chap{3}{\PP \VerseOne{1}“I am about to send
my messenger,
who will clear
the way
before
me. Indeed, the Lord
you
are seeking
will suddenly
come
to
his temple,
and the messenger
of the covenant,
whom
you
long for,
is certainly
coming,”
says
the {\ND{Lord}}
who rules over all.
\par }{\PP \VS{2}Who can
endure
the day
of his coming? Who can
keep standing
when
he appears? For
he will be like a refiner’s fire,
like
a launderer’s
soap.
\VS{3}He will act
like a refiner
and purifier
of silver
and will cleanse
the
Levites
and refine
them like gold
and silver.
Then
they will
offer the
{\ND{Lord}}
a proper
offering.
\VS{4}The offerings
of Judah
and Jerusalem
will be pleasing
to the
{\ND{Lord}}
as in former
times
and years
past.
\par }{\PP \VS{5}“I will come
to
you in judgment.
I will be
quick
to testify
against those who practice divination,
those who commit adultery,
those who break
promises,
and those who exploit
workers,
widows,
and orphans,
who refuse to help the immigrant
and in this way show
they do not
fear
me,” says
the {\ND{Lord}}
who rules over all.
\par }{\SH Resistance to the Lord through Selfishness
\par }{\PP \VS{6}“Since,
I,
the {\ND{Lord}}, do not
go back on my promises,
you,
sons
of Jacob,
have not
perished.
\VS{7}From
the days
of your ancestors
you have ignored my commandments
and have not
kept
them! Return
to me,
and I will return
to
you,” says
the {\ND{Lord}}
who rules over all.
“But you say,
‘How
should we return?’
\VS{8}Can a person
rob
God? You indeed
are
robbing
me, but you say,
‘How
are we robbing
you?’ In tithes
and contributions!
\VS{9}You are
bound
for judgment
because you are
robbing me – this whole nation is guilty.
\par }{\PP \VS{10}“Bring
the entire
tithe
into
the storehouse
so that there may be food
in my temple.
Test
me
in this
matter,” says
the {\ND{Lord}}
who rules over all,
“to see if
I will not
open
for you the
windows
of heaven
and pour out
for you a blessing
until
there is no
room for it all.
\VS{11}Then I will stop the plague
from ruining
your crops,
and the vine
will not
lose
its fruit
before harvest,”
says
the {\ND{Lord}}
who rules over all.
\VS{12}“All
nations
will call
you happy, for
you
indeed will live in a delightful
land,”
says
the {\ND{Lord}}
who rules over all.
\par }{\SH Resistance to the Lord through Self-sufficiency
\par }{\PP \VS{13}“You have criticized
me sharply,”
says
the {\ND{Lord}}, “but you ask, ‘How
have we criticized you?’
\VS{14}You have said,
‘It is useless
to serve
God.
How
have we
been helped
by keeping
his requirements
and going
about like mourners
before
the
{\ND{Lord}}
who rules over all?
\VS{15}So now
we
consider
the arrogant
to be happy; indeed,
those who practice
evil
are successful.
In fact, those who challenge
God
escape!’ ”
\par }{\PP \VS{16}Then
those who respected
the {\ND{Lord}}
spoke
to one
another,
and the
{\ND{Lord}}
took notice.
A scroll
was prepared
before
him in which were recorded the names
of those who respected
the {\ND{Lord}}
and honored
his name.
\VS{17}“They will belong
to me,” says
the {\ND{Lord}}
who rules over all,
“in the day
when I
prepare
my own special property.
I will spare
them as
a man
spares
his son
who serves him.
\VS{18}Then once
more you will see
that I make a distinction between
the righteous
and the wicked,
between
the one who serves
God
and the one who
does not.

\par }\Chap{4}{\PP \VerseOne{1}“For
indeed
the day
is coming,
burning
like a furnace,
and all
the arrogant
evildoers
will be
chaff.
The coming
day
will burn
them up,”
says
the
{\ND{Lord}}
who rules over all.
“It will not
leave
even a root
or branch.
\VS{2}But for you who respect
my name,
the sun
of vindication
will rise
with healing
wings,
and you will skip about
like calves
released
from the stall.
\VS{3}You will trample
on the wicked,
for
they will be
like ashes
under
the soles
of your feet
on the day
which
I am
preparing,”
says
the {\ND{Lord}}
who rules over all.
\par }{\SH Restoration through the Lord
\par }{\PP \VS{4}“Remember
the law
of my servant
Moses,
to whom
at Horeb
I gave
rules
and regulations
for all
Israel to obey.
\VS{5}Look,
I
will send
you Elijah
the prophet
before
the great
and terrible
day
of the {\ND{Lord}}
arrives.
\VS{6}He will encourage
fathers
and their children
to return
to me, so that I will not
come
and strike
the earth
with judgment.”
\par }