\NormalFont\ShortTitle{Matthew}
{\MT Matthew

\par }\ChapOne{1}{\SH The Genealogy of Jesus Christ
\par }{\PP \VerseOne{1}This is the record
of the genealogy
of Jesus
Christ,
the son
of David,
the son
of Abraham.
\par }{\PP \VS{2}Abraham
was the father
of Isaac,
Isaac
the father
of Jacob,
Jacob
the father
of Judah
and
his
brothers,
\VS{3}Judah
the father
of Perez
and
Zerah
(by
Tamar), Perez
the father
of Hezron,
Hezron
the father
of Ram,
\VS{4}Ram
the father
of Amminadab,
Amminadab
the father
of Nahshon,
Nahshon
the father
of Salmon,
\VS{5}Salmon
the father
of Boaz
(by
Rahab), Boaz
the father
of Obed
(by
Ruth), Obed
the father
of Jesse,
\VS{6}and
Jesse
the father
of David
the king.
\par }{\PP David
was the father
of Solomon
(by
the
wife of Uriah ),
\VS{7}Solomon
the father
of Rehoboam,
Rehoboam
the father
of Abijah,
Abijah
the father
of Asa,
\VS{8}Asa
the father
of Jehoshaphat,
Jehoshaphat
the father
of Joram,
Joram
the father
of Uzziah,
\VS{9}Uzziah
the father
of Jotham,
Jotham
the father
of Ahaz,
Ahaz
the father
of Hezekiah,
\VS{10}Hezekiah
the father
of Manasseh,
Manasseh
the father
of Amon,
Amon
the father
of Josiah,
\VS{11}and
Josiah
the father
of Jeconiah
and
his
brothers,
at
the time of the deportation
to Babylon.
\par }{\PP \VS{12}After
the deportation
to Babylon,
Jeconiah
became the father
of Shealtiel,
Shealtiel
the father
of Zerubbabel,
\VS{13}Zerubbabel
the father
of Abiud,
Abiud
the father
of Eliakim,
Eliakim
the father
of Azor,
\VS{14}Azor
the father
of Zadok,
Zadok
the father
of Achim,
Achim
the father
of Eliud,
\VS{15}Eliud
the father
of Eleazar,
Eleazar
the father
of Matthan,
Matthan
the father
of Jacob,
\VS{16}and Jacob
the father
of Joseph,
the
husband
of Mary,
by
whom
Jesus
was born,
who is called
Christ.
\par }{\PP \VS{17}So
all
the generations
from
Abraham
to David
are fourteen
generations,
and
from
David
to the deportation
to Babylon,
fourteen
generations,
and
from
the deportation
to Babylon
to Christ,
fourteen
generations.
\par }{\SH The Birth of Jesus Christ
\par }{\PP \VS{18}Now
the birth of Jesus
Christ
happened
this way.
While his
mother
Mary
was engaged
to Joseph,
but before
they came together,
she was found
to be pregnant
through
the Holy
Spirit.
\VS{19}Because
Joseph,
her
husband
to be, was a righteous
man, and
because he did not
want
to disgrace
her,
he intended
to divorce
her
privately.
\VS{20}When he
had contemplated
this,
an angel
of the Lord
appeared
to him
in
a dream
and said,
“Joseph,
son
of David,
do not
be afraid
to take
Mary
as your
wife,
because
the child conceived
in
her
is
from
the Holy
Spirit.
\VS{21}She will give birth
to a son
and
you will name
him
Jesus,
because
he
will save
his
people
from
their
sins.”
\VS{22}This
all
happened
so that
what was spoken
by
the Lord
through
the prophet
would be fulfilled:
\VS{23}“{\QT{
Look! The
virgin
will conceive
and
bear
a son,
and
they will call
him}}
{\QT{Emmanuel}},” which
means “{\IT{
God
with
us}}.”
\VS{24}When Joseph
awoke
from
sleep
he did
what
the angel
of the Lord
told
him.
He
took
his
wife,
\VS{25}but
did not
have marital
relations with her
until
she gave birth
to a son,
whom
he named
Jesus.

\par }\Chap{2}{\PP \VerseOne{1}After
Jesus
was born
in
Bethlehem
\VS{2}saying,
“Where
is
the one who is born
king
of the Jews? For
we saw
his
star
when it rose
and
have come
to worship
him.”
\VS{3}When King
Herod
heard
this he was alarmed,
and
all
Jerusalem
with
him.
\VS{4}After assembling
all
the chief priests
and
experts in
the law,
he asked
them
where
the Christ
was to be born.
\VS{5}“In
Bethlehem
of Judea,”
they said,
“for
it is written
this way
by
the prophet:
\par }{\Q \VS{6}‘{\QT{
And
you,
Bethlehem,
in the land
of Judah,}}
\par }{\Q {\QT{
are
in
no
way least
among the rulers
of Judah,}}
\par }{\Q {\QT{
for
out of
you
will come
a ruler
who
will shepherd
my
people
Israel}}.’ ”
\par }{\PP \VS{7}Then
Herod
privately
summoned
the wise men
and determined
from
them
when the star
had appeared.
\VS{8}He
sent
them
to
Bethlehem
and said,
“Go
and look
carefully
for
the child.
When
you find
him, inform
me
so that
I
can go
and worship
him as well.”
\VS{9}After
listening
to the king
they left,
and
once again the star
they saw
when it rose
led
them
until
it stopped
above
the place where
the child
was.
\VS{10}When
they saw
the star
they shouted
joyfully.
\VS{11}As
they came
into
the house
and saw
the child
with
Mary
his
mother,
they
bowed down
and
worshiped
him.
They opened
their
treasure
boxes
and gave him
gifts
of gold,
frankincense,
and
myrrh.
\VS{12}After being warned
in
a dream
not
to return
to
Herod,
they went back
by
another
route
to
their own
country.
\par }{\SH The Escape to Egypt
\par }{\PP \VS{13}After
they had gone, an angel
of the Lord
appeared
to Joseph
in
a dream
and said,
“Get up,
take
the
child
and
his
mother
and
flee
to
Egypt,
and
stay there
until
I tell
you,
for
Herod
is going
to look for
the
child
to kill
him.”
\VS{14}Then
he got up,
took
the child
and
his
mother
during
the night,
and
went
to
Egypt.
\VS{15}He stayed
there
until
Herod
died.
In this way
what was spoken
by
the Lord
through
the prophet
was fulfilled{\QT{
: “I called
my
Son
out of
Egypt}}.”
\par }{\PP \VS{16}When
Herod
saw
that
he had been tricked
by
the
wise men,
he became enraged.
He sent
men to kill
all
the
children
in
Bethlehem
and
throughout
the surrounding
region
from
the age of two
and
under,
according
to the
time
he had learned
from
the
wise men.
\VS{17}Then
what was spoken
by
Jeremiah
the prophet
was fulfilled:
\par }{\Q \VS{18}“{\QT{
A voice
was heard
in
Ramah,}}
\par }{\Q {\QT{
weeping
and
loud
wailing,}}
\par }{\Q {\QT{
Rachel
weeping
for her
children,}}
\par }{\Q {\QT{
and
she
did not
want
to be comforted,
because
they were gone.}}”
\par }{\SH The Return to Nazareth
\par }{\PP \VS{19}After
Herod
had died,
an angel
of the Lord
appeared
in
a dream
to Joseph
in
Egypt
\VS{20}saying,
“Get up,
take
the child
and
his
mother,
and
go
to
the land
of Israel,
for
those
who were seeking
the child’s
life
are dead.”
\VS{21}So
he got up
and took
the child
and
his
mother
and
returned
to
the land
of Israel.
\VS{22}But
when he heard
that
Archelaus
was reigning
over Judea
in place of
his
father
Herod,
he was afraid
to go
there.
After being warned
in
a dream,
he went
to
the regions
of Galilee.
\VS{23}He
came
to
a town
called
Nazareth
and lived
there. Then what
had been spoken
by
the prophets
was fulfilled,
that
Jesus would be called
a Nazarene.

\par }\Chap{3}{\PP \VerseOne{1}In
those
days
John
the Baptist
came
into the wilderness
of Judea
proclaiming,
\VS{2}“Repent,
for
the kingdom
of heaven
is near.”
\VS{3}For
he
is
the one about whom Isaiah
the prophet
had spoken:
\par }{\Q “{\QT{
The voice
of one shouting
in
the wilderness,}}
\par }{\Q ‘{\QT{
Prepare
the way
for the Lord,
make
his
paths
straight}}.’ ”
\par }{\PP \VS{4}Now
John
wore
clothing
made from
camel’s
hair
with a leather
belt
around
his
waist,
and
his
diet
consisted of
locusts
and
wild
honey.
\VS{5}Then
people from Jerusalem,
as well as all
Judea
and
all
the region around
the Jordan,
were going out
to
him,
\VS{6}and
he was baptizing
them in
the Jordan
River
as they confessed
their
sins.
\par }{\PP \VS{7}But
when he saw
many
Pharisees
and
Sadducees
coming
to
his baptism,
he said
to them,
“You offspring
of vipers! Who
warned
you
to flee
from
the coming
wrath?
\VS{8}Therefore
produce
fruit
that proves
your repentance,
\VS{9}and
don’t think
you can say
to yourselves,
‘We have
Abraham
as our father.’
For
I tell
you
that
God
can
raise up
children
for Abraham
from
these
stones!
\VS{10}Even
now
the ax
is laid
at
the root
of the
trees,
and every
tree
that does not
produce
good
fruit
will be cut down
and
thrown
into
the fire.
\par }{\PP \VS{11}“I
baptize
you
with
water,
for
repentance,
but
the one coming
after
me
is more powerful than
I
am
– I am
not
worthy
to carry
his
sandals.
He will baptize
you
with
the Holy
Spirit
and
fire.
\VS{12}His winnowing fork
is in
his
hand,
and
he will clean out his
threshing floor
and
will gather
his
wheat
into
the storehouse,
but
the chaff
he will burn up
with inextinguishable
fire.”
\par }{\SH The Baptism of Jesus
\par }{\PP \VS{13}Then
Jesus
came
from
Galilee
to
John
to be baptized
by
him
in
the Jordan River.
\VS{14}But
John tried to prevent
him,
saying,
“I
need
to be baptized
by
you,
and
yet you
come
to
me?”
\VS{15}So
Jesus
replied
to him,
“Let
it happen now,
for
it is
right for us
to fulfill
all
righteousness.”
Then
John yielded
to him.
\VS{16}After
Jesus
was baptized,
just as
he was coming up
out
of the
water,
the
heavens
opened
and
he saw
the Spirit
of God
descending
like
a dove
and coming
on
him.
\VS{17}And
a voice
from
heaven
said, “This
is
my
one dear
Son;
in
him I take great delight.”

\par }\Chap{4}{\PP \VerseOne{1}Then
Jesus
was led
by
the Spirit
into
the wilderness
to be tempted
by
the devil.
\VS{2}After he fasted
forty
days
and
forty
nights
he was famished.
\VS{3}The tempter
came
and
said
to him,
“If
you are
the Son
of God,
command
these
stones
to become
bread.”
\VS{4}But
he answered, “It is written, ‘{\QT{
Man
does not
live
by
bread
alone,
but
by
every
word
that comes from
the mouth
of God}}*.’ ”
\VS{5}Then
the devil
took
him
to
the holy
city,
had him
stand
on
the highest point
of
the temple,
\VS{6}and
said
to him,
“If
you are
the Son
of God,
throw
yourself
down.
For
it is written, ‘{\QT{
He will command
his
angels
concerning
you}}’
and ‘{\QT{
with
their hands
they will lift
you
up,
so that
you will
not
strike
your
foot
against
a stone}}.’ ”
\VS{7}Jesus
said
to him,
“Once again
it is written: ‘{\QT{
You are
not
to put
the Lord
your
God to the test}}.’ ”
\VS{8}Again,
the devil
took
him
to
a
very
high
mountain,
and
showed
him
all
the kingdoms
of
the world
and
their
grandeur.
\VS{9}And
he said
to him,
“I will give
you
all
these things
if
you throw
yourself to the ground
and worship
me.”
\VS{10}Then
Jesus
said
to him,
“Go away,
Satan! For
it is written: ‘{\QT{
You are to worship
the Lord
your
God
and
serve}}
only
{\QT{
him}}.’ ”
\VS{11}Then
the devil
left
him,
and
angels
came
and
began ministering
to his needs.
\par }{\SH Preaching in Galilee
\par }{\PP \VS{12}Now when
Jesus heard
that
John
had been imprisoned,
he went
into
Galilee.
\VS{13}While in Galilee, he moved
from Nazareth
by the sea,
in
the region
of Zebulun
and
Naphtali,
\VS{14}so that
what was spoken
by
Isaiah
the prophet
would be fulfilled:
\par }{\Q \VS{15}“{\QT{
Land
of Zebulun
and
land
of Naphtali,}}
\par }{\Q {\QT{
the way
by the sea,
beyond
the Jordan,
Galilee
of the Gentiles –}}
\par }{\Q \VS{16}{\QT{
the people
who sit
in
darkness
have seen
a great
light,}}
\par }{\Q {\QT{
and
on those
who sit
in
the region
and
shadow
of death
a light
has dawned}}.”
\par }{\PP \VS{17}From
that time
Jesus
began
to preach
this message: “Repent,
for
the kingdom
of heaven
is near.”
\par }{\SH The Call of the Disciples
\par }{\PP \VS{18}As
he was walking
by
the Sea
of Galilee
he saw
two
brothers,
Simon
(called
Peter) and
Andrew
his
brother,
casting
a net
into
the sea
(for
they
were fishermen).
\VS{19}He
said
to them,
“Follow
me,
and
I will turn
you
into fishers
of people.”
\VS{20}They left
their nets
immediately
and followed
him.
\VS{21}Going on
from there
he saw
two
other
brothers,
James
the son of Zebedee
and
John
his
brother,
in
a boat
with
Zebedee
their
father,
mending
their
nets.
Then
he called
them.
\VS{22}They immediately
left
the boat
and
their
father
and followed
him.
\par }{\SH Jesus’ Healing Ministry
\par }{\PP \VS{23}Jesus went throughout
all
of Galilee,
teaching
in
their
synagogues,
preaching
the gospel
of the kingdom,
and
healing
all
kinds of disease
and
sickness
among
the people.
\VS{24}So
a report
about him
spread
throughout
Syria.
People brought
to him
all
who suffered
with various
illnesses
and
afflictions,
those who had seizures,
paralytics,
and
those possessed by demons,
and
he healed
them.
\VS{25}And
large
crowds
followed
him
from
Galilee,
the Decapolis,
Jerusalem,
Judea,
and
beyond
the Jordan River.

\par }\Chap{5}{\PP \VerseOne{1}When he saw
the crowds,
he went up
the mountain.
After
he sat down
his
disciples
came
to him.
\VS{2}Then
he began to teach
them
by saying:
\par }{\Q \VS{3}“Blessed
are the
poor
in spirit,
for
the
kingdom
of heaven
belongs
to them.
\par }{\Q \VS{4}“Blessed
are those who mourn,
for
they
will be comforted.
\par }{\Q \VS{5}“Blessed
are the meek,
for
they
will inherit
the earth.
\par }{\Q \VS{6}“Blessed
are those who hunger
and
thirst
for righteousness,
for
they
will be satisfied.
\par }{\Q \VS{7}“Blessed
are the merciful,
for
they
will be shown mercy.
\par }{\Q \VS{8}“Blessed
are the pure
in heart,
for
they
will see
God.
\par }{\Q \VS{9}“Blessed
are the peacemakers,
for
they
will be called
the children
of God.
\par }{\Q \VS{10}“Blessed
are those who are persecuted
for
righteousness,
for
the kingdom
of heaven
belongs
to them.
\par }{\Q \VS{11}“Blessed
are you
when
people insult
you
and
persecute
you and
say
all kinds
of evil
things about
you
falsely
on account
of me.
\VS{12}Rejoice
and
be glad
because
your
reward
is great
in
heaven,
for
they persecuted
the prophets
before
you
in the same way.
\par }{\SH Salt and Light
\par }{\PP \VS{13}“You
are
the salt
of the earth.
But
if
salt
loses
its flavor,
how can
it be made salty
again? It is no longer
good
for
anything except
to be thrown
out
and trampled on
by
people.
\VS{14}You
are
the light
of the world.
A city
located
on
a hill
cannot
be hidden.
\VS{15}People do not
light
a lamp
and
put
it
under
a basket
but
on
a lampstand,
and
it gives light
to all
in
the house.
\VS{16}In the same way,
let your
light
shine
before
people,
so that
they can see
your
good
deeds
and
give honor
to your
Father
in
heaven.
\par }{\SH Fulfillment of the Law and Prophets
\par }{\PP \VS{17}“Do
not
think
that
I have come
to abolish
the law
or
the prophets.
I have
not
come
to abolish
these things but
to fulfill them.
\VS{18}I tell
you
the truth,
until
heaven
and
earth
pass away
not
the smallest letter
or
stroke
of a letter will pass
from
the law
until
everything
takes place.
\VS{19}So
anyone
who breaks
one
of the
least
of these
commands
and
teaches
others
to do so
will be called
least
in
the
kingdom
of heaven,
but
whoever
obeys
them and
teaches
others to do so will be called
great
in
the
kingdom
of heaven.
\VS{20}For
I tell
you,
unless
your
righteousness
goes
beyond
that of the experts in the law
and
the Pharisees,
you will
never
enter
the
kingdom
of heaven.
\par }{\SH Anger and Murder
\par }{\PP \VS{21}“You have heard
that
it was said
to an older
generation, ‘{\QT{
Do not
murder}},’ and
‘whoever
murders
will be
subjected
to judgment.’
\VS{22}But
I
say
to you
that
anyone
who is angry
with a brother
will be
subjected
to judgment.
And
whoever
insults
a brother
will be
brought
before the
council,
and
whoever
says
‘Fool’
will be
sent
to
fiery
hell.
\VS{23}So then,
if
you bring
your
gift
to
the altar
and there
remember
that
your
brother
has
something
against
you,
\VS{24}leave
your
gift
there
in front
of the altar.
First
go
and
be reconciled
to your
brother
and
then
come
and present
your
gift.
\VS{25}Reach agreement
quickly
with your
accuser
while
on
the
way
to court, or he may hand
you
over
to the
judge,
and
the
judge
hand you over to the warden,
and
you will be thrown
into
prison.
\VS{26}I tell
you
the truth,
you will
never
get out
of there
until
you have paid
the last
penny!
\par }{\SH Adultery
\par }{\PP \VS{27}“You have heard
that
it was said, ‘{\QT{
Do not
commit adultery}}.’
\VS{28}But
I
say
to you
that
whoever
looks at
a woman
to
desire
her
has
already
committed adultery
with her
in
his
heart.
\VS{29}If
your
right
eye
causes
you
to sin,
tear
it
out
and
throw
it away! It is better
to
lose
one
of your
members
than
to have your
whole
body
thrown
into
hell.
\VS{30}If
your
right
hand
causes
you
to sin,
cut
it
off
and
throw
it away! It is better
to
lose
one
of your
members
than
to have your
whole
body
go
into
hell.
\par }{\SH Divorce
\par }{\PP \VS{31}“It was said, ‘{\QT{
Whoever
divorces
his
wife
must give
her
a legal document}}.’
\VS{32}But
I
say
to you
that
everyone
who divorces
his
wife,
except
for immorality,
makes
her
commit adultery,
and
whoever
marries
a divorced woman
commits adultery.
\par }{\SH Oaths
\par }{\PP \VS{33}“Again,
you have heard
that
it was said
to an older generation, ‘{\BD{
{\QT{
{\IT{Do not}}}}
break
an oath,
but
fulfill
your
vows to the Lord
break
an oath,
but
fulfill
your
vows to the Lord}}.’
\VS{34}But
I
say
to you,
do not
take oaths
at all
– not
by
heaven,
because
it is
the throne
of God,
\VS{35}not
by
earth,
because
it is
his
footstool,
and not
by
Jerusalem,
because
it is
the city
of the great
King.
\VS{36}Do not
take an oath
by
your
head,
because
you are
not
able
to make
one
hair
white
or
black.
\VS{37}Let your
word
be
‘Yes,
yes’
or ‘No,
no.’
More
than this
is
from
the evil one.
\par }{\SH Retaliation
\par }{\PP \VS{38}“You have heard
that
it was said, ‘{\QT{
An eye
for an
eye
and
a tooth
for
a tooth}}.’
\VS{39}But
I
say
to you,
do not
resist
the evildoer.
But
whoever
strikes
you
on
the right
cheek,
turn
the other
to him as well.
\VS{40}And
if someone wants
to sue
you
and
to take
your
tunic,
give
him
your coat
also.
\VS{41}And
if
anyone forces
you
to go
one
mile,
go
with
him
two.
\VS{42}Give
to the one who asks
you,
and
do not
reject
the one who wants
to borrow
from
you.
\par }{\SH Love for Enemies
\par }{\PP \VS{43}“You have heard
that
it was said, ‘{\QT{
Love
your
neighbor}}’
and
‘hate
your
enemy.’
\VS{44}But
I
say
to you,
love
your
enemy
and
pray
for
those
who persecute
you,
\VS{45}so that
you may be
like
your
Father
in
heaven,
since
he causes
the sun
to rise
on
the evil
and
the good,
and
sends rain
on
the righteous
and
the unrighteous.
\VS{46}For
if
you love
those
who love
you,
what
reward
do you have? Even
the tax collectors
do
the same, don’t they?
\VS{47}And
if
you
only
greet
your
brothers,
what
more
do you do? Even
the Gentiles
do
the same, don’t they?
\VS{48}So then,
be
perfect,
as
your
heavenly
Father
is
perfect.

\par }\Chap{6}{\PP \VerseOne{1}“Be careful
not
to display
your
righteousness
merely
to be seen
by people.
Otherwise
you have
no
reward
with
your
Father
in
heaven.
\VS{2}Thus
whenever
you do
charitable giving,
do not
blow a trumpet
before
you,
as
the hypocrites
do
in
synagogues
and
on
streets
so that
people will praise
them. I tell
you
the truth,
they have
their
reward.
\VS{3}But
when you do
your
giving,
do not
let your
left hand
know
what
your
right hand
is doing,
\VS{4}so that
your
gift
may be in
secret.
And
your
Father,
who sees
in
secret,
will reward
you.
\par }{\SH Private Prayer
\par }{\PP \VS{5}“Whenever
you pray,
do not
be
like
the hypocrites,
because
they love
to pray
while standing
in
synagogues
and
on
street
corners
so that
people
can see
them. Truly
I say
to you,
they have
their
reward.
\VS{6}But
whenever
you
pray,
go
into
your
room,
close
the door,
and pray
to your
Father
in
secret.
And
your
Father,
who sees
in
secret,
will reward
you.
\VS{7}When
you pray,
do
not
babble repetitiously
like
the Gentiles,
because
they think
that
by their
many words
they will be heard.
\VS{8}Do
not
be like
them,
for
your
Father
knows
what
you need
before
you
ask
him.
\VS{9}So
pray
this way:

\par }{\Q Our
Father
in
heaven,
may
your
name
be honored,
\par }{\Q \VS{10}may
your
kingdom
come,
\par }{\Q may
your
will be
done
on
earth
as
it is in
heaven.
\par }{\Q \VS{11}Give
us
today
our
daily
bread,
\par }{\Q \VS{12}and
forgive
us
our
debts,
as
we
ourselves have forgiven
our
debtors.
\par }{\Q \VS{13}And
do
not
lead
us
into
temptation,
but
deliver
us
from
the evil one.
\par }{\PP \VS{14}“For
if
you forgive
others
their
sins,
your
heavenly
Father
will
also
forgive
you.
\VS{15}But
if
you do
not
forgive
others,
your
Father
will
not
forgive
you your
sins.
\par }{\SH Proper Fasting
\par }{\PP \VS{16}“When
you fast,
do not
look
sullen
like
the hypocrites,
for
they make
their
faces
unattractive
so that people
will see them fasting.
I tell
you
the truth,
they have
their
reward.
\VS{17}When you
fast,
put oil
on your
head
and
wash
your
face,
\VS{18}so that
it will not
be obvious
to others
when
you are fasting,
but
only to your
Father
who is in
secret.
And
your
Father,
who sees
in
secret,
will reward
you.
\par }{\SH Lasting Treasure
\par }{\PP \VS{19}“Do
not
accumulate
for yourselves
treasures
on
earth,
where
moth
and
rust
destroy
and
where
thieves
break in
and
steal.
\VS{20}But
accumulate
for yourselves
treasures
in
heaven,
where
moth
and rust
do not
destroy,
and
thieves
do
not
break in
and steal.
\VS{21}For
where
your
treasure
is,
there
your
heart
will be
also.
\par }{\PP \VS{22}“The eye
is
the lamp
of the body.
If
then
your
eye
is
healthy,
your
whole
body
will be
full of light.
\VS{23}But
if
your
eye
is
diseased,
your
whole
body
will be
full of darkness. If
then
the light
in
you
is
darkness,
how great
is the darkness!
\par }{\PP \VS{24}“No one
can
serve
two
masters,
for
either
he will hate
the one
and
love
the other,
or
he will be devoted
to the one
and
despise
the other.
You cannot
serve
God
and
money.
\par }{\SH Do Not Worry
\par }{\PP \VS{25}“Therefore
I tell
you,
do
not
worry
about your
life,
what
you will eat
or drink,
or
about your
body,
what
you will wear.
Isn’t
there more to life
than
food
and
more to the body
than clothing?
\VS{26}Look at
the birds
in the sky: They do
not
sow,
or
reap,
or
gather
into
barns,
yet
your
heavenly
Father
feeds
them.
Aren’t
you
more
valuable
than they are?
\VS{27}And
which
of
you
by worrying
can
add
even one
hour
to
his
life?
\VS{28}Why
do you worry
about
clothing? Think
about how the flowers
of the field
grow;
they do
not
work
or
spin.
\VS{29}Yet
I tell
you
that
not even
Solomon
in
all
his
glory
was clothed
like
one
of these!
\VS{30}And
if
this is how God clothes the wild grass,
which is
here today
and
tomorrow
is tossed
into
the fire
to heat the oven, won’t
he clothe
you even more,
you
people of little faith?
\VS{31}So
then, don’t
worry
saying,
‘What
will we eat?’ or
‘What
will we drink?’ or
‘What
will we wear?’
\VS{32}For
the unconverted
pursue
these things,
and your
heavenly
Father
knows
that
you need
them.
\VS{33}But
above all
pursue
his kingdom
and
righteousness,
and
all
these things
will be given
to you as well.
\VS{34}So then,
do
not
worry
about
tomorrow,
for
tomorrow
will worry
about itself.
Today
has enough
trouble
of its own.

\par }\Chap{7}{\PP \VerseOne{1}“Do
not
judge
so that
you will
not
be judged.
\VS{2}For
by
the standard
you judge
you will be judged,
and
the measure
you use
will be the measure
you receive.
\VS{3}Why
do you see
the speck
in
your
brother’s
eye,
but
fail
to see
the beam of wood
in
your own?
\VS{4}Or
how
can you say
to your
brother,
‘Let
me remove
the speck
from
your
eye,’
while there is a beam
in
your own?
\VS{5}You hypocrite! First
remove
the beam
from
your own
eye,
and
then
you can see clearly
to remove
the speck
from
your
brother’s
eye.
\VS{6}Do not
give
what is holy
to dogs
or
throw
your
pearls
before
pigs;
otherwise
they will trample
them
under
their
feet
and
turn around
and tear
you
to pieces.
\par }{\SH Ask, Seek, Knock
\par }{\PP \VS{7}“Ask
and
it will be given
to you;
seek
and
you will find;
knock
and
the door will be opened
for you.
\VS{8}For
everyone
who asks
receives,
and
the one who seeks
finds,
and
to the one who knocks,
the door will be opened.
\VS{9}Is there anyone
among
you
who,
if his
son
asks for
bread,
will give
him
a stone?
\VS{10}Or
if he asks for
a fish,
will give
him
a snake?
\VS{11}If
you then, although
you
are
evil,
know
how to give
good
gifts
to your
children,
how much
more
will
your
Father
in
heaven
give
good gifts
to those who ask
him!
\VS{12}In everything,
treat
others
as
you would want
them to
treat
you,
for
this
fulfills the law
and
the prophets.
\par }{\SH The Narrow Gate
\par }{\PP \VS{13}“Enter
through
the narrow
gate,
because
the gate is wide
and
the way
is spacious
that leads
to destruction,
and
there are
many
who enter
through
it.
\VS{14}But the gate
is narrow
and
the way
is difficult
that leads
to
life,
and
there are
few
who find
it.
\par }{\SH A Tree and Its Fruit
\par }{\PP \VS{15}“Watch
out
for false prophets,
who
come
to
you
in
sheep’s
clothing
but
inwardly
are
voracious
wolves.
\VS{16}You will recognize
them
by their
fruit.
Grapes
are
not
gathered
from
thorns
or
figs
from
thistles, are they?
\VS{17}In the same way,
every
good
tree
bears
good
fruit,
but
the bad
tree
bears
bad
fruit.
\VS{18}A good
tree
is not
able
to bear
bad
fruit,
nor
a bad
tree
to bear
good
fruit.
\VS{19}Every
tree
that does
not
bear
good
fruit
is cut down
and
thrown
into
the fire.
\VS{20}So
then,
you will recognize
them
by
their
fruit.
\par }{\SH Judgment of Pretenders
\par }{\PP \VS{21}“Not
everyone
who says
to me,
‘Lord,
Lord,’
will enter
into
the kingdom
of heaven
– only
the one who does
the will
of my
Father
in
heaven.
\VS{22}On
that
day,
many
will
say
to me,
‘Lord,
Lord,
didn’t
we prophesy
in your
name,
and
in your
name
cast out
demons
and
do
many
powerful deeds?’
\VS{23}Then
I will declare
to them,
‘I
never
knew
you.
Go away
from
me,
you lawbreakers!’
\par }{\SH Hearing and Doing
\par }{\PP \VS{24}“Everyone
who
hears
these
words
of mine
and
does
them
is like
a wise
man
who
built
his
house
on
rock.
\VS{25}The rain
fell,
the flood
came,
and
the winds
beat
against that house,
but it
did not
collapse
because
it had been founded
on
rock.
\VS{26}Everyone
who hears
these
words
of mine
and
does
not
do
them
is like
a foolish
man
who
built
his
house
on
sand.
\VS{27}The rain
fell,
the flood
came,
and
the winds
beat
against that
house,
and
it collapsed;
it was
utterly destroyed!”
\par }{\PP \VS{28}When
Jesus
finished
saying
these things,
the crowds
were amazed
by
his
teaching,
\VS{29}because
he
taught
them
like
one who had
authority,
not
like
their
experts in the law.

\par }\Chap{8}{\PP \VerseOne{1}After
he
came down
from
the mountain,
large
crowds
followed
him.
\VS{2}And
a leper
approached,
and bowed low
before him,
saying, “Lord,
if
you are willing,
you can
make
me
clean.”
\VS{3}He
stretched out
his hand
and touched
him
saying,
“I am willing.
Be clean!” Immediately
his
leprosy
was cleansed.
\VS{4}Then
Jesus
said
to him,
“See
that you do
not
speak
to anyone, but
go,
show
yourself
to a priest,
and
bring
the offering
that
Moses
commanded,
as
a testimony
to them.”
\par }{\SH Healing the Centurion’s Servant
\par }{\PP \VS{5}When
he
entered
Capernaum,
a centurion
came
to him
asking for help:
\VS{6}“Lord,
my
servant
is lying
at
home
paralyzed,
in terrible anguish.”
\VS{7}Jesus said
to him,
“I
will come
and heal
him.”
\VS{8}But
the centurion
replied, “Lord,
I am
not
worthy
to have
you come
under
my
roof.
Instead,
just
say
the word
and
my
servant
will be healed.
\VS{9}For
I
too
am
a man
under
authority,
with soldiers
under
me.
I say
to this one,
‘Go’
and
he goes,
and
to another
‘Come’
and
he comes,
and
to my
slave ‘Do
this’
and
he does it.”
\VS{10}When
Jesus
heard
this he was amazed
and
said
to those who followed
him, “I tell
you
the truth,
I have
not
found
such
faith
in anyone in
Israel!
\VS{11}I tell
you,
many
will come
from
the east
and
west
to share the banquet
with
Abraham,
Isaac,
and
Jacob
in
the kingdom
of heaven,
\VS{12}but
the sons
of the kingdom
will be thrown out
into
the outer
darkness,
where there
will be
weeping
and
gnashing
of teeth.”
\VS{13}Then
Jesus
said
to the centurion,
“Go;
just as
you believed,
it will be done
for you.”
And
the servant
was healed
at
that
hour.
\par }{\SH Healings at Peter’s House
\par }{\PP \VS{14}Now
when Jesus
entered
Peter’s
house,
he saw
his
mother-in-law
lying down,
sick with a fever.
\VS{15}He
touched
her
hand,
and
the fever
left
her.
Then
she got up
and
began to serve
them.
\VS{16}When
it was
evening,
many
demon-possessed
people were brought
to him.
He
drove out
the spirits
with a word,
and
healed
all
who were
sick.
\VS{17}In this way
what was spoken
by
Isaiah
the prophet
was fulfilled:
\par }{\Q “{\QT{
He took
our
weaknesses
and
carried
our diseases}}.”
\par }{\SH Challenging Professed Followers
\par }{\PP \VS{18}Now
when Jesus
saw
a large crowd
around
him,
he gave orders
to go
to
the other side of the lake.
\VS{19}Then
an expert in the law
came
to him
and said,
“Teacher,
I will follow
you
wherever
you go.”
\VS{20}Jesus
said
to him,
“Foxes
have
dens,
and
the birds
in the sky
have nests,
but
the Son
of Man
has
no
place
to lay
his head.”
\VS{21}Another
of the disciples
said
to him,
“Lord,
let
me
first
go
and
bury
my
father.”
\VS{22}But
Jesus
said
to him,
“Follow
me,
and
let
the dead
bury
their own
dead.”
\par }{\SH Stilling of a Storm
\par }{\PP \VS{23}As
he got
into
the boat,
his
disciples
followed
him.
\VS{24}And
a great
storm
developed
on
the sea
so
that the waves
began to swamp
the boat.
But
he
was asleep.
\VS{25}So
they came
and woke
him
up
saying,
“Lord,
save
us! We are about to die!”
\VS{26}But
he said
to them,
“Why
are you
cowardly,
you people of little faith?” Then
he got up
and rebuked
the winds
and
the sea,
and
it was
dead
calm.
\VS{27}And
the men
were amazed
and said, “What sort of person
is
this? Even
the winds
and
the sea
obey
him!”
\par }{\SH Healing the Gadarene Demoniacs
\par }{\PP \VS{28}When
he
came
to
the other side,
to
the region
of the Gadarenes,
two
demon-possessed men
coming
from
the tombs
met him. They were extremely
violent,
so that
no
one
was able
to pass
by
that
way.
\VS{29}They cried out,
“Son
of God,
leave us
alone! Have you come
here
to torment
us
before
the time?”
\VS{30}A large
herd
of pigs
was feeding
some distance
from
them.
\VS{31}Then
the demons
begged
him, “If
you drive
us
out,
send
us
into
the herd
of pigs.”
\VS{32}And
he said, “Go!” So
they came out
and went
into
the pigs,
and
the herd
rushed down
the steep slope
into
the lake
and
drowned
in
the water.
\VS{33}The
herdsmen
ran off,
went
into
the town,
and told
everything
that had happened to the demon-possessed men.
\VS{34}Then
the entire
town
came out
to
meet
Jesus.
And
when
they saw
him,
they begged
him to leave
their
region.

\par }\Chap{9}{\PP \VerseOne{1}After getting
into
a boat
he crossed
to the other side and
came
to
his own
town.
\VS{2}Just then some people brought
to him
a paralytic
lying
on
a stretcher.
When
Jesus
saw
their
faith,
he said
to the paralytic,
“Have courage,
son! Your
sins
are forgiven.”
\VS{3}Then
some
of the experts in the law
said
to
themselves,
“This man
is blaspheming!”
\VS{4}When
Jesus
saw
their
reaction
he said,
“Why
do you respond
with evil
in
your
hearts?
\VS{5}Which
is
easier,
to say,
‘Your
sins
are forgiven’
or
to say,
‘Stand up
and
walk’?
\VS{6}But
so that
you may know
that
the Son
of Man
has
authority
on
earth
to forgive
sins”
– then
he said
to the paralytic –
“Stand up,
take
your
stretcher,
and
go
home.”
\VS{7}And
he stood up
and went
home.
\VS{8}When
the crowd
saw
this, they were afraid
and
honored
God
who had given
such
authority
to men.
\par }{\SH The Call of Matthew; Eating with Sinners
\par }{\PP \VS{9}As Jesus
went on
from there,
he saw
a man
named
Matthew
sitting
at
the tax booth. “Follow
me,”
he said
to him.
And
he got up
and followed
him.
\VS{10}As Jesus
was having a meal
in
Matthew’s house,
many
tax collectors
and
sinners
came
and ate with
Jesus
and
his
disciples.
\VS{11}When
the Pharisees
saw
this they said
to his
disciples,
“Why
does
your
teacher
eat
with
tax collectors
and
sinners?”
\VS{12}When
Jesus heard
this he said,
“Those who are healthy
don’t
need
a physician,
but
those who are sick
do.
\VS{13}Go
and
learn
what
this saying means: ‘{\QT{
I want
mercy
and
not
sacrifice}}.’ For
I did
not
come
to call
the righteous,
but
sinners.”
\par }{\SH The Superiority of the New
\par }{\PP \VS{14}Then
John’s
disciples
came
to Jesus
and asked,
“Why
do
we
and
the Pharisees
fast
often, but
your
disciples
don’t
fast?”
\VS{15}Jesus
said
to them,
“The wedding guests
cannot
mourn
while
the bridegroom
is
with
them,
can they? But
the days
are coming
when
the bridegroom
will be taken
from
them,
and
then
they will fast.
\VS{16}No one
sews
a patch
of unshrunk cloth
on
an old
garment,
because
the patch
will pull away
from
the garment
and
the tear
will be
worse.
\VS{17}And no one
pours
new
wine
into
old
wineskins;
otherwise
the skins
burst
and
the wine
is spilled out
and
the skins
are destroyed.
Instead
they put
new
wine
into
new
wineskins
and
both
are preserved.”
\par }{\SH Restoration and Healing
\par }{\PP \VS{18}As he
was saying
these things,
a ruler
came,
bowed low
before him,
and said,
“My
daughter
has
just
died,
but
come
and lay
your
hand
on
her
and
she will live.”
\VS{19}Jesus
and
his
disciples
got up
and
followed
him.
\VS{20}But
a woman
who had been suffering from a hemorrhage
for twelve
years
came up
behind
him and touched
the edge
of his
cloak.
\VS{21}For
she kept saying
to
herself, “If
only
I touch
his
cloak,
I will be healed.”
\VS{22}But
when Jesus
turned
and
saw
her
he said,
“Have courage,
daughter! Your
faith
has made
you
well.”
And
the woman
was healed
from
that
hour.
\VS{23}When
Jesus
entered
the ruler’s
house
and
saw
the flute players
and
the disorderly
crowd,
\VS{24}he said,
“Go away,
for
the girl
is
not
dead
but
asleep.”
And
they began making fun
of him.
\VS{25}But
when
the crowd
had been put outside,
he went in
and gently took
her
by the hand,
and
the girl
got up.
\VS{26}And
the news
of this
spread
throughout
that
region.
\par }{\SH Healing the Blind and Mute
\par }{\PP \VS{27}As Jesus
went on
from there,
two
blind men
followed
him, shouting, “Have mercy
on us,
Son
of David!”
\VS{28}When he went
into
the house,
the blind men
came
to him.
Jesus
said
to them,
“Do you believe
that
I am able
to do
this?” They said
to him,
“Yes,
Lord.”
\VS{29}Then
he touched
their
eyes
saying,
“Let it be done
for you
according to
your
faith.”
\VS{30}And
their
eyes
were opened.
Then
Jesus
sternly warned
them,
“See
that no one
knows about this.”
\VS{31}But
they went out
and spread the news about
him
throughout
that
entire
region.
\par }{\PP \VS{32}As
they
were going away,
a man who could not talk
and was demon-possessed
was brought
to him.
\VS{33}After the demon
was cast out,
the man who had been mute
spoke.
The crowds
were amazed
and said,
“Never
has
anything like this
been seen
in
Israel!”
\VS{34}But
the Pharisees
said,
“By
the ruler
of demons
he casts out
demons.”
\par }{\SH Workers for the Harvest
\par }{\PP \VS{35}Then
Jesus
went throughout
all
the towns
and
villages,
teaching
in
their
synagogues,
preaching
the good news
of the kingdom,
and
healing
every kind
of disease
and
sickness.
\VS{36}When
he saw
the crowds,
he had compassion
on
them
because
they were
bewildered
and
helpless,
like
sheep
without
a shepherd.
\VS{37}Then
he said
to his
disciples,
“The harvest
is plentiful,
but
the workers
are few.
\VS{38}Therefore
ask
the Lord
of the harvest
to send out
workers
into
his
harvest.”

\par }\Chap{10}{\PP \VerseOne{1}Jesus called
his
twelve
disciples
and gave
them
authority
over unclean
spirits
so
they could cast
them
out
and
heal
every
kind of disease and
sickness.
\VS{2}Now
these
are
the names
of the twelve
apostles:
 first,
Simon
(called
Peter), and
Andrew
his
brother;
James
son of Zebedee
and
John
his
brother;
\VS{3}Philip
and
Bartholomew;
Thomas
and
Matthew
the tax collector;
James
the son of Alphaeus,
and
Thaddaeus;
\VS{4}Simon
the Zealot
and
Judas
Iscariot,
who betrayed
him.
\par }{\PP \VS{5}Jesus
sent out
these
twelve,
instructing
them
as follows: “Do not
go
to
Gentile
regions
and
do not
enter
any Samaritan
town.
\VS{6}Go
instead
to
the lost
sheep
of the house
of Israel.
\VS{7}As you
go,
preach
this message: ‘The kingdom
of heaven
is near!’
\VS{8}Heal
the sick,
raise
the dead,
cleanse
lepers,
cast out
demons.
Freely
you received,
freely
give.
\VS{9}Do not
take
gold,
silver,
or
copper
in
your
belts,
\VS{10}no
bag
for
the journey,
or
an extra
tunic,
or
sandals
or
staff,
for
the worker
deserves
his
provisions.
\VS{11}Whenever
you enter
a town
or
village,
find out
who
is
worthy
there
and stay
with
them
until
you leave.
\VS{12}As you enter
the house,
give
it
greetings.
\VS{13}And
if
the house
is
worthy,
let your
peace
come
on
it,
but
if
it is
not
worthy,
let your
peace
return
to
you.
\VS{14}And
if anyone
will
not
welcome
you
or
listen
to your
message,
shake
the dust
off
your
feet
as you leave
that house
or
that
town.
\VS{15}I tell
you
the truth,
it will be
more bearable
for the region
of Sodom
and
Gomorrah
on
the day
of judgment
than
for that
town!
\par }{\SH Persecution of Disciples
\par }{\PP \VS{16}“I
am sending
you
out
like
sheep
surrounded
by
wolves,
so
be
wise
as
serpents
and
innocent
as
doves.
\VS{17}Beware
of
people,
because
they will hand
you
over
to
councils
and
flog
you
in
their
synagogues.
\VS{18}And
you will be brought
before
governors
and
kings
because of
me,
as
a witness
to them
and
the Gentiles.
\VS{19}Whenever
they hand
you
over
for trial, do not
worry
about how
to speak
or
what
to say, for
what
you should say
will be given
to you
at
that
time.
\VS{20}For
it is
not
you
speaking,
but
the Spirit
of your
Father
speaking
through
you.
\par }{\PP \VS{21}“Brother
will hand over
brother
to
death,
and
a father
his child.
Children
will rise
against
parents
and
have
them
put to death.
\VS{22}And
you will be
hated
by
everyone
because of
my
name.
But
the one who endures
to
the end
will be saved.
\VS{23}Whenever
they persecute
you
in
one place,
flee
to
another.
I tell
you
the truth,
you will
not
finish
going through all the towns
of Israel
before
the Son
of Man
comes.
\par }{\PP \VS{24}“A disciple
is
not
greater than
his teacher,
nor
a slave
greater than
his
master.
\VS{25}It is enough
for the disciple
to
become
like
his
teacher,
and
the slave
like
his
master.
If
they have called
the head of the house
‘Beelzebul,’
how much
more
will they defame the members of his
household!
\par }{\SH Fear God, Not Man
\par }{\PP \VS{26}“Do not
be afraid
of them,
for
nothing
is
hidden
that
will
not
be revealed,
and
nothing is secret
that
will
not
be made known.
\VS{27}What
I say
to you
in
the dark,
tell
in
the light,
and
what
is whispered
in your ear,
proclaim
from
the housetops.
\VS{28}Do not
be afraid
of those who kill
the body
but
cannot
kill
the soul.
Instead,
fear
the one who is able
to destroy
both
soul
and
body
in
hell.
\VS{29}Aren’t
two
sparrows
sold
for a penny? Yet
not
one
of
them
falls
to
the ground
apart from
your
Father’s will.
\VS{30}Even all
the hairs
on your
head
are
numbered.
\VS{31}So
do not
be afraid;
you
are more valuable
than many
sparrows.
\par }{\PP \VS{32}“Whoever,
then,
acknowledges
me
before
people,
I
will acknowledge
before
my
Father
in
heaven.
\VS{33}But
whoever
denies
me
before
people,
I
will deny
him
also
before
my
Father
in
heaven.
\par }{\SH Not Peace, but a Sword
\par }{\PP \VS{34}“Do not
think
that
I have come
to bring
peace
to
the earth.
I have
not
come
to bring
peace
but
a sword.
\VS{35}For
I have come
to set
{\IT{
a man
against
his
father,
a daughter
against
her
mother,
and
a daughter-in-law
against
her
mother-in-law,}}
\VS{36}{\IT{
and
a man’s
enemies
will be the members of
his
household}}.
\par }{\PP \VS{37}“Whoever loves
father
or
mother
more than
me
is
not
worthy
of me,
and
whoever loves
son
or
daughter
more than
me
is
not
worthy
of me.
\VS{38}And
whoever
does not
take up
his
cross
and
follow
me
is
not
worthy
of me.
\VS{39}Whoever finds
his
life
will lose
it,
and
whoever loses
his
life
because of
me
will find
it.
\par }{\SH Rewards
\par }{\PP \VS{40}“Whoever receives
you
receives
me,
and
whoever receives
me
receives
the one who sent
me.
\VS{41}Whoever receives
a prophet
in
the name
of a prophet
will receive
a prophet’s
reward.
Whoever receives
a righteous person
in
the name
of a righteous person
will receive
a righteous person’s
reward.
\VS{42}And
whoever
gives
only
a cup
of cold water
to one
of these
little
ones in
the name
of a disciple,
I tell
you
the truth,
he will
never
lose
his
reward.”

\par }\Chap{11}{\PP \VerseOne{1}When
Jesus
had finished
instructing
his
twelve
disciples,
he went on
from there
to teach
and
preach
in
their
towns.
\par }{\SH Jesus and John the Baptist
\par }{\PP \VS{2}Now when John
heard
in
prison
about the deeds
Christ
had done, he sent
his
disciples to ask a question:
\VS{3}“Are you
the one who is
to come,
or
should we look
for another?”
\VS{4}Jesus
answered
them, “Go
tell
John
what
you hear
and
see:
\VS{5}The blind
see,
the lame
walk,
lepers
are cleansed,
the deaf
hear,
the dead
are raised,
and
the poor
have good news proclaimed to them.
\VS{6}Blessed
is
anyone who
takes no
offense
at
me.”
\par }{\PP \VS{7}While
they
were going away,
Jesus
began
to speak
to the crowd
about
John: “What
did you go out
into
the wilderness
to see? A reed
shaken
by
the wind?
\VS{8}What
did you go out
to see? A man
dressed
in
fancy clothes? Look,
those who wear
fancy clothes
are in
the homes
of kings!
\VS{9}What
did you go out
to see? A prophet? Yes,
I tell
you,
and
more than
a prophet.
\VS{10}This
is
the one about
whom
it is written:
\par }{\Q ‘{\QT{
Look,
I
am sending
my
messenger
ahead
of you,}}
\par }{\Q {\IT{
{\BD{
who
will prepare
your
way
before
you.}}}}’
\par }{\PP \VS{11}“I tell
you
the truth,
among those
born
of women,
no one
has arisen
greater than
John
the Baptist.
Yet the one who is least
in
the kingdom
of heaven
is greater than
he
is.
\VS{12}From
the days
of John
the Baptist
until
now
the kingdom
of heaven
has suffered violence,
and forceful people lay hold
of it.
\VS{13}For
all
the prophets
and
the law
prophesied
until
John appeared.
\VS{14}And
if
you are willing
to accept
it,
he is
Elijah,
who is
to come.
\VS{15}The one who has
ears
had better listen!
\par }{\PP \VS{16}“To what
should I compare
this
generation? They are
like
children
sitting
in
the marketplaces
who
call out
to one another,
\par }{\Q \VS{17}‘We played
the flute
for you,
yet
you
did not
dance;
\par }{\Q we wailed in mourning,
yet
you did
not
weep.’
\par }{\PI \VS{18}For
John
came
neither
eating
nor
drinking,
and
they say,
‘He has
a demon!’
\VS{19}The Son
of Man
came
eating
and
drinking,
and
they say,
‘Look at
him,
a glutton
and
a drunk,
a friend
of tax collectors
and
sinners!’ But
wisdom
is vindicated
by
her
deeds.”
\par }{\SH Woes on Unrepentant Cities
\par }{\PP \VS{20}Then
Jesus began
to criticize openly
the cities
in
which
he had done
many
of his
miracles,
because
they did
not
repent.
\VS{21}“Woe
to you,
Chorazin! Woe
to you,
Bethsaida! If
the miracles
done
in
you
had been done
in
Tyre
they would have repented
long ago in
sackcloth
and
ashes.
\VS{22}But
I tell
you,
it will be
more bearable
for Tyre
and
Sidon
on
the day
of judgment
than
for you!
\VS{23}And
you,
Capernaum,
will you be exalted
to
heaven? No,
you will be thrown down
to
Hades! For
if
the miracles
done
among
you
had been done
in
Sodom,
it would have continued
to
this day.
\VS{24}But
I tell
you,
it will be
more bearable
for
the region
of Sodom
on
the day
of judgment
than
for you!”
\par }{\SH Jesus’ Invitation
\par }{\PP \VS{25}At
that
time
Jesus
said, “I praise
you,
Father,
Lord
of heaven
and
earth,
because
you have hidden
these things
from
the wise
and
intelligent,
and
revealed
them
to little children.
\VS{26}Yes,
Father,
for
this
was
your
gracious will.
\VS{27}All things
have been handed over
to me
by
my
Father.
No one
knows
the Son
except
the Father,
and no one
knows
the Father
except
the Son
and
anyone
to whom
the Son
decides
to reveal him.
\VS{28}Come
to
me,
all
you who are weary
and
burdened,
and I
will give
you
rest.
\VS{29}Take
my
yoke
on
you
and
learn
from
me,
because
I am
gentle
and
humble
in heart,
and
you will find
rest
for your
souls.
\VS{30}For
my
yoke
is easy to bear,
and
my
load
is
not hard to carry.”

\par }\Chap{12}{\PP \VerseOne{1}At
that
time
Jesus
went
through
the grain fields
on a Sabbath.
His
disciples
were hungry,
and
they began
to pick
heads of wheat
and
eat them.
\VS{2}But
when
the Pharisees
saw
this they said
to him,
“Look,
your
disciples
are doing
what
is against
the law
to do
on
the Sabbath.”
\VS{3}He said
to them,
“Haven’t
you read
what
David
did
when
he
and
his
companions
were hungry –
\VS{4}how
he entered
the house
of God
and
they ate
the sacred
bread,
which
was
against
the law
for him
or
his
companions
to eat,
but
only
for the priests?
\VS{5}Or
have you
not
read
in
the law
that
the priests
in
the temple
desecrate
the Sabbath
and
yet are
not guilty?
\VS{6}I tell
you
that
something
greater than
the temple
is
here.
\VS{7}If
you had known
what
this means: ‘{\QT{
I want
mercy
and
not
sacrifice}},’ you would
not
have condemned
the innocent.
\VS{8}For
the Son
of Man
is
lord
of the Sabbath.”
\par }{\PP \VS{9}Then
Jesus left
that place
and entered
their
synagogue.
\VS{10}A man
was there who had
a withered
hand.
And
they asked
Jesus, “Is it
lawful
to heal
on the Sabbath?” so that
they could accuse
him.
\VS{11}He said
to them,
“Would
not any one of
you,
if he had
one
sheep
that fell
into
a pit
on the Sabbath,
take hold
of it
and
lift
it out?
\VS{12}How much
more valuable
is a person
than a sheep! So
it is lawful
to do
good
on the Sabbath.”
\VS{13}Then
he said
to the man,
“Stretch out
your
hand.”
He
stretched
it out
and
it was restored,
as healthy
as
the other.
\VS{14}But
the Pharisees
went out
and plotted
against
him,
as to how
they could assassinate
him.
\par }{\SH God’s Special Servant
\par }{\PP \VS{15}Now
when Jesus
learned
of this, he went away
from there.
Great crowds
followed
him,
and
he healed
them
all.
\VS{16}But
he sternly warned
them
not
to make
him
known.
\VS{17}This fulfilled
what was spoken
by
Isaiah
the prophet:
\par }{\Q \VS{18}“{\QT{
Here
is my
servant
whom
I have chosen,}}
\par }{\Q {\QT{
the one I
love,
in whom
I
take great delight}}.
\par }{\Q {\QT{
I will put
my
Spirit
on
him,
and
he will proclaim
justice
to the nations}}.
\par }{\Q \VS{19}{\QT{
He will
not
quarrel
or
cry out,}}
\par }{\Q {\QT{
nor
will
anyone
hear
his
voice
in
the streets}}.
\par }{\Q \VS{20}{\QT{
He will
not
break
a
bruised
reed
or extinguish
a smoldering wick,}}
\par }{\Q {\QT{
until
he brings
justice
to
victory}}.
\par }{\Q \VS{21}{\QT{
And
in his
name
the Gentiles
will hope}}.”
\par }{\SH Jesus and Beelzebul
\par }{\PP \VS{22}Then
they brought
to him
a demon-possessed man
who was blind
and
mute.
Jesus healed
him
so that
he could speak
and
see.
\VS{23}All
the crowds
were amazed
and
said,
“Could
this one
be
the Son
of David?”
\VS{24}But
when the Pharisees
heard
this they said,
“He
does
not
cast out
demons
except
by
the power of Beelzebul,
the ruler
of demons!”
\VS{25}Now when Jesus realized
what they
were thinking,
he said
to them, “Every
kingdom
divided
against
itself
is destroyed,
and
no
town
or
house
divided
against
itself
will stand.
\VS{26}So
if
Satan
casts out
Satan,
he is divided
against
himself.
How
then
will
his
kingdom
stand?
\VS{27}And
if
I
cast out
demons
by
Beelzebul,
by
whom
do your
sons
cast
them
out? For
this
reason they will be
your
judges.
\VS{28}But
if
I
cast out
demons
by
the Spirit
of God,
then
the kingdom
of God
has already overtaken
you.
\VS{29}How
else
can
someone
enter
a strong man’s
house
and
steal
his
property,
unless
he
first
ties up
the strong man? Then
he can thoroughly plunder
the house.
\VS{30}Whoever is
not
with
me
is
against
me,
and
whoever does
not
gather
with
me
scatters.
\VS{31}For
this
reason I tell
you,
people
will be forgiven
for every
sin
and
blasphemy,
but the
blasphemy
against the Spirit
will
not
be forgiven.
\VS{32}Whoever
speaks
a word
against
the Son
of Man
will be forgiven.
But
whoever
speaks
against
the Holy
Spirit
will
not
be forgiven,
either
in
this
age
or
in
the age to come.
\par }{\SH Trees and Their Fruit
\par }{\PP \VS{33}“Make
a tree
good
and
its
fruit
will be good,
or
make
a tree
bad
and
its
fruit
will be bad,
for
a tree
is known
by
its fruit.
\VS{34}Offspring
of vipers! How
are you able
to say
anything good,
since you are
evil? For
the mouth
speaks
from
what fills
the heart.
\VS{35}The good
person
brings
good things
out
of his good
treasury,
and
the evil
person
brings
evil things
out
of his evil
treasury.
\VS{36}I tell
you
that
on
the day
of judgment,
people
will give
an account
for
every
worthless
word
they speak.
\VS{37}For
by
your
words
you will be justified,
and
by
your
words
you will be condemned.”
\par }{\SH The Sign of Jonah
\par }{\PP \VS{38}Then
some
of the experts in the law
along with
some Pharisees
answered
him, “Teacher,
we want
to see
a sign
from
you.”
\VS{39}But
he answered
them, “An evil
and
adulterous
generation
asks for
a sign,
but
no
sign
will be given
to it
except
the sign
of the prophet
Jonah.
\VS{40}For
just as
Jonah
was
{\QT{
in
the belly
of the huge fish
for three
days
and
three
nights}}, so
the Son
of Man
will be
in
the heart
of the earth
for three
days
and
three
nights.
\VS{41}The people
of Nineveh
will stand up
at
the judgment
with
this
generation
and
condemn
it,
because
they repented
when Jonah
preached to them – and now, something greater than Jonah is here!
\VS{42}The queen
of the South
will rise up
at
the judgment
with
this
generation
and
condemn
it,
because
she came
from
the ends
of the earth
to hear
the wisdom
of Solomon
– and
now,
something greater than
Solomon
is here!
\par }{\SH The Return of the Unclean Spirit
\par }{\PP \VS{43}“When
an unclean
spirit
goes out
of
a person,
it passes
through
waterless
places
looking
for rest
but
does
not
find it.
\VS{44}Then
it says,
‘I will return
to
the home
I left.’
When
it returns,
it finds
the house empty,
swept clean,
and
put in order.
\VS{45}Then
it goes
and
brings
with
it
seven
other
spirits
more evil
than itself,
and
they go in
and live
there,
so the last state
of that
person
is
worse than
the first.
It will be
that way
for this
evil
generation as well!”
\par }{\SH Jesus’ True Family
\par }{\PP \VS{46}While Jesus
was
still
speaking
to the crowds,
his
mother
and
brothers
came and stood
outside,
asking
to speak
to him.
\VS{47}Someone told him, “Look, your mother and your brothers are standing outside wanting to speak to you.”
\VS{48}To the one
who had said
this, Jesus replied, “Who
is
my
mother
and
who
are
my
brothers?”
\VS{49}And
pointing
toward
his
disciples
he said,
“Here
are my
mother
and
my
brothers!
\VS{50}For
whoever
does
the will
of my
Father
in
heaven
is
my
brother
and
sister
and
mother.”

\par }\Chap{13}{\PP \VerseOne{1}On
that
day
after
Jesus
went out
of the house,
he sat
by
the lake.
\VS{2}And
such
a large
crowd
gathered
around him
that he got
into
a boat
to sit
while the whole
crowd
stood
on
the shore.
\VS{3}He
told
them
many things
in
parables,
saying: “Listen! A sower
went out
to sow.
\VS{4}And
as he sowed,
some seeds fell
along
the path,
and
the birds
came
and devoured
them.
\VS{5}Other
seeds fell
on
rocky ground
where
they did
not
have
much
soil.
They sprang up
quickly
because
the soil
was
not
deep.
\VS{6}But
when the sun
came up,
they were scorched,
and
because
they did
not
have
sufficient root,
they withered.
\VS{7}Other
seeds fell
among
the thorns,
and
they grew up
and
choked
them.
\VS{8}But
other
seeds fell
on
good
soil
and
produced
grain,
some
a hundred
times as much, some sixty,
and some thirty.
\VS{9}The one who has
ears
had better listen!”
\par }{\PP \VS{10}Then
the disciples
came
to him
and said,
“Why
do you speak
to them
in
parables?”
\VS{11}He replied, “You
have been given
the opportunity to know
the secrets
of the kingdom
of heaven,
but
they have
not.
\VS{12}For
whoever
has
will be given
more, and
will have an abundance.
But
whoever
does
not
have,
even
what
he has
will be taken
from
him.
\VS{13}For
this
reason I speak
to them
in
parables: Although
they see
they do
not
see,
and although
they hear
they do
not
hear
nor
do they understand.
\VS{14}And
concerning them
the prophecy
of Isaiah
is fulfilled
that says:
\par }{\Q ‘{\QT{
You will listen carefully
yet
will
never
understand,}}
\par }{\Q {\QT{
you will look closely
yet
will
never
comprehend}}.
\par }{\Q \VS{15}{\QT{
For
the heart
of this
people
has become dull;}}
\par }{\Q {\QT{
they are hard of hearing,}}
\par }{\Q {\QT{
and
they have shut
their
eyes,}}
\par }{\Q {\QT{
so that they would
not
see
with their eyes}}
\par }{\Q {\QT{
and
hear
with their ears}}
\par }{\Q {\QT{
and
understand
with their hearts}}
\par }{\Q {\QT{
and
turn,
and
I would heal
them}}.’
\par }{\PP \VS{16}“But
your
eyes
are blessed
because
they see,
and
your
ears
because
they hear.
\VS{17}For
I tell
you
the truth,
many
prophets
and
righteous people
longed
to see
what
you see
but
did
not
see
it, and
to hear
what
you hear
but
did
not
hear it.
\par }{\PP \VS{18}“So
listen
to the parable
of the sower:
\VS{19}When
anyone
hears
the word
about the kingdom
and
does
not
understand
it, the evil one
comes
and
snatches
what was sown
in
his
heart;
this
is
the seed sown
along
the path.
\VS{20}The seed sown
on
rocky ground
is
the person
who hears
the word
and
immediately
receives
it
with
joy.
\VS{21}But
he has
no
root
in
himself
and
does
not endure;
when trouble
or
persecution
comes
because
of the word,
immediately
he falls away.
\VS{22}The
seed sown
among
thorns
is
the person who hears
the word,
but
worldly
cares
and
the seductiveness
of wealth
choke
the word,
so
it produces nothing.
\VS{23}But
as for the seed sown
on
good
soil,
this
is
the person who hears
the word
and
understands.
He bears fruit,
yielding
a hundred,
sixty,
or thirty times what was sown.”
\par }{\SH The Parable of the Weeds
\par }{\PP \VS{24}He presented
them
with another
parable: “The kingdom
of heaven
is like
a person
who sowed
good
seed
in
his
field.
\VS{25}But
while everyone
was sleeping,
an enemy
came
and
sowed
weeds
among
the wheat
and
went away.
\VS{26}When
the plants
sprouted
and
bore
grain,
then
the weeds
also
appeared.
\VS{27}So
the slaves
of the owner
came
and said
to him,
‘Sir,
didn’t
you sow
good
seed
in
your
field? Then
where
did
the weeds
come from?’
\VS{28}He said,
‘An enemy
has done
this.’
So
the slaves replied,
‘Do you want
us to go
and gather
them?’
\VS{29}But
he said,
‘No,
since in
gathering
the weeds
you may uproot
the wheat
with
them.
\VS{30}Let
both
grow together
until
the harvest.
At
harvest
time
I will tell
the reapers,
“First
collect
the weeds
and
tie
them
in
bundles
to
be burned,
but
then gather
the wheat
into
my
barn.” ’ ”
\par }{\SH The Parable of the Mustard Seed
\par }{\PP \VS{31}He gave
them
another
parable: “The kingdom
of heaven
is
like
a mustard seed
that
a man
took
and sowed
in
his
field.
\VS{32}It is
the smallest
of all
the seeds,
but
when
it has grown
it is
the greatest
garden
plant and
becomes
a tree,
so that
the wild birds
come
and
nest
in
its
branches.”
\par }{\SH The Parable of the Yeast
\par }{\PP \VS{33}He told
them
another
parable: “The kingdom
of heaven
is
like
yeast
that
a woman
took
and mixed
with
three
measures
of flour
until
all
the dough had risen.”
\par }{\SH The Purpose of Parables
\par }{\PP \VS{34}Jesus
spoke
all
these things
in
parables
to the crowds;
he did
not
speak
to them
without
a parable.
\VS{35}This
fulfilled
what was spoken
by
the prophet:
\par }{\Q “{\QT{
I will open
my
mouth
in
parables,}}
\par }{\Q {\QT{
I will announce
what has been hidden
from
the foundation of the world}}.”
\par }{\SH Explanation for the Disciples
\par }{\PP \VS{36}Then
he left
the crowds
and went
into
the house.
And
his
disciples
came
to him
saying,
“Explain
to us
the parable
of the weeds
in the field.”
\VS{37}He
answered,
“The one who sowed
the good
seed
is
the Son
of Man.
\VS{38}The
field
is
the world
and
the good
seed
are
the people
of the kingdom.
The
weeds
are
the people
of the evil one,
\VS{39}and
the enemy
who sows
them
is
the devil.
The
harvest
is
the end
of the age,
and
the reapers
are
angels.
\VS{40}As
the weeds
are collected
and
burned
with fire,
so
it will be
at
the end
of the age.
\VS{41}The
Son
of Man
will send
his
angels,
and
they will gather
from
his
kingdom
everything
that causes sin
as well as
all lawbreakers.
\VS{42}They will
{\QT{
throw
them
into
the fiery
furnace}}, where there
will be
weeping
and
gnashing
of teeth.
\VS{43}Then
{\IT{
the righteous
will shine
like
the sun
in
the kingdom
of their
Father}}. The one who has
ears
had better listen!
\par }{\SH Parables on the Kingdom of Heaven
\par }{\PP \VS{44}“The kingdom
of heaven
is
like
a treasure,
hidden
in
a field,
that
a person
found
and
hid.
Then because of
joy
he went
and
sold
all
that he had
and
bought
that
field.
\par }{\PP \VS{45}“Again,
the kingdom
of heaven
is
like
a merchant
searching
for fine
pearls.
\VS{46}When
he found
a
pearl
of great value,
he went out
and sold
everything
he had
and
bought
it.
\par }{\PP \VS{47}“Again,
the kingdom
of heaven
is
like
a net
that was cast
into
the sea
that caught
all
kinds of fish.
\VS{48}When
it was
full,
they pulled
it ashore,
sat down,
and put
the good
fish into
containers
and
threw
the bad
away.
\VS{49}It will be
this way
at
the end
of the age.
Angels
will come
and
separate
the evil
from
the righteous
\VS{50}and
{\IT{
throw
them
into
the fiery
furnace}}, where there
will be
weeping
and
gnashing
of teeth.
\par }{\PP \VS{51}“Have you understood
all
these things?” They replied,
“Yes.”
\VS{52}Then he said
to them,
“Therefore
every
expert in the law
who has been trained
for the kingdom
of heaven
is
like
the owner of a house
who
brings out
of
his
treasure
what is new
and
old.”
\par }{\SH Rejection at Nazareth
\par }{\PP \VS{53}Now when
Jesus
finished
these
parables,
he moved on
from there.
\VS{54}Then
he came
to
his
hometown
and began to teach
the people
in
their
synagogue.
They
were astonished
and
said,
“Where
did this
man get such wisdom
and
miraculous powers?
\VS{55}Isn’t
this
the carpenter’s
son? Isn’t
his
mother
named
Mary? And
aren’t his
brothers
James,
Joseph,
Simon,
and Judas?
\VS{56}And
aren’t
all
his
sisters
here
with
us? Where
did he
get all
this?”
\VS{57}And
so they took offense
at
him.
But
Jesus
said
to them,
“A prophet
is
not
without honor
except
in
his hometown
and
in
his own
house.”
\VS{58}And
he did
not
do
many
miracles
there
because of
their
unbelief.

\par }\Chap{14}{\PP \VerseOne{1}At
that
time
Herod
the tetrarch
heard
reports
about Jesus,
\VS{2}and
he said
to his
servants,
“This
is
John
the Baptist.
He has been raised
from
the dead! And
because of
this,
miraculous powers
are at work
in
him.”
\VS{3}For
Herod
had arrested
John,
bound
him, and
put
him in
prison
on account of
Herodias,
his
brother
Philip’s
wife,
\VS{4}because
John
had repeatedly told
him,
“It is
not
lawful
for you
to have
her.”
\VS{5}Although Herod wanted
to kill
John,
he feared
the crowd
because
they accepted
John
as
a prophet.
\VS{6}But
on Herod’s
birthday,
the daughter
of Herodias
danced
before
them and
pleased
Herod,
\VS{7}so much that
he promised
with
an oath
to give
her
whatever
she asked.
\VS{8}Instructed
by
her
mother,
she said,
“Give
me
the head
of John
the Baptist
here
on
a platter.”
\VS{9}Although it grieved
the king,
because of
his oath
and
the dinner guests
he commanded
it to be given.
\VS{10}So
he sent
and had John
beheaded
in
the prison.
\VS{11}His
head
was brought
on
a platter
and
given
to the girl,
and
she brought
it to her
mother.
\VS{12}Then
John’s
disciples
came
and took
the body
and
buried
it
and
went
and told
Jesus.
\par }{\SH The Feeding of the Five Thousand
\par }{\PP \VS{13}Now
when Jesus
heard
this he went away
from there
privately in
a boat
to
an isolated
place.
But
when
the crowd
heard
about it, they followed
him
on foot
from
the towns.
\VS{14}As he got out
he saw
the large
crowd,
and
he had compassion
on
them
and
healed
their
sick.
\VS{15}When
evening
arrived,
his disciples
came
to him
saying,
“This is
an isolated
place
and
the hour
is
already
late.
Send
the crowds
away
so that
they can go
into
the villages
and buy
food
for themselves.”
\VS{16}But
he
replied,
“They
don’t
need
to go.
You
give
them
something to eat.”
\VS{17}They
said
to him,
“We have
here
only
five
loaves
and
two
fish.”
\VS{18}“Bring
them
here
to me,”
he replied.
\VS{19}Then
he instructed
the crowds
to sit down
on
the grass.
He took
the five
loaves
and
two
fish,
and looking up
to
heaven
he gave thanks
and
broke
the loaves.
He gave
them to the disciples,
who
in turn gave them to the
crowds.
\VS{20}They
all
ate
and
were satisfied,
and
they picked up
the broken pieces
left over,
twelve
baskets
full.
\VS{21}Not counting
women
and
children,
there were about
five thousand
men
who ate.
\par }{\SH Walking on Water
\par }{\PP \VS{22}Immediately
Jesus made
the disciples
get
into
the boat
and
go ahead
of him
to
the other side,
while
he dispersed
the crowds.
\VS{23}And
after he sent
the crowds
away,
he went up
the mountain
by
himself
to pray.
When
evening
came,
he was
there
alone.
\VS{24}Meanwhile
the boat,
already
far
from
land,
was taking
a beating
from
the waves
because
the wind
was
against it.
\VS{25}As the night
was ending,
Jesus came
to
them
walking
on
the sea.
\VS{26}When
the disciples
saw
him
walking
on
the water
they were terrified
and said,
“It’s
a ghost!” and
cried out
with
fear.
\VS{27}But
immediately
Jesus
spoke
to them: “Have courage! It is I.
Do
not
be afraid.”
\VS{28}Peter
said
to him, “Lord,
if
it is
you,
order
me
to come
to
you
on
the water.”
\VS{29}So
he said,
“Come.”
Peter
got
out
of the boat,
walked
on
the water,
and
came
toward
Jesus.
\VS{30}But
when
he saw
the strong wind
he became afraid.
And
starting
to sink,
he cried out, “Lord,
save
me!”
\VS{31}Immediately
Jesus
reached out
his hand
and caught
him,
saying
to him,
“You of little faith,
why
did you doubt?”
\VS{32}When
they
went up
into
the boat,
the wind
ceased.
\VS{33}Then those who were in
the boat
worshiped
him,
saying,
“Truly
you are
the Son
of God.”
\par }{\PP \VS{34}After
they had crossed over,
they came
to
land
at
Gennesaret.
\VS{35}When
the people
there
recognized
him,
they sent
word into
all
the surrounding area,
and
they brought
all
their
sick
to him.
\VS{36}They begged
him
if
they could
only
touch
the edge
of his
cloak,
and
all
who touched
it were healed.

\par }\Chap{15}{\PP \VerseOne{1}Then
Pharisees
and
experts in the law
came
from
Jerusalem
to Jesus
and said,
\VS{2}“Why
do
your
disciples
disobey
the tradition
of the elders? For
they don’t
wash
their hands
when
they eat.”
\VS{3}He answered
them, “And why
do
you
disobey
the commandment
of God
because
of your
tradition?
\VS{4}For
God
said, ‘{\QT{
Honor
your father
and
mother}}’
and ‘{\QT{
Whoever insults
his father
or
mother
must be put
to death}}.’
\VS{5}But
you say,
‘If someone
tells
his father
or
mother,
“Whatever help you would have received
from
me
is given to God,”
\VS{6}he does not need to honor his father.’ You have nullified
the word
of God
on account of
your
tradition.
\VS{7}Hypocrites! Isaiah
prophesied
correctly
about
you
when he said,
\par }{\Q \VS{8}‘{\QT{
This
people
honors
me with their lips,}}
\par }{\Q {\QT{
but
their
heart}}
{\QT{is}}
far
from
me,
\par }{\Q \VS{9}{\QT{
and they worship
me
in vain,}}
\par }{\Q {\QT{
teaching
as doctrines
the commandments
of men}}.’ ”
\par }{\SH True Defilement
\par }{\PP \VS{10}Then
he called
the crowd
to him and said, “Listen
and
understand.
\VS{11}What defiles
a person
is
not
what goes
into
the mouth;
it is what
comes out
of
the mouth
that
defiles
a person.”
\VS{12}Then
the disciples
came
to him
and said,
“Do you know
that
when the Pharisees
heard
this saying
they were offended?”
\VS{13}And
he replied, “Every
plant
that
my
heavenly
Father
did
not
plant
will be uprooted.
\VS{14}Leave
them! They are
blind
guides.
If
someone who is blind
leads
another who is blind,
both
will fall
into
a pit.”
\VS{15}But
Peter
said
to him,
“Explain
this parable
to us.”
\VS{16}Jesus said,
“Even
after all this, are
you
still so foolish?
\VS{17}Don’t
you understand
that
whatever
goes
into
the mouth
enters
the stomach
and
then passes out
into
the sewer?
\VS{18}But
the things that come out
of
the mouth
come
from
the heart,
and these things
defile
a person.
\VS{19}For
out
of the heart
come
evil
ideas,
murder,
adultery,
sexual immorality,
theft,
false testimony,
slander.
\VS{20}These
are
the things that defile
a person;
it is
not
eating
with unwashed
hands
that defiles
a person.”
\par }{\SH A Canaanite Woman’s Faith
\par }{\PP \VS{21}After going out
from there,
Jesus
went
to
the region
of Tyre
\VS{22}A Canaanite
woman
from
that
area
came
and cried out, “Have mercy
on me,
Lord,
Son
of David! My
daughter
is
horribly
demon-possessed!”
\VS{23}But
he did
not
answer
her
a word.
Then
his
disciples
came
and begged
him, “Send
her
away,
because
she keeps on crying out
after
us.”
\VS{24}So
he answered,
“I was sent
only
to
the lost
sheep
of
the house
of Israel.”
\VS{25}But
she came
and bowed down
before him
and said, “Lord,
help
me!”
\VS{26}“It is
not
right
to take
the children’s
bread
and
throw
it to the dogs,”
he said.
\VS{27}“Yes,
Lord,”
she replied, “but
even
the dogs
eat
the crumbs
that fall
from
their
masters’
table.”
\VS{28}Then
Jesus
answered
her,
“Woman,
your
faith
is great! Let what you want
be done
for you.”
And
her
daughter
was healed
from
that
hour.
\par }{\SH Healing Many Others
\par }{\PP \VS{29}When
he left
there,
Jesus
went
along
the Sea
of Galilee.
Then
he went up
a mountain,
where
he sat down.
\VS{30}Then
large
crowds
came
to him
bringing
with
them
the lame,
blind,
crippled,
mute,
and
many
others.
They laid
them
at
his
feet,
and
he healed
them.
\VS{31}As a result,
the crowd
was amazed
when
they saw
the mute
speaking,
the crippled healthy, the lame
walking,
and
the blind
seeing,
and
they praised
the God
of Israel.
\par }{\SH The Feeding of the Four Thousand
\par }{\PP \VS{32}Then
Jesus
called
the disciples
and said,
“I have compassion
on
the crowd,
because
they have
already
been here
with me
three
days
and
they have
nothing
to eat.
I
don’t
want
to send
them
away
hungry
since
they may faint
on
the way.”
\VS{33}The disciples
said
to him,
“Where
can we
get enough bread
in
this desolate place
to satisfy
so great
a crowd?”
\VS{34}Jesus
said
to them,
“How many
loaves
do you have?” They
replied,
“Seven
– and
a few
small fish.”
\VS{35}After instructing
the crowd
to sit down
on
the ground,
\VS{36}he took
the seven
loaves
and
the fish,
and
after giving thanks,
he broke
them and began giving
them to the disciples,
who
then gave them to the crowds.
\VS{37}They
all
ate
and
were satisfied,
and
they picked up
the broken pieces
left over,
seven
baskets
full.
\VS{38}Not counting
children
and
women,
there were
four thousand
men
who ate.
\VS{39}After sending away
the crowd,
he got
into
the boat
and
went
to
the region
of Magadan.

\par }\Chap{16}{\PP \VerseOne{1}Now when the Pharisees
and
Sadducees
came
to test
Jesus, they asked
him
to show
them
a sign
from
heaven.
\VS{2}He said,
“When evening
comes
you say,
‘It will be fair weather,
because
the sky
is red,’
\VS{3}and
in the morning,
‘It will be stormy
today,
because
the sky
is red
and darkening.’
You know
how to judge correctly
the appearance
of the sky,
but
you
cannot
evaluate
the signs
of the times.
\VS{4}A wicked
and
adulterous
generation
asks for
a sign,
but
no
sign
will be given
to it
except
the sign
of Jonah.”
Then
he left
them
and went away.
\par }{\SH The Yeast of the Pharisees and Sadducees
\par }{\PP \VS{5}When the disciples
went
to
the other side,
they forgot
to take
bread.
\VS{6}“Watch out,”
Jesus
said
to them,
“beware
of
the yeast
of the Pharisees
and
Sadducees.”
\VS{7}So
they began to discuss
this among
themselves,
saying,
“It is because
we brought
no
bread.”
\VS{8}When
Jesus
learned
of this, he said,
“You who have such little faith! Why
are you arguing
among
yourselves
about having
no
bread?
\VS{9}Do you still
not
understand? Don’t
you remember
the five
loaves
for the five thousand,
and
how many
baskets
you took up?
\VS{10}Or
the seven
loaves
for the four thousand
and
how many
baskets
you took up?
\VS{11}How
could you
not
understand
that
I was
not
speaking
to you
about
bread? But
beware
of
the yeast
of the Pharisees
and
Sadducees!”
\VS{12}Then
they understood
that
he had
not
told
them to be on guard
against
the yeast
in bread,
but
against
the teaching
of the Pharisees
and
Sadducees.
\par }{\SH Peter’s Confession
\par }{\PP \VS{13}When
Jesus
came
to
the area
of Caesarea
Philippi,
he asked
his
disciples, “Who
do
people
say
that the Son
of Man
is?”
\VS{14}They
answered,
“Some say John
the Baptist,
others
Elijah,
and
others
Jeremiah
or
one
of the prophets.”
\VS{15}He said
to them,
“But
who
do you say
that I
am?”
\VS{16}Simon
Peter
answered, “You
are
the Christ,
the Son
of the living
God.”
\VS{17}And
Jesus
answered
him, “You are
blessed,
Simon
son of Jonah,
because
flesh
and
blood
did
not
reveal
this to you,
but
my
Father
in
heaven!
\VS{18}And I
tell
you
that
you
are
Peter,
and
on
this
rock
I will build
my
church,
and
the gates
of Hades
will
not
overpower
it.
\VS{19}I will give
you
the keys
of the kingdom
of heaven.
Whatever
you bind
on
earth
will have been
bound
in
heaven,
and
whatever
you release
on
earth
will have been
released
in
heaven.”
\VS{20}Then
he instructed
his disciples
not
to tell
anyone
that
he
was
the Christ.
\par }{\SH First Prediction of Jesus’ Death and Resurrection
\par }{\PP \VS{21}From
that time
on Jesus
began
to show
his
disciples
that
he must
go
to
Jerusalem
and
suffer
many things
at the hands of
the elders,
chief priests,
and
experts in the law,
and
be killed,
and
on the third
day
be raised.
\VS{22}So
Peter
took
him
aside
and began
to rebuke
him: “God forbid,
Lord! This
must
not
happen
to you!”
\VS{23}But
he turned
and said
to Peter,
“Get
behind
me,
Satan! You are
a stumbling block
to me,
because
you are
not
setting
your mind
on God’s interests,
but
on man’s.”
\VS{24}Then
Jesus
said
to his
disciples,
“If
anyone
wants
to become
my
follower,
he must deny
himself,
take up
his
cross,
and
follow
me.
\VS{25}For
whoever
wants
to save
his
life
will lose
it,
but
whoever
loses
his
life
for
my
sake
will find
it.
\VS{26}For
what
does it benefit
a person
if
he gains
the whole
world
but
forfeits
his
life? Or
what
can a person
give
in exchange
for his
life?
\VS{27}For
the Son
of Man
will
come
with
his
angels
in
the glory
of his
Father,
and
then
{\IT{
he will reward
each person
according to
what he
has done}}.
\VS{28}I tell
you
the truth,
there are
some
standing
here
who
will
not
experience
death
before
they see
the Son
of Man
coming
in
his
kingdom.”

\par }\Chap{17}{\PP \VerseOne{1}Six
days
later
Jesus
took
with him Peter,
James,
and
John
the brother
of James,
and
led
them
privately
up
a high
mountain.
\VS{2}And
he was transfigured
before
them.
His
face
shone
like
the sun,
and
his
clothes
became
white
as
light.
\VS{3}Then
Moses
and
Elijah
also appeared
before them, talking
with
him.
\VS{4}So
Peter
said
to Jesus,
“Lord,
it is
good
for us
to be
here.
If
you want,
I will make
three
shelters –
one
for you,
one
for Moses,
and
one
for Elijah.”
\VS{5}While
he
was
still
speaking,
a bright
cloud
overshadowed
them,
and
a voice
from
the cloud
said, “This
is
my
one dear
Son,
in
whom
I take great delight.
Listen
to him!”
\VS{6}When
the disciples
heard
this, they were overwhelmed
with fear
and
threw
themselves down with
their
faces to the ground.
\VS{7}But
Jesus
came
and
touched
them.
“Get up,”
he said.
“Do
not
be afraid.”
\VS{8}When they
looked up,
all they saw
was Jesus
alone.
\par }{\PP \VS{9}As
they
were coming down
from
the mountain,
Jesus
commanded
them, “Do
not
tell
anyone
about the vision
until
the Son
of Man
is raised
from
the dead.”
\VS{10}The disciples
asked
him, “Why
then
do
the experts in the law
say
that
Elijah
must
come
first?”
\VS{11}He answered,
“Elijah
does
indeed
come
first and
will restore
all things.
\VS{12}And I
tell
you
that
Elijah
has
already
come.
Yet
they did
not
recognize
him,
but
did
to
him
whatever
they wanted.
In the same way,
the Son
of Man
will
suffer
at
their hands.”
\VS{13}Then
the disciples
understood
that
he was speaking
to them
about
John
the Baptist.
\par }{\SH The Disciples’ Failure to Heal
\par }{\PP \VS{14}When
they came
to
the crowd,
a man
came
to him,
knelt
before him,
\VS{15}and
said,
“Lord,
have mercy
on my
son,
because
he has seizures
and
suffers terribly,
for
he
often
falls
into
the fire
and
into
the water.
\VS{16}I brought
him
to your
disciples,
but
they were
not
able
to heal
him.”
\VS{17}Jesus
answered, “You unbelieving
and
perverse
generation! How much
longer must I be
with
you? How much longer
must I endure
you? Bring
him
here
to me.”
\VS{18}Then
Jesus
rebuked
the demon
and
it came out
of
him,
and
the boy
was healed
from
that
moment.
\VS{19}Then
the disciples
came
to Jesus
privately
and said,
“Why
couldn’t
we
cast
it
out?”
\VS{20}He told
them,
“It was because of
your
little faith.
I tell
you
the truth,
if
you have
faith
the size of a mustard seed,
you will say
to this
mountain,
‘Move
from here
to there,’
and
it will move;
nothing
will be impossible
for you.”
\VS{21}[[EMPTY]]
\par }{\SH \par }{\PP \VS{22}When
they gathered together
in
Galilee,
Jesus
told
them,
“The Son
of Man
is going
to be betrayed
into
the hands
of men.
\VS{23}They will kill
him,
and
on the third
day
he will be raised.”
And
they became
greatly
distressed.
\par }{\SH The Temple Tax
\par }{\PP \VS{24}After
they
arrived
in
Capernaum,
the collectors
of the temple tax
came
to Peter
and
said,
“Your
teacher
pays
the double drachma tax,
doesn’t he?”
\VS{25}He said,
“Yes.”
When
Peter came
into
the house,
Jesus
spoke
to him
first, “What
do
you
think,
Simon? From
whom do
earthly
kings
collect
tolls
or
taxes
– from
their
sons
or
from
foreigners?”
\VS{26}After
he said,
“From
foreigners,”
Jesus
said
to him,
“Then
the sons
are
free.
\VS{27}But
so that
we
don’t
offend
them,
go
to
the lake
and throw out
a hook.
Take
the first
fish
that comes up,
and
when
you open
its
mouth,
you will find
a four drachma coin.
Take
that
and give
it to them
for
me
and
you.”

\par }\Chap{18}{\PP \VerseOne{1}At
that
time
the disciples
came
to Jesus
saying,
“Who
is
the greatest
in
the kingdom
of heaven?”
\VS{2}He
called
a child,
had
him
stand
among
them,
\VS{3}and
said,
“I tell
you
the truth,
unless
you turn around
and
become
like
little children,
you will
never
enter
the kingdom
of heaven!
\VS{4}Whoever
then
humbles
himself
like
this little child
is
the greatest
in
the kingdom
of heaven.
\VS{5}And
whoever
welcomes
a child
like this
in
my
name
welcomes
me.
\par }{\PP \VS{6}“But
if anyone causes
one
of these
little
ones who believe
in
me
to sin, it would be better
for him
to have a huge millstone
hung
around
his
neck
and
to be drowned
in
the open
sea.
\VS{7}Woe
to the world
because of
stumbling blocks! It is necessary
that stumbling blocks
come,
but
woe
to the person
through
whom
they come.
\VS{8}If
your
hand
or
your
foot
causes
you
to sin,
cut
it
off
and
throw
it away. It is
better
for you
to enter
life
crippled
or
lame
than
to have
two
hands
or
two
feet
and be thrown
into
eternal
fire.
\VS{9}And
if
your
eye
causes
you
to sin,
tear
it
out
and
throw
it away.
It is
better
for you
to enter
into
life
with one eye
than
to have
two
eyes
and be thrown
into
fiery
hell.
\par }{\SH The Parable of the Lost Sheep
\par }{\PP \VS{10}“See
that you do
not
disdain
one
of these
little ones.
For
I tell
you
that
their
angels
in
heaven
always
see
the face
of my
Father
in
heaven.
\VS{11}[[EMPTY]]
\VS{12}What
do
you
think? If
someone
owns
a hundred
sheep
and
one
of
them
goes astray,
will he
not
leave
the ninety-nine
on
the mountains
and
go
look for
the one that went astray?
\VS{13}And
if
he finds
it,
I tell
you
the truth,
he will rejoice
more
over
it
than
over
the ninety-nine
that did
not
go astray.
\VS{14}In the same way,
your Father
in
heaven
is
not
willing
that
one of these
little
ones
be lost.
\par }{\SH Restoring Christian Relationships
\par }{\PP \VS{15}“If
your
brother
sins,
go
and show
him
his fault
when the two of you
are alone.
If
he listens
to you,
you have regained
your
brother.
\VS{16}But
if
he does
not
listen,
take
one
or
two
others with
you,
so
that
{\QT{
at
the testimony
of two
or
three
witnesses
every
matter
may be established}}.
\VS{17}If
he refuses to listen
to them,
tell
it to the church.
If
he refuses to listen
to the church,
treat him like
a Gentile
or
a tax collector.
\par }{\PP \VS{18}“I tell
you
the truth,
whatever
you bind
on
earth
will have been
bound
in
heaven,
and
whatever
you release
on
earth
will have been
released
in
heaven.
\VS{19}Again,
I tell
you
the truth,
if
two
of
you
on
earth
agree
about
whatever
you ask,
my
Father
in
heaven will do it for you.
\VS{20}For
where
two
or
three
are
assembled
in
my
name,
I am
there
among
them.”
\par }{\PP \VS{21}Then
Peter
came
to him
and said,
“Lord,
how many times
must I forgive
my
brother
who sins
against
me? As many as
seven times?”
\VS{22}Jesus
said
to him,
“Not
seven times,
I tell
you,
but
seventy-seven times!
\par }{\SH The Parable of the Unforgiving Slave
\par }{\PP \VS{23}“For
this
reason, the kingdom
of heaven
is like
a king
who
wanted
to settle
accounts
with
his
slaves.
\VS{24}As
he began settling
his
accounts,
a man
who owed ten thousand
talents
was brought
to
him.
\VS{25}Because he was
not
able
to repay
it, the lord
ordered
him
to be sold, along
with his wife,
children,
and
whatever
he possessed,
and
repayment to be made.
\VS{26}Then
the slave
threw himself
to the ground
before him,
saying, ‘Be patient
with
me,
and
I will repay
you
everything.’
\VS{27}The lord
had compassion
on that
slave
and released
him,
and
forgave
him
the debt.
\VS{28}After
he went out,
that
same slave
found
one
of his
fellow slaves
who
owed
him
one hundred
silver coins.
So
he grabbed
him
by the throat and started to choke
him, saying,
‘Pay back
what
you owe me!’
\VS{29}Then
his
fellow slave
threw himself down
and begged
him, ‘Be patient
with
me,
and
I will repay
you.’
\VS{30}But
he refused.
Instead,
he went out
and threw
him
in
prison
until
he repaid
the debt.
\VS{31}When
his
fellow slaves
saw
what had happened,
they were
very
upset
and
went
and told
their
lord
everything
that had taken place.
\VS{32}Then
his
lord
called
the first slave
and said
to him,
‘Evil
slave! I forgave
you all
that
debt
because
you
begged
me!
\VS{33}Should you
not
have shown mercy
to your
fellow slave,
just as
I
showed
it to you?’
\VS{34}And
in anger
his
lord
turned
him
over
to the prison guards to torture
him until
he repaid
all
he owed.
\VS{35}So
also
my
heavenly
Father
will do
to you,
if
each
of you does
not
forgive
your brother
from
your
heart.”

\par }\Chap{19}{\PP \VerseOne{1}Now when
Jesus
finished
these
sayings,
he left
Galilee
and
went
to
the region
of Judea
beyond
the Jordan River.
\VS{2}Large
crowds
followed
him,
and
he healed
them
there.
\par }{\PP \VS{3}Then
some Pharisees
came
to him
in order to test
him.
They asked,
“Is
it lawful
to divorce
a wife
for
any
cause?”
\VS{4}He answered,
“Have you
not
read
that
from
the beginning
the Creator
{\BD{
{\QT{
{\IT{made}}}}
them
male
and
female
them
male
and
female}},
\VS{5}and
said, ‘{\QT{
For
this
reason
a man
will leave
his father
and
mother
and
will be united
with his
wife,
and
the two
will become
one
flesh}}’?
\VS{6}So
they are
no longer
two,
but
one
flesh.
Therefore
what God
has joined together,
let
no
one
separate.”
\VS{7}They said
to him,
“Why
then
did Moses
command
us
{\IT{
to give
{\BD{
a certificate
of dismissal}}
and
to divorce}} her?”
\VS{8}Jesus said
to them,
“Moses
permitted
you
to divorce
your
wives
because of
your
hard hearts,
but
from
the beginning
it was
not
this way.
\VS{9}Now I say
to you
that
whoever
divorces
his
wife,
except
for
immorality,
and
marries
another
commits adultery.”
\VS{10}The disciples
said
to him,
“If
this
is
the case
of a husband
with
a wife,
it is better
not
to marry!”
\VS{11}He said
to them,
“Not
everyone
can accept
this statement,
except
those to whom
it has been given.
\VS{12}For
there are
some eunuchs
who
were that way
from
birth,
and
some who
were made eunuchs
by
others,
and
some who
became eunuchs
for the sake
of the kingdom
of heaven.
The one who is able
to accept
this should accept it.”
\par }{\SH Jesus and Little Children
\par }{\PP \VS{13}Then
little children
were brought
to him
for
him to lay
his hands
on them
and
pray.
But
the disciples
scolded
those who brought them.
\VS{14}But
Jesus
said,
“Let
the little children
come
to
me
and
do
not
try to stop
them,
for
the kingdom
of heaven
belongs to such
as these.”
\VS{15}And
he placed
his hands
on them
and went on
his way.
\par }{\SH The Rich Young Man
\par }{\PP \VS{16}Now
someone
came up
to him
and said,
“Teacher,
what
good thing
must I do
to gain
eternal
life?”
\VS{17}He said
to him,
“Why
do you ask
me
about
what is good? There is
only one
who is good.
But
if
you want
to enter
into
life,
keep
the commandments.”
\VS{18}“Which ones?” he asked.
Jesus
replied, “{\QT{
Do
not
murder,
do
not
commit adultery,
do
not
steal,
do
not
give false testimony,}}
\VS{19}{\QT{
honor
your father
and
mother,}}
and
{\BD{
{\QT{
{\IT{love}}}}
your
neighbor
as
yourself
your
neighbor
as
yourself}}.”
\VS{20}The young man
said
to him,
“I have wholeheartedly obeyed
all
these
laws. What
do I
still
lack?”
\VS{21}Jesus
said
to him,
“If
you wish
to be
perfect,
go
sell
your
possessions
and
give
the money to the poor,
and
you will have
treasure
in
heaven.
Then
come,
follow
me.”
\VS{22}But
when
the young man
heard
this
he went away
sorrowful,
for
he was
very
rich.
\par }{\PP \VS{23}Then Jesus
said
to his
disciples,
“I tell
you
the truth,
it will be hard
for a rich person
to enter
the kingdom
of heaven!
\VS{24}Again
I say,
it is
easier
for a camel
to go through
the eye
of a needle
than
for a rich person
to enter
into
the kingdom
of God.”
\VS{25}The disciples
were
greatly
astonished
when they heard
this and said,
“Then
who
can
be saved?”
\VS{26}Jesus
looked at
them
and replied,
“This
is
impossible
for
mere humans,
but
for
God
all things
are possible.”
\VS{27}Then
Peter
said
to him,
“Look,
we
have left
everything
to follow
you! What
then
will there be
for us?”
\VS{28}Jesus
said
to them,
“I tell
you
the truth: In
the age
when
all things are renewed, when the Son
of Man
sits
on
his
glorious
throne,
you
who have followed
me
will
also
sit
on
twelve
thrones,
judging
the twelve
tribes
of Israel.
\VS{29}And
whoever
has left
houses
or
brothers
or
sisters
or
father
or
mother
or
children
or
fields
for
my
sake
will receive
a hundred times
as much and
will inherit
eternal
life.
\VS{30}But
many
who are first
will be
last,
and
the last
first.

\par }\Chap{20}{\PP \VerseOne{1}“For
the kingdom
of heaven
is
like
a landowner
who
went out
early
in the morning
to hire
workers
for
his
vineyard.
\VS{2}And
after agreeing
with
the workers
for
the standard wage,
he sent
them
into
his
vineyard.
\VS{3}When
it was about
nine
o’clock
in the morning,
he went out
again and saw
others
standing around
in
the marketplace
without work.
\VS{4}He said
to them, ‘You
go
into
the vineyard
too, and
I will give
you
whatever
is
right.’
\VS{5}So
they went.
When
he went out
again
about
noon
and
three
o’clock
that afternoon,
he did
the same thing.
\VS{6}And
about
five o’clock
that afternoon
he went out
and found
others
standing around,
and
said
to them,
‘Why
are you standing
here
all
day
without work?’
\VS{7}They said
to him,
‘Because
no one
hired
us.’
He said
to them,
‘You go
and
work in
the vineyard too.’
\VS{8}When it was
evening
the owner
of the vineyard
said
to his
manager,
‘Call
the workers
and
give
the pay
starting
with
the last
hired until
the first.’
\VS{9}When those hired about
five
o’clock
came,
each
received
a full day’s pay.
\VS{10}And
when those hired first
came,
they thought
they would receive
more.
But each one
also
received
the standard wage.
\VS{11}When
they received
it, they began to complain
against
the landowner,
\VS{12}saying,
‘These
last
fellows worked one
hour,
and you have made
them
equal
to us
who bore
the hardship
and
burning heat
of the day.’
\VS{13}And
the landowner replied
to one
of them, ‘Friend,
I am
not
treating
you
unfairly.
Didn’t
you agree with
me
to work for the standard wage?
\VS{14}Take
what is yours
and
go.
I want
to give to this
last man
the same as
I gave
to you.
\VS{15}Am I
not
permitted
to do
what
I want
with
what belongs
to me? Or
are
you
envious
because
I
am
generous?’
\VS{16}So
the last
will be
first,
and
the first
last.”
\par }{\SH Third Prediction of Jesus’ Death and Resurrection
\par }{\PP \VS{17}As Jesus
was going
up
to
Jerusalem,
he took
the twelve
aside
privately
and
said
to them
on
the way,
\VS{18}“Look,
we are going up
to
Jerusalem,
and
the Son
of Man
will be handed over
to the chief priests
and
the experts in the law.
They will condemn
him
to death,
\VS{19}and
will turn
him
over
to the Gentiles
to
be mocked
and
flogged severely
and
crucified.
Yet
on the third
day,
he will be raised.”
\par }{\SH A Request for James and John
\par }{\PP \VS{20}Then
the mother
of the sons
of Zebedee
came
to him
with
her
sons,
and kneeling down
she asked
him
for a favor.
\VS{21}He said
to her,
“What
do you want?” She replied, “Permit these
two
sons
of mine
to sit,
one
at
your right hand
and
one
at
your
left,
in
your
kingdom.”
\VS{22}Jesus
answered,
“You
don’t
know
what
you are asking! Are you able
to drink
the cup
I
am about
to drink?” They said
to him,
“We are able.”
\VS{23}He told
them,
“You will drink
my
cup,
but
to sit
at
my
right
and at
my left
is
not
mine
to give.
Rather,
it is for those for whom
it has been prepared
by
my
Father.”
\par }{\PP \VS{24}Now when the other ten
heard
this, they were angry
with
the two
brothers.
\VS{25}But
Jesus
called
them
and said,
“You know
that
the rulers
of the Gentiles
lord it over
them,
and
those in high positions
use their authority
over them.
\VS{26}It must not
be
this way
among
you! Instead
whoever
wants
to be
great
among
you
must be
your
servant,
\VS{27}and
whoever
wants
to be
first
among
you
must be
your
slave –
\VS{28}just as
the Son
of Man
did
not
come
to be served
but
to serve,
and
to give
his
life
as a ransom
for
many.”
\par }{\SH Two Blind Men Healed
\par }{\PP \VS{29}As
they
were leaving
Jericho,
a
large
crowd
followed
them.
\VS{30}Two
blind men
were sitting
by
the road.
When
they heard
that
Jesus
was passing by,
they shouted, “Have mercy
on us,
Lord,
Son
of David!”
\VS{31}The crowd
scolded
them
to get
them to be quiet.
But
they shouted
even more loudly,
“Lord,
have mercy
on us,
Son
of David!”
\VS{32}Jesus
stopped,
called
them,
and
said,
“What
do you want
me to do
for you?”
\VS{33}They said
to him,
“Lord,
let
our
eyes
be opened.”
\VS{34}Moved with compassion,
Jesus
touched
their
eyes.
Immediately
they received
their sight
and
followed
him.

\par }\Chap{21}{\PP \VerseOne{1}Now
when
they approached
Jerusalem
and
came
to
Bethphage,
at
the Mount
of Olives,
Jesus
sent
two
disciples,
\VS{2}telling
them,
“Go
to
the village
ahead
of you.
Right away
you will find
a donkey
tied
there, and
a colt
with
her.
Untie
them and bring
them to me.
\VS{3}If
anyone
says
anything to you,
you are to say,
‘The Lord
needs
them,’
and
he will send
them
at once.”
\VS{4}This
took place
to
fulfill
what was spoken
by
the prophet:
\par }{\Q \VS{5}“{\BD{
{\IT{
Tell
the people
of Zion,}}}}
\par }{\Q ‘{\BD{
{\IT{
Look,
your
king
is coming
to you,}}}}
\par }{\Q {\BD{
{\IT{
unassuming
and
seated
on
a donkey,}}}}
\par }{\Q {\BD{
{\IT{
and
on
a colt,
the foal of
a donkey.}}}}’ ”
\par }{\PP \VS{6}So
the disciples
went
and
did
as
Jesus
had instructed
them.
\VS{7}They brought
the donkey
and
the colt
and
placed
their cloaks
on
them,
and
he sat
on
them.
\VS{8}A very large
crowd
spread
their
cloaks
on
the road.
Others
cut
branches
from
the trees
and
spread
them on
the road.
\VS{9}The crowds
that went ahead
of him
and
those following
kept shouting, “{\IT{Hosanna}}
to the Son
of David!
{\BD{
{\IT{
Blessed
is the one who comes
in
the name
of the Lord!}}}}
{\IT{Hosanna}}
in
the highest!”
\VS{10}As he
entered
Jerusalem
the whole
city
was thrown into an uproar,
saying,
“Who
is
this?”
\VS{11}And
the crowds
were saying,
“This
is
the prophet
Jesus,
from
Nazareth
in Galilee.”
\par }{\SH Cleansing the Temple
\par }{\PP \VS{12}Then
Jesus
entered
the temple area
and
drove out
all
those who were selling
and
buying
in
the temple courts,
and
turned over
the tables
of the money changers
and
the chairs
of those selling
doves.
\VS{13}And
he said
to them,
“It is written, ‘{\BD{
{\IT{
My
house
will be called
a house
of prayer,}}}}’
but
you
are turning
it
into
{\BD{
{\IT{
a den
of robbers}}}}!”
\par }{\PP \VS{14}The blind
and
lame
came
to him
in
the temple courts,
and
he healed
them.
\VS{15}But
when the chief priests
and
the experts in the law
saw
the wonderful
things
he did
and
heard the children
crying out
in
the temple courts, “Hosanna
to the Son
of David,”
they became indignant
\VS{16}and
said
to him,
“Do you hear
what
they
are saying?” Jesus
said
to them,
“Yes.
Have you
never
read, ‘{\BD{
{\IT{
Out
of
the mouths
of children
and
nursing infants
you have prepared
praise for yourself}}}}’?”
\VS{17}And
leaving
them,
he went out
of the city
to
Bethany
and
spent the night
there.
\par }{\SH The Withered Fig Tree
\par }{\PP \VS{18}Now
early in the morning,
as
he returned
to
the city,
he was hungry.
\VS{19}After noticing
a
fig tree
by
the road
he went
to
it,
but
found
nothing
on
it
except
leaves.
He
said
to it,
“Never again
will there be
fruit
from
you!” And
the fig tree
withered
at once.
\VS{20}When
the disciples
saw
it they were amazed,
saying,
“How
did
the fig tree
wither
so quickly?”
\VS{21}Jesus
answered
them,
“I tell
you
the truth,
if
you have
faith
and
do
not
doubt,
not
only
will you do
what was done
to the fig tree,
but
even if
you say
to this
mountain,
‘Be lifted up
and
thrown
into
the sea,’
it will happen.
\VS{22}And
whatever
you ask
in
prayer,
if you believe,
you will receive.”
\par }{\SH The Authority of Jesus
\par }{\PP \VS{23}Now
after
Jesus
entered
the temple courts,
the chief priests
and
elders
of the people
came up
to him
as he was teaching
and said,
“By
what
authority
are you doing
these things,
and
who
gave
you
this
authority?”
\VS{24}Jesus
answered
them,
“I
will
also
ask
you
one
question.
If
you answer
me
then I
will
also
tell
you
by
what
authority
I do
these things.
\VS{25}Where
did
John’s
baptism
come from? From
heaven
or
from
people?” They discussed
this among
themselves,
saying,
“If
we say,
‘From
heaven,’
he will say,
‘Then
why
did you
not
believe
him?’
\VS{26}But
if
we say,
‘From
people,’
we fear
the crowd,
for
they all
consider
John
to be
a prophet.”
\VS{27}So
they answered
Jesus, “We
don’t
know.”
Then he said
to them,
“Neither
will
I
tell
you
by
what
authority
I am doing
these things.
\par }{\SH The Parable of the Two Sons
\par }{\PP \VS{28}“What
do
you
think? A man
had
two
sons.
He went
to the first
and said,
‘Son,
go
and work
in
the vineyard
today.’
\VS{29}The boy answered, ‘I
will not.’
But later he had a change of heart and
went.
\VS{30}The father went
to the other son
and said
the same thing.
This boy answered, ‘I will,
sir,’ but did
not
go.
\VS{31}Which
of
the two
did
his father’s
will?” They said,
“The first.”
Jesus
said
to them,
“I tell
you
the truth,
tax collectors
and
prostitutes
will go ahead
of you
into
the kingdom
of God!
\VS{32}For
John
came
to
you
in
the way
of righteousness,
and
you did
not
believe
him.
But
the tax collectors
and
prostitutes
did believe.
Although you
saw
this, you did
not
later
change
your minds
and believe
him.
\par }{\SH The Parable of the Tenants
\par }{\PP \VS{33}“Listen
to another
parable: There
was
a landowner
who
planted
a vineyard.
He
put
a fence around
it,
dug
a pit for its winepress,
and
built
a watchtower.
Then
he leased
it
to tenant farmers
and
went on a journey.
\VS{34}When
the harvest
time
was near,
he sent
his
slaves
to
the tenants
to collect
his
portion of the crop.
\VS{35}But
the tenants
seized
his
slaves,
beat
one, killed
another, and stoned another.
\VS{36}Again
he sent
other
slaves,
more than
the first,
and
they treated
them
the same way.
\VS{37}Finally
he sent
his
son
to them,
saying,
‘They will respect
my
son.’
\VS{38}But
when
the tenants
saw
the son,
they said
to
themselves,
‘This
is
the heir.
Come,
let’s kill
him
and
get
his
inheritance!’
\VS{39}So
they seized
him,
threw
him out
of the vineyard,
and
killed him.
\VS{40}Now
when
the owner
of the vineyard
comes,
what
will he do
to those
tenants?”
\VS{41}They said
to him,
“He will
utterly
destroy
those evil men! Then he will lease
the vineyard
to other
tenants
who
will give
him
his
portion
at
the harvest.”
\par }{\PP \VS{42}Jesus
said
to them,
“Have you
never
read
in
the scriptures:
\par }{\Q ‘{\BD{
{\IT{
The stone
the builders
rejected
has become
the cornerstone}}}}.
\par }{\Q {\BD{
{\IT{
This
is from
the Lord,
and
it is
marvelous
in
our
eyes}}}}’?
\par }{\PP \VS{43}For
this
reason I tell
you
that
the kingdom
of God
will be taken
from
you
and
given
to a people
who will produce
its
fruit.
\VS{44}The one who falls
on
this
stone
will be broken to pieces,
and the one on
whom
it falls
will be crushed.”
\VS{45}When
the chief priests
and
the Pharisees
heard
his
parables,
they realized
that
he was speaking
about
them.
\VS{46}They wanted
to arrest
him,
but they were afraid
of the crowds,
because
the crowds regarded
him
as
a prophet.

\par }\Chap{22}{\PP \VerseOne{1}Jesus
spoke
to them again
in
parables,
saying:
\VS{2}“The kingdom
of heaven
can be compared
to a king
who
gave
a wedding banquet
for his
son.
\VS{3}He
sent
his
slaves
to summon
those who had been invited
to
the banquet,
but
they would
not
come.
\VS{4}Again
he sent
other
slaves,
saying,
‘Tell
those who have been invited,
“Look! The feast
I
have prepared
for you is ready. My
oxen
and
fattened cattle
have been slaughtered,
and
everything
is
ready.
Come
to
the wedding banquet.” ’
\VS{5}But
they were indifferent
and went away,
one to
his
farm,
another to
his
business.
\VS{6}The rest
seized
his
slaves,
insolently mistreated
them, and
killed them.
\VS{7}The
king
was furious! He
sent
his
soldiers,
and they put
those
murderers
to death
and
set
their
city
on fire.
\VS{8}Then
he said
to his
slaves,
‘The wedding
is
ready,
but
the ones who had been invited
were
not
worthy.
\VS{9}So
go
into
the main streets
and
invite
everyone
you find
to
the wedding banquet.’
\VS{10}And
those
slaves
went out
into
the streets
and gathered
all
they found,
both
bad
and
good,
and
the wedding hall was filled
with guests.
\VS{11}But
when
the king
came in to see
the wedding guests,
he saw
a man
there
who was
not
wearing
wedding
clothes.
\VS{12}And
he said
to him,
‘Friend,
how
did you get in
here
without
wedding
clothes?’ But
he had nothing to say.
\VS{13}Then
the king
said
to his attendants,
‘Tie
him
up
hand
and foot
and
throw
him
into
the outer
darkness,
where there
will be
weeping
and
gnashing
of teeth!’
\VS{14}For
many
are
called,
but
few
are chosen.”
\par }{\SH Paying Taxes to Caesar
\par }{\PP \VS{15}Then
the Pharisees
went out
and planned together
to entrap
him
with
his own words.
\VS{16}They sent
to him
their
disciples
along with
the Herodians,
saying,
“Teacher,
we know
that
you are
truthful,
and
teach
the way
of God
in accordance with
the truth.
You
do not
court
anyone’s
favor because
you show
no
partiality.
\VS{17}Tell
us
then,
what
do
you
think? Is it right
to pay
taxes
to Caesar
or
not?”
\par }{\PP \VS{18}But
Jesus
realized
their
evil intentions
and said,
“Hypocrites! Why
are you testing
me?
\VS{19}Show
me
the coin
used for the tax.”
So
they brought
him
a denarius.
\VS{20}Jesus said
to them,
“Whose
image
is this,
and
whose inscription?”
\VS{21}They replied, “Caesar’s.”
He said
to them, “Then
give
to Caesar
the things that are Caesar’s,
and
to God
the things that are God’s.”
\VS{22}Now
when
they heard
this they were stunned,
and
they left
him
and went away.
\par }{\SH Marriage and the Resurrection
\par }{\PP \VS{23}The same
day
Sadducees
(who say
there is
no
resurrection) came
to him
and
asked
him,
\VS{24}“Teacher,
Moses
said, ‘{\QT{
If
a man
dies
without
having
children,
his
brother
must marry
the widow
and
father
children}}
{\QT{
for his
brother}}.’
\VS{25}Now there were
seven
brothers
among
us.
The first one
married
and
died,
and
since he had
no
children
he left
his
wife
to his
brother.
\VS{26}The second
did the same,
and
the third,
down to
the seventh.
\VS{27}Last
of all,
the woman
died.
\VS{28}In
the resurrection,
therefore,
whose
wife of the seven
will
she
be? For
they
all
had married
her.”
\VS{29}Jesus
answered
them,
“You are deceived,
because you don’t
know
the scriptures
or
the power
of God.
\VS{30}For
in
the resurrection
they
neither
marry
nor
are given in marriage,
but
are
like
angels
in
heaven.
\VS{31}Now as for
the resurrection
of the dead,
have you
not
read
what was spoken
to you
by
God,
\VS{32}‘{\QT{
I
am
the God
of Abraham,
the God
of Isaac,
and
the God
of Jacob}}’? He is
not
the God
of the dead
but
of the living!”
\VS{33}When
the crowds
heard
this, they were amazed
at
his
teaching.
\par }{\SH The Greatest Commandment
\par }{\PP \VS{34}Now
when
the Pharisees
heard
that
he had silenced
the Sadducees,
they assembled together.
\VS{35}And
one
of
them,
an expert in religious law,
asked
him a question to test
him:
\VS{36}“Teacher,
which
commandment
in
the law
is the greatest?”
\VS{37}Jesus said
to him,
{\QT{
“ ‘Love
the Lord
your
God
with
all
your
heart,
with
all
your
soul,
and
with
all
your
mind}}.’
\VS{38}This
is
the first
and
greatest
commandment.
\VS{39}The second
is like
it{\QT{
: ‘Love
your
neighbor
as
yourself}}.’
\VS{40}All
the law
and
the prophets
depend
on
these
two
commandments.”
\par }{\SH The Messiah: David’s Son and Lord
\par }{\PP \VS{41}While
the Pharisees
were assembled,
Jesus
asked
them a question:
\VS{42}“What
do
you
think
about
the Christ? Whose
son
is he?” They said,
“The son of David.”
\VS{43}He said
to them,
“How
then
does David
by
the Spirit
call
him
‘Lord,’
saying,
\par }{\Q \VS{44}‘{\QT{
The Lord
said
to my
lord,}}
\par }{\Q “{\QT{
Sit
at
my
right hand,}}
\par }{\Q {\QT{
until
I put
your
enemies
under
your
feet}}” ’?
\par }{\PI \VS{45}If
David
then
calls
him
‘Lord,’
how can
he be
his
son?”
\VS{46}No
one was able
to answer
him
a word,
and from
that
day
on no one
dared
to question
him
any longer.

\par }\Chap{23}{\PP \VerseOne{1}Then
Jesus
said
to the crowds
and
to his
disciples,
\VS{2}“The experts in the law
and
the Pharisees
sit
on
Moses’
seat.
\VS{3}Therefore
pay
attention
to what they tell
you
and
do
it. But
do
not
do
what they
do,
for
they do
not
practice
what they teach.
\VS{4}They tie up
heavy
loads, hard to carry,
and
put
them on
men’s
shoulders,
but
they
themselves
are
not
willing
even to lift a finger
to move
them.
\VS{5}They do
all
their
deeds
to
be seen
by people,
for
they make
their
phylacteries
wide
and
their tassels
long.
\VS{6}They love
the place of honor
at
banquets
and
the best seats
in
the synagogues
\VS{7}and
elaborate greetings
in
the marketplaces,
and
to have
people
call
them ‘Rabbi.’
\VS{8}But
you
are not
to be called
‘Rabbi,’
for
you
have one
Teacher
and
you
are
all
brothers.
\VS{9}And
call
no
one your
‘father’
on
earth,
for
you
have one
Father,
who is in heaven.
\VS{10}Nor
are you to be called
‘teacher,’
for you
have
one
teacher, the Christ.
\VS{11}The greatest
among you
will be
your
servant.
\VS{12}And
whoever
exalts
himself
will be humbled,
and
whoever
humbles
himself
will be exalted.
\par }{\PP \VS{13}“But
woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You keep locking
people
out of the kingdom
of heaven! For
you
neither
enter
nor
permit
those trying to enter
to go in.
\VS{14}[[EMPTY]]
\par }{\PP \VS{15}“Woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You cross
land
and
sea
to make
one
convert,
and
when
you get one, you make
him
twice as much
a child
of hell
as yourselves!
\par }{\PP \VS{16}“Woe
to you,
blind
guides,
who
say,
‘Whoever
swears
by
the temple
is
bound by nothing.
But
whoever
swears
by
the gold
of the temple
is bound by the oath.’
\VS{17}Blind
fools! Which
is
greater,
the gold
or
the temple
that makes
the gold
sacred?
\VS{18}And,
‘Whoever
swears
by
the altar
is bound by nothing.
But
if anyone
swears
by
the gift
on
it
he is bound by the oath.’
\VS{19}You
are blind! For
which
is greater,
the gift
or
the altar
that makes
the gift
sacred?
\VS{20}So
whoever swears
by
the altar
swears
by
it
and
by
everything
on
it.
\VS{21}And
whoever swears
by
the temple
swears
by
it
and
the one who dwells
in
it.
\VS{22}And
whoever swears
by
heaven
swears
by
the throne
of God
and
the one who sits
on
it.
\par }{\PP \VS{23}“Woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You give a tenth
of mint,
dill,
and cumin,
yet you neglect
what is more important
in the
law
– justice,
mercy,
and
faithfulness! You should
have done
these things
without
neglecting the others.
\VS{24}Blind
guides! You strain out
a gnat
yet swallow
a camel!
\par }{\PP \VS{25}“Woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You clean
the outside
of the cup
and
the dish,
but
inside
they are full
of
greed
and
self-indulgence.
\VS{26}Blind
Pharisee! First
clean
the inside
of the cup,
so that the outside
may
become
clean too!
\par }{\PP \VS{27}“Woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You are like
whitewashed
tombs
that
look
beautiful
on the outside
but
inside
are full
of the bones
of the dead
and
of everything
unclean.
\VS{28}In the same way,
on the outside
you
look
righteous
to people,
but
inside
you are
full
of hypocrisy
and
lawlessness.
\par }{\PP \VS{29}“Woe
to you,
experts in the law
and
you Pharisees,
hypocrites! You build
tombs
for the prophets
and
decorate
the graves
of the righteous.
\VS{30}And
you say,
‘If
we had lived
in
the days
of our
ancestors,
we would
not
have participated
with them
in
shedding
the blood
of the prophets.’
\VS{31}By saying this
you testify
against yourselves
that
you are
descendants
of those who murdered
the prophets.
\VS{32}Fill up
then
the measure
of your
ancestors!
\VS{33}You snakes,
you offspring
of vipers! How
will you escape
being condemned
to hell?
\par }{\PP \VS{34}“For
this
reason I
am sending
you
prophets
and
wise men
and
experts in the law,
some of
whom
you will kill
and
crucify,
and
some
you will flog
in
your
synagogues
and
pursue
from
town
to
town,
\VS{35}so
that on
you
will come
all
the righteous
blood
shed
on
earth,
from
the blood
of righteous
Abel
to the blood
of Zechariah
son
of Barachiah,
whom
you murdered
between
the temple
and
the altar.
\VS{36}I tell
you
the truth,
this
generation
will be held responsible
for all
these things!
\par }{\SH Judgment on Israel
\par }{\PP \VS{37}“O Jerusalem,
Jerusalem,
you who kill
the prophets
and
stone
those who are sent
to
you! How often
I have longed
to gather
your
children
together
as
a hen
gathers
her
chicks
under
her wings,
but
you would
have none of it!
\VS{38}Look,
your
house
is left
to you
desolate!
\VS{39}For
I tell
you,
you will not see me from now until you say, ‘{\QT{Blessed is the one who comes in the name of the Lord!}}’ ”

\par }\Chap{24}{\PP \VerseOne{1}Now
as Jesus
was going out
of
the temple courts
and
walking away,
his
disciples
came
to show
him
the temple
buildings.
\VS{2}And
he said
to them, “Do
you see
all
these things? I tell
you
the truth,
not
one stone
will be left
on
another.
All will be torn down!”
\par }{\SH Signs of the End of the Age
\par }{\PP \VS{3}As
he
was sitting
on
the Mount
of Olives,
his disciples
came
to him
privately
and said,
“Tell
us,
when
will
these things
happen? And
what
will be the sign
of your
coming
and
of the end
of the age?”
\VS{4}Jesus
answered
them, “Watch out
that no
one
misleads
you.
\VS{5}For
many
will come
in
my
name,
saying,
‘I
am
the Christ,’
and
they will mislead
many.
\VS{6}You will
hear
of wars
and
rumors
of wars.
Make sure
that you are
not
alarmed,
for
this must
happen,
but
the end
is
still to come.
\VS{7}For
nation
will rise up
in arms against
nation,
and
kingdom
against
kingdom.
And
there will be
famines
and
earthquakes
in
various places.
\VS{8}All
these things
are the beginning
of birth pains.
\par }{\SH Persecution of Disciples
\par }{\PP \VS{9}“Then
they will hand
you
over
to
be persecuted
and
will kill
you.
You will be
hated
by
all
the nations
because
of my
name.
\VS{10}Then
many
will be led into sin,
and
they will betray
one another
and
hate
one another.
\VS{11}And
many
false prophets
will appear
and
deceive
many,
\VS{12}and
because
lawlessness
will increase so much,
the love
of many
will grow cold.
\VS{13}But
the person
who endures
to
the end
will be saved.
\VS{14}And
this
gospel
of the kingdom
will be preached
throughout
the whole
inhabited earth
as
a testimony
to all
the nations,
and
then
the end
will come.
\par }{\SH The Abomination of Desolation
\par }{\PP \VS{15}“So
when
you see
{\IT{
the abomination
of desolation}} –
spoken
about by
Daniel
the prophet
– standing
in
the holy
place
(let the reader
understand),
\VS{16}then
those in
Judea
must flee
to
the mountains.
\VS{17}The one on
the roof
must
not
come down
to take
anything out of
his
house,
\VS{18}and
the one in
the field
must
not
turn
back
to get
his
cloak.
\VS{19}Woe
to those who are pregnant
and
to those who are nursing
their babies
in
those
days!
\VS{20}Pray
that your
flight
may
not
be
in winter
or
on a Sabbath.
\VS{21}For
then
there will be
great
suffering
unlike
anything that has happened
from
the beginning
of the world
until
now,
or
ever will happen.
\VS{22}And
if
those
days
had
not
been cut short,
no one
would
be saved.
But for the sake
of
the elect
those
days
will be cut short.
\VS{23}Then
if
anyone
says
to you,
‘Look,
here
is the Christ!’ or
‘There
he is!’ do
not
believe him.
\VS{24}For
false messiahs
and
false prophets
will appear
and
perform
great
signs
and
wonders
to
deceive,
if
possible,
even
the elect.
\VS{25}Remember,
I have told
you
ahead of time.
\VS{26}So
then, if
someone says to you,
‘Look,
he is
in
the wilderness,’
do
not
go out,
or ‘Look,
he is in
the inner rooms,’
do
not
believe him.
\VS{27}For
just
like the lightning
comes
from
the east
and
flashes
to
the west,
so
the coming
of the Son
of Man
will be.
\VS{28}Wherever
the corpse
is,
there
the vultures
will gather.
\par }{\SH The Arrival of the Son of Man
\par }{\PP \VS{29}“Immediately
after
the suffering
of those
days,
{\IT{
the
sun
will be darkened,
and
the
moon
will
not
give
its
light;
the stars
will fall
from
heaven,
and
the powers
of heaven
will be shaken}}.
\VS{30}Then
the sign
of the Son
of Man
will appear
in
heaven,
and
all
the tribes
of the earth
will mourn.
They will see
{\IT{
the Son
of Man
arriving
on
the clouds
of heaven}}
with
power
and
great
glory.
\VS{31}And
he will send
his
angels
with
a loud
trumpet
blast,
and
they will gather
his
elect
from
the four
winds,
from
one end
of heaven
to
the other.
\par }{\SH The Parable of the Fig Tree
\par }{\PP \VS{32}“Learn
this parable
from
the fig tree: Whenever
its
branch
becomes
tender
and
puts out
its leaves,
you know
that
summer
is near.
\VS{33}So
also
you,
when
you see
all
these things,
know
that
he is
near, right
at
the door.
\VS{34}I tell
you
the truth,
this
generation
will
not
pass away
until
all
these things
take place.
\VS{35}Heaven
and
earth
will pass away,
but
my
words
will
never
pass away.
\par }{\SH Be Ready!
\par }{\PP \VS{36}“But
as for
that
day
and
hour
no one
knows it – not even the angels in heaven – except the Father alone.
\VS{37}For
just
like the days
of Noah
were, so
the coming
of the Son
of Man
will be.
\VS{38}For
in
those
days
before
the flood,
people were eating
and
drinking,
marrying
and
giving in marriage,
until
the day
Noah
entered
the ark.
\VS{39}And
they knew
nothing
until
the flood
came
and
took
them all
away.
It will be
the same at the coming
of the Son
of Man.
\VS{40}Then
there will be
two
men in
the field;
one
will be taken
and
one
left.
\VS{41}There will be
two
women grinding grain
with
a mill;
one
will be taken
and
one
left.
\par }{\PP \VS{42}“Therefore
stay alert,
because
you do
not
know
on what
day
your
Lord
will come.
\VS{43}But
understand
this: If
the owner of the house
had known
at what
time of night
the thief
was coming,
he would have been alert
and
would
not
have let
his
house
be broken into.
\VS{44}Therefore
you
also
must be
ready,
because
the Son
of Man
will come
at an hour
when you do
not
expect him.
\par }{\SH The Faithful and Wise Slave
\par }{\PP \VS{45}“Who
then
is
the faithful
and
wise
slave,
whom
the master
has put in charge
of
his
household,
to give
the other slaves their
food
at
the proper time?
\VS{46}Blessed
is that
slave
whom
the master
finds
at work
when he comes.
\VS{47}I tell
you
the truth,
the master will put
him
in charge
of
all
his
possessions.
\VS{48}But
if
that evil
slave
should say
to
himself, ‘My
master
is staying away a long time,’
\VS{49}and
he begins
to beat
his
fellow slaves
and
to eat
and
drink
with
drunkards,
\VS{50}then the master
of that
slave
will come
on
a day
when he does
not
expect
him and
at
an hour
he does
not
foresee,
\VS{51}and
will cut
him
in two, and
assign
him
a place with
the hypocrites,
where
there
will be
weeping
and
gnashing
of teeth.

\par }\Chap{25}{\PP \VerseOne{1}“At that time
the kingdom
of heaven
will be like
ten
virgins
who
took
their
lamps
and went out
to
meet
the bridegroom.
\VS{2}Five
of
the virgins
were
foolish,
and
five
were wise.
\VS{3}When the foolish ones
took
their
lamps,
they did not
take
extra olive oil
with
them.
\VS{4}But
the wise ones
took
flasks
of olive oil
with
their
lamps.
\VS{5}When
the bridegroom
was delayed a long time,
they
all
became drowsy
and
fell asleep.
\VS{6}But
at midnight
there was
a shout,
‘Look,
the bridegroom
is here! Come out
to
meet him.’
\VS{7}Then
all
the virgins
woke up
and
trimmed
their
lamps.
\VS{8}The foolish ones
said
to the wise,
‘Give
us
some of
your
oil,
because
our
lamps
are going out.’
\VS{9}‘No,’ they replied. ‘There won’t
be enough
for you
and
for us.
Go
instead
to
those who sell
oil and
buy
some for yourselves.’
\VS{10}But
while
they
had gone
to buy
it, the bridegroom
arrived, and
those who were ready
went inside
with
him
to
the wedding banquet.
Then
the door
was shut.
\VS{11}Later,
the other
virgins
came
too, saying,
‘Lord,
lord! Let
us
in!’
\VS{12}But
he replied, ‘I tell
you
the truth,
I do
not
know
you!’
\VS{13}Therefore
stay alert,
because
you do
not
know
the day
or
the hour.
\par }{\SH The Parable of the Talents
\par }{\PP \VS{14}“For
it is like
a man
going on a journey,
who summoned
his
slaves
and
entrusted
his
property
to them.
\VS{15}To one
he gave
five
talents,
to another
two,
and
to another
one,
each
according
to his
ability.
Then
he went on his journey.
\VS{16}The one who had received five
talents
went off right away
and put
his money to work
and
gained
five
more.
\VS{17}In the same way,
the one who had two
gained
two
more.
\VS{18}But
the one who had received
one talent
went out
and dug a hole
in the ground
and
hid
his
master’s
money in it.
\VS{19}After
a long
time,
the master
of those
slaves
came
and
settled
his
accounts
with
them.
\VS{20}The one who had received
the five
talents
came
and brought
five
more,
saying,
‘Sir,
you entrusted
me
with five
talents.
See,
I have gained
five
more.’
\VS{21}His
master
answered, ‘Well done,
good
and
faithful
slave! You have been
faithful
in
a few things.
I will put
you
in charge
of
many things.
Enter
into
the joy
of your
master.’
\VS{22}The one with the two
talents
also came
and
said,
‘Sir,
you entrusted
two
talents
to me.
See,
I have gained
two
more.’
\VS{23}His
master
answered,
‘Well done,
good
and
faithful
slave! You have been
faithful
with
a few things.
I will put
you
in charge
of many things.
Enter
into
the joy
of your
master.’
\VS{24}Then the one who had received
the one
talent
came
and said,
‘Sir,
I knew
that
you
were
a hard
man,
harvesting
where
you did
not
sow,
and
gathering
where
you did
not
scatter seed,
\VS{25}so
I was afraid,
and I went
and hid
your
talent
in
the ground.
See,
you have
what is yours.’
\VS{26}But
his
master
answered, ‘Evil
and
lazy
slave! So you knew
that
I harvest
where
I
didn’t
sow
and
gather
where
I
didn’t
scatter?
\VS{27}Then you
should
have deposited
my
money
with the bankers,
and
on my return
I
would
have received
my money
back with
interest!
\VS{28}Therefore
take
the talent
from
him
and
give
it to the one who has
ten.
\VS{29}For
the one who has
will be given
more,
and
he will have more than enough.
But
the one who does
not
have,
even
what
he has
will be taken
from
him.
\VS{30}And
throw
that worthless
slave
into
the outer
darkness,
where there
will be
weeping
and
gnashing
of teeth.’
\par }{\SH The Judgment
\par }{\PP \VS{31}“When
the Son
of Man
comes
in
his
glory
and
all
the angels
with
him,
then
he will sit
on
his
glorious
throne.
\VS{32}All
the nations
will be assembled
before
him,
and
he will separate
people
one
from
another
like
a shepherd
separates
the sheep
from
the goats.
\VS{33}He will put
the sheep
on
his
right
and
the goats
on
his left.
\VS{34}Then
the king
will say
to those on
his
right,
‘Come, you
who are blessed
by my
Father,
inherit
the kingdom
prepared
for you
from
the foundation
of the world.
\VS{35}For
I was hungry
and
you gave
me
food,
I was thirsty
and
you gave
me
something to drink,
I was
a stranger
and
you invited
me in,
\VS{36}I was naked
and
you gave
me
clothing,
I was sick
and
you took care
of me,
I was
in
prison
and
you visited
me.’
\VS{37}Then
the righteous
will answer
him, ‘Lord,
when
did we see
you
hungry
and
feed
you, or
thirsty
and
give
you something to drink?
\VS{38}When
did we see
you
a stranger
and
invite
you in, or
naked
and
clothe you?
\VS{39}When
did we see
you
sick
or
in
prison
and
visit
you?’
\VS{40}And
the king
will answer
them, ‘I tell
you
the truth,
just as
you did
it for one
of the least
of these
brothers
or sisters of mine,
you did
it for me.’
\par }{\PP \VS{41}“Then
he will say
to those on
his left,
‘Depart
from
me,
you accursed,
into
the eternal
fire
that has been prepared
for the devil
and
his
angels!
\VS{42}For
I was hungry
and
you gave
me
nothing
to eat,
I was thirsty
and
you gave
me
nothing
to drink.
\VS{43}I was
a stranger
and
you did
not
receive
me
as a guest, naked
and
you did
not
clothe
me,
sick
and
in
prison
and
you did
not
visit
me.’
\VS{44}Then
they
too
will answer, ‘Lord,
when
did we see
you
hungry
or
thirsty
or
a stranger
or
naked
or
sick
or
in
prison,
and
did not
give
you whatever
you
needed?’
\VS{45}Then
he will answer
them, ‘I tell
you
the truth,
just as
you did
not
do
it for one
of the least
of these,
you did
not
do
it for me.’
\VS{46}And
these
will depart
into
eternal
punishment,
but
the righteous
into
eternal
life.”

\par }\Chap{26}{\PP \VerseOne{1}When
Jesus
had finished
saying
all
these things,
he told
his
disciples,
\VS{2}“You know
that
after
two
days
the Passover
is coming,
and
the Son
of Man
will be handed over
to
be crucified.”
\VS{3}Then
the chief priests
and
the elders
of the people
met together
in
the palace
of the high priest,
who was named
Caiaphas.
\VS{4}They planned
to
arrest
Jesus
by stealth
and
kill him.
\VS{5}But
they said,
“Not
during
the feast,
so that
there won’t
be
a riot
among
the people.”
\par }{\SH Jesus’ Anointing
\par }{\PP \VS{6}Now while
Jesus
was
in
Bethany
at
the house
of Simon
the leper,
\VS{7}a woman
came
to him
with
an alabaster jar
of expensive
perfumed oil,
and
she poured
it on
his
head
as he was at the table.
\VS{8}When
the disciples
saw
this, they became indignant
and said,
“Why
this
waste?
\VS{9}It
could
have been sold
at a high price
and
the money given
to the poor!”
\VS{10}When
Jesus
learned
of this, he said
to them,
“Why
are you bothering
this woman? She has done
a good
service
for
me.
\VS{11}For
you will
always
have
the poor
with
you,
but
you will
not
always
have
me!
\VS{12}When
she poured
this
oil
on
my
body,
she did
it to
prepare
me
for burial.
\VS{13}I tell
you
the truth,
wherever
this
gospel
is proclaimed
in
the whole
world,
what
she
has done
will
also
be told
in
memory
of her.”
\par }{\SH The Plan to Betray Jesus
\par }{\PP \VS{14}Then
one
of the twelve,
the one named
Judas
Iscariot,
went
to
the chief priests
\VS{15}and said,
“What
will
you
give
me
to betray
him
into your hands?” So
they set out
thirty
silver coins
for him.
\VS{16}From
that time on,
Judas began looking for
an opportunity
to
betray
him.
\par }{\SH The Passover
\par }{\PP \VS{17}Now
on the first
day of the feast of Unleavened Bread
the disciples
came
to Jesus
and said, “Where
do you want
us to prepare
for you
to eat
the Passover?”
\VS{18}He said,
“Go
into
the city
to
a certain man
and
tell
him,
‘The Teacher
says,
“My
time
is
near.
I will observe
the Passover
with
my
disciples
at
your
house.” ’ ”
\VS{19}So
the disciples
did
as
Jesus
had instructed
them,
and
they prepared
the Passover.
\VS{20}When it was
evening,
he took
his place at the table
with
the twelve.
\VS{21}And
while
they
were eating
he said,
“I tell
you
the truth,
one
of
you
will betray
me.”
\VS{22}They became
greatly
distressed
and
each
one
began
to say
to him,
“Surely not
I,
Lord?”
\VS{23}He answered,
“The one
who has dipped
his hand
into
the bowl
with
me
will betray
me.
\VS{24}The Son
of Man
will go
as
it is written
about
him,
but woe
to that
man
by
whom
the Son
of Man
is betrayed! It would be
better
for him
if
he had
never
been born.”
\VS{25}Then
Judas,
the one who would betray
him,
said,
“Surely not
I,
Rabbi?” Jesus replied,
“You
have said it yourself.”
\par }{\SH The Lord’s Supper
\par }{\PP \VS{26}While
they
were eating,
Jesus
took
bread,
and
after giving thanks
he broke
it, gave
it to his disciples,
and said,
“Take,
eat,
this
is
my
body.”
\VS{27}And
after taking
the cup
and
giving thanks,
he gave
it to them,
saying,
“Drink
from
it,
all
of you,
\VS{28}for
this
is
my
blood,
the blood of the covenant,
that is poured out
for
many
for
the forgiveness
of sins.
\VS{29}I tell
you,
from
now on
I will
not
drink
of
this
fruit
of the vine
until
that
day
when
I drink
it
new
with
you
in
my
Father’s
kingdom.”
\VS{30}After singing a hymn,
they went out
to
the Mount
of Olives.
\par }{\SH The Prediction of Peter’s Denial
\par }{\PP \VS{31}Then
Jesus
said
to them,
“This
night
you
will
all
fall away
because
of me,
for
it is written:
\par }{\Q ‘{\QT{
I will strike
the shepherd,}}
\par }{\Q {\QT{
and
the sheep
of the flock
will be scattered.}}’
\par }{\PI \VS{32}But
after
I am raised,
I
will go ahead
of you
into
Galilee.”
\VS{33}Peter
said
to him,
“If
they
all
fall away
because of
you,
I
will
never
fall away!”
\VS{34}Jesus
said
to him,
“I tell
you
the truth,
on
this
night,
before
the rooster
crows,
you will deny
me
three times.”
\VS{35}Peter
said
to him,
“Even
if I
must
die
with
you,
I will
never
deny
you.”
And
all
the disciples
said
the same thing.
\par }{\SH Gethsemane
\par }{\PP \VS{36}Then
Jesus
went
with
them
to
a place
called
Gethsemane,
and
he said
to the disciples,
“Sit
here
while
I go over
there
and pray.”
\VS{37}He
took
with him Peter
and
the two
sons
of Zebedee,
and became
anguished
and
distressed.
\VS{38}Then
he said
to them,
“My
soul
is
deeply grieved,
even to the point
of death.
Remain
here
and
stay awake
with
me.”
\VS{39}Going
a little
farther,
he threw
himself down
with
his
face
to the ground and prayed, “My
Father,
if
possible,
let this
cup
pass
from
me! Yet
not
what
I
will,
but
what
you will.”
\VS{40}Then
he came
to
the disciples
and
found
them
sleeping.
He
said
to Peter,
“So,
couldn’t
you stay awake
with
me
for one
hour?
\VS{41}Stay awake
and
pray
that
you will
not
fall
into
temptation.
The spirit
is willing,
but
the flesh
is weak.”
\VS{42}He went away
a second time
and prayed, “My
Father,
if
this
cup cannot
be taken away
unless
I drink
it,
your
will
must be done.”
\VS{43}He
came
again
and found
them
sleeping;
they
could not keep
their
eyes
open.
\VS{44}So
leaving
them
again,
he went away
and prayed
for
the third time,
saying
the same thing
once more.
\VS{45}Then
he came
to
the disciples
and
said
to them,
“Are you still sleeping
and
resting? Look,
the hour
is approaching,
and
the Son
of Man
is betrayed
into
the hands
of sinners.
\VS{46}Get up,
let us go.
Look! My
betrayer
is approaching!”
\par }{\SH Betrayal and Arrest
\par }{\PP \VS{47}While
he
was
still
speaking,
Judas,
one
of the twelve,
arrived.
With
him
was a large
crowd
armed with
swords
and
clubs,
sent by the chief priests
and
elders
of the people.
\VS{48}(Now
the betrayer
had given
them
a sign,
saying,
“The one
I kiss
is
the man.
Arrest
him!”)
\VS{49}Immediately
he went up
to Jesus
and said,
“Greetings,
Rabbi,”
and
kissed
him.
\VS{50}Jesus
said
to him,
“Friend,
do what you are here
to do.” Then
they came
and took hold
of
Jesus
and
arrested
him.
\VS{51}But
one
of those with
Jesus
grabbed
his
sword,
drew
it
out,
and
struck
the high priest’s
slave,
cutting off
his
ear.
\VS{52}Then
Jesus
said
to him,
“Put
your
sword
back
in
its
place! For
all
who take
hold of the sword
will die
by
the sword.
\VS{53}Or
do you think
that
I cannot
call
on my
Father,
and
that he would send
me
more than
twelve
legions
of angels
right now?
\VS{54}How
then
would the scriptures
that say it must
happen
this way
be fulfilled?”
\VS{55}At
that
moment
Jesus
said
to the crowd,
“Have you come out
with
swords
and
clubs
to arrest
me
like
you would an outlaw? Day after day
I sat
teaching
in
the temple courts,
yet
you did
not
arrest
me.
\VS{56}But
this
has happened
so
that the scriptures
of the prophets
would be fulfilled.”
Then
all
the disciples
left
him
and fled.
\par }{\SH Condemned by the Sanhedrin
\par }{\PP \VS{57}Now
the ones who had arrested
Jesus
led
him to
Caiaphas,
the high priest,
in whose house
the experts in the law
and
the elders
had gathered.
\VS{58}But
Peter
was following
him
from
a distance,
all the way to
the high priest’s
courtyard.
After
going
in,
he sat
with
the guards
to see
the outcome.
\VS{59}The chief priests
and
the whole
Sanhedrin
were trying to find
false testimony
against
Jesus
so
that they could put
him
to death.
\VS{60}But
they did
not
find
anything, though many
false witnesses
came forward. Finally two came forward
\VS{61}and declared,
“This man
said,
‘I am able
to destroy
the temple
of God
and
rebuild
it in
three
days.’ ”
\VS{62}So
the high priest
stood up
and said
to him,
“Have you
no
answer? What
is this
that they are testifying against
you?”
\VS{63}But
Jesus
was silent.
The high priest
said
to him,
“I charge
you
under oath
by
the living
God,
tell
us
if
you
are
the Christ,
the Son
of God.”
\VS{64}Jesus
said
to him,
“You
have said
it yourself. But
I tell
you,
from
now on
you will see
the Son
of Man
{\IT{
sitting
at
the right hand}}
of the Power
and
{\IT{
coming
on
the clouds
of heaven}}.”
\VS{65}Then
the high priest
tore
his
clothes
and declared, “He has blasphemed! Why
do we
still
need
witnesses? Now
you have heard
the blasphemy!
\VS{66}What
is
your
verdict?” They answered,
“He is
guilty
and deserves death.”
\VS{67}Then
they spat
in
his
face
and
struck him with their
fists.
And
some slapped him,
\VS{68}saying,
“Prophesy
for us,
you Christ! Who
hit
you?”
\par }{\SH Peter’s Denials
\par }{\PP \VS{69}Now
Peter
was sitting
outside
in
the courtyard.
A
slave girl
came
to him
and
said,
“You
also were
with
Jesus
the Galilean.”
\VS{70}But
he denied
it in front of
them all: “I
don’t
know
what
you’re talking about!”
\VS{71}When
he went out
to
the gateway,
another slave girl
saw
him
and
said
to the people there,
“This man
was
with
Jesus
the Nazarene.”
\VS{72}He
denied
it again
with
an oath,
“I do
not
know
the man!”
\VS{73}After
a little
while,
those standing
there came up
to Peter
and said,
“You
really
are
one of
them too – even your accent gives you away!”
\VS{74}At that
he began
to curse,
and
he swore
with an oath, “I do
not
know
the man!” At that moment
a rooster
crowed.
\VS{75}Then
Peter
remembered
what Jesus
had said: “Before the rooster
crows,
you will deny
me
three times.”
And
he went
outside
and wept
bitterly.

\par }\Chap{27}{\PP \VerseOne{1}When it was
early in the morning,
all
the chief priests
and
the elders
of the people
plotted
against
Jesus
to
execute
him.
\VS{2}They tied
him
up,
led
him away,
and
handed
him over
to Pilate
the governor.
\par }{\SH Judas’ Suicide
\par }{\PP \VS{3}Now
when
Judas,
who had betrayed
him,
saw
that
Jesus had been condemned,
he regretted
what he had done and returned
the thirty
silver coins
to the chief priests
and
the elders,
\VS{4}saying,
“I have sinned
by betraying
innocent
blood!” But
they said,
“What
is that to
us? You
take care of it yourself!”
\VS{5}So
Judas threw
the silver coins
into
the temple
and
left.
Then he went out
and hanged himself.
\VS{6}The chief priests
took
the silver
and said,
“It is
not
lawful
to put
this
into
the temple treasury,
since
it is
blood
money.”
\VS{7}After consulting together
they bought
the Potter’s
Field
with
it,
as
a burial place
for foreigners.
\VS{8}For this reason
that
field
has been called
the “Field
of Blood”
to
this day.
\VS{9}Then
what was spoken
by
Jeremiah
the prophet
was fulfilled{\IT{
: “They took
the thirty
silver coins,
the price
of the one whose
price
had been set
by
the people
of Israel}},
\VS{10}{\IT{
and
they gave
them
for
the potter’s
field,
as
the Lord
commanded
me}}.”
\par }{\SH Jesus and Pilate
\par }{\PP \VS{11}Then Jesus
stood
before
the governor,
and
the governor
asked
him, “Are
you
the king
of the Jews?” Jesus
said,
“You
say so.”
\VS{12}But
when he
was accused
by
the chief priests
and
the elders,
he did
not
respond.
\VS{13}Then
Pilate
said
to him,
“Don’t
you hear
how many
charges they are bringing against
you?”
\VS{14}But
he did
not
answer
even
one
accusation,
so
that the governor
was
quite
amazed.
\par }{\PP \VS{15}During
the feast
the governor
was accustomed
to release
one
prisoner
to the crowd,
whomever
they wanted.
\VS{16}At that time
they had
in custody a notorious
prisoner
named
Jesus Barabbas.
\VS{17}So
after
they
had assembled,
Pilate
said
to them,
“Whom
do you want
me to release
for you,
Jesus Barabbas
or
Jesus
who is called
the Christ?”
\VS{18}(For
he knew
that
they had handed
him
over
because of
envy.)
\VS{19}As
he
was sitting
on
the judgment seat,
his
wife
sent a message
to
him: “Have nothing
to do with that innocent
man;
I have suffered
greatly
as a result of
a dream
about
him
today.”
\VS{20}But
the chief priests
and
the elders
persuaded
the crowds
to
ask for
Barabbas
and
to have
Jesus
killed.
\VS{21}The governor
asked
them,
“Which
of
the two
do you want
me to release
for you?” And
they said,
“Barabbas!”
\VS{22}Pilate
said
to them,
“Then what
should I do
with Jesus
who is called
the Christ?” They
all
said,
“Crucify him!”
\VS{23}He asked,
“Why? What
wrong
has he done?” But
they shouted
more insistently,
“Crucify him!”
\par }{\SH Jesus is Condemned and Mocked
\par }{\PP \VS{24}When
Pilate
saw
that
he could do
nothing,
but
that instead
a riot
was starting,
he took
some water,
washed
his hands
before
the crowd
and said,
“I am
innocent
of
this man’s blood.
You take care
of it
yourselves!”
\VS{25}In reply
all
the people
said,
“Let his
blood
be on
us
and
on
our
children!”
\VS{26}Then
he released
Barabbas
for them.
But
after he had Jesus
flogged,
he handed
him over
to
be crucified.
\VS{27}Then
the governor’s
soldiers
took
Jesus
into
the governor’s residence
and gathered
the whole
cohort
around
him.
\VS{28}They stripped
him
and put
a scarlet
robe
around him,
\VS{29}and
after braiding
a crown
of
thorns,
they put
it on
his
head.
They put a staff
in
his
right hand,
and
kneeling down
before
him,
they mocked
him: “Hail,
king
of the Jews!”
\VS{30}They spat
on
him
and took
the staff
and
struck
him
repeatedly
on
the head.
\VS{31}When
they had mocked
him,
they stripped
him
of the robe
and
put
his
own clothes
back on him.
Then
they led
him
away
to
crucify him.
\par }{\SH The Crucifixion
\par }{\PP \VS{32}As
they were going out,
they found
a man
from Cyrene
named
Simon,
whom
they forced
to
carry
his
cross.
\VS{33}They came
to
a place
called
Golgotha
(which
means
“Place
of the Skull”)
\VS{34}and offered
Jesus
wine
mixed
with
gall
to drink.
But
after tasting
it, he would
not
drink it.
\VS{35}When
they had crucified
him,
{\IT{
they divided
his
clothes
by throwing
dice}}.
\VS{36}Then
they sat down
and kept guard
over him
there.
\VS{37}Above
his
head
they put
the charge
against him,
which read: “This
is
Jesus,
the king
of the Jews.”
\VS{38}Then
two
outlaws
were crucified
with
him,
one
on
his right
and
one
on
his left.
\VS{39}Those who passed by
defamed
him,
shaking
their
heads
\VS{40}and
saying,
“You who can destroy
the temple
and
rebuild
it in
three
days,
save
yourself! If
you are
God’s
Son,
come down
from
the cross!”
\VS{41}In the same way
even
the chief priests
– together with
the experts in the law
and
elders –
were mocking him:
\VS{42}“He saved
others,
but he cannot
save
himself! He is
the king
of Israel! If he comes down
now
from
the cross,
we will believe
in
him!
\VS{43}{\IT{
He trusts
in
God
– let God, if
he wants
to, deliver
him
now}}
because
he said,
‘I am
God’s
Son’!”
\VS{44}The robbers
who were crucified
with
him
also spoke abusively
to him.
\par }{\SH Jesus’ Death
\par }{\PP \VS{45}Now
from
noon
until
three,
darkness
came
over
all
the land.
\VS{46}At about
three
o’clock
Jesus
shouted
with a loud
voice, “{\IT{
Eli,
Eli,
lema
sabachthani?
}}” that
is,
{\QT{
“My
God,
my
God,
why
have you forsaken
me}}?”
\VS{47}When some
of the bystanders
heard
it, they said,
“This man
is calling
for Elijah.”
\VS{48}Immediately
one
of
them
ran
and
got a
sponge,
filled
it with sour wine,
put
it on a stick,
and gave
it to him to drink.
\VS{49}But
the rest
said,
“Leave
him alone! Let’s see
if
Elijah
will come
to save
him.”
\VS{50}Then
Jesus
cried out
again
with a loud
voice
and gave up
his spirit.
\VS{51}Just
then the temple
curtain
was torn
in
two,
from
top
to
bottom.
The earth
shook
and
the rocks
were split apart.
\VS{52}And
tombs
were opened,
and
the bodies
of many
saints
who had died
were raised.
\VS{53}(They
came out
of
the tombs
after
his
resurrection
and went
into
the holy
city
and
appeared
to many people.)
\VS{54}Now
when the centurion
and
those with
him
who were guarding
Jesus
saw
the earthquake
and
what took place,
they were
extremely
terrified
and said,
“Truly
this one
was
God’s
Son!”
\VS{55}Many
women
who
had followed
Jesus
from
Galilee
and given
him
support
were
also
there,
watching
from
a distance.
\VS{56}Among
them
were
Mary
Magdalene,
Mary
the mother
of James
and
Joseph,
and
the mother
of the sons
of Zebedee.
\par }{\SH Jesus’ Burial
\par }{\PP \VS{57}Now when
it was
evening,
there came
a rich
man
from
Arimathea,
named
Joseph,
who
was
also a disciple
of Jesus.
\VS{58}He
went
to Pilate
and asked
for the body
of Jesus.
Then
Pilate
ordered
that it be given to him.
\VS{59}Joseph
took
the body,
wrapped
it
in
a clean
linen cloth,
\VS{60}and
placed
it
in
his
own new
tomb
that
he had cut
in
the rock.
Then
he rolled
a great
stone
across the entrance
of the tomb
and went away.
\VS{61}(Now
Mary
Magdalene
and
the other
Mary
were sitting
there,
opposite
the tomb.)
\par }{\SH The Guard at the Tomb
\par }{\PP \VS{62}The next day
(which
is
after
the day of preparation) the chief priests
and
the Pharisees
assembled
before
Pilate
\VS{63}and said,
“Sir,
we remember
that
while that
deceiver
was
still
alive
he said,
‘After
three
days
I will rise again.’
\VS{64}So
give orders
to secure
the tomb
until
the third
day.
Otherwise
his disciples
may come
and steal
his body
and
say
to the people,
‘He has been raised
from
the dead,’
and
the last
deception
will be
worse than
the first.”
\VS{65}Pilate
said
to them,
“Take
a guard of soldiers.
Go
and make
it as secure as
you can.”
\VS{66}So
they went
with the soldiers of the guard
and made
the tomb
secure
by sealing
the stone.

\par }\Chap{28}{\PP \VerseOne{1}Now
after
the Sabbath,
at dawn
on
the first
day of the week,
Mary
Magdalene
and
the other
Mary
went
to look at
the tomb.
\VS{2}Suddenly
there was
a severe
earthquake,
for
an angel
of the Lord
descending
from
heaven
came
and rolled away
the stone
and
sat
on
it.
\VS{3}His
appearance
was
like
lightning,
and
his
clothes
were white
as
snow.
\VS{4}The guards were shaken
and
became
like
dead
men because they were so afraid
of him.
\VS{5}But
the angel
said
to the women,
“Do
not
be afraid;
I know
that
you are looking for
Jesus,
who was crucified.
\VS{6}He is
not
here,
for
he has been raised,
just as
he said.
Come
and see
the place
where
he was lying.
\VS{7}Then
go
quickly
and tell
his
disciples,
‘He has been raised
from
the dead.
He
is going ahead
of you
into
Galilee.
You will see
him
there.’
Listen,
I have told
you!”
\VS{8}So
they left
the tomb
quickly,
with
fear
and
great
joy,
and ran
to tell
his
disciples.
\VS{9}But
Jesus
met
them,
saying,
“Greetings!” They came
to him, held on
to his
feet
and
worshiped
him.
\VS{10}Then
Jesus
said
to them,
“Do
not
be afraid.
Go
and tell
my
brothers
to
go
to
Galilee.
They will see
me
there.”
\par }{\SH The Guards’ Report
\par }{\PP \VS{11}While
they
were going,
some
of the guard
went
into
the city
and told
the chief priests
everything
that had happened.
\VS{12}After
they had assembled
with
the elders
and formed
a plan,
they gave
a large sum
of money
to the soldiers,
\VS{13}telling
them, “You are to say,
‘His
disciples
came
at night
and stole
his body
while
we
were asleep.’
\VS{14}If
this
matter is heard
before
the governor,
we
will satisfy
him and
keep
you
out
of trouble.”
\VS{15}So
they took
the money
and did
as
they were instructed.
And
this
story
is told among
the Jews
to
this day.
\par }{\SH The Great Commission
\par }{\PP \VS{16}So
the eleven
disciples
went
to
Galilee
to
the mountain
Jesus
had designated.
\VS{17}When
they saw
him,
they worshiped
him, but
some doubted.
\VS{18}Then
Jesus
came up
and said
to them, “All
authority
in
heaven
and
on
earth
has been given
to me.
\VS{19}Therefore
go
and make disciples
of all
nations,
baptizing
them
in
the name
of the Father
and
the Son
and
the Holy
Spirit,
\VS{20}teaching
them
to obey
everything
I have commanded
you.
And
remember,
I
am
with
you
always,
to
the end
of the age.”
\par }