\NormalFont\ShortTitle{Micah}
{\MT Micah

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \VerseOne{1}This
is the prophetic message
that
the {\ND{Lord}}
gave to
Micah
of Moresheth.
He delivered this message during
the reigns of Jotham,
Ahaz,
and Hezekiah,
kings
of Judah.
The prophecies
pertain to Samaria
\par }{\SH The Judge is Coming
\par }{\Q \VS{2}Listen,
all
you nations!
\par }{\Q Pay attention,
all inhabitants
of earth!
\par }{\Q The sovereign

{\ND{Lord}}
will testify
against you;
\par }{\Q the
{\ND{Lord}}
will accuse you from his majestic
palace.
\par }{\Q \VS{3}Look,
the {\ND{Lord}}
is coming out
of his dwelling place!
\par }{\Q He will descend
and march
on
the earth’s
mountaintops!
\par }{\Q \VS{4}The mountains
will
disintegrate
beneath
him,
\par }{\Q and the valleys
will be split
in two.

\par }{\Q The mountains will melt like wax
in a fire,
\par }{\Q the rocks will slide down like water
cascading
down a steep slope.
\par }{\Q \VS{5}All
this
is because of Jacob’s
rebellion
\par }{\Q and the sins
of the nation
of Israel.
\par }{\Q How has
Jacob
rebelled, you ask?

\par }{\Q Samaria
epitomizes
their rebellion!

\par }{\Q Where are Judah’s
pagan
worship centers,
you ask?

\par }{\Q They are right in Jerusalem!
\par }{\Q \VS{6}“I will turn
Samaria
into a heap of ruins
in an open field
–
\par }{\Q vineyards
will be planted
there!

\par }{\Q I will tumble
the rubble of her stone
walls down into the valley,
\par }{\Q and tear down
her fortifications to their foundations.
\par }{\Q \VS{7}All
her carved idols
will be smashed
to pieces;
\par }{\Q all
her metal cult statues
will be destroyed
by fire.
\par }{\Q I will make
a waste heap
of all
her images.
\par }{\Q Since
she gathered
the metal as a prostitute
collects her wages,
\par }{\Q the idols will become a prostitute’s
wages
again.”
\par }{\Q \VS{8}For
this
reason I will mourn
and wail;
\par }{\Q I will walk
around barefoot
and without my outer garments.
\par }{\Q I will howl
like a wild dog,
\par }{\Q and screech
like an owl.
\par }{\Q \VS{9}For
Samaria’s disease
is incurable.
\par }{\Q It has infected
Judah;
\par }{\Q it has spread
to
the leadership
of my people
\par }{\Q and has even
contaminated Jerusalem!
\par }{\Q \VS{10}Don’t
spread
the news
in Gath!

\par }{\Q Don’t
shed even a single tear!

\par }{\Q In Beth Leaphrah
sit
in the dust!
\par }{\Q \VS{11}Residents
of Shaphir,
pass
by in nakedness
and humiliation!

\par }{\Q The residents
of Zaanan
can’t
leave
their city.

\par }{\Q Beth Ezel
mourns,
\par }{\Q “He takes
from
you what he desires.”
\par }{\Q \VS{12}Indeed,
the residents
of Maroth
hope for something good
to happen,

\par }{\Q though
the
{\ND{Lord}}
has sent disaster
against the city
of Jerusalem.
\par }{\Q \VS{13}Residents
of Lachish,
hitch
the horses
to the chariots!
\par }{\Q You influenced
Daughter
Zion
to sin,
\par }{\Q for
Israel’s
rebellious
deeds can be traced back to you!
\par }{\Q \VS{14}Therefore
you will have to say
farewell
to Moresheth Gath.
\par }{\Q The residents
of Achzib
will be as disappointing
\par }{\Q as a dried up well to the kings
of Israel.
\par }{\Q \VS{15}Residents
of Mareshah,
a conqueror
will attack
you,

\par }{\Q the leaders
of Israel
shall flee
to Adullam.
\par }{\Q \VS{16}Shave your heads bald
as you mourn
for
the children
you love;
\par }{\Q shave
your foreheads
as bald
as an eagle,
\par }{\Q for
they are taken
from
you into exile.

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}Those who devise
sinful plans
are as good as dead,

\par }{\Q those who dream about doing
evil
as they lie in
bed.
\par }{\Q As soon as morning
dawns
they carry out
their plans,

\par }{\Q because
they have the power to do so.
\par }{\Q \VS{2}They confiscate
the fields
they desire,
\par }{\Q and seize
the houses
they want.

\par }{\Q They defraud
people of their
homes,
\par }{\Q and deprive people
of the land
they have inherited.
\par }{\Q \VS{3}Therefore
the

{\ND{Lord}}
says
this: “Look,
I am devising
disaster
for this
nation!

\par }{\Q It will be like
a yoke from
which
you cannot
free your neck.
\par }{\Q You will no
longer walk
proudly,
\par }{\Q for
it will be a time
of catastrophe.
\par }{\Q \VS{4}In that day
people will sing
this taunt song to you –
\par }{\Q they will mock you with this lament:
\par }{\Q ‘We are completely destroyed;
\par }{\Q they sell off the property of my people.
\par }{\Q How they remove it from me!
\par }{\Q They assign our fields to the conqueror.’
\par }{\Q \VS{5}Therefore
no
one will assign
you land
in the
{\ND{Lord}}’s
community.
\par }{\Q \VS{6}‘Don’t
preach
with such impassioned rhetoric,’ they say excitedly.
\par }{\Q ‘These
prophets should not preach
of such things;
\par }{\Q we will not
be overtaken by
humiliation.’
\par }{\Q \VS{7}Does the family
of Jacob
say,
\par }{\Q ‘The
{\ND{Lord}}’s
patience can’t be exhausted –
\par }{\Q he would never do such things’?
\par }{\Q To be sure, my commands bring a reward
\par }{\Q for those who obey them,
\par }{\Q \VS{8}but
you rise up
as an enemy
against
my people.
\par }{\Q You steal
a robe
from a friend,

\par }{\Q from those who pass by
peacefully
as if returning from
a war.
\par }{\Q \VS{9}You wrongly evict
widows among my people
from their cherished
homes.
\par }{\Q You defraud
their children
of their
prized inheritance.
\par }{\Q \VS{10}But you are the ones who will be forced
to leave!

\par }{\Q For
this
land is not
secure!

\par }{\Q Sin will thoroughly destroy it!
\par }{\Q \VS{11}If
a lying
windbag
should come
and say,

\par }{\Q ‘I’ll promise you blessings of wine
and beer,’
\par }{\Q he would be
just the right preacher
for these
people!
\par }{\SH The Lord Will Restore His People
\par }{\Q \VS{12}I will certainly gather
all
of you, O Jacob,
\par }{\Q I will certainly assemble
those Israelites
who remain.
\par }{\Q I will bring
them together
like sheep
in a fold,
\par }{\Q like a flock
in the middle
of a pasture;
\par }{\Q they will be so numerous
that they will make a lot of noise.
\par }{\Q \VS{13}The one who can break through barriers
will lead
them out
\par }{\Q they will break out,
pass through
the gate,
and leave.
\par }{\Q Their king
will advance
before
them,
\par }{\Q The
{\ND{Lord}}
himself will lead them.

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}I said,
\par }{\Q “Listen,
you leaders
of Jacob,
\par }{\Q you rulers
of the nation
of Israel!
\par }{\Q You ought to know
what is just,
\par }{\Q \VS{2}yet you hate
what is good,
\par }{\Q and love
what is evil.
\par }{\Q You flay
my people’s skin
\par }{\Q and rip the flesh
from their bones.
\par }{\Q \VS{3}You devour
my people’s
flesh,
\par }{\Q strip off
their skin,
\par }{\Q and crush
their bones.
\par }{\Q You chop
them up
like flesh in a pot – ‘span class=”s 01320” title=”01320” ’
\par }{\Q like meat
in
a kettle.
\par }{\PP \par }{\Q \VS{4}Someday
these sinners will cry
to
the {\ND{Lord}}
for help,

\par }{\Q but he will not
answer
them.
\par }{\Q He will hide
his
face
from them
at that time,
\par }{\Q because
they
have done such wicked
deeds.”
\par }{\Q \VS{5}This is what
the {\ND{Lord}}
says: “The prophets
who mislead
my people
\par }{\Q are as good as dead.

\par }{\Q If someone gives them enough to eat,
\par }{\Q they offer
an oracle
of peace.
\par }{\Q But if someone does not
give
them food,
\par }{\Q they are ready to declare
war
on him.
\par }{\Q \VS{6}Therefore
night
will fall, and you will receive no visions;
\par }{\Q it will grow dark,
and you will no longer be able to read the omens.
\par }{\Q The sun
will set
on
these prophets,
\par }{\Q and the daylight
will turn to darkness
over their heads.
\par }{\Q \VS{7}The prophets
will be ashamed;
\par }{\Q the omen
readers
will be humiliated.
\par }{\Q All
of them will cover their mouths,
\par }{\Q for
they will receive no
divine
oracles.”
\par }{\Q \VS{8}But
I am
full
of the courage
that the
{\ND{Lord}}’s
Spirit
gives,
\par }{\Q and have a strong
commitment to justice.
\par }{\Q This enables me to confront
Jacob
with its rebellion,
\par }{\Q and Israel
with its sin.
\par }{\Q \VS{9}Listen
to this,
you leaders
of the family
of Jacob,
\par }{\Q you rulers
of the nation
of Israel!
\par }{\Q You hate
justice
\par }{\Q and pervert
all
that is right.
\par }{\Q \VS{10}You build
Zion
through bloody
crimes,

\par }{\Q Jerusalem
through unjust violence.
\par }{\Q \VS{11}Her leaders
take bribes
when they decide
legal cases,

\par }{\Q her priests
proclaim
rulings for profit,
\par }{\Q and her prophets
read omens
for
pay.
\par }{\Q Yet they claim to trust
the {\ND{Lord}}
and say,
\par }{\Q “The
{\ND{Lord}}
is among
us.

\par }{\Q Disaster
will not
overtake us!”
\par }{\Q \VS{12}Therefore,
because
of you, Zion
will be plowed
up like a field,
\par }{\Q Jerusalem
will become a heap of ruins,
\par }{\Q and the Temple
Mount will become a hill
overgrown with brush!

\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}In the future
the
{\ND{Lord}}’s
Temple
Mount
will be
the most important
mountain
of all;

\par }{\Q it will be more prominent
than other hills.
\par }{\Q People
will stream to it.
\par }{\Q \VS{2}Many
nations
will come,
saying,
\par }{\Q “Come on! Let’s
go up
to
the
{\ND{Lord}}’s
mountain,
\par }{\Q to
the temple
of Jacob’s
God,
\par }{\Q so he can teach
us his commands
\par }{\Q and we
can live by his laws.”

\par }{\Q For
Zion
will be the source
of instruction;
\par }{\Q the
{\ND{Lord}}’s
teachings
will proceed from
Jerusalem.
\par }{\Q \VS{3}He will arbitrate
between
many
peoples
\par }{\Q and settle
disputes
between many distant
nations.
\par }{\Q They will beat
their swords
into plowshares,
\par }{\Q and their spears
into pruning hooks.
\par }{\Q Nations
will not
use
weapons
against
other nations,
\par }{\Q and they will no
longer
train
for war.
\par }{\Q \VS{4}Each
will sit
under
his own grapevine
\par }{\Q or under
his own fig tree
without
any fear.
\par }{\Q The
{\ND{Lord}}
who commands armies
has decreed it.
\par }{\Q \VS{5}Though
all
the nations
follow
their respective gods,
\par }{\Q we
will follow
the {\ND{Lord}}
our God
forever.
\par }{\SH Restoration Will Follow Crisis
\par }{\Q \VS{6}“In that day,”
says
the {\ND{Lord}}, “I will gather
the lame,
\par }{\Q and assemble
the outcasts
whom
I injured.
\par }{\Q \VS{7}I
will transform
the
lame
into the nucleus
of a new nation,
\par }{\Q and those
far off into a mighty
nation.
\par }{\Q The
{\ND{Lord}}
will reign
over them on
Mount
Zion,
\par }{\Q from
that day forward
and forevermore.”
\par }{\Q \VS{8}As for you,
watchtower
for the flock,
\par }{\Q fortress
of Daughter
Zion – ‘span class=”s 07223” title=”07223” ’
\par }{\Q your former
dominion
will be restored,
\par }{\PP \par }{\Q the sovereignty
that belongs to Daughter
Jerusalem.
\par }{\Q \VS{9}Jerusalem, why
are you now
shouting
so loudly?

\par }{\Q Has your king
disappeared?

\par }{\Q Has your wise leader
been destroyed?
\par }{\Q Is this why
pain
grips
you as if you were a woman in labor?
\par }{\Q \VS{10}Twist
and strain,
Daughter
Zion,
as if you were in labor!
\par }{\Q For
you will leave
the city
\par }{\Q and live
in the open field.
\par }{\Q You will go
to
Babylon,
\par }{\Q but there
you will be rescued.
\par }{\Q There
the {\ND{Lord}}
will deliver you
\par }{\Q from the power
of your enemies.
\par }{\Q \VS{11}Many
nations
have now
assembled
against
you.
\par }{\Q They say,
“Jerusalem must be desecrated,
\par }{\Q so we can gloat
over Zion!”
\par }{\Q \VS{12}But they
do not
know
what the
{\ND{Lord}}
is planning;
\par }{\Q they do not
understand
his strategy.
\par }{\Q He has gathered
them like stalks of grain
to be threshed at the threshing floor.
\par }{\Q \VS{13}“Get
up and thresh,
Daughter
Zion!
\par }{\Q For
I will give you iron
horns;
\par }{\Q I will give you bronze
hooves,
\par }{\Q and you will crush
many
nations.”
\par }{\Q You will devote
to the
{\ND{Lord}}
the spoils you take from them,
\par }{\Q and dedicate
their wealth
to the sovereign Ruler
of the whole
earth.

\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}But now
slash
yourself, daughter
surrounded by soldiers!

\par }{\Q We are besieged!
\par }{\Q With a scepter
they strike
Israel’s
ruler
\par }{\Q on
the side of his face.
\par }{\SH A King Will Come and a Remnant Will Prosper
\par }{\Q \VS{2}As
for you,
Bethlehem
Ephrathah,
\par }{\Q seemingly
insignificant
among the clans of Judah
–
\par }{\Q from
you a king will emerge
who will rule
over Israel
on
my behalf,

\par }{\Q one whose
origins
are in the distant past.
\par }{\Q \VS{3}So
the
{\ND{Lord}} will hand
the people of Israel over
to their enemies

\par }{\Q until
the time
when the woman in labor
gives birth.
\par }{\Q Then the rest
of the king’s countrymen
will return
\par }{\Q to be reunited
with the people of Israel.
\par }{\Q \VS{4}He will assume his post
and shepherd
the people by the
{\ND{Lord}}’s
strength,
\par }{\Q by the sovereign
authority
of the
{\ND{Lord}}
his God.
\par }{\Q They will live securely,
for
at that time
he will be honored
\par }{\Q even
in the distant regions
of the earth.
\par }{\Q \VS{5}He will give
us peace.
\par }{\Q Should the Assyrians
try to invade
our land
\par }{\Q and attempt to set foot
in our fortresses,
\par }{\Q we will send
against
them seven
shepherd-rulers,
\par }{\Q make that eight
commanders.
\par }{\Q \VS{6}They will rule
the land
of Assyria
with the sword,
\par }{\Q the land
of Nimrod
with a drawn sword.
\par }{\Q Our king will rescue
us from the Assyrians
\par }{\Q should they attempt to invade
our land
\par }{\Q and try to set foot
in our territory.
\par }{\Q \VS{7}Those survivors
from Jacob
will live

\par }{\Q in the midst
of many
nations.
\par }{\Q They will be
like the dew
the

{\ND{Lord}}
sends,
\par }{\Q like the rain
on
the grass,
\par }{\Q that
does not
hope
for men
to come
\par }{\Q or
wait
around for humans to arrive.
\par }{\Q \VS{8}Those survivors
from Jacob
will live among the nations,
\par }{\Q in the midst
of many
peoples.
\par }{\Q They will be
like a lion
among the animals
of the forest,
\par }{\Q like a young lion
among the flocks
of sheep,
\par }{\Q which
attacks when it
passes
through;
\par }{\Q it
rips
its prey and there is no
one to stop it.
\par }{\Q \VS{9}Lift
your hand
triumphantly
against your adversaries;
\par }{\Q may all
your enemies
be destroyed!
\par }{\SH The Lord Will Purify His People
\par }{\Q \VS{10}“In that
day,”
says
the {\ND{Lord}},
\par }{\Q “I will destroy
your horses
from your midst,
\par }{\Q and smash
your chariots.
\par }{\Q \VS{11}I will destroy
the cities
of your land,
\par }{\Q and tear
down all
your fortresses.
\par }{\Q \VS{12}I will remove the sorcery
that you practice,
\par }{\Q and you will no
longer have omen readers living among you.
\par }{\Q \VS{13}I will remove
your idols
and sacred pillars
from your midst;
\par }{\Q you will no
longer
worship
what your own hands
made.
\par }{\Q \VS{14}I will uproot
your images of Asherah
from your midst,
\par }{\Q and destroy
your idols.
\par }{\Q \VS{15}I will angrily
seek vengeance
\par }{\Q on the nations
that
do not
obey me.”

\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}Listen
to what
the {\ND{Lord}}
says:

\par }{\Q “Get
up! Defend
yourself
before the mountains!

\par }{\Q Present your case
before the hills!”
\par }{\Q \VS{2}Hear
the

{\ND{Lord}}’s
accusation,
you mountains,
\par }{\Q you enduring
foundations
of the earth!
\par }{\Q For
the
{\ND{Lord}}
has a case
against
his people;
\par }{\Q he has a dispute
with
Israel!
\par }{\Q \VS{3}“My people,
how
have I wronged
you?

\par }{\Q How
have I wearied
you? Answer me!
\par }{\Q \VS{4}In fact,
I brought you up
from the land
of Egypt,
\par }{\Q I delivered
you from that place of slavery.
\par }{\Q I sent
Moses,
Aaron,
and Miriam to lead you.
\par }{\Q \VS{5}My people,
recall
how
King
Balak
of Moab
planned
to harm you,

\par }{\Q how
Balaam
son
of Beor
responded
to him.
\par }{\Q Recall how you journeyed from
Shittim
to
Gilgal,
\par }{\Q so you might
acknowledge
that the
{\ND{Lord}}
has treated you fairly.”
\par }{\Q \VS{6}With what
should I enter
the
{\ND{Lord}}’s
presence?
\par }{\Q With what should I bow
before the sovereign
God?

\par }{\Q Should I enter
his presence with burnt offerings,
\par }{\Q with year-old
calves?
\par }{\Q \VS{7}Will the
{\ND{Lord}}
accept
a thousand
rams,
\par }{\Q or ten thousand
streams
of olive oil?
\par }{\Q Should I give
him my firstborn child
as payment
for my rebellion,
\par }{\Q my offspring
– my own flesh and blood
– for my sin?
\par }{\Q \VS{8}He has told
you, O man,
what
is good,
\par }{\Q and what
the {\ND{Lord}}
really wants
from
you:

\par }{\Q He wants you to promote
justice,
to be faithful,
\par }{\Q and to live
obediently
before your God.
\par }{\Q \VS{9}Listen! The
{\ND{Lord}}
is calling
to the city!
\par }{\Q It is wise
to respect
your authority,
O
{\ND{Lord}}!

\par }{\Q Listen,
O nation,
and those
assembled in the city!
\par }{\Q \VS{10}“I will
not overlook, O sinful
house,
the dishonest
gain
you have hoarded away,

\par }{\Q or the smaller-than-standard
measure
I hate so much.
\par }{\Q \VS{11}I
do
not condone
the use of rigged
scales,
\par }{\Q or a bag
of deceptive
weights.
\par }{\Q \VS{12}The city’s rich
men think nothing of resorting
to violence;
\par }{\Q her inhabitants
lie,
\par }{\Q their tongues
speak
deceptive
words.
\par }{\Q \VS{13}I
will strike
you brutally
\par }{\Q and destroy
you because
of your sin.
\par }{\Q \VS{14}You
will eat,
but not
be satisfied.
\par }{\Q Even if you have
the strength to overtake
some
prey,

\par }{\Q you will not
be able to carry
it
away;
\par }{\Q if you do happen to carry away
something,
\par }{\Q I will deliver
it over to the sword.
\par }{\Q \VS{15}You
will plant
crops, but will not
harvest
them;
\par }{\Q you
will squeeze
oil from the olives,
but you will have no
oil
to rub
on your bodies;

\par }{\Q you will squeeze juice
from the grapes, but you will have no
wine
to drink.
\par }{\Q \VS{16}You implement
the regulations
of Omri,
\par }{\Q and all
the practices
of Ahab’s
dynasty;
\par }{\Q you follow
their policies.
\par }{\Q Therefore I will make
you an appalling sight,
\par }{\Q the city’s inhabitants
will be taunted
derisively,
\par }{\Q and nations
will mock all of you.”

\par }\Chap{7}{\PP \par }{\Q \VerseOne{1}I am depressed!

\par }{\Q Indeed,
it is as if the summer fruit
has been gathered,
\par }{\Q and the grapes
have been harvested.

\par }{\Q There is no
grape
cluster
to eat,
\par }{\Q no fresh
figs that I crave
so much.
\par }{\Q \VS{2}Faithful men
have disappeared
from
the land;
\par }{\Q there are no
godly
men
left.

\par }{\Q They all
wait in ambush
so they can shed blood;
\par }{\Q they
hunt
their own brother
with a net.
\par }{\Q \VS{3}They are determined to be experts
at
doing
evil;
\par }{\Q government
officials
and judges
take bribes,
\par }{\Q prominent
men make demands,
\par }{\Q and they all
do
what
is necessary
to satisfy them.
\par }{\Q \VS{4}The best
of them is like a thorn;
\par }{\Q the most godly
among them are more dangerous than a row of thorn bushes.
\par }{\Q The day
you try to avoid by posting watchmen
–
\par }{\Q your appointed time of punishment
– is on the way,
\par }{\Q and then you will experience
confusion.
\par }{\Q \VS{5}Do not
rely on
a friend;
\par }{\Q do not
trust
a companion!
\par }{\Q Don’t
even share
secrets
with the one who lies
in your arms!
\par }{\Q \VS{6}For
a son
thinks
his father
is a fool,
\par }{\Q a daughter
challenges
her mother,
\par }{\Q and a daughter-in-law
her mother-in-law;
\par }{\Q a man’s
enemies
are his own servants.
\par }{\Q \VS{7}But I
will keep watching
for the
{\ND{Lord}};
\par }{\Q I will wait
for the God
who delivers
me.
\par }{\Q My God
will hear my lament.
\par }{\SH Jerusalem Will Be Vindicated
\par }{\Q \VS{8}My enemies,
do not
gloat
over me!
\par }{\Q Though
I have fallen,
I will get
up.
\par }{\Q Though
I sit
in darkness,
the {\ND{Lord}}
will be my light.
\par }{\Q \VS{9}I must endure
the
{\ND{Lord}}’s
anger,
\par }{\Q for
I have sinned
against him.
\par }{\Q But then he will defend
my cause,
\par }{\Q and accomplish
justice
on my behalf.
\par }{\Q He will lead me out
into the light;
\par }{\Q I
will experience firsthand
his deliverance.
\par }{\Q \VS{10}When my enemies
see
this, they will be covered
with shame.
\par }{\Q They say
to me,
“Where
is the
{\ND{Lord}}
your God?”
\par }{\Q I will gloat
over them.

\par }{\Q Then they will be
trampled down
\par }{\Q like mud
in the streets.
\par }{\Q \VS{11}It
will be a day
for rebuilding
your walls;
\par }{\Q in that day
your boundary will be extended.
\par }{\SH A Closing Prayer
\par }{\Q \VS{12}In that day
people
will come
to you

\par }{\Q from
Assyria
as far as
Egypt,
\par }{\Q from
Egypt
as far as the
Euphrates River,
\par }{\Q from the seacoasts
and the mountains.
\par }{\Q \VS{13}The earth
will become
desolate
\par }{\Q because
of what its inhabitants
have done.
\par }{\Q \VS{14}Shepherd
your people
with your shepherd’s rod,
\par }{\Q the flock
that belongs
to you,

\par }{\Q the one that lives
alone
in a thicket,
\par }{\Q in the midst
of a pastureland.
\par }{\Q Allow them to graze
in Bashan
and Gilead,
\par }{\Q as they did in the old
days.
\par }{\Q \VS{15}“As in the days
when you departed
from the land
of Egypt,
\par }{\Q I will show
you miraculous deeds.”
\par }{\Q \VS{16}Nations
will see
this and be disappointed
by all
their strength,
\par }{\Q they will put
their hands
over
their mouths,
\par }{\Q and act as if they were deaf.
\par }{\Q \VS{17}They will lick
the dust
like a snake,
\par }{\Q like serpents
crawling
on the ground.
\par }{\Q They will come trembling from their strongholds
\par }{\Q to
the {\ND{Lord}}
our God;
\par }{\Q they will be
terrified of you.
\par }{\Q \VS{18}There is no
other God
like
you!

\par }{\Q You forgive
sin
\par }{\Q and pardon
the rebellion
\par }{\Q of those who remain
among your people.
\par }{\Q You do not
remain angry
forever,
\par }{\Q but
delight
in showing
loyal love.
\par }{\Q \VS{19}You will once again
have mercy
on us;
\par }{\Q you will conquer
our evil deeds;
\par }{\Q you will hurl
our sins
into
the depths
of the sea.
\par }{\Q \VS{20}You will
be loyal
to Jacob
\par }{\Q and extend your loyal love
to Abraham,
\par }{\Q which
you promised
on oath to our ancestors
\par }{\PP in ancient
times.
\par }