\NormalFont\ShortTitle{Nahum}
{\MT Nahum

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \par }{\Q \VerseOne{1}The oracle
against Nineveh;
\par }{\Q the book
of the vision
of Nahum
the Elkoshite:
\par }{\SH God Takes Vengeance against His Enemies
\par }{\Q \VS{2}The
{\ND{Lord}}
is a zealous
and avenging
God;
\par }{\Q the
{\ND{Lord}}
is avenging
and very angry.
\par }{\Q The
{\ND{Lord}}
takes vengeance
against his foes;
\par }{\Q he sustains
his rage against his enemies.
\par }{\Q \VS{3}The
{\ND{Lord}}
is slow
to anger
but great
in power;
\par }{\Q the {\ND{Lord}}
will certainly
not
allow the wicked to go unpunished.
\par }{\SH The Divine Warrior Destroys His Enemies but Protects His People
\par }{\Q He marches out in the whirlwind
and the raging storm;
\par }{\Q dark storm clouds
billow
like dust
under his feet.
\par }{\Q \VS{4}He shouts
a battle cry against the sea
and makes it dry up;
\par }{\Q he makes all
the rivers
run dry.
\par }{\Q Bashan
and Carmel
wither;
\par }{\Q the blossom
of Lebanon
withers.
\par }{\Q \VS{5}The mountains
tremble
before him,

\par }{\Q the hills
convulse;
\par }{\Q the earth
is laid waste before
him,
\par }{\Q the world
and all
its inhabitants are laid waste.
\par }{\Q \VS{6}No one can withstand
his indignation!

\par }{\Q No one can resist
his fierce
anger!

\par }{\Q His wrath
is poured out
like volcanic fire,
\par }{\Q boulders
are broken up
as he approaches.
\par }{\Q \VS{7}The
{\ND{Lord}}
is good –
\par }{\Q indeed, he is a fortress
in time
of distress,
\par }{\Q and he protects
those who seek refuge in him.
\par }{\Q \VS{8}But with an overwhelming
flood
\par }{\Q he will make
a complete
end
of Nineveh;

\par }{\Q he will drive
his enemies
into darkness.
\par }{\SH Denunciation and Destruction of Nineveh
\par }{\Q \VS{9}Whatever
you plot
against the
{\ND{Lord}}, he will completely
destroy!

\par }{\Q Distress
will not
arise
a second time.
\par }{\Q \VS{10}Surely
they will be totally
consumed
\par }{\Q like entangled
thorn
bushes,

\par }{\Q like the drink
of drunkards,
\par }{\Q like very dry
stubble.
\par }{\Q \VS{11}From
you, O Nineveh, one has marched forth
who plots
evil against
the {\ND{Lord}},
\par }{\Q a wicked
military strategist.
\par }{\SH Oracle of Deliverance to Judah
\par }{\Q \VS{12}This is what
the {\ND{Lord}}
says:

\par }{\Q “Even though
they are powerful –
\par }{\Q and what is
more,
even though their army
is numerous –
\par }{\Q nevertheless, they will be destroyed and trickle away!
\par }{\Q Although I afflicted you,
\par }{\Q I will afflict you no more.
\par }{\Q \VS{13}And now,
I will break
Assyria’s
yoke bar from your neck;

\par }{\Q I will tear apart
the shackles
that are on you.”
\par }{\SH Oracle of Judgment against the King of Nineveh
\par }{\Q \VS{14}The
{\ND{Lord}}
has issued a decree
against
you:

\par }{\Q “Your dynasty
will come to an end.
\par }{\Q I will destroy
the idols
and images
in the temples
of your gods.
\par }{\Q I will desecrate
your grave
– because
you are accursed!”
\par }{\SH Proclamation of the Deliverance of Judah
\par }{\Q \VS{15} Look! A herald is running on
the mountains!
\par }{\Q A messenger is proclaiming
deliverance:

\par }{\Q “Celebrate
your sacred festivals,
O Judah!
\par }{\Q Fulfill
your sacred vows
to praise God!

\par }{\Q For
never
again
will
the wicked
Assyrians invade you,
\par }{\Q they have been completely destroyed.”

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}The watchmen of Nineveh shout:

\par }{\Q “An enemy who will scatter
you is marching out
to attack you!”

\par }{\Q “Guard
the rampart!

\par }{\Q Watch
the road!
\par }{\Q Prepare
yourselves
for battle!

\par }{\Q Muster
your mighty
strength!”
\par }{\Q \VS{2}For
the {\ND{Lord}}
will restore
the majesty
of Jacob,
\par }{\Q as well as the majesty
of Israel,
\par }{\Q though
their enemies have plundered
them

\par }{\Q and have destroyed
their fields.
\par }{\SH Prophetic Vision of the Fall of Nineveh
\par }{\Q \VS{3}The shields
of his warriors
are dyed red;
\par }{\Q the mighty
soldiers
are dressed in scarlet garments.
\par }{\Q The metal fittings
of the chariots
shine
\par }{\Q like fire
on the day
of battle;
\par }{\Q the soldiers brandish
their spears.
\par }{\Q \VS{4}The chariots
race madly
through the streets,
\par }{\Q they rush back and forth
in the broad plazas;
\par }{\Q they look like
lightning bolts,
\par }{\Q they dash here and there
like flashes of lightning.
\par }{\Q \VS{5}The commander orders
his officers;
\par }{\Q they stumble
as they advance;
\par }{\Q they rush
to the city wall
\par }{\Q and they set up
the covered siege tower.
\par }{\Q \VS{6}The sluice
gates
are opened;
\par }{\Q the royal palace
is deluged and dissolves.
\par }{\Q \VS{7}Nineveh
is taken into
exile
and is led away;
\par }{\Q her slave girls
moan
like doves
while they beat
their breasts.
\par }{\Q \VS{8}Nineveh
was like a pool
of water
throughout her days,
\par }{\Q but now her people are
running away;
\par }{\Q she cries out: “Stop! Stop!” –
\par }{\Q but no one turns back.
\par }{\Q \VS{9}Her conquerors cry out:

\par }{\Q “Plunder
the silver! Plunder
the gold!”
\par }{\Q There is no
end
to the treasure;
\par }{\Q riches
of every
kind of precious
thing.
\par }{\Q \VS{10}Destruction,
devastation,
and desolation!

\par }{\Q Their hearts
faint,
\par }{\Q their knees
tremble,
\par }{\Q each stomach
churns,
each
face
turns
pale!
\par }{\SH Taunt against the Once-Mighty Lion
\par }{\Q \VS{11}Where
now is the den
of the lions,
\par }{\Q the feeding place
of the young lions,
\par }{\Q where
the lion,
lioness,
and lion
cub
once prowled

\par }{\Q and no
one disturbed them?
\par }{\Q \VS{12}The lion
tore apart
as much prey as his cubs
needed
\par }{\Q and strangled
prey to provide food for his lionesses;
\par }{\Q he filled
his lairs
with prey
\par }{\Q and his dens
with torn flesh.
\par }{\SH Battle Cry of the Divine Warrior
\par }{\Q \VS{13}“I am
against
you!” declares
the {\ND{Lord}}
who commands armies:

\par }{\Q “I will burn
your chariots
with fire;
\par }{\Q the sword
will devour
your young lions;
\par }{\Q you will no longer
prey
upon the land;
\par }{\Q the voices
of your messengers
will no
longer
be heard.”

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}Woe
to the city
guilty of bloodshed!

\par }{\Q She is full of lies;
\par }{\Q she is filled with plunder;
\par }{\Q she has hoarded
her spoil!
\par }{\SH Portrayal of the Destruction of Nineveh
\par }{\Q \VS{2}The chariot
drivers will crack their whips;
\par }{\Q the chariot
wheels
will shake the ground;
\par }{\Q the chariot horses
will gallop;
\par }{\Q the war chariots
will bolt forward!
\par }{\Q \VS{3}The charioteers
will charge ahead;
\par }{\Q their swords
will flash
\par }{\Q and their spears
will glimmer!

\par }{\Q There will be many people
slain;
\par }{\Q there will be piles
of the dead,
\par }{\Q and countless
casualties –
\par }{\Q so many that people will stumble
over the corpses.
\par }{\SH Taunt against the Harlot City
\par }{\Q \VS{4}“Because you have acted like
a wanton
prostitute –
\par }{\Q a seductive
mistress who practices
sorcery,
\par }{\Q who enslaves
nations
by her harlotry,
\par }{\Q and entices peoples
by her sorcery –
\par }{\Q \VS{5}I am against
you,” declares
the {\ND{Lord}}
who commands armies.
\par }{\Q “I will strip off
your clothes!

\par }{\Q I will show
your nakedness
to the nations
\par }{\Q and your shame
to the kingdoms;
\par }{\Q \VS{6}I will pelt
you with filth;
\par }{\Q I will treat you with contempt;
\par }{\Q I will make
you a public spectacle.
\par }{\Q \VS{7}Everyone
who sees
you will turn away from
you in disgust;
\par }{\Q they will say,
‘Nineveh
has been devastated!
\par }{\Q Who
will lament for
her?’
\par }{\Q There will be
no one to comfort you!”
\par }{\SH Nineveh Will Suffer the Same Fate as Thebes
\par }{\Q \VS{8}You are no more secure
than Thebes –
\par }{\Q she was located
on the banks of the Nile;
\par }{\Q the waters
surrounded
her,
\par }{\Q her rampart
was the sea,
\par }{\Q the water
was her wall.
\par }{\Q \VS{9}Cush
and Egypt
had limitless
strength;

\par }{\Q Put
and the Libyans
were among her allies.
\par }{\Q \VS{10}Yet
she
went
into captivity
as an exile;
\par }{\Q even
her infants
were smashed
to pieces at the head
of every
street.
\par }{\Q They cast
lots
for her nobility;
\par }{\Q all
her dignitaries
were bound with chains.
\par }{\Q \VS{11}You
too
will act like drunkards;
\par }{\Q you
will go into hiding;
\par }{\Q you
too
will seek
refuge
from the enemy.
\par }{\SH The Assyrian Defenses Will Fail
\par }{\Q \VS{12}All
your fortifications
will be like fig trees
with
first-ripe fruit:

\par }{\Q If
they are shaken,
their figs will fall
into
the mouth
of the eater!
\par }{\Q \VS{13}Your warriors
will be like women
in your midst;
\par }{\Q the gates
of your land
will be wide open
to your enemies;
\par }{\Q fire
will consume
the bars of your gates.
\par }{\Q \VS{14}Draw
yourselves water
for a siege!

\par }{\Q Strengthen
your fortifications!
\par }{\Q Trample
the mud
and tread
the clay!
\par }{\Q Make mud bricks
to strengthen your walls!
\par }{\Q \VS{15}There
the fire
will consume
you;
\par }{\Q the sword
will cut
you down;
\par }{\Q it will devour
you like the young locust would.
\par }{\SH The Assyrian Defenders Will Flee
\par }{\Q Multiply yourself like the young locust;
\par }{\Q multiply yourself like the flying locust!
\par }{\Q \VS{16}Increase
your merchants
more than the stars
of heaven!
\par }{\Q They are like the young locust
which sheds
its skin
and flies away.
\par }{\Q \VS{17}Your courtiers
are like locusts,
\par }{\Q your officials
are like a swarm of locusts!
\par }{\Q They encamp in the walls
on a cold
day,
\par }{\Q yet when the sun
rises,
they fly away;
\par }{\Q and no
one knows
where
they are.
\par }{\SH Concluding Dirge
\par }{\Q \VS{18}Your shepherds
are sleeping,
O king
of Assyria!
\par }{\Q Your officers
are slumbering!

\par }{\Q Your people
are scattered
like sheep on
the mountains
\par }{\Q and there is no one
to regather them!
\par }{\Q \VS{19}Your destruction
is like an incurable
wound;
\par }{\Q your demise is like a fatal injury!

\par }{\Q All
who hear
what has happened
to you will clap
their hands
for
joy,

\par }{\PP for no
one
ever escaped
your endless
cruelty!
\par }