\NormalFont\ShortTitle{Obadiah}
{\MT Obadiah

\par }\OneChap {\SH God’s Judgment on Edom|Edom’s Approaching Destruction
\par }{\PP \par }{\Q \VerseOne{1}The vision
that Obadiah
saw.

\par }{\Q The Lord

{\ND{God}}
says
this
concerning Edom:
\par }{\SH God’s Judgment on Edom|Edom’s Approaching Destruction
\par }{\Q We have heard a report from the
{\ND{Lord}}.
\par }{\Q An envoy was sent among the nations, saying,
\par }{\Q “Arise! Let us make war against Edom!”
\par }{\Q \VS{2}The
{\ND{Lord}} says, “Look! I will
make you a weak
nation;
\par }{\Q you
will be greatly
despised!
\par }{\Q \VS{3}Your presumptuous
heart
has deceived you –
\par }{\Q you who reside in the safety of the rocky cliffs,
\par }{\Q whose home is high in the mountains.
\par }{\Q You think to yourself,
\par }{\Q ‘No one can bring me down to the ground!’
\par }{\Q \VS{4}Even if
you were to soar high
like an eagle,
\par }{\Q even if
you were to make
your nest
among
the stars,
\par }{\Q I can bring you down
even from there!” says
the {\ND{Lord}}.
\par }{\Q \VS{5}“If
thieves
came
to rob
you during the night,
\par }{\Q they would steal
only as much as they wanted!

\par }{\Q If
grape pickers
came
to harvest your vineyards,

\par }{\Q they would leave
some
behind for the poor!

\par }{\Q But
you will be totally destroyed!
\par }{\Q \VS{6}How
the people of Esau
will be thoroughly plundered!

\par }{\Q Their hidden valuables
will be ransacked!
\par }{\Q \VS{7}All
your allies
will force
you from your homeland!

\par }{\Q Your treaty partners
will deceive
you and overpower
you.
\par }{\Q Your trusted friends will set
an ambush
for
you
\par }{\Q that will take you by surprise!
\par }{\Q \VS{8}At that time,”
the {\ND{Lord}}
says,
\par }{\Q “I will destroy
the wise
sages of Edom!

\par }{\Q the advisers
from Esau’s
mountain!
\par }{\Q \VS{9}Your warriors
will be shattered,
O Teman,
\par }{\Q so
that everyone
will be destroyed
from Esau’s
mountain!
\par }{\SH Edom’s Treachery Against Judah
\par }{\Q \VS{10}“Because you violently
slaughtered your relatives,
the people of Jacob,
\par }{\Q shame
will cover
you, and you will be destroyed
forever.
\par }{\Q \VS{11}You stood
aloof
while strangers
took
his army
captive,
\par }{\Q and foreigners
advanced
to his gates.
\par }{\Q When they cast
lots
over
Jerusalem,
\par }{\Q you behaved as though
you
were in league
with them.
\par }{\Q \VS{12}You should not
have gloated
when
your relatives
suffered
calamity.

\par }{\Q You should not
have rejoiced
over the people
of Judah
when
they were destroyed.
\par }{\Q You should not
have boasted
when
they suffered adversity.
\par }{\Q \VS{13}You should not
have entered
the city
of my people
when
they experienced distress.
\par }{\Q You should not
have joined in gloating
over their misfortune
when
they suffered distress.
\par }{\Q You should not
have looted
their wealth
when
they endured distress.
\par }{\Q \VS{14}You should not
have stood
at
the fork in the road
to slaughter
those trying to escape.
\par }{\Q You should not
have captured
their refugees
when
they suffered adversity.
\par }{\SH The Coming Day of the Lord
\par }{\Q \VS{15}“For
the day
of the {\ND{Lord}}
is approaching
for all
the nations!

\par }{\Q Just
as you have done,
so it will be done
to you.
\par }{\Q You will get exactly what your deeds deserve.
\par }{\Q \VS{16}For
just
as you have drunk
on
my holy
mountain,
\par }{\Q so all
the nations
will drink
continually.
\par }{\Q They will drink,
and they will gulp down;
\par }{\Q they will be
as though they had never
been.
\par }{\Q \VS{17}But on Mount
Zion
there will be
a remnant
of those who escape,

\par }{\Q and it will be
a holy
place once again.
\par }{\Q The descendants
of Jacob
will conquer
\par }{\Q those who had conquered them.
\par }{\Q \VS{18}The descendants
of Jacob
will be
a fire,
\par }{\Q and the descendants
of Joseph
a flame.
\par }{\Q The descendants
of Esau
will be like stubble.
\par }{\Q They will burn them up
and devour
them.
\par }{\Q There will not
be
a single survivor
of the descendants
of Esau!”
\par }{\Q Indeed,
the {\ND{Lord}}
has spoken it.
\par }{\Q \VS{19}The people of the
Negev
will take possession
of Esau’s
mountain,
\par }{\Q and the people of the Shephelah
will take
\par }{\Q possession of the
land of the Philistines.
\par }{\Q They will also take possession
of the
territory
of Ephraim
and the
territory
of Samaria,
\par }{\Q and the people of Benjamin
will take possession of Gilead.
\par }{\Q \VS{20}The exiles
of this
fortress
of the people of Israel
\par }{\Q will take possession of what belongs
to
\par }{\Q the people of Canaan,
as far
as Zarephath,
\par }{\Q and the
exiles
of Jerusalem
who
are in Sepharad
\par }{\Q will take possession
of the towns
of the Negev.
\par }{\Q \VS{21}Those who have been delivered
will go up
on Mount
Zion
\par }{\Q in order
to rule over
Esau’s
mountain.
\par }{\PP Then
the {\ND{Lord}}
will reign as King!
\par }