\NormalFont\ShortTitle{Philemon}
{\MT Philemon

\par }\OneChap {\SH Salutation
\par }{\PP \VerseOne{1}From Paul,
a prisoner
of Christ
Jesus,
and
Timothy
our brother,
to Philemon,
our
dear friend
and
colaborer,
\VS{2}to Apphia
our sister,
to Archippus
our
fellow soldier,
and
to the church
that meets in your
house.
\VS{3}Grace
and
peace
to you
from
God
our
Father
and
the Lord
Jesus
Christ!
\par }{\SH Thanks for Philemon’s Love and Faith
\par }{\PP \VS{4}I
always
thank
my
God
as I remember
you
in
my
prayers,
\VS{5}because I hear
of your
faith
in
the Lord
Jesus
and
your love
for
all
the saints.
\VS{6}I pray that the faith
you
share
with
us
may deepen
your understanding
of every
blessing
that belongs to you in
Christ.
\VS{7}I have had
great
joy
and
encouragement
because of
your
love,
for
the hearts
of the saints
have been refreshed
through
you,
brother.
\par }{\SH Paul’s Request for Onesimus
\par }{\PP \VS{8}So, although
I have
quite a lot of
confidence
in
Christ
and could command
you
to do what is proper,
\VS{9}I would
rather
appeal
to you on the basis of
love
– I,
Paul,
an old man
and
even
now
a prisoner
for the sake of Christ
Jesus –
\VS{10}I am appealing
to you
concerning
my
child,
whose
spiritual father I have become
during
my imprisonment,
that is, Onesimus,
\VS{11}who was formerly
useless
to you,
but
is now
useful
to you
and
me.
\VS{12}I have sent
him
(who
is
my
very heart) back
to you.
\VS{13}I
wanted
to
keep
him so that
he could serve
me
in
your
place during
my imprisonment
for the sake of the gospel.
\VS{14}However,
without
your
consent
I did
not
want
to do
anything, so that
your
good deed
would not
be out of compulsion,
but
from your own willingness.
\VS{15}For
perhaps
it was for
this reason
that he was separated
from you for
a little while,
so that
you would have
him
back
eternally,
\VS{16}no longer
as
a slave,
but
more than
a slave,
as a dear
brother.
He is especially
so to me,
and
even
more
so to you
now, both
humanly speaking
and
in
the Lord.
\VS{17}Therefore
if
you regard
me
as a partner,
accept
him
as
you would me.
\VS{18}Now if
he has defrauded
you
of anything
or
owes
you anything, charge
what he owes
to me.
\VS{19}I,
Paul,
have written
this letter with my own
hand: I
will repay
it. I
could
also mention
that
you owe
me
your very self.
\VS{20}Yes,
brother,
let
me
have some benefit
from you
in
the Lord.
Refresh
my
heart
in
Christ.
\VS{21}Since I was confident
that you
would obey,
I wrote
to you,
because I knew
that
you would do
even
more than
what
I am asking you to do.
\VS{22}At the same time
also,
prepare
a place for me
to stay,
for
I hope
that
through
your
prayers
I will be given back
to you.
\par }{\SH Concluding Greetings
\par }{\PP \VS{23}Epaphras,
my
fellow prisoner
in
Christ
Jesus,
greets
you.
\VS{24}Mark,
Aristarchus,
Demas
and Luke,
my
colaborers, greet you too.
\VS{25}May the grace
of the Lord
Jesus
Christ
be with
your
spirit.
\par }