\NormalFont\ShortTitle{Proverbs}
{\MT Proverbs

\par }\ChapOne{1}{\SH Introduction to the Book
\par }{\PP \par }{\Q \VerseOne{1}The Proverbs
of Solomon
son
of David,
king
of Israel:
\par }{\Q \VS{2}To learn
wisdom
and moral instruction,
\par }{\Q and to discern
wise counsel.
\par }{\Q \VS{3}To receive
moral instruction
in skillful
living,

\par }{\Q in righteousness,
justice,
and equity.
\par }{\Q \VS{4}To impart shrewdness
to the morally naive,
\par }{\Q and a discerning plan to the young person.
\par }{\Q \VS{5}(Let the wise
also hear
and gain
instruction,
\par }{\Q and let the discerning
acquire
guidance! )
\par }{\Q \VS{6}To discern
the meaning of a proverb
and a parable,
\par }{\Q the sayings
of the wise
and their riddles.
\par }{\SH Introduction to the Theme of the Book
\par }{\Q \VS{7}Fearing
the {\ND{Lord}}
is the beginning
of moral knowledge,
\par }{\Q but fools
despise
wisdom
and instruction.
\par }{\Q \VS{8}Listen,
my child,
to the instruction
from your father,
\par }{\Q and do not
forsake
the teaching
from your mother.
\par }{\Q \VS{9}For
they
will be like an elegant
garland
on your head,
\par }{\Q and like pendants
around your neck.
\par }{\SH Admonition to Avoid Easy but Unjust Riches
\par }{\Q \VS{10}My child,
if
sinners
try to entice
you,
\par }{\Q do not
consent!
\par }{\Q \VS{11}If
they say,
“Come
with
us!
\par }{\Q We will lie in wait
to shed blood;
\par }{\Q we will ambush
an innocent
person capriciously.
\par }{\Q \VS{12}We will swallow
them alive
like Sheol,
\par }{\Q those full of vigor
like those
going down
to the Pit.
\par }{\Q \VS{13}We will seize all
kinds
of precious
wealth;
\par }{\Q we
will fill
our houses
with plunder.
\par }{\Q \VS{14}Join with us!

\par }{\Q We will all
share equally in what we steal.”
\par }{\Q \VS{15}My child,
do not
go
down their way,
\par }{\Q withhold
yourself
from their
path;
\par }{\Q \VS{16}for
they are eager
to inflict harm,
\par }{\Q and they
hasten
to shed
blood.
\par }{\Q \VS{17}Surely
it is futile
to spread
a net
\par }{\Q in plain
sight
of any
bird,
\par }{\Q \VS{18}but these
men lie in wait for their
own
blood,
\par }{\Q they ambush
their own lives!
\par }{\Q \VS{19}Such
are the ways
of all
who gain
profit unjustly;
\par }{\Q it takes
away
the life
of those who
obtain it!
\par }{\SH Warning Against Disregarding Wisdom
\par }{\Q \VS{20}Wisdom calls
out in the street,
\par }{\Q she shouts loudly
in the plazas;
\par }{\Q \VS{21}at the head
of the noisy
streets she calls,
\par }{\Q in the entrances
of the gates
in the city
she utters
her words:
\par }{\Q \VS{22}“How long
will you simpletons
love
naiveté?

\par }{\Q How long will mockers
delight
in mockery
\par }{\Q and fools
hate
knowledge?
\par }{\Q \VS{23}If only
you will respond
to my rebuke,
\par }{\Q then I
will pour out my thoughts
to you
\par }{\Q and I will make my
words
known to you.
\par }{\Q \VS{24}However, because
I called
but you refused
to listen,

\par }{\Q because I stretched
out my hand
but no
one paid attention,
\par }{\Q \VS{25}because you neglected
all
my advice,
\par }{\Q and did not
comply
with my rebuke,
\par }{\Q \VS{26}so I myself
will laugh
when disaster
strikes you,

\par }{\Q I will mock
when
what you dread
comes,
\par }{\Q \VS{27}when what you dread
comes
like a whirlwind,
\par }{\Q and disaster
strikes
you like a devastating storm,
\par }{\Q when distressing
trouble
comes
on you.
\par }{\Q \VS{28}Then
they will call
to me, but I will not
answer;
\par }{\Q they will diligently
seek me, but they will not
find me.
\par }{\Q \VS{29}Because
they hated
moral
knowledge,
\par }{\Q and did not
choose
to fear
the {\ND{Lord}},
\par }{\Q \VS{30}they did not
comply
with my advice,
\par }{\Q they spurned
all
my rebuke.
\par }{\Q \VS{31}Therefore they
will eat
from the fruit
of their way,
\par }{\Q and they will be stuffed
full
of their own counsel.
\par }{\Q \VS{32}For
the waywardness
of the
\par }{\Q simpletons
will kill
them,
\par }{\Q and the careless
ease of fools
will destroy them.
\par }{\Q \VS{33}But
the one who listens
to me will live
in security,
\par }{\Q and will be at ease
from the dread
of harm.

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}My child,
if
you receive
my words,
\par }{\Q and store
up my commands within you,
\par }{\Q \VS{2}by making your ear
attentive
to wisdom,
\par }{\Q and by turning
your heart
to understanding,
\par }{\Q \VS{3}indeed,
if
you call
out for discernment –
\par }{\Q raise
your voice
for understanding –
\par }{\Q \VS{4}if
you seek
it like silver,
\par }{\Q and search
for it like hidden treasure,
\par }{\Q \VS{5}then
you will understand
how to fear
the {\ND{Lord}},
\par }{\Q and you will discover
knowledge
about God.
\par }{\Q \VS{6}For
the {\ND{Lord}}
gives
wisdom,
\par }{\Q and from his mouth
comes knowledge
and understanding.
\par }{\Q \VS{7}He stores up
effective counsel
for the upright,
\par }{\Q and is like a shield
for those who live
with integrity,
\par }{\Q \VS{8}to guard
the paths
of the righteous
\par }{\Q and to protect
the way
of his pious ones.
\par }{\Q \VS{9}Then
you will understand
righteousness
and justice
\par }{\Q and equity
– every
good
way.
\par }{\Q \VS{10}For
wisdom
will enter
your heart,
\par }{\Q and moral
knowledge
will be attractive to you.
\par }{\Q \VS{11}Discretion
will protect
you,

\par }{\Q understanding
will guard you,
\par }{\Q \VS{12}to deliver
you from the way
of the wicked,
\par }{\Q from those
speaking
perversity,
\par }{\Q \VS{13}who leave
the upright
paths
\par }{\Q to walk
on the dark
ways,
\par }{\Q \VS{14}who delight
in doing
evil,
\par }{\Q they rejoice
in perverse
evil;
\par }{\Q \VS{15}whose
paths
are morally crooked,
\par }{\Q and who are devious
in their ways;
\par }{\Q \VS{16}to deliver
you from the adulteress,
\par }{\Q from the sexually loose
woman who
speaks flattering
words;
\par }{\Q \VS{17}who leaves
the husband
from her younger days,

\par }{\Q and forgets
her marriage covenant
made before
God.
\par }{\Q \VS{18}For
her house
sinks down
to
death,
\par }{\Q and her paths
lead to
the place of the departed spirits.
\par }{\Q \VS{19}None
who go
in to her will return,
\par }{\Q nor
will they reach
the paths
of life.
\par }{\Q \VS{20}So
you will walk
in the way
of good
people,

\par }{\Q and will keep
on the paths
of the righteous.
\par }{\Q \VS{21}For
the upright
will reside
in the land,
\par }{\Q and those with integrity
will remain in it,
\par }{\Q \VS{22}but the wicked
will be removed
from the land,
\par }{\Q and the treacherous
will be torn away
from it.


\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}My child,
do not
forget
my teaching,
\par }{\Q but let
your heart
keep my commandments,
\par }{\Q \VS{2}for
they will provide a long
and full
life,
\par }{\Q and they will add
well-being to you.
\par }{\Q \VS{3}Do not
let truth
and mercy
leave
you;
\par }{\Q bind
them around your neck,
\par }{\Q write
them on
the tablet
of your heart.
\par }{\Q \VS{4}Then you will find
favor
and good
understanding,
\par }{\Q in the sight
of God
and people.
\par }{\Q \VS{5}Trust
in the
{\ND{Lord}}
with all
your heart,
\par }{\Q and do not
rely
on your own understanding.
\par }{\Q \VS{6}Acknowledge
him in all
your ways,
\par }{\Q and he
will make your paths
straight.
\par }{\Q \VS{7}Do not
be
wise
in your own estimation;
\par }{\Q fear
the
{\ND{Lord}}
and turn
away from evil.
\par }{\Q \VS{8}This will bring healing
to your body,
\par }{\Q and refreshment
to your inner self.
\par }{\Q \VS{9}Honor
the {\ND{Lord}}
from your wealth
\par }{\Q and from the first fruits
of all
your crops;
\par }{\Q \VS{10}then your barns
will be filled
completely,
\par }{\Q and your vats
will overflow
with new wine.
\par }{\Q \VS{11}My child,
do not
despise
discipline
from the
{\ND{Lord}},
\par }{\Q and do not
loathe
his rebuke.
\par }{\Q \VS{12}For
the {\ND{Lord}}
disciplines
those he loves,
\par }{\Q just as
a father
disciplines the son
in whom he delights.
\par }{\SH Blessings of Obtaining Wisdom
\par }{\Q \VS{13}Blessed
is the one
who finds
wisdom,
\par }{\Q and the one
who obtains
understanding.
\par }{\Q \VS{14}For
her benefit
is more profitable
than silver,
\par }{\Q and her gain
is better than gold.
\par }{\Q \VS{15}She is
more precious
than rubies,
\par }{\Q and none
of the things you desire
can compare with her.
\par }{\Q \VS{16}Long
life
is in her right hand;
\par }{\Q in her left hand
are riches
and honor.
\par }{\Q \VS{17}Her
ways
are very pleasant,
\par }{\Q and all
her paths
are peaceful.
\par }{\Q \VS{18}She is like a tree
of life
to those who obtain
her,

\par }{\Q and everyone who grasps hold
of her will be blessed.
\par }{\Q \VS{19}By wisdom
the {\ND{Lord}}
laid the foundation
of the earth;
\par }{\Q he established
the heavens
by understanding.
\par }{\Q \VS{20}By his knowledge
the primordial sea
was broken
open,

\par }{\Q and the clouds
drip down
dew.
\par }{\Q \VS{21}My child,
do not
let
them escape
from your sight;
\par }{\Q safeguard
sound wisdom
and discretion.
\par }{\Q \VS{22}So they will
give life
to you,

\par }{\Q and grace
to adorn your neck.
\par }{\Q \VS{23}Then
you will walk
on your
way
with security,
\par }{\Q and you will not
stumble.
\par }{\Q \VS{24}When
you lie
down you will not
be filled
with fear;
\par }{\Q when you lie
down your sleep
will be pleasant.
\par }{\Q \VS{25}You will not
be afraid
of sudden
disaster,
\par }{\Q or when
destruction
overtakes
the wicked;
\par }{\Q \VS{26}for
the {\ND{Lord}}
will be
the source of your confidence,
\par }{\Q and he will guard
your foot
from being caught in a trap.
\par }{\SH Wisdom Demonstrated in Relationships with People
\par }{\Q \VS{27}Do not
withhold
good
from those who need
it,

\par }{\Q when
you have the ability to help.
\par }{\Q \VS{28}Do not
say
to your neighbor,
“Go! Return
tomorrow
\par }{\Q and I will give
it,” when you have
it with you at the time.
\par }{\Q \VS{29}Do not
plot
evil
against
your neighbor
\par }{\Q when
he dwells
by you unsuspectingly.
\par }{\Q \VS{30}Do not
accuse
anyone
without legitimate cause,
\par }{\Q if
he has not
treated
you wrongly.
\par }{\Q \VS{31}Do not
envy
a violent
man,
\par }{\Q and do not
choose
to imitate any
of his ways;
\par }{\Q \VS{32}for
one who goes astray
is an abomination
to the
{\ND{Lord}},
\par }{\Q but he reveals his intimate counsel
to the upright.
\par }{\Q \VS{33}The
{\ND{Lord}}’s
curse
is on the household
of the wicked,
\par }{\Q but he blesses
the home
of the righteous.
\par }{\Q \VS{34}Although
he is
scornful
to arrogant scoffers,
\par }{\Q yet he shows
favor
to the humble.
\par }{\Q \VS{35}The wise
inherit
honor,
\par }{\Q but he holds fools
up
to public
contempt.


\par }\Chap{4}{\PP \par }{\Q \VerseOne{1}Listen,
children,
to a father’s
instruction,
\par }{\Q and pay attention
so that
you may gain
discernment.
\par }{\Q \VS{2}Because
I give
you good
instruction,
\par }{\Q do not
forsake
my teaching.
\par }{\Q \VS{3}When
I was a son
to my father,
\par }{\Q a tender
only child
before
my mother,
\par }{\Q \VS{4}he taught
me, and he said
to me:
\par }{\Q “Let your heart
lay hold
of my words;
\par }{\Q keep
my commands
so that you will live.
\par }{\Q \VS{5}Acquire
wisdom,
acquire
understanding;
\par }{\Q do not
forget
and do not
turn aside
from the words
I speak.
\par }{\Q \VS{6}Do not
forsake
wisdom, and she will protect
you;
\par }{\Q love
her, and she will guard you.
\par }{\Q \VS{7}Wisdom
is supreme –
so acquire
wisdom,
\par }{\Q and whatever
you acquire,
acquire
understanding!
\par }{\Q \VS{8}Esteem
her highly and she will exalt
you;
\par }{\Q she will honor
you if
you embrace her.
\par }{\Q \VS{9}She will place
a fair
garland
on your head;
\par }{\Q she will bestow
a beautiful
crown on you.”
\par }{\Q \VS{10}Listen,
my child,
and accept
my words,
\par }{\Q so that the years
of your life
will be many.
\par }{\Q \VS{11}I will
guide
you in the way
of wisdom
\par }{\Q and I will
lead you in upright
paths.
\par }{\Q \VS{12}When you walk,
your steps
will not
be hampered,
\par }{\Q and when you
run,
you will not
stumble.
\par }{\Q \VS{13}Hold
on to instruction,
do not
let
it go;
\par }{\Q protect
it, because
it is
your life.
\par }{\Q \VS{14}Do not
enter
the path
of the wicked
\par }{\Q or
walk
in the way
of those who are evil.
\par }{\Q \VS{15}Avoid
it, do not
go on
it;
\par }{\Q turn away
from it, and go
on.
\par }{\Q \VS{16}For
they cannot
sleep
unless
they cause harm;
\par }{\Q they are robbed
of sleep
until they make someone
stumble.
\par }{\Q \VS{17}For
they eat
bread
gained from wickedness
\par }{\Q and drink
wine
obtained from violence.
\par }{\Q \VS{18}But the path
of the righteous
is like the bright
morning light,
\par }{\Q growing
brighter
and brighter until
full
day.
\par }{\Q \VS{19}The way
of the wicked
is like gloomy
darkness;

\par }{\Q they do not
know
what
causes them to stumble.
\par }{\Q \VS{20}My child,
pay attention
to my words;
\par }{\Q listen attentively
to my sayings.
\par }{\Q \VS{21}Do not
let them depart
from your sight,
\par }{\Q guard
them within
your heart;
\par }{\Q \VS{22}for
they are life
to those
who find
them
\par }{\Q and healing
to one’s entire
body.
\par }{\Q \VS{23}Guard
your heart
with all
vigilance,
\par }{\Q for
from
it are the sources
of life.
\par }{\Q \VS{24}Remove
perverse
speech from your mouth;
\par }{\Q keep devious
talk far
from
your lips.
\par }{\Q \VS{25}Let
your eyes
look directly
in front of you
\par }{\Q and let
your gaze
look straight
before you.
\par }{\Q \VS{26}Make
the path
for your feet
level,

\par }{\Q so that
all
your ways
may be established.
\par }{\Q \VS{27}Do not
turn
to the right
or to the left;
\par }{\Q turn
yourself
away from evil.


\par }\Chap{5}{\PP \par }{\Q \VerseOne{1}My child,
be attentive
to my wisdom,
\par }{\Q pay
close attention
to my understanding,
\par }{\Q \VS{2}in order to safeguard
discretion,
\par }{\Q and that your lips
may guard
knowledge.
\par }{\Q \VS{3}For
the lips
of the adulterous woman
drip
honey,
\par }{\Q and her seductive words
are smoother
than olive oil,
\par }{\Q \VS{4}but in the end
she is bitter
as wormwood,
\par }{\Q sharp
as a two-edged
sword.
\par }{\Q \VS{5}Her feet
go down
to death;
\par }{\Q her steps
lead straight
to the grave.
\par }{\Q \VS{6}Lest
she should make level
the path leading
to life,

\par }{\Q her paths are unstable
but she
does not
know it.
\par }{\Q \VS{7}So now,
children,
listen
to me;
\par }{\Q do not
turn aside
from the words
I speak.
\par }{\Q \VS{8}Keep yourself far
from her,
\par }{\Q and do
not
go near
the door
of her house,
\par }{\Q \VS{9}lest
you give
your vigor
to others
\par }{\Q and your years
to a cruel person,
\par }{\Q \VS{10}lest
strangers
devour your strength,
\par }{\Q and your labor benefit
another man’s
house.
\par }{\Q \VS{11}And at the end
of your life you will groan
\par }{\Q when your flesh
and your body
are wasted away.
\par }{\Q \VS{12}And you will say,
“How
I hated
discipline!
\par }{\Q My heart
spurned
reproof!
\par }{\Q \VS{13}For I did not
obey
my teachers
\par }{\Q and I did not
heed
my instructors.
\par }{\Q \VS{14}I almost
came to complete ruin
\par }{\Q in the midst
of the whole
congregation!”
\par }{\Q \VS{15}Drink
water
from your own cistern
\par }{\Q and running
water from your own well.
\par }{\Q \VS{16}Should your springs
be dispersed
outside,
\par }{\Q your streams
of water
in the wide plazas?
\par }{\Q \VS{17}Let them be
for yourself alone,
\par }{\Q and not
for strangers
with you.
\par }{\Q \VS{18}May
your fountain
be blessed,
\par }{\Q and may you rejoice
in your young
wife –
\par }{\Q \VS{19}a loving
doe,
a graceful
deer;
\par }{\Q may her breasts
satisfy
you at all
times,
\par }{\Q may you be captivated
by her love
always.
\par }{\Q \VS{20}But why
should you be captivated,
my son,
by an adulteress,
\par }{\Q and embrace
the bosom
of a
different woman?
\par }{\Q \VS{21}For
the ways
of a person
are in front
of the
{\ND{Lord}}’s
eyes,
\par }{\Q and the
{\ND{Lord}}
weighs
all
that person’s paths.
\par }{\Q \VS{22}The
wicked
will be captured
by his own iniquities,
\par }{\Q and he will be held
by the cords
of his own sin.
\par }{\Q \VS{23}He
will die
because
there was no
discipline;
\par }{\Q because
of the greatness
of his folly
he will reel.


\par }\Chap{6}{\PP \par }{\Q \VerseOne{1}My child,
if
you have made a pledge
for your neighbor,
\par }{\Q and have become
a guarantor
for a stranger,
\par }{\Q \VS{2}if you have been ensnared
by the words
you have uttered,
\par }{\Q and have been caught
by the words
you have spoken,
\par }{\Q \VS{3}then,
my child,
do
this
in order to deliver
yourself,

\par }{\Q because
you have fallen
into your neighbor’s
power:

\par }{\Q go,
humble yourself,
\par }{\Q and appeal firmly
to your neighbor.
\par }{\Q \VS{4}Permit
no
sleep
to your eyes
\par }{\Q or slumber
to your eyelids.
\par }{\Q \VS{5}Deliver
yourself like a gazelle
from a snare,
\par }{\Q and like a bird
from the trap
of the fowler.
\par }{\Q \VS{6}Go
to
the ant,
you sluggard;
\par }{\Q observe
its ways
and be wise!
\par }{\Q \VS{7}It has
no
commander,
\par }{\Q overseer,
or ruler,
\par }{\Q \VS{8}yet it prepares
its food
in the summer;
\par }{\Q it gathers
at the harvest
what it will eat.
\par }{\Q \VS{9}How long,
you sluggard,
will you lie
there?
\par }{\Q When
will you rise
from your sleep?
\par }{\Q \VS{10}A little
sleep,
a little
slumber,
\par }{\Q a little
folding
of the hands
to relax,
\par }{\Q \VS{11}and your poverty
will come
like a robber,
\par }{\Q and your need
like an armed
man.
\par }{\Q \VS{12}A
worthless
and wicked
person
\par }{\Q walks
around saying
perverse things;
\par }{\Q \VS{13}he winks
with his eyes,
\par }{\Q signals
with his feet,
\par }{\Q and points
with his fingers;
\par }{\Q \VS{14}he plots
evil
with perverse
thoughts in his heart,
\par }{\Q he spreads
contention
at all
times.
\par }{\Q \VS{15}Therefore,
his disaster
will come
suddenly;
\par }{\Q in an instant
he will be broken,
and there will be no
remedy.
\par }{\Q \VS{16}There are six
things
that the
{\ND{Lord}}
hates,
\par }{\Q even seven
things that are an abomination to him:
\par }{\Q \VS{17}haughty
eyes,
a lying
tongue,
\par }{\Q and hands
that shed
innocent
blood,
\par }{\Q \VS{18}a heart
that devises
wicked
plans,
\par }{\Q feet
that are swift
to run
to evil,
\par }{\Q \VS{19}a false
witness
who pours out
lies,
\par }{\Q and a person who spreads
discord
among
family members.
\par }{\Q \VS{20}My child,
guard
the commands
of your father
\par }{\Q and do not
forsake
the instruction
of your mother.
\par }{\Q \VS{21}Bind
them on
your heart
continually;
\par }{\Q fasten
them around
your neck.
\par }{\Q \VS{22}When you walk
about, they will guide
you;
\par }{\Q when you lie
down, they will watch
over
you;
\par }{\Q when you wake
up, they will talk to you.
\par }{\Q \VS{23}For
the commandments
are like a lamp,
\par }{\Q instruction
is like a light,
\par }{\Q and rebukes
of discipline
are like the road
leading to life,
\par }{\Q \VS{24}by keeping
you from the evil
woman,
\par }{\Q from the smooth
tongue
of the loose woman.
\par }{\Q \VS{25}Do not
lust in your heart
for her beauty,
\par }{\Q and do not
let her captivate you with her alluring eyes;
\par }{\Q \VS{26}for
on account
of a prostitute
one is brought down to
a loaf
of bread,
\par }{\Q but the wife
of another man
preys
on your precious
life.
\par }{\Q \VS{27}Can a man
hold
fire
against his chest
\par }{\Q without
burning
his clothes?
\par }{\Q \VS{28}Can
a man
walk
on
hot coals
\par }{\Q without
scorching
his feet?
\par }{\Q \VS{29}So
it is with the one who has sex
with
his neighbor’s
wife;
\par }{\Q no one
who touches
her will escape punishment.
\par }{\Q \VS{30}People do not
despise
a thief
when
he steals
\par }{\Q to fulfill
his need
when
he is hungry.
\par }{\Q \VS{31}Yet if he is caught
he must repay
seven times
over,
\par }{\Q he might even have to give
all
the wealth
of his house.
\par }{\Q \VS{32}A man who commits adultery
with a woman
lacks
wisdom,
\par }{\Q whoever does it destroys
his own life.
\par }{\Q \VS{33}He will be
beaten
and despised,
\par }{\Q and his reproach
will not
be wiped away;
\par }{\Q \VS{34}for
jealousy
kindles a husband’s
rage,
\par }{\Q and he will not
show mercy
when
he takes revenge.
\par }{\Q \VS{35}He will not
consider
any
compensation;
\par }{\Q he will not
be willing,
even if
you multiply
the compensation.


\par }\Chap{7}{\PP \par }{\Q \VerseOne{1}My child,
keep
my words
\par }{\Q and treasure up my commands
in your own keeping.
\par }{\Q \VS{2}Keep
my commands
so that you may live,
\par }{\Q and obey
my instruction
as your most prized possession.
\par }{\Q \VS{3}Bind
them on
your forearm;
\par }{\Q write
them on
the tablet
of your heart.
\par }{\Q \VS{4}Say
to wisdom,
“You
are my sister,”
\par }{\Q and call
understanding
a close relative,
\par }{\Q \VS{5}so that they may keep
you from the adulterous
woman,
\par }{\Q from the loose
woman who flatters
you with her words.
\par }{\Q \VS{6}For
at the window
of my house
\par }{\Q through my window
lattice
I looked out
\par }{\Q \VS{7}and I saw
among the naive
–
\par }{\Q I discerned
among the youths –
\par }{\Q a young man
who lacked
wisdom.
\par }{\Q \VS{8}He was passing
by the street
near
her corner,
\par }{\Q making
his way along the road
to her
house
\par }{\Q \VS{9}in the twilight,
the evening,
\par }{\Q in the dark
of the night.
\par }{\Q \VS{10}Suddenly
a woman
came out to meet
him!
\par }{\Q She was dressed
like a prostitute
and with secret intent.
\par }{\Q \VS{11}(She
is loud
and rebellious,
\par }{\Q she does not
remain
at home –
\par }{\Q \VS{12}at one time
outside,
at another
in the wide plazas,
\par }{\Q and by
every
corner
she lies in wait.)
\par }{\Q \VS{13}So she grabbed
him and kissed
him,
\par }{\Q and with a bold
expression
she said to him,
\par }{\Q \VS{14}“I have fresh meat at home;

\par }{\Q today
I have fulfilled
my vows!
\par }{\Q \VS{15}That is why
I came out
to meet
you,
\par }{\Q to look
for you, and I found you!
\par }{\Q \VS{16}I have spread
my bed
with elegant coverings,
\par }{\Q with richly
colored
fabric
from Egypt.
\par }{\Q \VS{17}I have perfumed
my bed
\par }{\Q with myrrh,
aloes,
and cinnamon.
\par }{\Q \VS{18}Come,
let’s drink deeply
of lovemaking
until
morning,
\par }{\Q let’s delight ourselves
with sexual intercourse.
\par }{\Q \VS{19}For
my husband
is not
at home;
\par }{\Q he has gone
on a journey
of some distance.
\par }{\Q \VS{20}He has taken
a bag
of money
with him;

\par }{\Q he will not return
until the end of the month.”
\par }{\Q \VS{21}She persuaded
him with persuasive
words;

\par }{\Q with her smooth
talk
she compelled him.
\par }{\Q \VS{22}Suddenly
he went
after
her
\par }{\Q like an ox
that goes
to
the slaughter,
\par }{\Q like a stag prancing
into
a trapper’s snare
\par }{\Q \VS{23}till
an arrow
pierces
his liver –
\par }{\Q like a bird
hurrying
into
a trap,
\par }{\Q and he does not
know
that
it will cost him his life.
\par }{\Q \VS{24}So now,
sons,
listen
to me,
\par }{\Q and pay attention
to the words
I speak.
\par }{\Q \VS{25}Do not
let your heart
turn aside
to her
ways
–
\par }{\Q do not
wander
into her pathways;
\par }{\Q \VS{26}for
she has brought
down
many
fatally
wounded,
\par }{\Q and all
those she has slain are many.
\par }{\Q \VS{27}Her house
is the way
to the grave,
\par }{\Q going down
to
the chambers
of death.


\par }\Chap{8}{\PP \par }{\Q \VerseOne{1}Does not
wisdom
call
out?
\par }{\Q Does not understanding
raise
her voice?
\par }{\Q \VS{2}At the top
of the elevated
places along the way,
\par }{\Q at the intersection of the paths
she takes her stand;
\par }{\Q \VS{3}beside
the gates
opening
into the city,
\par }{\Q at the entrance
of the doorways
she cries out:
\par }{\Q \VS{4}“To
you, O people,
I call
out,
\par }{\Q and my voice
calls to
all mankind.
\par }{\Q \VS{5}You who
are naive,
discern
wisdom!
\par }{\Q And you fools,
understand
discernment!
\par }{\Q \VS{6}Listen,
for
I will speak
excellent things,
\par }{\Q and my lips
will utter
what is right.
\par }{\Q \VS{7}For
my mouth
speaks
truth,
\par }{\Q and my lips
hate
wickedness.
\par }{\Q \VS{8}All
the words
of my mouth
are righteous;
\par }{\Q there is nothing
in them twisted or
crooked.
\par }{\Q \VS{9}All
of them are clear
to the discerning
\par }{\Q and upright
to those who find
knowledge.
\par }{\Q \VS{10}Receive
my instruction
rather than
silver,
\par }{\Q and knowledge
rather than choice
gold.
\par }{\Q \VS{11}For
wisdom
is better
than rubies,
\par }{\Q and desirable
things cannot
be compared to her.
\par }{\Q \VS{12}“I,
wisdom,
live
with prudence,
\par }{\Q and I find
knowledge
and discretion.
\par }{\Q \VS{13}The fear
of the {\ND{Lord}}
is to hate
evil;
\par }{\Q I hate
arrogant
pride
and the evil
way
\par }{\Q and perverse utterances.
\par }{\Q \VS{14}Counsel
and sound wisdom
belong to me;

\par }{\Q I
possess understanding
and might.
\par }{\Q \VS{15}Kings
reign
by means of me,
\par }{\Q and potentates
decree righteousness;
\par }{\Q \VS{16}by me princes
rule,
\par }{\Q as well as nobles
and all
righteous judges.
\par }{\Q \VS{17}I
love
those who love
me,
\par }{\Q and those who seek
me find me.
\par }{\Q \VS{18}Riches
and honor
are with
me,
\par }{\Q long-lasting
wealth
and righteousness.
\par }{\Q \VS{19}My fruit
is better
than the purest
gold,
\par }{\Q and what I produce
is better than choice
silver.
\par }{\Q \VS{20}I walk
in
the path
of righteousness,
\par }{\Q in
the pathway
of justice,
\par }{\Q \VS{21}that I may cause those who love
me to inherit
wealth,
\par }{\Q and that I may
fill
their treasuries.
\par }{\Q \VS{22}The
{\ND{Lord}}
created
me as the beginning
of his works,
\par }{\Q before his deeds of long ago.
\par }{\Q \VS{23}From eternity
I was appointed,
\par }{\Q from the beginning,
from before
the world existed.
\par }{\Q \VS{24}When there were no
deep
oceans
I was born,
\par }{\Q when there were no
springs
overflowing
with water;
\par }{\Q \VS{25}before
the mountains
were set in place –
\par }{\Q before the hills – I was born,
\par }{\Q \VS{26}before
he made
the earth
and its fields,
\par }{\Q or the beginning
of the dust
of the world.
\par }{\Q \VS{27}When he established
the heavens,
I
was there;
\par }{\Q when he marked
out the horizon
over
the face
of the deep,
\par }{\Q \VS{28}when he established
the clouds
above,
\par }{\Q when the fountains
of the deep
grew strong,
\par }{\Q \VS{29}when he gave
the sea
his decree
\par }{\Q that the waters
should not
pass
over his command,
\par }{\Q when he marked
out the foundations
of the earth,
\par }{\Q \VS{30}then
I was beside
him as a master craftsman,
\par }{\Q and I was his delight
day
by day,
\par }{\Q rejoicing
before
him at all
times,
\par }{\Q \VS{31}rejoicing
in the habitable part
of his earth,
\par }{\Q and delighting
in its people.
\par }{\Q \VS{32}“So now,
children,
listen
to me;
\par }{\Q blessed
are those who keep
my ways.
\par }{\Q \VS{33}Listen
to my instruction
so that you may be wise,
\par }{\Q and do not
neglect it.
\par }{\Q \VS{34}Blessed
is the one
who listens
to me,
\par }{\Q watching
at my doors
day
by day,
\par }{\Q waiting
beside my doorway.
\par }{\Q \VS{35}For
the one who finds
me finds
life
\par }{\Q and receives
favor
from the
{\ND{Lord}}.
\par }{\Q \VS{36}But the one who does not find me brings
harm to himself;
\par }{\Q all
who hate
me love
death.”


\par }\Chap{9}{\PP \par }{\Q \VerseOne{1}Wisdom
has built
her house;
\par }{\Q she has carved out
its seven
pillars.
\par }{\Q \VS{2}She has prepared
her meat,
she has mixed
her wine;
\par }{\Q she also
has arranged
her table.
\par }{\Q \VS{3}She has sent
out her female
servants;
\par }{\Q she calls
out on
the highest places
of the city.
\par }{\Q \VS{4}“Whoever
is naive,
let him turn
in here,”
\par }{\Q she says to those who lack
understanding.
\par }{\Q \VS{5}“Come,
eat
some of my food,
\par }{\Q and drink
some of the wine
I have mixed.
\par }{\Q \VS{6}Abandon
your foolish ways
so that you may live,
\par }{\Q and proceed
in the way
of understanding.”
\par }{\Q \VS{7}Whoever corrects
a mocker
is asking
for insult;
\par }{\Q whoever reproves
a wicked person
receives abuse.
\par }{\Q \VS{8}Do not
reprove
a mocker
or
he will hate
you;
\par }{\Q reprove
a wise
person and he will love you.
\par }{\Q \VS{9}Give
instruction to a wise person,
and he will become wiser
still;
\par }{\Q teach
a righteous
person and he will add
to his learning.
\par }{\Q \VS{10}The beginning
of wisdom
is to fear
the
{\ND{Lord}},

\par }{\Q and acknowledging
the Holy One
is understanding.
\par }{\Q \VS{11}For because
of me your days
will be many,
\par }{\Q and years
will be added
to your life.
\par }{\Q \VS{12}If
you are wise,
you are wise
to your own advantage,

\par }{\Q but if you are
a mocker,
you alone
must bear it.
\par }{\Q \VS{13}The woman
called Folly
is brash,

\par }{\Q she
is naive
and does not
know
anything.
\par }{\Q \VS{14}So she sits
at the door
of her house,
\par }{\Q on
a seat
at the highest
point of the city,
\par }{\Q \VS{15}calling out
to those who are passing
by her in the way,
\par }{\Q who go straight
on their way.
\par }{\Q \VS{16}“Whoever
is simple,
let him turn
in here,”
\par }{\Q she says to those who lack
understanding.
\par }{\Q \VS{17}“Stolen
waters
are sweet,
\par }{\Q and food
obtained in secret
is pleasant!”
\par }{\Q \VS{18}But they do not
realize
that
the dead
are there,
\par }{\Q that her guests are in the depths of the grave.


\par }\Chap{10}{\PP \par }{\Q \VerseOne{1}The Proverbs
of Solomon:
\par }{\Q A wise
child
makes a father
rejoice,
\par }{\Q but a foolish
child is a grief
to his mother.
\par }{\Q \VS{2}Treasures
gained by wickedness
do not
profit,
\par }{\Q but righteousness
delivers from mortal danger.
\par }{\Q \VS{3}The
{\ND{Lord}}
satisfies
the appetite
of the righteous,
\par }{\Q but he thwarts
the craving
of the wicked.
\par }{\Q \VS{4}The one who
is lazy
becomes
poor,

\par }{\Q but
the one who
works diligently
becomes wealthy.
\par }{\Q \VS{5}The one who gathers
crops in the summer
is a wise
son,
\par }{\Q but
the one who sleeps
during the harvest
\par }{\Q is a son
who brings shame to himself.
\par }{\Q \VS{6}Blessings
are on the head
of the righteous,
\par }{\Q but the speech
of the wicked
conceals
violence.
\par }{\Q \VS{7}The memory
of the righteous
is a blessing,
\par }{\Q but the reputation
of the wicked
will rot.
\par }{\Q \VS{8}The wise
person
accepts
instructions,
\par }{\Q but the one who speaks
foolishness
will come to ruin.
\par }{\Q \VS{9}The one who conducts
himself in integrity
will live
securely,
\par }{\Q but the one who behaves perversely
will be found out.
\par }{\Q \VS{10}The one who winks
his eye
causes
trouble,
\par }{\Q and the one who speaks
foolishness
will come to ruin.
\par }{\Q \VS{11}The teaching
of the righteous
is a fountain
of life,
\par }{\Q but
the speech
of the wicked
conceals
violence.
\par }{\Q \VS{12}Hatred
stirs up
dissension,
\par }{\Q but
love
covers
all
transgressions.
\par }{\Q \VS{13}Wisdom
is found
in the words
of the discerning
person,

\par }{\Q but the one who lacks
wisdom
will be disciplined.
\par }{\Q \VS{14}Those who are wise
store
up knowledge,
\par }{\Q but foolish
speech
leads to imminent destruction.
\par }{\Q \VS{15}The wealth
of a rich
person is like a fortified
city,
\par }{\Q but the poor
are brought to ruin
by their poverty.
\par }{\Q \VS{16}The reward
which the righteous
receive is life;
\par }{\Q the recompense
which the wicked
receive is judgment.
\par }{\Q \VS{17}The one who heeds
instruction
is on the way
to life,
\par }{\Q but the one who rejects rebuke
goes astray.
\par }{\Q \VS{18}The one who conceals
hatred
utters
lies,
\par }{\Q and the one who spreads
slander
is certainly a fool.
\par }{\Q \VS{19}When
words
abound,
transgression
is inevitable,
\par }{\Q but the one who restrains
his words
is wise.
\par }{\Q \VS{20}What
the righteous
say is like the best
silver,
\par }{\Q but what
the wicked
think
is of little value.
\par }{\Q \VS{21}The teaching
of the righteous
feeds
many,
\par }{\Q but fools
die
for lack
of wisdom.
\par }{\Q \VS{22}The blessing
from the
{\ND{Lord}}
makes a person rich,
\par }{\Q and he adds
no sorrow to it.
\par }{\Q \VS{23}Carrying out
a wicked scheme
is enjoyable
to a fool,
\par }{\Q and so is wisdom
for the one
who has discernment.
\par }{\Q \VS{24}What the wicked
fears
will come
on him;
\par }{\Q what the righteous
desire
will be granted.
\par }{\Q \VS{25}When the storm
passes
through, the wicked
are swept
away,

\par }{\Q but the righteous
are an everlasting
foundation.
\par }{\Q \VS{26}Like vinegar
to the teeth
and like smoke
to the eyes,
\par }{\Q so
is the sluggard
to those who send him.
\par }{\Q \VS{27}Fearing
the {\ND{Lord}}
prolongs
life,

\par }{\Q but the life
span
of the wicked
will be shortened.
\par }{\Q \VS{28}The hope
of the righteous
is joy,
\par }{\Q but the expectation
of the wicked
will remain unfulfilled.
\par }{\Q \VS{29}The way
of the {\ND{Lord}}
is like a stronghold
for the upright,
\par }{\Q but it is destruction
to evildoers.
\par }{\Q \VS{30}The righteous
will never
be moved,
\par }{\Q but the wicked
will not
inhabit
the land.
\par }{\Q \VS{31}The speech
of the righteous
bears
the fruit
of wisdom,
\par }{\Q but the one who speaks
perversion
will be destroyed.
\par }{\Q \VS{32}The lips
of the righteous
know
what is pleasing,
\par }{\Q but the speech
of the wicked
is perverse.


\par }\Chap{11}{\PP \par }{\Q \VerseOne{1}The
{\ND{Lord}}
abhors
dishonest
scales,
\par }{\Q but an accurate
weight is his delight.
\par }{\Q \VS{2}When
pride
comes,
then comes
disgrace,
\par }{\Q but with
humility
comes wisdom.
\par }{\Q \VS{3}The integrity
of the upright
guides
them,

\par }{\Q but the crookedness
of the unfaithful
destroys them.
\par }{\Q \VS{4}Wealth
does not
profit
in the day
of wrath,
\par }{\Q but righteousness
delivers from mortal danger.
\par }{\Q \VS{5}The righteousness
of the blameless
will make straight
their way,
\par }{\Q but the wicked
person will fall
by his own wickedness.
\par }{\Q \VS{6}The righteousness
of the upright
will deliver
them,

\par }{\Q but the faithless
will be captured
by their own desires.
\par }{\Q \VS{7}When a wicked
person
dies,
his expectation
perishes,
\par }{\Q and the hope
of his strength
perishes.
\par }{\Q \VS{8}The righteous
person is delivered
out of trouble,
\par }{\Q and the wicked
turns up
in his stead.
\par }{\Q \VS{9}With his speech the godless
person destroys
his neighbor,
\par }{\Q but by knowledge
the righteous
will be delivered.
\par }{\Q \VS{10}When the righteous
do well,
the city
rejoices;
\par }{\Q when the wicked
perish,
there is joy.
\par }{\Q \VS{11}A city is exalted
by the blessing
provided
from the upright,
\par }{\Q but it is destroyed
by the counsel
of the wicked.
\par }{\Q \VS{12}The one who denounces
his neighbor
lacks
wisdom,
\par }{\Q but the one
who has discernment
keeps silent.
\par }{\Q \VS{13}The one who goes
about slandering
others reveals
secrets,
\par }{\Q but the one
who is trustworthy
conceals
a matter.
\par }{\Q \VS{14}When there is no
guidance
a nation
falls,
\par }{\Q but there is success
in the abundance
of counselors.
\par }{\Q \VS{15}The one who puts up security
for
a stranger
will surely have trouble,
\par }{\Q but
whoever avoids
shaking hands will be secure.
\par }{\Q \VS{16}A generous
woman
gains
honor,
\par }{\Q and ruthless
men seize
wealth.
\par }{\Q \VS{17}A kind
person
benefits
himself,
\par }{\Q but a cruel
person brings himself
trouble.
\par }{\Q \VS{18}The wicked
person earns
deceitful
wages,

\par }{\Q but the one who sows
righteousness
reaps a genuine
reward.
\par }{\Q \VS{19}True
righteousness
leads to life,
\par }{\Q but the one who pursues
evil
pursues it to his own death.
\par }{\Q \VS{20}The
{\ND{Lord}}
abhors
those who are perverse
in heart,
\par }{\Q but those who are blameless
in their ways
are his delight.
\par }{\Q \VS{21}Be assured
that the evil
person will certainly be punished,

\par }{\Q but the descendants
of the righteous
will not
suffer unjust judgment.
\par }{\Q \VS{22}Like a gold
ring
in a pig’s
snout
\par }{\Q is a beautiful
woman
who rejects
discretion.
\par }{\Q \VS{23}What the righteous
desire
leads only
to good,
\par }{\Q but what the wicked
hope
for leads to wrath.
\par }{\Q \VS{24}One person is
generous
and yet grows more
wealthy,
\par }{\Q but
another withholds
more than he should
and comes to poverty.
\par }{\Q \VS{25}A generous
person
will be enriched,
\par }{\Q and the one who provides water
for others
will himself be satisfied.
\par }{\Q \VS{26}People
will curse
the one who withholds
grain,
\par }{\Q but they will praise
the one
who sells it.
\par }{\Q \VS{27}The one who diligently
seeks good
seeks
favor,
\par }{\Q but the one who searches
for evil
– it will come to him.
\par }{\Q \VS{28}The one who trusts
in his riches
will fall,
\par }{\Q but the righteous
will flourish
like a green leaf.
\par }{\Q \VS{29}The one who troubles
his family
will inherit
nothing,

\par }{\Q and the fool
will be a servant
to the wise
person.
\par }{\Q \VS{30}The fruit
of the righteous
is like a tree
producing life,
\par }{\Q and the one who wins
souls
is wise.
\par }{\Q \VS{31}If
the righteous
are recompensed
on earth,
\par }{\Q how much
more the wicked
sinner!

\par }\Chap{12}{\PP \par }{\Q \VerseOne{1}The one who loves
discipline
loves
knowledge,
\par }{\Q but the one who hates
reproof is stupid.
\par }{\Q \VS{2}A good
person obtains
favor
from the
{\ND{Lord}},
\par }{\Q but the
{\ND{Lord}}
condemns
a person
with wicked schemes.
\par }{\Q \VS{3}No
one
can be established
through wickedness,
\par }{\Q but a righteous
root
cannot
be moved.
\par }{\Q \VS{4}A noble
wife
is the crown
of her husband,
\par }{\Q but the wife who acts shamefully
is like rottenness
in his bones.
\par }{\Q \VS{5}The plans
of the righteous
are just;
\par }{\Q the counsels
of the wicked
are deceitful.
\par }{\Q \VS{6}The words
of the wicked
lie in wait
to shed innocent blood,
\par }{\Q but the words
of the upright
will deliver them.
\par }{\Q \VS{7}The wicked
are overthrown
and perish,

\par }{\Q but
the righteous
household
will stand.
\par }{\Q \VS{8}A person is praised
in accordance with
his wisdom,
\par }{\Q but the one
who has a twisted
mind
is
despised.
\par }{\Q \VS{9}Better
is a person of humble standing who nevertheless
has a servant,
\par }{\Q than
one who pretends
to be somebody important yet has no
food.
\par }{\Q \VS{10}A righteous
person cares
for the life
of his animal,
\par }{\Q but
even the most compassionate
acts of the wicked
are cruel.
\par }{\Q \VS{11}The one who works
his field
will have plenty
of food,
\par }{\Q but
whoever chases
daydreams
lacks
wisdom.
\par }{\Q \VS{12}The wicked
person desires
a stronghold,
\par }{\Q but the righteous
root
endures.
\par }{\Q \VS{13}The evil
person is ensnared by the transgression
of his speech,
\par }{\Q but the righteous
person escapes
out
of trouble.
\par }{\Q \VS{14}A person
will be satisfied
with good
from the fruit
of his words,
\par }{\Q and the work
of his hands
will be rendered to him.
\par }{\Q \VS{15}The way
of a fool
is right
in his own opinion,
\par }{\Q but the one who listens
to advice
is wise.
\par }{\Q \VS{16}A fool’s
annoyance
is known
at
once,

\par }{\Q but the prudent overlooks an insult.
\par }{\Q \VS{17}The faithful
witness
tells what is right,
\par }{\Q but a false
witness
speaks deceit.
\par }{\Q \VS{18}Speaking
recklessly
is like the thrusts
of a sword,
\par }{\Q but the words
of the wise
bring healing.
\par }{\Q \VS{19}The one who tells
the truth
will endure
forever,
\par }{\Q but the one who lies
will last only for a moment.
\par }{\Q \VS{20}Deceit
is in the heart
of those who plot
evil,
\par }{\Q but those who promote
peace
have joy.
\par }{\Q \VS{21}The righteous
do not
encounter
any
harm,
\par }{\Q but the wicked
are filled
with calamity.
\par }{\Q \VS{22}The
{\ND{Lord}}
abhors
a person who lies,
\par }{\Q but those who deal
truthfully
are his delight.
\par }{\Q \VS{23}The shrewd
person
conceals
knowledge,
\par }{\Q but foolish
people
publicize
folly.
\par }{\Q \VS{24}The diligent
person
will rule,
\par }{\Q but the slothful
will become
a slave.
\par }{\Q \VS{25}Anxiety
in a person’s
heart
weighs
him down,
\par }{\Q but an encouraging
word
brings him joy.
\par }{\Q \VS{26}The righteous
person is cautious
in his friendship,
\par }{\Q but the way
of the wicked
leads them astray.
\par }{\Q \VS{27}The lazy person
does not
roast
his prey,
\par }{\Q but personal possessions
are precious
to the diligent.
\par }{\Q \VS{28}In the path
of righteousness
there is life,
\par }{\Q but another path
leads
to death.

\par }\Chap{13}{\PP \par }{\Q \VerseOne{1}A wise
son
accepts his father’s
discipline,
\par }{\Q but a scoffer
does not
listen
to rebuke.
\par }{\Q \VS{2}From the fruit
of his speech
a person
eats
good
things,

\par }{\Q but the faithless
desire
the fruit of violence.
\par }{\Q \VS{3}The one who guards
his words
guards
his life,
\par }{\Q but whoever is talkative
will come to ruin.
\par }{\Q \VS{4}The appetite
of the sluggard
craves
but gets nothing,
\par }{\Q but the desire
of the diligent
will be abundantly satisfied.
\par }{\Q \VS{5}The righteous
person hates
anything
false,
\par }{\Q but the wicked person
acts in shameful
disgrace.
\par }{\Q \VS{6}Righteousness
guards
the one who lives with integrity,
\par }{\Q but wickedness
overthrows
the sinner.
\par }{\Q \VS{7}There is
one who pretends
to be rich
and yet has nothing;
\par }{\Q another pretends
to be poor
and yet possesses
great
wealth.
\par }{\Q \VS{8}The ransom
of a person’s
life
is his wealth,
\par }{\Q but the poor
person hears
no
threat.
\par }{\Q \VS{9}The light
of the righteous
shines brightly,

\par }{\Q but
the lamp
of the wicked
goes out.
\par }{\Q \VS{10}With pride
comes only
contention,
\par }{\Q but wisdom
is with
the well-advised.
\par }{\Q \VS{11}Wealth
gained quickly
will dwindle
away,

\par }{\Q but the one who gathers
it little by
little
will become rich.
\par }{\Q \VS{12}Hope
deferred
makes the heart
sick,
\par }{\Q but a longing
fulfilled
is like a tree
of life.
\par }{\Q \VS{13}The one who despises
instruction
will pay the penalty,
\par }{\Q but whoever esteems
instruction
will be rewarded.
\par }{\Q \VS{14}Instruction
from the wise
is like a life-giving
fountain,
\par }{\Q to turn
a person from deadly
snares.
\par }{\Q \VS{15}Keen
insight
wins
favor,
\par }{\Q but the conduct
of the unfaithful
is harsh.
\par }{\Q \VS{16}Every
shrewd
person acts
with knowledge,
\par }{\Q but a fool
displays
his folly.
\par }{\Q \VS{17}An unreliable
messenger
falls
into trouble,
\par }{\Q but a faithful
envoy
brings healing.
\par }{\Q \VS{18}The one who neglects
discipline
ends
up
in poverty
and shame,
\par }{\Q but the one who accepts
reproof
is honored.
\par }{\Q \VS{19}A desire
fulfilled is sweet
to the soul,
\par }{\Q but fools
abhor
turning away from
evil.
\par }{\Q \VS{20}The one who associates
with
the wise
grows wise,
\par }{\Q but a companion
of fools
suffers harm.
\par }{\Q \VS{21}Calamity
pursues
sinners,
\par }{\Q but prosperity
rewards
the righteous.
\par }{\Q \VS{22}A benevolent
person leaves an inheritance
for his grandchildren,
\par }{\Q but the wealth
of a sinner
is stored
up for the righteous.
\par }{\Q \VS{23}There
is abundant
food
in the field
of the poor,
\par }{\Q but it is
swept
away by injustice.
\par }{\Q \VS{24}The one who spares
his rod
hates
his child,
\par }{\Q but the one who loves
his child
is diligent
in disciplining him.
\par }{\Q \VS{25}The righteous
has enough food
to satisfy
his appetite,
\par }{\Q but the belly
of the wicked
lacks food.

\par }\Chap{14}{\PP \par }{\Q \VerseOne{1}Every wise
woman
builds
her household,
\par }{\Q but a foolish
woman
tears
it down with her own hands.
\par }{\Q \VS{2}The one who walks
in his uprightness
fears
the {\ND{Lord}},

\par }{\Q but the one who is perverted
in his ways
despises him.
\par }{\Q \VS{3}In the speech
of a fool
is a rod
for his back,

\par }{\Q but
the words
of the wise
protect them.
\par }{\Q \VS{4}Where there are no
oxen,
the feeding trough
is clean,
\par }{\Q but an abundant
harvest
is produced by strong
oxen.
\par }{\Q \VS{5}A truthful
witness
does not
lie,
\par }{\Q but a false
witness
breathes
out lies.
\par }{\Q \VS{6}The scorner
seeks
wisdom
but finds none,
\par }{\Q but understanding
is easy
for a discerning person.
\par }{\Q \VS{7}Leave
the presence
of a foolish
person,
\par }{\Q or you will not
understand
wise counsel.
\par }{\Q \VS{8}The wisdom
of the shrewd
person is to discern
his way,
\par }{\Q but the folly
of fools
is deception.
\par }{\Q \VS{9}Fools
mock
at reparation,
\par }{\Q but among
the upright
there is favor.
\par }{\Q \VS{10}The heart
knows
its own
bitterness,
\par }{\Q and with its joy
no
one else
can share.
\par }{\Q \VS{11}The household
of the wicked
will be destroyed,
\par }{\Q but the tent
of the upright
will flourish.
\par }{\Q \VS{12}There
is a way
that seems right
to a person,
\par }{\Q but its end
is the way
that leads to death.
\par }{\Q \VS{13}Even
in laughter
the heart
may ache,
\par }{\Q and the end
of joy
may be grief.
\par }{\Q \VS{14}The backslider
will be paid back from his own ways,
\par }{\Q but
a good
person
will be rewarded for his.
\par }{\Q \VS{15}A naive person
believes
everything,
\par }{\Q but the shrewd
person discerns
his steps.
\par }{\Q \VS{16}A wise
person is cautious
and turns
from evil,
\par }{\Q but a fool
throws off restraint
and is overconfident.
\par }{\Q \VS{17}A person who has a quick temper
does
foolish
things,
\par }{\Q and a person
with crafty schemes
is hated.
\par }{\Q \VS{18}The naive
inherit
folly,
\par }{\Q but the shrewd
are crowned
with knowledge.
\par }{\Q \VS{19}Those who are evil
will bow
before
those who are good,
\par }{\Q and the wicked
will bow at the gates
of the righteous.
\par }{\Q \VS{20}A poor
person is disliked
even
by his neighbors,
\par }{\Q but those who
love
the rich
are many.
\par }{\Q \VS{21}The one who despises
his neighbor
sins,
\par }{\Q but whoever is kind
to the needy
is blessed.
\par }{\Q \VS{22}Do not
those who devise
evil
go astray?
\par }{\Q But those who plan good
exhibit
faithful covenant love.
\par }{\Q \VS{23}In all
hard work
there is profit,
\par }{\Q but merely talking
about it
only
brings poverty.
\par }{\Q \VS{24}The crown
of the wise
is their riches,
\par }{\Q but the folly
of fools
is folly.
\par }{\Q \VS{25}A truthful
witness
rescues
lives,
\par }{\Q but the one who breathes
lies
brings deception.
\par }{\Q \VS{26}In the fear
of the
{\ND{Lord}}
one has strong
confidence,
\par }{\Q and it will be
a refuge
for his children.
\par }{\Q \VS{27}The fear
of the {\ND{Lord}}
is like a life-giving
fountain,
\par }{\Q to turn
people from deadly
snares.
\par }{\Q \VS{28}A king’s
glory
is the abundance
of people,
\par }{\Q but the lack
of subjects
is the ruin
of a ruler.
\par }{\Q \VS{29}The one who is slow
to anger
has great
understanding,
\par }{\Q but the one who has a quick
temper
exalts
folly.
\par }{\Q \VS{30}A tranquil
spirit revives
the body,
\par }{\Q but envy
is rottenness
to the bones.
\par }{\Q \VS{31}The one who oppresses
the poor
insults
his Creator,
\par }{\Q but whoever shows favor
to the needy
honors him.
\par }{\Q \VS{32}The wicked
will be thrown
down in his trouble,

\par }{\Q but
the righteous
have refuge
even in the threat
of death.
\par }{\Q \VS{33}Wisdom
rests
in the heart
of the discerning;
\par }{\Q it
is known
even in the heart
of fools.
\par }{\Q \VS{34}Righteousness
exalts
a nation,
\par }{\Q but sin
is a disgrace
to any people.
\par }{\Q \VS{35}The king
shows favor
to a wise
servant,
\par }{\Q but his wrath
falls on one who acts shamefully.

\par }\Chap{15}{\PP \par }{\Q \VerseOne{1}A gentle
response
turns
away anger,
\par }{\Q but a harsh word
stirs
up
wrath.
\par }{\Q \VS{2}The tongue
of the wise
treats
knowledge
correctly,

\par }{\Q but the mouth
of the fool
spouts
out folly.
\par }{\Q \VS{3}The eyes
of the {\ND{Lord}}
are in every
place,
\par }{\Q keeping watch
on those who are evil
and those who are good.
\par }{\Q \VS{4}Speech that heals
is like a life-giving
tree,
\par }{\Q but a perverse
tongue
breaks
the spirit.
\par }{\Q \VS{5}A fool
rejects
his father’s
discipline,
\par }{\Q but whoever heeds
reproof
shows good sense.
\par }{\Q \VS{6}In the house
of the righteous
is abundant
wealth,
\par }{\Q but
the income
of the wicked
brings trouble.
\par }{\Q \VS{7}The lips
of the wise
spread
knowledge,
\par }{\Q but not
so
the heart
of fools.
\par }{\Q \VS{8}The
{\ND{Lord}}
abhors
the sacrifices
of the wicked,
\par }{\Q but the prayer
of the upright
pleases him.
\par }{\Q \VS{9}The
{\ND{Lord}}
abhors
the way
of the wicked,
\par }{\Q but he loves
those who pursue
righteousness.
\par }{\Q \VS{10}Severe discipline
is for the one who abandons
the way;
\par }{\Q the one who hates
reproof
will die.
\par }{\Q \VS{11}Death
and Destruction
are before
the
{\ND{Lord}} –
\par }{\Q how
much
more the hearts
of humans!
\par }{\Q \VS{12}The scorner
does not
love
one who corrects
him;

\par }{\Q he will not
go
to
the wise.
\par }{\Q \VS{13}A joyful
heart
makes the face
cheerful,
\par }{\Q but by a painful
heart
the spirit
is broken.
\par }{\Q \VS{14}The discerning
heart
seeks
knowledge,
\par }{\Q but the mouth of fools
feeds
on folly.
\par }{\Q \VS{15}All
the days
of the afflicted
are bad,
\par }{\Q but one with a cheerful
heart
has a continual
feast.
\par }{\Q \VS{16}Better
is little
with the fear
of the {\ND{Lord}}
\par }{\Q than great
wealth
and turmoil with it.
\par }{\Q \VS{17}Better
a meal
of vegetables
where there is love
\par }{\Q than a fattened
ox
where there
is hatred.
\par }{\Q \VS{18}A quick-tempered
person
stirs up
dissension,
\par }{\Q but one who is slow
to anger
calms
a quarrel.
\par }{\Q \VS{19}The way
of the sluggard
is like a hedge
of thorns,
\par }{\Q but the path
of the upright
is like a highway.
\par }{\Q \VS{20}A wise
child
brings joy
to his father,
\par }{\Q but a foolish
person
despises
his mother.
\par }{\Q \VS{21}Folly
is a joy
to one who lacks
sense,
\par }{\Q but one
who has understanding
follows
an upright course.
\par }{\Q \VS{22}Plans
fail
when there is no
counsel,
\par }{\Q but with abundant
advisers
they are established.
\par }{\Q \VS{23}A person has joy
in giving an appropriate answer,
\par }{\Q and a word
at the right time –
how
good it is!
\par }{\Q \VS{24}The path
of life
is upward
for the wise
person,

\par }{\Q to keep him from
going downward
to Sheol.
\par }{\Q \VS{25}The
{\ND{Lord}}
tears down the house
of the proud,
\par }{\Q but he maintains
the boundaries
of the widow.
\par }{\Q \VS{26}The
{\ND{Lord}}
abhors
the plans
of the wicked,
\par }{\Q but pleasant
words
are pure.
\par }{\Q \VS{27}The one who is greedy
for gain
troubles
his household,
\par }{\Q but whoever hates
bribes
will live.
\par }{\Q \VS{28}The heart
of the righteous
considers
how to answer,
\par }{\Q but the mouth
of the wicked
pours
out evil things.
\par }{\Q \VS{29}The
{\ND{Lord}}
is far
from the wicked,
\par }{\Q but he hears
the prayer
of the righteous.
\par }{\Q \VS{30}A bright
look
brings joy
to the heart,
\par }{\Q and good
news
gives health
to the body.
\par }{\Q \VS{31}The person
who hears
the reproof
that leads to life
\par }{\Q is at home
among
the wise.
\par }{\Q \VS{32}The one who refuses
correction
despises
himself,
\par }{\Q but whoever hears
reproof
acquires
understanding.
\par }{\Q \VS{33}The fear
of the {\ND{Lord}}
provides wise
instruction,
\par }{\Q and before
honor
comes humility.

\par }\Chap{16}{\PP \par }{\Q \VerseOne{1}The intentions
of the heart
belong to a man,
\par }{\Q but the answer
of the tongue
comes from the
{\ND{Lord}}.
\par }{\Q \VS{2}All
a person’s
ways
seem
right in his own opinion,
\par }{\Q but the
{\ND{Lord}}
evaluates
the motives.
\par }{\Q \VS{3}Commit
your works
to
the {\ND{Lord}},
\par }{\Q and your plans
will be established.
\par }{\Q \VS{4}The
{\ND{Lord}}
works
everything
for its own ends –
\par }{\Q even
the wicked
for the day
of disaster.
\par }{\Q \VS{5}The
{\ND{Lord}}
abhors
every
arrogant person;

\par }{\Q rest assured
that they will not
go unpunished.
\par }{\Q \VS{6}Through loyal love
and truth
iniquity
is appeased;
\par }{\Q through fearing
the {\ND{Lord}}
one avoids
evil.
\par }{\Q \VS{7}When a person’s
ways
are pleasing
to the
{\ND{Lord}},

\par }{\Q he even
reconciles
his enemies
to himself.
\par }{\Q \VS{8}Better
to have a little
with righteousness
\par }{\Q than to have abundant
income
without
justice.
\par }{\Q \VS{9}A person
plans
his course,

\par }{\Q but
the {\ND{Lord}}
directs
his steps.
\par }{\Q \VS{10}The divine verdict
is in the words
of the king,
\par }{\Q his pronouncements
must not
act
treacherously
against justice.
\par }{\Q \VS{11}Honest
scales
and balances
are from the
{\ND{Lord}};
\par }{\Q all
the weights
in the bag
are his handiwork.
\par }{\Q \VS{12}Doing
wickedness
is an abomination
to kings,
\par }{\Q because
a throne
is established
in righteousness.
\par }{\Q \VS{13}The delight
of kings
is righteous
counsel,
\par }{\Q and they love
the one who speaks
uprightly.
\par }{\Q \VS{14}A king’s
wrath
is like a messenger
of death,
\par }{\Q but
a wise
person appeases it.
\par }{\Q \VS{15}In the light
of the king’s
face
there is life,
\par }{\Q and his favor
is like the clouds
of the spring rain.
\par }{\Q \VS{16}How
much better
it is to acquire
wisdom
than gold;
\par }{\Q to acquire
understanding
is more desirable
than silver.
\par }{\Q \VS{17}The highway
of the upright
is to turn
away from evil;
\par }{\Q the one who guards
his way
safeguards
his life.
\par }{\Q \VS{18}Pride
goes before
destruction,
\par }{\Q and a haughty
spirit
before
a fall.
\par }{\Q \VS{19}It is better
to be lowly
in spirit
with
the afflicted
\par }{\Q than to share
the spoils
with
the proud.
\par }{\Q \VS{20}The one who deals wisely
in a matter
will find
success,
\par }{\Q and blessed
is the one who trusts
in the
{\ND{Lord}}.
\par }{\Q \VS{21}The one who is wise
in heart
is called
discerning,
\par }{\Q and kind
speech
increases
persuasiveness.
\par }{\Q \VS{22}Insight
is like a life-giving
fountain
to the one who possesses
it,
\par }{\Q but folly
leads to the discipline
of fools.
\par }{\Q \VS{23}A wise
person’s heart
makes his speech
wise
\par }{\Q and it adds
persuasiveness
to
his words.
\par }{\Q \VS{24}Pleasant
words
are like a honeycomb,
\par }{\Q sweet
to the soul
and healing
to the bones.
\par }{\Q \VS{25}There
is a way
that seems right
to a person,
\par }{\Q but its end
is the way
that leads to death.
\par }{\Q \VS{26}A laborer’s
appetite
works
on his behalf,

\par }{\Q for
his hunger
urges him to work.
\par }{\Q \VS{27}A wicked
scoundrel
digs up
evil,
\par }{\Q and his slander
is like a scorching
fire.
\par }{\Q \VS{28}A perverse
person spreads
dissension,
\par }{\Q and a gossip
separates
the closest
friends.
\par }{\Q \VS{29}A violent
person
entices
his neighbor,
\par }{\Q and leads
him down a path
that is terrible.
\par }{\Q \VS{30}The one who winks
his eyes
devises
perverse things,
\par }{\Q and one who compresses
his lips
brings about
evil.
\par }{\Q \VS{31}Gray hair
is like a crown
of glory;
\par }{\Q it is attained
in the path
of righteousness.
\par }{\Q \VS{32}Better
to be slow
to anger
than to be a mighty warrior,
\par }{\Q and one who controls
his temper
is better than one who captures
a city.
\par }{\Q \VS{33}The dice
are thrown into the lap,

\par }{\Q but their every
decision
is from the
{\ND{Lord}}.

\par }\Chap{17}{\PP \par }{\Q \VerseOne{1}Better
is a dry
crust of bread
where there is quietness
\par }{\Q than a house
full
of feasting
with strife.
\par }{\Q \VS{2}A servant
who acts wisely
will rule
\par }{\Q over an heir who behaves
shamefully,
\par }{\Q and will share
the inheritance
along with the relatives.
\par }{\Q \VS{3}The crucible
is for refining silver
and the furnace
is for gold,
\par }{\Q likewise the
{\ND{Lord}}
tests
hearts.
\par }{\Q \VS{4}One who acts wickedly
pays attention
to
evil
counsel;
\par }{\Q a liar
listens
to
a malicious
tongue.
\par }{\Q \VS{5}The one who mocks
the poor
insults
his Creator;
\par }{\Q whoever rejoices
over disaster
will not
go unpunished.
\par }{\Q \VS{6}Grandchildren
are like a crown
to the elderly,
\par }{\Q and the glory
of children
is their parents.
\par }{\Q \VS{7}Excessive
speech
is not
becoming
for a fool;
\par }{\Q how much
less are lies
for a ruler!
\par }{\Q \VS{8}A bribe
works like a charm
for the one
who offers it;

\par }{\Q in whatever
he does
he succeeds.
\par }{\Q \VS{9}The one
who forgives
an offense
seeks
love,
\par }{\Q but whoever repeats
a matter
separates
close friends.
\par }{\Q \VS{10}A rebuke makes
a greater impression
on a discerning
person
\par }{\Q than a hundred
blows
on a fool.
\par }{\Q \VS{11}An evil
person seeks
only
rebellion,
\par }{\Q and so a cruel
messenger
will be sent against him.
\par }{\Q \VS{12}It is better for a person to meet
a mother bear
being robbed of her cubs,
\par }{\Q than to encounter
a fool
in his folly.
\par }{\Q \VS{13}As
for the one who repays
evil
for
good,
\par }{\Q evil
will not
leave
his house.
\par }{\Q \VS{14}Starting
a quarrel
is like letting out
water;
\par }{\Q stop
it before
strife
breaks out!
\par }{\Q \VS{15}The one
who acquits
the guilty
and the one who condemns
the innocent –
\par }{\Q both
of them are an abomination
to the
{\ND{Lord}}.
\par }{\Q \VS{16}Of what
use is money
in the hand
of a fool,
\par }{\Q since
he has no
intention
of acquiring
wisdom?
\par }{\Q \VS{17}A friend
loves
at all
times,
\par }{\Q and a relative
is born
to help in adversity.
\par }{\Q \VS{18}The one
who lacks
wisdom
strikes
hands
in pledge,
\par }{\Q and puts up financial security for his neighbor.
\par }{\Q \VS{19}The one who loves
a quarrel
loves
transgression;
\par }{\Q whoever builds
his gate
high seeks
destruction.
\par }{\Q \VS{20}The one who has a perverse
heart
does not
find
good,
\par }{\Q and the one who is deceitful
in speech
falls
into trouble.
\par }{\Q \VS{21}Whoever brings
a fool
into the world does so to his grief,
\par }{\Q and the father
of a fool
has no
joy.
\par }{\Q \VS{22}A cheerful
heart
brings good
healing,
\par }{\Q but a crushed
spirit
dries up
the bones.
\par }{\Q \VS{23}A wicked
person receives
a bribe
secretly
\par }{\Q to pervert
the ways
of justice.
\par }{\Q \VS{24}Wisdom
is directly
in front
of the discerning
person,
\par }{\Q but the eyes
of a fool
run to the ends
of the earth.
\par }{\Q \VS{25}A foolish
child
is a grief
to his father,
\par }{\Q and bitterness
to the mother who bore him.
\par }{\Q \VS{26}It is
terrible
to punish
a righteous
person,
\par }{\Q and to flog
honorable men is wrong.
\par }{\Q \VS{27}The truly wise person
restrains
his words,
\par }{\Q and the one
who stays calm
is discerning.
\par }{\Q \VS{28}Even
a fool
who remains silent
is considered
wise,
\par }{\Q and the one who holds
his tongue
is deemed discerning.

\par }\Chap{18}{\PP \par }{\Q \VerseOne{1}One who has isolated
himself seeks
his own desires;
\par }{\Q he rejects
all
sound judgment.
\par }{\Q \VS{2}A fool
takes no
pleasure
in understanding
\par }{\Q but only
in disclosing
what is on his mind.
\par }{\Q \VS{3}When a wicked
person arrives,
contempt
shows up
with
him,
\par }{\Q and with shame
comes a reproach.
\par }{\Q \VS{4}The words
of a person’s
mouth
are like deep
waters,
\par }{\Q and the fountain
of wisdom
is like a flowing
brook.
\par }{\Q \VS{5}It
is terrible
to show partiality
to the wicked,
\par }{\Q by depriving
a righteous
man of justice.
\par }{\Q \VS{6}The lips
of a fool
enter
into strife,
\par }{\Q and his mouth
invites
a flogging.
\par }{\Q \VS{7}The mouth
of a fool
is his ruin,
\par }{\Q and his lips
are a snare
for his life.
\par }{\Q \VS{8}The words
of a gossip
are like choice morsels;
\par }{\Q they
go down
into the person’s innermost being.
\par }{\Q \VS{9}The one who is slack
in his work
\par }{\Q is a brother
to one who
destroys.
\par }{\Q \VS{10}The name
of the
{\ND{Lord}}
is like a strong
tower;
\par }{\Q the righteous
person runs
to it and is set safely on high.
\par }{\Q \VS{11}The wealth
of a rich
person is like
a strong
city,
\par }{\Q and it is like
a high
wall
in his imagination.
\par }{\Q \VS{12}Before
destruction
the heart
of a person
is proud,
\par }{\Q but humility
comes before
honor.
\par }{\Q \VS{13}The one who gives an answer
before
he listens –
\par }{\Q that is
his folly
and his shame.
\par }{\Q \VS{14}A person’s
spirit
sustains
him through sickness
–
\par }{\Q but who can
bear
a crushed
spirit?
\par }{\Q \VS{15}The discerning
person
acquires
knowledge,
\par }{\Q and the wise
person seeks
knowledge.
\par }{\Q \VS{16}A person’s
gift
makes room
for him,
\par }{\Q and leads
him before
important people.
\par }{\Q \VS{17}The first
to state his case
seems right,
\par }{\Q until his opponent begins
to cross-examine him.
\par }{\Q \VS{18}A toss of a coin
ends
disputes,
\par }{\Q and settles the issue
between
strong opponents.
\par }{\Q \VS{19}A relative
offended is harder
to reach than a strong
city,
\par }{\Q and disputes
are like the barred
gates of a fortified citadel.
\par }{\Q \VS{20}From the fruit
of a person’s
mouth
his stomach
is satisfied,
\par }{\Q with the product
of his lips
is he satisfied.
\par }{\Q \VS{21}Death
and life
are in the power
of the tongue,
\par }{\Q and those who love
its use will eat
its fruit.
\par }{\Q \VS{22}The one who finds
a wife
finds
what is enjoyable,
\par }{\Q and receives
a pleasurable
gift from the
{\ND{Lord}}.
\par }{\Q \VS{23}A poor person makes
supplications,
\par }{\Q but
a rich
man answers
harshly.
\par }{\Q \VS{24}A person
who has friends
may be harmed
by them,

\par }{\Q but there
is a friend
who sticks
closer than a brother.

\par }\Chap{19}{\PP \par }{\Q \VerseOne{1}Better
is a poor person
who walks
in his integrity
\par }{\Q than one who is perverse
in his speech
and is a fool.
\par }{\Q \VS{2}It is dangerous
to have zeal
without
knowledge,
\par }{\Q and the one who acts hastily
makes poor choices.
\par }{\Q \VS{3}A person’s
folly
subverts
his way,
\par }{\Q and his heart
rages
against
the {\ND{Lord}}.
\par }{\Q \VS{4}Wealth
adds
many
friends,
\par }{\Q but a poor
person is separated
from his friend.
\par }{\Q \VS{5}A false
witness
will not
go unpunished,
\par }{\Q and the one who spouts
out lies
will not
escape punishment.
\par }{\Q \VS{6}Many
people
entreat
the favor
of a generous
person,

\par }{\Q and everyone
is the friend
of the person
who gives gifts.
\par }{\Q \VS{7}All
the relatives
of a poor
person hate
him;

\par }{\Q how much
more do his friends
avoid him –
\par }{\Q he pursues them with words, but they do not respond.
\par }{\Q \VS{8}The one who acquires
wisdom
loves
himself;
\par }{\Q the one who preserves
understanding
will prosper.
\par }{\Q \VS{9}A false
witness
will not
go unpunished,
\par }{\Q and the one who spouts out
lies
will perish.
\par }{\Q \VS{10}Luxury
is not
appropriate
for a fool;
\par }{\Q how much
less for
a servant
to rule
over princes!
\par }{\Q \VS{11}A person’s
wisdom
makes him slow
to anger,
\par }{\Q and it is his glory
to overlook
an offense.
\par }{\Q \VS{12}A king’s
wrath
is like the roar
of a lion,
\par }{\Q but
his favor
is like dew
on
the grass.
\par }{\Q \VS{13}A foolish
child
is the ruin of his father,
\par }{\Q and a contentious
wife
is like a constant
dripping.
\par }{\Q \VS{14}A house
and wealth
are inherited
from parents,
\par }{\Q but a prudent
wife
is from the
{\ND{Lord}}.
\par }{\Q \VS{15}Laziness
brings on
a deep sleep,
\par }{\Q and the idle
person
will go hungry.
\par }{\Q \VS{16}The one who obeys
commandments
guards
his life;
\par }{\Q the one who despises
his ways
will die.
\par }{\Q \VS{17}The one who is gracious
to the poor
lends
to the
{\ND{Lord}},
\par }{\Q and the
{\ND{Lord}} will repay
him for his good deed.
\par }{\Q \VS{18}Discipline
your child,
for
there is
hope,
\par }{\Q but do not
set
your heart
on
causing his death.
\par }{\Q \VS{19}A person with great
anger bears
the penalty,
\par }{\Q but if
you deliver
him from it once, you will
have to do it again.
\par }{\Q \VS{20}Listen
to advice
and receive
discipline,
\par }{\Q that you may
become wise
by the end of your life.
\par }{\Q \VS{21}There are many
plans
in a person’s
mind,
\par }{\Q but
it is
the counsel
of the
{\ND{Lord}}
which will stand.
\par }{\Q \VS{22}What is desirable
for a person
is to show loyal love,
\par }{\Q and a poor
person
is better
than a liar.
\par }{\Q \VS{23}Fearing
the {\ND{Lord}}
leads to life,
\par }{\Q and one who does so will live
satisfied;
he will not
be afflicted by
calamity.
\par }{\Q \VS{24}The sluggard
plunges
his hand
in the dish,
\par }{\Q and he will not
even
bring it back
to
his mouth!
\par }{\Q \VS{25}Flog
a scorner,
and as a result the simpleton
will learn prudence;
\par }{\Q correct
a discerning person,
and as a result he will understand
knowledge.
\par }{\Q \VS{26}The one who robs
his father
and chases away
his mother
\par }{\Q is a son
who brings shame
and disgrace.
\par }{\Q \VS{27}If you stop
listening
to instruction,
my child,
\par }{\Q you will stray
from the words
of knowledge.
\par }{\Q \VS{28}A crooked
witness
scorns
justice,
\par }{\Q and the mouth
of the wicked
devours
iniquity.
\par }{\Q \VS{29}Judgments
are prepared
for scorners,
\par }{\Q and floggings
for the backs
of fools.

\par }\Chap{20}{\PP \par }{\Q \VerseOne{1}Wine
is a mocker
and strong drink
is a brawler;
\par }{\Q whoever goes astray by
them is not
wise.
\par }{\Q \VS{2}The king’s
terrifying
anger is like the roar
of a lion;
\par }{\Q whoever provokes
him sins
against himself.
\par }{\Q \VS{3}It is an honor
for a person
to cease
from strife,
\par }{\Q but every
fool
quarrels.
\par }{\Q \VS{4}The sluggard
will not
plow
during the planting season,
\par }{\Q so at harvest
time he looks
for the crop but has nothing.
\par }{\Q \VS{5}Counsel
in a person’s
heart
is like
deep
water,
\par }{\Q but an understanding
person draws it out.
\par }{\Q \VS{6}Many
people
profess
their loyalty,
\par }{\Q but a faithful
person –
who
can find?
\par }{\Q \VS{7}The righteous
person behaves
in integrity;
\par }{\Q blessed
are his children
after him.
\par }{\Q \VS{8}A king
sitting
on
the throne
to judge
\par }{\Q separates
out all
evil
with his eyes.
\par }{\Q \VS{9}Who can
say, “I have kept
my heart
clean;
\par }{\Q I am pure from my sin”?
\par }{\Q \VS{10}Diverse weights
and diverse measures –
\par }{\Q the {\ND{Lord}}
abhors
both of them.
\par }{\Q \VS{11}Even
a young man
is known
by his actions,
\par }{\Q whether
his activity
is pure
and whether
it is right.
\par }{\Q \VS{12}The ear
that hears
and the eye
that sees –
\par }{\Q the {\ND{Lord}}
has made
them both.
\par }{\Q \VS{13}Do not
love
sleep,
lest
you become impoverished;
\par }{\Q open
your eyes
so that you might be satisfied
with food.
\par }{\Q \VS{14}“It’s worthless! It’s worthless!” says
the buyer,
\par }{\Q but when he goes
on his way, he boasts.
\par }{\Q \VS{15}There
is gold,
and an abundance
of rubies,
\par }{\Q but
words
of knowledge
are like a precious
jewel.
\par }{\Q \VS{16}Take
a man’s garment
when
he has given security
for
a stranger,
\par }{\Q and when he gives surety
for strangers,
hold him in pledge.
\par }{\Q \VS{17}Bread
gained by deceit
tastes sweet
to a person,
\par }{\Q but afterward
his mouth
will be filled
with gravel.
\par }{\Q \VS{18}Plans
are established
by counsel,
\par }{\Q so make
war
with guidance.
\par }{\Q \VS{19}The one who goes
about gossiping
reveals
secrets;
\par }{\Q therefore do not
associate
with someone who is always opening his mouth.
\par }{\Q \VS{20}The one who curses
his father
and his mother,
\par }{\Q his lamp
will be extinguished
in the blackest
darkness.
\par }{\Q \VS{21}An inheritance
gained easily
in the beginning
\par }{\Q will not
be blessed
in the end.
\par }{\Q \VS{22}Do not
say, “I will pay back
evil!”
\par }{\Q Wait
for the
{\ND{Lord}},
so that he may vindicate you.
\par }{\Q \VS{23}The
{\ND{Lord}}
abhors
differing
weights,
\par }{\Q and dishonest
scales are wicked.
\par }{\Q \VS{24}The steps
of a person
are ordained by the
{\ND{Lord}} –
\par }{\Q so how
can anyone understand
his own way?
\par }{\Q \VS{25}It is a snare
for a person
to rashly
cry, “Holy!”
\par }{\Q and only afterward
to consider
what he has vowed.
\par }{\Q \VS{26}A wise
king
separates
out the wicked;
\par }{\Q he turns
the threshing wheel
over them.
\par }{\Q \VS{27}The human
spirit
is like the lamp
of the {\ND{Lord}},
\par }{\Q searching
all
his innermost parts.
\par }{\Q \VS{28}Loyal love
and truth
preserve
a king,
\par }{\Q and his throne
is upheld
by loyal love.
\par }{\Q \VS{29}The glory
of young men
is their strength,
\par }{\Q and the splendor
of old men
is gray hair.
\par }{\Q \VS{30}Beatings
and wounds
cleanse away
evil,
\par }{\Q and floggings
cleanse the innermost being.

\par }\Chap{21}{\PP \par }{\Q \VerseOne{1}The king’s
heart
is in the hand
of the {\ND{Lord}}
like channels
of water;
\par }{\Q he turns it wherever
he wants.
\par }{\Q \VS{2}All
of a person’s
ways
seem right
in his own opinion,
\par }{\Q but the
{\ND{Lord}}
evaluates
the motives.
\par }{\Q \VS{3}To do
righteousness
and justice
\par }{\Q is more acceptable
to the
{\ND{Lord}}
than sacrifice.
\par }{\Q \VS{4}Haughty
eyes
and a proud
heart
–
\par }{\Q the agricultural product
of the wicked
is sin.
\par }{\Q \VS{5}The plans
of the diligent
lead only
to plenty,
\par }{\Q but everyone
who is hasty
comes only
to poverty.
\par }{\Q \VS{6}Making
a fortune
by
a lying
tongue
is like a vapor
driven
back and forth;

\par }{\Q they seek
death.
\par }{\Q \VS{7}The violence
done by the wicked
will drag
them away
\par }{\Q because
they refuse
to do
what is right.
\par }{\Q \VS{8}The way
of the guilty
person
is devious,

\par }{\Q but as for the pure,
his way is upright.
\par }{\Q \VS{9}It is better
to live
on
a corner
of the housetop
\par }{\Q than in a house
in company
with a quarrelsome
wife.
\par }{\Q \VS{10}The appetite
of the wicked
desires
evil;
\par }{\Q his neighbor
is shown no
favor
in his eyes.
\par }{\Q \VS{11}When
a scorner
is punished,
the naive
becomes wise;
\par }{\Q when
a wise
person is instructed,
he gains
knowledge.
\par }{\Q \VS{12}The Righteous
One considers
the house
of the wicked;
\par }{\Q he overthrows
the wicked
to their ruin.
\par }{\Q \VS{13}The one who shuts
his ears
to the cry
of the poor,
\par }{\Q he too
will cry out
and will not
be answered.
\par }{\Q \VS{14}A gift
given in secret
subdues
anger,
\par }{\Q and a bribe
given secretly
subdues strong
wrath.
\par }{\Q \VS{15}Doing
justice
brings
joy
to the righteous
\par }{\Q and terror
to those who do
evil.
\par }{\Q \VS{16}The one
who wanders
from the way
of wisdom
\par }{\Q will end up in the company
of the departed.
\par }{\Q \VS{17}The one who
loves
pleasure
will be a poor person;

\par }{\Q whoever
loves
wine
and anointing oil
will not
be rich.
\par }{\Q \VS{18}The wicked
become a ransom
for the righteous,
\par }{\Q and the faithless
are taken in the place
of the upright.
\par }{\Q \VS{19}It is better
to live
in a desert
land
\par }{\Q than with a quarrelsome
and easily-provoked
woman.
\par }{\Q \VS{20}There is desirable
treasure
and olive oil
in the dwelling
of the wise,
\par }{\Q but a foolish
person
devours all he has.
\par }{\Q \VS{21}The one who pursues
righteousness
and love
\par }{\Q finds
life,
bounty, and honor.
\par }{\Q \VS{22}The wise
person can scale
the city
of the mighty
\par }{\Q and bring down
the stronghold
in which they trust.
\par }{\Q \VS{23}The one who guards
his mouth
and his tongue
\par }{\Q keeps
his life
from troubles.
\par }{\Q \VS{24}A proud
and arrogant
person, whose name
is “Scoffer,”
\par }{\Q acts
with overbearing
pride.
\par }{\Q \VS{25}What the sluggard
desires
will kill
him,

\par }{\Q for
his hands
refuse to work.
\par }{\Q \VS{26}All
day
long he craves
greedily,
\par }{\Q but the righteous
gives
and does not
hold back.
\par }{\Q \VS{27}The wicked
person’s sacrifice
is an abomination;
\par }{\Q how much
more when
he brings
it with evil intent!
\par }{\Q \VS{28}A lying
witness
will perish,
\par }{\Q but the one
who reports accurately
speaks
forever.
\par }{\Q \VS{29}A wicked
person
shows boldness
with his face,
\par }{\Q but
as for the upright,
he
discerns
his ways.
\par }{\Q \VS{30}There is no
wisdom
and there is no
understanding,
\par }{\Q and there is no
counsel
against
the
{\ND{Lord}}.
\par }{\Q \VS{31}A horse
is prepared
for the day
of battle,
\par }{\Q but the victory
is from the
{\ND{Lord}}.

\par }\Chap{22}{\PP \par }{\Q \VerseOne{1}A good name
is to be chosen
rather than great
wealth,
\par }{\Q good
favor
more than silver
or gold.
\par }{\Q \VS{2}The rich
and the poor
meet
together;

\par }{\Q the {\ND{Lord}}
is the creator of them both.
\par }{\Q \VS{3}A shrewd
person sees
danger
and hides
himself,
\par }{\Q but the naive
keep right on going
and suffer for it.
\par }{\Q \VS{4}The reward
for humility
and fearing
the {\ND{Lord}} ‘span class=”s 06239” title=”06239” ’
\par }{\Q is riches
and honor
and life.
\par }{\PP \par }{\Q \VS{5}Thorns
and snares
are in the path
of the perverse,
\par }{\Q but the one who guards
himself
keeps far
from them.
\par }{\Q \VS{6}Train
a child
in the way
that
he should go,

\par }{\Q and when he is old
he will not
turn
from it.
\par }{\Q \VS{7}The rich
rule over
the poor,
\par }{\Q and the borrower
is servant
to the lender.
\par }{\Q \VS{8}The one who sows
iniquity
will reap
trouble,
\par }{\Q and the rod
of his fury
will end.
\par }{\Q \VS{9}A generous
person will be blessed,
\par }{\Q for
he gives
some of his food
to the poor.
\par }{\Q \VS{10}Drive
out
the scorner
and contention
will
leave;
\par }{\Q strife
and insults will cease.
\par }{\Q \VS{11}The one who loves
a pure
heart

\par }{\Q and whose speech
is gracious –
the king
will be his friend.
\par }{\Q \VS{12}The eyes
of the {\ND{Lord}}
guard
knowledge,
\par }{\Q but he overthrows
the words
of the faithless person.
\par }{\Q \VS{13}The sluggard
says, “There is a lion
outside!
\par }{\Q I will be killed
in the middle
of the streets!”
\par }{\Q \VS{14}The mouth
of an adulteress
is like a deep
pit;
\par }{\Q the one against whom the
{\ND{Lord}}
is angry
will
fall into it.
\par }{\Q \VS{15}Folly
is bound up
in the heart
of a child,
\par }{\Q but the rod
of discipline
will drive it far
from him.
\par }{\Q \VS{16}The one who oppresses
the poor
to increase
his own gain
\par }{\Q and the one who gives
to the rich –
both end up only
in poverty.
\par }{\SH The Sayings of the Wise
\par }{\Q \VS{17}Incline
your ear
and listen
to the words
of the wise,
\par }{\Q and apply
your heart
to my instruction.
\par }{\Q \VS{18}For
it is pleasing
if
you keep
these sayings within you,
\par }{\Q and they are ready
on
your lips.
\par }{\Q \VS{19}So that your confidence
may be
in the
{\ND{Lord}},
\par }{\Q I am making them known
to you today –
even
you.
\par }{\Q \VS{20}Have I not
written
thirty sayings
for you,
\par }{\Q sayings
of counsel and knowledge,
\par }{\Q \VS{21}to show
you true
and reliable
words,
\par }{\Q so that you may give
accurate
answers
to those who sent you?
\par }{\Q \VS{22}Do not
exploit
a poor
person because
he is
poor
\par }{\Q and do not
crush
the needy
in court,
\par }{\Q \VS{23}for
the
{\ND{Lord}}
will plead
their case
\par }{\Q and will rob those who are robbing
them.
\par }{\Q \VS{24}Do not
make friends
with
an angry
person,
\par }{\Q and do not
associate
with
a wrathful
person,
\par }{\Q \VS{25}lest
you learn
his ways
\par }{\Q and entangle
yourself
in a snare.
\par }{\Q \VS{26}Do not
be
one who strikes
hands
in pledge
\par }{\Q or who puts up security for debts.
\par }{\Q \VS{27}If
you do
not
have enough
to pay,
\par }{\Q your bed
will be taken
right out from under you!
\par }{\Q \VS{28}Do not
move
an ancient
boundary
stone

\par }{\Q which
was
put in place by your ancestors.
\par }{\Q \VS{29}Do you see
a person
skilled
in his work?
\par }{\Q He will take his position
before
kings;
\par }{\Q he will not
take his position
before
obscure people.

\par }\Chap{23}{\PP \par }{\Q \VerseOne{1}When
you sit
down to eat
with
a ruler,
\par }{\Q consider
carefully
what
is before you,
\par }{\Q \VS{2}and put
a knife
to your throat
\par }{\Q if
you
possess
a large appetite.
\par }{\Q \VS{3}Do not
crave
that ruler’s delicacies,
\par }{\Q for that
food
is deceptive.
\par }{\Q \VS{4}Do not
wear
yourself out
to become rich;
\par }{\Q be wise
enough to restrain yourself.
\par }{\Q \VS{5}When you gaze
upon riches,
they are gone,
\par }{\Q for they surely
make
wings
for themselves,
\par }{\Q and fly
off into the sky
like an eagle!
\par }{\Q \VS{6}Do not
eat
the food
of a stingy
person,

\par }{\Q do not
crave
his delicacies;
\par }{\Q \VS{7}for
he is like
someone calculating
the cost in his mind.

\par }{\Q “Eat
and drink,”
he says
to you,
\par }{\Q but his heart
is not
with you;
\par }{\Q \VS{8}you will vomit
up the little bit
you have eaten,
\par }{\Q and will have wasted
your pleasant
words.
\par }{\Q \VS{9}Do not
speak
in the ears
of a fool,
\par }{\Q for
he will despise
the wisdom
of your words.
\par }{\Q \VS{10}Do not
move
an ancient
boundary
stone,
\par }{\Q or
take over
the fields
of the fatherless,
\par }{\Q \VS{11}for
their Protector
is strong;
\par }{\Q he
will plead
their case
against you.
\par }{\Q \VS{12}Apply
your heart
to instruction
\par }{\Q and your ears
to the words
of knowledge.
\par }{\Q \VS{13}Do not
withhold
discipline
from a child;
\par }{\Q even if
you strike
him with the rod,
he will not
die.
\par }{\Q \VS{14}If you
strike
him with the rod,
\par }{\Q you will deliver
him from death.
\par }{\Q \VS{15}My child,
if
your heart
is wise,
\par }{\Q then my heart
also
will be glad;
\par }{\Q \VS{16}my soul
will rejoice
\par }{\Q when your lips
speak
what is right.
\par }{\Q \VS{17}Do not
let your heart
envy
sinners,
\par }{\Q but rather
be zealous in fearing
the {\ND{Lord}}
all
the time.
\par }{\Q \VS{18}For
surely
there is
a future,
\par }{\Q and your hope
will not
be cut off.
\par }{\Q \VS{19}Listen,
my child,
and be wise,
\par }{\Q and guide
your heart
on the right way.
\par }{\Q \VS{20}Do not
spend time among drunkards,
\par }{\Q among those who eat too much
meat,
\par }{\Q \VS{21}because
drunkards
and gluttons
become impoverished,
\par }{\Q and drowsiness
clothes
them with rags.
\par }{\Q \VS{22}Listen
to your father
who begot
you,
\par }{\Q and do not
despise
your mother
when
she is old.
\par }{\Q \VS{23}Acquire
truth
and do not
sell it –
\par }{\Q wisdom, and discipline, and understanding.
\par }{\Q \VS{24}The father
of a righteous
person will rejoice
greatly;

\par }{\Q whoever
fathers
a wise
child
will have joy
in him.
\par }{\Q \VS{25}May your father
and your mother
have joy;
\par }{\Q may
she who bore
you rejoice.
\par }{\Q \VS{26}Give
me your heart,
my son,
\par }{\Q and let your eyes
observe
my ways;
\par }{\Q \VS{27}for
a prostitute is like a deep
pit;
\par }{\Q a harlot
is like a narrow
well.
\par }{\Q \VS{28}Indeed,
she
lies in wait
like a robber,
\par }{\Q and increases
the unfaithful
among men.
\par }{\Q \VS{29}Who
has woe? Who
has sorrow?
\par }{\Q Who
has contentions? Who
has complaints?
\par }{\Q Who
has wounds
without cause? Who
has dullness
of the eyes?
\par }{\Q \VS{30}Those who linger
over
wine,
\par }{\Q those who go
looking for
mixed wine.
\par }{\Q \VS{31}Do not
look
on the wine
when
it is red,
\par }{\Q when
it sparkles
in the cup,
\par }{\Q when it goes down
smoothly.
\par }{\Q \VS{32}Afterward
it bites
like a snake,
\par }{\Q and stings
like a viper.
\par }{\Q \VS{33}Your eyes
will see
strange
things,

\par }{\Q and your mind
will speak
perverse things.
\par }{\Q \VS{34}And you will be
like one who lies down
in the midst
of the sea,
\par }{\Q and like one who lies down on
the top
of the rigging.
\par }{\Q \VS{35}You will say, “They
have struck
me, but
I am
not
harmed!
\par }{\Q They
beat me, but
I did not
know
it!

\par }{\Q When
will I awake? I will look
for another drink.”

\par }\Chap{24}{\PP \par }{\Q \VerseOne{1}Do not
envy
evil
people,
\par }{\Q do not
desire
to be
with them;
\par }{\Q \VS{2}for
their hearts
contemplate
violence,
\par }{\Q and their lips
speak
harm.
\par }{\Q \VS{3}By wisdom
a house
is built,
\par }{\Q and through understanding
it is established;
\par }{\Q \VS{4}by knowledge
its rooms
are filled
\par }{\Q with all
kinds
of precious
and pleasing treasures.
\par }{\Q \VS{5}A wise
warrior
is strong,
\par }{\Q and a man
of knowledge
makes his strength
stronger;
\par }{\Q \VS{6}for
with guidance
you wage
your war,
\par }{\Q and with numerous
advisers
there is victory.
\par }{\Q \VS{7}Wisdom is unattainable
for a fool;
\par }{\Q in court
he does not
open
his mouth.
\par }{\Q \VS{8}The one who plans
to do evil
\par }{\Q will be called
a scheming person.
\par }{\Q \VS{9}A foolish
scheme
is sin,
\par }{\Q and the scorner
is an abomination
to people.
\par }{\Q \VS{10}If you faint
in the day
of trouble,
\par }{\Q your strength is small!
\par }{\Q \VS{11}Deliver
those being taken
away to death,
\par }{\Q and hold back
those slipping to the slaughter.
\par }{\Q \VS{12}If
you say,
“But we
did not
know
about this,”
\par }{\Q does not
the one who evaluates
hearts
consider?
\par }{\Q Does not the one who guards
your life
know?
\par }{\Q Will he not repay
each person
according to his deeds?
\par }{\Q \VS{13}Eat
honey,
my child,
for
it is good,
\par }{\Q and honey from the honeycomb
is sweet
to your taste.
\par }{\Q \VS{14}Likewise,
know
that wisdom
is sweet to your soul;
\par }{\Q if
you find
it, you will have a future,
\par }{\Q and your hope
will not
be cut off.
\par }{\Q \VS{15}Do not
lie in wait
like the wicked
against the place
where the righteous
live;
\par }{\Q do not
assault
his home.
\par }{\Q \VS{16}Although
a righteous
person may fall
seven
times, he gets up again,
\par }{\Q but the wicked
will be brought down
by calamity.
\par }{\Q \VS{17}Do not
rejoice
when your enemy
falls,
\par }{\Q and when he stumbles
do not
let
your heart
rejoice,
\par }{\Q \VS{18}lest
the
{\ND{Lord}}
see
it, and be displeased,
\par }{\Q and turn
his wrath
away from him.
\par }{\Q \VS{19}Do not
fret
because of evil
people
\par }{\Q or
be envious
of wicked people,
\par }{\Q \VS{20}for
the evil
person has no
future,
\par }{\Q and the lamp
of the wicked
will be extinguished.
\par }{\Q \VS{21}Fear
the {\ND{Lord}}, my child,
as well as the king,
\par }{\Q and do not
associate
with
rebels,
\par }{\Q \VS{22}for
suddenly
their destruction
will overtake them,

\par }{\Q and who
knows
the ruinous
judgment both
the
{\ND{Lord}} and the king can bring?
\par }{\SH Further Sayings of the Wise
\par }{\Q \VS{23}These
sayings also
are from the wise:
\par }{\Q To show partiality
in judgment
is terrible:
\par }{\Q \VS{24}The one who says
to the guilty, “You
are innocent,”
\par }{\Q peoples
will curse
him, and nations
will denounce him.
\par }{\Q \VS{25}But
there will be delight for those who convict
the guilty,

\par }{\Q and a pleasing blessing
will come
on them.
\par }{\Q \VS{26}Like a kiss
on the lips
\par }{\Q is the one who gives an honest
answer.
\par }{\Q \VS{27}Establish
your work
outside
and get your fields
ready;
\par }{\Q afterward
build
your house.
\par }{\Q \VS{28}Do not
be
a witness
against your neighbor
without cause,
\par }{\Q and do not deceive
with your words.
\par }{\Q \VS{29}Do not
say,
“I will do
to him just as
he has done
to me;
\par }{\Q I
will pay him back
according
to what he has done.”
\par }{\Q \VS{30}I passed
by the field
of a sluggard,
\par }{\Q by
the vineyard
of one
who lacks
wisdom.
\par }{\Q \VS{31}I saw
that thorns
had grown up
all
over it,
\par }{\Q the ground was covered
with weeds,
\par }{\Q and its stone
wall
was broken down.
\par }{\Q \VS{32}When
I saw this,
I
gave careful consideration
to it;

\par }{\Q I received
instruction
from what
I saw:
\par }{\Q \VS{33}“A little
sleep,
a little
slumber,
\par }{\Q a little
folding
of the hands
to relax,
\par }{\Q \VS{34}and your poverty
will come
like a bandit,
\par }{\Q and your need
like an armed robber.”

\par }\Chap{25}{\PP \par }{\Q \VerseOne{1}These
also
are proverbs
of Solomon,
\par }{\Q which
the men
of King
Hezekiah
of Judah
copied:
\par }{\Q \VS{2}It is the glory
of God
to conceal
a matter,
\par }{\Q and it is the glory
of a king
to search
out a matter.
\par }{\Q \VS{3}As the heaven
is high
and the earth
is deep
\par }{\Q so the hearts
of kings
are unsearchable.
\par }{\Q \VS{4}Remove
the dross
from the silver,
\par }{\Q and material
for the silversmith
will emerge;
\par }{\Q \VS{5}remove
the wicked
from before
the king,
\par }{\Q and his throne
will be established
in righteousness.
\par }{\Q \VS{6}Do not
honor yourself
before
the king,
\par }{\Q and do not
stand
in the place
of great men;
\par }{\Q \VS{7}for
it is better
for him to say
to you, “Come up
here,”
\par }{\Q than to put you lower
before
a prince,
\par }{\Q whom
your eyes
have seen.
\par }{\Q \VS{8}Do not
go out
hastily
to litigation,
\par }{\Q or
what
will you do
afterward
\par }{\Q when your neighbor
puts you to shame?
\par }{\Q \VS{9}When you argue
a case
with
your neighbor,
\par }{\Q do not
reveal
the secret
of another person,
\par }{\Q \VS{10}lest
the one who hears
it put you to shame
\par }{\Q and your infamy
will never
go away.
\par }{\Q \VS{11}Like apples
of gold
in settings
of silver,
\par }{\Q so is a word
skillfully
spoken.
\par }{\Q \VS{12}Like an earring
of gold
and an ornament
of fine gold,
\par }{\Q so is a wise
reprover
to the ear
of the one who listens.
\par }{\Q \VS{13}Like the cold
of snow
in the time
of harvest,
\par }{\Q so is a faithful
messenger
to those who send
him,
\par }{\Q for he refreshes
the heart
of his masters.
\par }{\Q \VS{14}Like cloudy skies and wind
that produce no
rain,
\par }{\Q so is
the one
who boasts of a gift
not given.
\par }{\Q \VS{15}Through patience
a ruler
can be persuaded,
\par }{\Q and a soft
tongue
can break
a bone.
\par }{\Q \VS{16}When you find
honey,
eat
only what is sufficient for
you,
\par }{\Q lest
you become stuffed
with it and vomit it up.
\par }{\Q \VS{17}Don’t set foot
too frequently
in your neighbor’s
house,
\par }{\Q lest
he become weary
of you and hate you.
\par }{\Q \VS{18}Like a club
or a sword
or a sharp
arrow,
\par }{\Q so is the one
who testifies
against his neighbor
as a false
witness.
\par }{\Q \VS{19}Like a bad
tooth
or a foot
out of joint,
\par }{\Q so is confidence
in an unfaithful person
at the time
of trouble.
\par }{\Q \VS{20}Like one who takes
off a garment
on a cold
day,
\par }{\Q or like vinegar
poured on
soda,
\par }{\Q so is one who sings
songs
to a heavy heart.
\par }{\Q \VS{21}If
your enemy is hungry,
give him
food
to eat,
\par }{\Q and if
he is thirsty,
give him water
to drink,
\par }{\Q \VS{22}for
you
will heap
coals
of fire on
his head,
\par }{\Q and the
{\ND{Lord}}
will reward you.
\par }{\Q \VS{23}The north
wind
brings forth
rain,
\par }{\Q and a gossiping
tongue
brings forth an angry look.
\par }{\Q \VS{24}It is better
to live
on
a corner
of the housetop
\par }{\Q than in a house
in company
with a quarrelsome
wife.
\par }{\Q \VS{25}Like cold
water
to a weary
person,
\par }{\Q so is good
news
from a distant
land.
\par }{\Q \VS{26}Like a muddied
spring
and a polluted
well,
\par }{\Q so is a righteous
person who gives
way before
the wicked.
\par }{\Q \VS{27}It is not
good
to eat
too much
honey,
\par }{\Q nor is it honorable for people to seek
their own glory.
\par }{\Q \VS{28}Like a city
that is broken down
and without
a wall,
\par }{\Q so is a person
who
cannot
control
his temper.

\par }\Chap{26}{\PP \par }{\Q \VerseOne{1}Like snow
in summer
or rain
in harvest,
\par }{\Q so
honor
is not
fitting
for a fool.
\par }{\Q \VS{2}Like a fluttering bird
or like a flying
swallow,
\par }{\Q so
a curse
without cause
does not
come to rest.
\par }{\Q \VS{3}A whip
for the horse
and a bridle
for the donkey,
\par }{\Q and a rod
for the backs
of fools!
\par }{\Q \VS{4}Do not
answer
a fool
according to his folly,
\par }{\Q lest
you
yourself
also be like him.
\par }{\Q \VS{5}Answer
a fool
according to his folly,
\par }{\Q lest
he be
wise
in his own estimation.
\par }{\Q \VS{6}Like cutting
off the feet
or drinking
violence,
\par }{\Q so
is sending
a message
by the hand
of a fool.
\par }{\Q \VS{7}Like
legs
that hang limp
from the lame,
\par }{\Q so is a proverb
in the mouth
of fools.
\par }{\Q \VS{8}Like tying
a stone
in a sling,
\par }{\Q so
is giving
honor
to a fool.
\par }{\Q \VS{9}Like a thorn
that goes
into the hand
of a drunkard,
\par }{\Q so is a proverb
in the mouth
of a fool.
\par }{\Q \VS{10}Like an archer who wounds at random,

\par }{\Q so
is the one who hires
a fool
or hires
any passer-by.
\par }{\Q \VS{11}Like a dog
that returns
to its vomit,
\par }{\Q so a fool
repeats
his folly.
\par }{\Q \VS{12}Do you see
a man
wise
in his own eyes?

\par }{\Q There is more hope
for a fool
than for him.
\par }{\Q \VS{13}The sluggard
says, “There is a lion
in the road!
\par }{\Q A lion
in the streets!”
\par }{\Q \VS{14}Like a door
that turns
on
its hinges,
\par }{\Q so a sluggard
turns on
his bed.
\par }{\Q \VS{15}The sluggard
plunges
his hand
in the dish;
\par }{\Q he is too lazy
to bring it back
to
his mouth.
\par }{\Q \VS{16}The sluggard
is wiser
in his own estimation
\par }{\Q than seven
people who respond
with good sense.
\par }{\Q \VS{17}Like one who grabs
a wild dog
by the ears,
\par }{\Q so is the person passing
by who becomes furious over
a quarrel
not his own.
\par }{\Q \VS{18}Like a madman
who shoots
\par }{\Q firebrands
and deadly
arrows,
\par }{\Q \VS{19}so
is a person
who deceives
his neighbor,
\par }{\Q and says,
“Was I
not
only joking?”
\par }{\Q \VS{20}Where there is no wood,
a fire
goes out,
\par }{\Q and where there is no
gossip,
contention
ceases.
\par }{\Q \VS{21}Like charcoal
is to burning coals,
and wood
to fire,
\par }{\Q so is a contentious
person
to kindle
strife.
\par }{\Q \VS{22}The words
of a gossip
are like delicious morsels;
\par }{\Q they
go down
into a person’s innermost being.
\par }{\Q \VS{23}Like a coating of glaze
over earthenware
\par }{\Q are
fervent
lips
with an
evil
heart.
\par }{\Q \VS{24}The one who hates
others disguises
it with his lips,
\par }{\Q but he stores up deceit
within him.
\par }{\Q \VS{25}When
he speaks
graciously,
do not
believe
him,

\par }{\Q for
there are seven
abominations within him.
\par }{\Q \VS{26}Though his hatred
may be concealed
by deceit,
\par }{\Q his evil
will be uncovered
in the assembly.
\par }{\Q \VS{27}The one who digs
a pit
will fall
into it;
\par }{\Q the one who rolls
a stone
– it will come back on him.
\par }{\Q \VS{28}A lying
tongue
hates
those crushed
by it,
\par }{\Q and a flattering
mouth
works
ruin.

\par }\Chap{27}{\PP \par }{\Q \VerseOne{1}Do not
boast
about tomorrow;
\par }{\Q for
you do not
know
what
a day
may bring forth.
\par }{\Q \VS{2}Let another
praise
you, and not
your own mouth;
\par }{\Q someone else,
and not
your own lips.
\par }{\Q \VS{3}A stone
is heavy
and sand
is weighty,
\par }{\Q but vexation
by a fool
is more burdensome
than the two of them.
\par }{\Q \VS{4}Wrath
is cruel
and anger
is overwhelming,
\par }{\Q but who
can stand
before
jealousy?
\par }{\Q \VS{5}Better
is open
rebuke
\par }{\Q than hidden
love.
\par }{\Q \VS{6}Faithful
are the wounds
of a friend,
\par }{\Q but the kisses
of an enemy
are excessive.
\par }{\Q \VS{7}The one whose appetite
is satisfied
loathes
honey,
\par }{\Q but to the hungry
mouth
every
bitter thing
is sweet.
\par }{\Q \VS{8}Like a bird
that wanders
from
its nest,
\par }{\Q so
is a person
who wanders
from his home.
\par }{\Q \VS{9}Ointment
and incense
make
the heart
rejoice,
\par }{\Q likewise the sweetness
of one’s friend
from sincere
counsel.
\par }{\Q \VS{10}Do not
forsake
your friend
and your father’s
friend,
\par }{\Q and do not
enter
your brother’s
house
in the day
of your disaster;
\par }{\Q a neighbor
nearby
is better
than a brother
far away.
\par }{\Q \VS{11}Be wise,
my son,
and make my heart
glad,
\par }{\Q so that I may answer
anyone who taunts me.
\par }{\Q \VS{12}A shrewd
person sees
danger
and hides
himself,
\par }{\Q but the naive
keep right on going
and suffer for it.
\par }{\Q \VS{13}Take
a man’s garment
when
he has given security
for
a stranger,
\par }{\Q and when he gives surety
for a stranger,
hold him in pledge.
\par }{\Q \VS{14}If
someone blesses
his neighbor
with a loud
voice
early
in the morning,
\par }{\Q it
will be counted as
a curse to him.
\par }{\Q \VS{15}A continual
dripping
on a rainy
day
\par }{\Q and a contentious
wife
are alike.
\par }{\Q \VS{16}Whoever hides
her hides
the wind
\par }{\Q or grasps
oil
with his right hand.
\par }{\Q \VS{17}As iron
sharpens
iron,
\par }{\Q so a person
sharpens
his friend.
\par }{\Q \VS{18}The one who tends
a fig tree
will eat
its fruit,
\par }{\Q and whoever
takes care
of his master
will be honored.
\par }{\Q \VS{19}As in water
the face
is reflected as a face,
\par }{\Q so
a person’s
heart
reflects the person.
\par }{\Q \VS{20}As Death
and Destruction
are never
satisfied,
\par }{\Q so the eyes
of a person
are never
satisfied.
\par }{\Q \VS{21}As the crucible
is for silver
and the furnace
is for gold,
\par }{\Q so a person
is proved by the praise he receives.
\par }{\Q \VS{22}If
you should pound
the fool
in the mortar
\par }{\Q among
the grain
with the pestle,
\par }{\Q his foolishness
would not
depart
from him.
\par }{\Q \VS{23}Pay careful attention
to the condition
of your flocks,
\par }{\Q give careful
attention
to your herds,
\par }{\Q \VS{24}for
riches
do not
last forever,
\par }{\Q nor does
a crown
last from generation
to generation.
\par }{\Q \VS{25}When the hay
is removed
and new grass
appears,
\par }{\Q and the grass
from the hills
is gathered in,
\par }{\Q \VS{26}the lambs
will be for your clothing,
\par }{\Q and the goats
will be for the price
of a field.
\par }{\Q \VS{27}And there will be enough
goat’s
milk
for your food,
\par }{\Q for the food
of your household,
\par }{\Q and for the sustenance
of your servant girls.

\par }\Chap{28}{\PP \par }{\Q \VerseOne{1}The wicked
person flees when
there is no
one pursuing,
\par }{\Q but the righteous
person is as confident
as a lion.
\par }{\Q \VS{2}When a country
is rebellious
it has many
princes,
\par }{\Q but by someone
who is discerning
and knowledgeable
order is maintained.
\par }{\Q \VS{3}A poor person
who oppresses
the weak
\par }{\Q is like a driving
rain
without
food.
\par }{\Q \VS{4}Those who forsake
the law
praise
the wicked,
\par }{\Q but those who keep
the law
contend with them.
\par }{\Q \VS{5}Evil
people
do not
understand
justice,
\par }{\Q but those who seek
the
{\ND{Lord}}
understand
it all.
\par }{\Q \VS{6}A poor person
who walks
in his integrity
is better
\par }{\Q than one who is perverse
in his ways
even though he is rich.
\par }{\Q \VS{7}The one who keeps
the law
is a discerning
child,
\par }{\Q but a companion
of gluttons
brings shame
to his parents.
\par }{\Q \VS{8}The one who increases
his wealth
by increasing
interest
\par }{\Q gathers
it for someone who is gracious
to the needy.
\par }{\Q \VS{9}The one who turns
away his ear
from hearing
the law,
\par }{\Q even
his prayer
is an abomination.
\par }{\Q \VS{10}The one who leads
the upright
astray in an evil
way
\par }{\Q will himself
fall
into his own pit,
\par }{\Q but the blameless
will inherit
what is good.
\par }{\Q \VS{11}A rich
person
is wise
in his own eyes,
\par }{\Q but a discerning
poor
person can evaluate him properly.
\par }{\Q \VS{12}When the righteous
rejoice,
great
is the glory,
\par }{\Q but when the wicked
rise
to power, people
are sought out.
\par }{\Q \VS{13}The one who covers
his transgressions
will not
prosper,
\par }{\Q but
whoever confesses
them and forsakes
them will find mercy.
\par }{\Q \VS{14}Blessed
is the one
who is always
cautious,
\par }{\Q but whoever hardens
his heart
will fall
into evil.
\par }{\Q \VS{15}Like a roaring lion
or a roving
bear,
\par }{\Q so
is a wicked
ruler
over
a poor
people.
\par }{\Q \VS{16}The prince
who is a great
oppressor
lacks
wisdom,
\par }{\Q but the one who hates
unjust gain
will prolong
his days.
\par }{\Q \VS{17}The one
who is tormented
by the murder
of another will flee
to the pit;
\par }{\Q let no
one support him.
\par }{\Q \VS{18}The one
who walks
blamelessly
will be delivered,
\par }{\Q but whoever is perverse
in his ways
will fall
at once.
\par }{\Q \VS{19}The one who
works
his land
will be satisfied
with food,
\par }{\Q but whoever chases
daydreams
will have his fill
of poverty.
\par }{\Q \VS{20}A faithful
person
will have an abundance
of blessings,
\par }{\Q but the one who hastens
to gain riches
will not
go unpunished.
\par }{\Q \VS{21}To show partiality
is terrible,
\par }{\Q for a person will transgress
over
the smallest piece
of bread.
\par }{\Q \VS{22}The stingy person
hastens
after riches
\par }{\Q and does not
know
that
poverty
will overtake him.
\par }{\Q \VS{23}The one who reproves
another
will in the end
find
more favor
\par }{\Q than the one who flatters
with the tongue.
\par }{\Q \VS{24}The one who robs
his father
and mother
and says,
“There is no
transgression,”
\par }{\Q is a companion
to the one
who destroys.
\par }{\Q \VS{25}The greedy
person
stirs
up dissension,
\par }{\Q but the one who trusts
in the
{\ND{Lord}}
will prosper.
\par }{\Q \VS{26}The one who trusts
in his own heart
is
a fool,
\par }{\Q but the one who walks
in wisdom
will escape.
\par }{\Q \VS{27}The one who gives
to the poor
will not
lack,
\par }{\Q but whoever shuts
his eyes
to them will receive many
curses.
\par }{\Q \VS{28}When the wicked
gain control,
people
hide
themselves,

\par }{\Q but when they perish,
the righteous
increase.

\par }\Chap{29}{\PP \par }{\Q \VerseOne{1}The one
who stiffens
his neck
after numerous rebukes
\par }{\Q will suddenly
be
destroyed without
remedy.
\par }{\Q \VS{2}When the righteous
become numerous,
the people
rejoice;
\par }{\Q when the wicked
rule,
the people
groan.
\par }{\Q \VS{3}The man
who loves
wisdom
brings joy
to his father,
\par }{\Q but whoever associates
with prostitutes
wastes
his wealth.
\par }{\Q \VS{4}A king
brings stability
to a land
by
justice,
\par }{\Q but one
who exacts tribute
tears it down.
\par }{\Q \VS{5}The one
who flatters
his neighbor
\par }{\Q spreads
a net
for his steps.
\par }{\Q \VS{6}In the transgression
of an evil
person there is a snare,
\par }{\Q but a righteous
person can sing
and rejoice.
\par }{\Q \VS{7}The righteous
person cares
for the legal
rights of the poor;
\par }{\Q the wicked
does not
understand
such knowledge.
\par }{\Q \VS{8}Scornful
people
inflame
a city,
\par }{\Q but those who are wise
turn away
wrath.
\par }{\Q \VS{9}If
a wise
person goes
to court with
a foolish
person,
\par }{\Q there is no
peace whether he is angry or laughs.
\par }{\Q \VS{10}Bloodthirsty people
hate
someone with integrity;
\par }{\Q as for the upright,
they seek
his life.
\par }{\Q \VS{11}A fool
lets
fly with all
his temper,
\par }{\Q but a wise
person keeps
it back.
\par }{\Q \VS{12}If a ruler
listens
to lies,
\par }{\Q all
his ministers
will be wicked.
\par }{\Q \VS{13}The poor
person
and the oppressor
have this in common:

\par }{\Q the {\ND{Lord}}
gives light
to the eyes
of them both.
\par }{\Q \VS{14}If a king
judges
the poor
in truth,
\par }{\Q his throne
will be established
forever.
\par }{\Q \VS{15}A rod
and reproof
impart
wisdom,
\par }{\Q but a child
who is unrestrained
brings shame
to his mother.
\par }{\Q \VS{16}When
the wicked
increase,
transgression
increases,
\par }{\Q but the righteous
will see
their downfall.
\par }{\Q \VS{17}Discipline
your child,
and he will give
you rest;
\par }{\Q he will bring you happiness.
\par }{\Q \VS{18}When there is no
prophetic
vision
the people
cast off restraint,

\par }{\Q but
the one who keeps
the law,
blessed is he!
\par }{\Q \VS{19}A servant
cannot
be corrected
by words,
\par }{\Q for although
he understands,
there is no
answer.
\par }{\Q \VS{20}Do you see
someone
who is hasty
in his words?

\par }{\Q There is more hope
for a fool
than for him.
\par }{\Q \VS{21}If someone pampers
his servant
from youth,
\par }{\Q he will be
a weakling
in the end.
\par }{\Q \VS{22}An
angry
person stirs
up dissension,
\par }{\Q and a wrathful
person is abounding
in transgression.
\par }{\Q \VS{23}A person’s
pride
will bring
him low,
\par }{\Q but one who has a lowly
spirit
will gain
honor.
\par }{\Q \VS{24}Whoever shares
with
a thief
is his own enemy;

\par }{\Q he hears
the oath
to testify, but
does not
talk.
\par }{\Q \VS{25}The fear
of people
becomes a snare,
\par }{\Q but whoever trusts
in the
{\ND{Lord}}
will be set on high.
\par }{\Q \VS{26}Many
people seek
the face
of a ruler,
\par }{\Q but it is from the
{\ND{Lord}}
that one
receives justice.
\par }{\Q \VS{27}An unjust person
is an abomination
to the righteous,
\par }{\Q and the one
who lives
an upright
life is an abomination
to the wicked.

\par }\Chap{30}{\PP \par }{\Q \VerseOne{1}The words
of Agur,
the son
of Jakeh;
an oracle:

\par }{\Q This man
says
to Ithiel,
to Ithiel
and to Ukal:
\par }{\Q \VS{2}Surely
I am
more brutish
than any other human being,

\par }{\Q and I
do not
have human
understanding;
\par }{\Q \VS{3}I have not
learned
wisdom,
\par }{\Q nor do I have knowledge
of the Holy One.
\par }{\Q \VS{4}Who
has ascended
into heaven,
and then descended?

\par }{\Q Who
has gathered up
the winds
in his fists?

\par }{\Q Who
has bound up
the waters
in his cloak?

\par }{\Q Who
has established
all
the ends
of the earth?

\par }{\Q What
is his name,
and what
is his son’s
name? – if you know!
\par }{\Q \VS{5}Every
word
of God
is purified;
\par }{\Q he is
like a shield
for those who take refuge in him.
\par }{\Q \VS{6}Do not
add
to his words,
\par }{\Q lest
he reprove
you, and prove you to be a liar.
\par }{\Q \VS{7}Two
things I ask
from you;

\par }{\Q do not
refuse
me
before
I die:
\par }{\Q \VS{8}Remove falsehood
and lies
far
from
me;
\par }{\Q do not
give
me poverty
or riches,
\par }{\Q feed
me with my allotted portion
of bread,
\par }{\Q \VS{9}lest
I become satisfied
and act deceptively
\par }{\Q and say,
“Who
is the
{\ND{Lord}}?”
\par }{\Q Or lest
I become poor
and steal
\par }{\Q and demean
the name
of my God.
\par }{\Q \VS{10}Do not
slander
a servant
to
his master,
\par }{\Q lest
he curse
you, and you are found guilty.
\par }{\Q \VS{11}There is a generation
who curse
their fathers
\par }{\Q and do not
bless
their mothers.
\par }{\Q \VS{12}There is a generation
who are pure
in their own eyes
\par }{\Q and yet are not
washed
from their filthiness.
\par }{\Q \VS{13}There is a generation
whose eyes
are so lofty,

\par }{\Q and whose eyelids
are lifted
up disdainfully.
\par }{\Q \VS{14}There is a generation
whose teeth
are like swords
\par }{\Q and whose molars
are like knives
\par }{\Q to devour
the poor
from the earth
\par }{\Q and the needy
from among the human race.
\par }{\Q \VS{15}The leech
has two
daughters:


\par }{\Q “Give! Give!”

\par }{\Q There are three
things
that are never
satisfied,
\par }{\Q four
that never
say,
“Enough” –
\par }{\Q \VS{16}the grave,
the barren
womb,
\par }{\Q land
that is not
satisfied
with water,
\par }{\Q and fire
that never
says,
“Enough!”
\par }{\Q \VS{17}The eye
that mocks
at a father
\par }{\Q and despises
obeying
a mother
–
\par }{\Q the ravens
of the valley
will peck
it out
\par }{\Q and the young
vultures
will eat it.
\par }{\Q \VS{18}There are three
things that are
too wonderful
for me,
\par }{\Q four
that I do not
understand:
\par }{\Q \VS{19}the way
of an eagle
in the sky,
\par }{\Q the way
of a snake
on
a rock,
\par }{\Q the way
of a ship
in the sea,
\par }{\Q and the way
of a man
with a woman.
\par }{\Q \VS{20}This
is the way
of an adulterous
woman:
\par }{\Q she eats
and wipes
her mouth
\par }{\Q and says,
“I have not
done
wrong.”
\par }{\Q \VS{21}Under
three
things the earth
trembles,
\par }{\Q and under
four
things it cannot
bear up:
\par }{\Q \VS{22}under
a servant
who becomes
king,
\par }{\Q under a fool
who is stuffed
with food,
\par }{\Q \VS{23}under
an unloved
woman who is married,
\par }{\Q and under a female servant
who dispossesses
her mistress.
\par }{\Q \VS{24}There are four
things on earth
that are
small,
\par }{\Q but they
are exceedingly
wise:
\par }{\Q \VS{25}ants
are creatures with little strength,
\par }{\Q but
they prepare
their food
in the summer;
\par }{\Q \VS{26}rock badgers
are creatures with little power,
\par }{\Q but
they make
their homes
in the crags;
\par }{\Q \VS{27}locusts
have no
king,
\par }{\Q but they all
go forward
by ranks;
\par }{\Q \VS{28}a lizard
you can catch with the hand,
\par }{\Q but it
gets
into the palaces
of the king.
\par }{\Q \VS{29}There
are three
things that are magnificent
in their step,
\par }{\Q four
things that move
about magnificently:
\par }{\Q \VS{30}a lion,
mightiest
of the beasts,
\par }{\Q who does not
retreat
from anything;
\par }{\Q \VS{31}a strutting rooster,
a male goat,
\par }{\Q and a king
with
his army around him.
\par }{\Q \VS{32}If
you have done foolishly
by exalting
yourself

\par }{\Q or if
you have planned
evil,
\par }{\Q put your hand
over your mouth!
\par }{\Q \VS{33}For
as the churning
of milk
produces
butter
\par }{\Q and as punching
the nose
produces
blood,
\par }{\Q so stirring
up anger
produces
strife.


\par }\Chap{31}{\PP \par }{\Q \VerseOne{1}The words
of King
Lemuel,
\par }{\Q an oracle
that
his mother
taught him:
\par }{\Q \VS{2}O my son,
O son
of my womb,
\par }{\Q O son
of my vows,
\par }{\Q \VS{3}Do not
give
your strength
to women,
\par }{\Q nor your ways
to that which ruins
kings.
\par }{\Q \VS{4}It is not
for kings,
O Lemuel,
\par }{\Q it is not
for kings
to drink
wine,
\par }{\Q or
for rulers
to crave strong drink,
\par }{\Q \VS{5}lest
they drink
and forget
what is decreed,
\par }{\Q and remove
from all
the poor
their legal rights.
\par }{\Q \VS{6}Give
strong drink
to the one who is perishing,
\par }{\Q and wine
to those who
are bitterly distressed;
\par }{\Q \VS{7}let them drink
and forget
their poverty,
\par }{\Q and remember
their misery
no
more.
\par }{\Q \VS{8}Open
your mouth
on
behalf of those unable to speak,
\par }{\Q for the legal rights
of all
the dying.
\par }{\Q \VS{9}Open
your mouth,
judge
in righteousness,
\par }{\Q and plead
the cause of the poor
and needy.
\par }{\SH The Wife of Noble Character
\par }{\Q \VS{10}Who
can find
a wife
of noble character?

\par }{\Q For her value
is far more
than rubies.
\par }{\Q \VS{11}The heart
of her husband
has confidence
in her,
\par }{\Q and he has no
lack
of gain.
\par }{\Q \VS{12}She brings
him good
and not
evil
\par }{\Q all
the days
of her life.
\par }{\Q \VS{13}She obtains
wool
and flax,
\par }{\Q and she is pleased
to work with her hands.
\par }{\Q \VS{14}She is like
the merchant
ships;
\par }{\Q she brings
her food
from afar.
\par }{\Q \VS{15}She
also gets up
while it is still
night,
\par }{\Q and provides
food for her household
and a portion
to her female servants.
\par }{\Q \VS{16}She considers
a field
and buys
it;
\par }{\Q from her own income
she plants
a vineyard.
\par }{\Q \VS{17}She begins
her work
vigorously,
\par }{\Q and she strengthens
her arms.
\par }{\Q \VS{18}She knows
that
her merchandise
is good,
\par }{\Q and her lamp
does not
go out
in the night.
\par }{\Q \VS{19}Her hands
take hold of the distaff,
\par }{\Q and her hands
grasp the spindle.
\par }{\Q \VS{20}She extends
her hand
to the poor,
\par }{\Q and reaches
out her hand
to the needy.
\par }{\Q \VS{21}She is not
afraid
of the snow
for her household,
\par }{\Q for
all
of her household
are clothed
with scarlet.
\par }{\Q \VS{22}She makes for herself coverlets;
\par }{\Q her clothing
is fine linen
and purple.
\par }{\Q \VS{23}Her husband
is well-known
in the city gate
\par }{\Q when he sits
with
the elders
of the land.
\par }{\Q \VS{24}She makes linen
garments and sells
them,
\par }{\Q and supplies
the merchants
with sashes.
\par }{\Q \VS{25}She is clothed
with strength
and honor,
\par }{\Q and she can laugh
at the time to come.
\par }{\Q \VS{26}She opens
her mouth
with wisdom,
\par }{\Q and loving
instruction
is on
her tongue.
\par }{\Q \VS{27}She watches
over the ways
of her household,
\par }{\Q and does not
eat
the bread
of idleness.
\par }{\Q \VS{28}Her children
rise up
and call her blessed,
\par }{\Q her husband
also praises her:
\par }{\Q \VS{29}“Many
daughters
have done
valiantly,
\par }{\Q but you
surpass
them all!”
\par }{\Q \VS{30}Charm
is deceitful
and beauty
is fleeting,
\par }{\Q but a woman
who fears
the
{\ND{Lord}}
will be praised.
\par }{\Q \VS{31}Give
her credit
for what she has accomplished,
\par }{\PP and let her works
praise
her in the city gates.
\par }