\NormalFont\ShortTitle{Psalms}
{\MT Psalm
\par }{\SH \par }{\MS Book 1 \newline  (Psalms 1-41)Psalm 1

\par }\PsalmChap{1}{\PP \par }{\Q \VerseOne{1}How blessed
is the one
who does not
follow
the advice
of the wicked,
\par }{\Q or stand
in
the pathway
with sinners,
\par }{\Q or
sit in
the assembly of scoffers!
\par }{\Q \VS{2}Instead
he finds pleasure
in obeying
the
{\ND{Lord}}’s
commands;

\par }{\Q he meditates
on his commands day
and night.
\par }{\Q \VS{3}He is like
a tree
planted
by flowing
streams;
\par }{\Q it yields
its fruit
at the proper time,
\par }{\Q and its
leaves
never
fall off.

\par }{\Q He succeeds
in everything
he attempts.
\par }{\Q \VS{4}Not
so
with the wicked!
\par }{\Q Instead
they are like wind-driven
chaff.
\par }{\Q \VS{5}For this reason
the wicked
cannot
withstand
judgment,
\par }{\Q nor can sinners
join the assembly
of the godly.
\par }{\Q \VS{6}Certainly
the {\ND{Lord}}
guards
the way
of the godly,
\par }{\Q but the way
of the wicked
ends in destruction.

\par }\PsalmChap{2}{\PP \par }{\Q \VerseOne{1}Why
do the nations
rebel?

\par }{\Q Why are the countries
devising
plots that will fail?
\par }{\Q \VS{2}The kings
of the earth
form a united
front;

\par }{\Q the rulers
collaborate
\par }{\Q against
the {\ND{Lord}}
and his anointed king.
\par }{\Q \VS{3}They say, “Let’s tear
off the shackles
they’ve
put
on us!

\par }{\Q Let’s
free ourselves from
their ropes!”
\par }{\Q \VS{4}The one enthroned in
heaven
laughs
in disgust;

\par }{\Q the Lord
taunts them.
\par }{\Q \VS{5}Then
he angrily
speaks
to
them
\par }{\Q and terrifies
them in his rage, saying,
\par }{\Q \VS{6}“I myself
have installed
my king
\par }{\Q on
Zion,
my holy
hill.”
\par }{\Q \VS{7}The king
says, “I will announce
the
{\ND{Lord’s}}
decree.
He said
to me:

\par }{\Q ‘You
are my son! This very day
I
have become your father!
\par }{\Q \VS{8}Ask
me,
\par }{\Q and I will give
you the nations
as your inheritance,
\par }{\Q the ends
of the earth
as your personal property.
\par }{\Q \VS{9}You will break
them with an iron
scepter;
\par }{\Q you will smash
them like a potter’s
jar!’ ”
\par }{\Q \VS{10}So now,
you kings,
do what is wise;
\par }{\Q you rulers
of the earth,
submit to correction!
\par }{\Q \VS{11}Serve
the {\ND{Lord}}
in fear!
\par }{\Q Repent
in terror!
\par }{\Q \VS{12}Give sincere homage!

\par }{\Q Otherwise
he will be angry,
\par }{\Q and you will die
because
of your behavior,
\par }{\Q when his anger
quickly
ignites.
\par }{\Q How blessed
are all
who take shelter in him!


\par }\PsalmChap{3}{\D 
A psalm
of David,
written when he fled
from his son
Absalom.
\par }{\Q \VerseOne{1}{\ND{Lord}}, how
numerous
are my enemies!
\par }{\Q Many
attack me.
\par }{\Q \VS{2}Many
say
about me,
\par }{\Q “God
will not
deliver
him.” (Selah)
\par }{\Q \VS{3}But you,

{\ND{Lord}}, are a shield
that protects me;
\par }{\Q you are my glory
and the one who restores me.
\par }{\Q \VS{4}To
the
{\ND{Lord}}
I cried
out,

\par }{\Q and he answered
me from his holy
hill.
(Selah)
\par }{\Q \VS{5}I
rested
and slept;
\par }{\Q I awoke,
for
the {\ND{Lord}}
protects me.
\par }{\Q \VS{6}I am not
afraid
of the multitude
of people
\par }{\Q who attack me from all directions.
\par }{\Q \VS{7}Rise up,

{\ND{Lord}}!
\par }{\Q Deliver
me, my God!
\par }{\Q Yes,
you will strike
all
my enemies
on the jaw;
\par }{\Q you will break
the teeth
of the wicked.
\par }{\Q \VS{8}The
{\ND{Lord}}
delivers;
\par }{\Q you show favor
to your people.
(Selah)


\par }\PsalmChap{4}{\D 
For the music director,
to be accompanied by stringed instruments;
a psalm
of David.
\par }{\Q \VerseOne{1}When I call
out, answer
me,
\par }{\Q O God
who vindicates me!

\par }{\Q Though I am hemmed
in, you will lead me into a wide, open place.

\par }{\Q Have mercy
on me and respond
to my prayer!
\par }{\Q \VS{2}You men,
how
long will you try to turn my honor
into shame?

\par }{\Q How long will you love
what is worthless
\par }{\Q and search
for what is deceptive? (Selah)
\par }{\Q \VS{3}Realize
that
the {\ND{Lord}}
shows
the godly
special favor;

\par }{\Q the {\ND{Lord}}
responds when
I cry
out to him.
\par }{\Q \VS{4}Tremble
with fear and do not
sin!

\par }{\Q Meditate
as you lie in bed,
and repent
of your ways! (Selah)
\par }{\Q \VS{5}Offer
the prescribed
sacrifices
\par }{\Q and trust
in the
{\ND{Lord}}!
\par }{\Q \VS{6}Many
say,
“Who can
show
us anything good?”
\par }{\Q Smile
upon us,
{\ND{Lord}}!
\par }{\Q \VS{7}You make
me happier
\par }{\Q than those who have abundant
grain
and wine.
\par }{\Q \VS{8}I will lie
down and sleep
peacefully,
\par }{\Q for
you,

{\ND{Lord}}, make me safe
and secure.


\par }\PsalmChap{5}{\D 
For the music director,
to
be accompanied by wind instruments;
a psalm
of David.
\par }{\Q \VerseOne{1}Listen
to what I say,

{\ND{Lord}}!
\par }{\Q Carefully consider
my complaint!
\par }{\Q \VS{2}Pay attention
to my cry
for help,
\par }{\Q my king
and my God,
\par }{\Q for
I am praying
to you!
\par }{\Q \VS{3}{\ND{Lord}}, in the morning
you will hear
me;

\par }{\Q in the morning
I will present
my case
to you and then wait expectantly for an answer.
\par }{\Q \VS{4}Certainly
you are not
a God
who approves
of evil;
\par }{\Q evil people
cannot
dwell
with you.
\par }{\Q \VS{5}Arrogant people
cannot
stand
in your presence;
\par }{\Q you hate
all
who behave
wickedly.
\par }{\Q \VS{6}You destroy
liars;
\par }{\Q the {\ND{Lord}}
despises
violent
and deceitful
people.
\par }{\Q \VS{7}But as for me,
because of your great
faithfulness
I will enter
your house;
\par }{\Q I will bow
down toward
your holy
temple
as I worship you.
\par }{\Q \VS{8}{\ND{Lord}}, lead
me in your righteousness
\par }{\Q because
of those who wait to ambush
me,

\par }{\Q remove the obstacles
in the way in which you are guiding me!
\par }{\Q \VS{9}For
they do
not
speak
the truth;

\par }{\Q their stomachs
are like the place of destruction,

\par }{\Q their throats
like an open
grave,
\par }{\Q their tongues
like a steep slope leading into it.
\par }{\Q \VS{10}Condemn
them, O God!
\par }{\Q May their own schemes
be their downfall!

\par }{\Q Drive
them away
because of their many
acts of insurrection,
\par }{\Q for
they have rebelled against you.
\par }{\Q \VS{11}But may all
who take shelter
in you be happy!

\par }{\Q May they continually
shout
for joy!

\par }{\Q Shelter
them so that those who are loyal
to you
may rejoice!
\par }{\Q \VS{12}Certainly
you
reward
the godly,

{\ND{Lord}}.
\par }{\Q Like a shield
you protect
them in your good favor.


\par }\PsalmChap{6}{\D 
For the music director,
to be accompanied by stringed instruments,
according to the
{\IT{
sheminith}}
style; a psalm
of David.
\par }{\Q \VerseOne{1}{\ND{Lord}}, do not
rebuke
me in your anger!
\par }{\Q Do not
discipline
me in your raging fury!
\par }{\Q \VS{2}Have mercy
on me,
{\ND{Lord}}, for
I am
frail!
\par }{\Q Heal
me,
{\ND{Lord}}, for
my bones
are shaking!
\par }{\Q \VS{3}I am absolutely
terrified,
\par }{\Q and you,

{\ND{Lord}} –
how long will this continue?
\par }{\Q \VS{4}Relent,

{\ND{Lord}}, rescue
me!

\par }{\Q Deliver
me because
of your faithfulness!
\par }{\Q \VS{5}For
no
one remembers you in the realm
of death,

\par }{\Q In Sheol
who
gives you thanks?
\par }{\Q \VS{6}I am exhausted
as I groan;
\par }{\Q all
night
long I drench
my bed
in tears;

\par }{\Q my tears
saturate
the cushion
beneath me.
\par }{\Q \VS{7}My eyes
grow dim
from suffering;
\par }{\Q they grow
weak because of all
my enemies.
\par }{\Q \VS{8}Turn
back from me,
all
you who behave
wickedly,
\par }{\Q for
the {\ND{Lord}}
has heard
the sound
of my weeping!
\par }{\Q \VS{9}The
{\ND{Lord}}
has heard
my appeal for mercy;
\par }{\Q the {\ND{Lord}}
has accepted
my prayer.
\par }{\Q \VS{10}May all
my enemies
be humiliated
and absolutely terrified!

\par }{\Q May they turn back
and be suddenly
humiliated!


\par }\PsalmChap{7}{\D 
A musical composition
by David,
which
he sang
to the
{\ND{Lord}}
concerning
a Benjaminite
named Cush.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}
my God,
in you I have taken shelter.
\par }{\Q Deliver
me from all
who chase
me! Rescue me!
\par }{\Q \VS{2}Otherwise
they will rip
me
to shreds
like a lion;
\par }{\Q they will tear
me to bits
and no one
will be able to rescue me.
\par }{\Q \VS{3}O
{\ND{Lord}}
my God,
if
I have done
what they say,

\par }{\Q or
am guilty
of unjust actions,
\par }{\Q \VS{4}or
have wronged
my ally,
\par }{\Q or helped
his lawless
enemy,
\par }{\Q \VS{5}may an enemy
relentlessly
chase
me and catch
me;

\par }{\Q may he trample
me to death

\par }{\Q and leave me lying dishonored
in the dust.
(Selah)
\par }{\Q \VS{6}Stand
up angrily,

{\ND{Lord}}!
\par }{\Q Rise
up with raging fury
against my enemies!

\par }{\Q Wake
up for my sake and execute
the judgment
you have decreed for them!
\par }{\Q \VS{7}The countries
are assembled
all around
you;

\par }{\Q take once
more your rightful place over them!
\par }{\Q \VS{8}The
{\ND{Lord}}
judges
the nations.
\par }{\Q Vindicate
me,
{\ND{Lord}}, because I am innocent,
\par }{\Q because I am blameless, O Exalted One!
\par }{\Q \VS{9}May
the evil deeds
of the wicked
come to an end!

\par }{\Q But make the innocent
secure,
\par }{\Q O righteous
God,
\par }{\Q you who examine
inner thoughts
and motives!
\par }{\Q \VS{10}The Exalted God
is my
shield,
\par }{\Q the one who delivers
the morally
upright.
\par }{\Q \VS{11}God
is a just
judge;
\par }{\Q he
is
angry
throughout
the day.
\par }{\Q \VS{12}If
a person does not
repent,
God sharpens his sword
\par }{\Q and prepares
to shoot his bow.
\par }{\Q \VS{13}He prepares
to use deadly
weapons
against him;

\par }{\Q he gets
ready to shoot flaming
arrows.
\par }{\Q \VS{14}See
the one who is pregnant
with wickedness,
\par }{\Q who conceives
destructive plans,
\par }{\Q and gives birth
to harmful
lies –

\par }{\Q \VS{15}he digs
a pit
\par }{\Q and then falls
into the hole
he has made.
\par }{\Q \VS{16}He becomes
the victim
of his own destructive plans
\par }{\Q and the violence
he intended for others falls
on
his own head.
\par }{\Q \VS{17}I will thank
the {\ND{Lord}}
for his justice;
\par }{\Q I will sing praises
to the sovereign

{\ND{Lord}}!


\par }\PsalmChap{8}{\D 
For the music director,
according
{\IT{to the
gittith}} style;
a psalm
of David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}},
our Lord,
\par }{\Q how
magnificent
is your reputation
throughout
the earth!
\par }{\Q You
reveal your majesty
in the heavens
above!
\par }{\Q \VS{2}From the mouths
of children
and nursing
babies
\par }{\Q you have ordained
praise on account of your
adversaries,
\par }{\Q so that you might
put an end
to the vindictive
enemy.
\par }{\Q \VS{3}When
I look
up at the heavens,
which your fingers
made,
\par }{\Q and see the moon
and the stars,
which
you set
in place,
\par }{\Q \VS{4}Of what
importance is the human
race, that
you should notice
them?
\par }{\Q Of what importance is mankind,
that
you should pay attention to them,
\par }{\Q \VS{5}and make them a little
less
than the heavenly
beings?

\par }{\Q You grant mankind
honor
and majesty;
\par }{\Q \VS{6}you appoint them to rule
over your
creation;
\par }{\Q you have placed
everything
under
their
authority,
\par }{\Q \VS{7}including all
the sheep
and cattle,
\par }{\Q as well
as the wild
animals,
\par }{\Q \VS{8}the birds
in the sky,
the fish
in the sea
\par }{\Q and everything that moves through
the currents
of the seas.
\par }{\Q \VS{9}O
{\ND{Lord}}, our Lord,
\par }{\Q how
magnificent
is your reputation
throughout
the earth!


\par }\PsalmChap{9}{\D 
For the music director;
{\IT{according to the
alumoth-labben}}
style; a psalm
of David.
\par }{\Q \VerseOne{1}I will thank
the {\ND{Lord}}
with all
my heart!
\par }{\Q I will tell
about all
your amazing deeds!
\par }{\Q \VS{2}I will be happy
and rejoice
in you!
\par }{\Q I will sing praises
to you, O sovereign
One!
\par }{\Q \VS{3}When my enemies
turn
back,
\par }{\Q they trip
and are defeated
before you.
\par }{\Q \VS{4}For
you defended
my just
cause;
\par }{\Q from your throne
you pronounced
a just decision.
\par }{\Q \VS{5}You terrified
the nations
with your battle cry;

\par }{\Q you destroyed
the wicked;
\par }{\Q you permanently
wiped out
all memory of them.
\par }{\Q \VS{6}The enemy’s
cities have been reduced
to permanent
ruins;
\par }{\Q you destroyed
their cities;
\par }{\Q all memory
of the enemies
has perished.
\par }{\Q \VS{7}But the
{\ND{Lord}}
rules forever;
\par }{\Q he reigns
in
a just manner.
\par }{\Q \VS{8}He judges
the world
fairly;
\par }{\Q he
makes just legal
decisions
for the nations.
\par }{\Q \VS{9}Consequently
the {\ND{Lord}}
provides
safety
for the oppressed;

\par }{\Q he provides
safety in times
of trouble.
\par }{\Q \VS{10}Your loyal followers trust
in you,
\par }{\Q for
you,
{\ND{Lord}}, do not
abandon
those who seek your help.
\par }{\Q \VS{11}Sing praises
to the
{\ND{Lord}}, who rules
in Zion!
\par }{\Q Tell
the nations
what he has done!
\par }{\Q \VS{12}For
the one
who takes revenge
against murderers
took notice
of the oppressed;
\par }{\Q he did not
overlook
their cry for help
\par }{\Q \VS{13}when they prayed:

\par }{\Q “Have mercy
on me,
{\ND{Lord}}!
\par }{\Q See
how I am oppressed
by those who hate
me,

\par }{\Q O one who can snatch
me away
from the gates
of death!
\par }{\Q \VS{14}Then
I will
tell
about all
your praiseworthy
acts;

\par }{\Q in the gates
of Daughter
Zion
I will rejoice
because of your deliverance.”
\par }{\Q \VS{15}The nations
fell
into the pit
they had made;
\par }{\Q their feet
were caught
in the net
they had hidden.
\par }{\Q \VS{16}The
{\ND{Lord}}
revealed
himself;
\par }{\Q he accomplished
justice;
\par }{\Q the wicked
were ensnared
by their own actions. (Higgaion.
Selah)
\par }{\Q \VS{17}The wicked
are turned back
and sent to Sheol;
\par }{\Q this is the destiny of all
the nations
that ignore
God,
\par }{\Q \VS{18}for
the needy
are not
permanently
ignored,
\par }{\Q the hopes
of the oppressed
are not forever
dashed.
\par }{\Q \VS{19}Rise
up,
{\ND{Lord}}!

\par }{\Q Don’t
let men
be defiant!

\par }{\Q May
the nations
be judged
in your presence!
\par }{\Q \VS{20}Terrify
them,
{\ND{Lord}}!

\par }{\Q Let the nations
know
they
are
mere
mortals! (Selah)


\par }\PsalmChap{10}{\PP \par }{\Q \VerseOne{1}Why,

{\ND{Lord}}, do you stand
far
off?
\par }{\Q Why do you pay no attention during
times
of trouble?
\par }{\Q \VS{2}The wicked
arrogantly
chase
the oppressed;
\par }{\Q the oppressed are trapped
by the schemes
the wicked
have dreamed up.
\par }{\Q \VS{3}Yes,
the wicked
man boasts because he gets
what he wants;
\par }{\Q the one who robs
others curses
and rejects
the {\ND{Lord}}.
\par }{\Q \VS{4}The wicked
man is so arrogant
he always thinks,
\par }{\Q “God
won’t
hold me accountable;
he doesn’t care.”
\par }{\Q \VS{5}He is secure at all
times.
\par }{\Q He has no regard for your commands;

\par }{\Q he disdains
all
his enemies.
\par }{\Q \VS{6}He says
to himself,
\par }{\Q “I will never be upended,
\par }{\Q because I experience
no
calamity.”
\par }{\Q \VS{7}His mouth
is full
of curses
and deceptive,
harmful words;
\par }{\Q his tongue
injures
and destroys.
\par }{\Q \VS{8}He waits
in ambush
near the villages;
\par }{\Q in hidden places
he kills
the innocent.
\par }{\Q His eyes
look for
some unfortunate victim.
\par }{\Q \VS{9}He lies in ambush
in a hidden place,
like a lion
in a thicket;
\par }{\Q he lies in ambush,
waiting to catch
the oppressed;
\par }{\Q he catches
the oppressed
by pulling
in his net.
\par }{\Q \VS{10}His victims
are crushed
and beaten down;
\par }{\Q they are trapped
in his sturdy nets.
\par }{\Q \VS{11}He says
to himself,
\par }{\Q “God
overlooks
it;
\par }{\Q he does not
pay attention;
\par }{\Q he never notices.”
\par }{\Q \VS{12}Rise
up,
{\ND{Lord}}!

\par }{\Q O God,
strike
him down!

\par }{\Q Do not
forget
the oppressed!
\par }{\Q \VS{13}Why
does the wicked
man reject
God?

\par }{\Q He says
to himself, “You will not
hold me accountable.”
\par }{\Q \VS{14}You have taken notice,
\par }{\Q for
you
always see
one who inflicts
pain
and suffering.
\par }{\Q The unfortunate victim
entrusts
his cause
to
you;

\par }{\Q you
deliver
the fatherless.
\par }{\Q \VS{15}Break
the arm
of the wicked
and evil
man!
\par }{\Q Hold him accountable
for his wicked
deeds,

\par }{\Q which he thought you would not discover.
\par }{\Q \VS{16}The
{\ND{Lord}}
rules
forever!

\par }{\Q The nations
are driven out
of his land.
\par }{\Q \VS{17}{\ND{Lord}}, you have heard
the request
of the oppressed;
\par }{\Q you make them feel secure
because you listen to their prayer.
\par }{\Q \VS{18}You defend
the fatherless
and oppressed,

\par }{\Q so that mere mortals
may
no longer
terrorize them.


\par }\PsalmChap{11}{\D 
For the music director;
by David.
\par }{\Q \VerseOne{1}In the
{\ND{Lord}}
I have taken shelter.
\par }{\Q How
can you say
to me,
\par }{\Q “Flee
to a mountain
like a bird!
\par }{\Q \VS{2}For
look,
the wicked
prepare their
bows,
\par }{\Q they put
their arrows
on
the strings,
\par }{\Q to shoot
in the darkness
at the morally
upright.
\par }{\Q \VS{3}When
the foundations
are destroyed,
\par }{\Q what can
the godly
accomplish?”
\par }{\Q \VS{4}The
{\ND{Lord}}
is in his holy
temple;
\par }{\Q the
{\ND{Lord}}’s
throne
is in heaven.
\par }{\Q His eyes
watch;
\par }{\Q his eyes
examine
all people.
\par }{\Q \VS{5}The
{\ND{Lord}}
approves
of the godly,
\par }{\Q but he hates
the wicked
and those who love
to do violence.
\par }{\Q \VS{6}May the
{\ND{Lord}} rain down
burning coals
and brimstone
on
the wicked!
\par }{\Q A whirlwind is what they deserve!
\par }{\Q \VS{7}Certainly
the {\ND{Lord}}
is just;

\par }{\Q he rewards
godly deeds;
\par }{\Q the upright
will experience his favor.


\par }\PsalmChap{12}{\D 
For the music director;
according
{\IT{to the
sheminith}}
style; a psalm
of David.
\par }{\Q \VerseOne{1}Deliver,

{\ND{Lord}}!
\par }{\Q For
the godly
have disappeared;
\par }{\Q people
of integrity
have vanished.
\par }{\Q \VS{2}People lie
to one
another;
\par }{\Q they flatter
and deceive.
\par }{\Q \VS{3}May the
{\ND{Lord}}
cut off
all
flattering
lips,
\par }{\Q and the tongue
that boasts!
\par }{\Q \VS{4}They say, “We speak
persuasively;
\par }{\Q we know how to flatter and boast.
\par }{\Q Who
is our master?”
\par }{\Q \VS{5}“Because of the violence
done to the oppressed,
\par }{\Q because of the painful
cries of the needy,
\par }{\Q I will spring into action,”
says
the {\ND{Lord}}.
\par }{\Q “I will provide
the safety
they so desperately desire.”
\par }{\Q \VS{6}The
{\ND{Lord}}’s
words
are absolutely reliable.
\par }{\Q They are as untainted
as silver
purified
in a furnace
on the ground,
\par }{\Q where it is thoroughly
refined.
\par }{\Q \VS{7}You,

{\ND{Lord}}, will protect
them;

\par }{\Q you will continually
shelter
each one from
these evil people,
\par }{\Q \VS{8}for the wicked
seem to be everywhere,
\par }{\Q when people
promote
evil.


\par }\PsalmChap{13}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}How long,

{\ND{Lord}}, will you continue
to
ignore
me?

\par }{\Q How long will you pay no attention to me?
\par }{\Q \VS{2}How
long
must I worry,
\par }{\Q and suffer
in broad daylight?

\par }{\Q How
long
will my enemy
gloat
over me?
\par }{\Q \VS{3}Look
at me! Answer
me, O
{\ND{Lord}}
my God!
\par }{\Q Revive
me,
or
else I will die!
\par }{\Q \VS{4}Then my enemy
will say, “I have defeated
him!”
\par }{\Q Then my foes
will rejoice
because
I am
upended.
\par }{\Q \VS{5}But I
trust
in your faithfulness.
\par }{\Q May I rejoice
because of your deliverance!
\par }{\Q \VS{6}I will sing
praises to the
{\ND{Lord}}
\par }{\Q when
he vindicates me.


\par }\PsalmChap{14}{\D 
For the music director;
by David.
\par }{\Q \VerseOne{1}Fools
say
to themselves, “There is no
God.”
\par }{\Q They sin and commit
evil deeds;
\par }{\Q none
of them does what
is right.
\par }{\Q \VS{2}The
{\ND{Lord}}
looks
down from heaven
at the human
race,

\par }{\Q to see
if there is anyone
who is wise
and seeks
God.
\par }{\Q \VS{3}Everyone
rejects
God;

\par }{\Q they are all
morally corrupt.
\par }{\Q None
of them does
what is right,
\par }{\Q not
even
one!
\par }{\Q \VS{4}All
those who behave
wickedly
do not
understand –

\par }{\Q those who devour my people as if they were eating bread,
\par }{\Q and do not call out to the
{\ND{Lord}}.
\par }{\Q \VS{5}They are absolutely
terrified,
\par }{\Q for
God
defends
the godly.
\par }{\Q \VS{6}You want to humiliate
the oppressed,
\par }{\Q even though
the {\ND{Lord}}
is their shelter.
\par }{\Q \VS{7}I wish
the deliverance
of Israel
would come from Zion!
\par }{\Q When the
{\ND{Lord}}
restores the well-being
of his people,
\par }{\Q may
Jacob
rejoice,
\par }{\Q may Israel
be happy!


\par }\PsalmChap{15}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}{\ND{Lord}}, who
may be a guest
in your home?

\par }{\Q Who
may live
on your holy
hill?
\par }{\Q \VS{2}Whoever lives
a blameless
life,

\par }{\Q does what
is right,
\par }{\Q and speaks
honestly.
\par }{\Q \VS{3}He does not
slander,
\par }{\Q or
do
harm
to others,

\par }{\Q or
insult
his neighbor.
\par }{\Q \VS{4}He despises
a reprobate,
\par }{\Q but honors
the
{\ND{Lord}}’s
loyal followers.
\par }{\Q He makes firm commitments
and does not
renege
on his promise.
\par }{\Q \VS{5}He does not
charge
interest
when he lends his money.
\par }{\Q He does not
take
bribes
to testify against
the innocent.
\par }{\Q The one who lives
like this
will never
be upended.


\par }\PsalmChap{16}{\D 
A prayer
of David.
\par }{\Q \VerseOne{1}Protect
me, O God,
for
I have taken shelter in you.
\par }{\Q \VS{2}I say
to the
{\ND{Lord}}, “You are the Lord,
\par }{\Q my only source of well-being.”
\par }{\Q \VS{3}As
for God’s chosen people
who are
in the land,
\par }{\Q and the leading officials
I admired so much –
\par }{\Q \VS{4}their troubles
multiply,
\par }{\Q they desire
other
gods.

\par }{\Q I will not
pour out
drink offerings
of blood
to their gods,

\par }{\Q nor
will I make vows
in the name
of their gods.
\par }{\Q \VS{5}{\ND{Lord}}, you give me stability
and prosperity;
\par }{\Q you
make
my future secure.
\par }{\Q \VS{6}It is
as
if I have been given fertile fields
\par }{\Q or received a beautiful tract
of land.
\par }{\Q \VS{7}I will praise
the {\ND{Lord}}
who
guides
me;
\par }{\Q yes,
during the night
I reflect
and learn.
\par }{\Q \VS{8}I constantly
trust
in the
{\ND{Lord}};

\par }{\Q because
he is at my right hand,
I will not
be upended.
\par }{\Q \VS{9}So
my heart
rejoices
\par }{\Q and I am happy;
\par }{\Q My
life is safe.
\par }{\Q \VS{10}You will not
abandon
me
to Sheol;
\par }{\Q you will not
allow
your faithful follower
to see
the Pit.
\par }{\Q \VS{11}You lead
me in the path
of life;
\par }{\Q I experience absolute
joy
in your presence;
\par }{\Q you always give me sheer delight.


\par }\PsalmChap{17}{\D 
A prayer
of David.
\par }{\Q \VerseOne{1}{\ND{Lord}}, consider
my just cause!

\par }{\Q Pay attention
to my cry for help!
\par }{\Q Listen
to the prayer
\par }{\Q I sincerely
offer!
\par }{\Q \VS{2}Make
a just decision
on my behalf!

\par }{\Q Decide what is right!
\par }{\Q \VS{3}You have scrutinized
my inner motives;
\par }{\Q you have examined
me during the night.
\par }{\Q You have carefully evaluated
me, but you find
no
sin.
\par }{\Q I am determined
I will say
nothing sinful.
\par }{\Q \VS{4}As for the actions
of people –
\par }{\Q just as you have commanded,
\par }{\Q I
have not followed
in the footsteps
of violent men.
\par }{\Q \VS{5}I
carefully obey your commands;

\par }{\Q I do not
deviate from them.
\par }{\Q \VS{6}I
call
to you for
you will answer
me, O God.
\par }{\Q Listen
to me!

\par }{\Q Hear
what I say!
\par }{\Q \VS{7}Accomplish awesome,
faithful
deeds,

\par }{\Q you who powerfully
deliver
those who look to you for protection from their enemies.
\par }{\Q \VS{8}Protect me as you would protect
the pupil
of your eye!

\par }{\Q Hide
me in the shadow
of your wings!
\par }{\Q \VS{9}Protect me from
the wicked
men who
attack
me,
\par }{\Q my enemies
who crowd around me for the kill.
\par }{\Q \VS{10}They are calloused;
\par }{\Q they speak
arrogantly.
\par }{\Q \VS{11}They attack
me, now
they surround
me;

\par }{\Q they intend
to throw
me to the ground.
\par }{\Q \VS{12}He is like
a lion
that wants to tear
its prey to bits,

\par }{\Q like a young lion
crouching
in hidden places.
\par }{\Q \VS{13}Rise
up,
{\ND{Lord}}!
\par }{\Q Confront
him! Knock
him down!

\par }{\Q Use your sword
to rescue
me
from the wicked man!
\par }{\Q \VS{14}{\ND{Lord}}, use your power
to deliver me from these murderers,
\par }{\Q from the murderers
of this world!

\par }{\Q They enjoy prosperity;
\par }{\Q you overwhelm
them with the riches
they desire.

\par }{\Q They have many
children,
\par }{\Q and leave
their wealth
to their offspring.
\par }{\Q \VS{15}As for me, because I
am innocent
I will see
your face;
\par }{\Q when I awake you will reveal yourself to me.


\par }\PsalmChap{18}{\D 
For the music director;
by the
{\ND{Lord}}’s
servant
David,
who
sang
to the
{\ND{Lord}}
the
words
of this
song
when
the {\ND{Lord}}
rescued
him from the power
of all
his enemies,
including Saul.
\par }{\Q \VerseOne{1}He said:

\par }{\Q “I love
you,
{\ND{Lord}}, my source of strength!
\par }{\Q \VS{2}The
{\ND{Lord}}
is my high ridge,
my stronghold,
my deliverer.
\par }{\Q My God
is my rocky
summit where I take shelter,
\par }{\Q my shield,
the horn
that saves
me, and my refuge.
\par }{\Q \VS{3}I called
to the
{\ND{Lord}}, who is worthy of praise,
\par }{\Q and I was delivered
from
my enemies.
\par }{\Q \VS{4}The waves
of death
engulfed
me,
\par }{\Q the currents
of chaos
overwhelmed me.
\par }{\Q \VS{5}The ropes
of Sheol
tightened
around
me,

\par }{\Q the snares
of death trapped me.
\par }{\Q \VS{6}In my distress
I called
to the
{\ND{Lord}};
\par }{\Q I cried out
to
my God.
\par }{\Q From his heavenly temple
he heard
my voice;
\par }{\Q he listened to my cry
for help.
\par }{\Q \VS{7}The earth
heaved
and shook;
\par }{\Q the roots
of the mountains
trembled;
\par }{\Q they heaved
because
he was angry.
\par }{\Q \VS{8}Smoke
ascended
from his nose;
\par }{\Q fire
devoured
as it came from
his mouth;
\par }{\Q he hurled down fiery coals.
\par }{\Q \VS{9}He made
the sky
sink
as he descended;
\par }{\Q a thick cloud
was under
his feet.
\par }{\Q \VS{10}He mounted
a winged
angel and flew;
\par }{\Q he glided
on
the wings
of the wind.
\par }{\Q \VS{11}He shrouded
himself in darkness,
\par }{\Q in thick
rain
clouds.
\par }{\Q \VS{12}From the brightness
in front
of him came
\par }{\Q hail
and fiery
coals.
\par }{\Q \VS{13}The
{\ND{Lord}}
thundered
in the sky;
\par }{\Q the sovereign
One shouted.
\par }{\Q \VS{14}He shot
his arrows
and scattered
them,

\par }{\Q many
lightning
bolts
and routed them.
\par }{\Q \VS{15}The depths
of the sea were exposed;
\par }{\Q the inner regions
of the world
were uncovered
\par }{\Q by
your battle cry,

{\ND{Lord}},
\par }{\Q by
the powerful breath
from your nose.
\par }{\Q \VS{16}He reached
down from above
and took
hold of me;
\par }{\Q he pulled
me from the surging
water.
\par }{\Q \VS{17}He rescued me from my strong
enemy,
\par }{\Q from those who hate
me,
\par }{\Q for
they were too
strong for me.
\par }{\Q \VS{18}They confronted
me in my day
of calamity,
\par }{\Q but the
{\ND{Lord}}
helped me.
\par }{\Q \VS{19}He brought
me out
into a wide open place;
\par }{\Q he delivered
me because
he was pleased with me.
\par }{\Q \VS{20}The
{\ND{Lord}}
repaid
me for my godly
deeds;

\par }{\Q he rewarded
my blameless
behavior.
\par }{\Q \VS{21}For
I have obeyed
the
{\ND{Lord}}’s
commands;
\par }{\Q I have not
rebelled
against my God.
\par }{\Q \VS{22}For
I am aware
of all
his regulations,
\par }{\Q and I do not
reject
his rules.
\par }{\Q \VS{23}I was
innocent
before him,
\par }{\Q and kept
myself from sinning.
\par }{\Q \VS{24}The
{\ND{Lord}}
rewarded
me for my godly
deeds;

\par }{\Q he took notice of my blameless
behavior.
\par }{\Q \VS{25}You prove
to be loyal
to one who is faithful;
\par }{\Q you prove
to be trustworthy
to one who is
innocent.
\par }{\Q \VS{26}You prove
to be
reliable to
one who is blameless,
\par }{\Q but
you prove to be deceptive
to one who is perverse.
\par }{\Q \VS{27}For
you
deliver
oppressed
people,
\par }{\Q but you bring down those who have a proud look.
\par }{\Q \VS{28}Indeed,
you
are my lamp,

{\ND{Lord}}.

\par }{\Q My God
illuminates
the darkness around me.
\par }{\Q \VS{29}Indeed,
with your help I can charge against an army;
\par }{\Q by my God’s
power I can jump over
a wall.
\par }{\Q \VS{30}The one true God
acts in a faithful manner;
\par }{\Q the
{\ND{Lord}}’s
promise
is reliable;
\par }{\Q he is
a shield
to all
who take shelter in him.
\par }{\Q \VS{31}Indeed,
who
is God
besides
the {\ND{Lord}}?
\par }{\Q Who
is a protector
besides
our God?
\par }{\Q \VS{32}The one true God
gives
me strength;
\par }{\Q he removes
the obstacles
in my way.
\par }{\Q \VS{33}He gives me the agility
of a deer;
\par }{\Q he enables me to negotiate
the rugged terrain.
\par }{\Q \VS{34}He trains
my hands
for battle;
\par }{\Q my arms can bend
even the strongest
bow.
\par }{\Q \VS{35}You give
me your protective
shield;
\par }{\Q your right hand
supports
me;

\par }{\Q your willingness
to help enables me to prevail.
\par }{\Q \VS{36}You widen
my path;
\par }{\Q my feet
do not
slip.
\par }{\Q \VS{37}I chase
my enemies
and catch
them;
\par }{\Q I do not
turn back
until
I wipe them out.
\par }{\Q \VS{38}I beat
them to death;

\par }{\Q they fall
at my feet.
\par }{\Q \VS{39}You give
me strength
for battle;
\par }{\Q you make my foes kneel
before me.
\par }{\Q \VS{40}You make
my enemies
retreat;
\par }{\Q I destroy
those who hate
me.
\par }{\Q \VS{41}They cry out,
but there is no
one to help
them;

\par }{\Q they cry out to the
{\ND{Lord}}, but he does not
answer them.
\par }{\Q \VS{42}I grind
them as fine windblown dust;
\par }{\Q I beat them underfoot
like clay
in the streets.
\par }{\Q \VS{43}You rescue
me from a hostile
army;
\par }{\Q you make
me a
leader
of nations;
\par }{\Q people
over whom I had no
authority are now my subjects.
\par }{\Q \VS{44}When they hear
of my exploits, they submit
to me.

\par }{\Q Foreigners
are powerless before me;
\par }{\Q \VS{45}foreigners
lose their courage;
\par }{\Q they shake with fear as
they leave their
strongholds.
\par }{\Q \VS{46}The
{\ND{Lord}}
is alive!

\par }{\Q My protector
is praiseworthy!

\par }{\Q The God
who delivers
me is exalted as king!
\par }{\Q \VS{47}The one true God
completely vindicates
me;

\par }{\Q he makes
nations
submit to me.
\par }{\Q \VS{48}He delivers
me from my enemies;
\par }{\Q you snatch
me away from those who attack me;

\par }{\Q you rescue
me from violent
men.
\par }{\Q \VS{49}So
I will give you thanks
before the nations,
O
{\ND{Lord}}!
\par }{\Q I will sing praises
to you!
\par }{\Q \VS{50}He gives
his chosen king
magnificent
victories;
\par }{\Q he
is faithful
to his chosen ruler,
\par }{\Q to David
and his descendants
forever.”


\par }\PsalmChap{19}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}The heavens
declare
the glory
of God;
\par }{\Q the sky displays his handiwork.
\par }{\Q \VS{2}Day
after day
it speaks out;
\par }{\Q night
after night
it reveals
his greatness.
\par }{\Q \VS{3}There is no
actual speech
or word,
\par }{\Q nor is its voice
literally
heard.
\par }{\Q \VS{4}Yet its voice echoes
throughout
the earth;
\par }{\Q its words
carry to the distant horizon.
\par }{\Q In the sky he has pitched
a tent
for the sun.
\par }{\Q \VS{5}Like a bridegroom
it emerges
from its chamber;
\par }{\Q like a strong man it enjoys
running
its course.
\par }{\Q \VS{6}It emerges
from the distant horizon,
\par }{\Q and goes from one end
of the sky
to the other;
\par }{\Q nothing
can escape
its heat.
\par }{\Q \VS{7}The law
of the {\ND{Lord}}
is perfect
\par }{\Q and preserves
one’s life.
\par }{\Q The rules
set down by the
{\ND{Lord}}
are reliable
\par }{\Q and impart wisdom
to the inexperienced.
\par }{\Q \VS{8}The
{\ND{Lord}}’s
precepts
are fair
\par }{\Q and make one joyful.
\par }{\Q The
{\ND{Lord}}’s
commands
are pure
\par }{\Q and give insight
for life.
\par }{\Q \VS{9}The commands to fear
the {\ND{Lord}}
are right
\par }{\Q and endure
forever.
\par }{\Q The judgments
given by the
{\ND{Lord}}
are trustworthy
\par }{\Q and absolutely
just.
\par }{\Q \VS{10}They are of greater value
than gold,
\par }{\Q than even a great
amount of pure gold;
\par }{\Q they bring greater delight than honey,
\par }{\Q than even the sweetest
honey
from a honeycomb.
\par }{\Q \VS{11}Yes,
your servant
finds moral
guidance there;

\par }{\Q those who obey
them receive a rich
reward.
\par }{\Q \VS{12}Who
can know
all his errors?

\par }{\Q Please do not punish
me for sins I am unaware of.
\par }{\Q \VS{13}Moreover,
keep
me
from committing flagrant sins;
\par }{\Q do not
allow such sins to control
me.

\par }{\Q Then
I will be blameless,
\par }{\Q and innocent
of blatant
rebellion.
\par }{\Q \VS{14}May
my words
and my
thoughts
\par }{\Q be acceptable
in your sight,
\par }{\Q O
{\ND{Lord}}, my sheltering rock
and my redeemer.


\par }\PsalmChap{20}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}May the
{\ND{Lord}}
answer
you when
you are in trouble;
\par }{\Q may
the God
of Jacob
make
you secure!
\par }{\Q \VS{2}May he send
you help
from his temple;
\par }{\Q from Zion
may he give you support!
\par }{\Q \VS{3}May
he take notice
of your offerings;
\par }{\Q may he accept
your burnt sacrifice! (Selah)
\par }{\Q \VS{4}May
he grant
your heart’s desire;
\par }{\Q may he bring all
your plans
to pass!
\par }{\Q \VS{5}Then we will shout
for joy over your victory;
\par }{\Q we will rejoice in the name
of our God!
\par }{\Q May
the {\ND{Lord}}
grant all
your requests!
\par }{\Q \VS{6}Now
I am sure
that
the {\ND{Lord}}
will deliver
his chosen king;
\par }{\Q he will intervene
for him from his holy
heavenly
temple,

\par }{\Q and display
his mighty
ability to deliver.
\par }{\Q \VS{7}Some trust
in chariots
and others
in horses,
\par }{\Q but we
depend
on the
{\ND{Lord}}
our God.
\par }{\Q \VS{8}They
will fall
down,

\par }{\Q but we
will stand
firm.
\par }{\Q \VS{9}The
{\ND{Lord}}
will deliver
the king;
\par }{\Q he will answer
us when
we call to him for help!


\par }\PsalmChap{21}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}O

{\ND{Lord}}, the king
rejoices
in the strength
you give;

\par }{\Q he takes
great
delight
in the deliverance you provide.
\par }{\Q \VS{2}You grant
him his heart’s
desire;
\par }{\Q you do not
refuse
his request.
(Selah)
\par }{\Q \VS{3}For
you bring him rich
blessings;
\par }{\Q you place
a golden
crown
on his head.
\par }{\Q \VS{4}He asked
you
to sustain
his life,
\par }{\Q and you have granted
him long
life
and an
enduring dynasty.
\par }{\Q \VS{5}Your deliverance
brings him great
honor;
\par }{\Q you give him majestic
splendor.
\par }{\Q \VS{6}For
you grant
him lasting
blessings;
\par }{\Q you give him great joy
by allowing
him
into your presence.
\par }{\Q \VS{7}For
the king
trusts
in the
{\ND{Lord}},
\par }{\Q and because of the sovereign
{\SC{
Lord}}’s faithfulness
he is not
upended.
\par }{\Q \VS{8}You prevail
over all
your enemies;
\par }{\Q your power
is too great for those who hate you.
\par }{\Q \VS{9}You burn
them up
like a fiery
furnace
when you appear;
\par }{\Q the {\ND{Lord}}
angrily
devours
them;

\par }{\Q the fire
consumes them.
\par }{\Q \VS{10}You destroy
their offspring
from the earth,
\par }{\Q their descendants
from among the human race.
\par }{\Q \VS{11}Yes,
they intend
to do
you harm;
\par }{\Q they dream up
a scheme,
but they do not succeed.
\par }{\Q \VS{12}For
you make
them retreat
\par }{\Q when you shoot your arrows at them.
\par }{\Q \VS{13}Rise
up, O
{\ND{Lord}}, in strength!

\par }{\Q We
will sing
and praise your power!


\par }\PsalmChap{22}{\D 
For the music director;
according to the tune “Morning Doe;”
a psalm
of David.
\par }{\Q \VerseOne{1}My God,
my God,
why
have you abandoned
me?

\par }{\Q I groan
in prayer, but help seems
far away.
\par }{\Q \VS{2}My God,
I cry out
during the day,
\par }{\Q but you do not
answer,
\par }{\Q and during the night
my prayers
do not let up.
\par }{\Q \VS{3}You
are holy;
\par }{\Q you sit
as king receiving the praises
of Israel.
\par }{\Q \VS{4}In you our ancestors
trusted;
\par }{\Q they trusted
in you and you rescued them.
\par }{\Q \VS{5}To
you they cried
out, and they were saved;
\par }{\Q in you they trusted
and they were not
disappointed.
\par }{\Q \VS{6}But I
am a worm,
not
a man;
\par }{\Q people
insult
me and despise me.
\par }{\Q \VS{7}All
who see
me taunt
me;
\par }{\Q they mock
me and shake
their heads.
\par }{\Q \VS{8}They say,

\par }{\Q “Commit
yourself to
the {\ND{Lord}}!
\par }{\Q Let the
{\ND{Lord}} rescue
him!
\par }{\Q Let the
{\ND{Lord}} deliver
him, for
he delights in him.”
\par }{\Q \VS{9}Yes,
you
are the one who brought me out from the womb
\par }{\Q and made me feel
secure
on
my mother’s
breasts.
\par }{\Q \VS{10}I have been dependent on you since birth;
\par }{\Q from
the time I came out
of my mother’s
womb
you
have been my God.
\par }{\Q \VS{11}Do not
remain far
away from
me,
\par }{\Q for
trouble
is near
and I have no
one to help me.
\par }{\Q \VS{12}Many
bulls
surround
me;
\par }{\Q powerful bulls
of Bashan
hem me in.
\par }{\Q \VS{13}They open
their mouths
to devour me

\par }{\Q like a roaring
lion
that rips its prey.
\par }{\Q \VS{14}My strength drains
away like water;
\par }{\Q all
my bones
are dislocated;
\par }{\Q my heart
is like wax;
\par }{\Q it melts away
inside me.
\par }{\Q \VS{15}The roof of my mouth is as dry
as a piece of pottery;
\par }{\Q my tongue
sticks
to my gums.
\par }{\Q You set
me in the dust
of death.
\par }{\Q \VS{16}Yes,
wild dogs
surround me –
\par }{\Q a gang of evil men crowd around me;
\par }{\Q like a lion they pin my hands and feet.
\par }{\Q \VS{17}I can count
all
my bones;
\par }{\Q my enemies are
gloating
over me in triumph.
\par }{\Q \VS{18}They are dividing
up my clothes
among themselves;
\par }{\Q they are rolling
dice
for
my garments.
\par }{\Q \VS{19}But you,
O
{\ND{Lord}}, do not
remain far away!
\par }{\Q You are my source of strength! Hurry
and help me!
\par }{\Q \VS{20}Deliver
me from the sword!
\par }{\Q Save my life
from the claws
of the wild dogs!
\par }{\Q \VS{21}Rescue
me from the mouth
of the lion,
\par }{\Q and from the horns
of the wild oxen!

\par }{\Q You have answered me!
\par }{\Q \VS{22}I will declare
your name
to my countrymen!

\par }{\Q In the middle
of the assembly
I will praise you!
\par }{\Q \VS{23}You loyal followers
of the {\ND{Lord}}, praise
him!
\par }{\Q All
you descendants
of Jacob,
honor
him!
\par }{\Q All
you descendants
of Israel,
stand in awe of him!
\par }{\Q \VS{24}For
he did not
despise
or
detest
the suffering
of the oppressed;
\par }{\Q he did not
ignore
him;

\par }{\Q when he cried out
to
him, he responded.
\par }{\Q \VS{25}You are the reason I offer praise
in the great
assembly;
\par }{\Q I will fulfill my promises
before
the
{\ND{Lord}}’s loyal followers.
\par }{\Q \VS{26}Let the oppressed
eat
and be filled!

\par }{\Q Let those who seek
his help praise
the {\ND{Lord}}!
\par }{\Q May you live
forever!
\par }{\Q \VS{27}Let
all
the people of the earth
acknowledge
the {\ND{Lord}}
and turn
to
him!

\par }{\Q Let all
the nations
worship you!
\par }{\Q \VS{28}For
the {\ND{Lord}}
is king
\par }{\Q and rules
over the nations.
\par }{\Q \VS{29}All
of the thriving
people of the earth
will join the celebration
and worship;
\par }{\Q all
those who are descending
into the grave
will bow
before
him,
\par }{\Q including those who cannot
preserve
their lives.
\par }{\Q \VS{30}A whole generation
will serve
him;
\par }{\Q they will tell
the next generation
about the sovereign Lord.
\par }{\Q \VS{31}They will come
and tell
about his saving deeds;
\par }{\Q they will tell a future
generation
what
he has accomplished.


\par }\PsalmChap{23}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}
is my shepherd,
\par }{\Q I lack
nothing.
\par }{\Q \VS{2}He takes
me to lush
pastures,
\par }{\Q he leads
me to refreshing
water.
\par }{\Q \VS{3}He restores
my strength.
\par }{\Q He leads
me down the right
paths
\par }{\Q for
the sake of his reputation.
\par }{\Q \VS{4}Even
when
I must walk through
the darkest
valley,
\par }{\Q I fear
no
danger,
\par }{\Q for
you
are with me;
\par }{\Q your rod
and your staff
reassure me.
\par }{\Q \VS{5}You prepare
a feast
before
me

\par }{\Q in plain sight
of my enemies.
\par }{\Q You refresh
my head
with oil;
\par }{\Q my cup
is completely full.
\par }{\Q \VS{6}Surely
your goodness
and faithfulness
will pursue
me all
my days,
\par }{\Q and I will live in
the
{\ND{Lord}}’s
house
for the rest
of my life.

\par }\PsalmChap{24}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}
owns the earth
and all it contains,
\par }{\Q the world
and all who live in it.
\par }{\Q \VS{2}For
he set its foundation upon
the seas,
\par }{\Q and established
it upon
the ocean currents.
\par }{\Q \VS{3}Who
is allowed to ascend
the mountain
of the {\ND{Lord}}?

\par }{\Q Who
may go up
to his holy
dwelling place?
\par }{\Q \VS{4}The one whose deeds
are blameless
\par }{\Q and whose motives
are pure,
\par }{\Q who
does not
lie,
\par }{\Q or make promises
with no intention of keeping them.
\par }{\Q \VS{5}Such godly
people are rewarded
by the
{\ND{Lord}},

\par }{\Q and vindicated
by the God
who delivers them.
\par }{\Q \VS{6}Such
purity characterizes
the people who seek
his favor,
\par }{\Q Jacob’s
descendants, who pray to him. (Selah)
\par }{\Q \VS{7}Look
up, you gates!
\par }{\Q Rise
up, you eternal
doors!
\par }{\Q Then the majestic
king
will enter!
\par }{\Q \VS{8}Who
is this
majestic
king?

\par }{\Q The
{\ND{Lord}}
who is strong
and mighty!
\par }{\Q The
{\ND{Lord}}
who is mighty
in battle!
\par }{\Q \VS{9}Look
up, you gates!
\par }{\Q Rise
up, you eternal
doors!
\par }{\Q Then the majestic
king
will enter!
\par }{\Q \VS{10}Who
is this
majestic
king?
\par }{\Q The
{\ND{Lord}}
who commands armies!

\par }{\Q He is
the majestic
king! (Selah)


\par }\PsalmChap{25}{\D 
By David.
\par }{\Q \VerseOne{1}O

{\ND{Lord}}, I
come before
you in prayer.
\par }{\Q \VS{2}My God,
I trust
in you.
\par }{\Q Please do not
let me be humiliated;
\par }{\Q do not
let my enemies
triumphantly rejoice over me!
\par }{\Q \VS{3}Certainly
none
who rely on
you will be humiliated.
\par }{\Q Those who deal in treachery
will be thwarted
and humiliated.
\par }{\Q \VS{4}Make me understand
your ways,
O
{\ND{Lord}}!
\par }{\Q Teach
me your paths!
\par }{\Q \VS{5}Guide
me into your truth
and teach
me.
\par }{\Q For
you
are the God
who delivers
me;
\par }{\Q on you I rely
all
day long.
\par }{\Q \VS{6}Remember
your compassionate
and faithful
deeds, O
{\ND{Lord}},
\par }{\Q for
you have always
acted in this manner.
\par }{\Q \VS{7}Do not
hold
against me the sins
of my youth
or my rebellious
acts!
\par }{\Q Because you
are faithful
to me, extend
to me
your favor,
O
{\ND{Lord}}!
\par }{\Q \VS{8}The
{\ND{Lord}}
is
both
kind and fair;
\par }{\Q that is why
he teaches
sinners
the right way to live.
\par }{\Q \VS{9}May he show
the humble
what is right!

\par }{\Q May he teach
the humble
his way!
\par }{\Q \VS{10}The
{\ND{Lord}}
always proves
faithful
and reliable
\par }{\Q to those who follow
the demands
of his covenant.
\par }{\Q \VS{11}For the sake
of your reputation,
O
{\ND{Lord}},
\par }{\Q forgive
my sin,
because
it is
great.
\par }{\Q \VS{12}The
{\ND{Lord}}
shows
his faithful followers
\par }{\Q the way
they should
live.
\par }{\Q \VS{13}They experience
his favor;

\par }{\Q their descendants
inherit
the land.
\par }{\Q \VS{14}The
{\ND{Lord}}’s
loyal followers
receive his guidance,
\par }{\Q and he reveals his covenantal
demands to them.
\par }{\Q \VS{15}I continually
look
to
the {\ND{Lord}}
for help,

\par }{\Q for
he
will free
my feet
from the enemy’s net.
\par }{\Q \VS{16}Turn
toward me
and have mercy
on me,
\par }{\Q for
I am
alone
and oppressed!
\par }{\Q \VS{17}Deliver
me
from my distress;
\par }{\Q rescue me from my suffering!
\par }{\Q \VS{18}See
my pain
and suffering!
\par }{\Q Forgive
all
my sins!
\par }{\Q \VS{19}Watch
my enemies,
for
they outnumber
me;
\par }{\Q they hate
me and want to harm
me.
\par }{\Q \VS{20}Protect
me
and deliver
me!
\par }{\Q Please do not
let me be humiliated,
\par }{\Q for
I have taken shelter in you!
\par }{\Q \VS{21}May integrity
and godliness
protect
me,
\par }{\Q for
I rely on you!
\par }{\Q \VS{22}O God,
rescue
Israel
\par }{\Q from all
their distress!


\par }\PsalmChap{26}{\D 
By David.
\par }{\Q \VerseOne{1}Vindicate
me, O
{\ND{Lord}},
\par }{\Q for
I
have integrity,
\par }{\Q and I trust
in the
{\ND{Lord}}
without
wavering.
\par }{\Q \VS{2}Examine
me, O
{\ND{Lord}}, and test
me!
\par }{\Q Evaluate
my inner thoughts
and motives!
\par }{\Q \VS{3}For
I am ever aware
of your faithfulness,
\par }{\Q and your loyalty continually motivates me.
\par }{\Q \VS{4}I do not
associate
with
deceitful
men,
\par }{\Q or consort
with
those who are dishonest.
\par }{\Q \VS{5}I hate
the mob
of evil
men,
\par }{\Q and do not
associate
with
the wicked.
\par }{\Q \VS{6}I maintain
a pure
lifestyle,

\par }{\Q so I can appear
before your altar,
O
{\ND{Lord}},
\par }{\Q \VS{7}to give
you thanks,
\par }{\Q and to tell
about all
your amazing deeds.
\par }{\Q \VS{8}O
{\ND{Lord}}, I love
the temple
where you live,
\par }{\Q the place
where your splendor
is revealed.
\par }{\Q \VS{9}Do not
sweep
me away
with
sinners,
\par }{\Q or execute
me along with
violent
people,
\par }{\Q \VS{10}who
are always ready to do wrong
\par }{\Q or offer a bribe.
\par }{\Q \VS{11}But I
have integrity!

\par }{\Q Rescue
me and have mercy on me!
\par }{\Q \VS{12}I am safe,

\par }{\Q and among the worshipers
I will praise
the {\ND{Lord}}.


\par }\PsalmChap{27}{\D 
By David.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}
delivers
and vindicates
me!

\par }{\Q I fear
no one!

\par }{\Q The
{\ND{Lord}}
protects
my life!
\par }{\Q I am afraid
of no one!
\par }{\Q \VS{2}When
evil
men attack
me

\par }{\Q to devour
my flesh,
\par }{\Q when my adversaries
and enemies
attack me,

\par }{\Q they
stumble
and fall.
\par }{\Q \VS{3}Even
when an army
is deployed
against
me,
\par }{\Q I do not
fear.
\par }{\Q Even when
war
is imminent,
\par }{\Q I
remain confident.
\par }{\Q \VS{4}I have asked
the {\ND{Lord}}
for one thing –
\par }{\Q this is what I desire!
\par }{\Q I want to live in the
{\ND{Lord}}’s house all the days of my life,
\par }{\Q so I can gaze at the splendor of the
{\ND{Lord}}
\par }{\Q and contemplate in his temple.
\par }{\Q \VS{5}He will surely
give me shelter
in the day
of danger;
\par }{\Q he will hide
me in his home;
\par }{\Q he will place me on an inaccessible
rocky summit.
\par }{\Q \VS{6}Now
I will triumph
\par }{\Q over
my enemies
who surround
me!

\par }{\Q I will offer
sacrifices
in his dwelling
place and shout
for joy!

\par }{\Q I will sing
praises
to the
{\ND{Lord}}!
\par }{\Q \VS{7}Hear
me, O
{\ND{Lord}}, when
I cry
out!
\par }{\Q Have mercy
on me and answer me!
\par }{\Q \VS{8}My heart
tells
me to pray to you,

\par }{\Q and I do pray
to you, O
{\ND{Lord}}.
\par }{\Q \VS{9}Do not
reject
me!

\par }{\Q Do not
push
your servant
away
in anger!
\par }{\Q You are my deliverer!

\par }{\Q Do
not
forsake
or
abandon
me,
\par }{\Q O God
who vindicates me!
\par }{\Q \VS{10}Even if
my father
and mother
abandoned
me,

\par }{\Q the {\ND{Lord}}
would take me in.
\par }{\Q \VS{11}Teach
me how you want
me to live;
\par }{\Q lead
me along a level
path
because
of those who wait to ambush me!
\par }{\Q \VS{12}Do not
turn
me
over
to my enemies,
\par }{\Q for
false
witnesses
who want
to destroy
me testify against me.
\par }{\Q \VS{13}Where would I be if
I did not believe
I would experience
\par }{\Q the
{\ND{Lord}}’s
favor
in the land
of the living?
\par }{\Q \VS{14}Rely on
the {\ND{Lord}}!
\par }{\Q Be strong
and confident!

\par }{\Q Rely on
the {\ND{Lord}}!


\par }\PsalmChap{28}{\D 
By David.
\par }{\Q \VerseOne{1}To
you, O
{\ND{Lord}}, I cry
out!
\par }{\Q My protector,
do not
ignore
me!

\par }{\Q If you do not
respond
to me,
\par }{\Q I will join those who are descending
into the grave.
\par }{\Q \VS{2}Hear
my plea
for mercy when I cry out
to you for help,
\par }{\Q when I lift
my hands
toward
your holy
temple!
\par }{\Q \VS{3}Do not
drag
me away with
evil men,
\par }{\Q with
those who behave
wickedly,
\par }{\Q who talk
so friendly
to
their neighbors,
\par }{\Q while they plan to
harm them!
\par }{\Q \VS{4}Pay
them
back for their evil
deeds!
\par }{\Q Pay
them
back
for what
they do!
\par }{\Q Punish them!
\par }{\Q \VS{5}For
they do not
understand
the
{\ND{Lord}}’s actions,
\par }{\Q or the way he carries out justice.

\par }{\Q The
{\ND{Lord}}
will permanently demolish them.
\par }{\Q \VS{6}The
{\ND{Lord}}
deserves
praise,

\par }{\Q for
he has heard
my plea for mercy!
\par }{\Q \VS{7}The
{\ND{Lord}}
strengthens
and protects
me;

\par }{\Q I trust
in him
with all my heart.
\par }{\Q I am rescued and my heart
is full of joy;

\par }{\Q I will sing
to him in gratitude.
\par }{\Q \VS{8}The
{\ND{Lord}}
strengthens
his people;

\par }{\Q he protects
and delivers
his chosen king.
\par }{\Q \VS{9}Deliver
your people!
\par }{\Q Empower
the
nation that belongs
to you!

\par }{\Q Care for them like a shepherd
and carry
them in your arms
at
all
times!


\par }\PsalmChap{29}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}Acknowledge
the {\ND{Lord}}, you heavenly beings,
\par }{\Q acknowledge
the
{\ND{Lord}}’s
majesty
and power!
\par }{\Q \VS{2}Acknowledge
the majesty
of the
{\ND{Lord}}’s
reputation!

\par }{\Q Worship
the {\ND{Lord}}
in holy
attire!
\par }{\Q \VS{3}The
{\ND{Lord}}’s
shout
is heard over
the water;
\par }{\Q the majestic
God
thunders,
\par }{\Q the {\ND{Lord}}
appears over
the surging
water.
\par }{\Q \VS{4}The
{\ND{Lord}}’s
shout
is powerful,
\par }{\Q the
{\ND{Lord}}’s
shout
is majestic.
\par }{\Q \VS{5}The
{\ND{Lord}}’s
shout
breaks
the cedars,
\par }{\Q the {\ND{Lord}}
shatters
the cedars
of Lebanon.
\par }{\Q \VS{6}He makes Lebanon
skip
like
a calf
\par }{\Q and Sirion
like
a young
ox.
\par }{\Q \VS{7}The
{\ND{Lord}}’s
shout strikes
with flaming
fire.
\par }{\Q \VS{8}The
{\ND{Lord}}’s
shout
shakes
the wilderness,
\par }{\Q the {\ND{Lord}}
shakes
the wilderness
of Kadesh.
\par }{\Q \VS{9}The
{\ND{Lord}}’s
shout
bends
the large trees

\par }{\Q and strips
the leaves from the forests.
\par }{\Q Everyone
in his temple
says,
“Majestic!”
\par }{\Q \VS{10}The
{\ND{Lord}}
sits enthroned over the engulfing waters,
\par }{\Q the {\ND{Lord}}
sits
enthroned
as the eternal
king.
\par }{\Q \VS{11}The
{\ND{Lord}}
gives
his people
strength;
\par }{\Q the {\ND{Lord}}
grants
his people
security.


\par }\PsalmChap{30}{\D 
A psalm
– a song
used at the dedication
of the temple;
by David.
\par }{\Q \VerseOne{1}I will praise
you, O
{\ND{Lord}}, for
you lifted
me up,

\par }{\Q and did not
allow my enemies
to gloat over me.
\par }{\Q \VS{2}O
{\ND{Lord}}
my God,
\par }{\Q I cried out
to
you and you healed me.
\par }{\Q \VS{3}O
{\ND{Lord}}, you pulled
me up from
Sheol;
\par }{\Q you rescued
me from
among those descending into the grave.
\par }{\Q \VS{4}Sing
to the
{\ND{Lord}}, you faithful followers
of his;
\par }{\Q give thanks
to his holy
name.
\par }{\Q \VS{5}For
his anger
lasts only a brief moment,
\par }{\Q and his good favor
restores one’s life.
\par }{\Q One may experience
sorrow
during the night,
\par }{\Q but joy
arrives in the morning.
\par }{\Q \VS{6}In my self-confidence
I
said,
\par }{\Q “I will never
be upended.”
\par }{\Q \VS{7}O
{\ND{Lord}}, in your good favor
you made
me secure.
\par }{\Q Then you rejected
me and I was
terrified.
\par }{\Q \VS{8}To
you, O
{\ND{Lord}}, I cried
out;
\par }{\Q I begged the Lord
for mercy:
\par }{\Q \VS{9}“What
profit
is there in taking
my life,

\par }{\Q in my descending
into the Pit?

\par }{\Q Can the dust
of the grave praise
you?
\par }{\Q Can it declare
your loyalty?
\par }{\Q \VS{10}Hear,
O
{\ND{Lord}}, and have mercy
on me!
\par }{\Q O
{\ND{Lord}}, deliver me!”
\par }{\Q \VS{11}Then you turned
my lament
into dancing;
\par }{\Q you removed
my sackcloth
and covered
me with joy.
\par }{\Q \VS{12}So
now my heart
will sing
to you and not
be silent;
\par }{\Q O
{\ND{Lord}}
my God,
I will always
give thanks to you.


\par }\PsalmChap{31}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}In you, O
{\ND{Lord}}, I have taken shelter!
\par }{\Q Never
let
me be humiliated!
\par }{\Q Vindicate
me by rescuing me!
\par }{\Q \VS{2}Listen
to
me!

\par }{\Q Quickly
deliver
me!
\par }{\Q Be
my protector
and refuge,
\par }{\Q a stronghold
where I can be safe!
\par }{\Q \VS{3}For
you are
my high ridge
and my stronghold;
\par }{\Q for the sake
of your own reputation
you lead
me and guide me.
\par }{\Q \VS{4}You will free
me from the net
they hid
for me,
\par }{\Q for
you
are my place of refuge.
\par }{\Q \VS{5}Into your hand
I entrust
my life;

\par }{\Q you will rescue
me, O
{\ND{Lord}}, the faithful
God.
\par }{\Q \VS{6}I hate
those who serve
worthless
idols,
\par }{\Q but I
trust
in
the {\ND{Lord}}.
\par }{\Q \VS{7}I will be happy
and rejoice
in your faithfulness,
\par }{\Q because you notice
my pain
\par }{\Q and you are aware
of how distressed I am.
\par }{\Q \VS{8}You do not
deliver
me over to the power
of the enemy;
\par }{\Q you enable
me to stand
in a wide open place.
\par }{\Q \VS{9}Have mercy
on me, for
I am in distress!
\par }{\Q My eyes
grow dim
from suffering.
\par }{\Q I have lost my strength.
\par }{\Q \VS{10}For
my life
nears its end
in pain;
\par }{\Q my years
draw to a close as I groan.
\par }{\Q My strength
fails
me because of my sin,
\par }{\Q and my bones
become brittle.
\par }{\Q \VS{11}Because of all
my enemies,
people disdain
me;

\par }{\Q my neighbors
are appalled
by my suffering –
\par }{\Q those who know me are horrified by my condition;
\par }{\Q those who see me in the street run away from me.
\par }{\Q \VS{12}I am forgotten,
like a dead
man no one thinks
about;

\par }{\Q I am regarded
as worthless, like a broken
jar.
\par }{\Q \VS{13}For
I hear
what so many
are saying,

\par }{\Q the terrifying news
that comes from every direction.
\par }{\Q When they plot
together
against
me,
\par }{\Q they figure out
how they can take
my life.
\par }{\Q \VS{14}But I
trust
in you, O
{\ND{Lord}}!
\par }{\Q I declare,
“You are
my God!”
\par }{\Q \VS{15}You determine my
destiny!


\par }{\Q Rescue
me from the power
of my enemies
and those who chase me.
\par }{\Q \VS{16}Smile
on
your servant!
\par }{\Q Deliver
me because of your faithfulness!
\par }{\Q \VS{17}O
{\ND{Lord}}, do not
let me be humiliated,
\par }{\Q for
I call
out to you!
\par }{\Q May evil
men be humiliated!
\par }{\Q May they go wailing
to the grave!
\par }{\Q \VS{18}May lying
lips
be silenced
–
\par }{\Q lips
that speak
defiantly
against
the innocent
\par }{\Q with arrogance
and contempt!
\par }{\Q \VS{19}How
great
is your favor,
\par }{\Q which
you store up
for your loyal followers!

\par }{\Q In plain sight of everyone you bestow
it on those who take shelter
in you.
\par }{\Q \VS{20}You hide
them with you, where they are safe
from the attacks
of men;
\par }{\Q you conceal
them in a shelter,
where they are safe from slanderous
attacks.
\par }{\Q \VS{21}The
{\ND{Lord}}
deserves
praise

\par }{\Q for
he demonstrated his amazing
faithfulness
to me when I was besieged by enemies.
\par }{\Q \VS{22}I
jumped to conclusions
and said,
\par }{\Q “I am cut off from your presence!”

\par }{\Q But
you heard
my plea
for mercy when I cried out
to you for help.
\par }{\Q \VS{23}Love
the {\ND{Lord}}, all
you faithful
followers
of his!
\par }{\Q The
{\ND{Lord}}
protects
those who have integrity,
\par }{\Q but he pays back
in
full
the one who acts
arrogantly.
\par }{\Q \VS{24}Be strong
and confident,
\par }{\Q all
you who wait
on the
{\ND{Lord}}!


\par }\PsalmChap{32}{\D 
By David;
a well-written song.
\par }{\Q \VerseOne{1}How blessed
is the one whose rebellious acts
are forgiven,
\par }{\Q whose sin is pardoned!
\par }{\Q \VS{2}How blessed
is the one
whose wrongdoing
the {\ND{Lord}}
does not
punish,

\par }{\Q in whose spirit
there is no
deceit.
\par }{\Q \VS{3}When
I refused to confess my sin,

\par }{\Q my whole body wasted away,
\par }{\Q while I groaned
in pain all
day long.
\par }{\Q \VS{4}For
day
and night
you tormented
me;

\par }{\Q you tried
to destroy
me in the intense heat
of summer.
(Selah)
\par }{\Q \VS{5}Then I confessed
my sin;
\par }{\Q I no
longer covered
up my wrongdoing.
\par }{\Q I said,
“I will confess
my rebellious
acts to the
{\ND{Lord}}.”
\par }{\Q And then you
forgave
my sins.
(Selah)
\par }{\Q \VS{6}For this
reason every one
of your faithful followers
should pray
to
you
\par }{\Q while there is a window of opportunity.
\par }{\Q Certainly when the surging
water
rises,
\par }{\Q it will not
reach them.
\par }{\Q \VS{7}You
are my hiding
place;
\par }{\Q you protect
me from distress.
\par }{\Q You surround
me with shouts
of joy from those celebrating deliverance.
(Selah)
\par }{\Q \VS{8}I will instruct
and teach
you about
how
you should live.
\par }{\Q I will advise
you as I look you in
the eye.
\par }{\Q \VS{9}Do not
be
like an unintelligent
horse
or mule,
\par }{\Q which will not
obey
you
\par }{\Q unless they are controlled
by
a bridle
and bit.
\par }{\Q \VS{10}An evil
person suffers
much pain,
\par }{\Q but
the
{\ND{Lord}}’s
faithfulness
overwhelms
the one who trusts in him.
\par }{\Q \VS{11}Rejoice
in the
{\ND{Lord}}
and be happy,
you who are godly!
\par }{\Q Shout
for joy, all
you who are morally
upright!


\par }\PsalmChap{33}{\PP \par }{\Q \VerseOne{1}You godly
ones, shout
for joy because of the
{\ND{Lord}}!
\par }{\Q It is appropriate
for the morally upright
to offer him praise.
\par }{\Q \VS{2}Give thanks
to the
{\ND{Lord}}
with the harp!
\par }{\Q Sing
to him to the accompaniment
of a ten-stringed
instrument!
\par }{\Q \VS{3}Sing
to him a new
song!

\par }{\Q Play
skillfully
as you shout out your praises to him!
\par }{\Q \VS{4}For
the
{\ND{Lord}}’s
decrees
are just,
\par }{\Q and everything
he does is fair.
\par }{\Q \VS{5}The
{\ND{Lord}}
promotes
equity
and justice;
\par }{\Q the
{\ND{Lord}}’s
faithfulness
extends
throughout the earth.
\par }{\Q \VS{6}By the
{\ND{Lord}}’s
decree the heavens
were made;
\par }{\Q by a mere
word from his mouth
all
the stars
in the sky were created.
\par }{\Q \VS{7}He piles up
the water
of the sea;
\par }{\Q he puts
the oceans
in storehouses.
\par }{\Q \VS{8}Let the whole
earth
fear
the {\ND{Lord}}!
\par }{\Q Let all
who live
in the world
stand in awe of him!
\par }{\Q \VS{9}For
he spoke,
and it came into existence,
\par }{\Q he issued the decree,
and it stood firm.
\par }{\Q \VS{10}The
{\ND{Lord}}
frustrates
the decisions
of the nations;
\par }{\Q he nullifies
the plans
of the peoples.
\par }{\Q \VS{11}The
{\ND{Lord}}’s
decisions
stand
forever;
\par }{\Q his plans
abide
throughout
the ages.
\par }{\Q \VS{12}How blessed
is the nation
whose
God
is the
{\ND{Lord}},
\par }{\Q the people
whom he has chosen
to be his special possession.
\par }{\Q \VS{13}The
{\ND{Lord}}
watches
from heaven;
\par }{\Q he sees
all
people.
\par }{\Q \VS{14}From the place
where he lives
he looks
carefully
\par }{\Q at all
the earth’s
inhabitants.
\par }{\Q \VS{15}He is the one who forms
every human heart,
\par }{\Q and takes note
of all
their actions.
\par }{\Q \VS{16}No
king
is delivered
by
his vast
army;
\par }{\Q a warrior
is not
saved
by his great
might.
\par }{\Q \VS{17}A horse
disappoints
those who trust in it for victory;
\par }{\Q despite its great
strength,
it cannot
deliver.
\par }{\Q \VS{18}Look,
the {\ND{Lord}}
takes notice
of his loyal followers,
\par }{\Q those who wait
for him to demonstrate his faithfulness
\par }{\Q \VS{19}by saving
their lives
from death
\par }{\Q and sustaining
them during times of famine.
\par }{\Q \VS{20}We wait
for the
{\ND{Lord}};
\par }{\Q he is
our deliverer
and shield.
\par }{\Q \VS{21}For
our hearts
rejoice
in him,
\par }{\Q for
we trust
in his holy
name.
\par }{\Q \VS{22}May
we experience your faithfulness,
O
{\ND{Lord}},

\par }{\Q for we wait for you.


\par }\PsalmChap{34}{\D 
Written by David,
when he pretended
to be insane before
Abimelech,
causing the king to send
him away.
\par }{\Q \VerseOne{1}I will praise
the {\ND{Lord}}
at all
times;
\par }{\Q my mouth
will continually
praise him.
\par }{\Q \VS{2}I will boast
in the
{\ND{Lord}};
\par }{\Q let
the oppressed
hear
and rejoice!
\par }{\Q \VS{3}Magnify
the {\ND{Lord}}
with
me!
\par }{\Q Let’s praise
his name
together!
\par }{\Q \VS{4}I sought
the
{\ND{Lord}}’s
help and he answered
me;
\par }{\Q he delivered
me from
all
my fears.
\par }{\Q \VS{5}Those who look
to him
for help are happy;
\par }{\Q their faces
are not
ashamed.
\par }{\Q \VS{6}This
oppressed
man cried
out and the
{\ND{Lord}}
heard;
\par }{\Q he saved
him from all
his troubles.
\par }{\Q \VS{7}The
{\ND{Lord}}’s
angel
camps
around
\par }{\Q the
{\ND{Lord}}’s loyal followers
and delivers them.
\par }{\Q \VS{8}Taste
and see
that
the {\ND{Lord}}
is good!
\par }{\Q How blessed
is the one
who takes shelter in him!
\par }{\Q \VS{9}Remain loyal
to the
{\ND{Lord}}, you chosen people
of his,

\par }{\Q for
his loyal followers
lack
nothing!
\par }{\Q \VS{10}Even
young lions
sometimes
lack food and are hungry,
\par }{\Q but those who seek
the {\ND{Lord}}
lack
no
good
thing.
\par }{\Q \VS{11}Come
children! Listen
to me!
\par }{\Q I will teach
you what it means to fear
the {\ND{Lord}}.
\par }{\Q \VS{12}Do you want
to really
live?

\par }{\Q Would you love
to live a long,
happy life?
\par }{\Q \VS{13}Then make sure
you don’t speak
evil
words
\par }{\Q or use deceptive
speech!
\par }{\Q \VS{14}Turn away
from evil
and do
what is right!

\par }{\Q Strive
for peace
and promote it!
\par }{\Q \VS{15}The
{\ND{Lord}}
pays attention
to
the godly
\par }{\Q and hears
their cry for help.
\par }{\Q \VS{16}But the
{\ND{Lord}}
opposes
evildoers
\par }{\Q and wipes
out all memory
of them from
the earth.
\par }{\Q \VS{17}The godly cry out
and the
{\ND{Lord}}
hears;
\par }{\Q he saves
them from all
their troubles.
\par }{\Q \VS{18}The
{\ND{Lord}}
is near
the brokenhearted;
\par }{\Q he delivers
those who are discouraged.
\par }{\Q \VS{19}The godly
face many
dangers,
\par }{\Q but the
{\ND{Lord}}
saves them from each one of them.
\par }{\Q \VS{20}He protects
all
his bones;
\par }{\Q not
one
of them
is broken.
\par }{\Q \VS{21}Evil
people
self-destruct;
\par }{\Q those who hate
the godly
are punished.
\par }{\Q \VS{22}The
{\ND{Lord}}
rescues
his servants;
\par }{\Q all
who take shelter
in him escape punishment.


\par }\PsalmChap{35}{\D 
By David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, fight
those who fight
with me!
\par }{\Q Attack
those who attack me!
\par }{\Q \VS{2}Grab
your small shield
and large shield,
\par }{\Q and rise
up to help me!
\par }{\Q \VS{3}Use
your spear
and lance
against
those who chase
me!
\par }{\Q Assure
me with these words: “I am
your deliverer!”
\par }{\Q \VS{4}May those who seek
my life
be embarrassed
and humiliated!
\par }{\Q May those who
plan
to harm
me be turned back
and ashamed!
\par }{\Q \VS{5}May they be
like wind-driven
chaff,
\par }{\Q as the
{\ND{Lord}}’s
angel
attacks them!
\par }{\Q \VS{6}May
their path
be dark
and slippery,
\par }{\Q as the
{\ND{Lord}}’s
angel
chases them!
\par }{\Q \VS{7}I did not harm
them, but they hid
a net
to catch me
\par }{\Q and dug
a pit
to trap me.
\par }{\Q \VS{8}Let destruction
take
them by surprise!

\par }{\Q Let the net
they hid
catch
them!
\par }{\Q Let them fall
into destruction!
\par }{\Q \VS{9}Then I
will rejoice
in the
{\ND{Lord}}
\par }{\Q and be happy
because of his deliverance.
\par }{\Q \VS{10}With all
my strength
I will say,
\par }{\Q “O
{\ND{Lord}}, who
can compare
to you?
\par }{\Q You rescue
the oppressed
from
those who try to overpower
them;

\par }{\Q the oppressed
and needy
from those who try to rob them.”
\par }{\Q \VS{11}Violent
men perjure
themselves,

\par }{\Q and falsely
accuse me.
\par }{\Q \VS{12}They repay
me evil
for
the good
I have done;

\par }{\Q I am overwhelmed
with sorrow.
\par }{\Q \VS{13}When they were sick,
I
wore
sackcloth,
\par }{\Q and refrained from eating food.
\par }{\Q (If I
am lying, may my prayers
go unanswered!)
\par }{\Q \VS{14}I mourned for them as I would for a friend
or my brother.
\par }{\Q I bowed down
in sorrow
as if I were mourning
for my mother.
\par }{\Q \VS{15}But when I stumbled,
they rejoiced
and gathered together;
\par }{\Q they gathered together
to ambush me.

\par }{\Q They tore
at me without
stopping to rest.
\par }{\Q \VS{16}When I tripped, they taunted
me relentlessly,

\par }{\Q and tried to bite me.
\par }{\Q \VS{17}O Lord,
how long
are you going to just stand there and watch
this?

\par }{\Q Rescue
me
from their destructive attacks;
\par }{\Q guard my life
from the young lions!
\par }{\Q \VS{18}Then I will give you thanks
in the great
assembly;
\par }{\Q I will praise
you before a large crowd
of people!
\par }{\Q \VS{19}Do not
let those who are my enemies
for no reason
gloat
over me!
\par }{\Q Do not let those who hate
me without cause
carry out their wicked schemes!
\par }{\Q \VS{20}For
they do not
try to make
peace
with others,

\par }{\Q but plan
ways
to deceive
those who are unsuspecting.
\par }{\Q \VS{21}They are ready
to devour
me;

\par }{\Q they say,
“Aha! Aha! We’ve got you!”
\par }{\Q \VS{22}But you take notice,

{\ND{Lord}}!
\par }{\Q O Lord,
do not
remain far
away from me!
\par }{\Q \VS{23}Rouse
yourself, wake
up and vindicate
me!

\par }{\Q My God
and Lord,
defend my just cause!
\par }{\Q \VS{24}Vindicate
me by your justice,
O
{\ND{Lord}}
my God!
\par }{\Q Do not
let them gloat over me!
\par }{\Q \VS{25}Do not
let them say
to themselves, “Aha! We have what we wanted!”

\par }{\Q Do not
let them say,
“We have devoured him!”
\par }{\Q \VS{26}May those who want to harm
me be totally
embarrassed
and ashamed!

\par }{\Q May those who arrogantly
taunt
me be covered with shame
and humiliation!
\par }{\Q \VS{27}May those who desire
my vindication
shout for joy
and rejoice!
\par }{\Q May they continually
say, “May the
{\ND{Lord}}
be praised,
for he wants
his servant
to be secure.”
\par }{\Q \VS{28}Then
I will tell others
about your justice,
\par }{\Q and praise
you all
day long.


\par }\PsalmChap{36}{\D 
For the music director;
written by the
{\ND{Lord}}’s
servant,
David;
an oracle.
\par }{\Q \VerseOne{1}An evil man
is rebellious
to the core.
\par }{\Q He does not
fear
God,
\par }{\Q \VS{2}for
he is too
proud
\par }{\Q to recognize
and give
up his sin.
\par }{\Q \VS{3}The words
he speaks
are sinful
and deceitful;
\par }{\Q he does not care
about doing
what is wise and right.
\par }{\Q \VS{4}He plans
ways to sin
while he lies in
bed;
\par }{\Q he is committed
to a sinful
lifestyle;
\par }{\Q he does not
reject
what is evil.
\par }{\Q \VS{5}O
{\ND{Lord}}, your loyal love
reaches to the sky;
\par }{\Q your faithfulness
to the clouds.
\par }{\Q \VS{6}Your justice
is like the highest
mountains,
\par }{\Q your fairness
like the deepest sea;
\par }{\Q you preserve
mankind
and the animal kingdom.
\par }{\Q \VS{7}How
precious
is your loyal love,
O God!
\par }{\Q The human
race
finds shelter
under your wings.
\par }{\Q \VS{8}They are filled with
food
from your house,
\par }{\Q and you allow them to drink
from the river
of your delicacies.
\par }{\Q \VS{9}For
you
are the one who gives
\par }{\Q and sustains
life.
\par }{\Q \VS{10}Extend
your loyal love
to your faithful followers,
\par }{\Q and vindicate
the morally
upright!
\par }{\Q \VS{11}Do not
let arrogant
men overtake
me,
\par }{\Q or
let evil men
make
me homeless!
\par }{\Q \VS{12}I can see the evildoers! They have fallen!

\par }{\Q They have been knocked
down and are unable
to get up!


\par }\PsalmChap{37}{\D 
By David.
\par }{\Q \VerseOne{1}Do not
fret
when wicked
men seem to succeed!

\par }{\Q Do not
envy
evildoers!
\par }{\Q \VS{2}For
they will quickly
dry up
like grass,
\par }{\Q and wither away
like plants.
\par }{\Q \VS{3}Trust
in the
{\ND{Lord}}
and do
what is right!
\par }{\Q Settle
in the land
and maintain
your integrity!
\par }{\Q \VS{4}Then you will take delight
in the
{\ND{Lord}},

\par }{\Q and he will answer
your prayers.
\par }{\Q \VS{5}Commit
your future to the
{\ND{Lord}}!

\par }{\Q Trust
in him,
and he will act
on your behalf.
\par }{\Q \VS{6}He will vindicate
you in broad daylight,
\par }{\Q and publicly defend your just cause.
\par }{\Q \VS{7}Wait patiently
for the
{\ND{Lord}}!

\par }{\Q Wait confidently
for him!
\par }{\Q Do not
fret
over the apparent success
of a sinner,
\par }{\Q a man
who carries
out wicked schemes!
\par }{\Q \VS{8}Do not be angry
and frustrated!

\par }{\Q Do not
fret! That only
leads to trouble!
\par }{\Q \VS{9}Wicked
men will be wiped
out,

\par }{\Q but those who rely
on the
{\ND{Lord}}
are
the ones who will possess
the land.
\par }{\Q \VS{10}Evil men will
soon
disappear;

\par }{\Q you will stare
at the spot where
they once
were, but they will be gone.
\par }{\Q \VS{11}But the oppressed
will possess
the land
\par }{\Q and enjoy
great
prosperity.
\par }{\Q \VS{12}Evil men
plot
against the godly
\par }{\Q and viciously
attack them.
\par }{\Q \VS{13}The Lord
laughs
in disgust at them,
\par }{\Q for
he knows that
their day
is coming.
\par }{\Q \VS{14}Evil
men draw
their swords
\par }{\Q and prepare
their bows,
\par }{\Q to bring down
the oppressed
and needy,
\par }{\Q and to slaughter
those who are godly.
\par }{\Q \VS{15}Their swords
will pierce
their own hearts,
\par }{\Q and their bows
will be broken.
\par }{\Q \VS{16}The little
bit that a godly
man owns is better
than
\par }{\Q the wealth
of many
evil men,
\par }{\Q \VS{17}for
evil
men will lose
their power,
\par }{\Q but the
{\ND{Lord}}
sustains
the godly.
\par }{\Q \VS{18}The
{\ND{Lord}}
watches
over the innocent
day
by day

\par }{\Q and they possess a permanent
inheritance.
\par }{\Q \VS{19}They will not
be ashamed
when hard
times
come;

\par }{\Q when famine
comes
they will have enough to eat.
\par }{\Q \VS{20}But
evil
men will die;
\par }{\Q the
{\ND{Lord}}’s
enemies
will be incinerated –
\par }{\Q they will go up in smoke.
\par }{\Q \VS{21}Evil
men borrow,
but do not
repay
their debt,
\par }{\Q but the godly
show compassion
and are generous.
\par }{\Q \VS{22}Surely
those favored by
the
{\ND{Lord}} will possess
the land,
\par }{\Q but those rejected
by him will be wiped out.
\par }{\Q \VS{23}The
{\ND{Lord}}
grants
success to the one
\par }{\Q whose behavior he finds commendable.
\par }{\Q \VS{24}Even if
he trips,
he will not
fall
headlong,
\par }{\Q for
the {\ND{Lord}}
holds
his hand.
\par }{\Q \VS{25}I was once young,
now I am old.
\par }{\Q I have never
seen
a godly
man abandoned,
\par }{\Q or his children
forced to search
for food.
\par }{\Q \VS{26}All
day
long he shows compassion
and lends
to others,

\par }{\Q and his children
are blessed.
\par }{\Q \VS{27}Turn
away from evil! Do
what is right!

\par }{\Q Then you will enjoy
lasting
security.
\par }{\Q \VS{28}For
the {\ND{Lord}}
promotes
justice,
\par }{\Q and never
abandons
his faithful followers.
\par }{\Q They are permanently
secure,
\par }{\Q but the children
of evil
men are wiped out.
\par }{\Q \VS{29}The godly
will possess
the land
\par }{\Q and will dwell
in it permanently.
\par }{\Q \VS{30}The godly
speak
wise
words
\par }{\Q and promote
justice.
\par }{\Q \VS{31}The law
of their God
controls
their thinking;

\par }{\Q their feet do not
slip.
\par }{\Q \VS{32}Evil
men set an ambush
for the godly
\par }{\Q and try
to kill them.
\par }{\Q \VS{33}But the
{\ND{Lord}}
does not
surrender
the godly,
\par }{\Q or
allow them to be condemned in a court of law.
\par }{\Q \VS{34}Rely
on the
{\ND{Lord}}! Obey
his commands!

\par }{\Q Then he will permit
you to possess
the land;
\par }{\Q you will see
the demise of evil men.
\par }{\Q \VS{35}I have seen
ruthless
evil
men

\par }{\Q growing
in influence, like a green
tree grows in its native soil.
\par }{\Q \VS{36}But then one
passes
by, and suddenly
they have disappeared!

\par }{\Q I looked
for them, but
they could not
be found.
\par }{\Q \VS{37}Take note
of the one who has integrity! Observe
the godly!

\par }{\Q For
the one
who promotes peace
has a future.
\par }{\Q \VS{38}Sinful rebels
are totally
destroyed;
\par }{\Q evil
men have no future.
\par }{\Q \VS{39}But the
{\ND{Lord}}
delivers
the godly;
\par }{\Q he protects
them in times
of trouble.
\par }{\Q \VS{40}The
{\ND{Lord}}
helps
them and rescues
them;
\par }{\Q he rescues
them from
evil
men and delivers
them,

\par }{\Q for
they seek his protection.


\par }\PsalmChap{38}{\D 
A psalm
of David,
written
to get God’s attention.

\par }{\Q \VerseOne{1}O

{\ND{Lord}}, do not
continue
to rebuke
me in your anger!
\par }{\Q Do not continue to punish me in your raging fury!
\par }{\Q \VS{2}For
your arrows
pierce
me,
\par }{\Q and your hand
presses me down.
\par }{\Q \VS{3}My whole
body
is sick
because of your judgment;

\par }{\Q I
am deprived
of health
because
of my sin.
\par }{\Q \VS{4}For
my sins
overwhelm
me;

\par }{\Q like a heavy
load,
they are too much for me to bear.
\par }{\Q \VS{5}My wounds
are infected
and starting
to smell,

\par }{\Q because
of my foolish sins.
\par }{\Q \VS{6}I am dazed
and completely humiliated;
\par }{\Q all
day
long I walk
around mourning.
\par }{\Q \VS{7}For
I am overcome
with shame
\par }{\Q and my whole body
is sick.
\par }{\Q \VS{8}I am numb with pain
and severely
battered;
\par }{\Q I groan loudly
because of the anxiety
I feel.
\par }{\Q \VS{9}O Lord,
you understand my heart’s
desire;
\par }{\Q my groaning
is not
hidden
from you.
\par }{\Q \VS{10}My heart
beats quickly;
\par }{\Q my strength
leaves
me;
\par }{\Q I can hardly
see.
\par }{\Q \VS{11}Because of my condition,
even my friends
and acquaintances
keep their distance;
\par }{\Q my neighbors stand
far away.
\par }{\Q \VS{12}Those who seek
my life
try to entrap
me;

\par }{\Q those who want
to harm
me speak
destructive
words;
\par }{\Q all
day
long they say deceitful things.
\par }{\Q \VS{13}But I am
like a deaf man – I hear nothing;
\par }{\Q I am like a mute who cannot speak.
\par }{\Q \VS{14}I am like
a man
who
cannot
hear
\par }{\Q and is incapable
of arguing
his defense.
\par }{\Q \VS{15}Yet
I wait
for
you, O
{\ND{Lord}}!
\par }{\Q You
will respond,
O Lord,
my God!
\par }{\Q \VS{16}I have prayed
for deliverance, because
otherwise
they will gloat
over me;

\par }{\Q when my foot
slips
they will arrogantly taunt me.
\par }{\Q \VS{17}For
I
am about
to stumble,
\par }{\Q and I am in
constant
pain.
\par }{\Q \VS{18}Yes,
I confess
my wrongdoing,
\par }{\Q and I am concerned
about my sins.
\par }{\Q \VS{19}But those who are my enemies
for no reason are numerous;
\par }{\Q those who hate
me without cause outnumber me.
\par }{\Q \VS{20}They repay
me evil
for the good I have done;
\par }{\Q though I have tried to do good
to them,
they hurl accusations
at me.
\par }{\Q \VS{21}Do not
abandon
me, O
{\ND{Lord}}!
\par }{\Q My God,
do not
remain far
away from me!
\par }{\Q \VS{22}Hurry
and help
me, O Lord,
my deliverer!


\par }\PsalmChap{39}{\D 
For the music director,
Jeduthun;
a psalm
of David.
\par }{\Q \VerseOne{1}I decided, “I will watch what I say
\par }{\Q and make sure
I do not
sin
with my
tongue.
\par }{\Q I will put a muzzle
over
my mouth
\par }{\Q while
in
the
presence of
an evil man.”
\par }{\Q \VS{2}I was stone
silent;
\par }{\Q I held back
the urge to speak.

\par }{\Q My frustration grew;
\par }{\Q \VS{3}my anxiety
intensified.
\par }{\Q As I thought about
it, I became impatient.
\par }{\Q Finally I spoke
these words:
\par }{\Q \VS{4}“O

{\ND{Lord}}, help me understand
my mortality
\par }{\Q and the brevity
of life!

\par }{\Q Let me realize
how quickly my life will pass!
\par }{\Q \VS{5}Look,
you make
my days
short-lived,
\par }{\Q and my life span
is nothing
from your perspective.
\par }{\Q Surely
all
people,
even those who
seem secure, are nothing but vapor.
\par }{\Q \VS{6}Surely
people
go
through life as mere ghosts.
\par }{\Q Surely they accumulate
worthless
wealth
\par }{\Q without
knowing
who
will eventually haul it away.”
\par }{\Q \VS{7}But now,
O Lord,
upon what
am I relying?
\par }{\Q You are my only hope!
\par }{\Q \VS{8}Deliver
me from all
my sins of rebellion!
\par }{\Q Do not
make
me the object of
fools’
insults!
\par }{\Q \VS{9}I am silent
and cannot
open
my mouth
\par }{\Q because
of what you
have done.
\par }{\Q \VS{10}Please stop
wounding
me!

\par }{\Q You have almost beaten
me
to death!
\par }{\Q \VS{11}You severely discipline
people
for their sins;
\par }{\Q like a moth
you slowly devour
their strength.
\par }{\Q Surely
all
people
are a mere vapor.
(Selah)
\par }{\Q \VS{12}Hear
my prayer,
O
{\ND{Lord}}!
\par }{\Q Listen
to
my cry for help!
\par }{\Q Do not
ignore
my sobbing!

\par }{\Q For
I am
dependent on you, like one residing outside his native land;
\par }{\Q I am at your mercy, just as all
my ancestors were.
\par }{\Q \VS{13}Turn
your angry gaze away
from
me, so I can be happy
\par }{\Q before
I pass
away.


\par }\PsalmChap{40}{\D 
For the music director;
By David,
a psalm.
\par }{\Q \VerseOne{1}I relied
completely on
the {\ND{Lord}},
\par }{\Q and he turned
toward me
\par }{\Q and heard
my cry for help.
\par }{\Q \VS{2}He lifted
me out
of the watery
pit,
\par }{\Q out of the slimy
mud.
\par }{\Q He placed
my feet
on
a rock
\par }{\Q and gave me secure
footing.
\par }{\Q \VS{3}He gave
me reason
to sing
a new
song,

\par }{\Q praising
our God.
\par }{\Q May many
see
what
God has done,
\par }{\Q so that they might
swear allegiance to him and trust
in the
{\ND{Lord}}!
\par }{\Q \VS{4}How blessed
is the one
who
trusts
in the
{\ND{Lord}}

\par }{\Q and does not
seek help
from
the proud
or from liars!
\par }{\Q \VS{5}O
{\ND{Lord}}, my God,
you have
accomplished
many
things;
\par }{\Q you
have done amazing
things and carried out
your purposes for us.
\par }{\Q No one can
thwart you!

\par }{\Q I want to declare them
and talk about them,
\par }{\Q but they
are too
numerous to recount!
\par }{\Q \VS{6}Receiving sacrifices
and offerings
are not
your primary
concern.

\par }{\Q You make that quite clear to me!

\par }{\Q You do not
ask
for burnt sacrifices and sin offerings.
\par }{\Q \VS{7}Then
I say,
\par }{\Q “Look! I come!
\par }{\Q What is written
in the scroll
pertains to me.
\par }{\Q \VS{8}I want
to do
what pleases
you, my God.
\par }{\Q Your law
dominates my thoughts.”
\par }{\Q \VS{9}I have told
the great
assembly
about your justice.
\par }{\Q Look! I spare
no
words!

\par }{\Q O
{\ND{Lord}}, you
know this is true.
\par }{\Q \VS{10}I have not
failed
to tell about
your justice;
\par }{\Q I spoke about your reliability
and deliverance;
\par }{\Q I have not
neglected
to tell
the great
assembly
about your loyal love
and faithfulness.
\par }{\Q \VS{11}O
{\ND{Lord}}, you
do not
withhold
your compassion
from
me.
\par }{\Q May your loyal love
and faithfulness
continually
protect me!
\par }{\Q \VS{12}For
innumerable
dangers
surround
me.
\par }{\Q My sins
overtake
me
\par }{\Q so I am unable
to see;
\par }{\Q they outnumber
the hairs
of my head
\par }{\Q so my strength
fails me.
\par }{\Q \VS{13}Please be willing, O
{\ND{Lord}}, to rescue
me!
\par }{\Q O
{\ND{Lord}}, hurry
and help me!
\par }{\Q \VS{14}May those who are trying to snatch away
my life
\par }{\Q be totally
embarrassed
and ashamed!

\par }{\Q May those who want
to harm
me
\par }{\Q be turned
back
and ashamed!
\par }{\Q \VS{15}May those
who say to me, “Aha! Aha!”
\par }{\Q be humiliated
and disgraced!
\par }{\Q \VS{16}May all
those who seek
you be happy
and rejoice
in you!
\par }{\Q May those who love
to experience your deliverance
say
continually,
\par }{\Q “May the
{\ND{Lord}}
be praised!”
\par }{\Q \VS{17}I am
oppressed
and needy!

\par }{\Q May the Lord
pay attention
to me!

\par }{\Q You
are my helper
and my deliverer!
\par }{\Q O my God,
do not
delay!


\par }\PsalmChap{41}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}How blessed
is the one who treats
the poor
properly!

\par }{\Q When
trouble
comes,
the {\ND{Lord}}
delivers him.
\par }{\Q \VS{2}May the
{\ND{Lord}}
protect
him and save his life!

\par }{\Q May he be blessed
in the land!
\par }{\Q Do not
turn
him over
to his enemies!
\par }{\Q \VS{3}The
{\ND{Lord}}
supports
him on
his sickbed;
\par }{\Q you completely
heal
him from his illness.
\par }{\Q \VS{4}As for me, I
said:

\par }{\Q “O
{\ND{Lord}}, have mercy
on me!
\par }{\Q Heal
me,
for
I have sinned against you!
\par }{\Q \VS{5}My enemies
ask
this cruel
question about me,

\par }{\Q ‘When
will he finally die
and be forgotten?’
\par }{\Q \VS{6}When
someone comes
to visit,
he pretends
to be friendly;
\par }{\Q he thinks
of ways to defame
me,

\par }{\Q and when he leaves
he slanders me.
\par }{\Q \VS{7}All
who hate
me whisper
insults about me to one another;

\par }{\Q they plan
ways to harm me.
\par }{\Q \VS{8}They say,
\par }{\Q ‘An awful
disease overwhelms
him,

\par }{\Q and now that he is bed-ridden
he will never recover.’
\par }{\Q \VS{9}Even
my close
friend
whom
I trusted,
\par }{\Q he who shared
meals
with me, has turned against me.
\par }{\Q \VS{10}As for you,
O
{\ND{Lord}}, have mercy
on me and raise
me up,
\par }{\Q so I can pay
them back!”
\par }{\Q \VS{11}By this
I know
that
you are pleased
with me,
\par }{\Q for
my enemy
does not
triumph
over me.
\par }{\Q \VS{12}As for me,
you uphold
me because of my integrity;
\par }{\Q you allow me permanent
access
to your presence.
\par }{\Q \VS{13}The
{\ND{Lord}}
God
of Israel
deserves
praise

\par }{\Q in the future
and forevermore!

\par }{\Q We agree! We agree!


\par }\PsalmChap{42}{\D 
For the music director;
a well-written song
by the Korahites.
\par }{\Q \VerseOne{1}As a deer
longs
for streams
of water,
\par }{\Q so
I long for
you, O God!
\par }{\Q \VS{2}I thirst
for God,
\par }{\Q for the living
God.
\par }{\Q I say, “When
will I be able
to go
and appear
in God’s
presence?”
\par }{\Q \VS{3}I cannot eat,
I weep
day
and night;
\par }{\Q all
day
long they say
to me, “Where
is your God?”
\par }{\Q \VS{4}I will remember
and weep!

\par }{\Q For
I was once walking along with the great throng
to
the temple
of God,
\par }{\Q shouting
and giving thanks
along with the crowd
as we celebrated the holy festival.
\par }{\Q \VS{5}Why
are you depressed,
O my soul?

\par }{\Q Why are you upset?

\par }{\Q Wait
for God!
\par }{\Q For
I will again
give thanks
\par }{\Q to my
God for his saving
intervention.
\par }{\Q \VS{6}I am depressed,
\par }{\Q so
I will pray to you while I am trapped here in the region
of the upper Jordan,
\par }{\Q from Hermon,
from Mount
Mizar.
\par }{\Q \VS{7}One deep stream
calls
out to another
at the sound
of your waterfalls;
\par }{\Q all
your billows
and waves
overwhelm me.
\par }{\Q \VS{8}By day
the {\ND{Lord}}
decrees
his loyal love,
\par }{\Q and by night
he gives me a song,
\par }{\Q a prayer
to the living
God.
\par }{\Q \VS{9}I will pray to God,
my high ridge:

\par }{\Q “Why
do you ignore
me?
\par }{\Q Why
must I walk
around mourning
\par }{\Q because my enemies
oppress me?”
\par }{\Q \VS{10}My enemies’
taunts
cut
into me to the bone,
\par }{\Q as they say
to me
all
day
long, “Where
is your God?”
\par }{\Q \VS{11}Why
are you depressed,
O my soul?

\par }{\Q Why
are you upset?

\par }{\Q Wait
for God!
\par }{\Q For
I will again
give thanks
\par }{\Q to my
God
for his saving
intervention.


\par }\PsalmChap{43}{\PP \par }{\Q \VerseOne{1}Vindicate
me, O God!
\par }{\Q Fight for me
against an ungodly
nation!
\par }{\Q Deliver
me from deceitful
and evil
men!
\par }{\Q \VS{2}For
you
are the God
who shelters
me.

\par }{\Q Why
do you reject
me?

\par }{\Q Why
must I walk
around mourning
\par }{\Q because my enemies
oppress me?
\par }{\Q \VS{3}Reveal
your light
and your faithfulness!
\par }{\Q They
will lead
me,

\par }{\Q they will escort
me back to
your holy
hill,
\par }{\Q and to
the place where you live.
\par }{\Q \VS{4}Then I will go
to
the altar
of God,
\par }{\Q to
the God
who gives me ecstatic
joy,
\par }{\Q so that I express my thanks
to you, O God,
my God,
with a harp.
\par }{\Q \VS{5}Why
are you depressed,
O my soul?

\par }{\Q Why
are you upset?

\par }{\Q Wait
for God!
\par }{\Q For
I will again
give thanks
\par }{\Q to my
God
for his saving
intervention.


\par }\PsalmChap{44}{\D 
For the music director;
by the Korahites,
a well-written song.
\par }{\Q \VerseOne{1}O God,
we have clearly
heard;
\par }{\Q our ancestors
have told
us
\par }{\Q what
you did
in their days,
\par }{\Q in ancient
times.
\par }{\Q \VS{2}You,
by your power,
defeated nations
and settled our fathers on their land;

\par }{\Q you crushed the people
living there
and enabled our ancestors to occupy it.
\par }{\Q \VS{3}For
they did not
conquer
the land
by their swords,
\par }{\Q and they did not
prevail
by their strength,
\par }{\Q but rather
by your power,
strength
and good favor,
\par }{\Q for
you were partial to them.
\par }{\Q \VS{4}You
are my king,
O God!
\par }{\Q Decree
Jacob’s
deliverance!
\par }{\Q \VS{5}By your power we will drive back
our enemies;
\par }{\Q by your strength
we will trample down
our foes!
\par }{\Q \VS{6}For
I do not
trust
in my bow,
\par }{\Q and I do not
prevail
by my sword.
\par }{\Q \VS{7}For
you deliver
us from our enemies;
\par }{\Q you humiliate
those who hate us.
\par }{\Q \VS{8}In God
I boast
all
day
long,
\par }{\Q and we will continually
give thanks
to your name.
(Selah)
\par }{\Q \VS{9}But you rejected
and embarrassed
us!
\par }{\Q You did not
go
into battle
with our armies.
\par }{\Q \VS{10}You made us retreat
from
the enemy.
\par }{\Q Those who hate
us take whatever they want from us.
\par }{\Q \VS{11}You handed
us over like sheep
to be eaten;
\par }{\Q you scattered
us among the nations.
\par }{\Q \VS{12}You sold
your people
for a pittance;
\par }{\Q you did not
ask a high price for them.
\par }{\Q \VS{13}You made
us an object of disdain
to our neighbors;
\par }{\Q those who live on our borders
taunt
and insult us.
\par }{\Q \VS{14}You made
us an object of ridicule
among the nations;
\par }{\Q foreigners
treat us with contempt.
\par }{\Q \VS{15}All
day
long I feel humiliated
\par }{\Q and am overwhelmed
with shame,
\par }{\Q \VS{16}before
the vindictive
enemy
\par }{\Q who ridicules
and insults me.
\par }{\Q \VS{17}All
this
has happened
to us, even though we have not
rejected
you

\par }{\Q or
violated
your covenant with us.
\par }{\Q \VS{18}We have not
been unfaithful,
\par }{\Q nor have we disobeyed
your commands.
\par }{\Q \VS{19}Yet
you have battered
us, leaving us a heap of ruins
overrun by wild dogs;
\par }{\Q you have covered
us with
darkness.
\par }{\Q \VS{20}If
we had rejected
our God,
\par }{\Q and spread out
our hands
in prayer to another
god,
\par }{\Q \VS{21}would not
God
discover
it,
\par }{\Q for
he knows
one’s
thoughts?
\par }{\Q \VS{22}Yet because
of you we are killed
all
day
long;
\par }{\Q we are treated
like sheep
at the slaughtering block.
\par }{\Q \VS{23}Rouse
yourself! Why
do you sleep,
O Lord?
\par }{\Q Wake
up! Do not
reject
us forever!
\par }{\Q \VS{24}Why
do you look the other way,

\par }{\Q and ignore
the way we are oppressed
and mistreated?
\par }{\Q \VS{25}For
we lie in
the dirt,
\par }{\Q with
our bellies
pressed to the ground.
\par }{\Q \VS{26}Rise
up and help
us!
\par }{\Q Rescue
us because
of your loyal love!


\par }\PsalmChap{45}{\D 
For the music director;
according to
the tune of “Lilies;”
by the Korahites,
a well-written poem,
a love
song.
\par }{\Q \VerseOne{1}My heart
is stirred
by a beautiful
song.
\par }{\Q I say,
“I
have composed
this special song
for the king;
\par }{\Q my tongue
is as skilled
as the stylus
of an experienced scribe.”
\par }{\Q \VS{2}You are the most handsome
of all men!

\par }{\Q You speak
in an impressive
and fitting manner!

\par }{\Q For this reason
God
grants you continual
blessings.
\par }{\Q \VS{3}Strap
your sword
to your thigh,
O warrior!

\par }{\Q Appear in your majestic
splendor!
\par }{\Q \VS{4}Appear in your majesty
and be victorious!

\par }{\Q Ride
forth for the sake of what is right,

\par }{\Q on behalf
of justice!

\par }{\Q Then your right hand
will accomplish
mighty acts!
\par }{\Q \VS{5}Your arrows
are sharp
\par }{\Q and penetrate
the hearts
of the king’s
enemies.
\par }{\Q Nations
fall at your feet.
\par }{\Q \VS{6}Your throne,
O God,
is permanent.
\par }{\Q The scepter
of your kingdom
is a scepter
of justice.
\par }{\Q \VS{7}You love
justice
and hate
evil.
\par }{\Q For this reason
God,
your God
has anointed
you

\par }{\Q with the oil
of joy,
elevating
you above
your companions.
\par }{\Q \VS{8}All
your garments
are perfumed with myrrh,
aloes,
and cassia.
\par }{\Q From
the luxurious palaces
comes the music of stringed
instruments that makes you happy.
\par }{\Q \VS{9}Princesses
are among your honored
guests,
\par }{\Q your bride stands
at your right hand,
wearing jewelry made with gold
from Ophir.
\par }{\Q \VS{10}Listen,
O princess!

\par }{\Q Observe
and pay attention!

\par }{\Q Forget
your homeland
and
your family!
\par }{\Q \VS{11}Then the king
will be attracted
by your beauty.
\par }{\Q After all,
he is your master! Submit to him!
\par }{\Q \VS{12}Rich people
from Tyre
\par }{\Q will seek your favor
by bringing
a gift.
\par }{\Q \VS{13}The princess
looks absolutely
magnificent,
\par }{\Q decked out in pearls and clothed
in a brocade trimmed
with gold.
\par }{\Q \VS{14}In embroidered robes
she is escorted to
the king.
\par }{\Q Her attendants,
the maidens of honor
who follow
her,
\par }{\Q are led before you.
\par }{\Q \VS{15}They are bubbling
with joy
as they walk in procession
\par }{\Q and enter
the royal
palace.
\par }{\Q \VS{16}Your sons
will carry
on the dynasty of your ancestors;
\par }{\Q you will make
them princes
throughout
the land.
\par }{\Q \VS{17}I will proclaim
your greatness
through the coming years,
\par }{\Q then
the nations
will praise
you forever.


\par }\PsalmChap{46}{\D 
For the music director;
by the Korahites;
according to
{\IT{the
alamoth}}
style; a song.
\par }{\Q \VerseOne{1}God
is our strong
refuge;
\par }{\Q he is truly
our helper
in times of trouble.
\par }{\Q \VS{2}For this reason
we do not
fear
when
the earth
shakes,
\par }{\Q and the mountains
tumble into
the depths
of the sea,
\par }{\Q \VS{3}when its waves
crash
and foam,
\par }{\Q and the mountains
shake
before the surging
sea. (Selah)
\par }{\Q \VS{4}The river’s
channels
bring joy
to the city
of God,
\par }{\Q the special, holy
dwelling place
of the sovereign One.
\par }{\Q \VS{5}God
lives within
it, it cannot
be moved.
\par }{\Q God
rescues
it at
the break
of dawn.
\par }{\Q \VS{6}Nations
are in uproar,
kingdoms
are overthrown.
\par }{\Q God gives
a shout,
the earth
dissolves.
\par }{\Q \VS{7}The
{\ND{Lord}}
who commands armies
is on our
side!

\par }{\Q The God
of Jacob
is our protector! (Selah)
\par }{\Q \VS{8}Come! Witness the exploits
of the {\ND{Lord}},
\par }{\Q who
brings
devastation
to the earth!
\par }{\Q \VS{9}He brings
an end
to wars
throughout
the earth;
\par }{\Q he shatters
the bow
and breaks
the spear;
\par }{\Q he burns
the shields
with fire.
\par }{\Q \VS{10}He says, “Stop your striving
and recognize
that
I am
God!
\par }{\Q I will be exalted
over the nations! I will be exalted
over the earth!”
\par }{\Q \VS{11}The
{\ND{Lord}}
who commands armies
is on our
side!

\par }{\Q The God
of Jacob
is our protector! (Selah)


\par }\PsalmChap{47}{\D 
For the music director;
by the Korahites;
a psalm.
\par }{\Q \VerseOne{1}All
you nations,
clap
your hands!
\par }{\Q Shout
out to God
in celebration!
\par }{\Q \VS{2}For
the sovereign

{\ND{Lord}}
is awe-inspiring;
\par }{\Q he is the great
king
who rules the whole
earth!
\par }{\Q \VS{3}He subdued
nations
beneath
us

\par }{\Q and countries
under
our feet.
\par }{\Q \VS{4}He picked
out for us a special
land

\par }{\Q to be a source of pride
for Jacob,
whom
he loves.
(Selah)
\par }{\Q \VS{5}God
has ascended
his throne amid
loud shouts;

\par }{\Q the {\ND{Lord}}
has ascended his throne amid the blaring
of ram’s horns.
\par }{\Q \VS{6}Sing
to God! Sing!
\par }{\Q Sing
to our king! Sing!
\par }{\Q \VS{7}For
God
is king
of the whole
earth!
\par }{\Q Sing
a well-written song!
\par }{\Q \VS{8}God
reigns
over
the nations!
\par }{\Q God
sits
on
his holy
throne!
\par }{\Q \VS{9}The nobles
of the nations
assemble,
\par }{\Q along with the people
of the God
of Abraham,
\par }{\Q for
God
has authority over the rulers
of the earth.
\par }{\Q He is highly
exalted!


\par }\PsalmChap{48}{\D 
A song,
a psalm
by the Korahites.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}
is great
and certainly worthy
of praise
\par }{\Q in the city
of our God,
his holy
hill.
\par }{\Q \VS{2}It is lofty
and pleasing to look
at,

\par }{\Q a source of joy
to the whole
earth.
\par }{\Q Mount
Zion
resembles the peaks
of Zaphon;
\par }{\Q it is the city
of the great
king.
\par }{\Q \VS{3}God
is in its fortresses;
\par }{\Q he reveals
himself as its defender.
\par }{\Q \VS{4}For
look,
the kings
assemble;
\par }{\Q they advance
together.
\par }{\Q \VS{5}As soon as they
see,
they are shocked;
\par }{\Q they are terrified,
they quickly retreat.
\par }{\Q \VS{6}Look at them shake uncontrollably,
\par }{\Q like
a woman
writhing
in childbirth.
\par }{\Q \VS{7}With an east
wind
\par }{\Q you shatter
the large ships.
\par }{\Q \VS{8}We heard
about God’s mighty
deeds, now we have seen
them,

\par }{\Q in the city
of the
{\ND{Lord}}, the invincible
Warrior,

\par }{\Q in the city
of our God.
\par }{\Q God
makes
it permanently
secure. (Selah)
\par }{\Q \VS{9}We reflect
on your loyal love,
O God,
\par }{\Q within
your temple.
\par }{\Q \VS{10}The praise
you receive
as far away as the ends
of the earth
\par }{\Q is worthy
of your reputation,
O God.
\par }{\Q You execute
justice!
\par }{\Q \VS{11}Mount
Zion
rejoices;
\par }{\Q the towns
of Judah
are happy,
\par }{\Q because
of your acts of judgment.
\par }{\Q \VS{12}Walk around
Zion! Encircle
it!
\par }{\Q Count
its towers!
\par }{\Q \VS{13}Consider
its defenses!

\par }{\Q Walk through
its fortresses,
\par }{\Q so
you can tell
the next generation about it!
\par }{\Q \VS{14}For
God,
our God,
is our defender
forever!

\par }{\Q He guides us!


\par }\PsalmChap{49}{\D 
For the music director,
a psalm
by the Korahites.
\par }{\Q \VerseOne{1}Listen
to this,
all
you nations!
\par }{\Q Pay attention,
all
you inhabitants
of the world!
\par }{\Q \VS{2}Pay attention,
all
you people,
\par }{\Q both
rich
and poor!
\par }{\Q \VS{3}I will declare
a wise
saying;
\par }{\Q I will share my profound thoughts.
\par }{\Q \VS{4}I will learn
a song that imparts wisdom;
\par }{\Q I will then sing
my insightful song
to the accompaniment of a harp.
\par }{\Q \VS{5}Why
should I be afraid
in times
of trouble,
\par }{\Q when the sinful
deeds of deceptive men threaten
to overwhelm me?
\par }{\Q \VS{6}They trust
in their wealth
\par }{\Q and boast
in their great
riches.
\par }{\Q \VS{7}Certainly
a man cannot
rescue
his brother;
\par }{\Q he cannot pay
God
an adequate
ransom price
\par }{\Q \VS{8}(the ransom
price
for a human life
is too high,
\par }{\Q and people go to their final
destiny),
\par }{\Q \VS{9}so that he might continue
to live
forever
\par }{\Q and not
experience
death.
\par }{\Q \VS{10}Surely
one sees
that even wise people
die;
\par }{\Q fools
and spiritually insensitive
people all
pass away
\par }{\Q and leave
their wealth
to others.
\par }{\Q \VS{11}Their
grave becomes
their
permanent residence,
\par }{\Q their eternal
dwelling place.
\par }{\Q They name
their lands
after themselves,
\par }{\Q \VS{12}but, despite
their wealth, people
do not
last,

\par }{\Q they are like
animals
that perish.
\par }{\Q \VS{13}This
is the destiny
of fools,

\par }{\Q and of those who approve of their philosophy.
(Selah)
\par }{\Q \VS{14}They will travel to Sheol
like sheep,
\par }{\Q with death
as their shepherd.
\par }{\Q The godly
will rule over
them when the day of vindication dawns;
\par }{\Q Sheol
will consume their bodies
and they will no longer live
in impressive houses.
\par }{\Q \VS{15}But
God
will rescue
my life
from the power
of Sheol;
\par }{\Q certainly
he will pull
me to safety. (Selah)
\par }{\Q \VS{16}Do not
be afraid
when
a man
becomes
rich
\par }{\Q and his wealth multiplies!
\par }{\Q \VS{17}For
he will take
nothing
with him when he dies;
\par }{\Q his wealth
will not
follow
him down
into the grave.
\par }{\Q \VS{18}He pronounces this
blessing
on himself
while he is alive:
\par }{\Q “May men praise
you, for
you have done well!”
\par }{\Q \VS{19}But he will join
his ancestors;
\par }{\Q they will never
again see
the light of day.
\par }{\Q \VS{20}Wealthy
people
do not
understand;
\par }{\Q they are like
animals
that perish.


\par }\PsalmChap{50}{\D 
A psalm
by Asaph.
\par }{\Q \VerseOne{1}El,
God,
the {\ND{Lord}}
speaks,
\par }{\Q and summons
the earth
to come from the east
and west.
\par }{\Q \VS{2}From Zion,
the most
beautiful
of all places,

\par }{\Q God
comes in splendor.
\par }{\Q \VS{3}Our God
approaches
and is not
silent;
\par }{\Q consuming
fire
goes ahead
of him
\par }{\Q and all around
him a storm rages.
\par }{\Q \VS{4}He summons
the heavens
above,
\par }{\Q as well as the earth,
so that he might judge
his people.
\par }{\Q \VS{5}He says:

\par }{\Q “Assemble
my covenant people
before me,

\par }{\Q those who ratified
a covenant
with me by
sacrifice!”
\par }{\Q \VS{6}The heavens
declare
his fairness,
\par }{\Q for
God
is judge.
(Selah)
\par }{\Q \VS{7}He says:

\par }{\Q “Listen
my people! I am speaking!
\par }{\Q Listen Israel! I am accusing
you!

\par }{\Q I am
God,
your God!
\par }{\Q \VS{8}I am not
condemning
you because of your sacrifices,
\par }{\Q or because of your burnt sacrifices
that you continually offer me.
\par }{\Q \VS{9}I do not
need to take
a bull
from your household
\par }{\Q or goats
from your sheepfolds.
\par }{\Q \VS{10}For
every
wild animal
in the forest
belongs to me,
\par }{\Q as well as the cattle
that graze on a thousand
hills.
\par }{\Q \VS{11}I
keep track
of every
bird
in the hills,
\par }{\Q and the insects
of the field
are mine.
\par }{\Q \VS{12}Even if
I were hungry,
I would not
tell
you,
\par }{\Q for
the world
and all it contains belong to me.
\par }{\Q \VS{13}Do I eat
the flesh
of bulls?
\par }{\Q Do I drink
the blood
of goats?
\par }{\Q \VS{14}Present
to God
a thank-offering!
\par }{\Q Repay
your vows
to the sovereign One!
\par }{\Q \VS{15}Pray
to me when
you are in trouble!

\par }{\Q I will deliver
you, and you will honor me!”
\par }{\Q \VS{16}God
says
this to the evildoer:

\par }{\Q “How
can you declare
my commands,
\par }{\Q and talk
about my covenant?
\par }{\Q \VS{17}For you
hate
instruction
\par }{\Q and reject
my words.
\par }{\Q \VS{18}When
you see
a thief,
you join
him;
\par }{\Q you associate
with
men
who are unfaithful
to their wives.
\par }{\Q \VS{19}You do damage
with words,
\par }{\Q and use
your tongue
to deceive.
\par }{\Q \VS{20}You plot against
your brother;
\par }{\Q you slander
your own brother.
\par }{\Q \VS{21}When you did
these
things, I was silent,
\par }{\Q so you thought
I was exactly like
you.

\par }{\Q But now
I will
condemn
you
\par }{\Q and state
my case against you!
\par }{\Q \VS{22}Carefully
consider
this,
you who reject
God!

\par }{\Q Otherwise
I will rip you to shreds
\par }{\Q and no
one will be able to rescue you.
\par }{\Q \VS{23}Whoever presents a thank-offering
honors
me.

\par }{\Q To whoever obeys my commands,
I will reveal
my power
to deliver.”


\par }\PsalmChap{51}{\D 
For the music director;
a psalm
of David,
written when Nathan
the prophet
confronted
him after David’s affair
with
Bathsheba.
\par }{\Q \VerseOne{1}Have mercy
on me, O God,
because of your loyal love!
\par }{\Q Because of your great
compassion,
wipe away
my rebellious acts!
\par }{\Q \VS{2}Wash
away my wrongdoing!

\par }{\Q Cleanse
me of my sin!
\par }{\Q \VS{3}For
I
am aware
of my rebellious
acts;
\par }{\Q I am forever
conscious of my sin.
\par }{\Q \VS{4}Against you – you above all – I have sinned;
\par }{\Q I have done what is evil in your sight.
\par }{\Q So you are just when you confront me;
\par }{\Q you are right when you condemn me.
\par }{\Q \VS{5}Look,
I was guilty of sin
from birth,
\par }{\Q a sinner
the moment my mother
conceived me.
\par }{\Q \VS{6}Look,
you desire
integrity in the inner man;
\par }{\Q you want me to possess
wisdom.
\par }{\Q \VS{7}Sprinkle
me with water
and I will be pure;
\par }{\Q wash
me and I will be whiter
than snow.
\par }{\Q \VS{8}Grant
me the ultimate joy
of being forgiven!

\par }{\Q May the bones
you crushed
rejoice!
\par }{\Q \VS{9}Hide
your face
from my sins!
\par }{\Q Wipe
away all
my guilt!
\par }{\Q \VS{10}Create
for me a pure
heart,
O God!

\par }{\Q Renew
a resolute
spirit
within me!
\par }{\Q \VS{11}Do not
reject
me!

\par }{\Q Do not
take
your Holy
Spirit
away from me!
\par }{\Q \VS{12}Let me again
experience the joy
of your deliverance!
\par }{\Q Sustain
me by giving me the desire to obey!
\par }{\Q \VS{13}Then I will teach
rebels
your merciful ways,
\par }{\Q and sinners
will turn
to you.
\par }{\Q \VS{14}Rescue
me from the guilt of murder,
O God,
the God
who delivers
me!
\par }{\Q Then my tongue
will shout
for joy because of your deliverance.
\par }{\Q \VS{15}O Lord,
give me the words!

\par }{\Q Then
my mouth
will praise you.
\par }{\Q \VS{16}Certainly
you do not
want
a sacrifice,
or else I would
offer it;

\par }{\Q you do not
desire a burnt sacrifice.
\par }{\Q \VS{17}The sacrifices
God
desires are a humble spirit –
\par }{\Q O God,
a humble and repentant
heart
you will
not
reject.
\par }{\Q \VS{18}Because you favor
Zion,
do what
is good for her!

\par }{\Q Fortify
the walls
of Jerusalem!
\par }{\Q \VS{19}Then
you will accept
the proper
sacrifices,
burnt sacrifices
and whole
offerings;
\par }{\Q then
bulls
will be sacrificed
on
your altar.


\par }\PsalmChap{52}{\D 
For the music director;
a well-written song
by David.
It was written when Doeg
the Edomite
went
and informed
Saul: “David
has arrived
at
the home
of Ahimelech.”
\par }{\Q \VerseOne{1}Why
do you boast
about your evil
plans, O powerful man?
\par }{\Q God’s
loyal love
protects me all
day long!
\par }{\Q \VS{2}Your tongue
carries out your destructive
plans;
\par }{\Q it is as effective as a sharp
razor,
O deceiver.
\par }{\Q \VS{3}You love
evil
more than good,
\par }{\Q lies
more than speaking
the truth.
(Selah)
\par }{\Q \VS{4}You love
to use all
the words
that destroy,
\par }{\Q and the tongue
that deceives.
\par }{\Q \VS{5}Yet
God
will make you a permanent heap of ruins.

\par }{\Q He will
scoop
you up and remove you from
your home;
\par }{\Q he will
uproot
you from
the land
of the living.
(Selah)
\par }{\Q \VS{6}When the godly
see
this, they will be
filled with awe,
\par }{\Q and will mock the evildoer, saying:
\par }{\Q \VS{7}“Look,
here is the man
who would not
make
God
his protector!
\par }{\Q He trusted
in his great
wealth
\par }{\Q and was confident
about his plans to destroy others.”
\par }{\Q \VS{8}But I am
like a flourishing
olive tree
in the house
of God;
\par }{\Q I continually
trust
in God’s
loyal love.
\par }{\Q \VS{9}I will continually
thank
you when
you execute
judgment;

\par }{\Q I will rely
on you,
for
your loyal followers
know
you are good.


\par }\PsalmChap{53}{\D 
For the music director;
{\IT{according to the
machalath}}
style; a well-written song
by David.
\par }{\Q \VerseOne{1}Fools
say
to themselves, “There is no
God.”
\par }{\Q They sin and commit
evil
deeds;

\par }{\Q none
of them does what
is right.
\par }{\Q \VS{2}God
looks
down from heaven
at the human
race,

\par }{\Q to see
if there is anyone
who is wise
and seeks
God.
\par }{\Q \VS{3}Everyone
rejects
God;

\par }{\Q they are all
morally corrupt.
\par }{\Q None
of them does
what is right,
\par }{\Q not
even
one!
\par }{\Q \VS{4}All
those who behave
wickedly
do not
understand –
\par }{\Q those who devour
my people
as if they were eating
bread,
\par }{\Q and do not
call
out to God.
\par }{\Q \VS{5}They are absolutely
terrified,
\par }{\Q even by things
that do not
normally
cause fear.
\par }{\Q For
God
annihilates
those
who attack you.

\par }{\Q You are able to humiliate
them because
God
has rejected them.
\par }{\Q \VS{6}I wish
the deliverance
of Israel
would come from Zion!
\par }{\Q When God
restores the well-being
of his people,
\par }{\Q may
Jacob
rejoice,
\par }{\Q may Israel
be happy!


\par }\PsalmChap{54}{\D 
For the music director,
to be accompanied by stringed instruments;
a well-written song
by David.
It was written when the Ziphites
came
and informed
Saul: “David
is hiding
with
us.”

\par }{\Q \VerseOne{1}O God,
deliver
me by your name!

\par }{\Q Vindicate
me by your power!
\par }{\Q \VS{2}O God,
listen
to my prayer!
\par }{\Q Pay
attention
to what I say!
\par }{\Q \VS{3}For
foreigners
attack me;

\par }{\Q ruthless
men, who do not
respect
God,
seek
my life.
(Selah)
\par }{\Q \VS{4}Look,
God
is my deliverer!

\par }{\Q The Lord
is among those who
support
me.
\par }{\Q \VS{5}May
those who wait
to ambush me
be repaid
for their evil!

\par }{\Q As a demonstration
of your faithfulness,
destroy them!
\par }{\Q \VS{6}With a freewill
offering I will sacrifice
to you!
\par }{\Q I will give thanks
to your name,
O
{\ND{Lord}}, for
it is good!
\par }{\Q \VS{7}Surely
he rescues
me from
all
trouble,
\par }{\Q and I triumph
over my enemies.


\par }\PsalmChap{55}{\D 
For the music director,
to be accompanied by stringed instruments;
a well-written song
by David.
\par }{\Q \VerseOne{1}Listen,
O God,
to my prayer!
\par }{\Q Do not
ignore
my appeal for mercy!
\par }{\Q \VS{2}Pay attention
to me and answer
me!
\par }{\Q I am so upset
and distressed,
I am beside myself,
\par }{\Q \VS{3}because of what
the enemy
says,

\par }{\Q and because of how the wicked
pressure
me,

\par }{\Q for
they hurl
trouble
down upon me

\par }{\Q and angrily
attack me.
\par }{\Q \VS{4}My heart
beats
violently within
me;
\par }{\Q the horrors
of death
overcome me.
\par }{\Q \VS{5}Fear
and panic
overpower
me;

\par }{\Q terror
overwhelms me.
\par }{\Q \VS{6}I say, “I wish
I had wings
like a dove!
\par }{\Q I would fly
away and settle in a safe place!
\par }{\Q \VS{7}Look,
I will escape
to a distant
place;
\par }{\Q I will stay
in the wilderness.
(Selah)
\par }{\Q \VS{8}I will hurry off
to a place
that is safe
\par }{\Q from the strong wind
and the gale.”
\par }{\Q \VS{9}Confuse
them, O Lord!
\par }{\Q Frustrate
their plans!

\par }{\Q For
I see
violence
and conflict
in the city.
\par }{\Q \VS{10}Day
and night
they walk around
on
its walls,
\par }{\Q while wickedness
and destruction
are within it.
\par }{\Q \VS{11}Disaster
is within
it;
\par }{\Q violence
and deceit
do not
depart
from its public square.
\par }{\Q \VS{12}Indeed,
it is not
an enemy
who insults
me,
\par }{\Q or else I could bear
it;
\par }{\Q it is not
one who hates
me who arrogantly
taunts me,

\par }{\Q or else I could hide
from him.
\par }{\Q \VS{13}But it is you,
a man
like me,
\par }{\Q my close friend
in whom I confided.
\par }{\Q \VS{14}We would share personal
thoughts
with each other;

\par }{\Q in God’s
temple
we would walk
together
among the crowd.
\par }{\Q \VS{15}May death destroy
them!

\par }{\Q May they go down
alive
into Sheol!

\par }{\Q For
evil
is in their dwelling place
and in their midst.
\par }{\Q \VS{16}As for me, I
will call
out to
God,
\par }{\Q and the
{\ND{Lord}}
will deliver me.
\par }{\Q \VS{17}During the evening,
morning,
and noontime
\par }{\Q I will lament
and moan,
\par }{\Q and he will hear
me.
\par }{\Q \VS{18}He will rescue
me and protect
me from those who attack me,

\par }{\Q even though
they greatly outnumber
me.
\par }{\Q \VS{19}God,
the one who has
reigned
as king from long ago,
\par }{\Q will hear
and humiliate
them. (Selah)
\par }{\Q They refuse
to change,
\par }{\Q and do not
fear
God.
\par }{\Q \VS{20}He attacks
his friends;
\par }{\Q he breaks
his solemn promises to them.
\par }{\Q \VS{21}His words are as smooth
as butter,
\par }{\Q but he harbors animosity
in his heart.
\par }{\Q His words
seem softer
than oil,
\par }{\Q but they
are really like sharp swords.
\par }{\Q \VS{22}Throw
your burden
upon
the {\ND{Lord}},
\par }{\Q and he will sustain
you.

\par }{\Q He will never
allow
the godly
to be upended.
\par }{\Q \VS{23}But you,
O God,
will bring them down
to the deep Pit.
\par }{\Q Violent
and deceitful
people
will not
live even half a normal
lifespan.
\par }{\Q But as for me, I
trust in you.


\par }\PsalmChap{56}{\D 
For the music director;
{\IT{according to the
yonath-elem-rechovim}}
style; a prayer
of David,
written when the
Philistines
captured
him in Gath.
\par }{\Q \VerseOne{1}Have mercy
on me, O God,
for
men
are attacking
me!

\par }{\Q All
day
long hostile enemies
are tormenting me.
\par }{\Q \VS{2}Those who anticipate
my defeat
attack me all
day
long.
\par }{\Q Indeed,
many
are fighting
against me, O Exalted One.
\par }{\Q \VS{3}When
I am
afraid,
\par }{\Q I
trust in you.
\par }{\Q \VS{4}In God
– I boast
in his promise –
\par }{\Q in God
I trust,
I am not
afraid.
\par }{\Q What
can mere
men do to me?
\par }{\Q \VS{5}All
day
long they cause me trouble;

\par }{\Q they make a habit
of plotting
my demise.
\par }{\Q \VS{6}They
stalk
and lurk;
\par }{\Q they
watch
my every step,

\par }{\Q as they prepare to take my life.
\par }{\Q \VS{7}Because
they are bent on violence,
do not let them escape!

\par }{\Q In your anger
bring down
the nations,
O God!
\par }{\Q \VS{8}You keep
track
of my misery.
\par }{\Q Put
my tears
in your leather container!

\par }{\Q Are they not
recorded in your scroll?
\par }{\Q \VS{9}My enemies
will turn
back
when
I cry
out to you for help;

\par }{\Q I know
that
God is on my side.
\par }{\Q \VS{10}In God
– I boast
in his promise –
\par }{\Q in the
{\ND{Lord}}
–
I boast
in his promise –
\par }{\Q \VS{11}in God
I trust,
I am not
afraid.
\par }{\Q What
can mere men
do to me?
\par }{\Q \VS{12}I
am obligated
to fulfill
the vows
I made to you, O God;
\par }{\Q I will give you the thank-offerings you deserve,
\par }{\Q \VS{13}when
you deliver
my life
from death.
\par }{\Q You keep my feet
from stumbling,
\par }{\Q so that I might serve
God
as I enjoy life.


\par }\PsalmChap{57}{\D 
For the music director;
{\IT{according to the
al-tashcheth}} style;
a prayer
of David,
written when he fled
from Saul
into the cave.
\par }{\Q \VerseOne{1}Have mercy
on me, O God! Have mercy
on me!
\par }{\Q For
in you I have taken shelter.
\par }{\Q In the shadow
of your wings
I take shelter
\par }{\Q until
trouble passes.
\par }{\Q \VS{2}I cry
out for help to the sovereign
God,
\par }{\Q to the God
who vindicates me.
\par }{\Q \VS{3}May he send
help from heaven
and deliver
me

\par }{\Q from my enemies
who hurl
insults! (Selah)
\par }{\Q May God
send
his loyal love
and faithfulness!
\par }{\Q \VS{4}I
am surrounded by lions;
\par }{\Q I lie down
among
those who want to devour
me;

\par }{\Q men
whose teeth
are spears
and arrows,
\par }{\Q whose tongues
are a sharp
sword.
\par }{\Q \VS{5}Rise
up above
the sky,
O God!
\par }{\Q May your splendor
cover the whole
earth!
\par }{\Q \VS{6}They have prepared
a net
to trap
me;
\par }{\Q I am discouraged.
\par }{\Q They have dug
a pit
for me.

\par }{\Q They will fall
into
it! (Selah)
\par }{\Q \VS{7}I am determined,
O God! I am determined!
\par }{\Q I will sing and praise you!
\par }{\Q \VS{8}Awake,
my
soul!

\par }{\Q Awake,
O stringed instrument
and harp!
\par }{\Q I will wake
up at dawn!
\par }{\Q \VS{9}I will give you thanks
before the nations,
O Master!
\par }{\Q I will sing praises
to you before foreigners!
\par }{\Q \VS{10}For
your loyal love
extends
beyond
the sky,
\par }{\Q and your faithfulness
reaches
the clouds.
\par }{\Q \VS{11}Rise
up above
the sky,
O God!
\par }{\Q May your splendor
cover the whole
earth!


\par }\PsalmChap{58}{\D 
For the music director;
{\IT{according to the
al-tashcheth}} style;
a prayer
of David.
\par }{\Q \VerseOne{1}Do you rulers really
pronounce
just
decisions?

\par }{\Q Do you judge
people
fairly?
\par }{\Q \VS{2}No! You plan
how to do what
is unjust;
\par }{\Q you deal
out violence
in the earth.
\par }{\Q \VS{3}The wicked
turn aside
from birth;
\par }{\Q liars
go astray
as soon as they are born.
\par }{\Q \VS{4}Their venom
is like
that of a snake,
\par }{\Q like
a deaf
serpent
that does not
hear,
\par }{\Q \VS{5}that
does not
respond
to the magicians,
\par }{\Q or to a
skilled snake-charmer.
\par }{\Q \VS{6}O God,
break
the teeth
in their mouths!
\par }{\Q Smash
the jawbones
of the lions,
O
{\ND{Lord}}!
\par }{\Q \VS{7}Let
them disappear
like
water
that flows
away!

\par }{\Q Let them wither like
grass!
\par }{\Q \VS{8}Let them be like
a snail
that melts away
as it moves
along!

\par }{\Q Let them be like stillborn
babies
that never
see
the sun!
\par }{\Q \VS{9}Before
the kindling
is even placed under your pots,
\par }{\Q he will sweep it away
along with both the raw
and cooked meat.
\par }{\Q \VS{10}The godly
will rejoice
when
they see
vengeance
carried
out;
\par }{\Q they will bathe
their feet
in the blood
of the wicked.
\par }{\Q \VS{11}Then observers
will say,
\par }{\Q “Yes
indeed, the godly
are rewarded!

\par }{\Q Yes indeed,
there is
a God
who judges
in the earth!”


\par }\PsalmChap{59}{\D 
For the music director;
{\IT{according to the
al-tashcheth}} style;
a prayer
of David,
written when Saul
sent
men to surround
his house
and murder him.
\par }{\Q \VerseOne{1}Deliver
me from my enemies,
my God!
\par }{\Q Protect
me from those who attack me!
\par }{\Q \VS{2}Deliver
me from evildoers!

\par }{\Q Rescue
me from violent
men!
\par }{\Q \VS{3}For
look,
they wait to ambush
me;
\par }{\Q powerful men stalk
me,
\par }{\Q but not
because
I have rebelled
or
sinned,
O
{\ND{Lord}}.
\par }{\Q \VS{4}Though I have done nothing wrong, they
are anxious
to attack.

\par }{\Q Spring
into action
and help
me! Take notice of me!
\par }{\Q \VS{5}You,
O
{\ND{Lord}}
God,
the invincible
warrior, the God
of Israel,
\par }{\Q rouse
yourself and punish
all
the nations!
\par }{\Q Have no
mercy
on any
treacherous
evildoers! (Selah)
\par }{\Q \VS{6}They return
in the evening;
\par }{\Q they growl
like a dog
\par }{\Q and prowl around outside
the city.
\par }{\Q \VS{7}Look,
they hurl insults
at me
\par }{\Q and openly threaten to kill me,
\par }{\Q for
they say,

\par }{\Q “Who
hears?”
\par }{\Q \VS{8}But you,
O
{\ND{Lord}}, laugh
in disgust at them;

\par }{\Q you taunt
all
the nations.
\par }{\Q \VS{9}You are my source
of strength! I will wait for
you!

\par }{\Q For
God
is my refuge.
\par }{\Q \VS{10}The God
who loves
me
will help me;
\par }{\Q God
will enable
me to triumph over my enemies.
\par }{\Q \VS{11}Do not
strike
them dead suddenly,
\par }{\Q because then
my people
might
forget
the lesson.

\par }{\Q Use your power to make them homeless vagabonds
and then bring them down,
\par }{\Q O Lord
who shields us!
\par }{\Q \VS{12}They speak
sinful
words.
\par }{\Q So let them be trapped
by their own pride
\par }{\Q and by the curses
and lies
they speak!
\par }{\Q \VS{13}Angrily
wipe
them out! Wipe
them out so they vanish!
\par }{\Q Let them
know
that
God
rules
\par }{\Q in Jacob
and to the ends
of the earth! (Selah)
\par }{\Q \VS{14}They return
in the evening;
\par }{\Q they growl
like a dog
\par }{\Q and prowl around outside
the city.
\par }{\Q \VS{15}They
wander
around looking for something to eat;
\par }{\Q they refuse to sleep
until they are full.
\par }{\Q \VS{16}As for me, I
will sing
about your strength;
\par }{\Q I will praise
your loyal love
in the morning.
\par }{\Q For
you are my refuge
\par }{\Q and my place of shelter
when
I face trouble.
\par }{\Q \VS{17}You are my source of strength! I will sing praises
to you!

\par }{\Q For
God
is my refuge,
the God
who loves me.


\par }\PsalmChap{60}{\D 
For the music director;
according
{\IT{to the
shushan-eduth}} style;
a prayer
of David
written to instruct
others. It was written when he fought
against
Aram Naharaim
and Aram-Zobah.
That was when Joab
turned back
and struck down
12,000
Edomites
in the
Valley
of Salt.
\par }{\Q \VerseOne{1}O God,
you have rejected
us.

\par }{\Q You suddenly turned on
us in your anger.
\par }{\Q Please restore us!
\par }{\Q \VS{2}You made the earth
quake;
you split
it open.

\par }{\Q Repair
its breaches,
for
it is ready to fall.
\par }{\Q \VS{3}You have made your people
experience
hard
times;

\par }{\Q you have made us drink
intoxicating
wine.
\par }{\Q \VS{4}You have given
your loyal followers
a rallying flag,
\par }{\Q so that they might seek safety
from the bow. (Selah)
\par }{\Q \VS{5}Deliver
by your power
and answer
me,

\par }{\Q so that
the ones you love
may be safe.
\par }{\Q \VS{6}God
has spoken
in his sanctuary:

\par }{\Q “I will triumph! I will parcel
out Shechem;
\par }{\Q the Valley
of Succoth
I will measure off.
\par }{\Q \VS{7}Gilead
belongs to me,
\par }{\Q as does Manasseh!

\par }{\Q Ephraim
is my helmet,
\par }{\Q Judah
my royal scepter.
\par }{\Q \VS{8}Moab
is my washbasin.
\par }{\Q I will make
Edom
serve
me.

\par }{\Q I will shout in triumph
over
Philistia.”
\par }{\Q \VS{9}Who
will lead me into
the fortified
city?
\par }{\Q Who
will bring me
to Edom?
\par }{\Q \VS{10}Have you not
rejected
us, O
God?
\par }{\Q O God,
you do not
go
into battle
with our armies.
\par }{\Q \VS{11}Give
us help
against the enemy,
\par }{\Q for any help
men
might offer is futile.
\par }{\Q \VS{12}By God’s
power
we will conquer;
\par }{\Q he
will trample down
our enemies.


\par }\PsalmChap{61}{\D 
For the music director;
to be played on a stringed instrument;
written by David.
\par }{\Q \VerseOne{1}O God,
hear
my cry for help!
\par }{\Q Pay attention
to my prayer!
\par }{\Q \VS{2}From the most remote
place on earth
\par }{\Q I call
out to you in my despair.
\par }{\Q Lead
me up
to an inaccessible
rocky summit!
\par }{\Q \VS{3}Indeed,
you are my shelter,
\par }{\Q a strong
tower
that protects
me from the enemy.
\par }{\Q \VS{4}I will be a permanent
guest
in your home;
\par }{\Q I will find shelter
in
the protection
of your wings.
(Selah)
\par }{\Q \VS{5}For
you,
O God,
hear
my vows;
\par }{\Q you grant
me the reward
that belongs
to your loyal followers.
\par }{\Q \VS{6}Give
the king
long life!
\par }{\Q Make
his lifetime
span
several
generations!
\par }{\Q \VS{7}May he reign forever
before
God!
\par }{\Q Decree that your loyal love
and faithfulness
should protect him.
\par }{\Q \VS{8}Then
I will sing praises
to your name
continually,
\par }{\Q as I fulfill
my vows
day
after day.


\par }\PsalmChap{62}{\D 
For the music director,
Jeduthun;
a psalm
of David.
\par }{\Q \VerseOne{1}For God
alone
I patiently wait;
\par }{\Q he is the one who delivers
me.
\par }{\Q \VS{2}He alone
is my protector
and deliverer.
\par }{\Q He is my refuge;
I will not
be upended.
\par }{\Q \VS{3}How
long
will you threaten
a man?
\par }{\Q All
of you are murderers,
\par }{\Q as dangerous as a leaning
wall
or an unstable
fence.
\par }{\Q \VS{4}They spend all their time
planning how
to bring
him down.

\par }{\Q They love to use deceit;

\par }{\Q they pronounce blessings
with their
mouths,
\par }{\Q but inwardly
they utter curses.
(Selah)
\par }{\Q \VS{5}Patiently wait
for God
alone, my soul!

\par }{\Q For
he is the one who gives me confidence.
\par }{\Q \VS{6}He alone
is my protector
and deliverer.
\par }{\Q He is my refuge;
I will not
be upended.
\par }{\Q \VS{7}God
delivers
me and exalts
me;
\par }{\Q God
is my
strong
protector
and my shelter.
\par }{\Q \VS{8}Trust
in him at all
times,
you people!
\par }{\Q Pour
out your hearts
before
him!

\par }{\Q God
is our shelter! (Selah)
\par }{\Q \VS{9}Men
are nothing but
a mere breath;
\par }{\Q human
beings are unreliable.
\par }{\Q When they are weighed in the scales,
\par }{\Q all of them together
are lighter than air.
\par }{\Q \VS{10}Do not
trust
in what you can gain by oppression!

\par }{\Q Do not
put false confidence in what you can gain by robbery!

\par }{\Q If wealth
increases,
do not
become attached to it!
\par }{\Q \VS{11}God
has declared
one
principle;
\par }{\Q two
principles I have heard:

\par }{\Q God
is strong,
\par }{\Q \VS{12}and you, O Lord,
demonstrate loyal love.
\par }{\Q For
you
repay
men
for what they do.


\par }\PsalmChap{63}{\D 
A psalm
of David,
written when
he was in the Judean
wilderness.
\par }{\Q \VerseOne{1}O God,
you
are my God! I long
for you!

\par }{\Q My soul
thirsts
for you,
\par }{\Q my flesh
yearns
for you,
\par }{\Q in a dry
and parched
land
where there is no
water.
\par }{\Q \VS{2}Yes,
in the sanctuary
I have
seen
you,

\par }{\Q and witnessed
your power
and splendor.
\par }{\Q \VS{3}Because
experiencing
your loyal love
is better than
life
itself,
\par }{\Q my lips
will praise you.
\par }{\Q \VS{4}For this reason
I will praise
you while I live;
\par }{\Q in your name
I will lift
up my hands.
\par }{\Q \VS{5}As
if with choice meat
you satisfy
my soul.
\par }{\Q My mouth
joyfully
praises you,
\par }{\Q \VS{6}whenever
I remember
you on
my bed,
\par }{\Q and think about
you during the nighttime
hours.
\par }{\Q \VS{7}For
you are
my deliverer;
\par }{\Q under your wings
I rejoice.
\par }{\Q \VS{8}My soul
pursues
you;

\par }{\Q your right hand
upholds me.
\par }{\Q \VS{9}Enemies seek
to destroy
my life,
\par }{\Q but they
will descend
into the depths
of the earth.
\par }{\Q \VS{10}Each
one will be handed
over to the sword;
\par }{\Q their corpses
will be eaten by jackals.
\par }{\Q \VS{11}But the king
will rejoice
in God;
\par }{\Q everyone
who takes oaths
in his name will boast,
\par }{\Q for
the mouths
of those who speak
lies will be shut up.


\par }\PsalmChap{64}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}Listen
to me, O God,
as I offer
my lament!
\par }{\Q Protect
my life
from the enemy’s
terrifying attacks.
\par }{\Q \VS{2}Hide
me from the plots
of evil
men,
\par }{\Q from the crowd
of evildoers.
\par }{\Q \VS{3}They sharpen
their tongues
like a sword;
\par }{\Q they aim
their arrow,
a slanderous charge,
\par }{\Q \VS{4}in order to shoot
down the innocent in secluded places.
\par }{\Q They shoot
at him
suddenly
and are unafraid of retaliation.
\par }{\Q \VS{5}They encourage
one another to carry
out their evil
deed.
\par }{\Q They plan how to hide
snares,
\par }{\Q and boast, “Who
will see them?”
\par }{\Q \VS{6}They devise
unjust
schemes;
\par }{\Q they disguise
a well-conceived plot.
\par }{\Q Man’s
inner
thoughts cannot be discovered.
\par }{\Q \VS{7}But God
will shoot
at them;
\par }{\Q suddenly
they will be
wounded
by an arrow.
\par }{\Q \VS{8}Their
slander
will bring about their demise.
\par }{\Q All
who see
them will shudder,
\par }{\Q \VS{9}and all
people
will fear.
\par }{\Q They will proclaim
what God
has done,
\par }{\Q and reflect
on his deeds.
\par }{\Q \VS{10}The godly
will rejoice
in the
{\ND{Lord}}
\par }{\Q and take shelter
in him.
\par }{\Q All
the morally
upright
will boast.


\par }\PsalmChap{65}{\D 
For the music director;
a psalm
of David,
a song.
\par }{\Q \VerseOne{1}Praise
awaits
you, O God,
in Zion.
\par }{\Q Vows
made to you are fulfilled.
\par }{\Q \VS{2}You hear
prayers;
\par }{\Q all
people approach
you.
\par }{\Q \VS{3}Our record
of sins
overwhelms
me,
\par }{\Q but you
forgive
our acts of rebellion.
\par }{\Q \VS{4}How blessed
is the one whom you choose,
\par }{\Q and allow
to live
in your palace courts.
\par }{\Q May we be satisfied
with the good things
of your house
–
\par }{\Q your holy
palace.
\par }{\Q \VS{5}You answer
our prayers by performing
awesome
acts of deliverance,
\par }{\Q O God,
our savior.

\par }{\Q All
the ends
of the earth
trust
in you,

\par }{\Q as well
as those
living across
the wide
seas.
\par }{\Q \VS{6}You created
the mountains
by your power,
\par }{\Q and demonstrated
your strength.
\par }{\Q \VS{7}You calm
the raging
seas
\par }{\Q and their roaring
waves,
\par }{\Q as well as the commotion
made by the nations.
\par }{\Q \VS{8}Even those living
in the most remote areas
are awestruck
by your acts;
\par }{\Q you cause those living
in the east
and west
to praise you.
\par }{\Q \VS{9}You visit
the earth
and give it rain;
\par }{\Q you make it rich
and fertile
\par }{\Q with overflowing streams
full
of water.
\par }{\Q You provide
grain
for them,

\par }{\Q for
you prepare the earth to yield its crops.
\par }{\Q \VS{10}You saturate
its furrows,
\par }{\Q and soak
its plowed ground.
\par }{\Q With rain showers
you soften its soil,
\par }{\Q and make its crops
grow.
\par }{\Q \VS{11}You crown
the year
with your good
blessings,

\par }{\Q and you leave
abundance
in your wake.
\par }{\Q \VS{12}The pastures
in the wilderness
glisten
with moisture,

\par }{\Q and the hills
are clothed
with joy.
\par }{\Q \VS{13}The meadows
are clothed
with sheep,
\par }{\Q and the valleys
are covered
with grain.
\par }{\Q They shout
joyfully, yes,
they sing.


\par }\PsalmChap{66}{\D 
For the music director;
a song,
a psalm.
\par }{\Q \VerseOne{1}Shout
out praise to God,
all
the earth!
\par }{\Q \VS{2}Sing praises
about the majesty
of his reputation!

\par }{\Q Give
him the honor
he deserves!
\par }{\Q \VS{3}Say
to God:
\par }{\Q “How
awesome
are your deeds!
\par }{\Q Because of your great
power
your enemies
cower in fear before you.
\par }{\Q \VS{4}All
the earth
worships
you
\par }{\Q and sings praises
to you!
\par }{\Q They sing praises
to your name!” (Selah)
\par }{\Q \VS{5}Come
and witness
God’s
exploits!

\par }{\Q His acts
on
behalf of people
are awesome!
\par }{\Q \VS{6}He turned
the sea
into dry land;
\par }{\Q they passed
through the river
on foot.
\par }{\Q Let us rejoice
in him there!
\par }{\Q \VS{7}He rules
by his power
forever;
\par }{\Q he watches
the nations.
\par }{\Q Stubborn
rebels
should not
exalt
themselves. (Selah)
\par }{\Q \VS{8}Praise
our God,
you nations!
\par }{\Q Loudly
proclaim
his praise!
\par }{\Q \VS{9}He preserves
our lives
\par }{\Q and does not
allow
our feet
to slip.
\par }{\Q \VS{10}For
you, O God,
tested
us;
\par }{\Q you purified
us like refined
silver.
\par }{\Q \VS{11}You led
us into a trap;
\par }{\Q you caused
us to suffer.
\par }{\Q \VS{12}You allowed men
to ride over
our heads;
\par }{\Q we passed through
fire
and water,
\par }{\Q but you brought
us out
into a wide open place.
\par }{\Q \VS{13}I will enter
your temple
with burnt sacrifices;
\par }{\Q I will fulfill
the vows I made to you,
\par }{\Q \VS{14}which
my lips
uttered
\par }{\Q and my mouth
spoke
when I was in trouble.
\par }{\Q \VS{15}I will offer up to you fattened
animals as burnt sacrifices,
\par }{\Q along with
the smell
of sacrificial rams.
\par }{\Q I will offer
cattle
and goats.
(Selah)
\par }{\Q \VS{16}Come! Listen,
all
you who are loyal
to God!

\par }{\Q I will declare
what
he has done
for me.
\par }{\Q \VS{17}I cried out
to
him for help

\par }{\Q and praised
him with
my tongue.
\par }{\Q \VS{18}If
I had harbored sin
in my heart,
\par }{\Q the Lord
would not
have listened.
\par }{\Q \VS{19}However,
God
heard;
\par }{\Q he listened
to my prayer.
\par }{\Q \VS{20}God
deserves
praise,

\par }{\Q for he did not
reject
my prayer
\par }{\Q or abandon his love
for me!


\par }\PsalmChap{67}{\D 
For the music director;
to be accompanied
by stringed instruments;
a psalm,
a song.
\par }{\Q \VerseOne{1}May God
show us his favor
and bless
us!

\par }{\Q May he smile
on us! (Selah)
\par }{\Q \VS{2}Then those living
on earth
will know
what you are like;
\par }{\Q all
nations
will know how you deliver your people.
\par }{\Q \VS{3}Let the nations
thank
you, O God!
\par }{\Q Let all
the nations
thank you!
\par }{\Q \VS{4}Let foreigners
rejoice
and celebrate!
\par }{\Q For
you execute
justice
among the nations,
\par }{\Q and govern
the people
living on earth.
(Selah)
\par }{\Q \VS{5}Let the nations
thank
you, O God!
\par }{\Q Let all
the nations
thank you!
\par }{\Q \VS{6}The earth
yields
its crops.
\par }{\Q May God,
our God,
bless us!
\par }{\Q \VS{7}May God
bless
us!

\par }{\Q Then
all
the ends
of the earth
will give him the honor he deserves.


\par }\PsalmChap{68}{\D 
For the music director;
by David,
a psalm,
a song.
\par }{\Q \VerseOne{1}God
springs into action!

\par }{\Q His enemies
scatter;
\par }{\Q his adversaries
run
from him.
\par }{\Q \VS{2}As smoke
is driven away
by the wind, so you drive them away.
\par }{\Q As wax
melts
before
fire,
\par }{\Q so the wicked
are destroyed
before
God.
\par }{\Q \VS{3}But the godly
are happy;
\par }{\Q they rejoice
before
God
\par }{\Q and are overcome
with joy.
\par }{\Q \VS{4}Sing
to God! Sing praises
to his name!
\par }{\Q Exalt
the one who rides
on the clouds!

\par }{\Q For the
{\ND{Lord}}
is his name!

\par }{\Q Rejoice
before him!
\par }{\Q \VS{5}He is a father
to the fatherless
\par }{\Q and an advocate
for widows.
\par }{\Q God
rules
from his holy palace.
\par }{\Q \VS{6}God
settles
those who have been deserted
in their own homes;
\par }{\Q he frees
prisoners
and grants them prosperity.
\par }{\Q But
sinful rebels
live
in the desert.
\par }{\Q \VS{7}O God,
when you lead
your people
into battle,

\par }{\Q when you march
through the desert,
(Selah)
\par }{\Q \VS{8}the earth
shakes,
\par }{\Q yes,
the heavens
pour down rain
\par }{\Q before
God, the God
of Sinai,
\par }{\Q before
God,
the God
of Israel.
\par }{\Q \VS{9}O God,
you cause abundant showers
to fall on your chosen
people.

\par }{\Q When they are tired,
you
sustain them,
\par }{\Q \VS{10}for you live
among
them.

\par }{\Q You sustain
the oppressed
with your good
blessings, O God.
\par }{\Q \VS{11}The Lord
speaks;
\par }{\Q many, many women
spread the good news.
\par }{\Q \VS{12}Kings
leading
armies
run away
– they run away!

\par }{\Q The lovely
lady
of the house
divides
up the loot.
\par }{\Q \VS{13}When
you lie
down among
the sheepfolds,
\par }{\Q the wings
of the dove
are covered
with silver
\par }{\Q and with glittering
gold.
\par }{\Q \VS{14}When
the sovereign
judge
scatters
kings,
\par }{\Q let it snow
on Zalmon!
\par }{\Q \VS{15}The mountain
of Bashan
is a towering mountain;
\par }{\Q the mountain
of Bashan
is a mountain with many peaks.
\par }{\Q \VS{16}Why
do you look with envy,
O mountains
with many peaks,
\par }{\Q at the mountain
where God
has decided to live?

\par }{\Q Indeed
the {\ND{Lord}}
will live
there permanently!
\par }{\Q \VS{17}God
has countless
chariots;
\par }{\Q they number in the thousands.
\par }{\Q The Lord
comes from Sinai
in holy splendor.
\par }{\Q \VS{18}You ascend
on high,
\par }{\Q you have taken many
captives.
\par }{\Q You receive
tribute
from men,
\par }{\Q including even
sinful rebels.
\par }{\Q Indeed the
{\ND{Lord}}
God
lives there!
\par }{\Q \VS{19}The Lord
deserves
praise!

\par }{\Q Day
after day
he carries
our burden,
\par }{\Q the God
who delivers
us. (Selah)
\par }{\Q \VS{20}Our God
is a God
who delivers;
\par }{\Q the
{\ND{Lord}}, the sovereign Lord,
can rescue
from death.
\par }{\Q \VS{21}Indeed
God
strikes
the heads
of his enemies,
\par }{\Q the hairy
foreheads
of those who persist in rebellion.
\par }{\Q \VS{22}The Lord
says,
\par }{\Q “I will retrieve them from Bashan,
\par }{\Q I will bring
them back
from the depths
of the sea,
\par }{\Q \VS{23}so
that your feet
may stomp
in their blood,
\par }{\Q and your dogs
may eat their portion of the enemies’
corpses.”
\par }{\Q \VS{24}They see
your processions,
O God
–
\par }{\Q the processions
of my God,
my king,
who marches along in holy splendor.
\par }{\Q \VS{25}Singers
walk in front;
\par }{\Q musicians
follow
playing
their stringed instruments,

\par }{\Q in the midst
of young women playing
tambourines.
\par }{\Q \VS{26}In your large assemblies
praise
God,
\par }{\Q the {\ND{Lord}}, in the assemblies
of Israel!
\par }{\Q \VS{27}There
is little
Benjamin,
their ruler,
\par }{\Q and the princes
of Judah
in their robes,
\par }{\Q along with the princes
of Zebulun
and the princes
of Naphtali.
\par }{\Q \VS{28}God
has decreed
that you will be powerful.
\par }{\Q O God,
you who
have acted
on our behalf, demonstrate
your power,
\par }{\Q \VS{29}as you come out of your temple
in Jerusalem!

\par }{\Q Kings
bring tribute to you.
\par }{\Q \VS{30}Sound your battle cry against
the wild beast
of the reeds,
\par }{\Q and the nations
that assemble
like a herd of calves
led by bulls!

\par }{\Q They humble themselves
and offer gold and silver
as tribute.

\par }{\Q God scatters
the nations
that like
to do battle.
\par }{\Q \VS{31}They
come
with red cloth from
Egypt,
\par }{\Q Ethiopia
voluntarily offers
tribute
to God.
\par }{\Q \VS{32}O kingdoms
of the earth,
sing
to God!
\par }{\Q Sing praises
to the Lord,
(Selah)
\par }{\Q \VS{33}to the one who rides
through the sky
from ancient times!

\par }{\Q Look! He thunders loudly.
\par }{\Q \VS{34}Acknowledge
God’s
power,

\par }{\Q his sovereignty
over
Israel,
\par }{\Q and the power
he reveals
in the skies!
\par }{\Q \VS{35}You are awe-inspiring,
O God,
as you emerge from your holy temple!

\par }{\Q It is
the God
of Israel
who gives
the people
power
and strength.
\par }{\Q God
deserves praise!


\par }\PsalmChap{69}{\D 
For the music director;
according to
the tune of “Lilies;”
by David.
\par }{\Q \VerseOne{1}Deliver
me, O God,
\par }{\Q for
the water
has reached
my neck.
\par }{\Q \VS{2}I sink
into the deep
mire
\par }{\Q where there is no
solid ground;
\par }{\Q I am in
deep
water,
\par }{\Q and the current
overpowers me.
\par }{\Q \VS{3}I am
exhausted
from shouting for help;
\par }{\Q my throat
is sore;
\par }{\Q my eyes
grow tired of looking
for my God.
\par }{\Q \VS{4}Those who hate
me without cause
are more numerous
than the hairs
of my head.
\par }{\Q Those who want
to destroy
me, my enemies
for no reason,
outnumber me.

\par }{\Q They make me repay
what
I did not
steal!
\par }{\Q \VS{5}O God,
you
are aware
of my foolish
sins;

\par }{\Q my guilt
is not
hidden from you.
\par }{\Q \VS{6}Let
none who rely
on you be disgraced
because of me,
\par }{\Q O sovereign

{\ND{Lord}}
and king!

\par }{\Q Let
none
who seek
you be ashamed
because of me,
\par }{\Q O God
of Israel!
\par }{\Q \VS{7}For
I suffer
humiliation
for your sake

\par }{\Q and am thoroughly
disgraced.
\par }{\Q \VS{8}My own brothers
treat
me like a stranger;
\par }{\Q they act as if I were a foreigner.
\par }{\Q \VS{9}Certainly
zeal
for your house
consumes
me;
\par }{\Q I endure the insults
of those who insult you.
\par }{\Q \VS{10}I weep
and refrain from eating food,
\par }{\Q which causes
others to insult me.
\par }{\Q \VS{11}I
wear
sackcloth
\par }{\Q and they
ridicule me.
\par }{\Q \VS{12}Those
who sit
at the city gate
gossip about me;
\par }{\Q drunkards
mock me in their songs.
\par }{\Q \VS{13}O
{\ND{Lord}}, may you hear my
prayer
and be favorably
disposed to me!

\par }{\Q O God,
because of your great
loyal love,
\par }{\Q answer
me with your faithful
deliverance!
\par }{\Q \VS{14}Rescue
me from the mud! Don’t
let me sink!
\par }{\Q Deliver
me from those who hate
me,
\par }{\Q from the deep
water!
\par }{\Q \VS{15}Don’t
let the current
overpower
me!
\par }{\Q Don’t
let the deep
swallow
me up!
\par }{\Q Don’t
let the pit
devour me!
\par }{\Q \VS{16}Answer
me, O
{\ND{Lord}}, for
your loyal love
is good!

\par }{\Q Because of your great
compassion,
turn
toward me!
\par }{\Q \VS{17}Do not
ignore
your servant,
\par }{\Q for
I am in trouble! Answer
me right away!
\par }{\Q \VS{18}Come near
me
and redeem
me!

\par }{\Q Because
of my enemies,
rescue me!
\par }{\Q \VS{19}You
know
how I am insulted,
humiliated
and disgraced;
\par }{\Q you can see all
my enemies.
\par }{\Q \VS{20}Their insults
are painful
and make me lose heart;
\par }{\Q I look for
sympathy,
but
receive none,
\par }{\Q for comforters,
but find
none.
\par }{\Q \VS{21}They put
bitter poison
into my food,
\par }{\Q and to quench my thirst
they give me vinegar
to drink.
\par }{\Q \VS{22}May
their dining
table
become a trap
before
them!
\par }{\Q May it be a snare for that group of friends!
\par }{\Q \VS{23}May their eyes
be blinded!

\par }{\Q Make them shake violently!
\par }{\Q \VS{24}Pour
out your judgment
on
them!
\par }{\Q May your raging
anger
overtake them!
\par }{\Q \VS{25}May
their camp
become
desolate,
\par }{\Q their tents uninhabited!
\par }{\Q \VS{26}For
they harass
the one whom
you
discipline;
\par }{\Q they spread
the news
about
the suffering
of those whom you punish.
\par }{\Q \VS{27}Hold
them accountable
for
all their sins!

\par }{\Q Do not
vindicate them!
\par }{\Q \VS{28}May their names be deleted
from the scroll
of the living!

\par }{\Q Do not
let their names be listed
with
the godly!
\par }{\Q \VS{29}I am
oppressed
and suffering!
\par }{\Q O God,
deliver
and protect me!
\par }{\Q \VS{30}I
will sing
praises
to God’s
name!

\par }{\Q I will magnify
him as I give him
thanks!
\par }{\Q \VS{31}That will please
the {\ND{Lord}}
more than an ox
or a bull
\par }{\Q with horns
and hooves.
\par }{\Q \VS{32}The oppressed
look on – let them rejoice!
\par }{\Q You who seek God, may you be encouraged!
\par }{\Q \VS{33}For
the {\ND{Lord}}
listens
to
the needy;
\par }{\Q he does not
despise
his captive people.
\par }{\Q \VS{34}Let the heavens
and the earth
praise
him,
\par }{\Q along with the seas
and everything
that swims in them!
\par }{\Q \VS{35}For
God
will deliver
Zion
\par }{\Q and rebuild
the cities
of Judah,
\par }{\Q and his people will again live
in them and possess Zion.
\par }{\Q \VS{36}The descendants
of his servants
will inherit
it,
\par }{\Q and those who are loyal
to him
will live in it.


\par }\PsalmChap{70}{\D 
For the music director;
by David;
written
to get God’s attention.
\par }{\Q \VerseOne{1}O God,
please be willing to rescue
me!

\par }{\Q O
{\ND{Lord}}, hurry
and help me!
\par }{\Q \VS{2}May those who
are trying
to take my life
\par }{\Q be embarrassed
and ashamed!

\par }{\Q May those who want
to harm
me
\par }{\Q be turned back
and ashamed!
\par }{\Q \VS{3}May
those
who say,
“Aha! Aha!”
\par }{\Q be driven back and disgraced!
\par }{\Q \VS{4}May all
those who seek
you be happy
and rejoice
in you!
\par }{\Q May those who love
to experience your deliverance
say
continually,
\par }{\Q “May God
be praised!”
\par }{\Q \VS{5}I am
oppressed
and needy!

\par }{\Q O God,
hurry
to me!

\par }{\Q You
are my helper
and my deliverer!
\par }{\Q O

{\ND{Lord}}, do not
delay!


\par }\PsalmChap{71}{\PP \par }{\Q \VerseOne{1}In you, O
{\ND{Lord}}, I have taken shelter!
\par }{\Q Never
let
me be humiliated!
\par }{\Q \VS{2}Vindicate
me by rescuing
me!

\par }{\Q Listen
to
me! Deliver me!
\par }{\Q \VS{3}Be
my protector
and refuge,
\par }{\Q a stronghold where I can be safe!

\par }{\Q For
you are
my high ridge
and my stronghold.
\par }{\Q \VS{4}My God,
rescue
me from the power
of the wicked,
\par }{\Q from the hand
of the cruel
oppressor!
\par }{\Q \VS{5}For
you
give me confidence,
O Lord;
\par }{\Q O
{\ND{Lord}}, I have trusted
in you since I was young.
\par }{\Q \VS{6}I have leaned
on
you since birth;
\par }{\Q you
pulled
me from my mother’s
womb.
\par }{\Q I praise
you continually.
\par }{\Q \VS{7}Many
are appalled
when
they see me,

\par }{\Q but you
are my secure
shelter.
\par }{\Q \VS{8}I praise
you constantly
\par }{\Q and speak
of your splendor
all
day long.
\par }{\Q \VS{9}Do not
reject
me in my old age!

\par }{\Q When
my strength
fails,
do not
abandon me!
\par }{\Q \VS{10}For
my enemies
talk about me;
\par }{\Q those waiting
for a chance
to kill me plot my demise.
\par }{\Q \VS{11}They say, “God
has abandoned
him.
\par }{\Q Run
and seize
him, for
there is no
one who will rescue him!”
\par }{\Q \VS{12}O God,
do not
remain far
away from
me!
\par }{\Q My God,
hurry
and help me!
\par }{\Q \VS{13}May my accusers
be humiliated
and defeated!
\par }{\Q May those who want
to harm
me be covered with
scorn
and disgrace!
\par }{\Q \VS{14}As for me, I
will wait
continually,
\par }{\Q and will continue
to praise you.
\par }{\Q \VS{15}I will tell
about your justice,
\par }{\Q and all
day
long proclaim
your salvation,
\par }{\Q though
I cannot
fathom
its full extent.
\par }{\Q \VS{16}I will come
and tell about the mighty
acts of the sovereign

{\ND{Lord}}.
\par }{\Q I will proclaim
your justice
– yours alone.
\par }{\Q \VS{17}O God,
you have taught
me since I was young,
\par }{\Q and I am still declaring
your amazing deeds.
\par }{\Q \VS{18}Even
when
I am old
and gray,
\par }{\Q O God,
do not
abandon
me,
\par }{\Q until
I tell
the next generation
about your strength,
\par }{\Q and those coming
after me about your power.
\par }{\Q \VS{19}Your justice,
O God,
extends
to the skies above;
\par }{\Q you have
done
great
things.

\par }{\Q O God,
who can
compare to you?
\par }{\Q \VS{20}Though you have allowed me to experience
much
trouble
and distress,
\par }{\Q revive
me once again!

\par }{\Q Bring me up
once again
from the depths
of the earth!
\par }{\Q \VS{21}Raise
me to a position of great
honor!

\par }{\Q Turn
and comfort me!
\par }{\Q \VS{22}I
will express my thanks
to you with
a stringed instrument,
\par }{\Q praising
your faithfulness,
O my God!
\par }{\Q I will sing praises
to you accompanied by a harp,
\par }{\Q O Holy One
of Israel!
\par }{\Q \VS{23}My lips
will shout
for joy! Yes,
I will sing your praises!
\par }{\Q I will praise
you when you rescue me!
\par }{\Q \VS{24}All
day
long my tongue
will also
tell
about your justice,
\par }{\Q for
those who want
to harm
me will be embarrassed
and ashamed.


\par }\PsalmChap{72}{\D 
For Solomon.
\par }{\Q \VerseOne{1}O God,
grant the king
the ability to make
just decisions!

\par }{\Q Grant
the king’s
son
the ability to make fair decisions!
\par }{\Q \VS{2}Then he will judge
your people
fairly,
\par }{\Q and your oppressed
ones equitably.
\par }{\Q \VS{3}The mountains
will bring
news of peace
to the people,
\par }{\Q and the hills
will announce justice.
\par }{\Q \VS{4}He
will defend
the oppressed
among the people;
\par }{\Q he will deliver
the children
of the poor
\par }{\Q and crush
the oppressor.
\par }{\Q \VS{5}People will fear
you as long as the sun
and moon
remain in the sky,
\par }{\Q for generation
after generation.
\par }{\Q \VS{6}He will descend
like rain
on
the mown
grass,

\par }{\Q like showers
that drench
the earth.
\par }{\Q \VS{7}During
his days
the godly
will flourish;
\par }{\Q peace
will prevail as long as
the moon
remains in the sky.
\par }{\Q \VS{8}May he rule
from sea
to
sea,
\par }{\Q and from the Euphrates
River to
the ends
of the earth!
\par }{\Q \VS{9}Before
him the coastlands
will bow down,
\par }{\Q and his enemies
will lick
the dust.
\par }{\Q \VS{10}The kings
of Tarshish
and the coastlands
will offer
gifts;
\par }{\Q the kings
of Sheba
and Seba
will bring
tribute.
\par }{\Q \VS{11}All
kings
will bow
down to him;
\par }{\Q all
nations
will serve him.
\par }{\Q \VS{12}For
he will rescue
the needy
when they cry out
for help,
\par }{\Q and the oppressed
who have no defender.
\par }{\Q \VS{13}He will take pity
on
the poor
and needy;
\par }{\Q the lives
of the needy
he will save.
\par }{\Q \VS{14}From harm
and violence
he will defend
them;

\par }{\Q he will value
their lives.
\par }{\Q \VS{15}May he live! May they offer
him gold
from Sheba!

\par }{\Q May they continually
pray
for
him!
\par }{\Q May they pronounce blessings
on him all
day long!
\par }{\Q \VS{16}May there be
an abundance
of grain
in the earth;
\par }{\Q on the tops
of the mountains
may it
sway!

\par }{\Q May
its fruit
trees flourish
like the forests of Lebanon!

\par }{\Q May
its crops be as abundant as the grass
of the earth!
\par }{\Q \VS{17}May
his fame
endure!

\par }{\Q May his dynasty last
as long as
the sun
remains in the sky!

\par }{\Q May they use his name
when they formulate their blessings!

\par }{\Q May all
nations
consider him to be favored by God!
\par }{\Q \VS{18}The
{\ND{Lord}}
God,
the God
of Israel,
deserves
praise!

\par }{\Q He alone
accomplishes
amazing things!
\par }{\Q \VS{19}His glorious
name
deserves
praise forevermore!
\par }{\Q May his majestic splendor
fill
the whole
earth!
\par }{\Q We agree! We agree!
\par }{\Q \VS{20}This collection
of the prayers
of David
son
of Jesse ends here.


\par }\PsalmChap{73}{\D 
A psalm
by Asaph.
\par }{\Q \VerseOne{1}Certainly
God
is good
to Israel,
\par }{\Q and to those whose motives
are pure!
\par }{\Q \VS{2}But as for me,
my feet
almost
slipped;
\par }{\Q my feet
almost
slid out from under me.
\par }{\Q \VS{3}For
I envied
those who are proud,
\par }{\Q as I observed
the prosperity
of the wicked.
\par }{\Q \VS{4}For
they suffer
no
pain;
\par }{\Q their bodies
are strong
and well-fed.
\par }{\Q \VS{5}They are immune
to the trouble
common to
men;
\par }{\Q they do not
suffer
as other men do.
\par }{\Q \VS{6}Arrogance
is their necklace,
\par }{\Q and violence
their clothing.
\par }{\Q \VS{7}Their prosperity
causes
them to do
wrong;

\par }{\Q their thoughts
are sinful.
\par }{\Q \VS{8}They mock
and say
evil things;
\par }{\Q they proudly
threaten
violence.
\par }{\Q \VS{9}They speak
as if they rule in heaven,
\par }{\Q and lay claim
to the earth.
\par }{\Q \VS{10}Therefore
they
have more than enough
food to eat,
\par }{\Q and even suck up
the water of the sea.
\par }{\Q \VS{11}They say,
“How
does God
know
what we do?
\par }{\Q Is
the sovereign one
aware of what goes on?”
\par }{\Q \VS{12}Take a good look! This
is what the wicked
are like,

\par }{\Q those who always
have it so easy
and get richer
and richer.
\par }{\Q \VS{13}I concluded, “Surely
in vain
I have kept my motives
pure
\par }{\Q and maintained
a pure
lifestyle.
\par }{\Q \VS{14}I suffer
all
day
long,
\par }{\Q and am punished
every morning.”
\par }{\Q \VS{15}If
I had publicized
these thoughts,

\par }{\Q I would have betrayed
your loyal followers.
\par }{\Q \VS{16}When I tried
to make sense
of this,
\par }{\Q it
was troubling
to me.
\par }{\Q \VS{17}Then I entered
the precincts of God’s
temple,
\par }{\Q and understood
the destiny of the wicked.
\par }{\Q \VS{18}Surely
you put
them in slippery places;
\par }{\Q you bring
them down
to ruin.
\par }{\Q \VS{19}How
desolate
they become
in a mere moment!
\par }{\Q Terrifying judgments
make their demise
complete!
\par }{\Q \VS{20}They are like a dream
after one wakes up.
\par }{\Q O Lord,
when you awake
you will despise
them.
\par }{\Q \VS{21}Yes,
my spirit
was bitter,
\par }{\Q and my insides
felt sharp pain.
\par }{\Q \VS{22}I
was ignorant
and lacked insight;
\par }{\Q I was as senseless
as an animal before you.
\par }{\Q \VS{23}But I
am continually
with
you;
\par }{\Q you hold
my right
hand.
\par }{\Q \VS{24}You guide me by your wise advice,
\par }{\Q and then
you will lead
me to a position
of honor.
\par }{\Q \VS{25}Whom
do I have in heaven
but you?
\par }{\Q I desire
no
one but you on earth.
\par }{\Q \VS{26}My flesh
and my heart
may grow
weak,

\par }{\Q but God
always
protects
my heart
and gives me stability.
\par }{\Q \VS{27}Yes,
look! Those far from you
die;
\par }{\Q you destroy
everyone
who is unfaithful to you.
\par }{\Q \VS{28}But as for me, God’s
presence
is
all
I
need.

\par }{\Q I have made the sovereign

{\ND{Lord}}
my shelter,
\par }{\Q as I declare
all
the things you have done.


\par }\PsalmChap{74}{\D 
A well-written song
by Asaph.
\par }{\Q \VerseOne{1}Why,
O God,
have you permanently
rejected
us?

\par }{\Q Why does your anger
burn
against the sheep
of your pasture?
\par }{\Q \VS{2}Remember
your people
whom you acquired
in ancient
times,
\par }{\Q whom you rescued
so they could be your very own nation,
\par }{\Q as well as Mount
Zion,
where you dwell!
\par }{\Q \VS{3}Hurry
and look at the permanent ruins,
\par }{\Q and all
the damage
the enemy has done to the temple!
\par }{\Q \VS{4}Your enemies
roar
in the middle
of your sanctuary;
\par }{\Q they set up
their battle flags.
\par }{\Q \VS{5}They invade
like lumberjacks
\par }{\Q swinging
their axes
in a thick forest.
\par }{\Q \VS{6}And now they are tearing down
all
its engravings
\par }{\Q with axes
and crowbars.
\par }{\Q \VS{7}They set
your sanctuary
on fire;
\par }{\Q they desecrate
your dwelling place
by knocking
it to the ground.
\par }{\Q \VS{8}They say
to themselves,
\par }{\Q “We will oppress
all
of them.”

\par }{\Q They burn down
all
the places where
people worship God
in the land.
\par }{\Q \VS{9}We do not
see
any signs
of God’s presence;

\par }{\Q there
are no
longer
any prophets
\par }{\Q and we have no one
to tell
us how
long this will last.
\par }{\Q \VS{10}How long,
O God,
will the adversary
hurl insults?
\par }{\Q Will the enemy
blaspheme
your
name
forever?
\par }{\Q \VS{11}Why
do you remain inactive?
\par }{\Q Intervene and destroy him!
\par }{\Q \VS{12}But God
has been my king
from ancient
times,
\par }{\Q performing
acts of deliverance
on the earth.
\par }{\Q \VS{13}You
destroyed
the sea
by your strength;
\par }{\Q you shattered
the heads
of the sea monster
in
the water.
\par }{\Q \VS{14}You
crushed
the heads
of Leviathan;
\par }{\Q you fed
him to the people
who live along the coast.
\par }{\Q \VS{15}You
broke
open the spring
and the stream;
\par }{\Q you
dried
up perpetually flowing
rivers.
\par }{\Q \VS{16}You established
the cycle of day
and night;
\par }{\Q you
put the moon
and sun in place.
\par }{\Q \VS{17}You
set
up all
the boundaries
of the earth;
\par }{\Q you
created
the cycle of summer
and winter.
\par }{\Q \VS{18}Remember
how the enemy
hurls insults,
O
{\ND{Lord}},

\par }{\Q and how a foolish
nation
blasphemes your name!
\par }{\Q \VS{19}Do not
hand
the life
of your dove
over to a wild animal!
\par }{\Q Do not
continue
to disregard
the lives
of your oppressed people!
\par }{\Q \VS{20}Remember
your covenant
promises,

\par }{\Q for
the dark regions
of the earth
are full
of places
where violence rules.
\par }{\Q \VS{21}Do not
let the afflicted
be turned back
in shame!
\par }{\Q Let the oppressed
and poor
praise
your name!
\par }{\Q \VS{22}Rise
up, O God! Defend
your honor!

\par }{\Q Remember
how fools
insult
you all
day long!
\par }{\Q \VS{23}Do not
disregard
what your enemies
say,

\par }{\Q or the unceasing shouts
of those who defy you.


\par }\PsalmChap{75}{\D 
For the music director;
{\IT{according to the
al-tashcheth}} style;
a psalm
of Asaph;
a song.
\par }{\Q \VerseOne{1}We give thanks
to you, O God! We give thanks!
\par }{\Q You reveal
your presence;
\par }{\Q people tell
about your amazing deeds.
\par }{\Q \VS{2}God says,
\par }{\Q “At the appointed
times,

\par }{\Q I
judge
fairly.
\par }{\Q \VS{3}When the earth
and all
its inhabitants
dissolve
in fear,

\par }{\Q I
make
its pillars
secure.” (Selah)
\VS{4}\par }{\Q I say
to the proud,
“Do not
be proud,”
\par }{\Q and to the wicked,
“Do not
be so confident
of victory!
\par }{\Q \VS{5}Do not
be so certain
you have won!

\par }{\Q Do not speak
with your head held so high!
\par }{\Q \VS{6}For
victory
does not
come from the east or west,
\par }{\Q or from the wilderness.
\par }{\Q \VS{7}For
God
is the judge!

\par }{\Q He brings one
down and exalts
another.
\par }{\Q \VS{8}For
the {\ND{Lord}}
holds in his hand
a cup
full
\par }{\Q of foaming wine
mixed
with spices,

\par }{\Q and pours it out.
\par }{\Q Surely
all
the wicked
of the earth
\par }{\Q will slurp it up
and drink
it
to its
very last drop.”
\par }{\Q \VS{9}As for me, I
will continually
tell
what you have done;

\par }{\Q I will sing praises
to the God
of Jacob!
\par }{\Q \VS{10}God says,

\par }{\Q “I will bring down
all
the power
of the wicked;
\par }{\Q the godly
will be
victorious.”


\par }\PsalmChap{76}{\D 
For the music director;
to be accompanied by stringed instruments;
a psalm
of Asaph,
a song.
\par }{\Q \VerseOne{1}God
has revealed
himself in Judah;
\par }{\Q in Israel
his reputation
is great.
\par }{\Q \VS{2}He lives
in Salem;
\par }{\Q he dwells
in Zion.
\par }{\Q \VS{3}There
he shattered
the arrows,
\par }{\Q the shield,
the sword,
and the rest of the weapons of war.
(Selah)
\par }{\Q \VS{4}You
shine
brightly and reveal your majesty,
\par }{\Q as you descend from the hills
where you killed your prey.
\par }{\Q \VS{5}The bravehearted
were plundered;

\par }{\Q they “fell asleep.”
\par }{\Q All
the warriors were helpless.
\par }{\Q \VS{6}At the sound of your battle cry,
O God
of Jacob,
\par }{\Q both rider
and horse
“fell asleep.”
\par }{\Q \VS{7}You
are awesome! Yes, you!
\par }{\Q Who can
withstand
your intense anger?
\par }{\Q \VS{8}From heaven
you announced
what their punishment
would be.

\par }{\Q The earth
was afraid
and silent
\par }{\Q \VS{9}when God
arose
to execute judgment,
\par }{\Q and to deliver
all
the oppressed
of the earth.
(Selah)
\par }{\Q \VS{10}Certainly
your angry
judgment upon men
will bring
you praise;
\par }{\Q you reveal
your anger
in full measure.
\par }{\Q \VS{11}Make vows
to the
{\ND{Lord}}
your God
and repay
them!
\par }{\Q Let all
those who surround
him bring tribute
to the awesome one!
\par }{\Q \VS{12}He humbles
princes;
\par }{\Q the kings
of the earth
regard him as
awesome.


\par }\PsalmChap{77}{\D 
For the music director,
Jeduthun;
a psalm
of Asaph.
\par }{\Q \VerseOne{1}I will cry
out to
God
and call for help!
\par }{\Q I will cry out
to
God
and he will pay attention
to me.
\par }{\Q \VS{2}In my time
of trouble
I sought
the Lord.
\par }{\Q I kept my hand raised
in prayer throughout
the night.
\par }{\Q I
refused
to be comforted.
\par }{\Q \VS{3}I said, “I will remember
God
while I groan;
\par }{\Q I will think
about him while my strength leaves
me.” (Selah)
\par }{\Q \VS{4}You held
my eyelids
open;
\par }{\Q I was troubled
and could not
speak.
\par }{\Q \VS{5}I thought
about the days
of old,
\par }{\Q about ancient
times.
\par }{\Q \VS{6}I said,
“During the night
I will remember
the song
I
once sang;
\par }{\Q I
will think
very carefully.”
\par }{\Q I
tried to make sense
of what
was happening.
\par }{\Q \VS{7}I asked, “Will the Lord
reject
me
forever?
\par }{\Q Will he never
again
show me
his favor?
\par }{\Q \VS{8}Has his loyal love
disappeared
forever?
\par }{\Q Has his promise
failed
forever?
\par }{\Q \VS{9}Has God
forgotten
to be merciful?
\par }{\Q Has his anger
stifled
his compassion?”
\par }{\Q \VS{10}Then I said,
“I am sickened
by the thought
\par }{\Q that the sovereign
One might become inactive.
\par }{\Q \VS{11}I will remember
the works
of the
{\ND{Lord}}.
\par }{\Q Yes,
I will remember
the amazing
things you did long ago!
\par }{\Q \VS{12}I will think
about all
you have done;
\par }{\Q I will reflect upon
your deeds!”
\VS{13}\par }{\Q O God,
your deeds
are extraordinary!

\par }{\Q What
god
can compare
to our great
God?
\par }{\Q \VS{14}You
are the God
who does
amazing
things;
\par }{\Q you have revealed
your strength
among the nations.
\par }{\Q \VS{15}You delivered
your people
by your strength –
\par }{\Q the children
of Jacob
and Joseph.
(Selah)
\par }{\Q \VS{16}The waters
saw
you, O God,
\par }{\Q the waters
saw
you and trembled.
\par }{\Q Yes,
the depths
of the sea shook with fear.
\par }{\Q \VS{17}The clouds
poured down
rain;
\par }{\Q the skies
thundered.
\par }{\Q Yes,
your arrows
flashed about.
\par }{\Q \VS{18}Your thunderous
voice
was heard in the wind;
\par }{\Q the lightning bolts
lit
up the world;
\par }{\Q the earth
trembled
and shook.
\par }{\Q \VS{19}You walked through
the sea;
\par }{\Q you passed through
the surging
waters,
\par }{\Q but left no
footprints.
\par }{\Q \VS{20}You led
your people
like a flock
of sheep,
\par }{\Q by the hand
of Moses
and Aaron.


\par }\PsalmChap{78}{\D 
A well-written song
by Asaph.
\par }{\Q \VerseOne{1}Pay attention,
my people,
to my instruction!
\par }{\Q Listen
to the words
I speak!
\par }{\Q \VS{2}I will sing
a song that imparts wisdom;
\par }{\Q I will make
insightful observations
about
the past.
\par }{\Q \VS{3}What
we have heard
and learned –
\par }{\Q that which our ancestors
have told us –
\par }{\Q \VS{4}we will not
hide
from their descendants.
\par }{\Q We will tell
the next
generation
\par }{\Q about the
{\ND{Lord}}’s
praiseworthy acts,
\par }{\Q about his strength
and the amazing things
he has
done.
\par }{\Q \VS{5}He established
a rule
in Jacob;
\par }{\Q he set up
a law
in Israel.
\par }{\Q He commanded
our ancestors
\par }{\Q to make his deeds known
to their descendants,
\par }{\Q \VS{6}so that the next generation,
children
yet to be born,
\par }{\Q might
know
about them.
\par }{\Q They will grow up
and tell
their descendants about them.
\par }{\Q \VS{7}Then
they will place
their confidence
in God.
\par }{\Q They will not
forget
the works
of God,
\par }{\Q and they will obey his commands.
\par }{\Q \VS{8}Then they will not
be
like their ancestors,
\par }{\Q who were a stubborn
and rebellious
generation,
\par }{\Q a generation
that was not
committed
\par }{\Q and faithful
to God.
\par }{\Q \VS{9}The Ephraimites
were armed
with bows,
\par }{\Q but they retreated
in the day
of battle.
\par }{\Q \VS{10}They did not
keep
their covenant
with God,
\par }{\Q and they refused
to obey
his law.
\par }{\Q \VS{11}They forgot
what
he had done,
\par }{\Q the amazing
things he had
shown them.
\par }{\Q \VS{12}He did
amazing
things in the sight
of their ancestors,
\par }{\Q in the land
of Egypt,
in the region
of Zoan.
\par }{\Q \VS{13}He divided
the sea
and led them across
it;
\par }{\Q he made the water
stand
in a heap.
\par }{\Q \VS{14}He led
them with a cloud
by day,
\par }{\Q and with the light
of a fire
all
night long.
\par }{\Q \VS{15}He broke
open rocks
in the wilderness,
\par }{\Q and gave them enough
water
to fill the depths of the sea.
\par }{\Q \VS{16}He caused
streams
to flow from the rock,
\par }{\Q and made the water
flow
like rivers.
\par }{\Q \VS{17}Yet
they continued
to sin
against him,
\par }{\Q and rebelled
against the sovereign
One in the desert.
\par }{\Q \VS{18}They willfully
challenged
God
\par }{\Q by asking
for food
to satisfy their appetite.
\par }{\Q \VS{19}They insulted
God,
saying,
\par }{\Q “Is God
really able
to give
us food
in the wilderness?
\par }{\Q \VS{20}Yes,
he struck
a rock
and water
flowed out,
\par }{\Q streams
gushed forth.
\par }{\Q But can
he also
give
us food?
\par }{\Q Will he provide
meat
for his people?”
\par }{\Q \VS{21}When the
{\ND{Lord}}
heard
this,
he was furious.
\par }{\Q A fire
broke out against Jacob,
\par }{\Q and his anger
flared
up
against Israel,
\par }{\Q \VS{22}because
they did not
have faith
in God,
\par }{\Q and did not
trust
his ability to deliver them.
\par }{\Q \VS{23}He gave a command
to the clouds
above,
\par }{\Q and opened
the doors
in the sky.
\par }{\Q \VS{24}He rained
down manna
for them to eat;
\par }{\Q he gave
them the grain
of heaven.
\par }{\Q \VS{25}Man
ate
the food
of the mighty
ones.

\par }{\Q He sent
them
more than enough
to eat.
\par }{\Q \VS{26}He brought
the east wind
through the sky,
\par }{\Q and by his strength
led
forth the south wind.
\par }{\Q \VS{27}He rained down
meat
on
them like dust,
\par }{\Q birds
as numerous as the sand
on the seashores.
\par }{\Q \VS{28}He caused them to fall
right in the middle
of their camp,
\par }{\Q all around
their homes.
\par }{\Q \VS{29}They ate
until they were
stuffed;
\par }{\Q he gave
them what they
desired.
\par }{\Q \VS{30}They were not
yet filled up,
\par }{\Q their food
was still
in their mouths,
\par }{\Q \VS{31}when the anger
of God
flared up against
them.
\par }{\Q He killed
some of the strongest
of them;
\par }{\Q he brought the young men
of Israel
to their knees.
\par }{\Q \VS{32}Despite all
this,
they continued
to sin,
\par }{\Q and did not
trust
him to do amazing things.
\par }{\Q \VS{33}So he caused
them to die
unsatisfied
\par }{\Q and filled with terror.
\par }{\Q \VS{34}When
he struck
them down, they sought
his favor;

\par }{\Q they turned back
and longed
for God.
\par }{\Q \VS{35}They remembered
that
God
was their protector,
\par }{\Q and that the sovereign
God
was their deliverer.
\par }{\Q \VS{36}But they deceived
him with their words,
\par }{\Q and lied to him.
\par }{\Q \VS{37}They were not
really committed
to
him,

\par }{\Q and they were unfaithful
to his covenant.
\par }{\Q \VS{38}Yet he is
compassionate.
\par }{\Q He forgives
sin
and does not
destroy.
\par }{\Q He often
holds back
his anger,
\par }{\Q and does not
stir
up his fury.
\par }{\Q \VS{39}He remembered
that
they were
made of flesh,
\par }{\Q and were
like a wind
that blows
past and does not
return.
\par }{\Q \VS{40}How
often they rebelled
against him in the wilderness,
\par }{\Q and insulted
him in the desert!
\par }{\Q \VS{41}They again
challenged
God,
\par }{\Q and offended
the Holy One
of Israel.
\par }{\Q \VS{42}They did not
remember
what he had
done,

\par }{\Q how he delivered
them from
the enemy,
\par }{\Q \VS{43}when
he performed
his awesome
deeds in Egypt,
\par }{\Q and his acts of judgment
in the region
of Zoan.
\par }{\Q \VS{44}He turned
their rivers
into blood,
\par }{\Q and they could
not drink
from their streams.
\par }{\Q \VS{45}He sent
swarms
of biting
insects against them,

\par }{\Q as well as frogs
that overran their land.
\par }{\Q \VS{46}He gave
their crops
to the grasshopper,
\par }{\Q the fruit of their labor
to the locust.
\par }{\Q \VS{47}He destroyed
their vines
with hail,
\par }{\Q and their sycamore-fig trees
with driving rain.
\par }{\Q \VS{48}He rained
hail
down
on their cattle,
\par }{\Q and hurled lightning bolts
down on their livestock.
\par }{\Q \VS{49}His raging
anger
lashed out
against them,

\par }{\Q He sent
fury,
rage,
and trouble
\par }{\Q as messengers
who bring disaster.
\par }{\Q \VS{50}He sent
his anger
in full force;
\par }{\Q he did not
spare
them from death;
\par }{\Q he handed
their lives
over
to destruction.
\par }{\Q \VS{51}He struck
down all
the firstborn
in Egypt,
\par }{\Q the firstfruits
of their reproductive power
in the tents
of Ham.
\par }{\Q \VS{52}Yet
he brought out his people
like sheep;
\par }{\Q he led
them through the wilderness
like a flock.
\par }{\Q \VS{53}He guided
them safely
along,
\par }{\Q while the sea
covered
their enemies.
\par }{\Q \VS{54}He brought
them to
the border
of his holy land,
\par }{\Q to this
mountainous land
which his right hand
acquired.
\par }{\Q \VS{55}He drove
the nations
out from before
them;
\par }{\Q he assigned
them their tribal
allotments
\par }{\Q and allowed the tribes
of Israel
to settle down.
\par }{\Q \VS{56}Yet they challenged
and defied
the sovereign
God,
\par }{\Q and did not
obey his commands.
\par }{\Q \VS{57}They were unfaithful
and acted
as treacherously
as their ancestors;
\par }{\Q they were as unreliable
as a malfunctioning
bow.
\par }{\Q \VS{58}They made him angry
with their pagan shrines,
\par }{\Q and made him jealous
with their idols.
\par }{\Q \VS{59}God
heard
and was angry;
\par }{\Q he completely
rejected
Israel.
\par }{\Q \VS{60}He abandoned
the sanctuary
at Shiloh,
\par }{\Q the tent
where
he lived
among men.
\par }{\Q \VS{61}He allowed
the symbol of his strong
presence to be captured;
\par }{\Q he gave the symbol of his splendor
into the hand
of the enemy.
\par }{\Q \VS{62}He delivered
his people
over to the sword,
\par }{\Q and was angry
with his chosen nation.
\par }{\Q \VS{63}Fire
consumed
their young men,
\par }{\Q and their virgins
remained unmarried.
\par }{\Q \VS{64}Their priests
fell
by the sword,
\par }{\Q but their widows
did not
weep.
\par }{\Q \VS{65}But then
the Lord
awoke
from his sleep;
\par }{\Q he was like a warrior
in a drunken rage.
\par }{\Q \VS{66}He drove
his enemies
back;
\par }{\Q he made them a permanent
target
for insults.
\par }{\Q \VS{67}He rejected
the tent
of Joseph;
\par }{\Q he did not
choose
the tribe
of Ephraim.
\par }{\Q \VS{68}He chose
the tribe
of Judah,
\par }{\Q and Mount
Zion,
which
he loves.
\par }{\Q \VS{69}He made his sanctuary
as enduring
as the heavens above;

\par }{\Q as secure as the earth,
which he established
permanently.
\par }{\Q \VS{70}He chose
David,
his servant,
\par }{\Q and took
him from the sheepfolds.
\par }{\Q \VS{71}He took
him away from following
the mother sheep,
\par }{\Q and made him the shepherd
of Jacob,
his people,
\par }{\Q and of Israel,
his chosen nation.
\par }{\Q \VS{72}David cared for
them with pure motives;
\par }{\Q he led
them with skill.


\par }\PsalmChap{79}{\D 
A psalm
of Asaph.
\par }{\Q \VerseOne{1}O God,
foreigners
have invaded
your chosen land;
\par }{\Q they have polluted
your holy
temple
\par }{\Q and turned
Jerusalem
into a heap
of ruins.
\par }{\Q \VS{2}They have given
the
corpses
of your servants
\par }{\Q to the birds
of the sky;
\par }{\Q the flesh
of your loyal followers
\par }{\Q to the beasts
of the earth.
\par }{\Q \VS{3}They have made
their blood
flow
like water
\par }{\Q all around
Jerusalem,
and there is no
one to bury them.
\par }{\Q \VS{4}We have become
an object of disdain
to our neighbors;
\par }{\Q those who live on our borders
taunt
and insult us.
\par }{\Q \VS{5}How
long
will this go on, O
{\ND{Lord}}?

\par }{\Q Will you stay angry
forever?
\par }{\Q How long will your rage
burn
like
fire?
\par }{\Q \VS{6}Pour out
your anger
on
the nations
that
do not
acknowledge
you,

\par }{\Q on
the kingdoms
that
do not
pray
to you!
\par }{\Q \VS{7}For
they have devoured
Jacob
\par }{\Q and destroyed
his home.
\par }{\Q \VS{8}Do not
hold
us accountable
for the sins
of earlier
generations!

\par }{\Q Quickly
send
your compassion
our way,

\par }{\Q for
we are in serious trouble!
\par }{\Q \VS{9}Help
us, O God,
our deliverer!
\par }{\Q For the sake
of your glorious
reputation,
rescue
us!
\par }{\Q Forgive
our sins
for the sake
of your reputation!
\par }{\Q \VS{10}Why
should the nations
say,
“Where
is their God?”
\par }{\Q Before our very eyes
may the shed
blood
of your servants
\par }{\Q be avenged
among the nations!
\par }{\Q \VS{11}Listen to the painful cries
of the prisoners!

\par }{\Q Use your great
strength
to set free
those condemned
to die!
\par }{\Q \VS{12}Pay back
our neighbors
in full!

\par }{\Q May they be insulted
the same way they insulted
you, O Lord!
\par }{\Q \VS{13}Then we,
your people,
the sheep
of your pasture,
\par }{\Q will continually
thank
you.

\par }{\Q We will tell
coming generations
of your praiseworthy acts.


\par }\PsalmChap{80}{\D 
For the music director;
according to
{\IT{the
shushan-eduth}} style;
a psalm
of Asaph.
\par }{\Q \VerseOne{1}O shepherd
of Israel,
pay attention,
\par }{\Q you who lead
Joseph
like a flock
of sheep!
\par }{\Q You who
sit enthroned
above the winged
angels, reveal your splendor!
\par }{\Q \VS{2}In the sight
of Ephraim,
Benjamin,
and Manasseh
reveal
your power!
\par }{\Q Come
and deliver us!
\par }{\Q \VS{3}O God,
restore
us!
\par }{\Q Smile
on us! Then we will be delivered!
\par }{\Q \VS{4}O
{\ND{Lord}}
God,
invincible
warrior!

\par }{\Q How long
will you remain angry
at your people
while they pray to you?
\par }{\Q \VS{5}You have given
them tears
as food;
\par }{\Q you have made them drink
tears
by the measure.
\par }{\Q \VS{6}You have made
our neighbors
dislike
us,

\par }{\Q and our enemies
insult us.
\par }{\Q \VS{7}O God,
invincible
warrior, restore
us!
\par }{\Q Smile
on us! Then we will be delivered!
\par }{\Q \VS{8}You uprooted a vine
from
Egypt;
\par }{\Q you drove
out nations
and transplanted it.
\par }{\Q \VS{9}You cleared
the ground for
it;

\par }{\Q it took
root,
\par }{\Q and filled
the land.
\par }{\Q \VS{10}The mountains
were covered
by its shadow,
\par }{\Q the highest
cedars
by its branches.
\par }{\Q \VS{11}Its branches
reached
the Mediterranean
Sea,
\par }{\Q and its shoots
the Euphrates River.
\par }{\Q \VS{12}Why
did you break down
its walls,
\par }{\Q so that all
who pass by
pluck its fruit?
\par }{\Q \VS{13}The wild boars
of the forest
ruin it;

\par }{\Q the insects
of the field
feed on it.
\par }{\Q \VS{14}O God,
invincible
warrior, come back!
\par }{\Q Look
down from heaven
and take notice!
\par }{\Q Take care
of this
vine,
\par }{\Q \VS{15}the root
your right hand
planted,
\par }{\Q the shoot
you made to grow!
\par }{\Q \VS{16}It is burned
and cut down.
\par }{\Q They die
because you are displeased with them.
\par }{\Q \VS{17}May
you give support
to the one
you have chosen,
\par }{\Q to the one
whom you raised up for yourself!
\par }{\Q \VS{18}Then we will not
turn away
from
you.
\par }{\Q Revive
us and we will pray
to you!
\par }{\Q \VS{19}O
{\ND{Lord}}
God,
invincible
warrior, restore
us!
\par }{\Q Smile
on us! Then we will be delivered!


\par }\PsalmChap{81}{\D 
For the music director;
{\IT{according to the
gittith}} style;
by Asaph.
\par }{\Q \VerseOne{1}Shout
for joy to God,
our source of strength!
\par }{\Q Shout
out to the God
of Jacob!
\par }{\Q \VS{2}Sing
a song
and play
the tambourine,
\par }{\Q the pleasant
sounding harp,
and the ten-stringed instrument!
\par }{\Q \VS{3}Sound the ram’s horn
on the day of the new moon,
\par }{\Q and on the day
of the full moon
when our festival begins.
\par }{\Q \VS{4}For
observing the festival is a requirement
for Israel;
\par }{\Q it is
an ordinance
given by
the God
of Jacob.
\par }{\Q \VS{5}He decreed
it as a regulation
in Joseph,
\par }{\Q when he attacked
the land
of Egypt.
\par }{\Q I heard
a voice
I did not
recognize.
\par }{\Q \VS{6}It said: “I removed
the burden
from his shoulder;
\par }{\Q his hands
were released
from holding
the basket.
\par }{\Q \VS{7}In your distress
you called out
and I rescued
you.
\par }{\Q I answered
you from a dark
thundercloud.
\par }{\Q I tested
you at
the waters
of Meribah.
(Selah)
\par }{\Q \VS{8}I said, ‘Listen,
my people!
\par }{\Q I will warn
you!
\par }{\Q O Israel,
if
only you would obey me!
\par }{\Q \VS{9}There must be
no other
god
among you.
\par }{\Q You must not
worship
a foreign
god.
\par }{\Q \VS{10}I am
the {\ND{Lord}}, your God,
\par }{\Q the one who brought
you out of the land
of Egypt.
\par }{\Q Open
your mouth
wide and I will fill it!’
\par }{\Q \VS{11}But my people
did not
obey
me;
\par }{\Q Israel
did not
submit to me.
\par }{\Q \VS{12}I gave
them over to their stubborn
desires;
\par }{\Q they did
what seemed
right to them.
\par }{\Q \VS{13}If only
my people
would obey
me!

\par }{\Q If only Israel
would keep my commands!
\par }{\Q \VS{14}Then
I would quickly
subdue
their enemies,
\par }{\Q and attack
their adversaries.”
\par }{\Q \VS{15}(May those who hate
the {\ND{Lord}}
cower
in fear before him!
\par }{\Q May they be
permanently
humiliated!)
\par }{\Q \VS{16}“I would feed
Israel the best
wheat,
\par }{\Q and would satisfy
your appetite with honey
from the rocky cliffs.”


\par }\PsalmChap{82}{\D 
A psalm
of Asaph.
\par }{\Q \VerseOne{1}God
stands
in the assembly
of El;
\par }{\Q in the midst
of the gods
he renders judgment.
\par }{\Q \VS{2}He says, “How long
will you make
unjust
legal decisions
\par }{\Q and show favoritism
to the wicked? (Selah)
\par }{\Q \VS{3}Defend
the cause of the poor
and the fatherless!

\par }{\Q Vindicate
the oppressed
and suffering!
\par }{\Q \VS{4}Rescue
the poor
and needy!
\par }{\Q Deliver
them from the power
of the wicked!
\par }{\Q \VS{5}They neither
know
nor
understand.
\par }{\Q They stumble
around in the dark,
\par }{\Q while all
the foundations
of the earth
crumble.
\par }{\Q \VS{6}I
thought, ‘You are gods;
\par }{\Q all
of you
are sons
of the Most High.’
\par }{\Q \VS{7}Yet
you will die
like mortals;
\par }{\Q you will fall
like all the other
rulers.”
\par }{\Q \VS{8}Rise
up, O God,
and execute judgment
on the earth!
\par }{\Q For
you
own
all
the nations.


\par }\PsalmChap{83}{\D 
A song,
a psalm
of Asaph.
\par }{\Q \VerseOne{1}O God,
do not
be silent!
\par }{\Q Do not
ignore
us! Do not
be inactive,
O God!
\par }{\Q \VS{2}For
look,
your enemies
are making a commotion;
\par }{\Q those who hate
you are hostile.
\par }{\Q \VS{3}They carefully plot against
your people,
\par }{\Q and make plans to harm the ones you cherish.
\par }{\Q \VS{4}They say,
“Come on, let’s
annihilate
them so they are no longer a nation!

\par }{\Q Then the name
of Israel
will be remembered
no
more.”
\par }{\Q \VS{5}Yes,
they devise
a unified
strategy;
\par }{\Q they form
an alliance
against you.
\par }{\Q \VS{6}It includes the tents
of Edom
and the Ishmaelites,
\par }{\Q Moab
and the Hagrites,
\par }{\Q \VS{7}Gebal,
Ammon,
and Amalek,
\par }{\Q Philistia
and the inhabitants
of Tyre.
\par }{\Q \VS{8}Even
Assyria
has allied
with
them,
\par }{\Q lending its strength
to the descendants
of Lot.
(Selah)
\par }{\Q \VS{9}Do
to them
as you did to Midian –
\par }{\Q as you did to Sisera
and Jabin
at the Kishon
River!
\par }{\Q \VS{10}They were destroyed
at Endor;
\par }{\Q their corpses were like manure
on the ground.
\par }{\Q \VS{11}Make
their nobles
like Oreb
and Zeeb,
\par }{\Q and all
their rulers
like Zebah
and Zalmunna,
\par }{\Q \VS{12}who
said, “Let’s take
over the pastures
of God!”
\par }{\Q \VS{13}O my God,
make
them like dead thistles,
\par }{\Q like dead weeds blown
away by the wind!
\par }{\Q \VS{14}Like the fire
that burns down
the forest,
\par }{\Q or the flames
that consume
the mountainsides,
\par }{\Q \VS{15}chase
them with your gale winds,
\par }{\Q and terrify
them with your windstorm.
\par }{\Q \VS{16}Cover
their faces
with shame,
\par }{\Q so they might
seek
you, O

{\ND{Lord}}.
\par }{\Q \VS{17}May they be humiliated
and continually
terrified!

\par }{\Q May they die
in shame!
\par }{\Q \VS{18}Then they will know
that
you
alone
are the
{\ND{Lord}},

\par }{\Q the sovereign
king over
all
the earth.


\par }\PsalmChap{84}{\D 
For the music director;
according
{\IT{to the
gittith}} style;
written by the Korahites,
a psalm.
\par }{\Q \VerseOne{1}How
lovely
is the place where you live,
\par }{\Q O
{\ND{Lord}}
who rules over all!
\par }{\Q \VS{2}I desperately
want to be

\par }{\Q in the courts
of the
{\ND{Lord}}’s
temple.

\par }{\Q My heart
and my entire
being shout
for joy
\par }{\Q to
the living
God.
\par }{\Q \VS{3}Even
the birds
find
a home
there,
\par }{\Q and the swallow
builds a nest,
\par }{\Q where
she can protect her young
\par }{\Q near your altars,
O
{\ND{Lord}}
who rules over all,
\par }{\Q my king
and my God.
\par }{\Q \VS{4}How blessed
are those who live
in your temple
\par }{\Q and praise
you continually! (Selah)
\par }{\Q \VS{5}How blessed
are those who
find their strength
in you,
\par }{\Q and long to travel the roads that lead to your temple!
\par }{\Q \VS{6}As they pass
through the Baca
Valley,
\par }{\Q he provides
a spring
for them.

\par }{\Q The rain even
covers
it with pools of water.
\par }{\Q \VS{7}They are sustained
as they travel along;

\par }{\Q each one appears
before
God
in Zion.
\par }{\Q \VS{8}O
{\ND{Lord}}, sovereign
God,
\par }{\Q hear
my prayer!
\par }{\Q Listen,
O God
of Jacob! (Selah)
\par }{\Q \VS{9}O God,
take notice
of our shield!

\par }{\Q Show
concern
for your chosen king!
\par }{\Q \VS{10}Certainly
spending
just one day
in your temple courts
is better
\par }{\Q than spending
a thousand
elsewhere.
\par }{\Q I would rather stand at the entrance to the temple
of my God
\par }{\Q than live in the tents
of the wicked.
\par }{\Q \VS{11}For
the {\ND{Lord}}
God
is our sovereign protector.
\par }{\Q The
{\ND{Lord}}
bestows
favor
and honor;
\par }{\Q he withholds
no
good
thing from those who have integrity.
\par }{\Q \VS{12}O
{\ND{Lord}}
who rules over all,

\par }{\Q how blessed
are those
who trust in you!


\par }\PsalmChap{85}{\D 
For the music director;
written by the Korahites,
a psalm.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, you showed favor
to your land;
\par }{\Q you restored
the well-being
of Jacob.
\par }{\Q \VS{2}You pardoned
the wrongdoing
of your people;
\par }{\Q you forgave
all
their sin.
(Selah)
\par }{\Q \VS{3}You withdrew
all
your fury;
\par }{\Q you turned back
from your raging
anger.
\par }{\Q \VS{4}Restore
us, O God
our deliverer!
\par }{\Q Do not
be displeased
with us!
\par }{\Q \VS{5}Will you stay mad
at us forever?
\par }{\Q Will you remain
angry
throughout
future generations?
\par }{\Q \VS{6}Will you
not
revive
us once
more?
\par }{\Q Then your people
will rejoice in you!
\par }{\Q \VS{7}O
{\ND{Lord}}, show
us your loyal love!
\par }{\Q Bestow
on us your deliverance!
\par }{\Q \VS{8}I will listen
to what
God
the {\ND{Lord}}
says.
\par }{\Q For
he will make
peace
with
his people,
his faithful followers.
\par }{\Q Yet they must not
return
to their foolish ways.
\par }{\Q \VS{9}Certainly
his loyal followers
will soon
experience his deliverance;
\par }{\Q then his splendor
will again appear in our land.
\par }{\Q \VS{10}Loyal love
and faithfulness
meet;
\par }{\Q deliverance
and peace
greet each other with a kiss.
\par }{\Q \VS{11}Faithfulness
grows
from the ground,
\par }{\Q and deliverance
looks
down from the sky.
\par }{\Q \VS{12}Yes,
the {\ND{Lord}}
will bestow
his good
blessings,

\par }{\Q and our land
will yield
its crops.
\par }{\Q \VS{13}Deliverance
goes
before
him,
\par }{\Q and prepares
a pathway for him.


\par }\PsalmChap{86}{\D 
A prayer
of David.
\par }{\Q \VerseOne{1}Listen
O
{\ND{Lord}}! Answer
me!
\par }{\Q For
I
am oppressed
and needy.
\par }{\Q \VS{2}Protect
me,
for
I am
loyal!
\par }{\Q O
my God,
deliver
your servant,
who trusts in you!
\par }{\Q \VS{3}Have mercy
on me, O Lord,
\par }{\Q for
I cry
out to you all
day long!
\par }{\Q \VS{4}Make your servant
glad,
\par }{\Q for
to
you, O Lord,
I pray!
\par }{\Q \VS{5}Certainly
O
Lord,
you are
kind
and forgiving,
\par }{\Q and show great
faithfulness
to all
who cry out to you.
\par }{\Q \VS{6}O
{\ND{Lord}}, hear
my prayer!
\par }{\Q Pay attention
to my plea
for mercy!
\par }{\Q \VS{7}In my time
of trouble
I cry
out to you,
\par }{\Q for
you will answer me.
\par }{\Q \VS{8}None
can compare
to you among the gods,
O Lord!
\par }{\Q Your exploits are incomparable!
\par }{\Q \VS{9}All
the nations,
whom
you created,
\par }{\Q will come
and worship
you, O Lord.
\par }{\Q They will honor
your name.
\par }{\Q \VS{10}For
you
are great
and do
amazing
things.
\par }{\Q You
alone
are
God.
\par }{\Q \VS{11}O
{\ND{Lord}}, teach
me how
you want me to live!

\par }{\Q Then I will obey
your commands.
\par }{\Q Make me wholeheartedly
committed to you!
\par }{\Q \VS{12}O Lord,
my God,
I will give you thanks
with my whole
heart!
\par }{\Q I will honor
your name
continually!
\par }{\Q \VS{13}For
you will extend your great
loyal love
to me,

\par }{\Q and will deliver
my life
from the depths
of Sheol.
\par }{\Q \VS{14}O God,
arrogant men
attack
me;

\par }{\Q a gang
of ruthless
men, who do not
respect
you, seek
my life.
\par }{\Q \VS{15}But you,
O Lord,
are a compassionate
and merciful
God.
\par }{\Q You are patient
and demonstrate great
loyal love
and faithfulness.
\par }{\Q \VS{16}Turn
toward me
and have mercy
on me!
\par }{\Q Give
your servant
your strength!
\par }{\Q Deliver
your slave!
\par }{\Q \VS{17}Show
me evidence
of your favor!

\par }{\Q Then those who hate
me will see
it and be ashamed,
\par }{\Q for
you,
O
{\ND{Lord}}, will help
me and comfort me.


\par }\PsalmChap{87}{\D 
Written by the Korahites;
a psalm,
a song.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}’s city
is in the holy
hills.
\par }{\Q \VS{2}The
{\ND{Lord}}
loves
the gates
of Zion
\par }{\Q more than all
the dwelling places
of Jacob.
\par }{\Q \VS{3}People say wonderful
things about
you,

\par }{\Q O city
of God.
(Selah)
\par }{\Q \VS{4}I mention
Rahab
and Babylon
to my followers.
\par }{\Q Here
are Philistia
and Tyre,
along with
Ethiopia.
\par }{\Q It is said of them, “This
one was born
there.”
\par }{\Q \VS{5}But it is said
of Zion’s
residents,

\par }{\Q “Each
one
of these was born
in her,
\par }{\Q and the sovereign One
makes
her secure.”
\par }{\Q \VS{6}The
{\ND{Lord}}
writes in
the census
book of the nations,
\par }{\Q “This
one was born
there.”
(Selah)
\par }{\Q \VS{7}As for the singers,
as well as the pipers
–
\par }{\Q all of them sing within your walls.


\par }\PsalmChap{88}{\D 
A song,
a psalm
written by the Korahites;
for the music director;
according to
{\IT{the
machalath-leannoth}} style;
a well-written song
by Heman
the Ezrachite.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}
God
who delivers
me!

\par }{\Q By day
I cry out
\par }{\Q and at night
I pray before you.
\par }{\Q \VS{2}Listen to my prayer!

\par }{\Q Pay attention
to my cry for help!
\par }{\Q \VS{3}For
my life
is filled
with troubles
\par }{\Q and I am ready to enter
Sheol.
\par }{\Q \VS{4}They treat
me like
those who descend
into the grave.
\par }{\Q I am
like a helpless
man,
\par }{\Q \VS{5}adrift
among the dead,
\par }{\Q like
corpses
lying
in the grave,
\par }{\Q whom
you remember
no
more,
\par }{\Q and who are
cut off
from your power.
\par }{\Q \VS{6}You place me in
the lowest
regions
of the pit,
\par }{\Q in the dark places, in
the watery
depths.
\par }{\Q \VS{7}Your anger
bears
down on me,
\par }{\Q and you overwhelm me with all
your waves.
(Selah)
\par }{\Q \VS{8}You cause
those who know
me
to keep their distance;
\par }{\Q you make
me an appalling
sight to them.
\par }{\Q I am trapped
and cannot
get free.
\par }{\Q \VS{9}My eyes
grow weak
because of oppression.
\par }{\Q I call
out to you, O
{\ND{Lord}}, all
day
long;
\par }{\Q I spread
out my hands in prayer to you.
\par }{\Q \VS{10}Do
you accomplish
amazing things
for the dead?
\par }{\Q Do
the departed spirits
rise up
and give you thanks? (Selah)
\par }{\Q \VS{11}Is your loyal love
proclaimed
in the grave,
\par }{\Q or your faithfulness
in the place of the dead?
\par }{\Q \VS{12}Are your amazing
deeds experienced
in the dark
region,

\par }{\Q or your deliverance
in the land
of oblivion?
\par }{\Q \VS{13}As
for me,
I
cry out
to you, O
{\ND{Lord}};
\par }{\Q in the morning
my prayer
confronts you.
\par }{\Q \VS{14}O

{\ND{Lord}}, why
do you reject
me,
\par }{\Q and pay no attention to me?
\par }{\Q \VS{15}I am
oppressed
and have been on
the verge
of death since
my youth.
\par }{\Q I have been
subjected
to your horrors
and am numb
with pain.
\par }{\Q \VS{16}Your anger
overwhelms
me;

\par }{\Q your terrors
destroy me.
\par }{\Q \VS{17}They surround
me like water
all
day
long;
\par }{\Q they join forces and encircle me.
\par }{\Q \VS{18}You cause
my friends
and neighbors
to keep their distance;

\par }{\Q those who know
me leave me alone in the darkness.


\par }\PsalmChap{89}{\D 
A well-written song
by Ethan
the Ezrachite.
\par }{\Q \VerseOne{1}I will sing
continually about the
{\ND{Lord}}’s
faithful
deeds;
\par }{\Q to future
generations
I will proclaim
your faithfulness.
\par }{\Q \VS{2}For
I say,
“Loyal love
is permanently
established;
\par }{\Q in the skies
you set up
your faithfulness.”
\par }{\Q \VS{3}The
{\ND{Lord}} said,

\par }{\Q “I have made
a covenant
with my chosen one;
\par }{\Q I have made a promise on oath
to David,
my servant:
\par }{\Q \VS{4}‘I will give
you an
eternal
dynasty
\par }{\Q and establish
your
throne
throughout future generations.’ ”
(Selah)
\par }{\Q \VS{5}O

{\ND{Lord}}, the heavens
praise
your amazing
deeds,
\par }{\Q as well as your faithfulness
in the angelic
assembly.
\par }{\Q \VS{6}For
who
in the skies
can compare
to the
{\ND{Lord}}?
\par }{\Q Who is like
the {\ND{Lord}}
among the heavenly beings,
\par }{\Q \VS{7}a God
who is honored
in the great
angelic
assembly,
\par }{\Q and more
awesome
than all
who surround him?
\par }{\Q \VS{8}O
{\ND{Lord}}, sovereign
God!

\par }{\Q Who
is strong
like
you, O
{\ND{Lord}}?
\par }{\Q Your faithfulness
surrounds you.
\par }{\Q \VS{9}You
rule
over the proud
sea.
\par }{\Q When its waves
surge,
you
calm them.
\par }{\Q \VS{10}You
crushed
the Proud One
and killed
it;

\par }{\Q with your strong
arm
you scattered
your enemies.
\par }{\Q \VS{11}The heavens
belong to you, as does
the earth.
\par }{\Q You
made
the world
and all it contains.
\par }{\Q \VS{12}You
created
the north
and the south.
\par }{\Q Tabor
and Hermon
rejoice
in your name.
\par }{\Q \VS{13}Your arm
is powerful,
\par }{\Q your hand
strong,
\par }{\Q your right hand
victorious.
\par }{\Q \VS{14}Equity
and justice
are the foundation
of your throne.
\par }{\Q Loyal love
and faithfulness
characterize your rule.
\par }{\Q \VS{15}How blessed
are the people
who worship
you!

\par }{\Q O
{\ND{Lord}}, they experience
your favor.
\par }{\Q \VS{16}They rejoice
in your name
all
day
long,
\par }{\Q and are vindicated
by your justice.
\par }{\Q \VS{17}For
you give them splendor
and strength.
\par }{\Q By your
favor
we are victorious.
\par }{\Q \VS{18}For
our shield
belongs to the
{\ND{Lord}},
\par }{\Q our king
to the Holy One
of Israel.
\par }{\Q \VS{19}Then
you spoke
through a vision
to your faithful followers
and said:
\par }{\Q “I have energized
a warrior;
\par }{\Q I have raised up
a young man
from the people.
\par }{\Q \VS{20}I have discovered
David,
my servant.
\par }{\Q With my holy
oil
I have anointed him as king.
\par }{\Q \VS{21}My hand
will support
him,

\par }{\Q and my arm
will strengthen him.
\par }{\Q \VS{22}No
enemy
will be able to exact
tribute from him;

\par }{\Q a violent oppressor
will not
be able to humiliate him.
\par }{\Q \VS{23}I will crush
his enemies
before
him;
\par }{\Q I will strike
down those who hate him.
\par }{\Q \VS{24}He will experience my faithfulness
and loyal love,
\par }{\Q and by my name
he will win
victories.
\par }{\Q \VS{25}I will place
his hand over the sea,
\par }{\Q his right hand
over the rivers.
\par }{\Q \VS{26}He
will call
out to me,
\par }{\Q ‘You
are my father,
my God,
and the protector
who delivers me.’
\par }{\Q \VS{27}I
will appoint
him to be my firstborn son,
\par }{\Q the most
exalted of the earth’s
kings.
\par }{\Q \VS{28}I will always
extend
my loyal love
to him,
\par }{\Q and my covenant
with him is secure.
\par }{\Q \VS{29}I
will give
him an eternal
dynasty,
\par }{\Q and make his throne
as enduring
as the skies above.
\par }{\Q \VS{30}If
his sons
reject
my law
\par }{\Q and disobey my regulations,
\par }{\Q \VS{31}if
they break
my rules
\par }{\Q and do not
keep
my commandments,
\par }{\Q \VS{32}I will punish
their rebellion
by beating them with a club,
\par }{\Q their sin
by inflicting them with bruises.
\par }{\Q \VS{33}But I will not
remove
my loyal love
from
him,
\par }{\Q nor
be unfaithful
to my promise.
\par }{\Q \VS{34}I will not
break
my covenant
\par }{\Q or
go back
on what I promised.
\par }{\Q \VS{35}Once
and for all I have vowed
by my own holiness,
\par }{\Q I will never
deceive
David.
\par }{\Q \VS{36}His dynasty
will last
forever.
\par }{\Q His throne
will endure before
me, like the sun,
\par }{\Q \VS{37}it will remain
stable,
like the moon,
\par }{\Q his throne
will endure
like the skies.”
(Selah)
\par }{\Q \VS{38}But you
have spurned
and rejected
him;
\par }{\Q you are angry
with
your chosen king.
\par }{\Q \VS{39}You have repudiated
your covenant
with your servant;
\par }{\Q you have thrown
his crown
to the ground.
\par }{\Q \VS{40}You have broken down
all
his walls;
\par }{\Q you have made
his strongholds
a heap of ruins.
\par }{\Q \VS{41}All
who pass
by have robbed
him;
\par }{\Q he has become
an object of disdain
to his neighbors.
\par }{\Q \VS{42}You have allowed
his adversaries
to be victorious,
\par }{\Q and all
his enemies
to rejoice.
\par }{\Q \VS{43}You turn back
his sword
from the adversary,
\par }{\Q and have not
sustained
him in battle.
\par }{\Q \VS{44}You have brought to an end
his splendor,
\par }{\Q and have knocked his throne
to the ground.
\par }{\Q \VS{45}You have cut short
his youth,
\par }{\Q and have covered
him with shame.
(Selah)
\par }{\Q \VS{46}How
long,
O
{\ND{Lord}}, will this last?
\par }{\Q Will you remain hidden
forever?


\par }{\Q Will your anger
continue to burn
like
fire?
\par }{\Q \VS{47}Take note
of my brief lifespan!

\par }{\Q Why
do you make
all
people
so mortal?
\par }{\Q \VS{48}No
man
can live
on without
experiencing
death,
\par }{\Q or deliver
his life
from the power
of Sheol.
(Selah)
\par }{\Q \VS{49}Where
are your earlier
faithful
deeds, O Lord,
\par }{\Q the ones performed in accordance with your reliable
oath
to David?
\par }{\Q \VS{50}Take note,
O Lord,
of the way your servants
are taunted,
\par }{\Q and of how I must bear
so
many
insults
from people!
\par }{\Q \VS{51}Your enemies,
O
{\ND{Lord}}, hurl insults;
\par }{\Q they insult
your chosen king as they dog his footsteps.
\VS{52}\par }{\Q The
{\ND{Lord}}
deserves
praise forevermore!
\par }{\Q We agree! We agree!


\par }\PsalmChap{90}{\D 
A prayer
of Moses,
the man
of God.
\par }{\Q \VerseOne{1}O Lord,
you
have been our protector
through all generations!
\par }{\Q \VS{2}Even before
the mountains
came into existence,
\par }{\Q or you brought the world
into being,

\par }{\Q you were the eternal
God.
\par }{\Q \VS{3}You make mankind
return
to the dust,
\par }{\Q and say,
“Return,
O people!”
\par }{\Q \VS{4}Yes,
in your eyes
a thousand
years
\par }{\Q are like yesterday
that
quickly passes,
\par }{\Q or like one of the divisions of the nighttime.
\par }{\Q \VS{5}You bring their lives
to an end
and they “fall
asleep.”
\par }{\Q In the morning
they are like the grass
that sprouts up;
\par }{\Q \VS{6}in the morning
it glistens
and sprouts
up;
\par }{\Q at evening
time it withers
and dries up.
\par }{\Q \VS{7}Yes,
we are consumed
by your anger;
\par }{\Q we are terrified
by your wrath.
\par }{\Q \VS{8}You are aware
of our sins;
\par }{\Q you even know about our hidden sins.
\par }{\Q \VS{9}Yes,
throughout all
our days
we experience
your raging fury;
\par }{\Q the years
of our lives pass quickly,
like
a sigh.
\par }{\Q \VS{10}The days
of our lives
add up to seventy
years,
\par }{\Q or eighty,
if
one is especially strong.
\par }{\Q But even one’s best
years
are marred by trouble
and oppression.
\par }{\Q Yes,
they pass
quickly
and we fly away.
\par }{\Q \VS{11}Who can
really fathom
the intensity
of your
anger?

\par }{\Q Your raging fury
causes people to fear you.
\par }{\Q \VS{12}So teach
us to consider
our mortality,
\par }{\Q so that
we might live
wisely.
\par }{\Q \VS{13}Turn back
toward us, O
{\ND{Lord}}!
\par }{\Q How long
must this suffering
last?

\par }{\Q Have pity
on
your servants!
\par }{\Q \VS{14}Satisfy
us in the morning
with your loyal love!
\par }{\Q Then we will shout
for joy and be happy
all
our days!
\par }{\Q \VS{15}Make us happy
in proportion to the days
you have afflicted
us,
\par }{\Q in proportion to the years
we have experienced
trouble!
\par }{\Q \VS{16}May your servants
see your work!

\par }{\Q May their sons
see
your majesty!
\par }{\Q \VS{17}May
our sovereign
God
extend his favor
to
us!

\par }{\Q Make our endeavors
successful!

\par }{\Q Yes, make them successful!


\par }\PsalmChap{91}{\PP \par }{\Q \VerseOne{1}As for you, the one who lives
in the shelter
of the sovereign
One,

\par }{\Q and resides
in the protective shadow
of the mighty king –
\par }{\Q \VS{2}I say
this about the
{\ND{Lord}}, my shelter
and my stronghold,
\par }{\Q my God
in whom I trust –
\par }{\Q \VS{3}he
will certainly
rescue
you from the snare
of the hunter
\par }{\Q and from the destructive
plague.
\par }{\Q \VS{4}He will shelter
you with his wings;
\par }{\Q you will find safety
under
his wings.
\par }{\Q His faithfulness
is like a shield
or a protective wall.
\par }{\Q \VS{5}You need not
fear
the terrors
of the night,
\par }{\Q the arrow
that flies
by day,
\par }{\Q \VS{6}the plague
that comes
in the darkness,
\par }{\Q or the disease
that comes
at noon.
\par }{\Q \VS{7}Though a thousand
may fall
beside
you,
\par }{\Q and a multitude
on your right side,
\par }{\Q it will not
reach you.
\par }{\Q \VS{8}Certainly
you will see
it with your very own eyes
–
\par }{\Q you will see
the wicked
paid back.
\par }{\Q \VS{9}For
you
have taken refuge
in the
{\ND{Lord}},
\par }{\Q my shelter,
the sovereign One.
\par }{\Q \VS{10}No
harm
will overtake
you;
\par }{\Q no
illness
will come near
your home.
\par }{\Q \VS{11}For
he will order
his angels
\par }{\Q to protect
you in all
you do.
\par }{\Q \VS{12}They will lift
you up in their hands,
\par }{\Q so you
will not
slip
and fall on
a stone.
\par }{\Q \VS{13}You will subdue
a lion
and a snake;
\par }{\Q you will trample
underfoot
a young lion
and a serpent.
\par }{\Q \VS{14}The
{\ND{Lord}} says,

\par }{\Q “Because
he is devoted
to me,
I will deliver
him;
\par }{\Q I will protect
him because
he is loyal
to me.
\par }{\Q \VS{15}When he calls
out to me, I
will answer
him.
\par }{\Q I
will be with him
when he is in trouble;
\par }{\Q I
will rescue
him
and bring him honor.
\par }{\Q \VS{16}I will satisfy
him with long
life,
\par }{\Q and will let him see
my salvation.


\par }\PsalmChap{92}{\D 
A psalm;
a song
for the Sabbath
day.
\par }{\Q \VerseOne{1}It is fitting
to thank
the {\ND{Lord}},
\par }{\Q and to sing praises
to your name,
O sovereign One!
\par }{\Q \VS{2}It is fitting to proclaim your loyal love
in the morning,
\par }{\Q and your faithfulness
during the night,
\par }{\Q \VS{3}to the accompaniment of a ten-stringed
instrument
and a lyre,
\par }{\Q to the accompaniment
of the meditative tone of the harp.
\par }{\Q \VS{4}For
you, O
{\ND{Lord}}, have
made me happy
by your
work.
\par }{\Q I will sing
for joy because of what
you have done.
\par }{\Q \VS{5}How
great
are your works,
O
{\ND{Lord}}!
\par }{\Q Your plans are
very
intricate!
\par }{\Q \VS{6}The spiritually insensitive
do not
recognize
this;
\par }{\Q the fool
does not
understand
this.
\par }{\Q \VS{7}When the wicked
sprout
up like
grass,
\par }{\Q and all
the evildoers
glisten,
\par }{\Q it is so that they may be annihilated.
\par }{\Q \VS{8}But you,
O
{\ND{Lord}}, reign
forever!
\par }{\Q \VS{9}Indeed,
look
at your enemies,
O
{\ND{Lord}}!
\par }{\Q Indeed,
look
at how your enemies
perish!
\par }{\Q All
the evildoers
are scattered!
\par }{\Q \VS{10}You exalt
my horn
like that of a wild ox.
\par }{\Q I am covered
with fresh
oil.
\par }{\Q \VS{11}I gloat
in triumph over those who tried to ambush me;
\par }{\Q I hear
the defeated cries of the evil foes who attacked me.
\par }{\Q \VS{12}The godly
grow
like a palm tree;
\par }{\Q they grow
high like a cedar
in Lebanon.
\par }{\Q \VS{13}Planted
in the
{\ND{Lord}}’s
house,
\par }{\Q they grow
in the courts
of our God.
\par }{\Q \VS{14}They bear fruit
even when
they are old;
\par }{\Q they are
filled with vitality
and have many leaves.
\par }{\Q \VS{15}So they proclaim
that
the {\ND{Lord}}, my protector,
\par }{\Q is just and never
unfair.


\par }\PsalmChap{93}{\PP \par }{\Q \VerseOne{1}The
{\ND{Lord}}
reigns!
\par }{\Q He is robed
in majesty,
\par }{\Q the {\ND{Lord}}
is robed,
\par }{\Q he wears
strength
around his waist.

\par }{\Q Indeed,
the world
is established,
it cannot
be moved.
\par }{\Q \VS{2}Your throne
has been secure
from
ancient
times;
\par }{\Q you have always been king.
\par }{\Q \VS{3}The waves
roar,
O
{\ND{Lord}},
\par }{\Q the waves
roar,
\par }{\Q the waves
roar
and crash.
\par }{\Q \VS{4}Above the sound
of the surging
water,
\par }{\Q and the mighty
waves
of the sea,
\par }{\Q the {\ND{Lord}}
sits enthroned
in majesty.
\par }{\Q \VS{5}The rules
you set down
are completely
reliable.
\par }{\Q Holiness
aptly adorns
your house,
O
{\ND{Lord}}, forever.


\par }\PsalmChap{94}{\PP \par }{\Q \VerseOne{1}O
{\ND{Lord}}, the God
who avenges!
\par }{\Q O God
who avenges,
reveal
your splendor!
\par }{\Q \VS{2}Rise
up, O judge
of the earth!
\par }{\Q Pay
back
the proud!
\par }{\Q \VS{3}O

{\ND{Lord}}, how long
will the wicked,
\par }{\Q how long
will the wicked
celebrate?
\par }{\Q \VS{4}They spew
out threats and speak
defiantly;
\par }{\Q all
the evildoers
boast.
\par }{\Q \VS{5}O
{\ND{Lord}}, they crush
your people;
\par }{\Q they oppress
the nation
that belongs to you.
\par }{\Q \VS{6}They kill
the widow
and the one residing outside his native land,
\par }{\Q and they murder
the fatherless.
\par }{\Q \VS{7}Then they say,
“The
{\ND{Lord}}
does not
see
this;
\par }{\Q the God
of Jacob
does not
take notice of it.”
\par }{\Q \VS{8}Take notice
of this, you ignorant
people!

\par }{\Q You fools,
when
will you ever
understand?
\par }{\Q \VS{9}Does
the one who makes
the human ear
not
hear?
\par }{\Q Does
the one who forms
the human eye
not
see?
\par }{\Q \VS{10}Does
the one who disciplines
the nations
not
punish?
\par }{\Q He is the one who
imparts
knowledge
to human beings!
\par }{\Q \VS{11}The
{\ND{Lord}}
knows
that
\par }{\Q peoples’
thoughts
are morally bankrupt.
\par }{\Q \VS{12}How blessed
is the one
whom
you instruct,
O
{\ND{Lord}},
\par }{\Q the one whom
you teach
from your law,
\par }{\Q \VS{13}in order to protect
him from times
of trouble,
\par }{\Q until
the wicked
are destroyed.
\par }{\Q \VS{14}Certainly
the {\ND{Lord}}
does not
forsake
his people;
\par }{\Q he does not
abandon
the nation that belongs to him.
\par }{\Q \VS{15}For
justice
will prevail,
\par }{\Q and all
the morally
upright
will be vindicated.
\par }{\Q \VS{16}Who
will rise
up to defend me against
the wicked?
\par }{\Q Who
will stand
up for me against
the evildoers?
\par }{\Q \VS{17}If
the {\ND{Lord}}
had not helped
me,
\par }{\Q I would have laid down in the silence
of death.
\par }{\Q \VS{18}If
I say,
“My foot
is slipping,”
\par }{\Q your loyal love,
O
{\ND{Lord}}, supports me.
\par }{\Q \VS{19}When worries
threaten
to overwhelm
me,

\par }{\Q your soothing touch
makes
me
happy.
\par }{\Q \VS{20}Cruel
rulers are not your allies,
\par }{\Q those who make
oppressive
laws.
\par }{\Q \VS{21}They conspire
against
the blameless,
\par }{\Q and condemn
to death
the innocent.
\par }{\Q \VS{22}But the
{\ND{Lord}}
will protect
me,

\par }{\Q and my God
will shelter me.
\par }{\Q \VS{23}He will pay them
back
for their
sin.
\par }{\Q He will destroy
them because
of their
evil;
\par }{\Q the {\ND{Lord}}
our God
will destroy them.


\par }\PsalmChap{95}{\PP \par }{\Q \VerseOne{1}Come! Let’s
sing
for joy to the
{\ND{Lord}}!
\par }{\Q Let’s shout
out praises to our protector
who delivers us!
\par }{\Q \VS{2}Let’s
enter his presence
with thanksgiving!
\par }{\Q Let’s shout
out to him in celebration!
\par }{\Q \VS{3}For
the {\ND{Lord}}
is a great
God,
\par }{\Q a great
king
who is superior
to all
gods.
\par }{\Q \VS{4}The depths
of the earth
are in his hand,
\par }{\Q and the mountain
peaks belong to him.
\par }{\Q \VS{5}The sea
is
his, for he made
it.
\par }{\Q His hands
formed
the dry land.
\par }{\Q \VS{6}Come! Let’s bow down and worship!

\par }{\Q Let’s kneel
before
the {\ND{Lord}}, our creator!
\par }{\Q \VS{7}For
he is
our God;
\par }{\Q we
are the people
of his pasture,
\par }{\Q the sheep
he owns.
\par }{\Q Today,
if
only you would obey him!
\par }{\Q \VS{8}He says, “Do not
be stubborn
like they were at Meribah,
\par }{\Q like they were that day
at Massah
in the wilderness,
\par }{\Q \VS{9}where
your ancestors
challenged
my authority,

\par }{\Q and tried
my patience, even
though they had seen
my work.
\par }{\Q \VS{10}For forty
years
I was continually disgusted
with that generation,
\par }{\Q and I said,
‘These
people
desire
to go astray;
\par }{\Q they
do not
obey
my commands.’
\par }{\Q \VS{11}So I made a vow
in my anger,
\par }{\Q ‘They will never
enter
into
the resting place I had set aside for them.’ ”


\par }\PsalmChap{96}{\PP \par }{\Q \VerseOne{1}Sing
to the
{\ND{Lord}}
a new
song!

\par }{\Q Sing
to the
{\ND{Lord}}, all
the earth!
\par }{\Q \VS{2}Sing
to the
{\ND{Lord}}! Praise
his name!
\par }{\Q Announce
every day
how he delivers!
\par }{\Q \VS{3}Tell
the nations
about his splendor!
\par }{\Q Tell all
the nations
about his amazing deeds!
\par }{\Q \VS{4}For
the {\ND{Lord}}
is great
and certainly worthy
of praise;
\par }{\Q he is
more
awesome
than all
gods.
\par }{\Q \VS{5}For
all
the gods
of the nations
are worthless,
\par }{\Q but the
{\ND{Lord}}
made
the sky.
\par }{\Q \VS{6}Majestic
splendor
emanates
from him;

\par }{\Q his sanctuary
is firmly established
and beautiful.
\par }{\Q \VS{7}Ascribe
to the
{\ND{Lord}}, O families
of the nations,
\par }{\Q ascribe
to the
{\ND{Lord}}
splendor
and strength!
\par }{\Q \VS{8}Ascribe
to the
{\ND{Lord}}
the splendor
he deserves!

\par }{\Q Bring
an offering
and enter
his courts!
\par }{\Q \VS{9}Worship
the {\ND{Lord}}
in holy
attire!


\par }{\Q Tremble
before
him, all
the earth!
\par }{\Q \VS{10}Say
among the nations,
“The
{\ND{Lord}}
reigns!
\par }{\Q The world
is established,
it cannot
be moved.
\par }{\Q He judges
the nations
fairly.”
\par }{\Q \VS{11}Let
the sky
rejoice,
and the earth
be happy!
\par }{\Q Let
the sea
and everything in it shout!
\par }{\Q \VS{12}Let the fields
and everything
in them celebrate!
\par }{\Q Then
let the trees
of the forest
shout with joy
\par }{\Q \VS{13}before
the {\ND{Lord}}, for
he comes!
\par }{\Q For
he comes
to judge
the earth!
\par }{\Q He judges
the world
fairly,
\par }{\Q and the nations
in accordance with his justice.


\par }\PsalmChap{97}{\PP \par }{\Q \VerseOne{1}The
{\ND{Lord}}
reigns!
\par }{\Q Let
the earth
be happy!
\par }{\Q Let the many
coastlands
rejoice!
\par }{\Q \VS{2}Dark
clouds
surround
him;
\par }{\Q equity
and justice
are the foundation
of his throne.
\par }{\Q \VS{3}Fire
goes
before
him;
\par }{\Q on
every side
it burns up
his enemies.
\par }{\Q \VS{4}His lightning
bolts light
up the world;
\par }{\Q the earth
sees
and trembles.
\par }{\Q \VS{5}The mountains
melt
like wax
before
the {\ND{Lord}},
\par }{\Q before
the Lord
of the whole
earth.
\par }{\Q \VS{6}The sky
declares
his justice,
\par }{\Q and all
the nations
see
his splendor.
\par }{\Q \VS{7}All
who worship
idols
are ashamed,
\par }{\Q those who boast
about worthless idols.
\par }{\Q All
the gods
bow down before him.
\par }{\Q \VS{8}Zion
hears
and rejoices,
\par }{\Q the towns
of Judah
are happy,
\par }{\Q because
of your judgments,
O
{\ND{Lord}}.
\par }{\Q \VS{9}For
you,
O
{\ND{Lord}}, are the sovereign
king over
the whole
earth;
\par }{\Q you are elevated
high above
all
gods.
\par }{\Q \VS{10}You who love
the {\ND{Lord}}, hate
evil!
\par }{\Q He protects
the lives
of his faithful followers;
\par }{\Q he delivers
them from the power
of the wicked.
\par }{\Q \VS{11}The godly
bask
in the light;
\par }{\Q the morally
upright
experience joy.
\par }{\Q \VS{12}You godly
ones, rejoice
in the
{\ND{Lord}}!
\par }{\Q Give thanks
to his holy
name.


\par }\PsalmChap{98}{\D 
A psalm.
\par }{\Q \VerseOne{1}Sing
to the
{\ND{Lord}}
a new
song,
\par }{\Q for
he performs
amazing
deeds!
\par }{\Q His right hand
and his mighty arm
\par }{\Q accomplish
deliverance.
\par }{\Q \VS{2}The
{\ND{Lord}}
demonstrates
his power to deliver;
\par }{\Q in the sight
of the nations
he reveals
his justice.
\par }{\Q \VS{3}He remains
loyal
and faithful
to the family
of Israel.
\par }{\Q All
the ends
of the earth
see
our God
deliver us.
\par }{\Q \VS{4}Shout
out praises to the
{\ND{Lord}}, all
the earth!
\par }{\Q Break out in a joyful
shout
and sing!
\par }{\Q \VS{5}Sing
to the
{\ND{Lord}}
accompanied by a harp,
\par }{\Q accompanied by a harp
and the sound
of music!
\par }{\Q \VS{6}With trumpets
and the blaring
of the ram’s horn,
\par }{\Q shout
out praises before
the king,
the {\ND{Lord}}!
\par }{\Q \VS{7}Let
the sea
and everything
in it shout,
\par }{\Q along with the world
and those who live in it!
\par }{\Q \VS{8}Let the rivers
clap
their hands!
\par }{\Q Let the mountains
sing
in unison
\par }{\Q \VS{9}before
the {\ND{Lord}}!
\par }{\Q For
he comes
to judge
the earth!
\par }{\Q He judges
the world
fairly,
\par }{\Q and the nations
in a just manner.


\par }\PsalmChap{99}{\PP \par }{\Q \VerseOne{1}The
{\ND{Lord}}
reigns!
\par }{\Q The nations
tremble.
\par }{\Q He sits enthroned
above the winged
angels;
\par }{\Q the earth shakes.
\par }{\Q \VS{2}The
{\ND{Lord}}
is elevated
in Zion;
\par }{\Q he is
exalted
over
all
the nations.
\par }{\Q \VS{3}Let them praise
your great
and awesome
name!
\par }{\Q He is
holy!
\par }{\Q \VS{4}The king
is strong;
\par }{\Q he loves
justice.
\par }{\Q You
ensure
that legal decisions will be made
fairly;
\par }{\Q you
promote
justice
and equity in Jacob.
\par }{\Q \VS{5}Praise
the {\ND{Lord}}
our God!
\par }{\Q Worship
before his footstool!
\par }{\Q He
is holy!
\par }{\Q \VS{6}Moses
and Aaron
were among his priests;
\par }{\Q Samuel
was one of those who prayed
to him.

\par }{\Q They prayed
to the
{\ND{Lord}}
and he
answered them.
\par }{\Q \VS{7}He spoke
to them
from a pillar
of cloud;
\par }{\Q they obeyed
his regulations
and the ordinance
he gave them.
\par }{\Q \VS{8}O
{\ND{Lord}}
our God,
you
answered
them.
\par }{\Q They
found you to be a forgiving
God,
\par }{\Q but also one who punished
their sinful deeds.
\par }{\Q \VS{9}Praise
the {\ND{Lord}}
our God!
\par }{\Q Worship
on his holy
hill,
\par }{\Q for
the {\ND{Lord}}
our God
is holy!


\par }\PsalmChap{100}{\D 
A thanksgiving
psalm.
\par }{\Q \VerseOne{1}Shout
out praises to the
{\ND{Lord}}, all
the earth!
\par }{\Q \VS{2}Worship
the {\ND{Lord}}
with joy!
\par }{\Q Enter
his presence
with joyful singing!
\par }{\Q \VS{3}Acknowledge
that
the {\ND{Lord}}
is
God!
\par }{\Q He made
us and we belong to him;

\par }{\Q we
are his people,
the sheep
of his pasture.
\par }{\Q \VS{4}Enter
his gates
with thanksgiving,
\par }{\Q and his courts
with praise!
\par }{\Q Give him thanks!
\par }{\Q Praise
his name!
\par }{\Q \VS{5}For
the {\ND{Lord}}
is good.
\par }{\Q His loyal love
endures,
\par }{\Q and he is faithful
through all generations.


\par }\PsalmChap{101}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}I will sing
about loyalty
and justice!
\par }{\Q To you, O
{\ND{Lord}}, I will sing praises!
\par }{\Q \VS{2}I will walk in
the way
of integrity.
\par }{\Q When
will you come
to
me?
\par }{\Q I will conduct
my business with integrity
in the midst
of my palace.
\par }{\Q \VS{3}I will not
even consider
doing what
is dishonest.
\par }{\Q I hate
doing
evil;
\par }{\Q I will have no
part of it.
\par }{\Q \VS{4}I will have nothing
to do with a perverse
person;

\par }{\Q I
will not permit
evil.
\par }{\Q \VS{5}I will destroy
anyone who slanders
his neighbor
in secret.
\par }{\Q I will not
tolerate
anyone who has a cocky
demeanor
and an arrogant attitude.
\par }{\Q \VS{6}I will favor
the honest
people of the land,
\par }{\Q and allow them to live
with me.
\par }{\Q Those who walk
in the way
of integrity
will attend me.
\par }{\Q \VS{7}Deceitful
people will not
live
in
my palace.
\par }{\Q Liars
will not
be welcome
in my presence.
\par }{\Q \VS{8}Each morning
I will destroy
all
the wicked
people in the land,
\par }{\Q and remove
all
evildoers
from the city
of the {\ND{Lord}}.


\par }\PsalmChap{102}{\D 
The prayer
of an oppressed
man, as he grows faint
and pours out
his lament
before
the {\ND{Lord}}.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, hear
my prayer!
\par }{\Q Pay attention to my cry
for help!
\par }{\Q \VS{2}Do not
ignore
me in my time
of trouble!

\par }{\Q Listen
to
me!

\par }{\Q When
I call
out to you, quickly
answer me!
\par }{\Q \VS{3}For
my days
go up
in smoke,
\par }{\Q and my bones
are charred
like a fireplace.
\par }{\Q \VS{4}My heart
is parched
and withered
like grass,
\par }{\Q for
I am unable
to eat
food.
\par }{\Q \VS{5}Because of the anxiety that makes me groan,
\par }{\Q my bones
protrude from my skin.
\par }{\Q \VS{6}I
am like an owl
in the wilderness;
\par }{\Q I am like a screech owl
among the ruins.
\par }{\Q \VS{7}I stay awake;
\par }{\Q I am like
a solitary
bird
on
a roof.
\par }{\Q \VS{8}All
day
long my enemies
taunt
me;
\par }{\Q those who mock
me use my name in their curses.
\par }{\Q \VS{9}For
I eat
ashes
as if they were bread,
\par }{\Q and mix
my drink
with my tears,
\par }{\Q \VS{10}because
of your anger
and raging
fury.
\par }{\Q Indeed,
you pick
me up and throw me away.
\par }{\Q \VS{11}My days
are coming to an end,
\par }{\Q and I am
withered
like grass.
\par }{\Q \VS{12}But you,
O
{\ND{Lord}}, rule forever,
\par }{\Q and your reputation
endures.
\par }{\Q \VS{13}You
will rise
up and have compassion
on Zion.
\par }{\Q For
it is time
to have mercy
on her,
\par }{\Q for
the appointed
time has come.
\par }{\Q \VS{14}Indeed,
your servants
take delight
in her stones,
\par }{\Q and feel compassion
for the dust of her ruins.
\par }{\Q \VS{15}The nations
will respect
the
reputation
of the {\ND{Lord}},

\par }{\Q and all
the kings
of the earth
will respect his splendor,
\par }{\Q \VS{16}when
the {\ND{Lord}}
rebuilds
Zion,
\par }{\Q and reveals
his splendor,
\par }{\Q \VS{17}when
he responds
to
the prayer
of the destitute,
\par }{\Q and does not
reject
their request.
\par }{\Q \VS{18}The account of his intervention will be recorded
for future generations;
\par }{\Q people
yet to be born
will praise
the
{\ND{Lord}}.
\par }{\Q \VS{19}For
he will look
down from his sanctuary
above;
\par }{\Q from heaven
the {\ND{Lord}}
will look
toward earth,
\par }{\Q \VS{20}in order to hear
the painful
cries of the prisoners,
\par }{\Q and to set free
those condemned to die,
\par }{\Q \VS{21}so they may proclaim
the name
of the {\ND{Lord}}
in Zion,
\par }{\Q and praise
him in Jerusalem,
\par }{\Q \VS{22}when the nations
gather
together,
\par }{\Q and the kingdoms
pay tribute
to the
{\ND{Lord}}.
\par }{\Q \VS{23}He has taken away
my strength
in the middle of life;

\par }{\Q he has cut short
my days.
\par }{\Q \VS{24}I say,
“O my God,
please do not
take me away in the middle
of my life!

\par }{\Q You endure through all generations.
\par }{\Q \VS{25}In earlier
times you established
the earth;
\par }{\Q the skies
are your handiwork.
\par }{\Q \VS{26}They
will perish,
\par }{\Q but you
will endure.
\par }{\Q They will wear out
like a garment;
\par }{\Q like clothes
you will remove
them and they will disappear.
\par }{\Q \VS{27}But you
remain;

\par }{\Q your years
do not come to an end.
\par }{\Q \VS{28}The children
of your servants
will settle
down here,
\par }{\Q and their descendants
will live securely
in your presence.”


\par }\PsalmChap{103}{\D 
By David.
\par }{\Q \VerseOne{1}Praise
the {\ND{Lord}}, O my soul!
\par }{\Q With all
that is within
me, praise his holy
name!
\par }{\Q \VS{2}Praise
the {\ND{Lord}}, O my soul!
\par }{\Q Do not
forget
all
his kind deeds!
\par }{\Q \VS{3}He
is the one who forgives
all
your sins,
\par }{\Q who heals
all
your diseases,
\par }{\Q \VS{4}who delivers
your life
from the Pit,
\par }{\Q who crowns
you with his loyal love
and compassion,
\par }{\Q \VS{5}who satisfies
your life with good
things,
\par }{\Q so your youth
is renewed
like an eagle’s.
\par }{\Q \VS{6}The
{\ND{Lord}}
does
what is fair,
\par }{\Q and executes justice
for all
the oppressed.
\par }{\Q \VS{7}The
{\ND{Lord}} revealed
his faithful
acts to Moses,
\par }{\Q his deeds
to the Israelites.
\par }{\Q \VS{8}The
{\ND{Lord}}
is compassionate
and merciful;
\par }{\Q he is patient
and demonstrates great
loyal love.
\par }{\Q \VS{9}He does not
always
accuse,
\par }{\Q and does not
stay angry.
\par }{\Q \VS{10}He does not
deal with us as our sins
deserve;

\par }{\Q he does
not
repay us as our misdeeds deserve.
\par }{\Q \VS{11}For
as the skies
are high
above
the earth,
\par }{\Q so his loyal love
towers
over
his faithful followers.
\par }{\Q \VS{12}As far
as the eastern
horizon is from the west,
\par }{\Q so he removes
the guilt of our rebellious
actions from us.
\par }{\Q \VS{13}As a father
has compassion
on
his children,
\par }{\Q so the
{\ND{Lord}}
has compassion
on
his faithful followers.
\par }{\Q \VS{14}For
he knows
what we are made of;

\par }{\Q he realizes
we are made of clay.
\par }{\Q \VS{15}A person’s
life
is like grass.
\par }{\Q Like a flower
in the field
it flourishes,
\par }{\Q \VS{16}but when
the hot wind
blows by,
it disappears,
\par }{\Q and one can
no
longer
even spot
the place where it once grew.
\par }{\Q \VS{17}But the
{\ND{Lord}}
continually
shows loyal love
to his faithful followers,
\par }{\Q and is faithful
to their descendants,
\par }{\Q \VS{18}to those who keep
his covenant,
\par }{\Q who are careful
to obey
his commands.
\par }{\Q \VS{19}The
{\ND{Lord}}
has established
his throne
in heaven;
\par }{\Q his kingdom
extends
over everything.
\par }{\Q \VS{20}Praise
the {\ND{Lord}}, you angels
of his,
\par }{\Q you powerful
warriors
who carry
out his decrees
\par }{\Q and obey
his orders!
\par }{\Q \VS{21}Praise
the {\ND{Lord}}, all
you warriors
of his,

\par }{\Q you servants
of his who carry
out his desires!
\par }{\Q \VS{22}Praise
the {\ND{Lord}}, all
that he has made,
\par }{\Q in all
the regions
of his kingdom!
\par }{\Q Praise
the {\ND{Lord}}, O my soul!


\par }\PsalmChap{104}{\PP \par }{\Q \VerseOne{1}Praise
the {\ND{Lord}}, O my soul!
\par }{\Q O
{\ND{Lord}}
my God,
you are magnificent.
\par }{\Q You are robed
in splendor
and majesty.
\par }{\Q \VS{2}He covers
himself with light
as if it were a garment.
\par }{\Q He stretches
out the skies
like a tent curtain,
\par }{\Q \VS{3}and lays the beams
of the upper rooms
of his palace on the rain clouds.
\par }{\Q He makes
the clouds
his chariot,
\par }{\Q and travels along
on
the wings
of the wind.
\par }{\Q \VS{4}He makes
the winds
his messengers,
\par }{\Q and the flaming
fire
his attendant.
\par }{\Q \VS{5}He established
the earth
on
its foundations;
\par }{\Q it will never be upended.
\par }{\Q \VS{6}The watery deep
covered
it like a garment;
\par }{\Q the waters
reached above
the mountains.
\par }{\Q \VS{7}Your shout made the waters retreat;
\par }{\Q at
the sound
of your thunderous
voice
they hurried off –
\par }{\Q \VS{8}as the mountains
rose up,
\par }{\Q and the valleys
went down
–
\par }{\Q to
the place
you appointed
for them.
\par }{\Q \VS{9}You set
up a boundary
for them that they could not
cross,
\par }{\Q so that they would not
cover
the earth
again.
\par }{\Q \VS{10}He turns
springs
into streams;
\par }{\Q they flow between
the mountains.
\par }{\Q \VS{11}They provide water
for all
the animals
in the field;
\par }{\Q the wild donkeys
quench
their thirst.
\par }{\Q \VS{12}The birds
of the sky
live
beside
them;
\par }{\Q they chirp
among
the bushes.
\par }{\Q \VS{13}He waters
the mountains
from the upper
rooms of his palace;

\par }{\Q the earth
is full
of the fruit
you cause to grow.
\par }{\Q \VS{14}He provides grass
for the cattle,
\par }{\Q and crops
for
people
to cultivate,
\par }{\Q so they can produce
food
from
the ground,
\par }{\Q \VS{15}as well as wine
that makes people
feel
so good,
\par }{\Q and so they can have oil
to make their faces
shine,
\par }{\Q as well as food
that sustains
people’s
lives.
\par }{\Q \VS{16}The trees
of the {\ND{Lord}}
receive all the rain they need,
\par }{\Q the cedars
of Lebanon
which
he planted,
\par }{\Q \VS{17}where
the birds
make nests,
\par }{\Q near the evergreens
in which the herons live.
\par }{\Q \VS{18}The wild goats
live in the high
mountains;
\par }{\Q the rock badgers
find safety
in the cliffs.
\par }{\Q \VS{19}He made
the moon
to mark the months,
\par }{\Q and the sun
sets
according to a regular schedule.
\par }{\Q \VS{20}You make
it dark
and night
comes,
\par }{\Q during which
all
the beasts
of the forest
prowl around.
\par }{\Q \VS{21}The lions
roar
for prey,
\par }{\Q seeking
their food
from God.
\par }{\Q \VS{22}When the sun
rises,
they withdraw
\par }{\Q and sleep
in
their dens.
\par }{\Q \VS{23}Men
then go out
to do their work,
\par }{\Q and labor
away until
evening.
\par }{\Q \VS{24}How
many
living things you have made,
O
{\ND{Lord}}!

\par }{\Q You have exhibited
great skill
in making
all
of them;

\par }{\Q the earth
is full of the living things you have made.
\par }{\Q \VS{25}Over here
is the deep,
wide
sea,
\par }{\Q which
teems
with innumerable
swimming creatures,
\par }{\Q living things
both small
and large.
\par }{\Q \VS{26}The ships
travel
there,
\par }{\Q and over here swims the whale
you made
to play in it.
\par }{\Q \VS{27}All
of your creatures
wait for you
\par }{\Q to provide them
with food
on a regular basis.
\par }{\Q \VS{28}You give
food to them and they
receive
it;
\par }{\Q you open
your hand
and they are filled
with food.
\par }{\Q \VS{29}When you ignore
them, they panic.
\par }{\Q When you take away
their life’s breath,
they die
\par }{\Q and return
to
dust.
\par }{\Q \VS{30}When you send
your life-giving breath,
they are created,
\par }{\Q and you replenish
the surface
of the ground.
\par }{\Q \VS{31}May
the splendor
of the {\ND{Lord}}
endure!

\par }{\Q May the
{\ND{Lord}}
find pleasure
in the living things he has made!
\par }{\Q \VS{32}He looks
down on the earth
and it shakes;
\par }{\Q he touches
the mountains
and they start to smolder.
\par }{\Q \VS{33}I will sing
to the
{\ND{Lord}}
as long as I live;
\par }{\Q I will sing
praise to my God
as long as I exist!
\par }{\Q \VS{34}May my thoughts
be pleasing
to him!
\par }{\Q I
will rejoice
in the
{\ND{Lord}}.
\par }{\Q \VS{35}May sinners
disappear from
the earth,
\par }{\Q and the wicked
vanish!
\par }{\Q Praise
the {\ND{Lord}}, O my soul!
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{105}{\PP \par }{\Q \VerseOne{1}Give thanks
to the
{\ND{Lord}}!
\par }{\Q Call
on his name!
\par }{\Q Make known
his accomplishments
among the nations!
\par }{\Q \VS{2}Sing
to him!
\par }{\Q Make music
to him!
\par }{\Q Tell about all
his miraculous deeds!
\par }{\Q \VS{3}Boast
about his holy
name!
\par }{\Q Let the hearts
of those who seek
the {\ND{Lord}}
rejoice!
\par }{\Q \VS{4}Seek
the {\ND{Lord}}
and the strength
he gives!
\par }{\Q Seek
his presence
continually!
\par }{\Q \VS{5}Recall
the miraculous
deeds he performed,
\par }{\Q his mighty acts
and the judgments
he decreed,
\par }{\Q \VS{6}O children
of Abraham,
God’s servant,
\par }{\Q you descendants
of Jacob,
God’s chosen ones!
\par }{\Q \VS{7}He is
the {\ND{Lord}}
our God;
\par }{\Q he carries out judgment
throughout
the earth.
\par }{\Q \VS{8}He always
remembers
his covenantal
decree,
\par }{\Q the promise
he made to a thousand
generations –
\par }{\Q \VS{9}the promise
he made
to Abraham,
\par }{\Q the promise he made by oath
to Isaac!
\par }{\Q \VS{10}He gave
it to Jacob
as a decree,
\par }{\Q to Israel
as a lasting
promise,
\par }{\Q \VS{11}saying,
“To you I will give
the land
of Canaan
\par }{\Q as the portion
of your inheritance.”
\par }{\Q \VS{12}When
they were few
in number,
\par }{\Q just a
very few,
and resident aliens within it,
\par }{\Q \VS{13}they wandered
from nation
to
nation,
\par }{\Q and from one kingdom
to
another.
\par }{\Q \VS{14}He let
no
one
oppress
them;
\par }{\Q he disciplined
kings
for their sake,
\par }{\Q \VS{15}saying, “Don’t
touch
my chosen ones!
\par }{\Q Don’t
harm
my prophets!”
\par }{\Q \VS{16}He called
down a famine
upon
the earth;
\par }{\Q he cut off
all
the food
supply.
\par }{\Q \VS{17}He sent
a man
ahead of them –
\par }{\Q Joseph was sold as a servant.
\par }{\Q \VS{18}The shackles
hurt
his feet;
\par }{\Q his neck was placed
in an iron
collar,
\par }{\Q \VS{19}until
the time
when his prediction
came
true.
\par }{\Q The
{\ND{Lord}}’s
word
proved him right.
\par }{\Q \VS{20}The king
authorized
his release;
\par }{\Q the ruler
of nations
set him free.
\par }{\Q \VS{21}He put
him
in charge of his palace,
\par }{\Q and made him manager
of all
his property,
\par }{\Q \VS{22}giving him authority to imprison
his officials
\par }{\Q and to teach his advisers.
\par }{\Q \VS{23}Israel
moved
to Egypt;
\par }{\Q Jacob
lived
for a time in the land
of Ham.
\par }{\Q \VS{24}The
{\ND{Lord}} made his people
very
fruitful,
\par }{\Q and made them more numerous
than their enemies.
\par }{\Q \VS{25}He caused
them to hate
his people,
\par }{\Q and to mistreat
his servants.
\par }{\Q \VS{26}He sent
his servant
Moses,
\par }{\Q and Aaron,
whom
he had chosen.
\par }{\Q \VS{27}They executed
his miraculous
signs
among them,

\par }{\Q and his amazing
deeds in the land
of Ham.
\par }{\Q \VS{28}He made
it dark;
\par }{\Q they did not
disobey
his orders.
\par }{\Q \VS{29}He turned
their water
into blood,
\par }{\Q and killed
their fish.
\par }{\Q \VS{30}Their land
was overrun
by frogs,
\par }{\Q which even got into the rooms
of their kings.
\par }{\Q \VS{31}He ordered
flies
to come;
\par }{\Q gnats
invaded their whole
territory.
\par }{\Q \VS{32}He sent
hail
along with the rain;
\par }{\Q there was lightning
in their land.
\par }{\Q \VS{33}He destroyed
their vines
and fig trees,
\par }{\Q and broke
the trees
throughout their territory.
\par }{\Q \VS{34}He ordered
locusts
to come,
\par }{\Q innumerable
grasshoppers.
\par }{\Q \VS{35}They ate
all
the vegetation
in their land,
\par }{\Q and devoured
the crops
of their fields.
\par }{\Q \VS{36}He struck
down all
the firstborn
in their land,
\par }{\Q the firstfruits
of their reproductive power.
\par }{\Q \VS{37}He brought his people out
enriched with silver
and gold;
\par }{\Q none
of his tribes
stumbled.
\par }{\Q \VS{38}Egypt
was happy
when they left,
\par }{\Q for
they were
afraid of them.
\par }{\Q \VS{39}He spread
out a cloud
for a cover,
\par }{\Q and provided a fire
to light
up the night.
\par }{\Q \VS{40}They asked
for food, and he sent
quails;
\par }{\Q he satisfied
them with food
from the sky.
\par }{\Q \VS{41}He opened
up a rock
and water
flowed
out;
\par }{\Q a river
ran through dry regions.
\par }{\Q \VS{42}Yes,
he remembered
the sacred
promise
\par }{\Q he made to Abraham
his servant.
\par }{\Q \VS{43}When he led
his people
out,
they rejoiced;
\par }{\Q his chosen ones
shouted
with joy.
\par }{\Q \VS{44}He handed
the territory
of nations
over to them,
\par }{\Q and they
took possession
of what
other peoples
had produced,
\par }{\Q \VS{45}so that they might keep
his commands
\par }{\Q and obey his laws.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{106}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Give thanks
to the
{\ND{Lord}}, for
he is good,
\par }{\Q and his loyal love
endures!
\par }{\Q \VS{2}Who can
adequately
recount the
{\ND{Lord}}’s
mighty acts,
\par }{\Q or relate
all
his praiseworthy deeds?
\par }{\Q \VS{3}How blessed
are those who promote
justice,
\par }{\Q and do
what is right
all
the time!
\par }{\Q \VS{4}Remember
me, O
{\ND{Lord}}, when you show favor
to your people!
\par }{\Q Pay attention
to me, when you deliver,
\par }{\Q \VS{5}so I may see
the prosperity
of your chosen ones,
\par }{\Q rejoice
along with your nation,
\par }{\Q and boast
along with
the people who belong to you.
\par }{\Q \VS{6}We have sinned
like our ancestors;
\par }{\Q we have done wrong,
we have done evil.
\par }{\Q \VS{7}Our ancestors
in Egypt
failed
to appreciate
your miraculous
deeds,
\par }{\Q they failed
to remember
your many
acts of loyal love,
\par }{\Q and they rebelled
at the sea,
by the Red
Sea.
\par }{\Q \VS{8}Yet he delivered
them for the sake
of his reputation,
\par }{\Q that
he might reveal
his power.
\par }{\Q \VS{9}He shouted
at the Red
Sea
and it dried
up;
\par }{\Q he led
them through the deep
water as if it were a desert.
\par }{\Q \VS{10}He delivered
them from the power
of the one who hated
them,
\par }{\Q and rescued
them from the power
of the enemy.
\par }{\Q \VS{11}The water
covered
their enemies;
\par }{\Q not
even one
of them
survived.
\par }{\Q \VS{12}They believed
his promises;
\par }{\Q they sang
praises to him.
\par }{\Q \VS{13}They quickly
forgot
what he had done;
\par }{\Q they did not
wait
for his instructions.
\par }{\Q \VS{14}In the wilderness
they had an insatiable
craving
for meat;

\par }{\Q they challenged
God
in the desert.
\par }{\Q \VS{15}He granted
their
request,
\par }{\Q then struck
them with a disease.
\par }{\Q \VS{16}In the camp
they resented
Moses,
\par }{\Q and Aaron,
the
{\ND{Lord}}’s
holy priest.
\par }{\Q \VS{17}The earth
opened
up
and swallowed
Dathan;
\par }{\Q it engulfed
the group
led by Abiram.
\par }{\Q \VS{18}Fire
burned
their group;
\par }{\Q the flames
scorched
the wicked.
\par }{\Q \VS{19}They made
an image of a calf
at Horeb,
\par }{\Q and worshiped
a metal idol.
\par }{\Q \VS{20}They traded
their majestic
God

\par }{\Q for the image
of an ox
that eats
grass.
\par }{\Q \VS{21}They rejected
the God
who delivered
them,
\par }{\Q the one who performed
great
deeds in Egypt,
\par }{\Q \VS{22}amazing
feats in the land
of Ham,
\par }{\Q mighty acts
by the Red
Sea.
\par }{\Q \VS{23}He threatened
to destroy
them,
\par }{\Q but
Moses,
his chosen one,
interceded
with him

\par }{\Q and turned back
his destructive
anger.
\par }{\Q \VS{24}They rejected
the fruitful
land;
\par }{\Q they did not
believe
his promise.
\par }{\Q \VS{25}They grumbled
in their tents;
\par }{\Q they did not
obey
the {\ND{Lord}}.
\par }{\Q \VS{26}So he made a solemn
vow

\par }{\Q that he would make them
die
in the desert,
\par }{\Q \VS{27}make their descendants
die
among the nations,
\par }{\Q and scatter
them among foreign lands.
\par }{\Q \VS{28}They worshiped
Baal of Peor,
\par }{\Q and ate
sacrifices offered
to the dead.
\par }{\Q \VS{29}They made the
{\ND{Lord}} angry
by their actions,
\par }{\Q and a plague
broke out among them.
\par }{\Q \VS{30}Phinehas
took a stand
and intervened,
\par }{\Q and the plague
subsided.
\par }{\Q \VS{31}This brought
him a reward,
\par }{\Q an eternal
gift.
\par }{\Q \VS{32}They made him angry
by
the waters
of Meribah,
\par }{\Q and Moses
suffered because of them,
\par }{\Q \VS{33}for
they aroused
his temper,
\par }{\Q and he spoke
rashly.
\par }{\Q \VS{34}They did not
destroy
the nations,
\par }{\Q as
the
{\ND{Lord}}
had commanded
them to do.
\par }{\Q \VS{35}They mixed
in with the nations
\par }{\Q and learned
their ways.
\par }{\Q \VS{36}They worshiped
their idols,
\par }{\Q which became
a snare
to them.
\par }{\Q \VS{37}They sacrificed
their sons
and daughters
to demons.
\par }{\Q \VS{38}They shed
innocent
blood
–
\par }{\Q the blood
of their sons
and daughters,
\par }{\Q whom
they sacrificed
to the idols
of Canaan.
\par }{\Q The land
was polluted
by bloodshed.
\par }{\Q \VS{39}They were defiled
by their deeds,
\par }{\Q and unfaithful
in their actions.
\par }{\Q \VS{40}So the
{\ND{Lord}}
was angry
with his people
\par }{\Q and despised
the people who belong to him.
\par }{\Q \VS{41}He handed
them over
to the nations,
\par }{\Q and those who hated
them ruled over them.
\par }{\Q \VS{42}Their enemies
oppressed
them;
\par }{\Q they were subject
to their authority.
\par }{\Q \VS{43}Many
times
he delivered
them,
\par }{\Q but they
had
a rebellious
attitude,

\par }{\Q and degraded
themselves by their sin.
\par }{\Q \VS{44}Yet he took notice
of their distress,
\par }{\Q when he heard
their cry for help.
\par }{\Q \VS{45}He remembered
his covenant
with them,
\par }{\Q and relented
because of his great
loyal love.
\par }{\Q \VS{46}He caused
all
their conquerors
\par }{\Q to have pity on them.
\par }{\Q \VS{47}Deliver
us, O
{\ND{Lord}}, our God!
\par }{\Q Gather
us from
among the nations!
\par }{\Q Then we will give thanks
to your holy
name,
\par }{\Q and boast
about your praiseworthy deeds.
\par }{\Q \VS{48}The
{\ND{Lord}}
God
of Israel
deserves
praise,

\par }{\Q in the future
and forevermore.
\par }{\Q Let all
the people
say, “We agree! Praise
the
{\ND{Lord}}!”


\par }\PsalmChap{107}{\PP \par }{\Q \VerseOne{1}Give thanks
to the
{\ND{Lord}}, for
he is good,
\par }{\Q and his loyal love
endures!
\par }{\Q \VS{2}Let those delivered
by the
{\ND{Lord}}
speak out,

\par }{\Q those whom
he delivered
from the power
of the enemy,
\par }{\Q \VS{3}and gathered
from
foreign lands,
\par }{\Q from east
and west,
\par }{\Q from north
and south.
\par }{\Q \VS{4}They wandered
through the wilderness
on a desert
road;
\par }{\Q they found
no
city
in which to live.
\par }{\Q \VS{5}They were hungry
and thirsty;
\par }{\Q they fainted
from exhaustion.
\par }{\Q \VS{6}They cried out
to
the
{\ND{Lord}}
in their distress;
\par }{\Q he delivered
them
from their troubles.
\par }{\Q \VS{7}He led them on a level
road,
\par }{\Q that they might find a city
in which to live.
\par }{\Q \VS{8}Let them give thanks
to the
{\ND{Lord}}
for his loyal love,
\par }{\Q and for the amazing
things he has done for people!
\par }{\Q \VS{9}For
he has satisfied
those who
thirst,
\par }{\Q and those who
hunger
he has filled
with food.
\par }{\Q \VS{10}They sat
in utter darkness,
\par }{\Q bound
in painful
iron chains,
\par }{\Q \VS{11}because
they had rebelled
against God’s
commands,

\par }{\Q and rejected
the instructions
of the sovereign king.
\par }{\Q \VS{12}So he used suffering
to humble
them;

\par }{\Q they stumbled
and no
one helped them up.
\par }{\Q \VS{13}They cried
out to
the
{\ND{Lord}}
in their distress;
\par }{\Q he delivered
them
from their troubles.
\par }{\Q \VS{14}He brought
them out
of the utter darkness,
\par }{\Q and tore off
their shackles.
\par }{\Q \VS{15}Let them give thanks
to the
{\ND{Lord}}
for his loyal love,
\par }{\Q and for the amazing
things he has done for people!
\par }{\Q \VS{16}For
he shattered
the bronze
gates,
\par }{\Q and hacked
through the iron
bars.
\par }{\Q \VS{17}They acted like fools
in their rebellious
ways,
\par }{\Q and suffered
because of their sins.
\par }{\Q \VS{18}They lost
their appetite
for all
food,
\par }{\Q and they drew near
the gates
of death.
\par }{\Q \VS{19}They cried
out to
the
{\ND{Lord}}
in their distress;
\par }{\Q he delivered
them
from their troubles.
\par }{\Q \VS{20}He sent
them an assuring word
and healed
them;
\par }{\Q he rescued
them from the pits where they were trapped.
\par }{\Q \VS{21}Let them give thanks
to the
{\ND{Lord}}
for his loyal love,
\par }{\Q and for the amazing
things he has done for people!
\par }{\Q \VS{22}Let them present thank
offerings,
\par }{\Q and loudly
proclaim
what he has done!
\VS{23}\par }{\Q Some traveled
on the sea
in ships,
\par }{\Q and carried
cargo
over the vast
waters.
\par }{\Q \VS{24}They
witnessed
the acts
of the {\ND{Lord}},
\par }{\Q his amazing
feats on the deep water.
\par }{\Q \VS{25}He gave the order
for a windstorm,
\par }{\Q and it stirred up
the waves of the sea.
\par }{\Q \VS{26}They reached up
to the sky,
\par }{\Q then dropped into
the depths.
\par }{\Q The sailors’
strength left them because the danger was so great.
\par }{\Q \VS{27}They swayed
and staggered
like a drunk,
\par }{\Q and all
their skill
proved ineffective.
\par }{\Q \VS{28}They cried out
to
the
{\ND{Lord}}
in their distress;
\par }{\Q he delivered them
from their troubles.
\par }{\Q \VS{29}He calmed
the storm,
\par }{\Q and the waves
grew silent.
\par }{\Q \VS{30}The sailors rejoiced
because
the waves grew quiet,
\par }{\Q and he led
them to
the harbor
they desired.
\par }{\Q \VS{31}Let them give thanks
to the
{\ND{Lord}}
for his loyal love,
\par }{\Q and for the amazing
things he has done for people!
\par }{\Q \VS{32}Let them exalt
him in the assembly
of the people!
\par }{\Q Let them praise
him in
the place where
the leaders
preside!
\par }{\Q \VS{33}He turned
streams
into a desert,
\par }{\Q springs
of water
into arid land,
\par }{\Q \VS{34}and a fruitful
land
into a barren
place,

\par }{\Q because of the sin of its inhabitants.
\par }{\Q \VS{35}As for his people, he turned
a desert
into a pool
of water,
\par }{\Q and a dry
land
into springs
of water.
\par }{\Q \VS{36}He allowed the hungry
to settle
there,
\par }{\Q and they established
a city
in which to live.
\par }{\Q \VS{37}They cultivated
fields,
\par }{\Q and planted
vineyards,
\par }{\Q which yielded
a harvest
of fruit.
\par }{\Q \VS{38}He blessed
them so
that they became very
numerous.
\par }{\Q He would not
allow their cattle
to decrease in number.
\par }{\Q \VS{39}As for their
enemies, they decreased
in number and were beaten down,
\par }{\Q because of painful
distress and suffering.
\par }{\Q \VS{40}He would pour
contempt
upon
princes,
\par }{\Q and he made them wander
in a wasteland
with no
road.
\par }{\Q \VS{41}Yet he protected
the needy
from oppression,
\par }{\Q and cared
for his families
like a flock of sheep.
\par }{\Q \VS{42}When the godly
see
this, they rejoice,
\par }{\Q and every
sinner
shuts
his mouth.
\par }{\Q \VS{43}Whoever
is wise,
let him take note
of these
things!
\par }{\Q Let them consider
the
{\ND{Lord}}’s
acts of loyal love!


\par }\PsalmChap{108}{\D 
A song,
a psalm
of David.
\par }{\Q \VerseOne{1}I am determined,
O God!
\par }{\Q I will sing
and praise
you with my whole heart.
\par }{\Q \VS{2}Awake,
O stringed instrument
and harp!
\par }{\Q I will wake
up at dawn!
\par }{\Q \VS{3}I will give you thanks
before the nations,
O
{\ND{Lord}}!
\par }{\Q I will sing praises
to you before foreigners!
\par }{\Q \VS{4}For
your loyal love
extends
beyond the sky,
\par }{\Q and your faithfulness
reaches
the clouds.
\par }{\Q \VS{5}Rise
up above
the sky,
O God!
\par }{\Q May your splendor
cover the whole
earth!
\par }{\Q \VS{6}Deliver
by your power
and answer
me,
\par }{\Q so that
the ones you love
may be safe.
\par }{\Q \VS{7}God
has spoken
in his sanctuary:

\par }{\Q “I will triumph! I will parcel
out Shechem,
\par }{\Q the valley
of Succoth
I will measure off.
\par }{\Q \VS{8}Gilead
belongs to me,
\par }{\Q as does Manasseh!

\par }{\Q Ephraim
is my helmet,
\par }{\Q Judah
my royal scepter.
\par }{\Q \VS{9}Moab
is my wash basin.
\par }{\Q I will make
Edom
serve
me.

\par }{\Q I will shout in triumph
over
Philistia.”
\par }{\Q \VS{10}Who
will lead me into
the fortified
city?
\par }{\Q Who
will bring me
to Edom?
\par }{\Q \VS{11}Have you not
rejected
us, O God?
\par }{\Q O God,
you do not
go
into battle with our armies.
\par }{\Q \VS{12}Give
us help
against the enemy,
\par }{\Q for any help
men
might offer is futile.
\par }{\Q \VS{13}By God’s
power
we will conquer;
\par }{\Q he
will trample down
our enemies.


\par }\PsalmChap{109}{\D 
For the music director,
a psalm
of David.
\par }{\Q \VerseOne{1}O God
whom I praise,
do not
ignore me!
\par }{\Q \VS{2}For
they say
cruel
and deceptive
things to me;
\par }{\Q they lie to me.
\par }{\Q \VS{3}They surround
me and say hateful
things;
\par }{\Q they attack
me for no reason.
\par }{\Q \VS{4}They repay
my love
with accusations,
\par }{\Q but I
continue to pray.
\par }{\Q \VS{5}They repay
me evil
for
good,
\par }{\Q and hate
for
love.
\VS{6}\par }{\Q Appoint
an evil
man to testify against him!

\par }{\Q May an accuser
stand
at his right side!
\par }{\Q \VS{7}When he is judged,
he will be found guilty!

\par }{\Q Then his prayer
will be regarded as sinful.
\par }{\Q \VS{8}May
his days
be few!

\par }{\Q May another
take
his job!
\par }{\Q \VS{9}May
his children
be fatherless,
\par }{\Q and his wife
a widow!
\par }{\Q \VS{10}May
his children
roam
around begging,
\par }{\Q asking
for handouts
as they leave their ruined home!
\par }{\Q \VS{11}May
the creditor
seize all
he owns!
\par }{\Q May
strangers
loot
his property!
\par }{\Q \VS{12}May no
one show him kindness!

\par }{\Q May no
one have compassion
on his fatherless children!
\par }{\Q \VS{13}May
his descendants be
cut off!

\par }{\Q May the memory
of them be wiped out
by the time the next generation
arrives!
\par }{\Q \VS{14}May his ancestors’
sins
be remembered
by the
{\ND{Lord}}!
\par }{\Q May his mother’s
sin
not
be forgotten!
\par }{\Q \VS{15}May the
{\ND{Lord}}
be
constantly
aware
of them,

\par }{\Q and cut off
the memory
of his children from the earth!
\par }{\Q \VS{16}For
he never
bothered
to show
kindness;
\par }{\Q he harassed
the oppressed
and needy,
\par }{\Q and killed
the disheartened.
\par }{\Q \VS{17}He loved
to curse others, so those
curses
have come
upon him.

\par }{\Q He had
no
desire
to bless anyone, so he has experienced
no blessings.
\par }{\Q \VS{18}He made
cursing
a way of life,
\par }{\Q so curses poured
into his stomach like water
\par }{\Q and seeped
into his bones
like oil.
\par }{\Q \VS{19}May
a curse attach itself to him, like a garment
one puts
on,

\par }{\Q or a belt
one wears
continually!
\par }{\Q \VS{20}May the
{\ND{Lord}}
repay
my accusers
in this
way,
\par }{\Q those who say
evil
things about me!
\par }{\Q \VS{21}O sovereign
{\SC{
Lord}},

\par }{\Q intervene
on my behalf
for the sake
of your reputation!

\par }{\Q Because
your loyal love
is good,
deliver me!
\par }{\Q \VS{22}For
I am
oppressed
and needy,
\par }{\Q and my heart
beats
violently within me.
\par }{\Q \VS{23}I am fading
away
like a shadow
at the end of the day;

\par }{\Q I am shaken
off like a locust.
\par }{\Q \VS{24}I am so starved
my knees
shake;
\par }{\Q I
have turned into skin and bones.
\par }{\Q \VS{25}I am
disdained
by them.
\par }{\Q When they
see
me, they shake
their heads.
\par }{\Q \VS{26}Help
me, O
{\ND{Lord}}
my God!
\par }{\Q Because you are faithful
to me, deliver me!
\par }{\Q \VS{27}Then they will realize
this
is your work,

\par }{\Q and that you,

{\ND{Lord}}, have accomplished it.
\par }{\Q \VS{28}They curse,
but you
will bless.
\par }{\Q When they attack,
they will be humiliated,
\par }{\Q but your servant
will rejoice.
\par }{\Q \VS{29}My accusers
will be covered
with shame,
\par }{\Q and draped
in humiliation
as if it were
a robe.
\par }{\Q \VS{30}I will thank
the {\ND{Lord}}
profusely,
\par }{\Q in the middle
of a crowd
I will praise him,
\par }{\Q \VS{31}because
he stands
at the right hand
of the needy,
\par }{\Q to deliver
him from those who threaten
his life.


\par }\PsalmChap{110}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}Here is the
{\ND{Lord}}’s
proclamation
to my lord:

\par }{\Q “Sit down
at my right hand
until
I make
your enemies
your footstool!”
\par }{\Q \VS{2}The
{\ND{Lord}}
extends
your dominion
from Zion.
\par }{\Q Rule
in the midst
of your enemies!
\par }{\Q \VS{3}Your people
willingly follow
you when
you go into battle.
\par }{\Q On the holy
hills
at sunrise
the dew
of your youth belongs to you.
\par }{\Q \VS{4}The
{\ND{Lord}}
makes this promise on oath
and will not
revoke
it:

\par }{\Q “You
are an eternal
priest
after the pattern
of Melchizedek.”
\par }{\Q \VS{5}O sovereign
{\SC{
Lord}}, at your right hand
\par }{\Q he strikes down
kings
in the day
he unleashes his anger.
\par }{\Q \VS{6}He executes
judgment against the nations;
\par }{\Q he fills
the valleys with corpses;
\par }{\Q he shatters
their heads
over
the vast
battlefield.
\par }{\Q \VS{7}From the stream
along the road
he drinks;
\par }{\Q then
he lifts
up his head.


\par }\PsalmChap{111}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q I will give thanks
to the
{\ND{Lord}}
with my whole
heart,
\par }{\Q in the assembly
of the godly
and the congregation.
\par }{\Q \VS{2}The
{\ND{Lord}}’s
deeds are
great,
\par }{\Q eagerly
awaited
by all
who desire them.
\par }{\Q \VS{3}His
work
is majestic
and glorious,
\par }{\Q and his faithfulness
endures
forever.
\par }{\Q \VS{4}He does
amazing
things that will be remembered;
\par }{\Q the {\ND{Lord}}
is merciful
and compassionate.
\par }{\Q \VS{5}He gives
food
to his faithful followers;
\par }{\Q he always
remembers
his covenant.
\par }{\Q \VS{6}He announced
that he would do mighty
deeds
for his people,
\par }{\Q giving
them a land that belonged
to other nations.
\par }{\Q \VS{7}His acts
are characterized by faithfulness
and justice;
\par }{\Q all
his precepts
are reliable.
\par }{\Q \VS{8}They are
forever
firm,
\par }{\Q and should be
faithfully
and properly carried out.
\par }{\Q \VS{9}He delivered
his people;
\par }{\Q he ordained
that his covenant
be observed
forever.
\par }{\Q His name
is holy
and awesome.
\par }{\Q \VS{10}To obey
the {\ND{Lord}}
is the fundamental principle
for wise living;
\par }{\Q all
who carry out
his precepts acquire good
moral insight.
\par }{\Q He will receive praise
forever.


\par }\PsalmChap{112}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q How blessed
is the one
who obeys
the {\ND{Lord}},
\par }{\Q who takes great
delight
in keeping
his commands.
\par }{\Q \VS{2}His descendants
will be
powerful
on the earth;
\par }{\Q the godly
will be blessed.
\par }{\Q \VS{3}His house
contains wealth
and riches;
\par }{\Q his integrity
endures.
\par }{\Q \VS{4}In the darkness
a light
shines
for the godly,
\par }{\Q for each one who is merciful,
compassionate,
and just.
\par }{\Q \VS{5}It
goes well
for the one
who generously
lends
money,
\par }{\Q and conducts
his business honestly.
\par }{\Q \VS{6}For
he will never
be upended;
\par }{\Q others
will always
remember
one who is just.
\par }{\Q \VS{7}He does not
fear
bad
news.
\par }{\Q He is confident;
he trusts
in the
{\ND{Lord}}.
\par }{\Q \VS{8}His resolve
is firm; he will not
succumb
to fear
\par }{\Q before
he looks in triumph
on his enemies.
\par }{\Q \VS{9}He generously
gives
to the needy;
\par }{\Q his integrity
endures.
\par }{\Q He will be vindicated
and honored.
\par }{\Q \VS{10}When the wicked
see
this, they will worry;
\par }{\Q they will grind their teeth
in frustration
and melt away;
\par }{\Q the desire
of the wicked
will perish.


\par }\PsalmChap{113}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Praise,
you servants
of the {\ND{Lord}},
\par }{\Q praise
the name
of the {\ND{Lord}}!
\par }{\Q \VS{2}May
the
{\ND{Lord}}’s
name
be praised
\par }{\Q now
and forevermore!
\par }{\Q \VS{3}From east
to
west
\par }{\Q the
{\ND{Lord}}’s
name
is deserving of praise.
\par }{\Q \VS{4}The
{\ND{Lord}}
is exalted
over
all
the nations;
\par }{\Q his splendor
reaches beyond the sky.
\par }{\Q \VS{5}Who can compare
to the
{\ND{Lord}}
our God,
\par }{\Q who sits
on a high throne?
\par }{\Q \VS{6}He bends
down to look
\par }{\Q at the sky
and the earth.
\par }{\Q \VS{7}He raises
the poor
from the dirt,
\par }{\Q and lifts up
the needy
from the garbage pile,
\par }{\Q \VS{8}that he might seat
him with
princes,
\par }{\Q with
the princes
of his people.
\par }{\Q \VS{9}He makes
the barren
woman of the family
\par }{\Q a happy
mother
of children.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{114}{\PP \par }{\Q \VerseOne{1}When Israel
left
Egypt,
\par }{\Q when the family
of Jacob
left a foreign nation
behind,
\par }{\Q \VS{2}Judah
became his sanctuary,
\par }{\Q Israel
his kingdom.
\par }{\Q \VS{3}The sea
looked
and fled;
\par }{\Q the Jordan River
turned
back.
\par }{\Q \VS{4}The mountains
skipped
like rams,
\par }{\Q the hills
like lambs.
\par }{\Q \VS{5}Why
do you flee,
O sea?
\par }{\Q Why do you turn
back,
O Jordan River?
\par }{\Q \VS{6}Why do you skip
like rams,
O mountains,
\par }{\Q like lambs,
O hills?
\par }{\Q \VS{7}Tremble,
O earth,
before
the Lord
–
\par }{\Q before
the God
of Jacob,
\par }{\Q \VS{8}who turned
a rock
into a pool
of water,
\par }{\Q a hard rock
into springs
of water!


\par }\PsalmChap{115}{\PP \par }{\Q \VerseOne{1}Not
to us, O
{\ND{Lord}}, not
to us!
\par }{\Q But
to your name
bring honor,
\par }{\Q for the sake of your loyal love
and faithfulness.
\par }{\Q \VS{2}Why
should the nations
say,
\par }{\Q “Where
is their God?”
\par }{\Q \VS{3}Our God
is in heaven!
\par }{\Q He does whatever
he pleases!
\par }{\Q \VS{4}Their idols
are made of silver
and gold
–
\par }{\Q they are man-made.
\par }{\Q \VS{5}They have mouths,
but cannot
speak,
\par }{\Q eyes,
but cannot
see,
\par }{\Q \VS{6}ears,
but cannot
hear,
\par }{\Q noses,
but cannot
smell,
\par }{\Q \VS{7}hands,
but cannot
touch,
\par }{\Q feet,
but cannot
walk.
\par }{\Q They cannot
even clear
their throats.
\par }{\Q \VS{8}Those who make
them will end up like
them,
\par }{\Q as will everyone
who
trusts in them.
\par }{\Q \VS{9}O Israel,
trust
in the
{\ND{Lord}}!
\par }{\Q He is
their deliverer
and protector.
\par }{\Q \VS{10}O family
of Aaron,
trust
in the
{\ND{Lord}}!
\par }{\Q He is
their deliverer
and protector.
\par }{\Q \VS{11}You loyal followers
of the {\ND{Lord}}, trust
in the
{\ND{Lord}}!
\par }{\Q He is
their deliverer
and protector.
\par }{\Q \VS{12}The
{\ND{Lord}}
takes notice
of us, he will bless –
\par }{\Q he will bless
the
family
of Israel,
\par }{\Q he will bless
the
family
of Aaron.
\par }{\Q \VS{13}He will bless
his loyal followers,
\par }{\Q both young
and old.
\par }{\Q \VS{14}May he increase
your numbers,
\par }{\Q yours and your children’s!
\par }{\Q \VS{15}May you
be blessed
by the
{\ND{Lord}},
\par }{\Q the creator
of heaven
and earth!
\par }{\Q \VS{16}The heavens
belong to the
{\ND{Lord}},

\par }{\Q but the earth
he has given
to mankind.
\par }{\Q \VS{17}The dead
do not
praise
the
{\ND{Lord}},
\par }{\Q nor
do any
of those who descend
into the silence of death.
\par }{\Q \VS{18}But we
will praise
the
{\ND{Lord}}
\par }{\Q now
and forevermore.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{116}{\PP \par }{\Q \VerseOne{1}I love
the {\ND{Lord}}
\par }{\Q because
he heard
my plea for mercy,
\par }{\Q \VS{2}and listened
to me.

\par }{\Q As long
as I live, I will call to him when I need help.
\par }{\Q \VS{3}The ropes
of death
tightened around
me,

\par }{\Q the snares
of Sheol
confronted
me.
\par }{\Q I was confronted
with trouble
and sorrow.
\par }{\Q \VS{4}I called
on the name
of the {\ND{Lord}},
\par }{\Q “Please

{\ND{Lord}}, rescue
my life!”
\par }{\Q \VS{5}The
{\ND{Lord}}
is merciful
and fair;
\par }{\Q our God
is compassionate.
\par }{\Q \VS{6}The
{\ND{Lord}}
protects
the untrained;
\par }{\Q I was in serious
trouble and he delivered me.
\par }{\Q \VS{7}Rest
once
more, my soul,
\par }{\Q for
the {\ND{Lord}}
has vindicated you.
\par }{\Q \VS{8}Yes,
{\SC{
Lord}}, you rescued
my life
from death,
\par }{\Q and kept my
feet
from
stumbling.
\par }{\Q \VS{9}I will serve
the {\ND{Lord}}
\par }{\Q in the land
of the living.
\par }{\Q \VS{10}I had faith
when
I said,
\par }{\Q “I
am severely
oppressed.”
\par }{\Q \VS{11}I
rashly
declared,
\par }{\Q “All
men
are liars.”
\par }{\Q \VS{12}How
can I repay
the {\ND{Lord}}
\par }{\Q for all
his acts of kindness
to me?
\par }{\Q \VS{13}I will celebrate
my deliverance,
\par }{\Q and call
on the name
of the {\ND{Lord}}.
\par }{\Q \VS{14}I will fulfill
my vows
to the
{\ND{Lord}}
\par }{\Q before
all
his people.
\par }{\Q \VS{15}The
{\ND{Lord}}
values
\par }{\Q the lives of his faithful followers.
\par }{\Q \VS{16}Yes,

{\ND{Lord}}! I am
indeed
your servant;
\par }{\Q I am
your lowest
slave.
\par }{\Q You saved me from death.
\par }{\Q \VS{17}I will present a thank
offering to you,
\par }{\Q and call
on the name
of the {\ND{Lord}}.
\par }{\Q \VS{18}I will fulfill
my vows
to the
{\ND{Lord}}
\par }{\Q before
all
his people,
\par }{\Q \VS{19}in the courts
of the
{\ND{Lord}}’s
temple,
\par }{\Q in your midst,
O Jerusalem.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{117}{\PP \par }{\Q \VerseOne{1}Praise
the {\ND{Lord}}, all
you nations!
\par }{\Q Applaud
him, all
you foreigners!
\par }{\Q \VS{2}For
his loyal love towers
over
us,
\par }{\Q and the
{\ND{Lord}}’s
faithfulness
endures.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{118}{\PP \par }{\Q \VerseOne{1}Give thanks
to the
{\ND{Lord}}, for
he is good
\par }{\Q and his loyal love
endures!
\par }{\Q \VS{2}Let
Israel
say,
\par }{\Q “Yes,
his loyal love
endures!”
\par }{\Q \VS{3}Let
the family
of Aaron
say,
\par }{\Q “Yes,
his loyal love
endures!”
\par }{\Q \VS{4}Let
the loyal followers
of the {\ND{Lord}}
say,
\par }{\Q “Yes,
his loyal love
endures!”
\par }{\Q \VS{5}In
my distress
I cried out
to the
{\ND{Lord}}.
\par }{\Q The
{\ND{Lord}}
answered
me and put me in a wide open place.
\par }{\Q \VS{6}The
{\ND{Lord}}
is on my side, I am not
afraid!
\par }{\Q What
can people
do to me?
\par }{\Q \VS{7}The
{\ND{Lord}}
is on my side as my helper.
\par }{\Q I
look
in triumph
on those who hate me.
\par }{\Q \VS{8}It is better
to take shelter
in the
{\ND{Lord}}
\par }{\Q than to trust
in people.
\par }{\Q \VS{9}It is better
to take shelter
in the
{\ND{Lord}}
\par }{\Q than to trust
in princes.
\par }{\Q \VS{10}All
the nations
surrounded
me.

\par }{\Q Indeed, in the name
of the {\ND{Lord}}
I pushed them away.
\par }{\Q \VS{11}They surrounded
me, yes,
they surrounded
me.
\par }{\Q Indeed, in the name
of the {\ND{Lord}}
I pushed them away.
\par }{\Q \VS{12}They surrounded
me like bees.
\par }{\Q But they disappeared
as quickly as a fire
among thorns.
\par }{\Q Indeed,
in the name
of the {\ND{Lord}}
I pushed them away.
\par }{\Q \VS{13}“You aggressively
attacked
me and tried to knock me down,
\par }{\Q but the
{\ND{Lord}}
helped me.
\par }{\Q \VS{14}The
{\ND{Lord}}
gives me strength
and protects
me;

\par }{\Q he has become
my deliverer.”
\par }{\Q \VS{15}They celebrate
deliverance
in the tents
of the godly.
\par }{\Q The
{\ND{Lord}}’s
right hand
conquers,
\par }{\Q \VS{16}the
{\ND{Lord}}’s
right hand
gives victory,
\par }{\Q the
{\ND{Lord}}’s
right hand
conquers.
\par }{\Q \VS{17}I will not
die,
but
live,
\par }{\Q and I will proclaim
what the
{\ND{Lord}}
has done.
\par }{\Q \VS{18}The
{\ND{Lord}}
severely
punished
me,
\par }{\Q but he did not
hand
me over to death.
\par }{\Q \VS{19}Open
for me the gates
of the just
king’s temple!

\par }{\Q I will enter
through them and give thanks
to the
{\ND{Lord}}.
\par }{\Q \VS{20}This
is the
{\ND{Lord}}’s
gate
–
\par }{\Q the godly
enter through it.
\par }{\Q \VS{21}I will give you thanks,
for
you answered
me,
\par }{\Q and have become
my deliverer.
\par }{\Q \VS{22}The stone
which
the builders
discarded
\par }{\Q has become
the cornerstone.
\par }{\Q \VS{23}This
is the
{\ND{Lord}}’s
work.
\par }{\Q We consider
it amazing!
\par }{\Q \VS{24}This
is the day
the {\ND{Lord}}
has brought
about.

\par }{\Q We will be happy
and rejoice in it.
\par }{\Q \VS{25}Please

{\ND{Lord}}, deliver!
\par }{\Q Please

{\ND{Lord}}, grant us
success!
\par }{\Q \VS{26}May
the one
who comes
in the name
of the {\ND{Lord}}
be blessed!
\par }{\Q We will pronounce blessings
on you in the
{\ND{Lord}}’s
temple.
\par }{\Q \VS{27}The
{\ND{Lord}}
is God
and he has delivered us.

\par }{\Q Tie
the offering
with ropes
\par }{\Q to
the horns
of the altar!
\par }{\Q \VS{28}You are my God
and I will give you thanks!
\par }{\Q You are my God
and I will praise you!
\par }{\Q \VS{29}Give thanks
to the
{\ND{Lord}}, for
he is good
\par }{\Q and his loyal love
endures!


\par }\PsalmChap{119}{\SH 
{\IT{א (Alef}})
\par }{\Q \VerseOne{1}How
blessed are those whose
actions are
blameless,

\par }{\Q who
obey the
law of the
{\ND{Lord}}.
\par }{\Q \VS{2}How blessed
are those who observe
his rules,
\par }{\Q and seek
him with all
their heart,
\par }{\Q \VS{3}who, moreover,
do no
wrong,
\par }{\Q but follow
in his footsteps.
\par }{\Q \VS{4}You
demand
that your precepts
\par }{\Q be carefully kept.
\par }{\Q \VS{5}If only
I were predisposed
\par }{\Q to keep
your statutes!
\par }{\Q \VS{6}Then
I would not
be ashamed,
\par }{\Q if I were focused
on all
your commands.
\par }{\Q \VS{7}I will give you
sincere
thanks,
\par }{\Q when I learn
your just regulations.
\par }{\Q \VS{8}I will keep
your statutes.
\par }{\Q Do not
completely
abandon me!
\par }{\SH ב ({\IT{Bet}})
\par }{\Q \VS{9}How can a young person maintain a pure life?
\par }{\Q By guarding it according to your instructions!
\par }{\Q \VS{10}With all
my heart
I seek
you.
\par }{\Q Do not
allow me to stray
from your commands!
\par }{\Q \VS{11}In my heart
I store
up your words,
\par }{\Q so I might
not
sin against you.
\par }{\Q \VS{12}You deserve praise,
O

{\ND{Lord}}!
\par }{\Q Teach
me your statutes!
\par }{\Q \VS{13}With my lips
I proclaim
\par }{\Q all
the regulations
you have revealed.
\par }{\Q \VS{14}I rejoice
in the lifestyle
prescribed by
your rules
\par }{\Q as if they were riches of all
kinds.
\par }{\Q \VS{15}I will meditate
on your precepts
\par }{\Q and focus
on your behavior.
\par }{\Q \VS{16}I find delight
in your statutes;
\par }{\Q I do not
forget
your instructions.
\par }{\SH ג ({\IT{Gimel}})
\par }{\Q \VS{17}Be kind to your servant!
\par }{\Q Then I will live and keep your instructions.
\par }{\Q \VS{18}Open
my eyes
so I can truly
see

\par }{\Q the marvelous
things in your law!
\par }{\Q \VS{19}I am
like a foreigner
in this land.
\par }{\Q Do not
hide
your commands
from me!
\par }{\Q \VS{20}I
desperately long
to know
\par }{\Q your regulations
at all
times.
\par }{\Q \VS{21}You reprimand
arrogant people.
\par }{\Q Those who stray
from your commands
are doomed.
\par }{\Q \VS{22}Spare
me shame
and humiliation,
\par }{\Q for
I observe
your rules.
\par }{\Q \VS{23}Though
rulers
plot and slander
me,

\par }{\Q your servant
meditates
on your statutes.
\par }{\Q \VS{24}Yes,
I find delight
in your rules;
\par }{\Q they give me guidance.
\par }{\SH ד ({\IT{Dalet}})
\par }{\Q \VS{25}I collapse in the dirt.
\par }{\Q Revive me with your word!
\par }{\Q \VS{26}I told
you about my ways
and you answered
me.
\par }{\Q Teach
me your statutes!
\par }{\Q \VS{27}Help me to understand
what your precepts
mean!

\par }{\Q Then I can meditate
on your marvelous
teachings.
\par }{\Q \VS{28}I collapse
from grief.
\par }{\Q Sustain
me by your word!
\par }{\Q \VS{29}Remove
me from
the path
of deceit!

\par }{\Q Graciously
give me your law!
\par }{\Q \VS{30}I choose
the path
of faithfulness;
\par }{\Q I am committed
to your regulations.
\par }{\Q \VS{31}I hold fast
to your rules.
\par }{\Q O
{\ND{Lord}}, do not
let me be ashamed!
\par }{\Q \VS{32}I run
along the path
of your commands,
\par }{\Q for
you enable me to do so.
\par }{\SH ה ({\IT{He}})
\par }{\Q \VS{33}Teach me, O
{\ND{Lord}}, the lifestyle prescribed by your statutes,
\par }{\Q so that I might observe it continually.
\par }{\Q \VS{34}Give me understanding
so that I might observe
your law,
\par }{\Q and keep
it with all
my heart.
\par }{\Q \VS{35}Guide
me in the path
of your commands,
\par }{\Q for
I delight to walk in it.
\par }{\Q \VS{36}Give
me
a desire
for your rules,
\par }{\Q rather
than for wealth gained unjustly.
\par }{\Q \VS{37}Turn
my eyes
away from what is worthless!

\par }{\Q Revive me with your word!
\par }{\Q \VS{38}Confirm
to your servant
your promise,
\par }{\Q which
you made to the one who honors you.
\par }{\Q \VS{39}Take away
the insults
that
I dread!

\par }{\Q Indeed,
your regulations
are good.
\par }{\Q \VS{40}Look,
I long
for your precepts.
\par }{\Q Revive
me with your deliverance!
\par }{\SH ו ({\IT{Vav}})
\par }{\Q \VS{41}May I experience your loyal love, O
{\ND{Lord}},
\par }{\Q and your deliverance, as you promised.
\par }{\Q \VS{42}Then
I will have a reply
for
the one who insults
me,

\par }{\Q for
I trust
in your word.
\par }{\Q \VS{43}Do not
completely deprive me
of a truthful
testimony,
\par }{\Q for
I await
your justice.
\par }{\Q \VS{44}Then I will keep
your law
continually
\par }{\Q now and for all time.
\par }{\Q \VS{45}I will be secure,
\par }{\Q for
I seek
your precepts.
\par }{\Q \VS{46}I will speak
about your regulations
before
kings
\par }{\Q and not
be ashamed.
\par }{\Q \VS{47}I will find delight
in your commands,
\par }{\Q which
I love.
\par }{\Q \VS{48}I will lift
my hands
to
your commands,
\par }{\Q which
I love,
\par }{\Q and I will meditate
on your statutes.
\par }{\SH ז ({\IT{Zayin}})
\par }{\Q \VS{49}Remember your word to your servant,
\par }{\Q for you have given me hope.
\par }{\Q \VS{50}This
is what comforts
me in my trouble,
\par }{\Q for
your promise
revives me.
\par }{\Q \VS{51}Arrogant
people do nothing but scoff at
me.

\par }{\Q Yet I do not
turn aside
from your law.
\par }{\Q \VS{52}I remember
your ancient
regulations,
\par }{\Q O
{\ND{Lord}}, and console myself.
\par }{\Q \VS{53}Rage
takes hold
of me because of the wicked,
\par }{\Q those who reject
your law.
\par }{\Q \VS{54}Your statutes
have been my songs
\par }{\Q in the house
where I live.
\par }{\Q \VS{55}I remember
your name
during the night,
O
{\ND{Lord}},
\par }{\Q and I will keep
your law.
\par }{\Q \VS{56}This
has
been
my practice,
\par }{\Q for
I observe
your precepts.
\par }{\SH ח ({\IT{Khet}})
\par }{\Q \VS{57}The
{\ND{Lord}} is my source of security.
\par }{\Q I have determined to follow your instructions.
\par }{\Q \VS{58}I seek
your favor with all
my heart.
\par }{\Q Have mercy
on me as you promised!
\par }{\Q \VS{59}I consider
my
actions
\par }{\Q and follow
your rules.
\par }{\Q \VS{60}I keep
your commands
\par }{\Q eagerly
and without
delay.
\par }{\Q \VS{61}The ropes
of the wicked
tighten
around me,
\par }{\Q but
I do not
forget
your law.
\par }{\Q \VS{62}In the middle
of the night
I arise
to thank
you
\par }{\Q for your just regulations.
\par }{\Q \VS{63}I am
a friend
to all
your loyal followers,
\par }{\Q and to those
who keep
your precepts.
\par }{\Q \VS{64}O
{\ND{Lord}}, your loyal love
fills
the earth.
\par }{\Q Teach
me your statutes!
\par }{\SH ט ({\IT{Tet}})
\par }{\Q \VS{65}You are good to your servant,
\par }{\Q O
{\ND{Lord}}, just as you promised.
\par }{\Q \VS{66}Teach
me proper
discernment
and understanding!
\par }{\Q For
I consider your commands
to be reliable.
\par }{\Q \VS{67}Before
I was afflicted
I
used to stray
off,

\par }{\Q but now
I keep
your instructions.
\par }{\Q \VS{68}You are good
and you do good.
\par }{\Q Teach
me your statutes!
\par }{\Q \VS{69}Arrogant
people smear
my reputation with lies,
\par }{\Q but I
observe
your precepts
with all
my heart.
\par }{\Q \VS{70}Their hearts
are calloused,
\par }{\Q but I
find delight
in your law.
\par }{\Q \VS{71}It was good
for
me to suffer,
\par }{\Q so that
I might
learn
your statutes.
\par }{\Q \VS{72}The law
you have revealed
is more important
to me
\par }{\Q than thousands
of pieces of gold
and silver.
\par }{\SH י ({\IT{Yod}})
\par }{\Q \VS{73}Your hands made me and formed me.
\par }{\Q Give me understanding so that I might learn your commands.
\par }{\Q \VS{74}Your loyal followers
will be glad
when
they see
me,

\par }{\Q for
I find hope
in your word.
\par }{\Q \VS{75}I know,

{\ND{Lord}}, that
your regulations are just.
\par }{\Q You disciplined
me because of your faithful devotion to me.
\par }{\Q \VS{76}May
your loyal love
console
me,
\par }{\Q as you promised
your servant.
\par }{\Q \VS{77}May I experience
your compassion,
so I might live!
\par }{\Q For
I find delight
in your law.
\par }{\Q \VS{78}May the arrogant
be humiliated,
for
they have slandered
me!

\par }{\Q But I
meditate
on your precepts.
\par }{\Q \VS{79}May your loyal followers
turn
to me,
\par }{\Q those who know
your rules.
\par }{\Q \VS{80}May I be
fully
committed
to your statutes,
\par }{\Q so that I might
not
be ashamed.
\par }{\SH כ ({\IT{Kaf}})
\par }{\Q \VS{81}I desperately long for your deliverance.
\par }{\Q I find hope in your word.
\par }{\Q \VS{82}My eyes
grow tired
as I wait for your promise to be fulfilled.
\par }{\Q I say, “When
will you comfort me?”
\par }{\Q \VS{83}For
I am like a wineskin
dried up in smoke.
\par }{\Q I do not
forget
your statutes.
\par }{\Q \VS{84}How
long
must your servant
endure
this?

\par }{\Q When will you judge
those who pursue me?
\par }{\Q \VS{85}The arrogant
dig
pits
to trap me,

\par }{\Q which
violates your law.
\par }{\Q \VS{86}All
your commands
are reliable.
\par }{\Q I am pursued
without reason.
Help me!
\par }{\Q \VS{87}They have almost
destroyed
me here on the earth,
\par }{\Q but I
do not
reject
your precepts.
\par }{\Q \VS{88}Revive
me with your loyal love,
\par }{\Q that I might keep
the rules
you have revealed.
\par }{\SH ל ({\IT{Lamed}})
\par }{\Q \VS{89}O
{\ND{Lord}}, your instructions endure;
\par }{\Q they stand secure in heaven.
\par }{\Q \VS{90}You demonstrate
your faithfulness
to all
generations.
\par }{\Q You established
the earth
and it stood firm.
\par }{\Q \VS{91}Today
they stand
firm by your decrees,
\par }{\Q for
all
things are your servants.
\par }{\Q \VS{92}If
I had not
found encouragement
in your law,
\par }{\Q I would have died
in my sorrow.
\par }{\Q \VS{93}I will never
forget
your precepts,
\par }{\Q for
by them you have revived me.
\par }{\Q \VS{94}I
belong to you. Deliver
me!
\par }{\Q For
I seek
your precepts.
\par }{\Q \VS{95}The wicked
prepare
to kill
me,

\par }{\Q yet I concentrate
on your rules.
\par }{\Q \VS{96}I realize
that everything
has its limits,
\par }{\Q but your commands
are beyond full
comprehension.
\par }{\SH מ ({\IT{Mem}})
\par }{\Q \VS{97}O how I love your law!
\par }{\Q All day long I meditate on it.
\par }{\Q \VS{98}Your commandments
make me wiser
than my enemies,
\par }{\Q for
I am always aware of them.
\par }{\Q \VS{99}I have more insight
than all
my teachers,
\par }{\Q for
I meditate
on your rules.
\par }{\Q \VS{100}I am more discerning
than those older
than I,
\par }{\Q for
I observe
your precepts.
\par }{\Q \VS{101}I stay away
from the evil
path,
\par }{\Q so that I might
keep
your instructions.
\par }{\Q \VS{102}I do not
turn aside
from your regulations,
\par }{\Q for
you
teach me.
\par }{\Q \VS{103}Your words
are sweeter
\par }{\Q in my mouth
than honey!
\par }{\Q \VS{104}Your precepts
give me discernment.
\par }{\Q Therefore
I hate
all
deceitful
actions.
\par }{\SH נ ({\IT{Nun}})
\par }{\Q \VS{105}Your word is a lamp to walk by,
\par }{\Q and a light to illumine my path.
\par }{\Q \VS{106}I have vowed and solemnly
sworn
\par }{\Q to keep
your just regulations.
\par }{\Q \VS{107}I am suffering terribly.
\par }{\Q O
{\ND{Lord}}, revive
me with your word!
\par }{\Q \VS{108}O
{\ND{Lord}}, please
accept
the freewill
offerings of my praise!

\par }{\Q Teach
me your regulations!
\par }{\Q \VS{109}My life
is in continual
danger,
\par }{\Q but I do not
forget
your law.
\par }{\Q \VS{110}The wicked
lay a trap
for me,
\par }{\Q but I do not
wander
from your precepts.
\par }{\Q \VS{111}I claim
your rules
as my permanent
possession,
\par }{\Q for
they
give me joy.
\par }{\Q \VS{112}I am determined
to obey
your statutes
\par }{\Q at all times,
to the very end.
\par }{\SH 
{\BD{ס (}}{\IT{Samek}}{\BD{)}}
\par }{\Q \VS{113}I hate people with divided loyalties,
\par }{\Q but I love your law.
\par }{\Q \VS{114}You are my hiding
place and my shield.
\par }{\Q I find hope
in your word.
\par }{\Q \VS{115}Turn
away from me,
you evil
men,
\par }{\Q so that I can observe
the commands
of my God.
\par }{\Q \VS{116}Sustain
me as you promised,
so that I will live.
\par }{\Q Do not
disappoint me!
\par }{\Q \VS{117}Support
me, so that I will be delivered.
\par }{\Q Then I will focus
on your statutes
continually.
\par }{\Q \VS{118}You despise
all
who stray
from your statutes,
\par }{\Q for
they are deceptive
and unreliable.
\par }{\Q \VS{119}You remove
all
the wicked
of the earth
like slag.
\par }{\Q Therefore
I love
your rules.
\par }{\Q \VS{120}My body
trembles
because I fear
you;

\par }{\Q I am afraid
of your judgments.
\par }{\SH ע ({\IT{Ayin}})
\par }{\Q \VS{121}I do what is fair and right.
\par }{\Q Do not abandon me to my oppressors!
\par }{\Q \VS{122}Guarantee
the welfare
of your servant!

\par }{\Q Do not
let the arrogant
oppress me!
\par }{\Q \VS{123}My
eyes
grow tired
as I wait for your deliverance,
\par }{\Q for your reliable
promise to be fulfilled.
\par }{\Q \VS{124}Show
your servant
your loyal love!

\par }{\Q Teach
me your statutes!
\par }{\Q \VS{125}I am
your servant.
Give me insight,
\par }{\Q so that I
can understand
your rules.
\par }{\Q \VS{126}It is time
for the
{\ND{Lord}} to act –
\par }{\Q they break your law!
\par }{\Q \VS{127}For this reason
I love
your commands
\par }{\Q more than gold,
even purest gold.
\par }{\Q \VS{128}For this reason
I carefully follow all
your precepts.
\par }{\Q I hate
all
deceitful
actions.
\par }{\SH פ ({\IT{Pe}})
\par }{\Q \VS{129}Your rules are marvelous.
\par }{\Q Therefore I observe them.
\par }{\Q \VS{130}Your instructions
are a doorway
through which light
shines.

\par }{\Q They give insight
to the untrained.
\par }{\Q \VS{131}I open my mouth
and pant,
\par }{\Q because I long for
your commands.
\par }{\Q \VS{132}Turn
toward
me and extend mercy
to me,
\par }{\Q as you typically
do to your loyal
followers.
\par }{\Q \VS{133}Direct
my steps by
your word!

\par }{\Q Do not
let any
sin
dominate me!
\par }{\Q \VS{134}Deliver
me from oppressive
men,
\par }{\Q so that I can keep
your precepts.
\par }{\Q \VS{135}Smile
on your servant!
\par }{\Q Teach
me your statutes!
\par }{\Q \VS{136}Tears
stream
down
from my eyes,
\par }{\Q because
people do not
keep
your law.
\par }{\SH צ ({\IT{Tsade}})
\par }{\Q \VS{137}You are just, O
{\ND{Lord}},
\par }{\Q and your judgments are fair.
\par }{\Q \VS{138}The rules
you impose
are just,
\par }{\Q and absolutely
reliable.
\par }{\Q \VS{139}My zeal
consumes
me,
\par }{\Q for
my enemies
forget
your instructions.
\par }{\Q \VS{140}Your word
is absolutely
pure,
\par }{\Q and your servant
loves it!
\par }{\Q \VS{141}I am
insignificant
and despised,
\par }{\Q yet I do not
forget
your precepts.
\par }{\Q \VS{142}Your justice
endures,
\par }{\Q and your law
is reliable.
\par }{\Q \VS{143}Distress
and hardship
confront
me,
\par }{\Q yet I find delight
in your commands.
\par }{\Q \VS{144}Your rules
remain
just.
\par }{\Q Give me insight
so that I can live.
\par }{\SH ק ({\IT{Qof}})
\par }{\Q \VS{145}I cried out with all my heart, “Answer me, O
{\ND{Lord}}!
\par }{\Q I will observe your statutes.”
\par }{\Q \VS{146}I cried
out to you, “Deliver
me,
\par }{\Q so that I can keep
your rules.”
\par }{\Q \VS{147}I am up before
dawn
crying for help.
\par }{\Q I find hope
in your word.
\par }{\Q \VS{148}My
eyes
anticipate
the nighttime
hours,
\par }{\Q so that I can meditate on
your word.
\par }{\Q \VS{149}Listen
to me because of your loyal love!
\par }{\Q O
{\ND{Lord}}, revive
me, as you typically do!
\par }{\Q \VS{150}Those who
are eager
to do wrong draw near;
\par }{\Q they are far
from your law.
\par }{\Q \VS{151}You are near,
O
{\ND{Lord}},
\par }{\Q and all
your commands
are reliable.
\par }{\Q \VS{152}I learned
long ago
that
\par }{\Q you ordained
your rules
to last.
\par }{\SH ר ({\IT{Resh}})
\par }{\Q \VS{153}See my pain and rescue me!
\par }{\Q For I do not forget your law.
\par }{\Q \VS{154}Fight
for me and defend
me!

\par }{\Q Revive
me with your word!
\par }{\Q \VS{155}The wicked
have no chance
for deliverance,
\par }{\Q for
they do not
seek
your statutes.
\par }{\Q \VS{156}Your compassion
is great,
O
{\ND{Lord}}.
\par }{\Q Revive
me, as you typically do!
\par }{\Q \VS{157}The enemies
who chase
me are numerous.
\par }{\Q Yet I do not
turn aside
from your rules.
\par }{\Q \VS{158}I take note
of the treacherous
and despise
them,
\par }{\Q because they do not
keep
your instructions.
\par }{\Q \VS{159}See
how I love
your precepts!
\par }{\Q O
{\ND{Lord}}, revive
me with your loyal love!
\par }{\Q \VS{160}Your instructions
are totally reliable;
\par }{\Q all
your just regulations endure.
\par }{\SH שׂ/שׁ ({\IT{Sin/Shin}})
\par }{\Q \VS{161}Rulers pursue me for no reason,
\par }{\Q yet I am more afraid of disobeying your instructions.
\par }{\Q \VS{162}I
rejoice
in your instructions,
\par }{\Q like one who finds
much
plunder.
\par }{\Q \VS{163}I hate
and despise
deceit;
\par }{\Q I love
your law.
\par }{\Q \VS{164}Seven
times a day
I praise
you
\par }{\Q because
of your just regulations.
\par }{\Q \VS{165}Those who love
your law
are completely
secure;
\par }{\Q nothing
causes them to stumble.
\par }{\Q \VS{166}I hope
for your deliverance,
O
{\ND{Lord}},
\par }{\Q and I obey your commands.
\par }{\Q \VS{167}I keep
your
rules;
\par }{\Q I love
them greatly.
\par }{\Q \VS{168}I keep
your precepts
and rules,
\par }{\Q for
you are aware
of everything I do.
\par }{\SH ת ({\IT{Tav}})
\par }{\Q \VS{169}Listen to my cry for help, O
{\ND{Lord}}!
\par }{\Q Give me insight by your word!
\par }{\Q \VS{170}Listen to my appeal for mercy!
\par }{\Q Deliver
me, as you promised.
\par }{\Q \VS{171}May praise
flow freely
from my lips,
\par }{\Q for
you teach
me your statutes.
\par }{\Q \VS{172}May
my tongue
sing
about your instructions,
\par }{\Q for
all
your commands
are just.
\par }{\Q \VS{173}May
your hand
help
me,
\par }{\Q for
I choose
to obey your precepts.
\par }{\Q \VS{174}I long
for your deliverance,
O
{\ND{Lord}};
\par }{\Q I find delight
in your law.
\par }{\Q \VS{175}May I live
and praise
you!
\par }{\Q May your regulations
help me!
\par }{\Q \VS{176}I have wandered
off like a lost
sheep.
\par }{\Q Come looking
for your servant,
\par }{\Q for
I do not
forget
your commands.














\par }\PsalmChap{120}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}In my distress
I cried
out
\par }{\Q to
the {\ND{Lord}}
and he answered me.
\par }{\Q \VS{2}I said, “O
{\ND{Lord}},
rescue
me

\par }{\Q from those
who lie
with their lips
\par }{\Q and those who deceive
with their tongue.
\par }{\Q \VS{3}How
will
he severely
punish you,
\par }{\Q you deceptive
talker?
\par }{\Q \VS{4}Here’s how! With the sharp
arrows
of warriors,
\par }{\Q with
arrowheads
forged over the hot coals.
\par }{\Q \VS{5}How miserable
I am!

\par }{\Q For
I have lived temporarily
in Meshech;
\par }{\Q I have resided among
the tents
of Kedar.
\par }{\Q \VS{6}For too long
I have had to reside
\par }{\Q with
those who hate
peace.
\par }{\Q \VS{7}I
am committed to peace,
\par }{\Q but when I speak,
they
want to make war.


\par }\PsalmChap{121}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}I look
up toward
the hills.
\par }{\Q From where
does my help
come?
\par }{\Q \VS{2}My help
comes
from the
{\ND{Lord}},

\par }{\Q the Creator
of heaven
and earth!
\par }{\Q \VS{3}May he not
allow
your foot
to slip!
\par }{\Q May your protector
not
sleep!
\par }{\Q \VS{4}Look! Israel’s
protector
\par }{\Q does not
sleep
or
slumber!
\par }{\Q \VS{5}The
{\ND{Lord}}
is your protector;
\par }{\Q the {\ND{Lord}}
is the shade
at your right hand.
\par }{\Q \VS{6}The sun
will not
harm
you by day,
\par }{\Q or the moon
by night.
\par }{\Q \VS{7}The
{\ND{Lord}}
will protect
you from all
harm;
\par }{\Q he will protect
your life.
\par }{\Q \VS{8}The
{\ND{Lord}}
will protect
you in all you do,
\par }{\Q now
and forevermore.


\par }\PsalmChap{122}{\D 
A song
of ascents,
by David.
\par }{\Q \VerseOne{1}I was glad
because they said
to me,
\par }{\Q “We will go
to the
{\ND{Lord}}’s
temple.”
\par }{\Q \VS{2}Our feet
are standing
\par }{\Q inside your gates,
O Jerusalem.
\par }{\Q \VS{3}Jerusalem
is a city
designed
\par }{\Q to accommodate an assembly.
\par }{\Q \VS{4}The tribes
go up
there,
\par }{\Q the tribes
of the
{\ND{Lord}},
\par }{\Q where it is required that Israel
\par }{\Q give thanks
to the name
of the {\ND{Lord}}.
\par }{\Q \VS{5}Indeed,
the leaders sit
there
on thrones
and make legal decisions,
\par }{\Q on the thrones
of the house
of David.
\par }{\Q \VS{6}Pray
for the peace
of Jerusalem!
\par }{\Q May those who love
her prosper!
\par }{\Q \VS{7}May there be
peace
inside your defenses,
\par }{\Q and prosperity
inside your fortresses!
\par }{\Q \VS{8}For the sake
of my brothers
and my neighbors
\par }{\Q I will say,
“May there be peace in you!”
\par }{\Q \VS{9}For the sake
of the temple
of the {\ND{Lord}}
our God
\par }{\Q I will pray
for you to prosper.


\par }\PsalmChap{123}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}I look
up toward you,
\par }{\Q the one enthroned in
heaven.
\par }{\Q \VS{2}Look,
as the eyes
of servants
look to
the hand
of their master,
\par }{\Q as the eyes
of a female servant
look to
the hand
of her mistress,
\par }{\Q so
my eyes
will look to
the {\ND{Lord}}, our God,
until
he shows us favor.
\par }{\Q \VS{3}Show us favor,
O
{\ND{Lord}}, show us favor!
\par }{\Q For
we have had our fill
of humiliation,
and then some.
\par }{\Q \VS{4}We
have had our fill
\par }{\Q of the taunts
of the self-assured,
\par }{\Q of the contempt
of the proud.


\par }\PsalmChap{124}{\D 
A song
of ascents,
by David.
\par }{\Q \VerseOne{1}“If
the {\ND{Lord}}
had not been on our side” –
\par }{\Q let Israel say this! –
\par }{\Q \VS{2}if
the {\ND{Lord}}
had not been
on
our side,
\par }{\Q when men
attacked
us,
\par }{\Q \VS{3}they would have swallowed
us alive,
\par }{\Q when their anger
raged against us.
\par }{\Q \VS{4}The water
would have overpowered
us;
\par }{\Q the current
would have overwhelmed
us.
\par }{\Q \VS{5}The raging
water
\par }{\Q would have overwhelmed
us.
\par }{\Q \VS{6}The
{\ND{Lord}}
deserves praise,
\par }{\Q for he did not
hand
us over as prey
to their teeth.
\par }{\Q \VS{7}We escaped
with our lives,
like a bird
from
a hunter’s
snare.
\par }{\Q The snare
broke,
and we
escaped.
\par }{\Q \VS{8}Our deliverer
is the
{\ND{Lord}},

\par }{\Q the Creator
of heaven
and earth.


\par }\PsalmChap{125}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}Those who
trust
in the
{\ND{Lord}}
are like Mount
Zion;
\par }{\Q it cannot
be upended
and will endure
forever.
\par }{\Q \VS{2}As the mountains
surround
Jerusalem,
\par }{\Q so the
{\ND{Lord}}
surrounds
his people,
\par }{\Q now
and forevermore.
\par }{\Q \VS{3}Indeed,
the scepter
of a wicked
king will
not
settle
\par }{\Q upon
the allotted
land of the godly.
\par }{\Q Otherwise
the godly
might
\par }{\Q do what is wrong.
\par }{\Q \VS{4}Do good,
O
{\ND{Lord}}, to those who are good,
\par }{\Q to the morally
upright!
\par }{\Q \VS{5}As for those who are bent on
traveling
a sinful path,
\par }{\Q may the
{\ND{Lord}}
remove them, along with
those who behave
wickedly!

\par }{\Q May Israel
experience peace!


\par }\PsalmChap{126}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}When the
{\ND{Lord}}
restored
the well-being of Zion,
\par }{\Q we thought we were
dreaming.
\par }{\Q \VS{2}At that time
we laughed
loudly
\par }{\Q and shouted for joy.
\par }{\Q At that time
the nations
said,
\par }{\Q “The
{\ND{Lord}}
has accomplished
great
things
for these people.”
\par }{\Q \VS{3}The
{\ND{Lord}}
did
indeed accomplish
great
things
for us.
\par }{\Q We were happy.
\par }{\Q \VS{4}O
{\ND{Lord}}, restore
our well-being,
\par }{\Q just as the streams
in the arid south are replenished.
\par }{\Q \VS{5}Those who shed tears
as they plant
\par }{\Q will shout
for joy when they reap the harvest.
\par }{\Q \VS{6}The one who weeps
as he walks
along, carrying
his bag
of seed,
\par }{\Q will certainly
come
in with a shout of joy,
carrying
his sheaves of grain.


\par }\PsalmChap{127}{\D 
A song
of ascents,
by Solomon.
\par }{\Q \VerseOne{1}If
the {\ND{Lord}}
does not
build
a house,
\par }{\Q then those who build
it work
in vain.
\par }{\Q If
the {\ND{Lord}}
does not
guard
a city,
\par }{\Q then the watchman
stands guard
in vain.
\par }{\Q \VS{2}It is vain
for you to rise
early,
come home
late,
\par }{\Q and work so hard
for your food.
\par }{\Q Yes,
he can provide
for those whom he loves
even when they sleep.
\par }{\Q \VS{3}Yes,
sons
are a gift
from the
{\ND{Lord}},
\par }{\Q the fruit
of the womb
is a reward.
\par }{\Q \VS{4}Sons
born during one’s youth
\par }{\Q are like
arrows
in a warrior’s
hand.
\par }{\Q \VS{5}How blessed
is the man
who fills
his quiver
with them!
\par }{\Q They
will not
be put to shame
when
they confront
enemies
at the city gate.


\par }\PsalmChap{128}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}How blessed
is every
one of the
{\ND{Lord}}’s
loyal followers,
\par }{\Q each one who keeps
his commands!
\par }{\Q \VS{2}You will eat
what
you worked
so hard to grow.
\par }{\Q You will be blessed
and secure.
\par }{\Q \VS{3}Your wife
will be like a fruitful
vine
\par }{\Q in the inner rooms
of your house;
\par }{\Q your children
will be like olive
branches,
\par }{\Q as they sit all around
your table.
\par }{\Q \VS{4}Yes
indeed,
the man
who fears
the {\ND{Lord}}
\par }{\Q will be blessed
in this way.
\par }{\Q \VS{5}May the
{\ND{Lord}}
bless
you from Zion,
\par }{\Q that you might see
Jerusalem
prosper
\par }{\Q all
the days
of your life,
\par }{\Q \VS{6}and that you might see
your grandchildren.
\par }{\Q May Israel
experience peace!


\par }\PsalmChap{129}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}“Since my youth
they have often
attacked me,”
\par }{\Q let
Israel
say.
\par }{\Q \VS{2}“Since my youth
they have
often
attacked me,
\par }{\Q but they have not
defeated me.
\par }{\Q \VS{3}The plowers
plowed
my back;
\par }{\Q they made their furrows
long.
\par }{\Q \VS{4}The
{\ND{Lord}}
is just;
\par }{\Q he cut
the ropes
of the wicked.”
\par }{\Q \VS{5}May all
who hate
Zion
\par }{\Q be humiliated
and turned
back!
\par }{\Q \VS{6}May they be
like the grass
on the rooftops
\par }{\Q which withers
before
one can even pull
it up,
\par }{\Q \VS{7}which cannot
fill
the reaper’s
hand,
\par }{\Q or the lap
of the one who gathers the grain!
\par }{\Q \VS{8}Those who pass
by will not
say,
\par }{\Q “May you experience the
{\ND{Lord}}’s
blessing!
\par }{\Q We pronounce
a blessing on
you in the name
of the {\ND{Lord}}.”


\par }\PsalmChap{130}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}From the deep
water I cry
out to you, O
{\ND{Lord}}.
\par }{\Q \VS{2}O Lord,
listen
to me!

\par }{\Q Pay attention
to my plea for mercy!
\par }{\Q \VS{3}If
you, O
{\ND{Lord}}, were to keep track
of sins,
\par }{\Q O Lord,
who
could stand before you?
\par }{\Q \VS{4}But
you are willing
to forgive,
\par }{\Q so that you might
be honored.
\par }{\Q \VS{5}I rely
on the
{\ND{Lord}},
\par }{\Q I rely
on him with my whole being;
\par }{\Q I wait for his assuring
word.
\par }{\Q \VS{6}I yearn
for the Lord,
\par }{\Q more than watchmen
do
for the morning,
\par }{\Q yes, more than watchmen do
for the morning.
\par }{\Q \VS{7}O Israel,
hope
in the
{\ND{Lord}},
\par }{\Q for
the {\ND{Lord}}
exhibits loyal love,
\par }{\Q and is more
than willing to deliver.
\par }{\Q \VS{8}He
will deliver
Israel
\par }{\Q from all
the consequences of their sins.


\par }\PsalmChap{131}{\D 
A song
of ascents,
by David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, my heart
is not
proud,
\par }{\Q nor
do I have a haughty
look.
\par }{\Q I do not
have great
aspirations,
\par }{\Q or concern myself with things
that are beyond me.
\par }{\Q \VS{2}Indeed I am composed
and quiet,
\par }{\Q like a young child carried
by its
mother;
\par }{\Q I am content
like the young child
I carry.
\par }{\Q \VS{3}O Israel,
hope
in the
{\ND{Lord}}
\par }{\Q now
and forevermore!


\par }\PsalmChap{132}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, for David’s
sake remember
\par }{\Q all
his strenuous effort,
\par }{\Q \VS{2}and how
he made a vow
to the
{\ND{Lord}},
\par }{\Q and swore
an oath to the powerful ruler
of Jacob.
\par }{\Q \VS{3}He said, “I will not
enter
my own home,
\par }{\Q or
get into
my bed.
\par }{\Q \VS{4}I will not
allow
my eyes
to sleep,
\par }{\Q or my eyelids
to slumber,
\par }{\Q \VS{5}until
I find
a place
for the
{\ND{Lord}},
\par }{\Q a fine dwelling place
for the powerful ruler
of Jacob.”
\par }{\Q \VS{6}Look,
we heard
about it in Ephrathah,
\par }{\Q we found
it in the territory
of Jaar.
\par }{\Q \VS{7}Let us go
to his dwelling place!
\par }{\Q Let us worship
before his footstool!
\par }{\Q \VS{8}Ascend,
O
{\ND{Lord}}, to your resting place,
\par }{\Q you
and the ark
of your strength!
\par }{\Q \VS{9}May your priests
be clothed
with integrity!

\par }{\Q May your loyal followers
shout for joy!
\par }{\Q \VS{10}For the sake of
David,
your servant,
\par }{\Q do not
reject
your chosen king!
\par }{\Q \VS{11}The
{\ND{Lord}}
made
a reliable
promise
to David;
\par }{\Q he will not
go back
on his word.

\par }{\Q He said, “I will place one of your descendants
on your throne.
\par }{\Q \VS{12}If
your sons
keep
my covenant
\par }{\Q and the rules
I
teach
them,
\par }{\Q their sons
will also
sit
on your throne
forever.”
\par }{\Q \VS{13}Certainly
the {\ND{Lord}}
has chosen
Zion;
\par }{\Q he decided
to make it his home.
\par }{\Q \VS{14}He said, “This
will be my resting place
forever;
\par }{\Q I will live
here,
for
I have chosen it.
\par }{\Q \VS{15}I will abundantly
supply
what she needs;

\par }{\Q I will give her
poor
all the food
they need.
\par }{\Q \VS{16}I will protect
her priests,
\par }{\Q and her godly people
will shout
exuberantly.
\par }{\Q \VS{17}There
I will make
David
strong;
\par }{\Q I have determined
that my chosen king’s
dynasty will continue.
\par }{\Q \VS{18}I will humiliate
his enemies,
\par }{\Q and his crown
will shine.


\par }\PsalmChap{133}{\D 
A song
of ascents,
by David.
\par }{\Q \VerseOne{1}Look! How
good
and how
pleasant
it is
\par }{\Q when brothers
live
together!
\par }{\Q \VS{2}It is like fine
oil
poured
on
the head
\par }{\Q which flows down
the beard –
\par }{\Q Aaron’s
beard,
\par }{\Q and then flows
down
his garments.
\par }{\Q \VS{3}It is like the dew
of Hermon,
\par }{\Q which flows down
upon
the hills
of Zion.
\par }{\Q Indeed
that is where
the {\ND{Lord}}
has decreed
\par }{\Q a blessing will be available – eternal life.


\par }\PsalmChap{134}{\D 
A song
of ascents.
\par }{\Q \VerseOne{1}Attention! Praise
the {\ND{Lord}},
\par }{\Q all
you servants
of the {\ND{Lord}},
\par }{\Q who serve
in the
{\ND{Lord}}’s
temple
during the night.
\par }{\Q \VS{2}Lift
your hands
toward the sanctuary
\par }{\Q and praise
the {\ND{Lord}}!
\par }{\Q \VS{3}May the
{\ND{Lord}}, the Creator
of heaven
and earth,
\par }{\Q bless
you from Zion!


\par }\PsalmChap{135}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Praise
the name
of the {\ND{Lord}}!
\par }{\Q Offer praise,
you servants
of the {\ND{Lord}},
\par }{\Q \VS{2}who serve
in the
{\ND{Lord}}’s
temple,
\par }{\Q in the courts
of the temple
of our God.
\par }{\Q \VS{3}Praise
the
{\ND{Lord}}, for
the {\ND{Lord}}
is good!
\par }{\Q Sing praises
to his name,
for
it is pleasant!
\par }{\Q \VS{4}Indeed,
the
{\ND{Lord}}
has chosen
Jacob
for himself,
\par }{\Q Israel
to be his special possession.
\par }{\Q \VS{5}Yes,
I
know
the {\ND{Lord}}
is great,
\par }{\Q and our Lord
is superior to all
gods.
\par }{\Q \VS{6}He does whatever
he pleases
\par }{\Q in heaven
and on earth,
\par }{\Q in the seas
and all
the ocean depths.
\par }{\Q \VS{7}He causes
the clouds to arise from
the end
of the earth,
\par }{\Q makes lightning
bolts accompany the rain,
\par }{\Q and brings
the wind
out
of his storehouses.
\par }{\Q \VS{8}He struck
down the firstborn
of Egypt,
\par }{\Q including both men
and animals.
\par }{\Q \VS{9}He performed
awesome
deeds and acts of judgment
\par }{\Q in your midst,
O Egypt,
\par }{\Q against Pharaoh
and all
his servants.
\par }{\Q \VS{10}He defeated
many
nations,
\par }{\Q and killed
mighty
kings –
\par }{\Q \VS{11}Sihon,
king
of the Amorites,
\par }{\Q and Og,
king
of Bashan,
\par }{\Q and all
the kingdoms
of Canaan.
\par }{\Q \VS{12}He gave
their land
as an inheritance,
\par }{\Q as an inheritance
to Israel
his people.
\par }{\Q \VS{13}O
{\ND{Lord}}, your name
endures,
\par }{\Q your reputation, O
{\ND{Lord}}, lasts.
\par }{\Q \VS{14}For
the {\ND{Lord}}
vindicates
his people,
\par }{\Q and has compassion
on
his servants.
\par }{\Q \VS{15}The nations’
idols are made of silver
and gold,
\par }{\Q they are man-made.
\par }{\Q \VS{16}They have mouths,
but cannot
speak,
\par }{\Q eyes,
but cannot
see,
\par }{\Q \VS{17}and ears,
but cannot hear.
\par }{\Q Indeed,
they
cannot breathe.
\par }{\Q \VS{18}Those who make
them will end up like
them,
\par }{\Q as will everyone
who
trusts in them.
\par }{\Q \VS{19}O family
of Israel,
praise
the {\ND{Lord}}!
\par }{\Q O family
of Aaron,
praise
the {\ND{Lord}}!
\par }{\Q \VS{20}O family
of Levi,
praise
the {\ND{Lord}}!
\par }{\Q You loyal followers
of the {\ND{Lord}}, praise
the {\ND{Lord}}!
\par }{\Q \VS{21}The
{\ND{Lord}}
deserves
praise in Zion –
\par }{\Q he who dwells
in Jerusalem.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{136}{\PP \par }{\Q \VerseOne{1}Give thanks
to the
{\ND{Lord}},
for
he is good,
\par }{\Q for
his loyal love
endures.
\par }{\Q \VS{2}Give thanks
to the God
of gods,
\par }{\Q for
his loyal love
endures.
\par }{\Q \VS{3}Give thanks
to the Lord
of lords,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{4}to the one who performs
magnificent,
amazing
deeds all
by himself,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{5}to the one who used
wisdom
to make
the heavens,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{6}to the one who spread
out the earth
over
the water,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{7}to the one who made
the great
lights,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{8}the sun
to rule
by day,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{9}the moon
and stars
to rule
by night,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{10}to the one who struck
down the firstborn
of Egypt,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{11}and led
Israel
out from their midst,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{12}with a strong
hand
and an outstretched
arm,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{13}to the one who divided
the Red
Sea
in two,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{14}and led Israel
through
its midst,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{15}and tossed
Pharaoh
and his army
into the Red
Sea,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{16}to the one who led
his people
through the wilderness,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{17}to the one who struck
down great
kings,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{18}and killed
powerful
kings,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{19}Sihon,
king
of the Amorites,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{20}Og,
king
of Bashan,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{21}and gave
their land
as an inheritance,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{22}as an inheritance
to Israel
his servant,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{23}to the one who remembered
us when we were down,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{24}and snatched
us away from our enemies,
\par }{\Q for
his loyal love
endures,
\par }{\Q \VS{25}to the one who gives
food
to all
living
things,

\par }{\Q for
his loyal love
endures.
\par }{\Q \VS{26}Give thanks
to the God
of heaven,
\par }{\Q for
his loyal love
endures!


\par }\PsalmChap{137}{\PP \par }{\Q \VerseOne{1}By the rivers
of Babylon
\par }{\Q we sit
down and weep
\par }{\Q when we remember
Zion.
\par }{\Q \VS{2}On
the poplars
in her midst
\par }{\Q we hang
our harps,
\par }{\Q \VS{3}for
there
our captors
ask
us to compose songs;
\par }{\Q those who mock
us demand that we be happy,
saying:

\par }{\Q “Sing
for us a song
about Zion!”
\par }{\Q \VS{4}How
can we sing
a song
to the
{\ND{Lord}}
\par }{\Q in a foreign
land?
\par }{\Q \VS{5}If
I forget
you, O Jerusalem,
\par }{\Q may my right hand
be crippled!
\par }{\Q \VS{6}May my tongue
stick
to the roof of my mouth,
\par }{\Q if
I do not
remember
you,
\par }{\Q and do not
give Jerusalem
priority
\par }{\Q over
whatever gives me the most joy.
\par }{\Q \VS{7}Remember,
O
{\ND{Lord}}, what the Edomites
did
\par }{\Q on the day
Jerusalem
fell.
\par }{\Q They said,
“Tear
it down, tear
it down,

\par }{\Q right to its very
foundation!”
\par }{\Q \VS{8}O daughter
Babylon,
soon to be devastated!

\par }{\Q How blessed
will be the one who repays
you
\par }{\Q for what you dished
out to us!
\par }{\Q \VS{9}How blessed
will be the one who grabs
your babies
\par }{\Q and smashes
them on
a rock!


\par }\PsalmChap{138}{\D 
By David.
\par }{\Q \VerseOne{1}I will give you thanks
with all
my heart;
\par }{\Q before
the heavenly
assembly I will sing praises to you.
\par }{\Q \VS{2}I will bow
down toward
your holy
temple,
\par }{\Q and give thanks
to your name,
\par }{\Q because
of your loyal love
and faithfulness,
\par }{\Q for
you have exalted
your promise
above
the entire
sky.
\par }{\Q \VS{3}When
I cried
out for help, you answered
me.
\par }{\Q You made me bold
and energized me.
\par }{\Q \VS{4}Let all
the kings
of the earth
give thanks
to you, O
{\ND{Lord}},
\par }{\Q when
they hear
the words
you speak.
\par }{\Q \VS{5}Let them sing
about
the
{\ND{Lord}}’s
deeds,
\par }{\Q for
the
{\ND{Lord}}’s
splendor
is magnificent.
\par }{\Q \VS{6}Though
the {\ND{Lord}}
is exalted,
he takes note of the lowly,
\par }{\Q and recognizes
the proud
from far away.
\par }{\Q \VS{7}Even when I must
walk
in the midst
of danger,
you revive
me.
\par }{\Q You oppose
my angry
enemies,
\par }{\Q and your right hand
delivers me.
\par }{\Q \VS{8}The
{\ND{Lord}}
avenges
me.

\par }{\Q O
{\ND{Lord}}, your loyal love
endures.
\par }{\Q Do
not
abandon
those whom you have made!


\par }\PsalmChap{139}{\D 
For the music director,
a psalm
of David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, you examine
me and know.
\par }{\Q \VS{2}You
know
when I sit
down and when I get
up;
\par }{\Q even from far
away you understand
my motives.
\par }{\Q \VS{3}You carefully observe
me when I travel
or when I lie down
to rest;

\par }{\Q you are aware
of everything I do.
\par }{\Q \VS{4}Certainly
my tongue
does not
frame
a word
\par }{\Q without you, O
{\ND{Lord}}, being thoroughly aware of it.
\par }{\Q \VS{5}You squeeze
me in from
behind
and in front;
\par }{\Q you place
your hand
on me.
\par }{\Q \VS{6}Your knowledge
is beyond
my comprehension;
\par }{\Q it is so far beyond me,
I am unable
to fathom it.
\par }{\Q \VS{7}Where
can I go
to escape your spirit?
\par }{\Q Where
can I flee to escape
your presence?
\par }{\Q \VS{8}If
I were to ascend
to heaven,
you would be there.
\par }{\Q If
I were to sprawl out
in Sheol,
there you would be.
\par }{\Q \VS{9}If I were to fly
away on the wings
of the dawn,
\par }{\Q and settle
down on the other side
of the sea,
\par }{\Q \VS{10}even
there
your hand
would guide
me,
\par }{\Q your right hand
would grab hold of me.
\par }{\Q \VS{11}If I were to say,
“Certainly
the darkness
will cover
me,

\par }{\Q and the light
will turn
to night all around me,”
\par }{\Q \VS{12}even
the darkness
is not
too dark
for you to see,

\par }{\Q and the night
is as bright as
day;
\par }{\Q darkness
and light are the same to you.
\par }{\Q \VS{13}Certainly
you
made
my mind
and heart;
\par }{\Q you wove
me together in my mother’s
womb.
\par }{\Q \VS{14}I will give you thanks
because
your deeds
are awesome
and amazing.
\par }{\Q You knew
me thoroughly;
\par }{\Q \VS{15}my bones were not
hidden
from
you,
\par }{\Q when I was made
in secret
\par }{\Q and sewed
together in the depths
of the earth.
\par }{\Q \VS{16}Your eyes
saw
me when I was inside the womb.
\par }{\Q All
the days
ordained
for me
\par }{\Q were recorded
in your scroll
\par }{\Q before one of them came into existence.
\par }{\Q \VS{17}How
difficult it is for me to fathom
your thoughts
about me, O God!

\par }{\Q How
vast is their sum total!
\par }{\Q \VS{18}If I tried to count
them,
\par }{\Q they would outnumber the grains
of sand.
\par }{\Q Even if I finished counting
them,
\par }{\Q I would still
have to contend with you.
\par }{\Q \VS{19}If
only
you would kill the wicked,
O God!
\par }{\Q Get away
from
me, you violent
men!
\par }{\Q \VS{20}They rebel
against you and act
deceitfully;
\par }{\Q your enemies
lie.
\par }{\Q \VS{21}O
{\ND{Lord}}, do I not
hate
those who hate
you,
\par }{\Q and despise
those who oppose you?
\par }{\Q \VS{22}I absolutely
hate
them,
\par }{\Q they have become
my enemies!
\par }{\Q \VS{23}Examine
me, and probe
my thoughts!

\par }{\Q Test
me, and know
my concerns!
\par }{\Q \VS{24}See
if
there is any idolatrous tendency
in me,
\par }{\Q and lead
me in the reliable ancient
path!


\par }\PsalmChap{140}{\D 
For the music director;
a psalm
of David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, rescue
me from wicked
men!


\par }{\Q Protect
me from violent
men,
\par }{\Q \VS{2}who
plan
ways to harm
me.

\par }{\Q All
day
long they stir up conflict.
\par }{\Q \VS{3}Their tongues
wound like
a serpent;
\par }{\Q a viper’s
venom is behind their lips.
(Selah)
\par }{\Q \VS{4}O
{\ND{Lord}}, shelter
me from the power
of the wicked!
\par }{\Q Protect
me from violent
men,
\par }{\Q who
plan
to knock me over.
\par }{\Q \VS{5}Proud
men hide
a snare
for me;
\par }{\Q evil men spread
a net
by
the path;
\par }{\Q they set
traps
for me. (Selah)
\par }{\Q \VS{6}I say
to the
{\ND{Lord}}, “You are my God.”
\par }{\Q O
{\ND{Lord}}, pay attention
to my plea for mercy!
\par }{\Q \VS{7}O sovereign
{\SC{
Lord}}, my strong
deliverer,
\par }{\Q you shield
my head
in the day
of battle.
\par }{\Q \VS{8}O
{\ND{Lord}}, do not
let
the wicked
have their way!

\par }{\Q Do not
allow
their plan
to succeed
when they attack! (Selah)
\par }{\Q \VS{9}As for the heads
of those who surround me –
\par }{\Q may the harm done by their lips overwhelm them!
\par }{\Q \VS{10}May he rain down
fiery coals
upon
them!
\par }{\Q May he throw
them into the fire!
\par }{\Q From bottomless pits
they will not
escape.
\par }{\Q \VS{11}A slanderer
will not
endure on
the earth;
\par }{\Q calamity
will hunt
down a violent
man
and strike him down.
\par }{\Q \VS{12}I know
that
the {\ND{Lord}}
defends
the cause
of the oppressed
\par }{\Q and vindicates
the poor.
\par }{\Q \VS{13}Certainly
the godly
will give thanks
to your name;
\par }{\Q the morally upright
will live
in your presence.


\par }\PsalmChap{141}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, I cry
out to you. Come quickly
to me!
\par }{\Q Pay attention
to me when
I cry out to you!
\par }{\Q \VS{2}May you accept
my prayer
like incense,
\par }{\Q my uplifted
hands
like the evening
offering!
\par }{\Q \VS{3}O
{\ND{Lord}}, place a guard
on my mouth!
\par }{\Q Protect
the opening
of my lips!
\par }{\Q \VS{4}Do not
let me
have
evil
desires,

\par }{\Q or participate
in sinful activities
\par }{\Q with
men
who behave
wickedly.
\par }{\Q I will not
eat
their delicacies.
\par }{\Q \VS{5}May the godly
strike
me in love
and correct
me!
\par }{\Q May my head
not
refuse
choice
oil!

\par }{\Q Indeed,
my prayer
is a witness against their evil deeds.
\par }{\Q \VS{6}They will be thrown down
the side of a cliff
by
their judges.
\par }{\Q They will listen
to my words,
for
they are pleasant.
\par }{\Q \VS{7}As
when one plows
and breaks up the soil,

\par }{\Q so our bones
are scattered
at the mouth
of Sheol.
\par }{\Q \VS{8}Surely
I am looking
to
you, O
sovereign

{\ND{Lord}}.
\par }{\Q In you I take shelter.
\par }{\Q Do not
expose
me to danger!
\par }{\Q \VS{9}Protect
me from the snare
they have laid
for me,
\par }{\Q and the traps
the evildoers
have set.
\par }{\Q \VS{10}Let the wicked
fall
into their own nets,
\par }{\Q while I escape.


\par }\PsalmChap{142}{\D 
A well-written song
by David,
when
he was in the cave;
a prayer.
\par }{\Q \VerseOne{1}To
the {\ND{Lord}}
I cry
out;

\par }{\Q to
the {\ND{Lord}}
I plead for mercy.
\par }{\Q \VS{2}I pour out
my lament
before
him;
\par }{\Q I tell
him about my troubles.
\par }{\Q \VS{3}Even when my strength
leaves
me,

\par }{\Q you
watch
my footsteps.
\par }{\Q In the path
where
I walk
\par }{\Q they have hidden
a trap for me.
\par }{\Q \VS{4}Look
to the right
and see!
\par }{\Q No
one cares about
me.

\par }{\Q I have nowhere
to run;
\par }{\Q no
one is concerned about
my life.
\par }{\Q \VS{5}I cry
out to
you, O
{\ND{Lord}};
\par }{\Q I say,
“You
are my shelter,
\par }{\Q my security
in the land
of the living.”
\par }{\Q \VS{6}Listen
to
my cry
for help,
\par }{\Q for
I am in serious
trouble!

\par }{\Q Rescue
me from those who chase
me,
\par }{\Q for
they are stronger than I am.
\par }{\Q \VS{7}Free
me from prison,
\par }{\Q that I
may give thanks
to your name.
\par }{\Q Because of me the godly
will assemble,
\par }{\Q for
you will vindicate me.


\par }\PsalmChap{143}{\D 
A psalm
of David.
\par }{\Q \VerseOne{1}O
{\ND{Lord}}, hear
my prayer!
\par }{\Q Pay attention
to my plea
for help!
\par }{\Q Because of your faithfulness
and justice,
answer me!
\par }{\Q \VS{2}Do not
sit
in judgment
on your servant,
\par }{\Q for
no
one alive
is innocent
before you.
\par }{\Q \VS{3}Certainly
my enemies
chase
me.
\par }{\Q They smash
me into the ground.
\par }{\Q They force me
to live in
dark regions,
\par }{\Q like those who have been dead
for ages.
\par }{\Q \VS{4}My strength leaves
me;

\par }{\Q I am absolutely
shocked.
\par }{\Q \VS{5}I recall
the old
days;
\par }{\Q I meditate
on all
you have done;
\par }{\Q I reflect on your accomplishments.
\par }{\Q \VS{6}I spread
my hands
out to
you in prayer;

\par }{\Q my soul
thirsts for you in a parched
land.
\par }{\Q \VS{7}Answer
me quickly,

{\ND{Lord}}!
\par }{\Q My strength is fading.
\par }{\Q Do not
reject
me,

\par }{\Q or
I will join those descending
into the grave.
\par }{\Q \VS{8}May I hear
about your loyal love
in the morning,
\par }{\Q for
I trust
in you.
\par }{\Q Show
me the way
I should
go,
\par }{\Q because
I long for you.
\par }{\Q \VS{9}Rescue
me from my enemies,
O
{\ND{Lord}}!
\par }{\Q I run to you for protection.
\par }{\Q \VS{10}Teach
me to do
what pleases
you,

\par }{\Q for
you
are my God.
\par }{\Q May your kind
presence

\par }{\Q lead
me into a level
land.
\par }{\Q \VS{11}O
{\ND{Lord}}, for the sake
of your reputation,
revive
me!

\par }{\Q Because of your justice,
rescue
me from trouble!
\par }{\Q \VS{12}As a demonstration
of your loyal love,
destroy my enemies!
\par }{\Q Annihilate
all
who threaten
my life,
\par }{\Q for
I am
your servant.


\par }\PsalmChap{144}{\D 
By David.
\par }{\Q \VerseOne{1}The
{\ND{Lord}}, my protector,
deserves praise –
\par }{\Q the one who trains my hands for battle,
\par }{\Q and my fingers for war,
\par }{\Q \VS{2}who loves
me and is my stronghold,
\par }{\Q my refuge
and my deliverer,
\par }{\Q my shield
and the one in whom I take shelter,
\par }{\Q who makes
nations
submit to me.
\par }{\Q \VS{3}O
{\ND{Lord}}, of what importance
is the human
race, that
you should notice
them?
\par }{\Q Of what importance is mankind, that you should be concerned about them?
\par }{\Q \VS{4}People
are like
a vapor,
\par }{\Q their days
like a shadow
that disappears.
\par }{\Q \VS{5}O
{\ND{Lord}}, make the sky
sink
and come down!

\par }{\Q Touch
the mountains
and make them smolder!
\par }{\Q \VS{6}Hurl
lightning
bolts and scatter
them!
\par }{\Q Shoot
your arrows and rout them!
\par }{\Q \VS{7}Reach
down from
above!
\par }{\Q Grab me and rescue
me from the surging
water,
\par }{\Q from the power
of foreigners,
\par }{\Q \VS{8}who
speak
lies,
\par }{\Q and make false promises.
\par }{\Q \VS{9}O God,
I will sing
a new
song
to you!
\par }{\Q Accompanied by a ten-stringed
instrument,
I will sing praises to you,
\par }{\Q \VS{10}the one who delivers
kings,
\par }{\Q and rescued David
his servant
from a deadly
sword.
\par }{\Q \VS{11}Grab
me and rescue
me from the power
of foreigners,
\par }{\Q who
speak
lies,
\par }{\Q and make false promises.
\par }{\Q \VS{12}Then
our sons
will be like plants,
\par }{\Q that quickly grow
to full size.

\par }{\Q Our daughters
will be like corner pillars,
\par }{\Q carved
like
those in a palace.
\par }{\Q \VS{13}Our storehouses
will be full,
\par }{\Q providing
all kinds
of food.

\par }{\Q Our sheep
will multiply
by the thousands
\par }{\Q and fill our pastures.
\par }{\Q \VS{14}Our cattle
will be weighted down
with produce.

\par }{\Q No one
will break through
our walls,
\par }{\Q no one
will be taken captive,
\par }{\Q and there will be no
terrified cries
in our city squares.
\par }{\Q \VS{15}How blessed
are the people
who experience these things!

\par }{\Q How blessed
are the people
whose God
is the
{\ND{Lord}}!


\par }\PsalmChap{145}{\D 
A psalm of praise,
by David.
\par }{\Q \VerseOne{1}I
will extol
you, my God,
O king!
\par }{\Q I will praise
your name
continually!
\par }{\Q \VS{2}Every
day
I
will praise
you!
\par }{\Q I will praise
your name
continually!
\par }{\Q \VS{3}The
{\ND{Lord}}
is great
and certainly worthy
of praise!
\par }{\Q No
one can fathom
his greatness!
\par }{\Q \VS{4}One
generation
will praise
your deeds
to another,
\par }{\Q and tell
about your mighty acts!
\par }{\Q \VS{5}I
will focus on
your honor
and majestic
splendor,
\par }{\Q and your
amazing
deeds!
\par }{\Q \VS{6}They will proclaim
the power
of your awesome
acts!
\par }{\Q I will declare
your great deeds!
\par }{\Q \VS{7}They will talk about the fame
of your great
kindness,
\par }{\Q and sing
about your justice.
\par }{\Q \VS{8}The
{\ND{Lord}}
is merciful
and compassionate;
\par }{\Q he is patient
and demonstrates great
loyal love.
\par }{\Q \VS{9}The
{\ND{Lord}}
is good
to all,
\par }{\Q and has compassion
on
all
he has made.
\par }{\Q \VS{10}All
he has made
will give thanks
to the
{\ND{Lord}}.
\par }{\Q Your loyal followers
will praise you.
\par }{\Q \VS{11}They will proclaim the splendor
of your kingdom;
\par }{\Q they will tell
about your power,
\par }{\Q \VS{12}so that
mankind
might acknowledge
your mighty
acts,
\par }{\Q and the majestic
splendor
of your kingdom.
\par }{\Q \VS{13}Your kingdom
is an eternal
kingdom,
\par }{\Q and your dominion
endures
through
all
generations.
\VS{14}\par }{\Q The
{\ND{Lord}}
supports
all
who fall,
\par }{\Q and lifts
up all
who are bent over.
\par }{\Q \VS{15}Everything
looks
to you in anticipation,
\par }{\Q and you
provide
them
with food
on a
regular basis.
\par }{\Q \VS{16}You open
your hand,
\par }{\Q and fill
every
living
thing with the food they desire.
\par }{\Q \VS{17}The
{\ND{Lord}}
is just
in all
his actions,
\par }{\Q and exhibits love
in all
he does.
\par }{\Q \VS{18}The
{\ND{Lord}}
is near
all
who cry
out to him,
\par }{\Q all
who
cry
out to him sincerely.
\par }{\Q \VS{19}He satisfies the desire
of his loyal followers;
\par }{\Q he hears
their cry for help
and delivers them.
\par }{\Q \VS{20}The
{\ND{Lord}}
protects
those who love
him,
\par }{\Q but he destroys
all
the wicked.
\par }{\Q \VS{21}My mouth
will praise
the {\ND{Lord}}.

\par }{\Q Let all
who live
praise
his holy
name
forever!


\par }\PsalmChap{146}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Praise
the {\ND{Lord}}, O my soul!
\par }{\Q \VS{2}I will praise
the {\ND{Lord}}
as long as I live!
\par }{\Q I will sing praises
to my God
as long as I exist!
\par }{\Q \VS{3}Do not
trust
in princes,
\par }{\Q or in human
beings, who cannot
deliver!
\par }{\Q \VS{4}Their life’s breath
departs,
they return
to the ground;
\par }{\Q on that
day
their plans
die.
\par }{\Q \VS{5}How blessed
is the one whose helper
is the God
of Jacob,
\par }{\Q whose hope
is in
the {\ND{Lord}}
his God,
\par }{\Q \VS{6}the one who made
heaven
and earth,
\par }{\Q the sea,
and all
that
is in them,
\par }{\Q who remains
forever
faithful,
\par }{\Q \VS{7}vindicates
the oppressed,
\par }{\Q and gives
food
to the hungry.
\par }{\Q The
{\ND{Lord}}
releases
the imprisoned.
\par }{\Q \VS{8}The
{\ND{Lord}}
gives sight
to the blind.
\par }{\Q The
{\ND{Lord}}
lifts
up all who are bent
over.

\par }{\Q The
{\ND{Lord}}
loves
the godly.
\par }{\Q \VS{9}The
{\ND{Lord}}
protects
those residing outside their native land;
\par }{\Q he lifts up the fatherless
and the widow,
\par }{\Q but
he opposes
the wicked.
\par }{\Q \VS{10}The
{\ND{Lord}}
rules
forever,
\par }{\Q your God,
O Zion,
throughout
the generations
to come!

\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{147}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}},
\par }{\Q for
it is good
to sing praises
to our God!
\par }{\Q Yes,
praise
is pleasant
and appropriate!
\par }{\Q \VS{2}The
{\ND{Lord}}
rebuilds
Jerusalem,
\par }{\Q and gathers
the exiles of Israel.
\par }{\Q \VS{3}He heals
the brokenhearted,
\par }{\Q and bandages
their wounds.
\par }{\Q \VS{4}He counts
the number
of the stars;
\par }{\Q he names
all of them.
\par }{\Q \VS{5}Our Lord
is great
and has awesome
power;
\par }{\Q there is no
limit
to his wisdom.
\par }{\Q \VS{6}The
{\ND{Lord}}
lifts
up the oppressed,
\par }{\Q but knocks
the wicked
to the ground.
\par }{\Q \VS{7}Offer
to the
{\ND{Lord}}
a song of thanks!

\par }{\Q Sing praises
to our God
to the accompaniment of a harp!
\par }{\Q \VS{8}He covers
the sky
with clouds,
\par }{\Q provides
the earth
with rain,
\par }{\Q and causes grass
to grow
on the hillsides.
\par }{\Q \VS{9}He gives
food
to the animals,
\par }{\Q and to the young
ravens
when they chirp.
\par }{\Q \VS{10}He
is not enamored
with the strength
of a horse,
\par }{\Q nor is he impressed by the warrior’s strong legs.
\par }{\Q \VS{11}The
{\ND{Lord}}
takes delight
in his faithful followers,
\par }{\Q and in those who wait
for his loyal love.
\par }{\Q \VS{12}Extol
the {\ND{Lord}}, O Jerusalem!
\par }{\Q Praise
your God,
O Zion!
\par }{\Q \VS{13}For
he makes the bars
of your gates
strong.
\par }{\Q He blesses
your children
within you.
\par }{\Q \VS{14}He brings
peace
to your territory.
\par }{\Q He abundantly provides
for you the best
grain.
\par }{\Q \VS{15}He sends
his command
through the earth;
\par }{\Q swiftly his order reaches
its destination.
\par }{\Q \VS{16}He sends
the snow
that is white like wool;
\par }{\Q he spreads
the frost
that is white like ashes.
\par }{\Q \VS{17}He throws
his hailstones
like crumbs.
\par }{\Q Who can
withstand
the cold wind he sends?
\par }{\Q \VS{18}He then
orders
it all to melt;

\par }{\Q he
breathes
on it, and the water
flows.
\par }{\Q \VS{19}He proclaims
his word
to Jacob,
\par }{\Q his statutes
and regulations
to Israel.
\par }{\Q \VS{20}He has not
done
so
with any
other nation;
\par }{\Q they are not
aware
of his regulations.
\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{148}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Praise
the {\ND{Lord}}
from
the sky!
\par }{\Q Praise
him in the heavens!
\par }{\Q \VS{2}Praise
him, all
his angels!

\par }{\Q Praise
him, all
his heavenly assembly!
\par }{\Q \VS{3}Praise
him, O sun
and moon!
\par }{\Q Praise
him, all
you shiny
stars!
\par }{\Q \VS{4}Praise
him, O highest
heaven,
\par }{\Q and you waters
above
the sky!
\par }{\Q \VS{5}Let them praise
the name
of the {\ND{Lord}},
\par }{\Q for
he
gave the command
and they came into existence.
\par }{\Q \VS{6}He established
them so they would endure;
\par }{\Q he issued a
decree
that will
not be revoked.
\par }{\Q \VS{7}Praise
the {\ND{Lord}}
from
the earth,
\par }{\Q you sea creatures
and all
you ocean depths,
\par }{\Q \VS{8}O fire
and hail,
snow
and clouds,
\par }{\Q O stormy
wind
that carries
out his orders,
\par }{\Q \VS{9}you mountains
and all
you hills,
\par }{\Q you fruit
trees
and all
you cedars,
\par }{\Q \VS{10}you animals
and all
you cattle,
\par }{\Q you creeping things
and birds,
\par }{\Q \VS{11}you kings
of the earth
and all
you nations,
\par }{\Q you princes
and all
you leaders
on the earth,
\par }{\Q \VS{12}you young men
and young women,
\par }{\Q you elderly,
along with
you children!
\par }{\Q \VS{13}Let them praise
the name
of the {\ND{Lord}},
\par }{\Q for
his name
alone
is exalted;
\par }{\Q his majesty
extends over
the earth
and sky.
\par }{\Q \VS{14}He has made his people
victorious,
\par }{\Q and given all
his loyal followers
reason to praise
–
\par }{\Q the Israelites,
the people
who are close
to him.

\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{149}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Sing
to the
{\ND{Lord}}
a new
song!
\par }{\Q Praise
him in the assembly
of the godly!
\par }{\Q \VS{2}Let
Israel
rejoice
in their Creator!
\par }{\Q Let the people of Zion
delight
in their king!
\par }{\Q \VS{3}Let them praise
his name
with dancing!
\par }{\Q Let them sing praises
to him to the accompaniment of the tambourine
and harp!
\par }{\Q \VS{4}For
the {\ND{Lord}}
takes delight
in his people;
\par }{\Q he exalts
the oppressed
by delivering them.
\par }{\Q \VS{5}Let the godly
rejoice
because of their vindication!

\par }{\Q Let them shout
for joy upon
their beds!
\par }{\Q \VS{6}May they praise
God
\par }{\Q while they hold a two-edged
sword
in their hand,
\par }{\Q \VS{7}in order
to take revenge
on the nations,
\par }{\Q and punish
foreigners.
\par }{\Q \VS{8}They bind
their kings
in chains,
\par }{\Q and their nobles
in iron
shackles,
\par }{\Q \VS{9}and execute
the judgment
to which their enemies have been sentenced.
\par }{\Q All
his loyal followers
will be vindicated.

\par }{\Q Praise
the
{\ND{Lord}}!


\par }\PsalmChap{150}{\PP \par }{\Q \VerseOne{1}Praise
the
{\ND{Lord}}!
\par }{\Q Praise
God
in his sanctuary!
\par }{\Q Praise
him in the sky,
which testifies to his strength!
\par }{\Q \VS{2}Praise
him for his mighty
acts!
\par }{\Q Praise
him for his surpassing
greatness!
\par }{\Q \VS{3}Praise
him with the blast
of the horn!
\par }{\Q Praise
him with the lyre
and the harp!
\par }{\Q \VS{4}Praise
him with the tambourine
and with dancing!
\par }{\Q Praise
him with stringed
instruments and the flute!
\par }{\Q \VS{5}Praise
him with loud
cymbals!
\par }{\Q Praise
him with clanging
cymbals!
\par }{\Q \VS{6}Let everything
that has breath
praise
the
{\ND{Lord}}!
\par }{\PP Praise
the
{\ND{Lord}}!

\par }