\NormalFont\ShortTitle{Revelation}
{\MT Revelation

\par }\ChapOne{1}{\SH The Prologue
\par }{\PP \VerseOne{1}The revelation
of Jesus
Christ,
which
God
gave
him
to show
his
servants
what
must
happen
very soon.
He
made
it clear
by sending
his
angel
to his
servant
John,
\VS{2}who
then testified
to everything
that he saw
concerning the word
of God
and
the testimony
about Jesus
Christ.
\VS{3}Blessed
is the one who reads
the words
of this prophecy
aloud, and blessed are those who hear
and
obey
the things written
in
it,
because
the time
is near!
\par }{\PP \VS{4}From John,
to the seven
churches
that are in
the province of Asia: Grace
and
peace
to you
from
“he who is,”
and
who was,
and
who is still to come,
and
from
the seven
spirits
who
are before
his
throne,
\VS{5}and
from
Jesus
Christ
– the faithful
witness,
the firstborn
from among the dead,
the ruler
over the kings
of the earth.
To the one who loves
us
and
has set
us
free
from
our
sins
at
the cost of his
own blood
\VS{6}and
has appointed
us
as a kingdom,
as priests
serving his
God
and
Father –
to him
be the glory
and
the power
for
ever and ever!
Amen.
\par }{\Q \VS{7}(Look!
{\IT{
He is returning
with
the clouds}},
\par }{\Q and
{\IT{
every
eye
will see
him,}}
\par }{\Q {\IT{
even}}
{\IT{
those who
pierced
him}},

\par }{\Q and
all
the tribes
on the earth
will mourn
because of
him.
\par }{\Q This will certainly come to pass!
Amen.)
\par }{\PP \VS{8}“I
am
the Alpha
and
the Omega,”
says
the Lord
God –
the one who is,
and
who was,
and
who is still to come –
the All-Powerful!
\par }{\PP \VS{9}I,
John,
your
brother
and
the one who shares
with you in
the persecution,
kingdom,
and
endurance
that are in
Jesus,
was
on
the island
called
Patmos
because of
the word
of God
and
the testimony
about Jesus.
\VS{10}I was
in
the Spirit
on
the Lord’s
Day
when
I heard
behind
me
a loud
voice
like
a trumpet,
\VS{11}saying: “Write
in
a book
what
you see
and
send
it to the seven
churches
– to
Ephesus,
Smyrna,
Pergamum,
Thyatira,
Sardis,
Philadelphia,
and
Laodicea.”
\par }{\PP \VS{12}I turned
to see
whose voice
was
speaking
to
me,
and
when
I did
so, I saw
seven
golden
lampstands,
\VS{13}and
in
the midst
of the lampstands
was one
{\IT{
like
a son
of man}}.
He was dressed in
a robe extending down to
his feet
and
he wore a wide
golden
belt
around
his chest.
\VS{14}His
head
and
hair
were as white
as
wool,
even as white
as
snow,
and
his
eyes
were like
a fiery
flame.
\VS{15}His
feet
were like
polished bronze
refined
in
a furnace,
and
his
voice
was like
the roar
of many
waters.
\VS{16}He held
seven
stars
in
his
right
hand,
and
a sharp
double-edged
sword
extended out
of
his
mouth.
His
face shone
like
the sun
shining
at
full strength.
\VS{17}When
I saw
him
I fell down
at
his
feet
as though
I were dead,
but
he placed
his
right hand
on
me
and said: “Do
not
be afraid! I
am
the first
and
the last,
\VS{18}and
the one who lives! I was
dead,
but
look,
now I am
alive
– forever
and ever
– and
I hold
the keys
of death
and
of Hades!
\VS{19}Therefore
write
what
you saw,
what
is,
and
what
will
be
after
these things.
\VS{20}The
mystery
of the seven
stars
that
you saw
in
my
right hand
and
the seven
golden
lampstands
is this: The seven
stars
are
the angels
of the
seven
churches
and
the seven
lampstands
are
the seven
churches.

\par }\Chap{2}{\PP \VerseOne{1}“To the angel
of the church
in
Ephesus,
write
the following:
\par }{\PP “This is the solemn pronouncement
of the one who has a firm grasp
on the seven
stars
in
his
right hand –
the one who walks
among
the seven
golden
lampstands:
\VS{2}‘I know
your
works
as well as
your
labor
and
steadfast endurance,
and
that
you cannot
tolerate
evil.
You have
even
put to the test
those who refer
to themselves
as apostles
(but
are
not), and
have discovered
that they
are false.
\VS{3}I am also aware that you have persisted steadfastly,
endured
much for the sake of
my
name,
and
have
not
grown weary.
\VS{4}But
I have
this against
you: You have departed
from your
first
love!
\VS{5}Therefore,
remember
from what high state
you have fallen
and
repent! Do the deeds
you did
at the first;
if
not,
I will come
to you
and
remove
your
lampstand
from
its
place
– that is, if
you do
not
repent.
\VS{6}But
you do have
this
going for you: You hate
what the Nicolaitans
practice –
practices
I also
hate.
\VS{7}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.
To the one who conquers,
I will permit
him
to eat
from
the tree
of life
that
is
in
the paradise
of God.’
\par }{\SH To the Church in Smyrna
\par }{\PP \VS{8}“To the angel
of the church
in
Smyrna
write
the following:
\par }{\PP “This is the solemn pronouncement
of the one who is the first
and
the last,
the one who
was
dead,
but
came to life:
\VS{9}‘I know
the distress you
are suffering
and
your poverty
(but
you are
rich). I also
know the slander
against you by
those who call
themselves
Jews
and
really are
not,
but
are
a synagogue
of Satan.
\VS{10}Do
not
be afraid
of the things
you are about
to suffer.
The devil
is about
to have some of
you
thrown
into
prison
so
you may be tested,
and
you will experience
suffering
for ten
days.
Remain
faithful
even to the point
of death,
and
I will give
you
the crown
that is life itself.
\VS{11}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.
The one who conquers
will
in no way
be harmed
by
the second
death.’
\par }{\SH To the Church in Pergamum
\par }{\PP \VS{12}“To the angel
of the church
in
Pergamum
write
the following:
\par }{\PP “This is the solemn pronouncement
of the one who has
the sharp
double-edged
sword:
\VS{13}‘I know
where
you live
– where
Satan’s
throne
is. Yet
you continue to cling
to my
name
and
you have
not
denied
your faith
in me,
even
in
the days
of Antipas,
my
faithful
witness,
who
was killed
in
your
city where
Satan
lives.
\VS{14}But
I have
a few things
against
you: You
have
some people there
who follow
the teaching
of Balaam,
who
instructed
Balak
to put
a stumbling block
before
the people
of Israel
so they would eat
food sacrificed to idols
and
commit sexual immorality.
\VS{15}In the same way,
there are also some among you
who follow
the teaching
of the Nicolaitans.
\VS{16}Therefore,
repent! If
not,
I will come
against you
quickly
and
make war
against
those people
with
the sword
of my
mouth.
\VS{17}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.
To the one who conquers,
I will give
him
some of the hidden
manna,
and
I will give
him
a white
stone,
and
on
that stone
will be written
a new
name
that
no one
can understand
except
the one who receives it.’
\par }{\SH To the Church in Thyatira
\par }{\PP \VS{18}“To the angel
of the church
in
Thyatira
writethe following:
\par }{\PP “
This
is the solemn pronouncement
of the Son
of God,
the one who has
eyes
like
a fiery
flame
and
whose
feet
are like
polished bronze:
\VS{19}‘I know
your
deeds: your
love,
faith,
service,
and
steadfast endurance.
In fact, your
more recent
deeds
are greater than
your earlier ones.
\VS{20}But
I have
this against
you: You tolerate
that
woman
Jezebel,
who calls
herself
a prophetess,
and
by her teaching
deceives
my
servants
to commit sexual immorality
and
to eat
food sacrificed to idols.
\VS{21}I have given
her
time
to
repent,
but
she is
not
willing
to repent
of
her
sexual immorality.
\VS{22}Look! I am throwing
her
onto
a bed of violent illness,
and
those who commit adultery
with
her
into
terrible
suffering,
unless
they repent
of
her
deeds.
\VS{23}Furthermore,
I will strike
her
followers
with
a deadly disease,
and then
all
the churches
will know
that
I
am
the one who searches
minds
and
hearts.
I will repay
each one
of you
what
your
deeds deserve.
\VS{24}But
to the rest
of you
in
Thyatira,
all
who do
not
hold
to this
teaching
(who
have
not
learned
the so-called “deep secrets
of Satan”), to you I say: I do
not
put
any additional
burden
on
you.
\VS{25}However,
hold on
to what
you have
until
I come.
\VS{26}And
to the one who conquers
and
who continues
in my
deeds
until
the end,
I will give
him
authority
over
the nations –
\par }{\Q \VS{27}{\QT{
he will rule
them
with
an iron
rod}}*
\par }{\Q and like
clay
jars
he will break them to pieces,
\par }{\PP \VS{28}just as I have received the right to rule from my Father – and I will give him the morning star.
\VS{29}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.’

\par }\Chap{3}{\PP \VerseOne{1}“To the angel
of the church
in
Sardis
write the following:
\par }{\PP “This
is the solemn pronouncement
of the one who holds
the seven
spirits
of God
and
the seven
stars: ‘I know
your
deeds,
that
you have
a reputation
that
you are alive,
but
in reality you are
dead.
\VS{2}Wake up
then, and
strengthen
what remains
that
was about
to die,
because
I have
not
found
your
deeds
complete
in the sight of
my
God.
\VS{3}Therefore,
remember
what
you received
and
heard,
and
obey
it, and
repent.
If
you do
not
wake up,
I will come
like
a thief,
and
you will
never
know
at what
hour
I will come
against
you.
\VS{4}But
you have
a few
individuals
in
Sardis
who
have
not
stained
their
clothes,
and
they will walk
with
me
dressed in
white,
because
they are
worthy.
\VS{5}The one who conquers
will be dressed
like them in
white
clothing,
and
I will
never
erase
his
name
from
the book
of life,
but
will declare
his
name
before
my
Father
and
before
his
angels.
\VS{6}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.’
\par }{\SH To the Church in Philadelphia
\par }{\PP \VS{7}“To the angel
of the church
in
Philadelphia
write the following:
\par }{\PP “This
is the solemn pronouncement
of the Holy One,
the True One,
who holds
the key
of David,
who opens
doors no one
can shut,
and
shuts
doors no one
can open:
\VS{8}‘I know
your
deeds.
(Look! I have put
in front of
you
an open
door
that
no one
can
shut.) I know that
you have
little
strength,
but
you have obeyed
my
word
and
have
not
denied
my
name.
\VS{9}Listen! I am going to make
those people from
the synagogue
of Satan
– who say
they are Jews
yet are
not,
but
are lying
– Look,
I will make
them
come
and
bow down
at
your
feet
and
acknowledge
that
I
have loved
you.
\VS{10}Because
you have kept
my
admonition
to endure steadfastly,
I
will
also
keep
you
from
the hour
of testing
that is about
to come
on
the whole
world
to test
those who live
on
the earth.
\VS{11}I am coming
soon.
Hold on
to what
you have
so that
no one
can take away
your
crown.
\VS{12}The one who conquers
I will make
a pillar
in
the
temple
of my
God,
and
he will
never
depart
from it. I will write
on
him
the
name
of my
God
and
the
name
of the
city
of my
God
(the
new
Jerusalem
that comes down
out of
heaven
from
my
God), and
my
new
name as well.
\VS{13}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.’
\par }{\SH To the Church in Laodicea
\par }{\PP \VS{14}“To the angel
of the church
in
Laodicea
write
the following:
\par }{\PP “This is the solemn pronouncement
of the Amen,
the faithful
and
true
witness,
the originator of God’s creation:
\VS{15}‘I know
your
deeds,
that
you are
neither
cold
nor
hot.
I wish
you were
either cold
or
hot!
\VS{16}So
because
you are
lukewarm,
and
neither
hot
nor
cold,
I am going to
vomit
you
out of
my
mouth!
\VS{17}Because
you say,
“I am
rich
and
have acquired great wealth,
and
need
nothing,”
but
do
not
realize
that
you
are
wretched,
pitiful,
poor,
blind,
and
naked,
\VS{18}take my advice
and buy
gold
from
me
refined
by
fire
so
you can become rich! Buy from me white
clothing
so
you can be clothed
and
your
shameful
nakedness
will
not
be exposed,
and buy eye salve
to put on
your
eyes
so
you can see!
\VS{19}All those
I love,
I rebuke
and
discipline.
So
be earnest
and
repent!
\VS{20}Listen! I am standing
at
the door
and
knocking! If
anyone
hears
my
voice
and
opens
the door
I will come
into
his home
and
share a meal
with
him,
and
he
with
me.
\VS{21}I will grant
the one
who conquers
permission to sit
with
me
on
my
throne,
just as
I too
conquered
and
sat down
with
my
Father
on
his
throne.
\VS{22}The one who has
an ear
had better hear
what
the Spirit
says
to the churches.’ ”

\par }\Chap{4}{\PP \VerseOne{1}After
these things
I looked,
and
there was a door
standing open
in
heaven! And
the first
voice
I had heard
speaking
to me
like
a trumpet
said: “Come up
here
so
that I can show
you
what
must
happen
after
these things.”
\VS{2}Immediately
I was
in
the Spirit,
and
a throne
was standing
in
heaven
with someone seated
on
it!
\VS{3}And
the one seated
on it was like
jasper
and
carnelian
in appearance,
and
a rainbow
looking
like
it was made of emerald
encircled
the throne.
\VS{4}In a circle around
the throne
were twenty-four
other thrones,
and
seated
on
those thrones
were twenty-four
elders.
They were dressed
in white
clothing
and
had golden
crowns
on
their
heads.
\VS{5}From
the throne
came out
flashes of lightning
and roaring
and
crashes of thunder.
Seven
flaming
torches,
which
are
the seven
spirits
of God,
were burning
in front of
the throne
\VS{6}and
in front of
the throne
was something like
a sea
of glass,
like
crystal.
\par }{\PP In
the middle
of the throne
and
around
the throne
were four
living creatures
full
of eyes
in front
and
in back.
\VS{7}The first
living creature
was like
a lion,
the second
creature
like
an ox,
the third
creature
had
a face
like
a man’s,
and
the fourth
creature
looked like
an eagle
flying.
\VS{8}Each
one
of the four
living creatures
had
six
wings
and was full
of eyes
all around
and
inside.
They
never
rest
day
or
night,
saying:
\par }{\Q “{\QT{
Holy
Holy
Holy
is the Lord
God,
the All-Powerful}},
\par }{\Q Who was
and
who is,
and
who is still to come!”
\par }{\PP \VS{9}And
whenever
the living creatures
give
glory,
honor,
and
thanks
to the one who sits
on
the throne,
who lives
forever
and ever,
\VS{10}the twenty-four
elders
throw
themselves to the ground
before
the one who sits
on
the throne
and
worship
the one who lives
forever
and ever,
and
they offer
their
crowns
before
his throne,
saying:
\par }{\Q \VS{11}“You are
worthy,
our
Lord
and
God,
\par }{\Q to receive
glory
and
honor
and
power,
\par }{\Q since
you
created
all things,
\par }{\Q and
because of
your
will
they existed
and
were created!”

\par }\Chap{5}{\PP \VerseOne{1}Then
I saw
in
the right hand
of the one who was seated
on
the throne
a scroll
written
on the front
and
back
and sealed
with seven
seals.
\VS{2}And
I saw
a powerful
angel
proclaiming
in
a loud
voice: “Who
is worthy
to open
the scroll
and
to break
its
seals?”
\VS{3}But
no one
in
heaven
or
on
earth
or
under
the earth
was able
to open
the scroll
or
look into
it.
\VS{4}So
I
began weeping
bitterly
because
no one
was found
who was worthy
to open
the scroll
or
to look into
it.
\VS{5}Then
one
of
the elders
said
to me,
“Stop
weeping! Look,
the Lion
of the tribe
of Judah,
the root
of David,
has conquered;
thus he can open
the scroll
and
its
seven
seals.”
\par }{\PP \VS{6}Then
I saw
standing
in
the middle
of the throne
and
of the four
living creatures,
and
in
the middle
of the elders,
a Lamb
that appeared to have been killed.
He had
seven
horns
and
seven
eyes,
which
are
the seven
spirits
of God
sent out
into
all
the earth.
\VS{7}Then
he came
and
took
the scroll from
the right hand
of the one who was seated
on
the throne,
\VS{8}and
when
he had taken
the scroll,
the four
living creatures
and
the twenty-four
elders
threw
themselves to the ground
before
the Lamb.
Each of them
had
a harp
and
golden
bowls
full
of incense
(which
are
the prayers
of the saints).
\VS{9}They were singing
a new
song:

\par }{\Q “You are
worthy
to take
the scroll
\par }{\Q and
to open
its
seals
\par }{\Q because
you were killed,
\par }{\Q and
at
the cost of your own
blood
you have purchased
for God
\par }{\Q persons from
every
tribe,
language,
people,
and
nation.
\par }{\Q \VS{10}You have appointed
them
as a kingdom
and
priests
to serve our
God,
and
they will reign
on
the earth.”
\par }{\PP \VS{11}Then
I looked
and
heard
the voice
of many
angels
in a circle around
the throne,
as well as the living creatures
and
the elders.
Their
number
was ten thousand times ten thousand – thousands times thousands –
\VS{12}all of whom were singing
in a loud
voice:
\par }{\Q “Worthy
is
the lamb
who was killed
\par }{\Q to receive
power
and
wealth
\par }{\Q and
wisdom
and
might
\par }{\Q and
honor
and
glory
and
praise!”
\par }{\PP \VS{13}Then
I heard
every
creature
– in
heaven,
on
earth,
under
the earth,
in
the sea,
and
all
that is
in
them
– singing:

\par }{\Q “To the one seated
on
the
throne
and
to the
Lamb
\par }{\Q be praise,
honor,
glory,
and
ruling power
forever
and
ever!”
\par }{\PP \VS{14}And
the four
living creatures
were saying
“Amen,”
and
the elders
threw
themselves to the ground
and
worshiped.

\par }\Chap{6}{\PP \VerseOne{1}I looked on
when
the Lamb
opened
one
of
the seven
seals,
and
I heard
one
of
the four
living creatures
saying
with
a thunderous
voice, “Come!”
\VS{2}So
I looked,
and
here came
a white
horse! The one who rode
it
had
a bow,
and
he was given
a crown,
and
as
a conqueror
he rode out
to conquer.
\par }{\PP \VS{3}Then
when
the Lamb opened
the second
seal,
I heard
the second
living creature
saying,
“Come!”
\VS{4}And
another
horse,
fiery red,
came out,
and
the one who rode
it
was granted permission
to take
peace
from
the earth,
so that
people would butcher
one another,
and
he was given
a huge
sword.
\par }{\PP \VS{5}Then
when
the Lamb opened
the third
seal
I heard
the third
living creature
saying,
“Come!” So
I looked,
and
here came
a black
horse! The one who rode
it
had
a balance scale
in
his
hand.
\VS{6}Then
I heard
something like
a voice
from
among
the four
living creatures
saying,
“A quart
of wheat
will cost a day’s pay
and
three
quarts
of barley
will cost a day’s pay.
But
do
not
damage
the olive oil
and
the wine!”
\par }{\PP \VS{7}Then
when
the Lamb opened
the fourth
seal
I heard
the voice
of the fourth
living creature
saying,
“Come!”
\VS{8}So
I looked
and
here came
a pale green
horse! The name
of the one who rode
it
was Death,
and
Hades
followed
right behind.
They
were given
authority
over
a fourth
of the
earth,
to kill
its population with
the sword,
famine,
and
disease,
and
by
the
wild animals
of the
earth.
\par }{\PP \VS{9}Now
when
the Lamb opened
the fifth
seal,
I saw
under
the altar
the souls
of those who had been violently killed
because
of the word
of God
and
because
of the testimony
they had given.
\VS{10}They cried out
with a loud
voice, “How long,
Sovereign Master,
holy
and
true,
before you judge
those who live
on
the earth
and
avenge
our
blood?”
\VS{11}Each
of them
was given
a long
white
robe
and
they
were told
to
rest
for a little
longer,
until
the full number was reached
of both their
fellow servants
and
their
brothers
who were going
to be killed
just as they had been.
\par }{\PP \VS{12}Then
I looked
when
the Lamb opened
the sixth
seal,
and
a huge
earthquake
took place;
the sun
became
as black
as
sackcloth
made of hair,
and
the full
moon
became
blood red;
\VS{13}and
the stars
in the sky
fell
to
the earth
like
a fig tree
dropping
its
unripe figs
when shaken
by
a fierce
wind.
\VS{14}The sky
was split apart
like
a scroll
being rolled up,
and
every
mountain
and
island
was moved
from
its
place.
\VS{15}Then
the kings
of the earth,
the very important people,
the generals,
the rich,
the powerful,
and
everyone,
slave
and
free,
hid
themselves
in
the caves
and
among
the rocks
of the mountains.
\VS{16}They said
to the mountains
and
to the rocks,
“Fall
on
us
and
hide
us
from
the face
of the one who is seated
on
the throne
and
from
the wrath
of the Lamb,
\VS{17}because
the great
day
of their
wrath
has come,
and
who
is able
to withstand it?”

\par }\Chap{7}{\PP \VerseOne{1}After
this
I saw
four
angels
standing
at
the four
corners
of the earth,
holding back
the four
winds
of the earth
so
no
wind
could blow
on
the earth,
on
the sea,
or
on
any
tree.
\VS{2}Then
I saw
another
angel
ascending
from
the east,
who had
the seal
of the living
God.
He
shouted out
with a loud
voice
to the four
angels
who
had been given permission
to damage
the earth
and
the sea:
\VS{3}“Do
not
damage
the earth
or
the sea
or
the trees
until
we have put a seal
on
the foreheads
of the servants
of our
God.”
\VS{4}Now
I heard
the number
of those who were marked
with the seal, one hundred
and forty-four
thousand,
sealed
from
all
the tribes
of
the people
of Israel:
\par }{\Q \VS{5}From
the tribe
of Judah,
twelve
thousand
were sealed,
\par }{\Q from
the tribe
of Reuben,
twelve
thousand,
\par }{\Q from
the tribe
of Gad,
twelve
thousand,
\par }{\Q \VS{6}from
the tribe
of Asher,
twelve
thousand,
\par }{\Q from
the tribe
of Naphtali,
twelve
thousand,
\par }{\Q from
the tribe
of Manasseh,
twelve
thousand,
\par }{\Q \VS{7}from
the tribe
of Simeon,
twelve
thousand,
\par }{\Q from
the tribe
of Levi,
twelve
thousand,
\par }{\Q from
the tribe
of Issachar,
twelve
thousand,
\par }{\Q \VS{8}from
the tribe
of Zebulun,
twelve
thousand,
\par }{\Q from
the tribe
of Joseph,
twelve
thousand,
\par }{\Q from
the tribe
of Benjamin,
twelve
thousand
were sealed.
\par }{\PP \VS{9}After
these things
I looked,
and
here
was an enormous
crowd
that
no one
could
count,
made up of persons from
every
nation,
tribe,
people,
and
language,
standing
before
the throne
and
before
the Lamb
dressed
in long
white
robes,
and
with palm branches
in
their
hands.
\VS{10}They were shouting out
in a loud
voice,
\par }{\Q “Salvation
belongs to our
God,
\par }{\Q to the one seated
on
the throne,
and
to the Lamb!”
\par }{\PP \VS{11}And
all
the angels
stood
there in a circle around
the throne
and
around the elders
and
the four
living creatures,
and
they threw
themselves down
with
their
faces
to the ground
before
the throne
and
worshiped
God,
\VS{12}saying,
\par }{\Q “Amen! Praise
and
glory,
\par }{\Q and
wisdom
and
thanksgiving,
\par }{\Q and
honor
and
power
and
strength
\par }{\Q be to our
God
for
ever
and ever.
Amen!”
\par }{\PP \VS{13}Then
one
of the elders
asked
me,
“These
dressed in
long
white
robes
– who
are they
and
where
have they come from?”
\VS{14}So
I said
to him,
“My
lord,
you
know
the answer.” Then
he said
to me,
“These
are
the ones who have come
out of
the great
tribulation.
They have washed
their
robes
and
made
them
white
in
the blood
of the Lamb!
\VS{15}For
this reason
they are
before
the throne
of God,
and
they serve
him
day
and
night
in
his
temple,
and
the one seated
on
the throne
will shelter
them.
\VS{16}{\IT{
They will
never
go hungry
or
be thirsty
again,
and the sun
will not
beat down
on
them,
nor
any
burning heat}},
\VS{17}because
the Lamb
in the middle
of the throne
will shepherd
them
and
lead
them
to springs
of
living
water,
{\IT{
and
God
will wipe away
every
tear
from
their
eyes}}.”

\par }\Chap{8}{\PP \VerseOne{1}Now
when
the Lamb opened
the seventh
seal
there was
silence
in
heaven
for about
half an hour.
\VS{2}Then
I saw
the seven
angels
who
stand
before
God,
and
seven
trumpets
were given
to them.
\VS{3}Another
angel
holding
a golden
censer
came
and
was stationed
at
the altar.
A large amount
of incense
was given
to him
to
offer up,
with the prayers
of all
the saints,
on
the golden
altar
that is before
the throne.
\VS{4}The smoke
coming from
the incense,
along with the prayers
of the saints,
ascended before
God
from
the angel’s
hand.
\VS{5}Then
the angel
took
the censer,
filled
it
with
fire
from the altar,
and
threw
it on
the earth,
and
there were
crashes of thunder,
roaring,
flashes of lightning,
and
an earthquake.
\par }{\PP \VS{6}Now
the seven
angels
holding
the seven
trumpets
prepared
to
blow
them.
\par }{\PP \VS{7}The first angel
blew
his trumpet,
and
there was
hail
and
fire
mixed
with
blood,
and
it was thrown
at
the earth
so that
a third
of the earth
was burned up,
a third
of the trees
were burned up,
and
all
the green
grass
was burned up.
\par }{\PP \VS{8}Then
the second
angel
blew
his trumpet,
and
something like
a great
mountain
of burning
fire
was thrown
into
the sea.
A third
of the sea
became
blood,
\VS{9}and
a third
of the creatures
living
in
the sea
died,
and
a third
of the ships
were completely destroyed.
\par }{\PP \VS{10}Then
the third
angel
blew
his trumpet,
and
a huge
star
burning
like
a torch
fell
from
the sky;
it landed
on
a third
of the rivers
and
on
the springs
of water.
\VS{11}(Now
the name
of the star
is
Wormwood.) So
a third
of the waters
became
wormwood,
and
many
people
died
from
these waters
because
they were poisoned.
\par }{\PP \VS{12}Then
the fourth
angel
blew
his trumpet,
and
a third
of the
sun
was struck,
and
a third
of the
moon,
and
a third
of the
stars,
so that
a third
of them
were darkened.
And
there was no
light
for a third of the day
and for a third
of the
night
likewise.
\VS{13}Then
I looked,
and
I heard
an
eagle
flying
directly overhead,
proclaiming
with a loud
voice,
“Woe! Woe! Woe
to those who live
on
the earth
because of
the remaining
sounds
of the trumpets
of the three
angels
who are about
to blow them!”

\par }\Chap{9}{\PP \VerseOne{1}Then
the fifth
angel
blew
his trumpet,
and
I saw
a star
that had fallen
from
the sky
to
the earth,
and
he was given
the key
to the shaft
of the abyss.
\VS{2}He
opened
the shaft
of the abyss
and
smoke
rose out
of
it
like
smoke
from a giant
furnace.
The sun
and
the air
were darkened
with
smoke
from the shaft.
\VS{3}Then
out
of the smoke
came
locusts
onto
the earth,
and
they
were given
power
like
that of the scorpions
of the earth.
\VS{4}They
were told
not
to damage
the grass
of the earth,
or
any
green plant
or
tree,
but
only
those people
who
did
not
have
the seal
of God
on
their forehead.
\VS{5}The locusts
were
not
given permission
to
kill
them,
but
only to
torture
them for five
months,
and
their
torture
was like
that
of a scorpion
when
it stings
a person.
\VS{6}In
those
days
people
will seek
death,
but
will
not
be able to find
it;
they will long
to die,
but
death
will flee
from
them.
\par }{\PP \VS{7}Now
the locusts
looked
like
horses
equipped
for
battle.
On
their
heads
were something like
crowns
similar to
gold,
and
their
faces
looked like
men’s
faces.
\VS{8}They had
hair
like
women’s
hair,
and
their
teeth
were
like
lions’ teeth.
\VS{9}They had
breastplates
like
iron
breastplates,
and
the sound
of their
wings
was like
the noise
of many
horse-drawn
chariots
charging
into
battle.
\VS{10}They have
tails
and
stingers
like
scorpions,
and
their
ability
to injure
people
for five
months
is in
their
tails.
\VS{11}They have
as king
over
them
the angel
of the abyss,
whose
name
in Hebrew
is
{\IT{
Abaddon}}, and
in
Greek,
{\IT{
Apollyon}}.
\par }{\PP \VS{12}The first
woe
has passed,
but two
woes
are still
coming
after
these things!
\par }{\PP \VS{13}Then
the sixth
angel
blew
his trumpet,
and
I heard
a single
voice
coming from
the horns
on the golden
altar
that is before
God,
\VS{14}saying
to the sixth
angel,
the one holding
the trumpet,
“Set free
the four
angels
who are bound
at
the great
river
Euphrates!”
\VS{15}Then
the four
angels
who had been prepared
for
this hour,
day,
month,
and
year
were set free
to
kill
a third
of humanity.
\VS{16}The number
of soldiers
on horseback
was two hundred
million;
I heard
their
number.
\VS{17}Now this is what the horses
and
their
riders
looked like
in
my vision: The riders had
breastplates
that were fiery red, dark blue,
and
sulfurous
yellow
in color. The heads
of the horses
looked like
lions’
heads,
and
fire,
smoke,
and
sulfur
came out
of
their
mouths.
\VS{18}A third
of humanity
was killed
by
these
three
plagues,
that is, by
the fire,
the smoke,
and
the sulfur
that came out
of
their
mouths.
\VS{19}For
the power
of the horses
resides
in
their
mouths
and
in
their
tails,
because
their
tails
are like
snakes,
having
heads
that inflict injuries.
\VS{20}The rest
of humanity,
who had
not
been killed
by
these
plagues,
did
not
repent
of
the works
of their
hands,
so that
they did
not stop
worshiping
demons
and
idols
made of gold,
silver,
bronze,
stone,
and
wood
– idols
that cannot
see
or
hear
or
walk about.
\VS{21}Furthermore,
they did
not
repent
of
their
murders,
of
their
magic spells,
of
their
sexual immorality,
or
of
their
stealing.

\par }\Chap{10}{\PP \VerseOne{1}Then
I saw
another
powerful
angel
descending
from
heaven,
wrapped
in a cloud,
with a rainbow
above
his
head;
his
face
was like
the sun
and
his
legs
were like
pillars
of fire.
\VS{2}He
held
in
his
hand
a little scroll
that was open,
and
he put
his
right
foot
on
the sea
and
his left
on
the land.
\VS{3}Then
he shouted
in a loud
voice
like
a lion
roaring,
and
when
he shouted,
the seven
thunders
sounded
their
voices.
\VS{4}When
the seven
thunders
spoke,
I was preparing
to write,
but
just then I heard
a voice
from
heaven
say,
“Seal up
what
the seven
thunders
spoke
and
do
not
write
it
down.”
\VS{5}Then
the angel
I saw
standing
on
the sea
and
on
the land
raised
his
right
hand
to
heaven
\VS{6}and
swore
by
the one who lives
forever
and ever,
who
created
heaven
and
what is in
it,
and
the earth
and
what is in
it,
and
the
sea
and
what is in
it,
“There will be
no more
delay!
\VS{7}But
in
the days
when
the seventh
angel
is about
to blow
his trumpet,
the mystery
of God
is completed,
just as
he has proclaimed
to his
servants
the prophets.”
\VS{8}Then
the voice
I had heard
from
heaven
began to speak
to
me
again, “Go
and take
the open
scroll
in
the hand
of the angel
who is standing
on
the sea
and
on
the land.”
\VS{9}So
I went
to
the angel
and asked
him
to give
me
the little scroll.
He said
to me,
“Take
the scroll and eat
it.
It will make
your
stomach
bitter,
but
it will be
as sweet as
honey
in
your
mouth.”
\VS{10}So
I took
the little scroll
from
the angel’s
hand
and
ate
it,
and
it did taste as sweet
as
honey
in
my
mouth,
but
when
I had eaten
it,
my
stomach
became bitter.
\VS{11}Then
they told
me: “You
must
prophesy
again
about
many
peoples,
nations,
languages,
and
kings.”

\par }\Chap{11}{\PP \VerseOne{1}Then
a measuring rod
like
a staff
was given
to me,
and I was told, “Get up
and
measure
the temple
of God,
and
the altar,
and
the ones who worship
there.
\VS{2}But
do
not
measure
the outer
courtyard
of the temple;
leave
it
out,
because
it has been given
to the Gentiles,
and
they will trample
on the holy
city
for forty-two
months.
\VS{3}And
I will grant
my
two
witnesses
authority to prophesy
for 1,260
days,
dressed
in sackcloth.
\VS{4}(These
are
the two
olive trees
and
the two
lampstands
that stand
before
the Lord
of the earth.)
\VS{5}If
anyone
wants
to harm
them,
fire
comes out
of
their
mouths
and
completely consumes
their
enemies.
If
anyone
wants
to harm
them,
they must
be killed
this way.
\VS{6}These
two have
the power
to close up
the sky
so
that it does
not
rain
during the time
they
are prophesying.
They have
power
to turn
the waters
to
blood
and
to strike
the earth
with
every
kind of plague
whenever
they want.
\VS{7}When
they have completed
their
testimony,
the beast
that comes up
from
the abyss
will make
war
on
them
and
conquer
them
and
kill
them.
\VS{8}Their
corpses
will lie in
the street
of the great
city
that
is
symbolically
called
Sodom
and
Egypt,
where
their
Lord
was
also
crucified.
\VS{9}For three
and
a half
days
those from
every people,
tribe,
nation,
and
language
will look at
their
corpses,
because they will
not
permit
them to be placed
in
a tomb.
\VS{10}And
those who live
on
the earth
will rejoice
over
them
and
celebrate,
even
sending
gifts
to each other,
because
these
two
prophets
had tormented
those who live
on
the earth.
\VS{11}But
after
three
and
a half
days
a breath
of life
from
God
entered
them,
and
they stood
on
their
feet,
and
tremendous
fear
seized
those who were watching
them.
\VS{12}Then
they heard
a loud
voice
from
heaven
saying
to them: “Come up
here!” So
the two prophets went up
to
heaven
in
a cloud
while their
enemies
stared at
them.
\VS{13}Just then
a major
earthquake
took place
and
a tenth
of the city
collapsed;
seven
thousand
people
were killed
in
the earthquake,
and
the rest
were terrified
and
gave
glory
to the God
of heaven.
\par }{\PP \VS{14}The second
woe
has come and gone;
the third
is coming
quickly.
\par }{\SH The Seventh Trumpet
\par }{\PP \VS{15}Then
the seventh
angel
blew
his trumpet,
and
there were
loud
voices
in
heaven
saying:
\par }{\Q “The kingdom
of the world
\par }{\Q has become
the kingdom of our
Lord
\par }{\Q and
of his
Christ,
\par }{\Q and
he will reign
for
ever
and ever.”
\par }{\PP \VS{16}Then
the twenty-four
elders
who are seated
on
their
thrones
before
God
threw
themselves down
with
their
faces
to the ground
and
worshiped
God
\VS{17}with these words:

\par }{\Q “We give
you
thanks,
Lord
God,
the All-Powerful,
\par }{\Q the one who is
and
who was,
\par }{\Q because
you have taken
your
great
power
\par }{\Q and
begun to reign.
\par }{\Q \VS{18}The nations
were enraged,
\par }{\Q but
your
wrath
has come,
\par }{\Q and
the
time
has come for the
dead
to be judged,
\par }{\Q and
the time has come to give
to your
servants,
\par }{\Q the
prophets,
their reward,
\par }{\Q as well as to the saints
\par }{\Q and
to those who revere
your
name,
both small
and
great,
\par }{\Q and
the time has come to destroy
those who destroy
the earth.”
\par }{\PP \VS{19}Then
the temple
of God
in
heaven
was opened
and
the ark
of his
covenant
was visible
within
his
temple.
And
there were
flashes of lightning,
roaring,
crashes of thunder,
an earthquake,
and
a great
hailstorm.

\par }\Chap{12}{\PP \VerseOne{1}Then
a great
sign
appeared
in
heaven: a woman
clothed with
the sun,
and
with the moon
under
her
feet,
and
on
her
head
was a crown
of twelve
stars.
\VS{2}She was
pregnant
and
was screaming
in labor pains,
struggling
to give birth.
\VS{3}Then
another
sign
appeared
in
heaven: a
huge
red
dragon
that had
seven
heads
and
ten
horns,
and
on
its
heads
were seven
diadem crowns.
\VS{4}Now
the dragon’s
tail
swept away
a third
of the stars
in heaven
and
hurled
them
to
the
earth.
Then
the dragon
stood
before
the woman
who was about
to give birth,
so that
he might devour
her
child
as soon as
it was born.
\VS{5}So
the woman gave birth
to a son,
a male child,
who
is going
{\IT{
to rule over
all
the nations
with
an iron
rod}}. Her
child
was suddenly caught up
to
God
and
to
his
throne,
\VS{6}and
she
fled
into
the wilderness
where
a place
had been prepared
for her by God,
so
she could be taken care of for 1,260 days.
\par }{\SH War in Heaven
\par }{\PP \VS{7}Then
war
broke out
in
heaven: Michael
and
his
angels
fought
against
the dragon,
and
the dragon
and
his
angels
fought back.
\VS{8}But
the dragon was
not
strong
enough to prevail, so there was no longer
any place
left
in
heaven for him and his angels.
\VS{9}So that huge
dragon
– the ancient
serpent,
the one called
the devil
and
Satan,
who deceives
the whole
world
– was thrown down
to
the earth,
and
his
angels
along with
him.
\VS{10}Then
I heard
a loud
voice
in
heaven
saying,
\par }{\Q “The salvation
and
the power
\par }{\Q and
the
kingdom
of our
God,
\par }{\Q and
the ruling authority
of his
Christ,
have
now
come,
\par }{\Q because
the accuser
of our
brothers and sisters,
\par }{\Q the one who accuses
them
day
and
night
before
our
God,
\par }{\Q has been thrown down.
\par }{\Q \VS{11}But
they
overcame
him
\par }{\Q by
the blood
of the Lamb
\par }{\Q and
by
the word
of their
testimony,
\par }{\Q and
they did
not
love
their
lives
so much that they were afraid to
die.
\par }{\Q \VS{12}Therefore
you
heavens
rejoice,
and
all who reside
in
them!
\par }{\Q But woe
to the earth
and
the sea
\par }{\Q because
the devil
has come down
to
you!
\par }{\Q He is filled
with terrible
anger,
\par }{\Q for he knows
that
he
only has
a little
time!”
\par }{\PP \VS{13}Now
when
the dragon
realized
that
he had been thrown down
to
the earth,
he pursued
the woman
who
had given birth
to the male child.
\VS{14}But
the woman
was given
the two
wings
of a giant
eagle
so that
she could fly out
into
the wilderness,
to
the place
God prepared for her,
where
she is taken care of – away from the presence of the serpent – for a time, times, and half a time.
\VS{15}Then
the serpent
spouted
water
like
a river
out
of his
mouth
after
the woman
in an attempt
to
sweep
her
away by a flood,
\VS{16}but
the earth
came to
her
rescue;
the ground
opened up
and swallowed
the river
that
the dragon
had spewed
from
his
mouth.
\VS{17}So
the dragon
became enraged
at
the woman
and
went away
to make
war
on
the rest
of her
children,
those who keep
God’s
commandments
and
hold
to the testimony
about Jesus. And the dragon stood on the sand of the seashore.

\par }\Chap{13}{\PP \VerseOne{1}Then
I saw
a beast
coming up
out of
the sea.
It had
ten
horns
and
seven
heads,
and
on
its
horns
were ten
diadem crowns,
and
on
its
heads
a blasphemous
name.
\VS{2}Now
the beast
that
I saw
was
like
a leopard,
but
its
feet
were like
a bear’s,
and
its
mouth
was like
a lion’s
mouth.
The dragon
gave
the beast his
power,
his
throne,
and
great
authority to rule.
\VS{3}One
of
the beast’s
heads
appeared to have been killed,
but
the lethal wound
had been healed.
And
the whole
world
followed
the beast
in amazement;
\VS{4}they worshiped
the dragon
because
he had given
ruling authority
to the beast,
and
they worshiped
the beast
too, saying: “Who
is like
the beast?” and
“Who
is able
to make war
against
him?”
\VS{5}The beast
was given
a mouth
speaking
proud
words and
blasphemies,
and
he was permitted
to exercise
ruling authority
for forty-two
months.
\VS{6}So
the beast opened
his
mouth
to
blaspheme
against
God
– to blaspheme
both his
name
and
his
dwelling place,
that is, those who dwell
in
heaven.
\VS{7}The beast
was permitted
to go
to war
against
the saints
and
conquer
them.
He
was given
ruling authority
over
every
tribe,
people,
language,
and
nation,
\VS{8}and
all
those who live
on
the earth
will worship
the beast,
everyone whose name
has
not
been written
since
the foundation
of the world
in
the book
of life
belonging to the Lamb
who was killed.
\VS{9}If
anyone
has
an ear,
he had better listen!
\par }{\Q \VS{10}If
anyone
is meant for
captivity,
\par }{\Q into
captivity
he will go.
\par }{\Q If
anyone
is to be killed
by
the sword,
\par }{\Q then by
the sword
he must
be killed.
\par }{\PP This requires steadfast endurance
and
faith
from the saints.
\par }{\PP \VS{11}Then
I saw
another
beast
coming up
from
the earth.
He
had
two
horns
like
a lamb,
but
was speaking
like
a dragon.
\VS{12}He exercised
all
the ruling authority
of the first
beast
on
his
behalf,
and
made
the earth
and
those who inhabit
it
worship
the first
beast,
the one whose
lethal wound
had been healed.
\VS{13}He performed
momentous
signs,
even
making
fire
come down
from
heaven
in front of
people
\VS{14}and,
by
the signs
he was permitted
to perform
on behalf
of the
beast,
he deceived
those who live
on
the earth.
He told
those who live
on
the earth
to make
an image
to the beast
who
had
been wounded
by the
sword,
but
still lived.
\VS{15}The second beast
was empowered
to give
life
to the image
of the first beast
so that
it could speak,
and
could cause
all
those who did
not
worship
the image
of the beast
to be killed.
\VS{16}He
also
caused
everyone
(small
and
great,
rich
and
poor,
free
and
slave
) to
obtain
a mark
on
their
right
hand
or
on
their
forehead.
\VS{17}Thus
no
one
was allowed
to buy
or
sell
things unless
he bore
the mark
of the beast
– that is, his
name
or
his number.
\VS{18}This calls
for wisdom: Let the one who has
insight
calculate
the beast’s
number,
for
it is
man’s
number,
and
his
number
is 666.

\par }\Chap{14}{\PP \VerseOne{1}Then
I looked,
and
here
was the Lamb
standing
on
Mount
Zion,
and
with
him
were one hundred
and forty-four
thousand,
who had
his
name
and
his
Father’s
name
written
on
their
foreheads.
\VS{2}I
also
heard
a sound
coming out of
heaven
like
the sound
of many
waters
and
like
the sound
of loud
thunder.
Now
the sound
I heard
was like
that made by harpists
playing
their
harps,
\VS{3}and
they were singing
a new
song
before
the throne
and
before
the four
living creatures
and
the elders.
No one
was able
to learn
the song
except
the one hundred
and forty-four
thousand
who had been redeemed
from
the earth.
\par }{\PP \VS{4}These
are
the ones who have
not
defiled
themselves with
women,
for
they are
virgins.
These
are the ones who follow
the Lamb
wherever
he goes.
These
were redeemed
from
humanity
as firstfruits
to God
and
to the Lamb,
\VS{5}and
no
lie
was found
on
their
lips;
they are
blameless.
\par }{\SH Three Angels and Three Messages
\par }{\PP \VS{6}Then
I saw
another
angel
flying
directly
overhead,
and he had
an eternal
gospel
to proclaim
to
those who live
on
the earth
– to
every
nation,
tribe,
language,
and
people.
\VS{7}He declared
in
a loud
voice: “Fear
God
and
give
him
glory,
because
the hour
of his
judgment
has arrived,
and
worship
the one who made
heaven
and
earth,
the sea
and
the springs
of water!”
\par }{\PP \VS{8}A second
angel
followed
the first, declaring: “Fallen,
fallen
is Babylon
the great
city! She made
all
the nations
drink
of
the wine
of her
immoral
passion.”
\par }{\PP \VS{9}A third
angel
followed
the first two,
declaring
in
a loud
voice: “If
anyone
worships
the beast
and
his
image,
and
takes
the mark
on
his
forehead
or
his
hand,
\VS{10}that person
will
also
drink
of
the wine
of God’s
anger
that has been mixed
undiluted
in
the cup
of his
wrath,
and
he will be tortured
with
fire
and
sulfur
in front of
the holy
angels
and
in front of
the Lamb.
\VS{11}And
the smoke
from their
torture
will go up
forever
and
ever,
and those who worship
the beast
and
his
image
will have
no
rest
day
or
night,
along with anyone
who receives
the mark
of his
name.”
\VS{12}This requires
the steadfast endurance
of the saints
– those who obey
God’s
commandments
and
hold to their faith
in Jesus.
\par }{\PP \VS{13}Then
I heard
a voice
from
heaven
say,
“Write
this:
\par }{\Q ‘Blessed
are the dead,
\par }{\Q those who die
in
the Lord
from
this moment
on!’ ”
\par }{\PP “Yes,”
says
the Spirit,
“so
they can rest
from
their
hard work,
because
their
deeds
will follow
them.”
\par }{\PP \VS{14}Then
I looked,
and
a white
cloud
appeared,
and
seated
{\IT{
on
the cloud
was one like
a son
of man!}}
He had
a golden
crown
on
his
head
and
a sharp
sickle
in
his
hand.
\VS{15}Then
another
angel
came out
of
the temple,
shouting
in
a loud
voice
to the one seated
on
the cloud,
“Use
your
sickle
and
start to reap,
because
the time
to reap
has come,
since
the earth’s
harvest
is ripe!”
\VS{16}So the one seated
on
the cloud
swung
his
sickle
over
the earth,
and
the earth
was reaped.
\par }{\PP \VS{17}Then
another
angel
came out
of
the temple
in
heaven,
and
he
too had
a sharp
sickle.
\VS{18}Another
angel,
who was
in charge
of
the fire,
came
from
the altar
and
called
in a loud
voice
to the angel who had
the sharp
sickle,
“Use
your
sharp
sickle
and
gather
the clusters of grapes
off the
vine
of the
earth,
because
its
grapes
are now ripe.”
\VS{19}So
the angel
swung
his
sickle
over
the earth
and
gathered
the grapes
from the vineyard of the earth
and
tossed
them into
the great
winepress
of the wrath
of God.
\VS{20}Then
the winepress
was stomped
outside
the city,
and
blood
poured out
of
the winepress
up to the height of
horses’
bridles
for
a distance
of almost two hundred miles.

\par }\Chap{15}{\PP \VerseOne{1}Then
I saw
another
great
and
astounding
sign
in
heaven: seven
angels
who have
seven
final
plagues
(they are final because
in
them
God’s
anger
is completed).
\par }{\PP \VS{2}Then
I saw
something like
a sea
of glass
mixed
with fire,
and
those who had conquered
the beast
and
his
image
and
the number
of his
name.
They were standing
by
the
sea
of glass,
holding
harps
given to them by God.
\VS{3}They sang
the song
of Moses
the
servant
of God
and
the song
of the Lamb:

\par }{\Q “Great
and
astounding
are your
deeds,
\par }{\Q Lord
God,
the All-Powerful!

\par }{\Q Just
and
true
are your
ways,
\par }{\Q King
over the nations!
\par }{\Q \VS{4}Who
will
not
fear
you, O Lord,
\par }{\Q and
glorify
your
name,
because
you alone
are holy?

\par }{\Q All
nations
will come
and
worship
before
you
\par }{\Q for
your
righteous acts
have been revealed.”
\par }{\PP \VS{5}After
these things
I looked,
and
the temple
(the tent
of the testimony) was opened
in
heaven,
\VS{6}and
the seven
angels
who had
the seven
plagues
came out
of
the temple,
dressed in
clean
bright
linen, wearing
wide golden
belts
around
their chests.
\VS{7}Then
one
of
the four
living creatures
gave
the seven
angels
seven
golden
bowls
filled
with the wrath
of God
who lives
forever
and ever,
\VS{8}and
the temple
was filled
with smoke
from
God’s
glory
and
from
his
power.
Thus
no one
could
enter
the temple
until
the seven
plagues
from the seven
angels
were completed.

\par }\Chap{16}{\PP \VerseOne{1}Then
I heard
a loud
voice
from
the temple
declaring
to the seven
angels: “Go
and
pour out
on
the earth
the seven
bowls
containing God’s
wrath.”
\VS{2}So
the first
angel went
and poured out
his
bowl
on
the earth.
Then
ugly
and
painful
sores
appeared
on
the people
who had
the mark
of the beast
and
who worshiped
his
image.
\par }{\PP \VS{3}Next, the second
angel poured out
his
bowl
on
the sea
and
it turned
into blood,
like
that of a corpse,
and
every
living
creature
that was in
the sea
died.
\par }{\PP \VS{4}Then
the third
angel poured out
his
bowl
on
the rivers
and
the springs
of water,
and
they turned into
blood.
\VS{5}Now
I heard
the angel
of the waters
saying:

\par }{\Q “You are
just –
the one who is
and
who was,
\par }{\Q the Holy One
– because
you have passed
these
judgments,
\par }{\Q \VS{6}because
they poured out
the blood
of your saints
and
prophets,
\par }{\Q so
you have given
them
blood
to drink.
They got
what they deserved!”
\par }{\PP \VS{7}Then
I heard
the altar
reply, “Yes,
Lord
God,
the All-Powerful,
your
judgments
are true
and
just!”
\par }{\PP \VS{8}Then
the fourth
angel poured out
his
bowl
on
the sun,
and
it was permitted
to scorch
people
with
fire.
\VS{9}Thus
people
were scorched
by the terrible
heat,
yet
they blasphemed
the name
of God,
who has
ruling authority
over
these
plagues,
and
they would
not
repent
and give
him
glory.
\par }{\PP \VS{10}Then
the fifth
angel poured out
his
bowl
on
the throne
of the beast
so that
darkness
covered
his
kingdom,
and
people began to bite
their
tongues
because of
their pain.
\VS{11}They blasphemed
the God
of heaven
because of
their
sufferings
and
because of
their
sores,
but
nevertheless
they still refused to repent
of
their
deeds.
\par }{\PP \VS{12}Then
the sixth
angel poured out
his
bowl
on
the great
river
Euphrates
and
dried up
its
water
to
prepare
the way
for the kings
from
the east.
\VS{13}Then
I saw
three
unclean
spirits
that looked like
frogs
coming out
of the mouth
of the dragon,
out of
the mouth
of the beast,
and
out of
the mouth
of the false prophet.
\VS{14}For
they are
the spirits
of the demons
performing
signs
who
go out
to
the kings
of the earth
to bring them
together
for
the battle
that will take place on the great
day
of God,
the All-Powerful.
\par }{\Q \VS{15}(Look! I will come
like
a thief!
\par }{\Q Blessed
is the one who stays alert
and
does not lose
his
clothes
so that
he will
not
have to walk around
naked
and
his
shameful condition
be seen.)
\par }{\PP \VS{16}Now
the spirits gathered
the kings and their armies
to
the place
that is called
Armageddon
in Hebrew.
\par }{\PP \VS{17}Finally the seventh
angel poured out
his
bowl
into
the air
and
a loud
voice
came out
of
the temple
from
the throne,
saying: “It is done!”
\VS{18}Then
there were
flashes of lightning,
roaring,
and crashes of thunder,
and
there was
a tremendous
earthquake
– an earthquake
unequaled
since
humanity
has been
on
the earth,
so tremendous
was that earthquake.
\VS{19}The great
city
was split
into
three
parts
and
the cities
of the
nations
collapsed.
So
Babylon
the great
was remembered
before
God,
and was given
the
cup
filled with the wine
made of God’s
furious
wrath.
\VS{20}Every
island
fled away
and
no
mountains
could be found.
\VS{21}And
gigantic
hailstones,
weighing
about
a hundred pounds each,
fell
from
heaven
on
people,
but
they blasphemed
God
because of
the plague
of hail,
since it
was
so horrendous.

\par }\Chap{17}{\PP \VerseOne{1}Then
one
of the seven
angels
who had
the seven
bowls
came
and
spoke
to
me. “Come,”
he said,
“I will show
you
the condemnation and punishment
of the great
prostitute
who sits
on
many
waters,
\VS{2}with
whom
the kings
of the earth
committed sexual immorality
and
the earth’s
inhabitants
got drunk
with
the wine
of her
immorality.”
\VS{3}So
he carried
me
away
in
the Spirit
to
a wilderness,
and
there I saw
a woman
sitting
on
a scarlet
beast
that was full
of blasphemous
names
and had
seven
heads
and
ten
horns.
\VS{4}Now
the woman
was
dressed in
purple
and
scarlet
clothing, and
adorned
with gold,
precious
stones,
and
pearls.
She held
in
her
hand
a golden
cup
filled
with detestable things
and
unclean things
from her
sexual immorality.
\VS{5}On
her
forehead
was written
a name,
a mystery: “Babylon
the Great,
the Mother
of prostitutes
and
of the detestable things
of the earth.”
\VS{6}I saw
that the woman
was drunk
with
the blood
of the saints
and
the blood
of those who testified
to Jesus.
I was
greatly
astounded
when
I saw
her.
\VS{7}But
the angel
said
to me,
“Why
are you astounded? I
will interpret
for you
the mystery
of the woman
and
of the beast
with
the seven
heads
and
ten
horns
that carries
her.
\VS{8}The
beast
you saw
was,
and
is not,
but
is
about
to come up
from
the abyss
and
then go
to
destruction.
The inhabitants
of
the
earth
– all those whose
names
have
not
been written
in
the
book
of life
since
the foundation
of the world
– will be astounded
when
they see
that
the beast
was,
and
is not,
but is
to come.
\VS{9}(This requires a mind
that has
wisdom.) The seven
heads
are
seven
mountains
the woman
sits
on. They are also seven kings:
\VS{10}five
have fallen;
one
is,
and the other
has
not yet
come,
but
whenever
he does come,
he must
remain
for only a brief time.
\VS{11}The beast
that
was,
and
is
not,
is
himself
an eighth
king and
yet is
one of
the seven,
and
is going
to
destruction.
\VS{12}The ten
horns
that
you saw
are
ten
kings
who
have
not yet
received
a kingdom,
but
will receive
ruling authority
as
kings
with
the beast
for one
hour.
\VS{13}These kings
have
a single
intent,
and
they will give
their
power
and
authority
to the beast.
\VS{14}They
will make war
with
the Lamb,
but
the Lamb
will conquer
them,
because
he is
Lord
of lords
and
King
of kings,
and
those accompanying
the Lamb
are the called,
chosen,
and
faithful.”
\par }{\PP \VS{15}Then
the angel said
to me,
“The waters
you saw
(where
the prostitute
is seated) are
peoples,
multitudes,
nations,
and
languages.
\VS{16}The ten
horns
that
you saw,
and
the beast
– these
will hate
the prostitute
and
make
her
desolate
and
naked.
They will consume
her
flesh
and
burn
her
up
with
fire.
\VS{17}For
God
has put
into
their
minds
to carry out
his
purpose
by making
a
decision
to give
their
royal power
to the beast
until
the words
of God
are fulfilled.
\VS{18}As for the woman
you saw,
she is
the great
city
that has
sovereignty
over
the kings
of the earth.”

\par }\Chap{18}{\PP \VerseOne{1}After
these things
I saw
another
angel,
who possessed
great
authority,
coming down
out of
heaven,
and
the earth
was lit up
by
his
radiance.
\VS{2}He shouted
with
a powerful
voice:
\par }{\Q “Fallen,
fallen,
is Babylon
the great!
\par }{\Q She has become
a lair
for demons,
\par }{\Q a haunt
for every
unclean
spirit,
\par }{\Q a haunt
for every
unclean bird,
\par }{\Q a haunt for every unclean
and
detested beast.
\par }{\Q \VS{3}For
all
the nations
have fallen
from
\par }{\Q the wine
of her
immoral
passion,
\par }{\Q and
the kings
of the
earth
have committed sexual immorality
with
her,
\par }{\Q and
the merchants
of the earth
have gotten rich
from
the power
of her
sensual behavior.”
\par }{\PP \VS{4}Then
I heard
another
voice
from
heaven
saying,
“Come out
of
her,
my
people,
so
you will
not
take part in
her
sins
and
so
you will
not
receive
her
plagues,
\VS{5}because
her
sins
have piled up
all the way to
heaven
and
God
has remembered
her
crimes.
\VS{6}Repay
her
the same way
she
repaid
others; pay
her back
double
corresponding to
her
deeds.
In
the cup
she mixed,
mix
double the amount
for her.
\VS{7}As much as
she exalted
herself
and
lived in sensual luxury,
to this extent
give
her
torment
and
grief
because
she said
to
herself, ‘I rule
as queen
and
am
no
widow;
I will
never
experience
grief!’
\VS{8}For
this reason,
she will experience
her
plagues
in
a single
day: disease,
mourning,
and
famine,
and
she will be burned down
with
fire,
because
the Lord
God
who judges
her
is powerful!”
\par }{\PP \VS{9}Then
the kings
of the earth
who committed immoral acts
with
her
and
lived in sensual luxury
with her will weep
and
wail
for
her
when
they see
the smoke
from the fire that burns
her
up.
\VS{10}They will stand
a long way off
because
they are afraid
of her
torment,
and will say,
\par }{\Q “Woe,
woe,
O great
city,
\par }{\Q Babylon
the powerful
city!
\par }{\Q For
in a single
hour
your
doom
has come!”
\par }{\PP \VS{11}Then
the merchants
of the earth
will weep
and
mourn
for
her
because
no one
buys
their
cargo
any longer –
\VS{12}cargo
such as gold,
silver,
precious
stones,
pearls,
fine linen,
purple cloth,
silk,
scarlet cloth,
all sorts
of things made of citron wood,
all sorts
of objects
made of ivory,
all sorts
of things
made of
expensive
wood,
bronze,
iron
and
marble,
\VS{13}cinnamon,
spice, incense,
perfumed ointment,
frankincense,
wine,
olive oil
and
costly flour,
wheat,
cattle
and
sheep,
horses
and
four-wheeled carriages,
slaves
and
human
lives.
\par }{\Q \VS{14}(The ripe fruit
you
greatly desired
\par }{\Q has gone
from
you,
\par }{\Q and
all
your luxury
and
splendor
\par }{\Q have gone
from
you
–
\par }{\Q they will
never
ever be found again!)
\par }{\PP \VS{15}The merchants
who sold these things,
who got rich
from
her,
will stand
a long way off
because
they are afraid
of her
torment.
They will weep
and
mourn,
\VS{16}saying,
\par }{\Q “Woe,
woe,
O great
city
–
\par }{\Q dressed in
fine linen,
purple
and
scarlet clothing,
\par }{\Q and
adorned
with
gold,
precious
stones,
and
pearls –
\par }{\Q \VS{17}because
in a single
hour
such great
wealth
has been destroyed!”
\par }{\PP And
every
ship’s captain,
and
all
who sail
along
the coast –
seamen,
and
all
who make their living
from
the sea,
stood
a long way off
\VS{18}and
began to shout
when
they saw
the smoke
from the fire that burned
her
up, “Who
is like
the great
city?”
\VS{19}And
they threw
dust
on
their
heads
and
were shouting
with weeping
and
mourning,
\par }{\Q “Woe,
Woe,
O great
city
–
\par }{\Q in
which
all
those who had
ships
on the sea
got rich
from
her
wealth
–
\par }{\Q because
in a single
hour
she has been destroyed!”
\par }{\Q \VS{20}(Rejoice
over
her,
O heaven,
\par }{\Q and
you saints
and
apostles
and
prophets,
\par }{\Q for
God
has pronounced
judgment
against
her
on your behalf!)
\par }{\PP \VS{21}Then
one
powerful
angel
picked up
a stone
like
a huge
millstone,
threw
it into
the sea,
and said,
\par }{\Q “With this kind of
sudden violent
force

\par }{\Q Babylon
the great
city
will be thrown down
\par }{\Q and
it will
never
be found
again!
\par }{\Q \VS{22}And
the sound
of the harpists,
musicians,
\par }{\Q flute players,
and
trumpeters
\par }{\Q will
never
be heard
in
you
again.
\par }{\Q No
craftsman
who practices
any
trade
\par }{\Q will ever be found
in
you
again;
\par }{\Q the noise
of a mill
will
never
be heard
in
you
again.
\par }{\Q \VS{23}Even
the light
from a lamp
\par }{\Q will
never
shine
in
you
again!
\par }{\Q The voices
of the bridegroom
and
his bride
\par }{\Q will
never
be heard
in
you
again.
\par }{\Q For
your
merchants
were
the tycoons
of the world,
\par }{\Q because
all
the nations
were deceived
by
your
magic spells!
\par }{\Q \VS{24}The blood
of the saints
and
prophets
was found
in
her,
\par }{\Q along with the blood of all
those who had been killed
on
the earth.”

\par }\Chap{19}{\PP \VerseOne{1}After
these things
I heard
what sounded like
the loud voice
of a vast
throng
in
heaven,
saying,
\par }{\Q “Hallelujah! Salvation
and
glory
and
power
belong to our
God,
\par }{\Q \VS{2}because
his
judgments
are true
and
just.
\par }{\Q For
he has judged
the great
prostitute
\par }{\Q who
corrupted
the earth
with
her
sexual immorality,
\par }{\Q and
has avenged
the blood
of his
servants
poured out by
her own
hands!”
\par }{\PP \VS{3}Then
a second time
the crowd shouted,
“Hallelujah!” The smoke
rises
from her
forever
and ever.
\VS{4}The twenty-four
elders
and
the four
living creatures
threw
themselves to the ground
and
worshiped
God,
who was seated
on
the throne,
saying: “Amen! Hallelujah!”
\par }{\PP \VS{5}Then
a voice
came
from
the throne,
saying:
\par }{\Q “Praise
our
God
\par }{\Q all
you his
servants,
\par }{\Q and all you who fear
Him,
\par }{\Q both the small
and
the great!”
\par }{\SH The Wedding Celebration of the Lamb
\par }{\PP \VS{6}Then
I heard
what sounded like the voice
of a vast
throng,
like
the roar
of many
waters
and
like
loud
crashes
of thunder.
They were shouting:

\par }{\Q “Hallelujah!
\par }{\Q For
the Lord
our
God,
the All-Powerful,
reigns!
\par }{\Q \VS{7}Let us rejoice
and
exult
\par }{\Q and
give
him
glory,
\par }{\Q because
the wedding celebration
of the Lamb
has come,
\par }{\Q and
his
bride
has made
herself
ready.
\par }{\Q \VS{8}She was permitted
to
be dressed in
bright,
clean,
fine linen”
(for
the fine linen
is
the righteous deeds
of the saints).
\par }{\PP \VS{9}Then
the angel said
to me,
“Write
the following: Blessed
are those who are invited
to
the banquet
at the wedding celebration
of the Lamb!” He
also
said
to me,
“These
are
the true
words
of God.”
\VS{10}So
I threw
myself down
at his
feet
to worship
him,
but
he said,
“Do
not
do
this! I am
only a fellow servant
with you
and
your
brothers
who hold
to the testimony
about Jesus.
Worship
God,
for
the testimony
about Jesus
is
the
spirit
of prophecy.”
\par }{\SH The Son of God Goes to War
\par }{\PP \VS{11}Then
I saw
heaven
opened
and
here came
a white
horse! The one riding
it
was called
“Faithful”
and
“True,”
and
with
justice
he judges
and
goes to war.
\VS{12}His
eyes
are like a fiery
flame
and
there are many
diadem crowns
on
his
head.
He has
a name
written
that
no one
knows
except
himself.
\VS{13}He
is dressed in
clothing
dipped
in blood,
and
he is called
the Word
of God.
\VS{14}The armies
that are in
heaven,
dressed in
white,
clean,
fine linen,
were following
him
on
white
horses.
\VS{15}From
his
mouth
extends
a sharp
sword,
so that
with
it
he can strike
the nations.
{\BD{
{\IT{
He
will rule
them
with
an iron
rod}}}},
and
he
stomps
the winepress
of the
furious
wrath
of God,
the All-Powerful.
\VS{16}He
has
a name
written
on
his clothing
and
on
his
thigh: “King
of kings
and
Lord
of lords.”
\par }{\PP \VS{17}Then
I saw
one
angel
standing
in
the sun,
and
he shouted
in
a loud
voice
to all
the birds
flying
high in the sky:

\par }{\Q “Come,
gather around
for
the great
banquet
of God,
\par }{\Q \VS{18}to
eat
your fill of the flesh
of kings,
\par }{\Q the flesh
of generals,
\par }{\Q the flesh
of powerful
people,
\par }{\Q the flesh
of horses
and
those who ride
them,
\par }{\Q and
the flesh
of all people,
both
free
and
slave,
\par }{\Q and
small
and
great!”
\par }{\PP \VS{19}Then
I saw
the beast
and
the kings
of the earth
and
their
armies
assembled
to do
battle
with
the one who rode
the horse
and
with
his
army.
\VS{20}Now
the beast
was seized,
and
along with
him
the
false prophet
who had performed
the
signs
on
his
behalf –
signs by
which
he deceived
those who had received
the
mark
of the
beast
and
those who worshiped
his
image.
Both
of them were thrown
alive
into
the
lake
of fire
burning
with
sulfur.
\VS{21}The others
were killed
by
the sword
that extended
from
the mouth
of the one who rode
the horse,
and
all
the birds
gorged
themselves with
their
flesh.

\par }\Chap{20}{\PP \VerseOne{1}Then
I saw
an angel
descending
from
heaven,
holding
in
his
hand
the key
to the abyss
and
a huge
chain.
\VS{2}He
seized
the dragon
– the ancient
serpent,
who
is
the devil
and
Satan
– and
tied
him
up
for a thousand
years.
\VS{3}The angel then
threw
him
into
the abyss
and
locked
and
sealed
it so that
he could
not
deceive
the nations
until
the one thousand
years
were finished.
(After
these things
he must
be released
for a brief period
of time.)
\par }{\PP \VS{4}Then
I saw
thrones
and
seated
on
them
were those who had been given
authority to judge.
I also saw the
souls
of those who had been beheaded
because of
the
testimony
about Jesus
and
because
of the
word
of God.
These
had
not
worshiped
the
beast
or
his
image
and
had refused to receive
his mark
on
their
forehead
or hand.
They came to life
and
reigned
with
Christ
for a thousand
years.
\VS{5}(The rest
of the dead
did
not
come to life
until
the thousand
years
were finished.) This
is the first
resurrection.
\VS{6}Blessed
and
holy
is the one who takes
part
in
the first
resurrection.
The second
death
has
no
power
over them, but
they will be
priests
of God
and
of Christ,
and
they will reign
with
him
for a thousand
years.
\par }{\SH Satan’s Final Defeat
\par }{\PP \VS{7}Now
when
the thousand
years
are finished,
Satan
will be released
from
his
prison
\VS{8}and
will go out
to deceive
the nations
at
the four
corners
of the earth,
Gog
and
Magog,
to bring
them
together
for
the
battle.
They are as numerous
as
the grains of sand
in the sea.
\VS{9}They went up
on
the broad
plain of the earth
and
encircled
the camp
of the saints
and
the beloved
city,
but
fire
came down
from
heaven
and
devoured
them
completely.
\VS{10}And
the devil
who deceived
them
was thrown
into
the lake
of fire
and
sulfur,
where
the beast
and
the false prophet
are too, and
they will be tormented
there day
and
night
forever
and ever.
\par }{\SH The Great White Throne
\par }{\PP \VS{11}Then
I saw
a large
white
throne
and
the one who was seated
on
it;
the earth
and
the heaven
fled
from
his presence,
and
no
place
was found
for them.
\VS{12}And
I saw
the dead,
the great
and
the small,
standing
before
the throne.
Then
books
were opened,
and
another
book
was opened
– the book of life.
So
the dead
were judged
by
what was written
in
the books,
according
to their
deeds.
\VS{13}The sea
gave up
the dead
that were in
it,
and
Death
and
Hades
gave up
the dead
that were in
them,
and
each one
was judged
according
to his
deeds.
\VS{14}Then
Death
and
Hades
were thrown
into
the lake
of fire.
This
is
the second
death
– the lake
of fire.
\VS{15}If
anyone’s
name was
not
found
written
in
the book
of life,
that person was thrown
into
the lake
of fire.

\par }\Chap{21}{\PP \VerseOne{1}Then
I saw
a new
heaven
and
a new
earth,
for
the first
heaven
and
earth
had ceased
to exist, and
the sea
existed
no
more.
\VS{2}And
I saw
the holy
city
– the new
Jerusalem
– descending
out of
heaven
from
God,
made ready
like
a bride
adorned
for her
husband.
\VS{3}And
I heard
a loud
voice
from
the
throne
saying: “Look! The residence
of God
is among
human beings.
He will live
among
them,
and
they will be
his
people,
and
God
himself
will be
with
them.
\VS{4}He will wipe away
every
tear
from
their
eyes,
and
death
will
not
exist
any more
– or
mourning,
or
crying,
or
pain,
for the former things
have ceased
to exist.”
\par }{\PP \VS{5}And
the one seated
on
the throne
said: “Look! I am making
all things
new!” Then
he said
to me, “Write
it down,
because
these
words
are
reliable
and
true.”
\VS{6}He
also
said
to me,
“It is done! I
am the Alpha
and
the Omega,
the beginning
and
the end.
To the one who is thirsty
I
will give
water free of charge
from
the spring
of the water
of life.
\VS{7}The one who conquers
will inherit
these things,
and
I will be
his
God
and
he
will be
my
son.
\VS{8}But
to the cowards,
unbelievers,
detestable persons,
murderers,
the sexually immoral,
and
those who practice magic spells,
idol worshipers,
and
all
those who lie, their place will be in
the lake
that burns
with fire
and
sulfur.
That is
the second
death.”
\par }{\SH The New Jerusalem Descends
\par }{\PP \VS{9}Then
one
of the seven
angels
who had
the seven
bowls
full
of the seven
final
plagues
came
and
spoke
to me,
saying,
“Come,
I will show
you
the bride,
the wife
of the Lamb!”
\VS{10}So
he took
me
away
in
the Spirit
to
a huge,
majestic
mountain
and
showed
me
the holy
city,
Jerusalem,
descending
out
of heaven
from
God.
\VS{11}The city possesses
the glory
of God;
its
brilliance
is like
a precious
jewel,
like
a stone
of crystal-clear
jasper.
\VS{12}It has
a massive,
high
wall
with
twelve
gates,
with twelve
angels
at
the gates,
and
the names
of the twelve
tribes
of the nation
of Israel
are
written on the gates.
\VS{13}There are three
gates
on
the east side,
three
gates
on
the north side,
three
gates
on
the south side
and
three
gates
on
the west side.
\VS{14}The wall
of the city
has
twelve
foundations,
and
on
them
are the twelve
names
of the twelve
apostles
of the Lamb.
\par }{\PP \VS{15}The angel who spoke
to
me
had
a golden
measuring
rod
with which to
measure
the city
and
its
foundation
stones and
wall.
\VS{16}Now
the city
is laid out
as a square,
its
length
and width
the same. He
measured
the city
with the measuring rod
at
fourteen hundred miles
(its
length
and
width
and
height
are
equal).
\VS{17}He
also
measured
its
wall,
one hundred
forty-four
cubits
according to human
measurement,
which
is
also the angel’s.
\VS{18}The city’s wall
is made
of jasper
and
the city
is pure
gold,
like
transparent
glass.
\VS{19}The foundations
of the city’s
wall
are decorated
with every kind
of precious
stone.
The first
foundation
is jasper,
the second
sapphire,
the third
agate,
the fourth
emerald,
\VS{20}the fifth
onyx,
the sixth
carnelian,
the seventh
chrysolite,
the eighth
beryl,
the ninth
topaz,
the tenth
chrysoprase,
the eleventh
jacinth,
and the twelfth
amethyst.
\VS{21}And
the twelve
gates
are twelve
pearls
– each
one
of the gates
is
made from
just one
pearl! The main street
of the city
is pure
gold,
like
transparent
glass.
\par }{\PP \VS{22}Now
I saw
no
temple
in
the city,
because
the Lord
God
– the All-Powerful –
and
the Lamb
are
its
temple.
\VS{23}The city
does
not
need
the sun
or
the moon
to
shine
on it,
because
the glory
of God
lights
it
up,
and
its
lamp
is the Lamb.
\VS{24}The nations
will walk
by
its
light
and
the kings
of the earth
will bring
their
grandeur
into
it.
\VS{25}Its
gates
will
never
be closed
during the day
(and there will be
no
night
there).
\VS{26}They will bring
the grandeur
and
the wealth
of the nations
into
it,
\VS{27}but
nothing
ritually unclean
will ever enter
into
it,
nor
anyone
who does
what is detestable
or
practices falsehood,
but
only those whose names are written
in
the Lamb’s
book
of life.

\par }\Chap{22}{\PP \VerseOne{1}Then
the angel showed
me
the river
of the water
of life
– water as clear as
crystal
– pouring out
from
the throne
of God
and
of the Lamb,
\VS{2}flowing down
the middle
of the city’s main street.
On each side
of the river
is the tree
of life
producing
twelve
kinds of fruit,
yielding
its
fruit
every
month of the year.
Its
leaves
are for
the healing
of the nations.
\VS{3}And
there will
no longer
be
any
curse,
and
the throne
of God
and
the Lamb
will be
in
the city.
His
servants
will worship
him,
\VS{4}and
they will see
his
face,
and
his
name
will be on
their
foreheads.
\VS{5}Night
will be
no
more,
and
they will
not
need
the light
of a lamp
or
the light
of the sun,
because
the Lord
God
will shine
on
them,
and
they will reign
forever
and ever.
\par }{\SH A Final Reminder
\par }{\PP \VS{6}Then
the angel said
to me,
“These
words
are reliable
and
true.
The Lord,
the God
of the spirits
of the prophets,
has sent
his
angel
to show
his
servants
what
must
happen
soon.”
\par }{\Q \VS{7}(Look! I am coming
soon!
\par }{\Q Blessed
is the one who keeps
the words
of the prophecy
expressed in this
book.)
\par }{\PP \VS{8}I,
John,
am the one who heard
and
saw
these things,
and
when
I heard
and
saw
them, I threw
myself down
to worship
at
the feet
of the angel
who was showing
them
to me.
\VS{9}But
he said
to me,
“Do
not
do
this! I am
a fellow servant
with you
and
with your
brothers
the prophets,
and
with those who obey
the words
of this
book.
Worship
God!”
\VS{10}Then
he said
to me,
“Do
not
seal up
the words
of the prophecy
contained in this
book,
because
the time
is
near.
\VS{11}The evildoer
must continue to do evil,
and
the one who is morally filthy
must continue to be filthy.
The one who is righteous
must continue to act
righteously,
and
the one who is holy
must continue to be holy.”
\par }{\Q \VS{12}(Look! I am coming
soon,
\par }{\Q and
my
reward
is with
me
to pay
each one
according to
what he
has
done!
\par }{\Q \VS{13}I
am the Alpha
and
the Omega,
\par }{\Q the first
and
the last,
\par }{\Q the beginning
and
the end!)
\par }{\PP \VS{14}Blessed
are those who wash
their
robes
so
they
can have
access
to
the tree
of life
and
can enter
into
the city
by the gates.
\VS{15}Outside
are the dogs
and
the sorcerers
and
the sexually immoral,
and
the murderers,
and
the idolaters
and
everyone
who loves
and
practices
falsehood!
\par }{\PP \VS{16}“I,
Jesus,
have sent
my
angel
to testify
to you
about these things
for
the churches.
I
am
the root
and
the descendant
of David,
the bright
morning
star!”
\VS{17}And
the Spirit
and
the bride
say,
“Come!” And
let the one who hears
say: “Come!” And
let the one who is thirsty
come;
let the one who wants
it take
the water
of life
free of charge.
\par }{\PP \VS{18}I
testify
to the one who hears
the words
of the prophecy
contained in this
book: If
anyone
adds
to
them,
God
will add
to
him
the plagues
described
in
this
book.
\VS{19}And
if
anyone
takes away
from
the words
of this
book
of prophecy,
God
will take away
his
share
in
the
tree
of life
and
in
the holy
city
that are described
in
this
book.
\par }{\PP \VS{20}The one who testifies
to these things
says, “Yes,
I am coming
soon!” Amen! Come,
Lord
Jesus!
\VS{21}The grace
of the Lord
Jesus
be with
all.
\par }