\NormalFont\ShortTitle{Ruth}
{\MT Ruth

\par }\ChapOne{1}{\SH A Family Tragedy: Famine and Death
\par }{\PP \VerseOne{1}During
the time
of the judges
there was
a famine
in the land
of Judah. So
a man
from Bethlehem
in Judah
went to live as a resident foreigner
in the region
of Moab,
along with his wife
and two
sons.
\VS{2}(Now the man’s
name
was Elimelech,
his wife
was Naomi,
and his two
sons
were Mahlon
and Kilion.
They were of the clan of Ephrath
from Bethlehem
in Judah.) They entered
the region
of Moab
and settled
there.
\VS{3}Sometime later Naomi’s
husband
Elimelech
died,
so she and her two
sons
were left alone.
\VS{4}So
her sons married
Moabite
women.
(One
was named
Orpah
and the other Ruth.) And they continued to live
there
about ten
years.
\VS{5}Then Naomi’s
two
sons, Mahlon
and Kilion,
also died.
So the woman
was left all alone – bereaved of her two children as well as her husband!
\VS{6}So she
decided
to return
home from the region
of Moab,
accompanied by her daughters-in-law,
because
while she was living in Moab
she had heard
that
the
{\ND{Lord}}
had shown concern
for his people,
reversing the famine by providing
abundant crops.
\par }{\SH Ruth Returns with Naomi
\par }{\PP \VS{7}Now as she and her two
daughters-in-law
began to leave
the place
where
she had been living to return
to
the land
of Judah,
\VS{8}Naomi
said
to her two
daughters-in-law,
“Listen to me! Each of you should
return
to your mother’s
home! May
the {\ND{Lord}}
show you
the same kind of devotion
that you have shown
to your deceased
husbands and to me!
\VS{9}May
the {\ND{Lord}}
enable each of you to find
security
in the home
of a new husband!” Then she
kissed
them goodbye
and they
wept
loudly.
\VS{10}But they said
to her, “No! We
will return
with
you to your people.”
\par }{\PP \VS{11}But Naomi
replied,
“Go back
home, my daughters! There is no reason
for you to return
to Judah with
me! I am no longer
capable of giving birth
to sons
who might become
your husbands!
\VS{12}Go
back
home, my daughters! For
I am too old
to get married
again. Even if I thought
that there was
hope
that I could get married tonight
and conceive
sons,
\VS{13}surely
you would not want to wait
until
they were old enough
to marry! Surely
you would not
remain
unmarried
all that time! No, my daughters,
you must not return with me. For
my intense suffering
is too much
for you to bear.
For
the {\ND{Lord}}
is afflicting me!”
\par }{\PP \VS{14}Again
they wept
loudly.
Then Orpah
kissed
her mother-in-law
goodbye, but Ruth
clung tightly to her.
\VS{15}So Naomi said,
“Look,
your sister-in-law
is returning
to
her people
and to her
god.
Follow
your sister-in-law
back home!”
\VS{16}But Ruth
replied,
\par }{\Q “Stop
urging
me to abandon
you!

\par }{\Q For
wherever
you go,
I will go.
\par }{\Q Wherever
you live,
I will live.
\par }{\Q Your people
will become my people,
\par }{\Q and your God
will become my God.
\par }{\Q \VS{17}Wherever
you die,
I will die
– and there
I will be buried.
\par }{\Q May
the

{\ND{Lord}}
punish me severely
if
I do
not keep my promise!

\par }{\Q Only death
will be able to separate me from you!”
\par }{\PP \VS{18}When Naomi realized
that
Ruth was determined
to go
with
her, she stopped
trying to dissuade
her.
\VS{19}So
the two
of them journeyed
together until
they arrived
in Bethlehem.
\par }{\SH Naomi and Ruth Arrive in Bethlehem
\par }{\PP When
they entered
Bethlehem,
the whole
village
was excited about their
arrival. The women of the village
said, “Can this
be Naomi?”
\VS{20}But she replied
to
them, “Don’t
call
me ‘Naomi’! Call
me ‘Mara’
because
the Sovereign One
has treated me very
harshly.
\VS{21}I
left
here full,
but the
{\ND{Lord}}
has caused me to return
empty-handed.
Why
do you call
me ‘Naomi,’
seeing that the
{\ND{Lord}}
has opposed me,
and the Sovereign One
has caused me to suffer?”
\VS{22}So Naomi
returned,
accompanied by her Moabite
daughter-in-law
Ruth,
who came back
with
her from the region
of Moab.
(Now they
arrived
in Bethlehem
at the beginning
of the barley
harvest.)

\par }\Chap{2}{\PP \VerseOne{1}Now Naomi
had a relative
on her husband’s
side of the family
named
Boaz.
He was a wealthy,
prominent
man
from the clan of Elimelech.
\VS{2}One day Ruth
the Moabite
said to
Naomi,
“Let
me go
to the fields
so I can gather
grain
behind
whoever
permits
me to do so.” Naomi replied,
“You may go,
my daughter.”
\VS{3}So Ruth went
and gathered
grain in the fields
behind
the harvesters.
Now she just happened
to end up
in the portion
of the field
belonging to Boaz,
who
was from the clan
of Elimelech.
\par }{\SH Boaz and Ruth Meet
\par }{\PP \VS{4}Now at
that very moment,
Boaz
arrived
from Bethlehem
and greeted
the harvesters,
“May the
{\ND{Lord}}
be with
you!” They replied, “May the
{\ND{Lord}}
bless you!”
\VS{5}Boaz
asked
his servant
in charge
of the harvesters,
“To whom
does this young woman belong?”
\VS{6}The servant
in charge
of the harvesters
replied,
“She’s the young
Moabite
woman who came back
with
Naomi
from the region
of Moab.
\VS{7}She asked, ‘May
I follow
the harvesters
and gather
grain among the bundles?’ Since she arrived
she has been working hard
from this
morning
until
now –
except for sitting
in the resting hut
a short time.”
\par }{\PP \VS{8}So Boaz
said
to
Ruth,
“Listen
carefully, my dear! Do not
leave
to gather
grain in another
field.
You need
not
go beyond
the limits
of this
field. You may
go along beside
my female workers.
\VS{9}Take note
of the field
where
the men are harvesting
and follow
behind
with the female workers. I will tell
the men
to leave
you alone.
When you are thirsty,
you may go
to
the water jars
and drink
some of
the water the servants
draw.”
\par }{\PP \VS{10}Ruth knelt
before
him with her forehead to the ground
and said
to him,
“Why
are you so kind
and so attentive
to me, even though I am
a foreigner?”
\VS{11}Boaz
replied
to her, “I have been given
a full report
of all
that
you have done
for your mother-in-law
following
the death
of your husband
– how you left
your father
and your mother,
as well as your homeland,
and came
to
live among people
you did not
know
previously.
\VS{12}May the
{\ND{Lord}}
reward
your efforts! May your acts of kindness be repaid
fully
by
the {\ND{Lord}}
God
of Israel,
from whom
you have sought
protection!”
\VS{13}She said,
“You really
are being kind
to me, sir,
for
you have reassured
and encouraged
me, your servant,
even though I am
not
one
of your servants!”
\par }{\PP \VS{14}Later during the mealtime
Boaz
said
to her, “Come
here
and have some food! Dip
your bread
in the vinegar!” So she sat down
beside
the harvesters.
Then he handed her some
roasted grain.
She ate
until she was full
and saved the rest.
\VS{15}When she got
up to gather
grain, Boaz
told
his male servants,
“Let her gather grain
even
among
the bundles! Don’t chase her
off!
\VS{16}Make sure you pull out
ears of grain
for her and drop
them so she can gather
them up.
Don’t tell
her not to!”
\VS{17}So she gathered
grain in the field
until
evening.
When she threshed
what she had
gathered,
it came to about thirty pounds
of barley!
\par }{\SH Ruth Returns to Naomi
\par }{\PP \VS{18}She carried
it back
to town,
and her mother-in-law
saw
how much grain she had
gathered.
Then Ruth
gave
her the roasted grain she had
saved
from mealtime.
\VS{19}Her mother-in-law
asked
her, “Where
did you gather
grain today? Where
did
you work? May the one who took notice of you be
rewarded!” So Ruth
told
her mother-in-law
with
whom
she had worked.
She said,
“The name
of the man
with
whom
I worked
today
is Boaz.”
\VS{20}Naomi
said
to her daughter-in-law,
“May he be rewarded
by the
{\ND{Lord}}
because
he has
shown
loyalty
to the living
on behalf
of the dead!” Then Naomi
said
to her, “This man
is
a close relative
of ours; he is
our guardian.”
\VS{21}Ruth
the Moabite
replied,
“He even
told
me,
‘You may
go along beside
my servants
until
they have
finished
gathering
all
my harvest!’ ”
\VS{22}Naomi
then said
to
her daughter-in-law
Ruth,
“It is good,
my daughter,
that
you should go out
to work with
his female
servants. That way you will not
be harmed,
which could happen in another
field.”
\VS{23}So Ruth worked beside Boaz’s
female servants,
gathering
grain until
the end
of the barley
harvest
as well as the wheat
harvest.
After that she stayed
home with
her mother-in-law.

\par }\Chap{3}{\PP \VerseOne{1}At that time, Naomi,
her mother-in-law,
said
to her, “My daughter,
I must
find
a home
for you so
you will be secure.
\VS{2}Now
Boaz,
with
whose female servants
you worked, is
our close relative.
Look,
tonight
he
is winnowing
barley
at the threshing floor.
\VS{3}So bathe
yourself, rub on some perfumed oil,
and get
dressed up.
Then go down
to the threshing floor.
But don’t
let the man
know
you’re there until
he finishes
his meal.
\VS{4}When he gets ready to go
to sleep,
take careful notice
of the place
where
he lies down.
Then go,
uncover
his legs,
and lie down
beside him. He
will tell
you what
you should do.”
\VS{5}Ruth replied
to
Naomi, “I will do everything
you have
told me to do.”
\par }{\SH Ruth Visits Boaz
\par }{\PP \VS{6}So she went down
to the threshing floor
and did
everything
her mother-in-law
had instructed her to do.
\VS{7}When Boaz
had finished
his meal
and was feeling
satisfied,
he lay down to sleep
at the far end
of the grain heap.
Then Ruth crept up
quietly,
uncovered
his legs,
and lay down beside him.
\VS{8}In the middle
of the night
he was startled
and turned over. Now he saw
a woman
lying
beside him!
\VS{9}He said,
“Who
are you?” She replied,
“I am
Ruth,
your servant.
Marry
your servant,
for
you
are a guardian of the family interests.”
\VS{10}He said,
“May you
be rewarded
by the
{\ND{Lord}}, my dear! This act of devotion
is greater than
what you did before.
For you have not sought to marry one of the young men,
whether
rich
or
poor.
\VS{11}Now,
my dear,
don’t
worry! I intend
to do
for
you everything
you propose,
for
everyone
in the village
knows
that
you
are a worthy
woman.
\VS{12}Now
yes,
it is true
that
I am
a guardian,
but there
is another guardian
who is
a closer relative
than I am.
\VS{13}Remain
here tonight.
Then
in the morning,
if
he agrees
to marry you, fine, let him
do so. But if
he does not
want
to do so,
I
promise,
as surely
as the
{\ND{Lord}}
lives, to marry
you. Sleep
here
until
morning.”
\VS{14}So she slept
beside
him until
morning.
She woke up
while it was still
dark. Boaz thought, “No
one
must know
that
a woman
visited
the threshing floor.”
\VS{15}Then he said,
“Hold out
the shawl
you are wearing and grip
it tightly.” As she held
it tightly, he measured out
about sixty
pounds of barley
into the shawl and put
it on
her shoulders. Then he went
into town,
\VS{16}and she returned
to
her mother-in-law.
\par }{\SH Ruth Returns to Naomi
\par }{\PP When Ruth
returned to her mother-in-law, Naomi asked, “How did things
turn out for you, my daughter?” Ruth told
her about all
the man
had
done for her.
\VS{17}She said,
“He gave
me these
sixty
pounds of barley,
for
he said
to me, ‘Do not
go
to your mother-in-law
empty-handed.’ ”
\VS{18}Then Naomi said,
“Stay
put, my daughter,
until
you know
how
the matter
turns
out. For
the man
will not
rest
until he has
taken care
of the matter
today.”

\par }\Chap{4}{\PP \VerseOne{1}Now Boaz
went up
to the village gate
and sat
there.
Then
along came
the guardian
whom
Boaz
had mentioned
to Ruth! Boaz said,
“Come
here
and sit down,
‘John Doe’!” So
he came
and sat down.
\VS{2}Boaz chose
ten
of the village
leaders
and said,
“Sit
down here!” So they sat down.
\VS{3}Then Boaz said
to the guardian, “Naomi,
who has returned
from the region
of Moab,
is selling
the portion
of land
that belongs
to our relative
Elimelech.
\VS{4}So
I am
legally informing
you: Acquire
it before
those sitting
here and before
the leaders
of my people! If
you want to exercise
your right to redeem it,
then do so.
But if
not,
then tell
me
so
I will
know.
For
you possess
the first option
to redeem it;
I am
next
in line
after
you.” He replied,
“I
will redeem it.”
\VS{5}Then Boaz
said,
“When
you acquire
the field
from
Naomi,
you
must
also
acquire
Ruth
the Moabite,
the wife
of our deceased
relative,
in order to preserve
his family name
by raising
up
a descendant who will inherit his property.”
\VS{6}The guardian
said,
“Then I am unable
to redeem
it,
for I would
ruin
my own inheritance
in that case. You
may exercise
my redemption option,
for
I am unable
to redeem it.”
\VS{7}(Now this
used to be
the customary way to finalize
a transaction
involving
redemption
in
Israel: A man
would remove
his sandal
and give
it to the other party.
This
was a legally binding act
in Israel.)
\VS{8}So the guardian
said
to Boaz,
“You may acquire
it,” and he removed his sandal.
\VS{9}Then Boaz
said
to the leaders
and all
the people,
“You
are witnesses
today
that
I have acquired
from Naomi
all
that belonged
to Elimelech,
Kilion,
and Mahlon.
\VS{10}I
have also
acquired
Ruth
the Moabite,
the wife
of Mahlon,
as my wife
to raise
up a descendant who will inherit
his property
so
the name
of the deceased
might not
disappear from
among his
relatives
and from his village.
You
are witnesses
today.”
\VS{11}All
the people
who
were at the gate
and the elders
replied,
“We are witnesses.
May the
{\ND{Lord}}
make
the
woman
who is entering
your home
like Rachel
and Leah,
both
of whom
built
up the
house
of Israel! May you prosper in Ephrathah
and become famous
in Bethlehem.
\VS{12}May your family
become
like the family
of Perez –
whom
Tamar
bore
to Judah
– through the descendants
the {\ND{Lord}}
gives
you by this young woman.”
\par }{\SH A Grandson is Born to Naomi
\par }{\PP \VS{13}So Boaz
married
Ruth
and had sexual relations
with her.
The
{\ND{Lord}}
enabled
her to conceive
and she gave birth
to a son.
\VS{14}The village women
said
to
Naomi,
“May the
{\ND{Lord}}
be praised
because he has
not
left you without a guardian
today! May he become famous
in Israel!
\VS{15}He will encourage
you and provide
for you when you are old,
for
your daughter-in-law,
who
loves
you, has given him birth.
She is better
to you than seven
sons!”
\VS{16}Naomi
took
the child
and placed
him on her lap;
she became
his caregiver.
\VS{17}The neighbor
women named
him, saying,
“A son
has been born
to Naomi.”
They named
him Obed.
Now he
became the father
of Jesse
– David’s
father!
\par }{\SH Epilogue: Obed in the Genealogy of David
\par }{\PP \VS{18}These
are the descendants
of Perez: Perez
was the father
of Hezron,
\VS{19}Hezron
was the father
of Ram,
Ram
was the father
of Amminadab,
\VS{20}Amminadab
was the father
of Nachshon,
Nachshon
was the father
of Salmah,
\VS{21}Salmon
was the father
of Boaz,
Boaz
was the father
of Obed,
\VS{22}Obed
was the father
of Jesse,
and Jesse
was the father
of David.
\par }