\NormalFont\ShortTitle{Song of Solomon}
{\MT Song of Songs

\par }\ChapOne{1}{\SH Title/Superscription
\par }{\PP \par }{\Q \VerseOne{1}Solomon’s
Most Excellent
Love Song.
\par }{\SH The Desire for Love
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved}}}}
to Her Lover:}}
\par }{\Q \VS{2}Oh, how I wish you would kiss
me passionately!

\par }{\Q For
your lovemaking is
more delightful than
wine.
\par }{\Q \VS{3}The fragrance
of your colognes
is delightful;
\par }{\Q your name
is like the finest perfume.
\par }{\Q No wonder
the young women
adore you!
\par }{\Q \VS{4}Draw
me after
you; let
us hurry!

\par }{\Q May the king
bring
me into his bedroom chambers!

\par }{\SH {\QT{The Maidens to the Lover:}}*
\par }{\Q We will rejoice
and delight in you;

\par }{\Q we will praise your love
more than wine.
\par }{\Q {\QT{The Beloved to Her Lover:}}
\par }{\Q How rightly
the young women adore you!
\par }{\SH The Country Maiden and the Daughters of Jerusalem
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to the Maidens:}}}}}}
\par }{\Q \VS{5}I
am dark
but lovely,
O maidens
of Jerusalem,
\par }{\Q dark like the tents
of Qedar,
\par }{\Q lovely like the tent curtains
of Salmah.
\par }{\Q \VS{6}Do not
stare
at me because I am
dark,
\par }{\Q for the sun
has burned
my skin.

\par }{\Q My brothers
were angry
with me;
\par }{\Q they made
me the keeper
of the vineyards.
\par }{\Q Alas, my own vineyard
I could not
keep!
\par }{\SH The Shepherd and the Shepherdess
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{7}Tell
me, O you whom my heart loves,
\par }{\Q where do
you pasture your sheep?
\par }{\Q Where do
you rest
your sheep during the midday
heat?
\par }{\Q Tell me lest I wander
around

\par }{\Q beside the flocks
of your companions!
\par }{\SH 
{\QT{
The Lover to His Beloved:}}
\par }{\Q \VS{8}If
you do not
know,
O most beautiful
of women,
\par }{\Q simply follow
the tracks
of my flock,
\par }{\Q and pasture
your little lambs
\par }{\Q beside
the tents
of the shepherds.
\par }{\SH The Beautiful Mare and the Fragrant Myrrh
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{9}O my beloved,
you are like
a mare
\par }{\Q among Pharaoh’s
stallions.
\par }{\Q \VS{10}Your cheeks
are beautiful
with ornaments;
\par }{\Q your neck
is lovely with strings of jewels.
\par }{\Q \VS{11}We will make
for you gold
ornaments
\par }{\Q studded
with
silver.
\par }{\SH 
{\QT{
The Beloved about Her Lover:}}
\par }{\Q \VS{12}While
the king
was at his banqueting table,
\par }{\Q my nard
gave
forth its fragrance.
\par }{\Q \VS{13}My beloved
is like a fragrant pouch
of myrrh
\par }{\Q spending the night
between
my breasts.
\par }{\Q \VS{14}My beloved
is like a cluster
of henna
blossoms
\par }{\Q in the vineyards
of En-Gedi.
\par }{\SH Mutual Praise and Admiration
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{15}Oh,
how beautiful
you are, my beloved!

\par }{\Q Oh,
how beautiful
you are!
\par }{\Q Your eyes
are like doves!
\par }{\SH 
{\QT{
The Beloved to Her Lover:}}
\par }{\Q \VS{16}Oh, how handsome
you are, my lover!

\par }{\Q Oh, how
delightful
you are!
\par }{\Q The lush foliage is our canopied bed;
\par }{\Q \VS{17}the cedars
are the beams
of our bedroom chamber;
\par }{\Q the pines
are the rafters of our bedroom.

\par }\Chap{2}{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VerseOne{1}I
am a meadow flower
from Sharon,
\par }{\Q a lily
from the valleys.
\par }{\SH 
{\QT{
The Lover to His Beloved:}}
\par }{\Q \VS{2}Like a lily
among
the thorns,
\par }{\Q so
is my darling
among
the maidens.
\par }{\SH 
{\QT{
The Beloved about Her Lover:}}
\par }{\Q \VS{3}Like an apple tree
among the trees
of the forest,
\par }{\Q so
is my beloved
among
the young men.
\par }{\Q I delight
to sit
in his shade,
\par }{\Q and his fruit
is sweet
to my taste.
\par }{\SH The Banquet Hall for the Love-Sick
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VS{4}He brought
me into
the banquet hall,
\par }{\Q and he looked at me lovingly.
\par }{\Q \VS{5}Sustain
me with raisin cakes,
\par }{\Q refresh
me with apples,
\par }{\Q for
I
am faint
with love.
\par }{\Q {\IT{The Double Refrain: Embracing and Adjuration}}
\par }{\Q \VS{6}His left hand
caresses
my head,
\par }{\Q and his right hand
stimulates me.
\par }{\SH 
{\QT{
The
Beloved
to the Maidens:}}
\par }{\Q \VS{7}I adjure
you, O maidens
of Jerusalem,
\par }{\Q by the gazelles
and by the young does
of the open fields:

\par }{\Q Do
not
awaken or
arouse
love
until
it pleases!
\par }{\SH The Arrival of the Lover
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VS{8}Listen! My lover
is approaching!

\par }{\Q Look! Here
he comes,
\par }{\Q leaping
over
the mountains,
\par }{\Q bounding
over
the hills!
\par }{\Q \VS{9}My lover
is like
a gazelle
or
a young
stag.
\par }{\Q Look! There
he stands
behind
our wall,
\par }{\Q gazing
through
the window,
\par }{\Q peering
through
the lattice.
\par }{\SH The Season of Love and the Song of the Turtle-Dove
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover}}}}
to His Beloved:}}
\par }{\Q \VS{10}My lover
spoke to me, saying:
\par }{\Q “Arise,
my darling;
\par }{\Q My beautiful
one, come away with me!
\par }{\Q \VS{11}Look! The winter
has passed,
\par }{\Q the winter
rains
are over and gone.
\par }{\Q \VS{12}The pomegranates
have appeared
in the land,
\par }{\Q the time
for pruning and singing
has come;
\par }{\Q the voice
of the turtledove
is heard
in our land.
\par }{\Q \VS{13}The fig tree
has budded,
\par }{\Q the vines
have blossomed
and give
off their fragrance.
\par }{\Q Arise,
come away
my darling;
\par }{\Q my beautiful
one, come away with me!”
\par }{\SH The Dove in the Clefts of En-Gedi
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{14}O my dove,
in the clefts
of the rock,
\par }{\Q in the hiding places
of the mountain crags,
\par }{\Q let me see
your face,
\par }{\Q let me hear
your voice;
\par }{\Q for
your voice
is sweet,
\par }{\Q and your face
is lovely.
\par }{\SH The Foxes in the Vineyard
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{15}Catch
the foxes
for us,
\par }{\Q the little
foxes,
\par }{\Q that ruin
the vineyards –
\par }{\Q for our vineyard
is in bloom.
\par }{\SH Poetic Refrain: Mutual Possession
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover}}}}:}}
\par }{\Q \VS{16}My lover is mine and I
am his;
\par }{\Q he grazes
among the lilies.
\par }{\SH The Gazelle and the Rugged Mountains
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{17}Until
the dawn
arrives
and the shadows
flee,
\par }{\Q turn,
my beloved
–
\par }{\Q be like
a gazelle
or
a young
stag
\par }{\Q on
the mountain
gorges.

\par }\Chap{3}{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VerseOne{1}All night
long on my bed
\par }{\Q I longed
for my lover.
\par }{\Q I longed for him but he never
appeared.
\par }{\Q \VS{2}“I will arise
and look
all around
throughout the town,
\par }{\Q and throughout the streets
and squares;
\par }{\Q I will search
for my beloved.”
\par }{\Q I searched
for him but I did not
find him.
\par }{\Q \VS{3}The night watchmen
found me – the ones who guard the city walls.
\par }{\Q “Have you seen my beloved?”
\par }{\Q \VS{4}Scarcely
had I passed
them
by
\par }{\Q when
I found
my beloved!
\par }{\Q I held
onto him tightly and would not
let him go
\par }{\Q until
I brought
him to
my mother’s
house,
\par }{\Q to
the bedroom chamber
of the one who conceived me.
\par }{\SH 
{\IT{
The
Adjuration Refrain}}
\par }{\SH {\QT{The Beloved to the Maidens:}}

\par }{\Q \VS{5}I
admonish you,
O maidens of
Jerusalem,
\par }{\Q by
the gazelles and
by the young does of
the open fields:
\par }{\Q “Do
not
awake
or arouse
love
until
it
pleases!”
\par }{\SH The Royal Wedding Procession
\par }{\SH 
{\BD{
{\QT{
{\IT{The Speaker:}}}}}}
\par }{\Q \VS{6}Who
is this
coming up
from
the desert
\par }{\Q like a column
of smoke,
\par }{\Q like a fragrant billow
of myrrh
and frankincense,
\par }{\Q every kind
of fragrant powder
of the traveling merchants?
\par }{\Q \VS{7}Look! It is Solomon’s
portable couch!

\par }{\Q It is surrounded
by sixty
warriors,
\par }{\Q some of Israel’s
mightiest warriors.
\par }{\Q \VS{8}All
of them are skilled
with a sword,
\par }{\Q well-trained
in the art of warfare.
\par }{\Q Each
has his sword
at his side,
\par }{\Q to guard against
the terrors
of the night.
\par }{\Q \VS{9}King
Solomon
made
a sedan
chair for himself
\par }{\Q of wood
imported from Lebanon.
\par }{\Q \VS{10}Its posts
were made
of silver;
\par }{\Q its back
was made of gold.
\par }{\Q Its seat
was upholstered with purple
wool;

\par }{\Q its interior
was inlaid
with leather
by the maidens
of Jerusalem.
\par }{\Q \VS{11}Come out,
O maidens
of Zion,
\par }{\Q and gaze
upon King
Solomon!
\par }{\Q He is wearing the crown
with which his mother
crowned
him
\par }{\Q on his wedding
day,
\par }{\Q on the most joyous
day
of his life!

\par }\Chap{4}{\SH 
{\BD{
{\QT{
{\IT{The Lover}}}}

to His Beloved:}}
\par }{\Q \VerseOne{1}Oh, you are beautiful,
my darling!

\par }{\Q Oh,
you are beautiful!
\par }{\Q Your eyes
behind
your veil
are like doves.
\par }{\Q Your hair
is like a flock
of female goats
\par }{\Q descending
from Mount
Gilead.
\par }{\Q \VS{2}Your teeth
are like a flock
of newly-shorn
sheep
\par }{\Q coming up
from
the washing
place;

\par }{\Q each of them
has a twin,
\par }{\Q and not
one of them is missing.
\par }{\Q \VS{3}Your lips
are like a scarlet
thread;
\par }{\Q your mouth
is lovely.
\par }{\Q Your forehead
behind
your veil
\par }{\Q is like a slice
of pomegranate.
\par }{\Q \VS{4}Your neck
is like the tower
of David
\par }{\Q built
with courses of stones;
\par }{\Q one thousand
shields
are hung
on it –
\par }{\Q all shields of valiant warriors.
\par }{\Q \VS{5}Your two
breasts
are like two
fawns,
\par }{\Q twins
of the gazelle
\par }{\Q grazing
among the lilies.
\par }{\Q \VS{6}Until
the dawn arrives
\par }{\Q and the shadows
flee,
\par }{\Q I will go
up to
the mountain
of myrrh,
\par }{\Q and to
the hill
of frankincense.
\par }{\Q \VS{7}You are altogether
beautiful,
my darling!
\par }{\Q There is no
blemish in you!
\par }{\SH The Wedding Night: Beautiful as Lebanon
\par }{\Q \VS{8}Come
with
me from Lebanon,
my bride,
\par }{\Q come
with
me from Lebanon.
\par }{\Q Descend
from the crest
of Amana,
\par }{\Q from the top
of Senir,
the summit of Hermon,
\par }{\Q from the lions’
dens
\par }{\Q and the mountain
haunts of the leopards.
\par }{\Q \VS{9}You have stolen my heart,
my sister,
my bride!
\par }{\Q You have stolen my heart
with one
glance of your eyes,
\par }{\Q with one
jewel of your necklace.
\par }{\Q \VS{10}How
delightful
is your love, my
sister,
my bride!
\par }{\Q How
much
better is
your love than
wine;
\par }{\Q the fragrance
of your perfume
is better than
any
spice!
\par }{\Q \VS{11}Your lips
drip
sweetness like the honeycomb,
my bride,
\par }{\Q honey
and milk
are under
your tongue.
\par }{\Q The fragrance
of your garments
is like the fragrance
of Lebanon.
\par }{\SH The Wedding Night: The Delightful Garden
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{12}You are a locked
garden,
my sister,
my bride;
\par }{\Q you are an enclosed
spring,
a sealed-up
fountain.
\par }{\Q \VS{13}Your shoots
are a royal garden
full of pomegranates
\par }{\Q with
choice
fruits:
\par }{\Q henna
with
nard,
\par }{\Q \VS{14}nard
and saffron;
\par }{\Q calamus
and cinnamon
with
every kind
of spice,
\par }{\Q myrrh
and aloes
with
all
the finest
spices.
\par }{\Q \VS{15}You are a garden
spring,
\par }{\Q a well
of fresh
water
flowing
down from
Lebanon.
\par }{\SH 
{\QT{
The Beloved to Her Lover:}}
\par }{\Q \VS{16}Awake,
O north
wind; come,
O south wind!
\par }{\Q Blow
on my garden
so that its fragrant
spices
may send out their sweet smell.

\par }{\Q May my beloved
come
into his garden
\par }{\Q and eat
its delightful
fruit!


\par }\Chap{5}{\SH 
{\QT{
The Lover to His Beloved:}}
\par }{\Q \VerseOne{1}I have entered
my garden,
O my sister,
my bride;
\par }{\Q I have gathered
my myrrh
with
my balsam spice.
\par }{\Q I have eaten
my honeycomb
and my honey;
\par }{\Q I have drunk
my wine
and my milk!
\par }{\SH {\QT{The Poet to the Couple:}}

\par }{\Q Eat,
friends,
and drink!

\par }{\Q Drink freely,
O lovers!
\par }{\SH The Trials of Love: The Beloved’s Dream of Losing Her Lover
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VS{2}I
was asleep,
but my mind
was dreaming.
\par }{\Q Listen! My lover
is knocking
at the door!

\par }{\SH {\QT{The Lover to His Beloved:}}
\par }{\Q “Open
for me, my sister,
my darling,
\par }{\Q my dove,
my flawless
one!
\par }{\Q My head
is drenched
with dew,
\par }{\Q my hair
with the dampness
of the night.”
\par }{\SH 
{\QT{
The Beloved to Her Lover:}}
\par }{\Q \VS{3}“I have already
taken off
my robe
– must I put
it on
again?
\par }{\Q I have already
washed
my feet
– must I soil them again?”
\par }{\Q \VS{4}My lover
thrust
his hand
through the hole,
\par }{\Q and my feelings
were stirred for him.
\par }{\Q \VS{5}I
arose
to open
for my beloved;
\par }{\Q my hands
dripped
with myrrh
–
\par }{\Q my fingers
flowed
with myrrh
\par }{\Q on
the handles
of the lock.
\par }{\Q \VS{6}I
opened
for my beloved,
\par }{\Q but my lover
had already
turned and gone
away.

\par }{\Q I fell into despair
when
he departed.
\par }{\Q I looked
for him but did not
find
him;
\par }{\Q I called
him but he did not
answer me.
\par }{\Q \VS{7}The watchmen
found
me as they made their rounds
in the city.
\par }{\Q They beat
me, they bruised
me;
\par }{\Q they took away
my cloak,
those watchmen
on
the walls!
\par }{\SH The Triumph of Love: The Beloved Praises Her Lover
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to the Maidens}}}}:}}
\par }{\Q \VS{8}O maidens
of Jerusalem, I command you –
\par }{\Q If you find my beloved, what will you tell him?
\par }{\Q Tell him that I am lovesick!
\par }{\SH 
{\QT{
The Maidens to The Beloved:}}
\par }{\Q \VS{9}Why
is your beloved
better than others,

\par }{\Q O
most beautiful
of women?
\par }{\Q Why
is
your beloved
better than others,
\par }{\Q that you would command us in this manner?
\par }{\SH 
{\QT{
The Beloved to the Maidens:}}
\par }{\Q \VS{10}My beloved
is dazzling
and ruddy;
\par }{\Q he stands out
in comparison to all other men.
\par }{\Q \VS{11}His head
is like the most pure gold.
\par }{\Q His hair
is curly –
black
like a raven.
\par }{\Q \VS{12}His eyes
are like doves
by
streams
of water,
\par }{\Q washed
in milk,
mounted
like
jewels.
\par }{\Q \VS{13}His cheeks
are like garden beds
full of balsam trees
yielding perfume.
\par }{\Q His lips
are like lilies
dripping
with drops
of myrrh.
\par }{\Q \VS{14}His arms
are like rods
of gold
set
with chrysolite.
\par }{\Q His abdomen
is like polished
ivory
inlaid
with sapphires.
\par }{\Q \VS{15}His legs
are like pillars
of marble
set
on
bases
of pure gold.
\par }{\Q His appearance
is like Lebanon,
choice
as its cedars.
\par }{\Q \VS{16}His mouth
is very sweet;
\par }{\Q he is totally desirable.
\par }{\Q This
is my beloved!
\par }{\Q This
is my companion,
O maidens
of Jerusalem!

\par }\Chap{6}{\SH 
{\BD{
{\IT{The Maidens to the Beloved:}}}}
\par }{\Q \VerseOne{1}Where
has your beloved
gone,
\par }{\Q O most beautiful
among women?
\par }{\Q Where
has your beloved
turned?
\par }{\Q Tell us,
that we may seek
him with you.
\par }{\SH 
{\BD{
{\IT{
The Beloved
to the Maidens:}}}}
\par }{\Q \VS{2}My beloved has gone down
to his garden,
\par }{\Q to the flowerbeds
of balsam spices,
\par }{\Q to graze
in the gardens,
\par }{\Q and to gather
lilies.
\par }{\SH Poetic Refrain: Mutual Possession
\par }{\SH 
{\BD{
{\IT{The Beloved about Her Lover:}}}}
\par }{\Q \VS{3}I
am my lover’s
and my lover
is mine;

\par }{\Q he grazes
among the lilies.
\par }{\SH The Renewal of Love
\par }{\SH 
{\BD{
{\IT{The Lover to His Beloved:}}}}
\par }{\Q \VS{4}My darling,
you
are as beautiful
as Tirzah,
\par }{\Q as lovely
as Jerusalem,
\par }{\Q as awe-inspiring
as bannered armies!
\par }{\Q \VS{5}Turn
your eyes
away from me
–
\par }{\Q they
overwhelm
me!
\par }{\Q Your hair
is like a flock
of goats
\par }{\Q descending
from
Mount Gilead.
\par }{\Q \VS{6}Your teeth
are like a flock
of sheep
\par }{\Q coming up
from
the washing;
\par }{\Q each
has its twin;
\par }{\Q not one of them is missing.
\par }{\Q \VS{7}Like a slice
of pomegranate
\par }{\Q is your forehead
behind
your veil.
\par }{\Q \VS{8}There may be sixty
queens,
\par }{\Q and eighty
concubines,
\par }{\Q and young women
without
number.
\par }{\Q \VS{9}But she is unique!

\par }{\Q My dove,
my perfect
one!
\par }{\Q She
is the special daughter of her mother,
\par }{\Q she is the favorite
of the one who bore
her.
\par }{\Q The maidens
saw
her and complimented
her;

\par }{\Q the queens
and concubines
praised her:
\par }{\Q \VS{10}“Who
is this
who appears
like
the dawn?

\par }{\Q Beautiful
as the moon,
bright
as the sun,
\par }{\Q awe-inspiring
as the stars in procession?”
\par }{\SH The Return to the Vineyards
\par }{\SH 
{\BD{
{\IT{The Lover to His Beloved:}}}}
\par }{\Q \VS{11}I went down
to
the orchard
of walnut trees,
\par }{\Q to look
for the blossoms
of the valley,
\par }{\Q to see
if the vines
had budded
\par }{\Q or if the pomegranates
were in bloom.
\VS{12}\par }{\Q I was beside
myself
with joy!

\par }{\Q There please give
me your myrrh,
\par }{\Q O daughter of my princely
people.
\par }{\SH The Love Song and Dance
\par }{\SH 
{\BD{
{\IT{The Lover to His Beloved:}}}}
\par }{\Q \VS{13}Turn , turn, O Perfect One!
\par }{\Q Turn, turn, that I may stare at you!
\par }{\SH {\BD{
{\IT{The Beloved to Her Lover:}}}}
\par }{\Q Why do you gaze upon the Perfect One
\par }{\Q like the dance of the Mahanaim?

\par }\Chap{7}{\SH 
{\QT{The Lover to His Beloved:}}
\par }{\Q \VerseOne{1}How beautiful are your sandaled feet,
\par }{\Q O nobleman’s daughter!
\par }{\Q The curves of your thighs are like jewels,
\par }{\Q the work of the hands of a master craftsman.
\par }{\Q \VS{2}Your navel
is a round
mixing bowl –
\par }{\Q may it never
lack
mixed wine!

\par }{\Q Your belly
is a mound
of wheat,
\par }{\Q encircled
by lilies.
\par }{\Q \VS{3}Your two
breasts
are like two
fawns,
\par }{\Q twins
of a gazelle.
\par }{\Q \VS{4}Your neck
is like a tower
made of ivory.
\par }{\Q Your eyes
are the pools
in
Heshbon
\par }{\Q by the gate
of Bath-Rabbim.
\par }{\Q Your nose
is like the tower
of Lebanon
\par }{\Q overlooking
Damascus.
\par }{\Q \VS{5}Your head
crowns you like Mount Carmel.
\par }{\Q The locks
of your hair
are like royal tapestries – ‘span class=”s 04428” title=”04428” ’
\par }{\Q the king
is held captive
in its tresses!
\par }{\PP \par }{\Q \VS{6}How
beautiful
you are! How
lovely,
\par }{\Q O love,
with your delights!
\par }{\SH The Palm Tree and the Palm Tree Climber
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{7}Your stature
is like a palm
tree,

\par }{\Q and your breasts
are like clusters of grapes.
\par }{\Q \VS{8}I want
to climb
the palm tree,
\par }{\Q and take hold
of its fruit stalks.
\par }{\Q May
your breasts
be
like the clusters
of grapes,
\par }{\Q and may the fragrance
of your breath
be like apricots!
\par }{\Q \VS{9}May your mouth
be like the best
wine,
\par }{\Q flowing smoothly
for my beloved,
\par }{\Q gliding
gently
over our lips
as we
sleep together.
\par }{\SH Poetic Refrain: Mutual Possession
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VS{10}I
am my beloved’s,
\par }{\Q and he desires me!
\par }{\SH The Journey to the Countryside
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{11}Come,
my beloved,
let us go
to the countryside;
\par }{\Q let us spend
the night in the villages.
\par }{\Q \VS{12}Let us rise early
to go to the vineyards,
\par }{\Q to see
if
the vines
have budded,
\par }{\Q to see if their blossoms
have opened,
\par }{\Q if the pomegranates
are in bloom
–
\par }{\Q there
I will give
you my love.
\par }{\Q \VS{13}The mandrakes
send out
their fragrance;
\par }{\Q over
our door
is every
delicacy,
\par }{\Q both
new
and old,
which I have stored up
for you, my lover.

\par }\Chap{8}{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VerseOne{1}Oh, how I wish you were my little brother,
\par }{\Q nursing
at my mother’s
breasts;
\par }{\Q if I saw
you outside,
I could kiss you –
\par }{\Q surely no one would despise me!
\par }{\Q \VS{2}I would lead
you and bring
you to
my mother’s
house,
\par }{\Q the one who taught
me.

\par }{\Q I would give you spiced
wine
to drink,
\par }{\Q the nectar
of my pomegranates.
\par }{\SH Double Refrain: Embracing and Adjuration
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved about Her Lover:}}}}}}
\par }{\Q \VS{3}His left hand
caresses
my head,
\par }{\Q and his right hand
stimulates me.
\par }{\SH 
{\QT{
The Beloved to the Maidens:}}
\par }{\Q \VS{4}I admonish
you, O maidens
of Jerusalem:
\par }{\Q “Do
not arouse
or awaken love
until
it pleases!”
\par }{\SH The Awakening of Love
\par }{\SH 
{\BD{
{\QT{
{\IT{The Maidens about His Beloved:}}}}}}
\par }{\Q \VS{5}Who
is this
coming up
from
the desert,
\par }{\Q leaning
on
her beloved?
\par }{\SH {\QT{The Beloved}}
to Her Lover:
\par }{\Q Under
the apple tree
I aroused
you;

\par }{\Q there
your mother
conceived
you,
\par }{\Q there
she who bore you was in labor
of childbirth.
\par }{\SH The Nature of True Love
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{6}Set
me like a cylinder seal
over
your heart,
\par }{\Q like a signet
on
your arm.
\par }{\Q For
love
is as strong
as death,
\par }{\Q passion
is as unrelenting
as Sheol.
\par }{\Q Its flames burst forth,
\par }{\Q it is a blazing
flame.
\par }{\Q \VS{7}Surging
waters
cannot
quench
love;
\par }{\Q floodwaters
cannot
overflow
it.
\par }{\Q If
someone
were to offer
all
his possessions
to buy love,
\par }{\Q the offer would be utterly
despised.
\par }{\SH The Brother’s Plan and the Sister’s Reward
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved’s Brothers:}}}}}}
\par }{\Q \VS{8}We have a little
sister,
\par }{\Q and as yet she has no
breasts.
\par }{\Q What
shall we do
for our sister
\par }{\Q on the day
when she is spoken for?
\par }{\Q \VS{9}If
she is
a wall,
\par }{\Q we will build
on
her a battlement
of silver;
\par }{\Q but if
she is a door,
\par }{\Q we will barricade
her with boards
of cedar.
\par }{\SH 
{\QT{
The Beloved:}}
\par }{\Q \VS{10}I
was a wall,
\par }{\Q and my breasts
were like fortress towers.
\par }{\Q Then
I found
favor
in his eyes.
\par }{\SH Solomon’s Vineyard and the Beloved’s Vineyard
\par }{\SH 
{\BD{
{\QT{
{\IT{The Beloved to Her Lover:}}}}}}
\par }{\Q \VS{11}Solomon
had
a vineyard
at Baal-Hamon;
\par }{\Q he leased
out the
vineyard
to those who maintained
it.
\par }{\Q Each
was to bring
a thousand
shekels
of silver
for its fruit.
\par }{\Q \VS{12}My vineyard,
which belongs to me, is at my disposal
alone.

\par }{\Q The thousand
shekels belong to you, O Solomon,
\par }{\Q and two hundred
shekels belong to those who maintain
it for its fruit.
\par }{\SH Epilogue: The Lover’s Request and His Beloved’s Invitation
\par }{\SH 
{\BD{
{\QT{
{\IT{The Lover to His Beloved:}}}}}}
\par }{\Q \VS{13}O you who stay
in the gardens,
\par }{\Q my companions
are listening attentively
for your voice;
\par }{\Q let me be the one to hear it!
\par }{\SH 
{\QT{
The Beloved to Her Lover:}}
\par }{\Q \VS{14}Make haste,
my beloved!
\par }{\Q Be like
a gazelle
or
a young
stag
\par }{\PP on
the mountains
of spices.
\par }