\NormalFont\ShortTitle{Zechariah}
{\MT Zechariah

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \VerseOne{1}In the eighth
month
of Darius’
second
year,
the word
of the {\ND{Lord}}
came to
the prophet
Zechariah,
son
of Berechiah
son
of Iddo,
as follows:
\par }{\PP \VS{2}The
{\ND{Lord}}
was very angry
with your ancestors.
\VS{3}Therefore
say
to the
people: The
{\ND{Lord}}
who rules over all
says, “Turn
to me,”
says
the {\ND{Lord}}
who rules over all,
“and I will turn
to
you,” says
the {\ND{Lord}}
who rules over all.
\VS{4}“Do not
be
like your ancestors,
to whom
the former
prophets
called
out, saying,
‘The

{\ND{Lord}}
who rules over all
says, “Turn
now
from your evil wickedness,” ’
but they would
by
no
means
obey
me,”
says
the {\ND{Lord}}.
\VS{5}“As for your ancestors,
where
are they? And did the prophets
live
forever?
\VS{6}But
have my words
and statutes,
which
I commanded
my servants
the prophets,
not
outlived
your fathers? Then they paid
attention and confessed,
‘The
{\ND{Lord}}
who rules over all
has indeed done
what he said he would do
to us, because
of our
sinful ways.’ ”
\par }{\SH The Introduction to the Visions
\par }{\PP \VS{7}On the twenty-fourth
day
of the eleventh
month,
the month
{\IT{
Shebat}}, in Darius’
second
year,
the word
of the {\ND{Lord}}
came
to
the prophet
Zechariah
son
of Berechiah
son
of Iddo,
as follows:
\par }{\SH The Content of the First Vision
\par }{\PP \VS{8}I
was attentive
that night
and saw
a man
seated on
a red
horse
that
stood
among
some myrtle
trees in the ravine.
Behind him
were red,
sorrel,
and white
horses.
\par }{\SH The Interpretation of the First Vision
\par }{\PP \VS{9}Then I asked one nearby, “What
are these,
sir?” The angelic
messenger
who replied
to me
said,
“I
will show
you what
these are.”
\VS{10}Then
the man
standing
among
the myrtle
trees spoke up and said,
“These
are the ones whom
the {\ND{Lord}}
has sent
to walk
about on the earth.”
\VS{11}The riders
then agreed with the angel
of the
{\ND{Lord}}, who was standing
among
the myrtle
trees, “We have been walking
about on
the earth,
and now
everything
is at rest and quiet.”
\VS{12}The angel
of the {\ND{Lord}}
then asked,
“{\ND{Lord}}
who rules over all,
how long
before you
have compassion
on Jerusalem
and the other cities
of Judah
which
you have been so angry
with for these
seventy
years?”
\VS{13}The
{\ND{Lord}}
then addressed
good,
comforting
words
to the angelic messenger
who was speaking to me.
\VS{14}Turning to me,
the messenger
then
said,
“Cry
out that the

{\ND{Lord}}
who rules over all
says, ‘I am very much
moved for Jerusalem
and for Zion.
\VS{15}But I am
greatly
displeased
with the nations
that take my grace
for granted. I
was a little
displeased
with them,
but
they
have only made things
worse for themselves.
\par }{\SH The Oracle of Response
\par }{\PP \VS{16}“‘Therefore,’
says
the {\ND{Lord}}, ‘I have become compassionate
toward Jerusalem
and will rebuild
my temple
in it,’ says
the {\ND{Lord}}
who rules over all.
‘Once more a surveyor’s measuring line
will be stretched
out over
Jerusalem.’
\VS{17}Speak up again
with the message
of the {\ND{Lord}}
who rules over all: ‘My cities
will once more
overflow
with prosperity,
and once more
the {\ND{Lord}}
will comfort
Zion
and validate
his choice
of Jerusalem.’ ”
\par }{\SH Vision Two: The Four Horns
\par }{\PP \VS{18} Once again I looked
and this time I saw
four
horns.
\VS{19}So I asked
the angelic
messenger
who spoke
with me, “What
are these?” He replied,
“These
are the horns
that
have scattered
Judah,
Israel,
and Jerusalem.”
\VS{20}Next
the {\ND{Lord}}
showed
me four
blacksmiths.
\VS{21}I asked,
“What
are these
going
to do?” He answered,
“These
horns
are the ones that
have scattered
Judah
so that there
is no
one
to be seen. But the blacksmiths
have come
to terrify
Judah’s enemies and cut off the
horns
of the nations
that have thrust themselves against
the land
of Judah
in order to scatter its people.”

\par }\Chap{2}{\PP \VerseOne{1}I looked
again, and there
was a man
with a measuring
line
in his hand.
\VS{2}I asked,
“Where
are you
going?” He replied,
“To
measure
Jerusalem
in order to determine
its width
and its length.”
\VS{3}At this point
the angelic messenger
who spoke
to me went out,
and another
messenger
came
to meet him
\VS{4}and said
to him,
“Hurry,
speak
to
this
young man
as follows: ‘Jerusalem
will no longer be enclosed by walls
because of the multitude
of people
and animals
there.
\VS{5}But I
(the
{\ND{Lord}}
says) will be
a wall
of fire
surrounding
Jerusalem and the source of glory
in her midst.’ ”
\par }{\PP \VS{6}“You there! Flee
from the northland!” says
the {\ND{Lord}}, “for
like the four
winds
of heaven
I have scattered
you,” says
the {\ND{Lord}}.
\VS{7}“Escape,
Zion,
you who live
among the Babylonians!”
\VS{8}For
the

{\ND{Lord}}
who rules over all
says to me that for his own
glory
he has sent
me to
the nations
that plundered you – for anyone who touches you touches the pupil of his eye.
\VS{9}“I am
about to punish
them in such a way,” he says, “that they will be
looted
by
their own slaves.”
Then you will know
that
the {\ND{Lord}}
who rules over all
has sent me.
\par }{\PP \VS{10}“Sing
out and be happy,
Zion
my daughter! For
look,
I have come;
I will settle
in your midst,”
says
the {\ND{Lord}}.
\VS{11}“Many
nations
will join
themselves to
the {\ND{Lord}}
on the day
of salvation, and they will also be
my people.
Indeed, I will settle
in the midst
of you all.” Then you will know
that
the {\ND{Lord}}
who rules over all
has sent
me to you.
\VS{12}The
{\ND{Lord}}
will take possession
of Judah
as his portion
in the holy
land
and he will choose
Jerusalem
once again.
\VS{13}Be silent
in the
{\ND{Lord}}’s
presence,
all
people everywhere,
for
he is being moved to action
in his holy
dwelling place.

\par }\Chap{3}{\PP \VerseOne{1}Next I saw
Joshua
the high
priest
standing
before
the angel
of the {\ND{Lord}}, with Satan
standing
at his right hand
to accuse him.
\VS{2}The
{\ND{Lord}}
said
to
Satan,
“May the
{\ND{Lord}}
rebuke
you, Satan! May
the {\ND{Lord}}, who has chosen
Jerusalem,
rebuke you! Isn’t
this
man like a burning stick snatched
from
the fire?”
\VS{3}Now Joshua
was dressed
in filthy
clothes
as he stood
there before
the angel.
\VS{4}The angel
spoke
up to
those standing
all around, “Remove
his filthy
clothes.”
Then he said
to
Joshua, “I have freely
forgiven
your iniquity
and will dress
you in fine clothing.”
\VS{5}Then I spoke up, “Let a clean
turban
be put
on
his head.”
So they put
a clean
turban
on
his head
and clothed
him, while the angel
of the {\ND{Lord}}
stood nearby.
\VS{6}Then the angel
of the
{\ND{Lord}}
exhorted
Joshua
solemnly:
\VS{7}“The
{\ND{Lord}}
who rules over all
says,
‘If
you live
and work
according to my requirements,
you
will be able to
preside
over my temple
and attend
to my courtyards,
and I will allow
you to come and go
among
these
others who are standing by you.
\VS{8}Listen
now,
Joshua
the high
priest,
both you
and your colleagues
who are sitting
before
you, all of you
are a symbol
that
I am about to introduce
my servant,
the Branch.
\VS{9}As for
the stone
I have set
before
Joshua
– on
the one
stone
there are seven
eyes.
I am
about to engrave
an inscription
on it,’ says
the {\ND{Lord}}
who rules over all,
‘to the effect that I will remove
the
iniquity
of this land
in a single
day.
\VS{10}In that day,’
says
the {\ND{Lord}}
who rules over all,
‘everyone
will invite
his friend
to fellowship
under
his vine
and under
his fig tree.’ ”

\par }\Chap{4}{\PP \VerseOne{1}The angelic messenger
who had been
speaking
with me then returned
and woke me, as a person
is wakened
from sleep.
\VS{2}He asked
me,
“What
do you
see?” I replied, “I see
a menorah
of pure gold
with a receptacle
at
the top
and seven
lamps,
with fourteen
pipes
going to the lamps.
\VS{3}There are also two
olive
trees beside
it, one
on
the right
of the receptacle
and the other
on
the left.”
\VS{4}Then
I asked
the messenger
who spoke
with me, “What
are these,
sir?”
\VS{5}He
replied,
“Don’t
you know
what
these
are?” So I responded,
“No,
sir.”
\VS{6}Therefore
he told
me,
“These signify
the word
of the {\ND{Lord}}
to
Zerubbabel: ‘Not
by strength
and not
by power,
but
by my Spirit,’
says
the {\ND{Lord}}
who rules over all.”
\par }{\SH Oracle of Response
\par }{\PP \VS{7}“What
are you,
you great
mountain? Because
of Zerubbabel
you will become a level plain! And he will bring forth
the temple capstone
with shoutings
of ‘Grace! Grace!’ because of this.”
\VS{8}Moreover, the word
of the {\ND{Lord}}
came
to
me as follows:
\VS{9}“The hands
of Zerubbabel
have laid the foundations
of this
temple,
and his hands
will complete
it.” Then you will know
that
the {\ND{Lord}}
who rules over all
has sent
me to you.
\VS{10}For
who
dares
make light of small
beginnings? These seven
eyes will joyfully
look
on the tin
tablet
in Zerubbabel’s
hand.
(These
are the eyes
of the {\ND{Lord}}, which constantly range across
the whole
earth.)
\par }{\PP \VS{11}Next
I asked
the messenger,
“What
are these
two
olive
trees on
the right
and the left
of the menorah?”
\VS{12}Before he could reply
I asked
again,
“What
are these two
extensions
of the olive
trees, which
are emptying
out the golden
oil
through
the two
golden
pipes?”
\VS{13}He replied,
“Don’t
you know
what
these
are?” And I said,
“No,
sir.”
\VS{14}So he said,
“These
are the two
anointed
ones who stand
by the Lord
of the whole
earth.”

\par }\Chap{5}{\PP \VerseOne{1}Then I turned
to look,
and there
was a flying
scroll!
\VS{2}Someone
asked
me,
“What
do you
see?” I replied,
“I
see
a flying
scroll
thirty feet
long
and fifteen feet
wide.”
\VS{3}The speaker
went
on
to
say,
“This
is a curse
traveling
across the
whole
earth.
For
example, according to the
curse whoever
steals
will be removed from the community; or on
the other
hand (according to the curse) whoever
swears falsely will suffer the same fate.”
\VS{4}“I will send it out,”
says
the
{\ND{Lord}}
who rules over all,
“and it will enter
the house
of the thief
and of the person who swears
falsely
in my name.
It will land
in the middle
of his house
and destroy
both timber
and stones.”
\par }{\SH Vision Seven: The Ephah
\par }{\PP \VS{5}After
this
the angelic
messenger
who had been speaking
to me went out
and said,
“Look,
see
what
is leaving.”
\VS{6}I asked,
“What
is it?” And he replied,
“It is a basket
for measuring grain that is moving
away from
here.” Moreover, he said,
“This
is their ‘eye’
throughout all
the earth.”
\VS{7}Then
a round
lead
cover was raised up, revealing
a woman
sitting
inside
the basket.
\VS{8}He then said,
“This
woman represents wickedness,”
and he pushed
her down
into
the basket
and placed
the lead
cover
on
top.
\VS{9}Then I looked
again and saw
two
women
going forth
with the wind
in their wings
(they
had wings
like those
of a stork) and they lifted up
the
basket
between
the earth
and the sky.
\VS{10}I asked
the messenger
who was speaking
to me, “Where
are they
taking
the basket?”
\VS{11}He replied,
“To
build
a temple
for her in the land
of Babylonia.
When it
is finished,
she will be placed
there
in her own residence.”

\par }\Chap{6}{\PP \VerseOne{1}Once more I looked,
and this time I saw
four
chariots
emerging
from between
two
mountains
of bronze.
\VS{2}Harnessed to the first
chariot
were red
horses,
to the second
black
horses,
\VS{3}to the third
white
horses,
and to the fourth
spotted
horses,
all of them strong.
\VS{4}Then
I asked
the angelic
messenger
who was speaking
with me, “What
are these,
sir?”
\VS{5}The messenger
replied,
“These
are the four
spirits
of heaven
that have been presenting
themselves
before the Lord
of all
the earth.
\VS{6}The chariot with the black
horses
is going
to
the north
country
and the white
ones are going
after
them, but the spotted
ones are going
to
the south
country.
\VS{7}All these strong ones
are scattering;
they have sought
permission to go
and walk
about over the earth.”
The Lord had said,
“Go! Walk
about over the earth!” So they are doing so.
\VS{8}Then he cried out
to
me, “Look! The ones going
to
the northland
have brought me peace
about the northland.”
\par }{\SH A Concluding Oracle
\par }{\PP \VS{9}The word
of the {\ND{Lord}}
came to me
as follows:
\VS{10}“Choose
some people from among the exiles,
namely, Heldai,
Tobijah,
and Jedaiah,
all of whom
have come
from Babylon,
and when you
have done so go
to the house
of Josiah
son
of Zephaniah.
\VS{11}Then take
some silver
and gold
to make
a crown
and set
it on the head
of Joshua
son
of Jehozadak,
the high
priest.
\VS{12}Then say
to him,
‘The
{\ND{Lord}}
who rules over all
says,
“Look
– here is the man
whose name
is Branch,
who will sprout
up from his place
and build
the
temple
of the {\ND{Lord}}.
\VS{13}Indeed, he will build
the
temple
of the {\ND{Lord}},
and he
will be clothed
in splendor,
sitting
as king on
his throne.
Moreover, there will be a priest
with him on
his throne
and they will see eye
to eye on everything.
\VS{14}The crown
will then be turned over to Helem,
Tobijah,
Jedaiah,
and Hen
son
of Zephaniah
as a memorial
in the temple
of the {\ND{Lord}}.
\VS{15}Then those who are far
away will come
and build
the temple
of the {\ND{Lord}}
so that you may know
that
the {\ND{Lord}}
who rules over all
has sent
me to
you. This will all come to
pass if
you completely obey
the voice
of the {\ND{Lord}}
your God.” ’ ”

\par }\Chap{7}{\PP \VerseOne{1}In King
Darius’
fourth
year,
on the fourth
{\IT{day of
Kislev}},
the ninth
month,
the word
of the {\ND{Lord}}
came to
Zechariah.
\VS{2}Now the people of Bethel
had sent
Sharezer
and Regem-Melech
and their companions
to seek the
{\ND{Lord}}’s favor
\VS{3}by asking
both the priests
of the temple
of the
{\ND{Lord}}
who rules over all
and the prophets,
“Should we weep
in the fifth
month,
fasting
as we have
done
over
the years?”
\VS{4}The word
of the {\ND{Lord}}
who rules over all
then
came to me,
\VS{5}“Speak to
all
the people
and priests
of the land
as follows: ‘When
you fasted
and lamented
in the fifth
and seventh
months through all these
seventy
years,
did you truly fast for me – for me, indeed?
\VS{6}And now when
you eat
and drink,
are you not
doing so for yourselves?’ ”
\VS{7}Should you not
have obeyed the words
that
the {\ND{Lord}}
cried
out through
the former
prophets
when
Jerusalem
was peacefully
inhabited
and her surrounding
cities,
the Negev,
and the Shephelah
were also populated?
\par }{\PP \VS{8}Again
the word
of the {\ND{Lord}}
came to
Zechariah:
\VS{9}“The
{\ND{Lord}}
who rules over all
said,
‘Exercise
true
judgment
and show
brotherhood
and compassion
to
each
other.
\VS{10}You must not
oppress
the widow,
the orphan,
the foreigner,
or the poor,
nor should
anyone secretly
plot
evil
against his fellow
human being.’
\par }{\PP \VS{11}“But they refused
to pay attention,
turning away stubbornly
and stopping
their ears
so they could not hear.
\VS{12}Indeed, they made
their heart
as hard as diamond,
so that they could not obey
the Torah
and the
other words
the

{\ND{Lord}}
who rules over all
had
sent
by his Spirit
through
the former
prophets.
Therefore, the
{\ND{Lord}}
who rules over all
had poured out great
wrath.
\par }{\PP \VS{13}“‘It then came
about that just
as I cried
out, but they would not
obey,
so
they will cry
out, but I will not
listen,’
the
{\ND{Lord}}
{\ND{Lord}}
who rules over all
had said.
\VS{14}‘Rather, I will sweep
them away in a storm
into
all
the nations
they are not
familiar
with.’ Thus the land
had become desolate
because
of them, with no one crossing through
or returning,
for they had made
the fruitful
land
a waste.”

\par }\Chap{8}{\PP \VerseOne{1}Then
the word
of the {\ND{Lord}}
who rules over all
came to me as follows:
\VS{2}“The

{\ND{Lord}}
who rules over all
says, ‘I am very much
concerned for Zion;
indeed, I am so concerned
for her that my rage
will fall
on those who hurt her.’
\VS{3}The Lord
says,
‘I have returned
to
Zion
and will live
within
Jerusalem.
Now Jerusalem
will be called
“truthful
city,”
“mountain
of the {\ND{Lord}}
who rules over all,”
“holy
mountain.” ’
\VS{4}Moreover,
the {\ND{Lord}}
who rules over all
says,
‘Old men
and women
will once more
live
in the plazas
of Jerusalem,
each
one leaning
on a cane
because of advanced
age.
\VS{5}And the streets
of the city
will
be full
of boys
and girls
playing.
\VS{6}And,’ says
the
{\ND{Lord}}
who rules over all,
‘though
such a thing may seem
to be difficult
in
the opinion
of the small community of those
days,
will it also
appear difficult
to me?’ asks
the
{\ND{Lord}}
who rules over all.
\par }{\PP \VS{7}“The
{\ND{Lord}}
who rules over all
asserts,
‘I am about to save
my people
from the lands
of the east
and the west.
\VS{8}And I will bring
them
to settle
within
Jerusalem.
They will be
my people,
and I
will be
their God,
in truth
and righteousness.’
\par }{\PP \VS{9}“The

{\ND{Lord}}
who rules over all
also says, ‘Gather strength,
you who are
listening
to these
words
today
from the mouths
of the prophets
who
were
there at
the founding
of the house
of the {\ND{Lord}}
who rules over all,
so that the temple
might be built.
\VS{10}Before
that
time
there was
no
compensation
for man
or animal,
nor
was there any relief from adversity
for those who came
and went,
because I had pitted
everybody
– each one – against everyone else.
\VS{11}But I will be different
now
to this
remnant
of my people
from the way
I
was in those days,’
says
the {\ND{Lord}}
who rules over all,
\VS{12}‘for
there
will be
a peaceful
time of sowing,
the vine
will produce
its fruit
and the ground
its yield,
and the skies
will
rain down dew.
Then I will allow
the
remnant
of my people
to possess all
these things.
\VS{13}And it will
come about that just
as you (both Judah
and Israel) were a curse
to the nations,
so
I will save
you and you will be
a blessing.
Do not
be afraid! Instead, be strong!’
\par }{\PP \VS{14}“For
the

{\ND{Lord}}
who rules over all
says, ‘As I had
planned
to hurt
you when your fathers
made me angry,’
says
the {\ND{Lord}}
who rules over all,
‘and I was not
sorry,
\VS{15}so,
to the contrary, I have planned
in these
days
to do good
to Jerusalem
and Judah
– do not
fear!
\VS{16}These
are the things
you must do: Speak
the truth,
each
of you,
to one another.
Practice
true
and righteous
judgment
in your courts.
\VS{17}Do not
plan
evil
in your hearts
against one another.
Do not
favor
a false
oath
– these
are all
things that
I hate,’
says
the {\ND{Lord}}.”
\par }{\PP \VS{18}The word
of the {\ND{Lord}}
who rules over all
came
to
me as follows:
\VS{19}“The

{\ND{Lord}}
who rules over all
says,
‘The fast
of the fourth,
fifth,
seventh,
and tenth
months will become
joyful
and happy,
pleasant
feasts
for the house
of Judah,
so love
truth
and peace.’
\VS{20}The

{\ND{Lord}}
who rules over all
says, ‘It will someday
come
to pass that people
– residents
of many
cities – will come.
\VS{21}The inhabitants
of one
will go
to
another
and say,
“Let’s
go
up at once to ask the favor
of the

{\ND{Lord}}, to seek
the

{\ND{Lord}}
who rules over all.
Indeed,
I’ll
go with you.” ’
\VS{22}Many
peoples
and powerful
nations
will come
to Jerusalem
to seek
the {\ND{Lord}}
who rules over all
and to ask his favor.
\VS{23}The
{\ND{Lord}}
who rules over all
says,
‘In those
days
ten
people
from all
languages
and nations
will grasp hold of – indeed, grab – the robe of one Jew and say, “Let us go with you, for we have heard that God is with you.” ’ ”

\par }\Chap{9}{\PP \VerseOne{1}An oracle
of the word
of the
{\ND{Lord}}
concerning the land
of Hadrach,
with its focus
on Damascus:
\par }{\PP The eyes
of all
humanity,
especially of the tribes
of Israel,
are toward the
{\ND{Lord}},
\VS{2}as are those of Hamath
also,
which adjoins
Damascus, and Tyre
though
they consider themselves to be very
wise.
\VS{3}Tyre
built
herself a fortification
and piled up
silver
like dust
and gold
like the mud
of the streets!
\VS{4}Nevertheless
the Lord
will evict her and shove
her fortifications
into the sea
– she
will be consumed
by fire.
\VS{5}Ashkelon
will see
and be afraid;
Gaza
will be in great
anguish,
as will Ekron,
for
her hope
will have been dried up.
Gaza
will lose
her king,
and Ashkelon
will no
longer be inhabited.
\VS{6}A mongrel
people will live in
Ashdod,
for I will greatly humiliate
the Philistines.
\VS{7}I will take away
their abominable
religious practices;
then those who survive
will become a community of believers in our God,
like
a clan in Judah,
and Ekron
will be like the Jebusites.
\VS{8}Then I will surround
my temple
to protect it like a guard
from anyone crossing back
and forth;
so no one
will cross over
against
them anymore
as an oppressor,
for
now
I myself have seen it.
\par }{\Q \VS{9}Rejoice
greatly,
daughter
of Zion!
\par }{\Q Shout,
daughter
of Jerusalem!
\par }{\Q Look! Your king
is coming
to you:
\par }{\Q he is legitimate
and victorious,
\par }{\Q humble
and riding
on
a donkey –
\par }{\Q on
a young donkey,
the foal
of a female donkey.
\par }{\Q \VS{10}I will remove
the chariot
from Ephraim
\par }{\Q and the warhorse
from Jerusalem,
\par }{\Q and the battle
bow
will be removed.
\par }{\Q Then he will announce
peace
to the nations.
\par }{\Q His dominion
will be from sea
to
sea
\par }{\Q and from the Euphrates River
to
the ends
of the earth.
\par }{\PP \VS{11}Moreover, as
for you,
because of our covenant
relationship secured with blood,
I will release
your prisoners
from the waterless
pit.
\VS{12}Return
to the stronghold,
you prisoners,
with hope;
today
I declare
that I will return
double what was taken from you.
\VS{13}I
will bend
Judah
as my bow;
I will load
the bow with
Ephraim,
my arrow! I will stir
up your sons,
Zion,
against
yours, Greece,
and I will make
you, Zion, like a warrior’s
sword.
\par }{\PP \VS{14}Then the
{\ND{Lord}}
will appear
above them, and his arrow
will shoot forth
like lightning;
the Lord

{\ND{God}}
will blow
the trumpet
and will sally
forth on the southern
storm
winds.
\VS{15}The
{\ND{Lord}}
who rules over all
will guard
them, and they will prevail
and overcome
with sling
stones.
Then they will drink,
and will become noisy
like
drunkards,
full
like the sacrificial basin
or like the corners
of the altar.
\VS{16}On that day
the {\ND{Lord}}
their God
will deliver
them as the flock
of his people,
for
they are the precious stones
of a crown
sparkling
over
his land.
\VS{17}How
precious
and fair! Grain
will make the young men
flourish
and new wine
the young women.

\par }\Chap{10}{\PP \VerseOne{1}Ask
the
{\ND{Lord}}
for rain
in the season
of the late spring rains –
the
{\ND{Lord}}
who causes thunderstorms
– and he will give
everyone
showers
of rain
and green growth
in the field.
\VS{2}For
the household gods
have spoken
wickedness,
the soothsayers
have seen
a lie,
and as for the dreamers,
they have disclosed
emptiness
and give comfort
in
vain.
Therefore
the people set out
like
sheep
and become scattered
because
they have no
shepherd.
\VS{3}I am enraged
at the shepherds
and will punish
the lead-goats.
\par }{\PP For
the {\ND{Lord}}
who rules over all
has brought blessing
to his flock,
the
house
of Judah,
and will transform
them into his majestic
warhorse.
\VS{4}From
him will come the cornerstone,
the wall peg,
the battle
bow,
and every
ruler.
\VS{5}And they will be
like warriors
trampling
the mud
of the streets
in battle.
They will fight,
for
the {\ND{Lord}}
will be with
them, and will defeat
the enemy cavalry.
\par }{\PP \VS{6}“I (says the
{\ND{Lord}}) will strengthen
the kingdom
of Judah
and deliver
the people
of Joseph
and will bring them back
because
of my compassion
for them. They will be
as though I had
never
rejected
them, for
I am
the {\ND{Lord}}
their God
and therefore I will hear them.
\VS{7}The Ephraimites
will be
like warriors
and will rejoice
as if
they had drunk wine.
Their children
will see
it and rejoice;
they will celebrate
in the things of the
{\ND{Lord}}.
\VS{8}I will signal
for them and gather
them, for
I have already redeemed
them; then they will become
as
numerous
as
they were before.
\VS{9}Though I scatter
them among
the nations,
they will remember
in far-off places
– they and their children
will sprout forth and return.
\VS{10}I will bring them back
from Egypt
and gather
them from Assyria.
I will bring
them to
the lands
of Gilead
and Lebanon,
for there will not
be enough room for them in their own land.
\VS{11}The
{\ND{Lord}} will cross
the sea
of storms
and will calm
its
turbulence.
The depths
of the Nile
will dry up,
the pride
of Assyria
will be humbled,
and the domination
of Egypt
will be no more.
\VS{12}Thus I will strengthen
them by my power, and they will walk
about in my name,”
says
the {\ND{Lord}}.

\par }\Chap{11}{\PP \par }{\Q \VerseOne{1}Open
your gates,
Lebanon,
\par }{\Q so that the fire
may consume
your cedars.
\par }{\Q \VS{2}Howl,
fir
tree,
\par }{\Q because
the cedar
has
fallen;
\par }{\Q the majestic
trees have been destroyed.
\par }{\Q Howl,
oaks
of Bashan,
\par }{\Q because
the impenetrable
forest
has fallen.
\par }{\Q \VS{3}Listen
to the howling
of shepherds,
\par }{\Q because
their magnificence
has been destroyed.
\par }{\Q Listen
to the roaring
of young lions,
\par }{\Q because
the thickets
of the Jordan
have been devastated.
\par }{\PP \VS{4}The
{\ND{Lord}}
my God
says this: “Shepherd
the flock
set aside for slaughter.
\VS{5}Those who
buy
them slaughter
them and are not
held guilty;
those who sell
them say,
‘Blessed
be the
{\ND{Lord}}, for I am rich.’
Their own shepherds
have no
compassion for them.
\VS{6}Indeed,
I will no
longer
have compassion
on
the people
of the land,”
says
the {\ND{Lord}}, “but
instead I
will turn every last
person
over
to his neighbor
and his
king.
They will devastate
the
land,
and I will not
deliver it from them.”
\par }{\PP \VS{7}So I began to shepherd
the
flock
destined
for slaughter,
the most afflicted
of all the flock.
Then I took
two
staffs,
calling
one
“Pleasantness”
and the other
“Binders,”
and I tended
the
flock.
\VS{8}Next I eradicated
the three
shepherds
in one
month,
for I ran out
of patience
with them and, indeed, they detested
me as well.
\VS{9}I then said,
“I will not
shepherd
you. What is to die,
let it die,
and what is to be eradicated,
let it be eradicated.
As for those who survive,
let them eat
each other’s
flesh!”
\par }{\PP \VS{10}Then I took
my staff
“Pleasantness”
and cut
it in two to annul
my covenant
that I had
made
with
all
the people.
\VS{11}So it was
annulled
that very day,
and then
the most afflicted
of the flock
who kept
faith with me knew
that that
was the word
of the {\ND{Lord}}.
\par }{\PP \VS{12}Then I said
to
them, “If
it seems
good
to you, pay
me my wages,
but if
not,
forget
it.” So they weighed out
my payment
– thirty
pieces of silver.
\VS{13}The
{\ND{Lord}}
then said
to
me, “Throw
to
the potter
that exorbitant
sum
at which
they valued
me!” So
I took
the thirty
pieces of silver
and threw
them to
the
potter
at the temple
of the
{\ND{Lord}}.
\VS{14}Then I cut
the second
staff
“Binders”
in two in order to annul
the covenant of brotherhood
between
Judah
and Israel.
\par }{\PP \VS{15}Again
the {\ND{Lord}}
said
to me,
“Take
up once more
the equipment
of a foolish
shepherd.
\VS{16}Indeed,
I
am about to raise up
a shepherd
in the land
who will not
take heed
to the sheep headed to slaughter,
will not
seek
the scattered,
and will not
heal
the injured.
Moreover, he will not
nourish
the one that is healthy
but instead will eat
the meat
of the fat
sheep and tear off
their hooves.
\par }{\Q \VS{17}Woe
to the worthless
shepherd
\par }{\Q who abandons
the flock!
\par }{\Q May a sword
fall on
his arm
and his right
eye!
\par }{\Q May his
arm
wither
completely away,
\par }{\Q and his right
eye
become completely
blind!”

\par }\Chap{12}{\PP \VerseOne{1}The revelation
of the word
of the {\ND{Lord}}
concerning
Israel: The
{\ND{Lord}} –
he who stretches
out the heavens
and lays the foundations
of the earth,
who forms
the human
spirit
within
a person –
says,
\VS{2}“I
am about to make
Jerusalem
a cup
that brings dizziness
to all
the surrounding
nations;
indeed, Judah
will also
be included
when Jerusalem
is besieged.
\VS{3}Moreover,
on that
day
I will make
Jerusalem
a heavy burden
for all
the nations,
and all
who try to carry
it will be seriously injured;
yet all
the peoples
of the earth
will be assembled
against it.
\VS{4}In that day,”
says
the {\ND{Lord}}, “I will strike
every
horse
with confusion
and its rider
with madness.
I will pay close attention to
the house
of Judah,
but will strike
all
the horses
of the nations
with blindness.
\VS{5}Then the leaders
of Judah
will say to themselves, ‘The inhabitants
of Jerusalem
are a means of strength
to us through their God,
the {\ND{Lord}}
who rules over all.’
\VS{6}On that day
I will make
the leaders
of Judah
like an igniter
among sticks
and a burning
torch
among sheaves,
and they will burn up
all
the surrounding
nations right
and left.
Then the people
of Jerusalem
will settle
once more
in their place,
the city of Jerusalem.
\VS{7}The
{\ND{Lord}}
also will deliver
the homes
of Judah
first,
so that
the splendor
of the kingship
of David
and of the people
of Jerusalem
may not
exceed
that of Judah.
\VS{8}On that day
the {\ND{Lord}}
himself
will defend
the inhabitants
of Jerusalem,
so that
the weakest
among them will be like mighty
David,
and the dynasty
of David
will be like God,
like the angel
of the {\ND{Lord}}
before them.
\VS{9}So on that day
I will set
out to destroy
all
the nations
that come
against
Jerusalem.”
\par }{\PP \VS{10}“I will pour
out on
the kingship
of David
and the population
of Jerusalem
a spirit
of grace
and supplication
so
that
they will
look
to
me, the one they
have pierced. They will lament for him
as
one laments
for
an only son,
and there will
be a bitter
cry for him
like the bitter
cry for a firstborn.
\VS{11}On that day
the lamentation
in Jerusalem
will be
as great
as the lamentation
at Hadad-Rimmon
in the plain
of Megiddo.
\VS{12}The land
will mourn,
clan
by clan
– the clan
of the royal household
of David
by itself
and their wives
by themselves;
the clan
of the family
of Nathan
by itself
and their wives
by themselves;
\VS{13}the clan
of the descendants
of Levi
by itself
and their wives
by themselves;
and the clan
of the Shimeites
by itself
and their wives
by themselves –
\VS{14}all
the clans
that remain,
each separately
with their
wives.”

\par }\Chap{13}{\PP \VerseOne{1}“In that day
there will be
a fountain
opened
up for the dynasty
of David
and the people
of Jerusalem
to cleanse them from sin
and impurity.
\VS{2}And also on that day,”
says
the {\ND{Lord}}
who rules over all,
“I will remove
the
names
of the idols
from
the land
and they will never
again
be remembered.
Moreover,
I will remove
the
prophets
and the
unclean
spirit
from
the land.
\VS{3}Then,
if
anyone
prophesies
in spite
of this,
his father
and mother
to whom he was born
will
say
to him,
‘You cannot
live,
for
you lie
in the name
of the {\ND{Lord}}.’ Then
his father
and mother
to whom
he was born
will run him through with a sword when he prophesies.
\par }{\PP \VS{4}“Therefore,
on that
day
each
prophet
will be ashamed
of his vision
when he prophesies
and will no longer
wear
the hairy
garment
of a prophet to
deceive the people.
\VS{5}Instead he will say,
‘I
am no
prophet
– indeed,
I
am a farmer,
for
a man
has made me his indentured
servant since my youth.’
\VS{6}Then someone will ask him,
‘What
are these
wounds
on your chest?’ and he will answer,
‘Some that
I received
in the house
of my friends.’
\par }{\Q \VS{7}“Awake,
sword,
against
my shepherd,
\par }{\Q against
the man
who is my associate,”
\par }{\Q says
the
{\ND{Lord}}
who rules over all.
\par }{\Q Strike
the
shepherd
that the flock
may be scattered;
\par }{\Q I will turn
my hand
against
the insignificant ones.
\par }{\Q \VS{8}It will happen
in all
the land,
says
the {\ND{Lord}},
\par }{\Q that two-thirds
of the people
in it will be cut off
and die,
\par }{\Q but one-third
will be left in it.
\par }{\Q \VS{9}Then I will bring
the remaining third
into the fire;
\par }{\Q I will refine
them like silver
is refined
\par }{\Q and will test
them like gold
is tested.
\par }{\Q They will call
on my name
and I
will answer;
\par }{\Q I will say,
‘These are my people,’
\par }{\Q and they will say,
‘The
{\ND{Lord}}
is my God.’ ”

\par }\Chap{14}{\PP \VerseOne{1}A day
of the {\ND{Lord}}
is about
to come
when your possessions will be divided
as plunder
in your midst.
\VS{2}For I will gather
all
the nations
against Jerusalem
to
wage war;
the city
will be taken,
its houses
plundered,
and the women
raped.
Then
half
of the city
will go into exile,
but the remainder
of the people
will not
be taken away.
\par }{\PP \VS{3}Then
the
{\ND{Lord}}
will go to battle and fight
against those
nations,
just as he fought
battles
in ancient
days.
\VS{4}On that day
his feet
will stand
on
the Mount
of Olives
which
lies to the east
of Jerusalem,
and the Mount
of Olives
will be split
in half
from east
to west,
leaving a great
valley.
Half
the mountain
will move
northward
and the other half
southward.
\VS{5}Then you will escape
through my mountain
valley,
for
the mountains
will extend to
Azal.
Indeed, you will flee
as
you fled
from the earthquake
in the days
of King
Uzziah
of Judah.
Then
the {\ND{Lord}}
my God
will come
with all
his holy ones
with him.
\VS{6}On that
day
there will be
no
light
– the sources of light
in the heavens will congeal.
\VS{7}It will happen
in one
day
(a day known
to the
{\ND{Lord}}); not
in the day
or
the night,
but in the evening
there will be
light.
\VS{8}Moreover,
on that day
living
waters
will flow out
from Jerusalem,
half
of them to
the eastern
sea
and half
of them to
the western sea;
it will
happen both in summer
and in winter.
\par }{\PP \VS{9}The
{\ND{Lord}}
will then be
king
over
all
the earth.
In that day
the {\ND{Lord}}
will be
seen as one
with a single
name.
\VS{10}All
the land
will change
and become like the Arabah
from Geba
to Rimmon,
south
of Jerusalem; and Jerusalem
will be raised up
and will stay
in its own place
from the Benjamin
Gate
to
the site
of the First
Gate
and on to
the Corner
Gate,
and from the Tower
of Hananel
to
the royal
winepresses.
\VS{11}And people will settle
there, and there will no
longer
be the threat of divine
extermination
– Jerusalem
will dwell in security.
\par }{\PP \VS{12}But this
will be
the nature of the plague
with which
the {\ND{Lord}}
will strike
all
the nations
that have
fought
against
Jerusalem: Their flesh
will decay
while
they stand
on
their feet,
their eyes
will rot
away in
their sockets,
and their tongues
will dissolve
in their mouths.
\VS{13}On that day
there will be
great
confusion
from the
{\ND{Lord}}
among them; they will seize
each
other
and attack
one another
violently.
\VS{14}Moreover,
Judah
will fight
at Jerusalem,
and the wealth
of all
the surrounding
nations
will be gathered up –
gold,
silver,
and clothing
in great
abundance.
\VS{15}This is
the kind of plague
that will devastate horses,
mules,
camels,
donkeys,
and all
the other animals
in those
camps.
\par }{\PP \VS{16}Then
all
who survive
from all
the nations
that came
to attack
Jerusalem
will go up
annually
to worship
the King,
the {\ND{Lord}}
who rules over all,
and to observe
the Feast
of Tabernacles.
\VS{17}But if
any of the nations anywhere
on earth
refuse
to go up
to
Jerusalem
to worship
the King,
the {\ND{Lord}}
who rules over all,
they will
get no
rain.
\VS{18}If
the Egyptians
will not
do so, they will get
no rain – instead there will be the kind of plague which the
{\ND{Lord}} inflicts on any nations that do not go up to celebrate the Feast of Tabernacles.
\VS{19}This
will be
the punishment
of Egypt
and of all
nations
that
do not
go up
to celebrate
the Feast
of Tabernacles.
\par }{\PP \VS{20}On that day
the bells
of the horses
will bear the inscription “{\SC{Holy}}
{\SC{to the Lord}}.” The cooking pots
in the
{\ND{Lord}}’s
temple
will be
as holy as the bowls
in front
of the altar.
\VS{21}Every
cooking pot
in Jerusalem
and Judah
will become
holy
in the sight of the
{\ND{Lord}}
who rules over all,
so that all
who offer sacrifices
may come
and use
some of them
to boil
their sacrifices in them. On that day there will no
longer
be
a Canaanite
in the house
of the
{\ND{Lord}}
who rules over all.
\par }