\NormalFont\ShortTitle{Zephaniah}
{\MT Zephaniah

\par }\ChapOne{1}{\SH Introduction
\par }{\PP \VerseOne{1}This
is the prophetic message
that the
{\ND{Lord}}
gave to
Zephaniah
son
of Cushi,
son
of Gedaliah,
son
of Amariah,
son
of Hezekiah.
Zephaniah delivered this
message during
the reign of King
Josiah
son
of Amon
of Judah:
\par }{\SH The Lord’s Day of Judgment is Approaching
\par }{\Q \VS{2}“I will destroy everything
from
the face
of the earth,”
says
the {\ND{Lord}}.
\par }{\Q \VS{3}“I will destroy
people
and animals;
\par }{\Q I will destroy
the birds
in the sky
\par }{\Q and the fish
in the sea.
\par }{\Q (The idolatrous images of these creatures
will be destroyed along with
evil people.)

\par }{\Q I will remove
humanity
from
the
face
of the earth,”
says
the {\ND{Lord}}.
\par }{\Q \VS{4}“I will attack
Judah
\par }{\Q and all
who live
in Jerusalem.
\par }{\Q I will remove
from
this
place
every trace
of Baal
worship,

\par }{\Q as well as the very memory
of the pagan priests.
\par }{\Q \VS{5}I
will remove those who worship
the stars in the sky
from their rooftops,
\par }{\Q those who swear
allegiance to the
{\ND{Lord}}
while taking oaths in the name of their ‘king,’
\par }{\Q \VS{6}and those who turn their backs
on the
{\ND{Lord}}
\par }{\Q and do not
want
the
{\ND{Lord}}’s
help or
guidance.”
\par }{\Q \VS{7}Be silent
before
the Lord

{\ND{God}},

\par }{\Q for
the
{\ND{Lord}}’s
day
of judgment is almost here.
\par }{\Q The
{\ND{Lord}}
has prepared
a sacrificial meal;
\par }{\Q he has ritually purified
his guests.
\par }{\Q \VS{8}“On the day
of the
{\ND{Lord}}’s
sacrificial
meal,
\par }{\Q I will punish
the princes
and the king’s
sons,
\par }{\Q and all
who wear
foreign
styles of clothing.
\par }{\Q \VS{9}On
that day
I will punish
all
who leap
over
the threshold,
\par }{\Q who fill
the house
of their master
with wealth taken by violence
and deceit.
\par }{\Q \VS{10}On that day,”
says
the {\ND{Lord}},
\par }{\Q “a loud
cry
will go up from the Fish
Gate,
\par }{\Q wailing
from
the city’s newer district,
\par }{\Q and a loud
crash
from the hills.
\par }{\Q \VS{11}Wail,
you who live in
the market district,
\par }{\Q for
all
the merchants
will disappear
\par }{\Q and those who
count
money
will be removed.
\par }{\Q \VS{12}At that time
I will search
through Jerusalem
with lamps.
\par }{\Q I will punish
the people
who are entrenched
in their
sin,

\par }{\Q those who think
to themselves,

\par }{\Q ‘The
{\ND{Lord}}
neither
rewards
nor
punishes.’
\par }{\Q \VS{13}Their wealth
will be
stolen
\par }{\Q and their houses
ruined!
\par }{\Q They will not
live
in the houses
they have built,
\par }{\Q nor
will they drink
the
wine
from the vineyards
they have planted.
\par }{\Q \VS{14}The
{\ND{Lord}}’s
great
day
of judgment is almost
here;
\par }{\Q it is approaching
very
rapidly!
\par }{\Q There
will be a bitter
sound
on
the
{\ND{Lord}}’s
day
of judgment;
\par }{\Q at that time warriors will cry out in battle.
\par }{\Q \VS{15}That day
will be a day
of God’s anger,
\par }{\Q a day
of distress
and hardship,
\par }{\Q a day
of devastation
and ruin,
\par }{\Q a day
of darkness
and gloom,
\par }{\Q a day
of clouds
and dark skies,
\par }{\Q \VS{16}a day
of trumpet
blasts
and battle cries.

\par }{\Q Judgment will fall on
the fortified
cities
and the high
corner
towers.
\par }{\Q \VS{17}I will bring distress
on the people
\par }{\Q and they will stumble
like blind
men,
\par }{\Q for
they have sinned
against the
{\ND{Lord}}.
\par }{\Q Their blood
will be poured
out like dirt;
\par }{\Q their flesh
will be scattered like manure.
\par }{\Q \VS{18}Neither
their silver
nor their gold
will be able
to deliver
them
\par }{\Q in the day
of the
{\ND{Lord}}’s
angry
judgment.
\par }{\Q The whole
earth
will be consumed
by his fiery
wrath.
\par }{\Q Indeed,
he will bring
terrifying
destruction
on all
who live
on the earth.”

\par }\Chap{2}{\PP \par }{\Q \VerseOne{1}Bunch
yourselves together like straw, you undesirable
nation,
\par }{\Q \VS{2}before
God’s decree
becomes reality
and the day
of opportunity disappears
like windblown chaff,
\par }{\Q before
the
{\ND{Lord}}’s
raging
anger
overtakes you –
\par }{\Q before the day of the
{\ND{Lord}}’s angry judgment overtakes you!
\par }{\Q \VS{3}Seek
the
{\SC{‘span class=”s H3068” ’Lord}}’s
favor, all
you humble
people of the land
who
have
obeyed his commands!

\par }{\Q Strive
to do what
is right!
Strive
to be humble!

\par }{\Q Maybe
you will be protected
on the day
of the
{\ND{Lord}}’s
angry judgment.
\par }{\SH Judgment on Surrounding Nations
\par }{\Q \VS{4}Indeed,
Gaza
will be
deserted
\par }{\Q and Ashkelon
will become a heap of ruins.
\par }{\Q Invaders will drive
away the people of Ashdod
by noon,
\par }{\Q and Ekron
will be overthrown.
\par }{\Q \VS{5}Those who live
by the sea,
the people
who came from Crete,
are as good as dead.

\par }{\Q The
{\ND{Lord}}
has decreed your downfall, Canaan,
land
of the Philistines:
\par }{\Q “I will destroy
everyone who lives there!”
\par }{\Q \VS{6}The seacoast
will be
used as pasture
lands
by the shepherds
\par }{\Q and as pens
for their flocks.
\par }{\Q \VS{7}Those who are left
from the kingdom
of Judah
will take possession
of it.

\par }{\Q By the sea they will graze,
\par }{\Q in the houses
of Ashkelon
they will lie
down in the evening,
\par }{\Q for
the {\ND{Lord}}
their God
will intervene
for them
and restore their prosperity.
\par }{\Q \VS{8}“I have heard
Moab’s
taunts
\par }{\Q and the Ammonites’
insults.
\par }{\Q They taunted
my people
\par }{\Q and verbally harassed
those living in
Judah.
\par }{\Q \VS{9}Therefore,
as surely as I
live,”
says
the {\ND{Lord}}
who commands armies,
the God
of Israel,
\par }{\Q “be certain that
Moab
will become
like Sodom
\par }{\Q and the Ammonites
like Gomorrah.
\par }{\Q They will be overrun
by weeds,
\par }{\Q filled with salt
pits,
\par }{\Q and permanently
desolate.
\par }{\Q Those of my people
who are left
will plunder
their belongings;

\par }{\Q those who are left
in Judah
will take possession of their land.”
\par }{\Q \VS{10}This
is how they
will be repaid
for their arrogance,
\par }{\Q for
they taunted
and verbally harassed
the people
of the
{\ND{Lord}}
who commands armies.
\par }{\Q \VS{11}The
{\ND{Lord}}
will terrify
them,

\par }{\Q for
he will weaken
all
the gods
of the earth.
\par }{\Q All
the distant nations
will worship
the
{\ND{Lord}} in their own lands.
\par }{\Q \VS{12}“You
Ethiopians
will also
die by my sword!”
\par }{\Q \VS{13}The
{\ND{Lord}} will attack
the north
\par }{\Q and destroy
Assyria.
\par }{\Q He will make
Nineveh
a heap of ruins;
\par }{\Q it will be as barren
as the desert.
\par }{\Q \VS{14}Flocks and herds
will lie down
in the middle
of it,
\par }{\Q as well as
every kind
of wild animal.
\par }{\Q Owls
will sleep
in the tops of
its support pillars;
\par }{\Q they will hoot
through the windows.
\par }{\Q Rubble
will cover the thresholds;
\par }{\Q even
the cedar work
will be exposed to the elements.
\par }{\Q \VS{15}This
is how the once-proud
city will end up –
\par }{\Q the city that was so secure.
\par }{\Q She thought to herself, “I am unique! No one can compare to me!”
\par }{\Q What a heap of ruins she has become, a place where wild animals live!
\par }{\Q Everyone who passes by her taunts her and shakes his fist.

\par }\Chap{3}{\PP \par }{\Q \VerseOne{1}The filthy,
stained
city is as good as dead;
\par }{\Q the city
filled with oppressors is finished!
\par }{\Q \VS{2}She is disobedient;
\par }{\Q she refuses correction.
\par }{\Q She does not
trust
the {\ND{Lord}};
\par }{\Q she does not
seek the advice of her
God.
\par }{\Q \VS{3}Her
princes
are as
fierce as
roaring
lions;
\par }{\Q her rulers
are as hungry as
wolves
in the desert,

\par }{\Q who completely devour
their prey by morning.
\par }{\Q \VS{4}Her prophets
are proud;
\par }{\Q they are deceitful
men.
\par }{\Q Her priests
defile
what is holy;
\par }{\Q they break
God’s laws.
\par }{\Q \VS{5}The just

{\ND{Lord}}
resides within
her;
\par }{\Q he commits
no
unjust
acts.

\par }{\Q Every
morning
he reveals his justice.
\par }{\Q At dawn
he appears without
fail.
\par }{\Q Yet the unjust
know
no shame.
\par }{\SH The Lord’s Judgment will Purify
\par }{\Q \VS{6}“I destroyed
nations;
\par }{\Q their walled
cities
are in ruins.
\par }{\Q I turned their streets
into ruins;
\par }{\Q no one
passes
through them.
\par }{\Q Their cities
are desolate;
\par }{\Q no one lives there.
\par }{\Q \VS{7}I thought, ‘Certainly
you will respect
me!
\par }{\Q Now you will accept
correction!’

\par }{\Q If she had
done
so,
her home
would not
be destroyed

\par }{\Q by all
the punishments I have threatened.

\par }{\Q But
they eagerly sinned
\par }{\Q in everything they did.
\par }{\Q \VS{8}Therefore
you must wait
patiently for me,” says
the {\ND{Lord}},
\par }{\Q “for the day
when I attack
and take plunder.
\par }{\Q I have decided to gather
nations
together
\par }{\Q and assemble
kingdoms,
\par }{\Q so I can pour out
my fury
on them –
\par }{\Q all my raging anger.
\par }{\Q For the whole earth will be consumed
\par }{\Q by my fiery anger.
\par }{\Q \VS{9}Know for
sure that I will then
enable
\par }{\Q the nations
to give me acceptable
praise.
\par }{\Q All
of them will invoke
the
{\ND{Lord}}’s
name
when they pray,

\par }{\Q and will worship
him in unison.
\par }{\Q \VS{10}From beyond
the rivers
of Ethiopia,
\par }{\Q those who pray to me
will bring
me tribute.
\par }{\Q \VS{11}In that day
you will not
be ashamed
of all
your rebelliousness
against me,
\par }{\Q for
then
I will remove
from your midst
those who proudly
boast,

\par }{\Q and you will never
again
be arrogant
on my holy
hill.
\par }{\Q \VS{12}I will leave
in your midst
a humble
and meek
group of people,
\par }{\Q and they will find safety in
the
{\ND{Lord}}’s
presence.
\par }{\Q \VS{13}The Israelites
who remain
will not
act
deceitfully.
\par }{\Q They will not
lie,
\par }{\Q and a deceitful
tongue
will not
be found
in their mouth.
\par }{\Q Indeed, they
will graze peacefully like sheep
and lie
down;
\par }{\Q no
one will terrify them.”
\par }{\Q \VS{14}Shout
for joy, Daughter
Zion!

\par }{\Q Shout
out, Israel!
\par }{\Q Be happy
and boast
with all
your heart,
Daughter
Jerusalem!
\par }{\Q \VS{15}The
{\ND{Lord}}
has removed
the judgment
against you;

\par }{\Q he has turned
back your enemy.
\par }{\Q Israel’s
king,
the {\ND{Lord}}, is in your midst!
\par }{\Q You no
longer
need to fear
disaster.
\par }{\Q \VS{16}On that day
they will say
to Jerusalem,
\par }{\Q “Don’t
be afraid,
Zion!
\par }{\Q Your hands
must not
be paralyzed from panic!
\par }{\Q \VS{17}The
{\ND{Lord}}
your God
is in your midst;
\par }{\Q he is a warrior
who can deliver.
\par }{\Q He takes great delight
in
you;
\par }{\Q he renews
you by his love;
\par }{\Q he shouts
for joy
over you.”
\par }{\Q \VS{18}“As for those who grieve
because they cannot attend
the festivals
–
\par }{\Q I took them away from
you;
\par }{\Q they became
tribute
and were a source of shame to you.
\par }{\Q \VS{19}Look,
at that time
I will deal
with those
who mistreated
you.
\par }{\Q I will rescue
the lame
sheep

\par }{\Q and gather together
the scattered
sheep.
\par }{\Q I will take away their humiliation
\par }{\Q and make
the whole
earth
admire
and respect them.
\par }{\Q \VS{20}At that time
I will lead
you –
\par }{\Q at the time I gather you together.
\par }{\Q Be sure of this! I will make all the nations of the earth respect and admire you
\par }{\PP when you see me restore you,” says the
{\ND{Lord}}.
\par }